### 2014-05-06 OLD_1main.tex
Initial commit

\documentclass[conference]{IEEEtran}

\usepackage[pdftex]{graphicx}
\usepackage[cmex10]{amsmath}
\usepackage{algorithmic}
\usepackage{caption}
\usepackage{subfig}
\usepackage{url}
\usepackage{balance}

\begin{document}
\title{What Can Stack Traces Tell us about the Use of Exceptions in Java?}

\author{\IEEEauthorblockN{Roberta Coelho\IEEEauthorrefmark{1}, Georgios Gousios\IEEEauthorrefmark{2}, Arie van Deursen\IEEEauthorrefmark{2}, Lucas Mariano\IEEEauthorrefmark{1}}
\IEEEauthorblockA{\IEEEauthorrefmark{1}Federal University of Rio Grande do Norte\\
Natal, Brazil\\
Email: \texttt{\{roberta, lucasm\}@dimap.ufrn.br}}
\IEEEauthorblockA{\IEEEauthorrefmark{2}Delft University of Technology\\
Delft, The Netherlands\\
Email: \texttt{\{gousiosg,a.vandeursen\}@tudelft.nl}}
}

\newcommand{\todo}[1]{\textbf{TODO}\footnote{\textbf{TODO:} #1}}

% make the title area
\maketitle

\begin{abstract}
Exception handling is one of the most used techniques for structuring the error recovery code of a system. Although exception handling mechanisms have been embedded in several main stream programming languages (e.g. Java, C++, C\#), studies have shown that the exception handling code is often poorly understood and the least tested parts of systems. As a consequence, the exception handling code may make the program more complex and even less reliable, leading to the so-called uncaught exceptions, one of the main causes of system crashes. Stack traces are a useful source of information to better understand such system crashes. In this work, we performed the first large scale analysis of Java stack traces and pinpoint how they can reveal bad programming practices. We mined and analyzed the stack traces embedded in all issues of Java projects available GitHub. Overall, our research set includes 16.837 projects and 58.490 stack traces. Some patterns of exception (mis)-use were consistently detected in this study, such as: most of the stack traces were related to coding errors such as null pointers; unexpected wrappings such as Errors being wrapped in checked exceptions and re-signaled; and undocumented runtime exceptions signaled by third party code.
\end{abstract}

%\IEEEpeerreviewmaketitle

\section{Introduction}

Modern applications have to cope with an increasing number of abnormal computation states that arise as a consequence of faults in the application itself (e.g., access of null references), noisy user inputs or faults in underlying midleware or hardware. The exception handling mechanism ˜\cite{goodenough1975exception} is one of the most used schemes for detecting and recovering from such exceptional conditions. Although the exception handling mechanism have been embedded in several programming languages (e.g. Java, C++, C\#) and the target of some studies [][][][], the exception handling code is often generally poorly understood and the least tested parts of systems.

Often some languages constructions may lead the developers into believing that by just re-throwing the exceptions they can forget about the exceptional situations during the development of the "happy path". This "ignore-for-now" approach may turn the exception handling can become a kind of generalized "goto" mechanism ˜\cite{mandrioli1992advances} making the program more complex and even less reliable. This behaviour may lead to the well known uncaught exceptions problem ˜\cite{jo2004uncaught}, one of the main causes of application crashes.

Moreover, some languages use exceptions to represent coding errors (out-of-bounds array index, division-by-zero, access to a null) in this cases exceptions are implicitly signaled by the runtime environment when such condition happens. One of such implicit exceptions (caused by data conversion from 64-bit floating point to 16-bit signed integer) was the responsible for the failure on Ariane 5's first test flight ˜\cite{lions1996ariane} - leading to the rocket self-destructing 37 seconds after launch and a loss of 500 million dollars. Besides Ariane 5, several applications crash everyday due to uncaught exceptions ˜\cite{jo2004uncaught}. In other words, the consequence of a careless exception usage may have the opposite consequence of its initial intention of improving system robustness.

In Java, when the program fails due to an uncaught exception the program automatically terminates and the system prints the stack trace which includes a detail message and the stack trace when the exception happened. A typical Java stack trace consists of an ordered list of methods that were active on the call stack before the exception has occurred. Stack traces are a useful source of information about system crashes, and is often used to support developers in debugging [ref:do stacks help]. Some works have been proposed to extract stack trace information from crash and bug reports and use stack trace information to support bug classification and clustering [][][], debugging tasks [][][] and the automated bug fixing [] task [ref], the prediction of fault proneness of classes [][].

However, all these works focus on the analysis of a single system at a time. In our current context, on which we have plenty access to the information available on open-source repositories like GitHub a research questions arise: What the stack traces available in these repositories can tell us about the use of exceptions in Java programs?

To answer this question, we conducted a sequential mixed-methods exploratory study [ref-mixed methods] which focused on mining the stack trace information available on all Java projects on GitHub repository, and combining this information with bytecode analysis, source-code analysis and manual inspection.

To guide our exploratory study we firstly compiled a subset general guidelines proposed by Johua, Bertrand Myer and Rebeca (see Section X), and then using the mixed-methods described above the investigated whether we could discover whether those practices have been followed or not on the analyzed projects.A tool called ExceptionMiner, was developed in this work to mine stack trace information from Github issues and combine this information with source code and bytecode analysis.

This study was divided in two phases: in the fist phase the issues of all Java projects available on GitHub was analyzed and all available stack traces extracted using ExceptionMiner. In this first phase XXX projects were analyzed, overall containing XXXX issues, from which YYYY stack traces were extracted and used as a source of information for our analysis.

In the second phase of this study, we combined the information of the mined stack traces with sourcecode and bytecode analysis, and to do so we selected the subset of all Android applications availabe in GitHub, this subset is composed by 526 aplications (Sections X details how those aplications were selected and studied).

We choose the Android apps available GitHub mainly because this kind of application are known for dealing with several sources of exceptions: they deal with an io-intensive infrastructure (Network, Bluetooth, camera…), and their development model involve multiple threads (ICSE arie – android specification). Moreover they are based on a single platform which allow more specific analysis of the exception behaviour.

Some negative outcomes were consistently detected through this large scale analysis of exception stack traces in the wild, such as:

\begin{itemize}
\item  A multitude of NullPointerExceptions and JVM implicitly thrown exception was found on stack traces. approximately 80\% of all stacks were related to this kind of exceptions.
\item Odd exception chains were detected: ERROR  wrappred in a checked/runtime, chacked wrapped on Runtimes and ret-thrown.
 \item  Undocumented runtime exceptions thrown by  libraries and causing crashes.
\item the use of general exceptions with no semantic associated to its type - X\% of the stack traces were caused by pure Runtime exceptions.
\item  size of stack traces whose size contained aproximately 100 method calls. Which makes the handling difficult since contextual information is lost.
\end{itemize}

During this study we implemented an stack exception mining tool which was very useful when finding and characterizing these outcomes. The contributions of this study are as follows:
\begin{itemize}
\item  It performs the first large scale analysis of Java stack traces and how they can reveal bad-practices. 
\item  It briefly presents ExceptionMiner, a tool developed to support the analysis.
\end{itemize}

The contributions of this work allow for developers of robust Java applications to familiarize with the most common reported exceptions and help developers to avoid making them; (ii) designers of  languages to consider ways of reducing the abundance of NullPointerExceptions, (iii) and tool designers to consider developing tools that enable a cross-project defect analysis. Specially in the open-source environment faults are not to be hidden in a private bug issue,faults should be share and discussed, so a developer can learn from other projects mistakes, and currently the search facilities of repositories are very limited. The ExceptionMiner tool is the fist step in this direction.

The remainder of this paper is organized as follows. Section 2 presents a background on error handling mechanisms and Java exception handling constructs and practices. Section 3 defines the hypotheses and configuration of our exploratory study, the target applications and the evaluation procedures. Section 4 reports our analysis of the empirical data collected in this study. Section 6 provides further discussions and lessons learned. Section 7 describes the related work. Finally, Section 8 presents our conclusions and directions for future work.

%PS: OLD CONTENT OF THIS SECTION WERE REMOVED TO THE END of THIS PAPER.

\section{Background}

\subsection{Error handling in various programming languages}

Modern applications have to cope with an increasing number of abnormal computation states that arise as a consequence of faults in the application itself (e.g., access of null references), noisy user inputs or faults in underlying middleware or hardware.

Currently, the exception handling mechanism (Goodenough, 1975), imbedded in most of the main stream programming languages, is one of the most used schemes for detecting and recovering from such exceptional conditions.

\noindent\emph{Return codes}
The main disadvantage of return codes if the fact that the programmer can just ignore the return code and let the program execute in inconsistent state.

\noindent\emph{Exceptions}
 Exception Handling mechanisms are embedded in many mainstream programming languages, however they differ in the way exceptions are represented and handled in each language (Garcia et al). Considering for instance the need to declare exceptions on the method signatures, several approaches can be applied. In languages such as Lore, C++, and Arche fall although all exceptions are unchecked, the developer can optionally listing exception types on method signatures. In the other hand in Modula-3, Guide, and Extended Ada the programmer must explicitly list every exception that can be signaled. Hence, all exceptions are checked exception. Java is the only language that introduced a hybrid approach. In Java, a method can throw either an instance of RuntimeException or Error, which are both unchecked and not need to be listed on the methods signature or an instance of Exception which is therefore a checked exception and must be listed in the method signature.

\noindent\emph{Monads}
    In functional programming languages, side effects are undesirable and in
    the case of pure functional languages (e.g. Haskell) impossible. To encapsulate and 
    isolate side effects, such as exceptions, pure functional languages use monads. Monads 
    allow for composition of chains of computations that might throw recoverable exceptions 
    in a way that if an exception occurs, subsequent computations are abandoned and the 
    exception is returned as the computation result. In effect, monads wrap exceptions into
    values. An example of an exception handling monad is the \texttt{Try[T]} type in Scala.

\subsection{Exception handling in Java}

In Java, an exception is represented according to hierarchical structure illustrated in Figure 1  (Gosling et al., 1996). According to this structure every exception is an instance of the Throwable class, and can be of three kinds: checked exceptions (extends Exception), run-time exceptions (extends RuntimeException), and errors (extends Error).

\begin{figure}
\centering
\includegraphics[scale=0.5]{exception_hierarchy.png}
\caption{Exception Hierarchy in Java}
\label{fig:Exception Hierarchy in Java.}
\end{figure}

Checked exceptions must be declared as part of the method signature that may throw it. The use of checked exceptions allows the compiler to statically check that appropriate handlers are provided within the system. The use of checked exceptions, however, is costly to maintain (Dooren and Steegmans, 2005) since every method on the call chain of a method that raises a new exception should be updated to declare this exception (on the throws clause defined on its signature) or handle it.

On the other hand, both runtime exceptions and errors (which are also referred as unchecked exceptions) do not need to be declared on method signatures, but as a consequence there is very little that can be checked at compile time. The invalid operations in the program detected by the Java Virtual Machine are represented as runtime exceptions (e.g., NullPointerException). By convention an Error represent an unrecoverable condition which usually result from failures detected by the Java Virtual Machine, such as OutOfMemoryError - and should not be handled inside the application.

The client of a method cannot easily know which unchecked exceptions may be thrown by the method unless he/she carefully inspects the code of the method and the methods called from it – which can become a very time consuming or infeasible task. Moreover, the developer is not warned by the compiler if an unchecked exception is not handled inside the application. For that reason the Java specification suggests that most user-defined exceptions should be checked exceptions to take advantage of compile-time checks. However, there is a long-lasting debate about the use of checked and unchecked exceptions (links to debates). There are pros and cons associated to both uses.

\subsubsection{Exception Signaling and Handling}

In Java an exception can be signaled in one of the following circumstances [ref: java specification]:

\begin{itemize}
\item  explicitly signaled when a throw statement is executed;
\item  implicitly signaled by the JVM when the evaluation of an expression violates the normal semantics of the Java programming language.g., out-of-bounds array index, division-by-zero, access to a null reference; or 
\item implicitly signaled by the JVM due to an internal error or resource limitation (e.g., OutofMemoryError, StackOverflowError)
\end{itemize}

When an exception is signaled it causes a transfer of control from the point where the exception occurred to a point that can be specified by the programmer (exception handler). In Java it is represented by the try-catch block and is responsible for performing any recovery action necessary to bring the software system back to a normal state and, whenever this is not possible, to log the exception and abort the system in an expected safe way.

Exception Model. In Java, as in many languages, the search for the handler to deal with a raised exception occurs along the dynamic invocation chain. This is claimed to increase the software reusability, since the invoker of an operation can handle it in a wider context (Goodenough, 1975). If no handler in the chain is found the default handler associated to the Thread or ThreadGroup is invoked (see Java Specification for more detail) which usually terminates the system and print the stack trace on the console.

\subsubsection{Exception Wrapping}
A common handling technique for Java exceptions is wrapping. With wrapping, caught exceptions
are re-thrown by the catching method in the form of a more general exception type.
Exception wrapping is useful for restring the propagation of exception types towards
upper layer methods and for hiding the exception flow (and consequently the implementation)
of components from their users. Exception wrapping is implemented by special exception
constructor methods where wrapped exceptions are passed as arguments.

An example a multiple exception wrapping is depicted on the stack trace
 in Figure~\ref{fig:wrapping}. This stack traces was extracted from an issue of an Adroid application in Github.The bottom part of the stack trace is the \emph{root cause}, which indicates
the real reason the error was thrown (in this case, the computer run out of
memory). The top part of the stack trace indicates the location
of the exception manifestation; the stack trace is the result of no 
exception handling taking place at this part of the source code.
The intermediate blocks indicate how the exception was wrapped
while crossing API layers between the exception root location and the
exception manifestation. In all levels, the exception \emph{signaler}
method is the one that threw the exception.

\begin{figure*}
\centering
\includegraphics[scale=0.8]{wrappings}
\caption{Exception wrapping in Java stack traces}
\label{fig:wrapping}
\end{figure*}

\subsection{Exception Handling Best Practices}

Several general guidelines have been proposed on how to use exceptions in general [bertrand meyer] and in Java programs [specification-java] [Joshua] [Rebecca]. Although some of them may differ with respect to use of checked and unchecked exceptions, all them agree with the following:

\textbf{1. Exceptions representing coding errors should not be handled, then you should work to prevent them} [REBECA].  Since the implicit exceptions (out-of-bounds array index, divide-by- zero error, null pointer references) may occur in many points in the program  and recovery from them is usually difficult, they should not be handled inside the program. Otherwise a program handling such exceptions would be sprinkled with catches, and will simply be an irritation to programmers. Hence developers should use development and testing practices and tools in order to prevent them as much as possible.

\textbf{2. Throw exceptions appropriate to the abstraction.} Higher layers should catch exceptions thrown by lower-layers and in their place throw exceptions that can be explained in terms of the higher level abstraction. To do this the developer can perform an exception translation [ref] or exception chaining. Although this practice is superior to the carelessly propagation of exceptions it should not be overused. The caller should try to prevent, handle the exception and evaluate if the exception is appropriate to the higher level.

\textbf{3. Do not repeatedly re-throw exceptions, handle exceptions as close as possible to the problem} Although exceptions don’t cost anything until they’re raised, programs that frequently raise exceptions run more slowly. Moreover, close to the place that the exception was throw the caller usually have enough contextual information to perform  corrective actions. However, when the an exception is propagated far away from the method where it was signaled (i.e. further away from the problem) it is often be difficult to take meaningful recovery actions.

\textbf{4. Errors should not be handled.} They are reserved to the JVM to indicate resource deficiencies, invariant failures and other conditions that make impossible the program to recover.

\textbf{5. Use checked exceptions for recoverable conditions.} The developer should use checked exceptions for conditions from which the caller is expected to recover. By confronting the API user with a checked exception, the API designer is telling the client to handle the exceptional condition. <<<The client can explicitly ignore the exception (swallowing, or converting it to other kind) but it is usually a bad idea.>>>

\textbf{6. Document precisely all exceptions thrown by a method} This practice is particularly important for library methods [Joshua][rebeca]. The exceptions thrown by a method are an important part of its documentation required to use the method properly. As the checked exceptions should be part of the methods signature, the method caller is aware of the checked exceptions being thrown by a method. However, it is wise to document the explicitly-signaled unchecked exceptions as carefully as the checked exceptions (in the @throws tag in Javadoc). Doing so the clients of a method will be aware of the exceptions the method can throw. If the developer fails to do follow this practice (specially for library code) it will be difficult or even impossible for the caller to make effective use of such method.

These best practices guided our exploratory study described in the next sections. Based on information extracted from stack traces and on bytecode and source-code analysis, we investigated whether some characteristics could revealed whether these practices  have been obeyed. Next section discusses the study procedure and the practices' misuses that emerged from this study.

\section{Research Design}

We used the sequential mixed-methods approach, a procedure for collecting, analyzing, and integrating both quantitative and qualitative data at some stage of the research process within a single study for the purpose of gaining a better understanding of the use of exceptions in Java programs
based mainly on the information available on exception stack traces and source code. For specific research questions, we first explore the domain quantitatively, and then highlight interesting cases by exploring 
cases qualitatively.

Below, we present the research questions we investigate in this study and how we approach them:
\begin{description}

    \item[RQ1] What is the prevalence of stack traces on the reported issues of open-source Java programs?
    \item[RQ2] What are the most popular exceptions?
    \item[RQ3] What patterns of exception (mis)-use emerge from the stack trace analysis?

\end{description}

\textbf{RQ1} 
%What is the prevalence of exception-handling bugs on the reported issues of open-source Java programs? 
%This question should be answered in order to conduct this study. Hence if not enough exception stack traces %were reported we would not have enough information to investigate the use of best practices from stack trace information.
To assess the prevalence of exception handling bugs, we  mine and analyze statistically all issues 
from all Java projects in the Github repository.
In particular, we investigate such questions as 
what is the percentage of issues with stack traces and whether issues labeled by 
developers as defects include a stack trace.

\textbf{RQ2} To discover the most popular exceptions, we

\textbf{RQ3} Going in depth with analyzing the extracted
stack traces requires examining individual stack trace elements and classifying
them in categories such as Java-based exceptions, library exceptions or
application exceptions. Doing the classification on all stacktraces mined from Github
is not practical as one needs to know in advance the package names exported by the application
and those exported by the libraries the application uses.

To restrict the analysis domain, we focused our analysis on Android projects.
This decision was based on our experience
of analyzing stacktraces and was driven by the fact that most open source Android applications
make use of a rich, well-defined stack of {\sc api}s (Android) and a limited number of
shared external libraries. Moreover, Android applications are known~\todo{ref} for dealing with several sources of exceptions due to multithreaded execution and interaction with input/output resources.
%Since these applications share a common framework, and sometimes even libraries, we wanted to discover whether most of the application crashes were caused by exceptions on the application itself or exceptions signaled by the supporting framework or reused code - a study have show that there is a high degree of reuse between apps.

To answer research questions RQ2 and RQ3, we constructed a new dataset featuring
aand
analyzed it both statistically and using machine learning techniques (association rule 
mining) to

\textbf{RQ4} complete....

\section{Data and Tools}

We begun our investigation using the dataset provided by the GHTorrent project, 
an off-line mirror of the data offered
through the Github API. The Github API data come in two forms; a streaming data flow lists events, 
such as reporting an issue, happening on repositories in real time, while a static view contains 
the current state of entities. The data is stored in unprocessed format, in a MongoDB database, 
while metadata is extracted and stored in a MySQL relational database. The GHTorrent dataset covers
a broad range of development activities on Github, including issues. The project has been collecting data since February 2012. Up to Dec 2013, when we queried it to retrieve our bootstrapping 
sample, the project had collected XXX issues from YYY Java projects.

Issues on Github are different from issues on dedicated bug tracking tools such as Bugzilla
and Jira. The most important difference is that there are no predefined fields for 
bug reporters to select from (e.g. severity and priority); Github uses instead a more open
ended tagging system. Repositories are offered a pre-defined set of labels, but repository
owners can modify them at will. An issue can have an arbitrary set of labels attached.
This meant that our analysis could not be based on labels alone 
Moreover, issue and pull requests are dual on Github; all pull requests have a corresponding 
``backing'' issue. Therefore, we needed to exclude pull request based issues from our analysis.
Finally, the Github API allows the automated generation of issues for specific repositories, 
which automated tools often use to report crashes. In some cases, this led to high number
of issues that included stack traces (e.g. ). We identified and filtered those cases out.

After obtaining the bootstrap issue sample from GHTorrent, we analyzed it in the following
ways:

\begin{itemize}

    \item We obtained the issue labels for all issues. We were especially interested in 
    issues sporting the default ``defect'' label.

    \item We obtained project information for each issue, separated Android projects from
    the general population of Java projects and analyzed their source code (Section~\ref{sec:android}).

    \item We extracted stacktraces embedded in the issue body and classified the exceptions 
    (Section~\ref{sec:datasetcons}) using the custom ExceptionMiner tool
    (Section~\ref{sec:exceptionminer}).

    \item We qualitatively analyzed specific stack traces (Section~\ref{sec:qualitative}).

\end{itemize}

\subsection{Analysis of Android projects}
\label{sec:android}

%In the second phase of this study we selected a subset of Java applications to be analyzed in more
%detail.The subset chosen was comprised by the Android apps included on GitHub (until 23 Feb 2013).
To identify Android projects, we performed a case insensitive search for the 
term \textsf{android} in the repository's names and short descriptions.
The heuristic filtered XXX repositories on which XXX issues were reported. 
We performed further cleanup, inspecting the site of every Android app analysed in this 
study to make sure that they were indeed mobile apps. X apps were removed from the
set because although having Android on its description or name they were not Android apps (e.g. 
\textsf{selendroid} -- a tool for testing android apps).

This subset consisted initially in 2.542 Android apps, from which 589 apps had at least one issue containing a stack trace. From this set 443 were analyzed in this study. \todo{How did we select those?}

\todo{What other analysis did we do? Extract packages from source code?}

\subsection{Extracting exceptions with ExceptionMiner}
\label{sec:exceptionminer}

To extract exceptions from issues, we implemented ExceptionMiner, a modular mining tool
able to connect to various repositories (such as Google Code, GHTorrent or even directly to
Bugzilla), extract issues, mine stacktraces from them and classify them in predefined categories.
The ExceptionMiner works as follows.

\noindent\emph{Stacktrace Miner} The first step in the process is mining references to exceptions and 
stack traces embedded in issues, distilling the information that composes a stack trace and storing
the results in a relational database. The extracted stack trace information  the exception signaler, the root cause, the exception wrappers. The tool is based on a combination of regular expressions in with a parser\todo{What kind of parser? Need more info...}. In contrast to existing issue parsing 
solutions such as Infozilla, the parser created in this work can extract all causes related to an 
exception on a stack, and stack traces embedded in logs files.

%\begin{tiny}
%\begin{verbatim}

%03-01 W/dalvikvm(7924): threadid=1: thread exiting with uncaught exception (group=0x40adf210)
%03-01 15:55:01.609 (7924): FATAL EXCEPTION: main
%03-01  (7924): java.lang.RuntimeException: ...
%E/AndroidRuntime( 7924):  at android....performLaunchActivity(ActivityThread.java:1967)
%03-01 15:55:01.609 (7924):  at android....handleLaunchActivity(ActivityThread.java:1992)
%03-01 15:55:01.609:  at android.app.ActivityThread.access\$600(ActivityThread.java:127)
%E/AndroidRuntime( 7924):  at android....handleMessage(ActivityThread.java:1158)
%\end{verbatim}
%\end{tiny}

\noindent\emph{Exception Classifier} The next step in the ExceptionMiner process is classifying exceptions as either checked exceptions, runtime exceptions or an errors. The exception classifier
is based on bytecode analysis (based on ASM [REF]) that walks up the type hierarchy of a Java exception until it reaches a base exception type.

\noindent\emph{Exception Singaler Classifier} The final step is finding and classifying 
information about exception signalers. ExceptionMiner is initialized with an arbitrary 
set of packages for each classification category and pattern matches the signaler name
to classify it to one of the categories.

\subsection{Dataset Construction}
\label{sec:datasetcons}

We applied the ExceptionMiner tool on the whole set of issues in our bootstraping sample
and then independently on the set of issues from Android projects. 
In the first case the exception classifier was initialized with all the exceptions defined on
the Java SDK version 7, while the signaler classifier was configured with Java's package
names.
In addition to the above, for the Android dataset analysis, the exception classifier
was initialized with the exceptions of the Android 
platform (level X), and all the exceptions of the 443 Android projects selected for this study.
The signaler classifier was initialized with information
about all packages that compose Java Environment, all packages composing all versions of Android Platform (Operating System and Framework), all packages that compose the core libraries reused by the Android API and all packages of 443 out of the 529 mobile apps hosted in Github which defined at least one issue that contained a stack trace \todo{this list of all packages can be found at [site].}.
What we could not classify in one of the above categories was automatically labeled as a library.

\section{Study Operation and Results}
This section presents the results for each of the study phases, providing both a quantitative and qualitative analysis of the outcomes. Firstly (Section X) we present the results of inspecting all Stack Traces  of Java Programs considered in this study. The following section focuses....

%\subsection{Phase 1: Analyzing Pure Stack Trace Information of Java Programs}

\subsection{RQ1: What is the prevalence of exception-handling bugs on the reported issues of open-source Java programs?}

\input{stacktrace-stats}

\begin{table}
\centering
\begin{tabular}{p{1em}lrr}
\hline
 \bfseries{Metric} & & \bfseries{Java} & \bfseries{Android}  \\
\hline
 \multicolumn{2}{l}{\bfseries{Repositories}} & 16,837 & 2,542\\
    & incl. exception on 1+ issues & 4,777 (28,37\%) & 729 (28,68\%)\\
    & incl. stack trace on 1+ issue & 3,759 (22,33\%) & 589 (23,17\%)\\
    & incl. exception on 1+ issue labeled as defect & 2,020 (41,44\%) &  329 (47\%)\\
    & incl. stacktrace on 1+ issue labeled as defect  & 1,699 (34,86\%) & 293 (41,86\%)\\
 \multicolumn{2}{l}{\bfseries{Issues}} & 393,552 & 50,342\\
    & incl. exceptions & 54,113 (13,75\%) & 4,947 (9,83\%) \\
    & incl. stacktraces & 44,886 (11,41\%) & 3,693 (7,34\%) \\
    & labeled as defect & 55,228 & 8,751 \\
    & labeled as defect and incl. exception & 6,741 (12,21\%) & 1,719 (19,64\%) \\
    & labeled as defect and incl. stacktrace &  5,196 (9,41\%) & 1,508 (17,23\%)\\
\multicolumn{2}{l}{\bfseries{Exception Stack Traces}} &  & \\
    & on issues  & 58,490  & 4,714  \\
    & on issues labled as defect & 6,529 & 1,813 \\

\hline
\end{tabular}
\caption{Quantitative overview of the analyzed dataset}
\label{tab:overview}
\end{table}

%\begin{figure*}
%        \centering
%        \subfloat[Java\label{fig:stacks-java}]{
%                \includegraphics[scale=0.4]{java-stacks}
%        }
%        \subfloat[Android\label{fig:stacks-android}]{
%            \includegraphics[scale=0.4]{android-stacks}
%        }
%        \caption{Distribution of issues containing stack traces as 
%        percentage of total issues in Java (a) and Android (b) projects}
%        \label{fig:stacktraces}
%\end{figure*}

The use of the Github issue tracker in repositories hosted on Github is optional
and not widespread. 
As of March 2014, from the 332,864 non-fork Java repositories registered in the 
GHTorrent dataset, only 44,323 have received an issue report in their lifetime
while from those only 16,837 are being watched by external users.
To ensure that our work targets projects that are openly used by 
users other than their developers, we focused our analysis on just those
projects. Using the heuristics presented in Section~\ref{sec:postprocessing}, 
we identified 2,542 repositories featuring Android projects.
Moreover, a typical issue on Github sports a collection of labels. Labels
can be user specified, but it is common for repositories to use the default set
provided by the Github issue tracker. Consequently, we used the ``defect'' label
as a marker for issues that might contain stack traces.
Table I and Table~\ref{tab:overview} presents two views of the extracted dataset,
namely grouped by repository and grouped by issue.

At the repository level, issues that include exceptions and stack traces
are fairly common: 28\% and 22\% of the repositories include at least
one issue with either of the two. At the individual issue level, 
11\% and 13\% contain exceptions. The numbers are similar if we
isolate Android projects from the general population. This means
that focusing on Android, as we do in the next section, does not affect the 
generalizability of results at least at the sample representativeness  
level. The ``defect'' label adds context to the analyzed issues in the
Android case, as almost 20\% of the issues marked as defects include
a stack trace. This is not true for the general case however. We therefore
did not use this label for further filtering.

The results in Table~\ref{tab:overview} are cumulative; however, the numbers
are not the same among all projects. Figure~\ref{fig:stacktraces} 
presents boxplots of the range of values of the variables described in  
Table~\ref{tab:overview} across all projects in out dataset. 
There we can see that indeed there is significant variation among projects,
while the outliers lay at the bottom end of the spectrum, which means that for
some projects, we can find very limited number of stacktraces in their 
issue reports. Both Java and Android projects feature similar variation,
which again supports our Android focused analysis later on.

\noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
RQ1: 28\% of Java repositories on Github have issues that include
stacktraces; 11\% of all the issues in those repositories include
stacktraces. The Android issue sample is more homogenized.
}
\end{minipage}}

% \begin{itemize}

%     \item What are the most ``popular'' exceptions and what they can tell us?

%     \item Which elements are triggering most of the exceptions and what they 
%     can tell us about the exception structure?

%     \item What is the role of libraries and the underlying platform on crashes?

%     \item What is the role of the application itself on crashes?
% \end{itemize}

% \subsection{The rules of using checked and unchecked exceptions are being obeyed in mobile environment?}

\subsection{RQ2: What are the most popular exceptions?}

What are the main causes of exceptions on reported issues? To answer this question we mined the following information of every extracted stack trace: the top level exception being signaled, its signaler (the first method signature frame after the exception), the root exception (the exception which initiated the stack trace, defined on the bottom of the stack), and the root signaler. The intermediate elements that compose a stack trace as shown in Figure 1 (i.e., the intermediate causes and its execution stacks) were not be considered in the analysis.

Table~\ref{tab:toptenpopular} illustrates the top 10 most signaled exceptions - which represents X\% of the analyzed set. The table illustrates  (i) the number of times each exception appeared as the root cause of a stack trace (\#occurrences), (ii) the number of distinct projects reporting such causes on stack traces, and (iii) the exception popularity. The exception popularity metric was defined in this work as follows: (popularity = number of projects mentioning the exception in at least one stack/all projects that contain at least one stack).

\begin{table}
\begin{tabular}{lrr}
    \hline
    \bfseries{Exception} & \bfseries{Num Occurrencies} & \bfseries{Num Projects} \\
    \hline
      java.lang.NullPointerException &	13180 &	1671 \\
      java.lang.IllegalArgumentException &	1765 &	670 \\
      java.lang.ClassNotFoundException &	1166 &	506 \\
      java.lang.IllegalStateException &	11408 &	431 \\
      java.lang.RuntimeException &	1158 &	411 \\
      java.lang.ClassCastException &	886 &	408 \\
      java.lang.NoSuchMethodError &	838 &	370 \\
      java.lang.ArrayIndexOutOfBoundsException &	807 &	334 \\
      java.io.IOException &	646 &	316 \\
      java.lang.NoClassDefFoundError &	557 &	299  \\
      java.lang.OutOfMemoryError &	597 &	235 \\
    \hline
  \end{tabular}
\caption{Top 10 most popular exceptions on issues}
\label{tab:toptenpopular}
\end{table}

\textbf{TO DISCUSS: We can include a longer list order by the percentage of ocurrences until we reach 80\% of the 
exceptions. It means top 20 and includes some exceptions signaled by android.}

\textbf{TO DISCUSS: We can also include the top 10 excepption packages...}

We could observe that the NullPointerException was the one with the higher number of occurrences (22,5\% of all stacks were caused by NullPointerExceptions) and the one which had the highest popularity. We could also observe that most of the exceptions in the TOP 10 list (except java.lang.IllegalArgumentException, java.io.IOException and java.lang.RuntimeException) are exceptions implicitly signaled by Java Virtual Machine when an expression violates the normal semantics of the Java programming language (e.g. ArrayIndexOutOfBounds) or due to an internal error or resource limitation (OutOfMemoryError).

The high prevalence of NullPointerExceptions, and the other implicitly-thrown exceptions  is aligned with the findings other works [Sunghun et al.][Gordon fraser and][Jcrasher].[Android-predict]. For instance, Sunghun et al. showed that in Eclipse bug report system 38\% the bugs related to exception handling were caused by NullPointerException; other works on robustness testing [JCrasher, Evol;suite, Amiya] showed that most of the automatically detected bugs were due to NullPointerExceptions and implicitly-signaled of Java environment (as the ones found in this study).

To better understand the main causes of such exceptions we evaluated specification and semantics of the top 100 root exceptions reported on stack-issues, and classified them according to the the categories presented in Table~\ref{tophundrend}.

\begin{table}
  \centering
  \begin{tabular}{lrr}
    \hline
    \bfseries{Category} & \bfseries{Occurrences} & \bfseries{Num Projects} \\
    \hline
      java.lang and util programming logic & 40857	(80,59\%) &	4874 (52,62\%)
      \\
      Resources(NET, IO, FILE) & 4445 (8,77\%) &	2121 (22,90\%)\\
      Security & 351 (0,69\%) &	132	(1,43\%) \\
      Concurrency	& 633 (1,25\%) &	290	(3,13\%) \\
      Backward compatibility & 2112 (4,17\%) &	948	(10,23\%) \\
      Reflection &	299	(0,59\%) &	195	(2,11\%) \\
      Specific (GUI,FRAMEWORK) &	600	(1,18\%) &	153	(1,65\%)\\
      General (Error, Exception, Runtime)	& 1401 (2,76\%) & 550	(32,91\%)\\
    \hline
  \end{tabular}
  \caption{Top 100 exceptions were studied and classified according to its semantics in the following categories.}
  \label{tab:tophundrend}

\end{table}

As we can see from Table 4, the that programming errors are responsible for more than 80\% of the issues reporting a stack trace. Such high prevalence points to the fact that developers may not have been using tools [ref: tools to prevent null...] and practices to prevent from such mistakes or that it may be difficult to do so (specialy in the case of NullPointerExceptions).

\section{What kinds of exception (mis)-use emerge from stack trace analysis? }

To answer this question we analyzed information extracted from stack traces in combination with source code and bytecode analysis. As mentioned before (Section X) in the second phase of this exploratory study we focused on 443 Android projects. We downloaded the latest master version of each Android app, extracted its packages, and the user defined exceptions and run a bytecode analysis tool to classify exception mentioned on at least one stack trace as Runtime, Checked, or Error.

Since we downloaded the latest version of each app, when an exception had been removed from the latest version we used the search facilities of Github to locate the exception code. The ones that could not be found, even through this search interface, were remained as UNDEFINED and represented X\% of the exceptions being analyzed. Besides downloading and analyzing the code of every user defined exception, we also download and analyzed the exception defined on reused libraries mentioned on stacks and every exception defined on Java Environment and Android development kit and android platform. After the massive mining effort we could come up with the findings presented in Table~\ref{typeroottab} and Figure~\ref{typeroot}.

Next sections details what patterns of exception (mis)-use emerge from this analysis.

\subsection{A multitude of NullPointerExceptions and other implicitly signaled exceptions signaled by Third Party}

After extracting every package related to each exception signaler (see Section III-B) we could classify every stack trace based on its root signaler as presented in Table~\ref{signalers}.

\begin{table}
  \centering
  \begin{tabular}{rp{22em}}
    \hline
    \bfseries{Type of signaler} & \bfseries{Description} \\
    \hline
    \bfseries{android} &	When the exception signaler is a method defined in Android Platform.\\
    \bfseries{app} &	If the exception signaler was defined inside an application class.\\
    \bfseries{libcore} &	If the application was signaled by one of the libraries reused by Android Platform. The complete list of core libraries can be found here\\
    \bfseries{lib} &	If the exceptions was not defined by any of the elements above.\\
    \bfseries{java} &	If the application signaler and all the methods that composes the exception are defined on java. packages\\
    \hline
  \end{tabular}
  \caption{Sources of exceptions in Android}
  \label{tab:signalers}
\end{table}

For each exception we could discover: the number of times it occurred on stacks, the number of distinct projects such stacks were defined (and consequently its popularity) and the of times it was signaled by different signaler types. We could calculate the popularity of each root cause, considering the number of distinct projects they were found and the whole set of projects analyzed with at least one stack trace (443). Table ~\ref{overviewandroid} and Figuere~\ref{androidsignaler} presents the mined data.

\begin{table*}
  \centering
  \begin{tabular}{rcccccccc}
    \hline
    \bfseries{Root exception} & \bfseries{Projects} & \bfseries{Popularity} & \bfseries{Occurences}	& \texttt{android} & \texttt{libcore} & \texttt{app} & \texttt{lib} & \texttt{java} \\
    \hline
    java.lang.NullPointerException	& 238	& 53,72\%	& 1244 (X\%)	& 461	& 18	& 506	& 257	& 2 \\
    java.lang.IllegalArgumentException & 100 & 22,57\% & 275  (X\%) & 166 & 3 & 52 & 54 & 0 \\
    java.lang.RuntimeException & 96 & 21,67\% & 251 (X\%) & 182 & 0 & 46 & 23 & 0 \\
    java.lang.IllegalStateException & 93 & 20,99\% & 250 (X\%) & 164 & 10 & 30 & 45 & 1 \\
    java.lang.OutOfMemoryError & 50 & 11,29\% & 166 (X\%) & 117 & 13 & 8 & 23 & 5 \\
    java.lang.IndexOutOfBoundsException & 48 & 10,84\% & 135 (X\%) & 51 & 0 & 63 & 20 & 1 \\ 
    java.lang.NumberFormatException & 36 & 8,13\% & 114 (X\%) & 17 & 4 & 81 & 8 & 4 \\
    java.lang.ClassCastException & 44 & 9,93\% & 113 (X\%) & 50 & 0 & 38 & 25 & 0 \\
    java.lang.NoClassDefFoundError & 40 & 9,03\% & 74 (X\%) & 9 & 0 & 27 & 37 & 1\\
    java.io.IOException & 32 & 7,22\% & 70 (X\%) & 14 & 7 & 28 & 20 & 1\\
    android.database.sqlite.SQLiteException & 28 & 6,32\% & 63 (X\%) & 63 & 0 & 0 & 0 & 0\\
    java.lang.SecurityException & 21 & 4,74\% & 63 (X\%) & 60 & 0 & 0 & 3 & 0\\
    java.lang.ClassNotFoundException & 40 & 9,03\% & 62 (X\%) & 46 & 0 & 3 & 13 & 0\\
    java.lang.ArrayIndexOutOfBoundsException & 30 & 6,77\% & 60 (X\%) & 15 & 0 & 27 & 18 & 0\\
    java.lang.NoSuchMethodError & 38 & 8,58\% & 58 (X\%) & 11 & 0 & 31 & 16 &
    0\\
    java.util.ConcurrentModificationException & 17 & 3,84\% & 54 (X\%) & 4 & 0 & 35 & 15 & 0\\
    \hline
  \end{tabular}
\caption{Root Exceptions occurrences and popularity.}
\label{tab:overviewandroid}
\end{table*}

\begin{figure}
\centering
\includegraphics[width=\hsize]{pic4.png}
\caption{Exception distribution per signaler type}
\label{fig:androidsignaler}
\end{figure}

We could also observe that the NullPointerException is still the exception with higher number of occurrences. The NullPoiterExceptions are mainly signaled inside Android platform and inside application code. Although we also find NullPointerExceptions being signaled from third-party libraries. Specially when dealing the reusable code, the methods should make explicit the contrat of use, and usually instead of rethrowing a NullPointerException, the resused method should rethrow an IllegalArgumentException, and document it on its signature. The same applies for android, since adroid platform offers a set of methods that will be reused by apps. Making such exception handling behaviour explicit, many of those bugs could be prevented/avoided.

Another observation was the high prevalence of RuntimeExceptions being thrown by the Android platform. We could observe that most of the RuntimeExceptions were signaled by Android Platform (182 out of 251). Such exception is to general that we can know what caused the exception if we read the exception message on the stack trace. Inspecting the code of Android Platform we could observe that is is almost an exception handling policy to convert different kinds of application (both checked and unchecked) in a RuntimeException. The code snipet below was extracted from Android platform level X and illustrate this:

If we focus on the top-level exceptions only, in other words, the ones wrapping the root exceptions. We can observe that the number of occurrences and the popularity of RuntimeException considerably increases when compared to the number of RuntimeExceptions being signaled as a root exception.

TO DECIDE: Do we need to include the same Table showing Top Level Exceptions occurrences and popularity???

Manually inspecting the code on which RuntimeExceptions  were signaled as root exceptions and also as top level exception we could observe that RuntimeExceptions are strongly used by Android to wrap several kinds of exceptions. This wrapping behaviour may happen for 2 main reasons:  to by-pass Java checked exception rules, or simply because there is no reasonable way of handling the exception, and hence regardless the kind of exception being signaled, there is only one kind of handling (present an error message and abort/restart the application). This  shades a light in the need of robustness and fault tolerant solutions to deal with exceptional situations other than just presenting a error window.

\textbf{We could observe in this study that the NullPointerException and other other implicitly signaled exceptions represented aproximatey 80\% of all exceptions reported on issues in both Java and Android subset.
For such exceptions - which represent programming bugs or resource limitations - there is usually no proper handling besides presenting an error message to the user and restarting the application - only high fault tolerant systems need to provide solutions to handle them [REF: Java Specification]. Hence such exceptions should be prevented through the adoption of tools and techniques instead of handled in runtime. Some research works have proposed NullPointer analysis tools to statically find such potential null references [ref-ref-daniele], or testing tools targeting this kind of bug or even program transformation approaches to prevent such null references [ref]. However, we can observe that in practice such strategies are not been used. Maybe because the developers are not aware of the extent of the problem – as this study revealed that most of the crash reports were related to NullPointerException. This finding points to a broader discussion on the use of exception and how difficult is to prevent NullPointerExceptions in Java detailed in Section X.}

\subsection{Uncaught Checked Exceptions}
\begin{table}
\centering
\includegraphics[width=\hsize]{table7.png}
\caption{Types of Root Exceptions.}
\label{fig:typeroottab}
\end{table}

\begin{figure}
\centering
\includegraphics[width=\hsize]{pic2.png}
\caption{Distribution of Checked and Unchedked...}
\label{fig:typeroot}
\end{figure}

We could observe for instance that most of the exceptions reported were RUNTIME xxxx, and most of such exceptions were signaled by ANDROID environment and the application itself. <complete>

If we consider the top level exceptions can see from Table ~\ref{wrappingtop}.

\begin{table}
\centering
\includegraphics[width=\hsize]{table8.png}
\caption{Types of Top Level Exceptions.}
\label{fig:typetop}
\end{table}

\subsubsection{Undocumented Runtime Exceptions signaled by Third-Party code}
We could observe that most of the exceptions signaled by Android libcore where mostly checked exceptions, so they define a precise exception behaviour to its clients. However, most of the exceptions signaled by other third-party libraries reused by apps signaled mostly Runtime exceptions.

We did a sampling study on the signaler which explicitly signaled Runtime exceptions. Our goal was to observe the whether the runtime exceptions were as carefully documented as the checked exceptions signaled by the methods. In our analysis none of the inspected methods included a documentation to it on Javadoc annotation or in the throws declaration. The lack of care developers have while documenting Runtime exceptions is aligned with the results of other studies such as the one from Cabral and Marcques who observed that ..... [ref: Cabral at al]

The undocumented RuntimeExceptions is indeed a serious problem once it makes very difficult or even impossible for the  client method to protect against such undocumented runtime exceptions.

\textbf{When libraries are used, the developer does not have access to their source code and thus needs to rely on the library documentation about the runtime exceptions that should be thrown – which, more often than not, are neither complete nor precise (Thomas, 2002; Sacramento et al., 2006). 
As a consequence, unhandled unchecked exceptions can be seen as one of the major sources of bugs in current Java systems (Jo et al., 2004). }

\subsubsection{Odd Wrappings}
The difference on the types of top level to root exceptions stimulated us to inspect the kinds of wrappings that happened on stacks based on the kinds of exceptions being wrapped.  Table ~\ref{wrappingandroid} presents the wrappings found in this study. As we can see this analysis revealed unexpected wrappings such as CHECKED wrapping an  ERROR; ERROR wrapping a CHECKED exceptop; RUNTIME - ERROR; ERROR wrapping RUNTIME.

\begin{figure}
\centering
\includegraphics[width=\hsize]{table9.png}
\caption{Kinds of wrappings found on the stacks of 443 ANDROID projects found on Github}
\label{fig:wrappingandroid}
\end{figure}

Java is the only language that provides a hybrid exception handling mechanism which offers different types to represent different exception behaviors. According to Java specification Errors represent XXXX and for that reason should not be handled most of the application since they represent. Checked exceptions represent situation from which a recovering is possible. Runtime exceptions represent.... Although the use of Runtime and Checked exceptions are not a consensus the meaning of each type of exception should be .... 
Since there is no way of enforcing such conventions during program development,  we could observe that, in practice, in some situations (illustrated by the Odd-Wrappings) these three types of exceptions were used interchangeably. This interchangeability makes the behaviour of the exception handling code more complex and less reliable. For instance, considering a situation were an OutOfMemoryError (a situation that should not be handled) was wrapped in a checked exception (a situation that must he handled), in this case this wrapping is telling to the caller that it should handle

One may think that these strange wrappings may be a characteristic of this analyzed Android apps. However, since some of the wrapping were related to static blocks and reflection library, and both also happen in other Java projects, we can foresee that these wrappings also happens. But to get a brief  notion on the size of this scenarios we re-run the analysis for the whole java set – reusing the same exception classification of the initial analysis. Since this set contains XX distinct exceptions from which XXX were previously classified in the previous study, we could observe more instances of UNDEFINED exceptions. We could actually see that these strange wrappings also happened in Java projects as illustrated in Table ~\ref{wrappingjava}. 
\begin{table}
\centering
\includegraphics[width=\hsize]{table10.png}
\caption{Kinds of wrappings found on the stacks on Java projects}
\label{fig:wrappingjava}
\end{table}

TO DISCUSS: We could also not include this table anymore, once Android was chosen lets stick to it until the end.
TO INCLUDE Table X: Kinds of wrappings found on the stacks of XXX Java projects found on github.

\subsubsection{Long Stack Traces}
One of the general guidelines on exception handling says that the developer should not repeatedly re-throw exceptions, (s)he should handle exceptions as close as possible to the problem. To investigate this practice we analyzed the number of wrappings on each stack trace and also the number of frames composing them. The resuls of this analysis are presented in Table X.

\begin{table}
\centering
\includegraphics[scale=0.5]{tablesize.png}
\caption{Kinds of wrappings found on the stacks on Java projects}
\label{fig:wrappingjava}
\end{table}

Exception wrappings happened in approximately X\% of all stack traces. We could observe that in many cases (xxx) the number of wrappings was higher than XX.

Moreover, another interesting observation is related to the size of the stacks (i.e. the number of frames composing them) which sometimes exceeded 100 frames (in XXX stack traces) the corresponding execution trace was caused by one of the two reasons:  recursive calls - ;exceptions signaled by a method defined deep down in a reused framework. Every time a third party code is reused together with the desired functionality comes exceptions (checked or runtime) related to it. We could observe that XX\% of stack traces whose size was higher or equal 50 were stack traces associated to Runtime checked exceptions.

\section{Discussion}
\subsection{How can the developers protect their apps from crashing? General Guidelines to Protect your application from crashing}
A Fault is not something to be hidden, but to be shared and discussed, so even those who did not fail can learn from it
\subsubsection{Step 1: Understand your own failures}
The ExceptionMiner tool can provide interesting information for Java/Android developers concerting the following:
Perform a single-project crash analysis to discover: (i) hub-signaler; (ii) exception-authority; (iii) modules that are causing most of the crashes.

\subsubsection{Step 2: Learn from other systems failures}
In an environment were development data is hugely available ExceptionMiner makes use of this information to come up with interesting information....
- perform cross project crash analysis to discover:
-the exception interface of methods that have already caused crash in at least one system.
- Extract exception interface from stack traces defined on repos. Nost of them are Runtime and not documented. learn from other people's mistakes.
- which kinds of exceptions are causing most of the crashes - the ones developer should take care.
TODO: refers to paper DO NOT TRUST PROGRAMMERS DOCUMENTING RUNTIME …
- use this information to develop your own code.
A better support to understand and prevent such bugs is needed. Our works is
just a first step in de direction.

\subsubsection{Step 3: Exception-aware Software Development}
TODO: refer to Bertrand Meyer paper: If the developer does not no the contract of a method, it leads to defensive programming that makes the program more complex and becomes a source of errors.
TODO: best practicess.... Joshua
TODO: exceptions and contracts

\section{Threats to Validity}

\subsection{Access to Crash Logs}
\subsection{Data Extraction from Natural language Artifacts}

-Data cleaning removing invalid and duplicated issues.
-Not every crash information is specified on an issues....

\subsection{Construction, Internal and External Threats}

\section{Related Work}

In this section, we present works we believe are directly related to our own, distributed in four categories: (i) works that use the information available on stack traces; (ii) empirical studies on the usage of Java exceptions and its fault proneness; (iii) tools to extract stack traces information from natural language artifacts (i.e., issues and emails);  and (iv) empirical studies involving Android apps.

\textit{Analysis of Java Exception Usage.} Since the manual analysis of the Java exception-flow can easily become infeasible Robillard and Murphy ~\cite{} developed a tool called Jex that analyzes the flow of exceptions in Java source code. Based on java source code this tool performs dataflow analysis in order to find the propagation paths of checked and unchecked exception types. After this tool other tools have been proposed to support the static analysis of exception flows [Chen et al][Coelho et al]~\cite{}~\cite{}. The main limitation off all tools is the number of false positives inherent to static analysis solutions, which can lead to a high number of exception flows, specially if considering Java Enviroment exceptions and exceptions signaled from libraries. This was the fist of some tools proposed to statically analyzed the exception flow of exceptions. Cabral and Marques ~\cite{cabral2007exception} analyzed the exception handling code of 32 open-source systems, both for Java and .NET. The goal of their study was to identify how exceptions were handled in different categories of systems. They examined the exception handlers and the respective actions taken on them. They observed that the action handlers were very simple (e.g., logging and present a message to the user). In our work differs from previous works since in our study we tried to identify bad practices from the combined use of stack traces extracted from issues and bytecode and source code analysis. Finally, Reimer and Srinivasan ~\cite{reimer2003analyzing} listed a set of bad practices on exception handling that hinder software maintainability and robustness. Their differs from ours it was solely based on their own experience on working with Java enterprise applications and focus on source code examples of bad patterns on ways of handling exceptions such as exception swallowing, log verbosity.

\textit{Analysis and Use of Stack Trace Information.} Some works have investigated the usage of stack trace information on bug classification [][][], debugging [], automated bug fixing [], and prediction of class fault proneness[][]. Kim et al ~\cite{kim2011crash} use of crash graphs (an aggregated form of multiple stack traces available in crash reports) to detect duplicate crash reports and to predict if a given crash will be fixed. Dhaliwal et al ~\cite{} analyzed crash reports collected for Mozilla Firefox and proposed a crash grouping approach based on the Levenshtein distance. This work showed that a crash grouping approach can reduce bug fixing time in aproximately 5\%. Wang et al ~\cite{} proposes an approach to define correlated crash types automatically. Moreover it also provides a bug localization method to locate and rank files related to the bug described on a stack trace.
Schroter et al ~\cite{schroter2010stack} conducted an empirical study on the usefulness of stack traces for bug fixing. The study used the development data of Eclipse project and could observe that developers fixed the bugs faster when failing stack traces were included on bug issues. This study is aligned with the one performed by Bettenburg et al.  ~\cite{bettenburg2008makes} which evaluated what content could make a good bug report, and stack traces were mentioned as the second most important content for developers - just after the "steps to reproduce" information. Sinha et al ~\cite{sinha2009fault} proposed an approach for locating and repairing faults that are caused by JVM implicitly signaled exceptions related to coding errors such as NullPointerExceptions. This approach uses the information available on the stack trace to guide a dataflow analysis, to locate the source statement responsible for the exception. Kim at al. ~\cite{kim2013predicting} proposed an approach to predict the crash-proneness of methods based information extracted from stack traces and methods' bytecode operations. They used InfoZilla to extracted stack trace information from the bug reports of two open-source projects, i.e. AspectJ and Eclipse. They could observe that most of the stack traces were related to NullPointerException and other JMV implicitly thrown exceptions had the higher prevalence on the analyzed set of stacks.

\textit{Extracting Stack Traces from Natural language artifacts.} Nowadays many software vendors embed automatic crash reporting tools in their software systems. Hence, whenever a software crashes this tool sends a detailed crash report to its vendors. Moreover, there are also find third party software solutions specialized in bug reporting, most of them for the increasing marked of mobile apps ~\cite{BugSe14,BugSn14,Googl14,Acra14}. Hence most of the works described previously used the  information provided by crass reporting tools. Two works have proposed tools extract stack traces from natural language artifacts such as issues and emails: Infozilla [ref] is based on a set of regular expressions that extract an the set of frames related to a stack trace, the main limitation of this solution is that it is not able to extract stack traces embedded on verbose log files (i.e., on which we can find log text mixed with exception frames). Alberto and Lanza [ref] propose a solution to recognize stack trace frames from development emails and relate it to code artifacts (i.e. classes) mentioned on the stack trace.

\textit{Empirical studies using Android apps.} Ruiz et al.~\cite{Ruiz12} investigated the degree of reuse across applications in Android Market, the study showed that almost 23\% of the classes inherited
from a base class in the Android API, and that 217 mobile apps were reused
completely by another mobile app. Pathak et al.~\cite{Patha11} analyzed bug
reports and developers discussions of Android platform and found out that
approximately 20\% of energy-related bugs in Android occurred after an OS
update. McDonnell et al.~\cite{McDon13} conducted a case study of the
co-evolution behavior of Android API and 10 dependent applications using the
version history data found in github. The study found that approximately 25\% of
all methods in the client code used the Android API, and that the methods
reusing fast-evolving APIs were more defect prone then others. Vásquez et
al.~\cite{Linar13} analyzed approximately 7K free Android apps and observed that
the last successful apps used Android APIs that were on average 300\% more
change-prone than the APIs used by the most successful apps. Pingyu and
Elbaum~\cite{Zhang12} analyzed bug reports of 5 Android applications an observed
that 29\% had to do with poor exceptional handling code.  Our work differs from
the others as it aims at distilling the information of bug reports describing
uncaught exceptions created for mobile applications in Github and Googlecode, in
order to get a first view of what is causing the crashes across applications
available and what the characteristics of the stack traces can tell us about the
exception structure of those applications. Moreover this work\ldots.

\section{Conclusion}
That's a great paper with profound insights.

Nowadays many software vendors embed automatic crash reporting tools in their software systems. Hence, whenever a software crashes this tool sends a detailed crash report to its vendors. Moreover, we can also find third party sofware solutions specialized in bug reporting for different kinds of systems specially for the increasing marked of mobile apps ~\cite{BugSe14,BugSn14,Googl14,Acra14}. There is planty of information to be mined...

\section*{Acknowledgment}
This work is partially supported by the NWO TestRoots project (639.022.314).

\bibliographystyle{IEEEtran}
\bibliography{android-stacks}

% that's all folks
\end{document}

### 2014-05-06 OLD_2main.tex
Initial commit

\documentclass[conference]{IEEEtran}

\usepackage[pdftex]{graphicx}
\usepackage[cmex10]{amsmath}
\usepackage{algorithmic}
\usepackage{caption}
\usepackage{subfig}
\usepackage{url}
\usepackage{balance}

\begin{document}
\title{What Can Stack Traces Tell us about the Use of Exceptions in Java?}

\author{\IEEEauthorblockN{Roberta Coelho\IEEEauthorrefmark{1}, Georgios Gousios\IEEEauthorrefmark{2}, Arie van Deursen\IEEEauthorrefmark{2}, Lucas Mariano\IEEEauthorrefmark{1}}
\IEEEauthorblockA{\IEEEauthorrefmark{1}Federal University of Rio Grande do Norte\\
Natal, Brazil\\
Email: \texttt{\{roberta, lucasm\}@dimap.ufrn.br}}
\IEEEauthorblockA{\IEEEauthorrefmark{2}Delft University of Technology\\
Delft, The Netherlands\\
Email: \texttt{\{gousiosg,a.vandeursen\}@tudelft.nl}}
}

\newcommand{\todo}[1]{\textbf{TODO}\footnote{\textbf{TODO:} #1}}

% make the title area
\maketitle

\begin{abstract}
In this work, we perform the first large scale analysis of Java stack traces and pinpoint how they can reveal bad programming practices. We mined and analyzed the stack traces embedded in all issues of Java projects available GitHub. Overall, our research set includes 356,057 issues define at 16,836 projects (from which 28,800 stack traces were extracted). From this set, the stack traces of 482 Android projects were investigated in more detail, in combination with source code and bytecode analysis. In this study some patterns of exception (mis)-use were consistently detected such as: unexpected wrappings (e.g., Errors being wrapped in checked exceptions) - revealing that Java hybrid exception model is not fully used according to its purpose;  undocumented runtime exceptions signaled by third party code - which makes almost impossible for library clients to protect against such exceptions; and a high prevalence of java.lang exceptions reported on issues - representing approximately 50\% of the analyzed issues.

\end{abstract}

%\IEEEpeerreviewmaketitle

\section{Introduction}

Modern applications have to cope with an increasing number of abnormal computation states that arise as a consequence of faults in the application itself (e.g., access of null references), noisy user inputs or faults in underlying middleware or hardware. The exception handling mechanism~\cite{goodenough1975exception} is one of the most used schemes for detecting and recovering from such exceptional conditions. Although the exception handling mechanism have been embedded in several programming languages (e.g. Java, C++, C\#) and the target of several studies (e.g.~\cite{miller1997issues,robillard2000designing,shah2010understanding,garcia2007extracting,garcia2001comparative,cabral2007exception,coelho2011unveiling}), the exception handling code is often generally poorly understood and the least tested part of software systems.

Often some languages constructions may lead the developers into believing that by just re-throwing the exceptions they can forget about the exceptional situations during the development of the "happy path". This "ignore-for-now" approach may turn the exception handling into a generalized "goto" mechanism~\cite{mandrioli1992advances} making the program more complex and even less reliable. This behaviour may lead to the well known uncaught exceptions problem~\cite{jo2004uncaught}, one of the main causes of application crashes.

Moreover, some languages use exceptions to represent coding errors (out-of-bounds array index, division-by-zero, access to a null). In these cases exceptions are implicitly signaled by the runtime environment when such condition happens. One of such implicit exceptions (caused by data conversion from a 64-bit floating point to a 16-bit signed integer) was the responsible for the famous failure on Ariane 5's first test flight~\cite{lions1996ariane} --- leading to the rocket self-destructing 37 seconds after launch and a loss of 500 million dollars. Besides Ariane 5, several applications crash everyday due to uncaught exceptions~\cite{jo2004uncaught}. The consequences of not careful exception usage may have the opposite effect of its initial intention of improving system robustness.

In Java, when the program fails due to an uncaught exception, it automatically terminates, while the system prints a
stack trace to the console. A stack trace includes a detail message and the execution stack frame when the exception occurred. A typical Java stack trace consists of an ordered list of methods that were active on the call stack before the exception has occurred. Stack traces are a useful source of information about system crashes, and is often used to support developers in debugging~\cite{schroter2010stack}. Moreover, information from stack traces combined with crash and bug reports can be used to support bug classification and clustering~\cite{wang2013improving, kim2011crash, dhaliwal2011classifying}, fault-proneness prediction models~\cite{kim2013predicting} and even automated bug fixing~\cite{sinha2009fault} tools.

However, all these works focus on the analysis of a single system at a time. In our current context, on which we have plenty access to the information available on open-source repositories like GitHub initial questions arise: What is the prevalence of exception stack traces on reported issues? What are the common characteristics of such stack traces?

%What the stack traces available in these repositories can tell us about the use of exceptions in Java programs?

%To guide the our study we also compile general guidelines on how to use exceptions proposed by Gosling~\cite{gosling2000java}, Wirfs-Brock~\cite{wirfs2006toward} and Bloch~\cite{bloch2008effective} and then we investigate: what stack traces can tell us about the adherence to such practices in Java programs?

To answer these questions, we conduct an exploratory study of stack traces in the wild. Using a custom tool called ExceptionMiner, that was specifically developed to for this study, we mined stack traces from issues across all Java projects on GitHub. Overall, we analyzed 356,057 issues from 16,836 projects, from which 28,800 stack traces were extracted and used as a source of information for our analysis. The stack trace analysis  was augmented by additional information extracted using byte code and source code analyses done on
a carefully selected subset of our initial sample (482 Android projects).
%This study was divided in two phases: in the fist phase all the issues of all Java projects available on GitHub were analyzed and the embedded stack traces extracted using ExceptionMiner.

To guide this exploration we compiled general guidelines on how to use exceptions proposed by Gosling~\cite{gosling2000java}, Wirfs-Brock~\cite{wirfs2006toward} and Bloch~\cite{bloch2008effective} and then we investigate what stack traces can tell us about the adherence to such practices in Java programs.
.%~\todo{This should be our motivation, as this is our biggest contribution. I already changed the text. Lets discuss about this new version}

%GGG: I think this is too much detail for the introduction
%In the second phase of this study, we combined the information of the mined stack traces with source code and bytecode analysis. To do so we selected the subset of 482 Android repositories. The Android repositories were chosen because this kind of application is known for dealing with several sources of exceptions, as they are usually io-intensive (i.e., deals with Network, Bluetooth, camera, and other resources), and multi-htreaded. Moreover, they are based on a single platform which allowed more fine-grained analysis of their exception behaviour.

This study is built on the following assumption: the stack traces reported on issues contain relevant information related to system crashes. Several studies and techniques have also based on this assumption (e.g.,~\cite{sinha2009fault}~\cite{dhaliwal2011classifying}~\cite{kim2013predicting}) and we build on it as well. Some outcomes were consistently detected through this large scale analysis of exception stack traces in the wild, such as:

\begin{itemize}
\item  A multitude undocumented runtime exceptions signaled by third party code (i.e., libraries and Android platform).
\item  Odd exception chains (e.g., Error wrapping checked and runtime exceptions and vice-versa). As an indication that the purpose of Java hybrid exception model may not have been adequately used.
\item  Stack traces caused by uncaught checked exceptions.IMPROVE.

\end{itemize}

Hence, the contributions of this study are as follows:
\begin{itemize}
\item  It performs the first large scale analysis of Java stack traces and how they can reveal bad-practices on the use of exceptions in Java. 
\item  It briefly introduces ExceptionMiner, a tool developed to support the analysis.
\end{itemize}

The contributions of this work allow for developers of robust Java applications to familiarize with the most common reported exceptions mis-uses which is the first step to help developers to avoid making them; (ii) designers of languages to consider ways of reducing the abundance of NullPointerExceptions, (iii) and tool designers to consider developing tools that enable a cross-project defect analysis. We strongly believe that, specially in the open-source environment, faults are not to be hidden in a private bug issue. Faults should be share and discussed, so that a developer can learn from other projects mistakes. Currently, the search facilities of repositories are very limited, and the ExceptionMiner tool is a contribution in this direction.

%GGG: Personally, I hate this paragraph in all papers as it usually reads like a space-filler
The remainder of this paper is organized as follows. Section 2 presents a background on error handling mechanisms and Java exception handling constructs and practices. Section 3 defines research questions, the configuration of our exploratory study, the target applications and the evaluation procedures. Section 4 reports our analysis of the empirical data collected relating to the first research questions. Section X is dedicated to provide insights for the 3rd research question. Section 6 provides further discussions and lessons learned and discusses the threats to validity associated to this study. Finally Section 7 describes the related work, and Section 8 presents our conclusions and directions for future work.

\section{Background}

\subsection{Error handling in programming languages}

Techniques for error detection and handling are no more an optional add-on but a fundamental part of the system~\cite{bruntink2006discovering}. Several approaches exists for detecting errors during system execution and responding to them 
accordingly, such as the return-code idiom (very popular in C programs and operating systems), and the exception handling mechanism, embedded in many main stream programming languages.

The main disadvantage of return codes is the fact that the programmer can just ignore the return code and let the program execute in inconsistent state. The exception mechanism~\cite{goodenough1975exception} was proposed as a way of preventing errors from being ignored, and was embedded in many mainstream programming languages. Among some similarities they mainly differ in the way exceptions are represented and handled in each language~\cite{garcia2001comparative}:

\noindent\emph{Checked Approach}. In languages such Modula-3, Guide, and Extended Ada all exceptions are checked, which means that the compiler statically checks if appropriate handlers are provided within the system. Accordingly, the programmer must explicitly specify every exception that can be signaled to support the compile time checking. This approach is known to be costly to maintain (Dooren and Steegmans, 2005) since exception updates are cascading; if a new exception type is added to a method signature then all other methods directly using this method must either change their 
signatures to include the new exception type or handle the new exception.

\noindent\emph{Unchecked Approach}. In languages such as Lore, C++, and Arche all exceptions are unchecked but the developer can optionally list exception types on method signatures. However, the developer is not warned by the compiler if an unchecked exception is not handled inside the application. Such an approach may hinder robustness because the client of a method cannot easily know which unchecked exceptions may be thrown, unless the code of the method and the methods called from it is inspected, which can be a very time consuming or infeasible task.

\noindent\emph{Hybrid Approach}. In Java a method can throw either checked or unchecked exceptions. Java is the only language that allows this hybrid approach.

\subsection{Exception handling in Java}
\label{sec:extypes}

In Java, an exception is represented according to the class hierarchy shown in Figure~\ref{fig:exchier}. 
According to this structure every exception is an instance of the Throwable class, and can be of three kinds: checked exceptions (extends Exception), run-time exceptions (extends RuntimeException), and errors (extends Error)~\cite{gosling2000java}.

\begin{figure}
\centering
\includegraphics[width=\hsize]{new2_hierarchy.png}
\caption{Exception Hierarchy in Java}
\label{fig:exchier}
\end{figure}

Neither run-time exceptions nor errors need to be specified on method signatures, and hence both are referred as unchecked exceptions. The Java Virtual Machine represents as runtime exceptions invalid operations detected in the program (out-of-bounds array index, divide-by-zero error, null pointer references) from which most of the programs are not expected to recover. By convention an Error represents an unrecoverable condition which usually results from failures detected by the Java Virtual Machine, such as OutOfMemoryError and normally should not be handled inside the application.

The Java specification~\cite{gosling2000java} suggests that user-defined exceptions should be checked exceptions. The main reason is because doing so the callers of a method will know about the exceptions that a method can throw, and so they can decide what to do about them. However, there is a long-lasting debate about the use of checked and unchecked exceptions in Java~\cite{javatut,stackoverlow,debate}\footnote{152 questions in Stackoverlow are related to this debate}, since there are pros and cons associated to each of them.

\subsubsection{Exception Throwing and Handling}

In Java an exception can be signaled in one of the following circumstances~\cite{gosling2000java}:

\begin{itemize}
\item  explicitly thrown when a throw statement is executed;
\item  implicitly thrown by the JVM when the evaluation of an expression violates the normal semantics of language (e.g., out-of-bounds array index, division-by-zero, access to a null reference); or 
\item implicitly thrown by the JVM due to an internal error or resource limitation (e.g., OutofMemoryError).
\end{itemize}
%\todo{This sub-section is quite similar to the section introduction, shouldn't we merge them?}
When an exception is signaled it causes a transfer of control from the point where the exception occurred to a point that can be specified by the programmer (exception handler). In Java it is represented by the try-catch block and is responsible for performing any recovery action necessary to bring the software system back to a normal state and, whenever this is not possible, to log the exception and abort the system in an expected safe way.

In Java, as in many languages, the search for the handler to deal with a raised exception occurs along the dynamic invocation chain. This is claimed to increase the reusability, since the invoker of an operation can handle it in a wider context ~\cite{goodenough1975exception}. If no handler is found in the chain, the default handler associated to the Thread or ThreadGroup is invoked, which usually terminates the system and prints the exception stack trace.

\subsubsection{Exception Chaining}
Exception Chaining is a common way of propagating exceptions across components of different abstraction levels. It is implemented by special exception constructor which receives another exception as a parameter. An example a multiple exception chaining is depicted on the stack trace in Figure~\ref{fig:wrapping}. 
%This stack trace was extracted from an issue of an Android application in Github.

The bottom part of the stack trace is the \emph{root cause}, which indicates the first reason for the error to be thrown (in this case, the computer run out of memory). The top part of the stack trace indicates the location of the exception manifestation (which will call \emph{top level exception} along this paper). The execution flow  between the root cause and the exception manifestation may include other intermediate exception wrappings. In all levels, the exception \emph{signaler}, is the method that threw the exception, represented on the stack trace as the first method call below the exception declaration.

\begin{figure}
\centering
\includegraphics[scale=0.6]{wrappings}
\caption{Example of an Exception stack trace in Java.}
\label{fig:wrapping}
\end{figure}

\subsection{Exception Handling Best Practices}
\label{sec:best}
%\todo{This section feels a bit disconnected.}
Several general guidelines have been proposed on how to use exceptions~\cite{mandrioli1992advances,gosling2000java,wirfs2006toward, bloch2008effective}. To guide our investigation, we compile general guidelines related to exception types and the propagation of exception in Java programs \footnote{We also compiled guidelines related to exception handling but they are out of the scope of this paper, since they do not relate to information available on stack traces.}.

\noindent\emph{Meaning of Exception Types}

\textbf{1) Use checked exceptions to represent recoverable conditions}(~\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,bloch2008effective}). The developer should use checked exceptions for conditions from which the caller is expected to recover. By confronting the API user with a checked exception, the API designer is telling the client to handle the exceptional condition. The client can explicitly ignore the exception (swallowing, or converting it to other type) at
the expense of the program's robustness~\cite{gosling2000java}.

\textbf{2) Error represents an unrecoverable condition detected by the JVM which should not he handled}~\cite{gosling2000java}. It results from failures detected by the Java Virtual Machine which indicate resource deficiencies, invariant failures or other conditions that make impossible the program to recover.

\noindent\emph{Exception Throwing}

\textbf{3) A method should throw exceptions that precisely define the exceptional condition}~\cite{gosling2000java,bloch2008effective}. To do so, developers should either try to reuse the exception types already defined in the JVM or they should create a specific exception. Throwing general types such as a pure Exception or a RuntimeException is hence considered a bad practice.

\textbf{4) Do not repeatedly re-throw exceptions, handle exceptions as close as possible to the problem}~\cite{wirfs2006toward}. Programs that frequently throw exceptions lose in performance ~\cite{wirfs2006toward,gosling2000java}. Moreover, close to the place that the exception was throw the caller usually has enough contextual information to perform  corrective actions. When the exception is propagated far away from the method where it was signaled (i.e. further away from the problem) it can be difficult to take meaningful recovery actions.

\noindent\emph{Exception Documentation}

\textbf{5) Document all exceptions thrown by a method}I~\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,bloch2008effective}). The exceptions thrown by a method are an important part of methods interface, and is required to use the method properly. The checked exceptions are already part of the  methods signature, and the method caller is aware of the checked exceptions being thrown by it. According to~\cite{bloch2008effective}, it is also wise to document the explicitly signaled runtime exceptions\footnote{This excludes the implicitly signaled JVM runtime exceptions related to coding errors} as carefully as checked exceptions. Doing so, the clients of a method will be aware of the exceptions the method can throw. If the developer fails to do follow this practice (specially when developing library code) it will be difficult or even impossible for the caller to make effective use of such method ~\cite{wirfs2006toward, bloch2008effective}.

These best practices guided our exploratory study described in the next sections. Based on information extracted from stack traces (complemented by bytecode and source-code analysis), we investigate whether stack characteristics can reveal whether these practices have been obeyed. Next section discusses the study procedure and the practices' misuses that emerged from this study.

\section{Study Design}

This work describes an exploratory study which was based on a sequential mixed-methods approach~\cite{ivankova2006using} for collecting, analyzing, and integrating both quantitative and qualitative analysis along the stages of the research process. The main goal of this study was to gain a better understanding of the use of exceptions in Java programs based mainly on the information available on exception stack traces. More specifically, we aimed at answering the following research questions:

\begin{description}
    \item[RQ1] What is the prevalence of stack traces on  reported issues?
    %\item[RQ2] What are the most common exceptions?
    \item[RQ2] What are the common characteristics of such stack traces?
    \item[RQ3] What patterns of exception (mis)-use emerge from the stack trace analysis?
    %~\todo{Perhaps rephrase this to: Are the identified best practices followed in practice?}
\end{description}

For RQ1 and RQ2, we explore the domain quantitatively, and use theirs answers as a basis to investigate the third research question. RQ3 aimed at finding whether of exception (mis)-uses could be revealed through stack trace analysis. To support this investigation, source code and bytecode analysis was used to leverage the  understanding of stack traces. For RQ3 we investigate quantitatively and highlight interesting cases by exploring cases qualitatively. Figure~\ref{fig:overviewfig} presents an overview of the mixed-methods approach conducted to answer these questions.

\begin{figure*}
\centering
\includegraphics[width=\hsize]{overview.pdf}
\caption{Overview of the study based on mixed approaches.}
\label{fig:overviewfig}
\end{figure*}

We begun our investigation using the dataset provided by the GHTorrent project~\cite{Gousi13}, an off-line mirror of the data offered through the Github API. The Github API data come in two forms; a streaming data flow lists events, such as reporting an issue, happening on repositories in real time, while a static view contains the current state of entities. The data is stored in unprocessed format, in a MongoDB database, while metadata is extracted and stored in a relational database. The GHTorrent dataset covers a broad range of development activities on Github, including issues. The project has been collecting data since February 2012. Up to Dec 2013, when we queried it to retrieve our bootstrapping sample which included: 356.057 issues from 16.836 Java projects (step 1).\footnote{A Java project is specified in GitHub with a specific label.}

After obtaining the bootstrap issue sample from GHTorrent, each issue was processed by ExceptionMiner tool developed in this work which extracted every exception defined on issues or on issue comments (step 2). Since this initial set of Java project is too diverse (it includes libraries, web-based or stand-alone applications), in the second phase of this study we restrict the analysis domain to Android projects (step 3). This decision was driven by the fact that most open source Android applications are known~\cite{ref} for dealing with several sources of exceptions due to multi-threaded execution and interaction with input/output resources; moreover they share a common well-defined platform and a limited set of libraries.

To support a deeper investigation of the stack traces of Android apps, every exception defined on a stack trace was then classified according to its type (e.g. Error, Exception or RuntimeException) (step 4). The exception classifier was initialized with all the exceptions defined on the Java SDK version 7, the exceptions of the Android (Platform and OS) (level 19), and all the exceptions of the 482 Android projects selected for this study, extracted in the previous step. Moreover, all the packages associated to exceptions and signalers were also extracted to support further analysis (step 5). The signaler classifier is initialized with information about all packages found on Java SDK version 7, and Android Platform and OS (level 19). Finally, additional meta-data is combined to each stack trace (step 6) which used together with manual inspections (step 7) enabled to provide answers to RQ3.

\subsection{Github Data}
\label{sec:git}

\noindent\emph{Kinds of Analyzed Issues.} Issues on Github are different from issues on dedicated bug tracking tools such as Bugzilla and Jira. The most important difference is that there are no predefined fields for bug reporters to select from (e.g. severity and priority); Github uses instead a more open
ended tagging system. Repositories are offered a pre-defined set of labels, but repository
owners can modify them at will. An issue can have an arbitrary set of labels attached. Hence, this study does not restrict the analysis of stack traces defined on defect-issues, based on the assumption that: regardless the type of issue where its is defined the stack trace information contain relevant information concerning the exception use. Moreover, issue and pull requests are dual on Github; all pull requests have a corresponding ``backing'' issue - which are automatically generated. Therefore, we needed to exclude pull request based issues from our analysis. Finally, the Github API allows the automated generation of issues for specific repositories, which automated tools often use to report crashes. In some cases, this led to high number of issues that included stack traces. We identified and filtered those cases out (e.g., pullWifi project was responsible for almost 50\% of all issues created on GitHub, it was removed from our analysis and also from GitHub).

\subsection{Android Apps Selection}
\label{sec:android}

%In the second phase of this study we selected a subset of Java applications to be analyzed in more
%detail.The subset chosen was comprised by the Android apps included on GitHub (until 23 Feb 2013).
To identify Android projects, we performed a case insensitive search for the 
term \textsf{android} in the repository's names and short descriptions.
The heuristic filtered 2.542 repositories, from which 589 apps had at least one issue containing a stack trace.

Then we performed further cleanup, inspecting the site of every Android reporting at least one stack trace, to make sure that they represented real mobile apps. During this clean up 106 apps were removed because they were either example projects (i.e., toy projects) or tools to support Android development (e.g. selendroid, roboeletric -- tools to support the testing of Android apps). The filtered set consisted of 482 apps, from which approximately 50\% are also hosted in Google Market Place. For each one of the analyzed projects, we downloaded the source code and using custom scripts, we extracted their package hierarchy (recursive list of of Java packages) and their custom exception types as described next.

\subsection{The ExceptionMiner Tool}
\label{sec:exceptionminer}

To extract exceptions from issues, we implemented ExceptionMiner, a modular mining tool
able to connect to various repositories (such as Google Code, GHTorrent or even directly to
Bugzilla), extract issues, mine stack traces from them and classify them in predefined categories. The main components of ExceptionMiner are as follows:

\noindent\emph{StackTraceMiner and Distiller} The first step in the process is mining references to exceptions and 
stack traces embedded in issues, distilling the information that composes a stack trace and storing
the results in a relational database. The extracted stack trace information  the exception signaler, 
the root cause, the exception wrappers. The tool is based on a combination of a regular
expression based parser with heuristics able to identify and filter exception names 
and stack traces inline with text. In contrast to existing issue parsing 
solutions such as Infozilla, the parser created in this work can extract all causes related to an 
exception on a stack, and stack traces embedded in logs files.

%\begin{tiny}
%\begin{verbatim}

%03-01 W/dalvikvm(7924): threadid=1: thread exiting with uncaught exception (group=0x40adf210)
%03-01 15:55:01.609 (7924): FATAL EXCEPTION: main
%03-01  (7924): java.lang.RuntimeException: ...
%E/AndroidRuntime( 7924):  at android....performLaunchActivity(ActivityThread.java:1967)
%03-01 15:55:01.609 (7924):  at android....handleLaunchActivity(ActivityThread.java:1992)
%03-01 15:55:01.609:  at android.app.ActivityThread.access\$600(ActivityThread.java:127)
%E/AndroidRuntime( 7924):  at android....handleMessage(ActivityThread.java:1158)
%\end{verbatim}
%\end{tiny}

\noindent\emph{Exception Classifier} The next step in the ExceptionMiner process is classifying exceptions 
as either checked exceptions, runtime exceptions or an errors. The exception classifier
is based on bytecode analysis (based on Design Wizard~\cite{}) that walks up the type hierarchy of a Java exception until it reaches a base exception type.

\noindent\emph{Exception Signaler Classifier} After its initialization (see Section 
set of packages for each classification category and pattern matches the signaler name
to classify it to one of the categories.

%\section{Study Operation and Results}
Next section presents the results for each of the study phases, providing both a quantitative and qualitative analysis of the outcomes.

%\subsection{Phase 1: Analyzing Pure Stack Trace Information of Java Programs}

\section{RQ1: What is the prevalence of stack traces on reported issues?}

Prevalence or prevalence proportion, a term often used in epidemiology, represents the proportion of a population found to have a particular condition (e.g. a disease) over the total number of people studied. In our work, we used this term to evaluate a condition (i.e., presence of a stack trace) over issues reported on GitHub repositories.

\input{stacktrace-stats}

\begin{table}
\centering
\begin{tabular}{p{1em}lrr}
\hline
 \multicolumn{2}{l}{\bfseries{Java repository-level metrics}} & \\
\hline
\multicolumn{2}{l}{\bfseries{Repositories}} & 16,836\\
    & incl. exception on 1+ issues & 4,776 (28,37\%) \\
    & incl. stack trace on 1+ issues & 3,758 (22,32\%) \\
    & incl. exception on 1+ defect issues & 2,019 (41,44\%)\\
    & incl. stack trace on 1+ defect issues  & 1,698 (34,84\%) \\

\hline
\end{tabular}
\caption{Projects x Stack Traces}
\label{tab:proj}
\end{table}

\begin{table}
\centering
\begin{tabular}{p{1em}lrr}
\hline
\multicolumn{2}{l}{\bfseries{Issue-level metrics}} & \bfseries{Java} & \bfseries{Android Subset}\\
\hline
   \multicolumn{2}{l}{\bfseries{Repositories}} & 16,836 & 482\\
%   \multicolumn{2}{l}{\bfseries{ }} &  &  \\  
 \multicolumn{2}{l}{\bfseries{Issues}} & 356,057 & 32,582\\
    & incl. exceptions & 30,236 (8,5\%) & 3,956 (12,1\%) \\
    & incl. stack traces & 21,013 (5,9\%) & 3,101 (7,3\%) \\
%   \multicolumn{2}{l}{\bfseries{ }} &  &  \\   
 \multicolumn{2}{l}{\bfseries{Defect Issues}} & 55,226 & 2,376 \\  
    & incl. exception & 6,741 (12,2\%) & 1,557 (65,6\%) \\
    & incl. stack trace &  5,196 (9,4\%) & 1,393 (58,7\%)\\
%\multicolumn{2}{l}{\bfseries{ }} &  &  \\  
\multicolumn{2}{l}{\bfseries{Exception Stack Traces}} &  & \\
    & on issues  & 28,800 & 4,208  \\
    & on issues labeled as defect & 6,529 & 1,699 \\
\hline
\end{tabular}
\caption{Quantitative overview of the analyzed dataset}
\label{tab:overview}
\end{table}

The use of the Github issue tracker in repositories hosted on Github is optional
and not widespread. As of March 2014, from the 332,864 non-fork Java repositories registered in the 
GHTorrent dataset, only 44,323 have received an issue report in their lifetime while from those only 16,837 are being watched by external users. To ensure that our work targets projects that are openly used by 
users other than their developers, we focused our analysis on just those
projects. Using the heuristics presented in Section~\ref{sec:android}, 
we selected 482 repositories featuring Android projects. Table~\ref{tab:proj} and 
Table~\ref{tab:overview} presents two views of the extracted dataset,
namely grouped by repository and grouped by issue.

From Table~\ref{tab:overview}, we can see that repositories that include at least one exception or stack trace on an issue 
are fairly common: 28\% and 22\%  respectively. Considering only the exceptions and stack traces reported on defect issues we find out that 41\% of the repositories reported at least one exception in a defect issue, and approximately 35\% of the projects mentioned one stack trace in at least one defect issue.

At the individual issue level, 
8,5\% and 5,9\% of all issues reported in Java repositories contain the full name of an exception or a stack trace respectively. The numbers are similar if we isolate Android projects from the general population (12,1\% and 7,3\%). Considering only the defect-issues however, in the
Android repositories as almost 58,7\% of defect-issues include a stack trace. This is not true for the general case however.

\input{android-stack-stats}

%Based on the assumption presented in Section III, we therefore did not use this label for further filtering.

%The results in Table~\ref{tab:overview} are cumulative; however, the numbers
%are not the same among all projects. Figure~\ref{fig:stack traces} 
%presents the descriptive statistics of the range of values of the metrics described in  
%Table~\ref{tab:overview} across all projects in our dataset. 
%There we can see that indeed there is significant variation among projects,
%while the outliers lay at the bottom end of the spectrum, which means that for
%some projects, we can find very limited number of stack traces in their 
%issue reports. Both Java and Android projects feature similar variation,
%which again supports our Android focused analysis later on.

\noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
RQ1: 28,37\% of Java repositories on Github have at least 1 issue that includes a stack trace. 9,4\% of the defect issues of such repositories mentioned at least one stack trace. Considering only defect-issues of Android subset 58,7\% mentioned at least 1 stack trace.
}
\end{minipage}}

% \begin{itemize}

%     \item What are the most ``popular'' exceptions and what they can tell us?

%     \item Which elements are triggering most of the exceptions and what they 
%     can tell us about the exception structure?

%     \item What is the role of libraries and the underlying platform on crashes?

%     \item What is the role of the application itself on crashes?
% \end{itemize}

% \subsection{The rules of using checked and unchecked exceptions are being obeyed in mobile environment?}

\section{RQ2: What are the common characteristics of such stack traces?}

To answer this question, we mined the following information of every extracted stack trace: the top level exception being signaled, its signaler (the first method signature frame after the exception), the root exception (the exception which initiated the stack trace, defined on the bottom of the stack), and the root signaler. The intermediate elements that compose a stack trace as shown in Figure 1 (i.e., the intermediate causes and its execution stacks) were not be considered in the analysis.~\todo{Move this to the previous section? IMO, It is better to keep all facts about how we did the
research in one place}

Based on the assumption that, regardless the issue type every exception stack trace contains relevant information concerting the exception structure of projects analyzed we opted for not restricting the analysis only on defect issues~\footnote{We conducted the same analysis on the defect issues and the top exceptions are similar to the ones mentioned on defect issues. Due to space limitation we limit to present the general analysis here. More detailed analysis on the defect issues can be found here: \url{www.dimap.ufrn.br/~roberta/icsm2014}}.

\subsection{Top 10 most common exceptions in Java Projects}

Table~\ref{tab:toptenjava} illustrates the top 10 most common root causes which represents more than 50\% of the analyzed set. The table illustrates  (i) the number of times each exception appeared as the root cause of a stack trace (\#occurrences), (ii) the number of distinct projects reporting such causes on stack traces, and (iii) the exception popularity. The exception popularity metric was defined in this work as follows: \emph{popularity = number of projects mentioning the exception in at least one stack/all projects that contain at least one stack}.

\begin{table}
\begin{tabular}{lrrrr}
    \hline
    \bfseries{Exception} &  \multicolumn{2}{c}{\bfseries{Occurrences}} &  \multicolumn{2}{c}{\bfseries{Projects}}\\
    & \bfseries{\#} &  \bfseries{\%} & \bfseries{\# } & \bfseries{\% } \\
    \hline
java.lang.NullPointerException	& 7578	& 26,31\%	& 1671	& 44,47 \\
java.lang.IllegalArgumentException &	1573 &	5,46\% &	670	& 17,83\% \\
java.lang.ClassNotFoundException &	1203 &	4,18\%	& 514 & 13,68\% \\
java.lang.RuntimeException &	1093	& 3,80\% &	411	& 10,94\% \\
java.lang.IllegalStateException	& 964 &	3,35\%	&  431 &	11,47\% \\
java.lang.ClassCastException &	886	& 3,08\% & 408 & 10,86\% \\ 
java.lang.ArrayIndexOutOfBoundsException &	834	& 2,90\% & 345 & 9,18\% \\
java.lang.NoSuchMethodError	& 822 &	2,85\% & 369 & 9,82\% \\
java.io.IOException &	656 &	2,28\%	& 337 &	8,97\% \\
java.lang.OutOfMemoryError	& 601 &	2,09\% & 235 & 6,25\% \\
%java.lang.NoClassDefFoundError &	554	& 1,92\% &	299 &	7,96\% \\
%java.util.ConcurrentModificationException	& 465 &	1,61\% & 180 & 4,79\% \\
%java.lang.IndexOutOfBoundsException	& 445 &	1,55\% &	205 &	5,46\% \\
\hline
All	 &	23956 &	100\%	& 3758 & 100\%	 \\
\hline
  \end{tabular}
\caption{Top 10 most popular exceptions on issues}
\label{tab:toptenjava}
\end{table}

We can observe that the NullPointerException was the one with the higher number of occurrences (22,5\% of all stacks were caused by NullPointerExceptions) and the one which had the highest popularity across projects. Actually, most of the exceptions in the TOP 10 list (except java.lang.IllegalArgumentException, java.io.IOException and java.lang.RuntimeException) are exceptions implicitly signaled by Java Virtual Machine, when an expression violates the normal semantics of the Java programming language (e.g. ArrayIndexOutOfBounds) or due to an internal error or resource limitation (OutOfMemoryError).

The high prevalence of NullPointerExceptions, and the other implicitly-thrown exceptions  is aligned with the findings other works [Sunghun et al.][Gordon fraser and][Jcrasher].[Android-predict]~\todo{refs}. For instance, Sunghun et al. showed that in Eclipse bug report system 38\% the bugs related to exception handling were caused by NullPointerException; other works on robustness testing [JCrasher, Evol;suite, Amiya]~\todo{refs} showed that most of the automatically detected bugs were due to NullPointerExceptions and implicitly-signaled of Java environment (as the ones found in this study).

\subsection{Top 10 most common exceptions in Android Projects}

As detailed in Section III, 443 Android projects were carefully selected to enable a deaper investigation of stack trace information, by leveraging stack trace information with the source code and bytecode analysis. For this subset of Android projects we could extract every package related to each exception signaler, and based on this information we could classify every stack trace based on its root signaler as presented in Table~\ref{signalers}.

\begin{table}
  \centering
  \begin{tabular}{rp{22em}}
    \hline
    \bfseries{Type of signaler} & \bfseries{Description} \\
    \hline
    \bfseries{android} & When the exception is explicitly thrown by a method defined in Android Platform, in a the JDK library used by it.\\
    \bfseries{app} &	If the exception signaler was explicitly thrown by an application method or in a the JDK library used by it.\\
    \bfseries{libcore} &	If the application was signaled by one of the libraries reused by Android Platform (i.e., org.apache.harmony, org.w3c.dom, sun.misc, org.apache.http, org.json, org.xml) besides the JDK. \\
    \bfseries{lib} &	If the exceptions was not defined by any of the elements above.\\
    \bfseries{java} &	If the application signaler and all the methods that composes the exception are defined in a the JDK library method.\\
    \hline
  \end{tabular}
  \caption{Sources of exceptions in Android}
  \label{tab:signalers}
\end{table}

For each exception we could discover: the number of times it occurred on stacks, the number of distinct projects such stacks were defined (and consequently its popularity) and the of times it was signaled by different signaler types. We could calculate the popularity of each root cause, considering the number of distinct projects they were found and the whole set of projects analyzed with at least one stack trace (443). Table~\ref{toptenandroid} and Figuere~\ref{androidsignaler} presents the mined data.

\begin{table*}
  \centering
  \begin{tabular}{rccccccccc}
    \hline
    \bfseries{Root Exception} &  \multicolumn{2}{c}{\bfseries{Occurrences}} &  \multicolumn{2}{c}{\bfseries{Projects}} & \textsf{android} & \textsf{libcore} & \textsf{app} & \textsf{lib} & \textsf{java} \\
    & \bfseries{\#} &  \bfseries{\%} & \bfseries{\# } & \bfseries{\% } &&&&&\\
    \hline
java.lang.NullPointerException	            & 1225  &   30,08\% &	254	&	52,70\%	&	473	&	18	&	595	&	137	&	2	\\
java.lang.IllegalStateException	            & 234	&   5,75\%  &	99	&	20,54\%	&	165	&	12	&	36	&	20	&	1	\\
java.lang.IllegalArgumentException	        & 255	&   6,26\%  &	94	&	19,50\%	&	146	&	6	&	64	&	39	&	0	\\
java.lang.RuntimeException	                & 232	&   5,70\%  &	91	&	18,88\%	&	167	&	1	&	47	&	17	&	0	\\
java.lang.OutOfMemoryError	                & 180	&   4,42\%  &   56	&	11,62\%	&	121	&	15	&	17	&	23	&	4	\\
java.lang.NoClassDefFoundError	            & 73	&   1,79\%  &	52	&	10,79\%	&	9	&	0	&	37	&	26	&	1	\\
java.lang.ClassCastException	            & 94	&   2,31\%  &	49	&	10,17\%	&	43	&	0	&	40	&	11	&	0	\\
java.lang.IndexOutOfBoundsException	        & 127	&   3,12\%  &	47	&	9,75\%	&	47	&	0	&	71	&	8	&	1	\\
java.lang.NoSuchMethodError	                & 57	&   1,40\%  &	40	&	8,30\%	&	9	&	0	&	39	&	9	&	0	\\
java.util.ConcurrentModificationException   & 54	&   1,33\%  &	38	&	7,88\%	&	5	&	0	&	43	&	6	&	0	\\

    \hline
  \end{tabular}
\caption{Root Exceptions occurrences and popularity.}
\label{tab:toptenandroid}
\end{table*}

\begin{figure}
\centering
\includegraphics[width=\hsize]{top_exceptios_android.pdf}
\caption{Top exceptions in Android repositories.}
\label{fig:androidsignaler}
\end{figure}

%\textbf{
Coding errors are the most common root causes for most types of signalers. We could observe that the NullPointerException is still the exception with higher number of occurrences. The NullPoiterExceptions are mainly signaled inside Android platform and inside application code. Although we also find NullPointerExceptions being signaled from third-party libraries. Regarding reusable code, there is no consensus if it s a good or bad practice to re-throw NullPointerException. Some prefer to encapsulate such exception on IllegalArgumentException. But others~\cite{joshuabloch} argue that the NullPointerException  explicits the cause of the problem and hence should not be wrapped.

\textbf{We could observe in this study that the NullPointerException and other other implicitly signaled exceptions represented most of the exceptions reported on issues in both Java and Android subset. For such exceptions - which represent programming bugs or resource limitations - there is usually no proper handling besides presenting an error message to the user and restarting the application - only high fault tolerant systems need to provide solutions to handle them [REF: Java Specification].}

\subsection{Classification of the main causes of top exceptions}

To better understand the main causes of such exceptions we evaluated semantics related to the top 100 root causes reported on stack-issues of both Java dataset and Android subset (which correspond to approximately 77\% of all exceptions reported reported on Java stack-issues and 95\% of all exceptions reported on Android stack-issues), and classified them according to the the categories presented in Table~\ref{tophundrend}.

\begin{table}
  \centering
  \begin{tabular}{lrr}
    \hline
    \bfseries{Category} & \bfseries{Java} & \bfseries{Android Subset} \\
    \hline
      Programming logic (java.lang and util) & 12625	(41,10\%)	& 2235 (55,64\%)\\
      Resources (IO) & 12440 (40,50\%) & 727 (18,10\%) \\ 
      Security & 220	(0,72\%) & 165 (4,11\%)\\
      Concurrency	& 633 (2,06\%) & 116	(2,89\%)\\
      Backward compatibility & 1580 (5,14\%) & 219 (5,45\%) \\
      Reflection &1413 (4,60\%)   & 91 (2,27\%)\\
      Specific (GUI,FRAMEWORK) &	340	(1,11\%) & 197 (4,90\%)\\
      General (Error, Exception, Runtime) & 1468 (4,78\%) & 267 (6,65\%)\\
    \hline
  \end{tabular}
  \caption{Characterization of the 100 root causes.}
  \label{tab:tophundrend}
\end{table}

As we can see from Table 4, the that programming errors are the causes of most of the exception stack traces reported on issues. In the Android subset we could find more issues related to Backward compatibility and security issues than on the general data set.

\subsection{Sizes of Stack traces}

Concerning the size of the stacks (i.e. the number of frames composing them), we could observe that  sometimes it exceeded 100 frames (158 stack frames). Manually inspecting the corresponding execution trace was caused by one of the two reasons: recursive calls; exceptions signaled by a method defined deep down in a reused framework; exceptions successively wrapped. Regardless the reason, in exception flows involving many methods, it is piratically impossible to handle the exception a way different than existing the  application, because there is hardly enough contextual information to perform any other recovery action.

TODO: Include a Histogram here.

%Considering only the packages of each exception we could observe that 81,28\% of all analyzed set of stack traces reported exceptions defined by java.lang (see Table XXX).

%\begin{table}
% \centering
%\begin{tabular}{lrrrr}
%    \hline
% \bfseries{rank} & \bfseries{package} & \bfseries{occurr} &  \bfseries{\%occur} & % \bfseries{projects} \\
%     \hline
%1 &java.lang &	19471 &	81,28\%	& 3080 \\
%2 &	java.io	 & 1222	& 5,10\% & 509 \\
%3 &	java.net &	740	& 3,09\% &	311 \\
%4 &	java.util &	689	& 2,88\% &	287 \\
%5 &	java.lang.reflect & 115 &	0,48\%	& 78 \\
%\hline
%  \end{tabular}
%\caption{Top 5 most popular packages associated to the root exceptions of stack traces}
%\label{tab:toptenpopular2}
%\end{table}

\section{RQ3.  What kinds of exception (mis)-use emerge from stack trace analysis? }

To answer this question we analyzed general characteristics of stack traces reported on Java repositories (e.g., exceptions being thrown presented in Tables~\ref{toptenjava} and ~\ref{toptenandroid}), and took a deeper look at the stack traces of the Android repositories combining them with with source code and byte code analyses. Doing so, we mined some characteristics of stacks that points to best practices violations.

\subsection{General exceptions being thrown.}

From Tables Tables~\ref{toptenjava}, ~\ref{toptenandroid} we can observe that, in both Java and Android related issues, direct instances java.lang.RuntimeException were thrown  (in 10,94\% of Java repositories and 18,88\% of Android repositories respectively). Moreover, after inspecting the top 100 most thrown exceptions (see ~\ref{tophundrend}) we could observe that stack traces on which direct instances of java.lang.Exception and java.lang.Error were thrown. Throwing general exceptions is usually considered a bad practice, because the type does not carry enough information to identify the cause of the exceptional behaviour. In such cases, developers need to relying on exception message which can be neither complete nor precise ~\cite{gosling2000java}.

In Table ~\ref{toptenandroid} we can observe a high prevalence of RuntimeExceptions being thrown by the Android platform (167 out of 232).  Manually inspecting the code on which RuntimeExceptions were thrown in Android Platform we could observe that direct instances of java.lang.RuntimeException they were thrown for 2 main reasons: (i) when there was no reasonable way of handling the exception, which means that regardless the kind of exception the only possible action was to abort the application; and when wrapping exceptions signaled by application methods. Since Android platform is heavily based on "call backs", when the application does not handle its exceptions, they scape to Android platform which does not have contextual information to adequately handle the exception and simply terminates the application. What is apparently a bad practice of signaling general exceptions in this context it was and intended design decision.

%Even an instance of Throwable as signaled in one project. It points to a the "dont worry from know behaviour".

%\subsection{Combining Stack Trace Information with Exception Type Information}

\subsection{Undocumented runtime exceptions signaled by third party code}

In this study we also classified each stack trace according to the type of the root exception been signaled. Table~\ref{typeroottab} presents the types of root causes of all stack traces reported on the Android repositories.

\begin{table}
\centering
\begin{tabular}{lcccccc}
    \hline
    \bfseries{Type} & \bfseries{Android} & \bfseries{Lib} & \bfseries{Libcore} & \bfseries{Java} & \bfseries{App} & \bfseries{All}\\
    \hline

Runtime	&	1101	&	56	&	1446	&	423	&	10	&	3036	\\
Error	&	148	&	34	&	168	&	114	&	9	&	473	\\
Checked	&	115	&	241	&	139	&	176	&	12	&	683	\\
Undefined	&	1	&	0	&	6	&	8	&	0	&	15	\\
Throwable	&	0	&	0	&	1	&	0	&	0	&	1	\\

    \hline
 All  &	1365	&	331	&	1760	&	721	&	31	&	4208	\\
    \hline
  \end{tabular}
\caption{Types of Root Exceptions.}
  \label{fig:typeroottab}
\end{table}

\begin{figure}
\centering
\includegraphics[scale=0.3]{chart_exceptiontypes.pdf}
\caption{Distribution of Checked and Unchedked exceptions.}
\label{fig:typeroot}
\end{figure}

%TO DO: INCLUDE A NEW TABLE WITH NON JAVA.LANG RUNTIME EXCEPTIONS. Point to we cannot trust developers to document.

As we can see from Table ~ref{typeroottab}, that most of the exceptions signaled by third-party code (e.g., Android and libs) were runtime exceptions. 
When an exception is signaled by third party code, it can be due to 2 reasons: the precondition associated to third party's code was not obeyed; or there is a coding error on third party's code. Only manual code inspection can reveal the reason. However, regardless the reason it is a good practice that all the exceptions signaled by a third part code are properly documented. Manually inspecting a sample of stack traces caused by runtime exceptions thrown by library code (approximately 10\%) we could observe that none of the runtime exceptions signaled by third part code were documented in Javadoc - this analysis did not consider the exceptions implicitly signaled by JVM (e.g. NullPoiterException). The undocumented runtime exceptions are a serious problem as it makes very difficult or even impossible for the  client method to protect against  undocumented runtime exceptions signaled by library code. As a consequence, the undocumented runtime exception may remain uncaught leading to  system crashes.

%This result is aligned with the results of other study conducted by from Cabral and Marcques ~\cite{} which also observe that runtime exceptions are most often not documented.

%\textbf{When libraries are used, the developer does not have access to their source code and thus needs to rely on the library documentation about the runtime exceptions that should be thrown – which, more often than not, are neither complete nor precise (Thomas, 2002; Sacramento et al., 2006). As a consequence, unhandled unchecked exceptions can be seen as one of the major sources of bugs in current Java systems (Jo et al., 2004). }

On the other hand, we could observe that most of the exceptions signaled by the set of libraries reused by Android platform (i.e., org.apache.harmony, org.w3c.dom, sun.misc, org.apache.http, org.json, org.xml, javax.) referred in this work as libcore, were checked exceptions. 
It is pointed as a good practice for libraries (see Section ~\ref{best}) because by using checked exceptions libraries can define a precise exception interface to its clients. Such libraries are heavily used in several projects, and such precise exception interface may be related to the libraries maturity.

\subsection{Uncaught Checked Exceptions}

Although most of the reported were runtime, we could also find checked exceptions as the root causes of stack traces reported on issues. An exception can remain uncaught in two circumstances: if all methods on the execution trace on which this exception flows explicitly specifies the exception, or if the checked exception is wrapped in a runtime and than re-thrown.Since checked exceptions should be used to represent recoverable conditions, ignoring a checked exception is considered a bad practice. When the developer is confronted with a checked exception, the designer of the API is telling him to handle the exceptional condition. To better understand what was causing checked exceptions to scape from being handled we investigated the kinds of wrappings that happened on stacks the stacks.  Next sections presents the wrappings found and discuss about them.

%\noindent\emph{3) Odd Wrappings}
\subsection{Odd Wrappings}

 Table~\ref{wrappingandroid} presents the wrappings found in this study for all stack traces found in Android repositories. As we can see, some of the checked exceptions where indeed wrapped in runtime exceptions or even errors, while most of them were not wrapped along the stack trace (represented as the table lines on which the root cause is none). This analysis also revealed unexpected wrappings such as: Error wrapping a Checked exception; a Runtime wrapping and Error; and an Error wrapping a Runtime.

\begin{table}
\centering
\begin{tabular}{llll}
    \hline
    \bfseries{Exception Type} & \bfseries{Root Cause} & \bfseries{Occurrences} & \bfseries{Projects} \\
    \hline
    Runtime	&	None	&	2360	&	359	\\
Runtime	&	Runtime	&	560	&	182	\\
Checked	&	None	&	422	&	117	\\
Error	&	None	&	381	&	141	\\
Runtime	&	Checked	&	109	&	70	\\
Checked	&	Checked	&	98	&	42	\\
Runtime	&	Error	&	55	&	38	\\
Checked	&	Runtime	&	22	&	12	\\
Error	&	Error	&	15	&	14	\\
Undefined	&	None	&	15	&	10	\\
Error	&	Runtime	&	13	&	6	\\
Checked	&	Error	&	8	&	7	\\
Error	&	Checked	&	7	&	4	\\
Throwable	&	Runtime	&	3	&	1	\\
Undefined	&	Runtime	&	3	&	3	\\
Throwable	&	None	&	1	&	1	\\
Undefined	&	Error	&	1	&	1	\\
    \hline
  \end{tabular}
\caption{Kinds of wrappings found on the stacks of 443 ANDROID projects found on Github}
\label{fig:wrappingandroid}
\end{table}

Java is the only language that provides a hybrid exception handling mechanism which offers different types to represent different exception behaviors (see Section ~\ref{extypes}) . According to Java specification Errors should not be handled inside the system since they usually represent unrecoverable conditions detected by the JVM such as OutOfMemoryError. Checked exceptions should represent recoverable conditions and  Runtime exceptions represent are usually related to coding errors from which the developer is not expected to recover from.

%Table X illustrated examples of such wrappings.
%TODO INCLUIR ESTA TABLEA

Since there is no way of enforcing such conventions during program development,  we could observe the stack trace analysis combined with types these three types of exceptions were used interchangeably. This interchangeability makes the behaviour of the exception handling code more complex and less reliable. For instance, considering a situation were an OutOfMemoryError (a situation that should not be handled) was wrapped in a checked exception (a situation that must he handled), in this case this wrapping is telling to the caller that it should handled.

\section{Threats to Validity}

%\subsection{Access to Crash Logs}

%The data

%\subsection{Data Extraction from Natural language Artifacts}

%-Data cleaning removing invalid and duplicated issues.
%-Not every crash information is specified on an issues....

%\subsection{Construct, Internal and External Threats}

\subsection{Internal Validity}
We used a heuristics-based parser to mine exceptions from issues.  
Our parsing strategy was conservative by default; for example, we only
considered exception names using a fully qualified class name as valid
exception identifiers, while, in many cases, developers use the exception
name in issue description. Conservative parsing may minimize false positives,
which was our initial target, but also tends to increase false negatives,
which means that some cases may have not been identified as exceptions or
stack traces. Our limited manual inspection did not reveal such cases.

\subsection{External Validity}
The results presented here were based on mining issues on Github through the GHTorrent
dataset. While comprehensive and extensive, GHTorrent is not an exact replica of
Github, so several issues might be left out. Due to the way data collection works with
GHTorrent, for projects that are relatively inactive, GHTorrent might have not collected a 
significant proportion of their issues. We did not investigate the extend of this threat on 
our sample.

\subsection{Construct Validity}
Parts of our analysis are based on the availability of stack traces on issues on
Github. In using this dataset, we make an underlying assumption: the
stack traces reported on issues are representative of real crashes in 
the applications. The assumption is impossible to mitigate without access to 
the full set of crash data per application. Services exist to collect all
crash data from applications, so access to such a dataset will allow for
a thorough replication of our analysis, perhaps at the individual application 
level.

\section{Related Work}

In this section, we present works thaat are related to our own, distributed in four categories: (i) works that use the information available on stack traces; (ii) empirical studies on the usage of Java exceptions and its fault proneness; (iii) tools to extract stack traces information from natural language artifacts (i.e., issues and emails);  and (iv) empirical studies involving Android apps.

\textit{Analysis of Java Exception Usage.} Since the manual analysis of the Java exception-flow can easily become infeasible Robillard and Murphy~\cite{Robil00} developed a tool called Jex that analyzes the flow of exceptions in Java source code. Based on java source code this tool performs dataflow analysis in order to find the propagation paths of checked and unchecked exception types. After this tool other tools have been proposed to support the static analysis of exception flows ~\cite{fu2007exception} ~\cite{coelho2008assessing}. The main limitation off all tools is the number of false positives inherent to static analysis solutions, which can lead to a high number of exception flows, specially if considering Java Environment exceptions and exceptions signaled from libraries. This was the fist of some tools proposed to statically analyzed the exception flow of exceptions. Cabral and Marques~\cite{cabral2007exception} analyzed the exception handling code of 32 open-source systems, both for Java and .NET. The goal of their study was to identify how exceptions were handled in different categories of systems. They examined the exception handlers and the respective actions taken on them. They observed that the action handlers were very simple (e.g., logging and present a message to the user). In our work differs from previous works since in our study we tried to identify bad practices from the combined use of stack traces extracted from issues and bytecode and source code analysis. Finally, Reimer and Srinivasan~\cite{reimer2003analyzing} listed a set of bad practices on exception handling that hinder software maintainability and robustness. Their differs from ours it was solely based on their own experience on working with Java enterprise applications and focus on source code examples of bad patterns on ways of handling exceptions such as exception swallowing, log verbosity.

\textit{Analysis and Use of Stack Trace Information.} Some works have investigated the usage of stack trace information to support bug classification and clustering~\cite{wang2013improving, kim2011crash, dhaliwal2011classifying}, fault-proneness prediction models~\cite{kim2013predicting} and even automated bug fixing~\cite{sinha2009fault} tools.Kim et al~\cite{kim2011crash} use of crash graphs (an aggregated form of multiple stack traces available in crash reports) to detect duplicate crash reports and to predict if a given crash will be fixed. Dhaliwal et al~\cite{} analyzed crash reports collected for Mozilla Firefox and proposed a crash grouping approach based on the Levenshtein distance. This work showed that a crash grouping approach can reduce bug fixing time in approximately 5\%. Wang et al~\cite{} proposes an approach to define correlated crash types automatically. Moreover it also provides a bug localization method to locate and rank files related to the bug described on a stack trace.
Schroter et al~\cite{schroter2010stack} conducted an empirical study on the usefulness of stack traces for bug fixing. The study used the development data of Eclipse project and could observe that developers fixed the bugs faster when failing stack traces were included on bug issues. This study is aligned with the one performed by Bettenburg et al. ~\cite{bettenburg2008makes} which evaluated what content could make a good bug report, and stack traces were mentioned as the second most important content for developers - just after the "steps to reproduce" information. Sinha et al~\cite{sinha2009fault} proposed an approach for locating and repairing faults that are caused by JVM implicitly signaled exceptions related to coding errors such as NullPointerExceptions. This approach uses the information available on the stack trace to guide a dataflow analysis, to locate the source statement responsible for the exception. Kim at al.~\cite{kim2013predicting} proposed an approach to predict the crash-proneness of methods based information extracted from stack traces and methods' bytecode operations. They used InfoZilla to extracted stack trace information from the bug reports of two open-source projects, i.e. AspectJ and Eclipse. They could observe that most of the stack traces were related to NullPointerException and other JMV implicitly thrown exceptions had the higher prevalence on the analyzed set of stacks.

\textit{Extracting Stack Traces from Natural language artifacts.} Nowadays many software vendors embed automatic crash reporting tools in their software systems. Hence, whenever a software crashes this tool sends a detailed crash report to its vendors. Moreover, there are also find third party software solutions specialized in bug reporting, most of them for the increasing marked of mobile apps~\cite{BugSe14,BugSn14,Googl14,Acra14}. Hence most of the works described previously used the  information provided by crass reporting tools. Two works have proposed tools extract stack traces from natural language artifacts such as issues and emails: Infozilla ~\cite{bettenburg2008extracting} is based on a set of regular expressions that extract an the set of frames related to a stack trace, the main limitation of this solution is that it is not able to extract stack traces embedded on verbose log files (i.e., on which we can find log text mixed with exception frames). Alberto and Lanza ~\cite{bacchelli2012content}  propose a solution to recognize stack trace frames from development emails and relate it to code artifacts (i.e. classes) mentioned on the stack trace.

\textit{Empirical studies using Android apps.} Ruiz et al.~\cite{Ruiz12} investigated the degree of reuse across applications in Android Market, the study showed that almost 23\% of the classes inherited
from a base class in the Android API, and that 217 mobile apps were reused
completely by another mobile app. Pathak et al.~\cite{Patha11} analyzed bug
reports and developers discussions of Android platform and found out that
approximately 20\% of energy-related bugs in Android occurred after an OS
update. McDonnell et al.~\cite{McDon13} conducted a case study of the
co-evolution behavior of Android API and 10 dependent applications using the
version history data found in github. The study found that approximately 25\% of
all methods in the client code used the Android API, and that the methods
reusing fast-evolving APIs were more defect prone then others. Vásquez et
al.~\cite{Linar13} analyzed approximately 7K free Android apps and observed that
the last successful apps used Android APIs that were on average 300\% more
change-prone than the APIs used by the most successful apps. Pingyu and
Elbaum~\cite{Zhang12} analyzed bug reports of 5 Android applications an observed
that 29\% had to do with poor exceptional handling code.  Our work differs from
the others as it aims at distilling the information of bug reports describing
uncaught exceptions created for mobile applications in Github and Googlecode, in
order to get a first view of what is causing the crashes across applications
available and what the characteristics of the stack traces can tell us about the
exception structure of those applications. Moreover this work\ldots.

\section{Conclusion}
% The line below was meant to be a joke :-)
%Georgious wrote: That's a great paper with profound insights.

In this work, we perform the first large scale analysis of Java stack traces and pinpoint how they can reveal bad programming practices. We mined and analyzed the stack traces embedded in all issues of Java projects available GitHub. Overall, our research set includes 356,057 issues define at 16,836 projects (from which 28,800 stack traces were extracted). From this set, the stack traces of 482 Android projects were investigated in more detail, in combination with source code and bytecode analysis. In this study some patterns of exception (mis)-use were consistently detected such as: unexpected wrappings (e.g., Errors being wrapped in checked exceptions) - revealing that Java hybrid exception model is not fully used according to its purpose;  undocumented runtime exceptions signaled by third party code - which makes almost impossible for library clients to protect against such exceptions; and a high prevalence of java.lang exceptions reported on issues - representing approximately 50\% of the analyzed issues.

%Nowadays many software vendors embed automatic crash reporting tools in their software systems. Hence, whenever a software crashes this tool sends a detailed crash report to its vendors. Moreover, we can also find third party sofware solutions specialized in bug reporting for different kinds of systems specially for the increasing marked of mobile apps~\cite{BugSe14,BugSn14,Googl14,Acra14}. There is planty of information to be mined...

\section*{Acknowledgment}
This work is partially supported by: CNPq - Proc. 484209/2013-2 and the NWO TestRoots project (639.022.314).

\bibliographystyle{IEEEtran}
\bibliography{android-stacks}

% that's all folks
\end{document}

### 2014-05-06 android-stack-stats.tex
Initial commit

% latex table generated in R 2.15.3 by xtable 1.7-1 package
% Tue May  6 11:34:13 2014
\begin{table*}[ht]
\centering
\begin{tabular}{rp{16em}rrrrc}
  \hline
Feature & Description & quant\_5 & mean & median & quant\_95 & histogram \\ 
  \hline
all & All issues recorded for the project. & 2.00 & 67.70 & 20.00 & 248.00 & \includegraphics[scale = 0.1, clip = true, trim= 50px 60px 50px 60px]{hist-5d4a87daa44c63097df9bb1f3506d541.pdf} \\ 
  with\_exception & All issues featuring an exception. & 1.00 & 8.16 & 2.00 & 34.00 & \includegraphics[scale = 0.1, clip = true, trim= 50px 60px 50px 60px]{hist-bf73235d0503f937ef6f36c28cf1ba0d.pdf} \\ 
  with\_stack & All issues featuring a stack trace. & 1.00 & 6.44 & 2.00 & 27.00 & \includegraphics[scale = 0.1, clip = true, trim= 50px 60px 50px 60px]{hist-528cb6be67074e6e2c22d11245920e13.pdf} \\ 
  defect\_issues & All issues labeled as defect. & 0.00 & 13.28 & 1.00 & 52.00 & \includegraphics[scale = 0.1, clip = true, trim= 50px 60px 50px 60px]{hist-eb1797e86ef23b2a07907e5ccbeff7b9.pdf} \\ 
  defect\_with\_exception & All issues labeled labeled as defect and including an exception. & 0.00 & 3.25 & 0.00 & 13.00 & \includegraphics[scale = 0.1, clip = true, trim= 50px 60px 50px 60px]{hist-f1a879616aa3e0995db35c6c02bc8520.pdf} \\ 
  defect\_with\_stack & All issues labeled labeled as defect and including a stacktrace & 0.00 & 2.92 & 0.00 & 10.00 & \includegraphics[scale = 0.1, clip = true, trim= 50px 60px 50px 60px]{hist-49dbfa02c366cab8b489f33014de587d.pdf} \\ 
   \hline
\end{tabular}
\caption{Descriptive statistics for the analyzed issue dataset for Android projects. Historgrams are in log scale.} 
\label{tab:android-stack-stats}
\end{table*}

### 2014-05-06 android-stacks.tex
Initial commit

--- 2014-05-06 OLD_2main.tex
@@
 \item  Odd exception chains (e.g., Error wrapping checked and runtime exceptions and vice-versa). As an indication that the purpose of Java hybrid exception model may not have been adequately used.
-\item  Stack traces caused by uncaught checked exceptions.IMPROVE.
-
+\item  Stack traces caused by uncaught checked exceptions.
 \end{itemize}
@@
 
-\textbf{5) Document all exceptions thrown by a method}I~\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,bloch2008effective}).
+\textbf{5) Document all exceptions thrown by a method}~\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,bloch2008effective}).
 The exceptions thrown by a method are an important part of methods interface, and is required to use the method properly.
@@
 
-\noindent\emph{Exception Signaler Classifier} After its initialization (see Section 
-set of packages for each classification category and pattern matches the signaler name
-to classify it to one of the categories.
+\noindent\emph{Exception Signaler Classifier} After its initialization it classifies each signaler according to its origin (i.e.
+Android Plaform, Aplication, Library, Libcore, or Java) based on pattern matching between the signaler name
+and the packages associated to each category.
 
@@
 To answer this question, we mined the following information of every extracted stack trace: the top level exception being signaled, its signaler (the first method signature frame after the exception), the root exception (the exception which initiated the stack trace, defined on the bottom of the stack), and the root signaler.
-The intermediate elements that compose a stack trace as shown in Figure 1 (i.e., the intermediate causes and its execution stacks) were not be considered in the analysis.~\todo{Move this to the previous section?
-IMO, It is better to keep all facts about how we did the
-research in one place}
+The intermediate elements that compose a stack trace as shown in Figure 1 (i.e., the intermediate causes and its execution stacks) were not be considered in the analysis.
+%~\todo{Move this to the previous section? IMO, It is better to keep all facts about how we did the research in one place}
 
@@
 
-As detailed in Section III, 443 Android projects were carefully selected to enable a deaper investigation of stack trace information, by leveraging stack trace information with the source code and bytecode analysis.
+As detailed in Section III, 482 Android projects were carefully selected to enable a deaper investigation of stack trace information, by leveraging stack trace information with the source code and bytecode analysis.
 For this subset of Android projects we could extract every package related to each exception signaler, and based on this information we could classify every stack trace based on its root signaler as presented in Table~\ref{signalers}.
@@
 For each exception we could discover: the number of times it occurred on stacks, the number of distinct projects such stacks were defined (and consequently its popularity) and the of times it was signaled by different signaler types.
-We could calculate the popularity of each root cause, considering the number of distinct projects they were found and the whole set of projects analyzed with at least one stack trace (443).
+We could calculate the popularity of each root cause, considering the number of distinct projects they were found and the whole set of projects analyzed with at least one stack trace (482).
 Table~\ref{toptenandroid} and Figuere~\ref{androidsignaler} presents the mined data.
@@
 
-\textbf{We could observe in this study that the NullPointerException and other other implicitly signaled exceptions represented most of the exceptions reported on issues in both Java and Android subset.
+%commented only for submission:
+%\textbf{We could observe in this study that the NullPointerException and other other implicitly signaled exceptions represented most of the exceptions reported on issues in both Java and Android subset.
 For such exceptions - which represent programming bugs or resource limitations - there is usually no proper handling besides presenting an error message to the user and restarting the application - only high fault tolerant systems need to provide solutions to handle them [REF: Java Specification].}
@@
 
-TODO: Include a Histogram here.
+%TODO: Include a Histogram here.
 
@@
 
-To answer this question we analyzed general characteristics of stack traces reported on Java repositories (e.g., exceptions being thrown presented in Tables~\ref{toptenjava} and ~\ref{toptenandroid}), and took a deeper look at the stack traces of the Android repositories combining them with with source code and byte code analyses.
+To answer this question we analyzed general characteristics of stack traces reported on Java repositories (e.g., exceptions being thrown presented in Tables~\ref{tab:toptenjava} and ~\ref{tab:toptenandroid}), and took a deeper look at the stack traces of the Android repositories combining them with with source code and byte code analyses.
 Doing so, we mined some characteristics of stacks that points to best practices violations.
@@
 
-From Tables Tables~\ref{toptenjava}, ~\ref{toptenandroid} we can observe that, in both Java and Android related issues, direct instances java.lang.RuntimeException were thrown  (in 10,94\% of Java repositories and 18,88\% of Android repositories respectively).
+From Tables~\ref{tab:toptenjava}, ~\ref{tab:toptenandroid} we can observe that, in both Java and Android related issues, direct instances java.lang.RuntimeException were thrown  (in 10,94\% of Java repositories and 18,88\% of Android repositories respectively).
 Moreover, after inspecting the top 100 most thrown exceptions (see ~\ref{tophundrend}) we could observe that stack traces on which direct instances of java.lang.Exception and java.lang.Error were thrown.
@@
 
-\subsection{Undocumented runtime exceptions signaled by third party code}
+%todo{DISCOMMENT AFTER SUBMISSION}
+%\subsection{Undocumented runtime exceptions signaled by third party code}
+
+%In this study we also classified each stack trace according to the type of the root exception been signaled.
+Table~\ref{typeroottab} presents the types of root causes of all stack traces reported on the Android repositories.
+
+%TO DO: INCLUDE A NEW TABLE WITH NON JAVA.LANG RUNTIME EXCEPTIONS. Point to we cannot trust developers to document.
+
+%todo{discomment after submission}
+%As we can see from Table ~ref{typeroottab}, that most of the exceptions signaled by third-party code (e.g., Android and libs) were runtime exceptions.
+When an exception is signaled by third party code, it can be due to 2 reasons: the precondition associated to third party's code was not obeyed; or there is a coding error on third party's code.
+Only manual code inspection can reveal the reason.
+However, regardless the reason it is a good practice that all the exceptions signaled by a third part code are properly documented.
+Manually inspecting a sample of stack traces caused by runtime exceptions thrown by library code (approximately 10\%) we could observe that none of the runtime exceptions signaled by third part code were documented in Javadoc - this analysis did not consider the exceptions implicitly signaled by JVM (e.g.
+NullPoiterException).
+The undocumented runtime exceptions are a serious problem as it makes very difficult or even impossible for the  client method to protect against  undocumented runtime exceptions signaled by library code.
+As a consequence, the undocumented runtime exception may remain uncaught leading to  system crashes.
+
+%This result is aligned with the results of other study conducted by from Cabral and Marcques ~\cite{} which also observe that runtime exceptions are most often not documented.
+
+%\textbf{When libraries are used, the developer does not have access to their source code and thus needs to rely on the library documentation about the runtime exceptions that should be thrown – which, more often than not, are neither complete nor precise (Thomas, 2002; Sacramento et al., 2006).
+As a consequence, unhandled unchecked exceptions can be seen as one of the major sources of bugs in current Java systems (Jo et al., 2004).
+}
+
+%On the other hand, we could observe that most of the exceptions signaled by the set of libraries reused by Android platform (i.e., org.apache.harmony, org.w3c.dom, sun.misc, org.apache.http, org.json, org.xml, javax.) referred in this work as libcore, were checked exceptions.
+It is pointed as a good practice for libraries (see Section ~\ref{best}) because by using checked exceptions libraries can define a precise exception interface to its clients.
+Such libraries are heavily used in several projects, and such precise exception interface may be related to the libraries maturity.
+
+\subsection{Uncaught Checked Exceptions}
 
@@
 \end{figure}
-
-%TO DO: INCLUDE A NEW TABLE WITH NON JAVA.LANG RUNTIME EXCEPTIONS. Point to we cannot trust developers to document.
-
-As we can see from Table ~ref{typeroottab}, that most of the exceptions signaled by third-party code (e.g., Android and libs) were runtime exceptions. 
-When an exception is signaled by third party code, it can be due to 2 reasons: the precondition associated to third party's code was not obeyed; or there is a coding error on third party's code.
-Only manual code inspection can reveal the reason.
-However, regardless the reason it is a good practice that all the exceptions signaled by a third part code are properly documented.
-Manually inspecting a sample of stack traces caused by runtime exceptions thrown by library code (approximately 10\%) we could observe that none of the runtime exceptions signaled by third part code were documented in Javadoc - this analysis did not consider the exceptions implicitly signaled by JVM (e.g.
-NullPoiterException).
-The undocumented runtime exceptions are a serious problem as it makes very difficult or even impossible for the  client method to protect against  undocumented runtime exceptions signaled by library code.
-As a consequence, the undocumented runtime exception may remain uncaught leading to  system crashes.
-
-%This result is aligned with the results of other study conducted by from Cabral and Marcques ~\cite{} which also observe that runtime exceptions are most often not documented.
-
-%\textbf{When libraries are used, the developer does not have access to their source code and thus needs to rely on the library documentation about the runtime exceptions that should be thrown – which, more often than not, are neither complete nor precise (Thomas, 2002; Sacramento et al., 2006).
-As a consequence, unhandled unchecked exceptions can be seen as one of the major sources of bugs in current Java systems (Jo et al., 2004).
-}
-
-On the other hand, we could observe that most of the exceptions signaled by the set of libraries reused by Android platform (i.e., org.apache.harmony, org.w3c.dom, sun.misc, org.apache.http, org.json, org.xml, javax.) referred in this work as libcore, were checked exceptions.
-It is pointed as a good practice for libraries (see Section ~\ref{best}) because by using checked exceptions libraries can define a precise exception interface to its clients.
-Such libraries are heavily used in several projects, and such precise exception interface may be related to the libraries maturity.
-
-\subsection{Uncaught Checked Exceptions}
 
@@
   \end{tabular}
-\caption{Kinds of wrappings found on the stacks of 443 ANDROID projects found on Github}
+\caption{Kinds of wrappings found on the stacks of 482 ANDROID projects found on Github}
 \label{fig:wrappingandroid}
@@
 available and what the characteristics of the stack traces can tell us about the
-exception structure of those applications. Moreover this work\ldots.
+exception structure of those applications. %Moreover this work\ldots.
 

### 2014-05-06 java-descr-stats.tex
Initial commit

% latex table generated in R 2.15.3 by xtable 1.7-1 package
% Mon Apr 28 15:44:09 2014
\begin{table*}[ht]
\centering
\begin{tabular}{rp{16em}rrrrc}
  \hline
Feature & Description & quant\_5 & mean & median & quant\_95 & histogram \\ 
  \hline
loc & Lines of code. & 267.00 & 16211.15 & 4878.00 & 61719.60 & \includegraphics[scale = 0.1, clip = true, trim= 50px 60px 50px 60px]{hist-4fea7249a109b9e1d6791b4c1ebc1f38.pdf} \\ 
  ehloc & Lines of code devoted to exception handling. & 1.00 & 753.12 & 167.00 & 2759.00 & \includegraphics[scale = 0.1, clip = true, trim= 50px 60px 50px 60px]{hist-40db73293da278b6983709bc45b14f55.pdf} \\ 
  num\_classes & Number of user defined exception types. & 3.00 & 173.95 & 59.00 & 681.60 & \includegraphics[scale = 0.1, clip = true, trim= 50px 60px 50px 60px]{hist-027fd281b3f6bc7386445836eee5d87e.pdf} \\ 
  userdefined\_exceptions & Number of classes. & 0.00 & 3.09 & 0.00 & 14.00 & \includegraphics[scale = 0.1, clip = true, trim= 50px 60px 50px 60px]{hist-135e03395e7521809f10de2e03c70b19.pdf} \\ 
  num\_methods & Number of methods. & 12.00 & 1220.10 & 364.00 & 4564.00 & \includegraphics[scale = 0.1, clip = true, trim= 50px 60px 50px 60px]{hist-b6b0d4aa4024bfd6f76543e50b59b4a4.pdf} \\ 
  num\_methods\_throw & Number of methods that throw an exception. & 0.00 & 161.19 & 26.00 & 679.20 & \includegraphics[scale = 0.1, clip = true, trim= 50px 60px 50px 60px]{hist-762ba48e13fc4bdec8910cf5d8e488c1.pdf} \\ 
  num\_public\_methods & Number of public methods. & 8.00 & 907.93 & 260.00 & 3578.20 & \includegraphics[scale = 0.1, clip = true, trim= 50px 60px 50px 60px]{hist-6a589f87539a5232bab5b7cb76849876.pdf} \\ 
  num\_public\_methods\_throws & Number of public methods that throw an exception. & 0.00 & 114.16 & 16.00 & 507.80 & \includegraphics[scale = 0.1, clip = true, trim= 50px 60px 50px 60px]{hist-abd19a466fed24b2e5ab69fd7b2523df.pdf} \\ 
  num\_nonpublic\_methods & Number of private/protected/package methods. & 3.00 & 312.17 & 85.00 & 1281.60 & \includegraphics[scale = 0.1, clip = true, trim= 50px 60px 50px 60px]{hist-5ffac6360ae85e10530d281920f084d5.pdf} \\ 
  num\_nonpublic\_methods\_throws & Number of private/protected/package methods that throw an exception. & 0.00 & 47.02 & 7.00 & 169.00 & \includegraphics[scale = 0.1, clip = true, trim= 50px 60px 50px 60px]{hist-3ed6823532ea8acc01d29f33cf9379a8.pdf} \\ 
  num\_throw\_statements & Number of throw statements. & 0.00 & 109.38 & 18.00 & 437.60 & \includegraphics[scale = 0.1, clip = true, trim= 50px 60px 50px 60px]{hist-b25888a559b579f477c029e39bd403d8.pdf} \\ 
  num\_try & Number of try blocks. & 0.00 & 119.08 & 30.00 & 446.00 & \includegraphics[scale = 0.1, clip = true, trim= 50px 60px 50px 60px]{hist-37304fd9b59fe8289145f1008ad89ad9.pdf} \\ 
  num\_catch & Number of catch blocks. & 0.00 & 119.32 & 32.00 & 451.30 & \includegraphics[scale = 0.1, clip = true, trim= 50px 60px 50px 60px]{hist-afc847f0b6862e86823253de78fc0d16.pdf} \\ 
  throw\_on\_catch & Number of try blocks that throw an exception. & 0.00 & 29.24 & 4.00 & 127.00 & \includegraphics[scale = 0.1, clip = true, trim= 50px 60px 50px 60px]{hist-98dd667c113add4bb8ccce60ef1cb8d7.pdf} \\ 
  return\_on\_catch & Number of catch blocks with an explicit return statement. & 0.00 & 11.91 & 2.00 & 50.00 & \includegraphics[scale = 0.1, clip = true, trim= 50px 60px 50px 60px]{hist-f9a7775b7b8e5425ef21fcd168e80196.pdf} \\ 
  return\_on\_finall & Number of finally blocks with an explicit return statement. & 0.00 & 0.44 & 0.00 & 0.00 & \includegraphics[scale = 0.1, clip = true, trim= 50px 60px 50px 60px]{hist-5b9c377bd3fc66fbdbd3e345bf1f1665.pdf} \\ 
  log\_on\_finally & Number of finally blocks including logging statements. & 0.00 & 0.76 & 0.00 & 3.00 & \includegraphics[scale = 0.1, clip = true, trim= 50px 60px 50px 60px]{hist-5f1e0aa338cc8aba188c8b993766730e.pdf} \\ 
  specific\_action\_onfinally & Number of finally blocks including resource cleanup actions. & 0.00 & 28.81 & 0.00 & 84.00 & \includegraphics[scale = 0.1, clip = true, trim= 50px 60px 50px 60px]{hist-b7598a0e0c1a6992089680d29db9c202.pdf} \\ 
   \hline
\end{tabular}
\caption{Descriptive statistics for the Java dataset. Historgrams are in log scale.} 
\label{tab:java-descr-stats}
\end{table*}

### 2014-05-06 stacktrace-stats.tex
Initial commit

% latex table generated in R 2.15.3 by xtable 1.7-1 package
% Mon Apr 28 15:44:10 2014
\begin{table*}[ht]
\centering
\begin{tabular}{rp{27em}rrrrc}
  \hline
Feature & Description & quant\_5 & mean & median & quant\_95 & histogram \\ 
  \hline
all & All issues recorded for the project. & 1.00 & 23.37 & 3.00 & 80.00 & \includegraphics[scale = 0.1, clip = true, trim= 50px 60px 50px 60px]{hist-31b77719e3773ad7ac2073785de62a5f.pdf} \\ 
  with\_exception & All issues featuring an exception. & 0.00 & 3.21 & 0.00 & 6.00 & \includegraphics[scale = 0.1, clip = true, trim= 50px 60px 50px 60px]{hist-0c1305fc40f695dbdd1f4696db24f147.pdf} \\ 
  with\_stack & All issues featuring a stack trace. & 0.00 & 2.67 & 0.00 & 4.00 & \includegraphics[scale = 0.1, clip = true, trim= 50px 60px 50px 60px]{hist-896524802e915aa3f01ef1ecd2489e7e.pdf} \\ 
  defect\_issues & All issues labeled as defect. & 0.00 & 5.51 & 0.00 & 12.00 & \includegraphics[scale = 0.1, clip = true, trim= 50px 60px 50px 60px]{hist-5d0ab95cd3e300c21b0a40d6bec58e4c.pdf} \\ 
  defect\_with\_exception & All issues labeled labeled as defect and including an exception. & 0.00 & 1.82 & 0.00 & 1.00 & \includegraphics[scale = 0.1, clip = true, trim= 50px 60px 50px 60px]{hist-502028d1a7e2e77414e2f9f63e0bcd50.pdf} \\ 
  defect\_with\_stack & All issues labeled labeled as defect and including a stacktrace & 0.00 & 1.73 & 0.00 & 1.00 & \includegraphics[scale = 0.1, clip = true, trim= 50px 60px 50px 60px]{hist-dab0764147debd99066c84814d46d50e.pdf} \\ 
   \hline
\end{tabular}
\caption{Descriptive statistics for the analyzed issue dataset. Historgrams are in log scale.} 
\label{tab:stacktrace-stats}
\end{table*}

### 2014-05-07 android-stacks.tex
test
Split lines in 80 characters blocks

@@
 
-\author{\IEEEauthorblockN{Roberta Coelho\IEEEauthorrefmark{1}, Georgios Gousios\IEEEauthorrefmark{2}, Arie van Deursen\IEEEauthorrefmark{2}, Lucas Mariano\IEEEauthorrefmark{1}}
+\author{\IEEEauthorblockN{Roberta Coelho\IEEEauthorrefmark{1},
+Georgios Gousios\IEEEauthorrefmark{2}, Arie van Deursen\IEEEauthorrefmark{2},
+Lucas Mariano\IEEEauthorrefmark{1}}
 \IEEEauthorblockA{\IEEEauthorrefmark{1}Federal University of Rio Grande do Norte\\
@@
 Delft, The Netherlands\\
-Email: \texttt{\{gousiosg,a.vandeursen\}@tudelft.nl}}
+Email: \texttt{\{g.gousios,arie.vandeursen\}@tudelft.nl}}
 }
@@
 \begin{abstract}
-In this work, we perform the first large scale analysis of Java stack traces and pinpoint how they can reveal bad programming practices.
-We mined and analyzed the stack traces embedded in all issues of Java projects available GitHub.
-Overall, our research set includes 356,057 issues define at 16,836 projects (from which 28,800 stack traces were extracted).
-From this set, the stack traces of 482 Android projects were investigated in more detail, in combination with source code and bytecode analysis.
-In this study some patterns of exception (mis)-use were consistently detected such as: unexpected wrappings (e.g., Errors being wrapped in checked exceptions) - revealing that Java hybrid exception model is not fully used according to its purpose;  undocumented runtime exceptions signaled by third party code - which makes almost impossible for library clients to protect against such exceptions; and a high prevalence of java.lang exceptions reported on issues - representing approximately 50\% of the analyzed issues.
+
+In this work, we perform the first large scale analysis of Java stack traces and
+pinpoint how they can reveal bad programming practices. We mined and analyzed
+the stack traces embedded in all issues of Java projects available GitHub.
+Overall, our research set includes 356,057 issues define at 16,836 projects
+(from which 28,800 stack traces were extracted). From this set, the stack traces
+of 482 Android projects were investigated in more detail, in combination with
+source code and bytecode analysis. In this study some patterns of exception
+(mis)-use were consistently detected such as: unexpected wrappings (e.g., Errors
+being wrapped in checked exceptions) - revealing that Java hybrid exception
+model is not fully used according to its purpose;  undocumented runtime
+exceptions signaled by third party code - which makes almost impossible for
+library clients to protect against such exceptions; and a high prevalence of
+java.lang exceptions reported on issues - representing approximately 50\% of the
+analyzed issues.
 
@@
 
-Modern applications have to cope with an increasing number of abnormal computation states that arise as a consequence of faults in the application itself (e.g., access of null references), noisy user inputs or faults in underlying middleware or hardware.
-The exception handling mechanism~\cite{goodenough1975exception} is one of the most used schemes for detecting and recovering from such exceptional conditions.
-Although the exception handling mechanism have been embedded in several programming languages (e.g.
-Java, C++, C\#) and the target of several studies (e.g.~\cite{miller1997issues,robillard2000designing,shah2010understanding,garcia2007extracting,garcia2001comparative,cabral2007exception,coelho2011unveiling}), the exception handling code is often generally poorly understood and the least tested part of software systems.
-
-Often some languages constructions may lead the developers into believing that by just re-throwing the exceptions they can forget about the exceptional situations during the development of the "happy path".
-This "ignore-for-now" approach may turn the exception handling into a generalized "goto" mechanism~\cite{mandrioli1992advances} making the program more complex and even less reliable.
-This behaviour may lead to the well known uncaught exceptions problem~\cite{jo2004uncaught}, one of the main causes of application crashes.
-
-Moreover, some languages use exceptions to represent coding errors (out-of-bounds array index, division-by-zero, access to a null).
-In these cases exceptions are implicitly signaled by the runtime environment when such condition happens.
-One of such implicit exceptions (caused by data conversion from a 64-bit floating point to a 16-bit signed integer) was the responsible for the famous failure on Ariane 5's first test flight~\cite{lions1996ariane} --- leading to the rocket self-destructing 37 seconds after launch and a loss of 500 million dollars.
-Besides Ariane 5, several applications crash everyday due to uncaught exceptions~\cite{jo2004uncaught}.
-The consequences of not careful exception usage may have the opposite effect of its initial intention of improving system robustness.
-
-In Java, when the program fails due to an uncaught exception, it automatically terminates, while the system prints a
-stack trace to the console.
-A stack trace includes a detail message and the execution stack frame when the exception occurred.
-A typical Java stack trace consists of an ordered list of methods that were active on the call stack before the exception has occurred.
-Stack traces are a useful source of information about system crashes, and is often used to support developers in debugging~\cite{schroter2010stack}.
-Moreover, information from stack traces combined with crash and bug reports can be used to support bug classification and clustering~\cite{wang2013improving, kim2011crash, dhaliwal2011classifying}, fault-proneness prediction models~\cite{kim2013predicting} and even automated bug fixing~\cite{sinha2009fault} tools.
-
-However, all these works focus on the analysis of a single system at a time.
-In our current context, on which we have plenty access to the information available on open-source repositories like GitHub initial questions arise: What is the prevalence of exception stack traces on reported issues?
-What are the common characteristics of such stack traces?
-
-%What the stack traces available in these repositories can tell us about the use of exceptions in Java programs?
-
-%To guide the our study we also compile general guidelines on how to use exceptions proposed by Gosling~\cite{gosling2000java}, Wirfs-Brock~\cite{wirfs2006toward} and Bloch~\cite{bloch2008effective} and then we investigate: what stack traces can tell us about the adherence to such practices in Java programs?
-
-To answer these questions, we conduct an exploratory study of stack traces in the wild.
-Using a custom tool called ExceptionMiner, that was specifically developed to for this study, we mined stack traces from issues across all Java projects on GitHub.
-Overall, we analyzed 356,057 issues from 16,836 projects, from which 28,800 stack traces were extracted and used as a source of information for our analysis.
-The stack trace analysis  was augmented by additional information extracted using byte code and source code analyses done on
-a carefully selected subset of our initial sample (482 Android projects).
-%This study was divided in two phases: in the fist phase all the issues of all Java projects available on GitHub were analyzed and the embedded stack traces extracted using ExceptionMiner.
-
-To guide this exploration we compiled general guidelines on how to use exceptions proposed by Gosling~\cite{gosling2000java}, Wirfs-Brock~\cite{wirfs2006toward} and Bloch~\cite{bloch2008effective} and then we investigate what stack traces can tell us about the adherence to such practices in Java programs.
-.%~\todo{This should be our motivation, as this is our biggest contribution. I already changed the text. Lets discuss about this new version}
-
-%GGG: I think this is too much detail for the introduction
-%In the second phase of this study, we combined the information of the mined stack traces with source code and bytecode analysis.
-To do so we selected the subset of 482 Android repositories.
-The Android repositories were chosen because this kind of application is known for dealing with several sources of exceptions, as they are usually io-intensive (i.e., deals with Network, Bluetooth, camera, and other resources), and multi-htreaded.
-Moreover, they are based on a single platform which allowed more fine-grained analysis of their exception behaviour.
-
-This study is built on the following assumption: the stack traces reported on issues contain relevant information related to system crashes.
-Several studies and techniques have also based on this assumption (e.g.,~\cite{sinha2009fault}~\cite{dhaliwal2011classifying}~\cite{kim2013predicting}) and we build on it as well.
-Some outcomes were consistently detected through this large scale analysis of exception stack traces in the wild, such as:
+Modern applications have to cope with an increasing number of abnormal
+computation states that arise as a consequence of faults in the application
+itself (e.g., access of null references), noisy user inputs or faults in
+underlying middleware or hardware. The exception handling
+mechanism~\cite{goodenough1975exception} is one of the most used schemes for
+detecting and recovering from such exceptional conditions. Although the
+exception handling mechanism have been embedded in several programming languages
+(e.g. Java, C++, C\#) and the target of several studies
+(e.g.~\cite{miller1997issues,robillard2000designing,shah2010understanding,
+garcia2007extracting,garcia2001comparative,cabral2007exception,coelho2011unveiling}),
+the exception handling code is often generally poorly understood and the least
+tested part of software systems.
+
+Often some languages constructions may lead the developers into believing that
+by just re-throwing the exceptions they can forget about the exceptional
+situations during the development of the "happy path". This "ignore-for-now"
+approach may turn the exception handling into a generalized "goto"
+mechanism~\cite{mandrioli1992advances} making the program more complex and even
+less reliable. This behaviour may lead to the well known uncaught exceptions
+problem~\cite{jo2004uncaught}, one of the main causes of application crashes.
+
+Moreover, some languages use exceptions to represent coding errors
+(out-of-bounds array index, division-by-zero, access to a null). In these cases
+exceptions are implicitly signaled by the runtime environment when such
+condition happens. One of such implicit exceptions (caused by data conversion
+from a 64-bit floating point to a 16-bit signed integer) was the responsible for
+the famous failure on Ariane 5's first test flight~\cite{lions1996ariane} ---
+leading to the rocket self-destructing 37 seconds after launch and a loss of 500
+million dollars. Besides Ariane 5, several applications crash everyday due to
+uncaught exceptions~\cite{jo2004uncaught}. The consequences of not careful
+exception usage may have the opposite effect of its initial intention of
+improving system robustness.
+
+In Java, when the program fails due to an uncaught exception, it automatically
+terminates, while the system prints a stack trace to the console. A stack trace
+includes a detail message and the execution stack frame when the exception
+occurred. A typical Java stack trace consists of an ordered list of methods that
+were active on the call stack before the exception has occurred. Stack traces
+are a useful source of information about system crashes, and is often used to
+support developers in debugging~\cite{schroter2010stack}. Moreover, information
+from stack traces combined with crash and bug reports can be used to support bug
+classification and clustering~\cite{wang2013improving, kim2011crash,
+dhaliwal2011classifying}, fault-proneness prediction
+models~\cite{kim2013predicting} and even automated bug
+fixing~\cite{sinha2009fault} tools.
+
+However, all these works focus on the analysis of a single system at a time. In
+our current context, on which we have plenty access to the information available
+on open-source repositories like GitHub initial questions arise: What is the
+prevalence of exception stack traces on reported issues? What are the common
+characteristics of such stack traces?
+
+%What the stack traces available in these repositories can tell us about the use
+%of exceptions in Java programs?
+
+%To guide the our study we also compile general guidelines on how to use
+%exceptions proposed by Gosling~\cite{gosling2000java},
+%Wirfs-Brock~\cite{wirfs2006toward} and Bloch~\cite{bloch2008effective} and then
+%we investigate: what stack traces can tell us about the adherence to such
+%practices in Java programs?
+
+To answer these questions, we conduct an exploratory study of stack traces in
+the wild. Using a custom tool called ExceptionMiner, that was specifically
+developed to for this study, we mined stack traces from issues across all Java
+projects on GitHub. Overall, we analyzed 356,057 issues from 16,836 projects,
+from which 28,800 stack traces were extracted and used as a source of
+information for our analysis. The stack trace analysis  was augmented by
+additional information extracted using byte code and source code analyses done
+on a carefully selected subset of our initial sample (482 Android projects).
+%This study was divided in two phases: in the fist phase all the issues of all
+%Java projects available on GitHub were analyzed and the embedded stack traces
+%extracted using ExceptionMiner.
+
+To guide this exploration we compiled general guidelines on how to use
+exceptions proposed by Gosling~\cite{gosling2000java},
+Wirfs-Brock~\cite{wirfs2006toward} and Bloch~\cite{bloch2008effective} and then
+we investigate what stack traces can tell us about the adherence to such
+practices in Java programs.
+%~\todo{This should be our motivation, as this is
+%our biggest contribution. I already changed the text. Lets discuss about this
+%new version}
+
+%GGG: I think this is too much detail for the introduction In the second phase
+%of this study, we combined the information of the mined stack traces with
+%source code and bytecode analysis. To do so we selected the subset of 482
+%Android repositories. The Android repositories were chosen because this kind of
+%application is known for dealing with several sources of exceptions, as they
+%are usually io-intensive (i.e., deals with Network, Bluetooth, camera, and
+%other resources), and multi-htreaded. Moreover, they are based on a single
+%platform which allowed more fine-grained analysis of their exception behaviour.
+
+This study is built on the following assumption: the stack traces reported on
+issues contain relevant information related to system crashes. Several studies
+and techniques have also based on this assumption
+(e.g.,~\cite{sinha2009fault}~\cite{dhaliwal2011classifying}~\cite{kim2013predicting})
+and we build on it as well. Some outcomes were consistently detected through
+this large scale analysis of exception stack traces in the wild, such as:
 
 \begin{itemize}
-\item  A multitude undocumented runtime exceptions signaled by third party code (i.e., libraries and Android platform).
-\item  Odd exception chains (e.g., Error wrapping checked and runtime exceptions and vice-versa). As an indication that the purpose of Java hybrid exception model may not have been adequately used.
+
+  \item  A multitude undocumented runtime exceptions signaled by third party
+    code (i.e., libraries and Android platform).
+
+  \item  Odd exception chains (e.g., Error wrapping checked and runtime
+    exceptions and vice-versa). As an indication that the purpose of Java hybrid
+    exception model may not have been adequately used.
+
   \item  Stack traces caused by uncaught checked exceptions.
+
 \end{itemize}
@@
 \begin{itemize}
-\item  It performs the first large scale analysis of Java stack traces and how they can reveal bad-practices on the use of exceptions in Java. 
-\item  It briefly introduces ExceptionMiner, a tool developed to support the analysis.
+
+  \item  It performs the first large scale analysis of Java stack traces and how
+    they can reveal bad-practices on the use of exceptions in Java.
+
+  \item  It briefly introduces ExceptionMiner, a tool developed to support the
+    analysis.
+
 \end{itemize}
 
-The contributions of this work allow for developers of robust Java applications to familiarize with the most common reported exceptions mis-uses which is the first step to help developers to avoid making them; (ii) designers of languages to consider ways of reducing the abundance of NullPointerExceptions, (iii) and tool designers to consider developing tools that enable a cross-project defect analysis.
-We strongly believe that, specially in the open-source environment, faults are not to be hidden in a private bug issue.
-Faults should be share and discussed, so that a developer can learn from other projects mistakes.
-Currently, the search facilities of repositories are very limited, and the ExceptionMiner tool is a contribution in this direction.
+The contributions of this work allow for developers of robust Java applications
+to familiarize with the most common reported exceptions mis-uses which is the
+first step to help developers to avoid making them; (ii) designers of languages
+to consider ways of reducing the abundance of NullPointerExceptions, (iii) and
+tool designers to consider developing tools that enable a cross-project defect
+analysis. We strongly believe that, specially in the open-source environment,
+faults are not to be hidden in a private bug issue. Faults should be share and
+discussed, so that a developer can learn from other projects mistakes.
+Currently, the search facilities of repositories are very limited, and the
+ExceptionMiner tool is a contribution in this direction.
 
 %GGG: Personally, I hate this paragraph in all papers as it usually reads like a space-filler
-The remainder of this paper is organized as follows.
-Section 2 presents a background on error handling mechanisms and Java exception handling constructs and practices.
-Section 3 defines research questions, the configuration of our exploratory study, the target applications and the evaluation procedures.
-Section 4 reports our analysis of the empirical data collected relating to the first research questions.
-Section X is dedicated to provide insights for the 3rd research question.
-Section 6 provides further discussions and lessons learned and discusses the threats to validity associated to this study.
-Finally Section 7 describes the related work, and Section 8 presents our conclusions and directions for future work.
+The remainder of this paper is organized as follows. Section 2 presents a
+background on error handling mechanisms and Java exception handling constructs
+and practices. Section 3 defines research questions, the configuration of our
+exploratory study, the target applications and the evaluation procedures.
+Section 4 reports our analysis of the empirical data collected relating to the
+first research questions. Section X is dedicated to provide insights for the 3rd
+research question. Section 6 provides further discussions and lessons learned
+and discusses the threats to validity associated to this study. Finally Section
+7 describes the related work, and Section 8 presents our conclusions and
+directions for future work.
 
@@
 
-Techniques for error detection and handling are no more an optional add-on but a fundamental part of the system~\cite{bruntink2006discovering}.
-Several approaches exists for detecting errors during system execution and responding to them
-accordingly, such as the return-code idiom (very popular in C programs and operating systems), and the exception handling mechanism, embedded in many main stream programming languages.
-
-The main disadvantage of return codes is the fact that the programmer can just ignore the return code and let the program execute in inconsistent state.
-The exception mechanism~\cite{goodenough1975exception} was proposed as a way of preventing errors from being ignored, and was embedded in many mainstream programming languages.
-Among some similarities they mainly differ in the way exceptions are represented and handled in each language~\cite{garcia2001comparative}:
-
-\noindent\emph{Checked Approach}.
-In languages such Modula-3, Guide, and Extended Ada all exceptions are checked, which means that the compiler statically checks if appropriate handlers are provided within the system.
-Accordingly, the programmer must explicitly specify every exception that can be signaled to support the compile time checking.
-This approach is known to be costly to maintain (Dooren and Steegmans, 2005) since exception updates are cascading; if a new exception type is added to a method signature then all other methods directly using this method must either change their
-signatures to include the new exception type or handle the new exception.
-
-\noindent\emph{Unchecked Approach}.
-In languages such as Lore, C++, and Arche all exceptions are unchecked but the developer can optionally list exception types on method signatures.
-However, the developer is not warned by the compiler if an unchecked exception is not handled inside the application.
-Such an approach may hinder robustness because the client of a method cannot easily know which unchecked exceptions may be thrown, unless the code of the method and the methods called from it is inspected, which can be a very time consuming or infeasible task.
-
-\noindent\emph{Hybrid Approach}. In Java a method can throw either checked or unchecked exceptions. Java is the only language that allows this hybrid approach.
-
-\subsection{Exception handling in Java}
-\label{sec:extypes}
-
-In Java, an exception is represented according to the class hierarchy shown in Figure~\ref{fig:exchier}. 
-According to this structure every exception is an instance of the Throwable class, and can be of three kinds: checked exceptions (extends Exception), run-time exceptions (extends RuntimeException), and errors (extends Error)~\cite{gosling2000java}.
-
-\begin{figure}
-\centering
-\includegraphics[width=\hsize]{new2_hierarchy.png}
-\caption{Exception Hierarchy in Java}
-\label{fig:exchier}
-\end{figure}
-
-Neither run-time exceptions nor errors need to be specified on method signatures, and hence both are referred as unchecked exceptions.
-The Java Virtual Machine represents as runtime exceptions invalid operations detected in the program (out-of-bounds array index, divide-by-zero error, null pointer references) from which most of the programs are not expected to recover.
-By convention an Error represents an unrecoverable condition which usually results from failures detected by the Java Virtual Machine, such as OutOfMemoryError and normally should not be handled inside the application.
-
-The Java specification~\cite{gosling2000java} suggests that user-defined exceptions should be checked exceptions.
-The main reason is because doing so the callers of a method will know about the exceptions that a method can throw, and so they can decide what to do about them.
-However, there is a long-lasting debate about the use of checked and unchecked exceptions in Java~\cite{javatut,stackoverlow,debate}\footnote{152 questions in Stackoverlow are related to this debate}, since there are pros and cons associated to each of them.
+Techniques for error detection and handling are no more an optional add-on but a
+fundamental part of the system~\cite{bruntink2006discovering}. Several
+approaches exists for detecting errors during system execution and responding to
+them accordingly, such as the return-code idiom (very popular in C programs and
+operating systems), and the exception handling mechanism, embedded in many main
+stream programming languages.
+
+The main disadvantage of return codes is the fact that the programmer can just
+ignore the return code and let the program execute in inconsistent state. The
+exception mechanism~\cite{goodenough1975exception} was proposed as a way of
+preventing errors from being ignored, and was embedded in many mainstream
+programming languages. Among some similarities they mainly differ in the way
+exceptions are represented and handled in each
+language~\cite{garcia2001comparative}:
+
+\noindent\emph{Checked Approach}. In languages such Modula-3, Guide, and
+Extended Ada all exceptions are checked, which means that the compiler
+statically checks if appropriate handlers are provided within the system.
+Accordingly, the programmer must explicitly specify every exception that can be
+signaled to support the compile time checking. This approach is known to be
+costly to maintain (Dooren and Steegmans, 2005) since exception updates are
+cascading; if a new exception type is added to a method signature then all other
+methods directly using this method must either change their signatures to
+include the new exception type or handle the new exception.
+
+\noindent\emph{Unchecked Approach}. In languages such as Lore, C++, and Arche
+all exceptions are unchecked but the developer can optionally list exception
+types on method signatures. However, the developer is not warned by the compiler
+if an unchecked exception is not handled inside the application. Such an
+approach may hinder robustness because the client of a method cannot easily know
+which unchecked exceptions may be thrown, unless the code of the method and the
+methods called from it is inspected, which can be a very time consuming or
+infeasible task.
+
+\noindent\emph{Hybrid Approach}. In Java a method can throw either checked or
+unchecked exceptions. Java is the only language that allows this hybrid
+approach.
+
+\subsection{Exception handling in Java} \label{sec:extypes}
+
+In Java, an exception is represented according to the class hierarchy shown in
+Figure~\ref{fig:exchier}.  According to this structure every exception is an
+instance of the Throwable class, and can be of three kinds: checked exceptions
+(extends Exception), run-time exceptions (extends RuntimeException), and errors
+(extends Error)~\cite{gosling2000java}.
+
+\begin{figure} \centering \includegraphics[width=\hsize]{new2_hierarchy.png}
+  \caption{Exception Hierarchy in Java} \label{fig:exchier} \end{figure}
+
+Neither run-time exceptions nor errors need to be specified on method
+signatures, and hence both are referred as unchecked exceptions. The Java
+Virtual Machine represents as runtime exceptions invalid operations detected in
+the program (out-of-bounds array index, divide-by-zero error, null pointer
+references) from which most of the programs are not expected to recover. By
+convention an Error represents an unrecoverable condition which usually results
+from failures detected by the Java Virtual Machine, such as OutOfMemoryError and
+normally should not be handled inside the application.
+
+The Java specification~\cite{gosling2000java} suggests that user-defined
+exceptions should be checked exceptions. The main reason is because doing so the
+callers of a method will know about the exceptions that a method can throw, and
+so they can decide what to do about them. However, there is a long-lasting
+debate about the use of checked and unchecked exceptions in
+Java~\cite{javatut,stackoverlow,debate}\footnote{152 questions in Stackoverlow
+are related to this debate}, since there are pros and cons associated to each of
+them.
 
@@
 
-In Java an exception can be signaled in one of the following circumstances~\cite{gosling2000java}:
-
-\begin{itemize}
-\item  explicitly thrown when a throw statement is executed;
-\item  implicitly thrown by the JVM when the evaluation of an expression violates the normal semantics of language (e.g., out-of-bounds array index, division-by-zero, access to a null reference); or 
-\item implicitly thrown by the JVM due to an internal error or resource limitation (e.g., OutofMemoryError).
-\end{itemize}
-%\todo{This sub-section is quite similar to the section introduction, shouldn't we merge them?}
-When an exception is signaled it causes a transfer of control from the point where the exception occurred to a point that can be specified by the programmer (exception handler).
-In Java it is represented by the try-catch block and is responsible for performing any recovery action necessary to bring the software system back to a normal state and, whenever this is not possible, to log the exception and abort the system in an expected safe way.
-
-In Java, as in many languages, the search for the handler to deal with a raised exception occurs along the dynamic invocation chain.
-This is claimed to increase the reusability, since the invoker of an operation can handle it in a wider context ~\cite{goodenough1975exception}.
-If no handler is found in the chain, the default handler associated to the Thread or ThreadGroup is invoked, which usually terminates the system and prints the exception stack trace.
-
-\subsubsection{Exception Chaining}
-Exception Chaining is a common way of propagating exceptions across components of different abstraction levels.
-It is implemented by special exception constructor which receives another exception as a parameter.
-An example a multiple exception chaining is depicted on the stack trace in Figure~\ref{fig:wrapping}.
-%This stack trace was extracted from an issue of an Android application in Github.
-
-The bottom part of the stack trace is the \emph{root cause}, which indicates the first reason for the error to be thrown (in this case, the computer run out of memory).
-The top part of the stack trace indicates the location of the exception manifestation (which will call \emph{top level exception} along this paper).
-The execution flow  between the root cause and the exception manifestation may include other intermediate exception wrappings.
-In all levels, the exception \emph{signaler}, is the method that threw the exception, represented on the stack trace as the first method call below the exception declaration.
-
-\begin{figure}
-\centering
-\includegraphics[scale=0.6]{wrappings}
+In Java an exception can be signaled in one of the following
+circumstances~\cite{gosling2000java}:
+
+\begin{itemize} \item  explicitly thrown when a throw statement is executed;
+  \item  implicitly thrown by the JVM when the evaluation of an expression
+    violates the normal semantics of language (e.g., out-of-bounds array index,
+    division-by-zero, access to a null reference); or \item implicitly thrown by
+    the JVM due to an internal error or resource limitation (e.g.,
+    OutofMemoryError).  \end{itemize}
+%\todo{This sub-section is quite similar to the section introduction, shouldn't
+%we merge them?}
+When an exception is signaled it causes a transfer of control from the point
+where the exception occurred to a point that can be specified by the programmer
+(exception handler). In Java it is represented by the try-catch block and is
+responsible for performing any recovery action necessary to bring the software
+system back to a normal state and, whenever this is not possible, to log the
+exception and abort the system in an expected safe way.
+
+In Java, as in many languages, the search for the handler to deal with a raised
+exception occurs along the dynamic invocation chain. This is claimed to increase
+the reusability, since the invoker of an operation can handle it in a wider
+context ~\cite{goodenough1975exception}. If no handler is found in the chain,
+the default handler associated to the Thread or ThreadGroup is invoked, which
+usually terminates the system and prints the exception stack trace.
+
+\subsubsection{Exception Chaining} Exception Chaining is a common way of
+propagating exceptions across components of different abstraction levels. It is
+implemented by special exception constructor which receives another exception as
+a parameter. An example a multiple exception chaining is depicted on the stack
+trace in Figure~\ref{fig:wrapping}.
+%This stack trace was extracted from an issue of an Android application in
+%Github.
+
+The bottom part of the stack trace is the \emph{root cause}, which indicates the
+first reason for the error to be thrown (in this case, the computer run out of
+memory). The top part of the stack trace indicates the location of the exception
+manifestation (which will call \emph{top level exception} along this paper). The
+execution flow  between the root cause and the exception manifestation may
+include other intermediate exception wrappings. In all levels, the exception
+\emph{signaler}, is the method that threw the exception, represented on the
+stack trace as the first method call below the exception declaration.
+
+\begin{figure} \centering \includegraphics[scale=0.6]{wrappings}
 \caption{Example of an Exception stack trace in Java.}
@@
 \label{sec:best}
+
 %\todo{This section feels a bit disconnected.}
-Several general guidelines have been proposed on how to use exceptions~\cite{mandrioli1992advances,gosling2000java,wirfs2006toward, bloch2008effective}.
-To guide our investigation, we compile general guidelines related to exception types and the propagation of exception in Java programs \footnote{We also compiled guidelines related to exception handling but they are out of the scope of this paper, since they do not relate to information available on stack traces.}.
+Several general guidelines have been proposed on how to use
+exceptions~\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,
+bloch2008effective}. To guide our investigation, we compile general guidelines
+related to exception types and the propagation of exception in Java programs
+\footnote{We also compiled guidelines related to exception handling but they are
+out of the scope of this paper, since they do not relate to information
+available on stack traces.}.
 
@@
 
-\textbf{1) Use checked exceptions to represent recoverable conditions}(~\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,bloch2008effective}).
-The developer should use checked exceptions for conditions from which the caller is expected to recover.
-By confronting the API user with a checked exception, the API designer is telling the client to handle the exceptional condition.
-The client can explicitly ignore the exception (swallowing, or converting it to other type) at
-the expense of the program's robustness~\cite{gosling2000java}.
-
-\textbf{2) Error represents an unrecoverable condition detected by the JVM which should not he handled}~\cite{gosling2000java}.
-It results from failures detected by the Java Virtual Machine which indicate resource deficiencies, invariant failures or other conditions that make impossible the program to recover.
+\textbf{1) Use checked exceptions to represent recoverable
+conditions}(~\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,bloch2008effective}).
+The developer should use checked exceptions for conditions from which the caller
+is expected to recover. By confronting the API user with a checked exception,
+the API designer is telling the client to handle the exceptional condition. The
+client can explicitly ignore the exception (swallowing, or converting it to
+other type) at the expense of the program's robustness~\cite{gosling2000java}.
+
+\textbf{2) Error represents an unrecoverable condition detected by the JVM which
+should not he handled}~\cite{gosling2000java}. It results from failures detected
+by the Java Virtual Machine which indicate resource deficiencies, invariant
+failures or other conditions that make impossible the program to recover.
 
@@
 
-\textbf{3) A method should throw exceptions that precisely define the exceptional condition}~\cite{gosling2000java,bloch2008effective}.
-To do so, developers should either try to reuse the exception types already defined in the JVM or they should create a specific exception.
-Throwing general types such as a pure Exception or a RuntimeException is hence considered a bad practice.
-
-\textbf{4) Do not repeatedly re-throw exceptions, handle exceptions as close as possible to the problem}~\cite{wirfs2006toward}.
-Programs that frequently throw exceptions lose in performance ~\cite{wirfs2006toward,gosling2000java}.
-Moreover, close to the place that the exception was throw the caller usually has enough contextual information to perform  corrective actions.
-When the exception is propagated far away from the method where it was signaled (i.e.
-further away from the problem) it can be difficult to take meaningful recovery actions.
+\textbf{3) A method should throw exceptions that precisely define the
+exceptional condition}~\cite{gosling2000java,bloch2008effective}. To do so,
+developers should either try to reuse the exception types already defined in the
+JVM or they should create a specific exception. Throwing general types such as a
+pure Exception or a RuntimeException is hence considered a bad practice.
+
+\textbf{4) Do not repeatedly re-throw exceptions, handle exceptions as close as
+possible to the problem}~\cite{wirfs2006toward}. Programs that frequently throw
+exceptions lose in performance ~\cite{wirfs2006toward,gosling2000java}.
+Moreover, close to the place that the exception was throw the caller usually has
+enough contextual information to perform  corrective actions. When the exception
+is propagated far away from the method where it was signaled (i.e. further away
+from the problem) it can be difficult to take meaningful recovery actions.
 
@@
 
-\textbf{5) Document all exceptions thrown by a method}~\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,bloch2008effective}).
-The exceptions thrown by a method are an important part of methods interface, and is required to use the method properly.
-The checked exceptions are already part of the  methods signature, and the method caller is aware of the checked exceptions being thrown by it.
-According to~\cite{bloch2008effective}, it is also wise to document the explicitly signaled runtime exceptions\footnote{This excludes the implicitly signaled JVM runtime exceptions related to coding errors} as carefully as checked exceptions.
-Doing so, the clients of a method will be aware of the exceptions the method can throw.
-If the developer fails to do follow this practice (specially when developing library code) it will be difficult or even impossible for the caller to make effective use of such method ~\cite{wirfs2006toward, bloch2008effective}.
-
-These best practices guided our exploratory study described in the next sections.
-Based on information extracted from stack traces (complemented by bytecode and source-code analysis), we investigate whether stack characteristics can reveal whether these practices have been obeyed.
-Next section discusses the study procedure and the practices' misuses that emerged from this study.
+\textbf{5) Document all exceptions thrown by a
+method}~\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,bloch2008effective}).
+The exceptions thrown by a method are an important part of methods interface,
+and is required to use the method properly. The checked exceptions are already
+part of the  methods signature, and the method caller is aware of the checked
+exceptions being thrown by it. According to~\cite{bloch2008effective}, it is
+also wise to document the explicitly signaled runtime exceptions\footnote{This
+excludes the implicitly signaled JVM runtime exceptions related to coding
+errors} as carefully as checked exceptions. Doing so, the clients of a method
+will be aware of the exceptions the method can throw. If the developer fails to
+do follow this practice (specially when developing library code) it will be
+difficult or even impossible for the caller to make effective use of such method
+~\cite{wirfs2006toward, bloch2008effective}.
+
+These best practices guided our exploratory study described in the next
+sections. Based on information extracted from stack traces (complemented by
+bytecode and source-code analysis), we investigate whether stack characteristics
+can reveal whether these practices have been obeyed. Next section discusses the
+study procedure and the practices' misuses that emerged from this study.
 
@@
 
-This work describes an exploratory study which was based on a sequential mixed-methods approach~\cite{ivankova2006using} for collecting, analyzing, and integrating both quantitative and qualitative analysis along the stages of the research process.
-The main goal of this study was to gain a better understanding of the use of exceptions in Java programs based mainly on the information available on exception stack traces.
-More specifically, we aimed at answering the following research questions:
+This work describes an exploratory study which was based on a sequential
+mixed-methods approach~\cite{ivankova2006using} for collecting, analyzing, and
+integrating both quantitative and qualitative analysis along the stages of the
+research process. The main goal of this study was to gain a better understanding
+of the use of exceptions in Java programs based mainly on the information
+available on exception stack traces. More specifically, we aimed at answering
+the following research questions:
 
 \begin{description}
+
   \item[RQ1] What is the prevalence of stack traces on  reported issues?
     %\item[RQ2] What are the most common exceptions?
+
   \item[RQ2] What are the common characteristics of such stack traces?
-    \item[RQ3] What patterns of exception (mis)-use emerge from the stack trace analysis?
-    %~\todo{Perhaps rephrase this to: Are the identified best practices followed in practice?}
+
+  \item[RQ3] What patterns of exception (mis)-use emerge from the stack trace
+    analysis?
+    %~\todo{Perhaps rephrase this to: Are the identified best practices followed
+    %in practice?}
+
 \end{description}
 
-For RQ1 and RQ2, we explore the domain quantitatively, and use theirs answers as a basis to investigate the third research question.
-RQ3 aimed at finding whether of exception (mis)-uses could be revealed through stack trace analysis.
-To support this investigation, source code and bytecode analysis was used to leverage the  understanding of stack traces.
-For RQ3 we investigate quantitatively and highlight interesting cases by exploring cases qualitatively.
-Figure~\ref{fig:overviewfig} presents an overview of the mixed-methods approach conducted to answer these questions.
+For RQ1 and RQ2, we explore the domain quantitatively, and use theirs answers as
+a basis to investigate the third research question. RQ3 aimed at finding whether
+of exception (mis)-uses could be revealed through stack trace analysis. To
+support this investigation, source code and bytecode analysis was used to
+leverage the  understanding of stack traces. For RQ3 we investigate
+quantitatively and highlight interesting cases by exploring cases qualitatively.
+Figure~\ref{fig:overviewfig} presents an overview of the mixed-methods approach
+conducted to answer these questions.
 
@@
 
-We begun our investigation using the dataset provided by the GHTorrent project~\cite{Gousi13}, an off-line mirror of the data offered through the Github API.
-The Github API data come in two forms; a streaming data flow lists events, such as reporting an issue, happening on repositories in real time, while a static view contains the current state of entities.
-The data is stored in unprocessed format, in a MongoDB database, while metadata is extracted and stored in a relational database.
-The GHTorrent dataset covers a broad range of development activities on Github, including issues.
-The project has been collecting data since February 2012.
-Up to Dec 2013, when we queried it to retrieve our bootstrapping sample which included: 356.057 issues from 16.836 Java projects (step 1).\footnote{A Java project is specified in GitHub with a specific label.}
-
-After obtaining the bootstrap issue sample from GHTorrent, each issue was processed by ExceptionMiner tool developed in this work which extracted every exception defined on issues or on issue comments (step 2).
-Since this initial set of Java project is too diverse (it includes libraries, web-based or stand-alone applications), in the second phase of this study we restrict the analysis domain to Android projects (step 3).
-This decision was driven by the fact that most open source Android applications are known~\cite{ref} for dealing with several sources of exceptions due to multi-threaded execution and interaction with input/output resources; moreover they share a common well-defined platform and a limited set of libraries.
-
-To support a deeper investigation of the stack traces of Android apps, every exception defined on a stack trace was then classified according to its type (e.g.
-Error, Exception or RuntimeException) (step 4).
-The exception classifier was initialized with all the exceptions defined on the Java SDK version 7, the exceptions of the Android (Platform and OS) (level 19), and all the exceptions of the 482 Android projects selected for this study, extracted in the previous step.
-Moreover, all the packages associated to exceptions and signalers were also extracted to support further analysis (step 5).
-The signaler classifier is initialized with information about all packages found on Java SDK version 7, and Android Platform and OS (level 19).
-Finally, additional meta-data is combined to each stack trace (step 6) which used together with manual inspections (step 7) enabled to provide answers to RQ3.
+We begun our investigation using the dataset provided by the GHTorrent
+project~\cite{Gousi13}, an off-line mirror of the data offered through the
+Github API. The Github API data come in two forms; a streaming data flow lists
+events, such as reporting an issue, happening on repositories in real time,
+while a static view contains the current state of entities. The data is stored
+in unprocessed format, in a MongoDB database, while metadata is extracted and
+stored in a relational database. The GHTorrent dataset covers a broad range of
+development activities on Github, including issues. The project has been
+collecting data since February 2012. Up to Dec 2013, when we queried it to
+retrieve our bootstrapping sample which included: 356.057 issues from 16.836
+Java projects (step 1).\footnote{A Java project is specified in GitHub with a
+specific label.}
+
+After obtaining the bootstrap issue sample from GHTorrent, each issue was
+processed by ExceptionMiner tool developed in this work which extracted every
+exception defined on issues or on issue comments (step 2). Since this initial
+set of Java project is too diverse (it includes libraries, web-based or
+stand-alone applications), in the second phase of this study we restrict the
+analysis domain to Android projects (step 3). This decision was driven by the
+fact that most open source Android applications are known~\cite{ref} for dealing
+with several sources of exceptions due to multi-threaded execution and
+interaction with input/output resources; moreover they share a common
+well-defined platform and a limited set of libraries.
+
+To support a deeper investigation of the stack traces of Android apps, every
+exception defined on a stack trace was then classified according to its type
+(e.g. Error, Exception or RuntimeException) (step 4). The exception classifier
+was initialized with all the exceptions defined on the Java SDK version 7, the
+exceptions of the Android (Platform and OS) (level 19), and all the exceptions
+of the 482 Android projects selected for this study, extracted in the previous
+step. Moreover, all the packages associated to exceptions and signalers were
+also extracted to support further analysis (step 5). The signaler classifier is
+initialized with information about all packages found on Java SDK version 7, and
+Android Platform and OS (level 19). Finally, additional meta-data is combined to
+each stack trace (step 6) which used together with manual inspections (step 7)
+enabled to provide answers to RQ3.
 
@@
 
-\noindent\emph{Kinds of Analyzed Issues.} Issues on Github are different from issues on dedicated bug tracking tools such as Bugzilla and Jira.
-The most important difference is that there are no predefined fields for bug reporters to select from (e.g.
-severity and priority); Github uses instead a more open
-ended tagging system. Repositories are offered a pre-defined set of labels, but repository
-owners can modify them at will.
-An issue can have an arbitrary set of labels attached.
-Hence, this study does not restrict the analysis of stack traces defined on defect-issues, based on the assumption that: regardless the type of issue where its is defined the stack trace information contain relevant information concerning the exception use.
-Moreover, issue and pull requests are dual on Github; all pull requests have a corresponding ``backing'' issue - which are automatically generated.
-Therefore, we needed to exclude pull request based issues from our analysis.
-Finally, the Github API allows the automated generation of issues for specific repositories, which automated tools often use to report crashes.
-In some cases, this led to high number of issues that included stack traces.
-We identified and filtered those cases out (e.g., pullWifi project was responsible for almost 50\% of all issues created on GitHub, it was removed from our analysis and also from GitHub).
+\noindent\emph{Kinds of Analyzed Issues.} Issues on Github are different from
+issues on dedicated bug tracking tools such as Bugzilla and Jira. The most
+important difference is that there are no predefined fields for bug reporters to
+select from (e.g. severity and priority); Github uses instead a more open ended
+tagging system. Repositories are offered a pre-defined set of labels, but
+repository owners can modify them at will. An issue can have an arbitrary set of
+labels attached. Hence, this study does not restrict the analysis of stack
+traces defined on defect-issues, based on the assumption that: regardless the
+type of issue where its is defined the stack trace information contain relevant
+information concerning the exception use. Moreover, issue and pull requests are
+dual on Github; all pull requests have a corresponding ``backing'' issue - which
+are automatically generated. Therefore, we needed to exclude pull request based
+issues from our analysis. Finally, the Github API allows the automated
+generation of issues for specific repositories, which automated tools often use
+to report crashes. In some cases, this led to high number of issues that
+included stack traces. We identified and filtered those cases out (e.g.,
+pullWifi project was responsible for almost 50\% of all issues created on
+GitHub, it was removed from our analysis and also from GitHub).
 
@@
 
-%In the second phase of this study we selected a subset of Java applications to be analyzed in more
-%detail.The subset chosen was comprised by the Android apps included on GitHub (until 23 Feb 2013).
+%In the second phase of this study we selected a subset of Java applications to
+%be analyzed in more detail.The subset chosen was comprised by the Android apps
+%included on GitHub (until 23 Feb 2013).
 To identify Android projects, we performed a case insensitive search for the
-term \textsf{android} in the repository's names and short descriptions.
-The heuristic filtered 2.542 repositories, from which 589 apps had at least one issue containing a stack trace.
-
-Then we performed further cleanup, inspecting the site of every Android reporting at least one stack trace, to make sure that they represented real mobile apps.
-During this clean up 106 apps were removed because they were either example projects (i.e., toy projects) or tools to support Android development (e.g.
-selendroid, roboeletric -- tools to support the testing of Android apps).
-The filtered set consisted of 482 apps, from which approximately 50\% are also hosted in Google Market Place.
-For each one of the analyzed projects, we downloaded the source code and using custom scripts, we extracted their package hierarchy (recursive list of of Java packages) and their custom exception types as described next.
+term \textsf{android} in the repository's names and short descriptions.  The
+heuristic filtered 2.542 repositories, from which 589 apps had at least one
+issue containing a stack trace.
+
+Then we performed further cleanup, inspecting the site of every Android
+reporting at least one stack trace, to make sure that they represented real
+mobile apps. During this clean up 106 apps were removed because they were either
+example projects (i.e., toy projects) or tools to support Android development
+(e.g. selendroid, roboeletric -- tools to support the testing of Android apps).
+The filtered set consisted of 482 apps, from which approximately 50\% are also
+hosted in Google Market Place. For each one of the analyzed projects, we
+downloaded the source code and using custom scripts, we extracted their package
+hierarchy (recursive list of of Java packages) and their custom exception types
+as described next.
 
@@
 
-To extract exceptions from issues, we implemented ExceptionMiner, a modular mining tool
-able to connect to various repositories (such as Google Code, GHTorrent or even directly to
-Bugzilla), extract issues, mine stack traces from them and classify them in predefined categories. The main components of ExceptionMiner are as follows:
-
-\noindent\emph{StackTraceMiner and Distiller} The first step in the process is mining references to exceptions and 
-stack traces embedded in issues, distilling the information that composes a stack trace and storing
-the results in a relational database. The extracted stack trace information  the exception signaler, 
-the root cause, the exception wrappers. The tool is based on a combination of a regular
-expression based parser with heuristics able to identify and filter exception names 
-and stack traces inline with text. In contrast to existing issue parsing 
-solutions such as Infozilla, the parser created in this work can extract all causes related to an 
-exception on a stack, and stack traces embedded in logs files.
+To extract exceptions from issues, we implemented ExceptionMiner, a modular
+mining tool able to connect to various repositories (such as Google Code,
+GHTorrent or even directly to Bugzilla), extract issues, mine stack traces from
+them and classify them in predefined categories. The main components of
+ExceptionMiner are as follows:
+
+\noindent\emph{StackTraceMiner and Distiller} The first step in the process is
+mining references to exceptions and stack traces embedded in issues, distilling
+the information that composes a stack trace and storing the results in a
+relational database. The extracted stack trace information  the exception
+signaler, the root cause, the exception wrappers. The tool is based on a
+combination of a regular expression based parser with heuristics able to
+identify and filter exception names and stack traces inline with text. In
+contrast to existing issue parsing solutions such as Infozilla, the parser
+created in this work can extract all causes related to an exception on a stack,
+and stack traces embedded in logs files.
 
@@
 
-\noindent\emph{Exception Classifier} The next step in the ExceptionMiner process is classifying exceptions 
-as either checked exceptions, runtime exceptions or an errors. The exception classifier
-is based on bytecode analysis (based on Design Wizard~\cite{}) that walks up the type hierarchy of a Java exception until it reaches a base exception type.
-
-\noindent\emph{Exception Signaler Classifier} After its initialization it classifies each signaler according to its origin (i.e.
-Android Plaform, Aplication, Library, Libcore, or Java) based on pattern matching between the signaler name
-and the packages associated to each category.
+\noindent\emph{Exception Classifier} The next step in the ExceptionMiner process
+is classifying exceptions as either checked exceptions, runtime exceptions or an
+errors. The exception classifier is based on bytecode analysis (based on Design
+Wizard~\cite{}) that walks up the type hierarchy of a Java exception until it
+reaches a base exception type.
+
+\noindent\emph{Exception Signaler Classifier} After its initialization it
+classifies each signaler according to its origin (i.e. Android Plaform,
+Aplication, Library, Libcore, or Java) based on pattern matching between the
+signaler name and the packages associated to each category.
 
 %\section{Study Operation and Results}
-Next section presents the results for each of the study phases, providing both a quantitative and qualitative analysis of the outcomes.
+Next section presents the results for each of the study phases, providing both a
+quantitative and qualitative analysis of the outcomes.
 
@@
 
-Prevalence or prevalence proportion, a term often used in epidemiology, represents the proportion of a population found to have a particular condition (e.g.
-a disease) over the total number of people studied.
-In our work, we used this term to evaluate a condition (i.e., presence of a stack trace) over issues reported on GitHub repositories.
+Prevalence or prevalence proportion, a term often used in epidemiology,
+represents the proportion of a population found to have a particular condition
+(e.g. a disease) over the total number of people studied. In our work, we used
+this term to evaluate a condition (i.e., presence of a stack trace) over issues
+reported on GitHub repositories.
 
@@
    \multicolumn{2}{l}{\bfseries{Repositories}} & 16,836 & 482\\
-%   \multicolumn{2}{l}{\bfseries{ }} &  &  \\  
  \multicolumn{2}{l}{\bfseries{Issues}} & 356,057 & 32,582\\
@@
     & incl. stack traces & 21,013 (5,9\%) & 3,101 (7,3\%) \\
-%   \multicolumn{2}{l}{\bfseries{ }} &  &  \\   
  \multicolumn{2}{l}{\bfseries{Defect Issues}} & 55,226 & 2,376 \\
@@
     & incl. stack trace &  5,196 (9,4\%) & 1,393 (58,7\%)\\
-%\multicolumn{2}{l}{\bfseries{ }} &  &  \\  
 \multicolumn{2}{l}{\bfseries{Exception Stack Traces}} &  & \\
@@
 The use of the Github issue tracker in repositories hosted on Github is optional
-and not widespread. As of March 2014, from the 332,864 non-fork Java repositories registered in the 
-GHTorrent dataset, only 44,323 have received an issue report in their lifetime while from those only 16,837 are being watched by external users.
-To ensure that our work targets projects that are openly used by
+and not widespread. As of March 2014, from the 332,864 non-fork Java
+repositories registered in the GHTorrent dataset, only 44,323 have received an
+issue report in their lifetime while from those only 16,837 are being watched by
+external users. To ensure that our work targets projects that are openly used by
 users other than their developers, we focused our analysis on just those
-projects. Using the heuristics presented in Section~\ref{sec:android}, 
-we selected 482 repositories featuring Android projects. Table~\ref{tab:proj} and 
-Table~\ref{tab:overview} presents two views of the extracted dataset,
-namely grouped by repository and grouped by issue.
-
-From Table~\ref{tab:overview}, we can see that repositories that include at least one exception or stack trace on an issue 
-are fairly common: 28\% and 22\%  respectively.
-Considering only the exceptions and stack traces reported on defect issues we find out that 41\% of the repositories reported at least one exception in a defect issue, and approximately 35\% of the projects mentioned one stack trace in at least one defect issue.
-
-At the individual issue level, 
-8,5\% and 5,9\% of all issues reported in Java repositories contain the full name of an exception or a stack trace respectively.
-The numbers are similar if we isolate Android projects from the general population (12,1\% and 7,3\%).
-Considering only the defect-issues however, in the
-Android repositories as almost 58,7\% of defect-issues include a stack trace. This is not true for the general case however.
+projects. Using the heuristics presented in Section~\ref{sec:android}, we
+selected 482 repositories featuring Android projects. Table~\ref{tab:proj} and
+Table~\ref{tab:overview} presents two views of the extracted dataset, namely
+grouped by repository and grouped by issue.
+
+From Table~\ref{tab:overview}, we can see that repositories that include at
+least one exception or stack trace on an issue are fairly common: 28\% and 22\%
+respectively. Considering only the exceptions and stack traces reported on
+defect issues we find out that 41\% of the repositories reported at least one
+exception in a defect issue, and approximately 35\% of the projects mentioned
+one stack trace in at least one defect issue.
+
+At the individual issue level, 8,5\% and 5,9\% of all issues reported in Java
+repositories contain the full name of an exception or a stack trace
+respectively. The numbers are similar if we isolate Android projects from the
+general population (12,1\% and 7,3\%). Considering only the defect-issues
+however, in the Android repositories as almost 58,7\% of defect-issues include a
+stack trace. This is not true for the general case however.
 
@@
 
-%The results in Table~\ref{tab:overview} are cumulative; however, the numbers
-%are not the same among all projects. Figure~\ref{fig:stack traces} 
-%presents the descriptive statistics of the range of values of the metrics described in  
-%Table~\ref{tab:overview} across all projects in our dataset. 
-%There we can see that indeed there is significant variation among projects,
-%while the outliers lay at the bottom end of the spectrum, which means that for
-%some projects, we can find very limited number of stack traces in their 
-%issue reports. Both Java and Android projects feature similar variation,
-%which again supports our Android focused analysis later on.
-
-\noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
-RQ1: 28,37\% of Java repositories on Github have at least 1 issue that includes a stack trace.
-9,4\% of the defect issues of such repositories mentioned at least one stack trace.
-Considering only defect-issues of Android subset 58,7\% mentioned at least 1 stack trace.
-}
+\noindent \fbox{
+\begin{minipage}{0.96\columnwidth} \emph{ RQ1: 28,37\% of Java
+  repositories on Github have at least 1 issue that includes a stack trace.
+  9,4\% of the defect issues of such repositories mentioned at least one stack
+  trace. Considering only defect-issues of Android subset 58,7\% mentioned at
+  least 1 stack trace.  }
 \end{minipage}}
 
-% \begin{itemize}
-
-%     \item What are the most ``popular'' exceptions and what they can tell us?
-
-%     \item Which elements are triggering most of the exceptions and what they 
-%     can tell us about the exception structure?
-
-%     \item What is the role of libraries and the underlying platform on crashes?
-
-%     \item What is the role of the application itself on crashes?
-% \end{itemize}
-
-% \subsection{The rules of using checked and unchecked exceptions are being obeyed in mobile environment?}
-
 \section{RQ2: What are the common characteristics of such stack traces?}
 
-To answer this question, we mined the following information of every extracted stack trace: the top level exception being signaled, its signaler (the first method signature frame after the exception), the root exception (the exception which initiated the stack trace, defined on the bottom of the stack), and the root signaler.
-The intermediate elements that compose a stack trace as shown in Figure 1 (i.e., the intermediate causes and its execution stacks) were not be considered in the analysis.
-%~\todo{Move this to the previous section? IMO, It is better to keep all facts about how we did the research in one place}
-
-Based on the assumption that, regardless the issue type every exception stack trace contains relevant information concerting the exception structure of projects analyzed we opted for not restricting the analysis only on defect issues~\footnote{We conducted the same analysis on the defect issues and the top exceptions are similar to the ones mentioned on defect issues.
-Due to space limitation we limit to present the general analysis here.
-More detailed analysis on the defect issues can be found here: \url{www.dimap.ufrn.br/~roberta/icsm2014}}.
+To answer this question, we mined the following information of every extracted
+stack trace: the top level exception being signaled, its signaler (the first
+method signature frame after the exception), the root exception (the exception
+which initiated the stack trace, defined on the bottom of the stack), and the
+root signaler. The intermediate elements that compose a stack trace as shown in
+Figure 1 (i.e., the intermediate causes and its execution stacks) were not be
+considered in the analysis.
+%~\todo{Move this to the previous section? IMO, It is better to keep all facts
+%about how we did the research in one place}
+
+Based on the assumption that, regardless the issue type every exception stack
+trace contains relevant information concerting the exception structure of
+projects analyzed we opted for not restricting the analysis only on defect
+issues~\footnote{We conducted the same analysis on the defect issues and the top
+exceptions are similar to the ones mentioned on defect issues. Due to space
+limitation we limit to present the general analysis here. More detailed analysis
+on the defect issues can be found here:
+\url{www.dimap.ufrn.br/~roberta/icsm2014}}.
 
@@
 
-Table~\ref{tab:toptenjava} illustrates the top 10 most common root causes which represents more than 50\% of the analyzed set.
-The table illustrates  (i) the number of times each exception appeared as the root cause of a stack trace (\#occurrences), (ii) the number of distinct projects reporting such causes on stack traces, and (iii) the exception popularity.
-The exception popularity metric was defined in this work as follows: \emph{popularity = number of projects mentioning the exception in at least one stack/all projects that contain at least one stack}.
+Table~\ref{tab:toptenjava} illustrates the top 10 most common root causes which
+represents more than 50\% of the analyzed set. The table illustrates  (i) the
+number of times each exception appeared as the root cause of a stack trace
+(\#occurrences), (ii) the number of distinct projects reporting such causes on
+stack traces, and (iii) the exception popularity. The exception popularity
+metric was defined in this work as follows: \emph{popularity = number of
+projects mentioning the exception in at least one stack/all projects that
+contain at least one stack}.
 
@@
 
-We can observe that the NullPointerException was the one with the higher number of occurrences (22,5\% of all stacks were caused by NullPointerExceptions) and the one which had the highest popularity across projects.
-Actually, most of the exceptions in the TOP 10 list (except java.lang.IllegalArgumentException, java.io.IOException and java.lang.RuntimeException) are exceptions implicitly signaled by Java Virtual Machine, when an expression violates the normal semantics of the Java programming language (e.g.
-ArrayIndexOutOfBounds) or due to an internal error or resource limitation (OutOfMemoryError).
-
-The high prevalence of NullPointerExceptions, and the other implicitly-thrown exceptions  is aligned with the findings other works [Sunghun et al.][Gordon fraser and][Jcrasher].[Android-predict]~\todo{refs}.
-For instance, Sunghun et al.
-showed that in Eclipse bug report system 38\% the bugs related to exception handling were caused by NullPointerException; other works on robustness testing [JCrasher, Evol;suite, Amiya]~\todo{refs} showed that most of the automatically detected bugs were due to NullPointerExceptions and implicitly-signaled of Java environment (as the ones found in this study).
+We can observe that the NullPointerException was the one with the higher number
+of occurrences (22,5\% of all stacks were caused by NullPointerExceptions) and
+the one which had the highest popularity across projects. Actually, most of the
+exceptions in the TOP 10 list (except java.lang.IllegalArgumentException,
+java.io.IOException and java.lang.RuntimeException) are exceptions implicitly
+signaled by Java Virtual Machine, when an expression violates the normal
+semantics of the Java programming language (e.g. ArrayIndexOutOfBounds) or due
+to an internal error or resource limitation (OutOfMemoryError).
+
+The high prevalence of NullPointerExceptions, and the other implicitly-thrown
+exceptions  is aligned with the findings other works [Sunghun et al.][Gordon
+fraser and][Jcrasher].[Android-predict]~\todo{refs}. For instance, Sunghun et
+al. showed that in Eclipse bug report system 38\% the bugs related to exception
+handling were caused by NullPointerException; other works on robustness testing
+[JCrasher, Evol;suite, Amiya]~\todo{refs} showed that most of the automatically
+detected bugs were due to NullPointerExceptions and implicitly-signaled of Java
+environment (as the ones found in this study).
 
@@
 
-As detailed in Section III, 482 Android projects were carefully selected to enable a deaper investigation of stack trace information, by leveraging stack trace information with the source code and bytecode analysis.
-For this subset of Android projects we could extract every package related to each exception signaler, and based on this information we could classify every stack trace based on its root signaler as presented in Table~\ref{signalers}.
+As detailed in Section III, 482 Android projects were carefully selected to
+enable a deaper investigation of stack trace information, by leveraging stack
+trace information with the source code and bytecode analysis. For this subset of
+Android projects we could extract every package related to each exception
+signaler, and based on this information we could classify every stack trace
+based on its root signaler as presented in Table~\ref{signalers}.
 
@@
 
-For each exception we could discover: the number of times it occurred on stacks, the number of distinct projects such stacks were defined (and consequently its popularity) and the of times it was signaled by different signaler types.
-We could calculate the popularity of each root cause, considering the number of distinct projects they were found and the whole set of projects analyzed with at least one stack trace (482).
-Table~\ref{toptenandroid} and Figuere~\ref{androidsignaler} presents the mined data.
+For each exception we could discover: the number of times it occurred on stacks,
+the number of distinct projects such stacks were defined (and consequently its
+popularity) and the of times it was signaled by different signaler types. We
+could calculate the popularity of each root cause, considering the number of
+distinct projects they were found and the whole set of projects analyzed with at
+least one stack trace (482). Table~\ref{toptenandroid} and
+Figuere~\ref{androidsignaler} presents the mined data.
 
@@
 %\textbf{
-Coding errors are the most common root causes for most types of signalers.
-We could observe that the NullPointerException is still the exception with higher number of occurrences.
-The NullPoiterExceptions are mainly signaled inside Android platform and inside application code.
-Although we also find NullPointerExceptions being signaled from third-party libraries.
-Regarding reusable code, there is no consensus if it s a good or bad practice to re-throw NullPointerException.
-Some prefer to encapsulate such exception on IllegalArgumentException.
-But others~\cite{joshuabloch} argue that the NullPointerException  explicits the cause of the problem and hence should not be wrapped.
+Coding errors are the most common root causes for most types of signalers. We
+could observe that the NullPointerException is still the exception with higher
+number of occurrences. The NullPoiterExceptions are mainly signaled inside
+Android platform and inside application code. Although we also find
+NullPointerExceptions being signaled from third-party libraries. Regarding
+reusable code, there is no consensus if it s a good or bad practice to re-throw
+NullPointerException. Some prefer to encapsulate such exception on
+IllegalArgumentException. But others~\cite{joshuabloch} argue that the
+NullPointerException  explicits the cause of the problem and hence should not be
+wrapped.
 
 %commented only for submission:
-%\textbf{We could observe in this study that the NullPointerException and other other implicitly signaled exceptions represented most of the exceptions reported on issues in both Java and Android subset.
-For such exceptions - which represent programming bugs or resource limitations - there is usually no proper handling besides presenting an error message to the user and restarting the application - only high fault tolerant systems need to provide solutions to handle them [REF: Java Specification].}
+\textbf{We could observe in this study that the NullPointerException and other
+other implicitly signaled exceptions represented most of the exceptions reported
+on issues in both Java and Android subset. For such exceptions - which represent
+programming bugs or resource limitations - there is usually no proper handling
+besides presenting an error message to the user and restarting the application -
+only high fault tolerant systems need to provide solutions to handle them [REF:
+Java Specification].}
 
@@
 
-To better understand the main causes of such exceptions we evaluated semantics related to the top 100 root causes reported on stack-issues of both Java dataset and Android subset (which correspond to approximately 77\% of all exceptions reported reported on Java stack-issues and 95\% of all exceptions reported on Android stack-issues), and classified them according to the the categories presented in Table~\ref{tophundrend}.
+To better understand the main causes of such exceptions we evaluated semantics
+related to the top 100 root causes reported on stack-issues of both Java dataset
+and Android subset (which correspond to approximately 77\% of all exceptions
+reported reported on Java stack-issues and 95\% of all exceptions reported on
+Android stack-issues), and classified them according to the the categories
+presented in Table~\ref{tophundrend}.
 
@@
       Backward compatibility                 & 1580 (5,14\%)    & 219 (5,45\%) \\
-      Reflection &1413 (4,60\%)   & 91 (2,27\%)\\
+      Reflection                             & 1413 (4,60\%)    & 91 (2,27\%)\\
       Specific (GUI,FRAMEWORK)               & 340 (1,11\%)     & 197 (4,90\%)\\
@@
 
-As we can see from Table 4, the that programming errors are the causes of most of the exception stack traces reported on issues.
-In the Android subset we could find more issues related to Backward compatibility and security issues than on the general data set.
+As we can see from Table 4, the that programming errors are the causes of most
+of the exception stack traces reported on issues. In the Android subset we could
+find more issues related to Backward compatibility and security issues than on
+the general data set.
 
@@
 
-Concerning the size of the stacks (i.e.
-the number of frames composing them), we could observe that  sometimes it exceeded 100 frames (158 stack frames).
-Manually inspecting the corresponding execution trace was caused by one of the two reasons: recursive calls; exceptions signaled by a method defined deep down in a reused framework; exceptions successively wrapped.
-Regardless the reason, in exception flows involving many methods, it is piratically impossible to handle the exception a way different than existing the  application, because there is hardly enough contextual information to perform any other recovery action.
+Concerning the size of the stacks (i.e. the number of frames composing them), we
+could observe that  sometimes it exceeded 100 frames (158 stack frames).
+Manually inspecting the corresponding execution trace was caused by one of the
+two reasons: recursive calls; exceptions signaled by a method defined deep down
+in a reused framework; exceptions successively wrapped. Regardless the reason,
+in exception flows involving many methods, it is piratically impossible to
+handle the exception a way different than existing the  application, because
+there is hardly enough contextual information to perform any other recovery
+action.
 
@@
 
-%Considering only the packages of each exception we could observe that 81,28\% of all analyzed set of stack traces reported exceptions defined by java.lang (see Table XXX).
+%Considering only the packages of each exception we could observe that 81,28\%
+%of all analyzed set of stack traces reported exceptions defined by java.lang
+%(see Table XXX).
 
@@
 %     \hline
-%1 &java.lang &	19471 &	81,28\%	& 3080 \\
+%1 & java.lang         & 19471 & 81,28\% & 3080 \\
 %2 & java.io           & 1222  & 5,10\%  & 509 \\
@@
 
-To answer this question we analyzed general characteristics of stack traces reported on Java repositories (e.g., exceptions being thrown presented in Tables~\ref{tab:toptenjava} and ~\ref{tab:toptenandroid}), and took a deeper look at the stack traces of the Android repositories combining them with with source code and byte code analyses.
-Doing so, we mined some characteristics of stacks that points to best practices violations.
+To answer this question we analyzed general characteristics of stack traces
+reported on Java repositories (e.g., exceptions being thrown presented in
+Tables~\ref{tab:toptenjava} and ~\ref{tab:toptenandroid}), and took a deeper
+look at the stack traces of the Android repositories combining them with with
+source code and byte code analyses. Doing so, we mined some characteristics of
+stacks that points to best practices violations.
 
@@
 
-From Tables~\ref{tab:toptenjava}, ~\ref{tab:toptenandroid} we can observe that, in both Java and Android related issues, direct instances java.lang.RuntimeException were thrown  (in 10,94\% of Java repositories and 18,88\% of Android repositories respectively).
-Moreover, after inspecting the top 100 most thrown exceptions (see ~\ref{tophundrend}) we could observe that stack traces on which direct instances of java.lang.Exception and java.lang.Error were thrown.
-Throwing general exceptions is usually considered a bad practice, because the type does not carry enough information to identify the cause of the exceptional behaviour.
-In such cases, developers need to relying on exception message which can be neither complete nor precise ~\cite{gosling2000java}.
-
-In Table ~\ref{toptenandroid} we can observe a high prevalence of RuntimeExceptions being thrown by the Android platform (167 out of 232).
-Manually inspecting the code on which RuntimeExceptions were thrown in Android Platform we could observe that direct instances of java.lang.RuntimeException they were thrown for 2 main reasons: (i) when there was no reasonable way of handling the exception, which means that regardless the kind of exception the only possible action was to abort the application; and when wrapping exceptions signaled by application methods.
-Since Android platform is heavily based on "call backs", when the application does not handle its exceptions, they scape to Android platform which does not have contextual information to adequately handle the exception and simply terminates the application.
-What is apparently a bad practice of signaling general exceptions in this context it was and intended design decision.
-
-%Even an instance of Throwable as signaled in one project. It points to a the "dont worry from know behaviour".
+From Tables~\ref{tab:toptenjava}, ~\ref{tab:toptenandroid} we can observe that,
+in both Java and Android related issues, direct instances
+java.lang.RuntimeException were thrown  (in 10,94\% of Java repositories and
+18,88\% of Android repositories respectively). Moreover, after inspecting the
+top 100 most thrown exceptions (see ~\ref{tophundrend}) we could observe that
+stack traces on which direct instances of java.lang.Exception and
+java.lang.Error were thrown. Throwing general exceptions is usually considered a
+bad practice, because the type does not carry enough information to identify the
+cause of the exceptional behaviour. In such cases, developers need to relying on
+exception message which can be neither complete nor precise
+~\cite{gosling2000java}.
+
+In Table ~\ref{toptenandroid} we can observe a high prevalence of
+RuntimeExceptions being thrown by the Android platform (167 out of 232).
+Manually inspecting the code on which RuntimeExceptions were thrown in Android
+Platform we could observe that direct instances of java.lang.RuntimeException
+they were thrown for 2 main reasons: (i) when there was no reasonable way of
+handling the exception, which means that regardless the kind of exception the
+only possible action was to abort the application; and when wrapping exceptions
+signaled by application methods. Since Android platform is heavily based on
+"call backs", when the application does not handle its exceptions, they scape to
+Android platform which does not have contextual information to adequately handle
+the exception and simply terminates the application. What is apparently a bad
+practice of signaling general exceptions in this context it was and intended
+design decision.
+
+%Even an instance of Throwable as signaled in one project. It points to a the
+%"dont worry from know behaviour".
 
@@
 
-%In this study we also classified each stack trace according to the type of the root exception been signaled.
-Table~\ref{typeroottab} presents the types of root causes of all stack traces reported on the Android repositories.
+%In this study we also classified each stack trace according to the type of the
+%root exception been signaled. Table~\ref{typeroottab} presents the types of
+%root causes of all stack traces reported on the Android repositories.
 
@@
 
-%todo{discomment after submission}
-%As we can see from Table ~ref{typeroottab}, that most of the exceptions signaled by third-party code (e.g., Android and libs) were runtime exceptions.
-When an exception is signaled by third party code, it can be due to 2 reasons: the precondition associated to third party's code was not obeyed; or there is a coding error on third party's code.
-Only manual code inspection can reveal the reason.
-However, regardless the reason it is a good practice that all the exceptions signaled by a third part code are properly documented.
-Manually inspecting a sample of stack traces caused by runtime exceptions thrown by library code (approximately 10\%) we could observe that none of the runtime exceptions signaled by third part code were documented in Javadoc - this analysis did not consider the exceptions implicitly signaled by JVM (e.g.
-NullPoiterException).
-The undocumented runtime exceptions are a serious problem as it makes very difficult or even impossible for the  client method to protect against  undocumented runtime exceptions signaled by library code.
-As a consequence, the undocumented runtime exception may remain uncaught leading to  system crashes.
-
-%This result is aligned with the results of other study conducted by from Cabral and Marcques ~\cite{} which also observe that runtime exceptions are most often not documented.
-
-%\textbf{When libraries are used, the developer does not have access to their source code and thus needs to rely on the library documentation about the runtime exceptions that should be thrown – which, more often than not, are neither complete nor precise (Thomas, 2002; Sacramento et al., 2006).
-As a consequence, unhandled unchecked exceptions can be seen as one of the major sources of bugs in current Java systems (Jo et al., 2004).
-}
-
-%On the other hand, we could observe that most of the exceptions signaled by the set of libraries reused by Android platform (i.e., org.apache.harmony, org.w3c.dom, sun.misc, org.apache.http, org.json, org.xml, javax.) referred in this work as libcore, were checked exceptions.
-It is pointed as a good practice for libraries (see Section ~\ref{best}) because by using checked exceptions libraries can define a precise exception interface to its clients.
-Such libraries are heavily used in several projects, and such precise exception interface may be related to the libraries maturity.
+%todo{discomment after submission} As we can see from Table ~ref{typeroottab},
+%that most of the exceptions signaled by third-party code (e.g., Android and
+%libs) were runtime exceptions. When an exception is signaled by third party
+%code, it can be due to 2 reasons: the precondition associated to third party's
+%code was not obeyed; or there is a coding error on third party's code. Only
+%manual code inspection can reveal the reason. However, regardless the reason it
+%is a good practice that all the exceptions signaled by a third part code are
+%properly documented. Manually inspecting a sample of stack traces caused by
+%runtime exceptions thrown by library code (approximately 10\%) we could observe
+%that none of the runtime exceptions signaled by third part code were documented
+%in Javadoc - this analysis did not consider the exceptions implicitly signaled
+%by JVM (e.g. NullPoiterException). The undocumented runtime exceptions are a
+%serious problem as it makes very difficult or even impossible for the  client
+%method to protect against  undocumented runtime exceptions signaled by library
+%code. As a consequence, the undocumented runtime exception may remain uncaught
+%leading to  system crashes.
+
+%This result is aligned with the results of other study conducted by from Cabral
+%and Marcques ~\cite{} which also observe that runtime exceptions are most often
+%not documented.
+
+%\textbf{When libraries are used, the developer does not have access to their
+%source code and thus needs to rely on the library documentation about the
+%runtime exceptions that should be thrown – which, more often than not, are
+%neither complete nor precise (Thomas, 2002; Sacramento et al., 2006). As a
+%consequence, unhandled unchecked exceptions can be seen as one of the major
+%sources of bugs in current Java systems (Jo et al., 2004). }
+
+%On the other hand, we could observe that most of the exceptions signaled by the
+%set of libraries reused by Android platform (i.e., org.apache.harmony,
+%org.w3c.dom, sun.misc, org.apache.http, org.json, org.xml, javax.) referred in
+%this work as libcore, were checked exceptions. It is pointed as a good practice
+%for libraries (see Section ~\ref{best}) because by using checked exceptions
+%libraries can define a precise exception interface to its clients. Such
+%libraries are heavily used in several projects, and such precise exception
+%interface may be related to the libraries maturity.
 
@@
 
-In this study we also classified each stack trace according to the type of the root exception been signaled.
-Table~\ref{typeroottab} presents the types of root causes of all stack traces reported on the Android repositories.
+In this study we also classified each stack trace according to the type of the
+root exception been signaled. Table~\ref{typeroottab} presents the types of root
+causes of all stack traces reported on the Android repositories.
 
@@
 Throwable & 0    & 0   & 1    & 0   & 0  & 1 \\
-
     \hline
@@
 
-Although most of the reported were runtime, we could also find checked exceptions as the root causes of stack traces reported on issues.
-An exception can remain uncaught in two circumstances: if all methods on the execution trace on which this exception flows explicitly specifies the exception, or if the checked exception is wrapped in a runtime and than re-thrown.Since checked exceptions should be used to represent recoverable conditions, ignoring a checked exception is considered a bad practice.
-When the developer is confronted with a checked exception, the designer of the API is telling him to handle the exceptional condition.
-To better understand what was causing checked exceptions to scape from being handled we investigated the kinds of wrappings that happened on stacks the stacks.
-Next sections presents the wrappings found and discuss about them.
+Although most of the reported were runtime, we could also find checked
+exceptions as the root causes of stack traces reported on issues. An exception
+can remain uncaught in two circumstances: if all methods on the execution trace
+on which this exception flows explicitly specifies the exception, or if the
+checked exception is wrapped in a runtime and than re-thrown.Since checked
+exceptions should be used to represent recoverable conditions, ignoring a
+checked exception is considered a bad practice. When the developer is confronted
+with a checked exception, the designer of the API is telling him to handle the
+exceptional condition. To better understand what was causing checked exceptions
+to scape from being handled we investigated the kinds of wrappings that happened
+on stacks the stacks.  Next sections presents the wrappings found and discuss
+about them.
 
@@
 
- Table~\ref{wrappingandroid} presents the wrappings found in this study for all stack traces found in Android repositories.
-As we can see, some of the checked exceptions where indeed wrapped in runtime exceptions or even errors, while most of them were not wrapped along the stack trace (represented as the table lines on which the root cause is none).
-This analysis also revealed unexpected wrappings such as: Error wrapping a Checked exception; a Runtime wrapping and Error; and an Error wrapping a Runtime.
+Table~\ref{wrappingandroid} presents the wrappings found in this study for all
+stack traces found in Android repositories. As we can see, some of the checked
+exceptions where indeed wrapped in runtime exceptions or even errors, while most
+of them were not wrapped along the stack trace (represented as the table lines
+on which the root cause is none). This analysis also revealed unexpected
+wrappings such as: Error wrapping a Checked exception; a Runtime wrapping and
+Error; and an Error wrapping a Runtime.
 
@@
 
-Java is the only language that provides a hybrid exception handling mechanism which offers different types to represent different exception behaviors (see Section ~\ref{extypes}) .
-According to Java specification Errors should not be handled inside the system since they usually represent unrecoverable conditions detected by the JVM such as OutOfMemoryError.
-Checked exceptions should represent recoverable conditions and  Runtime exceptions represent are usually related to coding errors from which the developer is not expected to recover from.
+Java is the only language that provides a hybrid exception handling mechanism
+which offers different types to represent different exception behaviors (see
+Section ~\ref{extypes}) . According to Java specification Errors should not be
+handled inside the system since they usually represent unrecoverable conditions
+detected by the JVM such as OutOfMemoryError. Checked exceptions should
+represent recoverable conditions and  Runtime exceptions represent are usually
+related to coding errors from which the developer is not expected to recover
+from.
 
@@
 
-Since there is no way of enforcing such conventions during program development,  we could observe the stack trace analysis combined with types these three types of exceptions were used interchangeably.
-This interchangeability makes the behaviour of the exception handling code more complex and less reliable.
-For instance, considering a situation were an OutOfMemoryError (a situation that should not be handled) was wrapped in a checked exception (a situation that must he handled), in this case this wrapping is telling to the caller that it should handled.
+Since there is no way of enforcing such conventions during program development,
+we could observe the stack trace analysis combined with types these three types
+of exceptions were used interchangeably. This interchangeability makes the
+behaviour of the exception handling code more complex and less reliable. For
+instance, considering a situation were an OutOfMemoryError (a situation that
+should not be handled) was wrapped in a checked exception (a situation that must
+he handled), in this case this wrapping is telling to the caller that it should
+handled.
 
@@
 
-%\subsection{Access to Crash Logs}
-
-%The data
-
-%\subsection{Data Extraction from Natural language Artifacts}
-
-%-Data cleaning removing invalid and duplicated issues.
-%-Not every crash information is specified on an issues....
-
-%\subsection{Construct, Internal and External Threats}
-
-\subsection{Internal Validity}
-We used a heuristics-based parser to mine exceptions from issues.  
-Our parsing strategy was conservative by default; for example, we only
-considered exception names using a fully qualified class name as valid
-exception identifiers, while, in many cases, developers use the exception
-name in issue description. Conservative parsing may minimize false positives,
-which was our initial target, but also tends to increase false negatives,
-which means that some cases may have not been identified as exceptions or
-stack traces. Our limited manual inspection did not reveal such cases.
-
-\subsection{External Validity}
-The results presented here were based on mining issues on Github through the GHTorrent
-dataset. While comprehensive and extensive, GHTorrent is not an exact replica of
-Github, so several issues might be left out. Due to the way data collection works with
-GHTorrent, for projects that are relatively inactive, GHTorrent might have not collected a 
-significant proportion of their issues. We did not investigate the extend of this threat on 
+\subsection{Internal Validity} We used a heuristics-based parser to mine
+exceptions from issues.  Our parsing strategy was conservative by default; for
+example, we only considered exception names using a fully qualified class name
+as valid exception identifiers, while, in many cases, developers use the
+exception name in issue description. Conservative parsing may minimize false
+positives, which was our initial target, but also tends to increase false
+negatives, which means that some cases may have not been identified as
+exceptions or stack traces. Our limited manual inspection did not reveal such
+cases.
+
+\subsection{External Validity} The results presented here were based on mining
+issues on Github through the GHTorrent dataset. While comprehensive and
+extensive, GHTorrent is not an exact replica of Github, so several issues might
+be left out. Due to the way data collection works with GHTorrent, for projects
+that are relatively inactive, GHTorrent might have not collected a significant
+proportion of their issues. We did not investigate the extend of this threat on
 our sample.
@@
 
-In this section, we present works thaat are related to our own, distributed in four categories: (i) works that use the information available on stack traces; (ii) empirical studies on the usage of Java exceptions and its fault proneness; (iii) tools to extract stack traces information from natural language artifacts (i.e., issues and emails);  and (iv) empirical studies involving Android apps.
-
-\textit{Analysis of Java Exception Usage.} Since the manual analysis of the Java exception-flow can easily become infeasible Robillard and Murphy~\cite{Robil00} developed a tool called Jex that analyzes the flow of exceptions in Java source code.
-Based on java source code this tool performs dataflow analysis in order to find the propagation paths of checked and unchecked exception types.
-After this tool other tools have been proposed to support the static analysis of exception flows ~\cite{fu2007exception} ~\cite{coelho2008assessing}.
-The main limitation off all tools is the number of false positives inherent to static analysis solutions, which can lead to a high number of exception flows, specially if considering Java Environment exceptions and exceptions signaled from libraries.
-This was the fist of some tools proposed to statically analyzed the exception flow of exceptions.
-Cabral and Marques~\cite{cabral2007exception} analyzed the exception handling code of 32 open-source systems, both for Java and .NET.
-The goal of their study was to identify how exceptions were handled in different categories of systems.
-They examined the exception handlers and the respective actions taken on them.
-They observed that the action handlers were very simple (e.g., logging and present a message to the user).
-In our work differs from previous works since in our study we tried to identify bad practices from the combined use of stack traces extracted from issues and bytecode and source code analysis.
-Finally, Reimer and Srinivasan~\cite{reimer2003analyzing} listed a set of bad practices on exception handling that hinder software maintainability and robustness.
-Their differs from ours it was solely based on their own experience on working with Java enterprise applications and focus on source code examples of bad patterns on ways of handling exceptions such as exception swallowing, log verbosity.
-
-\textit{Analysis and Use of Stack Trace Information.} Some works have investigated the usage of stack trace information to support bug classification and clustering~\cite{wang2013improving, kim2011crash, dhaliwal2011classifying}, fault-proneness prediction models~\cite{kim2013predicting} and even automated bug fixing~\cite{sinha2009fault} tools.Kim et al~\cite{kim2011crash} use of crash graphs (an aggregated form of multiple stack traces available in crash reports) to detect duplicate crash reports and to predict if a given crash will be fixed.
-Dhaliwal et al~\cite{} analyzed crash reports collected for Mozilla Firefox and proposed a crash grouping approach based on the Levenshtein distance.
-This work showed that a crash grouping approach can reduce bug fixing time in approximately 5\%.
-Wang et al~\cite{} proposes an approach to define correlated crash types automatically.
-Moreover it also provides a bug localization method to locate and rank files related to the bug described on a stack trace.
-Schroter et al~\cite{schroter2010stack} conducted an empirical study on the usefulness of stack traces for bug fixing.
-The study used the development data of Eclipse project and could observe that developers fixed the bugs faster when failing stack traces were included on bug issues.
-This study is aligned with the one performed by Bettenburg et al.
-~\cite{bettenburg2008makes} which evaluated what content could make a good bug report, and stack traces were mentioned as the second most important content for developers - just after the "steps to reproduce" information.
-Sinha et al~\cite{sinha2009fault} proposed an approach for locating and repairing faults that are caused by JVM implicitly signaled exceptions related to coding errors such as NullPointerExceptions.
-This approach uses the information available on the stack trace to guide a dataflow analysis, to locate the source statement responsible for the exception.
-Kim at al.~\cite{kim2013predicting} proposed an approach to predict the crash-proneness of methods based information extracted from stack traces and methods' bytecode operations.
-They used InfoZilla to extracted stack trace information from the bug reports of two open-source projects, i.e.
-AspectJ and Eclipse.
-They could observe that most of the stack traces were related to NullPointerException and other JMV implicitly thrown exceptions had the higher prevalence on the analyzed set of stacks.
-
-\textit{Extracting Stack Traces from Natural language artifacts.} Nowadays many software vendors embed automatic crash reporting tools in their software systems.
-Hence, whenever a software crashes this tool sends a detailed crash report to its vendors.
-Moreover, there are also find third party software solutions specialized in bug reporting, most of them for the increasing marked of mobile apps~\cite{BugSe14,BugSn14,Googl14,Acra14}.
-Hence most of the works described previously used the  information provided by crass reporting tools.
-Two works have proposed tools extract stack traces from natural language artifacts such as issues and emails: Infozilla ~\cite{bettenburg2008extracting} is based on a set of regular expressions that extract an the set of frames related to a stack trace, the main limitation of this solution is that it is not able to extract stack traces embedded on verbose log files (i.e., on which we can find log text mixed with exception frames).
-Alberto and Lanza ~\cite{bacchelli2012content}  propose a solution to recognize stack trace frames from development emails and relate it to code artifacts (i.e.
-classes) mentioned on the stack trace.
-
-\textit{Empirical studies using Android apps.} Ruiz et al.~\cite{Ruiz12} investigated the degree of reuse across applications in Android Market, the study showed that almost 23\% of the classes inherited
-from a base class in the Android API, and that 217 mobile apps were reused
-completely by another mobile app. Pathak et al.~\cite{Patha11} analyzed bug
-reports and developers discussions of Android platform and found out that
-approximately 20\% of energy-related bugs in Android occurred after an OS
-update. McDonnell et al.~\cite{McDon13} conducted a case study of the
-co-evolution behavior of Android API and 10 dependent applications using the
-version history data found in github. The study found that approximately 25\% of
-all methods in the client code used the Android API, and that the methods
-reusing fast-evolving APIs were more defect prone then others. Vásquez et
-al.~\cite{Linar13} analyzed approximately 7K free Android apps and observed that
-the last successful apps used Android APIs that were on average 300\% more
-change-prone than the APIs used by the most successful apps. Pingyu and
-Elbaum~\cite{Zhang12} analyzed bug reports of 5 Android applications an observed
-that 29\% had to do with poor exceptional handling code.  Our work differs from
-the others as it aims at distilling the information of bug reports describing
-uncaught exceptions created for mobile applications in Github and Googlecode, in
-order to get a first view of what is causing the crashes across applications
-available and what the characteristics of the stack traces can tell us about the
-exception structure of those applications. %Moreover this work\ldots.
+In this section, we present works thaat are related to our own, distributed in
+four categories: (i) works that use the information available on stack traces;
+(ii) empirical studies on the usage of Java exceptions and its fault proneness;
+(iii) tools to extract stack traces information from natural language artifacts
+(i.e., issues and emails);  and (iv) empirical studies involving Android apps.
+
+\textit{Analysis of Java Exception Usage.} Since the manual analysis of the Java
+exception-flow can easily become infeasible Robillard and Murphy~\cite{Robil00}
+developed a tool called Jex that analyzes the flow of exceptions in Java source
+code. Based on java source code this tool performs dataflow analysis in order to
+find the propagation paths of checked and unchecked exception types. After this
+tool other tools have been proposed to support the static analysis of exception
+flows ~\cite{fu2007exception} ~\cite{coelho2008assessing}. The main limitation
+off all tools is the number of false positives inherent to static analysis
+solutions, which can lead to a high number of exception flows, specially if
+considering Java Environment exceptions and exceptions signaled from libraries.
+This was the fist of some tools proposed to statically analyzed the exception
+flow of exceptions. Cabral and Marques~\cite{cabral2007exception} analyzed the
+exception handling code of 32 open-source systems, both for Java and .NET. The
+goal of their study was to identify how exceptions were handled in different
+categories of systems. They examined the exception handlers and the respective
+actions taken on them. They observed that the action handlers were very simple
+(e.g., logging and present a message to the user). In our work differs from
+previous works since in our study we tried to identify bad practices from the
+combined use of stack traces extracted from issues and bytecode and source code
+analysis. Finally, Reimer and Srinivasan~\cite{reimer2003analyzing} listed a set
+of bad practices on exception handling that hinder software maintainability and
+robustness. Their differs from ours it was solely based on their own experience
+on working with Java enterprise applications and focus on source code examples
+of bad patterns on ways of handling exceptions such as exception swallowing, log
+verbosity.
+
+\textit{Analysis and Use of Stack Trace Information.} Some works have
+investigated the usage of stack trace information to support bug classification
+and clustering~\cite{wang2013improving, kim2011crash, dhaliwal2011classifying},
+fault-proneness prediction models~\cite{kim2013predicting} and even automated
+bug fixing~\cite{sinha2009fault} tools.Kim et al~\cite{kim2011crash} use of
+crash graphs (an aggregated form of multiple stack traces available in crash
+reports) to detect duplicate crash reports and to predict if a given crash will
+be fixed. Dhaliwal et al~\cite{} analyzed crash reports collected for Mozilla
+Firefox and proposed a crash grouping approach based on the Levenshtein
+distance. This work showed that a crash grouping approach can reduce bug fixing
+time in approximately 5\%. Wang et al~\cite{} proposes an approach to define
+correlated crash types automatically. Moreover it also provides a bug
+localization method to locate and rank files related to the bug described on a
+stack trace.  Schroter et al~\cite{schroter2010stack} conducted an empirical
+study on the usefulness of stack traces for bug fixing. The study used the
+development data of Eclipse project and could observe that developers fixed the
+bugs faster when failing stack traces were included on bug issues. This study is
+aligned with the one performed by Bettenburg et al. ~\cite{bettenburg2008makes}
+which evaluated what content could make a good bug report, and stack traces were
+mentioned as the second most important content for developers - just after the
+"steps to reproduce" information. Sinha et al~\cite{sinha2009fault} proposed an
+approach for locating and repairing faults that are caused by JVM implicitly
+signaled exceptions related to coding errors such as NullPointerExceptions. This
+approach uses the information available on the stack trace to guide a dataflow
+analysis, to locate the source statement responsible for the exception. Kim at
+al.~\cite{kim2013predicting} proposed an approach to predict the crash-proneness
+of methods based information extracted from stack traces and methods' bytecode
+operations. They used InfoZilla to extracted stack trace information from the
+bug reports of two open-source projects, i.e. AspectJ and Eclipse. They could
+observe that most of the stack traces were related to NullPointerException and
+other JMV implicitly thrown exceptions had the higher prevalence on the analyzed
+set of stacks.
+
+\textit{Extracting Stack Traces from Natural language artifacts.} Nowadays many
+software vendors embed automatic crash reporting tools in their software
+systems. Hence, whenever a software crashes this tool sends a detailed crash
+report to its vendors. Moreover, there are also find third party software
+solutions specialized in bug reporting, most of them for the increasing marked
+of mobile apps~\cite{BugSe14,BugSn14,Googl14,Acra14}. Hence most of the works
+described previously used the  information provided by crass reporting tools.
+Two works have proposed tools extract stack traces from natural language
+artifacts such as issues and emails: Infozilla ~\cite{bettenburg2008extracting}
+is based on a set of regular expressions that extract an the set of frames
+related to a stack trace, the main limitation of this solution is that it is not
+able to extract stack traces embedded on verbose log files (i.e., on which we
+can find log text mixed with exception frames). Alberto and Lanza
+~\cite{bacchelli2012content}  propose a solution to recognize stack trace frames
+from development emails and relate it to code artifacts (i.e. classes) mentioned
+on the stack trace.
+
+\textit{Empirical studies using Android apps.} Ruiz et al.~\cite{Ruiz12}
+investigated the degree of reuse across applications in Android Market, the
+study showed that almost 23\% of the classes inherited from a base class in the
+Android API, and that 217 mobile apps were reused completely by another mobile
+app. Pathak et al.~\cite{Patha11} analyzed bug reports and developers
+discussions of Android platform and found out that approximately 20\% of
+energy-related bugs in Android occurred after an OS update. McDonnell et
+al.~\cite{McDon13} conducted a case study of the co-evolution behavior of
+Android API and 10 dependent applications using the version history data found
+in github. The study found that approximately 25\% of all methods in the client
+code used the Android API, and that the methods reusing fast-evolving APIs were
+more defect prone then others. Vásquez et al.~\cite{Linar13} analyzed
+approximately 7K free Android apps and observed that the last successful apps
+used Android APIs that were on average 300\% more change-prone than the APIs
+used by the most successful apps. Pingyu and Elbaum~\cite{Zhang12} analyzed bug
+reports of 5 Android applications an observed that 29\% had to do with poor
+exceptional handling code.  Our work differs from the others as it aims at
+distilling the information of bug reports describing uncaught exceptions created
+for mobile applications in Github and Googlecode, in order to get a first view
+of what is causing the crashes across applications available and what the
+characteristics of the stack traces can tell us about the exception structure of
+those applications. %Moreover this work\ldots.
 
 \section{Conclusion}
-% The line below was meant to be a joke :-)
-%Georgious wrote: That's a great paper with profound insights.
-
-In this work, we perform the first large scale analysis of Java stack traces and pinpoint how they can reveal bad programming practices.
-We mined and analyzed the stack traces embedded in all issues of Java projects available GitHub.
-Overall, our research set includes 356,057 issues define at 16,836 projects (from which 28,800 stack traces were extracted).
-From this set, the stack traces of 482 Android projects were investigated in more detail, in combination with source code and bytecode analysis.
-In this study some patterns of exception (mis)-use were consistently detected such as: unexpected wrappings (e.g., Errors being wrapped in checked exceptions) - revealing that Java hybrid exception model is not fully used according to its purpose;  undocumented runtime exceptions signaled by third party code - which makes almost impossible for library clients to protect against such exceptions; and a high prevalence of java.lang exceptions reported on issues - representing approximately 50\% of the analyzed issues.
-
-%Nowadays many software vendors embed automatic crash reporting tools in their software systems.
-Hence, whenever a software crashes this tool sends a detailed crash report to its vendors.
-Moreover, we can also find third party sofware solutions specialized in bug reporting for different kinds of systems specially for the increasing marked of mobile apps~\cite{BugSe14,BugSn14,Googl14,Acra14}.
-There is planty of information to be mined...
-
-\section*{Acknowledgment}
-This work is partially supported by: CNPq - Proc. 484209/2013-2 and the NWO TestRoots project (639.022.314).
+
+In this work, we perform the first large scale analysis of Java stack traces and
+pinpoint how they can reveal bad programming practices. We mined and analyzed
+the stack traces embedded in all issues of Java projects available GitHub.
+Overall, our research set includes 356,057 issues define at 16,836 projects
+(from which 28,800 stack traces were extracted). From this set, the stack traces
+of 482 Android projects were investigated in more detail, in combination with
+source code and bytecode analysis. In this study some patterns of exception
+(mis)-use were consistently detected such as: unexpected wrappings (e.g., Errors
+being wrapped in checked exceptions) - revealing that Java hybrid exception
+model is not fully used according to its purpose;  undocumented runtime
+exceptions signaled by third party code - which makes almost impossible for
+library clients to protect against such exceptions; and a high prevalence of
+java.lang exceptions reported on issues - representing approximately 50\% of the
+analyzed issues.
+
+%Nowadays many software vendors embed automatic crash reporting tools in their
+%software systems. Hence, whenever a software crashes this tool sends a detailed
+%crash report to its vendors. Moreover, we can also find third party sofware
+%solutions specialized in bug reporting for different kinds of systems specially
+%for the increasing marked of mobile apps~\cite{BugSe14,BugSn14,Googl14,Acra14}.
+%There is planty of information to be mined...
+
+\section*{Acknowledgment} This work is partially supported by: CNPq - Proc.
+484209/2013-2 and the NWO TestRoots project (639.022.314).
 

### 2014-05-07 android-stacks.tex
fixing table IX
small changes on paragraph

@@
 
-To answer this question we analyzed general characteristics of stack traces
-reported on Java repositories (e.g., exceptions being thrown presented in
-Tables~\ref{tab:toptenjava} and ~\ref{tab:toptenandroid}), and took a deeper
-look at the stack traces of the Android repositories combining them with with
-source code and byte code analyses. Doing so, we mined some characteristics of
-stacks that points to best practices violations.
+To answer this question we analyzed general characteristics of reported stack traces and, in the case of stack traces reported on Android repositories, we combine stack trace information with
+source code and byte code analyses to enable a more detailed analysis.
+Doing so, we mined some characteristics of stacks that points to best practices violations.
 
@@
 
-From Tables~\ref{tab:toptenjava}, ~\ref{tab:toptenandroid} we can observe that,
+From Tables~\ref{tab:toptenjava} and ~\ref{tab:toptenandroid} we can observe that,
 in both Java and Android related issues, direct instances
-java.lang.RuntimeException were thrown  (in 10,94\% of Java repositories and
-18,88\% of Android repositories respectively). Moreover, after inspecting the
+java.lang.RuntimeException were thrown  in 10,94\% of Java repositories and
+18,88\% of Android repositories respectively. Moreover, after inspecting the
 top 100 most thrown exceptions (see ~\ref{tophundrend}) we could observe that
 stack traces on which direct instances of java.lang.Exception and
-java.lang.Error were thrown. Throwing general exceptions is usually considered a
+java.lang.Error were thrown.
+
+Throwing general exceptions is usually considered a
 bad practice, because the type does not carry enough information to identify the
@@
     \hline
-    \bfseries{Type} & \bfseries{Android} & \bfseries{Lib} & \bfseries{Libcore} & \bfseries{Java} & \bfseries{App} & \bfseries{All}\\
+    \bfseries{Type} & \bfseries{Android} & \bfseries{Libcore} & \bfseries{App} & \bfseries{Lib} & \bfseries{Java} & \bfseries{All}\\
     \hline

### 2014-05-07 android-stacks.tex
Sections 1 and 2

--- 2014-05-07 android-stacks.tex
@@
 
-To answer this question, we mined the following information of every extracted
-stack trace: the top level exception being signaled, its signaler (the first
-method signature frame after the exception), the root exception (the exception
-which initiated the stack trace, defined on the bottom of the stack), and the
-root signaler. The intermediate elements that compose a stack trace as shown in
-Figure 1 (i.e., the intermediate causes and its execution stacks) were not be
-considered in the analysis.
+To answer this question, we mined from every extracted stack trace the top level
+exception being signaled, its signaler (the first method signature frame after
+the exception), the root exception (the exception which initiated the stack
+trace, defined on the bottom of the stack), and the root signaler. The
+intermediate elements that compose a stack trace as shown in
+Figure~\ref{fig:wrapping} (i.e., the intermediate causes and its execution
+stacks) were not be considered in the analysis.
 %~\todo{Move this to the previous section? IMO, It is better to keep all facts
@@
 
-Based on the assumption that, regardless the issue type every exception stack
-trace contains relevant information concerting the exception structure of
+Based on the assumption that regardless the issue type, every exception stack
+trace contains relevant information concerning the exception structure of
 projects analyzed we opted for not restricting the analysis only on defect
@@
 Table~\ref{tab:toptenjava} illustrates the top 10 most common root causes which
-represents more than 50\% of the analyzed set. The table illustrates  (i) the
+represents more than 50\% of the analyzed set. The table illustrates  i) the
 number of times each exception appeared as the root cause of a stack trace
-(\#occurrences), (ii) the number of distinct projects reporting such causes on
-stack traces, and (iii) the exception popularity. The exception popularity
-metric was defined in this work as follows: \emph{popularity = number of
-projects mentioning the exception in at least one stack/all projects that
-contain at least one stack}.
+(\#occurrences), ii) the number of distinct projects reporting such causes on
+stack traces, and iii) the exception popularity among projects.
 
@@
   \end{tabular}
-\caption{Top 10 most popular exceptions on issues}
+\caption{Top 10 most popular exceptions on issues in Java projects}
 \label{tab:toptenjava}
@@
 We can observe that the NullPointerException was the one with the higher number
-of occurrences (22,5\% of all stacks were caused by NullPointerExceptions) and
+of occurrences (26,3\% of all stacks were caused by NullPointerExceptions) and
 the one which had the highest popularity across projects. Actually, most of the
-exceptions in the TOP 10 list (except java.lang.IllegalArgumentException,
+exceptions in the top 10 list (except java.lang.IllegalArgumentException,
 java.io.IOException and java.lang.RuntimeException) are exceptions implicitly
-signaled by Java Virtual Machine, when an expression violates the normal
+thrown by Java Virtual Machine, when an expression violates the normal
 semantics of the Java programming language (e.g. ArrayIndexOutOfBounds) or due
@@
 The high prevalence of NullPointerExceptions, and the other implicitly-thrown
-exceptions  is aligned with the findings other works [Sunghun et al.][Gordon
-fraser and][Jcrasher].[Android-predict]~\todo{refs}. For instance, Sunghun et
-al. showed that in Eclipse bug report system 38\% the bugs related to exception
-handling were caused by NullPointerException; other works on robustness testing
-[JCrasher, Evol;suite, Amiya]~\todo{refs} showed that most of the automatically
+exceptions  is aligned with the findings other works~\cite{kim2013predicting,fraser20131600,csallner2004jcrasher}. For instance, Sunghun et
+al.~\cite{kim2013predicting} showed that in Eclipse bug report system 38\% the bugs related to exception handling were caused by NullPointerException; other works on robustness testing~\cite{maji2012empirical,csallner2004jcrasher}
+showed that most of the automatically
 detected bugs were due to NullPointerExceptions and implicitly-signaled of Java
@@
 
-As detailed in Section III, 482 Android projects were carefully selected to
-enable a deaper investigation of stack trace information, by leveraging stack
-trace information with the source code and bytecode analysis. For this subset of
-Android projects we could extract every package related to each exception
-signaler, and based on this information we could classify every stack trace
-based on its root signaler as presented in Table~\ref{signalers}.
+As detailed in Section~\ref{sec:android}, 482 Android projects were carefully
+selected to enable a deeper investigation of stack trace information, by
+leveraging stack trace information with the source code and bytecode analysis.
+For this subset of Android project, we could extract every package related to
+each exception signaler, and based on this information we could classify every
+stack trace based on its root signaler as presented in Table~\ref{tab:signalers}.
 
@@
 
-For each exception we could discover: the number of times it occurred on stacks,
+For each exception we could discover the number of times it occurred on stacks,
 the number of distinct projects such stacks were defined (and consequently its
@@
 distinct projects they were found and the whole set of projects analyzed with at
-least one stack trace (482). Table~\ref{toptenandroid} and
-Figuere~\ref{androidsignaler} presents the mined data.
+least one stack trace (482). Table~\ref{tab:toptenandroid} and
+Figure~\ref{fig:androidsignaler} presents the mined data.
 
@@
 
-%\textbf{
 Coding errors are the most common root causes for most types of signalers. We
@@
 number of occurrences. The NullPoiterExceptions are mainly signaled inside
-Android platform and inside application code. Although we also find
+Android platform and inside application code, although we also find
 NullPointerExceptions being signaled from third-party libraries. Regarding
-reusable code, there is no consensus if it s a good or bad practice to re-throw
-NullPointerException. Some prefer to encapsulate such exception on
-IllegalArgumentException. But others~\cite{joshuabloch} argue that the
-NullPointerException  explicits the cause of the problem and hence should not be
-wrapped.
+reusable code, there is no consensus if it is a good or bad practice to re-throw
+NullPointerException. Some prefer to encapsulate such an exception on
+IllegalArgumentException, while others~\cite{bloch2008effective} argue that the
+NullPointerException makes the cause of the problem explicit and hence 
+should not be wrapped.
 
 %commented only for submission:
-\textbf{We could observe in this study that the NullPointerException and other
+We could observe in this study that the NullPointerException and other
 other implicitly signaled exceptions represented most of the exceptions reported
-on issues in both Java and Android subset. For such exceptions - which represent
-programming bugs or resource limitations - there is usually no proper handling
+on issues in both Java and Android subset. For such exceptions, which represent
+programming bugs or resource limitations, there is usually no proper handling
 besides presenting an error message to the user and restarting the application -
-only high fault tolerant systems need to provide solutions to handle them [REF:
-Java Specification].}
+only high fault tolerant systems need to provide solutions to handle them.
+
+\noindent \fbox{
+\begin{minipage}{0.96\columnwidth} 
+\emph{RQ2: Most exceptions are due to programming mistakes
+or generally irrecoverable errors.\todo{Revise this}}
+\end{minipage}}
 

### 2014-05-07 android-stacks.tex
new table

--- 2014-05-07 android-stacks.tex
@@
 
-Throwing general exceptions is usually considered a
+Throwing general exceptions is  considered a
 bad practice, because the type does not carry enough information to identify the
@@
 Platform we could observe that direct instances of java.lang.RuntimeException
-they were thrown for 2 main reasons: (i) when there was no reasonable way of
-handling the exception, which means that regardless the kind of exception the
-only possible action was to abort the application; and when wrapping exceptions
+ were thrown due to two main reasons: (i) when there was no recovery action associated
+to the exceptional condition, which means that regardless the kind of exception the
+only possible action was to abort the application; and when translating exceptions
 signaled by application methods. Since Android platform is heavily based on
-"call backs", when the application does not handle its exceptions, they scape to
-Android platform which does not have contextual information to adequately handle
-the exception and simply terminates the application. What is apparently a bad
-practice of signaling general exceptions in this context it was and intended
-design decision.
+"call backs", when a application throws a runtime exception (i.e., one of the life cicle methods of an Activity) and does it, the exception scapes to compotenst of the Android platform 
+which do not have contextual information to adequately handle
+the exception and simply terminates the application.
+
+We could observe that although been known as a bad practice, the signaling general exceptions in this context it was an intended design decision when no specific recovery action can be defined for the exception.
 
@@
 
-Runtime   & 1101 & 56  & 1446 & 423 & 10 & 3036  \\
-Error     & 148  & 34  & 168  & 114 & 9  & 473 \\
-Checked   & 115  & 241 & 139  & 176 & 12 & 683 \\
+Runtime	&	1075	&	56	&	1446	&	374	&	10	&	2961	\\
+Error	&	144	&	34	&	168	&	105	&	9	&	460	\\
+Checked	&	110	&	230	&	139	&	145	&	12	&	636	\\
+Throwable	&	0	&	0	&	1	&	0	&	0	&	1	\\
 Undefined	&	1	&	0	&	6	&	8	&	0	&	15	\\
-Throwable & 0    & 0   & 1    & 0   & 0  & 1 \\
-    \hline
-All       &	1365	&	331	&	1760	&	721	&	31	&	4208	\\
+    \hline
+All		& 1330	&	320	&	1760	&	632	&	31	&	4073	\\
     \hline
@@
     \hline
-    \bfseries{Exception Type} & \bfseries{Root Cause} & \bfseries{Occurrences} & \bfseries{Projects} \\
-    \hline
-    Runtime   & None    & 2360 & 359 \\
+    \bfseries{Root Cause} & \bfseries{Wrapper Type} & \bfseries{Occurrences} & \bfseries{Projects} \\
+    \hline
+     Runtime    & -   &  2360 & 359 \\
+
     Runtime   & Runtime & 560  & 182 \\
-    Checked   & None    & 422  & 117 \\
-    Error     & None    & 381  & 141 \\
-    Runtime   & Checked & 109  & 70  \\
+
+ Checked    &   -  & 422  & 117 \\
+     Error    &   -   & 381  & 141 \\
+   Checked &  Runtime   & 109  & 70  \\
     Checked   & Checked & 98   & 42  \\
-    Runtime   & Error   & 55   & 38  \\
-    Checked   & Runtime & 22   & 12  \\
+   Error   &  Runtime   & 55   & 38  \\
+   Runtime &  Checked   & 22   & 12  \\
     Error     & Error   & 15   & 14  \\
-    Undefined & None    & 15   & 10  \\
-    Error     & Runtime & 13   & 6 \\
-    Checked   & Error   & 8    & 7 \\
-    Error     & Checked & 7    & 4 \\
-    Throwable & Runtime & 3    & 1 \\
-    Undefined & Runtime & 3    & 3 \\
-    Throwable & None    & 1    & 1 \\
-    Undefined & Error   & 1    & 1 \\
+  Undefined    &  -  &  15   & 10  \\
+   Runtime & Error     &  13   & 6 \\
+  Error   & Checked   &  8    & 7 \\
+    Checked &Error     &  7    & 4 \\
+    Runtime &Throwable &  3    & 1 \\
+   Runtime & Undefined &  3    & 3 \\
+   Throwable    & -  &  1    & 1 \\
+   Error   & Undefined &  1    & 1 \\
     \hline

### 2014-05-07 android-stacks.tex
Related work rework

--- 2014-05-07 android-stacks.tex
@@
 
-In this section, we present works thaat are related to our own, distributed in
-four categories: (i) works that use the information available on stack traces;
-(ii) empirical studies on the usage of Java exceptions and its fault proneness;
-(iii) tools to extract stack traces information from natural language artifacts
-(i.e., issues and emails);  and (iv) empirical studies involving Android apps.
+In this section, we present works that are related to our own, distributed in
+four categories: i) works that use the information available on stack traces;
+ii) empirical studies on the usage of Java exceptions and its fault proneness;
+iii) tools to extract stack traces information from natural language artifacts
+(i.e., issues and emails), and iv) empirical studies involving Android apps.
 
 \textit{Analysis of Java Exception Usage.} Since the manual analysis of the Java
-exception-flow can easily become infeasible Robillard and Murphy~\cite{Robil00}
-developed a tool called Jex that analyzes the flow of exceptions in Java source
-code. Based on java source code this tool performs dataflow analysis in order to
-find the propagation paths of checked and unchecked exception types. After this
-tool other tools have been proposed to support the static analysis of exception
-flows ~\cite{fu2007exception} ~\cite{coelho2008assessing}. The main limitation
-off all tools is the number of false positives inherent to static analysis
+exception flow can easily become infeasible, Robillard and Murphy~\cite{Robil00}
+employed dataflow analysis to find the propagation paths of checked and
+unchecked exception types. Modeled after Robillard and Murphy's work, other
+tools have been proposed to support the static analysis of exception
+flows~\cite{coelho2008assessing}. The main limitation off all
+static analysis tools is the number of false positives inherent to static analysis
 solutions, which can lead to a high number of exception flows, specially if
 considering Java Environment exceptions and exceptions signaled from libraries.
-This was the fist of some tools proposed to statically analyzed the exception
-flow of exceptions. Cabral and Marques~\cite{cabral2007exception} analyzed the
-exception handling code of 32 open-source systems, both for Java and .NET. The
-goal of their study was to identify how exceptions were handled in different
-categories of systems. They examined the exception handlers and the respective
-actions taken on them. They observed that the action handlers were very simple
-(e.g., logging and present a message to the user). In our work differs from
-previous works since in our study we tried to identify bad practices from the
-combined use of stack traces extracted from issues and bytecode and source code
-analysis. Finally, Reimer and Srinivasan~\cite{reimer2003analyzing} listed a set
-of bad practices on exception handling that hinder software maintainability and
-robustness. Their differs from ours it was solely based on their own experience
-on working with Java enterprise applications and focus on source code examples
-of bad patterns on ways of handling exceptions such as exception swallowing, log
-verbosity.
-
-\textit{Analysis and Use of Stack Trace Information.} Some works have
-investigated the usage of stack trace information to support bug classification
+
+Additionally, Cabral and Marques~\cite{cabral2007exception} analyzed the
+exception handling code of 32 open-source systems, both for Java and .NET. They
+observed that the action handlers were very simple (e.g., logging and present a
+message to the user). Reimer and Srinivasan~\cite{reimer2003analyzing} listed a
+set of bad practices on exception handling that hinder software maintainability
+and robustness, based on their own experience with Java enterprise applications.
+Our work differs from those two works as we tried to identify bad practices from
+the combined use of stack traces extracted from issues and bytecode and source
+code analysis.
+
+\textit{Analysis and Use of Stack Trace Information.} Several works have
+investigated the use of stack trace information to support bug classification
 and clustering~\cite{wang2013improving, kim2011crash, dhaliwal2011classifying},
 fault-proneness prediction models~\cite{kim2013predicting} and even automated
-bug fixing~\cite{sinha2009fault} tools.Kim et al~\cite{kim2011crash} use of
-crash graphs (an aggregated form of multiple stack traces available in crash
-reports) to detect duplicate crash reports and to predict if a given crash will
-be fixed. Dhaliwal et al~\cite{} analyzed crash reports collected for Mozilla
-Firefox and proposed a crash grouping approach based on the Levenshtein
-distance. This work showed that a crash grouping approach can reduce bug fixing
-time in approximately 5\%. Wang et al~\cite{} proposes an approach to define
-correlated crash types automatically. Moreover it also provides a bug
-localization method to locate and rank files related to the bug described on a
-stack trace.  Schroter et al~\cite{schroter2010stack} conducted an empirical
-study on the usefulness of stack traces for bug fixing. The study used the
-development data of Eclipse project and could observe that developers fixed the
-bugs faster when failing stack traces were included on bug issues. This study is
-aligned with the one performed by Bettenburg et al. ~\cite{bettenburg2008makes}
-which evaluated what content could make a good bug report, and stack traces were
-mentioned as the second most important content for developers - just after the
-"steps to reproduce" information. Sinha et al~\cite{sinha2009fault} proposed an
-approach for locating and repairing faults that are caused by JVM implicitly
-signaled exceptions related to coding errors such as NullPointerExceptions. This
-approach uses the information available on the stack trace to guide a dataflow
-analysis, to locate the source statement responsible for the exception. Kim at
-al.~\cite{kim2013predicting} proposed an approach to predict the crash-proneness
-of methods based information extracted from stack traces and methods' bytecode
-operations. They used InfoZilla to extracted stack trace information from the
-bug reports of two open-source projects, i.e. AspectJ and Eclipse. They could
-observe that most of the stack traces were related to NullPointerException and
-other JMV implicitly thrown exceptions had the higher prevalence on the analyzed
-set of stacks.
-
-\textit{Extracting Stack Traces from Natural language artifacts.} Nowadays many
-software vendors embed automatic crash reporting tools in their software
-systems. Hence, whenever a software crashes this tool sends a detailed crash
-report to its vendors. Moreover, there are also find third party software
-solutions specialized in bug reporting, most of them for the increasing marked
-of mobile apps~\cite{BugSe14,BugSn14,Googl14,Acra14}. Hence most of the works
-described previously used the  information provided by crass reporting tools.
-Two works have proposed tools extract stack traces from natural language
-artifacts such as issues and emails: Infozilla ~\cite{bettenburg2008extracting}
-is based on a set of regular expressions that extract an the set of frames
-related to a stack trace, the main limitation of this solution is that it is not
+bug fixing tools~\cite{sinha2009fault}. Kim et al.~\cite{kim2011crash} use an
+aggregated form of multiple stack traces available in crash reports to detect
+duplicate crash reports and to predict if a given crash will be fixed. Dhaliwal
+et al.~\cite{dhaliwal2011classifying} proposed a crash grouping approach that
+can reduce bug fixing time in approximately 5\%. Wang et
+al.~\cite{wang2013improving} propose an approach to identify correlated crash
+types and describe a fault localization method to locate and rank files related
+to the bug described on a stack trace. Schroter et al.~\cite{schroter2010stack}
+conducted an empirical study on the usefulness of stack traces for bug fixing
+and showed that developers fixed the bugs faster when failing stack traces were
+included on bug issues.  In a similar study, Bettenburg et
+al.~\cite{bettenburg2008makes} identify stack traces as the second most stack
+trace feature for developers.  Sinha et al.~\cite{sinha2009fault} proposed an
+approach that uses stack traces to guide a dataflow analysis for locating and
+repairing faults that are caused by the JVM implicitly signaled exceptions. Kim
+at al.~\cite{kim2013predicting} proposed an approach to predict the
+crash-proneness of methods based information extracted from stack traces and
+methods' bytecode operations.  They observed that most of the stack traces were
+related to NullPointerException and other JMV implicitly thrown exceptions had
+the higher prevalence on the analyzed set of stacks.
+
+\textit{Extracting Stack Traces from natural language artifacts.} 
+%Currently, many
+%software vendors embed automatic crash reporting tools in their software
+%systems. Moreover, third party crash collection services exist, most of them 
+%targeted for applications run on mobile phones~\cite{BugSe14,BugSn14,Googl14,Acra14}.
+Apart from bug reports, stack traces can be embedded in other forms of
+communication between developers, such as discussion logs and emails.
+Being intermixed with text makes the accurate extraction of stacktraces 
+an involved process.
+Infozilla~\cite{bettenburg2008extracting}
+is based on a set of regular expressions that extract a set of frames
+related to a stack trace. The main limitation of this solution is that it is not
 able to extract stack traces embedded on verbose log files (i.e., on which we
-can find log text mixed with exception frames). Alberto and Lanza
+can find log text mixed with exception frames). Bacchelli and Lanza
 ~\cite{bacchelli2012content} propose a solution to recognize stack trace frames
 from development emails and relate it to code artifacts (i.e. classes) mentioned
-on the stack trace.
+on the stack trace. In addition to those tools, ExceptionMiner is able to 
+both extract stack traces from natural language artifacts and to 
+classify them in a set of predefined categories.
 
@@
 distilling the information of bug reports describing uncaught exceptions created
-for mobile applications in Github and Googlecode, in order to get a first view
+for mobile applications in Github, in order to get a first view
 of what is causing the crashes across applications available and what the
 characteristics of the stack traces can tell us about the exception structure of
-those applications. %Moreover this work\ldots.
+those applications.
 

### 2014-05-07 scripts-data/boxplots.R
Scripts and data used for parts of the analysis

@@ -1,0 +1,28 @@
+library(ggplot2)
+library(reshape)
+
+data_android <- read.csv('android_repo_issues_metrics.txt', sep=';')
+data_android$with_exception = (data_android$with_exception/ data_android$all) * 100
+data_android$with_stack = (data_android$with_stack/ data_android$all) * 100
+data_android$defect_issues = (data_android$defect_issues/ data_android$all) * 100
+data_android$defect_with_exception = (data_android$defect_with_exception/ data_android$all) * 100
+data_android$defect_with_stack = (data_android$defect_with_stack/ data_android$all) * 100
+data_android <- data_android[ , -which(names(data_android) %in% c("all"))]
+android_melted <- melt(data_android)
+
+data_java <- read.csv('java_repo_issues_metrics.txt', sep=';')
+data_java$with_exception = (data_java$with_exception/ data_java$all) * 100
+data_java$with_stack = (data_java$with_stack/ data_java$all) * 100
+data_java$defect_issues = (data_java$defect_issues/ data_java$all) * 100
+data_java$defect_with_exception = (data_java$defect_with_exception/ data_java$all) * 100
+data_java$defect_with_stack = (data_java$defect_with_stack/ data_java$all) * 100
+data_java <- data_java[ , -which(names(data_java) %in% c("all"))]
+java_melted <- melt(data_java)
+
+pdf("android-stacks.pdf")
+ggplot(android_melted) + aes(x=variable, y = value) + geom_boxplot() + scale_y_log10(name = "Percentage (log)") +  scale_x_discrete(name="Variable") + theme(axis.text.x = element_text(angle = 90)) + ggtitle("Android Projects")
+dev.off()
+
+pdf("java-stacks.pdf")
+ggplot(java_melted) + aes(x=variable, y = value) + geom_boxplot() + scale_y_log10(name = "Percentage (log)") +  scale_x_discrete(name="Variable") + theme(axis.text.x = element_text(angle = 90)) + ggtitle("Java Projects")
+dev.off()

### 2014-05-07 scripts-data/crosscorrelation.R

@@ -1,0 +1,41 @@
+library(ellipse)
+library(ggplot2)
+library(reshape)
+library(xtable)
+
+issue_metrics <- read.csv('java_repo_issues_metrics.txt', sep=';')
+src_metrics <- read.csv('java_repo_src_metrics.txt')
+src_metrics$repo_java <- apply(src_metrics, 1, function(x){strsplit(x,
+                                                                    "/")[[1]][2]})
+src_metrics$repo_java <- as.factor(src_metrics$repo_java)
+
+combined<-merge(src_metrics, issue_metrics, by.x="repo_java", by.y="X.repo_java")
+
+used <- combined[ , -which(names(combined) %in% c('repo_java', 'repo'))]
+
+# Cross correlation ellipses
+ctab <- cor(used, method = "spearman", use='complete.obs')
+colorfun <- colorRamp(c("#ff0000","white","#0000ff"), space="rgb")
+pdf("cross-cor.pdf")
+plotcorr(ctab,col=rgb(colorfun((ctab+1)/2), maxColorValue=255),
+         outline = FALSE)
+dev.off()
+
+# Cross correlation table
+print(xtable(ctab,
+             caption="Cross correlation matrix (Spearman) between examined factors",
+             label="tab:crosscor"),
+      type = "latex",
+      size = "small",
+      file = "cross-cor.tex")
+
+# Cross correlation heatmap
+ctab.m <- melt(ctab)
+pdf("cross-cor-heat.pdf")
+ggplot(ctab.m, aes(X1, X2, fill = value)) +
+  geom_tile() +
+  scale_fill_gradient2(space = "Lab") +
+  theme(axis.title = element_blank(),
+        axis.text = element_text(size = 11),
+        axis.text.x = element_text(angle = -90, hjust = 0, vjust = 0.5))
+dev.off()

### 2014-05-07 scripts-data/exceptions-assoc-rules.R

@@ -1,0 +1,50 @@
+#install.packages('arules')
+#install.packages('arulesViz')
+
+library('arules')
+library('arulesViz')
+
+merge.dataframes <- function(dfs, min_num_rows = 1) {
+  Reduce(function(acc, x){
+    if (nrow(x) >= min_num_rows) {
+      rbind(acc, x)
+    } else {
+      printf("Warning: %s has less than %d rows (%d), skipping", project.name(x), min_num_rows, nrow(x))
+      acc
+    }
+  }, dfs)
+}
+
+files <- list.files(pattern= "*.csv")
+csvs <- Map(function(x){read.csv(x, sep = ";")}, files)
+all <- merge.dataframes(csvs)
+all <- all[,!(names(all) %in% c("stack"))]
+
+# assoc.rules.df <- all[,(names(all)) %in% c("exception_extends", "exception_from", 
+#                                            "main_extends", "main_from", 
+#                                            "mainsignaler_from", "root_extends", 
+#                                            "root_from")]
+
+#assoc.rules.df <- all[,(names(all)) %in% c("exception_extends", "signaler_from", 
+#                                            "root_extends", "rootsignaler_from")]
+
+assoc.rules.df <- all[,(names(all)) %in% c("main_extends", "mainsignaler_from", "exception_extends")]
+
+rules <- apriori(assoc.rules.df, 
+                 parameter = list(minlen=2, supp=0.005, conf=0.8),
+                 appearance = list(rhs=c("exception_extends=CHECKED", 
+                                         "exception_extends=ERROR", 
+                                         "exception_extends=RUNTIME", "exception_extends=UNDEFINED"), 
+                                   default="lhs"),
+                 control = list(verbose=F))
+
+rules.sorted <- sort(rules, by="lift")
+
+subset.matrix <- is.subset(rules.sorted, rules.sorted)
+subset.matrix[lower.tri(subset.matrix, diag=T)] <- NA
+redundant <- colSums(subset.matrix, na.rm=T) >= 1
+rules.pruned <- rules.sorted[!redundant]
+
+inspect(rules.pruned)
+
+plot(rules.pruned, method="graph", control=list(type="items", cex=0.6))

### 2014-05-07 scripts-data/parse_results.rb

@@ -1,0 +1,42 @@
+#!/usr/bin/env ruby
+
+require 'pp'
+
+fields = %w(repo loc num_classes num_public_methods public_with_explicit_throws
+public_with_explicit_throws_percent nonpublic_methods
+nonpublic_with_explicit_throws nonpublic_with_explicit_throws
+num_throw_statements num_try num_catch throw_on_catch return_on_catch
+log_on_catch num_finally throw_on_finally return_on_finally log_on_finally
+)
+
+puts fields.join(',')
+
+Dir['*.txt'].map do |file|
+  contents = File.open(file).read.split(/\n/)
+
+  next if contents.empty?
+
+  up_to = contents.find_index { |x| x.start_with? 'Most Frequently caught' }
+
+  contents[0..(up_to - 1)].reduce({}) do |acc, line|
+    unless line.start_with? '-----'
+
+      var, val = line.split(/:/)
+      var = var.strip.downcase
+      var = var.tr '-', ''
+      val = val.strip.downcase
+
+      if var == 'system'
+        var = 'repo'
+        val = val.split(/\//)[1]
+        val = val.tr '@', '/'
+      end
+      acc[var] = val
+    end
+    acc
+  end
+end.select do |x|
+  !x.nil?
+end.each do |x|
+  puts fields.map{ |f| x[f].to_s }.join(',')
+end

### 2014-05-07 scripts-data/projects.sh

@@ -1,0 +1,11 @@
+#!/usr/bin/env bash
+
+mkdir -p repo
+
+cat repos.txt|
+cut -f5,6 -d'/'|
+while read repo; do
+  repout=`echo $repo|tr '/' '@'`
+  git clone https://github.com/$repo.git repos/$repout
+  java -cp metrics.jar br.ufrn.exminer.sourcecode.metrics.CalculateEHMetrics repos/$repout $repout-result.txt
+done

### 2014-05-07 scripts-data/stack-sizes-plots.R

@@ -1,0 +1,8 @@
+library(ggplot2)
+library(reshape)
+library(plyr)
+
+data <- read.csv('stack-sizes.csv', sep=';')
+ggplot(data) + facet_grid(size ~ type) + geom_histogram() + aes(x = )
+
+

### 2014-05-07 scripts-data/summary-stats.R

@@ -1,0 +1,158 @@
+library(ellipse)
+library(ggplot2)
+library(reshape)
+library(xtable)
+library(digest)
+library(xtable)
+library(plyr)
+
+quant5 <- function(x){quantile(x, 0.05)}
+quant95 <- function(x){quantile(x, 0.95)}
+
+run <- function(all = data.frame(), descr.stats = data.frame(), 
+                ref="descr.stats", capt = "" ) {
+
+  descr.stats$Feature <- as.character(descr.stats$Feature)
+
+  descr.stats$quant_5 <- lapply(descr.stats$Feature, 
+                                function(x){quantile(all[,x], 0.05,na.rm = T)})
+  descr.stats$mean <- lapply(descr.stats$Feature, 
+                             function(x){mean(all[,x], na.rm = T)})
+  descr.stats$median <- lapply(descr.stats$Feature, 
+                               function(x){median(all[,x], na.rm = T)})
+  descr.stats$quant_95 <- lapply(descr.stats$Feature, 
+                                 function(x){quantile(all[,x], 0.95, na.rm = T)})
+
+  descr.stats$histogram <- lapply(descr.stats$Feature, function(x) {
+  
+    print(sprintf("Histogram for %s", x))
+    data <- all[, x]
+    unq <- digest(sprintf("%s.%s", ref, as.character(x)))
+    fname <- sprintf("hist-%s.pdf",unq)
+
+    par(mar=c(0,0,0,0))
+    plot.window(c(0,1),c(0,1),  xaxs='i', yaxs='i')
+    pdf(file = fname , width = 6, height = 3)
+    hist(log(data), probability = TRUE, col = "red", border = "white",
+         breaks = 10, xlab = "", ylab = "", axes = F, main = NULL)
+    dev.off()
+    sprintf("\\includegraphics[scale = 0.1, clip = true, trim= 50px 60px 50px 60px]{hist-%s.pdf}", unq)
+})
+
+  table <- xtable(descr.stats, label=sprintf("tab:%s", ref),
+                  caption= sprintf("%s%s", capt, " Historgrams are in log scale."),
+                  align = c("l", "r","p{16em}", rep("r", 4), "c"))
+
+  print.xtable(table, file = sprintf("%s.tex", ref),
+               floating.environment = "table*",
+               include.rownames = F, size = c(-2),
+               sanitize.text.function = 
+                  function(str)gsub("_","\\_",str,fixed=TRUE))
+}
+
+# Java src code statistics
[…]

### 2014-05-07 scripts-data/stack-sizes-plots.R
Stack and wrappings historgrams

@@ -3,6 +3,61 @@ library(reshape)
 library(plyr)
 
 data <- read.csv('stack-sizes.csv', sep=';')
-ggplot(data) + facet_grid(size ~ type) + geom_histogram() + aes(x = )
 
+multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
+  require(grid)
 
+  # Make a list from the ... arguments and plotlist
+  plots <- c(list(...), plotlist)
+
+  numPlots = length(plots)
+
+  # If layout is NULL, then use 'cols' to determine layout
+  if (is.null(layout)) {
+    # Make the panel
+    # ncol: Number of columns of plots
+    # nrow: Number of rows needed, calculated from # of cols
+    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
+                    ncol = cols, nrow = ceiling(numPlots/cols))
+  }
+
+ if (numPlots==1) {
+    print(plots[[1]])
+
+  } else {
+    # Set up the page
+    grid.newpage()
+    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
+
+    # Make each plot, in the correct location
+    for (i in 1:numPlots) {
+      # Get the i,j matrix positions of the regions that contain this subplot
+      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
+
+      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
+                                      layout.pos.col = matchidx$col))
+    }
+  }
+}
+
+p1 <- ggplot(subset(data, type == 'java')) + 
+      aes(x = size) + 
+      geom_histogram(bin = 5) +
+      ggtitle("Java")
+p2 <- ggplot(subset(data, type == 'android')) + 
+      aes(x = size) + 
+      geom_histogram(bin = 5) +
+      ggtitle("Android")
+
+p3 <- ggplot(subset(data, type == 'java')) + 
+      aes(x = wrappings) + 
+      geom_histogram() +
+      ggtitle("Java")
+p4 <- ggplot(subset(data, type == 'android')) + 
+      aes(x = wrappings) + 
+      geom_histogram() +
+      ggtitle("Android")
+
+pdf("stack-size-wrappings-hist.pdf")
+multiplot(p1, p3, p2, p4, cols = 2)
+dev.off()

### 2014-05-07 android-stacks.tex
Stack traces historgarms

@@
 
+\begin{figure}
+  \begin{center}
+    \includegraphics[scale=0.5]{stack-size-wrappings-hist.pdf}
+  \end{center}
+  \caption{Stack sizes (top) and wrappings (bottom) for Java (left) and Android (right) stack traces}
+  \label{fig:sizewrapbhists}
+\end{figure}
+
 Table~\ref{wrappingandroid} presents the wrappings found in this study for all
@@
 
-\subsection{Internal Validity} We used a heuristics-based parser to mine
+\noindent\emph{Internal Validity} We used a heuristics-based parser to mine
 exceptions from issues.  Our parsing strategy was conservative by default; for
@@
 
-\subsection{External Validity} The results presented here were based on mining
+\noindent\emph{External Validity} The results presented here were based on mining
 issues on Github through the GHTorrent dataset. While comprehensive and
@@
 
-\subsection{Construct Validity}
-Parts of our analysis are based on the availability of stack traces on issues on
+\noindent\emph{Construct Validity} Parts of our analysis are based on the availability of stack traces on issues on
 Github. In using this dataset, we make an underlying assumption: the

### 2014-05-07 android-stacks.tex
A bit more on RQ2

@@
 
-\input{android-stack-stats}
-
 %Based on the assumption presented in Section III, we therefore did not use this label for further filtering.
@@
 
-\noindent \fbox{
-\begin{minipage}{0.96\columnwidth} 
-\emph{RQ2: Most exceptions are due to programming mistakes
-or generally irrecoverable errors.\todo{Revise this}}
-\end{minipage}}
-
 \subsection{Classification of the main causes of top exceptions}
 
-To better understand the main causes of such exceptions we evaluated semantics
+To better understand the main causes of such exceptions, we evaluated semantics
 related to the top 100 root causes reported on stack-issues of both Java dataset
@@
 Android stack-issues), and classified them according to the the categories
-presented in Table~\ref{tophundrend}.
+presented in Table~\ref{tab:tophundrend}.
 
@@
 
-As we can see from Table 4, the that programming errors are the causes of most
-of the exception stack traces reported on issues. In the Android subset we could
-find more issues related to Backward compatibility and security issues than on
-the general data set.
+Programming errors are the causes of most of the exception stack traces reported
+on issues. In the Android subset we could find more issues related to Backward
+compatibility and security issues than on the general data set.
 
@@
 Manually inspecting the corresponding execution trace was caused by one of the
-two reasons: recursive calls; exceptions signaled by a method defined deep down
-in a reused framework; exceptions successively wrapped. Regardless the reason,
-in exception flows involving many methods, it is piratically impossible to
-handle the exception a way different than existing the  application, because
-there is hardly enough contextual information to perform any other recovery
-action.
-
-%TODO: Include a Histogram here.
+two reasons: recursive calls or exceptions signaled by a method defined deep
+down in a reused framework; exceptions successively wrapped. Irrespective of the
+reason, in exception flows involving many methods, it is practically impossible
+to handle an exception in a way other than exiting the application, as at the
+place the exception occurs there is hardly enough contextual information to
+perform recovery actions.
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[scale=0.5]{stack-size-wrappings-hist.pdf}
+  \end{center}
+  \caption{Stack sizes (top) and wrappings (bottom) for Java (left) and Android (right) stack traces}
+  \label{fig:sizewrapbhists}
+\end{figure}
 
@@
 
+\noindent \fbox{
+\begin{minipage}{0.96\columnwidth} 
+\emph{RQ2: Most exceptions are due to programming mistakes
+or generally irrecoverable errors.}
+\end{minipage}}
+
 \section{RQ3.  What kinds of exception (mis)-use emerge from stack trace analysis? }
@@
 \subsection{Odd Wrappings}
-
-\begin{figure}
-  \begin{center}
-    \includegraphics[scale=0.5]{stack-size-wrappings-hist.pdf}
-  \end{center}
-  \caption{Stack sizes (top) and wrappings (bottom) for Java (left) and Android (right) stack traces}
-  \label{fig:sizewrapbhists}
-\end{figure}
-
 Table~\ref{wrappingandroid} presents the wrappings found in this study for all
@@
 ii) empirical studies on the usage of Java exceptions and its fault proneness;
-iii) tools to extract stack traces information from natural language artifacts
-(i.e., issues and emails), and iv) empirical studies involving Android apps.
+and iii) tools to extract stack traces information from natural language artifacts
+(i.e., issues and emails)
+%, and iv) empirical studies involving Android apps.
 
@@
 considering Java Environment exceptions and exceptions signaled from libraries.
-
 Additionally, Cabral and Marques~\cite{cabral2007exception} analyzed the
@@
 
-\textit{Empirical studies using Android apps.} Ruiz et al.~\cite{Ruiz12}
-investigated the degree of reuse across applications in Android Market, the
-study showed that almost 23\% of the classes inherited from a base class in the
-Android API, and that 217 mobile apps were reused completely by another mobile
-app. Pathak et al.~\cite{Patha11} analyzed bug reports and developers
-discussions of Android platform and found out that approximately 20\% of
-energy-related bugs in Android occurred after an OS update. McDonnell et
-al.~\cite{McDon13} conducted a case study of the co-evolution behavior of
-Android API and 10 dependent applications using the version history data found
-in github. The study found that approximately 25\% of all methods in the client
-code used the Android API, and that the methods reusing fast-evolving APIs were
-more defect prone then others. Vásquez et al.~\cite{Linar13} analyzed
-approximately 7K free Android apps and observed that the last successful apps
-used Android APIs that were on average 300\% more change-prone than the APIs
-used by the most successful apps. Pingyu and Elbaum~\cite{Zhang12} analyzed bug
-reports of 5 Android applications an observed that 29\% had to do with poor
-exceptional handling code. Our work differs from the others as it aims at
-distilling the information of bug reports describing uncaught exceptions created
-for mobile applications in Github, in order to get a first view
-of what is causing the crashes across applications available and what the
-characteristics of the stack traces can tell us about the exception structure of
-those applications.
+%\textit{Empirical studies using Android apps.} Ruiz et al.~\cite{Ruiz12}
+%investigated the degree of reuse across applications in Android Market, the
+%study showed that almost 23\% of the classes inherited from a base class in the
+%Android API, and that 217 mobile apps were reused completely by another mobile
+%app. Pathak et al.~\cite{Patha11} analyzed bug reports and developers
+%discussions of Android platform and found out that approximately 20\% of
+%energy-related bugs in Android occurred after an OS update. McDonnell et
+%al.~\cite{McDon13} conducted a case study of the co-evolution behavior of
+%Android API and 10 dependent applications using the version history data found
+%in github. The study found that approximately 25\% of all methods in the client
+%code used the Android API, and that the methods reusing fast-evolving APIs were
+%more defect prone then others. Vásquez et al.~\cite{Linar13} analyzed
+%approximately 7K free Android apps and observed that the last successful apps
+%used Android APIs that were on average 300\% more change-prone than the APIs
+%used by the most successful apps. Pingyu and Elbaum~\cite{Zhang12} analyzed bug
+%reports of 5 Android applications an observed that 29\% had to do with poor
+%exceptional handling code. Our work differs from the others as it aims at
+%distilling the information of bug reports describing uncaught exceptions created
+%for mobile applications in Github, in order to get a first view
+%of what is causing the crashes across applications available and what the
+%characteristics of the stack traces can tell us about the exception structure of
+%those applications.
 

### 2014-05-07 stacktrace-stats.tex
A bit more on RQ2

@@
   \hline
+   \multicolumn{2}{l}{\bfseries{Java}}  \\
 all & All issues recorded for the project. & 1.00 & 23.37 & 3.00 & 80.00 & \includegraphics[scale = 0.1, clip = true, trim= 50px 60px 50px 60px]{hist-31b77719e3773ad7ac2073785de62a5f.pdf} \\ 
@@
    \hline
+   \multicolumn{2}{l}{\bfseries{Android}}\\
+   all & All issues recorded for the project. & 2.00 & 67.70 & 20.00 & 248.00 & \includegraphics[scale = 0.1, clip = true, trim= 50px 60px 50px 60px]{hist-5d4a87daa44c63097df9bb1f3506d541.pdf} \\ 
+  with\_exception & All issues featuring an exception.
+& 1.00 & 8.16 & 2.00 & 34.00 & \includegraphics[scale = 0.1, clip = true, trim= 50px 60px 50px 60px]{hist-bf73235d0503f937ef6f36c28cf1ba0d.pdf} \\
+  with\_stack & All issues featuring a stack trace. & 1.00 & 6.44 & 2.00 & 27.00 & \includegraphics[scale = 0.1, clip = true, trim= 50px 60px 50px 60px]{hist-528cb6be67074e6e2c22d11245920e13.pdf} \\ 
+  defect\_issues & All issues labeled as defect. & 0.00 & 13.28 & 1.00 & 52.00 & \includegraphics[scale = 0.1, clip = true, trim= 50px 60px 50px 60px]{hist-eb1797e86ef23b2a07907e5ccbeff7b9.pdf} \\ 
+  defect\_with\_exception & All issues labeled labeled as defect and including an exception.
+& 0.00 & 3.25 & 0.00 & 13.00 & \includegraphics[scale = 0.1, clip = true, trim= 50px 60px 50px 60px]{hist-f1a879616aa3e0995db35c6c02bc8520.pdf} \\
+  defect\_with\_stack & All issues labeled labeled as defect and including a stacktrace & 0.00 & 2.92 & 0.00 & 10.00 & \includegraphics[scale = 0.1, clip = true, trim= 50px 60px 50px 60px]{hist-49dbfa02c366cab8b489f33014de587d.pdf} \\
+ \hline
  \end{tabular}

### 2014-05-07 android-stacks.tex
removing duplicated tables.

@@
 
-\subsection{General exceptions being thrown.}
+\subsection{Direct instances of RuntimeExceptions been thrown.}
 
@@
 java.lang.RuntimeException were thrown  in 10,94\% of Java repositories and
-18,88\% of Android repositories respectively. Moreover, after inspecting the
-top 100 most thrown exceptions (see ~\ref{tophundrend}) we could observe that
-stack traces on which direct instances of java.lang.Exception and
-java.lang.Error were thrown.
-
-Throwing general exceptions is  considered a
-bad practice, because the type does not carry enough information to identify the
-cause of the exceptional behaviour. In such cases, developers need to relying on
-exception message which can be neither complete nor precise
+18,88\% of Android repositories respectively. 
+In the Android repositories, most of such exceptions  were  thrown by the Android 
+platform/os (167 out of 232).
+
+Throwing general exceptions, such as direct instances of RuntimeException, is  considered a
+bad practice, because the exception type does not carry enough information to identify the
+cause of the exceptional behaviour, and as a consequence developers need to relying on
+the exception message which can be neither complete nor precise
 ~\cite{gosling2000java}.
 
-In Table ~\ref{toptenandroid} we can observe a high prevalence of
-RuntimeExceptions being thrown by the Android platform (167 out of 232).
-Manually inspecting the code on which RuntimeExceptions were thrown in Android
-Platform we could observe that direct instances of java.lang.RuntimeException
- were thrown due to two main reasons: (i) when there was no recovery action associated
-to the exceptional condition, which means that regardless the kind of exception the
-only possible action was to abort the application; and when translating exceptions
-signaled by application methods. Since Android platform is heavily based on
-"call backs", when a application throws a runtime exception (i.e., one of the life cicle methods of an Activity) and does it, the exception scapes to compotenst of the Android platform 
-which do not have contextual information to adequately handle
-the exception and simply terminates the application.
-
-We could observe that although been known as a bad practice, the signaling general exceptions in this context it was an intended design decision when no specific recovery action can be defined for the exception.
-
-%Even an instance of Throwable as signaled in one project. It points to a the
-%"dont worry from know behaviour".
-
-%\subsection{Combining Stack Trace Information with Exception Type Information}
-
-%todo{DISCOMMENT AFTER SUBMISSION}
-%\subsection{Undocumented runtime exceptions signaled by third party code}
-
-%In this study we also classified each stack trace according to the type of the
-%root exception been signaled. Table~\ref{typeroottab} presents the types of
-%root causes of all stack traces reported on the Android repositories.
-
-%TO DO: INCLUDE A NEW TABLE WITH NON JAVA.LANG RUNTIME EXCEPTIONS. Point to we cannot trust developers to document.
-
-%todo{discomment after submission} As we can see from Table ~ref{typeroottab},
-%that most of the exceptions signaled by third-party code (e.g., Android and
-%libs) were runtime exceptions. When an exception is signaled by third party
-%code, it can be due to 2 reasons: the precondition associated to third party's
-%code was not obeyed; or there is a coding error on third party's code. Only
-%manual code inspection can reveal the reason. However, regardless the reason it
-%is a good practice that all the exceptions signaled by a third part code are
-%properly documented. Manually inspecting a sample of stack traces caused by
-%runtime exceptions thrown by library code (approximately 10\%) we could observe
-%that none of the runtime exceptions signaled by third part code were documented
-%in Javadoc - this analysis did not consider the exceptions implicitly signaled
-%by JVM (e.g. NullPoiterException). The undocumented runtime exceptions are a
-%serious problem as it makes very difficult or even impossible for the  client
-%method to protect against  undocumented runtime exceptions signaled by library
-%code. As a consequence, the undocumented runtime exception may remain uncaught
-%leading to  system crashes.
-
-%This result is aligned with the results of other study conducted by from Cabral
-%and Marcques ~\cite{} which also observe that runtime exceptions are most often
-%not documented.
-
-%\textbf{When libraries are used, the developer does not have access to their
+\subsection{Undocumente runtime exceptions been thrown.}
+
+Manualy inspecting the code of the methods throwing direct instances of RuntimeException revealed 
+another and even worse bad practice:  none of the analyzed methods 
+documented this exception as part of the exception interface (e.g., using @throws annotation).
+
+Such undocumented runtime exceptions represent a serious threat to system robustness, specially
+when such exceptions are thrown by reused third party code (e.g. libraries, or framework utility code).
+Since it is very difficult or even impossible for the client of such thrid party code to 
+protect the application against undocumented runtime exceptions, the
+ undocumented runtime exception may remain uncaught and lead to  system crashes.
+
+To better investigate the extent of this problem we classified each stack trace according to the type of the
+root exception. Table~\ref{typeroottab} presents the types of root causes of all stack traces reported on 
+the Android repositories.
+
+\begin{table}
+\centering
+\begin{tabular}{lcccccc}
+    \hline
+    \bfseries{Type} & \bfseries{Android} & \bfseries{Libcore} & \bfseries{App} & \bfseries{Lib} & \bfseries{Java} & \bfseries{All}\\
+    \hline
+
+Runtime	&	1075	&	56	&	1446	&	374	&	10	&	2961	\\
+Error	&	144	&	34	&	168	&	105	&	9	&	460	\\
+Checked	&	110	&	230	&	139	&	145	&	12	&	636	\\
+Throwable	&	0	&	0	&	1	&	0	&	0	&	1	\\
+Undefined	&	1	&	0	&	6	&	8	&	0	&	15	\\
+    \hline
+All		& 1330	&	320	&	1760	&	632	&	31	&	4073	\\
+    \hline
+  \end{tabular}
+\caption{Types of Root Exceptions.}
+  \label{fig:typeroottab}
+\end{table}
+
+\begin{figure}
+\centering
+\includegraphics[scale=0.3]{chart_exceptiontypes.pdf}
+\caption{Distribution of Checked and Unchedked exceptions.}
+\label{fig:typeroot}
+\end{figure}
+
+As we can see from Table ~ref{typeroottab},  most of the exceptions signaled by third-party code
+ (e.g., Android and libs) were runtime exceptions.
+
+ Manually inspecting the library code of methods signaling runtime exceptions 
+ - this analysis did not consider the exceptions implicitly signaled  by JVM (e.g. NullPoiterException) -
+ we could also observe that none of the runtime exceptions signaled by them were documented.
+Examples of such methos are listed on the Table X.
+
+This result is aligned with the results of other study conducted by from Cabral
+and Marcques ~\cite{} which ialso observe that runtime exceptions in .Net programs 
+are most often not documented.
+
+On the other hand, we could observe that most of the exceptions signaled by the
+set of libraries reused by Android platform (i.e., org.apache.harmony,
+org.w3c.dom, sun.misc, org.apache.http, org.json, org.xml, javax.) referred in
+this work as libcore, were checked exceptions. It is pointed as a good practice
+for libraries (see Section ~\ref{best}) because by using checked exceptions
+libraries can define a precise exception interface to its clients. Such
+libraries are heavily used in several projects, and such precise exception
+interface may be related to the libraries maturity.
+
+%\textbf{
+
+% the developer does not have access to their
 %source code and thus needs to rely on the library documentation about the
@@
 
+%Manually inspecting the code signalers which RuntimeExceptions were thrown in Android
+%Platform we could observe that direct instances of java.lang.RuntimeException
+% were thrown due to two main reasons: (i) when there was no recovery action associated
+%to the exceptional condition, which means that regardless the kind of exception the
+%only possible action was to abort the application; and when translating exceptions
+%signaled by application methods. Since Android platform is heavily based on
+%call backs", when a application throws a runtime exception (i.e., one of the life cicle 
+%methods of an Activity) and does it, the exception scapes to compotenst of the Android platform 
+%which do not have contextual information to adequately handle
+%the exception and simply terminates the application.
+
+%We could observe that although been known as a bad practice, the signaling general 
+%exceptions in this context it was an intended design decision when no specific recovery 
+%action can be defined for the exception.
+
+%Even an instance of Throwable as signaled in one project. It points to a the
+%"dont worry from know behaviour".
+
+%\subsection{Combining Stack Trace Information with Exception Type Information}
+
+%todo{DISCOMMENT AFTER SUBMISSION}
+%\subsection{Undocumented runtime exceptions signaled by third party code}
+
+%In this study we also classified each stack trace according to the type of the
+%root exception been signaled. Table~\ref{typeroottab} presents the types of
+%root causes of all stack traces reported on the Android repositories.
+
+%TO DO: INCLUDE A NEW TABLE WITH NON JAVA.LANG RUNTIME EXCEPTIONS. Point to we cannot trust developers to document.
+
+%This result is aligned with the results of other study conducted by from Cabral
+%and Marcques ~\cite{} which also observe that runtime exceptions are most often
+%not documented.
+
+%\textbf{When libraries are used, the developer does not have access to their
+%source code and thus needs to rely on the library documentation about the
+%runtime exceptions that should be thrown – which, more often than not, are
+%neither complete nor precise (Thomas, 2002; Sacramento et al., 2006). As a
+%consequence, unhandled unchecked exceptions can be seen as one of the major
+%sources of bugs in current Java systems (Jo et al., 2004). }
+
+%On the other hand, we could observe that most of the exceptions signaled by the
+%set of libraries reused by Android platform (i.e., org.apache.harmony,
+%org.w3c.dom, sun.misc, org.apache.http, org.json, org.xml, javax.) referred in
+%this work as libcore, were checked exceptions. It is pointed as a good practice
+%for libraries (see Section ~\ref{best}) because by using checked exceptions
+%libraries can define a precise exception interface to its clients. Such
+%libraries are heavily used in several projects, and such precise exception
+%interface may be related to the libraries maturity.
+
 \subsection{Uncaught Checked Exceptions}
@@
 causes of all stack traces reported on the Android repositories.
-
-\begin{table}
-\centering
-\begin{tabular}{lcccccc}
-    \hline
-    \bfseries{Type} & \bfseries{Android} & \bfseries{Libcore} & \bfseries{App} & \bfseries{Lib} & \bfseries{Java} & \bfseries{All}\\
-    \hline
-
-Runtime	&	1075	&	56	&	1446	&	374	&	10	&	2961	\\
-Error	&	144	&	34	&	168	&	105	&	9	&	460	\\
-Checked	&	110	&	230	&	139	&	145	&	12	&	636	\\
-Throwable	&	0	&	0	&	1	&	0	&	0	&	1	\\
-Undefined	&	1	&	0	&	6	&	8	&	0	&	15	\\
-    \hline
-All		& 1330	&	320	&	1760	&	632	&	31	&	4073	\\
-    \hline
-  \end{tabular}
-\caption{Types of Root Exceptions.}
-  \label{fig:typeroottab}
-\end{table}
-
-\begin{figure}
-\centering
-\includegraphics[scale=0.3]{chart_exceptiontypes.pdf}
-\caption{Distribution of Checked and Unchedked exceptions.}
-\label{fig:typeroot}
-\end{figure}
 

### 2014-05-07 android-stacks.tex
reducing backgounf section and introduction

@@
 %new version}
+%RSC: I AGREE.
 
@@
 %platform which allowed more fine-grained analysis of their exception behaviour.
+%RSC: I AGREE.
 
@@
 %GGG: Personally, I hate this paragraph in all papers as it usually reads like a space-filler
+%RSC: If needed we can remove it.
 The remainder of this paper is organized as follows. Section 2 presents a
@@
 
-\subsection{Error handling in programming languages}
-
-Techniques for error detection and handling are no more an optional add-on but a
+%\subsection{Error handling in programming languages}
+
+Techniques for error detection and handling are not  an optional add-on but a
 fundamental part of the system~\cite{bruntink2006discovering}. Several
-approaches exists for detecting errors during system execution and responding to
-them accordingly, such as the return-code idiom (very popular in C programs and
-operating systems), and the exception handling mechanism, embedded in many main
+approaches exists for detecting errors and responding to them, such as the return-code 
+idiom (very popular in C programs and operating systems), and the exception handling 
+mechanism, embedded in many main
 stream programming languages.
@@
 In Java, an exception is represented according to the class hierarchy shown in
-Figure~\ref{fig:exchier}.  According to this structure every exception is an
+Figure~\ref{fig:exchier}.  According to it every exception is an
 instance of the Throwable class, and can be of three kinds: checked exceptions
@@
 In Java an exception can be signaled in one of the following
-circumstances~\cite{gosling2000java}:
-
-\begin{itemize} \item  explicitly thrown when a throw statement is executed;
-  \item  implicitly thrown by the JVM when the evaluation of an expression
+circumstances~\cite{gosling2000java}: (i) explicitly thrown when a throw statement is executed; 
+(ii) implicitly thrown by the JVM when the evaluation of an expression
  violates the normal semantics of language (e.g., out-of-bounds array index,
-    division-by-zero, access to a null reference); or \item implicitly thrown by
+division-by-zero, access to a null reference); or (iii) implicitly thrown by
 the JVM due to an internal error or resource limitation (e.g.,
-    OutofMemoryError).  \end{itemize}
+OutofMemoryError).
+
 %\todo{This sub-section is quite similar to the section introduction, shouldn't
 %we merge them?}
+
 When an exception is signaled it causes a transfer of control from the point
 where the exception occurred to a point that can be specified by the programmer
-(exception handler). In Java it is represented by the try-catch block and is
-responsible for performing any recovery action necessary to bring the software
-system back to a normal state and, whenever this is not possible, to log the
-exception and abort the system in an expected safe way.
-
-In Java, as in many languages, the search for the handler to deal with a raised
-exception occurs along the dynamic invocation chain. This is claimed to increase
-the reusability, since the invoker of an operation can handle it in a wider
-context ~\cite{goodenough1975exception}. If no handler is found in the chain,
-the default handler associated to the Thread or ThreadGroup is invoked, which
-usually terminates the system and prints the exception stack trace.
-
-\subsubsection{Exception Chaining} Exception Chaining is a common way of
-propagating exceptions across components of different abstraction levels. It is
-implemented by special exception constructor which receives another exception as
-a parameter. An example a multiple exception chaining is depicted on the stack
-trace in Figure~\ref{fig:wrapping}.
+(exception handler) - in Java it is represented by the try-catch block .
+
+%and is
+%responsible for performing any recovery action necessary to bring the software
+%system back to a normal state and, whenever this is not possible, to log the
+%exception and abort the system in an pre-defined safe way.
+
+%In Java, as in many languages, 
+The search for the handler to deal with a raised
+exception occurs along the dynamic invocation chain. 
+%This is claimed to increase
+%the reusability, since the invoker of an operation can handle it in a wider
+%context ~\cite{goodenough1975exception}. 
+If no handler is found in the chain, the default handler associated to the 
+Thread or ThreadGroup is invoked, which usually terminates the system 
+and prints the exception stack trace.
+
+%\subsubsection{Exception Chaining} Exception Chaining 
 %This stack trace was extracted from an issue of an Android application in
 %Github.
+
+Figure~\ref{fig:wrapping} presents an example of a stack trace, on which multiple 
+exception chaining is depicted. Exception Chaining (or exception wrapping) is  
+common way of propagating exceptions  across components of different 
+abstraction levels. It is implemented by special exception 
+constructor which receives another exception as a parameter.
 
@@
 Such undocumented runtime exceptions represent a serious threat to system robustness, specially
-when such exceptions are thrown by reused third party code (e.g. libraries, or framework utility code).
-Since it is very difficult or even impossible for the client of such thrid party code to 
-protect the application against undocumented runtime exceptions, the
+when such exceptions are thrown then the third party code is invoked inside the application (e.g. libraries, or framework utility code).
+In such cases the client usualy do not have access to the source code, and in the absense of
+the exception documentation it is very difficult or even impossible for the client of such thrid party code to 
+protect the application against undocumented runtime exceptions. As a consequenve, the
  undocumented runtime exception may remain uncaught and lead to system crashes.
@@
 
- Manually inspecting the library code of methods signaling runtime exceptions 
- - this analysis did not consider the exceptions implicitly signaled  by JVM (e.g. NullPoiterException) -
- we could also observe that none of the runtime exceptions signaled by them were documented.
-Examples of such methos are listed on the Table X.
-
-This result is aligned with the results of other study conducted by from Cabral
-and Marcques ~\cite{} which ialso observe that runtime exceptions in .Net programs 
-are most often not documented.
-
-On the other hand, we could observe that most of the exceptions signaled by the
-set of libraries reused by Android platform (i.e., org.apache.harmony,
+If we filter the exceptions implicitly signaled by JVM (due to coding erros) the number of runtime
+exceptions thrown by third party code is XXX signaled by Android and YYY signaled by library code.
+
+Manually inspecting 10% of each set we could also observe that none of the methods throwing 
+runtime exceptions documented the exceptions on the method signature (as suggested by Boch
+ ~\cite{}) nor included a comment in Javadoc. This result is aligned with the results of other 
+study conducted by from Cabral and Marcques ~\cite{} which  observed that 
+runtime exceptions in .Net programs are most often not documented.
+
+On the other hand, we could observe that the checked exceptionsmost represent most of the  
+exceptions signaled by the set of libraries reused by Android platform (i.e., org.apache.harmony,
 org.w3c.dom, sun.misc, org.apache.http, org.json, org.xml, javax.) referred in
-this work as libcore, were checked exceptions. It is pointed as a good practice
+this work as libcore. It is pointed as a good practice
 for libraries (see Section ~\ref{best}) because by using checked exceptions
@@
 
-%\textbf{
-
-% the developer does not have access to their
-%source code and thus needs to rely on the library documentation about the
-%runtime exceptions that should be thrown – which, more often than not, are
-%neither complete nor precise (Thomas, 2002; Sacramento et al., 2006). As a
-%consequence, unhandled unchecked exceptions can be seen as one of the major
-%sources of bugs in current Java systems (Jo et al., 2004). }
-
-%On the other hand, we could observe that most of the exceptions signaled by the
-%set of libraries reused by Android platform (i.e., org.apache.harmony,
-%org.w3c.dom, sun.misc, org.apache.http, org.json, org.xml, javax.) referred in
-%this work as libcore, were checked exceptions. It is pointed as a good practice
-%for libraries (see Section ~\ref{best}) because by using checked exceptions
-%libraries can define a precise exception interface to its clients. Such
-%libraries are heavily used in several projects, and such precise exception
-%interface may be related to the libraries maturity.
-
-%Manually inspecting the code signalers which RuntimeExceptions were thrown in Android
-%Platform we could observe that direct instances of java.lang.RuntimeException
-% were thrown due to two main reasons: (i) when there was no recovery action associated
-%to the exceptional condition, which means that regardless the kind of exception the
-%only possible action was to abort the application; and when translating exceptions
-%signaled by application methods. Since Android platform is heavily based on
-%call backs", when a application throws a runtime exception (i.e., one of the life cicle 
-%methods of an Activity) and does it, the exception scapes to compotenst of the Android platform 
-%which do not have contextual information to adequately handle
-%the exception and simply terminates the application.
-
-%We could observe that although been known as a bad practice, the signaling general 
-%exceptions in this context it was an intended design decision when no specific recovery 
-%action can be defined for the exception.
-
-%Even an instance of Throwable as signaled in one project. It points to a the
-%"dont worry from know behaviour".
-
-%\subsection{Combining Stack Trace Information with Exception Type Information}
-
-%todo{DISCOMMENT AFTER SUBMISSION}
-%\subsection{Undocumented runtime exceptions signaled by third party code}
-
-%In this study we also classified each stack trace according to the type of the
-%root exception been signaled. Table~\ref{typeroottab} presents the types of
-%root causes of all stack traces reported on the Android repositories.
-
-%TO DO: INCLUDE A NEW TABLE WITH NON JAVA.LANG RUNTIME EXCEPTIONS. Point to we cannot trust developers to document.
-
-%This result is aligned with the results of other study conducted by from Cabral
-%and Marcques ~\cite{} which also observe that runtime exceptions are most often
-%not documented.
-
-%\textbf{When libraries are used, the developer does not have access to their
-%source code and thus needs to rely on the library documentation about the
-%runtime exceptions that should be thrown – which, more often than not, are
-%neither complete nor precise (Thomas, 2002; Sacramento et al., 2006). As a
-%consequence, unhandled unchecked exceptions can be seen as one of the major
-%sources of bugs in current Java systems (Jo et al., 2004). }
-
-%On the other hand, we could observe that most of the exceptions signaled by the
-%set of libraries reused by Android platform (i.e., org.apache.harmony,
-%org.w3c.dom, sun.misc, org.apache.http, org.json, org.xml, javax.) referred in
-%this work as libcore, were checked exceptions. It is pointed as a good practice
-%for libraries (see Section ~\ref{best}) because by using checked exceptions
-%libraries can define a precise exception interface to its clients. Such
-%libraries are heavily used in several projects, and such precise exception
-%interface may be related to the libraries maturity.
-
 \subsection{Uncaught Checked Exceptions}
 
-In this study we also classified each stack trace according to the type of the
-root exception been signaled. Table~\ref{typeroottab} presents the types of root
-causes of all stack traces reported on the Android repositories.
-
-Although most of the reported were runtime, we could also find checked
-exceptions as the root causes of stack traces reported on issues. An exception
+Although most of the  reported exceptions were runtime, we could also find checked
+exceptions as the root causes of stack traces reported on issues. A checked exception
 can remain uncaught in two circumstances: if all methods on the execution trace
 on which this exception flows explicitly specifies the exception, or if the
-checked exception is wrapped in a runtime and than re-thrown.Since checked
-exceptions should be used to represent recoverable conditions, ignoring a
-checked exception is considered a bad practice. When the developer is confronted
+checked exception is wrapped in a runtime and than re-thrown.
+
+Since checked exceptions should be used to represent recoverable conditions, ignoring a
+checked exception, is considered a bad practice. As mentioned before, when the developer is confronted
 with a checked exception, the designer of the API is telling him to handle the
@@
 of them were not wrapped along the stack trace (represented as the table lines
-on which the root cause is none). This analysis also revealed unexpected
-wrappings such as: Error wrapping a Checked exception; a Runtime wrapping and
-Error; and an Error wrapping a Runtime.
+on which the Wapper type is -).
 
@@
 
+This analysis also revealed unexpected wrappings such as: Error wrapping a Checked 
+exception; a Runtime wrapping and Error; and an Error wrapping a Runtime.
 Java is the only language that provides a hybrid exception handling mechanism
-which offers different types to represent different exception behaviors (see
-Section ~\ref{extypes}) . According to Java specification Errors should not be
+which offers  different types to represent different exception behaviors (i.e., error, 
+runtime and checked) (see Section ~\ref{extypes}) . According to Java specification Errors should not be
 handled inside the system since they usually represent unrecoverable conditions

### 2014-05-07 android-stacks.tex
mprovements on section III.

@@
 
-%\subsection{Error handling in programming languages}
+\subsection{Error handling in programming languages}
 
@@
 
-\subsection{Exception handling in Java} \label{sec:extypes}
+\subsection{Exception Handling in Java} \label{sec:extypes}
 
@@
 (exception handler) - in Java it is represented by the try-catch block .
-
-%and is
-%responsible for performing any recovery action necessary to bring the software
-%system back to a normal state and, whenever this is not possible, to log the
-%exception and abort the system in an pre-defined safe way.
-
-%In Java, as in many languages, 
 The search for the handler to deal with a raised
 exception occurs along the dynamic invocation chain. 
-%This is claimed to increase
-%the reusability, since the invoker of an operation can handle it in a wider
-%context ~\cite{goodenough1975exception}. 
 If no handler is found in the chain, the default handler associated to the 
 Thread or ThreadGroup is invoked, which usually terminates the system 
-and prints the exception stack trace.
-
-%\subsubsection{Exception Chaining} Exception Chaining 
-%This stack trace was extracted from an issue of an Android application in
-%Github.
-
-Figure~\ref{fig:wrapping} presents an example of a stack trace, on which multiple 
+and prints the exception stack trace. Figure~\ref{fig:wrapping} presents an 
+example of a stack trace, on which multiple 
 exception chaining is depicted. Exception Chaining (or exception wrapping) is  
@@
 constructor which receives another exception as a parameter.
+
+%\subsubsection{Exception Chaining} Exception Chaining 
+%This stack trace was extracted from an issue of an Android application in
+%Github.
 
@@
 %\todo{This section feels a bit disconnected.}
-Several general guidelines have been proposed on how to use
+Several general guidelines have been proposed on how to use Java
 exceptions~\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,
-bloch2008effective}. To guide our investigation, we compile general guidelines
-related to exception types and the propagation of exception in Java programs
+bloch2008effective}. 
+% To guide our investigation, we compile the guidelines
+A compiled list of the guidelines related to exception types and exception propagation in Java the following:
 \footnote{We also compiled guidelines related to exception handling but they are
 out of the scope of this paper, since they do not relate to information
-available on stack traces.}.
-
-\noindent\emph{Meaning of Exception Types}
-
-\textbf{1) Use checked exceptions to represent recoverable
+available on stack traces.}:
+
+%\noindent\emph{Meaning of Exception Types}
+
+\textbf{I-Use checked exceptions to represent recoverable
 conditions}(~\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,bloch2008effective}).
@@
 
-\textbf{2) Error represents an unrecoverable condition detected by the JVM which
+\textbf{II-Error represents an unrecoverable condition detected by the JVM which
 should not he handled}~\cite{gosling2000java}. It results from failures detected
@@
 
-\noindent\emph{Exception Throwing}
-
-\textbf{3) A method should throw exceptions that precisely define the
+%\noindent\emph{Exception Throwing}
+
+\textbf{III-A method should throw exceptions that precisely define the
 exceptional condition}~\cite{gosling2000java,bloch2008effective}. To do so,
@@
 
-\textbf{4) Do not repeatedly re-throw exceptions, handle exceptions as close as
-possible to the problem}~\cite{wirfs2006toward}. Programs that frequently throw
-exceptions lose in performance ~\cite{wirfs2006toward,gosling2000java}.
-Moreover, close to the place that the exception was throw the caller usually has
+\textbf{IV-Do not repeatedly re-throw exceptions, handle exceptions as close as
+possible to the problem}~\cite{wirfs2006toward}. 
+%Programs that frequently throw exceptions lose in performance ~\cite{wirfs2006toward,gosling2000java}.
+%Moreover, 
+Close to the place that the exception was throw the caller usually has
 enough contextual information to perform  corrective actions. When the exception
@@
 from the problem) it can be difficult to take meaningful recovery actions.
-
-\noindent\emph{Exception Documentation}
-
-\textbf{5) Document all exceptions thrown by a
+todo{RSC:this practice was not well explored, can be removed if we need space.}
+
+%\noindent\emph{Exception Documentation}
+\textbf{V-Document all exceptions thrown by a
 method}~\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,bloch2008effective}).
@@
 
-These best practices guided our exploratory study described in the next
-sections. Based on information extracted from stack traces (complemented by
+These best practices guided our exploratory study described next
+. Based on information extracted from stack traces (complemented by
 bytecode and source-code analysis), we investigate whether stack characteristics
-can reveal whether these practices have been obeyed. Next section discusses the
-study procedure and the practices' misuses that emerged from this study.
+can reveal whether these practices have been obeyed. 
+%Next section discusses the
+%study procedure and the practices' misuses that emerged from this study.
 
@@
 represents the proportion of a population found to have a particular condition
-(e.g. a disease) over the total number of people studied. In our work, we used
+(e.g. a disease) over the total number of the studied population. In our work, we used
 this term to evaluate a condition (i.e., presence of a stack trace) over issues
@@
 
-To answer this question, we mined from every extracted stack trace the top level
-exception being signaled, its signaler (the first method signature frame after
-the exception), the root exception (the exception which initiated the stack
-trace, defined on the bottom of the stack), and the root signaler. The
-intermediate elements that compose a stack trace as shown in
-Figure~\ref{fig:wrapping} (i.e., the intermediate causes and its execution
-stacks) were not be considered in the analysis.
+%To answer this question, we mined from every extracted stack trace the top level
+%exception being signaled, its signaler (the first method signature frame after
+%the exception), the root exception (the exception which initiated the stack
+%trace, defined on the bottom of the stack), and the root signaler. The
+%intermediate elements that compose a stack trace as shown in
+%Figure~\ref{fig:wrapping} (i.e., the intermediate causes and its execution
+%stacks) were not be considered in the analysis.
 %~\todo{Move this to the previous section? IMO, It is better to keep all facts
-%about how we did the research in one place}
+%about how we did the research in one place
+%RSC: I WAS ABOUT TO MOOVE BUT THERE WAS ALREADY A SIMILAR TEXT ON 
+%THE EXCEPTION DISTEILLER}
 
@@
 
-\subsection{Sizes of Stack traces}
-
-Concerning the size of the stacks (i.e. the number of frames composing them), we
-could observe that  sometimes it exceeded 100 frames (158 stack frames).
-Manually inspecting the corresponding execution trace was caused by one of the
-two reasons: recursive calls or exceptions signaled by a method defined deep
-down in a reused framework; exceptions successively wrapped. Irrespective of the
-reason, in exception flows involving many methods, it is practically impossible
-to handle an exception in a way other than exiting the application, as at the
-place the exception occurs there is hardly enough contextual information to
-perform recovery actions.
-
-\begin{figure}
-  \begin{center}
-    \includegraphics[scale=0.5]{stack-size-wrappings-hist.pdf}
-  \end{center}
-  \caption{Stack sizes (top) and wrappings (bottom) for Java (left) and Android (right) stack traces}
-  \label{fig:sizewrapbhists}
-\end{figure}
+%\subsection{Sizes of Stack traces}
+%Concerning the size of the stacks (i.e. the number of frames composing them), we
+%could observe that  sometimes it exceeded 100 frames (158 stack frames).
+%Manually inspecting the corresponding execution trace was caused by one of the
+%two reasons: recursive calls or exceptions signaled by a method defined deep
+%down in a reused framework; exceptions successively wrapped. Irrespective of the
+%reason, in exception flows involving many methods, it is practically impossible
+%to handle an exception in a way other than exiting the application, as at the
+%place the exception occurs there is hardly enough contextual information to
+%perform recovery actions.
 
@@
 root exception. Table~\ref{typeroottab} presents the types of root causes of all stack traces reported on 
-the Android repositories.
+the Android repositories. As we can see from Table ~ref{typeroottab},  most of the exceptions signaled by third-party code
+ (e.g., Android and libs) were runtime exceptions. If we filter the exceptions implicitly signaled by 
+JVM (due to coding erros) the number of runtime exceptions thrown by third party code is XXX 
+signaled by Android and YYY signaled by library code.
 
@@
 \end{figure}
-
-As we can see from Table ~ref{typeroottab},  most of the exceptions signaled by third-party code
- (e.g., Android and libs) were runtime exceptions.
-
-If we filter the exceptions implicitly signaled by JVM (due to coding erros) the number of runtime
-exceptions thrown by third party code is XXX signaled by Android and YYY signaled by library code.
 
@@
 %TODO INCLUIR ESTA TABLEA
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[scale=0.5]{only-wrap-histogram.pdf}
+  \end{center}
+  \caption{Stack sizes (top) and wrappings (bottom) for Java (left) and Android (right) stack traces}
+  \label{fig:sizewrapbhists}
+\end{figure}
 

### 2014-05-07 android-stacks.tex
smal inprovements.

@@
 \end{figure}
+
+We could also observe that in 205 stacks found in Android repository the exception was wrapped 
+more then once. Figure ~\ref{sizewrapbhists}  illustrates the number of wrappings per stack, both
+ in Java and Android studied repositories. One interesting wrapping found in ones Android repo was the 
+following: Runtime-Checked-Runtime-Checked-Runtime-Checked-Runtime.
 

### 2014-05-08 android-stacks.tex
changing and commenting pictures to gain space.

@@
     \hline
-    \bfseries{android} & When the exception is explicitly thrown by a method defined in Android Platform, in a the JDK library used by it.\\
-    \bfseries{app}     & If the exception signaler was explicitly thrown by an application method or in a the JDK library used by it.\\
-    \bfseries{libcore} & If the application was signaled by one of the libraries reused by Android Platform (i.e., org.apache.harmony, org.w3c.dom, sun.misc, org.apache.http, org.json, org.xml) besides the JDK.
-\\
-    \bfseries{lib}     & If the exceptions was not defined by any of the elements above.\\
-    \bfseries{java}    & If the application signaler and all the methods that composes the exception are defined in a the JDK library method.\\
+    \bfseries{android} & If the exception is thrown by a method defined in Android Platform or OS, or in a JDK library used by them.\\
+    \bfseries{app}     & If the exception is thrown by an application method or in a  DK library used by it.\\
+    \bfseries{libcore} & If the exception is thrown by one of the core libraries reused by Android (i.e., org.apache.harmony, org.w3c.dom, sun.misc, org.apache.http, org.json, org.xml). \\
+    \bfseries{lib}     & If the exception is thrown by a method that was not defined by any of the elements above.\\
+    \bfseries{java}    & If all methods on the stack trace are JDK library methods.\\
     \hline
@@
 
-\begin{figure}
-\centering
-\includegraphics[width=\hsize]{top_exceptios_android.pdf}
-\caption{Top exceptions in Android repositories.}
-\label{fig:androidsignaler}
-\end{figure}
+%\begin{figure}
+%\centering
+%\includegraphics[width=\hsize]{top_exceptios_android_new2.pdf}
+%\includegraphics[width=\hsize]{top_exceptions_android_new}
+%\caption{Top exceptions in Android repositories.}
+%\label{fig:androidsignaler}
+%\end{figure}
 
@@
 \centering
-\includegraphics[scale=0.3]{chart_exceptiontypes.pdf}
+%\includegraphics[width=\hsize]{chart_exceptiontypes.pdf}
+\includegraphics[width=\hsize]{exception_types.pdf}
 \caption{Distribution of Checked and Unchedked exceptions.}

### 2014-05-08 android-stacks.tex
small changes on identifications.

@@
 Georgios Gousios\IEEEauthorrefmark{2}, Arie van Deursen\IEEEauthorrefmark{2},
-Lucas Mariano\IEEEauthorrefmark{1}}
+Lucas Almeida\IEEEauthorrefmark{1}}
 \IEEEauthorblockA{\IEEEauthorrefmark{1}Federal University of Rio Grande do Norte\\
 Natal, Brazil\\
-Email: \texttt{\{roberta, lucasm\}@dimap.ufrn.br}}
+Email: \texttt{roberta@dimap.ufrn.br,lucas.almeida@lcc.ufrn.br}}
 \IEEEauthorblockA{\IEEEauthorrefmark{2}Delft University of Technology\\

### 2014-05-08 android-stacks.tex
Unifying the terminology.

@@
 
-Moreover, some languages use exceptions to represent coding errors
+Moreover, some languages use exceptions to represent programming mistakes
 (out-of-bounds array index, division-by-zero, access to a null). In these cases
@@
 
-  \item  A multitude undocumented runtime exceptions signaled by third party
+  \item  A multitude of programming mistakes as the main causes of stack traces.
+
+  \item Undocumented runtime exceptions signaled by third party
     code (i.e., libraries and Android platform).
@@
 
-\subsection{Error handling in programming languages}
+\subsection{Error Handling in Programming Languages}
 
@@
 idiom (very popular in C programs and operating systems), and the exception handling 
-mechanism, embedded in many main
-stream programming languages.
+mechanism, embedded in many main stream programming languages.
 
@@
 
-\subsubsection{Exception Throwing and Handling}
+%\subsubsection{Exception Throwing and Handling}
 
@@
 
-\subsection{Exception Handling Best Practices}
+\subsection{Exception Handling Best Practices in Java}
 \label{sec:best}
@@
 % To guide our investigation, we compile the guidelines
-A compiled list of the guidelines related to exception types and exception propagation in Java the following:
+A compiled list of the guidelines related to exception types and exception propagation in Java the following
 \footnote{We also compiled guidelines related to exception handling but they are
-out of the scope of this paper, since they do not relate to information
-available on stack traces.}:
+out of the scope of this paper.}:
 
@@
 
-\textbf{IV-Do not repeatedly re-throw exceptions, handle exceptions as close as
-possible to the problem}~\cite{wirfs2006toward}. 
+%\textbf{IV-Do not repeatedly re-throw exceptions, handle exceptions as close as
+%possible to the problem}~\cite{wirfs2006toward}. 
 %Programs that frequently throw exceptions lose in performance ~\cite{wirfs2006toward,gosling2000java}.
 %Moreover, 
-Close to the place that the exception was throw the caller usually has
-enough contextual information to perform  corrective actions. When the exception
-is propagated far away from the method where it was signaled (i.e. further away
-from the problem) it can be difficult to take meaningful recovery actions.
-todo{RSC:this practice was not well explored, can be removed if we need space.}
+%Close to the place that the exception was throw the caller usually has
+%enough contextual information to perform  corrective actions. When the exception
+%is propagated far away from the method where it was signaled (i.e. further away
+%#from the problem) it can be difficult to take meaningful recovery actions.
+%*~\todo{RSC:this practice was not well explored, can be removed if we need space.}
 
 %\noindent\emph{Exception Documentation}
-\textbf{V-Document all exceptions thrown by a
+\textbf{IV-Document all exceptions thrown by a
 method}~\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,bloch2008effective}).
@@
 also wise to document the explicitly signaled runtime exceptions\footnote{This
-excludes the implicitly signaled JVM runtime exceptions related to coding
-errors} as carefully as checked exceptions. Doing so, the clients of a method
+excludes the implicitly signaled JVM runtime exceptions related to programming
+mistakes} as carefully as checked exceptions. Doing so, the clients of a method
 will be aware of the exceptions the method can throw. If the developer fails to
@@
 
-Coding errors are the most common root causes for most types of signalers. We
+Programming mistakes are the most common root causes for most types of signalers. We
 could observe that the NullPointerException is still the exception with higher
@@
  (e.g., Android and libs) were runtime exceptions. If we filter the exceptions implicitly signaled by 
-JVM (due to coding erros) the number of runtime exceptions thrown by third party code is XXX 
+JVM (due to programming mistakes) the number of runtime exceptions thrown by third party code is XXX 
 signaled by Android and YYY signaled by library code.
@@
 Manually inspecting 10% of each set we could also observe that none of the methods throwing 
-runtime exceptions documented the exceptions on the method signature (as suggested by Boch
- ~\cite{}) nor included a comment in Javadoc. This result is aligned with the results of other 
-study conducted by from Cabral and Marcques ~\cite{} which  observed that 
+runtime exceptions documented the exceptions on the method signature (as suggested by Bloch
+ ~\cite{bloch2008effective}) nor included a comment in Javadoc. This result is aligned with the results of other 
+study conducted by from Sacramento et al ~\cite{sacramento2006unchecked} which observed that the
 runtime exceptions in .Net programs are most often not documented.
@@
 represent recoverable conditions and  Runtime exceptions represent are usually
-related to coding errors from which the developer is not expected to recover
+related to programming mistakes from which the developer is not expected to recover
 from.
@@
   \end{center}
-  \caption{Stack sizes (top) and wrappings (bottom) for Java (left) and Android (right) stack traces}
+  \caption{Number of  stack traces wrappings for Java and Android repositories. }
   \label{fig:sizewrapbhists}
@@
 
+IMPORTANT:  ~\todo{We need to update this data because it is still including PulWifi. Gerogios I sent the filtered data to you by email, and saved on file java-new-stack-sizes.csv in this repository.}
+
 We could also observe that in 205 stacks found in Android repository the exception was wrapped 
 more then once. Figure ~\ref{sizewrapbhists}  illustrates the number of wrappings per stack, both
- in Java and Android studied repositories. One interesting wrapping found in ones Android repo was the 
+ in Java and Android studied repositories. One interesting multiple-wrapping found in Android repo was the 
 following: Runtime-Checked-Runtime-Checked-Runtime-Checked-Runtime.
 
-Since there is no way of enforcing such conventions during program development,
+Since there is no way of enforcing Java exception type conventions during program development,
 we could observe the stack trace analysis combined with types these three types
@@
 handled.
+
+%\section{Discussions}
 

### 2014-05-09 scripts-data/stack-sizes-plots.R
Stack sizes hists without pulwifi

@@ -2,7 +2,7 @@ library(ggplot2)
 library(reshape)
 library(plyr)
 
-data <- read.csv('stack-sizes.csv', sep=';')
+data <- read.csv('stack-sizes.csv')
 
 multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
   require(grid)

### 2014-05-09 android-stacks.tex
Minor header fixes
Stack wrappings only histogram

@@
 
-%commented only for submission:
 We could observe in this study that the NullPointerException and other
@@
 
-\subsection{Direct instances of RuntimeExceptions been thrown.}
+\subsection{Direct instances of RuntimeExceptions being thrown.}
 
@@
 
-\subsection{Undocumente runtime exceptions been thrown.}
+\subsection{Undocumented runtime exceptions being thrown.}
 
@@
   \begin{center}
-    \includegraphics[scale=0.5]{only-wrap-histogram.pdf}
+    \includegraphics[scale=0.5]{stack-wrappings-hist.pdf}
   \end{center}
@@
 \end{figure}
-
-IMPORTANT:  ~\todo{We need to update this data because it is still including PulWifi. Gerogios I sent the filtered data to you by email, and saved on file java-new-stack-sizes.csv in this repository.}
 

### 2014-05-09 scripts-data/stack-sizes-plots.R

@@ -53,6 +53,7 @@ p3 <- ggplot(subset(data, type == 'java')) +
       aes(x = wrappings) + 
       geom_histogram() +
       ggtitle("Java")
+
 p4 <- ggplot(subset(data, type == 'android')) + 
       aes(x = wrappings) + 
       geom_histogram() +
@@ -61,3 +62,8 @@ p4 <- ggplot(subset(data, type == 'android')) +
 pdf("stack-size-wrappings-hist.pdf")
 multiplot(p1, p3, p2, p4, cols = 2)
 dev.off()
+
+pdf("stack-wrappings-hist.pdf")
+multiplot(p3, p4, cols = 1)
+dev.off()
+

### 2014-05-09 android-stacks.tex
Shorten references

@@
 (e.g. Java, C++, C\#) and the target of several studies
-(e.g.~\cite{miller1997issues,robillard2000designing,shah2010understanding,
+(e.g.~\cite{miller1997issues,Robil00,shah2010understanding,
 garcia2007extracting,garcia2001comparative,cabral2007exception,coelho2011unveiling}),

### 2014-05-09 android-stacks.tex
new statistics on undocumented runtime exceptions

--- 2014-05-09 android-stacks.tex
@@
 The remainder of this paper is organized as follows. Section 2 presents a
-background on error handling mechanisms and Java exception handling constructs
-and practices. Section 3 defines research questions, the configuration of our
-exploratory study, the target applications and the evaluation procedures.
-Section 4 reports our analysis of the empirical data collected relating to the
-first research questions. Section X is dedicated to provide insights for the 3rd
-research question. Section 6 provides further discussions and lessons learned
-and discusses the threats to validity associated to this study. Finally Section
+background on exception handling mechanism. Section 3 presents the study design.
+Section 4 reports study findings. Section 5 provides further discussions and insights.
+Section 6 presents the threats to validity associated to this study. Finally Section
 7 describes the related work, and Section 8 presents our conclusions and
@@
 
-In Java an exception can be signaled in one of the following
+In Java an exception can be thrown in one of the following
 circumstances~\cite{gosling2000java}: (i) explicitly thrown when a throw statement is executed; 
@@
 
-When an exception is signaled it causes a transfer of control from the point
-where the exception occurred to a point that can be specified by the programmer
-(exception handler) - in Java it is represented by the try-catch block .
-The search for the handler to deal with a raised
-exception occurs along the dynamic invocation chain. 
-If no handler is found in the chain, the default handler associated to the 
-Thread or ThreadGroup is invoked, which usually terminates the system 
-and prints the exception stack trace. Figure~\ref{fig:wrapping} presents an 
-example of a stack trace, on which multiple 
-exception chaining is depicted. Exception Chaining (or exception wrapping) is  
-common way of propagating exceptions  across components of different 
-abstraction levels. It is implemented by special exception 
-constructor which receives another exception as a parameter.
+%When an exception is signaled it causes a transfer of control from the point
+%where the exception occurred to a point that can be specified by the programmer
+%(exception handler) - in Java it is represented by the try-catch block .
+%The search for the handler to deal with a raised
+%exception occurs along the dynamic invocation chain. 
+%If no handler is found in the chain, the default handler associated to the 
+%Thread or ThreadGroup is invoked, which usually terminates the system 
+%and prints the exception stack trace.
 
@@
 
-The bottom part of the stack trace is the \emph{root cause}, which indicates the
+Exception chaining is a common way of propagating exceptions in Java programs, 
+which allows one exception to be wrapped in another exception and re-thrown. 
+Figure~\ref{fig:wrapping} presents a stack trace associated which illustrates an
+exception chaining. The bottom part of the stack trace is the \emph{root cause}, which indicates the
 first reason for the error to be thrown (in this case, the computer run out of
@@
   \begin{center}
-    \includegraphics[scale=0.5]{stack-wrappings-hist.pdf}
+    \includegraphics[scale=0.4]{stack-wrappings-hist.pdf}
   \end{center}

### 2014-05-09 stacktrace-stats.tex
Save some space in the descriptive stats table

--- 2014-05-06 stacktrace-stats.tex
@@
   \hline
-all & All issues recorded for the project. & 1.00 & 23.37 & 3.00 & 80.00 & \includegraphics[scale = 0.1, clip = true, trim= 50px 60px 50px 60px]{hist-31b77719e3773ad7ac2073785de62a5f.pdf} \\ 
+   \multicolumn{2}{l}{\bfseries{Java}}  \\
+all & All issues recorded for the project. & 1.00 & 23.37 & 3.00 & 80.00 & \includegraphics[scale = 0.08, clip = true, trim= 50px 70px 50px 60px]{hist-31b77719e3773ad7ac2073785de62a5f.pdf} \\ 
   with\_exception & All issues featuring an exception.
-& 0.00 & 3.21 & 0.00 & 6.00 & \includegraphics[scale = 0.1, clip = true, trim= 50px 60px 50px 60px]{hist-0c1305fc40f695dbdd1f4696db24f147.pdf} \\
-  with\_stack & All issues featuring a stack trace. & 0.00 & 2.67 & 0.00 & 4.00 & \includegraphics[scale = 0.1, clip = true, trim= 50px 60px 50px 60px]{hist-896524802e915aa3f01ef1ecd2489e7e.pdf} \\ 
-  defect\_issues & All issues labeled as defect. & 0.00 & 5.51 & 0.00 & 12.00 & \includegraphics[scale = 0.1, clip = true, trim= 50px 60px 50px 60px]{hist-5d0ab95cd3e300c21b0a40d6bec58e4c.pdf} \\ 
+& 0.00 & 3.21 & 0.00 & 6.00 & \includegraphics[scale = 0.08, clip = true, trim= 50px 70px 50px 60px]{hist-0c1305fc40f695dbdd1f4696db24f147.pdf} \\
+  with\_stack & All issues featuring a stack trace. & 0.00 & 2.67 & 0.00 & 4.00 & \includegraphics[scale = 0.08, clip = true, trim= 50px 70px 50px 60px]{hist-896524802e915aa3f01ef1ecd2489e7e.pdf} \\ 
+  defect\_issues & All issues labeled as defect. & 0.00 & 5.51 & 0.00 & 12.00 & \includegraphics[scale = 0.08, clip = true, trim= 50px 70px 50px 60px]{hist-5d0ab95cd3e300c21b0a40d6bec58e4c.pdf} \\ 
   defect\_with\_exception & All issues labeled labeled as defect and including an exception.
-& 0.00 & 1.82 & 0.00 & 1.00 & \includegraphics[scale = 0.1, clip = true, trim= 50px 60px 50px 60px]{hist-502028d1a7e2e77414e2f9f63e0bcd50.pdf} \\
-  defect\_with\_stack & All issues labeled labeled as defect and including a stacktrace & 0.00 & 1.73 & 0.00 & 1.00 & \includegraphics[scale = 0.1, clip = true, trim= 50px 60px 50px 60px]{hist-dab0764147debd99066c84814d46d50e.pdf} \\
+& 0.00 & 1.82 & 0.00 & 1.00 & \includegraphics[scale = 0.08, clip = true, trim= 50px 70px 50px 60px]{hist-502028d1a7e2e77414e2f9f63e0bcd50.pdf} \\
+  defect\_with\_stack & All issues labeled labeled as defect and including a stacktrace & 0.00 & 1.73 & 0.00 & 1.00 & \includegraphics[scale = 0.08, clip = true, trim= 50px 70px 50px 60px]{hist-dab0764147debd99066c84814d46d50e.pdf} \\
+   \hline
+   \multicolumn{2}{l}{\bfseries{Android}}\\
+   all & All issues recorded for the project. & 2.00 & 67.70 & 20.00 & 248.00 & \includegraphics[scale = 0.08, clip = true, trim= 50px 70px 50px 60px]{hist-5d4a87daa44c63097df9bb1f3506d541.pdf} \\ 
+  with\_exception & All issues featuring an exception.
+& 1.00 & 8.16 & 2.00 & 34.00 & \includegraphics[scale = 0.08, clip = true, trim= 50px 70px 50px 60px]{hist-bf73235d0503f937ef6f36c28cf1ba0d.pdf} \\
+  with\_stack & All issues featuring a stack trace. & 1.00 & 6.44 & 2.00 & 27.00 & \includegraphics[scale = 0.08, clip = true, trim= 50px 70px 50px 60px]{hist-528cb6be67074e6e2c22d11245920e13.pdf} \\ 
+  defect\_issues & All issues labeled as defect. & 0.00 & 13.28 & 1.00 & 52.00 & \includegraphics[scale = 0.08, clip = true, trim= 50px 70px 50px 60px]{hist-eb1797e86ef23b2a07907e5ccbeff7b9.pdf} \\ 
+  defect\_with\_exception & All issues labeled labeled as defect and including an exception.
+& 0.00 & 3.25 & 0.00 & 13.00 & \includegraphics[scale = 0.08, clip = true, trim= 50px 70px 50px 60px]{hist-f1a879616aa3e0995db35c6c02bc8520.pdf} \\
+  defect\_with\_stack & All issues labeled labeled as defect and including a stacktrace & 0.00 & 2.92 & 0.00 & 10.00 & \includegraphics[scale = 0.08, clip = true, trim= 50px 70px 50px 60px]{hist-49dbfa02c366cab8b489f33014de587d.pdf} \\
  \hline

### 2014-05-09 android-stacks.tex
removing mistakes on paragraph wrapping

@@
 exceptions where indeed wrapped in runtime exceptions or even errors, while most
-of them were not wrapped along the stack trace (represented as the table lines
-on which the Wapper type is -).
+of them were not wrapped along the stack trace.
 

### 2014-05-09 android-stacks.tex
new content to undocumented section

@@
 
-Manualy inspecting the code of the methods throwing direct instances of RuntimeException revealed 
-another and even worse bad practice:  none of the analyzed methods 
-documented this exception as part of the exception interface (e.g., using @throws annotation).
+To better investigate the extent of this problem we classified each stack trace according to the type of the
+root exception. Table~\ref{typeroottab} presents the types of root causes of all stack traces reported on 
+the Android repositories. As we can see from Table ~ref{typeroottab},  most of the exceptions signaled by third-party code
+ (e.g., Android and libs) were runtime exceptions. After filtering all the exceptions implicitly signaled by 
+JVM (due to programming mistakes) and inspecting the signaler methods of such exceptions (i.e.,  direct an inderect 
+instances of java.lang.RuntimeException). We could observe observe that only 1 method out of 118 inspected signalers
+ (i.e., 0,8%) documented the explicitly thrown runtime exceptions in a Javadoc comment, and none of the analized 
+included the exception  as part of the exception interface of the method (i.e., using throws clause on method signature). 
+This result is aligned with the results of other  study conducted by from 
+Sacramento et al ~\cite{sacramento2006unchecked} which observed that the
+runtime exceptions in .Net programs are most often not documented.
 
@@
 the exception documentation it is very difficult or even impossible for the client of such thrid party code to 
-protect the application against undocumented runtime exceptions. As a consequenve, the
+protect the application against undocumented runtime exceptions. As a consequence, the
  undocumented runtime exception may remain uncaught and lead to system crashes.
-
-To better investigate the extent of this problem we classified each stack trace according to the type of the
-root exception. Table~\ref{typeroottab} presents the types of root causes of all stack traces reported on 
-the Android repositories. As we can see from Table ~ref{typeroottab},  most of the exceptions signaled by third-party code
- (e.g., Android and libs) were runtime exceptions. If we filter the exceptions implicitly signaled by 
-JVM (due to programming mistakes) the number of runtime exceptions thrown by third party code is XXX 
-signaled by Android and YYY signaled by library code.
 
@@
 
-Manually inspecting 10% of each set we could also observe that none of the methods throwing 
-runtime exceptions documented the exceptions on the method signature (as suggested by Bloch
- ~\cite{bloch2008effective}) nor included a comment in Javadoc. This result is aligned with the results of other 
-study conducted by from Sacramento et al ~\cite{sacramento2006unchecked} which observed that the
-runtime exceptions in .Net programs are most often not documented.
-
-On the other hand, we could observe that the checked exceptionsmost represent most of the  
-exceptions signaled by the set of libraries reused by Android platform (i.e., org.apache.harmony,
-org.w3c.dom, sun.misc, org.apache.http, org.json, org.xml, javax.) referred in
-this work as libcore. It is pointed as a good practice
+On the other hand, we could observe that most of the exceptions signaled by the set of 
+libraries reused by Android platform (i.e., org.apache.harmony,
+org.w3c.dom, sun.misc, org.apache.http, org.json, org.xml, javax - referred in
+this work as libcore) were checked exceptions. It is pointed as a good practice
 for libraries (see Section ~\ref{best}) because by using checked exceptions

### 2014-05-09 android-stacks.tex
more content to undocumented section

@@
 instances of java.lang.RuntimeException). We could observe observe that only 1 method out of 118 inspected signalers
- (i.e., 0,8%) documented the explicitly thrown runtime exceptions in a Javadoc comment, and none of the analized 
-included the exception  as part of the exception interface of the method (i.e., using throws clause on method signature). 
-This result is aligned with the results of other  study conducted by from 
+ (i.e., 0,8%) documented the explicitly thrown runtime exception in a Javadoc comment; and none of these methods
+included the runtime exception  (reported on the stack) as part of the exception interface of the method (i.e., using 
+throws clause on method signature). This result is aligned with the results of other  study conducted by from 
 Sacramento et al ~\cite{sacramento2006unchecked} which observed that the

### 2014-05-09 android-stacks.tex
Remove hassanboxes for the moment

@@
 
-\noindent \fbox{
-\begin{minipage}{0.96\columnwidth} \emph{ RQ1: 28,37\% of Java
-  repositories on Github have at least 1 issue that includes a stack trace.
-  9,4\% of the defect issues of such repositories mentioned at least one stack
-  trace. Considering only defect-issues of Android subset 58,7\% mentioned at
-  least 1 stack trace.  }
-\end{minipage}}
+%\noindent \fbox{
+%\begin{minipage}{0.96\columnwidth} \emph{ RQ1: 28,37\% of Java
+%  repositories on Github have at least 1 issue that includes a stack trace.
+%  9,4\% of the defect issues of such repositories mentioned at least one stack
+%  trace. Considering only defect-issues of Android subset 58,7\% mentioned at
+%  least 1 stack trace.  }
+%\end{minipage}}
 
@@
   \centering
-  \begin{tabular}{rp{22em}}
-    \hline
-    \bfseries{Type of signaler} & \bfseries{Description} \\
+  \begin{tabular}{rp{29em}}
+    \hline
+    \bfseries{Signaler} & \bfseries{Description} \\
     \hline
@@
 
-\noindent \fbox{
-\begin{minipage}{0.96\columnwidth} 
-\emph{RQ2: Most exceptions are due to programming mistakes
-or generally irrecoverable errors.}
-\end{minipage}}
+%\noindent \fbox{
+%\begin{minipage}{0.96\columnwidth} 
+%\emph{RQ2: Most exceptions are due to programming mistakes
+%or generally irrecoverable errors.}
+%\end{minipage}}
 

### 2014-05-09 android-stacks.tex
discussion - structue

--- 2014-05-09 android-stacks.tex
@@
 
-%\section{Discussions}
+\section{Discussions}
+
+The exception (mis-)use that emerged from this study shades light on two
+
+It is wise to document the explicitly signaled runtime exceptions as carefully as checked exceptions. Doing so, the clients of a method will be aware of the exceptions the method can throw.
+
+When developer fails to document the runtime exceptions thrown by a method (s)he will make difficult or even impossible for the caller to effective use of such method.
+
+Specially when developing library code, developers must carefully document the runtime exceptions, since there is no compilation checking
+
+have not been fully adopted, mainly there is no compilation or runtime tool to enforce such policies.
+noindent\emph{Undocumented runtime exceptions}
 

### 2014-05-09 android-stacks.tex
Consistency fixes

@@
 (mis)-use were consistently detected such as: unexpected wrappings (e.g., Errors
-being wrapped in checked exceptions) - revealing that Java hybrid exception
+being wrapped in checked exceptions), revealing that Java hybrid exception
 model is not fully used according to its purpose;  undocumented runtime
-exceptions signaled by third party code - which makes almost impossible for
+exceptions signaled by third party code, which makes almost impossible for
 library clients to protect against such exceptions; and a high prevalence of
-java.lang exceptions reported on issues - representing approximately 50\% of the
+java.lang exceptions reported on issues, representing approximately 50\% of the
 analyzed issues.
@@
 
-%\todo{This section feels a bit disconnected.}
 Several general guidelines have been proposed on how to use Java
@@
 bloch2008effective}. 
-% To guide our investigation, we compile the guidelines
-A compiled list of the guidelines related to exception types and exception propagation in Java the following
-\footnote{We also compiled guidelines related to exception handling but they are
-out of the scope of this paper.}:
+A compiled list of the guidelines related to exception types and exception
+propagation in Java the following: \footnote{We also compiled guidelines related
+to exception handling but they are out of the scope of this paper.}
 
@@
 \textbf{I-Use checked exceptions to represent recoverable
-conditions}(~\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,bloch2008effective}).
+conditions} (\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,bloch2008effective})
 The developer should use checked exceptions for conditions from which the caller
@@
 \textbf{II-Error represents an unrecoverable condition detected by the JVM which
-should not he handled}~\cite{gosling2000java}. It results from failures detected
+should not he handled} (\cite{gosling2000java}). It results from failures detected
 by the Java Virtual Machine which indicate resource deficiencies, invariant
@@
 \textbf{III-A method should throw exceptions that precisely define the
-exceptional condition}~\cite{gosling2000java,bloch2008effective}. To do so,
+exceptional condition} (\cite{gosling2000java,bloch2008effective}). To do so,
 developers should either try to reuse the exception types already defined in the
@@
 \textbf{IV-Document all exceptions thrown by a
-method}~\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,bloch2008effective}).
+method} (\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,bloch2008effective}).
 The exceptions thrown by a method are an important part of methods interface,
@@
 
-These best practices guided our exploratory study described next
-. Based on information extracted from stack traces (complemented by
+These best practices guided our exploratory study described next.
+Based on information extracted from stack traces (complemented by
 bytecode and source-code analysis), we investigate whether stack characteristics
@@
     analysis?
-    %~\todo{Perhaps rephrase this to: Are the identified best practices followed
-    %in practice?}
 
@@
 collecting data since February 2012. Up to Dec 2013, when we queried it to
-retrieve our bootstrapping sample which included: 356.057 issues from 16.836
+retrieve our bootstrapping sample, it included 356.057 issues from 16.836
 Java projects (step 1).\footnote{A Java project is specified in GitHub with a
@@
 analysis domain to Android projects (step 3). This decision was driven by the
-fact that most open source Android applications are known~\cite{ref} for dealing
+fact that most open source Android applications are known for dealing
 with several sources of exceptions due to multi-threaded execution and
@@
 information concerning the exception use. Moreover, issue and pull requests are
-dual on Github; all pull requests have a corresponding ``backing'' issue - which
+dual on Github; all pull requests have a corresponding ``backing'' issue which
 are automatically generated. Therefore, we needed to exclude pull request based
@@
 included stack traces. We identified and filtered those cases out (e.g.,
-pullWifi project was responsible for almost 50\% of all issues created on
+the \textsf{pullwifi} project was responsible for almost 50\% of all issues created on
 GitHub, it was removed from our analysis and also from GitHub).
@@
 example projects (i.e., toy projects) or tools to support Android development
-(e.g. selendroid, roboeletric -- tools to support the testing of Android apps).
+(e.g. \textsf{selendroid}, \textsf{roboeletric} --- tools to support the testing of Android apps).
 The filtered set consisted of 482 apps, from which approximately 50\% are also
@@
 errors. The exception classifier is based on bytecode analysis (based on Design
-Wizard~\cite{}) that walks up the type hierarchy of a Java exception until it
-reaches a base exception type.
+Wizard~\cite{Brunet09}) that walks up the type hierarchy of a Java exception
+until it reaches a base exception type.
 
@@
 projects analyzed we opted for not restricting the analysis only on defect
-issues~\footnote{We conducted the same analysis on the defect issues and the top
+issues.\footnote{We conducted the same analysis on the defect issues and the top
 exceptions are similar to the ones mentioned on defect issues. Due to space
@@
 on the defect issues can be found here:
-\url{www.dimap.ufrn.br/~roberta/icsm2014}}.
+\url{www.dimap.ufrn.br/~roberta/icsm2014}}
 
@@
 distinct projects they were found and the whole set of projects analyzed with at
-least one stack trace (482). Table~\ref{tab:toptenandroid} and
-Figure~\ref{fig:androidsignaler} presents the mined data.
+least one stack trace (482). Table~\ref{tab:toptenandroid} 
+%and Figure~\ref{fig:androidsignaler}
+presents the mined data.
 
@@
 programming bugs or resource limitations, there is usually no proper handling
-besides presenting an error message to the user and restarting the application -
+besides presenting an error message to the user and restarting the application ---
 only high fault tolerant systems need to provide solutions to handle them.
@@
 
-From Tables~\ref{tab:toptenjava} and ~\ref{tab:toptenandroid} we can observe that,
+From Tables~\ref{tab:toptenjava} and~\ref{tab:toptenandroid} we can observe that,
 in both Java and Android related issues, direct instances
@@
 In the Android repositories, most of such exceptions  were  thrown by the Android 
-platform/os (167 out of 232).
+platform/OS (167 out of 232).
 
@@
 To better investigate the extent of this problem we classified each stack trace according to the type of the
-root exception. Table~\ref{typeroottab} presents the types of root causes of all stack traces reported on 
+root exception. Table~\ref{tab:typeroottab} presents the types of root causes of all stack traces reported on 
 the Android repositories. As we can see from Table ~ref{typeroottab},  most of the exceptions signaled by third-party code
@@
   \end{tabular}
-\caption{Types of Root Exceptions.}
-  \label{fig:typeroottab}
+\caption{Types of root exceptions.}
+  \label{tab:typeroottab}
 \end{table}
@@
 libraries reused by Android platform (i.e., org.apache.harmony,
-org.w3c.dom, sun.misc, org.apache.http, org.json, org.xml, javax - referred in
+org.w3c.dom, sun.misc, org.apache.http, org.json, org.xml, javax -- referred in
 this work as libcore) were checked exceptions. It is pointed as a good practice
-for libraries (see Section ~\ref{best}) because by using checked exceptions
+for libraries (see Section~\ref{sec:best}) because by using checked exceptions
 libraries can define a precise exception interface to its clients. Such
@@
 \subsection{Odd Wrappings}
-Table~\ref{wrappingandroid} presents the wrappings found in this study for all
+Table~\ref{tab:wrappingandroid} presents the wrappings found in this study for all
 stack traces found in Android repositories. As we can see, some of the checked
@@
 \caption{Kinds of wrappings found on the stacks of 482 ANDROID projects found on Github}
-\label{fig:wrappingandroid}
+\label{tab:wrappingandroid}
 \end{table}
@@
 We could also observe that in 205 stacks found in Android repository the exception was wrapped 
-more then once. Figure ~\ref{sizewrapbhists}  illustrates the number of wrappings per stack, both
+more then once. Figure ~\ref{fig:sizewrapbhists}  illustrates the number of wrappings per stack, both
  in Java and Android studied repositories. One interesting multiple-wrapping found in Android repo was the 
@@
 (mis)-use were consistently detected such as: unexpected wrappings (e.g., Errors
-being wrapped in checked exceptions) - revealing that Java hybrid exception
+being wrapped in checked exceptions), revealing that Java hybrid exception
 model is not fully used according to its purpose; undocumented runtime
-exceptions signaled by third party code - which makes almost impossible for
+exceptions signaled by third party code, which makes almost impossible for
 library clients to protect against such exceptions; and a high prevalence of
-java.lang exceptions reported on issues - representing approximately 50\% of the
+java.lang exceptions reported on issues, representing approximately 50\% of the
 analyzed issues.
@@
 
-\section*{Acknowledgment} This work is partially supported by: CNPq - Proc.
+\section*{Acknowledgment} This work is partially supported by: CNPq -- Proc.
 484209/2013-2 and the NWO TestRoots project (639.022.314).

### 2014-05-09 android-stacks.tex

@@
 
-The exception (mis-)use that emerged from this study shades light on two
-
-It is wise to document the explicitly signaled runtime exceptions as carefully as checked exceptions. Doing so, the clients of a method will be aware of the exceptions the method can throw.
-
-When developer fails to document the runtime exceptions thrown by a method (s)he will make difficult or even impossible for the caller to effective use of such method.
-
-Specially when developing library code, developers must carefully document the runtime exceptions, since there is no compilation checking
-
-have not been fully adopted, mainly there is no compilation or runtime tool to enforce such policies.
-noindent\emph{Undocumented runtime exceptions}
+The exception (mis-)use patterns that emerged  in this study sheds enough light on 
+problems that may be causeing the lack of rocustness of current Java programs in the face of
+exceptional conditions: (i) the undocumented runtime exceptions thrown by library code - 
+documenting runtime exceptions is a tedious and error prone task, to help developers
+mitigating this problem tools should be developed to automate document runtime exceptions
+scaping from library code. (ii) the odd wrappings found in this study can lead to what we call
+an \emph{exception handling confusion problem} which may lead the problem to an
+unpredictable state while trying to handle an exception. To illustrate this problem, we can use
+ one of the examples found in this study: when the developer is 
+confronted with a checked exception, the designer of the API is telling him 
+to handle the exceptional condition, however we could observe that in some cases the 
+checked exception wrapped an OutOfMemoryError, which represent resource deficiency detected 
+by the JVM which make impossible the program to recover. Further investigation
+is needed on finding ways to help developers in deadling with such problem, either preventing 
+odd wrappings or enabling the developer to better dealing with them - when they are innevitable.
 

### 2014-05-09 android-stacks.tex
discussion finished v1

@@
 
-The exception (mis-)use patterns that emerged  in this study sheds enough light on 
-problems that may be causeing the lack of rocustness of current Java programs in the face of
-exceptional conditions: (i) the undocumented runtime exceptions thrown by library code - 
+The exception (mis-)use patterns that emerged  in this study sheds light on 
+problems that may be causing the lack of robustness of current Java programs in the face of
+exceptional conditions:
+ \noindent\emph{The undocumented runtime exceptions} thrown by library code -
 documenting runtime exceptions is a tedious and error prone task, to help developers
 mitigating this problem tools should be developed to automate document runtime exceptions
-scaping from library code. (ii) the odd wrappings found in this study can lead to what we call
-an \emph{exception handling confusion problem} which may lead the problem to an
-unpredictable state while trying to handle an exception. To illustrate this problem, we can use
+scaping from library code, few solutions in this directions have been proposed so far ~\cite{van2005combining}. 
+ \noindent\emph{The odd wrappings}  found in this study can lead to what we call
+an \emph{exception handling confusion problem} which can lead the program to
+an unpredictable state in the presence of exceptions. To illustrate this problem, we can use
  one of the examples found in this study: when the developer is 
@@
 checked exception wrapped an OutOfMemoryError, which represent resource deficiency detected 
-by the JVM which make impossible the program to recover. Further investigation
-is needed on finding ways to help developers in deadling with such problem, either preventing 
-odd wrappings or enabling the developer to better dealing with them - when they are innevitable.
+by the JVM which make impossible the program to recover. Further investigation is needed on finding 
+ways to help developers in deadling with such problem, either preventing odd wrappings or enabling 
+the developer to better dealing with them.
 

### 2014-05-09 android-stacks.tex
discussion finished v2

@@
 problems that may be causing the lack of robustness of current Java programs in the face of
-exceptional conditions:
+exceptional conditions.
  \noindent\emph{The undocumented runtime exceptions} thrown by library code -

### 2014-05-09 android-stacks.tex
discussion finished v3

@@
 The exception (mis-)use patterns that emerged  in this study sheds light on 
-problems that may be causing the lack of robustness of current Java programs in the face of
-exceptional conditions.
- \noindent\emph{The undocumented runtime exceptions} thrown by library code -
+problems that may be reducing the robustness of current Java programs in the face of
+exceptional conditions. \noindent\emph{The undocumented runtime exceptions} thrown by library code -
 documenting runtime exceptions is a tedious and error prone task, to help developers

### 2014-05-09 android-stacks.tex
new conclusion

@@
 
-In this work, we perform the first large scale analysis of Java stack traces and
-pinpoint how they can reveal bad programming practices. We mined and analyzed
-the stack traces embedded in all issues of Java projects available GitHub.
-Overall, our research set includes 356,057 issues define at 16,836 projects
-(from which 28,800 stack traces were extracted). From this set, the stack traces
-of 482 Android projects were investigated in more detail, in combination with
-source code and bytecode analysis. In this study some patterns of exception
+In this work, we mined and analyzed the stack traces embedded in all issues 
+of Java projects available GitHub focusing our attention on the stack traces reported on
+a subset of 482 Android projects. In this study the exception information available 
+on stack traces was used in combination with source code and bytecode analysis to 
+pin-point patterns of exceptions (mis-)use in Java. In this study some patterns of exception
 (mis)-use were consistently detected such as: unexpected wrappings (e.g., Errors
@@
 model is not fully used according to its purpose; undocumented runtime
-exceptions signaled by third party code, which makes almost impossible for
-library clients to protect against such exceptions; and a high prevalence of
-java.lang exceptions reported on issues, representing approximately 50\% of the
-analyzed issues.
+exceptions signaled by third party code (which makes almost impossible for
+library clients to protect against such exceptions); and a high prevalence of
+java.lang exceptions reported on issues (representing approximately 50\% of the
+analyzed issues). Such (mis)uses detected in out study can negatively affecting the 
+robustness of current Java software systems.  As a result, there is a need for both 
+improving the design of exception handling code in Java 
+and developing tools support to help developers dealing with such problems.
 

### 2014-05-09 android-stacks.tex
new conclusion v2

--- 2014-05-09 android-stacks.tex
@@
 
-In this work, we perform the first large scale analysis of Java stack traces and
-pinpoint how they can reveal bad programming practices. We mined and analyzed
-the stack traces embedded in all issues of Java projects available GitHub.
-Overall, our research set includes 356,057 issues define at 16,836 projects
-(from which 28,800 stack traces were extracted). From this set, the stack traces
-of 482 Android projects were investigated in more detail, in combination with
-source code and bytecode analysis. In this study some patterns of exception
+In this work we present an exploratory study on which we mined the stack 
+traces embedded in all issues of Java projects available GitHub, and analyzed the stack traces
+reported on a subset of 482 Android projects in more detail. In this study the information extracted 
+from stack traces was used in combination with source code and bytecode analysis to 
+pin-point patterns of exceptions (mis-)use in Java. And some patterns of exception
 (mis)-use were consistently detected such as: unexpected wrappings (e.g., Errors
@@
 model is not fully used according to its purpose; undocumented runtime
-exceptions signaled by third party code, which makes almost impossible for
-library clients to protect against such exceptions; and a high prevalence of
-java.lang exceptions reported on issues, representing approximately 50\% of the
-analyzed issues.
+exceptions signaled by third party code (which makes almost impossible for
+library clients to protect against such exceptions); and a high prevalence of
+java.lang exceptions reported on issues (representing approximately 50\% of the
+analyzed issues). Such (mis)uses detected in out study can negatively affect the 
+robustness of current Java software systems. Such results shed light to needs for  
+ developing tools support to help developers dealing with such problems and 
+improving the design of exception handling mechanism in Java.
 

### 2014-05-09 android-stacks.tex
small corrections 1

@@
 the stack traces embedded in all issues of Java projects available GitHub.
-Overall, our research set includes 356,057 issues define at 16,836 projects
-(from which 28,800 stack traces were extracted). From this set, the stack traces
+Overall, our research set includes 356,057 issues defined at 16,836 projects, 
+from which 28,800 stack traces were extracted. From this set, the stack traces
 of 482 Android projects were investigated in more detail, in combination with
@@
 model is not fully used according to its purpose;  undocumented runtime
-exceptions signaled by third party code, which makes almost impossible for
-library clients to protect against such exceptions; and a high prevalence of
-java.lang exceptions reported on issues, representing approximately 50\% of the
-analyzed issues.
+exceptions signaled by third party code, which is a serious threat to program
+dependability; and a high prevalence of java.lang exceptions reported on issues,  
+causing approximately 50\% of the analyzed stack traces.
 
@@
 
-Often some languages constructions may lead the developers into believing that
+Often some exception handling constructions may lead the developers into believing that
 by just re-throwing the exceptions they can forget about the exceptional
@@
 which offers  different types to represent different exception behaviors (i.e., error, 
-runtime and checked) (see Section ~\ref{extypes}) . According to Java specification Errors should not be
+runtime and checked) (see Section ~\ref{sec:best}) . According to Java specification Errors should not be
 handled inside the system since they usually represent unrecoverable conditions
@@
 behaviour of the exception handling code more complex and less reliable. For
-instance, considering a situation were an OutOfMemoryError (a situation that
+instance, considering a situation were an instance of OutOfMemoryError (a situation that
 should not be handled) was wrapped in a checked exception (a situation that must
@@
 In this work we present an exploratory study on which we mined the stack 
-traces embedded in all issues of Java projects available GitHub, and analyzed the stack traces
-reported on a subset of 482 Android projects in more detail. In this study the information extracted 
+traces embedded in all issues of Java projects available GitHub, and analyzed in mode detail thr
+stack traces reported on a subset of 482 Android projects. In this study, the information extracted 
 from stack traces was used in combination with source code and bytecode analysis to 

### 2014-05-09 android-stacks.tex
small corrections 2
small corrections 3

@@
 Natal, Brazil\\
-Email: \texttt{roberta@dimap.ufrn.br,lucas.almeida@lcc.ufrn.br}}
+Email: \texttt{roberta@dimap.ufrn.br,lucas.almeida@ppgsc.ufrn.br!}}
 \IEEEauthorblockA{\IEEEauthorrefmark{2}Delft University of Technology\\
@@
 limitation we limit to present the general analysis here. More detailed analysis
-on the defect issues can be found here:
-\url{www.dimap.ufrn.br/~roberta/icsm2014}}
+on the defect issues can be found at:
+\url{www.dimap.ufrn.br/~roberta/icsme2014}}
 
@@
 and iii) tools to extract stack traces information from natural language artifacts
-(i.e., issues and emails)
+(i.e., issues and emails).
 %, and iv) empirical studies involving Android apps.

### 2014-05-09 android-stacks.tex
small corrections 4

@@
 Natal, Brazil\\
-Email: \texttt{roberta@dimap.ufrn.br,lucas.almeida@ppgsc.ufrn.br!}}
+Email: \texttt{roberta@dimap.ufrn.br,lucas.almeida@ppgsc.ufrn.br}}
 \IEEEauthorblockA{\IEEEauthorrefmark{2}Delft University of Technology\\

### 2014-05-11 android-stacks.tex
monads
monads more ideas

@@
 ways to help developers in deadling with such problem, either preventing odd wrappings or enabling 
-the developer to better dealing with them.
+the developer to better dealing with them. Last but not least, the high number of null pointer exceptions 
+lead us to think on other pertinent research direction: whether language designs which avoid null pointers, 
+such as Monads ~cite{Walde95} (i.e., used in functional languages for values that may not be available 
+or computations that may fail) could improve the robustness of Java programs.
 

### 2014-05-11 android-stacks.tex
monads more ideas - reduced
Proof reading comments

@@
 detecting and recovering from such exceptional conditions. Although the
-exception handling mechanism have been embedded in several programming languages
-(e.g. Java, C++, C\#) and the target of several studies
+exception handling mechanism has been embedded in several programming languages
+(e.g. Java, C++, C\#) and was the target of several studies
 (e.g.~\cite{miller1997issues,Robil00,shah2010understanding,
@@
 
-Often some exception handling constructions may lead the developers into believing that
+Often exception handling constructions may lead the developers into believing that
 by just re-throwing the exceptions they can forget about the exceptional
@@
 Moreover, some languages use exceptions to represent programming mistakes
-(out-of-bounds array index, division-by-zero, access to a null). In these cases
+(out-of-bounds array index, division-by-zero, access to null pointers). In these cases,
 exceptions are implicitly signaled by the runtime environment when such
-condition happens. One of such implicit exceptions (caused by data conversion
+conditions occur. One of such implicit exceptions (caused by data conversion
 from a 64-bit floating point to a 16-bit signed integer) was the responsible for
@@
 million dollars. Besides Ariane 5, several applications crash everyday due to
-uncaught exceptions~\cite{jo2004uncaught}. The consequences of not careful
-exception usage may have the opposite effect of its initial intention of
-improving system robustness.
+uncaught exceptions~\cite{jo2004uncaught}. Not using exceptions carefully
+may have the opposite effect of the initial intention of improving
+system robustness.
 
@@
 information for our analysis. The stack trace analysis  was augmented by
-additional information extracted using byte code and source code analyses done
+additional information extracted using byte code and source code analysis done
 on a carefully selected subset of our initial sample (482 Android projects).
@@
 
-  \item  It briefly introduces ExceptionMiner, a tool developed to support the
+  \item  It introduces ExceptionMiner, a tool developed to support the
     analysis.
@@
 
-The contributions of this work allow for developers of robust Java applications
+The contributions of this work allow for developers of Java applications
 to familiarize with the most common reported exceptions mis-uses which is the
@@
 analysis. We strongly believe that, specially in the open-source environment,
-faults are not to be hidden in a private bug issue. Faults should be share and
+faults are not to be hidden in a private bug issue. Faults should be shared and
 discussed, so that a developer can learn from other projects mistakes.
@@
 debate about the use of checked and unchecked exceptions in
-Java~\cite{javatut,stackoverlow,debate}\footnote{152 questions in Stackoverlow
-are related to this debate}, since there are pros and cons associated to each of
+Java~\cite{javatut,stackoverlow,debate},\footnote{152 questions in Stackoverlow
+are related to this debate} since there are pros and cons associated to each of
 them.
@@
 bloch2008effective}. 
-A compiled list of the guidelines related to exception types and exception
-propagation in Java the following: \footnote{We also compiled guidelines related
+A list of the guidelines related to exception types and exception
+propagation in Java the following:\footnote{We also compiled guidelines related
 to exception handling but they are out of the scope of this paper.}
@@
 is expected to recover. By confronting the API user with a checked exception,
-the API designer is telling the client to handle the exceptional condition. The
+the API designer is forcing the client to handle the exceptional condition. The
 client can explicitly ignore the exception (swallowing, or converting it to
@@
 exceptions being thrown by it. According to~\cite{bloch2008effective}, it is
-also wise to document the explicitly signaled runtime exceptions\footnote{This
+also wise to document the explicitly thrown runtime exceptions\footnote{This
 excludes the implicitly signaled JVM runtime exceptions related to programming
 mistakes} as carefully as checked exceptions. Doing so, the clients of a method
-will be aware of the exceptions the method can throw. If the developer fails to
+will be aware of all the exceptions the method can throw. If the developer fails to
 do follow this practice (specially when developing library code) it will be
-difficult or even impossible for the caller to make effective use of such method
-~\cite{wirfs2006toward, bloch2008effective}.
+difficult or even impossible for the caller to make effective use of such 
+method~\cite{wirfs2006toward, bloch2008effective}.
 
@@
 mixed-methods approach~\cite{ivankova2006using} for collecting, analyzing, and
-integrating both quantitative and qualitative analysis along the stages of the
+integrating both quantitative and qualitative evidence along the stages of the
 research process. The main goal of this study was to gain a better understanding
@@
 For RQ1 and RQ2, we explore the domain quantitatively, and use theirs answers as
-a basis to investigate the third research question. RQ3 aimed at finding whether
-of exception (mis)-uses could be revealed through stack trace analysis. To
+a basis to investigate the third research question. RQ3 aims at examining whether
+exception (mis)-uses could be revealed through stack trace analysis. To
 support this investigation, source code and bytecode analysis was used to
-leverage the  understanding of stack traces. For RQ3 we investigate
-quantitatively and highlight interesting cases by exploring cases qualitatively.
+leverage the  understanding of stack traces. 
+%For RQ3 we investigate
+%quantitatively and highlight interesting cases by exploring cases qualitatively.
 Figure~\ref{fig:overviewfig} presents an overview of the mixed-methods approach
@@
 project~\cite{Gousi13}, an off-line mirror of the data offered through the
-Github API. The Github API data come in two forms; a streaming data flow lists
+Github API. The Github API data come in two flavours; a streaming data flow lists
 events, such as reporting an issue, happening on repositories in real time,
@@
 labels attached. Hence, this study does not restrict the analysis of stack
-traces defined on defect-issues, based on the assumption that: regardless the
+traces defined on defect-issues, based on the assumption that regardless of the
 type of issue where its is defined the stack trace information contain relevant
-information concerning the exception use. Moreover, issue and pull requests are
+information concerning the exception use. Moreover, issues and pull requests are
 dual on Github; all pull requests have a corresponding ``backing'' issue which
-are automatically generated. Therefore, we needed to exclude pull request based
+are automatically generated. Therefore, we excluded pull request based
 issues from our analysis. Finally, the Github API allows the automated
@@
 included stack traces. We identified and filtered those cases out (e.g.,
-the \textsf{pullwifi} project was responsible for almost 50\% of all issues created on
-GitHub, it was removed from our analysis and also from GitHub).
+the~\textsf{pullwifi} project was responsible for almost 50\% of all Android issues in our dataset).
 
@@
 \noindent\emph{Exception Classifier} The next step in the ExceptionMiner process
-is classifying exceptions as either checked exceptions, runtime exceptions or an
+is classifying exceptions as either checked exceptions, runtime exceptions or 
 errors. The exception classifier is based on bytecode analysis (based on Design
@@
 
-The use of the Github issue tracker in repositories hosted on Github is optional
-and not widespread. As of March 2014, from the 332,864 non-fork Java
-repositories registered in the GHTorrent dataset, only 44,323 have received an
-issue report in their lifetime while from those only 16,837 are being watched by
-external users. To ensure that our work targets projects that are openly used by
-users other than their developers, we focused our analysis on just those
-projects. Using the heuristics presented in Section~\ref{sec:android}, we
-selected 482 repositories featuring Android projects. Table~\ref{tab:proj} and
-Table~\ref{tab:overview} presents two views of the extracted dataset, namely
-grouped by repository and grouped by issue.
-
+%The use of the Github issue tracker in repositories hosted on Github is optional
+%and not widespread. As of March 2014, from the 332,864 non-fork Java
+%repositories registered in the GHTorrent dataset, only 44,323 have received an
+%issue report in their lifetime while from those only 16,837 are being watched by
+%external users. To ensure that our work targets projects that are openly used by
+%users other than their developers, we focused our analysis on just those
+%projects. Using the heuristics presented in Section~\ref{sec:android}, we
+%selected 482 repositories featuring Android projects. Table~\ref{tab:proj} and
+%Table~\ref{tab:overview} presents two views of the extracted dataset, namely
+%grouped by repository and grouped by issue.
+%
 From Table~\ref{tab:overview}, we can see that repositories that include at
@@
 NullPointerExceptions being signaled from third-party libraries. Regarding
-reusable code, there is no consensus if it is a good or bad practice to re-throw
-NullPointerException. Some prefer to encapsulate such an exception on
+reusable code, there is no consensus whether it is a good or bad practice to 
+re-throw
+a NullPointerException. Some prefer to encapsulate such an exception on
 IllegalArgumentException, while others~\cite{bloch2008effective} argue that the
@@
 programming bugs or resource limitations, there is usually no proper handling
-besides presenting an error message to the user and restarting the application ---
-only high fault tolerant systems need to provide solutions to handle them.
+besides presenting an error message to the user and restarting the application.
+Only high fault tolerant systems need to provide solutions to handle them.
 
@@
 cause of the exceptional behaviour, and as a consequence developers need to relying on
-the exception message which can be neither complete nor precise
-~\cite{gosling2000java}.
+the exception message which can be neither complete nor precise~\cite{gosling2000java}.
 
@@
 JVM (due to programming mistakes) and inspecting the signaler methods of such exceptions (i.e.,  direct an inderect 
-instances of java.lang.RuntimeException). We could observe observe that only 1 method out of 118 inspected signalers
- (i.e., 0,8%) documented the explicitly thrown runtime exception in a Javadoc comment; and none of these methods
+instances of java.lang.RuntimeException). We could observe that only 1 method out of 118 inspected signalers
+ (i.e., 0,8\%) documented the explicitly thrown runtime exception in a Javadoc comment; and none of these methods
 included the runtime exception  (reported on the stack) as part of the exception interface of the method (i.e., using 
@@
 
-Such undocumented runtime exceptions represent a serious threat to system robustness, specially
+Such undocumented runtime exceptions represent a threat to system robustness, specially
 when such exceptions are thrown then the third party code is invoked inside the application (e.g. libraries, or framework utility code).
-In such cases the client usualy do not have access to the source code, and in the absense of
-the exception documentation it is very difficult or even impossible for the client of such thrid party code to 
+In such cases the client usually do not have access to the source code, and in the absence of
+the exception documentation it is very difficult or even impossible for the client of such third party code to 
 protect the application against undocumented runtime exceptions. As a consequence, the
@@
 libraries are heavily used in several projects, and such precise exception
-interface may be related to the libraries maturity.
+interface may be related to the library's maturity.
 
@@
 on which this exception flows explicitly specifies the exception, or if the
-checked exception is wrapped in a runtime and than re-thrown.
-
-Since checked exceptions should be used to represent recoverable conditions, ignoring a
+checked exception is wrapped in a runtime and than re-thrown. Since checked exceptions 
+should be used to represent recoverable conditions, ignoring a
 checked exception, is considered a bad practice. As mentioned before, when the developer is confronted
@@
 which offers  different types to represent different exception behaviors (i.e., error, 
-runtime and checked) (see Section ~\ref{sec:best}) . According to Java specification Errors should not be
+runtime and checked) (see Section~\ref{sec:best}). According to Java specification Errors should not be
 handled inside the system since they usually represent unrecoverable conditions
@@
 
-\section{Discussions}
+\section{Discussion}
 
@@
 documenting runtime exceptions is a tedious and error prone task, to help developers
-mitigating this problem tools should be developed to automate document runtime exceptions
-scaping from library code, few solutions in this directions have been proposed so far ~\cite{van2005combining}. 
+mitigate this problem tools should be developed to automate the documentation of runtime exceptions
+scraping from library code, few solutions in this directions have been proposed so far ~\cite{van2005combining}. 
  \noindent\emph{The odd wrappings}  found in this study can lead to what we call
@@
 In this section, we present works that are related to our own, distributed in
-four categories: i) works that use the information available on stack traces;
+three categories: i) works that use the information available on stack traces;
 ii) empirical studies on the usage of Java exceptions and its fault proneness;
@@
 
-\textit{Analysis of Java Exception Usage.} Since the manual analysis of the Java
+\textit{Analysis of Java Exception Usage.} As the manual analysis of the Java
 exception flow can easily become infeasible, Robillard and Murphy~\cite{Robil00}
@@
 tools have been proposed to support the static analysis of exception
-flows~\cite{coelho2008assessing}. The main limitation off all
+flows~\cite{coelho2008assessing}. The main limitation of all
 static analysis tools is the number of false positives inherent to static analysis
@@
 able to extract stack traces embedded on verbose log files (i.e., on which we
-can find log text mixed with exception frames). Bacchelli and Lanza
-~\cite{bacchelli2012content} propose a solution to recognize stack trace frames
+can find log text mixed with exception frames). Bacchelli
+et al.~\cite{bacchelli2012content} propose a solution to recognize stack trace frames
 from development emails and relate it to code artifacts (i.e. classes) mentioned

### 2014-05-14 android-stacks.tex
Several smaller textual improvements to introduction
And removed paragraph about private issues,
as we don't look into those.

@@
 Often exception handling constructions may lead the developers into believing that
-by just re-throwing the exceptions they can forget about the exceptional
+by just re-throwing and possibly logging the exceptions they can forget about the exceptional
 situations during the development of the "happy path". This "ignore-for-now"
@@
 mechanism~\cite{mandrioli1992advances} making the program more complex and even
-less reliable. This behaviour may lead to the well known uncaught exceptions
-problem~\cite{jo2004uncaught}, one of the main causes of application crashes.
+less reliable. This behaviour may lead to the problem of uncaught exceptions~\cite{jo2004uncaught}, one of the main causes of application crashes.
 
 Moreover, some languages use exceptions to represent programming mistakes
-(out-of-bounds array index, division-by-zero, access to null pointers). In these cases,
+(array index out-of-bounds, division-by-zero, dereferencing null pointers). In these cases,
 exceptions are implicitly signaled by the runtime environment when such
 conditions occur. One of such implicit exceptions (caused by data conversion
-from a 64-bit floating point to a 16-bit signed integer) was the responsible for
-the famous failure on Ariane 5's first test flight~\cite{lions1996ariane} ---
+from a 64-bit floating point to a 16-bit signed integer) was responsible for
+the infamous failure on Ariane 5's first test flight~\cite{lions1996ariane} ---
 leading to the rocket self-destructing 37 seconds after launch and a loss of 500
@@
 terminates, while the system prints a stack trace to the console. A stack trace
-includes a detail message and the execution stack frame when the exception
+includes an explanatory message and the execution stack frame when the exception
 occurred. A typical Java stack trace consists of an ordered list of methods that
-were active on the call stack before the exception has occurred. Stack traces
-are a useful source of information about system crashes, and is often used to
+were active on the call stack before the exception occurred. Stack traces
+are a useful source of information about system crashes, and are often used to
 support developers in debugging~\cite{schroter2010stack}. Moreover, information
@@
 
-However, all these works focus on the analysis of a single system at a time. In
-our current context, on which we have plenty access to the information available
-on open-source repositories like GitHub initial questions arise: What is the
+However, all existing research has focused on the analysis of a single system at a time. In
+our current context, on which we have access to wealth of information available
+on open-source repositories like GitHub additional questions arise: What is the
 prevalence of exception stack traces on reported issues? What are the common
@@
 To answer these questions, we conduct an exploratory study of stack traces in
-the wild. Using a custom tool called ExceptionMiner, that was specifically
-developed to for this study, we mined stack traces from issues across all Java
+the wild. Using a custom tool called ExceptionMiner, which was specifically
+developed for this study, we mined stack traces from issues across all Java
 projects on GitHub. Overall, we analyzed 356,057 issues from 16,836 projects,
@@
 
-This study is built on the following assumption: the stack traces reported on
+This study builds upon the following assumption: the stack traces reported on
 issues contain relevant information related to system crashes. Several studies
-and techniques have also based on this assumption
-(e.g.,~\cite{sinha2009fault}~\cite{dhaliwal2011classifying}~\cite{kim2013predicting})
-and we build on it as well. Some outcomes were consistently detected through
-this large scale analysis of exception stack traces in the wild, such as:
+and techniques have also taken this assumption as starting point
+(e.g.,~\cite{sinha2009fault}~\cite{dhaliwal2011classifying}~\cite{kim2013predicting}). Some outcomes were consistently detected through
+our large scale analysis of exception stack traces in the wild, such as:
 
@@
 
-  \item  A multitude of programming mistakes as the main causes of stack traces.
+  \item  A multitude of programming mistakes which are the main causes of stack traces.
 
@@
   \item  Odd exception chains (e.g., Error wrapping checked and runtime
-    exceptions and vice-versa). As an indication that the purpose of Java hybrid
+    exceptions and vice-versa), indicating that the purpose of Java's hybrid
     exception model may not have been adequately used.
@@
 
-  \item  It introduces ExceptionMiner, a tool developed to support the
+  \item  It introduces ExceptionMiner, a tool developed to support cross-repository stack trace
     analysis.
@@
 
-The contributions of this work allow for developers of Java applications
+These contributions are relevant to:
+(i) developers of Java applications
 to familiarize with the most common reported exceptions mis-uses which is the
-first step to help developers to avoid making them; (ii) designers of languages
-to consider ways of reducing the abundance of NullPointerExceptions, (iii) and
+first step to help developers to avoid making them; 
+(ii) designers of languages
+to consider ways of reducing the abundance of NullPointerExceptions;
+and (iii)
 tool designers to consider developing tools that enable a cross-project defect
-analysis. We strongly believe that, specially in the open-source environment,
-faults are not to be hidden in a private bug issue. Faults should be shared and
-discussed, so that a developer can learn from other projects mistakes.
-Currently, the search facilities of repositories are very limited, and the
-ExceptionMiner tool is a contribution in this direction.
+analysis.
+
+% AVD: We have not looked at private issues, and the tool is described above.
+%We strongly believe that, specially in the open-source environment,
+%faults are not to be hidden in a private bug issue. Faults should be shared and
+%discussed, so that a developer can learn from other projects mistakes.
+%Currently, the search facilities of repositories are very limited, and the
+%ExceptionMiner tool is a contribution in this direction.
 
@@
 %RSC: If needed we can remove it.
-The remainder of this paper is organized as follows. Section 2 presents a
+This paper is organized as follows. Section 2 presents a
 background on exception handling mechanism. Section 3 presents the study design.

### 2014-05-14 android-stacks.tex

@@
 Techniques for error detection and handling are not  an optional add-on but a
-fundamental part of the system~\cite{bruntink2006discovering}. Several
-approaches exists for detecting errors and responding to them, such as the return-code 
-idiom (very popular in C programs and operating systems), and the exception handling 
-mechanism, embedded in many main stream programming languages.
+fundamental part of any system~\cite{bruntink2006discovering}. Several
+approaches exist for detecting errors and responding to them, such as the return-code 
+idiom (widely used in systems written in C), and the exception handling 
+mechanism, embedded in many main-stream programming languages.
 
@@
 
-In this section, we present works that are related to our own, distributed in
-three categories: i) works that use the information available on stack traces;
+In this section, we present work that is related to the present paper, divided into
+three categories: i) papers that use the information available on stack traces;
 ii) empirical studies on the usage of Java exceptions and its fault proneness;
 and iii) tools to extract stack traces information from natural language artifacts
-(i.e., issues and emails).
+(e.g., issues and emails).
 %, and iv) empirical studies involving Android apps.
@@
 
+\enlargethispage{-2\baselineskip}
+
 \section{Conclusion}
 
-In this work we present an exploratory study on which we mined the stack 
-traces embedded in all issues of Java projects available GitHub, and analyzed in mode detail thr
+In this paper we present an exploratory study in which we mined the stack 
+traces embedded in all issues of Java projects available GitHub.
+Furthermore, we analyzed in more detail the
 stack traces reported on a subset of 482 Android projects. In this study, the information extracted 
 from stack traces was used in combination with source code and bytecode analysis to 
-pin-point patterns of exceptions (mis-)use in Java. And some patterns of exception
-(mis)-use were consistently detected such as: unexpected wrappings (e.g., Errors
-being wrapped in checked exceptions), revealing that Java hybrid exception
+pin-point patterns of exception use and misuse in Java.
+Certain patterns of exception
+(mis)-use were consistently detected such as unexpected wrappings (e.g., Errors
+being wrapped in checked exceptions), revealing that  the Java hybrid exception
 model is not fully used according to its purpose; undocumented runtime
-exceptions signaled by third party code (which makes almost impossible for
-library clients to protect against such exceptions); and a high prevalence of
-java.lang exceptions reported on issues (representing approximately 50\% of the
-analyzed issues). Such (mis)uses detected in out study can negatively affect the 
-robustness of current Java software systems. Such results shed light to needs for  
+exceptions signaled by third party code (which makes it almost impossible for
+library clients to protect against such exceptions); 
+and a high prevalence of
+\texttt{java.lang} exceptions reported on issues (representing approximately 50\% of the
+analyzed issues). Such (mis)uses detected in our study can negatively affect the 
+robustness of current Java software systems. Our results call for  
  developing tools support to help developers dealing with such problems and 
-improving the design of exception handling mechanism in Java.
+improving the design of the exception handling mechanism in Java.
 

### 2014-06-27 android-stacks.tex
changing the focus of Introduction

@@
 dependability; and a high prevalence of java.lang exceptions reported on issues,  
-causing approximately 50\% of the analyzed stack traces.
+they were found in aproximately 50\% of all analyzed exception stack traces.
 
@@
 \section{Introduction}
+
+Along the recent years we have witnessed a rapid increase in 
+the number of open source software (oss) available on the Web.
+Since such projects make their control version system, mailing lists and 
+bug tracking systems (bts) publicly available, they represent a valuable 
+source of information for the "state of practice" on software development
+which has already started to be investigated [][][][][].
+
+In this scenario Github emerged as one of the most used repository hosting sites.
+It  alows developers to create and maintain opensource projects relying on Git distributed version 
+control and provides a set of tools which allow the information about developers 
+and their activities to be visible within and across open souce projects.
+
+Github is also known as a social coding platfotm [ref] as it ....
+
+Among those tools there is the Github issue tracking system, which makes
+issues related to new, features, bugs and general comments publicly available.
+
+The bug issues have been the focus of some research... In [ref] defined a 
+set of characteristics a good bug issue should have. The 2 most important information 
+are the: (1) steps to reproduce the bug; (2) and the failure stack trace.
+
+A typical stack trace consists of an ordered list of methods that
+were active on the call stack before the failure occurred. Stack traces
+are a useful source of information about system crashes, and are often used to
+support developers in debugging~\cite{schroter2010stack}. Moreover, information
+from stack traces combined with crash and bug reports can be used to support bug
+classification and clustering~\cite{wang2013improving, kim2011crash,
+dhaliwal2011classifying}, fault-proneness prediction
+models~\cite{kim2013predicting} and even automated bug
+fixing~\cite{sinha2009fault} tools.
+
+However, all existing research has focused on the analysis of a single system at a time. In
+our current context, on which we have access to wealth of information available
+on open-source repositories like GitHub additional questions arise: What are the common
+characteristics of such stack traces? Can developers benefit from other systems faults?
+
+-----
+Along the recent years we have witnessed a rapid increase in 
+the number of social applications on the Web. Applications which 
+allow their users to track the activities of a large number 
+of others regardless of his/her location and affiliation.
+
+This tendency can also be seen on the software development context 
+where Github emerged and became an extensively used social coding 
+platform. Github alows developers to create and maintain opensource 
+projects relying  on Git distributed version control, and provides a set 
+of tools which allow the information about developers and their activities
+ to be visible within and across open souce projects.
+
+Among those tools there is the Github issue tracking system, which makes
+issues related to new, features, bugs and general comments publicly available.
+
+-----
 
@@
 In Java, when the program fails due to an uncaught exception, it automatically
-terminates, while the system prints a stack trace to the console. A stack trace
+terminates, while the system prints a stack trace to the console.
+
+A stack trace
 includes an explanatory message and the execution stack frame when the exception

### 2014-07-01 android-stacks.tex
including crash/exception information on  Introduction 2

@@
 Along the recent years we have witnessed a rapid increase in 
-the number of open source software (oss) available on the Web.
-Since such projects make their control version system, mailing lists and 
-bug tracking systems (bts) publicly available, they represent a valuable 
+the number of open source software  available on the Web.
+Since such projects make their control version and 
+issue tracking systems publicly available, they represent a valuable 
 source of information for the "state of practice" on software development
-which has already started to be investigated [][][][][].
+which had already started to be investigated [][][][][].
 
@@
 
-Github is also known as a social coding platfotm [ref] as it ....
-
-Among those tools there is the Github issue tracking system, which makes
-issues related to new, features, bugs and general comments publicly available.
-
-The bug issues have been the focus of some research... In [ref] defined a 
-set of characteristics a good bug issue should have. The 2 most important information 
-are the: (1) steps to reproduce the bug; (2) and the failure stack trace.
+Among these tools there is the Github issue tracking system, which makes
+issues related to new features, improvements, bugs publicly available.
+
+Bug issues have called the attention at have been the target of several 
+researchs [][][]. A subtype of a bug issue are the ones related to system 
+crashes [complete]...
+
+In [ref] defined a set of characteristics a good bug issue should have. They showed that among 
+the most important relevant information are: (1) steps to reproduce the bug; 
+(2) and the failure stack trace.
 
@@
 
-However, all existing research has focused on the analysis of a single system at a time. In
-our current context, on which we have access to wealth of information available
-on open-source repositories like GitHub additional questions arise: What are the common
-characteristics of such stack traces? Can developers benefit from other systems faults?
+We could observe that existing research has focused on the analysis of a 
+single system at a time. Moreover, the search facilities of repositories are very limited, also
+enabling the analysis of a single project.
+
+Modern applications have to cope with similar and increasing number of abnormal
+computation states that arise as a consequence of faults in the application
+itself (e.g., access of null references), noisy user inputs or faults in
+underlying middleware or hardware. The exception handling mechanism~\cite{goodenough1975exception} is one of the most used schemes for detecting and recovering from such exceptional conditions.
+
+Although the exception handling mechanism has been embedded in several 
+programming languages (e.g. Java, C++, C\#) and was the target of several studies
+(e.g.~\cite{miller1997issues,Robil00,shah2010understanding, garcia2007extracting, garcia2001comparative,cabral2007exception,coelho2011unveiling}),
+the exception handling code is often generally poorly understood and the least
+tested part of software systems.
+
+Often exception handling constructions may lead the developers into believing that
+by just re-throwing and possibly logging the exceptions they can forget about the exceptional
+situations during the development of the "happy path". This "ignore-for-now"
+approach may turn the exception handling into a generalized "goto"
+mechanism~\cite{mandrioli1992advances} making the program more complex and even
+less reliable. This behaviour may lead to the problem of uncaught exceptions~\cite{jo2004uncaught}, one of the main causes of application crashes.
+
+Specially in the open-source environment, where faults are usualy not hidden in a private trackers,  it would be intresting if developers could learn from other projects mistakes.
+Moreover, interesting knowledge may emerge from a cross-project crash analysis.
+
+Hence, in such context where we have access to wealth of information available
+on open-source repositories such as GitHub, some questions arise: What are the common
+characteristics of Github crash reports? How can we benefit from such cross-project crash analysis?
+
+. The ExceptionMiner tool is a contribution in this direction (to enable a cross-project defect analysis).
 

### 2014-07-02 android-stacks.tex
Including content about CLU language in Section2

@@
 exception mechanism~\cite{goodenough1975exception} was proposed as a way of
-preventing errors from being ignored, and was embedded in many mainstream
-programming languages. Among some similarities they mainly differ in the way
-exceptions are represented and handled in each
-language~\cite{garcia2001comparative}:
-
-\noindent\emph{Checked Approach}. In languages such Modula-3, Guide, and
-Extended Ada all exceptions are checked, which means that the compiler
+preventing errors from being ignored and was embedded in many mainstream
+programming languages ~\cite{garcia2001comparative} since then.
+
+\noindent\emph{Checked Approach}. In languages such as CLU \footnote{In 
+CLU if a handler is found for a given exception, it is converted in a predened 
+default exception, called failure which is the only one which can be implicitly 
+propagated in this model.}, Modula-3, Guide, and Extended Ada all exceptions 
+are checked, which means that the compiler
 statically checks if appropriate handlers are provided within the system.
@@
 all exceptions are unchecked but the developer can optionally list exception
-types on method signatures. However, the developer is not warned by the compiler
-if an unchecked exception is not handled inside the application. Such an
-approach may hinder robustness because the client of a method cannot easily know
-which unchecked exceptions may be thrown, unless the code of the method and the
-methods called from it is inspected, which can be a very time consuming or
+types on method signatures. Although in some languages such exceptions can 
+be documented on the methods signature, the developer is not warned by the compiler
+if an unchecked exception is not handled inside the application. According to ~\cite{Robil00} 
+such an approach may hinder robustness because the client of a method cannot easily know
+which unchecked exceptions may be thrown (unless the code of the method and the
+methods called from it is inspected) which can be a very time consuming or
 infeasible task.
@@
 \noindent\emph{Hybrid Approach}. In Java a method can throw either checked or
-unchecked exceptions. Java is the only language that allows this hybrid
-approach.
+unchecked exceptions. Java is the only language that allows this hybrid approach,
+on which the developer can define exceptions of both types of exceptions to be thrown and
+handled in a single system.
 
@@
 represents more than 50\% of the analyzed set. The table illustrates  i) the
-number of times each exception appeared as the root cause of a stack trace
+number of times each exception appeared as the root cause reported on the exception stack trace
 (\#occurrences), ii) the number of distinct projects reporting such causes on

### 2014-07-02 android-stacks.tex
addressing minor changes

@@
 the information that composes a stack trace and storing the results in a
-relational database. The extracted stack trace information  the exception
-signaler, the root cause, the exception wrappers. The tool is based on a
-combination of a regular expression based parser with heuristics able to
-identify and filter exception names and stack traces inline with text. In
+relational database. Some of the attributes extracted from the stack traces were:
+ the exception signaler, the root cause, and the root cause wrappers. 
+The tool is based on a combination between a regular expression based parser 
+with heuristics able to identify and filter exception names and stack traces inline with text. In
 contrast to existing issue parsing solutions such as Infozilla, the parser
@@
 For each exception we could discover the number of times it occurred on stacks,
-the number of distinct projects such stacks were defined (and consequently its
+the number of distinct projects on which such stacks were defined (and consequently its
 popularity) and the of times it was signaled by different signaler types. We

### 2014-07-02 android-stacks.tex
improving process description

@@
 
-\noindent\emph{Checked Approach}. In languages such as CLU \footnote{In 
-CLU if a handler is found for a given exception, it is converted in a predened 
-default exception, called failure which is the only one which can be implicitly 
-propagated in this model.}, Modula-3, Guide, and Extended Ada all exceptions 
-are checked, which means that the compiler
-statically checks if appropriate handlers are provided within the system.
+In languages such as CLU, Modula-3, Guide, and Extended Ada the exceptions are 
+checked, which means that the 
+compiler statically checks if appropriate handlers are provided within the system. 
 Accordingly, the programmer must explicitly specify every exception that can be
-signaled to support the compile time checking. This approach is known to be
+signaled to support the compile time checking \footnote{In CLU,  if a handler is 
+found for a given exception, it is converted in a predened default exception, 
+called failure which is the only one which can be implicitly 
+propagated in this model.}. This approach is known to be
 costly to maintain (Dooren and Steegmans, 2005) since exception updates are
@@
 
-\noindent\emph{Unchecked Approach}. In languages such as Lore, C++, and Arche
-all exceptions are unchecked but the developer can optionally list exception
-types on method signatures. Although in some languages such exceptions can 
-be documented on the methods signature, the developer is not warned by the compiler
+In languages such as Lore, C++, and Arche all exceptions are unchecked but the developer can optionally list exception types on method signatures. Although exceptions can 
+be documented on methods signatures, the developer is not warned by the compiler
 if an unchecked exception is not handled inside the application. According to ~\cite{Robil00} 
@@
 
-\noindent\emph{Hybrid Approach}. In Java a method can throw either checked or
-unchecked exceptions. Java is the only language that allows this hybrid approach,
-on which the developer can define exceptions of both types of exceptions to be thrown and
-handled in a single system.
+In Java a method can throw either checked or unchecked exceptions.
+Java is the only language that allows this hybrid approach, on which the developer can define exceptions of both kinds to be thrown and
+handled in a single system. Next section details Java exception handling model.
 
@@
 
-%\subsubsection{Exception Chaining} Exception Chaining 
+%\subsubsection{Exception Chaining} Exception Wrapping 
 %This stack trace was extracted from an issue of an Android application in
@@
 
-Exception chaining is a common way of propagating exceptions in Java programs, 
+The Exception wrapping (also known as exception chaining) is a common way of 
+propagating exceptions in Java programs, 
 which allows one exception to be wrapped in another exception and re-thrown. 
 Figure~\ref{fig:wrapping} presents a stack trace associated which illustrates an
-exception chaining. The bottom part of the stack trace is the \emph{root cause}, which indicates the
+exception wrapping. The bottom part of the stack trace is the \emph{root cause}, which indicates the
 first reason for the error to be thrown (in this case, the computer run out of
@@
 
-For RQ1 and RQ2, we explore the domain quantitatively, and use theirs answers as
+For RQ1 and RQ2, we explore the domain quantitatively, and use their answers as
 a basis to investigate the third research question. RQ3 aims at examining whether
@@
 collecting data since February 2012. Up to Dec 2013, when we queried it to
-retrieve our bootstrapping sample, it included 356.057 issues from 16.836
-Java projects (step 1).\footnote{A Java project is specified in GitHub with a
+retrieve our sample, it included  16.836 Java projects (step 1) which overall contained
+356.057 issues.\footnote{A Java project is specified in GitHub with a
 specific label.}
 
-After obtaining the bootstrap issue sample from GHTorrent, each issue was
-processed by ExceptionMiner tool developed in this work which extracted every
-exception defined on issues or on issue comments (step 2). Since this initial
-set of Java project is too diverse (it includes libraries, web-based or
+After obtaining this sample of projects and associated issues, 
+the ExceptionMiner tool, developed in this work, performed a per-project 
+mining task as follows: for each project, the content of each issue and issue comments 
+were analized and all exceptions and exception stack traces were extracted (step 2). 
+Since this initial set of Java project is too diverse (it includes libraries, web-based or
 stand-alone applications), in the second phase of this study we restrict the
@@
 root exception. Table~\ref{tab:typeroottab} presents the types of root causes of all stack traces reported on 
-the Android repositories. As we can see from Table ~ref{typeroottab},  most of the exceptions signaled by third-party code
+the Android repositories. As we can see from Table ~\ref{tab:typeroottab},  most of the exceptions signaled by third-party code
  (e.g., Android and libs) were runtime exceptions. After filtering all the exceptions implicitly signaled by 
@@
 lead us to think on other pertinent research direction: whether language designs which avoid null pointers, 
-such as Monads ~cite{Walde95} (i.e., used in functional languages for values that may not be available 
+such as Monads ~\cite{Walde95} (i.e., used in functional languages for values that may not be available 
 or computations that may fail) could improve the robustness of Java programs.

### 2014-07-02 android-stacks.tex
addressing minor changes - reviewer 3

@@
 %\noindent\emph{Exception Documentation}
-\textbf{IV-Document all exceptions thrown by a
+\textbf{ IV-Document all exceptions thrown by a
 method} (\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,bloch2008effective}).
@@
 difficult or even impossible for the caller to make effective use of such 
-method~\cite{wirfs2006toward, bloch2008effective}.
+method~\cite{wirfs2006toward, bloch2008effective}. As mentioned in Section II, 
+in some languages such as C
 
@@
 
-We begun our investigation using the dataset provided by the GHTorrent
-project~\cite{Gousi13}, an off-line mirror of the data offered through the
-Github API. The Github API data come in two flavours; a streaming data flow lists
-events, such as reporting an issue, happening on repositories in real time,
-while a static view contains the current state of entities. The data is stored
-in unprocessed format, in a MongoDB database, while metadata is extracted and
-stored in a relational database. The GHTorrent dataset covers a broad range of
-development activities on Github, including issues. The project has been
-collecting data since February 2012. Up to Dec 2013, when we queried it to
-retrieve our sample, it included  16.836 Java projects (step 1) which overall contained
-356.057 issues.\footnote{A Java project is specified in GitHub with a
-specific label.}
-
-After obtaining this sample of projects and associated issues, 
-the ExceptionMiner tool, developed in this work, performed a per-project 
-mining task as follows: for each project, the content of each issue and issue comments 
-were analized and all exceptions and exception stack traces were extracted (step 2). 
-Since this initial set of Java project is too diverse (it includes libraries, web-based or
-stand-alone applications), in the second phase of this study we restrict the
-analysis domain to Android projects (step 3). This decision was driven by the
-fact that most open source Android applications are known for dealing
+%The GHTorrent dataset covers a broad range of
+% development activities on Github, including the all the issues per project. Such data 
+%is stored in unprocessed format, in a MongoDB database, while metadata is extracted 
+%and stored in a relational database.
+
+The mixed-methods approach is composed by the following steps:
+
+Step1. Retrieving the Dataset to be Analized. We begun our investigation using the dataset 
+provided by the GHTorrent project~\cite{Gousi13}, an off-line mirror of the data 
+offered through the Github API.  GHTorrent has been collecting data from Git since 
+February 2012. Up to Dec 2013, when we queried GHTorrent, it included 16.836 Java projects 
+which overall contained 356.057 issues.\footnote {A Java project is specified in GitHub with a
+specific label.} From GHTorrent we retrieved the metadata related to each Java project (e.g., 
+repository's names and short descriptions), and the content of every issue and issue comment 
+and stored in a relational database.
+
+Step 2. Extracting Exception Information from Issues. In this step, we used the ExceptionMiner tool,
+ developed in this work, to extract the exception names and exception stack traces of each
+ issue and issue comment - analyzing the information of each project at a time. This initial 
+set of analyzed Java projects included libraries, web-based, Android and stand-alone applications.
+
+Step 3. Restricting the Dataset to Android Projects. We select a subset of all Android 
+projects available on the dataset to perform a deeper analysis (which combined mined
+ information extrated from issues with source-code analysis and manual inspection 
+of projects resources described on next steps). This decision was driven by the
+fact that most open source Anedroid applications are known for dealing
 with several sources of exceptions due to multi-threaded execution and
@@
 well-defined platform and a limited set of libraries.
+To identify Android projects, we performed a case insensitive search for 
+the term \textsf{android} in the repository's names and short descriptions.  The
+heuristic filtered 2.542 repositories, from which 589 apps had at least one
+issue containing a stack trace. Then we performed further cleanup, inspecting the site of every Android
+reporting at least one stack trace, to make sure that they represented real
+mobile apps. During this clean up 106 apps were removed because they were either
+example projects (i.e., toy projects) or tools to support Android development
+(e.g. \textsf{selendroid}, \textsf{roboeletric} --- tools to support the testing of Android apps).
+The filtered set consisted of 482 apps, from which approximately 50\% are also
+hosted in Google Market Place.
+
+Step 4. Combining Information Mined From Issues with Bytecode-Code Analysis.
+
+Step 5
+. Combining Information Mined From Issues with Source-Code Analysis.
+For each one of the analyzed projects, we
+downloaded the source code and using custom scripts, we extracted their package
+hierarchy (recursive list of of Java packages) and their custom exception types
+as described next.
+
+Step 5. Combining Information Mined From Issues with Manual Inspection.
 

### 2014-07-07 android-stacks.tex
changing title. testing commit wih a different user
changes on Introduction

@@
 \begin{document}
-\title{What Can Stack Traces Tell us about the Use of Exceptions in Java?}
+\title{What Can Stack Traces Tell us about Exceptions in Java?}
 
@@
 source of information for the "state of practice" on software development
-which had already started to be investigated [][][][][].
+which had already started to be investigated [][][][][][I still need to collect these references].
 
@@
 
-Among these tools there is the Github issue tracking system, which makes
-issues related to new features, improvements, bugs publicly available.
-
-Bug issues have called the attention at have been the target of several 
-researchs [][][]. A subtype of a bug issue are the ones related to system 
-crashes [complete]...
+Bug issues have called the attention of researhc community and  have been the target of several 
+researches [][][]. A subtype of a bug issue are the ones related to system 
+crashes, also called crash reports such reports can be automatically generated 
+as the system crashes or can be manually reported by developers/users [complete]...
 

### 2014-07-07 android-stacks.tex
more changes on Introduction

@@
 
-In [ref] defined a set of characteristics a good bug issue should have. They showed that among 
-the most important relevant information are: (1) steps to reproduce the bug; 
-(2) and the failure stack trace.
-
-A typical stack trace consists of an ordered list of methods that
-were active on the call stack before the failure occurred. Stack traces
-are a useful source of information about system crashes, and are often used to
+One of the most important elements of a crash report genereted eigther manually
+or automatically is the failure stack trace [ref-paper], which tipically consists 
+of an ordered list of methods that were active on the call stack before the failure occurred.
+
+Modern applications have to cope with similar and increasing number of sources of failures
+ that arise as a consequence of faults in the application
+itself (e.g., access of null references), noisy user inputs or faults in
+underlying middleware or hardware.   The exception handling 
+mechanism~\cite{goodenough1975exception}, been embedded in several 
+programming languages (e.g. Java, C++, C\#), is one of the most used
+ schemes for detecting and recovering from such failures.
+
+Such exception stack traces are a useful source of 
+information about system crashes, and are often used to
 support developers in debugging~\cite{schroter2010stack}. Moreover, information
@@
 
-We could observe that existing research has focused on the analysis of a 
+However existing research has focused on the analysis of a 
 single system at a time. Moreover, the search facilities of repositories are very limited, also
 enabling the analysis of a single project.
-
-Modern applications have to cope with similar and increasing number of abnormal
-computation states that arise as a consequence of faults in the application
-itself (e.g., access of null references), noisy user inputs or faults in
-underlying middleware or hardware. The exception handling mechanism~\cite{goodenough1975exception} is one of the most used schemes for detecting and recovering from such exceptional conditions.
 
@@
 
------
-Along the recent years we have witnessed a rapid increase in 
-the number of social applications on the Web. Applications which 
-allow their users to track the activities of a large number 
-of others regardless of his/her location and affiliation.
-
-This tendency can also be seen on the software development context 
-where Github emerged and became an extensively used social coding 
-platform. Github alows developers to create and maintain opensource 
-projects relying  on Git distributed version control, and provides a set 
-of tools which allow the information about developers and their activities
- to be visible within and across open souce projects.
-
-Among those tools there is the Github issue tracking system, which makes
-issues related to new, features, bugs and general comments publicly available.
-
------
+....... OLD VERSION ......
 

### 2014-07-07 android-stacks.tex
even more changes on Introduction

@@
 
-Bug issues have called the attention of researhc community and  have been the target of several 
+Bug issues have called the attention of research community and have been the target of several 
 researches [][][]. A subtype of a bug issue are the ones related to system 
@@
 
-Modern applications have to cope with similar and increasing number of sources of failures
- that arise as a consequence of faults in the application
-itself (e.g., access of null references), noisy user inputs or faults in
-underlying middleware or hardware.   The exception handling 
-mechanism~\cite{goodenough1975exception}, been embedded in several 
-programming languages (e.g. Java, C++, C\#), is one of the most used
- schemes for detecting and recovering from such failures.
-
-Such exception stack traces are a useful source of 
+Such failure stack traces are a useful source of 
 information about system crashes, and are often used to
@@
 
-However existing research has focused on the analysis of a 
-single system at a time. Moreover, the search facilities of repositories are very limited, also
-enabling the analysis of a single project.
+The exception handling mechanism~\cite{goodenough1975exception},
+ is one of the most used schemes for detecting and recovering from such 
+failures. Exceptions may arrise as a consequence of faults in the application
+itself (e.g., access of null references), noisy user inputs or faults in
+underlying middleware or hardware.
+
+Although moder application have to deal with a similar and increasing number of 
+source of failures, existing research has focused on the analysis of a 
+single system at a time.
 

### 2014-07-07 android-stacks.tex
improving process description

@@
 on open-source repositories such as GitHub, some questions arise: What are the common
-characteristics of Github crash reports? How can we benefit from such cross-project crash analysis?
+characteristics of Github crash reports?
 
@@
 
-%\subsubsection{Exception Throwing and Handling}
-
-In Java an exception can be thrown in one of the following
+Hence, in Java an exception can be thrown in one of the following
 circumstances~\cite{gosling2000java}: (i) explicitly thrown when a throw statement is executed; 
@@
 
-%When an exception is signaled it causes a transfer of control from the point
-%where the exception occurred to a point that can be specified by the programmer
-%(exception handler) - in Java it is represented by the try-catch block .
-%The search for the handler to deal with a raised
-%exception occurs along the dynamic invocation chain. 
-%If no handler is found in the chain, the default handler associated to the 
-%Thread or ThreadGroup is invoked, which usually terminates the system 
-%and prints the exception stack trace.
-
-%\subsubsection{Exception Chaining} Exception Wrapping 
-%This stack trace was extracted from an issue of an Android application in
-%Github.
-
-The Exception wrapping (also known as exception chaining) is a common way of 
-propagating exceptions in Java programs, 
-which allows one exception to be wrapped in another exception and re-thrown. 
+When an exception is throw it causes a transfer of control from the point
+where the exception occurred to a point that can be specified by the programmer
+(exception handler) - in Java it is represented by the try-catch block .
+
+A common way of  propagating exceptions in Java programs, is the exception wrapping (also known as exception chaining), which allows one exception to be wrapped in another exception and re-thrown. 
 Figure~\ref{fig:wrapping} presents a stack trace associated which illustrates an
@@
 of the use of exceptions in Java programs based mainly on the information
-available on exception stack traces. More specifically, we aimed at answering
+available on exception stack traces. The exceptions available on exception stack trace
+are not limited to the exceptions expliciltly thrown by the Java developer but also include
+the exceptions implicitly thrown by the JVM during the program execution,
+due to a programming error (e.g., out-of-bounds array index, division-by-zero, 
+access to a null reference) or an internal error or resource limitation (e.g.,
+OutofMemoryError).
+
+ More specifically, we aimed at answering
 the following research questions:
@@
 
-  \item[RQ1] What is the prevalence of stack traces on  reported issues?
-    %\item[RQ2] What are the most common exceptions?
-
-  \item[RQ2] What are the common characteristics of such stack traces?
-
-  \item[RQ3] What patterns of exception (mis)-use emerge from the stack trace
+  \item[RQ1] What are the common characteristics of such stack traces?
+
+  \item[RQ2] What patterns of exception (mis)-use emerge from the stack trace
     analysis?
@@
 
-For RQ1 and RQ2, we explore the domain quantitatively, and use their answers as
-a basis to investigate the third research question. RQ3 aims at examining whether
-exception (mis)-uses could be revealed through stack trace analysis. To
-support this investigation, source code and bytecode analysis was used to
-leverage the  understanding of stack traces. 
+For RQ1 and RQ2, we explore the domain quantitatively and highlight interesting cases by 
+exploring cases qualitatively. To support this investigation, source code and bytecode 
+analysis were used in combination 
+with manual inspection to to everage the  understanding of stack traces.
+
 %For RQ3 we investigate
@@
 
-\subsection{Github Data}
+\subsection{Study Context and Github Data}
 \label{sec:git}
+
+ What is the prevalence of stack traces on reported issues?
+   What are the most common exceptions?
 
@@
 or computations that may fail) could improve the robustness of Java programs.
+
+TO DISCUSS: How can we benefit from such cross-project crash analysis?
 

### 2014-07-07 android-stacks.tex
more changes process description - section III and subsections

@@
 integrating both quantitative and qualitative evidence along the stages of the
-research process. The main goal of this study was to gain a better understanding
+research process.
+
+The main goal of this study was to gain a better understanding
 of the use of exceptions in Java programs based mainly on the information
-available on exception stack traces. The exceptions available on exception stack trace
-are not limited to the exceptions expliciltly thrown by the Java developer but also include
+available on exception stack traces embedded on Github issues.
+The exceptions available on exception stack trace are not limited to the exceptions 
+expliciltly thrown by the Java developer but also include
 the exceptions implicitly thrown by the JVM during the program execution,
@@
 exploring cases qualitatively. To support this investigation, source code and bytecode 
-analysis were used in combination 
-with manual inspection to to everage the  understanding of stack traces.
-
-%For RQ3 we investigate
-%quantitatively and highlight interesting cases by exploring cases qualitatively.
+analysis were used in combination with manual inspection to to everage the  understanding of stack traces.
+
+\subsection{Github Data}
+\label{sec:git}
+
+As mentioned before, this study amined at mining information available on GitHub issues,
+more specifically, extracting and distilling the stack traces embedded on GItHub issues.
+
+Issues on Github are different from issues on dedicated bug tracking tools such as 
+Bugzilla and Jira. The most important difference is that there are no predefined fields
+  (e.g. severity and priority). Instead, Github uses a more open ended tagging system, on which
+repositories are offered a pre-defined set of labels, but repository owners can modify 
+them at will. Therefore, an issue may have none ore an arbitrary set of labels depending 
+on which repository it was created.
+
+Therefore, this study is based on the assumption if an stack trace information embedded
+ in an issue, it contain relevant information concerning the exception use -  regardless the
+ lables associated to the issue.
+
+Moreover, issues and pull requests are dual on Github; all pull requests have a corresponding 
+``backing'' issue which
+are automatically generated. Therefore, we excluded such automatically generated
+issues from our analysis. Finally, the Github API allows the automated
+generation of issues for specific repositories, which automated tools often use
+to report crashes. In some cases, this led to high number of issues that
+included stack traces. We identified and filtered those cases out (e.g.,
+the~\textsf{pullwifi} project was responsible for almost 50\% of all Android issues in our dataset).
+
+This study used the dataset provided by the GHTorrent project~\cite{Gousi13}, an off-line mirror of the data 
+offered through the Github API.  GHTorrent has been collecting data from Git since 
+February 2012. Up to Dec 2013, and when we queried GHTorrent, it included 332,864 non-fork Java
+repositories \footnote {A Java project is specified in GitHub with a
+specific label.}  from which only 44,323 have contained at least issue report in their lifetime.
+From this set only 16,837 were watched by external users.
+To ensure that our work targets projects that are openly used by
+users other than their developers, we focused our analysis on just those projects.
+
+From this set of 16.836 Java projects which overall contained 356.057 issues, we retrieved the metadata related to each Java project (e.g., repository's names and short descriptions), and the content of every issue and issue comment and stored in a relational database.
+
+Using the heuristics presented in Section~\ref{sec:android}, we selected 482 repositories featuring Android projects on which we perform manual inspection was performed.
+
+\subsection{Study Context}
+
+Before investigating the research questions listed above we analyzed the prevalence of stack traces 
+on issues created on Github. Prevalence or prevalence proportion, a term often used in epidemiology,
+represents the proportion of a population found to have a particular condition
+(e.g. a disease) over the total number of the studied population. In our work, we used
+this term to evaluate a condition (i.e., presence of a stack trace) over issues
+reported on GitHub repositories.
+
+\input{stacktrace-stats}
+
+\begin{table}
+\centering
+\begin{tabular}{p{1em}lrr}
+\hline
+ \multicolumn{2}{l}{\bfseries{Java repository-level metrics}} & \\
+\hline
+\multicolumn{2}{l}{\bfseries{Repositories}} & 16,836\\
+    & incl. exception on 1+ issues & 4,776 (28,37\%) \\
+    & incl. stack trace on 1+ issues & 3,758 (22,32\%) \\
+    & incl. exception on 1+ defect issues & 2,019 (41,44\%)\\
+    & incl. stack trace on 1+ defect issues  & 1,698 (34,84\%) \\
+
+\hline
+\end{tabular}
+\caption{Projects x Stack Traces}
+\label{tab:proj}
+\end{table}
+
+\begin{table}
+\centering
+\begin{tabular}{p{1em}lrr}
+\hline
+\multicolumn{2}{l}{\bfseries{Issue-level metrics}} & \bfseries{Java} & \bfseries{Android Subset}\\
+\hline
+   \multicolumn{2}{l}{\bfseries{Repositories}} & 16,836 & 482\\
+ \multicolumn{2}{l}{\bfseries{Issues}} & 356,057 & 32,582\\
+    & incl. exceptions & 30,236 (8,5\%) & 3,956 (12,1\%) \\
+    & incl. stack traces & 21,013 (5,9\%) & 3,101 (7,3\%) \\
+ \multicolumn{2}{l}{\bfseries{Defect Issues}} & 55,226 & 2,376 \\
+    & incl. exception & 6,741 (12,2\%) & 1,557 (65,6\%) \\
+    & incl. stack trace &  5,196 (9,4\%) & 1,393 (58,7\%)\\
+\multicolumn{2}{l}{\bfseries{Exception Stack Traces}} &  & \\
+    & on issues  & 28,800 & 4,208  \\
+    & on issues labeled as defect & 6,529 & 1,699 \\
+\hline
+\end{tabular}
+\caption{Quantitative overview of the analyzed dataset}
+\label{tab:overview}
+\end{table}
+
+%The use of the Github issue tracker in repositories hosted on Github is optional
+%and not widespread. As of March 2014, from the 332,864 non-fork Java
+%repositories registered in the GHTorrent dataset, only 44,323 have received an
+%issue report in their lifetime while from those only 16,837 are being watched by
+%external users. To ensure that our work targets projects that are openly used by
+%users other than their developers, we focused our analysis on just those
+%projects. Using the heuristics presented in Section~\ref{sec:android}, we
+%selected 482 repositories featuring Android projects. Table~\ref{tab:proj} and
+%Table~\ref{tab:overview} presents two views of the extracted dataset, namely
+%grouped by repository and grouped by issue.
+%
+From Table~\ref{tab:overview}, we can see that repositories that include at
+least one exception or stack trace on an issue are fairly common: 28\% and 22\%
+respectively. Considering only the exceptions and stack traces reported on
+defect issues we find out that 41\% of the repositories reported at least one
+exception in a defect issue, and approximately 35\% of the projects mentioned
+one stack trace in at least one defect issue.
+
+At the individual issue level, 8,5\% and 5,9\% of all issues reported in Java
+repositories contain the full name of an exception or a stack trace
+respectively. The numbers are similar if we isolate Android projects from the
+general population (12,1\% and 7,3\%). Considering only the defect-issues
+however, in the Android repositories as almost 58,7\% of defect-issues include a
+stack trace. This is not true for the general case however.
+
+%Based on the assumption presented in Section III, we therefore did not use this label for further filtering.
+
+%\noindent \fbox{
+%\begin{minipage}{0.96\columnwidth} \emph{ RQ1: 28,37\% of Java
+%  repositories on Github have at least 1 issue that includes a stack trace.
+%  9,4\% of the defect issues of such repositories mentioned at least one stack
+%  trace. Considering only defect-issues of Android subset 58,7\% mentioned at
+%  least 1 stack trace.  }
+%\end{minipage}}
+
+\subsection{The Mining Process}
+\label{sec:miningproc}
+
 Figure~\ref{fig:overviewfig} presents an overview of the mixed-methods approach
@@
 
-\subsection{Study Context and Github Data}
-\label{sec:git}
-
- What is the prevalence of stack traces on reported issues?
-   What are the most common exceptions?
-
-\noindent\emph{Kinds of Analyzed Issues.} Issues on Github are different from
-issues on dedicated bug tracking tools such as Bugzilla and Jira. The most
-important difference is that there are no predefined fields for bug reporters to
-select from (e.g. severity and priority); Github uses instead a more open ended
-tagging system. Repositories are offered a pre-defined set of labels, but
-repository owners can modify them at will. An issue can have an arbitrary set of
-labels attached. Hence, this study does not restrict the analysis of stack
-traces defined on defect-issues, based on the assumption that regardless of the
-type of issue where its is defined the stack trace information contain relevant
-information concerning the exception use. Moreover, issues and pull requests are
-dual on Github; all pull requests have a corresponding ``backing'' issue which
-are automatically generated. Therefore, we excluded pull request based
-issues from our analysis. Finally, the Github API allows the automated
-generation of issues for specific repositories, which automated tools often use
-to report crashes. In some cases, this led to high number of issues that
-included stack traces. We identified and filtered those cases out (e.g.,
-the~\textsf{pullwifi} project was responsible for almost 50\% of all Android issues in our dataset).
+\subsection{The ExceptionMiner Tool}
+\label{sec:exceptionminer}
+
+To extract exceptions from issues, we implemented ExceptionMiner, a modular
+mining tool able to connect to various repositories (such as Google Code,
+GHTorrent or even directly to Bugzilla), extract issues, mine stack traces from
+them and classify them in predefined categories. The main components of
+ExceptionMiner are as follows:
+
+\noindent\emph{StackTraceMiner and Distiller} The first step in the process is
+mining references to exceptions and stack traces embedded in issues, distilling
+the information that composes a stack trace and storing the results in a
+relational database. Some of the attributes extracted from the stack traces were:
+ the exception signaler, the root cause, and the root cause wrappers. 
+The tool is based on a combination between a regular expression based parser 
+with heuristics able to identify and filter exception names and stack traces inline with text. In
+contrast to existing issue parsing solutions such as Infozilla, the parser
+created in this work can extract all causes related to an exception on a stack,
+and stack traces embedded in logs files.
+
+%\begin{tiny}
+%\begin{verbatim}
+
+%03-01 W/dalvikvm(7924): threadid=1: thread exiting with uncaught exception (group=0x40adf210)
+%03-01 15:55:01.609 (7924): FATAL EXCEPTION: main
+%03-01  (7924): java.lang.RuntimeException: ...
+%E/AndroidRuntime( 7924):  at android....performLaunchActivity(ActivityThread.java:1967)
+%03-01 15:55:01.609 (7924):  at android....handleLaunchActivity(ActivityThread.java:1992)
+%03-01 15:55:01.609:  at android.app.ActivityThread.access\$600(ActivityThread.java:127)
+%E/AndroidRuntime( 7924):  at android....handleMessage(ActivityThread.java:1158)
+%\end{verbatim}
+%\end{tiny}
+
+\noindent\emph{Exception Classifier} The next step in the ExceptionMiner process
+is classifying exceptions as either checked exceptions, runtime exceptions or 
+errors. The exception classifier is based on bytecode analysis (based on Design
+Wizard~\cite{Brunet09}) that walks up the type hierarchy of a Java exception
+until it reaches a base exception type.
+
+\noindent\emph{Exception Signaler Classifier} After its initialization it
+classifies each signaler according to its origin (i.e. Android Plaform,
+Aplication, Library, Libcore, or Java) based on pattern matching between the
+signaler name and the packages associated to each category.
+
+%\section{Study Operation and Results}
+Next section presents the results for each of the study phases, providing both a
+quantitative and qualitative analysis of the outcomes.
 
@@
 
-\subsection{The ExceptionMiner Tool}
-\label{sec:exceptionminer}
-
-To extract exceptions from issues, we implemented ExceptionMiner, a modular
-mining tool able to connect to various repositories (such as Google Code,
-GHTorrent or even directly to Bugzilla), extract issues, mine stack traces from
-them and classify them in predefined categories. The main components of
-ExceptionMiner are as follows:
-
-\noindent\emph{StackTraceMiner and Distiller} The first step in the process is
-mining references to exceptions and stack traces embedded in issues, distilling
-the information that composes a stack trace and storing the results in a
-relational database. Some of the attributes extracted from the stack traces were:
- the exception signaler, the root cause, and the root cause wrappers. 
-The tool is based on a combination between a regular expression based parser 
-with heuristics able to identify and filter exception names and stack traces inline with text. In
-contrast to existing issue parsing solutions such as Infozilla, the parser
-created in this work can extract all causes related to an exception on a stack,
-and stack traces embedded in logs files.
-
-%\begin{tiny}
-%\begin{verbatim}
-
-%03-01 W/dalvikvm(7924): threadid=1: thread exiting with uncaught exception (group=0x40adf210)
-%03-01 15:55:01.609 (7924): FATAL EXCEPTION: main
-%03-01  (7924): java.lang.RuntimeException: ...
-%E/AndroidRuntime( 7924):  at android....performLaunchActivity(ActivityThread.java:1967)
-%03-01 15:55:01.609 (7924):  at android....handleLaunchActivity(ActivityThread.java:1992)
-%03-01 15:55:01.609:  at android.app.ActivityThread.access\$600(ActivityThread.java:127)
-%E/AndroidRuntime( 7924):  at android....handleMessage(ActivityThread.java:1158)
-%\end{verbatim}
-%\end{tiny}
-
-\noindent\emph{Exception Classifier} The next step in the ExceptionMiner process
-is classifying exceptions as either checked exceptions, runtime exceptions or 
-errors. The exception classifier is based on bytecode analysis (based on Design
-Wizard~\cite{Brunet09}) that walks up the type hierarchy of a Java exception
-until it reaches a base exception type.
-
-\noindent\emph{Exception Signaler Classifier} After its initialization it
-classifies each signaler according to its origin (i.e. Android Plaform,
-Aplication, Library, Libcore, or Java) based on pattern matching between the
-signaler name and the packages associated to each category.
-
-%\section{Study Operation and Results}
-Next section presents the results for each of the study phases, providing both a
-quantitative and qualitative analysis of the outcomes.
-
 %\subsection{Phase 1: Analyzing Pure Stack Trace Information of Java Programs}
@@
 \section{RQ1: What is the prevalence of stack traces on reported issues?}
-
-Prevalence or prevalence proportion, a term often used in epidemiology,
-represents the proportion of a population found to have a particular condition
-(e.g. a disease) over the total number of the studied population. In our work, we used
-this term to evaluate a condition (i.e., presence of a stack trace) over issues
-reported on GitHub repositories.
-
-\input{stacktrace-stats}
-
-\begin{table}
-\centering
-\begin{tabular}{p{1em}lrr}
-\hline
- \multicolumn{2}{l}{\bfseries{Java repository-level metrics}} & \\
-\hline
-\multicolumn{2}{l}{\bfseries{Repositories}} & 16,836\\
-    & incl. exception on 1+ issues & 4,776 (28,37\%) \\
-    & incl. stack trace on 1+ issues & 3,758 (22,32\%) \\
-    & incl. exception on 1+ defect issues & 2,019 (41,44\%)\\
-    & incl. stack trace on 1+ defect issues  & 1,698 (34,84\%) \\
-
-\hline
-\end{tabular}
-\caption{Projects x Stack Traces}
-\label{tab:proj}
-\end{table}
-
-\begin{table}
-\centering
-\begin{tabular}{p{1em}lrr}
-\hline
-\multicolumn{2}{l}{\bfseries{Issue-level metrics}} & \bfseries{Java} & \bfseries{Android Subset}\\
-\hline
-   \multicolumn{2}{l}{\bfseries{Repositories}} & 16,836 & 482\\
- \multicolumn{2}{l}{\bfseries{Issues}} & 356,057 & 32,582\\
-    & incl. exceptions & 30,236 (8,5\%) & 3,956 (12,1\%) \\
-    & incl. stack traces & 21,013 (5,9\%) & 3,101 (7,3\%) \\
- \multicolumn{2}{l}{\bfseries{Defect Issues}} & 55,226 & 2,376 \\
-    & incl. exception & 6,741 (12,2\%) & 1,557 (65,6\%) \\
-    & incl. stack trace &  5,196 (9,4\%) & 1,393 (58,7\%)\\
-\multicolumn{2}{l}{\bfseries{Exception Stack Traces}} &  & \\
-    & on issues  & 28,800 & 4,208  \\
-    & on issues labeled as defect & 6,529 & 1,699 \\
-\hline
-\end{tabular}
-\caption{Quantitative overview of the analyzed dataset}
-\label{tab:overview}
-\end{table}
-
-%The use of the Github issue tracker in repositories hosted on Github is optional
-%and not widespread. As of March 2014, from the 332,864 non-fork Java
-%repositories registered in the GHTorrent dataset, only 44,323 have received an
-%issue report in their lifetime while from those only 16,837 are being watched by
-%external users. To ensure that our work targets projects that are openly used by
-%users other than their developers, we focused our analysis on just those
-%projects. Using the heuristics presented in Section~\ref{sec:android}, we
-%selected 482 repositories featuring Android projects. Table~\ref{tab:proj} and
-%Table~\ref{tab:overview} presents two views of the extracted dataset, namely
-%grouped by repository and grouped by issue.
-%
-From Table~\ref{tab:overview}, we can see that repositories that include at
-least one exception or stack trace on an issue are fairly common: 28\% and 22\%
-respectively. Considering only the exceptions and stack traces reported on
-defect issues we find out that 41\% of the repositories reported at least one
-exception in a defect issue, and approximately 35\% of the projects mentioned
-one stack trace in at least one defect issue.
-
-At the individual issue level, 8,5\% and 5,9\% of all issues reported in Java
-repositories contain the full name of an exception or a stack trace
-respectively. The numbers are similar if we isolate Android projects from the
-general population (12,1\% and 7,3\%). Considering only the defect-issues
-however, in the Android repositories as almost 58,7\% of defect-issues include a
-stack trace. This is not true for the general case however.
-
-%Based on the assumption presented in Section III, we therefore did not use this label for further filtering.
-
-%\noindent \fbox{
-%\begin{minipage}{0.96\columnwidth} \emph{ RQ1: 28,37\% of Java
-%  repositories on Github have at least 1 issue that includes a stack trace.
-%  9,4\% of the defect issues of such repositories mentioned at least one stack
-%  trace. Considering only defect-issues of Android subset 58,7\% mentioned at
-%  least 1 stack trace.  }
-%\end{minipage}}
 

### 2014-07-07 android-stacks.tex
restucturing process section in Github Data -  Stack traces on issues - Android Seclection

@@
 
-\subsection{Study Context}
+\subsection{Stack Traces on Issues}
 
@@
 
-\input{stacktrace-stats}
-
-\begin{table}
-\centering
-\begin{tabular}{p{1em}lrr}
-\hline
- \multicolumn{2}{l}{\bfseries{Java repository-level metrics}} & \\
-\hline
-\multicolumn{2}{l}{\bfseries{Repositories}} & 16,836\\
-    & incl. exception on 1+ issues & 4,776 (28,37\%) \\
-    & incl. stack trace on 1+ issues & 3,758 (22,32\%) \\
-    & incl. exception on 1+ defect issues & 2,019 (41,44\%)\\
-    & incl. stack trace on 1+ defect issues  & 1,698 (34,84\%) \\
-
-\hline
-\end{tabular}
-\caption{Projects x Stack Traces}
-\label{tab:proj}
-\end{table}
-
-\begin{table}
-\centering
-\begin{tabular}{p{1em}lrr}
-\hline
-\multicolumn{2}{l}{\bfseries{Issue-level metrics}} & \bfseries{Java} & \bfseries{Android Subset}\\
-\hline
-   \multicolumn{2}{l}{\bfseries{Repositories}} & 16,836 & 482\\
- \multicolumn{2}{l}{\bfseries{Issues}} & 356,057 & 32,582\\
-    & incl. exceptions & 30,236 (8,5\%) & 3,956 (12,1\%) \\
-    & incl. stack traces & 21,013 (5,9\%) & 3,101 (7,3\%) \\
- \multicolumn{2}{l}{\bfseries{Defect Issues}} & 55,226 & 2,376 \\
-    & incl. exception & 6,741 (12,2\%) & 1,557 (65,6\%) \\
-    & incl. stack trace &  5,196 (9,4\%) & 1,393 (58,7\%)\\
-\multicolumn{2}{l}{\bfseries{Exception Stack Traces}} &  & \\
-    & on issues  & 28,800 & 4,208  \\
-    & on issues labeled as defect & 6,529 & 1,699 \\
-\hline
-\end{tabular}
-\caption{Quantitative overview of the analyzed dataset}
-\label{tab:overview}
-\end{table}
-
-%The use of the Github issue tracker in repositories hosted on Github is optional
-%and not widespread. As of March 2014, from the 332,864 non-fork Java
-%repositories registered in the GHTorrent dataset, only 44,323 have received an
-%issue report in their lifetime while from those only 16,837 are being watched by
-%external users. To ensure that our work targets projects that are openly used by
-%users other than their developers, we focused our analysis on just those
-%projects. Using the heuristics presented in Section~\ref{sec:android}, we
-%selected 482 repositories featuring Android projects. Table~\ref{tab:proj} and
-%Table~\ref{tab:overview} presents two views of the extracted dataset, namely
-%grouped by repository and grouped by issue.
-%
-From Table~\ref{tab:overview}, we can see that repositories that include at
-least one exception or stack trace on an issue are fairly common: 28\% and 22\%
-respectively. Considering only the exceptions and stack traces reported on
-defect issues we find out that 41\% of the repositories reported at least one
-exception in a defect issue, and approximately 35\% of the projects mentioned
-one stack trace in at least one defect issue.
-
-At the individual issue level, 8,5\% and 5,9\% of all issues reported in Java
-repositories contain the full name of an exception or a stack trace
-respectively. The numbers are similar if we isolate Android projects from the
-general population (12,1\% and 7,3\%). Considering only the defect-issues
-however, in the Android repositories as almost 58,7\% of defect-issues include a
-stack trace. This is not true for the general case however.
-
-%Based on the assumption presented in Section III, we therefore did not use this label for further filtering.
-
-%\noindent \fbox{
-%\begin{minipage}{0.96\columnwidth} \emph{ RQ1: 28,37\% of Java
-%  repositories on Github have at least 1 issue that includes a stack trace.
-%  9,4\% of the defect issues of such repositories mentioned at least one stack
-%  trace. Considering only defect-issues of Android subset 58,7\% mentioned at
-%  least 1 stack trace.  }
-%\end{minipage}}
+Using the ExceptionMiner tool, developed in this work, we could extract the exception stack traces defined on issues and issue comments.
+Executing ExceptionMiner tool each issue and issue comment defined on the 16,836 projects considered in this study, we could observe that 3,758 projects (22,32\%) contained at least 1 issue on which a stack trace was found.
+
+Overall the number of issues on which a stack trace was found was 21,013 from which 28,800 stack traces were extracted and analyzed - some issues defined more than one stack trace.
+
+We then analyzed the lables associated to such issues, and we discovered that only 5,196 were labled as defect ~\footnote{the dedect lables considered in this work were every lable containing as substring: defect, bug, crash, failure, fail, exception} (which represent only 24,7% of the issues on which a stack trace was found), most of the issues on which stack traces were found contained no lable.
+Such observation supports our assumption to consider not only the issues related with defect issues on this analysis.
+
+\subsection{Android Apps Selection}
+\label{sec:android}
+
+As mentioned before, in this study we selected a subset of Java projects to
+be analyzed in more detail. The subset chosen was comprised by the Android apps
+included on GitHub (until 23 Feb 2013).
+To identify Android projects, we performed a case insensitive search for the
+term \textsf{android} in the repository's names and short descriptions.  The
+heuristic filtered 2.542 repositories, from which 589 apps had at least one
+issue containing a stack trace.
+
+Then we performed further cleanup, inspecting the site of every Android
+reporting at least one stack trace, to make sure that they represented real
+mobile apps. During this clean up 106 apps were removed because they were either
+example projects (i.e., toy projects) or tools to support Android development
+(e.g. \textsf{selendroid}, \textsf{roboeletric} --- tools to support the testing of Android apps).
+The filtered set consisted of 482 apps, from which approximately 50\% are also
+hosted in Google Market Place.
+
+This set of 482 projects contained overall 32,582 issues from which 4,208 stacktraces 
+were extracted. From this set of issues 44,9% were labled as defect.
 
@@
 
+For each one of the analyzed projects, we
+downloaded the source code and using custom scripts, we extracted their package
+hierarchy (recursive list of of Java packages) and their custom exception types
+as described next.
+
 Step 5
@@
 quantitative and qualitative analysis of the outcomes.
-
-\subsection{Android Apps Selection}
-\label{sec:android}
-
-%In the second phase of this study we selected a subset of Java applications to
-%be analyzed in more detail.The subset chosen was comprised by the Android apps
-%included on GitHub (until 23 Feb 2013).
-To identify Android projects, we performed a case insensitive search for the
-term \textsf{android} in the repository's names and short descriptions.  The
-heuristic filtered 2.542 repositories, from which 589 apps had at least one
-issue containing a stack trace.
-
-Then we performed further cleanup, inspecting the site of every Android
-reporting at least one stack trace, to make sure that they represented real
-mobile apps. During this clean up 106 apps were removed because they were either
-example projects (i.e., toy projects) or tools to support Android development
-(e.g. \textsf{selendroid}, \textsf{roboeletric} --- tools to support the testing of Android apps).
-The filtered set consisted of 482 apps, from which approximately 50\% are also
-hosted in Google Market Place. For each one of the analyzed projects, we
-downloaded the source code and using custom scripts, we extracted their package
-hierarchy (recursive list of of Java packages) and their custom exception types
-as described next.
 

### 2014-07-07 android-stacks.tex
mprovements on the Introduction

@@
 
-Bug issues have called the attention of research community and have been the target of several 
-researches [][][]. A subtype of a bug issue are the ones related to system 
-crashes, also called crash reports such reports can be automatically generated 
-as the system crashes or can be manually reported by developers/users [complete]...
-
-One of the most important elements of a crash report genereted eigther manually
-or automatically is the failure stack trace [ref-paper], which tipically consists 
-of an ordered list of methods that were active on the call stack before the failure occurred.
-
-Such failure stack traces are a useful source of 
+Bug issues have called the attention of the research community and have been the target of increasing 
+researches [][][]. A specific type of bug issue is the one related to system 
+crashes.  Such issues usually contain an exception stack trace [ref-paper], which tipically consists 
+of an ordered list of methods that were active on the call stack before the failure occurred [ref-bug-issues].
+
+Exception stack traces are a useful source of 
 information about system crashes, and are often used to
@@
 
-The exception handling mechanism~\cite{goodenough1975exception},
- is one of the most used schemes for detecting and recovering from such 
-failures. Exceptions may arrise as a consequence of faults in the application
+Although modern applications have to deal with increasinng and similar 
+sources of failures - that arise as a consequence of faults in the application
 itself (e.g., access of null references), noisy user inputs or faults in
-underlying middleware or hardware.
-
-Although moder application have to deal with a similar and increasing number of 
-source of failures, existing research has focused on the analysis of a 
-single system at a time.
-
-Although the exception handling mechanism has been embedded in several 
-programming languages (e.g. Java, C++, C\#) and was the target of several studies
-(e.g.~\cite{miller1997issues,Robil00,shah2010understanding, garcia2007extracting, garcia2001comparative,cabral2007exception,coelho2011unveiling}),
-the exception handling code is often generally poorly understood and the least
-tested part of software systems.
-
-Often exception handling constructions may lead the developers into believing that
-by just re-throwing and possibly logging the exceptions they can forget about the exceptional
-situations during the development of the "happy path". This "ignore-for-now"
-approach may turn the exception handling into a generalized "goto"
-mechanism~\cite{mandrioli1992advances} making the program more complex and even
-less reliable. This behaviour may lead to the problem of uncaught exceptions~\cite{jo2004uncaught}, one of the main causes of application crashes.
-
-Specially in the open-source environment, where faults are usualy not hidden in a private trackers,  it would be intresting if developers could learn from other projects mistakes.
-Moreover, interesting knowledge may emerge from a cross-project crash analysis.
-
-Hence, in such context where we have access to wealth of information available
-on open-source repositories such as GitHub, some questions arise: What are the common
-characteristics of Github crash reports?
-
-. The ExceptionMiner tool is a contribution in this direction (to enable a cross-project defect analysis).
-
-....... OLD VERSION ......
-
-Modern applications have to cope with an increasing number of abnormal
-computation states that arise as a consequence of faults in the application
-itself (e.g., access of null references), noisy user inputs or faults in
-underlying middleware or hardware. The exception handling
-mechanism~\cite{goodenough1975exception} is one of the most used schemes for
-detecting and recovering from such exceptional conditions. Although the
-exception handling mechanism has been embedded in several programming languages
-(e.g. Java, C++, C\#) and was the target of several studies
-(e.g.~\cite{miller1997issues,Robil00,shah2010understanding,
-garcia2007extracting,garcia2001comparative,cabral2007exception,coelho2011unveiling}),
-the exception handling code is often generally poorly understood and the least
-tested part of software systems.
-
-Often exception handling constructions may lead the developers into believing that
-by just re-throwing and possibly logging the exceptions they can forget about the exceptional
-situations during the development of the "happy path". This "ignore-for-now"
-approach may turn the exception handling into a generalized "goto"
-mechanism~\cite{mandrioli1992advances} making the program more complex and even
-less reliable. This behaviour may lead to the problem of uncaught exceptions~\cite{jo2004uncaught}, one of the main causes of application crashes.
-
-Moreover, some languages use exceptions to represent programming mistakes
-(array index out-of-bounds, division-by-zero, dereferencing null pointers). In these cases,
-exceptions are implicitly signaled by the runtime environment when such
-conditions occur. One of such implicit exceptions (caused by data conversion
-from a 64-bit floating point to a 16-bit signed integer) was responsible for
-the infamous failure on Ariane 5's first test flight~\cite{lions1996ariane} ---
-leading to the rocket self-destructing 37 seconds after launch and a loss of 500
-million dollars. Besides Ariane 5, several applications crash everyday due to
-uncaught exceptions~\cite{jo2004uncaught}. Not using exceptions carefully
-may have the opposite effect of the initial intention of improving
-system robustness.
-
-In Java, when the program fails due to an uncaught exception, it automatically
-terminates, while the system prints a stack trace to the console.
-
-A stack trace
-includes an explanatory message and the execution stack frame when the exception
-occurred. A typical Java stack trace consists of an ordered list of methods that
-were active on the call stack before the exception occurred. Stack traces
-are a useful source of information about system crashes, and are often used to
-support developers in debugging~\cite{schroter2010stack}. Moreover, information
-from stack traces combined with crash and bug reports can be used to support bug
-classification and clustering~\cite{wang2013improving, kim2011crash,
-dhaliwal2011classifying}, fault-proneness prediction
-models~\cite{kim2013predicting} and even automated bug
-fixing~\cite{sinha2009fault} tools.
-
-However, all existing research has focused on the analysis of a single system at a time. In
-our current context, on which we have access to wealth of information available
-on open-source repositories like GitHub additional questions arise: What is the
-prevalence of exception stack traces on reported issues? What are the common
-characteristics of such stack traces?
-
-%What the stack traces available in these repositories can tell us about the use
-%of exceptions in Java programs?
-
-%To guide the our study we also compile general guidelines on how to use
-%exceptions proposed by Gosling~\cite{gosling2000java},
-%Wirfs-Brock~\cite{wirfs2006toward} and Bloch~\cite{bloch2008effective} and then
-%we investigate: what stack traces can tell us about the adherence to such
-%practices in Java programs?
+underlying middleware or hardware -  on bug classificationthe current research on 
+exploring bug issue information are structured on  a system at a time basis.
+
+Specially  in a context where we have access to wealth of information available
+on open-source repositories such as GitHub, interesting knowledge may emerge 
+from a cross-project crash analysis. Hence such initial questions arise: What are the common
+characteristics of the stack traces reported on GitHub? Which patterns of exception (mis)use may emerge?
 
@@
 
-  \item  It performs the first large scale analysis of Java stack traces and how
+  \item  It performs a large scale analysis of Java stack traces and how
     they can reveal bad-practices on the use of exceptions in Java.
@@
 The main disadvantage of return codes is the fact that the programmer can just
-ignore the return code and let the program execute in inconsistent state. The
-exception mechanism~\cite{goodenough1975exception} was proposed as a way of
+ignore the return code and let the program execute in inconsistent state.
+
+The exception mechanism~\cite{goodenough1975exception} was proposed as a way of
 preventing errors from being ignored and was embedded in many mainstream
@@
 handled in a single system. Next section details Java exception handling model.
+
+.......
+
+Although exception handling mechanism is the most used technique to deal with such 
+failures - been embedded in several programming languages (e.g. Java, C++, C\#) -  
+the exception handling code is often generally poorly understood and the least
+tested part of software systems.This behaviour may lead to the problem of uncaught exceptions~\cite{jo2004uncaught}, one of the main causes of application crashes
+(e.g.~\cite{miller1997issues,Robil00,shah2010understanding, garcia2007extracting, garcia2001comparative,cabral2007exception,coelho2011unveiling}).
+
+........
 
@@
 more specifically, extracting and distilling the stack traces embedded on GItHub issues. 
-
 Issues on Github are different from issues on dedicated bug tracking tools such as 
@@
 them at will. Therefore, an issue may have none ore an arbitrary set of labels depending 
-on which repository it was created.
-
-Therefore, this study is based on the assumption if an stack trace information embedded
- in an issue, it contain relevant information concerning the exception use -  regardless the
- lables associated to the issue.
+on which repository it was created. Therefore, this study is based on the assumption 
+if an stack trace information embedded in an issue, it contain relevant information
+ concerning the exception use -  regardless the lables associated to the issue.
 
@@
 
-This study used the dataset provided by the GHTorrent project~\cite{Gousi13}, an off-line mirror of the data 
+\subsubsection{GHTorrent Dataset} This study used the dataset provided by the GHTorrent project~\cite{Gousi13}, an off-line mirror of the data 
 offered through the Github API.  GHTorrent has been collecting data from Git since 
@@
 From this set of 16.836 Java projects which overall contained 356.057 issues, we retrieved the metadata related to each Java project (e.g., repository's names and short descriptions), and the content of every issue and issue comment and stored in a relational database.
-
 Using the heuristics presented in Section~\ref{sec:android}, we selected 482 repositories featuring Android projects on which we perform manual inspection was performed.
 
-\subsection{Stack Traces on Issues}
+\subsubsection{Stack Traces on Issues}
 
@@
 Executing ExceptionMiner tool each issue and issue comment defined on the 16,836 projects considered in this study, we could observe that 3,758 projects (22,32\%) contained at least 1 issue on which a stack trace was found.
-
 Overall the number of issues on which a stack trace was found was 21,013 from which 28,800 stack traces were extracted and analyzed - some issues defined more than one stack trace.
 
-We then analyzed the lables associated to such issues, and we discovered that only 5,196 were labled as defect ~\footnote{the dedect lables considered in this work were every lable containing as substring: defect, bug, crash, failure, fail, exception} (which represent only 24,7% of the issues on which a stack trace was found), most of the issues on which stack traces were found contained no lable.
+We then analyzed the lables associated to such issues, and we discovered that only 5,196 were labled as defect ~\footnote{the dedect lables considered in this work were every lable containing as substring: defect, bug, crash, failure, fail, exception} (which represent only 24,7\% of the issues on which a stack trace was found), most of the issues on which stack traces were found contained no lable.
 Such observation supports our assumption to consider not only the issues related with defect issues on this analysis.
 
-\subsection{Android Apps Selection}
+\subsubsection{Android Apps Selection}
 \label{sec:android}
@@
 This set of 482 projects contained overall 32,582 issues from which 4,208 stacktraces 
-were extracted. From this set of issues 44,9% were labled as defect.
+were extracted. From this set of issues 44,9\% were labled as defect.
 
@@
 the information that composes a stack trace and storing the results in a
-relational database. Some of the attributes extracted from the stack traces were:
- the exception signaler, the root cause, and the root cause wrappers. 
+relational database. Some of the information extracted from the stack traces were:
+ the exception being thrown, its signaler, the exception root cause, its corresponding signaler,
+and packages related to each one of these fields.
 The tool is based on a combination between a regular expression based parser 
@@
 
-%\subsection{Phase 1: Analyzing Pure Stack Trace Information of Java Programs}
-
-\section{RQ1: What is the prevalence of stack traces on reported issues?}
-
-\section{RQ2: What are the common characteristics of such stack traces?}
+\section{RQ1: What are the common characteristics of such stack traces?}
 
@@
 
-\subsection{Top 10 most common exceptions in Java Projects}
-
-Table~\ref{tab:toptenjava} illustrates the top 10 most common root causes which
-represents more than 50\% of the analyzed set. The table illustrates  i) the
-number of times each exception appeared as the root cause reported on the exception stack trace
-(\#occurrences), ii) the number of distinct projects reporting such causes on
-stack traces, and iii) the exception popularity among projects.
+\subsection{Characterizing the Most Common Exceptions}
+
+After distilling the information available on the 28,800 stack traces were extracted from the 
+issues and issues comments of 16,836 Java projects: 607 distinct root causes were found, 
+and  2,026 distinct top level exceptions were found.
+
+From this set the java.lang.NullPointerException  was the root cause most 
+reported;  7578 stack traces had a java.lang.NullPointerException  as root cause,
+ which corresponds of 26,31\%  of all analyzed stack traces.
+
+If we perform a cross-project analysis, the java.lang.NullPointerException, we could observe that such 
+exception was reported on 1,671 projects (which represents 44,47% of all projects reporting at least one stack trace).
+
+The high prevalence of NullPointerExceptions is aligned with the findings other works~\cite{kim2013predicting,fraser20131600,csallner2004jcrasher}. For instance, Sunghun et
+al.~\cite{kim2013predicting} showed that in Eclipse bug report system 38\% the bugs related to exception handling were caused by NullPointerException; other works on robustness testing~\cite{maji2012empirical,csallner2004jcrasher} showed that most of the automatically detected bugs were due to NullPointerExceptions and implicitly-signaled of Java
+environment (as the ones found in this study).
+
+To better understand the root causes reported on stacks, we performed a categorization 
+study on which the top 100 root causes reported on stack-issues were manually inspected.
+The characterization process consisted of classifiying each exception in one of the
+chategories listed on Table X, based on the inspection of source code, Javadoc, 
+or additional documentation related to the exception.
 
 \begin{table}
-\begin{tabular}{lrrrr}
-    \hline
-    \bfseries{Exception} &  \multicolumn{2}{c}{\bfseries{Occurrences}} &  \multicolumn{2}{c}{\bfseries{Projects}}\\
-    & \bfseries{\#} &  \bfseries{\%} & \bfseries{\# } & \bfseries{\% } \\
-    \hline
-java.lang.NullPointerException             & 7578 & 26,31\% & 1671 & 44,47 \\
-java.lang.IllegalArgumentException         & 1573 & 5,46\%  & 670  & 17,83\% \\
-java.lang.ClassNotFoundException           & 1203 & 4,18\%  & 514  & 13,68\% \\
-java.lang.RuntimeException                 & 1093 & 3,80\%  & 411  & 10,94\% \\
-java.lang.IllegalStateException            & 964  & 3,35\%  & 431  & 11,47\% \\
-java.lang.ClassCastException               & 886  & 3,08\%  & 408  & 10,86\% \\
-java.lang.ArrayIndexOutOfBoundsException   & 834  & 2,90\%  & 345  & 9,18\% \\
-java.lang.NoSuchMethodError                & 822  & 2,85\%  & 369  & 9,82\% \\
-java.io.IOException                        & 656  & 2,28\%  & 337  & 8,97\% \\
-java.lang.OutOfMemoryError                 & 601  & 2,09\%  & 235  & 6,25\% \\
-%java.lang.NoClassDefFoundError            & 554  & 1,92\%  & 299  & 7,96\% \\
-%java.util.ConcurrentModificationException & 465  & 1,61\%  & 180  & 4,79\% \\
-%java.lang.IndexOutOfBoundsException       & 445  & 1,55\%  & 205  & 5,46\% \\
-\hline
-All	 &	23956 &	100\%	& 3758 & 100\%	 \\
+  \centering
+  \begin{tabular}{|p{2cm}| p{5cm}|}
+    \hline
+    \bfseries{Category} & \bfseries{Description} \\
+    \hline
+      Programming logic &  Exceptions implicitly  thrown by the JVM when the 
+evaluation of an expression violates the normal semantics of language (e.g., 
+out-of-bounds array index, division-by-zero, access to a null reference). 
+Such exceptions are usually defined on java.lang and java.util packages. \\ \hline
+      Resources (IO)                         & Exception releated to the manipulation of IO resources or device  (e.g., file, network) \\ \hline
+      Security                               & Exception releated to security issues (e.g. password validation, criptography) \\ \hline
+      Concurrency                            & Exceptions related to multi-threaded programming \\ \hline
+      Backward compatibility                 &  A list of exceptions related to backward compatibility - a complete list can be found here [ref]  \\ \hline
+      Reflection                             & Exceptions related to Java reflection library  \\ \hline
+      Specific                & Exceptions created for specific applications, frameworks or libraries \\ \hline
+      General (Exception, Error, Runtime)    & The top level exceptions in Java    \\ \hline
     \hline
   \end{tabular}
-\caption{Top 10 most popular exceptions on issues in Java projects}
-\label{tab:toptenjava}
+  \caption{Categories description.}
+  \label{tab:categories}
 \end{table}
 
-We can observe that the NullPointerException was the one with the higher number
-of occurrences (26,3\% of all stacks were caused by NullPointerExceptions) and
-the one which had the highest popularity across projects. Actually, most of the
-exceptions in the top 10 list (except java.lang.IllegalArgumentException,
-java.io.IOException and java.lang.RuntimeException) are exceptions implicitly
-thrown by Java Virtual Machine, when an expression violates the normal
-semantics of the Java programming language (e.g. ArrayIndexOutOfBounds) or due
-to an internal error or resource limitation (OutOfMemoryError).
-
-The high prevalence of NullPointerExceptions, and the other implicitly-thrown
-exceptions  is aligned with the findings other works~\cite{kim2013predicting,fraser20131600,csallner2004jcrasher}. For instance, Sunghun et
-al.~\cite{kim2013predicting} showed that in Eclipse bug report system 38\% the bugs related to exception handling were caused by NullPointerException; other works on robustness testing~\cite{maji2012empirical,csallner2004jcrasher}
-showed that most of the automatically
-detected bugs were due to NullPointerExceptions and implicitly-signaled of Java
-environment (as the ones found in this study).
-
-\subsection{Top 10 most common exceptions in Android Projects}
+The top 100 root causes correspond to approximately 77\% of all stack traces
+reported reported on Java stack-issues and 95\% of all exceptions reported on
+Android stack-issues. Table~\ref{tab:tophundrend} illustrates the result of such 
+classification.
+
+\begin{table}
+  \centering
+  \begin{tabular}{lrr}
+    \hline
+    \bfseries{Category} & \bfseries{Java} & \bfseries{Android Subset} \\
+    \hline
+      Programming logic (java.lang and util) & 12625  (41,10\%) & 2235 (55,64\%)\\
+      Resources (IO)                         & 12440 (40,50\%)  & 727 (18,10\%) \\
+      Security                               & 220  (0,72\%)    & 165 (4,11\%)\\
+      Concurrency                            & 633 (2,06\%)     & 116  (2,89\%)\\
+      Backward compatibility                 & 1580 (5,14\%)    & 219 (5,45\%) \\
+      Reflection                             & 1413 (4,60\%)    & 91 (2,27\%)\\
+      Specific (GUI,FRAMEWORK)               & 340 (1,11\%)     & 197 (4,90\%)\\
+      General (Error, Exception, Runtime)    & 1468 (4,78\%)    & 267 (6,65\%)\\
+    \hline
+  \end{tabular}
+  \caption{Characterization of the 100 root causes.}
+  \label{tab:tophundrend}
+\end{table}
+
+Programming errors are the causes of most of the exception stack traces reported
+on issues. In the Android subset we could find more issues related to Backward
+compatibility and security issues than on the general data set.
+
+\subsection{Investigating Android Subset}
 
@@
 Only high fault tolerant systems need to provide solutions to handle them.
-
-\subsection{Classification of the main causes of top exceptions}
-
-To better understand the main causes of such exceptions, we evaluated semantics
-related to the top 100 root causes reported on stack-issues of both Java dataset
-and Android subset (which correspond to approximately 77\% of all exceptions
-reported reported on Java stack-issues and 95\% of all exceptions reported on
-Android stack-issues), and classified them according to the the categories
-presented in Table~\ref{tab:tophundrend}.
-
-\begin{table}
-  \centering
-  \begin{tabular}{lrr}
-    \hline
-    \bfseries{Category} & \bfseries{Java} & \bfseries{Android Subset} \\
-    \hline
-      Programming logic (java.lang and util) & 12625  (41,10\%) & 2235 (55,64\%)\\
-      Resources (IO)                         & 12440 (40,50\%)  & 727 (18,10\%) \\
-      Security                               & 220  (0,72\%)    & 165 (4,11\%)\\
-      Concurrency                            & 633 (2,06\%)     & 116  (2,89\%)\\
-      Backward compatibility                 & 1580 (5,14\%)    & 219 (5,45\%) \\
-      Reflection                             & 1413 (4,60\%)    & 91 (2,27\%)\\
-      Specific (GUI,FRAMEWORK)               & 340 (1,11\%)     & 197 (4,90\%)\\
-      General (Error, Exception, Runtime)    & 1468 (4,78\%)    & 267 (6,65\%)\\
-    \hline
-  \end{tabular}
-  \caption{Characterization of the 100 root causes.}
-  \label{tab:tophundrend}
-\end{table}
-
-Programming errors are the causes of most of the exception stack traces reported
-on issues. In the Android subset we could find more issues related to Backward
-compatibility and security issues than on the general data set.
 
@@
 
-\section{RQ3.  What kinds of exception (mis)-use emerge from stack trace analysis? }
+\section{RQ2.  What kinds of exception (mis)-use emerge from stack trace analysis? }
 
@@
 
+\section {Old Introduction}
+
+....... OLD VERSION ......
+
+Modern applications have to cope with an increasing number of abnormal
+computation states that arise as a consequence of faults in the application
+itself (e.g., access of null references), noisy user inputs or faults in
+underlying middleware or hardware. The exception handling
+mechanism~\cite{goodenough1975exception} is one of the most used schemes for
+detecting and recovering from such exceptional conditions. Although the
+exception handling mechanism has been embedded in several programming languages
+(e.g. Java, C++, C\#) and was the target of several studies
+(e.g.~\cite{miller1997issues,Robil00,shah2010understanding,
+garcia2007extracting,garcia2001comparative,cabral2007exception,coelho2011unveiling}),
+the exception handling code is often generally poorly understood and the least
+tested part of software systems.
+
+Often exception handling constructions may lead the developers into believing that
+by just re-throwing and possibly logging the exceptions they can forget about the exceptional
+situations during the development of the "happy path". This "ignore-for-now"
+approach may turn the exception handling into a generalized "goto"
+mechanism~\cite{mandrioli1992advances} making the program more complex and even
+less reliable. This behaviour may lead to the problem of uncaught exceptions~\cite{jo2004uncaught}, one of the main causes of application crashes.
+
+Moreover, some languages use exceptions to represent programming mistakes
+(array index out-of-bounds, division-by-zero, dereferencing null pointers). In these cases,
+exceptions are implicitly signaled by the runtime environment when such
+conditions occur. One of such implicit exceptions (caused by data conversion
+from a 64-bit floating point to a 16-bit signed integer) was responsible for
+the infamous failure on Ariane 5's first test flight~\cite{lions1996ariane} ---
+leading to the rocket self-destructing 37 seconds after launch and a loss of 500
+million dollars. Besides Ariane 5, several applications crash everyday due to
+uncaught exceptions~\cite{jo2004uncaught}. Not using exceptions carefully
+may have the opposite effect of the initial intention of improving
+system robustness.
+
+In Java, when the program fails due to an uncaught exception, it automatically
+terminates, while the system prints a stack trace to the console. A stack trace
+includes an explanatory message and the execution stack frame when the exception
+occurred. A typical Java stack trace consists of an ordered list of methods that
+were active on the call stack before the exception occurred. Stack traces
+are a useful source of information about system crashes, and are often used to
+support developers in debugging~\cite{schroter2010stack}. Moreover, information
+from stack traces combined with crash and bug reports can be used to support bug
+classification and clustering~\cite{wang2013improving, kim2011crash,
+dhaliwal2011classifying}, fault-proneness prediction
+models~\cite{kim2013predicting} and even automated bug
+fixing~\cite{sinha2009fault} tools.
+
+However, all existing research has focused on the analysis of a single system at a time. In
+our current context, on which we have access to wealth of information available
+on open-source repositories like GitHub additional questions arise: What is the
+prevalence of exception stack traces on reported issues? What are the common
+characteristics of such stack traces?
+
+%What the stack traces available in these repositories can tell us about the use
+%of exceptions in Java programs?
+
+%To guide the our study we also compile general guidelines on how to use
+%exceptions proposed by Gosling~\cite{gosling2000java},
+%Wirfs-Brock~\cite{wirfs2006toward} and Bloch~\cite{bloch2008effective} and then
+%we investigate: what stack traces can tell us about the adherence to such
+%practices in Java programs?
+
+To answer these questions, we conduct an exploratory study of stack traces in
+the wild. Using a custom tool called ExceptionMiner, which was specifically
+developed for this study, we mined stack traces from issues across all Java
+projects on GitHub. Overall, we analyzed 356,057 issues from 16,836 projects,
+from which 28,800 stack traces were extracted and used as a source of
+information for our analysis. The stack trace analysis  was augmented by
+additional information extracted using byte code and source code analysis done
+on a carefully selected subset of our initial sample (482 Android projects).
+%This study was divided in two phases: in the fist phase all the issues of all
+%Java projects available on GitHub were analyzed and the embedded stack traces
+%extracted using ExceptionMiner.
+
+To guide this exploration we compiled general guidelines on how to use
+exceptions proposed by Gosling~\cite{gosling2000java},
+Wirfs-Brock~\cite{wirfs2006toward} and Bloch~\cite{bloch2008effective} and then
+we investigate what stack traces can tell us about the adherence to such
+practices in Java programs.
+%~\todo{This should be our motivation, as this is
+%our biggest contribution. I already changed the text. Lets discuss about this
+%new version}
+%RSC: I AGREE.
+
+%GGG: I think this is too much detail for the introduction In the second phase
+%of this study, we combined the information of the mined stack traces with
+%source code and bytecode analysis. To do so we selected the subset of 482
+%Android repositories. The Android repositories were chosen because this kind of
+%application is known for dealing with several sources of exceptions, as they
+%are usually io-intensive (i.e., deals with Network, Bluetooth, camera, and
+%other resources), and multi-htreaded. Moreover, they are based on a single
+%platform which allowed more fine-grained analysis of their exception behaviour.
+%RSC: I AGREE.
+
+This study builds upon the following assumption: the stack traces reported on
+issues contain relevant information related to system crashes. Several studies
+and techniques have also taken this assumption as starting point
+(e.g.,~\cite{sinha2009fault}~\cite{dhaliwal2011classifying}~\cite{kim2013predicting}). Some outcomes were consistently detected through
+our large scale analysis of exception stack traces in the wild, such as:
+
+\begin{itemize}
+
+  \item  A multitude of programming mistakes which are the main causes of stack traces.
+
+  \item Undocumented runtime exceptions signaled by third party
+    code (i.e., libraries and Android platform).
+
+  \item  Odd exception chains (e.g., Error wrapping checked and runtime
+    exceptions and vice-versa), indicating that the purpose of Java's hybrid
+    exception model may not have been adequately used.
+
+  \item  Stack traces caused by uncaught checked exceptions.
+
+\end{itemize}
+
+Hence, the contributions of this study are as follows:
+\begin{itemize}
+
+  \item  It performs the first large scale analysis of Java stack traces and how
+    they can reveal bad-practices on the use of exceptions in Java.
+
+  \item  It introduces ExceptionMiner, a tool developed to support cross-repository stack trace
+    analysis.
+
+\end{itemize}
+
+These contributions are relevant to:
+(i) developers of Java applications
+to familiarize with the most common reported exceptions mis-uses which is the
+first step to help developers to avoid making them; 
+(ii) designers of languages
+to consider ways of reducing the abundance of NullPointerExceptions;
+and (iii)
+tool designers to consider developing tools that enable a cross-project defect
+analysis.
+
+% AVD: We have not looked at private issues, and the tool is described above.
+%We strongly believe that, specially in the open-source environment,
+%faults are not to be hidden in a private bug issue. Faults should be shared and
+%discussed, so that a developer can learn from other projects mistakes.
+%Currently, the search facilities of repositories are very limited, and the
+%ExceptionMiner tool is a contribution in this direction.
+
+%GGG: Personally, I hate this paragraph in all papers as it usually reads like a space-filler
+%RSC: If needed we can remove it.
+This paper is organized as follows. Section 2 presents a
+background on exception handling mechanism. Section 3 presents the study design.
+Section 4 reports study findings. Section 5 provides further discussions and insights.
+Section 6 presents the threats to validity associated to this study. Finally Section
+7 describes the related work, and Section 8 presents our conclusions and
+directions for future work.
+
 \bibliographystyle{IEEEtran}

### 2014-07-07 android-stacks.tex
A new take on the introduction

--- 2014-07-07 android-stacks.tex
@@
 
+\subsection{Georgios's Introduction}
+
+Modern applications have to cope with an increasing number of abnormal
+computation states that arise as a consequence of faults in the application
+itself (e.g., access of null references), noisy user inputs or faults in
+underlying middleware or hardware. The exception handling
+mechanism~\cite{goodenough1975exception} is one of the most used schemes for
+detecting and recovering from such exceptional conditions. When an exception is
+caught by the program's exception handling code, the program has the opportunity
+to assess the conditions that caused the fault and either recover the
+computation or abort. An exception that is not caught by an exception handler
+leads to a program crash. In most programming languages with exceptions, a crash
+usually causes debugging information to emited by the language runtime to help
+the programmer identify and fix the reason of the crash. The debugging
+information is typical a stack trace, which includes an explanatory message and
+the execution stack frame when the exception occurred. In Java, stack trace
+consists of an ordered list of methods that were active on the call stack before
+the exception occurred.
+
+Although the exception handling mechanism has been the target of several studies
+(e.g.~\cite{miller1997issues,Robil00,shah2010understanding,
+garcia2007extracting,garcia2001comparative,cabral2007exception,coelho2011unveiling}),
+the exception handling code is often generally poorly understood and the least
+tested part of software systems~\cite{coelho2011unveiling}. Similarily to source
+code, current works focus their exception handling analysis on either static
+(source code-based) or dynamic (runtime-based) analysis methods. Neither method
+can explore the full execution space of a program and therefore they both fail
+to account for exceptions caused by rutime issues (e.g. resource depletion) or
+dynamically loaded code. Exception stack traces, when available, enable the
+post-mortem analysis of exception handling behaviours of programs and
+concequently the evaluation of exception handling mechanisms in programming
+languages.
+
+The availability of exception stack traces until recently was scarce, while the
+available data were only focused on few large scale projects. In the recent
+past, many open source software projects have their source code and issue
+repositories to Github. The concentration of projects and the existence of
+projects that mine such information, such as GHTorrent~\cite{Gousi13}, made full
+population studies possible.
+
+In this work, we perform the first post mortem analysis of exception handling in
+Java.  Using a custom tool called ExceptionMiner, which we develop specifically
+for this study, we mine stack traces from issues across all Java projects on
+GitHub. Overall, we analyzed 356,057 issues from 16,836 projects, from which
+28,800 stack traces were extracted and used as a source of information for our
+analysis. The stack trace analysis was augmented by additional information
+extracted using byte code and source code analysis done on a carefully selected
+subset of our initial sample (482 Android projects).
+To guide this exploration we compiled general guidelines on how to use
+exceptions proposed by Gosling~\cite{gosling2000java},
+Wirfs-Brock~\cite{wirfs2006toward} and Bloch~\cite{bloch2008effective} and then
+we investigate what stack traces can tell us about the adherence to such
+practices in Java programs.
+
+Some outcomes were consistently detected through
+our large scale analysis of exception stack traces in the wild, such as:
+
+\begin{itemize}
+
+  \item  A multitude of programming mistakes which are the main causes of stack traces.
+
+  \item Undocumented runtime exceptions signaled by third party
+    code (i.e., libraries and Android platform).
+
+  \item  Odd exception chains (e.g., Error wrapping checked and runtime
+    exceptions and vice-versa), indicating that the purpose of Java's hybrid
+    exception model may not have been adequately used.
+
+  \item  Stack traces caused by uncaught checked exceptions.
+
+\end{itemize}
+
+The contributions of this study are as follows:
+\begin{itemize}
+
+  \item  It performs the first large scale analysis of Java stack traces and how
+    they can reveal bad-practices on the use of exceptions in Java.
+
+  \item  It introduces ExceptionMiner, a tool developed to support cross-repository stack trace
+    analysis.
+
+\end{itemize}
+
+These contributions are relevant to:
+(i) developers of Java applications
+to familiarize with the most common reported exceptions mis-uses which is the
+first step to help developers to avoid making them;
+(ii) designers of languages
+to consider ways of reducing the abundance of NullPointerExceptions;
+and (iii)
+tool designers to consider developing tools that enable a cross-project defect
+analysis.
+
+% AVD: We have not looked at private issues, and the tool is described above.
+%We strongly believe that, specially in the open-source environment,
+%faults are not to be hidden in a private bug issue. Faults should be shared and
+%discussed, so that a developer can learn from other projects mistakes.
+%Currently, the search facilities of repositories are very limited, and the
+%ExceptionMiner tool is a contribution in this direction.
+
+%This paper is organized as follows. Section 2 presents a
+%background on exception handling mechanism. Section 3 presents the study design.
+%Section 4 reports study findings. Section 5 provides further discussions and insights.
+%Section 6 presents the threats to validity associated to this study. Finally Section
+%7 describes the related work, and Section 8 presents our conclusions and
+%directions for future work.
+
+\subsection{Remaining Introductions :-)}
+
 Along the recent years we have witnessed a rapid increase in 
@@
 
-Bug issues have called the attention of research community and have been the target of several 
+Bug issues have called the attention of researhc community and  have been the target of several 
 researches [][][]. A subtype of a bug issue are the ones related to system 
@@
 
-One of the most important elements of a crash report genereted eigther manually
-or automatically is the failure stack trace [ref-paper], which tipically consists 
-of an ordered list of methods that were active on the call stack before the failure occurred.
-
-Such failure stack traces are a useful source of 
-information about system crashes, and are often used to
+In [ref] defined a set of characteristics a good bug issue should have. They showed that among 
+the most important relevant information are: (1) steps to reproduce the bug; 
+(2) and the failure stack trace.
+
+A typical stack trace consists of an ordered list of methods that
+were active on the call stack before the failure occurred. Stack traces
+are a useful source of information about system crashes, and are often used to
 support developers in debugging~\cite{schroter2010stack}. Moreover, information
@@
 
-The exception handling mechanism~\cite{goodenough1975exception},
- is one of the most used schemes for detecting and recovering from such 
-failures. Exceptions may arrise as a consequence of faults in the application
+We could observe that existing research has focused on the analysis of a 
+single system at a time. Moreover, the search facilities of repositories are very limited, also
+enabling the analysis of a single project.
+
+Modern applications have to cope with similar and increasing number of abnormal
+computation states that arise as a consequence of faults in the application
 itself (e.g., access of null references), noisy user inputs or faults in
-underlying middleware or hardware.
-
-Although moder application have to deal with a similar and increasing number of 
-source of failures, existing research has focused on the analysis of a 
-single system at a time.
+underlying middleware or hardware. The exception handling mechanism~\cite{goodenough1975exception} is one of the most used schemes for detecting and recovering from such exceptional conditions.
 
@@
 on open-source repositories such as GitHub, some questions arise: What are the common
-characteristics of Github crash reports?
+characteristics of Github crash reports? How can we benefit from such cross-project crash analysis?
 
@@
 
-....... OLD VERSION ......
+-----
+Along the recent years we have witnessed a rapid increase in 
+the number of social applications on the Web. Applications which 
+allow their users to track the activities of a large number 
+of others regardless of his/her location and affiliation.
+
+This tendency can also be seen on the software development context 
+where Github emerged and became an extensively used social coding 
+platform. Github alows developers to create and maintain opensource 
+projects relying  on Git distributed version control, and provides a set 
+of tools which allow the information about developers and their activities
+ to be visible within and across open souce projects.
+
+Among those tools there is the Github issue tracking system, which makes
+issues related to new, features, bugs and general comments publicly available.
+
+-----
 
@@
 
-Hence, in Java an exception can be thrown in one of the following
+%\subsubsection{Exception Throwing and Handling}
+
+In Java an exception can be thrown in one of the following
 circumstances~\cite{gosling2000java}: (i) explicitly thrown when a throw statement is executed; 
@@
 
-When an exception is throw it causes a transfer of control from the point
-where the exception occurred to a point that can be specified by the programmer
-(exception handler) - in Java it is represented by the try-catch block .
-
-A common way of  propagating exceptions in Java programs, is the exception wrapping (also known as exception chaining), which allows one exception to be wrapped in another exception and re-thrown. 
+%When an exception is signaled it causes a transfer of control from the point
+%where the exception occurred to a point that can be specified by the programmer
+%(exception handler) - in Java it is represented by the try-catch block .
+%The search for the handler to deal with a raised
+%exception occurs along the dynamic invocation chain. 
+%If no handler is found in the chain, the default handler associated to the 
+%Thread or ThreadGroup is invoked, which usually terminates the system 
+%and prints the exception stack trace.
+
+%\subsubsection{Exception Chaining} Exception Wrapping 
+%This stack trace was extracted from an issue of an Android application in
+%Github.
+
+The Exception wrapping (also known as exception chaining) is a common way of 
+propagating exceptions in Java programs, 
+which allows one exception to be wrapped in another exception and re-thrown. 
 Figure~\ref{fig:wrapping} presents a stack trace associated which illustrates an
@@
 integrating both quantitative and qualitative evidence along the stages of the
-research process.
-
-The main goal of this study was to gain a better understanding
+research process. The main goal of this study was to gain a better understanding
 of the use of exceptions in Java programs based mainly on the information
-available on exception stack traces embedded on Github issues.
-The exceptions available on exception stack trace are not limited to the exceptions 
-expliciltly thrown by the Java developer but also include
-the exceptions implicitly thrown by the JVM during the program execution,
-due to a programming error (e.g., out-of-bounds array index, division-by-zero, 
-access to a null reference) or an internal error or resource limitation (e.g.,
-OutofMemoryError).
-
- More specifically, we aimed at answering
+available on exception stack traces. More specifically, we aimed at answering
 the following research questions:
@@
 
-  \item[RQ1] What are the common characteristics of such stack traces?
-
-  \item[RQ2] What patterns of exception (mis)-use emerge from the stack trace
+  \item[RQ1] What is the prevalence of stack traces on  reported issues?
+    %\item[RQ2] What are the most common exceptions?
+
+  \item[RQ2] What are the common characteristics of such stack traces?
+
+  \item[RQ3] What patterns of exception (mis)-use emerge from the stack trace
     analysis?
@@
 
-For RQ1 and RQ2, we explore the domain quantitatively and highlight interesting cases by 
-exploring cases qualitatively. To support this investigation, source code and bytecode 
-analysis were used in combination with manual inspection to to everage the  understanding of stack traces.
-
-\subsection{Github Data}
-\label{sec:git}
-
-As mentioned before, this study amined at mining information available on GitHub issues,
-more specifically, extracting and distilling the stack traces embedded on GItHub issues.
-
-Issues on Github are different from issues on dedicated bug tracking tools such as 
-Bugzilla and Jira. The most important difference is that there are no predefined fields
-  (e.g. severity and priority). Instead, Github uses a more open ended tagging system, on which
-repositories are offered a pre-defined set of labels, but repository owners can modify 
-them at will. Therefore, an issue may have none ore an arbitrary set of labels depending 
-on which repository it was created.
-
-Therefore, this study is based on the assumption if an stack trace information embedded
- in an issue, it contain relevant information concerning the exception use -  regardless the
- lables associated to the issue.
-
-Moreover, issues and pull requests are dual on Github; all pull requests have a corresponding 
-``backing'' issue which
-are automatically generated. Therefore, we excluded such automatically generated
-issues from our analysis. Finally, the Github API allows the automated
-generation of issues for specific repositories, which automated tools often use
-to report crashes. In some cases, this led to high number of issues that
-included stack traces. We identified and filtered those cases out (e.g.,
-the~\textsf{pullwifi} project was responsible for almost 50\% of all Android issues in our dataset).
-
-This study used the dataset provided by the GHTorrent project~\cite{Gousi13}, an off-line mirror of the data 
-offered through the Github API.  GHTorrent has been collecting data from Git since 
-February 2012. Up to Dec 2013, and when we queried GHTorrent, it included 332,864 non-fork Java
-repositories \footnote {A Java project is specified in GitHub with a
-specific label.}  from which only 44,323 have contained at least issue report in their lifetime.
-From this set only 16,837 were watched by external users.
-To ensure that our work targets projects that are openly used by
-users other than their developers, we focused our analysis on just those projects.
-
-From this set of 16.836 Java projects which overall contained 356.057 issues, we retrieved the metadata related to each Java project (e.g., repository's names and short descriptions), and the content of every issue and issue comment and stored in a relational database.
-
-Using the heuristics presented in Section~\ref{sec:android}, we selected 482 repositories featuring Android projects on which we perform manual inspection was performed.
-
-\subsection{Study Context}
-
-Before investigating the research questions listed above we analyzed the prevalence of stack traces 
-on issues created on Github. Prevalence or prevalence proportion, a term often used in epidemiology,
-represents the proportion of a population found to have a particular condition
-(e.g. a disease) over the total number of the studied population. In our work, we used
-this term to evaluate a condition (i.e., presence of a stack trace) over issues
-reported on GitHub repositories.
-
-\input{stacktrace-stats}
-
-\begin{table}
-\centering
-\begin{tabular}{p{1em}lrr}
-\hline
- \multicolumn{2}{l}{\bfseries{Java repository-level metrics}} & \\
-\hline
-\multicolumn{2}{l}{\bfseries{Repositories}} & 16,836\\
-    & incl. exception on 1+ issues & 4,776 (28,37\%) \\
-    & incl. stack trace on 1+ issues & 3,758 (22,32\%) \\
-    & incl. exception on 1+ defect issues & 2,019 (41,44\%)\\
-    & incl. stack trace on 1+ defect issues  & 1,698 (34,84\%) \\
-
-\hline
-\end{tabular}
-\caption{Projects x Stack Traces}
-\label{tab:proj}
-\end{table}
-
-\begin{table}
-\centering
-\begin{tabular}{p{1em}lrr}
-\hline
-\multicolumn{2}{l}{\bfseries{Issue-level metrics}} & \bfseries{Java} & \bfseries{Android Subset}\\
-\hline
-   \multicolumn{2}{l}{\bfseries{Repositories}} & 16,836 & 482\\
- \multicolumn{2}{l}{\bfseries{Issues}} & 356,057 & 32,582\\
-    & incl. exceptions & 30,236 (8,5\%) & 3,956 (12,1\%) \\
-    & incl. stack traces & 21,013 (5,9\%) & 3,101 (7,3\%) \\
- \multicolumn{2}{l}{\bfseries{Defect Issues}} & 55,226 & 2,376 \\
-    & incl. exception & 6,741 (12,2\%) & 1,557 (65,6\%) \\
-    & incl. stack trace &  5,196 (9,4\%) & 1,393 (58,7\%)\\
-\multicolumn{2}{l}{\bfseries{Exception Stack Traces}} &  & \\
-    & on issues  & 28,800 & 4,208  \\
-    & on issues labeled as defect & 6,529 & 1,699 \\
-\hline
-\end{tabular}
-\caption{Quantitative overview of the analyzed dataset}
-\label{tab:overview}
-\end{table}
-
-%The use of the Github issue tracker in repositories hosted on Github is optional
-%and not widespread. As of March 2014, from the 332,864 non-fork Java
-%repositories registered in the GHTorrent dataset, only 44,323 have received an
-%issue report in their lifetime while from those only 16,837 are being watched by
-%external users. To ensure that our work targets projects that are openly used by
-%users other than their developers, we focused our analysis on just those
-%projects. Using the heuristics presented in Section~\ref{sec:android}, we
-%selected 482 repositories featuring Android projects. Table~\ref{tab:proj} and
-%Table~\ref{tab:overview} presents two views of the extracted dataset, namely
-%grouped by repository and grouped by issue.
-%
-From Table~\ref{tab:overview}, we can see that repositories that include at
-least one exception or stack trace on an issue are fairly common: 28\% and 22\%
-respectively. Considering only the exceptions and stack traces reported on
-defect issues we find out that 41\% of the repositories reported at least one
-exception in a defect issue, and approximately 35\% of the projects mentioned
-one stack trace in at least one defect issue.
-
-At the individual issue level, 8,5\% and 5,9\% of all issues reported in Java
-repositories contain the full name of an exception or a stack trace
-respectively. The numbers are similar if we isolate Android projects from the
-general population (12,1\% and 7,3\%). Considering only the defect-issues
-however, in the Android repositories as almost 58,7\% of defect-issues include a
-stack trace. This is not true for the general case however.
-
-%Based on the assumption presented in Section III, we therefore did not use this label for further filtering.
-
-%\noindent \fbox{
-%\begin{minipage}{0.96\columnwidth} \emph{ RQ1: 28,37\% of Java
-%  repositories on Github have at least 1 issue that includes a stack trace.
-%  9,4\% of the defect issues of such repositories mentioned at least one stack
-%  trace. Considering only defect-issues of Android subset 58,7\% mentioned at
-%  least 1 stack trace.  }
-%\end{minipage}}
-
-\subsection{The Mining Process}
-\label{sec:miningproc}
-
+For RQ1 and RQ2, we explore the domain quantitatively, and use their answers as
+a basis to investigate the third research question. RQ3 aims at examining whether
+exception (mis)-uses could be revealed through stack trace analysis. To
+support this investigation, source code and bytecode analysis was used to
+leverage the  understanding of stack traces. 
+%For RQ3 we investigate
+%quantitatively and highlight interesting cases by exploring cases qualitatively.
 Figure~\ref{fig:overviewfig} presents an overview of the mixed-methods approach
@@
 
+\subsection{Github Data}
+\label{sec:git}
+
+\noindent\emph{Kinds of Analyzed Issues.} Issues on Github are different from
+issues on dedicated bug tracking tools such as Bugzilla and Jira. The most
+important difference is that there are no predefined fields for bug reporters to
+select from (e.g. severity and priority); Github uses instead a more open ended
+tagging system. Repositories are offered a pre-defined set of labels, but
+repository owners can modify them at will. An issue can have an arbitrary set of
+labels attached. Hence, this study does not restrict the analysis of stack
+traces defined on defect-issues, based on the assumption that regardless of the
+type of issue where its is defined the stack trace information contain relevant
+information concerning the exception use. Moreover, issues and pull requests are
+dual on Github; all pull requests have a corresponding ``backing'' issue which
+are automatically generated. Therefore, we excluded pull request based
+issues from our analysis. Finally, the Github API allows the automated
+generation of issues for specific repositories, which automated tools often use
+to report crashes. In some cases, this led to high number of issues that
+included stack traces. We identified and filtered those cases out (e.g.,
+the~\textsf{pullwifi} project was responsible for almost 50\% of all Android issues in our dataset).
+
+\subsection{Android Apps Selection}
+\label{sec:android}
+
+%In the second phase of this study we selected a subset of Java applications to
+%be analyzed in more detail.The subset chosen was comprised by the Android apps
+%included on GitHub (until 23 Feb 2013).
+To identify Android projects, we performed a case insensitive search for the
+term \textsf{android} in the repository's names and short descriptions.  The
+heuristic filtered 2.542 repositories, from which 589 apps had at least one
+issue containing a stack trace.
+
+Then we performed further cleanup, inspecting the site of every Android
+reporting at least one stack trace, to make sure that they represented real
+mobile apps. During this clean up 106 apps were removed because they were either
+example projects (i.e., toy projects) or tools to support Android development
+(e.g. \textsf{selendroid}, \textsf{roboeletric} --- tools to support the testing of Android apps).
+The filtered set consisted of 482 apps, from which approximately 50\% are also
+hosted in Google Market Place. For each one of the analyzed projects, we
+downloaded the source code and using custom scripts, we extracted their package
+hierarchy (recursive list of of Java packages) and their custom exception types
+as described next.
+
 \subsection{The ExceptionMiner Tool}
@@
 
-\subsection{Android Apps Selection}
-\label{sec:android}
-
-%In the second phase of this study we selected a subset of Java applications to
-%be analyzed in more detail.The subset chosen was comprised by the Android apps
-%included on GitHub (until 23 Feb 2013).
-To identify Android projects, we performed a case insensitive search for the
-term \textsf{android} in the repository's names and short descriptions.  The
-heuristic filtered 2.542 repositories, from which 589 apps had at least one
-issue containing a stack trace.
-
-Then we performed further cleanup, inspecting the site of every Android
-reporting at least one stack trace, to make sure that they represented real
-mobile apps. During this clean up 106 apps were removed because they were either
-example projects (i.e., toy projects) or tools to support Android development
-(e.g. \textsf{selendroid}, \textsf{roboeletric} --- tools to support the testing of Android apps).
-The filtered set consisted of 482 apps, from which approximately 50\% are also
-hosted in Google Market Place. For each one of the analyzed projects, we
-downloaded the source code and using custom scripts, we extracted their package
-hierarchy (recursive list of of Java packages) and their custom exception types
-as described next.
-
 %\subsection{Phase 1: Analyzing Pure Stack Trace Information of Java Programs}
@@
 \section{RQ1: What is the prevalence of stack traces on reported issues?}
+
+Prevalence or prevalence proportion, a term often used in epidemiology,
+represents the proportion of a population found to have a particular condition
+(e.g. a disease) over the total number of the studied population. In our work, we used
+this term to evaluate a condition (i.e., presence of a stack trace) over issues
+reported on GitHub repositories.
+
+\input{stacktrace-stats}
+
+\begin{table}
+\centering
+\begin{tabular}{p{1em}lrr}
+\hline
+ \multicolumn{2}{l}{\bfseries{Java repository-level metrics}} & \\
+\hline
+\multicolumn{2}{l}{\bfseries{Repositories}} & 16,836\\
+    & incl. exception on 1+ issues & 4,776 (28,37\%) \\
+    & incl. stack trace on 1+ issues & 3,758 (22,32\%) \\
+    & incl. exception on 1+ defect issues & 2,019 (41,44\%)\\
+    & incl. stack trace on 1+ defect issues  & 1,698 (34,84\%) \\
+
+\hline
+\end{tabular}
+\caption{Projects x Stack Traces}
+\label{tab:proj}
+\end{table}
+
+\begin{table}
+\centering
+\begin{tabular}{p{1em}lrr}
+\hline
+\multicolumn{2}{l}{\bfseries{Issue-level metrics}} & \bfseries{Java} & \bfseries{Android Subset}\\
+\hline
+   \multicolumn{2}{l}{\bfseries{Repositories}} & 16,836 & 482\\
+ \multicolumn{2}{l}{\bfseries{Issues}} & 356,057 & 32,582\\
+    & incl. exceptions & 30,236 (8,5\%) & 3,956 (12,1\%) \\
+    & incl. stack traces & 21,013 (5,9\%) & 3,101 (7,3\%) \\
+ \multicolumn{2}{l}{\bfseries{Defect Issues}} & 55,226 & 2,376 \\
+    & incl. exception & 6,741 (12,2\%) & 1,557 (65,6\%) \\
+    & incl. stack trace &  5,196 (9,4\%) & 1,393 (58,7\%)\\
+\multicolumn{2}{l}{\bfseries{Exception Stack Traces}} &  & \\
+    & on issues  & 28,800 & 4,208  \\
+    & on issues labeled as defect & 6,529 & 1,699 \\
+\hline
+\end{tabular}
+\caption{Quantitative overview of the analyzed dataset}
+\label{tab:overview}
+\end{table}
+
+%The use of the Github issue tracker in repositories hosted on Github is optional
+%and not widespread. As of March 2014, from the 332,864 non-fork Java
+%repositories registered in the GHTorrent dataset, only 44,323 have received an
+%issue report in their lifetime while from those only 16,837 are being watched by
+%external users. To ensure that our work targets projects that are openly used by
+%users other than their developers, we focused our analysis on just those
+%projects. Using the heuristics presented in Section~\ref{sec:android}, we
+%selected 482 repositories featuring Android projects. Table~\ref{tab:proj} and
+%Table~\ref{tab:overview} presents two views of the extracted dataset, namely
+%grouped by repository and grouped by issue.
+%
+From Table~\ref{tab:overview}, we can see that repositories that include at
+least one exception or stack trace on an issue are fairly common: 28\% and 22\%
+respectively. Considering only the exceptions and stack traces reported on
+defect issues we find out that 41\% of the repositories reported at least one
+exception in a defect issue, and approximately 35\% of the projects mentioned
+one stack trace in at least one defect issue.
+
+At the individual issue level, 8,5\% and 5,9\% of all issues reported in Java
+repositories contain the full name of an exception or a stack trace
+respectively. The numbers are similar if we isolate Android projects from the
+general population (12,1\% and 7,3\%). Considering only the defect-issues
+however, in the Android repositories as almost 58,7\% of defect-issues include a
+stack trace. This is not true for the general case however.
+
+%Based on the assumption presented in Section III, we therefore did not use this label for further filtering.
+
+%\noindent \fbox{
+%\begin{minipage}{0.96\columnwidth} \emph{ RQ1: 28,37\% of Java
+%  repositories on Github have at least 1 issue that includes a stack trace.
+%  9,4\% of the defect issues of such repositories mentioned at least one stack
+%  trace. Considering only defect-issues of Android subset 58,7\% mentioned at
+%  least 1 stack trace.  }
+%\end{minipage}}
 
@@
 or computations that may fail) could improve the robustness of Java programs.
-
-TO DISCUSS: How can we benefit from such cross-project crash analysis?
 

### 2014-07-07 android-stacks.tex

--- 2014-07-07 android-stacks.tex
@@
 \section{Introduction}
+
+\subsection{Georgios's Introduction}
+
+Modern applications have to cope with an increasing number of abnormal
+computation states that arise as a consequence of faults in the application
+itself (e.g., access of null references), noisy user inputs or faults in
+underlying middleware or hardware. The exception handling
+mechanism~\cite{goodenough1975exception} is one of the most used schemes for
+detecting and recovering from such exceptional conditions. When an exception is
+caught by the program's exception handling code, the program has the opportunity
+to assess the conditions that caused the fault and either recover the
+computation or abort. An exception that is not caught by an exception handler
+leads to a program crash. In most programming languages with exceptions, a crash
+usually causes debugging information to emited by the language runtime to help
+the programmer identify and fix the reason of the crash. The debugging
+information is typical a stack trace, which includes an explanatory message and
+the execution stack frame when the exception occurred. In Java, stack trace
+consists of an ordered list of methods that were active on the call stack before
+the exception occurred.
+
+Although the exception handling mechanism has been the target of several studies
+(e.g.~\cite{miller1997issues,Robil00,shah2010understanding,
+garcia2007extracting,garcia2001comparative,cabral2007exception,coelho2011unveiling}),
+the exception handling code is often generally poorly understood and the least
+tested part of software systems~\cite{coelho2011unveiling}. Similarily to source
+code, current works focus their exception handling analysis on either static
+(source code-based) or dynamic (runtime-based) analysis methods. Neither method
+can explore the full execution space of a program and therefore they both fail
+to account for exceptions caused by rutime issues (e.g. resource depletion) or
+dynamically loaded code. Exception stack traces, when available, enable the
+post-mortem analysis of exception handling behaviours of programs and
+concequently the evaluation of exception handling mechanisms in programming
+languages.
+
+The availability of exception stack traces until recently was scarce, while the
+available data were only focused on few large scale projects. In the recent
+past, many open source software projects have their source code and issue
+repositories to Github. The concentration of projects and the existence of
+projects that mine such information, such as GHTorrent~\cite{Gousi13}, made full
+population studies possible.
+
+In this work, we perform the first post mortem analysis of exception handling in
+Java.  Using a custom tool called ExceptionMiner, which we develop specifically
+for this study, we mine stack traces from issues across all Java projects on
+GitHub. Overall, we analyzed 356,057 issues from 16,836 projects, from which
+28,800 stack traces were extracted and used as a source of information for our
+analysis. The stack trace analysis was augmented by additional information
+extracted using byte code and source code analysis done on a carefully selected
+subset of our initial sample (482 Android projects).
+To guide this exploration we compiled general guidelines on how to use
+exceptions proposed by Gosling~\cite{gosling2000java},
+Wirfs-Brock~\cite{wirfs2006toward} and Bloch~\cite{bloch2008effective} and then
+we investigate what stack traces can tell us about the adherence to such
+practices in Java programs.
+
+Some outcomes were consistently detected through
+our large scale analysis of exception stack traces in the wild, such as:
+
+\begin{itemize}
+
+  \item  A multitude of programming mistakes which are the main causes of stack traces.
+
+  \item Undocumented runtime exceptions signaled by third party
+    code (i.e., libraries and Android platform).
+
+  \item  Odd exception chains (e.g., Error wrapping checked and runtime
+    exceptions and vice-versa), indicating that the purpose of Java's hybrid
+    exception model may not have been adequately used.
+
+  \item  Stack traces caused by uncaught checked exceptions.
+
+\end{itemize}
+
+The contributions of this study are as follows:
+\begin{itemize}
+
+  \item  It performs the first large scale analysis of Java stack traces and how
+    they can reveal bad-practices on the use of exceptions in Java.
+
+  \item  It introduces ExceptionMiner, a tool developed to support cross-repository stack trace
+    analysis.
+
+\end{itemize}
+
+These contributions are relevant to:
+(i) developers of Java applications
+to familiarize with the most common reported exceptions mis-uses which is the
+first step to help developers to avoid making them;
+(ii) designers of languages
+to consider ways of reducing the abundance of NullPointerExceptions;
+and (iii)
+tool designers to consider developing tools that enable a cross-project defect
+analysis.
+
+% AVD: We have not looked at private issues, and the tool is described above.
+%We strongly believe that, specially in the open-source environment,
+%faults are not to be hidden in a private bug issue. Faults should be shared and
+%discussed, so that a developer can learn from other projects mistakes.
+%Currently, the search facilities of repositories are very limited, and the
+%ExceptionMiner tool is a contribution in this direction.
+
+%This paper is organized as follows. Section 2 presents a
+%background on exception handling mechanism. Section 3 presents the study design.
+%Section 4 reports study findings. Section 5 provides further discussions and insights.
+%Section 6 presents the threats to validity associated to this study. Finally Section
+%7 describes the related work, and Section 8 presents our conclusions and
+%directions for future work.
+
+\subsection{Remaining Introductions :-)}
 

### 2014-07-07 android-stacks.tex
small corrections

--- 2014-07-07 android-stacks.tex
@@
 \begin{document}
-\title{What Can Stack Traces Tell us about Exceptions in Java?}
+\title{What Can Stack Traces Tell us about of the use Exceptions in Java?}
 
@@
 
-  \item  A multitude of programming mistakes which are the main causes of stack traces.
+  \item  A multitude of programming mistakes were reported as the main causes of stack traces:
+  41,1\%  (12,625)  of reported stack traces were caused by programming mistakes; and 26,31\% of analyzed stack traces reported a java.lang.NullPointerException as root cause;
+
+   \item 1,671 projects (44,47\%) contained at least one issue reporting a java.lang.NullPointerException as the root cause of a stack trace.
 
@@
     code (i.e., libraries and Android platform).
+This study found 118 distinct methods defined on third party code and  reported as the root signalers of runtime exceptions.
+Only 1 of such methods (0,8\%) documented the explicitly thrown runtime exception in a Javadoc comment; and none of these methods included the exception as part of the exception interface (i.e., using throws clause).
 
@@
 
-  \item  Stack traces caused by uncaught checked exceptions.
+  \item  Issues reporting uncaught checked exceptions.
 
@@
 
-Hence, the contributions of this study are as follows:
-\begin{itemize}
-
-  \item  It performs a large scale analysis of Java stack traces and how
-    they can reveal bad-practices on the use of exceptions in Java.
-
-  \item  It introduces ExceptionMiner, a tool developed to support cross-repository stack trace
-    analysis.
-
-\end{itemize}
+%Hence, the contributions of this study are as follows:
+%\begin{itemize}
+%  \item  It performs a large scale analysis of Java stack traces and how
+ %   they can reveal bad-practices on the use of exceptions in Java.
+% \item  It introduces ExceptionMiner, a tool developed to support cross-repository stack trace
+%    analysis.
+%\end{itemize}
 
@@
 mechanism, embedded in many main-stream programming languages.
-
 The main disadvantage of return codes is the fact that the programmer can just
 ignore the return code and let the program execute in inconsistent state. 
-
 The exception mechanism~\cite{goodenough1975exception} was proposed as a way of
@@
 
-In languages such as CLU, Modula-3, Guide, and Extended Ada the exceptions are 
+In languages such as CLU\footnote{In CLU,  if a handler is 
+found for a given exception, it is converted in a predened default exception, 
+called failure which is the only one which can be implicitly 
+propagated in this model [ref-clue].}, Modula-3, Guide, and Extended Ada the exceptions are 
 checked, which means that the 
@@
 Accordingly, the programmer must explicitly specify every exception that can be
-signaled to support the compile time checking \footnote{In CLU,  if a handler is 
-found for a given exception, it is converted in a predened default exception, 
-called failure which is the only one which can be implicitly 
-propagated in this model.}. This approach is known to be
+signaled to support the compile time checking. This approach is known to be
 costly to maintain (Dooren and Steegmans, 2005) since exception updates are
@@
 In Java a method can throw either checked or unchecked exceptions.
-Java is the only language that allows this hybrid approach, on which the developer can define exceptions of both kinds to be thrown and
-handled in a single system. Next section details Java exception handling model.
-
-.......
+Java is the only language that allows a hybrid approach on which the developer can define exceptions of both kinds to be thrown and handled in a single system.
+Next section details Java exception handling model.
 
 Although exception handling mechanism is the most used technique to deal with such 
-failures - been embedded in several programming languages (e.g. Java, C++, C\#) -  
+failures - been embedded in several programming languages as show above,   
 the exception handling code is often generally poorly understood and the least
@@
 
-........
-
-\subsection{Exception Handling in Java} \label{sec:extypes}
+\subsection{Checked an Unchecked Exceptions in Java} \label{sec:extypes}
 
@@
 Figure~\ref{fig:overviewfig} presents an overview of the mixed-methods approach
-conducted to answer these questions.
+conducted to answer these questions. The mixed-methods approach is composed by the following steps:
 
@@
 
-The mixed-methods approach is composed by the following steps:
-
-Step1. Retrieving the Dataset to be Analized. We begun our investigation using the dataset 
-provided by the GHTorrent project~\cite{Gousi13}, an off-line mirror of the data 
-offered through the Github API.  GHTorrent has been collecting data from Git since 
-February 2012. Up to Dec 2013, when we queried GHTorrent, it included 16.836 Java projects 
-which overall contained 356.057 issues.\footnote {A Java project is specified in GitHub with a
-specific label.} From GHTorrent we retrieved the metadata related to each Java project (e.g., 
+\textbf{Step1.} To obtain the Dataset to be analized, we queried GHTorrent as detailed in Section XXX. From GHTorrent we retrieved the metadata related to each Java project (e.g., 
 repository's names and short descriptions), and the content of every issue and issue comment 
@@
 
-Step 2. Extracting Exception Information from Issues. In this step, we used the ExceptionMiner tool,
+\textbf{Step 2.} Extracting Exception Information from Issues. In this step, we used the ExceptionMiner tool,
  developed in this work, to extract the exception names and exception stack traces of each
@@
 
-Step 3. Restricting the Dataset to Android Projects. We select a subset of all Android 
+\textbf{Step 3.} Restricting the Dataset to Android Projects. We select a subset of all Android 
 projects available on the dataset to perform a deeper analysis (which combined mined
@@
 
-Step 4. Combining Information Mined From Issues with Bytecode-Code Analysis.
-
+To support a deeper investigation of the stack traces of Android apps, every
+exception defined on a stack trace was then classified according to its type
+(e.g. Error, Exception or RuntimeException) and to its signaler. To do so, the following 
+steps were performed:
+
+\textbf{Step 4.}  Bytecode-Code analysis of JVM and Android Exceptions.
+We developerd a bytecode analysis tool based on DesignWizard to discover all exceptions thrown by the last version of SDK version 7, and Android Platform and OS (level 19).
+Every exception was extracted and classified as Runtime, Checked or Error based on the analysis of the exception hierarchy.
+Our analysis found XXX runtime exceptions in JVM and XX checked, on Android xxx checked and YYY runtime.
+
+\textbf{Step 5.}  Source-Code Analysis of Android Projects. 
 For each one of the analyzed projects, we
 downloaded the source code and using custom scripts, we extracted their package
-hierarchy (recursive list of of Java packages) and their custom exception types
-as described next.
-
-Step 5
-. Combining Information Mined From Issues with Source-Code Analysis.
-For each one of the analyzed projects, we
-downloaded the source code and using custom scripts, we extracted their package
-hierarchy (recursive list of of Java packages) and their custom exception types
-as described next.
-
-Step 5. Combining Information Mined From Issues with Manual Inspection.
-
-To support a deeper investigation of the stack traces of Android apps, every
-exception defined on a stack trace was then classified according to its type
-(e.g. Error, Exception or RuntimeException) (step 4). The exception classifier
-was initialized with all the exceptions defined on the Java SDK version 7, the
-exceptions of the Android (Platform and OS) (level 19), and all the exceptions
-of the 482 Android projects selected for this study, extracted in the previous
-step. Moreover, all the packages associated to exceptions and signalers were
-also extracted to support further analysis (step 5). The signaler classifier is
-initialized with information about all packages found on Java SDK version 7, and
-Android Platform and OS (level 19). Finally, additional meta-data is combined to
-each stack trace (step 6) which used together with manual inspections (step 7)
-enabled to provide answers to RQ3.
+hierarchy (recursive list of of Java packages) and their custom exception types.
+
+\textbf{Step 6.}  Manual Inspection.
+XXXX of the exceptions reported on stack traces  could not be found on JMV, SO or inside the last version of the project.
+Based on the full name of the exception, a manual inspection was performed on the artifacts associated to the project and libraries used by the project.
+When even on those sources the exception was not foung [grep-code and similar sites] the exception, in order to discover the exception type.
+Only 3 exceptions remained undefined and were removed from this study.
+
+OBS: THE PICTURE WILL BE UPDATED TO REFLECT THIS PROCESS DESCRIPTION.
 
@@
 If we perform a cross-project analysis, the java.lang.NullPointerException, we could observe that such 
-exception was reported on 1,671 projects (which represents 44,47% of all projects reporting at least one stack trace).
+exception was reported on 1,671 projects (which represents 44,47\% of all projects reporting at least one stack trace).
 
@@
  (e.g., Android and libs) were runtime exceptions. After filtering all the exceptions implicitly signaled by 
-JVM (due to programming mistakes) and inspecting the signaler methods of such exceptions (i.e.,  direct an inderect 
+JVM (due to programming mistakes) and inspecting the signaler methods of such exceptions (i.e.,  direct an indirect 
 instances of java.lang.RuntimeException). We could observe that only 1 method out of 118 inspected signalers
@@
 
-\begin{figure}
-  \begin{center}
-    \includegraphics[scale=0.4]{stack-wrappings-hist.pdf}
-  \end{center}
-  \caption{Number of stack traces wrappings for Java and Android repositories. }
-  \label{fig:sizewrapbhists}
-\end{figure}
+%\begin{figure}
+%  \begin{center}
+%    \includegraphics[scale=0.4]{stack-wrappings-hist.pdf}
+%  \end{center}
+%  \caption{Number of stack traces wrappings for Java and Android repositories. }
+%  \label{fig:sizewrapbhists}
+%\end{figure}
 

### 2014-07-07 android-stacks.tex
merging introductions - step 1

@@
 
-\subsection{Georgios's Introduction}
+\subsection{Georgios's and Roberta Merged Introduction}
 
@@
 detecting and recovering from such exceptional conditions. When an exception is
-caught by the program's exception handling code, the program has the opportunity
-to assess the conditions that caused the fault and either recover the
-computation or abort. An exception that is not caught by an exception handler
-leads to a program crash. In most programming languages with exceptions, a crash
-usually causes debugging information to emited by the language runtime to help
+caught inside the application by the program's exception handling code, 
+the application has the opportunity
+to assess the conditions that caused the fault and either recover from them or abort. 
+An exception that remains uncaught (is not caught by an exception handler)
+leads to a program crash. In most programming languages with embedded exception handling mechanism, when a crash happens usually the language runtime 
+agenerates debugging information to help
 the programmer identify and fix the reason of the crash. The debugging
@@
 the exception handling code is often generally poorly understood and the least
-tested part of software systems~\cite{coelho2011unveiling}. Similarily to source
-code, current works focus their exception handling analysis on either static
-(source code-based) or dynamic (runtime-based) analysis methods. Neither method
+tested part of software systems~\cite{coelho2011unveiling}.
+
+%Similarily to source
+%code, 
+%TODO GEORGIOUS I DID NOT UNDERESTA.ND THE IDEA OF THE PREVIOUS SENTENCE
+
+Such studies  perform the analysis of exception handling code based  either on static
+(source code-based) analysis or dynamic (runtime-based) testing methods. Neither of such methods, however,
 can explore the full execution space of a program and therefore they both fail
-to account for exceptions caused by rutime issues (e.g. resource depletion) or
+to account for exceptions implicitly signaled by rutime environment (e.g. resource depletion) or
 dynamically loaded code. Exception stack traces, when available, enable the
@@
 
-  \item  A multitude of programming mistakes which are the main causes of stack traces.
+   \item  A multitude of programming mistakes were reported as the main causes of stack traces:
+  41,1\%  (12,625)  of reported stack traces were caused by programming mistakes; and 26,31\% of analyzed stack traces reported a java.lang.NullPointerException as root cause;
+
+   \item 1,671 projects (44,47\%) contained at least one issue reporting a java.lang.NullPointerException as the root cause of a stack trace.
 

### 2014-07-07 android-stacks.tex
merging introductions - step 2

@@
 mechanism~\cite{goodenough1975exception} is one of the most used schemes for
-detecting and recovering from such exceptional conditions. When an exception is
-caught inside the application by the program's exception handling code, 
-the application has the opportunity
-to assess the conditions that caused the fault and either recover from them or abort. 
-An exception that remains uncaught (is not caught by an exception handler)
-leads to a program crash. In most programming languages with embedded exception handling mechanism, when a crash happens usually the language runtime 
-agenerates debugging information to help
-the programmer identify and fix the reason of the crash. The debugging
+detecting and recovering from such exceptional conditions.
+When an exception is thrown it can be caught inside the application (by the program's exception handling code) and the the application has the opportunity to assess the conditions that caused the fault and either recover from them or abort.
+If the exception that remains uncaught (is not caught by an exception handler) it
+leads to a program crash -  uncaught exceptions ~\cite{jo2004uncaught} are one 
+of the main causes of application crashes.
+
+ In most programming languages with embedded exception handling mechanism, when a crash happens usually the language runtime generates debugging information to help
+the programmer identify and fix the cause of the crash. The debugging
 information is typical a stack trace, which includes an explanatory message and
@@
 The availability of exception stack traces until recently was scarce, while the
-available data were only focused on few large scale projects. In the recent
-past, many open source software projects have their source code and issue
-repositories to Github. The concentration of projects and the existence of
-projects that mine such information, such as GHTorrent~\cite{Gousi13}, made full
-population studies possible.
+available data were only focused on few large scale projects.
+Along the recent years we have witnessed a rapid increase in the number of open source 
+projects that make their control version and issue tracking systems publicly 
+available in repository hosting sites such as GitHub.
+
+The concentration of open source projects and the existence of
+initiatives that mine such information, such as GHTorrent~\cite{Gousi13},
+ allow the information about projects to be visible within and across projects and
+ make a full population studies possible.
 
@@
 extracted using byte code and source code analysis done on a carefully selected
-subset of our initial sample (482 Android projects).
-To guide this exploration we compiled general guidelines on how to use
-exceptions proposed by Gosling~\cite{gosling2000java},
+subset of our initial sample (482 Android projects). Manual inspection was also 
+used in this study to leverage the understanding of stack traces and support 
+further discussions and insights.
+
+To guide this exploration we compiled general 
+guidelines on how to use exceptions proposed by Gosling~\cite{gosling2000java},
 Wirfs-Brock~\cite{wirfs2006toward} and Bloch~\cite{bloch2008effective} and then
 we investigate what stack traces can tell us about the adherence to such
-practices in Java programs.
-
-Some outcomes were consistently detected through
+practices in Java programs. Some outcomes were consistently detected through
 our large scale analysis of exception stack traces in the wild, such as:
@@
 
-\subsection{Remaining Introductions :-)}
+\subsection{Tentative  Introductions to chance the focus}
 
@@
 exploring cases qualitatively. To support this investigation, source code and bytecode 
-analysis were used in combination with manual inspection to to everage the  understanding of stack traces.
+analysis were used in combination with manual inspection to leverage the  understanding of stack traces and support further discussions and insights.
 

### 2014-07-07 android-stacks.tex
merging introductions - step 3
Giving the introduction some love

@@
 
-\subsection{Georgios's and Roberta Merged Introduction}
+\subsection{Georgios's and Roberta's Merged Introduction}
 
@@
 mechanism~\cite{goodenough1975exception} is one of the most used schemes for
-detecting and recovering from such exceptional conditions.
-When an exception is thrown it can be caught inside the application (by the program's exception handling code) and the the application has the opportunity to assess the conditions that caused the fault and either recover from them or abort.
-If the exception that remains uncaught (is not caught by an exception handler) it
-leads to a program crash -  uncaught exceptions ~\cite{jo2004uncaught} are one 
-of the main causes of application crashes.
-
- In most programming languages with embedded exception handling mechanism, when a crash happens usually the language runtime generates debugging information to help
-the programmer identify and fix the cause of the crash. The debugging
-information is typical a stack trace, which includes an explanatory message and
-the execution stack frame when the exception occurred. In Java, stack trace
-consists of an ordered list of methods that were active on the call stack before
-the exception occurred.
+detecting and recovering from such exceptional conditions. When an exception is
+thrown it can be caught inside the application (by the program's exception
+handling code) and the application has the opportunity to assess the conditions
+that caused the fault and either recover from them or abort. If the exception
+that remains uncaught, the program crashes. Uncaught exceptions are one of the
+main causes of application crashes~\cite{jo2004uncaught}.
+
+In most programming languages featuring exception handling, when a crash
+happens, the language runtime generates debugging information to help the
+programmer identify the cause of the crash. The debugging information is
+typically a stack trace, which includes an explanatory message and the execution
+stack frame when the exception occurred.
+%In Java, stack trace consists of an ordered list of methods that were active on
+%the call stack before the exception occurred.
 
@@
 garcia2007extracting,garcia2001comparative,cabral2007exception,coelho2011unveiling}),
-the exception handling code is often generally poorly understood and the least
-tested part of software systems~\cite{coelho2011unveiling}.
-
-%Similarily to source
-%code, 
-%TODO GEORGIOUS I DID NOT UNDERESTA.ND THE IDEA OF THE PREVIOUS SENTENCE
-
-Such studies  perform the analysis of exception handling code based  either on static
-(source code-based) analysis or dynamic (runtime-based) testing methods. Neither of such methods, however,
-can explore the full execution space of a program and therefore they both fail
-to account for exceptions implicitly signaled by rutime environment (e.g. resource depletion) or
-dynamically loaded code. Exception stack traces, when available, enable the
-post-mortem analysis of exception handling behaviours of programs and
-concequently the evaluation of exception handling mechanisms in programming
-languages.
+the exception handling code is generally poorly understood and often the least
+tested part of software systems~\cite{coelho2011unveiling}.  To analyze the
+exception handling code, current studies use either static (source code-based)
+analysis or dynamic (runtime-based) testing methods. Neither method can explore
+the full execution space of a program and therefore they both fail to account
+for exceptions implicitly signaled by the runtime environment (e.g. due to
+resource depletion) or dynamically loaded code. Exception stack traces, when
+available, enable the post-mortem analysis of exception handling behaviours of
+programs and consequently the evaluation of exception handling mechanisms in
+programming languages.
 
 The availability of exception stack traces until recently was scarce, while the
-available data were only focused on few large scale projects.
-Along the recent years we have witnessed a rapid increase in the number of open source 
-projects that make their control version and issue tracking systems publicly 
-available in repository hosting sites such as GitHub.
-
-The concentration of open source projects and the existence of
-initiatives that mine such information, such as GHTorrent~\cite{Gousi13},
- allow the information about projects to be visible within and across projects and
- make a full population studies possible.
+available data concerned a few large scale projects (e.g. the Mozilla
+crash report
+dataset\footnote{https://code.google.com/p/promisedata/issues/detail?id=88}).
+Along the recent years, we have witnessed a rapid increase in the number of open
+source projects that make their control version and issue tracking systems
+publicly available in repository hosting sites such as GitHub. The
+concentration of open source projects and the existence of initiatives that mine
+such information, such as GHTorrent~\cite{Gousi13}, allow the information about
+projects to be visible within and across projects and make a full population
+studies possible.
 
@@
 extracted using byte code and source code analysis done on a carefully selected
-subset of our initial sample (482 Android projects). Manual inspection was also 
-used in this study to leverage the understanding of stack traces and support 
+subset of our initial sample (482 Android projects). Manual inspection was
+used to leverage the understanding of stack traces and support
 further discussions and insights.
-
-To guide this exploration we compiled general 
-guidelines on how to use exceptions proposed by Gosling~\cite{gosling2000java},
-Wirfs-Brock~\cite{wirfs2006toward} and Bloch~\cite{bloch2008effective} and then
-we investigate what stack traces can tell us about the adherence to such
-practices in Java programs. Some outcomes were consistently detected through
-our large scale analysis of exception stack traces in the wild, such as:
+To guide our analysis, we compiled a set of best practices on how to use
+exceptions by composing relevant suggestions by Gosling~\cite{gosling2000java},
+Wirfs-Brock~\cite{wirfs2006toward} and Bloch~\cite{bloch2008effective}.  The
+comparison of those with the mined exceptions revealed several interesting
+findings, such as:
 
@@
 %directions for future work.
-
-\subsection{Tentative  Introductions to chance the focus}
-
-Along the recent years we have witnessed a rapid increase in 
-the number of open source software  available on the Web.
-Since such projects make their control version and 
-issue tracking systems publicly available, they represent a valuable 
-source of information for the "state of practice" on software development
-which had already started to be investigated [][][][][][I still need to collect these references].
-
-In this scenario Github emerged as one of the most used repository hosting sites.
-It  alows developers to create and maintain opensource projects relying on Git distributed version 
-control and provides a set of tools which allow the information about developers 
-and their activities to be visible within and across open souce projects.
-
-Bug issues have called the attention of the research community and have been the target of increasing 
-researches [][][]. A specific type of bug issue is the one related to system 
-crashes.  Such issues usually contain an exception stack trace [ref-paper], which tipically consists 
-of an ordered list of methods that were active on the call stack before the failure occurred [ref-bug-issues].
-
-Exception stack traces are a useful source of 
-information about system crashes, and are often used to
-support developers in debugging~\cite{schroter2010stack}. Moreover, information
-from stack traces combined with crash and bug reports can be used to support bug
-classification and clustering~\cite{wang2013improving, kim2011crash,
-dhaliwal2011classifying}, fault-proneness prediction
-models~\cite{kim2013predicting} and even automated bug
-fixing~\cite{sinha2009fault} tools.
-
-Although modern applications have to deal with increasinng and similar 
-sources of failures - that arise as a consequence of faults in the application
-itself (e.g., access of null references), noisy user inputs or faults in
-underlying middleware or hardware -  on bug classificationthe current research on 
-exploring bug issue information are structured on  a system at a time basis.
-
-Specially  in a context where we have access to wealth of information available
-on open-source repositories such as GitHub, interesting knowledge may emerge 
-from a cross-project crash analysis. Hence such initial questions arise: What are the common
-characteristics of the stack traces reported on GitHub? Which patterns of exception (mis)use may emerge?
-
-To answer these questions, we conduct an exploratory study of stack traces in
-the wild. Using a custom tool called ExceptionMiner, which was specifically
-developed for this study, we mined stack traces from issues across all Java
-projects on GitHub. Overall, we analyzed 356,057 issues from 16,836 projects,
-from which 28,800 stack traces were extracted and used as a source of
-information for our analysis. The stack trace analysis  was augmented by
-additional information extracted using byte code and source code analysis done
-on a carefully selected subset of our initial sample (482 Android projects).
-%This study was divided in two phases: in the fist phase all the issues of all
-%Java projects available on GitHub were analyzed and the embedded stack traces
-%extracted using ExceptionMiner.
-
-To guide this exploration we compiled general guidelines on how to use
-exceptions proposed by Gosling~\cite{gosling2000java},
-Wirfs-Brock~\cite{wirfs2006toward} and Bloch~\cite{bloch2008effective} and then
-we investigate what stack traces can tell us about the adherence to such
-practices in Java programs.
-%~\todo{This should be our motivation, as this is
-%our biggest contribution. I already changed the text. Lets discuss about this
-%new version}
-%RSC: I AGREE.
-
-%GGG: I think this is too much detail for the introduction In the second phase
-%of this study, we combined the information of the mined stack traces with
-%source code and bytecode analysis. To do so we selected the subset of 482
-%Android repositories. The Android repositories were chosen because this kind of
-%application is known for dealing with several sources of exceptions, as they
-%are usually io-intensive (i.e., deals with Network, Bluetooth, camera, and
-%other resources), and multi-htreaded. Moreover, they are based on a single
-%platform which allowed more fine-grained analysis of their exception behaviour.
-%RSC: I AGREE.
-
-This study builds upon the following assumption: the stack traces reported on
-issues contain relevant information related to system crashes. Several studies
-and techniques have also taken this assumption as starting point
-(e.g.,~\cite{sinha2009fault}~\cite{dhaliwal2011classifying}~\cite{kim2013predicting}). Some outcomes were consistently detected through
-our large scale analysis of exception stack traces in the wild, such as:
-
-\begin{itemize}
-
-  \item  A multitude of programming mistakes were reported as the main causes of stack traces:
-  41,1\%  (12,625)  of reported stack traces were caused by programming mistakes; and 26,31\% of analyzed stack traces reported a java.lang.NullPointerException as root cause;
-
-   \item 1,671 projects (44,47\%) contained at least one issue reporting a java.lang.NullPointerException as the root cause of a stack trace.
-
-  \item Undocumented runtime exceptions signaled by third party
-    code (i.e., libraries and Android platform).
-This study found 118 distinct methods defined on third party code and  reported as the root signalers of runtime exceptions.
-Only 1 of such methods (0,8\%) documented the explicitly thrown runtime exception in a Javadoc comment; and none of these methods included the exception as part of the exception interface (i.e., using throws clause).
-
-  \item  Odd exception chains (e.g., Error wrapping checked and runtime
-    exceptions and vice-versa), indicating that the purpose of Java's hybrid
-    exception model may not have been adequately used.
-
-  \item  Issues reporting uncaught checked exceptions.
-
-\end{itemize}
-
-%Hence, the contributions of this study are as follows:
-%\begin{itemize}
-%  \item  It performs a large scale analysis of Java stack traces and how
- %   they can reveal bad-practices on the use of exceptions in Java.
-% \item  It introduces ExceptionMiner, a tool developed to support cross-repository stack trace
-%    analysis.
-%\end{itemize}
-
-These contributions are relevant to:
-(i) developers of Java applications
-to familiarize with the most common reported exceptions mis-uses which is the
-first step to help developers to avoid making them; 
-(ii) designers of languages
-to consider ways of reducing the abundance of NullPointerExceptions;
-and (iii)
-tool designers to consider developing tools that enable a cross-project defect
-analysis.
-
-% AVD: We have not looked at private issues, and the tool is described above.
-%We strongly believe that, specially in the open-source environment,
-%faults are not to be hidden in a private bug issue. Faults should be shared and
-%discussed, so that a developer can learn from other projects mistakes.
-%Currently, the search facilities of repositories are very limited, and the
-%ExceptionMiner tool is a contribution in this direction.
-
-%GGG: Personally, I hate this paragraph in all papers as it usually reads like a space-filler
-%RSC: If needed we can remove it.
-This paper is organized as follows. Section 2 presents a
-background on exception handling mechanism. Section 3 presents the study design.
-Section 4 reports study findings. Section 5 provides further discussions and insights.
-Section 6 presents the threats to validity associated to this study. Finally Section
-7 describes the related work, and Section 8 presents our conclusions and
-directions for future work.
 
@@
 
+\subsection{Tentative  Introductions to chance the focus}
+
+Along the recent years we have witnessed a rapid increase in 
+the number of open source software  available on the Web.
+Since such projects make their control version and 
+issue tracking systems publicly available, they represent a valuable 
+source of information for the "state of practice" on software development
+which had already started to be investigated [][][][][][I still need to collect these references].
+
+In this scenario Github emerged as one of the most used repository hosting sites.
+It  alows developers to create and maintain opensource projects relying on Git distributed version 
+control and provides a set of tools which allow the information about developers 
+and their activities to be visible within and across open souce projects.
+
+Bug issues have called the attention of the research community and have been the target of increasing 
+researches [][][]. A specific type of bug issue is the one related to system 
+crashes.  Such issues usually contain an exception stack trace [ref-paper], which tipically consists 
+of an ordered list of methods that were active on the call stack before the failure occurred [ref-bug-issues].
+
+Exception stack traces are a useful source of 
+information about system crashes, and are often used to
+support developers in debugging~\cite{schroter2010stack}. Moreover, information
+from stack traces combined with crash and bug reports can be used to support bug
+classification and clustering~\cite{wang2013improving, kim2011crash,
+dhaliwal2011classifying}, fault-proneness prediction
+models~\cite{kim2013predicting} and even automated bug
+fixing~\cite{sinha2009fault} tools.
+
+Although modern applications have to deal with increasinng and similar 
+sources of failures - that arise as a consequence of faults in the application
+itself (e.g., access of null references), noisy user inputs or faults in
+underlying middleware or hardware -  on bug classificationthe current research on 
+exploring bug issue information are structured on  a system at a time basis.
+
+Specially  in a context where we have access to wealth of information available
+on open-source repositories such as GitHub, interesting knowledge may emerge 
+from a cross-project crash analysis. Hence such initial questions arise: What are the common
+characteristics of the stack traces reported on GitHub? Which patterns of exception (mis)use may emerge?
+
+To answer these questions, we conduct an exploratory study of stack traces in
+the wild. Using a custom tool called ExceptionMiner, which was specifically
+developed for this study, we mined stack traces from issues across all Java
+projects on GitHub. Overall, we analyzed 356,057 issues from 16,836 projects,
+from which 28,800 stack traces were extracted and used as a source of
+information for our analysis. The stack trace analysis  was augmented by
+additional information extracted using byte code and source code analysis done
+on a carefully selected subset of our initial sample (482 Android projects).
+%This study was divided in two phases: in the fist phase all the issues of all
+%Java projects available on GitHub were analyzed and the embedded stack traces
+%extracted using ExceptionMiner.
+
+To guide this exploration we compiled general guidelines on how to use
+exceptions proposed by Gosling~\cite{gosling2000java},
+Wirfs-Brock~\cite{wirfs2006toward} and Bloch~\cite{bloch2008effective} and then
+we investigate what stack traces can tell us about the adherence to such
+practices in Java programs.
+%~\todo{This should be our motivation, as this is
+%our biggest contribution. I already changed the text. Lets discuss about this
+%new version}
+%RSC: I AGREE.
+
+%GGG: I think this is too much detail for the introduction In the second phase
+%of this study, we combined the information of the mined stack traces with
+%source code and bytecode analysis. To do so we selected the subset of 482
+%Android repositories. The Android repositories were chosen because this kind of
+%application is known for dealing with several sources of exceptions, as they
+%are usually io-intensive (i.e., deals with Network, Bluetooth, camera, and
+%other resources), and multi-htreaded. Moreover, they are based on a single
+%platform which allowed more fine-grained analysis of their exception behaviour.
+%RSC: I AGREE.
+
+This study builds upon the following assumption: the stack traces reported on
+issues contain relevant information related to system crashes. Several studies
+and techniques have also taken this assumption as starting point
+(e.g.,~\cite{sinha2009fault}~\cite{dhaliwal2011classifying}~\cite{kim2013predicting}). Some outcomes were consistently detected through
+our large scale analysis of exception stack traces in the wild, such as:
+
+\begin{itemize}
+
+  \item  A multitude of programming mistakes were reported as the main causes of stack traces:
+  41,1\%  (12,625)  of reported stack traces were caused by programming mistakes; and 26,31\% of analyzed stack traces reported a java.lang.NullPointerException as root cause;
+
+   \item 1,671 projects (44,47\%) contained at least one issue reporting a java.lang.NullPointerException as the root cause of a stack trace.
+
+  \item Undocumented runtime exceptions signaled by third party
+    code (i.e., libraries and Android platform).
+This study found 118 distinct methods defined on third party code and  reported as the root signalers of runtime exceptions.
+Only 1 of such methods (0,8\%) documented the explicitly thrown runtime exception in a Javadoc comment; and none of these methods included the exception as part of the exception interface (i.e., using throws clause).
+
+  \item  Odd exception chains (e.g., Error wrapping checked and runtime
+    exceptions and vice-versa), indicating that the purpose of Java's hybrid
+    exception model may not have been adequately used.
+
+  \item  Issues reporting uncaught checked exceptions.
+
+\end{itemize}
+
+%Hence, the contributions of this study are as follows:
+%\begin{itemize}
+%  \item  It performs a large scale analysis of Java stack traces and how
+ %   they can reveal bad-practices on the use of exceptions in Java.
+% \item  It introduces ExceptionMiner, a tool developed to support cross-repository stack trace
+%    analysis.
+%\end{itemize}
+
+These contributions are relevant to:
+(i) developers of Java applications
+to familiarize with the most common reported exceptions mis-uses which is the
+first step to help developers to avoid making them; 
+(ii) designers of languages
+to consider ways of reducing the abundance of NullPointerExceptions;
+and (iii)
+tool designers to consider developing tools that enable a cross-project defect
+analysis.
+
+% AVD: We have not looked at private issues, and the tool is described above.
+%We strongly believe that, specially in the open-source environment,
+%faults are not to be hidden in a private bug issue. Faults should be shared and
+%discussed, so that a developer can learn from other projects mistakes.
+%Currently, the search facilities of repositories are very limited, and the
+%ExceptionMiner tool is a contribution in this direction.
+
+%GGG: Personally, I hate this paragraph in all papers as it usually reads like a space-filler
+%RSC: If needed we can remove it.
+This paper is organized as follows. Section 2 presents a
+background on exception handling mechanism. Section 3 presents the study design.
+Section 4 reports study findings. Section 5 provides further discussions and insights.
+Section 6 presents the threats to validity associated to this study. Finally Section
+7 describes the related work, and Section 8 presents our conclusions and
+directions for future work.
+
 \bibliographystyle{IEEEtran}

### 2014-07-08 android-stacks.tex
smal change on a subsection title
changing CLU description

@@
 
-In languages such as CLU\footnote{In CLU,  if a handler is 
-found for a given exception, it is converted in a predened default exception, 
-called failure which is the only one which can be implicitly 
-propagated in this model [ref-clue].}, Modula-3, Guide, and Extended Ada the exceptions are 
-checked, which means that the 
-compiler statically checks if appropriate handlers are provided within the system. 
+In languages such as Modula-3, Guide, and Extended Ada the exceptions are 
+checked, which means that the compiler statically checks if appropriate handlers are provided within the system. 
 Accordingly, the programmer must explicitly specify every exception that can be
@@
 
-In languages such as Lore, C++, and Arche all exceptions are unchecked but the developer can optionally list exception types on method signatures. Although exceptions can 
+In languages such as Lore, C++, and Arche all exceptions are unchecked but the developer can 
+optionally list exception types on method signatures. Although exceptions can 
 be documented on methods signatures, the developer is not warned by the compiler
@@
 
-In Java a method can throw either checked or unchecked exceptions.
-Java is the only language that allows a hybrid approach on which the developer can define exceptions of both kinds to be thrown and handled in a single system.
-Next section details Java exception handling model.
-
-Although exception handling mechanism is the most used technique to deal with such 
-failures - been embedded in several programming languages as show above,   
-the exception handling code is often generally poorly understood and the least
-tested part of software systems.This behaviour may lead to the problem of uncaught exceptions~\cite{jo2004uncaught}, one of the main causes of application crashes
-(e.g.~\cite{miller1997issues,Robil00,shah2010understanding, garcia2007extracting, garcia2001comparative,cabral2007exception,coelho2011unveiling}).
+Both Java [ref: Gosling et al. 1996] and CLU [Liskov and Snyder 1979] support the declaration of exceptions
+ in method signatures, hence the compiler can then check if handlers are provided to such exceptions.
+ However, this support is only partial because both languages also provides a form of unchecked exceptions.
+In CLU,  if a handler is found for a given exception, it is converted in a predened default exception, 
+called failure which is the only one which can be implicitly propagated in this model [ref-clue].
+ Java allows a hybrid approach on which the developer can define specific exceptions of both kinds 
+(checked or unchecked) to be thrown and handled inside the system.
+Therefore, in both languages the developer of a client module is not warned by the compiler of the possibility of 
+an unchecked exception being thrown.
+
+[REF]: Static Analysis to Support the Evolution of Exception Structure in Object-Oriented Systems
+
+Although exception handling mechanism has been embedded in several programming languages 
+as show above,   the exception handling code is often generally poorly understood and the least
+tested part of software systems.
+
+% This behaviour may lead to the problem of uncaught exceptions~\cite{jo2004uncaught}, one of the main causes of application crashes
+%(e.g.~\cite{miller1997issues,Robil00,shah2010understanding, garcia2007extracting, garcia2001comparative,cabral2007exception,coelho2011unveiling}).
 
@@
 
-\subsubsection{Stack Traces on Issues}
+\subsubsection{Exception Stack Traces on Issues}
 
@@
 Programming errors are the causes of most of the exception stack traces reported
-on issues. In the Android subset we could find more issues related to Backward
-compatibility and security issues than on the general data set.
-
-\subsection{Investigating Android Subset}
+on issues both in  general Java data set and Android subset. In the Android subset we could find 
+more stack traces reporting exceptions related to Backward
+compatibility and security concerns than on the general data set.
+
+TODO: Discuss about the interater agreement here...
+
+\subsection{Investigating The Android Subset}
 

### 2014-07-11 android-stacks.tex
changing java and CLU description

@@
 
-Both Java [ref: Gosling et al. 1996] and CLU [Liskov and Snyder 1979] support the declaration of exceptions
- in method signatures, hence the compiler can then check if handlers are provided to such exceptions.
+Both Java [ref: Gosling et al. 1996] and CLU [Liskov and Snyder 1979] support the checked exceptions,
+ hence the compiler can then check if handlers are provided to such exceptions.
  However, this support is only partial because both languages also provides a form of unchecked exceptions.
 In CLU,  if a handler is found for a given exception, it is converted in a predened default exception, 
-called failure which is the only one which can be implicitly propagated in this model [ref-clue].
- Java allows a hybrid approach on which the developer can define specific exceptions of both kinds 
-(checked or unchecked) to be thrown and handled inside the system.
-Therefore, in both languages the developer of a client module is not warned by the compiler of the possibility of 
-an unchecked exception being thrown.
+called failure which is the only one which can be implicitly propagated [ref-clue].
+ Java allows a hybrid approach on which the developer can define specific unchecked and checked exceptions,
+and use both kinds of exceptions inside a single system.
+In both languages the developer of a client module is not warned by the compiler if 
+ unchecked exceptions are thrown by a module.
 

### 2014-07-11 android-stacks.tex
Cross-coding of exception types

--- 2014-07-07 android-stacks.tex
@@
 
-To better understand the root causes reported on stacks, we performed a categorization 
-study on which the top 100 root causes reported on stack-issues were manually inspected.
-The characterization process consisted of classifiying each exception in one of the
-chategories listed on Table X, based on the inspection of source code, Javadoc, 
-or additional documentation related to the exception.
+To better understand the root causes reported on stacks, we categorized the top
+100 distinct exception types in in one of the chategories listed on
+Table~\ref{tab:categories}, based on the inspection of source code, Javadoc, or
+additional documentation related to the exception. To ensure the quality
+of the classification, three independent coders classified a randomly selected
+sample of 25 exception types (from the total 100) using the same categories;
+the interater agreement was high (96\%). We can therefore assume that the
+categories are descriptive enough for covering the whole range of exception
+types that can occur in real world Java projects.
 
@@
   \end{tabular}
-  \caption{Categories description.}
+  \caption{Description of categories.}
   \label{tab:categories}

### 2014-07-11 android-stacks.tex

--- 2014-07-11 android-stacks.tex
@@
 
-[REF]: Static Analysis to Support the Evolution of Exception Structure in Object-Oriented Systems
-
 Although exception handling mechanism has been embedded in several programming languages 
 as show above,   the exception handling code is often generally poorly understood and the least
-tested part of software systems.
+tested part of software systems. There is a long-lasting debate about the pros and cons of checked and unchecked exceptions 
+~\cite{javatut,stackoverlow,debate},\footnote{152 questions in Stackoverlow
+are related to this debate}. The debates, however, are based on the developers own experience in
+using both models. There is a lack of empirical study tackling the pros and cons of each model.
+
+[REF]: Static Analysis to Support the Evolution of Exception Structure in Object-Oriented Systems
 
@@
 callers of a method will know about the exceptions that a method can throw, and
-so they can decide what to do about them. However, there is a long-lasting
-debate about the use of checked and unchecked exceptions in
-Java~\cite{javatut,stackoverlow,debate},\footnote{152 questions in Stackoverlow
-are related to this debate} since there are pros and cons associated to each of
-them.
-
-Hence, in Java an exception can be thrown in one of the following
+so they can decide what to do about them.
+
+In Java an exception can be thrown in one of the following
 circumstances~\cite{gosling2000java}: (i) explicitly thrown when a throw statement is executed; 

### 2014-07-11 android-stacks.tex
ckecked and unckeded debate

@@
 
+[REF]: Static Analysis to Support the Evolution of Exception Structure in Object-Oriented Systems
+
 Although exception handling mechanism has been embedded in several programming languages 
 as show above,   the exception handling code is often generally poorly understood and the least
-tested part of software systems. There is a long-lasting debate about the pros and cons of checked and unchecked exceptions 
+tested part of software systems.
+
+There is a long-lasting debate about the pros and cons of checked and unchecked exceptions 
 ~\cite{javatut,stackoverlow,debate},\footnote{152 questions in Stackoverlow
@@
 using both models. There is a lack of empirical study tackling the pros and cons of each model.
-
-[REF]: Static Analysis to Support the Evolution of Exception Structure in Object-Oriented Systems
 
@@
 
-To better understand the root causes reported on stacks, we categorized the top
-100 distinct exception types in in one of the chategories listed on
-Table~\ref{tab:categories}, based on the inspection of source code, Javadoc, or
-additional documentation related to the exception. To ensure the quality
-of the classification, three independent coders classified a randomly selected
-sample of 25 exception types (from the total 100) using the same categories;
-the interater agreement was high (96\%). We can therefore assume that the
-categories are descriptive enough for covering the whole range of exception
-types that can occur in real world Java projects.
+To better understand the root causes reported on stacks, we performed a categorization 
+study on which the top 100 root causes reported on stack-issues were manually inspected.
+The characterization process consisted of classifiying each exception in one of the
+chategories listed on Table X, based on the inspection of source code, Javadoc, 
+or additional documentation related to the exception.
 
@@
   \end{tabular}
-  \caption{Description of categories.}
+  \caption{Categories description.}
   \label{tab:categories}

### 2014-07-11 android-stacks.tex
ckecked and unckeded debate improvements
including examples of strange wrappings.

@@
 
-[REF]: Static Analysis to Support the Evolution of Exception Structure in Object-Oriented Systems
-
 Although exception handling mechanism has been embedded in several programming languages 
 as show above,   the exception handling code is often generally poorly understood and the least
-tested part of software systems.
-
-There is a long-lasting debate about the pros and cons of checked and unchecked exceptions 
+tested part of software systems. There is a long-lasting debate about the pros and cons of checked and unchecked exceptions 
 ~\cite{javatut,stackoverlow,debate},\footnote{152 questions in Stackoverlow
@@
 using both models. There is a lack of empirical study tackling the pros and cons of each model.
+
+[REF]: Static Analysis to Support the Evolution of Exception Structure in Object-Oriented Systems
 
@@
 from.
+
+\subsubsection{Checked Exception wrapping Error}
+
+We could observe that in 4 projects this strange wrapping happend 7 times in total.
+
+\subsubsection{Error wrapping Checked Exception}
+
+We could observe that in 7 projects this strange wrapping happend 8 times in total.
+Inspecting the stack traces we could observe that 7 out of 8 wrappings were caused
+by the reflection java library on which "java.lang.OutOfMemoryError"
+"java.lang.NoSuchMethodError" were wrapped into a "java.lang.reflect.InvocationTargetException".
+
+In one of the cases, however, in project OpenDocument.droid a java.lang.Exception was 
+wrappings a java.lang.OutOfMemoryError.
+This general handlers is a known bad practice and here we could observe a bad consequence of
+such bad practice.
+
+\begin{table}
+\centering
+\begin{tabular}{llll}
+    \hline
+    \bfseries{Project} & \bfseries{Stars} & \bfseries{Watchers} & \bfseries{Forks} \\
+    \hline
+androidremocons &  9  &  2    & 10 \\
+android-timepicker &  73  &  13    & 32 \\
+PathfinderOpenReference &  23  &  12   & 15 \\
+ActionBarSherlock &  5926  &   767  & 3356 \\
+OpenDocument.droid &  40  &  9   & 12 \\
+TouchDB-Android &  -  &  -    & - \\
+ContractionTimer &  -  &  -    & - \\
+\hline
+  \end{tabular}
+\caption{Projects containing wrappings (Error-Checked)}
+\label{tab:wrappings2}
+\end{table}
+
+\subsubsection{Runtime Exception wrapping Error}
+
+\subsubsection{Error wrapping Runtime Exception}
+
+\subsubsection{Runtime wrapping Checked Exception}
+
+\subsubsection{Checked wrapping Runtime Exception}
 

### 2014-08-13 android-stacks.tex
including more detailed examples of strange wrappings.

--- 2014-07-11 android-stacks.tex
@@
 \begin{document}
-\title{What Can Stack Traces Tell us about of the use Exceptions in Java?}
+\title{ What can Stack Traces tell us about Java Exceptions? }
+
+%\title{A first Glance of what is causing Mobile Crashes: A GitHub and Googlecode Repository Mining Study }
 
@@
 \section{Introduction}
-
-\subsection{Georgios's and Roberta's Merged Introduction}
 
@@
 
-\subsection{Error Handling in Programming Languages}
+\subsection{Exception Handling in Programming Languages}
 
 Techniques for error detection and handling are not  an optional add-on but a
-fundamental part of any system~\cite{bruntink2006discovering}. Several
-approaches exist for detecting errors and responding to them, such as the return-code 
-idiom (widely used in systems written in C), and the exception handling 
-mechanism, embedded in many main-stream programming languages.
-The main disadvantage of return codes is the fact that the programmer can just
-ignore the return code and let the program execute in inconsistent state. 
+fundamental part of any system~\cite{bruntink2006discovering}. 
 The exception mechanism~\cite{goodenough1975exception} was proposed as a way of
@@
 In languages such as Modula-3, Guide, and Extended Ada the exceptions are 
-checked, which means that the compiler statically checks if appropriate handlers are provided within the system. 
-Accordingly, the programmer must explicitly specify every exception that can be
-signaled to support the compile time checking. This approach is known to be
+checked, which means that the compiler statically checks if appropriate handlers 
+are provided within the system. Accordingly, the programmer must explicitly specify 
+every exception that can be signaled to support the compile time checking. This approach is known to be
 costly to maintain (Dooren and Steegmans, 2005) since exception updates are
@@
 
-Both Java [ref: Gosling et al. 1996] and CLU [Liskov and Snyder 1979] support the checked exceptions,
- hence the compiler can then check if handlers are provided to such exceptions.
- However, this support is only partial because both languages also provides a form of unchecked exceptions.
-In CLU,  if a handler is found for a given exception, it is converted in a predened default exception, 
+Java [ref: Gosling et al. 1996] and CLU [Liskov and Snyder 1979] support the checked exceptions,
+ (therefore the compiler can then check if handlers are provided to such exceptions), however, this support is 
+only partial because both languages also provides a form of unchecked exceptions [ref: Robilard].
+In CLU,  if a handler is not found for a given exception, it is automaticaly converted in a predened default exception, 
 called failure which is the only one which can be implicitly propagated [ref-clue].
- Java allows a hybrid approach on which the developer can define specific unchecked and checked exceptions,
-and use both kinds of exceptions inside a single system.
-In both languages the developer of a client module is not warned by the compiler if 
- unchecked exceptions are thrown by a module.
-
-[REF]: Static Analysis to Support the Evolution of Exception Structure in Object-Oriented Systems
-
-Although exception handling mechanism has been embedded in several programming languages 
-as show above,   the exception handling code is often generally poorly understood and the least
-tested part of software systems.
+ In Java the developer can define specific unchecked and checked exceptions and use both kinds in a single system.
 
@@
 ~\cite{javatut,stackoverlow,debate},\footnote{152 questions in Stackoverlow
-are related to this debate}. The debates, however, are based on the developers own experience in
-using both models. There is a lack of empirical study tackling the pros and cons of each model.
+are related to this debate}. However, the debates are mainly based on the developers own experience in
+using both models, since there is still few empirical evidence of the pros and cons of each model.
+
+%Although exception handling mechanism has been embedded in several programming languages 
+%as show above,  the exception handling code is often poorly understood and the least
+%tested part of software systems [ref:POORLY].
+
+%[REF]: Static Analysis to Support the Evolution of Exception Structure in Object-Oriented Systems
 
@@
 (extends Exception), run-time exceptions (extends RuntimeException), and errors
-(extends Error)~\cite{gosling2000java}.
+(extends Error)~\cite{gosling2000java}. Neither run-time exceptions nor errors need to be specified on method
+signatures, and hence both are referred as unchecked exceptions.
 
@@
 
-Neither run-time exceptions nor errors need to be specified on method
-signatures, and hence both are referred as unchecked exceptions. The Java
-Virtual Machine represents as runtime exceptions invalid operations detected in
+The Java Virtual Machine represents as runtime exceptions invalid operations detected in
 the program (out-of-bounds array index, divide-by-zero error, null pointer
@@
 from failures detected by the Java Virtual Machine, such as OutOfMemoryError and
-normally should not be handled inside the application.
-
-The Java specification~\cite{gosling2000java} suggests that user-defined
-exceptions should be checked exceptions. The main reason is because doing so the
+normally should not be handled inside the application. The Java specification~\cite{gosling2000java} 
+suggests that user-defined exceptions should be checked exceptions. The main reason is because doing so the
 callers of a method will know about the exceptions that a method can throw, and
@@
 
-In Java an exception can be thrown in one of the following
-circumstances~\cite{gosling2000java}: (i) explicitly thrown when a throw statement is executed; 
+Hence, in Java an exception can be thrown in one of the following
+circumstances~\cite{gosling2000java}: 
+(i) explicitly thrown when a throw statement is executed; 
 (ii) implicitly thrown by the JVM when the evaluation of an expression
- violates the normal semantics of language (e.g., out-of-bounds array index,
-division-by-zero, access to a null reference); or (iii) implicitly thrown by
-the JVM due to an internal error or resource limitation (e.g.,
+ violates the normal semantics of language, also refered as a coding error
+ (e.g., out-of-bounds array index, division-by-zero, access to a null reference); or 
+(iii) implicitly thrown by the JVM due to an internal error or resource limitation (e.g.,
 OutofMemoryError).
@@
 
-When an exception is throw it causes a transfer of control from the point
+When an exception is throw, it causes a transfer of control from the point
 where the exception occurred to a point that can be specified by the programmer
-(exception handler) - in Java it is represented by the try-catch block .
-
-A common way of  propagating exceptions in Java programs, is the exception wrapping (also known as exception chaining), which allows one exception to be wrapped in another exception and re-thrown. 
+(exception handler) - in Java it is represented by the try-catch block . A common way of  
+propagating exceptions in Java programs is through exception wrapping
+ (also known as exception chaining). Exception wrapping allows one exception to be wrapped in another one and re-thrown. 
 Figure~\ref{fig:wrapping} presents a stack trace associated which illustrates an
@@
 
-\subsection{Exception Handling Best Practices in Java}
-\label{sec:best}
-
-Several general guidelines have been proposed on how to use Java
-exceptions~\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,
-bloch2008effective}. 
-A list of the guidelines related to exception types and exception
-propagation in Java the following:\footnote{We also compiled guidelines related
-to exception handling but they are out of the scope of this paper.}
-
-%\noindent\emph{Meaning of Exception Types}
-
-\textbf{I-Use checked exceptions to represent recoverable
-conditions} (\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,bloch2008effective})
-The developer should use checked exceptions for conditions from which the caller
-is expected to recover. By confronting the API user with a checked exception,
-the API designer is forcing the client to handle the exceptional condition. The
-client can explicitly ignore the exception (swallowing, or converting it to
-other type) at the expense of the program's robustness~\cite{gosling2000java}.
-
-\textbf{II-Error represents an unrecoverable condition detected by the JVM which
-should not he handled} (\cite{gosling2000java}). It results from failures detected
-by the Java Virtual Machine which indicate resource deficiencies, invariant
-failures or other conditions that make impossible the program to recover.
-
-%\noindent\emph{Exception Throwing}
-
-\textbf{III-A method should throw exceptions that precisely define the
-exceptional condition} (\cite{gosling2000java,bloch2008effective}). To do so,
-developers should either try to reuse the exception types already defined in the
-JVM or they should create a specific exception. Throwing general types such as a
-pure Exception or a RuntimeException is hence considered a bad practice.
-
-%\textbf{IV-Do not repeatedly re-throw exceptions, handle exceptions as close as
-%possible to the problem}~\cite{wirfs2006toward}. 
-%Programs that frequently throw exceptions lose in performance ~\cite{wirfs2006toward,gosling2000java}.
-%Moreover, 
-%Close to the place that the exception was throw the caller usually has
-%enough contextual information to perform  corrective actions. When the exception
-%is propagated far away from the method where it was signaled (i.e. further away
-%#from the problem) it can be difficult to take meaningful recovery actions.
-%*~\todo{RSC:this practice was not well explored, can be removed if we need space.}
-
-%\noindent\emph{Exception Documentation}
-\textbf{ IV-Document all exceptions thrown by a
-method} (\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,bloch2008effective}).
-The exceptions thrown by a method are an important part of methods interface,
-and is required to use the method properly. The checked exceptions are already
-part of the  methods signature, and the method caller is aware of the checked
-exceptions being thrown by it. According to~\cite{bloch2008effective}, it is
-also wise to document the explicitly thrown runtime exceptions\footnote{This
-excludes the implicitly signaled JVM runtime exceptions related to programming
-mistakes} as carefully as checked exceptions. Doing so, the clients of a method
-will be aware of all the exceptions the method can throw. If the developer fails to
-do follow this practice (specially when developing library code) it will be
-difficult or even impossible for the caller to make effective use of such 
-method~\cite{wirfs2006toward, bloch2008effective}. As mentioned in Section II, 
-in some languages such as C
-
-These best practices guided our exploratory study described next.
-Based on information extracted from stack traces (complemented by
-bytecode and source-code analysis), we investigate whether stack characteristics
-can reveal whether these practices have been obeyed. 
-%Next section discusses the
-%study procedure and the practices' misuses that emerged from this study.
-
 \section{Study Design}
@@
 
-The main goal of this study was to gain a better understanding
-of the use of exceptions in Java programs based mainly on the information
-available on exception stack traces embedded on Github issues.
-The exceptions available on exception stack trace are not limited to the exceptions 
+Based mainly on the information available on exception stack traces embedded 
+on Github issues, the goal of this study was to gain a better understanding 
+of exception-related system crashes.
+
+The exceptions available on exception stack traces are not limited to the exceptions 
 expliciltly thrown by the Java developer but also include
-the exceptions implicitly thrown by the JVM during the program execution,
-due to a programming error (e.g., out-of-bounds array index, division-by-zero, 
-access to a null reference) or an internal error or resource limitation (e.g.,
-OutofMemoryError).
+the exceptions implicitly thrown by the JVM due to programming errors (e.g., out-of-bounds array index, division-by-zero, 
+access to a null reference) or resource limitation (e.g., OutOfMemoryError).
 
@@
 
-  \item[RQ1] What are the common characteristics of such stack traces?
-
-  \item[RQ2] What patterns of exception (mis)-use emerge from the stack trace
+  \item[RQ1] What are the common characteristics of exception stack traces reported on issues?/What are the main causes of crashes reported on issues? (more explicit?)
+
+  \item[RQ2] What patterns of exception (mis)-use emerge from the exception stack trace
     analysis?
@@
 
+\subsection{Meaning of Exception Types and related Best Practices}
+\label{sec:best}
+
+Several general guidelines have been proposed on how to use Java
+exceptions~\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,
+bloch2008effective}. 
+We compiled a list of the guidelines related to exception types and exception
+propagation in Java the following:\footnote{We also compiled guidelines related
+to exception handling but they are out of the scope of this paper.}
+
+%\noindent\emph{Meaning of Exception Types}
+
+\textbf{I-Use checked exceptions to represent recoverable
+conditions} (\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,bloch2008effective})
+The developer should use checked exceptions for conditions from which the caller
+is expected to recover. By confronting the API user with a checked exception,
+the API designer is forcing the client to handle the exceptional condition. The
+client can explicitly ignore the exception (swallowing, or converting it to
+other type) at the expense of the program's robustness~\cite{gosling2000java}.
+
+\textbf{II-Error represents an unrecoverable condition detected by the JVM which
+should not he handled} (\cite{gosling2000java}). It results from failures detected
+by the Java Virtual Machine which indicate resource deficiencies, invariant
+failures or other conditions that make impossible the program to recover.
+
+%\noindent\emph{Exception Throwing}
+
+\textbf{III-A method should throw exceptions that precisely define the
+exceptional condition} (\cite{gosling2000java,bloch2008effective}). To do so,
+developers should either try to reuse the exception types already defined in the
+JVM or they should create a specific exception. Throwing general types such as a
+pure Exception or a RuntimeException is hence considered a bad practice.
+
+%\textbf{IV-Do not repeatedly re-throw exceptions, handle exceptions as close as
+%possible to the problem}~\cite{wirfs2006toward}. 
+%Programs that frequently throw exceptions lose in performance ~\cite{wirfs2006toward,gosling2000java}.
+%Moreover, 
+%Close to the place that the exception was throw the caller usually has
+%enough contextual information to perform  corrective actions. When the exception
+%is propagated far away from the method where it was signaled (i.e. further away
+%#from the problem) it can be difficult to take meaningful recovery actions.
+%*~\todo{RSC:this practice was not well explored, can be removed if we need space.}
+
+%\noindent\emph{Exception Documentation}
+\textbf{ IV-Document all exceptions thrown by a
+method} (\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,bloch2008effective}).
+The exceptions thrown by a method are an important part of methods interface,
+and is required to use the method properly. The checked exceptions are already
+part of the  methods signature, and the method caller is aware of the checked
+exceptions being thrown by it. According to~\cite{bloch2008effective}, it is
+also wise to document the explicitly thrown runtime exceptions\footnote{This
+excludes the implicitly signaled JVM runtime exceptions related to programming
+mistakes} as carefully as checked exceptions. Doing so, the clients of a method
+will be aware of all the exceptions the method can throw. If the developer fails to
+do follow this practice (specially when developing library code) it will be
+difficult or even impossible for the caller to make effective use of such 
+method~\cite{wirfs2006toward, bloch2008effective}. As mentioned in Section II, 
+in some languages such as C
+
+These best practices guided our exploratory study described next.
+Based on information extracted from stack traces (complemented by
+bytecode and source-code analysis), we investigate whether stack characteristics
+can reveal whether these practices have been obeyed. 
+%Next section discusses the
+%study procedure and the practices' misuses that emerged from this study.
+
 \section{RQ1: What are the common characteristics of such stack traces?}
@@
 Based on the assumption that regardless the issue type, every exception stack
-trace contains relevant information concerning the exception structure of
-projects analyzed we opted for not restricting the analysis only on defect
+trace contains relevant information concerning the exception structure of the
+projects analyzed, we opted for not restricting the analysis only on defect
 issues.\footnote{We conducted the same analysis on the defect issues and the top
@@
 
-After distilling the information available on the 28,800 stack traces were extracted from the 
-issues and issues comments of 16,836 Java projects: 607 distinct root causes were found, 
-and  2,026 distinct top level exceptions were found.
-
-From this set the java.lang.NullPointerException  was the root cause most 
-reported;  7578 stack traces had a java.lang.NullPointerException  as root cause,
- which corresponds of 26,31\%  of all analyzed stack traces.
-
-If we perform a cross-project analysis, the java.lang.NullPointerException, we could observe that such 
-exception was reported on 1,671 projects (which represents 44,47\% of all projects reporting at least one stack trace).
-
-The high prevalence of NullPointerExceptions is aligned with the findings other works~\cite{kim2013predicting,fraser20131600,csallner2004jcrasher}. For instance, Sunghun et
-al.~\cite{kim2013predicting} showed that in Eclipse bug report system 38\% the bugs related to exception handling were caused by NullPointerException; other works on robustness testing~\cite{maji2012empirical,csallner2004jcrasher} showed that most of the automatically detected bugs were due to NullPointerExceptions and implicitly-signaled of Java
+After distilling the information available on the 28,800 stack traces (extracted from the 
+issues and issues comments of 16,836 Java projects) we found 607 distinct exceptions
+reported as the root causes of stack traces.
+
+Table \ref{tab:toptenjava} illustrates the top 10 root couses found in the study ranked by the number of distinct
+project repositories on which they were reported. The table illustrates the
+number of times each exception was reported as the root cause of a stack trace
+, the number of distinct projects reporting such causes on stack traces, and the percentages of
+such in relation to all reported exceptions and analyzed projects respectaly.
+
+\begin{table}
+\begin{tabular}{lrrrr}
+    \hline
+    \bfseries{Exception} &  \multicolumn{2}{c}{\bfseries{Occurrences}} &  \multicolumn{2}{c}{\bfseries{Projects}}\\
+    & \bfseries{\#} &  \bfseries{\%} & \bfseries{\# } & \bfseries{\% } \\
+    \hline
+java.lang.NullPointerException             & 7578 & 26,31\% & 1671 & 44,47 \\
+java.lang.IllegalArgumentException         & 1573 & 5,46\%  & 670  & 17,83\% \\
+java.lang.ClassNotFoundException           & 1203 & 4,18\%  & 514  & 13,68\% \\
+java.lang.RuntimeException                 & 1093 & 3,80\%  & 411  & 10,94\% \\
+java.lang.IllegalStateException            & 964  & 3,35\%  & 431  & 11,47\% \\
+java.lang.ClassCastException               & 886  & 3,08\%  & 408  & 10,86\% \\
+java.lang.ArrayIndexOutOfBoundsException   & 834  & 2,90\%  & 345  & 9,18\% \\
+java.lang.NoSuchMethodError                & 822  & 2,85\%  & 369  & 9,82\% \\
+java.io.IOException                        & 656  & 2,28\%  & 337  & 8,97\% \\
+java.lang.OutOfMemoryError                 & 601  & 2,09\%  & 235  & 6,25\% \\
+%java.lang.NoClassDefFoundError            & 554  & 1,92\%  & 299  & 7,96\% \\
+%java.util.ConcurrentModificationException & 465  & 1,61\%  & 180  & 4,79\% \\
+%java.lang.IndexOutOfBoundsException       & 445  & 1,55\%  & 205  & 5,46\% \\
+\hline
+All	 &	23956 &	100\%	& 3758 & 100\%	 \\
+\hline
+  \end{tabular}
+\caption{Top 10 most popular exceptions on issues in Java projects}
+\label{tab:toptenjava}
+\end{table}
+
+We can observe that most of this set is composed by exceptions defined by java.lang,
+which are usually implicitly thrown by JMV dur to a programming mistake. From thi set
+the java.lang.NullPointerException  was the most reported root cause;  7578 stack traces 
+had a java.lang.NullPointerException  as root cause, which corresponds of 26,31\%  of 
+all analyzed stack traces. Considering the number of distinct project repositories, 
+the java.lang.NullPointerException, was reported on at least one issue of 1,671 projects 
+(which represents 44,47\% of all projects reporting at least one stack trace).
+
+The high prevalence of NullPointerExceptions is aligned with the findings other 
+works~\cite{kim2013predicting,fraser20131600,csallner2004jcrasher}. For instance, Sunghun et
+al.~\cite{kim2013predicting} showed that in Eclipse bug report system 38\% the bugs 
+related to exception handling were caused by NullPointerException; other works on robustness 
+testing~\cite{maji2012empirical,csallner2004jcrasher} showed that most of the automatically 
+detected bugs were due to NullPointerExceptions and implicitly-signaled of Java
 environment (as the ones found in this study).
 
-To better understand the root causes reported on stacks, we performed a categorization 
-study on which the top 100 root causes reported on stack-issues were manually inspected.
-The characterization process consisted of classifiying each exception in one of the
-chategories listed on Table X, based on the inspection of source code, Javadoc, 
-or additional documentation related to the exception.
+In Section XX we provide further discussion on the NullPointerException problem, pointing 
+to ways of how to deal with it and mitigate it.
+
+\subsubsection{Classifying Top 100 Exceptions}
+
+To get a broather view of the root causes of stack traces, we performed an empirical evaluation
+whose goal was to categorized the top 100 root causes.
+
+After a first inspection of the complete list of exceptions reported, a first set of exception 
+categories was defined. This list represent the most common terms reffered on exception
+names and javadoc documentation. A similar approach was addopted in [COELHO-ECOOP].
+The list of categories is illustrated in Table~\ref{tab:categories}. Hence in the second phase
+of the study, each exception was mapped to one of the categories previously defined - 
+this analysis was based on the inspection of source code, Javadoc, or additional 
+documentation related to the exception.
+
+To ensure the quality of the classification, three independent coders classified a randomly selected
+sample of 25 exception types (from the total 100) using the same categories;
+the interater agreement was high (96\%). We can therefore assume that the
+categories are descriptive enough for covering the whole range of exception
+types that can occur in real world Java projects.
 
@@
     \hline
-    \bfseries{Root Cause} & \bfseries{Wrapper Type} & \bfseries{Occurrences} & \bfseries{Projects} \\
-    \hline
-     Runtime    & -   &  2360 & 359 \\
-
+    \bfseries{Wrapper Type} & \bfseries{Root Cause} & \bfseries{Occurrences} & \bfseries{Projects} \\
+    \hline
+      none   & Runtime   &  2360 & 359 \\
+      none  &   Checked  & 422  & 117 \\
+      none  &   Error   & 381  & 141 \\
+      none &  Undefined  &  15   & 10  \\
+      none & Throwable  &  1    & 1 \\
+     \hline
       Runtime   & Runtime & 560  & 182 \\    
-
- Checked    &   -  & 422  & 117 \\
-     Error    &   -   & 381  & 141 \\
-   Checked &  Runtime   & 109  & 70  \\
       Checked   & Checked & 98   & 42  \\
-   Error   &  Runtime   & 55   & 38  \\
-   Runtime &  Checked   & 22   & 12  \\
       Error     & Error   & 15   & 14  \\
-  Undefined    &  -  &  15   & 10  \\
-   Runtime & Error     &  13   & 6 \\
-  Error   & Checked   &  8    & 7 \\
-    Checked &Error     &  7    & 4 \\
-    Runtime &Throwable &  3    & 1 \\
-   Runtime & Undefined &  3    & 3 \\
-   Throwable    & -  &  1    & 1 \\
-   Error   & Undefined &  1    & 1 \\
+    \hline
+      Runtime &  Checked   & 114  & 70  \\
+      Runtime   &  Error   & 57   & 38  \\
+      Checked &  Runtime   & 22   & 12  \\
+      Checked & Error    &  8    & 7 \\
+      Error & Checked     &  7    & 4 \\
+      Error & Runtime     &  13   & 6 \\
+      Throwable & Runtime &  3    & 1 \\
+      Undefined & Runtime &  3    & 3 \\
+      Undefined & Error &  1    & 1 \\
     \hline
@@
 
+\subsubsection{Runtime wrapping Checked Exception}
+
+Such wrapping was detected on distinct 70 projects. Inspecting the elements responsible for performing such wrapping we could observe that,
+most of such wrappings were perform on methods defined on Android.
+
+ANDROID	 \- 66
+APP	\- 29 	(HOW MANY?? AND WHICH??)
+LIB  \- 19 (HOW MANY AND WHICH?)
+
+Taking a closer look on the signalers reposponsible for wrapping we could observe that, aproximately 41%
+of all wrappings were performed by a single class, called ActivityThread, responsible for managing the life cycle 
+of an Android application. It performs the application callbacks.
+
+android.app.ActivityThread \- 47
+android.view.LayoutInflater \- 7
+org.gege.caldavsyncadapter.syncadapter.SyncAdapter \- 7
+org.springframework.web.client.RestTemplate \- 6
+android.os.AsyncTask \- 5
+
+Code snipet:
+
+android.app.ActivityThread.
+
+public final Activity startActivityNow(...) \{
+...
+return performLaunchActivity(r, null);
+\}
+
+private Activity performLaunchActivity(...) \{
+try \{
+.....
+\} catch (Exception e) {
+..
+throw new RuntimeException(
+"Unable to instantiate activity " + component
++ ": " + e.toString(), e);
+..
+\}
+\}
+
+The exceptions most wrapped on RuntimeExceptinos were:
+
+(java.lang.RuntimeException,java.lang.ClassNotFoundException)	32
+(java.lang.RuntimeException,java.lang.NoSuchMethodException)	6
+(java.lang.RuntimeException,java.lang.InstantiationException)	5
+(com.github.andlyticsproject.console.NetworkException,org.apache.http.client.HttpResponseException)	5
+(android.view.InflateException,java.lang.NoSuchMethodException)	4
+
+java.lang.RuntimeException	69 \- 61%
+android.view.InflateException	10 \- 8,7%
+com.github.andlyticsproject.console.NetworkException	7
+java.lang.IllegalStateException	6
+org.springframework.web.client.ResourceAccessException	6
+com.github.andlyticsproject.console.DevConsoleProtocolException	3
+
+\subsubsection{Runtime Exception wrapping Error}
+
+ We could find  57 stack traces on 38 projects on which an Error was wrapped in a Runtime Exception.
+
+ANDROID	50
+LIB 	6
+APP	1
+
+android.os.AsyncTask \-	25
+android.app.ActivityThread \- 15
+android.view.LayoutInflater \- 7
+org.asynctasktex.AsyncTaskExd \- 3
+android.support.v4.content.ModernAsyncTask \- 2
+
+EXAMPLE: android.os.AsyncTask
+
+@Override
+protected void done() \{
+...
+try \{
+...
+
+\} catch (InterruptedException e) \{
+android.util.Log.w(.., e);
+\} catch (ExecutionException e) \{
+throw new RuntimeException("An error occured while executing doInBackground()",
+e.getCause());
+\} catch (CancellationException e)  \{
+...
+\} catch (Throwable t) \{
+throw new RuntimeException("An error occured while executing "
++ "doInBackground()", t);
+\}
+...
+\}
+\}
+
+%SELECT (exception, root_cause) AS PAIR, COUNT (*) AS N FROM stacktracesummary where 
+%EXCEPTION_EXTENDS='RUNTIME' AND ROOT_EXTENDS='ERROR' GROUP BY PAIR ORDER BY N DESC;
+
+(java.lang.RuntimeException,java.lang.OutOfMemoryError) \- 30
+(java.lang.RuntimeException,java.lang.NoClassDefFoundError) \- 6
+(java.lang.RuntimeException,java.lang.UnsatisfiedLinkError) \- 5
+(android.view.InflateException,java.lang.OutOfMemoryError) \- 4
+(java.lang.RuntimeException,java.lang.StackOverflowError) \- 4
+
+TO INVESTIGATE: WHICH WERE THE LIBS AND THE APP? AND ITS REPUTATION?
+
+\subsubsection{Checked wrapping Runtime Exception}
+
+APP	9
+LIB	9
+ANDROID	4
+
+(com.octo.android.robospice.persistence.exception.CacheSavingException,java.lang.ClassCastException)	6
+(java.lang.reflect.InvocationTargetException,java.lang.NullPointerException)	3
+(java.lang.reflect.InvocationTargetException,android.content.res.ResourcesdNotFoundException)	2
+(java.sql.SQLException,android.database.sqlite.SQLiteDiskIOException)	2
+(java.lang.Exception,at.andiwand.odf2html.odf.IllegalMimeTypeException)	2
+
+The most frequent exception wrapping were were performed inside RoboSpice is an android library that supports the creation of asynchronous long running tasks.
+This library has 1.682 stars and 367 forks.
+
+Most of such wrappings however were caused by the reflection Java library, used by Android, Application and library classes.
+relaction methods convert every exception thata happens when invoking a reflected method into an INVOCATION TARGET EXCEPTION
+this conversion is performed inside native code.
+
+\subsubsection{Checked Exception wrapping Error}
+
+We could observe that in 7 projects this  wrapping happend 8 times in total.
+Inspecting the stack traces we could observe that 7 out of 8 wrappings were caused
+by the reflection java library on which java.lang.OutOfMemoryError
+java.lang.NoSuchMethodError were wrapped into a java.lang.reflect.InvocationTargetException.
+
+ANDROID	5
+APP	3
+
+(java.lang.reflect.InvocationTargetException,java.lang.OutOfMemoryError)	5
+(java.lang.Exception,java.lang.OutOfMemoryError)	1
+(java.lang.reflect.InvocationTargetException,java.lang.NoSuchMethodError)	1
+(java.lang.reflect.InvocationTargetException,java.lang.UnsatisfiedLinkError)	1
+
+All wrappings except 1 were caused by the reflection library.
+
+at java.lang.reflect.Constructor.constructNative(Native Method)
+NATIVE IMPLEMENTATION, WE DO NOT HAVE ACCESS TO THE SOURCE CODE… IT IS IN C…
+CHECKED EXCEPTIONS CAN BE THROWN BY NATIVE METHODS…. ANY CHANING IS PERFORMED?
+
+In one of the cases, however, in project OpenDocument.droid a java.lang.Exception was 
+wrapping a java.lang.OutOfMemoryError. General handlers are considered as bad practice, 
+but wrapphing any exception on a general Exception make the matters worse.
+
+\subsubsection{Error wrapping Checked Exception}
+
+%SELECT (exception, root_cause) AS PAIR, COUNT (*) AS N FROM stacktracesummary where 
+%EXCEPTION_EXTENDS='ERROR' AND ROOT_EXTENDS='CHECKED' GROUP BY PAIR ORDER BY N DESC;
+
+libcore	\- 3
+ JAVA ENV \- 	4
+
+(java.lang.NoClassDefFoundError,java.lang.ClassNotFoundException);4
+(java.lang.AssertionError,javax.crypto.ShortBufferException);3
+
+All ClassDefFound wrappings were caused by java.lang.Class methods such as 
+at java.lang.Class.getDeclaredFields(Native Method)
+at java.lang.Class.getDeclaredConstructors(Native Method). They are all NATIVE METHODS.
+
+\subsubsection{Error wrapping Runtime}
+
+APP	11
+ANDROID	1
+LIB	1
+
 %Table X illustrated examples of such wrappings.
@@
 Since there is no way of enforcing Java exception type conventions during program development,
-we could observe the stack trace analysis combined with types these three types
-of exceptions were used interchangeably. This interchangeability makes the
+the exceptions regardless its types can se used interchangeably. This interchangeability makes the
 behaviour of the exception handling code more complex and less reliable. For
-instance, considering a situation were an instance of OutOfMemoryError (a situation that
-should not be handled) was wrapped in a checked exception (a situation that must
-he handled), in this case this wrapping is telling to the caller that it should
+instance, considering a situation were an instance of OutOfMemoryError (a situation that usually cannot
+ be handled) was wrapped in a checked exception which tells the caller that such exception should
 handled.
@@
 problems that may be reducing the robustness of current Java programs in the face of
-exceptional conditions. \noindent\emph{The undocumented runtime exceptions} thrown by library code -
-documenting runtime exceptions is a tedious and error prone task, to help developers
+exceptional conditions.
+
+%\noindent\emph{The undocumented runtime exceptions} 
+\subsection{The undocumented runtime exceptions thrown by thrid party code} 
+Documenting runtime exceptions is a tedious and error prone task, to help developers
 mitigate this problem tools should be developed to automate the documentation of runtime exceptions
 scraping from library code, few solutions in this directions have been proposed so far ~\cite{van2005combining}.
- \noindent\emph{The odd wrappings}  found in this study can lead to what we call
+
+%\noindent\emph{The odd wrappings}  
+\subsection{Exception Wrappings: Language design purpose x real use}
+
+The exception wrappings were inicially proposed by Java [ref-specification] to ...
+
+But the types of wrappings found in this study pointing to the fact that some wrappings may have not 
+been used according to its intial purpose. It can sometimes be used to bypass the  
+language restrictions imposed by checked exceptions.
+
+When (mis)applied, the exception wrapping can lead to what we call
 an \emph{exception handling confusion problem} which can lead the program to
@@
 ways to help developers in deadling with such problem, either preventing odd wrappings or enabling 
-the developer to better dealing with them. Last but not least, the high number of null pointer exceptions 
-lead us to think on other pertinent research direction: whether language designs which avoid null pointers, 
-such as Monads ~\cite{Walde95} (i.e., used in functional languages for values that may not be available 
+the developer to better deal with them.
+
+%\noindent\emph{The high number of null pointer exceptions}
+
+\subsection{The high number of null pointer exceptions}  
+Last but not least, the high number of null pointer exceptions  lead us to think on other 
+pertinent research direction: (i) lightweight intra-method nullpointer analysissuch as the @Nullable;
+(ii) intermethod nullpointer analysis tools such as the one proposed by [DANIELE-RELATED-WORK];
+or (iii) whether language designs which avoid null pointers, such 
+as Monads ~\cite{Walde95} (i.e., used in functional languages for values that may not be available 
 or computations that may fail) could improve the robustness of Java programs. 
-
-TO DISCUSS: How can we benefit from such cross-project crash analysis?
+@NULABLE ANNOTATIONS.
+
+%\noindent\emph{Dependency injection - Inversion of Control -  callback and Exceptions }  
+\subsection{Dependency injection - Inversion of Control -  callback and Exceptions}
+
+%\noindent\emph{How can we benefit from such cross-project crash analysis?}  
+\subsection{The benefits of a cross-project crash analysis}
+
+When family of systems make use of a common core, or framework such systems can
+bennefit from a cross project crash analysis. The bug issues reported for a group of systems
+can identify vulnerabilities which were more difficult to identify in a single project analysis.
+
+The test case explosion problem... A test suite is always incomplete,
+one application could benefit from the usage scenarios on which such framework is applied.
+Framework testing -  what are the main challenges wor framework testing?
+A crossproject crash analysis based on repository mining techniques can 
+identify vulnerabilitues in pieces of code that are shared.
+
+Using exception miner tool we could identify the methods reponsible for most of the crashes - 
+such methods can be a target for manual inspections improvements, or better documentation
+(because sometimes an exception can be thrown because the user is not invoking the method
+in the proper way
+
+During such analysis we could also find the undocumented runtime exceptions 
+signaled by libraries.
+
+It allows the developers to benefit from the faults happening on different systems.
 
@@
 
+\begin{table}
+\centering
+\begin{tabular}{llll}
+    \hline
+    \bfseries{Project} & \bfseries{merging imeStars} & \bfseries{Watchers} & \bfseries{Forks} \\
+    \hline
+subsampling-scale-image-view  &  -  &  -    & - \\
+Android-PullToRefresh &  -  &  -    & - \\
+HackerNews &  -  &  -    & - \\
+PathfinderOpenReference  &  -  &  -    & - \\
+AndroidStaggeredGrid  &  -  &  -    & - \\
+HtmlSpanner  &  -  &  -    & - \\
+screen-notifications  &  -  &  -    & - \\
+appdatapreferences-android  &  -  &  -    & - \\
+Android-Universal-Image-Loader  &  -  &  -    & - \\
+Android-SDK  &  -  &  -    & - \\
+WebCachedImageView  &  -  &  -    & - \\
+Android-SkyDrive-Browser  &  -  &  -    & - \\
+android-pdfview  &  -  &  -    & - \\
+hn-android  &  -  &  -    & - \\
+Android-BitmapCache  &  -  &  -    & - \\
+cropper  &  -  &  -    & - \\
+WordPress-Android  &  -  &  -    & - \\
+changeloglib  &  -  &  -    & - \\
+cgeo  &  -  &  -    & - \\
+chromium-webview-UNDERLINE  &  -  &  -    & - \\
+AndroidSlidingUpPanel  &  -  &  -    & - \\
+DsaTab  &  -  &  -    & - \\
+box-android-sdk-v2  &  -  &  -    & - \\
+chromeview  &  -  &  -    & - \\
+andlytics  &  -  &  -    & - \\
+ContractionTimer  &  -  &  -    & - \\
+android  &  -  &  -    & - \\
+okhttp  &  -  &  -    & - \\
+ActionBarSherlock  &  -  &  -    & - \\
+mobile-android  &  -  &  -    & - \\
+android-timepicker  &  -  &  -    & - \\
+QuiltViewLibrary  &  -  &  -    & - \\
+congress-android  &  -  &  -    & - \\
+c-geo-opensource  &  -  &  -    & - \\
+TweetLanes  &  -  &  -    & - \\
+PageTurner  &  -  &  -    & - \\
+SWADroid  &  -  &  -    & - \\
+picasso  &  -  &  -    & - \\
+uservoice-android-sdk  &  -  &  -    & - \\
+\hline
+  \end{tabular}
+\caption{Projects containing wrappings Error-Runtime}
+\label{tab:wrappings3}
+\end{table}
+
+The methods performing such wrappings were the following:
+
+%source code of ActivityThread -  https://github.com/android/platform_frameworks_base/blob/master/core/java/android/app/ActivityThread.java
+
+\begin{table}
+\begin{tabular}{l}
+    \hline
+    \bfseries{root signaler -  occurences} \\
+      \hline
+android.app.ActivityThread.handleBindApplication – 2 \\
+android.app.ActivityThread.performLaunchActivity – 13 \\
+android.os.AsyncTaskd3.done – 25	 \\
+android.support.v4.content.ModernAsyncTaskd3.done - 2 \\
+android.view.LayoutInflater.createView - 7 \\
+android.view.Viewd1.onClick - 1 \\
+com.squareup.picasso.Requestd1.run - 1 \\
+me.openphoto.android.app.util.concurrent.AsyncTaskExd3.done - 1 \\
+org.asynctasktex.AsyncTaskExd3.done – 3 \\
+org.chromium.androidwebview.AwBrowserProcess.loadLibrary -1 \\
+org.gradle.groovy.scripts.internal.DefaultScriptRunnerFactorydScriptRunnerImpl.run - 1 \\
+\hline
+  \end{tabular}
+\caption{Top 10 most popular exceptions on issues in Java projects}
+\label{tab:toptenjava}
+\end{table}
+
 \section*{Acknowledgment} This work is partially supported by: CNPq -- Proc.
@@
 
-\section {Old Introduction}
-
-....... OLD VERSION ......
-
-Modern applications have to cope with an increasing number of abnormal
-computation states that arise as a consequence of faults in the application
-itself (e.g., access of null references), noisy user inputs or faults in
-underlying middleware or hardware. The exception handling
-mechanism~\cite{goodenough1975exception} is one of the most used schemes for
-detecting and recovering from such exceptional conditions. Although the
-exception handling mechanism has been embedded in several programming languages
-(e.g. Java, C++, C\#) and was the target of several studies
-(e.g.~\cite{miller1997issues,Robil00,shah2010understanding,
-garcia2007extracting,garcia2001comparative,cabral2007exception,coelho2011unveiling}),
-the exception handling code is often generally poorly understood and the least
-tested part of software systems.
-
-Often exception handling constructions may lead the developers into believing that
-by just re-throwing and possibly logging the exceptions they can forget about the exceptional
-situations during the development of the "happy path". This "ignore-for-now"
-approach may turn the exception handling into a generalized "goto"
-mechanism~\cite{mandrioli1992advances} making the program more complex and even
-less reliable. This behaviour may lead to the problem of uncaught exceptions~\cite{jo2004uncaught}, one of the main causes of application crashes.
-
-Moreover, some languages use exceptions to represent programming mistakes
-(array index out-of-bounds, division-by-zero, dereferencing null pointers). In these cases,
-exceptions are implicitly signaled by the runtime environment when such
-conditions occur. One of such implicit exceptions (caused by data conversion
-from a 64-bit floating point to a 16-bit signed integer) was responsible for
-the infamous failure on Ariane 5's first test flight~\cite{lions1996ariane} ---
-leading to the rocket self-destructing 37 seconds after launch and a loss of 500
-million dollars. Besides Ariane 5, several applications crash everyday due to
-uncaught exceptions~\cite{jo2004uncaught}. Not using exceptions carefully
-may have the opposite effect of the initial intention of improving
-system robustness.
-
-In Java, when the program fails due to an uncaught exception, it automatically
-terminates, while the system prints a stack trace to the console. A stack trace
-includes an explanatory message and the execution stack frame when the exception
-occurred. A typical Java stack trace consists of an ordered list of methods that
-were active on the call stack before the exception occurred. Stack traces
-are a useful source of information about system crashes, and are often used to
-support developers in debugging~\cite{schroter2010stack}. Moreover, information
-from stack traces combined with crash and bug reports can be used to support bug
-classification and clustering~\cite{wang2013improving, kim2011crash,
-dhaliwal2011classifying}, fault-proneness prediction
-models~\cite{kim2013predicting} and even automated bug
-fixing~\cite{sinha2009fault} tools.
-
-However, all existing research has focused on the analysis of a single system at a time. In
-our current context, on which we have access to wealth of information available
-on open-source repositories like GitHub additional questions arise: What is the
-prevalence of exception stack traces on reported issues? What are the common
-characteristics of such stack traces?
-
-%What the stack traces available in these repositories can tell us about the use
-%of exceptions in Java programs?
-
-%To guide the our study we also compile general guidelines on how to use
-%exceptions proposed by Gosling~\cite{gosling2000java},
-%Wirfs-Brock~\cite{wirfs2006toward} and Bloch~\cite{bloch2008effective} and then
-%we investigate: what stack traces can tell us about the adherence to such
-%practices in Java programs?
-
-To answer these questions, we conduct an exploratory study of stack traces in
-the wild. Using a custom tool called ExceptionMiner, which was specifically
-developed for this study, we mined stack traces from issues across all Java
-projects on GitHub. Overall, we analyzed 356,057 issues from 16,836 projects,
-from which 28,800 stack traces were extracted and used as a source of
-information for our analysis. The stack trace analysis  was augmented by
-additional information extracted using byte code and source code analysis done
-on a carefully selected subset of our initial sample (482 Android projects).
-%This study was divided in two phases: in the fist phase all the issues of all
-%Java projects available on GitHub were analyzed and the embedded stack traces
-%extracted using ExceptionMiner.
-
-To guide this exploration we compiled general guidelines on how to use
-exceptions proposed by Gosling~\cite{gosling2000java},
-Wirfs-Brock~\cite{wirfs2006toward} and Bloch~\cite{bloch2008effective} and then
-we investigate what stack traces can tell us about the adherence to such
-practices in Java programs.
-%~\todo{This should be our motivation, as this is
-%our biggest contribution. I already changed the text. Lets discuss about this
-%new version}
-%RSC: I AGREE.
-
-%GGG: I think this is too much detail for the introduction In the second phase
-%of this study, we combined the information of the mined stack traces with
-%source code and bytecode analysis. To do so we selected the subset of 482
-%Android repositories. The Android repositories were chosen because this kind of
-%application is known for dealing with several sources of exceptions, as they
-%are usually io-intensive (i.e., deals with Network, Bluetooth, camera, and
-%other resources), and multi-htreaded. Moreover, they are based on a single
-%platform which allowed more fine-grained analysis of their exception behaviour.
-%RSC: I AGREE.
-
-This study builds upon the following assumption: the stack traces reported on
-issues contain relevant information related to system crashes. Several studies
-and techniques have also taken this assumption as starting point
-(e.g.,~\cite{sinha2009fault}~\cite{dhaliwal2011classifying}~\cite{kim2013predicting}). Some outcomes were consistently detected through
-our large scale analysis of exception stack traces in the wild, such as:
-
-\begin{itemize}
-
-  \item  A multitude of programming mistakes which are the main causes of stack traces.
-
-  \item Undocumented runtime exceptions signaled by third party
-    code (i.e., libraries and Android platform).
-
-  \item  Odd exception chains (e.g., Error wrapping checked and runtime
-    exceptions and vice-versa), indicating that the purpose of Java's hybrid
-    exception model may not have been adequately used.
-
-  \item  Stack traces caused by uncaught checked exceptions.
-
-\end{itemize}
-
-Hence, the contributions of this study are as follows:
-\begin{itemize}
-
-  \item  It performs the first large scale analysis of Java stack traces and how
-    they can reveal bad-practices on the use of exceptions in Java.
-
-  \item  It introduces ExceptionMiner, a tool developed to support cross-repository stack trace
-    analysis.
-
-\end{itemize}
-
-These contributions are relevant to:
-(i) developers of Java applications
-to familiarize with the most common reported exceptions mis-uses which is the
-first step to help developers to avoid making them; 
-(ii) designers of languages
-to consider ways of reducing the abundance of NullPointerExceptions;
-and (iii)
-tool designers to consider developing tools that enable a cross-project defect
-analysis.
-
-% AVD: We have not looked at private issues, and the tool is described above.
-%We strongly believe that, specially in the open-source environment,
-%faults are not to be hidden in a private bug issue. Faults should be shared and
-%discussed, so that a developer can learn from other projects mistakes.
-%Currently, the search facilities of repositories are very limited, and the
-%ExceptionMiner tool is a contribution in this direction.
-
-%GGG: Personally, I hate this paragraph in all papers as it usually reads like a space-filler
-%RSC: If needed we can remove it.
-This paper is organized as follows. Section 2 presents a
-background on exception handling mechanism. Section 3 presents the study design.
-Section 4 reports study findings. Section 5 provides further discussions and insights.
-Section 6 presents the threats to validity associated to this study. Finally Section
-7 describes the related work, and Section 8 presents our conclusions and
-directions for future work.
-
-\subsection{Tentative  Introductions to chance the focus}
-
-Along the recent years we have witnessed a rapid increase in 
-the number of open source software  available on the Web.
-Since such projects make their control version and 
-issue tracking systems publicly available, they represent a valuable 
-source of information for the "state of practice" on software development
-which had already started to be investigated [][][][][][I still need to collect these references].
-
-In this scenario Github emerged as one of the most used repository hosting sites.
-It  alows developers to create and maintain opensource projects relying on Git distributed version 
-control and provides a set of tools which allow the information about developers 
-and their activities to be visible within and across open souce projects.
-
-Bug issues have called the attention of the research community and have been the target of increasing 
-researches [][][]. A specific type of bug issue is the one related to system 
-crashes.  Such issues usually contain an exception stack trace [ref-paper], which tipically consists 
-of an ordered list of methods that were active on the call stack before the failure occurred [ref-bug-issues].
-
-Exception stack traces are a useful source of 
-information about system crashes, and are often used to
-support developers in debugging~\cite{schroter2010stack}. Moreover, information
-from stack traces combined with crash and bug reports can be used to support bug
-classification and clustering~\cite{wang2013improving, kim2011crash,
-dhaliwal2011classifying}, fault-proneness prediction
-models~\cite{kim2013predicting} and even automated bug
-fixing~\cite{sinha2009fault} tools.
-
-Although modern applications have to deal with increasinng and similar 
-sources of failures - that arise as a consequence of faults in the application
-itself (e.g., access of null references), noisy user inputs or faults in
-underlying middleware or hardware -  on bug classificationthe current research on 
-exploring bug issue information are structured on  a system at a time basis.
-
-Specially  in a context where we have access to wealth of information available
-on open-source repositories such as GitHub, interesting knowledge may emerge 
-from a cross-project crash analysis. Hence such initial questions arise: What are the common
-characteristics of the stack traces reported on GitHub? Which patterns of exception (mis)use may emerge?
-
-To answer these questions, we conduct an exploratory study of stack traces in
-the wild. Using a custom tool called ExceptionMiner, which was specifically
-developed for this study, we mined stack traces from issues across all Java
-projects on GitHub. Overall, we analyzed 356,057 issues from 16,836 projects,
-from which 28,800 stack traces were extracted and used as a source of
-information for our analysis. The stack trace analysis  was augmented by
-additional information extracted using byte code and source code analysis done
-on a carefully selected subset of our initial sample (482 Android projects).
-%This study was divided in two phases: in the fist phase all the issues of all
-%Java projects available on GitHub were analyzed and the embedded stack traces
-%extracted using ExceptionMiner.
-
-To guide this exploration we compiled general guidelines on how to use
-exceptions proposed by Gosling~\cite{gosling2000java},
-Wirfs-Brock~\cite{wirfs2006toward} and Bloch~\cite{bloch2008effective} and then
-we investigate what stack traces can tell us about the adherence to such
-practices in Java programs.
-%~\todo{This should be our motivation, as this is
-%our biggest contribution. I already changed the text. Lets discuss about this
-%new version}
-%RSC: I AGREE.
-
-%GGG: I think this is too much detail for the introduction In the second phase
-%of this study, we combined the information of the mined stack traces with
-%source code and bytecode analysis. To do so we selected the subset of 482
-%Android repositories. The Android repositories were chosen because this kind of
-%application is known for dealing with several sources of exceptions, as they
-%are usually io-intensive (i.e., deals with Network, Bluetooth, camera, and
-%other resources), and multi-htreaded. Moreover, they are based on a single
-%platform which allowed more fine-grained analysis of their exception behaviour.
-%RSC: I AGREE.
-
-This study builds upon the following assumption: the stack traces reported on
-issues contain relevant information related to system crashes. Several studies
-and techniques have also taken this assumption as starting point
-(e.g.,~\cite{sinha2009fault}~\cite{dhaliwal2011classifying}~\cite{kim2013predicting}). Some outcomes were consistently detected through
-our large scale analysis of exception stack traces in the wild, such as:
-
-\begin{itemize}
-
-  \item  A multitude of programming mistakes were reported as the main causes of stack traces:
-  41,1\%  (12,625)  of reported stack traces were caused by programming mistakes; and 26,31\% of analyzed stack traces reported a java.lang.NullPointerException as root cause;
-
-   \item 1,671 projects (44,47\%) contained at least one issue reporting a java.lang.NullPointerException as the root cause of a stack trace.
-
-  \item Undocumented runtime exceptions signaled by third party
-    code (i.e., libraries and Android platform).
-This study found 118 distinct methods defined on third party code and  reported as the root signalers of runtime exceptions.
-Only 1 of such methods (0,8\%) documented the explicitly thrown runtime exception in a Javadoc comment; and none of these methods included the exception as part of the exception interface (i.e., using throws clause).
-
-  \item  Odd exception chains (e.g., Error wrapping checked and runtime
-    exceptions and vice-versa), indicating that the purpose of Java's hybrid
-    exception model may not have been adequately used.
-
-  \item  Issues reporting uncaught checked exceptions.
-
-\end{itemize}
-
-%Hence, the contributions of this study are as follows:
-%\begin{itemize}
-%  \item  It performs a large scale analysis of Java stack traces and how
- %   they can reveal bad-practices on the use of exceptions in Java.
-% \item  It introduces ExceptionMiner, a tool developed to support cross-repository stack trace
-%    analysis.
-%\end{itemize}
-
-These contributions are relevant to:
-(i) developers of Java applications
-to familiarize with the most common reported exceptions mis-uses which is the
-first step to help developers to avoid making them; 
-(ii) designers of languages
-to consider ways of reducing the abundance of NullPointerExceptions;
-and (iii)
-tool designers to consider developing tools that enable a cross-project defect
-analysis.
-
-% AVD: We have not looked at private issues, and the tool is described above.
-%We strongly believe that, specially in the open-source environment,
-%faults are not to be hidden in a private bug issue. Faults should be shared and
-%discussed, so that a developer can learn from other projects mistakes.
-%Currently, the search facilities of repositories are very limited, and the
-%ExceptionMiner tool is a contribution in this direction.
-
-%GGG: Personally, I hate this paragraph in all papers as it usually reads like a space-filler
-%RSC: If needed we can remove it.
-This paper is organized as follows. Section 2 presents a
-background on exception handling mechanism. Section 3 presents the study design.
-Section 4 reports study findings. Section 5 provides further discussions and insights.
-Section 6 presents the threats to validity associated to this study. Finally Section
-7 describes the related work, and Section 8 presents our conclusions and
-directions for future work.
-
 \bibliographystyle{IEEEtran}

### 2014-08-19 java-stacks.tex
This version of the paper discuss the results starting from Java.

--- 2014-08-13 android-stacks.tex
@@
 
+\subsubsection{Distiling Exception Stack Trace Information}
+
+\- root exception
+
+\- signaler
+\- method
+\- class
+\- package
+
+\- Exception
+
 \subsubsection{Android Apps Selection}
 \label{sec:android}
+
+ % SELECT id, exception, exception_extends, exception_from, exception_package, 
+ %      issueid_onexminer, main_cause, main_extends, main_from, main_package, 
+ %      main_pre_signaler_java, main_signaler, mainsignaler_class, mainsignaler_from, 
+ %      mainsignaler_method, mainsignaler_package, pre_signaler_java, 
+ %      repo, root_cause, root_extends, root_from, root_package, root_pre_signaler_java, 
+ %      root_signaler, rootsignaler_class, rootsignaler_from, rootsignaler_method, 
+ %      rootsignaler_package, signaler, signaler_class, signaler_from, 
+ %      signaler_method, signaler_package, stack
+ % FROM stacktracesummary;
 
@@
 
-\begin{figure}
-\centering
+%\begin{figure}
+%\centering
 %\includegraphics[width=\hsize]{chart_exceptiontypes.pdf}
-\includegraphics[width=\hsize]{exception_types.pdf}
-\caption{Distribution of Checked and Unchedked exceptions.}
-\label{fig:typeroot}
-\end{figure}
+%\includegraphics[width=\hsize]{exception_types.pdf}
+%\caption{Distribution of Checked and Unchedked exceptions.}
+%\label{fig:typeroot}
+%\end{figure}
 
@@
     \hline
-      none   & Runtime   &  2360 & 359 \\
-      none  &   Checked  & 422  & 117 \\
-      none  &   Error   & 381  & 141 \\
-      none &  Undefined  &  15   & 10  \\
-      none & Throwable  &  1    & 1 \\
-     \hline
-      Runtime   & Runtime & 560  & 182 \\    
-      Checked   & Checked & 98   & 42  \\
-      Error     & Error   & 15   & 14  \\
-    \hline
+      %none   & Runtime   &  2360 & 359 \\
+      %none  &   Checked  & 422  & 117 \\
+      %none  &   Error   & 381  & 141 \\
+      %none &  Undefined  &  15   & 10  \\
+      %none & Throwable  &  1    & 1 \\
+     %\hline
+      %Runtime   & Runtime & 560  & 182 \\    
+      %Checked   & Checked & 98   & 42  \\
+      %Error     & Error   & 15   & 14  \\
+    %\hline
       Runtime &  Checked   & 114  & 70  \\
@@
       Throwable & Runtime &  3    & 1 \\
-      Undefined & Runtime &  3    & 3 \\
-      Undefined & Error &  1    & 1 \\
+      %Undefined & Runtime &  3    & 3 \\
+      %Undefined & Error &  1    & 1 \\
     \hline
@@
 
-Such wrapping was detected on distinct 70 projects. Inspecting the elements responsible for performing such wrapping we could observe that,
+Such wrapping was detected on distinct 70 projects. Inspecting the 
+elements responsible for performing such wrapping we could observe that,
 most of such wrappings were perform on methods defined on Android.
 
-ANDROID	 \- 66
-APP	\- 29 	(HOW MANY?? AND WHICH??)
-LIB  \- 19 (HOW MANY AND WHICH?)
+\begin{table}
+\centering
+\begin{tabular}{ll}
+    \hline
+    \bfseries{signaler from} & \bfseries{Ocurrences} \\
+    \hline
+ANDROID	 & 66 \\
+APP	& 29 	(HOW MANY?? AND WHICH??) \\
+LIB  & 19 (HOW MANY AND WHICH?)\\
+\hline
+  \end{tabular}
+\caption{Elements responsible for the wrapping.}
+\label{tab:wrapping01}
+\end{table}
 
@@
 
-android.app.ActivityThread \- 47
-android.view.LayoutInflater \- 7
-org.gege.caldavsyncadapter.syncadapter.SyncAdapter \- 7
-org.springframework.web.client.RestTemplate \- 6
-android.os.AsyncTask \- 5
-
-Code snipet:
-
-android.app.ActivityThread.
-
-public final Activity startActivityNow(...) \{
+\begin{table}
+\centering
+\begin{tabular}{ll}
+ \bfseries{Signaler class} & \bfseries{Ocurrences} \\
+    \hline
+android.app.ActivityThread & 47 \\
+android.view.LayoutInflater & 7 \\
+org.gege.caldavsyncadapter.syncadapter.SyncAdapter & 7 \\
+org.springframework.web.client.RestTemplate & 6 \\
+android.os.AsyncTask & 5 \\
+\hline
+  \end{tabular}
+\caption{Elements tresponsible for the wrapping.}
+\label{tab:wrapping01}
+\end{table}
+
+Code snipet extracted from android.app.ActivityThread.
+
+{\footnotesize
+\begin{verbatim}
+
+private Activity performLaunchActivity(...) {
+  try {
    ...
-return performLaunchActivity(r, null);
-\}
-
-private Activity performLaunchActivity(...) \{
-try \{
-.....
-\} catch (Exception e) {
-..
-throw new RuntimeException(
-"Unable to instantiate activity " + component
-+ ": " + e.toString(), e);
-..
-\}
-\}
+  } catch (Exception e) {
+    ...
+    throw new RuntimeException(..., e);
+  }
+}
+\end{verbatim}
+}
 
@@
 
-(java.lang.RuntimeException,java.lang.ClassNotFoundException)	32
-(java.lang.RuntimeException,java.lang.NoSuchMethodException)	6
-(java.lang.RuntimeException,java.lang.InstantiationException)	5
-(com.github.andlyticsproject.console.NetworkException,org.apache.http.client.HttpResponseException)	5
-(android.view.InflateException,java.lang.NoSuchMethodException)	4
-
-java.lang.RuntimeException	69 \- 61%
-android.view.InflateException	10 \- 8,7%
-com.github.andlyticsproject.console.NetworkException	7
-java.lang.IllegalStateException	6
-org.springframework.web.client.ResourceAccessException	6
-com.github.andlyticsproject.console.DevConsoleProtocolException	3
+\begin{table}
+\centering
+\begin{tabular}{lll}
+    \hline
+    \bfseries{wrapper} & \bfseries{root} & \bfseries{\#} \\
+    \hline
+java.lang.RuntimeException & java.lang.ClassNotFoundException &	32 \\
+java.lang.RuntimeException & java.lang.NoSuchMethodException &	6 \\
+java.lang.RuntimeException & java.lang.InstantiationException) & 5 \\
+...andlyticsproject.NetworkException & org.apache...HttpResponseException &	5\\
+android.view.InflateException & java.lang.NoSuchMethodException & 4 \\
+\hline
+  \end{tabular}
+\caption{Elements tresponsible for the wrapping.}
+\label{tab:wrapping01}
+\end{table}
+
+\begin{table}
+\centering
+\begin{tabular}{lll}
+    \hline
+    \bfseries{Exception} & \bfseries{Ocurrences} \\
+    \hline
+java.lang.RuntimeException &	69 & 61\% \\
+android.view.InflateException	& 10 & 8,7\% \\
+com.github.andlyticsproject.console.NetworkException	& 7 & - \\
+java.lang.IllegalStateException	& 6 & - \\
+
+\hline
+  \end{tabular}
+\caption{Elements tresponsible for the wrapping.}
+\label{tab:wrapping01}
+\end{table}
 
@@
 
-ANDROID	50
-LIB 	6
-APP	1
-
-android.os.AsyncTask \-	25
-android.app.ActivityThread \- 15
-android.view.LayoutInflater \- 7
-org.asynctasktex.AsyncTaskExd \- 3
-android.support.v4.content.ModernAsyncTask \- 2
+\begin{table}
+\centering
+\begin{tabular}{ll}
+ \bfseries{Signaler from} & \bfseries{Ocurrences} \\
+    \hline
+ANDROID	& 50 \\
+LIB  & 6 \\
+APP	& 1 \\
+\hline
+  \end{tabular}
+\caption{Elements tresponsible for the wrapping.}
+\label{tab:wrapping01}
+\end{table}
+
+\begin{table}
+\centering
+\begin{tabular}{ll}
+ \bfseries{Signaler method} & \bfseries{Ocurrences} \\
+    \hline
+android.os.AsyncTask &	25 \\
+android.app.ActivityThread & 15 \\
+android.view.LayoutInflater & 7 \\
+org.asynctasktex.AsyncTaskExd & 3 \\
+android.support.v4.content.ModernAsyncTask & 2\\
+\hline
+  \end{tabular}
+\caption{Elements tresponsible for the wrapping.}
+\label{tab:wrapping01}
+\end{table}
 
@@
 
-@Override
-protected void done() \{
+{\footnotesize
+\begin{verbatim}
+protected void done() {
  ...
-try \{
+   try {
       ...
-
-\} catch (InterruptedException e) \{
-android.util.Log.w(.., e);
-\} catch (ExecutionException e) \{
-throw new RuntimeException("An error occured while executing doInBackground()",
-e.getCause());
-\} catch (CancellationException e)  \{
+   } catch (InterruptedException e) {
+      android.util.Log.w(..., e);
+   } catch (ExecutionException e) {
+      throw new RuntimeException("...",e.getCause());
+   } catch (CancellationException e) {
       ...
-\} catch (Throwable t) \{
-throw new RuntimeException("An error occured while executing "
-+ "doInBackground()", t);
-\}
+   } catch (Throwable t) {
+     throw new RuntimeException("...", t);
+   }
     ...
-\}
-\}
+ }
+}
+\end{verbatim}
+}
 
@@
 
-(java.lang.RuntimeException,java.lang.OutOfMemoryError) \- 30
-(java.lang.RuntimeException,java.lang.NoClassDefFoundError) \- 6
-(java.lang.RuntimeException,java.lang.UnsatisfiedLinkError) \- 5
-(android.view.InflateException,java.lang.OutOfMemoryError) \- 4
-(java.lang.RuntimeException,java.lang.StackOverflowError) \- 4
+\begin{table}
+\centering
+\begin{tabular}{ll}
+ \bfseries{Signaler method} & \bfseries{Ocurrences} \\
+    \hline
+java.lang.RuntimeException,java.lang.OutOfMemoryError & 30\\
+java.lang.RuntimeException,java.lang.NoClassDefFoundError & 6\\
+java.lang.RuntimeException,java.lang.UnsatisfiedLinkError) & 5\\
+android.view.InflateException,java.lang.OutOfMemoryError) & 4\\
+java.lang.RuntimeException,java.lang.StackOverflowError)  & 4\\
+\hline
+  \end{tabular}
+\caption{Elements tresponsible for the wrapping.}
+\label{tab:wrapping01}
+\end{table}
 
@@
 
-APP	9
-LIB	9
-ANDROID	4
-
-(com.octo.android.robospice.persistence.exception.CacheSavingException,java.lang.ClassCastException)	6
-(java.lang.reflect.InvocationTargetException,java.lang.NullPointerException)	3
-(java.lang.reflect.InvocationTargetException,android.content.res.ResourcesdNotFoundException)	2
-(java.sql.SQLException,android.database.sqlite.SQLiteDiskIOException)	2
-(java.lang.Exception,at.andiwand.odf2html.odf.IllegalMimeTypeException)	2
+\begin{table}
+\centering
+\begin{tabular}{ll}
+ \bfseries{Origin} & \bfseries{Ocurrences} \\
+    \hline
+APP	& 9 \\
+LIB	& 9 \\
+ANDROID &	4 \\
+\hline
+  \end{tabular}
+\caption{Elements tresponsible for the wrapping.}
+\label{tab:wrapping01}
+\end{table}
+\begin{table}
+\centering
+\begin{tabular}{lll}
+ \bfseries{Wrapper} & \bfseries{Root} & \bfseries{Ocurrences}\\
+ \hline
+...android.robospice.CacheSavingException,java.lang.ClassCastException & 6 \\
+java.lang.reflect.InvocationTargetException,java.lang.NullPointerException &  3 \\
+java.lang.reflect.InvocationTargetException,android...ResourcesdNotFoundException & 2 \\
+java.sql.SQLException,android.database.sqlite.SQLiteDiskIOException & 2 \\
+java.lang.Exception,at.andiwand.odf2html.odf.IllegalMimeTypeException & 	2 \\
+  \end{tabular}
+\caption{Elements tresponsible for the wrapping.}
+\label{tab:wrapping01}
+\end{table}
 
@@
 
-ANDROID	5
-APP	3
-
-(java.lang.reflect.InvocationTargetException,java.lang.OutOfMemoryError)	5
-(java.lang.Exception,java.lang.OutOfMemoryError)	1
-(java.lang.reflect.InvocationTargetException,java.lang.NoSuchMethodError)	1
-(java.lang.reflect.InvocationTargetException,java.lang.UnsatisfiedLinkError)	1
+\begin{table}
+\centering
+\begin{tabular}{ll}
+ \bfseries{Origin} & \bfseries{Ocurrences} \\
+    \hline
+ANDROID	& 5 \\
+APP	& 3 \\
+  \end{tabular}
+\caption{Elements tresponsible for the wrapping.}
+\label{tab:wrapping01}
+\end{table}
+
+\begin{table}
+\centering
+\begin{tabular}{lll}
+ \bfseries{Wrapper} & \bfseries{Root} & \bfseries{Ocurrences}\\
+ \hline
+java.lang.reflect.InvocationTargetException,java.lang.OutOfMemoryError & 5 \\
+java.lang.Exception,java.lang.OutOfMemoryError & 1 \\
+java.lang.reflect.InvocationTargetException,java.lang.NoSuchMethodError &	1 \\
+java.lang.reflect.InvocationTargetException,java.lang.UnsatisfiedLinkError&	1 \\
+  \end{tabular}
+\caption{Elements tresponsible for the wrapping.}
+\label{tab:wrapping01}
+\end{table}
 
 All wrappings except 1 were caused by the reflection library.
-
-at java.lang.reflect.Constructor.constructNative(Native Method)
-NATIVE IMPLEMENTATION, WE DO NOT HAVE ACCESS TO THE SOURCE CODE… IT IS IN C…
-CHECKED EXCEPTIONS CAN BE THROWN BY NATIVE METHODS…. ANY CHANING IS PERFORMED?
-
-In one of the cases, however, in project OpenDocument.droid a java.lang.Exception was 
+The method responsible for the wrappings (e.g., at java.lang.reflect.Constructor.constructNative(Native Method))
+were native methods written in C. In one of the cases, however, in project OpenDocument.droid a java.lang.Exception was 
 wrapping a java.lang.OutOfMemoryError. General handlers are considered as bad practice, 
-but wrapphing any exception on a general Exception make the matters worse.
+but wrapping an Error on a general Exception make the matters worse.
 
@@
 
-libcore	\- 3
- JAVA ENV \- 	4
-
-(java.lang.NoClassDefFoundError,java.lang.ClassNotFoundException);4
-(java.lang.AssertionError,javax.crypto.ShortBufferException);3
-
-All ClassDefFound wrappings were caused by java.lang.Class methods such as 
+\begin{table}
+\centering
+\begin{tabular}{ll}
+libcore	&  3 \\
+ JAVA ENV  & 	4 \\
+  \end{tabular}
+\caption{Elements tresponsible for the wrapping.}
+\label{tab:wrapping01}
+\end{table}
+
+\begin{table}
+\centering
+\begin{tabular}{lll}
+ \bfseries{Wrapper} & \bfseries{Root} & \bfseries{Ocurrences}\\
+ \hline
+java.lang.NoClassDefFoundError & java.lang.ClassNotFoundException & 4 \\
+java.lang.AssertionError & javax.crypto.ShortBufferException) & 3 \\
+ \end{tabular}
+\caption{Elements tresponsible for the wrapping.}
+\label{tab:wrapping01}
+\end{table}
+
+Some wrappings were performed by methods defined by java.lang.Class methods such as 
 at java.lang.Class.getDeclaredFields(Native Method)
-at java.lang.Class.getDeclaredConstructors(Native Method). They are all NATIVE METHODS.
+at java.lang.Class.getDeclaredConstructors(Native Method).  While were performed by javax.crypto package.
 
@@
 
-APP	11
-ANDROID	1
-LIB	1
+\begin{table}
+\centering
+\begin{tabular}{ll}
+APP & 11 \\
+ANDROID	& 1 \\
+LIB	& 1 \\
+  \end{tabular}
+\caption{Elements tresponsible for the wrapping.}
+\label{tab:wrapping01}
+\end{table}
+
+\begin{table}
+\centering
+\begin{tabular}{lll}
+ \bfseries{Wrapper} & \bfseries{Root} & \bfseries{Ocurrences}\\
+ \hline
+java.lang.ExceptionInInitializerError,java.lang.NullPointerException & 9 \\
+java.lang.ExceptionInInitializerError,java.lang.IllegalArgumentException &	2 \\
+java.lang.ExceptionInInitializerError,java.util.MissingResourceException) &	1 \\
+java.lang.ExceptionInInitializerError,org.andengine.util.exception.AndEngineRuntimeException &	1 \\
+ \end{tabular}
+\caption{Elements tresponsible for the wrapping.}
+\label{tab:wrapping01}
+\end{table}
+
+after manually inspecting the code associated to such wrappings we discovered that all of them were 
+caused by static initializers. If any exception is thrown in the context of a static initializer (i.e., static block) 
+it is converted into an ExceptionInitializerError on the point where the class is first used.
 
@@
 
+\subsubsection{Multiple Wrappings}
+
 We could also observe that in 205 stacks found in Android repository the exception was wrapped 
@@
 
-Since there is no way of enforcing Java exception type conventions during program development,
-the exceptions regardless its types can se used interchangeably. This interchangeability makes the
-behaviour of the exception handling code more complex and less reliable. For
-instance, considering a situation were an instance of OutOfMemoryError (a situation that usually cannot
- be handled) was wrapped in a checked exception which tells the caller that such exception should
-handled.
+\subsection{Checked and Unchecked Exceptions in JVM and Android and what they can tell us?}
+
+We could observe that some of the type-changing-wrappings were performed by JDK classes.
+Such finding lead us to think that the exception types have been used as intercheagable and may not being carrying 
+the meaning they were firstly designed to carry.
+
+The way the exceptions are being handle are more important that the type itself, it points to the fact already discussed 
+by other languages such as M\# blog [ref] and \"thinking in Java author\" and other blogs. It is all about the exception handling
+politics of a system and not the type itself. The exception type is not enought to ensure a more strict or robust development,
+since the checked exceptions may be silienced or converted to a checked exception. According to Java manual,
+they say that it can actually happen but it will be done intentionally and they advocate that it is a bad practice.
+
+Present the number of checke and unchecked exceptions in JVM, JDK, Android SO, and Android SDK.
+Discuss about the amount of checked and unchecked defined in each, and discuss about interesting exceptions that
+have the same name ex: classNotFound...
+
+\subsection{Top Signalers and what they can tell us?}
+
+Present the top signalers found in this study, and top pairs (signalers and exceptions), and discuss 
+about inversion of control and the difficult to adequately handle exceptions in this style of development.
+Inversion of control stimulate the use of unchecked exceptions since the "call back" do not know which exception to expect.
+
+{\tiny
+\begin{verbatim}
+
+%SELECT (mainsignaler_method, main_cause), count(distinct(repo)) as n
+%  FROM stacktracesummary group by (mainsignaler_method, main_cause) order by n desc;
+
+GITHUB:
+"(dalvik.system.BaseDexClassLoader.findClass,java.lang.ClassNotFoundException)";19
+"(android.app.Instrumentation.checkStartActivityResult,android.content.ActivityNotFoundException)";17
+"(android.graphics.BitmapFactory.nativeDecodeStream,java.lang.OutOfMemoryError)";16
+"(android.os.Parcel.readException,java.lang.SecurityException)";16
+"(android.view.WindowManagerImpl.findViewLocked,java.lang.IllegalArgumentException)";14
+"(android.graphics.BitmapFactory.nativeDecodeAsset,java.lang.OutOfMemoryError)";14
+"(dalvik.system.PathClassLoader.findClass,java.lang.ClassNotFoundException)";13
+"(android.support.v4.app.FragmentManagerImpl.checkStateLoss,java.lang.IllegalStateException)";12
+"(android.graphics.Bitmap.nativeCreate,java.lang.OutOfMemoryError)";11
+
+GOOGLECODE:
+"(android.app.Instrumentation.checkStartActivityResult,android.content.ActivityNotFoundException)";8
+"(dalvik.system.BaseDexClassLoader.findClass,java.lang.ClassNotFoundException)";6
+"(java.lang.ClassLoader.loadClass,java.lang.ClassNotFoundException)";6
+"(com.android.dx.dex.file.ClassDefsSection.add,java.lang.IllegalArgumentException)";6
+"(android.graphics.Bitmap.nativeCreate,java.lang.OutOfMemoryError)";6
+"(android.view.WindowManagerImpl.findViewLocked,java.lang.IllegalArgumentException)";5
+"(android.graphics.BitmapFactory.nativeDecodeStream,java.lang.OutOfMemoryError)";5
+"(junit.framework.Assert.fail,junit.framework.AssertionFailedError)";4
+"(android.graphics.BitmapFactory.nativeDecodeAsset,java.lang.OutOfMemoryError)";4
+"(dalvik.system.PathClassLoader.findClass,java.lang.ClassNotFoundException)";4
+
+COMBINED:
+"(dalvik.system.BaseDexClassLoader.findClass,java.lang.ClassNotFoundException)"; 25
+"(android.app.Instrumentation.checkStartActivityResult,android.content.ActivityNotFoundException)";25
+"(android.graphics.BitmapFactory.nativeDecodeStream,java.lang.OutOfMemoryError)";21
+"(android.view.WindowManagerImpl.findViewLocked,java.lang.IllegalArgumentException)";19
+"(dalvik.system.PathClassLoader.findClass,java.lang.ClassNotFoundException)";17
+"(android.os.Parcel.readException,java.lang.SecurityException)";16
+"(android.graphics.BitmapFactory.nativeDecodeAsset,java.lang.OutOfMemoryError)";14
+"(android.support.v4.app.FragmentManagerImpl.checkStateLoss,java.lang.IllegalStateException)";12
+"(android.graphics.Bitmap.nativeCreate,java.lang.OutOfMemoryError)";11
+
+%SELECT (mainsignaler_class, main_cause), count(distinct(repo)) as n
+ % FROM stacktracesummary group by (mainsignaler_class, main_cause) order by n desc;
+
+GITHUB:
+"(android.graphics.BitmapFactory,java.lang.OutOfMemoryError)";27
+"(dalvik.system.BaseDexClassLoader,java.lang.ClassNotFoundException)";19
+"(android.content.res.Resources,android.content.res.ResourcesdNotFoundException)";18
+"(android.app.Instrumentation,android.content.ActivityNotFoundException)";17
+"(libcore.io.Posix,libcore.io.ErrnoException)";17
+"(android.os.Parcel,java.lang.SecurityException)";16
+"(android.app.ActivityThread,java.lang.RuntimeException)";15
+"(android.support.v4.app.FragmentManagerImpl,java.lang.IllegalStateException)";14
+"(android.view.WindowManagerImpl,java.lang.IllegalArgumentException)";14
+"(android.graphics.Bitmap,java.lang.OutOfMemoryError)";13
+"(dalvik.system.PathClassLoader,java.lang.ClassNotFoundException)";13
+"(android.os.Parcel,java.lang.NullPointerException)";12
+"(android.support.v4.app.Fragment,java.lang.IllegalStateException)";11
+
+GOOGLECODE:
+"(android.app.Instrumentation,android.content.ActivityNotFoundException)";8
+"(android.graphics.BitmapFactory,java.lang.OutOfMemoryError)";7
+"(android.app.ActivityThread,java.lang.RuntimeException)";7
+"(java.lang.ClassLoader,java.lang.ClassNotFoundException)";6
+"(dalvik.system.BaseDexClassLoader,java.lang.ClassNotFoundException)";6
+"(android.graphics.Bitmap,java.lang.OutOfMemoryError)";6
+"(com.android.dx.dex.file.ClassDefsSection,java.lang.IllegalArgumentException)";6
+"(android.view.WindowManagerImpl,java.lang.IllegalArgumentException)";5
+"(android.database.sqlite.SQLiteDatabase,android.database.sqlite.SQLiteException)";4
+"(android.os.Parcel,java.lang.IllegalArgumentException)";4
+
+COMBINED:
+"(android.graphics.BitmapFactory,java.lang.OutOfMemoryError)"; 34
+"(dalvik.system.BaseDexClassLoader,java.lang.ClassNotFoundException)"; 27
+"(android.content.res.Resources,android.content.res.ResourcesdNotFoundException)";18
+"(android.app.Instrumentation,android.content.ActivityNotFoundException)";25
+"(libcore.io.Posix,libcore.io.ErrnoException)";17
+"(android.os.Parcel,java.lang.SecurityException)";20
+"(android.app.ActivityThread,java.lang.RuntimeException)"; 22
+"(android.support.v4.app.FragmentManagerImpl,java.lang.IllegalStateException)";14
+"(android.view.WindowManagerImpl,java.lang.IllegalArgumentException)";19
+"(android.graphics.Bitmap,java.lang.OutOfMemoryError)";19
+"(dalvik.system.PathClassLoader,java.lang.ClassNotFoundException)";13
+"(android.os.Parcel,java.lang.NullPointerException)";12
+"(android.support.v4.app.Fragment,java.lang.IllegalStateException)";11
+
+PACKAGE:
+
+GITHUB:
+"(android.graphics,java.lang.OutOfMemoryError)";35
+"(dalvik.system,java.lang.ClassNotFoundException)";32
+"(android.os,java.lang.RuntimeException)";27
+"(android.support.v4.app,java.lang.IllegalStateException)";27
+"(android.view,java.lang.IllegalArgumentException)";26
+"(android.database.sqlite,android.database.sqlite.SQLiteException)";24
+"(libcore.io,libcore.io.ErrnoException)";20
+"(android.widget,java.lang.NullPointerException)";19
+"(android.graphics,java.lang.RuntimeException)";18
+"(android.content.res,android.content.res.ResourcesdNotFoundException)";18
+
+GOOGLECODE:
+"(android.os,java.lang.RuntimeException)";9
+"(android.app,java.lang.RuntimeException)";9
+"(android.app,java.lang.NullPointerException)";9
+"(android.graphics,java.lang.OutOfMemoryError)";9
+"(android.app,android.content.ActivityNotFoundException)";8
+"(dalvik.system,java.lang.ClassNotFoundException)";8
+"(android.view,java.lang.IllegalArgumentException)";7
+"(java.lang,java.lang.ClassNotFoundException)";7
+"(android.app,java.lang.IllegalArgumentException)";7
+"(android.database.sqlite,android.database.sqlite.SQLiteException)";6
+
+% SELECT mainsignaler_method,count(distinct(repo)) as n, string_agg(distinct(main_cause), ',')
+% FROM stacktracesummary group by (mainsignaler_method) order by n desc;
+
+GITHUB:
+"android.os.Parcel.readException";28;"java.lang.IllegalArgumentException,java.lang.IllegalStateException,java.lang.NullPointerException,java.lang.SecurityException"
+"dalvik.system.BaseDexClassLoader.findClass";19;"java.lang.ClassNotFoundException"
+"android.app.Instrumentation.checkStartActivityResult";17;"android.content.ActivityNotFoundException"
+"android.graphics.BitmapFactory.nativeDecodeStream";16;"java.lang.OutOfMemoryError"
+"android.view.WindowManagerImpl.findViewLocked";15;"java.lang.IllegalArgumentException"
+"android.graphics.Bitmap.nativeCreate";15;"java.lang.IllegalArgumentException,java.lang.OutOfMemoryError"
+"android.graphics.BitmapFactory.nativeDecodeAsset";14;"java.lang.OutOfMemoryError,java.lang.RuntimeException"
+"dalvik.system.PathClassLoader.findClass";13;"java.lang.ClassNotFoundException"
+"android.support.v4.app.FragmentManagerImpl.checkStateLoss";12;"java.lang.IllegalStateException"
+"android.os.BinderProxy.transact";12;"android.os.DeadObjectException,android.os.TransactionTooLargeException,java.lang.RuntimeException"
+
+GOOGLECODE:
+"android.app.Instrumentation.checkStartActivityResult";8;"android.content.ActivityNotFoundException"
+"android.graphics.Bitmap.nativeCreate";8;"java.lang.IllegalArgumentException,java.lang.OutOfMemoryError,java.lang.RuntimeException"
+"java.lang.ClassLoader.loadClass";6;"java.lang.ClassNotFoundException,java.lang.IncompatibleClassChangeError,java.lang.UnsupportedClassVersionError"
+"dalvik.system.BaseDexClassLoader.findClass";6;"ClassNotFoundException,java.lang.ClassNotFoundException"
+"com.android.dx.dex.file.ClassDefsSection.add";6;"java.lang.IllegalArgumentException"
+"android.os.Parcel.readException";6;"Exception,java.lang.IllegalArgumentException,java.lang.IllegalStateException,java.lang.NullPointerException,java.lang.SecurityException"
+"android.view.WindowManagerImpl.findViewLocked";5;"java.lang.IllegalArgumentException"
+"android.os.BinderProxy.transact";5;"android.os.DeadObjectException,android.os.RemoteException,android.os.TransactionTooLargeException,java.lang.RuntimeException,java.lang.SecurityException"
+"android.app.ActivityThread.performLaunchActivity";5;"Error,java.lang.NullPointerException,java.lang.RuntimeException"
+"android.graphics.BitmapFactory.nativeDecodeStream";5;"java.lang.OutOfMemoryError"
+
+COMBINED:
+
+\end{verbatim}
+}
 
@@
 the developer to better deal with them.
+
+\subsubsection{Ways of Preventing Odd Wrappings}
+
+Since there is no way of enforcing Java exception type conventions during program development,
+the exceptions regardless its types can se used interchangeably. This interchangeability makes the
+behaviour of the exception handling code more complex and less reliable. For
+instance, considering a situation were an instance of OutOfMemoryError (a situation that usually cannot
+ be handled) was wrapped in a checked exception which tells the caller that such exception should
+handled.
 

### 2014-08-22 android-stacks.tex
Changing the focus of the paper to android only.

@@
 \begin{document}
-\title{ What can Stack Traces tell us about Java Exceptions? }
-
-%\title{A first Glance of what is causing Mobile Crashes: A GitHub and Googlecode Repository Mining Study }
+%\title{ What GitHub and Googlecode can tell us about the Causes of Crashes of Android Apps? }
+%\title{ A Mining Study the Googlecode can tell us about the Causes of Crashes of Android Apps? }
+
+\title{ What can GitHub and Googlecode Issues tell us about Exceptions in Android Apps? }
+
+%\title{On the Common Characteristics Android-Apps Crashes: Results of a GitHub and Googlecode Mining Study}
+
+%\title{On the Common Characteristics of Android-Apps Crashes: Results of a GitHub and Googlecode Mining Study}
+
+%\title{On the Common Characteristics of Android Crashes: Results of a GitHub and Googlecode Mining Study}
+
+%\title{A first Glance on Characteristics of Android Crashes: Results of a GitHub and Googlecode Mining Study}
 
@@
 
-In this work, we perform the first large scale analysis of Java stack traces and
-pinpoint how they can reveal bad programming practices. We mined and analyzed
-the stack traces embedded in all issues of Java projects available GitHub.
-Overall, our research set includes 356,057 issues defined at 16,836 projects, 
-from which 28,800 stack traces were extracted. From this set, the stack traces
-of 482 Android projects were investigated in more detail, in combination with
-source code and bytecode analysis. In this study some patterns of exception
-(mis)-use were consistently detected such as: unexpected wrappings (e.g., Errors
-being wrapped in checked exceptions), revealing that Java hybrid exception
-model is not fully used according to its purpose;  undocumented runtime
-exceptions signaled by third party code, which is a serious threat to program
-dependability; and a high prevalence of java.lang exceptions reported on issues,  
-they were found in aproximately 50\% of all analyzed exception stack traces.
+This paper reports a repository mining study which analyzed all the issues created on
+ Android projects available in GitHub (482 applications) and Google code (141 appplications)
+Overall, our research set includes XXXX issues from which XXX stack traces 
+were extracted. Such stack traces were distilled and analyzed in 
+combination with source code and bytecode analysis. In this study some patterns of exception
+(mis)-use were consistently detected such as: a high prevalence of NullPointerExceptions and 
+other java.lang exceptions reported on issues - they were found in aproximately 50\% of all analyzed 
+exception stack traces; undocumented runtime exceptions signaled by third party code - which is a serious threat to program
+dependability; unexpected wrappings (e.g., Errors being wrapped in checked exceptions), revealing that Java hybrid exception
+model is not fully used according to its purpose.
 
@@
 
-Modern applications have to cope with an increasing number of abnormal
-computation states that arise as a consequence of faults in the application
-itself (e.g., access of null references), noisy user inputs or faults in
-underlying middleware or hardware. The exception handling
-mechanism~\cite{goodenough1975exception} is one of the most used schemes for
-detecting and recovering from such exceptional conditions. When an exception is
-thrown it can be caught inside the application (by the program's exception
-handling code) and the application has the opportunity to assess the conditions
-that caused the fault and either recover from them or abort. If the exception
-that remains uncaught, the program crashes. Uncaught exceptions are one of the
-main causes of application crashes~\cite{jo2004uncaught}.
-
-In most programming languages featuring exception handling, when a crash
-happens, the language runtime generates debugging information to help the
-programmer identify the cause of the crash. The debugging information is
-typically a stack trace, which includes an explanatory message and the execution
-stack frame when the exception occurred.
-%In Java, stack trace consists of an ordered list of methods that were active on
-%the call stack before the exception occurred.
-
-Although the exception handling mechanism has been the target of several studies
-(e.g.~\cite{miller1997issues,Robil00,shah2010understanding,
-garcia2007extracting,garcia2001comparative,cabral2007exception,coelho2011unveiling}),
-the exception handling code is generally poorly understood and often the least
-tested part of software systems~\cite{coelho2011unveiling}.  To analyze the
-exception handling code, current studies use either static (source code-based)
-analysis or dynamic (runtime-based) testing methods. Neither method can explore
-the full execution space of a program and therefore they both fail to account
-for exceptions implicitly signaled by the runtime environment (e.g. due to
-resource depletion) or dynamically loaded code. Exception stack traces, when
-available, enable the post-mortem analysis of exception handling behaviours of
-programs and consequently the evaluation of exception handling mechanisms in
-programming languages.
-
-The availability of exception stack traces until recently was scarce, while the
+The pervasiveness of the World Wide Web and the astonishing increase in the
+number of smartphones have stimulated the creation of a new class of
+applications which extend the phones capabilities far beyond of the basic calls
+and textual messages.
+
+Such applications run over one or more platforms, among them the Android platform
+leads the smartphone market, comprising around XXXX [GoogleIO2014], thanks to its
+Java-based open-source model and hardware partners like Samsung, HTC, Motorola, Asus 
+(besides Nexus its own device) [ref].
+
+Such applications are intrinsically multi-threaded and the development model and used APIs 
+can be  slightly or drastically different and more error prone than tradicional development [arie-icse].
+
+Moreover, such applications have to cope with an increasing number of abnormal
+computation states. Besides the known source of failures (i.e., programmming errors, noisy 
+user inputs or faults in underlying middleware or hardware) such applications need to deal with abnormal states that 
+ arise as a consequence of:  memory and batery restrictions [ref], faults on third party services and libraries, 
+compatibility issues caused by constant platforms updates [ref], conection with devices such as GPS, camera.
+
+Therefore, techniques for error detection and handling are not  an optional add-on but a fundamental part of any system~\cite{bruntink2006discovering}.
+The Android Platform uses the exception handling mechanism embeeded
+ in Java to deal with such conditions. However, although the exception handling mechanism
+~\cite{goodenough1975exception} is one of the most used schemes for
+detecting and recovering from such exceptional conditions, studies (e.g.~\cite{miller1997issues,Robil00,shah2010understanding, garcia2007extracting,garcia2001comparative,cabral2007exception,coelho2011unveiling}),
+have shown that it is generally poorly understood and the less tested parts of the system ~\cite{coelho2011unveiling}.
+The consequence of the exception handling code may come the other way around making the code even less robust.
+
+Our conjecture is that several Android applicatons may be suffering from similar vulnerabilities on the exception
+handling code, and that such vulnerabiliries may be leading to application crashes.
+
+Untill recent years, the availability of issues reporting exceptional conditions was scarce, while the
 available data concerned a few large scale projects (e.g. the Mozilla
-crash report
-dataset\footnote{https://code.google.com/p/promisedata/issues/detail?id=88}).
+crash report dataset\footnote{https://code.google.com/p/promisedata/issues/detail?id=88}).
 Along the recent years, we have witnessed a rapid increase in the number of open
 source projects that make their control version and issue tracking systems
-publicly available in repository hosting sites such as GitHub. The
-concentration of open source projects and the existence of initiatives that mine
-such information, such as GHTorrent~\cite{Gousi13}, allow the information about
-projects to be visible within and across projects and make a full population
-studies possible.
-
-In this work, we perform the first post mortem analysis of exception handling in
-Java. Using a custom tool called ExceptionMiner, which we develop specifically
-for this study, we mine stack traces from issues across all Java projects on
-GitHub. Overall, we analyzed 356,057 issues from 16,836 projects, from which
-28,800 stack traces were extracted and used as a source of information for our
-analysis. The stack trace analysis was augmented by additional information
-extracted using byte code and source code analysis done on a carefully selected
-subset of our initial sample (482 Android projects). Manual inspection was
-used to leverage the understanding of stack traces and support
-further discussions and insights.
-To guide our analysis, we compiled a set of best practices on how to use
-exceptions by composing relevant suggestions by Gosling~\cite{gosling2000java},
-Wirfs-Brock~\cite{wirfs2006toward} and Bloch~\cite{bloch2008effective}.  The
-comparison of those with the mined exceptions revealed several interesting
-findings, such as:
+publicly available in repository hosting sites such as GitHub and Googlecode.
+
+% The
+%concentration of open source projects and the existence of initiatives that mine
+%such information, such as GHTorrent~\cite{Gousi13}, allow the information about
+%projects to be visible within and across projects and make a full population
+%studies possible.
+
+Among the issues defined on issue trackers, we can find issues reporting bugs, and system crashes 
+found as a result of application executions. Such issues usually contain exception stack traces [ref][ref]. 
+When available, such information can enable different kinds of post-mortem analysis, such as:
+: classify .... (get from previous versions).
+
+Such analysis focuses a system at a time, and none of them target a family of Android Application
+(which share a common core) looking for similar characteristics and vulnerabilitys on the crashes reported.
+
+In this work, we perform the first post mortem analysis of the exception stack traces embedded on issues
+of all Android applications available in GitHub and Googlecode. Using a custom tool called ExceptionMiner,
+ which we develop specifically for this study, we mine stack traces from issues across all Android projects on
+GitHub and Googlecode. Overall, we analyzed the XXXX issues extracted from on XXX Android projects defined in
+GitHub and XXX projects defined in Googlecode. From these issues  XXX stack traces were extracted and used as
+ a source of information for our analysis. The stack trace analysis was augmented by additional information
+extracted from JDK, Android SO and Android SDK using byte code and source code analysis - which was done 
+on a carefully selected subset of our initial sample (482 Android projects). Manual inspection was
+used to leverage the understanding of exception stack traces and support further discussions and insights.
+
+%To guide our analysis, we compiled a set of best practices on how to use
+%exceptions by composing relevant suggestions by Gosling~\cite{gosling2000java},
+%Wirfs-Brock~\cite{wirfs2006toward} and Bloch~\cite{bloch2008effective}.  The
+%comparison of those with the mined exceptions revealed several interesting
+%findings, such as:
+
+Some outcomes consistently detected during this study were the following:
 
@@
 
-  \item  Odd exception chains (e.g., Error wrapping checked and runtime
-    exceptions and vice-versa), indicating that the purpose of Java's hybrid
-    exception model may not have been adequately used.
-
-  \item  Issues reporting uncaught checked exceptions.
+   \item  Unexpected exception wrappings (e.g., Error wrapping checked and runtime
+    exceptions and vice-versa), indicating that the purpose of Java's hybrid exception model may not have been adequately used.
+
+   \item  The development model of Android platform (strongly based on system call-backs) encouraged 
+     the use of general RuntimeExceptions.
+
+  %\item  Issues reporting uncaught checked exceptions.
 
@@
 
-The contributions of this study are as follows:
+Although the study focused on Android opensource applications, some of the results are not restricted to this context, 
+for instance: some of the unexpected wrappings were performed by Java libraries and elements (e.g. reflection library and 
+the static blocks); and the undocumented runtime exceptions were also found in Java libraries beyonds the Android plaftorm.
+The study results are relevant to: (i) Android and Java developers to familiarize with the most common reported exceptions 
+mis-uses which is the first step to help developers to avoid making them; (ii) designers of languages to consider ways of reducing the abundance of NullPointerExceptions; and (iii) tool designers to consider developing tools that enable a cross-project defect analysis.
+
+Hence, the main contributions of this study are as follows:
 \begin{itemize}
 
-  \item  It performs the first large scale analysis of Java stack traces and how
-    they can reveal bad-practices on the use of exceptions in Java.
+  \item  It performs a large scale post morten analysis of exception stack traces of Android projects.
 
@@
 
+  \item  It presents the the exploratory study methodology to ..... (complete - view the reviewer comment)
+
 \end{itemize}
-
-These contributions are relevant to:
-(i) developers of Java applications
-to familiarize with the most common reported exceptions mis-uses which is the
-first step to help developers to avoid making them;
-(ii) designers of languages
-to consider ways of reducing the abundance of NullPointerExceptions;
-and (iii)
-tool designers to consider developing tools that enable a cross-project defect
-analysis.
 
@@
 
-%This paper is organized as follows. Section 2 presents a
-%background on exception handling mechanism. Section 3 presents the study design.
-%Section 4 reports study findings. Section 5 provides further discussions and insights.
-%Section 6 presents the threats to validity associated to this study. Finally Section
-%7 describes the related work, and Section 8 presents our conclusions and
-%directions for future work.
+The ramaining of this paper is organized as follows. Section 2 presents a
+background on Android Plaform and Java exception handling mechanism. 
+Section 3 presents the study design. Section 4 reports study findings. 
+Section 5 provides further discussions and insights.
+Section 6 presents the threats to validity associated to this study. Finally Section
+7 describes the related work, and Section 8 presents our conclusions and
+directions for future work.
 
@@
 
-\subsection{Exception Handling in Programming Languages}
-
-Techniques for error detection and handling are not  an optional add-on but a
-fundamental part of any system~\cite{bruntink2006discovering}. 
-The exception mechanism~\cite{goodenough1975exception} was proposed as a way of
-preventing errors from being ignored and was embedded in many mainstream
-programming languages ~\cite{garcia2001comparative} since then.
-
-In languages such as Modula-3, Guide, and Extended Ada the exceptions are 
-checked, which means that the compiler statically checks if appropriate handlers 
-are provided within the system. Accordingly, the programmer must explicitly specify 
-every exception that can be signaled to support the compile time checking. This approach is known to be
-costly to maintain (Dooren and Steegmans, 2005) since exception updates are
-cascading; if a new exception type is added to a method signature then all other
-methods directly using this method must either change their signatures to
-include the new exception type or handle the new exception.
-
-In languages such as Lore, C++, and Arche all exceptions are unchecked but the developer can 
-optionally list exception types on method signatures. Although exceptions can 
-be documented on methods signatures, the developer is not warned by the compiler
-if an unchecked exception is not handled inside the application. According to ~\cite{Robil00} 
-such an approach may hinder robustness because the client of a method cannot easily know
-which unchecked exceptions may be thrown (unless the code of the method and the
-methods called from it is inspected) which can be a very time consuming or
-infeasible task.
-
-Java [ref: Gosling et al. 1996] and CLU [Liskov and Snyder 1979] support the checked exceptions,
- (therefore the compiler can then check if handlers are provided to such exceptions), however, this support is 
-only partial because both languages also provides a form of unchecked exceptions [ref: Robilard].
-In CLU,  if a handler is not found for a given exception, it is automaticaly converted in a predened default exception, 
-called failure which is the only one which can be implicitly propagated [ref-clue].
- In Java the developer can define specific unchecked and checked exceptions and use both kinds in a single system.
-
-There is a long-lasting debate about the pros and cons of checked and unchecked exceptions 
-~\cite{javatut,stackoverlow,debate},\footnote{152 questions in Stackoverlow
-are related to this debate}. However, the debates are mainly based on the developers own experience in
-using both models, since there is still few empirical evidence of the pros and cons of each model.
-
-%Although exception handling mechanism has been embedded in several programming languages 
-%as show above,  the exception handling code is often poorly understood and the least
-%tested part of software systems [ref:POORLY].
-
-%[REF]: Static Analysis to Support the Evolution of Exception Structure in Object-Oriented Systems
-
-% This behaviour may lead to the problem of uncaught exceptions~\cite{jo2004uncaught}, one of the main causes of application crashes
-%(e.g.~\cite{miller1997issues,Robil00,shah2010understanding, garcia2007extracting, garcia2001comparative,cabral2007exception,coelho2011unveiling}).
-
-\subsection{Checked an Unchecked Exceptions in Java} \label{sec:extypes}
+\subsection{Java Exception Handling Mechanism} \label{sec:extypes}
 
@@
 Figure~\ref{fig:exchier}.  According to it every exception is an
-instance of the Throwable class, and can be of three kinds: checked exceptions
-(extends Exception), run-time exceptions (extends RuntimeException), and errors
-(extends Error)~\cite{gosling2000java}. Neither run-time exceptions nor errors need to be specified on method
+instance of the Throwable class, and can be of three kinds: the checked exceptions
+(extends Exception) - means that the compiler statically checks if appropriate handlers 
+are provided within the system, to enable such checking each checked exception must be specified 
+on the method signature;  the run-time exceptions (extends RuntimeException) and errors
+(extends Error)~\cite{gosling2000java}  -  which do not need to be specified on method
 signatures, and hence both are referred as unchecked exceptions.
@@
 
-\begin{figure} \centering \includegraphics[scale=0.6]{wrappings}
+\begin{figure} \centering \includegraphics[scale=0.]{stacktrace.png}
 \caption{Example of an Exception stack trace in Java.}
@@
 \end{figure}
+
+% \subsection{Public Project Repositories and Issue Trackers} 
+% ONE OF THE REVIEWERS COMPLAINED ABOUT THE ABSENCE OF EXPLANATIONS OF WHAT AN ISSU IS...
+% As mentioned before, currently more and more projects are becomming available in 
+% public project repositories such as GitHub and Google Code.
+% Such repositories provides facilities such as....
+% The issue trackers available in such projects ....
+
+\subsection{The Android Platform} \label{sec:extypes}
+
+\begin{figure} \centering \includegraphics[scale=0.8]{arch3.png}
+  \caption{Android Archtecture} \label{fig:exchier} \end{figure}
 
@@
 
-  \item[RQ2] What patterns of exception (mis)-use emerge from the exception stack trace
-    analysis?
+  \item[RQ2] What patterns of exception (mis)-use related to exception types and exception propagation emerge from the exception stack trace analysis?
 
@@
 
+\subsubsection{Google Code}
+
+Until January 2014 there were 788 android projects available in Google code repository from which 724 defined at least 1 issue and 183 projects 
+had at least one issue including a stack trace. To identify the Android projects on Googlecode we used the search facilities provided looking for 
+projects mentioning "Android" on the name of the project or its description. The repository page of each project was inspected, and from this set
+142 corresponded to Android applications.
+
+To mine the information available on such Google code projects we implemented a web crawler which navegated on each project and issue page
+and extracted the content of each issue to store on relational database.
+\footnote{Google code previoulsly provided a webservice to extract information about each project, but this service was removed in what Google called "clean-up action" which happened in June 2013.}
+
 \subsubsection{Exception Stack Traces on Issues}
@@
 on the defect issues can be found at:
-\url{www.dimap.ufrn.br/~roberta/icsme2014}}
+\url{www.dimap.ufrn.br/~roberta/android_repo_mining}}
 
@@
 
-After distilling the information available on the 28,800 stack traces (extracted from the 
-issues and issues comments of 16,836 Java projects) we found 607 distinct exceptions
-reported as the root causes of stack traces.
-
-Table \ref{tab:toptenjava} illustrates the top 10 root couses found in the study ranked by the number of distinct
-project repositories on which they were reported. The table illustrates the
-number of times each exception was reported as the root cause of a stack trace
-, the number of distinct projects reporting such causes on stack traces, and the percentages of
-such in relation to all reported exceptions and analyzed projects respectaly.
-
-\begin{table}
-\begin{tabular}{lrrrr}
-    \hline
-    \bfseries{Exception} &  \multicolumn{2}{c}{\bfseries{Occurrences}} &  \multicolumn{2}{c}{\bfseries{Projects}}\\
-    & \bfseries{\#} &  \bfseries{\%} & \bfseries{\# } & \bfseries{\% } \\
-    \hline
-java.lang.NullPointerException             & 7578 & 26,31\% & 1671 & 44,47 \\
-java.lang.IllegalArgumentException         & 1573 & 5,46\%  & 670  & 17,83\% \\
-java.lang.ClassNotFoundException           & 1203 & 4,18\%  & 514  & 13,68\% \\
-java.lang.RuntimeException                 & 1093 & 3,80\%  & 411  & 10,94\% \\
-java.lang.IllegalStateException            & 964  & 3,35\%  & 431  & 11,47\% \\
-java.lang.ClassCastException               & 886  & 3,08\%  & 408  & 10,86\% \\
-java.lang.ArrayIndexOutOfBoundsException   & 834  & 2,90\%  & 345  & 9,18\% \\
-java.lang.NoSuchMethodError                & 822  & 2,85\%  & 369  & 9,82\% \\
-java.io.IOException                        & 656  & 2,28\%  & 337  & 8,97\% \\
-java.lang.OutOfMemoryError                 & 601  & 2,09\%  & 235  & 6,25\% \\
-%java.lang.NoClassDefFoundError            & 554  & 1,92\%  & 299  & 7,96\% \\
-%java.util.ConcurrentModificationException & 465  & 1,61\%  & 180  & 4,79\% \\
-%java.lang.IndexOutOfBoundsException       & 445  & 1,55\%  & 205  & 5,46\% \\
-\hline
-All	 &	23956 &	100\%	& 3758 & 100\%	 \\
-\hline
-  \end{tabular}
-\caption{Top 10 most popular exceptions on issues in Java projects}
-\label{tab:toptenjava}
-\end{table}
-
-We can observe that most of this set is composed by exceptions defined by java.lang,
-which are usually implicitly thrown by JMV dur to a programming mistake. From thi set
-the java.lang.NullPointerException  was the most reported root cause;  7578 stack traces 
-had a java.lang.NullPointerException  as root cause, which corresponds of 26,31\%  of 
-all analyzed stack traces. Considering the number of distinct project repositories, 
-the java.lang.NullPointerException, was reported on at least one issue of 1,671 projects 
-(which represents 44,47\% of all projects reporting at least one stack trace).
-
-The high prevalence of NullPointerExceptions is aligned with the findings other 
-works~\cite{kim2013predicting,fraser20131600,csallner2004jcrasher}. For instance, Sunghun et
-al.~\cite{kim2013predicting} showed that in Eclipse bug report system 38\% the bugs 
-related to exception handling were caused by NullPointerException; other works on robustness 
-testing~\cite{maji2012empirical,csallner2004jcrasher} showed that most of the automatically 
-detected bugs were due to NullPointerExceptions and implicitly-signaled of Java
-environment (as the ones found in this study).
-
-In Section XX we provide further discussion on the NullPointerException problem, pointing 
-to ways of how to deal with it and mitigate it.
-
-\subsubsection{Classifying Top 100 Exceptions}
-
-To get a broather view of the root causes of stack traces, we performed an empirical evaluation
-whose goal was to categorized the top 100 root causes.
-
-After a first inspection of the complete list of exceptions reported, a first set of exception 
-categories was defined. This list represent the most common terms reffered on exception
-names and javadoc documentation. A similar approach was addopted in [COELHO-ECOOP].
-The list of categories is illustrated in Table~\ref{tab:categories}. Hence in the second phase
-of the study, each exception was mapped to one of the categories previously defined - 
-this analysis was based on the inspection of source code, Javadoc, or additional 
-documentation related to the exception.
-
-To ensure the quality of the classification, three independent coders classified a randomly selected
-sample of 25 exception types (from the total 100) using the same categories;
-the interater agreement was high (96\%). We can therefore assume that the
-categories are descriptive enough for covering the whole range of exception
-types that can occur in real world Java projects.
-
-\begin{table}
-  \centering
-  \begin{tabular}{|p{2cm}| p{5cm}|}
-    \hline
-    \bfseries{Category} & \bfseries{Description} \\
-    \hline
-      Programming logic &  Exceptions implicitly  thrown by the JVM when the 
-evaluation of an expression violates the normal semantics of language (e.g., 
-out-of-bounds array index, division-by-zero, access to a null reference). 
-Such exceptions are usually defined on java.lang and java.util packages. \\ \hline
-      Resources (IO)                         & Exception releated to the manipulation of IO resources or device  (e.g., file, network) \\ \hline
-      Security                               & Exception releated to security issues (e.g. password validation, criptography) \\ \hline
-      Concurrency                            & Exceptions related to multi-threaded programming \\ \hline
-      Backward compatibility                 &  A list of exceptions related to backward compatibility - a complete list can be found here [ref]  \\ \hline
-      Reflection                             & Exceptions related to Java reflection library  \\ \hline
-      Specific                & Exceptions created for specific applications, frameworks or libraries \\ \hline
-      General (Exception, Error, Runtime)    & The top level exceptions in Java    \\ \hline
-    \hline
-  \end{tabular}
-  \caption{Description of categories.}
-  \label{tab:categories}
-\end{table}
-
-The top 100 root causes correspond to approximately 77\% of all stack traces
-reported reported on Java stack-issues and 95\% of all exceptions reported on
-Android stack-issues. Table~\ref{tab:tophundrend} illustrates the result of such 
-classification.
-
-\begin{table}
-  \centering
-  \begin{tabular}{lrr}
-    \hline
-    \bfseries{Category} & \bfseries{Java} & \bfseries{Android Subset} \\
-    \hline
-      Programming logic (java.lang and util) & 12625  (41,10\%) & 2235 (55,64\%)\\
-      Resources (IO)                         & 12440 (40,50\%)  & 727 (18,10\%) \\
-      Security                               & 220  (0,72\%)    & 165 (4,11\%)\\
-      Concurrency                            & 633 (2,06\%)     & 116  (2,89\%)\\
-      Backward compatibility                 & 1580 (5,14\%)    & 219 (5,45\%) \\
-      Reflection                             & 1413 (4,60\%)    & 91 (2,27\%)\\
-      Specific (GUI,FRAMEWORK)               & 340 (1,11\%)     & 197 (4,90\%)\\
-      General (Error, Exception, Runtime)    & 1468 (4,78\%)    & 267 (6,65\%)\\
-    \hline
-  \end{tabular}
-  \caption{Characterization of the 100 root causes.}
-  \label{tab:tophundrend}
-\end{table}
-
-Programming errors are the causes of most of the exception stack traces reported
-on issues both in  general Java data set and Android subset. In the Android subset we could find 
-more stack traces reporting exceptions related to Backward
-compatibility and security concerns than on the general data set.
-
-TODO: Discuss about the interater agreement here...
-
-\subsection{Investigating The Android Subset}
-
-As detailed in Section~\ref{sec:android}, 482 Android projects were carefully
-selected to enable a deeper investigation of stack trace information, by
-leveraging stack trace information with the source code and bytecode analysis.
-For this subset of Android project, we could extract every package related to
-each exception signaler, and based on this information we could classify every
-stack trace based on its root signaler as presented in Table~\ref{tab:signalers}.
+After distilling the information available on the XXXX  stack traces, extracted from the 
+issues and issues comments of XXXX Android projects on GitHub and Googlecode,  
+we found XXXX  distinct exceptions reported as the root causes of stack traces.
+Table \ref{tab:toptenandroid} illustrates the top 10 root couses found in the study ranked by the number of distinct
+project repositories on which they were reported.
+
+\begin{table*}
+  \centering
+  \begin{tabular}{rcccccccc}
+    \hline
+    \bfseries{Root Exception} &  \multicolumn{2}{c}{\bfseries{Projects}} &  \multicolumn{2}{c}{\bfseries{Occurrences}} & \textsf{Android} & \textsf{Libcore} & \textsf{App} & \textsf{Lib} \\
+    & \bfseries{\#} &  \bfseries{\%} & \bfseries{\# } & \bfseries{\% } &&&&\\
+    \hline
+java.lang.NullPointerException            & 254 & 52,70\%  & 1225 & 30,08\% & 473 & 18 & 595 & 137 \\
+java.lang.IllegalStateException          & 99  & 20,54\%  & 234  & 5,75\%  & 165 & 12 & 36  & 20  \\
+java.lang.IllegalArgumentException          & 94  & 19,50\% & 255  & 6,26\% & 146 & 6  & 64  & 39   \\
+java.lang.RuntimeException                & 91  & 18,88\% & 232  & 5,70\%  & 167 & 1  & 47  & 17   \\
+java.lang.OutOfMemoryError                 & 56  & 11,62\% & 180  & 4,42\% & 121 & 15 & 17  & 23   \\
+java.lang.NoClassDefFoundError           & 52  & 10,79\%  & 73   & 1,79\%  & 9   & 0  & 37  & 26   \\
+java.lang.ClassCastException             & 49  & 10,17\%  & 94   & 2,31\%  & 43  & 0  & 40  & 11   \\
+java.lang.IndexOutOfBoundsException       & 47  & 9,75\%   & 127  & 3,12\% & 47  & 0  & 71  & 8    \\
+java.lang.NoSuchMethodError               & 40  & 8,30\%  & 57   & 1,40\%  & 9   & 0  & 39  & 9   \\
+java.util.ConcurrentModificationException & 38  & 7,88\% & 54   & 1,33\%   & 5   & 0  & 43  & 6   \\
+
+    \hline
+  \end{tabular}
+\caption{Root Exceptions occurrences and popularity.}
+\label{tab:toptenandroid}
+\end{table*}
+
+%\begin{figure}
+%\centering
+%\includegraphics[width=\hsize]{top_exceptios_android_new2.pdf}
+%\includegraphics[width=\hsize]{top_exceptions_android_new}
+%\caption{Top exceptions in Android repositories.}
+%\label{fig:androidsignaler}
+%\end{figure}
+
+Table \ref{tab:toptenandroid} illustrate the number of times each exception was reported as 
+the root cause of a stack trace, the number of distinct projects reporting such causes on stack traces, 
+and the percentages of such in relation to all reported exceptions and analyzed projects respectaly.
+The table also presents the whether the method responsible for signaling the exceptions were
+part of Android, the aplication itself, libcore, or a third-party library as presented in Table  \ref{tab:signalers}  \footnote{9 stack traces contained only elements of JVM and were discarded from the analysis}.
 
@@
     \bfseries{lib}     & If the exception is thrown by a method that was not defined by any of the elements above.\\
-    \bfseries{java}    & If all methods on the stack trace are JDK library methods.\\
     \hline
@@
 
-For each exception we could discover the number of times it occurred on stacks,
+For each exception we could discover the number of times it occurred on exception stack traces,
 the number of distinct projects on which such stacks were defined (and consequently its
@@
 presents the mined data.
-
-\begin{table*}
-  \centering
-  \begin{tabular}{rccccccccc}
-    \hline
-    \bfseries{Root Exception} &  \multicolumn{2}{c}{\bfseries{Occurrences}} &  \multicolumn{2}{c}{\bfseries{Projects}} & \textsf{android} & \textsf{libcore} & \textsf{app} & \textsf{lib} & \textsf{java} \\
-    & \bfseries{\#} &  \bfseries{\%} & \bfseries{\# } & \bfseries{\% } &&&&&\\
-    \hline
-java.lang.NullPointerException            & 1225 & 30,08\% & 254 & 52,70\% & 473 & 18 & 595 & 137 & 2 \\
-java.lang.IllegalStateException           & 234  & 5,75\%  & 99  & 20,54\% & 165 & 12 & 36  & 20  & 1 \\
-java.lang.IllegalArgumentException        & 255  & 6,26\%  & 94  & 19,50\% & 146 & 6  & 64  & 39  & 0 \\
-java.lang.RuntimeException                & 232  & 5,70\%  & 91  & 18,88\% & 167 & 1  & 47  & 17  & 0 \\
-java.lang.OutOfMemoryError                & 180  & 4,42\%  & 56  & 11,62\% & 121 & 15 & 17  & 23  & 4 \\
-java.lang.NoClassDefFoundError            & 73   & 1,79\%  & 52  & 10,79\% & 9   & 0  & 37  & 26  & 1 \\
-java.lang.ClassCastException              & 94   & 2,31\%  & 49  & 10,17\% & 43  & 0  & 40  & 11  & 0 \\
-java.lang.IndexOutOfBoundsException       & 127  & 3,12\%  & 47  & 9,75\%  & 47  & 0  & 71  & 8   & 1 \\
-java.lang.NoSuchMethodError               & 57   & 1,40\%  & 40  & 8,30\%  & 9   & 0  & 39  & 9   & 0 \\
-java.util.ConcurrentModificationException & 54   & 1,33\%  & 38  & 7,88\%  & 5   & 0  & 43  & 6   & 0 \\
-
-    \hline
-  \end{tabular}
-\caption{Root Exceptions occurrences and popularity.}
-\label{tab:toptenandroid}
-\end{table*}
-
-%\begin{figure}
-%\centering
-%\includegraphics[width=\hsize]{top_exceptios_android_new2.pdf}
-%\includegraphics[width=\hsize]{top_exceptions_android_new}
-%\caption{Top exceptions in Android repositories.}
-%\label{fig:androidsignaler}
-%\end{figure}
 
@@
 Only high fault tolerant systems need to provide solutions to handle them.
+
+----------
+
+We can observe that most of this set is composed by exceptions defined by java.lang,
+which are usually implicitly thrown by JMV dur to a programming mistake. From thi set
+the java.lang.NullPointerException  was the most reported root cause;  7578 stack traces 
+had a java.lang.NullPointerException  as root cause, which corresponds of 26,31\%  of 
+all analyzed stack traces. Considering the number of distinct project repositories, 
+the java.lang.NullPointerException, was reported on at least one issue of 1,671 projects 
+(which represents 44,47\% of all projects reporting at least one stack trace).
+
+The high prevalence of NullPointerExceptions is aligned with the findings other 
+works~\cite{kim2013predicting,fraser20131600,csallner2004jcrasher}. For instance, Sunghun et
+al.~\cite{kim2013predicting} showed that in Eclipse bug report system 38\% the bugs 
+related to exception handling were caused by NullPointerException; other works on robustness 
+testing~\cite{maji2012empirical,csallner2004jcrasher} showed that most of the automatically 
+detected bugs were due to NullPointerExceptions and implicitly-signaled of Java
+environment (as the ones found in this study).
+
+In Section XX we provide further discussion on the NullPointerException problem, pointing 
+to ways of how to deal with it and mitigate it.
+
+\subsection{Investigating The Android Subset}
+
+----------
+
+\subsection{Direct instances of RuntimeExceptions being thrown.}
+
+From Tables~\ref{tab:toptenjava} and~\ref{tab:toptenandroid} we can observe that,
+in both Java and Android related issues, direct instances
+java.lang.RuntimeException were thrown  in 10,94\% of Java repositories and
+18,88\% of Android repositories respectively. 
+In the Android repositories, most of such exceptions  were  thrown by the Android 
+platform/OS (167 out of 232).
+
+Throwing general exceptions, such as direct instances of RuntimeException, is  considered a
+bad practice, because the exception type does not carry enough information to identify the
+cause of the exceptional behaviour, and as a consequence developers need to relying on
+the exception message which can be neither complete nor precise~\cite{gosling2000java}.
+
+\subsubsection{Classifying Top 100 Exceptions}
+
+To get a broather view of the root causes of stack traces, we performed an empirical evaluation
+whose goal was to categorized the top 100 root causes.
+
+After a first inspection of the complete list of exceptions reported, a first set of exception 
+categories was defined. This list represent the most common terms reffered on exception
+names and javadoc documentation. A similar approach was addopted in [COELHO-ECOOP].
+The list of categories is illustrated in Table~\ref{tab:categories}. Hence in the second phase
+of the study, each exception was mapped to one of the categories previously defined - 
+this analysis was based on the inspection of source code, Javadoc, or additional 
+documentation related to the exception.
+
+To ensure the quality of the classification, three independent coders classified a randomly selected
+sample of 25 exception types (from the total 100) using the same categories;
+the interater agreement was high (96\%). We can therefore assume that the
+categories are descriptive enough for covering the whole range of exception
+types that can occur in real world Java projects.
+
+TODO: SEPARAR JAVA DE ANDROID
+
+\begin{table}
+  \centering
+  \begin{tabular}{|p{2cm}| p{5cm}|}
+    \hline
+    \bfseries{Category} & \bfseries{Description} \\
+    \hline
+      Programming logic &  Exceptions thrown by the JVM when the 
+evaluation of an expression violates the normal semantics of language (e.g., 
+out-of-bounds array index, division-by-zero, access to a null reference). \\ \hline
+      Resources (IO, Memory, Energy)  & Exception releated to the manipulation or restriction of resources (e.g., file, network, memory) \\ \hline
+      Security                               & Exception releated to security issues (e.g. password validation, criptography) \\ \hline
+      Concurrency                            & Exceptions related to multi-threaded programming \\ \hline
+      Backward compatibility                 &  A list of exceptions related to backward compatibility - a complete list can be found here [ref]  \\ \hline
+      Reflection                             & Exceptions related to Java reflection library  \\ \hline
+      Specific                & Exceptions created for specific applications, frameworks or libraries \\ \hline
+      General (Exception, Error, Runtime)    & The top level exceptions in Java    \\ \hline
+    \hline
+  \end{tabular}
+  \caption{Description of categories.}
+  \label{tab:categories}
+\end{table}
+
+The top 100 root causes correspond to approximately 77\% of all stack traces
+reported reported on Java stack-issues and 95\% of all exceptions reported on
+Android stack-issues. Table~\ref{tab:tophundrend} illustrates the result of such 
+classification.
+
+\begin{table}
+  \centering
+  \begin{tabular}{lrr}
+    \hline
+    \bfseries{Category} &  \bfseries{ Ocurrences (\%)} \\
+    \hline
+      Programming logic (java.lang and util) & 2235 (55,64\%)\\
+      Resources (IO)                         &  727 (18,10\%) \\
+      Security                               &  165 (4,11\%)\\
+      Concurrency                            & 116  (2,89\%)\\
+      Backward compatibility                 &  219 (5,45\%) \\
+      Reflection                             &  91 (2,27\%)\\
+      Specific (GUI,FRAMEWORK)               &  197 (4,90\%)\\
+      General (Error, Exception, Runtime)    &  267 (6,65\%)\\
+    \hline
+  \end{tabular}
+  \caption{Characterization of the 100 root causes.}
+  \label{tab:tophundrend}
+\end{table}
+
+Programming errors are the causes of most of the exception stack traces reported
+on issues both in  general Java data set and Android subset. In the Android subset we could find 
+more stack traces reporting exceptions related to Backward
+compatibility and security concerns than on the general data set.
+
+TODO: Discuss about the interater agreement here...
 
@@
 
-To answer this question we analyzed general characteristics of reported stack traces and, in the case of stack traces reported on Android repositories, we combine stack trace information with
-source code and byte code analyses to enable a more detailed analysis.
-Doing so, we mined some characteristics of stacks that points to best practices violations.
-
-\subsection{Direct instances of RuntimeExceptions being thrown.}
-
-From Tables~\ref{tab:toptenjava} and~\ref{tab:toptenandroid} we can observe that,
-in both Java and Android related issues, direct instances
-java.lang.RuntimeException were thrown  in 10,94\% of Java repositories and
-18,88\% of Android repositories respectively. 
-In the Android repositories, most of such exceptions  were  thrown by the Android 
-platform/OS (167 out of 232).
-
-Throwing general exceptions, such as direct instances of RuntimeException, is  considered a
-bad practice, because the exception type does not carry enough information to identify the
-cause of the exceptional behaviour, and as a consequence developers need to relying on
-the exception message which can be neither complete nor precise~\cite{gosling2000java}.
-
-\subsection{Undocumented runtime exceptions being thrown.}
-
-To better investigate the extent of this problem we classified each stack trace according to the type of the
-root exception. Table~\ref{tab:typeroottab} presents the types of root causes of all stack traces reported on 
-the Android repositories. As we can see from Table ~\ref{tab:typeroottab},  most of the exceptions signaled by third-party code
- (e.g., Android and libs) were runtime exceptions. After filtering all the exceptions implicitly signaled by 
-JVM (due to programming mistakes) and inspecting the signaler methods of such exceptions (i.e.,  direct an indirect 
-instances of java.lang.RuntimeException). We could observe that only 1 method out of 118 inspected signalers
- (i.e., 0,8\%) documented the explicitly thrown runtime exception in a Javadoc comment; and none of these methods
-included the runtime exception  (reported on the stack) as part of the exception interface of the method (i.e., using 
-throws clause on method signature). This result is aligned with the results of other  study conducted by from 
-Sacramento et al ~\cite{sacramento2006unchecked} which observed that the
-runtime exceptions in .Net programs are most often not documented.
-
-Such undocumented runtime exceptions represent a threat to system robustness, specially
-when such exceptions are thrown then the third party code is invoked inside the application (e.g. libraries, or framework utility code).
-In such cases the client usually do not have access to the source code, and in the absence of
-the exception documentation it is very difficult or even impossible for the client of such third party code to 
-protect the application against undocumented runtime exceptions. As a consequence, the
- undocumented runtime exception may remain uncaught and lead to system crashes.
-
-\begin{table}
-\centering
-\begin{tabular}{lcccccc}
-    \hline
-    \bfseries{Type} & \bfseries{Android} & \bfseries{Libcore} & \bfseries{App} & \bfseries{Lib} & \bfseries{Java} & \bfseries{All}\\
-    \hline
-
-Runtime	&	1075	&	56	&	1446	&	374	&	10	&	2961	\\
-Error	&	144	&	34	&	168	&	105	&	9	&	460	\\
-Checked	&	110	&	230	&	139	&	145	&	12	&	636	\\
-Throwable	&	0	&	0	&	1	&	0	&	0	&	1	\\
-Undefined	&	1	&	0	&	6	&	8	&	0	&	15	\\
-    \hline
-All		& 1330	&	320	&	1760	&	632	&	31	&	4073	\\
-    \hline
-  \end{tabular}
-\caption{Types of root exceptions.}
-  \label{tab:typeroottab}
-\end{table}
-
-\begin{figure}
-\centering
-%\includegraphics[width=\hsize]{chart_exceptiontypes.pdf}
-\includegraphics[width=\hsize]{exception_types.pdf}
-\caption{Distribution of Checked and Unchedked exceptions.}
-\label{fig:typeroot}
-\end{figure}
-
-On the other hand, we could observe that most of the exceptions signaled by the set of 
+To answer this question we: (i) distilled the information available on stack traces (i.e., exception signalers, root causes, exception wrappings); 
+(ii) mined the exception types related to each exception reported on stack traces using source code and byte code analyses; and finally checked 
+this information against the compiled best practices. Doing so, we mined some characteristics of exception stack traces that points to best practices violations.
+
+ Table~\ref{tab:typeroottab} presents the types of root causes of all stack traces reported on the analyzed issues. As we can see 
+from Table ~\ref{tab:typeroottab},  most of the reported exceptions were signaled by Android and libraries reused by Apps were runtime exceptions. 
+On the other hand, we could observe that most of the reported exceptions signaled by the set of 
 libraries reused by Android platform (i.e., org.apache.harmony,
@@
 
-\subsection{Uncaught Checked Exceptions}
-
-Although most of the  reported exceptions were runtime, we could also find checked
-exceptions as the root causes of stack traces reported on issues. A checked exception
-can remain uncaught in two circumstances: if all methods on the execution trace
-on which this exception flows explicitly specifies the exception, or if the
-checked exception is wrapped in a runtime and than re-thrown. Since checked exceptions 
-should be used to represent recoverable conditions, ignoring a
-checked exception, is considered a bad practice. As mentioned before, when the developer is confronted
-with a checked exception, the designer of the API is telling him to handle the
-exceptional condition. To better understand what was causing checked exceptions
-to scape from being handled we investigated the kinds of wrappings that happened
-on stacks the stacks.  Next sections presents the wrappings found and discuss
-about them.
-
-%\noindent\emph{3) Odd Wrappings}
-\subsection{Odd Wrappings}
+\begin{table}
+\centering
+\begin{tabular}{lccccc}
+    \hline
+    \bfseries{Root Cause Type} & \bfseries{Android} & \bfseries{Libcore} & \bfseries{App} & \bfseries{Lib}  & \bfseries{All}\\
+    \hline
+
+Runtime	&	1075	&	56	&	1446	&	374	&	2951\\
+Error	&	144	&	34	&	168	&	105	&		451	\\
+Checked	&	110	&	230	&	139	&	145	&	624	\\
+Throwable	&	0	&	0	&	1	&	0	&	1	\\
+Undefined	&	1	&	0	&	6	&	8	&	15	\\
+    \hline
+All		& 1330	&	320	&	1760	&	632	&	4042	\\
+    \hline
+  \end{tabular}
+\caption{Types of root exceptions.}
+  \label{tab:typeroottab}
+\end{table}
+
+%\begin{figure}
+%\centering
+%\includegraphics[width=\hsize]{chart_exceptiontypes.pdf}
+%\includegraphics[width=\hsize]{exception_types.pdf}
+%\caption{Distribution of Checked and Unchedked exceptions.}
+%\label{fig:typeroot}
+%\end{figure}
+
+% REMOVED BECAUSE WE DO NOT HAVE SPACE TO DISCUSS IT, AND IT IS NOT WELL
+% CONNECTED WITH PREVIOUS.
+%\subsection{Uncaught Checked Exceptions}
+%Although most of the  reported exceptions were runtime, we could also find checked
+%exceptions as the root causes of stack traces reported on issues. A checked exception
+%can remain uncaught in two circumstances: if all methods on the execution trace
+%on which this exception flows explicitly specifies the exception, or if the
+%checked exception is wrapped in a runtime and than re-thrown. Since checked exceptions 
+%should be used to represent recoverable conditions, ignoring a
+%checked exception, is considered a bad practice. As mentioned before, when the developer is confronted
+%with a checked exception, the designer of the API is telling him to handle the
+%exceptional condition. To better understand what was causing checked exceptions
+%to scape from being handled we investigated the kinds of wrappings that happened
+%on stacks the stacks.  Next sections presents the wrappings found and discuss
+%about them.
+
+\subsection{What can the Exception Wrappings tell us?}
+
 Table~\ref{tab:wrappingandroid} presents the wrappings found in this study for all
 stack traces found in Android repositories. As we can see, some of the checked
-exceptions where indeed wrapped in runtime exceptions or even errors, while most
+exceptions were indeed wrapped in runtime exceptions or even errors, while most
 of them were not wrapped along the stack trace.
 
-\begin{table}
-\centering
-\begin{tabular}{llll}
-    \hline
-    \bfseries{Wrapper Type} & \bfseries{Root Cause} & \bfseries{Occurrences} & \bfseries{Projects} \\
-    \hline
-      none   & Runtime   &  2360 & 359 \\
-      none  &   Checked  & 422  & 117 \\
-      none  &   Error   & 381  & 141 \\
-      none &  Undefined  &  15   & 10  \\
-      none & Throwable  &  1    & 1 \\
-     \hline
-      Runtime   & Runtime & 560  & 182 \\    
-      Checked   & Checked & 98   & 42  \\
-      Error     & Error   & 15   & 14  \\
-    \hline
-      Runtime &  Checked   & 114  & 70  \\
-      Runtime   &  Error   & 57   & 38  \\
-      Checked &  Runtime   & 22   & 12  \\
-      Checked & Error    &  8    & 7 \\
-      Error & Checked     &  7    & 4 \\
-      Error & Runtime     &  13   & 6 \\
-      Throwable & Runtime &  3    & 1 \\
-      Undefined & Runtime &  3    & 3 \\
-      Undefined & Error &  1    & 1 \\
+\begin{table*}
+  \centering
+  \begin{tabular}{rccccccc}
+    \hline
+    \bfseries{Wrapper Type}  &  \bfseries{Root Cause Type} &  \bfseries{\#Projects}  &  \bfseries{\#Occurrences} & \textsf{Android} & \textsf{Libcore} & \textsf{App} & \textsf{Lib}  \\
+    \hline
+
+%\begin{table}
+%\centering
+%\begin{tabular}{llll}
+%    \hline
+%    \bfseries{Wrapper Type} & \bfseries{Root Cause Type} & \bfseries{Occurrences} & \bfseries{Projects} \\
+%    \hline
+      %none   & Runtime    & 359 &  2360  & - &  - \\
+      %none  &   Checked   & 117 & 422 & - &  -   \\
+      %none  &   Error    & 141 & 381 & - &  -  \\
+      %none &  Undefined  & 10  &  15  & - &  -    \\
+      %none & Throwable   & 1  &  1 & - &  -   \\
+     %\hline
+      %Runtime   & Runtime  & 182 & 560 & - &  -  \\    
+      %Checked   & Checked   & 42 & 98  & - &  -  \\
+      %Error     & Error    & 14 & 15  & - &  -   \\
+    %\hline
+
+     Runtime &  Checked   & 70 & 114 & 0 &  0  & 0 &  0   \\
+      Runtime   &  Error   & 38  & 57 & 0 &  0  & 0 &  0       \\
+      Checked &  Runtime   & 12 & 22 & 0 &  0  & 0 &  0        \\
+      Checked & Error      & 7 &  8 & 0 &  0  & 0 &  0       \\
+      Error & Checked      & 4  &  7 & 0 &  0  & 0 &  0     \\
+      Error & Runtime     & 6  &  13   & 0 &  0  & 0 &  0    \
+    % Throwable & Runtime   & 1 &  3   & 0 &  0  & 0 &  0    \\
     \hline
@@
 \label{tab:wrappingandroid}
-\end{table}
+\end{table*}
 
@@
 
-Since there is no way of enforcing Java exception type conventions during program development,
-the exceptions regardless its types can se used interchangeably. This interchangeability makes the
-behaviour of the exception handling code more complex and less reliable. For
-instance, considering a situation were an instance of OutOfMemoryError (a situation that usually cannot
- be handled) was wrapped in a checked exception which tells the caller that such exception should
-handled.
-
-\section{Discussion}
+\subsection{What can the Signalers tell us?}
+
+After investigating the most popular exceptions on issues, we also investigated elements
+responsible for signaling most of the exceptions reported on issues.
+Such elements were directly involved on the reported crashes. We cannot
+directly blame such component for the crash since the client may not have been
+using the component according to its "contract", or an external event
+may have happened during the execution of such method.
+
+Regardless the reason, such analysis give as a clue to comon element involved on 
+reported crashes. Table X illustrates the most common root exception signalers
+on issues, toglether with the list of exceptions being thrown.
+
+%Present the top signalers found in this study, and top pairs (signalers and exceptions), and discuss 
+%about inversion of control and the difficult to adequately handle exceptions in this style of development.
+%Inversion of control stimulate the use of unchecked exceptions since the "call back" do not know which exception to expect.
+
+{\tiny
+\begin{verbatim}
+
+% SELECT mainsignaler_method,count(distinct(repo)) as n, string_agg(distinct(main_cause), ',')
+% FROM stacktracesummary group by (mainsignaler_method) order by n desc;
+
+GITHUB:
+"android.os.Parcel.readException";28;"Exception,java.lang.IllegalArgumentException,java.lang.IllegalStateException,java.lang.NullPointerException,java.lang.SecurityException"
+"dalvik.system.BaseDexClassLoader.findClass";19;"java.lang.ClassNotFoundException"
+---> "android.app.Instrumentation.checkStartActivityResult";17;"android.content.ActivityNotFoundException"
+"android.graphics.BitmapFactory.nativeDecodeStream";16;"java.lang.OutOfMemoryError"
+--->"android.view.WindowManagerImpl.findViewLocked";15;"java.lang.IllegalArgumentException"
+"android.graphics.Bitmap.nativeCreate";15;"java.lang.IllegalArgumentException,java.lang.OutOfMemoryError"
+"android.graphics.BitmapFactory.nativeDecodeAsset";14;"java.lang.OutOfMemoryError,java.lang.RuntimeException"
+"dalvik.system.PathClassLoader.findClass";13;"java.lang.ClassNotFoundException"
+---> "android.support.v4.app.FragmentManagerImpl.checkStateLoss";12;"java.lang.IllegalStateException"
+"android.os.BinderProxy.transact";12;"android.os.DeadObjectException,android.os.TransactionTooLargeException,java.lang.RuntimeException"
+"android.database.DatabaseUtils.readExceptionFromParcel";11;"android.database.sqlite.SQLiteDiskIOException,android.database.sqlite.SQLiteException,java.lang.IllegalArgumentException"
+"android.app.ActivityThread.performLaunchActivity";11;"java.lang.NullPointerException,java.lang.RuntimeException"
+--->"android.graphics.Canvas.throwIfRecycled";10;"java.lang.NullPointerException,java.lang.RuntimeException" (private method throwing a RuntimeException)
+--->"android.os.StrictModedAndroidBlockGuardPolicy.onNetwork";10;"android.os.NetworkOnMainThreadException" (public)
+---> "android.support.v4.app.Fragment.getResources";10;"java.lang.IllegalStateException"
+"android.os.AsyncTaskd3.done";10;"java.lang.RuntimeException"
+
+"The Canvas class holds the "draw" calls.
+To draw something, you need 4 basic components: A Bitmap to hold the pixels, a Canvas to host the draw calls (writing into the bitmap), a drawing primitive (e.g.
+Rect, Path, text, Bitmap), and a paint (to describe the colors and styles for the drawing)."
+
+"private static void More ...throwIfRecycled(Bitmap bitmap) {
+941         if (bitmap.isRecycled()) {
+942             throw new RuntimeException(
+943                         "Canvas: trying to use a recycled bitmap " + bitmap);
+944         }
+945     }"
+
+GOOGLECODE:
+"android.app.Instrumentation.checkStartActivityResult";8;"android.content.ActivityNotFoundException"
+"android.graphics.Bitmap.nativeCreate";8;"java.lang.IllegalArgumentException,java.lang.OutOfMemoryError,java.lang.RuntimeException"
+"java.lang.ClassLoader.loadClass";6;"java.lang.ClassNotFoundException,java.lang.IncompatibleClassChangeError,java.lang.UnsupportedClassVersionError"
+"dalvik.system.BaseDexClassLoader.findClass";6;"ClassNotFoundException,java.lang.ClassNotFoundException"
+"com.android.dx.dex.file.ClassDefsSection.add";6;"java.lang.IllegalArgumentException"
+"android.os.Parcel.readException";6;"Exception,java.lang.IllegalArgumentException,java.lang.IllegalStateException,java.lang.NullPointerException,java.lang.SecurityException"
+"android.view.WindowManagerImpl.findViewLocked";5;"java.lang.IllegalArgumentException"
+"android.os.BinderProxy.transact";5;"android.os.DeadObjectException,android.os.RemoteException,android.os.TransactionTooLargeException,java.lang.RuntimeException,java.lang.SecurityException"
+"android.app.ActivityThread.performLaunchActivity";5;"Error,java.lang.NullPointerException,java.lang.RuntimeException"
+"android.graphics.BitmapFactory.nativeDecodeStream";5;"java.lang.OutOfMemoryError"
+
+\end{verbatim}
+}
+
+\subsection{Undocumented runtime exceptions.} After filtering out all the exceptions implicitly 
+signaled by  JVM (due to programming mistakes) and inspecting the signaler methods of such exceptions (i.e.,  direct an indirect 
+instances of java.lang.RuntimeException). We could observe that only 1 method out of 118 inspected signalers
+ (i.e., 0,8\%) documented the explicitly thrown runtime exception in a Javadoc comment; and none of these methods
+included the runtime exception  (reported on the stack) as part of the exception interface of the method (i.e., using 
+throws clause on method signature). This result is aligned with the results of other  study conducted by from 
+Sacramento et al ~\cite{sacramento2006unchecked} which observed that the
+runtime exceptions in .Net programs are most often not documented.
+
+Such undocumented runtime exceptions represent a threat to system robustness, specially
+when such exceptions are thrown then the third party code is invoked inside the application (e.g. libraries, or framework utility code).
+In such cases the client usually do not have access to the source code, and in the absence of
+the exception documentation it is very difficult or even impossible for the client of such third party code to 
+protect the application against undocumented runtime exceptions. As a consequence, the
+ undocumented runtime exception may remain uncaught and lead to system crashes.
+
+\section{Discussions and Actionable Advices}
 
@@
 exceptional conditions.
+
+Discuss about the actionable advices propagated by TAO XIE.
 
@@
 
+\subsubsection{Ways of Preventing Odd Wrappings}
+Since there is no way of enforcing Java exception type conventions during program development,
+the exceptions regardless its types can se used interchangeably. This interchangeability makes the
+behaviour of the exception handling code more complex and less reliable. For
+instance, considering a situation were an instance of OutOfMemoryError (a situation that usually cannot
+ be handled) was wrapped in a checked exception which tells the caller that such exception should
+handled.
+
+\subsection{Checked and Unchecked Exceptions in JVM and Android and what they can tell us?}
+
+We could observe that some of the type-changing-wrappings were performed by JDK classes.
+Such finding lead us to think that the exception types have been used as intercheagable and may not being carrying 
+the meaning they were firstly designed to carry.
+
+The way the exceptions are being handle are more important that the type itself, it points to the fact already discussed 
+by other languages such as M\# blog [ref] and \"thinking in Java author\" and other blogs. It is all about the exception handling
+politics of a system and not the type itself. The exception type is not enought to ensure a more strict or robust development,
+since the checked exceptions may be silienced or converted to a checked exception. According to Java manual,
+they say that it can actually happen but it will be done intentionally and they advocate that it is a bad practice.
+
+Present the number of checke and unchecked exceptions in JVM, JDK, Android SO, and Android SDK.
+Discuss about the amount of checked and unchecked defined in each, and discuss about interesting exceptions that
+have the same name ex: classNotFound...
+
 %\noindent\emph{The high number of null pointer exceptions}
@@
 \subsection{The high number of null pointer exceptions}
-Last but not least, the high number of null pointer exceptions  lead us to think on other 
-pertinent research direction: (i) lightweight intra-method nullpointer analysissuch as the @Nullable;
+
+The null references was firstly introdduced by Tony Hoare in ALGOL W, which after some years he called his “one-billion-dollar mistake”:
+
+"I call it my billion-dollar mistake. It was the invention of the null reference in 1965 [...] I couldn’t resist the temptation to put 
+in a null reference, simply because it was so easy to implement. This has led to innumerable errors, vulnerabilities, and system 
+crashes, which have probably caused a billion dollars of pain and damage in the last forty year."
+
+ref: http://bertrandmeyer.com/category/computer-science/
+
+The high number of null pointer exceptions found in this mining study shed light into a problem
+that although known was not supported by large scale mining studies on the large.
+
+This observation emphasises the importance of solutions to avoid NullPointerExceptions, such as:
+(i) lightweight intra-method nullpointer analysis supported by Java8 @Nullable annotations;
 (ii) intermethod nullpointer analysis tools such as the one proposed by [DANIELE-RELATED-WORK];
@@
 or computations that may fail) could improve the robustness of Java programs.
-@NULABLE ANNOTATIONS.
+
+The @Nullable annotations are used by analyzers like the Checker Framework, FindBugs, Eclipse, 
+NetBeans, IntelliJ, and other commercial tools which run at compile time and detect potential null 
+pointer derecefences. Android Studio 0.5.5 Released support these annotations.
+
+(1) mostrar a utilizacao de @nullable em projetos android. rodar em alguns e mostrar.
+Android Project Unwritten Rules (http://blog.lemberg.co.uk/android-project-unwritten-rules)
+
+(2) dar foco em android
+
+(3) falar do modelo do ciclo de vida da aplicacao e da necessidade de usar excecoes Runtime.
+
+(4) Local onde tratar excecao - handler padrao - politica da aplicacao. tratamento difucultado multiplas threads e assyncronos.
 
@@
 It allows the developers to benefit from the faults happening on different systems.
+
+%\subsection{Actionable Advices}
+%The results of a mining study should be a set of actionalble advices (tao xie).
 
@@
 
-\textit{Analysis of Java Exception Usage.} As the manual analysis of the Java
-exception flow can easily become infeasible, Robillard and Murphy~\cite{Robil00}
-employed dataflow analysis to find the propagation paths of checked and
-unchecked exception types. Modeled after Robillard and Murphy's work, other
-tools have been proposed to support the static analysis of exception
-flows~\cite{coelho2008assessing}. The main limitation of all
-static analysis tools is the number of false positives inherent to static analysis
-solutions, which can lead to a high number of exception flows, specially if
-considering Java Environment exceptions and exceptions signaled from libraries.
-Additionally, Cabral and Marques~\cite{cabral2007exception} analyzed the
-exception handling code of 32 open-source systems, both for Java and .NET. They
-observed that the action handlers were very simple (e.g., logging and present a
-message to the user). Reimer and Srinivasan~\cite{reimer2003analyzing} listed a
-set of bad practices on exception handling that hinder software maintainability
-and robustness, based on their own experience with Java enterprise applications.
-Our work differs from those two works as we tried to identify bad practices from
-the combined use of stack traces extracted from issues and bytecode and source
-code analysis.
+\textit{Empirical studies using Android apps.} Ruiz et al.~\cite{Ruiz12}
+investigated the degree of reuse across applications in Android Market, the
+study showed that almost 23\% of the classes inherited from a base class in the
+Android API, and that 217 mobile apps were reused completely by another mobile
+app. Pathak et al.~\cite{Patha11} analyzed bug reports and developers
+discussions of Android platform and found out that approximately 20\% of
+energy-related bugs in Android occurred after an OS update. McDonnell et
+al.~\cite{McDon13} conducted a case study of the co-evolution behavior of
+Android API and 10 dependent applications using the version history data found
+in github. The study found that approximately 25\% of all methods in the client
+code used the Android API, and that the methods reusing fast-evolving APIs were
+more defect prone then others. Vásquez et al.~\cite{Linar13} analyzed
+approximately 7K free Android apps and observed that the last successful apps
+used Android APIs that were on average 300\% more change-prone than the APIs
+used by the most successful apps. Pingyu and Elbaum~\cite{Zhang12} analyzed bug
+reports of 5 Android applications an observed that 29\% had to do with poor
+exceptional handling code. Our work differs from the others as it aims at
+distilling the information of bug reports describing uncaught exceptions created
+for mobile applications in Github, in order to get a first view
+of what is causing the crashes across applications available and what the
+characteristics of the stack traces can tell us about the exception structure of
+those applications.
+
+\textit{Extracting Stack Traces from natural language artifacts.} 
+%Currently, many
+%software vendors embed automatic crash reporting tools in their software
+%systems. Moreover, third party crash collection services exist, most of them 
+%targeted for applications run on mobile phones~\cite{BugSe14,BugSn14,Googl14,Acra14}.
+Apart from bug reports, stack traces can be embedded in other forms of
+communication between developers, such as discussion logs and emails.
+Being intermixed with text makes the accurate extraction of stacktraces 
+an involved process.
+Infozilla~\cite{bettenburg2008extracting}
+is based on a set of regular expressions that extract a set of frames
+related to a stack trace. The main limitation of this solution is that it is not
+able to extract stack traces embedded on verbose log files (i.e., on which we
+can find log text mixed with exception frames). Bacchelli
+et al.~\cite{bacchelli2012content} propose a solution to recognize stack trace frames
+from development emails and relate it to code artifacts (i.e. classes) mentioned
+on the stack trace. In addition to those tools, ExceptionMiner is able to 
+both extract stack traces from natural language artifacts and to 
+classify them in a set of predefined categories.
+(MENCIONAR QUE X% DOAS EXCECOES ANALISADAS ESTAVAM EMBEBIDAS EM ALGUM TIPO DE ARQUIVO DE LOG - 
+LOGCAT DO ANDROID....)
 
@@
 
-\textit{Extracting Stack Traces from natural language artifacts.} 
-%Currently, many
-%software vendors embed automatic crash reporting tools in their software
-%systems. Moreover, third party crash collection services exist, most of them 
-%targeted for applications run on mobile phones~\cite{BugSe14,BugSn14,Googl14,Acra14}.
-Apart from bug reports, stack traces can be embedded in other forms of
-communication between developers, such as discussion logs and emails.
-Being intermixed with text makes the accurate extraction of stacktraces 
-an involved process.
-Infozilla~\cite{bettenburg2008extracting}
-is based on a set of regular expressions that extract a set of frames
-related to a stack trace. The main limitation of this solution is that it is not
-able to extract stack traces embedded on verbose log files (i.e., on which we
-can find log text mixed with exception frames). Bacchelli
-et al.~\cite{bacchelli2012content} propose a solution to recognize stack trace frames
-from development emails and relate it to code artifacts (i.e. classes) mentioned
-on the stack trace. In addition to those tools, ExceptionMiner is able to 
-both extract stack traces from natural language artifacts and to 
-classify them in a set of predefined categories.
-
-%\textit{Empirical studies using Android apps.} Ruiz et al.~\cite{Ruiz12}
-%investigated the degree of reuse across applications in Android Market, the
-%study showed that almost 23\% of the classes inherited from a base class in the
-%Android API, and that 217 mobile apps were reused completely by another mobile
-%app. Pathak et al.~\cite{Patha11} analyzed bug reports and developers
-%discussions of Android platform and found out that approximately 20\% of
-%energy-related bugs in Android occurred after an OS update. McDonnell et
-%al.~\cite{McDon13} conducted a case study of the co-evolution behavior of
-%Android API and 10 dependent applications using the version history data found
-%in github. The study found that approximately 25\% of all methods in the client
-%code used the Android API, and that the methods reusing fast-evolving APIs were
-%more defect prone then others. Vásquez et al.~\cite{Linar13} analyzed
-%approximately 7K free Android apps and observed that the last successful apps
-%used Android APIs that were on average 300\% more change-prone than the APIs
-%used by the most successful apps. Pingyu and Elbaum~\cite{Zhang12} analyzed bug
-%reports of 5 Android applications an observed that 29\% had to do with poor
-%exceptional handling code. Our work differs from the others as it aims at
-%distilling the information of bug reports describing uncaught exceptions created
-%for mobile applications in Github, in order to get a first view
-%of what is causing the crashes across applications available and what the
-%characteristics of the stack traces can tell us about the exception structure of
-%those applications.
+\textit{Tools Support for Detecting Faults on Java Exception Handling Code.}
+
+Manually analyzing the exception handling code looking for defects, can easily 
+become infeasible [ref]. Hence, some tools have been proposed to detect faults 
+on the exception handling code either statically (cabral, roberta,nelio,chang) 
+or dynamicaly through testing (roberta). Both approaches, however, have inherent
+ limitations. On the one hand, the static analysis tools needs to deal with the false 
+positives and cannot adequately deal with dynamically loaded code (e.g. reflection  
+[ref]). On the other hand the testing tools are llimited to the fact that the developer
+ needs to create a test to exercise every exception, which is also costly.... (olhar tese). 
+Moreover, both methods fail to account for exceptions implicitly signaled by the runtime 
+environment (e.g. due to resource depletion which can be signaled in every statement invocation).
+
+Robillard and Murphy~\cite{Robil00}
+employed dataflow analysis to find the propagation paths of checked and
+unchecked exception types. Modeled after Robillard and Murphy's work, other
+tools have been proposed to support the static analysis of exception
+flows~\cite{coelho2008assessing}. The main limitation of all
+static analysis tools is the number of false positives inherent to static analysis
+solutions, which can lead to a high number of exception flows, specially if
+considering Java Environment exceptions and exceptions signaled from libraries.
+
+Additionally, Cabral and Marques~\cite{cabral2007exception} analyzed the
+exception handling code of 32 open-source systems, both for Java and .NET. They
+observed that the action handlers were very simple (e.g., logging and present a
+message to the user). Reimer and Srinivasan~\cite{reimer2003analyzing} listed a
+set of bad practices on exception handling that hinder software maintainability
+and robustness, based on their own experience with Java enterprise applications.
+Our work differs from those two works as we tried to identify bad practices from
+the analysis of stack traces extracted from issues in combination with and bytecode and source
+code analysis.
 
@@
 
+\subsection{Exception Handling in Programming Languages}
+
+Techniques for error detection and handling are not  an optional add-on but a
+fundamental part of any system~\cite{bruntink2006discovering}. 
+The exception mechanism~\cite{goodenough1975exception} was proposed as a way of
+preventing errors from being ignored and was embedded in many mainstream
+programming languages ~\cite{garcia2001comparative} since then.
+
+In languages such as Modula-3, Guide, and Extended Ada the exceptions are 
+checked, which means that the compiler statically checks if appropriate handlers 
+are provided within the system. Accordingly, the programmer must explicitly specify 
+every exception that can be signaled to support the compile time checking. This approach is known to be
+costly to maintain (Dooren and Steegmans, 2005) since exception updates are
+cascading; if a new exception type is added to a method signature then all other
+methods directly using this method must either change their signatures to
+include the new exception type or handle the new exception.
+
+In languages such as Lore, C++, and Arche all exceptions are unchecked but the developer can 
+optionally list exception types on method signatures. Although exceptions can 
+be documented on methods signatures, the developer is not warned by the compiler
+if an unchecked exception is not handled inside the application. According to ~\cite{Robil00} 
+such an approach may hinder robustness because the client of a method cannot easily know
+which unchecked exceptions may be thrown (unless the code of the method and the
+methods called from it is inspected) which can be a very time consuming or
+infeasible task.
+
+Java [ref: Gosling et al. 1996] and CLU [Liskov and Snyder 1979] support the checked exceptions,
+ (therefore the compiler can then check if handlers are provided to such exceptions), however, this support is 
+only partial because both languages also provides a form of unchecked exceptions [ref: Robilard].
+In CLU,  if a handler is not found for a given exception, it is automaticaly converted in a predened default exception, 
+called failure which is the only one which can be implicitly propagated [ref-clue].
+ In Java the developer can define specific unchecked and checked exceptions and use both kinds in a single system.
+
+There is a long-lasting debate about the pros and cons of checked and unchecked exceptions 
+~\cite{javatut,stackoverlow,debate},\footnote{152 questions in Stackoverlow
+are related to this debate}. However, the debates are mainly based on the developers own experience in
+using both models, since there is still few empirical evidence of the pros and cons of each model.
+
+%Although exception handling mechanism has been embedded in several programming languages 
+%as show above,  the exception handling code is often poorly understood and the least
+%tested part of software systems [ref:POORLY].
+
+%[REF]: Static Analysis to Support the Evolution of Exception Structure in Object-Oriented Systems
+
+% This behaviour may lead to the problem of uncaught exceptions~\cite{jo2004uncaught}, one of the main causes of application crashes
+%(e.g.~\cite{miller1997issues,Robil00,shah2010understanding, garcia2007extracting, garcia2001comparative,cabral2007exception,coelho2011unveiling}).
+
 \section*{Acknowledgment} This work is partially supported by: CNPq -- Proc.

### 2014-08-23 android-stacks.tex
changing the focus. More inprovements in sec 2

@@
 
-\title{ What can GitHub and Googlecode Issues tell us about Exceptions in Android Apps? }
+\title{ What can GitHub and Googlecode Issues tell us about Exceptions in Android? }
 
@@
 
-\subsection{Java Exception Handling Mechanism} \label{sec:extypes}
+\subsection{The Android Platform} \label{sec:extypes}
+
+\begin{figure} \centering \includegraphics[scale=0.8]{arch3.png}
+  \caption{Android Archtecture} \label{fig:exchier} \end{figure}
+
+- opensource
+- elements of archtecture.
+- Inversion of control development model.
+- Activity onStart onResume onStop...
+- Acess and share resources...
+
+\subsection{Exceptions in Java} \label{sec:extypes}
 
@@
 instance of the Throwable class, and can be of three kinds: the checked exceptions
-(extends Exception) - means that the compiler statically checks if appropriate handlers 
-are provided within the system, to enable such checking each checked exception must be specified 
-on the method signature;  the run-time exceptions (extends RuntimeException) and errors
-(extends Error)~\cite{gosling2000java}  -  which do not need to be specified on method
-signatures, and hence both are referred as unchecked exceptions.
+(extends Exception), the runtime exceptions (extends RuntimeException) and errors
+(extends Error)~\cite{gosling2000java}.
+
+The checked exceptions must be specified on the method signature that may throw it, 
+and the compiler statically checks if appropriate handlers are provided within the system.
+Both runtime exceptions and errors are known as unchecked exceptions, because 
+they do not need to be specified on method signatures and do not trigger any 
+compile time checking.
+
+By convention an Error represents an unrecoverable condition which usually results
+from failures detected by the Java Virtual Machine, such as OutOfMemoryError and
+normally should not be handled inside the application. The user-defined exceptions 
+can be either checked or runtime. Although the Java specification~\cite{gosling2000java} 
+suggests that user-defined exceptions should be checked exception (because doing so 
+the callers of a method will know about the exceptions that a method can throw and so 
+they can decide what to do about them) there is a long lasting debate about the use of 
+checked and unchecked exceptions in Java.
 
@@
 
-The Java Virtual Machine represents as runtime exceptions invalid operations detected in
-the program (out-of-bounds array index, divide-by-zero error, null pointer
-references) from which most of the programs are not expected to recover. By
-convention an Error represents an unrecoverable condition which usually results
-from failures detected by the Java Virtual Machine, such as OutOfMemoryError and
-normally should not be handled inside the application. The Java specification~\cite{gosling2000java} 
-suggests that user-defined exceptions should be checked exceptions. The main reason is because doing so the
-callers of a method will know about the exceptions that a method can throw, and
-so they can decide what to do about them.
-
-Hence, in Java an exception can be thrown in one of the following
-circumstances~\cite{gosling2000java}: 
-(i) explicitly thrown when a throw statement is executed; 
-(ii) implicitly thrown by the JVM when the evaluation of an expression
- violates the normal semantics of language, also refered as a coding error
- (e.g., out-of-bounds array index, division-by-zero, access to a null reference); or 
-(iii) implicitly thrown by the JVM due to an internal error or resource limitation (e.g.,
-OutofMemoryError).
-
+The runtime exceptions, however are not only used to represent user-defined
+exceptional conditions, they are used by the JVM when a program violates 
+the semantic constraints of the Java programming language (out-of-bounds array index, divide-by-zero 
+error, null pointer references) from which most of the programs are not expected to recover. 
+\footnote{Some programming languages react to such errors by peremptorily terminating the program; 
+other programming languages allow similar approaches e.g.C\#; and others let the program to continue
+ its execution in some situations such as the out-of-bounds array index e.g., C++. }
+
+%In Java an exception can be thrown in one of the following
+%circumstances~\cite{gosling2000java}: 
+%(i) explicitly thrown when a throw statement is executed; 
+%(ii) implicitly thrown by the JVM when the evaluation of an expression
+% violates the normal semantics of language, also refered as a coding error
+% (e.g., out-of-bounds array index, division-by-zero, access to a null reference); or 
+%(iii) implicitly thrown by the JVM due to an internal error or resource limitation (e.g.,
+%OutofMemoryError). 
 %\todo{This sub-section is quite similar to the section introduction, shouldn't
@@
 
-When an exception is throw, it causes a transfer of control from the point
+When an exception is thrown, it causes a transfer of control from the point
 where the exception occurred to a point that can be specified by the programmer
@@
 propagating exceptions in Java programs is through exception wrapping
- (also known as exception chaining). Exception wrapping allows one exception to be wrapped in another one and re-thrown. 
-Figure~\ref{fig:wrapping} presents a stack trace associated which illustrates an
-exception wrapping. The bottom part of the stack trace is the \emph{root cause}, which indicates the
+ (also known as exception chaining). Exception wrapping allows one exception 
+to be wrapped in another one and re-thrown. Figure~\ref{fig:wrapping} presents 
+a stack trace associated which illustrates an exception wrapping. The bottom 
+part of the stack trace is the \emph{root cause}, which indicates the
 first reason for the error to be thrown (in this case, the computer run out of
@@
 
-\begin{figure} \centering \includegraphics[scale=0.]{stacktrace.png}
+\begin{figure} \centering \includegraphics[scale=0.8]{stacktrace.png}
 \caption{Example of an Exception stack trace in Java.}
@@
 
-\subsection{The Android Platform} \label{sec:extypes}
-
-\begin{figure} \centering \includegraphics[scale=0.8]{arch3.png}
-  \caption{Android Archtecture} \label{fig:exchier} \end{figure}
+\subsection{Related Best Practices}
+\label{sec:best}
+
+Several general guidelines have been proposed on how to use Java
+exceptions~\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,
+bloch2008effective}. 
+We compiled a list of the guidelines related to exception types and exception propagation in Java, summarized below:\footnote{We also compiled guidelines related
+to exception handling but they are out of the scope of this paper.}
+
+%\noindent\emph{Meaning of Exception Types}
+
+\textbf{I-Checked exceptions should be used to represent recoverable
+conditions} (\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,bloch2008effective})
+The developer should use checked exceptions for conditions from which the caller
+is expected to recover. By confronting the API user with a checked exception,
+the API designer is forcing the client to handle the exceptional condition. The
+client can explicitly ignore the exception (swallowing, or converting it to
+other type) at the expense of the program's robustness~\cite{gosling2000java}.
+
+\textbf{II-Error represents an unrecoverable condition detected by the JVM which
+should not he handled} (\cite{gosling2000java}). It results from failures detected
+by the Java Virtual Machine which indicate resource deficiencies, invariant
+failures or other conditions that make impossible the program to recover.
+
+%\noindent\emph{Exception Throwing}
+
+\textbf{III-A method should throw exceptions that precisely define the
+exceptional condition} (\cite{gosling2000java,bloch2008effective}). To do so,
+developers should either try to reuse the exception types already defined in the
+JVM or they should create a specific exception. Throwing general types such as a
+pure java.lang.Exception or a java.lang.RuntimeException is considered a bad practice.
+
+%\textbf{IV-Do not repeatedly re-throw exceptions, handle exceptions as close as
+%possible to the problem}~\cite{wirfs2006toward}. 
+%Programs that frequently throw exceptions lose in performance ~\cite{wirfs2006toward,gosling2000java}.
+%Moreover, 
+%Close to the place that the exception was throw the caller usually has
+%enough contextual information to perform  corrective actions. When the exception
+%is propagated far away from the method where it was signaled (i.e. further away
+%#from the problem) it can be difficult to take meaningful recovery actions.
+%*~\todo{RSC:this practice was not well explored, can be removed if we need space.}
+
+%\noindent\emph{Exception Documentation}
+%\textbf{IV- It is wise to document all exceptions thrown by a method}
+
+\textbf{IV- All exceptions explicitly thrown by a method should be documented.}
+(\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,bloch2008effective}).
+The exceptions thrown by a method are an important part of methods interface,
+and is required to use the method properly. The checked exceptions are already
+part of the  methods signature, and the method caller is aware of the checked
+exceptions being thrown by it. According to~\cite{bloch2008effective}, it is
+also wise to document the explicitly thrown runtime exceptions\footnote{This
+excludes the implicitly signaled JVM runtime exceptions related to programming
+mistakes} as carefully as checked exceptions. Doing so, the clients of a method
+will be aware of all the exceptions the method can throw. If the developer fails to
+do follow this practice (specially when developing library code) it will be
+difficult or even impossible for the caller to make effective use of such 
+method~\cite{wirfs2006toward, bloch2008effective} and protect its code
+against unforeseen exceptions.
+
+These best practices guided our exploratory study described next.
+Based on information extracted from stack traces (complemented by
+bytecode and source-code analysis), we investigate whether stack characteristics
+can reveal whether these practices have been obeyed. 
+%Next section discusses the
+%study procedure and the practices' misuses that emerged from this study.
 
@@
 
-\subsection{Meaning of Exception Types and related Best Practices}
-\label{sec:best}
-
-Several general guidelines have been proposed on how to use Java
-exceptions~\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,
-bloch2008effective}. 
-We compiled a list of the guidelines related to exception types and exception
-propagation in Java the following:\footnote{We also compiled guidelines related
-to exception handling but they are out of the scope of this paper.}
-
-%\noindent\emph{Meaning of Exception Types}
-
-\textbf{I-Use checked exceptions to represent recoverable
-conditions} (\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,bloch2008effective})
-The developer should use checked exceptions for conditions from which the caller
-is expected to recover. By confronting the API user with a checked exception,
-the API designer is forcing the client to handle the exceptional condition. The
-client can explicitly ignore the exception (swallowing, or converting it to
-other type) at the expense of the program's robustness~\cite{gosling2000java}.
-
-\textbf{II-Error represents an unrecoverable condition detected by the JVM which
-should not he handled} (\cite{gosling2000java}). It results from failures detected
-by the Java Virtual Machine which indicate resource deficiencies, invariant
-failures or other conditions that make impossible the program to recover.
-
-%\noindent\emph{Exception Throwing}
-
-\textbf{III-A method should throw exceptions that precisely define the
-exceptional condition} (\cite{gosling2000java,bloch2008effective}). To do so,
-developers should either try to reuse the exception types already defined in the
-JVM or they should create a specific exception. Throwing general types such as a
-pure Exception or a RuntimeException is hence considered a bad practice.
-
-%\textbf{IV-Do not repeatedly re-throw exceptions, handle exceptions as close as
-%possible to the problem}~\cite{wirfs2006toward}. 
-%Programs that frequently throw exceptions lose in performance ~\cite{wirfs2006toward,gosling2000java}.
-%Moreover, 
-%Close to the place that the exception was throw the caller usually has
-%enough contextual information to perform  corrective actions. When the exception
-%is propagated far away from the method where it was signaled (i.e. further away
-%#from the problem) it can be difficult to take meaningful recovery actions.
-%*~\todo{RSC:this practice was not well explored, can be removed if we need space.}
-
-%\noindent\emph{Exception Documentation}
-\textbf{ IV-Document all exceptions thrown by a
-method} (\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,bloch2008effective}).
-The exceptions thrown by a method are an important part of methods interface,
-and is required to use the method properly. The checked exceptions are already
-part of the  methods signature, and the method caller is aware of the checked
-exceptions being thrown by it. According to~\cite{bloch2008effective}, it is
-also wise to document the explicitly thrown runtime exceptions\footnote{This
-excludes the implicitly signaled JVM runtime exceptions related to programming
-mistakes} as carefully as checked exceptions. Doing so, the clients of a method
-will be aware of all the exceptions the method can throw. If the developer fails to
-do follow this practice (specially when developing library code) it will be
-difficult or even impossible for the caller to make effective use of such 
-method~\cite{wirfs2006toward, bloch2008effective}. As mentioned in Section II, 
-in some languages such as C
-
-These best practices guided our exploratory study described next.
-Based on information extracted from stack traces (complemented by
-bytecode and source-code analysis), we investigate whether stack characteristics
-can reveal whether these practices have been obeyed. 
-%Next section discusses the
-%study procedure and the practices' misuses that emerged from this study.
-
 \section{RQ1: What are the common characteristics of such stack traces?}
@@
 to ways of how to deal with it and mitigate it.
-
-\subsection{Investigating The Android Subset}
-
-----------
 
@@
   \centering
-  \begin{tabular}{rccccccc}
-    \hline
-    \bfseries{Wrapper Type}  &  \bfseries{Root Cause Type} &  \bfseries{\#Projects}  &  \bfseries{\#Occurrences} & \textsf{Android} & \textsf{Libcore} & \textsf{App} & \textsf{Lib}  \\
-    \hline
-
-%\begin{table}
-%\centering
-%\begin{tabular}{llll}
-%    \hline
-%    \bfseries{Wrapper Type} & \bfseries{Root Cause Type} & \bfseries{Occurrences} & \bfseries{Projects} \\
-%    \hline
-      %none   & Runtime    & 359 &  2360  & - &  - \\
-      %none  &   Checked   & 117 & 422 & - &  -   \\
-      %none  &   Error    & 141 & 381 & - &  -  \\
-      %none &  Undefined  & 10  &  15  & - &  -    \\
-      %none & Throwable   & 1  &  1 & - &  -   \\
-     %\hline
-      %Runtime   & Runtime  & 182 & 560 & - &  -  \\    
-      %Checked   & Checked   & 42 & 98  & - &  -  \\
-      %Error     & Error    & 14 & 15  & - &  -   \\
-    %\hline
-
+  \begin{tabular}{llcccccc}
+    \hline
+    \bfseries{Wrapper Type}  &  \bfseries{Root Cause Type} &  \bfseries{Projects}  &  \bfseries{Occurrences} & \textsf{Android} & \textsf{Libcore} & \textsf{App} & \textsf{Lib}  \\
+    \hline
       Runtime &  Checked   & 70 & 114 & 0 &  0  & 0 &  0 \\
@@
       Error & Checked      & 4  &  7 & 0 &  0  & 0 &  0     \\
-      Error & Runtime     & 6  &  13   & 0 &  0  & 0 &  0    \
-    % Throwable & Runtime   & 1 &  3   & 0 &  0  & 0 &  0    \\
-    \hline
-  \end{tabular}
-\caption{Kinds of wrappings found on the stacks of 482 ANDROID projects found on Github}
+      Error & Runtime     & 6  &  13   & 0 &  0  & 0 &  0    \\
+     Throwable & Runtime   & 1 &  3   & 0 &  0  & 0 &  0    \\
+    \hline
+  \end{tabular}
+\caption{Wrappings comprising different exception types.}
 \label{tab:wrappingandroid}
@@
 
-%\subsection{Actionable Advices}
+\subsection{Actionable Advices - The Unwrittern rules of Android development}
+
 %The results of a mining study should be a set of actionalble advices (tao xie).

### 2014-08-24 android-stacks.tex
changing the focus. More inprovements in sec 2

@@
 
+Android is an open source platform for mobile devices based on Linux kernel.
+The plaftom is also comprised of a core Java library, a set of libraries third-party libraries e.g.  SQLite,
+the Dalvik runtime environment, and the framework which provides higher-level services to the applications
+running on the platform e.g., controlling activities or providing data access.
+
+Android supports an event-driven development model. The main element responsible for 
+receiving the GUI events is the Activity, it is the unity of user interaction and execution in
+Android. Android also provides APIs for data storage, communication between Activities, 
+and other services. Moreover, it also provide frameworks for concurrent developement (e.g. AsyncTask) 
+which is higly encouraged in Android development due to plaftorm restrictions [shimidit and arie].
+\footnote{Other elements compose an application such as the Intents, resources
+but are out of the scope for this paper.}.
+
 \begin{figure} \centering \includegraphics[scale=0.8]{arch3.png}
   \caption{Android Archtecture} \label{fig:exchier} \end{figure}
-
-- opensource
-- elements of archtecture.
-- Inversion of control development model.
-- Activity onStart onResume onStop...
-- Acess and share resources...
 

### 2014-08-24 android-stacks.tex
changing the focus. More inprovements in sec 2

--- 2014-08-23 android-stacks.tex
@@
 
+Android is an open source platform for mobile devices based on Linux kernel.
+The plaftom is also comprised of a core Java library, a set of libraries third-party libraries e.g.  SQLite,
+the Dalvik runtime environment, and the Aplication Framework which provides higher-level services to the applications
+running on the platform e.g., controlling activities or providing data access.
+
+Android supports an event-driven development model. The main element responsible for 
+receiving the GUI events is called Activity, which is the unity of user interaction and execution in
+Android. Android also provides APIs for data storage, communication between Activities, 
+and other services. Moreover, it also provides frameworks for concurrent developement (e.g. AsyncTask) 
+which is highly needed due to the demand of responsiveness in such environment  - 
+mobile apps can easily become unresponsivedue to resurce limitation and excessive
+ network access [shimidit and arie].
+
 \begin{figure} \centering \includegraphics[scale=0.8]{arch3.png}
   \caption{Android Archtecture} \label{fig:exchier} \end{figure}
-
-- opensource
-- elements of archtecture.
-- Inversion of control development model.
-- Activity onStart onResume onStop...
-- Acess and share resources...
 

### 2014-08-25 android-stacks.tex
restructuring section 3

@@
 
-\textbf{I-Checked exceptions should be used to represent recoverable
+%\textbf{I
+\emph{I-Checked exceptions should be used to represent recoverable
 conditions} (\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,bloch2008effective})
@@
 
-\textbf{II-Error represents an unrecoverable condition detected by the JVM which
+\emph{II-Error represents an unrecoverable condition detected by the JVM which
 should not he handled} (\cite{gosling2000java}). It results from failures detected
@@
 
-\textbf{III-A method should throw exceptions that precisely define the
+\emph{III-A method should throw exceptions that precisely define the
 exceptional condition} (\cite{gosling2000java,bloch2008effective}). To do so,
@@
 
-\textbf{IV- All exceptions explicitly thrown by a method should be documented.}
+\emph{IV- All exceptions explicitly thrown by a method should be documented.}
 (\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,bloch2008effective}).
@@
 
-\section{Study Design}
-
-This work describes an exploratory study which was based on a sequential
-mixed-methods approach~\cite{ivankova2006using} for collecting, analyzing, and
-integrating both quantitative and qualitative evidence along the stages of the
-research process.
-
-Based mainly on the information available on exception stack traces embedded 
-on Github issues, the goal of this study was to gain a better understanding 
-of exception-related system crashes.
+\section{Empirical Study Design}
+
+The goal of this study was to extract nformation available on exception stack 
+traces embedded on issues reported on Android projects, looking for the common characteristics of
+those issues and investigating whether they could reveal common defficiencies
+on the exception handling code of both the applications and the underlaying 
+ framework.
+
+We choose Android apps because it is the only top platform that is open-source,
+and due to its increasing popularity it xxxx.
+
+Our study focused on free apps, since the information needed to perform our study
+cannot be retrieved from commercial apps because their issue reports and their
+source code is not publicly available. The Android open-source apps were also 
+target of other research studies [fse2013][ahimed].
+
+\subsection{Research Questions}
+In the context of our study we formulated the following reseach questions:
+
+\begin{description}
+
+  \item[RQ1] \noindent\emph{What are the common characteristics of exception stack traces reported on issues?} What are the main causes of crashes reported on issues? (more explicit?)
+
+  \item[RQ2] What patterns of exception (mis)-use related to exception types and exception propagation emerge from the exception stack trace analysis?
+
+\end{description}
+
+%Based mainly on the information available on exception stack traces embedded 
+%on issues, the goal of this study was to gain a better understanding 
+%of exception-related system crashes.
 
@@
 
- More specifically, we aimed at answering
-the following research questions:
-
-\begin{description}
-
-  \item[RQ1] What are the common characteristics of exception stack traces reported on issues?/What are the main causes of crashes reported on issues? (more explicit?)
-
-  \item[RQ2] What patterns of exception (mis)-use related to exception types and exception propagation emerge from the exception stack trace analysis?
-
-\end{description}
-
 For RQ1 and RQ2, we explore the domain quantitatively and highlight interesting cases by 
@@
 
-\subsection{Github Data}
+\subsection{Data Extraction Process}
+\label{sec:miningproc}
+
+The data used in our study to answer our research questions are: the issues related to each 
+Android project available in each repository; the source code and manifest file of each application;
+the bytecode of all exceptions defined and reused by Android platform.
+
+\subsubsection{Android Apps in GitHub}
 \label{sec:git}
@@
 
-\subsubsection{GHTorrent Dataset} This study used the dataset provided by the GHTorrent project~\cite{Gousi13}, an off-line mirror of the data 
+RESUMIR:
+This study used the dataset provided by the GHTorrent project~\cite{Gousi13}, an off-line mirror of the data 
 offered through the Github API.  GHTorrent has been collecting data from Git since 
@@
 
-\subsubsection{Google Code}
-
-Until January 2014 there were 788 android projects available in Google code repository from which 724 defined at least 1 issue and 183 projects 
-had at least one issue including a stack trace. To identify the Android projects on Googlecode we used the search facilities provided looking for 
-projects mentioning "Android" on the name of the project or its description. The repository page of each project was inspected, and from this set
-142 corresponded to Android applications.
-
-To mine the information available on such Google code projects we implemented a web crawler which navegated on each project and issue page
-and extracted the content of each issue to store on relational database.
-\footnote{Google code previoulsly provided a webservice to extract information about each project, but this service was removed in what Google called "clean-up action" which happened in June 2013.}
-
-\subsubsection{Exception Stack Traces on Issues}
-
-Before investigating the research questions listed above we analyzed the prevalence of stack traces 
-on issues created on Github. Prevalence or prevalence proportion, a term often used in epidemiology,
-represents the proportion of a population found to have a particular condition
-(e.g. a disease) over the total number of the studied population. In our work, we used
-this term to evaluate a condition (i.e., presence of a stack trace) over issues
-reported on GitHub repositories.
-
-Using the ExceptionMiner tool, developed in this work, we could extract the exception stack traces defined on issues and issue comments.
-Executing ExceptionMiner tool each issue and issue comment defined on the 16,836 projects considered in this study, we could observe that 3,758 projects (22,32\%) contained at least 1 issue on which a stack trace was found.
-Overall the number of issues on which a stack trace was found was 21,013 from which 28,800 stack traces were extracted and analyzed - some issues defined more than one stack trace.
-
-We then analyzed the lables associated to such issues, and we discovered that only 5,196 were labled as defect ~\footnote{the dedect lables considered in this work were every lable containing as substring: defect, bug, crash, failure, fail, exception} (which represent only 24,7\% of the issues on which a stack trace was found), most of the issues on which stack traces were found contained no lable.
-Such observation supports our assumption to consider not only the issues related with defect issues on this analysis.
-
-\subsubsection{Distiling Exception Stack Trace Information}
-
-\- root exception
-
-\- signaler
-\- method
-\- class
-\- package
-
-\- Exception
-
-\subsubsection{Android Apps Selection}
-\label{sec:android}
-
- % SELECT id, exception, exception_extends, exception_from, exception_package, 
- %      issueid_onexminer, main_cause, main_extends, main_from, main_package, 
- %      main_pre_signaler_java, main_signaler, mainsignaler_class, mainsignaler_from, 
- %      mainsignaler_method, mainsignaler_package, pre_signaler_java, 
- %      repo, root_cause, root_extends, root_from, root_package, root_pre_signaler_java, 
- %      root_signaler, rootsignaler_class, rootsignaler_from, rootsignaler_method, 
- %      rootsignaler_package, signaler, signaler_class, signaler_from, 
- %      signaler_method, signaler_package, stack
- % FROM stacktracesummary;
-
 As mentioned before, in this study we selected a subset of Java projects to
@@
 
-\subsection{The Mining Process}
-\label{sec:miningproc}
+\subsubsection{Android Apps in GoogleCode}
+
+Until January 2014 there were 788 android projects available in Google code repository from which 724 defined at least 1 issue and 183 projects 
+had at least one issue including a stack trace. To identify the Android projects on Googlecode we used the search facilities provided looking for 
+projects mentioning "Android" on the name of the project or its description. The repository page of each project was inspected, and from this set
+142 corresponded to Android applications.
+
+To mine the information available on such Google code projects we implemented a web crawler which navegated on each project and issue page
+and extracted the content of each issue to store on relational database.
+\footnote{Google code previoulsly provided a webservice to extract information about each project, but this service was removed in what Google called "clean-up action" which happened in June 2013.}
+
+\subsection{The ExceptionMiner Tool}
+\label{sec:exceptionminer}
+
+To extract exceptions from issues, we implemented ExceptionMiner, a modular
+mining tool able to connect to various repositories (such as Google Code,
+GHTorrent or even directly to Bugzilla), extract issues, mine stack traces from
+them and classify them in predefined categories. The main components of
+ExceptionMiner are as follows:
+
+\noindent\emph{Exception StackTrace Miner} The first step in the process is
+mining references to exceptions and stack traces embedded in issues, d
+GitHub Commector, and GoogleCode Issue Crawler.
+
+\noindent\emph{Exception Stack Trace Distiller} 
+Distills the information that composes a stack trace and storing the results in a
+relational database. Some of the information extracted from the stack traces were:
+ the exception being thrown, its signaler, the exception root cause, its corresponding signaler,
+and packages related to each one of these fields.
+The tool is based on a combination between a regular expression based parser 
+with heuristics able to identify and filter exception names and stack traces inline with text. In
+contrast to existing issue parsing solutions such as Infozilla, the parser
+created in this work can extract all causes related to an exception on a stack,
+and stack traces embedded in logs files.
+AS PRESENTED IN FIGURE X
+
+%\begin{tiny}
+%\begin{verbatim}
+
+%03-01 W/dalvikvm(7924): threadid=1: thread exiting with uncaught exception (group=0x40adf210)
+%03-01 15:55:01.609 (7924): FATAL EXCEPTION: main
+%03-01  (7924): java.lang.RuntimeException: ...
+%E/AndroidRuntime( 7924):  at android....performLaunchActivity(ActivityThread.java:1967)
+%03-01 15:55:01.609 (7924):  at android....handleLaunchActivity(ActivityThread.java:1992)
+%03-01 15:55:01.609:  at android.app.ActivityThread.access\$600(ActivityThread.java:127)
+%E/AndroidRuntime( 7924):  at android....handleMessage(ActivityThread.java:1158)
+%\end{verbatim}
+%\end{tiny}
+
+\noindent\emph{Exception Classifier} The next step in the ExceptionMiner process
+is classifying exceptions as either checked exceptions, runtime exceptions or 
+errors. The exception classifier is based on bytecode analysis (based on Design
+Wizard~\cite{Brunet09}) that walks up the type hierarchy of a Java exception
+until it reaches a base exception type.
+
+\noindent\emph{Exception Signaler Classifier} After its initialization it
+classifies each signaler according to its origin (i.e. Android Plaform,
+Aplication, Library, Libcore, or Java) based on pattern matching between the
+signaler name and the packages associated to each category.
+
+\subsubsection{The ExceptionMiner Tool}
+
+Before investigating the research questions listed above we analyzed the prevalence of stack traces 
+on issues created on Github. Prevalence or prevalence proportion, a term often used in epidemiology,
+represents the proportion of a population found to have a particular condition
+(e.g. a disease) over the total number of the studied population. In our work, we used
+this term to evaluate a condition (i.e., presence of a stack trace) over issues
+reported on GitHub repositories.
+
+Using the ExceptionMiner tool, developed in this work, we could extract the exception stack traces defined on issues and issue comments.
+Executing ExceptionMiner tool each issue and issue comment defined on the 16,836 projects considered in this study, we could observe that 3,758 projects (22,32\%) contained at least 1 issue on which a stack trace was found.
+Overall the number of issues on which a stack trace was found was 21,013 from which 28,800 stack traces were extracted and analyzed - some issues defined more than one stack trace.
+
+We then analyzed the lables associated to such issues, and we discovered that only 5,196 were labled as defect ~\footnote{the dedect lables considered in this work were every lable containing as substring: defect, bug, crash, failure, fail, exception} (which represent only 24,7\% of the issues on which a stack trace was found), most of the issues on which stack traces were found contained no lable.
+Such observation supports our assumption to consider not only the issues related with defect issues on this analysis.
+
+\subsubsection{Data Filtering and Manual Inspections}
+
+COMPLETE HERE.....
+
+\subsubsection{Database of Exception Stack Trace Information}
+\subsubsection{Distiling Exception Stack Trace Information}
+
+\- root exception
+\- signaler
+\- method
+\- class
+\- package
+\- Exception
+
+\subsection{Process}
+
+ESPALHAR ESTA SECAO NAS OUTRAS.
 
@@
 
-\begin{figure*}
-\centering
-\includegraphics[width=\hsize]{overview.pdf}
-\caption{Overview of the study based on mixed approaches.}
-\label{fig:overviewfig}
-\end{figure*}
+%\begin{figure*}
+%\centering
+%\includegraphics[width=\hsize]{overview.pdf}
+%\caption{Overview of the study based on mixed approaches.}
+%\label{fig:overviewfig}
+%\end{figure*}
 
@@
 
-OBS: THE PICTURE WILL BE UPDATED TO REFLECT THIS PROCESS DESCRIPTION.
-
-\subsection{The ExceptionMiner Tool}
-\label{sec:exceptionminer}
-
-To extract exceptions from issues, we implemented ExceptionMiner, a modular
-mining tool able to connect to various repositories (such as Google Code,
-GHTorrent or even directly to Bugzilla), extract issues, mine stack traces from
-them and classify them in predefined categories. The main components of
-ExceptionMiner are as follows:
-
-\noindent\emph{StackTraceMiner and Distiller} The first step in the process is
-mining references to exceptions and stack traces embedded in issues, distilling
-the information that composes a stack trace and storing the results in a
-relational database. Some of the information extracted from the stack traces were:
- the exception being thrown, its signaler, the exception root cause, its corresponding signaler,
-and packages related to each one of these fields.
-The tool is based on a combination between a regular expression based parser 
-with heuristics able to identify and filter exception names and stack traces inline with text. In
-contrast to existing issue parsing solutions such as Infozilla, the parser
-created in this work can extract all causes related to an exception on a stack,
-and stack traces embedded in logs files.
-
-%\begin{tiny}
-%\begin{verbatim}
-
-%03-01 W/dalvikvm(7924): threadid=1: thread exiting with uncaught exception (group=0x40adf210)
-%03-01 15:55:01.609 (7924): FATAL EXCEPTION: main
-%03-01  (7924): java.lang.RuntimeException: ...
-%E/AndroidRuntime( 7924):  at android....performLaunchActivity(ActivityThread.java:1967)
-%03-01 15:55:01.609 (7924):  at android....handleLaunchActivity(ActivityThread.java:1992)
-%03-01 15:55:01.609:  at android.app.ActivityThread.access\$600(ActivityThread.java:127)
-%E/AndroidRuntime( 7924):  at android....handleMessage(ActivityThread.java:1158)
-%\end{verbatim}
-%\end{tiny}
-
-\noindent\emph{Exception Classifier} The next step in the ExceptionMiner process
-is classifying exceptions as either checked exceptions, runtime exceptions or 
-errors. The exception classifier is based on bytecode analysis (based on Design
-Wizard~\cite{Brunet09}) that walks up the type hierarchy of a Java exception
-until it reaches a base exception type.
-
-\noindent\emph{Exception Signaler Classifier} After its initialization it
-classifies each signaler according to its origin (i.e. Android Plaform,
-Aplication, Library, Libcore, or Java) based on pattern matching between the
-signaler name and the packages associated to each category.
-
-%\section{Study Operation and Results}
+\subsection{Replication Package}
+All the data used in this study will be publicly available at http://.....
+Specifically we provide: (i) all issues related to Android projects found
+in GITHub and GoogleCode used in this study; (iii) all stack traces extracted
+from issues (ii) the results of manual inspection steps; (iv) the tool developed 
+in this work to support stack trace extraction and distilling.
+
 Next section presents the results for each of the study phases, providing both a
@@
 
-\subsection{Characterizing the Most Common Exceptions}
+\subsection{Characterizing the Most Common Root Exceptions}
 

### 2014-08-25 android-stacks.tex
best practices

@@
 
-\subsection{Exceptions in Java} \label{sec:extypes}
+\subsection{Exception Types in Java} \label{sec:extypes}
 
@@
 
-\subsection{Related Best Practices}
+\subsection{Best Practices}
 \label{sec:best}
 
-Several general guidelines have been proposed on how to use Java
+There is a long-lasting debate about the pros and cons of checked and unchecked exceptions 
+~\cite{javatut,stackoverlow,debate},\footnote{152 questions in Stackoverlow
+are related to this debate}. Such debates, however, are mainly based on the developers own experience in
+using both types - since there is of a lack empirical studies on their the pros and cons.
+
+Nevertheless, several general guidelines have been proposed on how to use Java
 exceptions~\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,
-bloch2008effective}. 
-We compiled a list of the guidelines related to exception types and exception propagation in Java, summarized below:\footnote{We also compiled guidelines related
+bloch2008effective}, regardless the exception type chosen. Such guidelines do not focus on 
+advocating any specific type but rather propose ways to effectively use each of them.
+ We compiled a list of the guidelines related to exception types in Java, 
+summarized below:\footnote{We also compiled guidelines related
 to exception handling but they are out of the scope of this paper.}
@@
 
-  \item[RQ2] What patterns of exception (mis)-use related to exception types and exception propagation emerge from the exception stack trace analysis?
+  \item[RQ2] \noindent\emph{What patterns of exception (mis)-use related to exception types and propagation emerge from the exception stack trace analysis?}
 

### 2014-08-25 android-stacks.tex
best practices more improvements

@@
 are related to this debate}. Such debates, however, are mainly based on the developers own experience in
-using both types - since there is of a lack empirical studies on their the pros and cons.
+using both approaches - since there is a lack of empirical studies on their the pros and cons.
 

### 2014-08-25 android-stacks.tex
Improving introduction

@@
 
-The pervasiveness of the World Wide Web and the astonishing increase in the
-number of smartphones have stimulated the creation of a new class of
-applications which extend the phones capabilities far beyond of the basic calls
-and textual messages.
-
-Such applications run over one or more platforms, among them the Android platform
-leads the smartphone market, comprising around XXXX [GoogleIO2014], thanks to its
-Java-based open-source model and hardware partners like Samsung, HTC, Motorola, Asus 
-(besides Nexus its own device) [ref].
-
-Such applications are intrinsically multi-threaded and the development model and used APIs 
-can be  slightly or drastically different and more error prone than tradicional development [arie-icse].
-
-Moreover, such applications have to cope with an increasing number of abnormal
-computation states. Besides the known source of failures (i.e., programmming errors, noisy 
-user inputs or faults in underlying middleware or hardware) such applications need to deal with abnormal states that 
- arise as a consequence of:  memory and batery restrictions [ref], faults on third party services and libraries, 
-compatibility issues caused by constant platforms updates [ref], conection with devices such as GPS, camera.
-
-Therefore, techniques for error detection and handling are not  an optional add-on but a fundamental part of any system~\cite{bruntink2006discovering}.
-The Android Platform uses the exception handling mechanism embeeded
- in Java to deal with such conditions. However, although the exception handling mechanism
+The pervasiveness of the World Wide Web and the astonishing increase 
+in the number of smartphones stimulated the creation of applications which extend the phones
+ capabilities far beyond of the basic calls and textual messages.
+
+If we look at global smartphone shipments the numbers are astoning;
+the mobile industry shiped over 315 million smartphones in the last quarter of 2013.
+Such increase on smartphones combined with the pervasiveness of the World Wide Web
+ have stimulated the creation of applications which extend the phones capabilities 
+far beyond of the basic calls and textual messages.
+
+In this context, Android  leads the smartphone market, having over then 1 billion 
+active users per month; thanks to its open-source model and hardware partners
+ like Samsung, HTC, Motorola, Asus (besides Nexus its own device) [ref].
+
+The number of apps built on top of Android increases in a daily rate <INCLUDE SOME STATISTICS>.
+
+%If on the one hand such applications are getting mode and more popular....
+%On the other hand... there is a need for realiability, as some core functionalities 
+%are moving to apps....such applications impose more challenges to software reliability
+
+%Such mobile platforms are intrinsically multi-threaded and support a development model 
+% slightly (or sometimes drastically) different from those adopted on Web development [arie-icse],
+
+%abnormal computation states
+
+Such applications have to cope with an increasing number of abnormal
+computation states, arising from: programmming errors (e.g. null dereferences);
+ noisy user inputs; faults in underlying middleware or hardware; 
+faults on third party services and libraries, compatibility issues [ref], 
+conection with devices e.g., as GPS, camera, among others; memory and batery restrictions
+among others.
+
+Therefore, techniques for error detection and handling are not  an optional add-on but a 
+fundamental part of such apps.
+% as any other modern system~\cite{bruntink2006discovering}. 
+The Android platform uses the exception handling mechanism embeeded
+ in Java to deal with such conditions. Although the exception handling mechanism
 ~\cite{goodenough1975exception} is one of the most used schemes for
@@
 
-Untill recent years, the availability of issues reporting exceptional conditions was scarce, while the
+Untill recent years, the availability of issues reporting abnormal
+computation states (e.g., crashes) was scarce, while the
 available data concerned a few large scale projects (e.g. the Mozilla
@@
 
-Among the issues defined on issue trackers, we can find issues reporting bugs, and system crashes 
-found as a result of application executions. Such issues usually contain exception stack traces [ref][ref]. 
+%Among the issues defined on issue trackers, we can find issues reporting bugs, and system crashes 
+%found as a result of application executions. Such issues usually contain exception stack traces [ref][ref].
+
 When available, such information can enable different kinds of post-mortem analysis, such as:
-: classify .... (get from previous versions).
-
-Such analysis focuses a system at a time, and none of them target a family of Android Application
-(which share a common core) looking for similar characteristics and vulnerabilitys on the crashes reported.
+: classify .... (get from previous versions). Previous studies on crash report analysis focused a system 
+at a time. Moreover, none of them target 
+a family of Android applications (which share a common core) extracting the characteristics on exceptional issues 
+and looking for similar vulnerabilities on applications and underlaying platform.
 
@@
 exceptions~\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,
-bloch2008effective}, regardless the exception type chosen. Such guidelines do not focus on 
+bloch2008effective} regardless the exception type used. Such guidelines do not focus on 
 advocating any specific type but rather propose ways to effectively use each of them.

### 2014-08-25 android-stacks.tex
More improvements on introduction

@@
 Our conjecture is that several Android applicatons may be suffering from similar vulnerabilities on the exception
-handling code, and that such vulnerabiliries may be leading to application crashes.
+handling code of the applications themselves or on the underlaying platform, and that such vulnerabiliries may be leading
+ to application crashes.
 
@@
 
+<SUCH ISSUES INCLUDE STACK TRACES. WHAT IS THE RELATION BETWEEN STACK TRACE AND CRASH>
+<stack traces are part of good issues....>
 When available, such information can enable different kinds of post-mortem analysis, such as:
@@
 
-In this work, we perform the first post mortem analysis of the exception stack traces embedded on issues
+In this work, we perform the first post mortem analysis of the exception stack traces embedded on the issues
 of all Android applications available in GitHub and Googlecode. Using a custom tool called ExceptionMiner,
@@
 %findings, such as:
+
+This study is built on the following assumption: the stack traces reported on issues contain 
+relevant information related to application crashes.
+Several studies and techniques have also based on this assumption (e.g.,~\cite{sinha2009fault}~\cite{dhaliwal2011classifying}~\cite{kim2013predicting}).
 

### 2014-08-26 android-stacks.tex
More improvements on introduction

@@
 
-The pervasiveness of the World Wide Web and the astonishing increase 
-in the number of smartphones stimulated the creation of applications which extend the phones
- capabilities far beyond of the basic calls and textual messages.
-
-If we look at global smartphone shipments the numbers are astoning;
-the mobile industry shiped over 315 million smartphones in the last quarter of 2013.
-Such increase on smartphones combined with the pervasiveness of the World Wide Web
- have stimulated the creation of applications which extend the phones capabilities 
-far beyond of the basic calls and textual messages.
-
-In this context, Android  leads the smartphone market, having over then 1 billion 
-active users per month; thanks to its open-source model and hardware partners
- like Samsung, HTC, Motorola, Asus (besides Nexus its own device) [ref].
-
-The number of apps built on top of Android increases in a daily rate <INCLUDE SOME STATISTICS>.
+Along the recent years, we have witnessed an astonishing increase in the number 
+mobile applications available. Such increase is motivated by pervasiveness of the World Wide Web 
+and the unprecedent increase of global smartphone shipments -
+the mobile industry shiped over 315 million smartphones in the last quarter of 2013 [ref-IO].
+
+%the mobile industry shiped over 315 million smartphones in the last quarter of 2013
+%The pervasiveness of the World Wide Web and the astonishing increase 
+%in the number of smartphones stimulated the creation of applications which extend the phones
+% capabilities far beyond of the basic calls and textual messages. 
+%f we look at global smartphone shipments the numbers are astoning;
+%the mobile industry shiped over 315 million smartphones in the last quarter of 2013 [ref-googleIO].
+%Such increase on smartphones combined with the pervasiveness of the World Wide Web
+%have stimulated the creation of applications which extend the phones capabilities 
+%far beyond of the basic calls and textual messages.
+
+In this context, Android is currently the leading plaftorm of such applictions - thanks to its open-source 
+model and hardware partners like Samsung, HTC, Motorola, Asus (besides Nexus its
+ own device) [ref].  Google estimates that there is over 1 billion Android active users per month [ref-IO],
+and the number of apps built on top of Android increases in a daily rate.
 
@@
 %are moving to apps....such applications impose more challenges to software reliability 
-
 %Such mobile platforms are intrinsically multi-threaded and support a development model 
 % slightly (or sometimes drastically) different from those adopted on Web development [arie-icse],  
-
 %abnormal computation states
 
+If on the one hand such applications extend the phones capabilities 
+far beyond of the basic calls and textual messages, on the other hand
+they have to face to new challenges imposed by mobile environmet.
 Such applications have to cope with an increasing number of abnormal
-computation states, arising from: programmming errors (e.g. null dereferences);
+computation states, arising from: programmming errors;
  noisy user inputs; faults in underlying middleware or hardware; 
 faults on third party services and libraries, compatibility issues [ref], 
-conection with devices e.g., as GPS, camera, among others; memory and batery restrictions
+conection with devices e.g., as GPS, camera, memory and batery restrictions
 among others.
@@
 have shown that it is generally poorly understood and the less tested parts of the system ~\cite{coelho2011unveiling}.
-The consequence of the exception handling code may come the other way around making the code even less robust.
-
-Our conjecture is that several Android applicatons may be suffering from similar vulnerabilities on the exception
-handling code of the applications themselves or on the underlaying platform, and that such vulnerabiliries may be leading
- to application crashes.
+As a consequence that the exception handling code may become  ilself a source of faults, 
+and may affect the system the other way around, making the system even less robust ~\cite{coelho2011unveiling}[ref FU CHEN].
+
+In the same rathe the number of number of mobile appsincreases, also increases 
+the number of users affected by their failures. Our conjecture is that several 
+Android applicatons may be suffering from similar vulnerabilities on the exception 
+handling code of the applications themselves or on the underlaying platform, and 
+that such vulnerabiliries may be leading to application crashes.
 
 Untill recent years, the availability of issues reporting abnormal
-computation states (e.g., crashes) was scarce, while the
-available data concerned a few large scale projects (e.g. the Mozilla
+computation states (e.g., crashes) was scarce. The
+available data comprised few large scale projects (e.g. the Mozilla
 crash report dataset\footnote{https://code.google.com/p/promisedata/issues/detail?id=88}).
-Along the recent years, we have witnessed a rapid increase in the number of open
+Recently we have observed a rapid increase in the number of open
 source projects that make their control version and issue tracking systems
@@
 <stack traces are part of good issues....>
-When available, such information can enable different kinds of post-mortem analysis, such as:
-: classify .... (get from previous versions). Previous studies on crash report analysis focused a system 
-at a time. Moreover, none of them target 
-a family of Android applications (which share a common core) extracting the characteristics on exceptional issues 
-and looking for similar vulnerabilities on applications and underlaying platform.
+
+When available, crash-related information added on issues and crash reports can enable different kinds of post-mortem analysis, such as
+ bug classification and clustering ~\cite{wang2013improving, kim2011crash, dhaliwal2011classifying},  the automated bug fixing ~\cite{sinha2009fault},  
+fault-proneness prediction models ~\cite{kim2013predicting} and can also be used to support developers in debugging ~\cite{schroter2010stack}.
+
+Previous studies on crash report analysis focused a system 
+at a time. Moreover, none of them target a family of Android applications (which share a common core) extracting 
+the characteristics on exceptional issues and looking for similar vulnerabilities on applications and underlaying platform.
 

### 2014-08-26 android-stacks.tex
More improvements on introduction

@@
 Along the recent years, we have witnessed an astonishing increase in the number 
-mobile applications available. Such increase is motivated by pervasiveness of the World Wide Web 
+mobile applications publicly available. Such increase is motivated by pervasiveness of the World Wide Web 
 and the unprecedent increase of global smartphone shipments -
@@
 
-In this context, Android is currently the leading plaftorm of such applictions - thanks to its open-source 
+In this context, Android has become the leading plaftorm of such applictions - thanks to its open-source 
 model and hardware partners like Samsung, HTC, Motorola, Asus (besides Nexus its
@@
 far beyond of the basic calls and textual messages, on the other hand
-they have to face to new challenges imposed by mobile environmet.
+they have to cope to new challenges imposed by mobile environmet.
 Such applications have to cope with an increasing number of abnormal
@@
 
-In the same rathe the number of number of mobile appsincreases, also increases 
-the number of users affected by their failures. Our conjecture is that several 
-Android applicatons may be suffering from similar vulnerabilities on the exception 
-handling code of the applications themselves or on the underlaying platform, and 
-that such vulnerabiliries may be leading to application crashes.
+Google estimates that Android has over 1 billion active users per month, 
+who pull out theirs phones and check it over 100 bilion times. 
+In the same rate the number of mobile users increase, also increases 
+the number of users affected by faulty apps and app crashes.
+
+%If we google for "Android app crashes"
+% we can find 1.200.000 pages about the subject on forums and blogs. In Feb. 2014
+%Gartner issue a report on mobile technologies and capabilities 
+%[ref-https://www.gartner.com/doc/2665315/top--mobile-technologies-capabilities], 
+%which emphasized the importance of monitoring application crashes.
+
+Our conjecture is that several Android applicatons may be suffering from similar 
+vulnerabilities on the exception handling code of the applications themselves or 
+on the underlaying platform, and that such vulnerabiliries may be leading to
+ application crashes -  as Uncaught Exceptions lead to app to crashes in
+Android and is one of the main causes of application crashes in Java [ref AST] .
 
@@
 computation states (e.g., crashes) was scarce. The
-available data comprised few large scale projects (e.g. the Mozilla
+available data comprised only few large scale projects (e.g. the Mozilla
 crash report dataset\footnote{https://code.google.com/p/promisedata/issues/detail?id=88}).
 Recently we have observed a rapid increase in the number of open
-source projects that make their control version and issue tracking systems
+source projects that make their control version as well as the issue tracking systems
 publicly available in repository hosting sites such as GitHub and Googlecode.
@@
 
-Previous studies on crash report analysis focused a system 
-at a time. Moreover, none of them target a family of Android applications (which share a common core) extracting 
-the characteristics on exceptional issues and looking for similar vulnerabilities on applications and underlaying platform.
+Most of such studies, however, focused the analysis on one or two system at a time. None 
+of them target crash-related information of family of Android applications 
+(which share a common core) aiming at  extracting their common characteristics 
+ and looking for similar vulnerabilities on applications and underlaying platform.
 
@@
     \hline
-    \bfseries{Wrapper Type}  &  \bfseries{Root Cause Type} &  \bfseries{Projects}  &  \bfseries{Occurrences} & \textsf{Android} & \textsf{Libcore} & \textsf{App} & \textsf{Lib}  \\
-    \hline
-      Runtime &  Checked   & 70 & 114 & 0 &  0  & 0 &  0 \\
-      Runtime   &  Error   & 38  & 57 & 0 &  0  & 0 &  0 \\
-      Checked &  Runtime   & 12 & 22 & 0 &  0  & 0 &  0 \\
-      Checked & Error      & 7 &  8 & 0 &  0  & 0 &  0 \\
-      Error & Checked      & 4  &  7 & 0 &  0  & 0 &  0     \\
-      Error & Runtime     & 6  &  13   & 0 &  0  & 0 &  0    \\
-     Throwable & Runtime   & 1 &  3   & 0 &  0  & 0 &  0    \\
+    \bfseries{Wrapper Type}  &  \bfseries{Root Cause Type} &  \bfseries{Projects}  &  \bfseries{Occurrences} & \textsf{Android} & \textsf{Java/Libcore} & \textsf{Lib} & \textsf{App}  \\
+    \hline
+      Runtime &  Checked   & 70 & 114 & 66 &  0  & 19 &  29 \\
+      Runtime   &  Error   & 38  & 57 & 50 &  0  & 6 &  1 \\
+      Checked &  Runtime   & 12 & 22 & 4 &  0  & 9 &  9 \\
+      Checked & Error      & 7 &  8 & 5 &  0  & 0 &  3 \\
+      Error & Checked      & 4  &  7 &  &  0  & 7 &  0     \\
+      Error & Runtime     & 6  &  13   & 1 &  1  & 0 &  11    \\
     \hline

### 2014-08-26 android-stacks.tex
Improving backgound and discussion

@@
 
-Android is an open source platform for mobile devices based on Linux kernel.
-The plaftom is also comprised of a core Java library, a set of libraries third-party libraries e.g.  SQLite,
-the Dalvik runtime environment, and the Aplication Framework which provides higher-level services to the applications
-running on the platform e.g., controlling activities or providing data access.
-
-Android supports an event-driven development model. The main element responsible for 
-receiving the GUI events is called Activity, which is the unity of user interaction and execution in
-Android. Android also provides APIs for data storage, communication between Activities, 
-and other services. Moreover, it also provides frameworks for concurrent developement (e.g. AsyncTask) 
-which is highly needed due to the demand of responsiveness in such environment  - 
-mobile apps can easily become unresponsivedue to resurce limitation and excessive
- network access [shimidit and arie].
+Android is an open source platform for mobile devices based on Linux kernel,
+which is also composed by a core Java library, a set of libraries third-party libraries (e.g.  SQLite),
+the Dalvik runtime environment, and the Aplication Framework (which provides higher-level services to the applications
+running on the platform).
+
+The Android framework supports an event-driven development model. 
+Among the basic elements that comprise an Android application, we can cite the
+Activity. Each application usually defines one Activity per application Window.
+ Each Activity defines a set of hook methods  (e.g., OnCreate(), onPause()) that are dispatched
+automatically by the Android framework (by inversion of
+control [ref]), in response to infrastructure and application events.
+
+Due to the high demand for responsiveness in mobile environment (mobile apps can easily become 
+unresponsivedue due to resurce limitation and excessive network access [shimidit and arie]) the
+Android framework also provides APIs to support concurrent developement (e.g. AsyncTask).
 

### 2014-08-26 android-stacks.tex
Improving backgound

@@
 Android is an open source platform for mobile devices based on Linux kernel,
-which is also composed by a core Java library, a set of libraries third-party libraries (e.g.  SQLite),
-the Dalvik runtime environment, and the Aplication Framework (which provides higher-level services to the applications
+which also contains by a core Java library, a set of libraries third-party libraries (e.g.  SQLite),
+the Dalvik runtime environment, and the Application Framework (which provides higher-level services to the applications
 running on the platform).
 
-The Android framework supports an event-driven development model. 
+The Android Application Framework supports an event-driven development model. 
 Among the basic elements that comprise an Android application, we can cite the
@@
 unresponsivedue due to resurce limitation and excessive network access [shimidit and arie]) the
-Android framework also provides APIs to support concurrent developement (e.g. AsyncTask).
+Android Application Framework also provides APIs to support concurrent developement (e.g. AsyncTask).
+As this framework is developed in Java, it reuses the embedded Java Exception model 
+to detect and handle abnormal computation states.
 

### 2014-08-26 android-stacks.tex
small changes on sections 2 and dicussions

@@
 Android is an open source platform for mobile devices based on Linux kernel,
-which also contains by a core Java library, a set of libraries third-party libraries (e.g.  SQLite),
+which also contains  a core Java library, a set of libraries third-party libraries (e.g.  SQLite),
 the Dalvik runtime environment, and the Application Framework (which provides higher-level services to the applications
@@
 
+<AVALIAR MOVER ESTE STACK TRACES PARA UMA FIGURA QUE MOSTR COMO ELE É QUEBRADO...
+PELO EXCEPTION MINER E AS INFORMAÇÕES VAO PARA O BANCO...>
+
 % \subsection{Public Project Repositories and Issue Trackers} 
@@
 
-The goal of this study was to extract nformation available on exception stack 
-traces embedded on issues reported on Android projects, looking for the common characteristics of
-those issues and investigating whether they could reveal common defficiencies
-on the exception handling code of both the applications and the underlaying 
+This study focus on the analysis of crash information (related to uncaught exceptions) of a family of Android applications
+aiming at  extracting their common characteristics and looking for similar vulnerabilities on applications and underlaying platform.
+
+In Java when an application crashes dut to an uncaught exception ... 
+Exception Stach trace is usually (procurar por artigo).
+
+Falar da assumption que stacj trá relacionada a crash?
+
+The goal of this study was to extract exception stack 
+traces embedded on issues of Android projects (hosted in GitHub and GoogleCode), looking for the 
+common characteristics of such stack-trace and investigating whether
+ they could reveal common defficiencies on the exception handling code of both the applications and the underlaying 
  framework.
 
-We choose Android apps because it is the only top platform that is open-source,
-and due to its increasing popularity it xxxx.
-
-Our study focused on free apps, since the information needed to perform our study
-cannot be retrieved from commercial apps because their issue reports and their
-source code is not publicly available. The Android open-source apps were also 
+Our study focused on open-source apps, since the information needed to perform our study
+cannot be retrieved from commercial apps, whose issue reports and 
+source code are not publicly available. The Android open-source apps were also 
 target of other research studies [fse2013][ahimed].
@@
 
-  \item[RQ1] \noindent\emph{What are the common characteristics of exception stack traces reported on issues?} What are the main causes of crashes reported on issues? (more explicit?)
-
-  \item[RQ2] \noindent\emph{What patterns of exception (mis)-use related to exception types and propagation emerge from the exception stack trace analysis?}
+  \item[RQ1] \noindent\emph{What are the common characteristics of exception stack traces reported on issues?} Some of the common characteristics to be investigated are the 
+ main causes reported on exception stack traces, the elements responsible for signaling such exceptions.
+
+  \item[RQ2] \noindent\emph{What patterns of exception (mis)-use related to exception types and propagation emerge from the exception stack trace analysis? Based on a set of best practices
+compiled in this study, looked for characteristics on stack traces that could pinpoint to practices that were not being obeyed.}
 
@@
 
-The exceptions available on exception stack traces are not limited to the exceptions 
-expliciltly thrown by the Java developer but also include
-the exceptions implicitly thrown by the JVM due to programming errors (e.g., out-of-bounds array index, division-by-zero, 
-access to a null reference) or resource limitation (e.g., OutOfMemoryError).
-
 For RQ1 and RQ2, we explore the domain quantitatively and highlight interesting cases by 
 exploring cases qualitatively. To support this investigation, source code and bytecode 
-analysis were used in combination with manual inspection to leverage the  understanding of stack traces and support further discussions and insights.
+analysis were used in combination with manual inspection to leverage the  understanding 
+of stack traces and support further discussions and insights.
 
@@
 Only high fault tolerant systems need to provide solutions to handle them.
+
+INSERIR EM ALGUM PONTO DO TEXTO:
+The exceptions available on exception stack traces are not limited to the exceptions 
+expliciltly thrown by the Java developer but also include
+the exceptions implicitly thrown by the JVM due to programming errors (e.g., out-of-bounds array index, division-by-zero, 
+access to a null reference) or resource limitation (e.g., OutOfMemoryError).
 
@@
 Discuss about the actionable advices propagated by TAO XIE.
+
+\subsection{Inversion of Control and Exception Handling}
+- Falar de Pytolomy e Ridresh rajan... desafios ja apontados....
 

### 2014-08-27 android-stacks.tex
changesl all over

@@
 
+In Java, when an application fails due to an uncaught exception, 
+it automatically terminates, while the system prints a stack trace to the console, or log file [REF-STACK-TRACE-CRASH]. 
+A stack trace includes a detail message and the execution stack frame when 
+the exception occurred. A typical Java stack trace consists of an ordered list of 
+methods that were active on the call stack before the exception has occurred [REF-STACKTRACE-MANUAL].
+
 Untill recent years, the availability of issues reporting abnormal
-computation states (e.g., crashes) was scarce. The
+computation states (e.g., exception stack traces) was scarce. The
 available data comprised only few large scale projects (e.g. the Mozilla
@@
 
-When available, crash-related information added on issues and crash reports can enable different kinds of post-mortem analysis, such as
- bug classification and clustering ~\cite{wang2013improving, kim2011crash, dhaliwal2011classifying},  the automated bug fixing ~\cite{sinha2009fault},  
-fault-proneness prediction models ~\cite{kim2013predicting} and can also be used to support developers in debugging ~\cite{schroter2010stack}.
+%Stack traces are a useful source of information about system crashes, and is often used to support developers in debugging~\cite{schroter2010stack}.
+Moreover, information %from stack traces combined with crash and bug reports can be used to support bug classification and clustering~\cite{wang2013improving, kim2011crash, %dhaliwal2011classifying}, fault-proneness prediction models~\cite{kim2013predicting} and even automated bug fixing~\cite{sinha2009fault} tools.
+
+When available, exception stack traces provides a useful source of information about system crashes and 
+ can enable different kinds of post-mortem analysis, such as:   debugging~\cite{schroter2010stack}, 
+ bug classification and clustering ~\cite{wang2013improving, kim2011crash, dhaliwal2011classifying},  
+automated bug fixing ~\cite{sinha2009fault} and fault-proneness prediction models analysis ~\cite{kim2013predicting}.
 
 Most of such studies, however, focused the analysis on one or two system at a time. None 
-of them target crash-related information of family of Android applications 
-(which share a common core) aiming at  extracting their common characteristics 
- and looking for similar vulnerabilities on applications and underlaying platform.
-
-In this work, we perform the first post mortem analysis of the exception stack traces embedded on the issues
-of all Android applications available in GitHub and Googlecode. Using a custom tool called ExceptionMiner,
+of them targeted a family of applications (sharing a common framework) aiming at extracting 
+the common characteristics of crash information and checking whether they could provide
+insights on  similar vulnerabilities on the applications and underlaying framework.
+
+In this work, we perform the first post mortem analysis of the exception stack traces embedded on all issues
+created on Android projects available in GitHub and Googlecode. Using a custom tool called ExceptionMiner,
  which we develop specifically for this study, we mine stack traces from issues across all Android projects on
 GitHub and Googlecode. Overall, we analyzed the XXXX issues extracted from on XXX Android projects defined in
-GitHub and XXX projects defined in Googlecode. From these issues  XXX stack traces were extracted and used as
+GitHub and XXX projects defined in Googlecode. From these issues,  XXX stack traces were extracted and used as
  a source of information for our analysis. The stack trace analysis was augmented by additional information
-extracted from JDK, Android SO and Android SDK using byte code and source code analysis - which was done 
-on a carefully selected subset of our initial sample (482 Android projects). Manual inspection was
-used to leverage the understanding of exception stack traces and support further discussions and insights.
+extracted from JDK, Android SO and Android SDK and the applications using byte code and source code analysis.
+ Manual inspection was used to leverage the understanding of exception stack traces and support further discussions and insights.
 
@@
 %findings, such as:
-
-This study is built on the following assumption: the stack traces reported on issues contain 
-relevant information related to application crashes.
-Several studies and techniques have also based on this assumption (e.g.,~\cite{sinha2009fault}~\cite{dhaliwal2011classifying}~\cite{kim2013predicting}).
 
@@
 
-Falar da assumption que stacj trá relacionada a crash?
-
 The goal of this study was to extract exception stack 
@@
 target of other research studies [fse2013][ahimed].
+
+Falar da assumption que stacj trá relacionada a crash?
+This study is built on the following assumption: the stack traces reported on issues contain 
+relevant information related to application crashes.
+Several studies and techniques have also based on this assumption (e.g.,~\cite{sinha2009fault}~\cite{dhaliwal2011classifying}~\cite{kim2013predicting}).
 
@@
 
-INSERIR EM ALGUM PONTO DO TEXTO:
-The exceptions available on exception stack traces are not limited to the exceptions 
-expliciltly thrown by the Java developer but also include
-the exceptions implicitly thrown by the JVM due to programming errors (e.g., out-of-bounds array index, division-by-zero, 
-access to a null reference) or resource limitation (e.g., OutOfMemoryError).
-
 ----------
@@
 
-\subsection{Direct instances of RuntimeExceptions being thrown.}
+\subsubsection{Direct instances of RuntimeExceptions being thrown.}
 
@@
 
-\subsubsection{Classifying Top 100 Exceptions}
+INSERIR EM ALGUM PONTO DO TEXTO:
+The exceptions available on exception stack traces are not limited to the exceptions 
+expliciltly thrown by the Java developer but also include
+the exceptions implicitly thrown by the JVM due to programming errors (e.g., out-of-bounds array index, division-by-zero, 
+access to a null reference) or resource limitation (e.g., OutOfMemoryError).
+
+\subsection{Classifying Top 100 Exceptions}
 
@@
 
+\subsection{Characterizing the Most Common Root Exceptions Signalers}
+<COMPLEMENT.... INCLUDE THE METHODS AND CLASSES RESPONSIBLE FOR SIGNALING
+MOST OF THE EXCEPTIONS AND THE EXCEPTIONS SIGNALED BY THEM. SIMILAR TO THE WRAPPINGS TABLES>
+
 \section{RQ2.  What kinds of exception (mis)-use emerge from stack trace analysis? }
 
-To answer this question we: (i) distilled the information available on stack traces (i.e., exception signalers, root causes, exception wrappings); 
-(ii) mined the exception types related to each exception reported on stack traces using source code and byte code analyses; and finally checked 
-this information against the compiled best practices. Doing so, we mined some characteristics of exception stack traces that points to best practices violations.
+To answer this question, besides distilling the information available on stack traces (i.e., exception signalers, root causes, exception wrappings),
+we mined the exception types (i.e., RuntimeException, checked Exception or Error) related to each exception reported on stack traces 
+using source code and byte code analyses supported by Exception Miner; and finally checked 
+this information against the compiled best practices. Doing so, we mined some characteristics 
+of exception stack traces that points to best practices violations.
 
@@
 
-\subsection{What can the Exception Wrappings tell us?}
-
+\subsection{What the Types of Root Exceptions can tell us?}
+\subsection{Exploring the Information Embeeded on Exception Types}
+
+<falar dos tipos mais comuns.... Runtime e NullPointer???>
+<LINKCAR COM ANALISE DOS SUPERTIPOS>
+<complementar falando do grafico acima>
+\subsubsection{Undocumented runtime exceptions.} After filtering out all the exceptions implicitly 
+signaled by  JVM (due to programming mistakes) and inspecting the signaler methods of such exceptions (i.e.,  direct an indirect 
+instances of java.lang.RuntimeException). We could observe that only 1 method out of 118 inspected signalers
+ (i.e., 0,8\%) documented the explicitly thrown runtime exception in a Javadoc comment; and none of these methods
+included the runtime exception  (reported on the stack) as part of the exception interface of the method (i.e., using 
+throws clause on method signature). This result is aligned with the results of other  study conducted by from 
+Sacramento et al ~\cite{sacramento2006unchecked} which observed that the
+runtime exceptions in .Net programs are most often not documented.
+
+Such undocumented runtime exceptions represent a threat to system robustness, specially
+when such exceptions are thrown then the third party code is invoked inside the application (e.g. libraries, or framework utility code).
+In such cases the client usually do not have access to the source code, and in the absence of
+the exception documentation it is very difficult or even impossible for the client of such third party code to 
+protect the application against undocumented runtime exceptions. As a consequence, the
+ undocumented runtime exception may remain uncaught and lead to system crashes.
+
+<fazer o link com wrappings>
+
+\subsection{What the Exception Wrappings can Tell us?}
+\subsection{Exploring the Information Embeeded on Exception Wrappings}
 Table~\ref{tab:wrappingandroid} presents the wrappings found in this study for all
@@
 
-Such wrapping was detected on distinct 70 projects. Inspecting the 
-elements responsible for performing such wrapping we could observe that,
-most of such wrappings were perform on methods defined on Android.
+Such wrapping was detected on 70 distinct projects. Inspecting the 
+elements responsible for performing such wrapping, we could observe that
+most of such wrappings were performed on methods defined on Android
+Platform (approximately 58%).
+
+%\begin{table}
+%\centering
+%\begin{tabular}{ll}
+%    \hline
+%   \bfseries{signaler from} & \bfseries{Ocurrences} \\
+%    \hline
+%ANDROID	 & 66 \\
+%APP	& 29 	(HOW MANY?? AND WHICH??) \\
+%LIB  & 19 (HOW MANY AND WHICH?)\\
+%\hline
+% \end{tabular}
+%\caption{Elements responsible for the wrapping.}
+%\label{tab:wrapping01}
+%\end{table}
+
+Taking a closer look on the methods reposponsible for such wrappings,
+ we could observe that aproximately 41% of them were performed by a single class,
+ named ActivityThreat. This class is responsible for managing the life cycle of an Android Activities.
 
@@
 \centering
-\begin{tabular}{ll}
-    \hline
-    \bfseries{signaler from} & \bfseries{Ocurrences} \\
-    \hline
-ANDROID	 & 66 \\
-APP	& 29 	(HOW MANY?? AND WHICH??) \\
-LIB  & 19 (HOW MANY AND WHICH?)\\
-\hline
-  \end{tabular}
-\caption{Elements responsible for the wrapping.}
-\label{tab:wrapping01}
-\end{table}
-
-Taking a closer look on the signalers reposponsible for wrapping we could observe that, aproximately 41%
-of all wrappings were performed by a single class, called ActivityThread, responsible for managing the life cycle 
-of an Android application. It performs the application callbacks.
-
-\begin{table}
-\centering
-\begin{tabular}{ll}
- \bfseries{Signaler class} & \bfseries{Ocurrences} \\
-    \hline
-android.app.ActivityThread & 47 \\
-android.view.LayoutInflater & 7 \\
-org.gege.caldavsyncadapter.syncadapter.SyncAdapter & 7 \\
-org.springframework.web.client.RestTemplate & 6 \\
-android.os.AsyncTask & 5 \\
+\begin{tabular}{lll}
+ \bfseries{Signaler class} &  \bfseries{Wrapper Type} & \bfseries{Ocurrences} \\
+    \hline
+android.app.ActivityThread & java.lang.RuntimeException & 47 \\
+android.view.LayoutInflater & android.view.InflateException &  7 \\
+android.os.AsyncTask & java.lang.RuntimeException &  5 \\
+Other platform classes & java.lang.RuntimeException &   7 \\
 \hline
@@
 
-Code snipet extracted from android.app.ActivityThread.
+Figure X presents a code snippet of ActivityThreat which is reponsible for (among other things)
+calling the Activities onCreate method. Hence, as we can see from lines 4-8,
+ in case any exception is thrown by an application Activity during its creation, 
+it is wrapped in a general Runtime exception and rethrown.
 
@@
 
-private Activity performLaunchActivity(...) {
-  try {
-   ...
-  } catch (Exception e) {
-    ...
-    throw new RuntimeException(..., e);
-  }
-}
+1. private Activity performLaunchActivity(...) {
+2.  try {
+3.   ...
+4.   ...callActivityOnCreate(...);
+5.  } catch (Exception e) {
+6.    ...
+7.   throw new RuntimeException(..., e);
+8.  }
+9.}
 \end{verbatim}
@@
 
-The exceptions most wrapped on RuntimeExceptinos were:
-
-\begin{table}
-\centering
-\begin{tabular}{lll}
-    \hline
-    \bfseries{wrapper} & \bfseries{root} & \bfseries{\#} \\
-    \hline
-java.lang.RuntimeException & java.lang.ClassNotFoundException &	32 \\
-java.lang.RuntimeException & java.lang.NoSuchMethodException &	6 \\
-java.lang.RuntimeException & java.lang.InstantiationException) & 5 \\
-...andlyticsproject.NetworkException & org.apache...HttpResponseException &	5\\
-android.view.InflateException & java.lang.NoSuchMethodException & 4 \\
-\hline
-  \end{tabular}
-\caption{Elements tresponsible for the wrapping.}
-\label{tab:wrapping01}
-\end{table}
-
-\begin{table}
-\centering
-\begin{tabular}{lll}
-    \hline
-    \bfseries{Exception} & \bfseries{Ocurrences} \\
-    \hline
-java.lang.RuntimeException &	69 & 61\% \\
-android.view.InflateException	& 10 & 8,7\% \\
-com.github.andlyticsproject.console.NetworkException	& 7 & - \\
-java.lang.IllegalStateException	& 6 & - \\
-
-\hline
-  \end{tabular}
-\caption{Elements tresponsible for the wrapping.}
-\label{tab:wrapping01}
-\end{table}
+The inversion of control makes it dificult to the application developer to associate 
+a specific handler to the exceptions thrown during the onCreate method (or any other Activity life cycle method).
+Since according to Java handling model the handler should me added on any method in the call chain,
+and in the inversion of control development model the developer usually does not have fully access to 
+the invocation call chain of a method in order to associate a handler, as a result the 
+exception may scape from any handling and remain uncaught.
+
+Difficulty of associating a specific handler to exceptions signaled by application methods called by the framework,
+may also be the reason behind the design desicion of encapsulatling any exception in a RuntimeException.
+As framework does not have enought contextual information the exception regardless its types,
+it encapsulates in a general type and re-throws. The only possible answer to such exceptions would be 
+to finish the application.
+
+Such design decision may lead the developers into believing that by just re-throwing 
+the exceptions they can forget about the exceptional situations during the development of the "happy path". 
+This "ignore-for-now" approach may turn the exception handling into a generalized "goto" mechanism~\cite{mandrioli1992advances}
+ making the program more complex and even less reliable. This behaviour may lead to the well known uncaught exceptions 
+problem~\cite{jo2004uncaught}, one of the main causes of application crashes.
+
+% Several types of exceptions were  wrapped the RuntimeExceptinos thrown by the 
+% ActivityThread. Table X illustrates them.
+% #\begin{table}
+% \centering
+% \begin{tabular}{lll}
+%    \hline
+%    \bfseries{wrapper} & \bfseries{root} & \bfseries{\#} \\
+%    \hline
+% java.lang.RuntimeException & java.lang.ClassNotFoundException &	32 \\
+% java.lang.RuntimeException & java.lang.NoSuchMethodException &	6 \\
+% java.lang.RuntimeException & java.lang.InstantiationException) & 5 \\
+% ...andlyticsproject.NetworkException & org.apache...HttpResponseException &	5\\
+% android.view.InflateException & java.lang.NoSuchMethodException & 4 \\
+% \hline
+%  \end{tabular}
+% \caption{Elements tresponsible for the wrapping.}
+% \label{tab:wrapping01}
+% \end{table}
+
+% \begin{table}
+% \centering
+% \begin{tabular}{lll}
+%   \hline
+%   \bfseries{Exception} & \bfseries{Ocurrences} \\
+%   \hline
+% java.lang.RuntimeException &	69 & 61\% \\
+% android.view.InflateException	& 10 & 8,7\% \\
+% com.github.andlyticsproject.console.NetworkException	& 7 & - \\
+% java.lang.IllegalStateException	& 6 & - \\
+% \hline
+%  \end{tabular}
+% \caption{Elements tresponsible for the wrapping.}
+% \label{tab:wrapping01}
+% \end{table}
 
@@
 
-\subsection{Undocumented runtime exceptions.} After filtering out all the exceptions implicitly 
-signaled by  JVM (due to programming mistakes) and inspecting the signaler methods of such exceptions (i.e.,  direct an indirect 
-instances of java.lang.RuntimeException). We could observe that only 1 method out of 118 inspected signalers
- (i.e., 0,8\%) documented the explicitly thrown runtime exception in a Javadoc comment; and none of these methods
-included the runtime exception  (reported on the stack) as part of the exception interface of the method (i.e., using 
-throws clause on method signature). This result is aligned with the results of other  study conducted by from 
-Sacramento et al ~\cite{sacramento2006unchecked} which observed that the
-runtime exceptions in .Net programs are most often not documented.
-
-Such undocumented runtime exceptions represent a threat to system robustness, specially
-when such exceptions are thrown then the third party code is invoked inside the application (e.g. libraries, or framework utility code).
-In such cases the client usually do not have access to the source code, and in the absence of
-the exception documentation it is very difficult or even impossible for the client of such third party code to 
-protect the application against undocumented runtime exceptions. As a consequence, the
- undocumented runtime exception may remain uncaught and lead to system crashes.
-
 \section{Discussions and Actionable Advices}
@@
 
-%\noindent\emph{The high number of null pointer exceptions}
+%\noindent\emph{The high number of NullPointerExceptions and other JVM-signaled Exceptions}
+
+Moreover, some languages use exceptions to represent coding errors (out-of-bounds array index, division-by-zero, access to a null).
+In these cases exceptions are implicitly signaled by the runtime environment when such condition happens.
+One of such implicit exceptions (caused by data conversion from a 64-bit floating point to a 16-bit signed integer) was the responsible for the famous failure on Ariane 5's first test flight~\cite{lions1996ariane} --- leading to the rocket self-destructing 37 seconds after launch and a loss of 500 million dollars.
+Besides Ariane 5, several applications crash everyday due to uncaught exceptions~\cite{jo2004uncaught}.
+The consequences of not careful exception usage may have the opposite effect of its initial intention of improving system robustness.
 
@@
 (4) Local onde tratar excecao - handler padrao - politica da aplicacao. tratamento difucultado multiplas threads e assyncronos.
+
+%\noindent\emph{Unwritten Rules of Android Exception Handling Developement}
 
@@
 our sample.
+<OLHAR ARTIGO DO FSE QUE TRABALHA COM OPENSOURCE.
+FALAR QUE DESCOBERTAS FORAM DE LIBS JAVA. APESAR DO ESTUDO NAO SER FEITO EM PROJS JAVA.....
+PODEMOS CONCLUIR ALGUMAS COISAS QUE SE APLICAM A TODOS JAVA>
 

### 2014-08-29 android-stacks.tex
changes along the paper, intro

@@
 were extracted. Such stack traces were distilled and analyzed in 
-combination with source code and bytecode analysis. In this study some patterns of exception
-(mis)-use were consistently detected such as: a high prevalence of NullPointerExceptions and 
-other java.lang exceptions reported on issues - they were found in aproximately 50\% of all analyzed 
-exception stack traces; undocumented runtime exceptions signaled by third party code - which is a serious threat to program
-dependability; unexpected wrappings (e.g., Errors being wrapped in checked exceptions), revealing that Java hybrid exception
-model is not fully used according to its purpose.
+combination with source code and bytecode analysis. This study 
+consistently identifies common characteristics of reported 
+exception stacks traces - e.g., the NullPointerException was the reported cause 
+on 30,08\% of all analyzed exception stack traces, and ewas mentioned in approximatelly 
+50% of all projects analyzed. Moreover,  the cross-project analysis 
+of exception stack traces enabled to consistenly reveal   
+fragilities of exception handling code: 
+(1) undocumented runtime exceptions signaled by third party code,
+(2) undocumente checked exceptions thrown by Android native code
+(3) unexpected exception wrappings (e.g., Errors being wrapped in checked exceptions) 
+revealing that Java hybrid exception model is not fully used according to its purpose.
+Such fragilities impose challenges to application robusness not only 
+in Android environemt but also to other systems based on Java exception model.
 
@@
 Along the recent years, we have witnessed an astonishing increase in the number 
-mobile applications publicly available. Such increase is motivated by pervasiveness of the World Wide Web 
+mobile applications being developed and some of them being largely used. 
+Such increase is motivated by pervasiveness of the World Wide Web 
 and the unprecedent increase of global smartphone shipments -
-the mobile industry shiped over 315 million smartphones in the last quarter of 2013 [ref-IO].
+the mobile industry shiped over 315 million smartphones in the last quarter of 2013 ~\cite{googleio}.
 
@@
 
-In this context, Android has become the leading plaftorm of such applictions - thanks to its open-source 
-model and hardware partners like Samsung, HTC, Motorola, Asus (besides Nexus its
- own device) [ref].  Google estimates that there is over 1 billion Android active users per month [ref-IO],
+In this context, Android has become a leading plaftorm for such applictions - thanks to its open-source 
+model and hardware partners like Samsung, HTC, Motorola, Asus (besides Nexus Google's
+ own device) ~\cite{gartner}.  Google estimates that there is over 1 billion Android active users per month ~\cite{googleio},
 and the number of apps built on top of Android increases in a daily rate.
+
+%ref: http://www.gartner.com/newsroom/id/2645115
+% Gartner, Inc. Press release, 2012. www.gartner.com/newsroom/id/2237315.
+% http://www.gartner.com/newsroom/id/2645115
 
@@
 
-If on the one hand such applications extend the phones capabilities 
+In the same rate the number of a users increase, also increases 
+the number of users affected by application faults and crashes.
+If on the one hand such applications extend phones capabilities 
 far beyond of the basic calls and textual messages, on the other hand
-they have to cope to new challenges imposed by mobile environmet.
+they have to face new threats to application robustness imposed by mobile environmet.
 Such applications have to cope with an increasing number of abnormal
@@
 among others.
+
+%Hence, in the same rate the number of mobile users increase, also increases 
+%the number of users affected by such threats.
+
+%faulty apps and app crashes.
 
@@
 As a consequence that the exception handling code may become ilself a source of faults, 
-and may affect the system the other way around, making the system even less robust ~\cite{coelho2011unveiling}[ref FU CHEN].
-
-Google estimates that Android has over 1 billion active users per month, 
-who pull out theirs phones and check it over 100 bilion times. 
-In the same rate the number of mobile users increase, also increases 
-the number of users affected by faulty apps and app crashes.
+and may affect the system the other way around, making the system even less robust ~\cite{robillard2003static,coelho2011unveiling}.
+
+% Google estimates that Android has over 1 billion active users per month, 
+% who pull out theirs phones and check it over 100 bilion times.
 
@@
 Our conjecture is that several Android applicatons may be suffering from similar 
-vulnerabilities on the exception handling code of the applications themselves or 
-on the underlaying platform, and that such vulnerabiliries may be leading to
- application crashes -  as Uncaught Exceptions lead to app to crashes in
-Android and is one of the main causes of application crashes in Java [ref AST] .
+fragilities on the exception handling code (of the applications themselves or 
+the underlaying platform), and that such fragilities may be leading to
+decrease the application robustness, leading for instance to the Uncaught Exception problem
+ are one of the main causes of application crashes ~\cite{jo2004uncaught} ~\cite{zhang2012amplifying}.
+
+%[ref AST].
 
 In Java, when an application fails due to an uncaught exception, 
-it automatically terminates, while the system prints a stack trace to the console, or log file [REF-STACK-TRACE-CRASH]. 
-A stack trace includes a detail message and the execution stack frame when 
-the exception occurred. A typical Java stack trace consists of an ordered list of 
-methods that were active on the call stack before the exception has occurred [REF-STACKTRACE-MANUAL].
+it automatically terminates, while the system prints a stack trace to the console, 
+or log file ~\cite{gosling2000java}.  A typical Java stack trace consists of an ordered list of 
+methods that were active on the call stack before the exception has occurred,  
+ and the fully qualified name of such exception.
 
@@
 Android Application Framework also provides APIs to support concurrent developement (e.g. AsyncTask).
-As this framework is developed in Java, it reuses the embedded Java Exception model 
-to detect and handle abnormal computation states.
+
+Android platform is mostly implemented in Java, uses JNI[ref] to access native C code, and 
+reuses the embedded Java Exception model to detect and handle abnormal computation states.
 
@@
 
-When an exception is thrown, it causes a transfer of control from the point
-where the exception occurred to a point that can be specified by the programmer
-(exception handler) - in Java it is represented by the try-catch block . A common way of  
-propagating exceptions in Java programs is through exception wrapping
+%When an exception is thrown, it causes a transfer of control from the point
+%where the exception occurred to a point that can be specified by the programmer
+%(exception handler) - in Java it is represented by the try-catch block .
+
+In Java, once an exception is thrown, the JVM looks for the nearest enclosing exception handler
+(in Java it is represented by the try-catch block), and unwinds the execution stack if necessary.
+ To look for the handler on the invocation stack is claimed to increase software reusability, 
+since the invoker of an operation can handle it in a wider context [ref-miller].
+
+INCLUIR REF:
+Miller, R.; Tripathi, A.: Issues with Exception Handling in Object-Oriented Systems. In:
+ECOOP’97. LNCS, vol. 1241, Springer (1997) 85–103.
+
+ A common way of  propagating exceptions in Java programs is through exception wrapping
  (also known as exception chaining). Exception wrapping allows one exception 
@@
 
+(1)
+
 The inversion of control makes it dificult to the application developer to associate 
@@
 android.support.v4.content.ModernAsyncTask & 2\\
+\hline
+  \end{tabular}
+\caption{Elements tresponsible for the wrapping.}
+\label{tab:wrapping01}
+\end{table}
+
+\begin{table}
+\centering
+\begin{tabular}{ll}
+ \bfseries{Method} & \bfseries{Project} & & \bfseries{Ocurrences}\\
+    \hline
+android.os.AsyncTask &	122 & 549  \\
+android.view.LayoutInflater & 63 & 163 \\
+android.support.v4.* & 100 & 512\\
 \hline

### 2014-08-29 android-stacks.tex
changes along the paper

@@
 %found as a result of application executions. Such issues usually contain exception stack traces [ref][ref]. 
-
-<SUCH ISSUES INCLUDE STACK TRACES. WHAT IS THE RELATION BETWEEN STACK TRACE AND CRASH>
-<stack traces are part of good issues....>
+%<SUCH ISSUES INCLUDE STACK TRACES. WHAT IS THE RELATION BETWEEN STACK TRACE AND CRASH>
+%<stack traces are part of good issues....>
 
@@
 
-When available, exception stack traces provides a useful source of information about system crashes and 
- can enable different kinds of post-mortem analysis, such as:   debugging~\cite{schroter2010stack}, 
+When available, exception stack traces provides a useful source of information about system crashes ~\cite{bettenburg2008makes} which 
+ can enable different kinds of post-mortem analysis to support:  debugging~\cite{schroter2010stack}, 
  bug classification and clustering ~\cite{wang2013improving, kim2011crash, dhaliwal2011classifying},  
-automated bug fixing ~\cite{sinha2009fault} and fault-proneness prediction models analysis ~\cite{kim2013predicting}.
-
-Most of such studies, however, focused the analysis on one or two system at a time. None 
-of them targeted a family of applications (sharing a common framework) aiming at extracting 
-the common characteristics of crash information and checking whether they could provide
-insights on  similar vulnerabilities on the applications and underlaying framework.
-
-In this work, we perform the first post mortem analysis of the exception stack traces embedded on all issues
+automated bug fixing ~\cite{sinha2009fault} and fault-proneness prediction models ~\cite{kim2013predicting}.
+
+Most of such studies, however, focused the analysis on one or two system at a time.
+This paper reports an exploratory study which targets a family of applications 
+(sharing the common Android framework) aiming at extracting 
+the common characteristics of exception stack traces reported to such systems to provide
+insights on similar fragilities on the exception related code of the applications and underlaying framework.
+
+To do so we perform a post mortem analysis of the exception stack traces embedded on all issues
 created on Android projects available in GitHub and Googlecode. Using a custom tool called ExceptionMiner,
@@
 
+%projects: 788	 - projects with issues - 724 - 
+%projects with issues (with exceptions): 243 
+%projects with stacks: 158  -  stacks: 3559 (defect)
+%Github: issues: 50342 - stacks: 3865 - projects: 2542 - projects with stacks: 293
+
 %To guide our analysis, we compiled a set of best practices on how to use
@@
 
-   \item  A multitude of programming mistakes were reported as the main causes of stack traces:
-  41,1\%  (12,625)  of reported stack traces were caused by programming mistakes; and 26,31\% of analyzed stack traces reported a java.lang.NullPointerException as root cause;
-
-   \item 1,671 projects (44,47\%) contained at least one issue reporting a java.lang.NullPointerException as the root cause of a stack trace.
-
-   \item Undocumented runtime exceptions signaled by third party
-    code (i.e., libraries and Android platform).
-This study found 118 distinct methods defined on third party code and  reported as the root signalers of runtime exceptions.
-Only 1 of such methods (0,8\%) documented the explicitly thrown runtime exception in a Javadoc comment; and none of these methods included the exception as part of the exception interface (i.e., using throws clause).
-
    \item  Unexpected exception wrappings (e.g., Error wrapping checked and runtime
@@
 
-   \item  The development model of Android platform (strongly based on system call-backs) encouraged 
-     the use of general RuntimeExceptions.
-
-  %\item  Issues reporting uncaught checked exceptions.
+   \item Undocumented runtime exceptions signaled by third party  code (i.e., libraries and Android platform). 
+This study found 118 distinct methods defined on third party code and  reported as the root signalers of 
+runtime exceptions. Only 1 of such methods (0,8\%) documented the explicitly thrown runtime exception 
+in a Javadoc comment; and none of these methods included the exception as part of the exception 
+interface (i.e., using throws clause).
+
+   \item Undocumented checked exception signaled by native Android library. This finding points to a fragility 
+on the way exceptions on JNI interface funcions, which can lead to uncaught exceptions that are difficult do 
+debug.
+
+  \item  A multitude of programming mistakes were reported as the main causes of stack traces:
+ XXX  of reported stack traces were caused by programming mistakes; and XXX of analyzed stack traces reported a java.lang.NullPointerException as root cause;
 
@@
 
-Although the study focused on Android opensource applications, some of the results are not restricted to this context, 
+Such fragilities represent obstacles to the development of dependable mobile apps as well as other Java-based systems
+as they share the same exception model. Although the study focused on Android opensource applications, some of the results are not restricted to this context, 
 for instance: some of the unexpected wrappings were performed by Java libraries and elements (e.g. reflection library and 
-the static blocks); and the undocumented runtime exceptions were also found in Java libraries beyonds the Android plaftorm.
-The study results are relevant to: (i) Android and Java developers to familiarize with the most common reported exceptions 
-mis-uses which is the first step to help developers to avoid making them; (ii) designers of languages to consider ways of reducing the abundance of NullPointerExceptions; and (iii) tool designers to consider developing tools that enable a cross-project defect analysis.
+the static blocks); and the undocumented runtime exceptions were also found in Java libraries besides the Android plaftorm.
+The study results are relevant to: (i) Android and Java developers to familiarize common characteristics and fragilities of exception-related code 
+which is the first step to help developers preventing and dealing with them; (ii) designers of languages and tools to consider ways of reducing 
+such fragilities in Android and Java environments.
 
@@
     analysis.
-
-  \item  It presents the the exploratory study methodology to ..... (complete - view the reviewer comment)
 

### 2014-08-29 android-stacks.tex
changes along the paper

@@
 
-Such fragilities represent obstacles to the development of dependable mobile apps as well as other Java-based systems
-as they share the same exception model. Although the study focused on Android opensource applications, some of the results are not restricted to this context, 
+Such fragilities represent obstacles to the development of dependable java-based applications as they make difficult and 
+sometimes almost impossible to protect the code against the Uncaught Exceptions that lead to system crashes.
+Although the study focused on Android opensource applications, some of the results are not restricted to this context, 
 for instance: some of the unexpected wrappings were performed by Java libraries and elements (e.g. reflection library and 

### 2014-08-29 android-stacks.tex
more fine grained in sec 2

@@
 which also contains  a core Java library, a set of libraries third-party libraries (e.g.  SQLite),
-the Dalvik runtime environment, and the Application Framework (which provides higher-level services to the applications
-running on the platform).
+the Dalvik runtime environment, and the Application Framework (which provides 
+higher-level services to the applications running on the platform) ~cite{andguide}.
+Android platform is mostly implemented in Java and uses JNI[ref] to access native C/C++ code. 
+And as mentioned before, it reuses the embedded Java Exception model to detect and handle 
+abnormal computation states.
 
@@
 automatically by the Android framework (by inversion of
-control [ref]), in response to infrastructure and application events.
+control), in response to infrastructure and application events.
 
 Due to the high demand for responsiveness in mobile environment (mobile apps can easily become 
-unresponsivedue due to resurce limitation and excessive network access [shimidit and arie]) the
-Android Application Framework also provides APIs to support concurrent developement (e.g. AsyncTask).
-
-Android platform is mostly implemented in Java, uses JNI[ref] to access native C code, and 
-reuses the embedded Java Exception model to detect and handle abnormal computation states.
+unresponsivedue due to resurce limitation and excessive network access) the
+Android Application Framework also provides APIs to support concurrent 
+developement (e.g. AsyncTask).
 

### 2014-08-30 android-stacks.tex
study description

@@
 
-  \item  It performs a large scale post morten analysis of exception stack traces of Android projects.
+  \item  It performs a large scale post morten analysis of exception stack traces reported on Android projects hosted in GitHub and Googlecode.
 
@@
 The ramaining of this paper is organized as follows. Section 2 presents a
-background on Android Plaform and Java exception handling mechanism. 
+background on Android Plaform and Java exception model. 
 Section 3 presents the study design. Section 4 reports study findings. 
@@
 Section 6 presents the threats to validity associated to this study. Finally Section
-7 describes the related work, and Section 8 presents our conclusions and
+7 describes the related work, and Section 8 concludes the paper and outlines
 directions for future work.
@@
 Among the basic elements that comprise an Android application, we can cite the
-Activity. Each application usually defines one Activity per application Window.
+Activity. Each application usually defines one Activity per application window.
  Each Activity defines a set of hook methods  (e.g., OnCreate(), onPause()) that are dispatched
@@
 
-Due to the high demand for responsiveness in mobile environment (mobile apps can easily become 
-unresponsivedue due to resurce limitation and excessive network access) the
-Android Application Framework also provides APIs to support concurrent 
-developement (e.g. AsyncTask).
+The Android Application Framework also provides several supporting APIs, such as 
+the APIs to support concurrent developement (e.g. AsyncTask). 
+Such APIs particularly useful due to the high demand for responsiveness 
+in mobile environment (mobile apps can easily become 
+unresponsivedue due to resurce limitation and excessive network access).
 
@@
 
-\subsection{Exception Types in Java} \label{sec:extypes}
+\subsection{Exception Model in Java} \label{sec:extypes}
 
@@
  To look for the handler on the invocation stack is claimed to increase software reusability, 
-since the invoker of an operation can handle it in a wider context [ref-miller].
-
-INCLUIR REF:
-Miller, R.; Tripathi, A.: Issues with Exception Handling in Object-Oriented Systems. In:
-ECOOP’97. LNCS, vol. 1241, Springer (1997) 85–103.
+since the invoker of an operation can handle it in a wider context ~/cite{miller1997issues}.
 
@@
 mistakes} as carefully as checked exceptions. Doing so, the clients of a method
-will be aware of all the exceptions the method can throw. If the developer fails to
-do follow this practice (specially when developing library code) it will be
-difficult or even impossible for the caller to make effective use of such 
-method~\cite{wirfs2006toward, bloch2008effective} and protect its code
+will be aware of all the exceptions the method can throw and can protect its code
 against unforeseen exceptions.
+
+If the developer fails to do follow such practices (specially when developing 
+library/reusable code) it will be difficult or even impossible for the caller to 
+make effective use of such method~\cite{wirfs2006toward, bloch2008effective},
+and protect the system against the uncaught exceptions - one of the main causes of 
+systems crashes.
 
@@
 can reveal whether these practices have been obeyed.
+
 %Next section discusses the
@@
 
-This study focus on the analysis of crash information (related to uncaught exceptions) of a family of Android applications
-aiming at  extracting their common characteristics and looking for similar vulnerabilities on applications and underlaying platform.
-
-In Java when an application crashes dut to an uncaught exception ... 
-Exception Stach trace is usually (procurar por artigo).
-
-The goal of this study was to extract exception stack 
-traces embedded on issues of Android projects (hosted in GitHub and GoogleCode), looking for the 
-common characteristics of such stack-trace and investigating whether
- they could reveal common defficiencies on the exception handling code of both the applications and the underlaying 
- framework.
+%This study focus on the analysis of crash information (related to uncaught exceptions) of a family of Android applications
+%aiming at  extracting their common characteristics and looking for similar vulnerabilities on applications and underlaying platform.
+%In Java when an application crashes dut to an uncaught exception ... 
+%Exception Stach trace is usually (procurar por artigo).
+
+The goal of this study was to extract exception stack traces embedded on issues of Android projects 
+(hosted in GitHub and GoogleCode), looking for its 
+common characteristics and investigating whether
+ they could reveal common fragilities on the exception code of both the 
+applications and the underlaying  framework.
 
@@
 source code are not publicly available. The Android open-source apps were also 
-target of other research studies [fse2013][ahimed].
-
-Falar da assumption que stacj trá relacionada a crash?
-This study is built on the following assumption: the stack traces reported on issues contain 
-relevant information related to application crashes.
-Several studies and techniques have also based on this assumption (e.g.,~\cite{sinha2009fault}~\cite{dhaliwal2011classifying}~\cite{kim2013predicting}).
-
-\subsection{Research Questions}
+target of other research studies ~\cite{Linar13,ahimed}.
+
+% Falar da assumption que stacj trá relacionada a crash?
+%This study is built on the following assumption: the stack traces reported on issues contain 
+%relevant information related to application crashes.
+Several studies and techniques have also based on this assumption %(e.g.,~\cite{sinha2009fault}~\cite{dhaliwal2011classifying}~\cite{kim2013predicting}).
+
+%\subsection{Research Questions}
 In the context of our study we formulated the following reseach questions:
@@
 
-  \item[RQ1] \noindent\emph{What are the common characteristics of exception stack traces reported on issues?} Some of the common characteristics to be investigated are the 
- main causes reported on exception stack traces, the elements responsible for signaling such exceptions.
-
-  \item[RQ2] \noindent\emph{What patterns of exception (mis)-use related to exception types and propagation emerge from the exception stack trace analysis? Based on a set of best practices
-compiled in this study, looked for characteristics on stack traces that could pinpoint to practices that were not being obeyed.}
+  \item[RQ1] \noindent\emph{What are the common characteristics of exception stack traces reported on issues?} 
+Some of the common characteristics to be investigated are the exceptions reported as the root cause on
+the exception stack trace, the origin of such exceptions (e.g., the application, a library, the Android platform).
+
+%Which fragilities on the exception-related code emerge from the exception stack trace analysis? 
+%Can such characteristics reveal common fragilities on the exception code of applications or the underlying framework
+
+  \item[RQ2] \noindent\emph{Which patterns of exception (mis)-use emerge?}
+Based on a set of best practices compiled in this study, the study investigates whether characteristics on exception stack traces  
+can pinpoint best practices violations. Chraracteristics of exception stack traces and detected violations can reveal fragilities
+ on the exception handling code, which threats the application dependability. }
 
@@
 
-The data used in our study to answer our research questions are: the issues related to each 
-Android project available in each repository; the source code and manifest file of each application;
-the bytecode of all exceptions defined and reused by Android platform.
+To answer the research questions the following information is needed: (i) the issues of reported on every Android project hosted on 
+GitHub and Googlecode; (ii) the stack traces embedded on those issues; (iii) the type of the exceptions
+ reported on the stack traces (e.g., error, runtime, checked); (iv) the origin of such exceptions 
+(e.g., the application, a library, the Android platform). Next sections describes how such information 
+was retrieved from GitHub and Googlecode, how the exceptions stack traces were extracted and distilled,
+ and how the type associated to each exception and its origens was identified.
+
+%mentioned 
+%The data used in our study to answer our research questions are: the issues related to each 
+%Android project available in each repository; the source code and manifest file of each application;
+%the bytecode of all exceptions defined and reused by Android platform.
 
@@
 
-As mentioned before, this study amined at mining information available on GitHub issues,
-more specifically, extracting and distilling the stack traces embedded on GItHub issues. 
-Issues on Github are different from issues on dedicated bug tracking tools such as 
+%As mentioned before, this study amined at mining information available on GitHub issues,
+%more specifically, extracting and distilling the stack traces embedded on GItHub issues.
+
+This study used the dataset provided by the GHTorrent project~\cite{Gousi13}, 
+an off-line mirror of the data  offered through the Github API.  
+To identify Android projects, we performed a case insensitive search for the
+term \textsf{android} in the repository's names and short descriptions.  
+Up to 23 Feb 2014,  when we queried GHTorrent, such heuristic could
+ find 2.542 repositories,  from which 589 apps had at least one
+issue containing a stack trace.
+
+Then we performed further cleanup, inspecting the site of every Android
+reporting at least one stack trace, to make sure that they represented real
+mobile apps. During this clean up 106 apps were removed because they were either
+example projects (i.e., toy projects) or tools to support Android development
+(e.g. \textsf{selendroid}, \textsf{roboeletric} --- tools to support the testing of Android apps).
+The filtered set consisted of 482 apps.
+
+This set of 482 projects contained overall 32,582 issues from which 4,208 stacktraces 
+were extracted. Issues on Github are different from issues on dedicated bug tracking tools such as 
 Bugzilla and Jira. The most important difference is that there are no predefined fields
@@
 if an stack trace information embedded in an issue, it contain relevant information
- concerning the exception use -  regardless the lables associated to the issue.
+that can reveal violation to best practices - regardless the lables associated to the issue.
 
 Moreover, issues and pull requests are dual on Github; all pull requests have a corresponding 
-``backing'' issue which
-are automatically generated. Therefore, we excluded such automatically generated
+``backing'' issue which are automatically generated. Therefore, we excluded such automatically generated
 issues from our analysis. Finally, the Github API allows the automated
@@
 
-RESUMIR:
-This study used the dataset provided by the GHTorrent project~\cite{Gousi13}, an off-line mirror of the data 
-offered through the Github API.  GHTorrent has been collecting data from Git since 
-February 2012. Up to Dec 2013, and when we queried GHTorrent, it included 332,864 non-fork Java
-repositories \footnote {A Java project is specified in GitHub with a
-specific label.}  from which only 44,323 have contained at least issue report in their lifetime.
-From this set only 16,837 were watched by external users.
-To ensure that our work targets projects that are openly used by
-users other than their developers, we focused our analysis on just those projects.
-
-From this set of 16.836 Java projects which overall contained 356.057 issues, we retrieved the metadata related to each Java project (e.g., repository's names and short descriptions), and the content of every issue and issue comment and stored in a relational database.
-Using the heuristics presented in Section~\ref{sec:android}, we selected 482 repositories featuring Android projects on which we perform manual inspection was performed.
-
-As mentioned before, in this study we selected a subset of Java projects to
-be analyzed in more detail. The subset chosen was comprised by the Android apps
-included on GitHub (until 23 Feb 2013).
-To identify Android projects, we performed a case insensitive search for the
-term \textsf{android} in the repository's names and short descriptions.  The
-heuristic filtered 2.542 repositories, from which 589 apps had at least one
-issue containing a stack trace.
-
-Then we performed further cleanup, inspecting the site of every Android
-reporting at least one stack trace, to make sure that they represented real
-mobile apps. During this clean up 106 apps were removed because they were either
-example projects (i.e., toy projects) or tools to support Android development
-(e.g. \textsf{selendroid}, \textsf{roboeletric} --- tools to support the testing of Android apps).
-The filtered set consisted of 482 apps, from which approximately 50\% are also
-hosted in Google Market Place.
-
-This set of 482 projects contained overall 32,582 issues from which 4,208 stacktraces 
-were extracted. From this set of issues 44,9\% were labled as defect.
-
 \subsubsection{Android Apps in GoogleCode}
 
-Until January 2014 there were 788 android projects available in Google code repository from which 724 defined at least 1 issue and 183 projects 
-had at least one issue including a stack trace. To identify the Android projects on Googlecode we used the search facilities provided looking for 
-projects mentioning "Android" on the name of the project or its description. The repository page of each project was inspected, and from this set
-142 corresponded to Android applications.
-
-To mine the information available on such Google code projects we implemented a web crawler which navegated on each project and issue page
-and extracted the content of each issue to store on relational database.
+Different from GitHub, Googlecode does not provide an API to access the information related
+ to hosted projects ~\footnote{Googlecode previoulsly provided a webservice but was deacrivated in June 2013.}.
+However many well used open source Android application such as K9mail, is hosted in Googlecode which stores
+the issue history of this and other apps are store there ~\footnote{K9mail moved to Github but advice their users to create issues on Googlecode as a way of not loosing the issue history.}.
+
+To overcome this limitation and mine the information available on such Googlecode 
+projects we implemented a Web crawler which navegated (incorporated in Exception Miner tool described next) 
+ on each project and issue page, extracting the content of each issue to store on relational database.
 \footnote{Google code previoulsly provided a webservice to extract information about each project, but this service was removed in what Google called "clean-up action" which happened in June 2013.}
+
+To overcome this limitation we needed to implement a Web Crawler (incorporated in Exception Miner tool described next) that navegated 
+through the web interface of Googlecode projects extracting all issues and issue comments and storing in a relational database for later analysis.
+
+To identify Android projects in Googlecode, we performed a similar heuristic, we performed a case insensitive search 
+(on Googlecode search interface) for the term \textsf{android}.  On January 2014, when we queried Googlecode, such heuristic could
+ find 788  projects, from which 724 defined at least 1 issue and, 183 projects 
+had at least one issue including a stack trace. Then we performed further cleanup, inspecting the site of every Android
+reporting at least one stack trace, and as a result we could identify 158 
+Android projects.
 
@@
 
-To extract exceptions from issues, we implemented ExceptionMiner, a modular
-mining tool able to connect to various repositories (such as Google Code,
+To support this study we developed ExceptionMiner, a modular mining tool able 
+to connect to various repositories (such as Google Code,
 GHTorrent or even directly to Bugzilla), extract issues, mine stack traces from
-them and classify them in predefined categories. The main components of
+them and distill exteption stack trace information. The main components of
 ExceptionMiner are as follows:
 
-\noindent\emph{Exception StackTrace Miner} The first step in the process is
+\noindent\emph{Repository Connectors} The first step in the process is
 mining references to exceptions and stack traces embedded in issues, d
@@
 
-\noindent\emph{Exception Classifier} The next step in the ExceptionMiner process
-is classifying exceptions as either checked exceptions, runtime exceptions or 
-errors. The exception classifier is based on bytecode analysis (based on Design
-Wizard~\cite{Brunet09}) that walks up the type hierarchy of a Java exception
-until it reaches a base exception type.
-
-\noindent\emph{Exception Signaler Classifier} After its initialization it
-classifies each signaler according to its origin (i.e. Android Plaform,
-Aplication, Library, Libcore, or Java) based on pattern matching between the
-signaler name and the packages associated to each category.
+\noindent\emph{Exception Classifier} To support a deeper investigation of the stack traces 
+every exception defined on a stack trace need to be classified according to its type
+(e.g. Error, Exception or RuntimeException). The exception classifier is based on 
+bytecode analysis (based on Design Wizard~\cite{Brunet09}) that walks up the 
+type hierarchy of a Java exception until it reaches a base exception type.
+
+\noindent\emph{Exception Signaler Classifier} 
+This module is responsible for classifying each signaler according 
+to its origin (i.e. Android Plaform, Aplication, Library, Libcore, or Java) 
+based on pattern matching between the signaler name and the packages associated to each category.
+TELL THAT ALL PACKAGES COMPOSISNG APPS .... WERE RECOVED....
+
+\textbf{Step 4.}  Bytecode-Code analysis of JVM and Android Exceptions.
+We developerd a bytecode analysis tool based on DesignWizard to discover all exceptions thrown by the last version of SDK version 7, and Android Platform and OS (level 19).
+Every exception was extracted and classified as Runtime, Checked or Error based on the analysis of the exception hierarchy.
+Our analysis found XXX runtime exceptions in JVM and XX checked, on Android xxx checked and YYY runtime.
+
+\textbf{Step 5.}  Source-Code Analysis of Android Projects. 
+For each one of the analyzed projects, we
+downloaded the source code and using custom scripts, we extracted their package
+hierarchy (recursive list of of Java packages) and their custom exception types.
+
+\textbf{Step 6.}  Manual Inspection.
+XXXX of the exceptions reported on stack traces  could not be found on JMV, SO or inside the last version of the project.
+Based on the full name of the exception, a manual inspection was performed on the artifacts associated to the project and libraries used by the project.
+When even on those sources the exception was not foung [grep-code and similar sites] the exception, in order to discover the exception type.
+Only 3 exceptions remained undefined and were removed from this study.
 
@@
 
-\subsubsection{Data Filtering and Manual Inspections}
-
+\subsubsection{The Minig Process}
+
+{MOSTRAR UMA FIGURA COM TODOS OS PASSOS E RESSALTAR OS PASSOS MANUAIS}
 COMPLETE HERE.....
@@
 conducted to answer these questions. The mixed-methods approach is composed by the following steps:
-
-%\begin{figure*}
-%\centering
-%\includegraphics[width=\hsize]{overview.pdf}
-%\caption{Overview of the study based on mixed approaches.}
-%\label{fig:overviewfig}
-%\end{figure*}
-
-%The GHTorrent dataset covers a broad range of
-% development activities on Github, including the all the issues per project. Such data 
-%is stored in unprocessed format, in a MongoDB database, while metadata is extracted 
-%and stored in a relational database.
-
-\textbf{Step1.} To obtain the Dataset to be analized, we queried GHTorrent as detailed in Section XXX. From GHTorrent we retrieved the metadata related to each Java project (e.g., 
-repository's names and short descriptions), and the content of every issue and issue comment 
-and stored in a relational database.
-
-\textbf{Step 2.} Extracting Exception Information from Issues. In this step, we used the ExceptionMiner tool,
- developed in this work, to extract the exception names and exception stack traces of each
- issue and issue comment - analyzing the information of each project at a time. This initial 
-set of analyzed Java projects included libraries, web-based, Android and stand-alone applications.
-
-\textbf{Step 3.} Restricting the Dataset to Android Projects. We select a subset of all Android 
-projects available on the dataset to perform a deeper analysis (which combined mined
- information extrated from issues with source-code analysis and manual inspection 
-of projects resources described on next steps). This decision was driven by the
-fact that most open source Anedroid applications are known for dealing
-with several sources of exceptions due to multi-threaded execution and
-interaction with input/output resources; moreover they share a common
-well-defined platform and a limited set of libraries.
-To identify Android projects, we performed a case insensitive search for 
-the term \textsf{android} in the repository's names and short descriptions.  The
-heuristic filtered 2.542 repositories, from which 589 apps had at least one
-issue containing a stack trace. Then we performed further cleanup, inspecting the site of every Android
-reporting at least one stack trace, to make sure that they represented real
-mobile apps. During this clean up 106 apps were removed because they were either
-example projects (i.e., toy projects) or tools to support Android development
-(e.g. \textsf{selendroid}, \textsf{roboeletric} --- tools to support the testing of Android apps).
-The filtered set consisted of 482 apps, from which approximately 50\% are also
-hosted in Google Market Place.
-
-To support a deeper investigation of the stack traces of Android apps, every
-exception defined on a stack trace was then classified according to its type
-(e.g. Error, Exception or RuntimeException) and to its signaler. To do so, the following 
-steps were performed:
-
-\textbf{Step 4.}  Bytecode-Code analysis of JVM and Android Exceptions.
-We developerd a bytecode analysis tool based on DesignWizard to discover all exceptions thrown by the last version of SDK version 7, and Android Platform and OS (level 19).
-Every exception was extracted and classified as Runtime, Checked or Error based on the analysis of the exception hierarchy.
-Our analysis found XXX runtime exceptions in JVM and XX checked, on Android xxx checked and YYY runtime.
-
-\textbf{Step 5.}  Source-Code Analysis of Android Projects. 
-For each one of the analyzed projects, we
-downloaded the source code and using custom scripts, we extracted their package
-hierarchy (recursive list of of Java packages) and their custom exception types.
-
-\textbf{Step 6.}  Manual Inspection.
-XXXX of the exceptions reported on stack traces  could not be found on JMV, SO or inside the last version of the project.
-Based on the full name of the exception, a manual inspection was performed on the artifacts associated to the project and libraries used by the project.
-When even on those sources the exception was not foung [grep-code and similar sites] the exception, in order to discover the exception type.
-Only 3 exceptions remained undefined and were removed from this study.
 

### 2014-08-30 android-stacks.tex
study description more changes

@@
 
-\begin{figure} \centering \includegraphics[scale=0.8]{arch3.png}
-  \caption{Android Archtecture} \label{fig:exchier} \end{figure}
+%\begin{figure} \centering \includegraphics[scale=0.8]{arch3.png}
+%  \caption{Android Archtecture} \label{fig:exchier} \end{figure}
 
@@
 can pinpoint best practices violations. Chraracteristics of exception stack traces and detected violations can reveal fragilities
- on the exception handling code, which threats the application dependability. }
+ on the exception handling code, which threats the application dependability.
 
@@
 them at will. Therefore, an issue may have none ore an arbitrary set of labels depending 
-on which repository it was created. Therefore, this study is based on the assumption 
-if an stack trace information embedded in an issue, it contain relevant information
-that can reveal violation to best practices - regardless the lables associated to the issue.
+on which repository it was created.
+
+Based on the assumption that regardless the issue type, every exception stack
+trace contains relevant information concerning the exception structure of the
+projects analyzed - and therefore can review fragilities on the exceptio related code -  
+we opted for not restricting the analysis only on defect
+issues.\footnote{We conducted the same analysis on the defect issues and the top
+exceptions are similar to the ones mentioned on defect issues. Due to space
+limitation we limit to present the general analysis here. More detailed analysis
+on the defect issues can be found at:
+\url{www.dimap.ufrn.br/~roberta/android_repo_mining}}
 
@@
 Different from GitHub, Googlecode does not provide an API to access the information related
- to hosted projects ~\footnote{Googlecode previoulsly provided a webservice but was deacrivated in June 2013.}.
+ to hosted projects \footnote{Googlecode previoulsly provided a webservice but was deacrivated in June 2013.}.
 However many well used open source Android application such as K9mail, is hosted in Googlecode which stores
@@
 
-\section{RQ1: What are the common characteristics of such stack traces?}
+\section{The Study Results}
 
@@
 
-Based on the assumption that regardless the issue type, every exception stack
-trace contains relevant information concerning the exception structure of the
-projects analyzed, we opted for not restricting the analysis only on defect
-issues.\footnote{We conducted the same analysis on the defect issues and the top
-exceptions are similar to the ones mentioned on defect issues. Due to space
-limitation we limit to present the general analysis here. More detailed analysis
-on the defect issues can be found at:
-\url{www.dimap.ufrn.br/~roberta/android_repo_mining}}
-
-\subsection{Characterizing the Most Common Root Exceptions}
+This section presents the results amed at answering the research 
+questions presented before.
+
+\subsection{The Most Common Root Exceptions }
 
@@
 The table also presents the whether the method responsible for signaling the exceptions were
-part of Android, the aplication itself, libcore, or a third-party library as presented in Table  \ref{tab:signalers}  \footnote{9 stack traces contained only elements of JVM and were discarded from the analysis}.
+part of Android, the aplication itself, libcore, or a third-party library as presented in Table  \ref{tab:signalers} 
+ \footnote{9 stack traces contained only elements of JVM and were discarded from the analysis}.
 
@@
 
+\subsubsection{What They can Tell us?}
+
 \subsubsection{Direct instances of RuntimeExceptions being thrown.}
@@
 
-\subsection{Classifying Top 100 Exceptions}
+\subsubsection{Classifying Top 100 Exceptions}
 
@@
 \centering
-\begin{tabular}{ll}
- \bfseries{Method} & \bfseries{Project} & & \bfseries{Ocurrences}\\
+\begin{tabular}{lll}
+ \bfseries{Method} & \bfseries{Project} & \bfseries{Ocurrences}\\
     \hline
@@
 code analysis.
+
+RESSALTAR LIMITACAO DE QUANTIDADE DE FALSOS POSITIVOS... E QUE OS STACKS SAO REAIS...
+REPRESENTAM FLUXOS REAIS.... POST MOSTERN ANALISYS OF REAL EXECUTIONS....
 

### 2014-08-31 android-stacks.tex
study results

@@
 type hierarchy of a Java exception until it reaches a base exception type.
+This module analyzed the exception hierachy of all exceptions thrown on:
+ the last version of SDK version 7, and Android Platform and OS (level 19).
 
@@
 This module is responsible for classifying each signaler according 
-to its origin (i.e. Android Plaform, Aplication, Library, Libcore, or Java) 
-based on pattern matching between the signaler name and the packages associated to each category.
-TELL THAT ALL PACKAGES COMPOSISNG APPS .... WERE RECOVED....
-
-\textbf{Step 4.}  Bytecode-Code analysis of JVM and Android Exceptions.
-We developerd a bytecode analysis tool based on DesignWizard to discover all exceptions thrown by the last version of SDK version 7, and Android Platform and OS (level 19).
-Every exception was extracted and classified as Runtime, Checked or Error based on the analysis of the exception hierarchy.
-Our analysis found XXX runtime exceptions in JVM and XX checked, on Android xxx checked and YYY runtime.
-
-\textbf{Step 5.}  Source-Code Analysis of Android Projects. 
-For each one of the analyzed projects, we
-downloaded the source code and using custom scripts, we extracted their package
-hierarchy (recursive list of of Java packages) and their custom exception types.
-
-\textbf{Step 6.}  Manual Inspection.
-XXXX of the exceptions reported on stack traces  could not be found on JMV, SO or inside the last version of the project.
-Based on the full name of the exception, a manual inspection was performed on the artifacts associated to the project and libraries used by the project.
-When even on those sources the exception was not foung [grep-code and similar sites] the exception, in order to discover the exception type.
-Only 3 exceptions remained undefined and were removed from this study.
-
-\subsubsection{The ExceptionMiner Tool}
-
-Before investigating the research questions listed above we analyzed the prevalence of stack traces 
-on issues created on Github. Prevalence or prevalence proportion, a term often used in epidemiology,
-represents the proportion of a population found to have a particular condition
-(e.g. a disease) over the total number of the studied population. In our work, we used
-this term to evaluate a condition (i.e., presence of a stack trace) over issues
-reported on GitHub repositories.
-
-Using the ExceptionMiner tool, developed in this work, we could extract the exception stack traces defined on issues and issue comments.
-Executing ExceptionMiner tool each issue and issue comment defined on the 16,836 projects considered in this study, we could observe that 3,758 projects (22,32\%) contained at least 1 issue on which a stack trace was found.
-Overall the number of issues on which a stack trace was found was 21,013 from which 28,800 stack traces were extracted and analyzed - some issues defined more than one stack trace.
-
-We then analyzed the lables associated to such issues, and we discovered that only 5,196 were labled as defect ~\footnote{the dedect lables considered in this work were every lable containing as substring: defect, bug, crash, failure, fail, exception} (which represent only 24,7\% of the issues on which a stack trace was found), most of the issues on which stack traces were found contained no lable.
-Such observation supports our assumption to consider not only the issues related with defect issues on this analysis.
-
-\subsubsection{The Minig Process}
-
-{MOSTRAR UMA FIGURA COM TODOS OS PASSOS E RESSALTAR OS PASSOS MANUAIS}
-COMPLETE HERE.....
-
-\subsubsection{Database of Exception Stack Trace Information}
-\subsubsection{Distiling Exception Stack Trace Information}
-
-\- root exception
-\- signaler
-\- method
-\- class
-\- package
-\- Exception
-
-\subsection{Process}
-
-ESPALHAR ESTA SECAO NAS OUTRAS.
-
-Figure~\ref{fig:overviewfig} presents an overview of the mixed-methods approach
-conducted to answer these questions. The mixed-methods approach is composed by the following steps:
+to its origin (i.e. Android Plaform, Android Libcore, Aplication, Library). 
+We downloaded the source code of Android Platform and OS (level 19)
+and this module extractes the packages hierarchies (recursive list of 
+of Java packages), that composes it. To discover the packages that compose 
+each application, this module extracted the manifest files of each Android apps
+ - which defines the main packages that compose the applications.
+Then based on pattern matching between the signaler name and the packages 
+ this module identifies the origin of the exception signalers.
+The exceptions were considered from libraries if their packages were not defined 
+on Android Platform, core libraries nor on applications.
+Table ~ref{signalers} summarizes the signaler classification.
+
+\begin{table}
+  \centering
+  \begin{tabular}{rp{29em}}
+    \hline
+    \bfseries{Signaler} & \bfseries{Description} \\
+    \hline
+    \bfseries{android} & If the exception is thrown by a method defined in Android Platform, or in a JDK library used by i.\\
+    \bfseries{app}     & If the exception is thrown by an application method or in a JDK library used by it.\\
+    \bfseries{libcore} & If the exception is thrown by one of the core libraries reused by Android (i.e., org.apache.harmony, org.w3c.dom, sun.misc, org.apache.http, org.json, org.xml). \\
+    \bfseries{lib}     & If the exception is thrown by a method that was not defined by any of the elements above.\\
+    \hline
+  \end{tabular}
+  \caption{Sources of exceptions in Android}
+  \label{tab:signalers}
+\end{table}
+
+The information extracted by each module was structured in a Database,
+which contained the distilled information of stack traces, exception types,
+and signalers origin (e.g., root exception and its type, root signaler and its origin,
+exception wrapper type, wrapper signaler).
+
+%\subsubsection{The ExceptionMiner Tool}
+%Using the ExceptionMiner tool, developed in this work, we could extract the exception stack traces defined 
+%on issues and issue comments. Executing ExceptionMiner tool each issue and issue comment defined on the 
+%16,836 projects considered in this study, we could observe that 3,758 projects (22,32\%) contained at 
+%least 1 issue on which a stack trace was found. Overall the number of issues on which a stack trace was
+% found was 21,013 from which 28,800 stack traces were extracted and analyzed - some issues defined 
+% more than one stack trace.
+
+%We then analyzed the lables associated to such issues, and we discovered that only 5,196 were
+% labled as defect ~\footnote{the dedect lables considered in this work were every lable containing 
+%as substring: defect, bug, crash, failure, fail, exception} (which represent only 24,7\% of the
+% issues on which a stack trace was found), most of the issues on which stack traces were found 
+%contained no lable. Such observation supports our assumption to consider not only the issues 
+% related with defect issues on this analysis.
+
+\subsubsection{Manual Inspections}
+This study was also based in a set of manual inspection tasks. Besides the 
+ identification of Android projects detailes previously. 
+Manual inspections were used discover the type of exceptions thrown by 
+the applications and libraries we manually inspect the sourcecode or javadoc
+ documentation of each of them. When the exception was removed from the 
+last app version, the previous versions were analyzed. When the exception was
+ not found on the previous versions, the exception was discarded.  Only 3 exceptions 
+remained undefined and were removed from this study.
+
+%{MOSTRAR UMA FIGURA COM TODOS OS PASSOS E RESSALTAR OS PASSOS MANUAIS}
+%COMPLETE HERE.....
+%\subsubsection{Database of Exception Stack Trace Information}
+%\subsubsection{Distiling Exception Stack Trace Information}
+%\subsection{Process}
+%ESPALHAR ESTA SECAO NAS OUTRAS.
+%Figure~\ref{fig:overviewfig} presents an overview of the mixed-methods approach
+%conducted to answer these questions. The mixed-methods approach is composed by the following steps:
 
@@
 
-\subsection{The Most Common Root Exceptions }
-
-After distilling the information available on the XXXX  stack traces, extracted from the 
-issues and issues comments of XXXX Android projects on GitHub and Googlecode,  
-we found XXXX  distinct exceptions reported as the root causes of stack traces.
-Table \ref{tab:toptenandroid} illustrates the top 10 root couses found in the study ranked by the number of distinct
-project repositories on which they were reported.
+\subsection{The Common Root Exceptions and What They can Tell us? }
+
+After distilling the information available on the exception stack traces we could find 
+the the exceptions commonly reported as the root causes of stack traces.
+Table \ref{tab:toptenandroid} illustrates the top 10 root couses found in the study,
+ ranked by the number of distinct project repositories on which they were reported. 
+Thes table shows: the number of times each exception was reported as 
+the root cause of a stack trace, the number of distinct projects reporting such causes on stack traces, 
+and the percentages of such in relation to all reported exceptions and analyzed projects respectaly.
+It also presents the whether the method responsible for signaling the exceptions were
+part of Android, the aplication itself, libcore, or a third-party library as presented in Table  \ref{tab:signalers} 
+ \footnote{9 stack traces contained only elements of JVM and were discarded from the analysis}.
+
+%For each exception we could discover the number of times it was reported as the root caused on exception stack traces,
+%the number of distinct projects on which such stacks were defined (and consequently its
+%popularity) and the of times it was signaled by different signaler types. We
+%could calculate the popularity of each root cause, considering the number of
+%distinct projects they were found and the whole set of projects analyzed with at
+%least one stack trace (482). Table~\ref{tab:toptenandroid} 
+%and Figure~\ref{fig:androidsignaler} presents the mined data.
 
@@
 
-Table \ref{tab:toptenandroid} illustrate the number of times each exception was reported as 
-the root cause of a stack trace, the number of distinct projects reporting such causes on stack traces, 
-and the percentages of such in relation to all reported exceptions and analyzed projects respectaly.
-The table also presents the whether the method responsible for signaling the exceptions were
-part of Android, the aplication itself, libcore, or a third-party library as presented in Table  \ref{tab:signalers} 
- \footnote{9 stack traces contained only elements of JVM and were discarded from the analysis}.
-
-\begin{table}
-  \centering
-  \begin{tabular}{rp{29em}}
-    \hline
-    \bfseries{Signaler} & \bfseries{Description} \\
-    \hline
-    \bfseries{android} & If the exception is thrown by a method defined in Android Platform or OS, or in a JDK library used by them.\\
-    \bfseries{app}     & If the exception is thrown by an application method or in a  DK library used by it.\\
-    \bfseries{libcore} & If the exception is thrown by one of the core libraries reused by Android (i.e., org.apache.harmony, org.w3c.dom, sun.misc, org.apache.http, org.json, org.xml). \\
-    \bfseries{lib}     & If the exception is thrown by a method that was not defined by any of the elements above.\\
-    \hline
-  \end{tabular}
-  \caption{Sources of exceptions in Android}
-  \label{tab:signalers}
-\end{table}
-
-For each exception we could discover the number of times it occurred on exception stack traces,
-the number of distinct projects on which such stacks were defined (and consequently its
-popularity) and the of times it was signaled by different signaler types. We
-could calculate the popularity of each root cause, considering the number of
-distinct projects they were found and the whole set of projects analyzed with at
-least one stack trace (482). Table~\ref{tab:toptenandroid} 
-%and Figure~\ref{fig:androidsignaler}
-presents the mined data.
-
-Programming mistakes are the most common root causes for most types of signalers. We
-could observe that the NullPointerException is still the exception with higher
-number of occurrences. The NullPoiterExceptions are mainly signaled inside
-Android platform and inside application code, although we also find
+INSERIR EM ALGUM PONTO DO TEXTO:
+
+We can observe that most of this set is composed by exceptions defined by java.lang,
+which are mostly not expliciltly thrown by the developer but include
+the exceptions implicitly thrown by the JVM due to programming errors 
+(e.g., out-of-bounds array index, division-by-zero, access to a null reference)
+ or resource limitation (e.g., OutOfMemoryError).
+For such exceptions, which represent programming bugs or resource limitations, 
+there is usually no proper handling besides presenting an error message to
+ the user and restarting the application. Only high fault tolerant systems need to 
+provide solutions to handle them. Hence, the best strategy to deal with them
+is try to prevent them.
+
+\subsubsection{The high number of NullPointerExceptions}
+From this set the java.lang.NullPointerException was the most reported root cause; 1225 stack traces 
+had a java.lang.NullPointerException  as root cause, which corresponds of 30,08\%  of 
+all analyzed stack traces. Considering the number of distinct project repositories, 
+the java.lang.NullPointerException, was reported on at least one issue of 99 projects 
+(which represents 20,54\% of all projects reporting at least one stack trace).
+
+ The NullPoiterExceptions are mainly signaled inside Android platform and inside 
+application code, although we also find
 NullPointerExceptions being signaled from third-party libraries. Regarding
 reusable code, there is no consensus whether it is a good or bad practice to 
-re-throw
-a NullPointerException. Some prefer to encapsulate such an exception on
+re-throw a NullPointerException. Some prefer to encapsulate such an exception on
 IllegalArgumentException, while others~\cite{bloch2008effective} argue that the
@@
 should not be wrapped.
-
-We could observe in this study that the NullPointerException and other
-other implicitly signaled exceptions represented most of the exceptions reported
-on issues in both Java and Android subset. For such exceptions, which represent
-programming bugs or resource limitations, there is usually no proper handling
-besides presenting an error message to the user and restarting the application.
-Only high fault tolerant systems need to provide solutions to handle them.
-
-----------
-
-We can observe that most of this set is composed by exceptions defined by java.lang,
-which are usually implicitly thrown by JMV dur to a programming mistake. From thi set
-the java.lang.NullPointerException  was the most reported root cause;  7578 stack traces 
-had a java.lang.NullPointerException  as root cause, which corresponds of 26,31\%  of 
-all analyzed stack traces. Considering the number of distinct project repositories, 
-the java.lang.NullPointerException, was reported on at least one issue of 1,671 projects 
-(which represents 44,47\% of all projects reporting at least one stack trace).
 
@@
 
-\subsubsection{What They can Tell us?}
-
 \subsubsection{Direct instances of RuntimeExceptions being thrown.}
@@
 the exception message which can be neither complete nor precise~\cite{gosling2000java}.
-
-INSERIR EM ALGUM PONTO DO TEXTO:
-The exceptions available on exception stack traces are not limited to the exceptions 
-expliciltly thrown by the Java developer but also include
-the exceptions implicitly thrown by the JVM due to programming errors (e.g., out-of-bounds array index, division-by-zero, 
-access to a null reference) or resource limitation (e.g., OutOfMemoryError).
 
@@
 types that can occur in real world Java projects.
-
-TODO: SEPARAR JAVA DE ANDROID
 
@@
 
-Programming errors are the causes of most of the exception stack traces reported
-on issues both in  general Java data set and Android subset. In the Android subset we could find 
-more stack traces reporting exceptions related to Backward
-compatibility and security concerns than on the general data set.
-
-TODO: Discuss about the interater agreement here...
+This classification also enable us to find stack traces reporting exceptions 
+directly related to concerns often mentioned on research
+works on Android Plaftorm, such as: Concurrency, Backward compatibility, Security, etc.
+Considering this broader set the programming mistakesstill represent most of the causes (55,64\%)
+of the analyzed exception stack traces.
+
+%TODO: Discuss about the interater agreement here...
 

### 2014-08-31 android-stacks.tex
discussions before wrappings

@@
 
-INSERIR EM ALGUM PONTO DO TEXTO:
-
 We can observe that most of this set is composed by exceptions defined by java.lang,
@@
 
-Throwing general exceptions, such as direct instances of RuntimeException, is  considered a
-bad practice, because the exception type does not carry enough information to identify the
-cause of the exceptional behaviour, and as a consequence developers need to relying on
+Throwing general exceptions, such as direct instances of RuntimeException, violates best
+practice III. It is considered a bad practice because exception type does not carry enough
+ information to identify the cause of the exceptional behaviour, and as a consequence developers need to relying on
 the exception message which can be neither complete nor precise~\cite{gosling2000java}.
@@
 To get a broather view of the root causes of stack traces, we performed an empirical evaluation
-whose goal was to categorized the top 100 root causes.
-
-After a first inspection of the complete list of exceptions reported, a first set of exception 
-categories was defined. This list represent the most common terms reffered on exception
-names and javadoc documentation. A similar approach was addopted in [COELHO-ECOOP].
-The list of categories is illustrated in Table~\ref{tab:categories}. Hence in the second phase
-of the study, each exception was mapped to one of the categories previously defined - 
-this analysis was based on the inspection of source code, Javadoc, or additional 
-documentation related to the exception.
-
-To ensure the quality of the classification, three independent coders classified a randomly selected
-sample of 25 exception types (from the total 100) using the same categories;
-the interater agreement was high (96\%). We can therefore assume that the
-categories are descriptive enough for covering the whole range of exception
-types that can occur in real world Java projects.
+whose goal was to categorized the top 100 root exceptions. The top 100 root causes correspond 
+to approximately 95\% of all exception types reported on the exception stack-issues.
+
+Such categorization was based known sources of faults mobile developement: concurrency ~cite{to_include}
+, backward compatibility ~cite{to_include}, security  ~cite{to_include} and resource management (IO, Memory, Batery) ~cite{to_include}.
+We added to this set categories to represent the exceptions caused by programming mistakes,
+ and general Java exceptions (whose types not convey any specific cause of failure), and 
+exceptions that represent specific  exceptional situations of apps, frameworks or libraries.
+The list of categories is illustrated in Table~\ref{tab:categories}.
 
@@
 out-of-bounds array index, division-by-zero, access to a null reference). \\ \hline
-      Resources (IO, Memory, Energy)  & Exception releated to the manipulation or restriction of resources (e.g., file, network, memory) \\ \hline
+      Resources (IO, Memory, Batery)  & Exception releated to the manipulation or restriction of resources (e.g., file, network, memory) \\ \hline
       Security                               & Exception releated to security issues (e.g. password validation, criptography) \\ \hline
@@
       Backward compatibility                 &  A list of exceptions related to backward compatibility - a complete list can be found here [ref]  \\ \hline
-      Reflection                             & Exceptions related to Java reflection library  \\ \hline
-      Specific                & Exceptions created for specific applications, frameworks or libraries \\ \hline
+      Specific Exceptions              & Exceptions created for specific applications, frameworks or libraries \\ \hline
       General (Exception, Error, Runtime)    & The top level exceptions in Java    \\ \hline
@@
 
-The top 100 root causes correspond to approximately 77\% of all stack traces
-reported reported on Java stack-issues and 95\% of all exceptions reported on
-Android stack-issues. Table~\ref{tab:tophundrend} illustrates the result of such 
-classification.
+Each top 100 root exceptions was mapped to one of these categories - 
+this analysis was based on the inspection of source code, Javadoc, or additional 
+documentation related to each exception.  Table~\ref{tab:tophundrend} 
+illustrates the result of such classification, the number of exception stack traces
+on which the exceptions of each category was found as the root cause.
 
@@
     \hline
-    \bfseries{Category} &  \bfseries{ Ocurrences (\%)} \\
-    \hline
-      Programming logic (java.lang and util) & 2235 (55,64\%)\\
-      Resources (IO)                         &  727 (18,10\%) \\
-      Security                               &  165 (4,11\%)\\
-      Concurrency                            & 116  (2,89\%)\\
-      Backward compatibility                 &  219 (5,45\%) \\
-      Reflection                             &  91 (2,27\%)\\
-      Specific (GUI,FRAMEWORK)               &  197 (4,90\%)\\
-      General (Error, Exception, Runtime)    &  267 (6,65\%)\\
+    \bfseries{Category} &  \bfseries{Occurence on Stacks (\%)} \\
+    \hline
+      Programming logic (java.lang and util) & 2326 (57,9\%)\\
+      Resources (IO)                         &  727 (18,1\%) \\
+      Security                               &  165 (4,1\%)\\
+      Concurrency                            & 116  (2,9\%)\\
+      Backward compatibility                 &  219 (5,5\%) \\
+      Specific Exceptions               &  197 (4,9\%)\\
+      General (Error, Exception, Runtime)    &  267 (6,7\%)\\
     \hline
@@
 
-This classification also enable us to find stack traces reporting exceptions 
-directly related to concerns often mentioned on research
-works on Android Plaftorm, such as: Concurrency, Backward compatibility, Security, etc.
-Considering this broader set the programming mistakesstill represent most of the causes (55,64\%)
-of the analyzed exception stack traces.
+To ensure the quality of the classification, three independent coders classified a randomly selected
+sample of 25 exception types (from the total 100) using the same categories;
+the interater agreement was high (96\%). We could also observe that
+We can therefore assume that the categories are descriptive 
+enough for covering the range of exception types analyzed.
+
+In this second analysis the programming mistakes were also the
+most mentioned root causes, followed by exceptions related to resource 
+management (which was expected since on mobile environment 
+resources are usualy scarse). We could observe that there were 
+stack traces related to the main concerns of mobile develoment.
 
 %TODO: Discuss about the interater agreement here...
-
 %\subsection{Sizes of Stack traces}
@@
 
-\subsection{Characterizing the Most Common Root Exceptions Signalers}
-<COMPLEMENT.... INCLUDE THE METHODS AND CLASSES RESPONSIBLE FOR SIGNALING
-MOST OF THE EXCEPTIONS AND THE EXCEPTIONS SIGNALED BY THEM. SIMILAR TO THE WRAPPINGS TABLES>
-
-\section{RQ2.  What kinds of exception (mis)-use emerge from stack trace analysis? }
-
-To answer this question, besides distilling the information available on stack traces (i.e., exception signalers, root causes, exception wrappings),
+\subsection{What the Types of Root Exceptions can tell us?}
+%\subsection{Exploring the Information Embeeded on Exception Types}
+%\subsection{The Exception Types and What They can Tell us? }
+
+Besides distilling the information available on stack traces (i.e., exception signalers, root causes, exception wrappings),
 we mined the exception types (i.e., RuntimeException, checked Exception or Error) related to each exception reported on stack traces 
-using source code and byte code analyses supported by Exception Miner; and finally checked 
-this information against the compiled best practices. Doing so, we mined some characteristics 
-of exception stack traces that points to best practices violations.
+using source code and byte code analyses supported by Exception Miner described before.
 
@@
 
-\subsection{What the Types of Root Exceptions can tell us?}
-\subsection{Exploring the Information Embeeded on Exception Types}
-
-<falar dos tipos mais comuns.... Runtime e NullPointer???>
-<LINKCAR COM ANALISE DOS SUPERTIPOS>
-<complementar falando do grafico acima>
-\subsubsection{Undocumented runtime exceptions.} After filtering out all the exceptions implicitly 
-signaled by  JVM (due to programming mistakes) and inspecting the signaler methods of such exceptions (i.e.,  direct an indirect 
-instances of java.lang.RuntimeException). We could observe that only 1 method out of 118 inspected signalers
+We could observe that most of the exceptions reported on issues were runtime, 
+mainly  signaled on the Android Platform and in the Application code. we could also find 
+runtime exceptions signaled from libraries.
+
+In order to evaluate if  best practice IV was being obeyed, we manualy inspected the   
+signature of each method signaling a runtime exception, that was defined by third 
+party code (lib or Android) and reused by an application.
+In this task we filtered out all the exceptions implicitly 
+signaled by  JVM (due to programming mistakes) since these exceptions should not be documented on the 
+method signature.
+
+We could observe that only 1 method out of 118 inspected signalers
  (i.e., 0,8\%) documented the explicitly thrown runtime exception in a Javadoc comment; and none of these methods
@@
 
-<fazer o link com wrappings>
-
-\subsection{What the Exception Wrappings can Tell us?}
+We could also observe checked exceptions (which according to Java specification represent
+recoverable conditions) as the root causes of reported exception stack traces.
+A checked exception can remain uncaught if it is declared by evey method on the call chain or if 
+it is wrapped in a runtime exception and rethrown. Next, section describes the kinds of 
+wrappings found in this study.
+
+\subsection{What the Cross-Type  Exception Wrappings can Tell us?}
+
 \subsection{Exploring the Information Embeeded on Exception Wrappings}
@@
 stack traces found in Android repositories. As we can see, some of the checked
-exceptions were indeed wrapped in runtime exceptions or even errors, while most
-of them were not wrapped along the stack trace.
+exceptions were indeed wrapped in runtime exceptions or even errors.
+
+%, while most of them were not wrapped along the stack trace.
 
@@
     \hline
+      %none   & Runtime   &  2360 & 359 \\
+      %none  &   Checked  & 422  & 117 \\
+      %none  &   Error   & 381  & 141 \\
+      %none &  Undefined  &  15   & 10  \\
+      %none & Throwable  &  1    & 1 \\
+     %\hline
+      %Runtime   & Runtime & 560  & 182 \\    
+      %Checked   & Checked & 98   & 42  \\
+      %Error     & Error   & 15   & 14  \\
+    %\hline
   \end{tabular}
@@
 
-This analysis also revealed unexpected wrappings such as: Error wrapping a Checked 
+This analysis revealed unexpected wrappings such as: Error wrapping a Checked 
 exception; a Runtime wrapping and Error; and an Error wrapping a Runtime.
@@
 related to programming mistakes from which the developer is not expected to recover
-from.
+from. Next sections discusses about such types of wrappings.
 
@@
 }
-
-(1)
 
@@
 
- We could find  57 stack traces on 38 projects on which an Error was wrapped in a Runtime Exception.
-
-\begin{table}
-\centering
-\begin{tabular}{ll}
- \bfseries{Signaler from} & \bfseries{Ocurrences} \\
-    \hline
-ANDROID	& 50 \\
-LIB  & 6 \\
-APP	& 1 \\
-\hline
-  \end{tabular}
-\caption{Elements tresponsible for the wrapping.}
-\label{tab:wrapping01}
-\end{table}
-
-\begin{table}
-\centering
-\begin{tabular}{ll}
- \bfseries{Signaler method} & \bfseries{Ocurrences} \\
-    \hline
-android.os.AsyncTask &	25 \\
-android.app.ActivityThread & 15 \\
-android.view.LayoutInflater & 7 \\
-org.asynctasktex.AsyncTaskExd & 3 \\
-android.support.v4.content.ModernAsyncTask & 2\\
-\hline
-  \end{tabular}
-\caption{Elements tresponsible for the wrapping.}
-\label{tab:wrapping01}
-\end{table}
-
-\begin{table}
-\centering
-\begin{tabular}{lll}
- \bfseries{Method} & \bfseries{Project} & \bfseries{Ocurrences}\\
-    \hline
-android.os.AsyncTask &	122 & 549  \\
-android.view.LayoutInflater & 63 & 163 \\
-android.support.v4.* & 100 & 512\\
-\hline
-  \end{tabular}
-\caption{Elements tresponsible for the wrapping.}
-\label{tab:wrapping01}
-\end{table}
-
-EXAMPLE: android.os.AsyncTask
+We could find  57 stack traces on 38 projects on which an Error was wrapped in a runtime exception.
+Manually inspecting each exception stack trace, we could observe that 44% of the wrappings were performed 
+by android.os.AsyncTask.
 
@@
 
-TO INVESTIGATE: WHICH WERE THE LIBS AND THE APP? AND ITS REPUTATION?
+%TO INVESTIGATE: WHICH WERE THE LIBS AND THE APP? AND ITS REPUTATION?
 
 \subsubsection{Checked wrapping Runtime Exception}
+
+COMPLETE
 
@@
 
-The most frequent exception wrapping were were performed inside RoboSpice is an android library that supports the creation of asynchronous long running tasks.
+The most frequent exception wrapping were were performed inside RoboSpice
+ is an android library that supports the creation of asynchronous long running tasks.
 This library has 1.682 stars and 367 forks.
 
-Most of such wrappings however were caused by the reflection Java library, used by Android, Application and library classes.
-relaction methods convert every exception thata happens when invoking a reflected method into an INVOCATION TARGET EXCEPTION
-this conversion is performed inside native code.
+Most of such wrappings however were caused by the reflection Java library, used by Android, 
+Application and library classes. Reflaction methods convert every exception thata happens 
+when invoking a reflected method into an InvoncationTargetException (this conversion is performed inside native code).
 

### 2014-08-31 android-stacks.tex
discussions and typos

@@
 This paper reports a repository mining study which analyzed all the issues created on
-Android projects available in GitHub (482 applications) and Google code (141 appplications)
+Android projects available in GitHub (482 applications) and Google code (141 applications)
 Overall, our research set includes XXXX issues from which XXX stack traces 
@@
 exception stacks traces - e.g., the NullPointerException was the reported cause 
-on 30,08\% of all analyzed exception stack traces, and ewas mentioned in approximatelly 
+on 30,08\% of all analyzed exception stack traces, and was mentioned in approximately 
 50% of all projects analyzed. Moreover,  the cross-project analysis 
-of exception stack traces enabled to consistenly reveal   
+of exception stack traces enabled to consistently reveal   
 fragilities of exception handling code: 
 (1) undocumented runtime exceptions signaled by third party code,
-(2) undocumente checked exceptions thrown by Android native code
+(2) undocumented checked exceptions thrown by Android native code
 (3) unexpected exception wrappings (e.g., Errors being wrapped in checked exceptions) 
 revealing that Java hybrid exception model is not fully used according to its purpose.
-Such fragilities impose challenges to application robusness not only 
-in Android environemt but also to other systems based on Java exception model.
+Such fragilities impose challenges to application robustness not only 
+in Android environment but also to other systems based on Java exception model.
 
@@
 Such increase is motivated by pervasiveness of the World Wide Web 
-and the unprecedent increase of global smartphone shipments -
-the mobile industry shiped over 315 million smartphones in the last quarter of 2013 ~\cite{googleio}.
-
-%the mobile industry shiped over 315 million smartphones in the last quarter of 2013
+and the unprecedented increase of global smartphone shipments -
+the mobile industry shipped over 315 million smartphones in the last quarter of 2013 ~\cite{googleio}.
+
+%the mobile industry shipped over 315 million smartphones in the last quarter of 2013
 %The pervasiveness of the World Wide Web and the astonishing increase 
@@
 % capabilities far beyond of the basic calls and textual messages. 
-%f we look at global smartphone shipments the numbers are astoning;
-%the mobile industry shiped over 315 million smartphones in the last quarter of 2013 [ref-googleIO].
+%f we look at global smartphone shipments the numbers are astonishng;
+%the mobile industry shipped over 315 million smartphones in the last quarter of 2013 [ref-googleIO].
 %Such increase on smartphones combined with the pervasiveness of the World Wide Web
@@
 
-In this context, Android has become a leading plaftorm for such applictions - thanks to its open-source 
+In this context, Android has become a leading platform for such applications - thanks to its open-source 
 model and hardware partners like Samsung, HTC, Motorola, Asus (besides Nexus Google's
@@
 %If on the one hand such applications are getting mode and more popular....
-%On the other hand... there is a need for realiability, as some core functionalities 
+%On the other hand... there is a need for reliability, as some core functionalities 
 %are moving to apps....such applications impose more challenges to software reliability 
@@
 far beyond of the basic calls and textual messages, on the other hand
-they have to face new threats to application robustness imposed by mobile environmet.
+they have to face new threats to application robustness imposed by mobile environment.
 Such applications have to cope with an increasing number of abnormal
-computation states, arising from: programmming errors;
+computation states, arising from: programming errors;
  noisy user inputs; faults in underlying middleware or hardware; 
 faults on third party services and libraries, compatibility issues [ref], 
-conection with devices e.g., as GPS, camera, memory and batery restrictions
+connection with devices e.g., as GPS, camera, memory and battery restrictions
 among others.
@@
 % as any other modern system~\cite{bruntink2006discovering}. 
-The Android platform uses the exception handling mechanism embeeded
+The Android platform uses the exception handling mechanism embedded
  in Java to deal with such conditions. Although the exception handling mechanism
@@
 have shown that it is generally poorly understood and the less tested parts of the system ~\cite{coelho2011unveiling}.
-As a consequence that the exception handling code may become ilself a source of faults, 
+As a consequence that the exception handling code may become itself a source of faults, 
 and may affect the system the other way around, making the system even less robust ~\cite{robillard2003static,coelho2011unveiling}.
@@
 % Google estimates that Android has over 1 billion active users per month, 
-% who pull out theirs phones and check it over 100 bilion times.
+% who pull out theirs phones and check it over 100 billion times.
 
@@
 
-Our conjecture is that several Android applicatons may be suffering from similar 
+Our conjecture is that several Android applications may be suffering from similar 
 fragilities on the exception handling code (of the applications themselves or 
@@
 
-Untill recent years, the availability of issues reporting abnormal
+Until recent years, the availability of issues reporting abnormal
 computation states (e.g., exception stack traces) was scarce. The
@@
    \item Undocumented checked exception signaled by native Android library. This finding points to a fragility 
-on the way exceptions on JNI interface funcions, which can lead to uncaught exceptions that are difficult do 
+on the way exceptions on JNI interface functions, which can lead to uncaught exceptions that are difficult do 
 debug.
@@
 sometimes almost impossible to protect the code against the Uncaught Exceptions that lead to system crashes.
-Although the study focused on Android opensource applications, some of the results are not restricted to this context, 
+Although the study focused on Android open-source applications, some of the results are not restricted to this context, 
 for instance: some of the unexpected wrappings were performed by Java libraries and elements (e.g. reflection library and 
-the static blocks); and the undocumented runtime exceptions were also found in Java libraries besides the Android plaftorm.
+the static blocks); and the undocumented runtime exceptions were also found in Java libraries besides the Android platform.
 The study results are relevant to: (i) Android and Java developers to familiarize common characteristics and fragilities of exception-related code 
@@
 
-The ramaining of this paper is organized as follows. Section 2 presents a
-background on Android Plaform and Java exception model. 
+The remaining of this paper is organized as follows. Section 2 presents a
+background on Android Platform and Java exception model. 
 Section 3 presents the study design. Section 4 reports study findings. 
@@
 the Dalvik runtime environment, and the Application Framework (which provides 
-higher-level services to the applications running on the platform) ~cite{andguide}.
+higher-level services to the applications running on the platform) ~\cite{andguide}.
 Android platform is mostly implemented in Java and uses JNI[ref] to access native C/C++ code. 
@@
 The Android Application Framework also provides several supporting APIs, such as 
-the APIs to support concurrent developement (e.g. AsyncTask). 
+the APIs to support concurrent development (e.g. AsyncTask). 
 Such APIs particularly useful due to the high demand for responsiveness 
 in mobile environment (mobile apps can easily become 
-unresponsivedue due to resurce limitation and excessive network access).
+unresponsive due to resource limitation and excessive network access).
 
 %\begin{figure} \centering \includegraphics[scale=0.8]{arch3.png}
-%  \caption{Android Archtecture} \label{fig:exchier} \end{figure}
+%  \caption{Android Architecture} \label{fig:exchier} \end{figure}
 
@@
 %(ii) implicitly thrown by the JVM when the evaluation of an expression
-% violates the normal semantics of language, also refered as a coding error
+% violates the normal semantics of language, also referred as a coding error
 % (e.g., out-of-bounds array index, division-by-zero, access to a null reference); or 
@@
 % ONE OF THE REVIEWERS COMPLAINED ABOUT THE ABSENCE OF EXPLANATIONS OF WHAT AN ISSU IS...
-% As mentioned before, currently more and more projects are becomming available in 
+% As mentioned before, currently more and more projects are becoming available in 
 % public project repositories such as GitHub and Google Code.
@@
 %aiming at  extracting their common characteristics and looking for similar vulnerabilities on applications and underlaying platform.
-%In Java when an application crashes dut to an uncaught exception ... 
+%In Java when an application crashes due to an uncaught exception ... 
 %Exception Stach trace is usually (procurar por artigo).
@@
 %\subsection{Research Questions}
-In the context of our study we formulated the following reseach questions:
+In the context of our study we formulated the following research questions:
 
@@
 Based on a set of best practices compiled in this study, the study investigates whether characteristics on exception stack traces  
-can pinpoint best practices violations. Chraracteristics of exception stack traces and detected violations can reveal fragilities
+can pinpoint best practices violations. Characteristics of exception stack traces and detected violations can reveal fragilities
  on the exception handling code, which threats the application dependability.
@@
 was retrieved from GitHub and Googlecode, how the exceptions stack traces were extracted and distilled,
- and how the type associated to each exception and its origens was identified.
+ and how the type associated to each exception and its origins was identified.
 
@@
 
-%As mentioned before, this study amined at mining information available on GitHub issues,
+%As mentioned before, this study mined information available on GitHub issues,
 %more specifically, extracting and distilling the stack traces embedded on GItHub issues.
@@
 
-This set of 482 projects contained overall 32,582 issues from which 4,208 stacktraces 
+This set of 482 projects contained overall 32,582 issues from which 4,208 exception stack traces 
 were extracted. Issues on Github are different from issues on dedicated bug tracking tools such as 
@@
 trace contains relevant information concerning the exception structure of the
-projects analyzed - and therefore can review fragilities on the exceptio related code -  
+projects analyzed - and therefore can review fragilities on the exception-related code -  
 we opted for not restricting the analysis only on defect
@@
 Different from GitHub, Googlecode does not provide an API to access the information related
- to hosted projects \footnote{Googlecode previoulsly provided a webservice but was deacrivated in June 2013.}.
+ to hosted projects \footnote{Googlecode previoulsly provided a Web service but was deactivated in June 2013.}.
 However many well used open source Android application such as K9mail, is hosted in Googlecode which stores
-the issue history of this and other apps are store there ~\footnote{K9mail moved to Github but advice their users to create issues on Googlecode as a way of not loosing the issue history.}.
+the issue history of this and other apps are store there ~\footnote{K9mail moved to Github but advice their 
+users to create issues on Googlecode as a way of not loosing the issue history.}.
 
 To overcome this limitation and mine the information available on such Googlecode 
-projects we implemented a Web crawler which navegated (incorporated in Exception Miner tool described next) 
+projects we implemented a Web crawler which navigated (incorporated in Exception Miner tool described next) 
  on each project and issue page, extracting the content of each issue to store on relational database.
-\footnote{Google code previoulsly provided a webservice to extract information about each project, but this service was removed in what Google called "clean-up action" which happened in June 2013.}
-
-To overcome this limitation we needed to implement a Web Crawler (incorporated in Exception Miner tool described next) that navegated 
+\footnote{Google code previously provided a Web service to extract information about each project, but this service was removed in what Google called "clean-up action" which happened in June 2013.}
+
+To overcome this limitation we needed to implement a Web Crawler (incorporated in Exception Miner tool described next) that navigated 
 through the web interface of Googlecode projects extracting all issues and issue comments and storing in a relational database for later analysis.
@@
 To support this study we developed ExceptionMiner, a modular mining tool able 
-to connect to various repositories (such as Google Code,
-GHTorrent or even directly to Bugzilla), extract issues, mine stack traces from
-them and distill exteption stack trace information. The main components of
+to connect to repositories, extract issues, mine stack traces from
+them and distill exception stack trace information. The main components of
 ExceptionMiner are as follows:
 
-\noindent\emph{Repository Connectors} The first step in the process is
-mining references to exceptions and stack traces embedded in issues, d
-GitHub Commector, and GoogleCode Issue Crawler.
+\noindent\emph{Repository Connectors}  This component is responsible to connect to
+ the host repositories in order to collect project issues. The GitHub Connector connected
+ to GiHub through GHTorrent. The GoogleCode connector comprised of a Web Crawler,
+ that traversed Googlecode Web interface in order to extract the project issues.
 
@@
 type hierarchy of a Java exception until it reaches a base exception type.
-This module analyzed the exception hierachy of all exceptions thrown on:
+This module analyzed the exception hierarchy of all exceptions thrown on:
  the last version of SDK version 7, and Android Platform and OS (level 19).
@@
 We downloaded the source code of Android Platform and OS (level 19)
-and this module extractes the packages hierarchies (recursive list of 
+and this module extracts the packages hierarchies (recursive list of 
 of Java packages), that composes it. To discover the packages that compose 
@@
 
-%We then analyzed the lables associated to such issues, and we discovered that only 5,196 were
+%We then analyzed the labels associated to such issues, and we discovered that only 5,196 were
 % labled as defect ~\footnote{the dedect lables considered in this work were every lable containing 
@@
 This study was also based in a set of manual inspection tasks. Besides the 
- identification of Android projects detailes previously. 
+ identification of Android projects details previously. 
 Manual inspections were used discover the type of exceptions thrown by 
@@
 
-This section presents the results amed at answering the research 
+This section presents the results aimed at answering the research 
 questions presented before.
@@
 the the exceptions commonly reported as the root causes of stack traces.
-Table \ref{tab:toptenandroid} illustrates the top 10 root couses found in the study,
+Table \ref{tab:toptenandroid} illustrates the top 10 root causes found in the study,
  ranked by the number of distinct project repositories on which they were reported. 
-Thes table shows: the number of times each exception was reported as 
+The table shows: the number of times each exception was reported as 
 the root cause of a stack trace, the number of distinct projects reporting such causes on stack traces, 
-and the percentages of such in relation to all reported exceptions and analyzed projects respectaly.
+and the percentages of such in relation to all reported exceptions and analyzed projects respectively.
 It also presents the whether the method responsible for signaling the exceptions were
-part of Android, the aplication itself, libcore, or a third-party library as presented in Table  \ref{tab:signalers} 
+part of Android, the application itself, libcore, or a third-party library as presented in Table  \ref{tab:signalers} 
  \footnote{9 stack traces contained only elements of JVM and were discarded from the analysis}.
@@
 We can observe that most of this set is composed by exceptions defined by java.lang,
-which are mostly not expliciltly thrown by the developer but include
+which are mostly not explicitly thrown by the developer but include
 the exceptions implicitly thrown by the JVM due to programming errors 
@@
 
- The NullPoiterExceptions are mainly signaled inside Android platform and inside 
+ The NullPointerExceptions are mainly signaled inside Android platform and inside 
 application code, although we also find
@@
 practice III. It is considered a bad practice because exception type does not carry enough
- information to identify the cause of the exceptional behaviour, and as a consequence developers need to relying on
+ information to identify the cause of the exceptional behavior, and as a consequence developers need to relying on
 the exception message which can be neither complete nor precise~\cite{gosling2000java}.
@@
 
-To get a broather view of the root causes of stack traces, we performed an empirical evaluation
+To get a broader view of the root causes of stack traces, we performed an empirical evaluation
 whose goal was to categorized the top 100 root exceptions. The top 100 root causes correspond 
@@
 
-Such categorization was based known sources of faults mobile developement: concurrency ~cite{to_include}
-, backward compatibility ~cite{to_include}, security  ~cite{to_include} and resource management (IO, Memory, Batery) ~cite{to_include}.
+Such categorization was based known sources of faults mobile development: concurrency [INCLUDEREF]
+, backward compatibility  [INCLUDEREF], security   [INCLUDEREF] and resource management (IO, Memory, Batery)  [INCLUDEREF].
 We added to this set categories to represent the exceptions caused by programming mistakes,
@@
 management (which was expected since on mobile environment 
-resources are usualy scarse). We could observe that there were 
+resources are usually scarce). We could observe that there were 
 stack traces related to the main concerns of mobile develoment.
@@
 \subsection{What the Types of Root Exceptions can tell us?}
-%\subsection{Exploring the Information Embeeded on Exception Types}
+%\subsection{Exploring the Information Embedded on Exception Types}
 %\subsection{The Exception Types and What They can Tell us? }
@@
 %\includegraphics[width=\hsize]{exception_types.pdf}
-%\caption{Distribution of Checked and Unchedked exceptions.}
+%\caption{Distribution of Checked and Unchecked exceptions.}
 %\label{fig:typeroot}
@@
 
-In order to evaluate if  best practice IV was being obeyed, we manualy inspected the   
+In order to evaluate if  best practice IV was being obeyed, we manually inspected the   
 signature of each method signaling a runtime exception, that was defined by third 
@@
 recoverable conditions) as the root causes of reported exception stack traces.
-A checked exception can remain uncaught if it is declared by evey method on the call chain or if 
-it is wrapped in a runtime exception and rethrown. Next, section describes the kinds of 
+A checked exception can remain uncaught if it is declared by every method on the call chain or if 
+it is wrapped in a runtime exception and re-thrown. Next, section describes the kinds of 
 wrappings found in this study.
@@
 
-\subsection{Exploring the Information Embeeded on Exception Wrappings}
+\subsection{Exploring the Information Embedded on Exception Wrappings}
 Table~\ref{tab:wrappingandroid} presents the wrappings found in this study for all
@@
       Checked & Error      & 7 &  8 & 5 &  0  & 0 &  3 \\
-      Error & Checked      & 4  &  7 &  &  0  & 7 &  0     \\
+      Error & Checked      & 4  &  7 & 0 &  7  & 0 &  0     \\
       Error & Runtime     & 6  &  13   & 1 &  1  & 0 &  11    \\
@@
 
-Taking a closer look on the methods reposponsible for such wrappings,
- we could observe that aproximately 41% of them were performed by a single class,
+Taking a closer look on the methods responsible for such wrappings,
+ we could observe that approximately 41% of them were performed by a single class,
  named ActivityThreat. This class is responsible for managing the life cycle of an Android Activities.
@@
 
-Figure X presents a code snippet of ActivityThreat which is reponsible for (among other things)
+Figure X presents a code snippet of ActivityThreat which is responsible for (among other things)
 calling the Activities onCreate method. Hence, as we can see from lines 4-8,
  in case any exception is thrown by an application Activity during its creation, 
-it is wrapped in a general Runtime exception and rethrown.
+it is wrapped in a general Runtime exception and re-thrown.
 
@@
 
-The inversion of control makes it dificult to the application developer to associate 
+The inversion of control makes it difficult to the application developer to associate 
 a specific handler to the exceptions thrown during the onCreate method (or any other Activity life cycle method).
@@
 Difficulty of associating a specific handler to exceptions signaled by application methods called by the framework,
-may also be the reason behind the design desicion of encapsulatling any exception in a RuntimeException.
-As framework does not have enought contextual information the exception regardless its types,
+may also be the reason behind the design decision of encapsulating any exception in a RuntimeException.
+As framework does not have enough contextual information the exception regardless its types,
 it encapsulates in a general type and re-throws. The only possible answer to such exceptions would be 
@@
 This "ignore-for-now" approach may turn the exception handling into a generalized "goto" mechanism~\cite{mandrioli1992advances}
- making the program more complex and even less reliable. This behaviour may lead to the well known uncaught exceptions 
+ making the program more complex and even less reliable. This behavior may lead to the well known uncaught exceptions 
 problem~\cite{jo2004uncaught}, one of the main causes of application crashes.
 
-% Several types of exceptions were  wrapped the RuntimeExceptinos thrown by the 
+% Several types of exceptions were  wrapped the RuntimeExceptions thrown by the 
 % ActivityThread. Table X illustrates them.
@@
 %  \end{tabular}
-% \caption{Elements tresponsible for the wrapping.}
+% \caption{Elements responsible for the wrapping.}
 % \label{tab:wrapping01}
@@
 Manually inspecting each exception stack trace, we could observe that 44% of the wrappings were performed 
-by android.os.AsyncTask.
+by android.os.AsyncTask. This class is responsible for dealing with asynchronous tasks in Android.
+The code snippet below shows that any throwable signaled in the context of an asynchronous 
+task is converted in to a runtime and re-thrown.
+
+ASSNCTASKS NON ANNOTATED METHODS…….
+SOME HAVE EXCEPTION INTERFACE 
+EXCEPTIONS ON MEHTODS SIGNATURE OTHERS NOT
+INCONSISTENCY…. DEPENDS ON DEVELOPERS 
+WORK INSTEAD OF TOOLS.
 
@@
 
-%SELECT (exception, root_cause) AS PAIR, COUNT (*) AS N FROM stacktracesummary where 
-%EXCEPTION_EXTENDS='RUNTIME' AND ROOT_EXTENDS='ERROR' GROUP BY PAIR ORDER BY N DESC;
+We could observe that several types of exceptions were encapsulated in 
+instances of Runtime. Table X shows some kinds of errors encapsulated a
+RuntimeException.
 
@@
 \begin{tabular}{ll}
- \bfseries{Signaler method} & \bfseries{Ocurrences} \\
-    \hline
-java.lang.RuntimeException,java.lang.OutOfMemoryError & 30\\
-java.lang.RuntimeException,java.lang.NoClassDefFoundError & 6\\
-java.lang.RuntimeException,java.lang.UnsatisfiedLinkError) & 5\\
-android.view.InflateException,java.lang.OutOfMemoryError) & 4\\
-java.lang.RuntimeException,java.lang.StackOverflowError)  & 4\\
+ \bfseries{Wrapper type - Root cause type} & \bfseries{Ocurrences} \\
+    \hline
+java.lang.RuntimeException - java.lang.OutOfMemoryError & 30\\
+java.lang.RuntimeException - java.lang.NoClassDefFoundError & 6\\
+java.lang.RuntimeException - java.lang.UnsatisfiedLinkError) & 5\\
+java.lang.RuntimeException -  java.lang.StackOverflowError)  & 4\\
 \hline
   \end{tabular}
-\caption{Elements tresponsible for the wrapping.}
+\caption{Elements responsible for the wrapping.}
 \label{tab:wrapping01}
@@
 
-%TO INVESTIGATE: WHICH WERE THE LIBS AND THE APP? AND ITS REPUTATION?
-
 \subsubsection{Checked wrapping Runtime Exception}
 
-COMPLETE
-
-\begin{table}
-\centering
-\begin{tabular}{ll}
- \bfseries{Origin} & \bfseries{Ocurrences} \\
-    \hline
-APP	& 9 \\
-LIB	& 9 \\
-ANDROID &	4 \\
-\hline
-  \end{tabular}
-\caption{Elements tresponsible for the wrapping.}
-\label{tab:wrapping01}
-\end{table}
+Manually inspecting the stack traces we could observe that most of the wrappings were performed 
+by RoboSpice (an android library that supports the creation of asynchronous long running tasks.
+This library has 1.682 stars and 367 forks) and by the reflection Java library (used by Android, 
+Application and library classes). Inspecting the code related to such wrappings we could observe that
+the reflection library convert every exception that happens 
+when invoking a reflected method into an InvoncationTargetException (this conversion is 
+performed inside native code).
+
 \begin{table}
@@
  \hline
-...android.robospice.CacheSavingException,java.lang.ClassCastException & 6 \\
-java.lang.reflect.InvocationTargetException,java.lang.NullPointerException &  3 \\
-java.lang.reflect.InvocationTargetException,android...ResourcesdNotFoundException & 2 \\
+...android.robospice.CacheSavingException - java.lang.ClassCastException & 6 \\
+java.lang.reflect.InvocationTargetException - java.lang.NullPointerException &  3 \\
+java.lang.reflect.InvocationTargetException - android...ResourcesdNotFoundException & 2 \\
 java.sql.SQLException,android.database.sqlite.SQLiteDiskIOException & 2 \\
@@
 
-The most frequent exception wrapping were were performed inside RoboSpice
- is an android library that supports the creation of asynchronous long running tasks.
-This library has 1.682 stars and 367 forks.
-
-Most of such wrappings however were caused by the reflection Java library, used by Android, 
-Application and library classes. Reflaction methods convert every exception thata happens 
-when invoking a reflected method into an InvoncationTargetException (this conversion is performed inside native code).
-
 \subsubsection{Checked Exception wrapping Error}
 
-We could observe that in 7 projects this  wrapping happend 8 times in total.
-Inspecting the stack traces we could observe that 7 out of 8 wrappings were caused
-by the reflection java library on which java.lang.OutOfMemoryError
-java.lang.NoSuchMethodError were wrapped into a java.lang.reflect.InvocationTargetException.
-
-\begin{table}
-\centering
-\begin{tabular}{ll}
- \bfseries{Origin} & \bfseries{Ocurrences} \\
-    \hline
-ANDROID	& 5 \\
-APP	& 3 \\
-  \end{tabular}
-\caption{Elements tresponsible for the wrapping.}
-\label{tab:wrapping01}
-\end{table}
+Almost all of these wrappings were also caused by the reflection library 
+used by Android applications. The method responsible for the wrappings
+were native methods written in C. Table  ~\ref{tab:wrapcheckerr} illustrates
+some oth these wrappings some of them are masking an OutOfMemoryError
+into a checked exception.
 
@@
 \caption{Elements tresponsible for the wrapping.}
-\label{tab:wrapping01}
+\label{tab:wrapcheckerr}
 \end{table}
 
-All wrappings except 1 were caused by the reflection library.
-The method responsible for the wrappings (e.g., at java.lang.reflect.Constructor.constructNative(Native Method))
-were native methods written in C. In one of the cases, however, in project OpenDocument.droid a java.lang.Exception was 
-wrapping a java.lang.OutOfMemoryError. General handlers are considered as bad practice, 
-but wrapping an Error on a general Exception make the matters worse.
-
 \subsubsection{Error wrapping Checked Exception}
 
-%SELECT (exception, root_cause) AS PAIR, COUNT (*) AS N FROM stacktracesummary where 
-%EXCEPTION_EXTENDS='ERROR' AND ROOT_EXTENDS='CHECKED' GROUP BY PAIR ORDER BY N DESC;
-
-\begin{table}
-\centering
-\begin{tabular}{ll}
-libcore	&  3 \\
- JAVA ENV  & 	4 \\
-  \end{tabular}
-\caption{Elements tresponsible for the wrapping.}
-\label{tab:wrapping01}
-\end{table}
+Some of these wrappings were performed by methods defined by java.lang.Class methods such as 
+at java.lang.Class.getDeclaredFields(Native Method)
+at java.lang.Class.getDeclaredConstructors(Native Method), 
+while others were performed by javax.crypto package. Table X illustrates
+the exception types involved in such wrappings.
 
@@
 
-Some wrappings were performed by methods defined by java.lang.Class methods such as 
-at java.lang.Class.getDeclaredFields(Native Method)
-at java.lang.Class.getDeclaredConstructors(Native Method).  While were performed by javax.crypto package.
-
 \subsubsection{Error wrapping Runtime}
 
-\begin{table}
-\centering
-\begin{tabular}{ll}
-APP & 11 \\
-ANDROID	& 1 \\
-LIB	& 1 \\
-  \end{tabular}
-\caption{Elements tresponsible for the wrapping.}
-\label{tab:wrapping01}
-\end{table}
+After manually inspecting the code associated to such wrappings we discovered that all of them were 
+caused by static initializers. If any exception is thrown in the context of a static initializer (i.e., static block) 
+it is converted into an ExceptionInitializerError on the point where the class is first used. Table X illustrates
+the exception types involved in such wrappings.
 
@@
 
-after manually inspecting the code associated to such wrappings we discovered that all of them were 
-caused by static initializers. If any exception is thrown in the context of a static initializer (i.e., static block) 
-it is converted into an ExceptionInitializerError on the point where the class is first used.
-
-%Table X illustrated examples of such wrappings.
-%TODO INCLUIR ESTA TABLEA
-
-%\begin{figure}
-%  \begin{center}
-%    \includegraphics[scale=0.4]{stack-wrappings-hist.pdf}
-%  \end{center}
-%  \caption{Number of stack traces wrappings for Java and Android repositories. }
-%  \label{fig:sizewrapbhists}
-%\end{figure}
-
 \subsubsection{Multiple Wrappings}
@@
 
-\subsection{What can the Signalers tell us?}
-
-After investigating the most popular exceptions on issues, we also investigated elements
-responsible for signaling most of the exceptions reported on issues.
-Such elements were directly involved on the reported crashes. We cannot
-directly blame such component for the crash since the client may not have been
-using the component according to its "contract", or an external event
-may have happened during the execution of such method.
-
-Regardless the reason, such analysis give as a clue to comon element involved on 
-reported crashes. Table X illustrates the most common root exception signalers
-on issues, toglether with the list of exceptions being thrown.
-
-%Present the top signalers found in this study, and top pairs (signalers and exceptions), and discuss 
-%about inversion of control and the difficult to adequately handle exceptions in this style of development.
-%Inversion of control stimulate the use of unchecked exceptions since the "call back" do not know which exception to expect.
-
-{\tiny
-\begin{verbatim}
-
-% SELECT mainsignaler_method,count(distinct(repo)) as n, string_agg(distinct(main_cause), ',')
-% FROM stacktracesummary group by (mainsignaler_method) order by n desc;
-
-GITHUB:
-"android.os.Parcel.readException";28;"Exception,java.lang.IllegalArgumentException,java.lang.IllegalStateException,java.lang.NullPointerException,java.lang.SecurityException"
-"dalvik.system.BaseDexClassLoader.findClass";19;"java.lang.ClassNotFoundException"
----> "android.app.Instrumentation.checkStartActivityResult";17;"android.content.ActivityNotFoundException"
-"android.graphics.BitmapFactory.nativeDecodeStream";16;"java.lang.OutOfMemoryError"
---->"android.view.WindowManagerImpl.findViewLocked";15;"java.lang.IllegalArgumentException"
-"android.graphics.Bitmap.nativeCreate";15;"java.lang.IllegalArgumentException,java.lang.OutOfMemoryError"
-"android.graphics.BitmapFactory.nativeDecodeAsset";14;"java.lang.OutOfMemoryError,java.lang.RuntimeException"
-"dalvik.system.PathClassLoader.findClass";13;"java.lang.ClassNotFoundException"
----> "android.support.v4.app.FragmentManagerImpl.checkStateLoss";12;"java.lang.IllegalStateException"
-"android.os.BinderProxy.transact";12;"android.os.DeadObjectException,android.os.TransactionTooLargeException,java.lang.RuntimeException"
-"android.database.DatabaseUtils.readExceptionFromParcel";11;"android.database.sqlite.SQLiteDiskIOException,android.database.sqlite.SQLiteException,java.lang.IllegalArgumentException"
-"android.app.ActivityThread.performLaunchActivity";11;"java.lang.NullPointerException,java.lang.RuntimeException"
---->"android.graphics.Canvas.throwIfRecycled";10;"java.lang.NullPointerException,java.lang.RuntimeException" (private method throwing a RuntimeException)
---->"android.os.StrictModedAndroidBlockGuardPolicy.onNetwork";10;"android.os.NetworkOnMainThreadException" (public)
----> "android.support.v4.app.Fragment.getResources";10;"java.lang.IllegalStateException"
-"android.os.AsyncTaskd3.done";10;"java.lang.RuntimeException"
-
-"The Canvas class holds the "draw" calls.
-To draw something, you need 4 basic components: A Bitmap to hold the pixels, a Canvas to host the draw calls (writing into the bitmap), a drawing primitive (e.g.
-Rect, Path, text, Bitmap), and a paint (to describe the colors and styles for the drawing)."
-
-"private static void More ...throwIfRecycled(Bitmap bitmap) {
-941         if (bitmap.isRecycled()) {
-942             throw new RuntimeException(
-943                         "Canvas: trying to use a recycled bitmap " + bitmap);
-944         }
-945     }"
-
-GOOGLECODE:
-"android.app.Instrumentation.checkStartActivityResult";8;"android.content.ActivityNotFoundException"
-"android.graphics.Bitmap.nativeCreate";8;"java.lang.IllegalArgumentException,java.lang.OutOfMemoryError,java.lang.RuntimeException"
-"java.lang.ClassLoader.loadClass";6;"java.lang.ClassNotFoundException,java.lang.IncompatibleClassChangeError,java.lang.UnsupportedClassVersionError"
-"dalvik.system.BaseDexClassLoader.findClass";6;"ClassNotFoundException,java.lang.ClassNotFoundException"
-"com.android.dx.dex.file.ClassDefsSection.add";6;"java.lang.IllegalArgumentException"
-"android.os.Parcel.readException";6;"Exception,java.lang.IllegalArgumentException,java.lang.IllegalStateException,java.lang.NullPointerException,java.lang.SecurityException"
-"android.view.WindowManagerImpl.findViewLocked";5;"java.lang.IllegalArgumentException"
-"android.os.BinderProxy.transact";5;"android.os.DeadObjectException,android.os.RemoteException,android.os.TransactionTooLargeException,java.lang.RuntimeException,java.lang.SecurityException"
-"android.app.ActivityThread.performLaunchActivity";5;"Error,java.lang.NullPointerException,java.lang.RuntimeException"
-"android.graphics.BitmapFactory.nativeDecodeStream";5;"java.lang.OutOfMemoryError"
-
-\end{verbatim}
-}
-
 \section{Discussions and Actionable Advices}
 
-The exception (mis-)use patterns that emerged  in this study sheds light on 
+The results of this study sheds light on 
 problems that may be reducing the robustness of current Java programs in the face of
@@
 
-Discuss about the actionable advices propagated by TAO XIE.
-
-\subsection{Inversion of Control and Exception Handling}
-- Falar de Pytolomy e Ridresh rajan... desafios ja apontados....
-
-%\noindent\emph{The undocumented runtime exceptions} 
-\subsection{The undocumented runtime exceptions thrown by thrid party code} 
-Documenting runtime exceptions is a tedious and error prone task, to help developers
-mitigate this problem tools should be developed to automate the documentation of runtime exceptions
-scraping from library code, few solutions in this directions have been proposed so far ~\cite{van2005combining}.
-
-%\noindent\emph{The odd wrappings}  
-\subsection{Exception Wrappings: Language design purpose x real use}
-
-The exception wrappings were inicially proposed by Java [ref-specification] to ...
-
-But the types of wrappings found in this study pointing to the fact that some wrappings may have not 
-been used according to its intial purpose. It can sometimes be used to bypass the  
-language restrictions imposed by checked exceptions.
-
-When (mis)applied, the exception wrapping can lead to what we call
-an \emph{exception handling confusion problem} which can lead the program to
+\subsection{Cross-Type Wrappings and Threats to Dependability}
+The cross-type wrappings detected in this study points to the fact that: (1) exception 
+wrappings may prevent exception types from being used according to its initial purpose
+ (e.g, Errors should represent situations that should not be handled); and (2) may  be used
+ to bypass the language restrictions imposed by checked exceptions  (e.g,
+ checked exceptions may be wrapped in runtime exceptions).
+
+Hence, when (mis)applied, the exception wrapping can make the exception handling code
+ more complex and lead to what we call an \emph{exception handling confusion problem} which can lead the program to
 an unpredictable state in the presence of exceptions. To illustrate this problem, we can use
@@
 checked exception wrapped an OutOfMemoryError, which represent resource deficiency detected 
-by the JVM which make impossible the program to recover. Further investigation is needed on finding 
-ways to help developers in deadling with such problem, either preventing odd wrappings or enabling 
-the developer to better deal with them.
-
-\subsubsection{Ways of Preventing Odd Wrappings}
-Since there is no way of enforcing Java exception type conventions during program development,
-the exceptions regardless its types can se used interchangeably. This interchangeability makes the
-behaviour of the exception handling code more complex and less reliable. For
-instance, considering a situation were an instance of OutOfMemoryError (a situation that usually cannot
- be handled) was wrapped in a checked exception which tells the caller that such exception should
-handled.
-
-\subsection{Checked and Unchecked Exceptions in JVM and Android and what they can tell us?}
-
-We could observe that some of the type-changing-wrappings were performed by JDK classes.
-Such finding lead us to think that the exception types have been used as intercheagable and may not being carrying 
-the meaning they were firstly designed to carry.
-
-The way the exceptions are being handle are more important that the type itself, it points to the fact already discussed 
-by other languages such as M\# blog [ref] and \"thinking in Java author\" and other blogs. It is all about the exception handling
-politics of a system and not the type itself. The exception type is not enought to ensure a more strict or robust development,
-since the checked exceptions may be silienced or converted to a checked exception. According to Java manual,
-they say that it can actually happen but it will be done intentionally and they advocate that it is a bad practice.
-
-Present the number of checke and unchecked exceptions in JVM, JDK, Android SO, and Android SDK.
-Discuss about the amount of checked and unchecked defined in each, and discuss about interesting exceptions that
-have the same name ex: classNotFound...
-
-%\noindent\emph{The high number of NullPointerExceptions and other JVM-signaled Exceptions}
-
-Moreover, some languages use exceptions to represent coding errors (out-of-bounds array index, division-by-zero, access to a null).
-In these cases exceptions are implicitly signaled by the runtime environment when such condition happens.
-One of such implicit exceptions (caused by data conversion from a 64-bit floating point to a 16-bit signed integer) was the responsible for the famous failure on Ariane 5's first test flight~\cite{lions1996ariane} --- leading to the rocket self-destructing 37 seconds after launch and a loss of 500 million dollars.
-Besides Ariane 5, several applications crash everyday due to uncaught exceptions~\cite{jo2004uncaught}.
-The consequences of not careful exception usage may have the opposite effect of its initial intention of improving system robustness.
-
-\subsection{The high number of null pointer exceptions}
-
-The null references was firstly introdduced by Tony Hoare in ALGOL W, which after some years he called his “one-billion-dollar mistake”:
+by the JVM which make impossible the program to recover.
+
+Currently there is no way of enforcing Java exception type conventions during program development.
+Hence, further investigation is needed on finding ways to help developers in dealing with such
+ problem, either preventing odd wrappings or enabling the developer to better deal with them or even
+researches on the real usefulness of Java hybrid exception model.
+
+\subsection{The Almost Infeasible Task of Preventing Uncaught Exceptions}
+
+In this study we could observe undocumented runtime exceptions thrown by third party code,
+and even undocumented checked exception thrown by Java supporting library.
+Such undocumented exceptions make difficult, and most of the times infeasible
+for the client code to protect against "unforeseen" situations that may happen 
+while calling a library code.
+
+On the other hand documenting runtime exceptions is a tedious and error prone task, to help developers
+mitigate this problem tools should be developed to automate the documentation of runtime exceptions
+scraping from library code, few solutions in this directions have been proposed so far ~\cite{van2005combining}.
+
+\subsection{The high number of NullPointerExceptions}
+
+Besides the explicitly thrown runtime exceptions, in this work we also observe a high number of 
+stack traces reporting coding errors (out-of-bounds array index, division-by-zero, access to a null).
+Such exceptions are implicitly thrown by JVM, and among them the NullPointerException 
+were the reported root cause in most of the analyzed exception stack traces.
+
+The null references was firstly introduced by Tony Hoare in ALGOL W, which after some years he called his “one-billion-dollar mistake”:
 
@@
 
-ref: http://bertrandmeyer.com/category/computer-science/
-
 The high number of null pointer exceptions found in this mining study shed light into a problem
-that although known was not supported by large scale mining studies on the large.
-
-This observation emphasises the importance of solutions to avoid NullPointerExceptions, such as:
-(i) lightweight intra-method nullpointer analysis supported by Java8 @Nullable annotations;
-(ii) intermethod nullpointer analysis tools such as the one proposed by [DANIELE-RELATED-WORK];
+that although known was not supported by large scale mining studies.
+
+This observation emphasizes the importance of solutions to avoid NullPointerExceptions, such as:
+(i) lightweight intra-method null pointer analysis supported by Java8 @Nullable annotations;
+(ii) inter-method null pointer analysis tools such as the one proposed by [DANIELE-RELATED-WORK];
 or (iii) whether language designs which avoid null pointers, such 
@@
 NetBeans, IntelliJ, and other commercial tools which run at compile time and detect potential null 
-pointer derecefences. Android Studio 0.5.5 Released support these annotations.
-
-(1) mostrar a utilizacao de @nullable em projetos android. rodar em alguns e mostrar.
-Android Project Unwritten Rules (http://blog.lemberg.co.uk/android-project-unwritten-rules)
-
-(2) dar foco em android
-
-(3) falar do modelo do ciclo de vida da aplicacao e da necessidade de usar excecoes Runtime.
-
-(4) Local onde tratar excecao - handler padrao - politica da aplicacao. tratamento difucultado multiplas threads e assyncronos.
+pointer dereferences. Android Studio 0.5.5 Released support these annotations.
+
+%(1) mostrar a utilizacao de @nullable em projetos android. rodar em alguns e mostrar.
+%Android Project Unwritten Rules (http://blog.lemberg.co.uk/android-project-unwritten-rules)
+%(2) dar foco em android
+%(3) falar do modelo do ciclo de vida da aplicacao e da necessidade de usar excecoes Runtime.
+%(4) Local onde tratar excecao - handler padrao - politica da aplicacao. tratamento difucultado multiplas threads e assyncronos.
+
+\subsection{General Solution: a General Handler?}
+One may think that the solution for the uncaught exceptions that lead to the system crash,
+is to define a general handler, which is responsible for handling any exception that is not
+adequately handled inside the applications. Although this solution may prevent 
+the system of abruptly crashing, such general handler will not have enough
+contextual information to adequately handle the exception, 
+and the main recovery task will be to present a message to the user
+ and restart the application.
 
 %\noindent\emph{Unwritten Rules of Android Exception Handling Developement}
-
 %\noindent\emph{Dependency injection - Inversion of Control -  callback and Exceptions }  
-\subsection{Dependency injection - Inversion of Control -  callback and Exceptions}
-
+%\subsection{Dependency injection - Inversion of Control -  callback and Exceptions}  
 %\noindent\emph{How can we benefit from such cross-project crash analysis?}
+
 \subsection{The benefits of a cross-project crash analysis}
@@
 When family of systems make use of a common core, or framework such systems can
-bennefit from a cross project crash analysis. The bug issues reported for a group of systems
+benefit from a cross project crash analysis. The bug issues reported for a group of systems
 can identify vulnerabilities which were more difficult to identify in a single project analysis.
@@
 one application could benefit from the usage scenarios on which such framework is applied.
-Framework testing -  what are the main challenges wor framework testing?
-A crossproject crash analysis based on repository mining techniques can 
-identify vulnerabilitues in pieces of code that are shared.
-
-Using exception miner tool we could identify the methods reponsible for most of the crashes - 
+Framework testing -  what are the main challenges on framework testing?
+A cross-project crash analysis based on repository mining techniques can 
+identify vulnerabilities in pieces of code that are shared.
+
+Using exception miner tool we could identify the methods responsible for most of the crashes - 
 such methods can be a target for manual inspections improvements, or better documentation
@@
 
-\subsection{Actionable Advices - The Unwrittern rules of Android development}
-
-%The results of a mining study should be a set of actionalble advices (tao xie).
+\subsection{Actionable Advices - The Unwritten rules of Android development}
+
+%The results of a mining study should be a set of actionable advices (tao xie).
 
@@
 %software systems. Hence, whenever a software crashes this tool sends a detailed
-%crash report to its vendors. Moreover, we can also find third party sofware
+%crash report to its vendors. Moreover, we can also find third party software
 %solutions specialized in bug reporting for different kinds of systems specially
 %for the increasing marked of mobile apps~\cite{BugSe14,BugSn14,Googl14,Acra14}.
-%There is planty of information to be mined...
-
-\begin{table}
-\centering
-\begin{tabular}{llll}
-    \hline
-    \bfseries{Project} & \bfseries{merging imeStars} & \bfseries{Watchers} & \bfseries{Forks} \\
-    \hline
-subsampling-scale-image-view  &  -  &  -    & - \\
-Android-PullToRefresh &  -  &  -    & - \\
-HackerNews &  -  &  -    & - \\
-PathfinderOpenReference  &  -  &  -    & - \\
-AndroidStaggeredGrid  &  -  &  -    & - \\
-HtmlSpanner  &  -  &  -    & - \\
-screen-notifications  &  -  &  -    & - \\
-appdatapreferences-android  &  -  &  -    & - \\
-Android-Universal-Image-Loader  &  -  &  -    & - \\
-Android-SDK  &  -  &  -    & - \\
-WebCachedImageView  &  -  &  -    & - \\
-Android-SkyDrive-Browser  &  -  &  -    & - \\
-android-pdfview  &  -  &  -    & - \\
-hn-android  &  -  &  -    & - \\
-Android-BitmapCache  &  -  &  -    & - \\
-cropper  &  -  &  -    & - \\
-WordPress-Android  &  -  &  -    & - \\
-changeloglib  &  -  &  -    & - \\
-cgeo  &  -  &  -    & - \\
-chromium-webview-UNDERLINE  &  -  &  -    & - \\
-AndroidSlidingUpPanel  &  -  &  -    & - \\
-DsaTab  &  -  &  -    & - \\
-box-android-sdk-v2  &  -  &  -    & - \\
-chromeview  &  -  &  -    & - \\
-andlytics  &  -  &  -    & - \\
-ContractionTimer  &  -  &  -    & - \\
-android  &  -  &  -    & - \\
-okhttp  &  -  &  -    & - \\
-ActionBarSherlock  &  -  &  -    & - \\
-mobile-android  &  -  &  -    & - \\
-android-timepicker  &  -  &  -    & - \\
-QuiltViewLibrary  &  -  &  -    & - \\
-congress-android  &  -  &  -    & - \\
-c-geo-opensource  &  -  &  -    & - \\
-TweetLanes  &  -  &  -    & - \\
-PageTurner  &  -  &  -    & - \\
-SWADroid  &  -  &  -    & - \\
-picasso  &  -  &  -    & - \\
-uservoice-android-sdk  &  -  &  -    & - \\
-\hline
-  \end{tabular}
-\caption{Projects containing wrappings Error-Runtime}
-\label{tab:wrappings3}
-\end{table}
-
-The methods performing such wrappings were the following:
-
-%source code of ActivityThread -  https://github.com/android/platform_frameworks_base/blob/master/core/java/android/app/ActivityThread.java
-
-\begin{table}
-\begin{tabular}{l}
-    \hline
-    \bfseries{root signaler -  occurences} \\
-      \hline
-android.app.ActivityThread.handleBindApplication – 2 \\
-android.app.ActivityThread.performLaunchActivity – 13 \\
-android.os.AsyncTaskd3.done – 25	 \\
-android.support.v4.content.ModernAsyncTaskd3.done - 2 \\
-android.view.LayoutInflater.createView - 7 \\
-android.view.Viewd1.onClick - 1 \\
-com.squareup.picasso.Requestd1.run - 1 \\
-me.openphoto.android.app.util.concurrent.AsyncTaskExd3.done - 1 \\
-org.asynctasktex.AsyncTaskExd3.done – 3 \\
-org.chromium.androidwebview.AwBrowserProcess.loadLibrary -1 \\
-org.gradle.groovy.scripts.internal.DefaultScriptRunnerFactorydScriptRunnerImpl.run - 1 \\
-\hline
-  \end{tabular}
-\caption{Top 10 most popular exceptions on issues in Java projects}
-\label{tab:toptenjava}
-\end{table}
-
-\subsection{Exception Handling in Programming Languages}
-
-Techniques for error detection and handling are not  an optional add-on but a
-fundamental part of any system~\cite{bruntink2006discovering}. 
-The exception mechanism~\cite{goodenough1975exception} was proposed as a way of
-preventing errors from being ignored and was embedded in many mainstream
-programming languages ~\cite{garcia2001comparative} since then.
-
-In languages such as Modula-3, Guide, and Extended Ada the exceptions are 
-checked, which means that the compiler statically checks if appropriate handlers 
-are provided within the system. Accordingly, the programmer must explicitly specify 
-every exception that can be signaled to support the compile time checking. This approach is known to be
-costly to maintain (Dooren and Steegmans, 2005) since exception updates are
-cascading; if a new exception type is added to a method signature then all other
-methods directly using this method must either change their signatures to
-include the new exception type or handle the new exception.
-
-In languages such as Lore, C++, and Arche all exceptions are unchecked but the developer can 
-optionally list exception types on method signatures. Although exceptions can 
-be documented on methods signatures, the developer is not warned by the compiler
-if an unchecked exception is not handled inside the application. According to ~\cite{Robil00} 
-such an approach may hinder robustness because the client of a method cannot easily know
-which unchecked exceptions may be thrown (unless the code of the method and the
-methods called from it is inspected) which can be a very time consuming or
-infeasible task.
-
-Java [ref: Gosling et al. 1996] and CLU [Liskov and Snyder 1979] support the checked exceptions,
- (therefore the compiler can then check if handlers are provided to such exceptions), however, this support is 
-only partial because both languages also provides a form of unchecked exceptions [ref: Robilard].
-In CLU,  if a handler is not found for a given exception, it is automaticaly converted in a predened default exception, 
-called failure which is the only one which can be implicitly propagated [ref-clue].
- In Java the developer can define specific unchecked and checked exceptions and use both kinds in a single system.
-
-There is a long-lasting debate about the pros and cons of checked and unchecked exceptions 
-~\cite{javatut,stackoverlow,debate},\footnote{152 questions in Stackoverlow
-are related to this debate}. However, the debates are mainly based on the developers own experience in
-using both models, since there is still few empirical evidence of the pros and cons of each model.
-
-%Although exception handling mechanism has been embedded in several programming languages 
-%as show above,  the exception handling code is often poorly understood and the least
-%tested part of software systems [ref:POORLY].
-
-%[REF]: Static Analysis to Support the Evolution of Exception Structure in Object-Oriented Systems
-
-% This behaviour may lead to the problem of uncaught exceptions~\cite{jo2004uncaught}, one of the main causes of application crashes
-%(e.g.~\cite{miller1997issues,Robil00,shah2010understanding, garcia2007extracting, garcia2001comparative,cabral2007exception,coelho2011unveiling}).
+%There is plenty of information to be mined...
 

### 2014-08-31 android-stacks.tex
all over

@@
 on 30,08\% of all analyzed exception stack traces, and was mentioned in approximately 
-50% of all projects analyzed. Moreover,  the cross-project analysis 
+50\% of all projects analyzed. Moreover,  the cross-project analysis 
 of exception stack traces enabled to consistently reveal   
@@
   \item  A multitude of programming mistakes were reported as the main causes of stack traces:
- XXX  of reported stack traces were caused by programming mistakes; and XXX of analyzed stack traces reported a java.lang.NullPointerException as root cause;
+ XXX  of reported stack traces were caused by programming mistakes; and XXX of analyzed stack
+ traces reported a java.lang.NullPointerException as root cause;
 
@@
 sometimes almost impossible to protect the code against the Uncaught Exceptions that lead to system crashes.
-Although the study focused on Android open-source applications, some of the results are not restricted to this context, 
-for instance: some of the unexpected wrappings were performed by Java libraries and elements (e.g. reflection library and 
+Although the study focused on Android open-source applications, these findings are not restricted to this context, since 
+the above mentioned fragilities may happen in any Java-based system. 
+For instance: the unexpected wrappings were performed by Java libraries and elements (e.g. reflection library and 
 the static blocks); and the undocumented runtime exceptions were also found in Java libraries besides the Android platform.
@@
 to be wrapped in another one and re-thrown. Figure~\ref{fig:wrapping} presents 
-a stack trace associated which illustrates an exception wrapping. The bottom 
+a stack trace  illustrates an exception wrapping. The bottom 
 part of the stack trace is the \emph{root cause}, which indicates the
@@
 memory). The top part of the stack trace indicates the location of the exception
-manifestation (which will call \emph{top level exception} along this paper). The
+manifestation (which will call \emph{top level wrapper} along this paper). The
 execution flow  between the root cause and the exception manifestation may
@@
 which contained the distilled information of stack traces, exception types,
-and signalers origin (e.g., root exception and its type, root signaler and its origin,
-exception wrapper type, wrapper signaler).
+and signalers origin (e.g., root cause, root cause type, root cause signaler and its origin,
+exception wrappers,  types, and its signalers).
 
@@
  checked exceptions may be wrapped in runtime exceptions).
+
+\begin{figure} \centering \includegraphics[scale=0.4]{codeexample2.pdf}
+\caption{Code snipets illustrating fragilities on the exception handling code} \label{fig:exchier} \end{figure}
 

### 2014-08-31 android-stacks.tex
adding more data on google code

@@
 
-\title{ What can GitHub and Googlecode Issues tell us about Exceptions in Android? }
+\title{ What can GitHub and Googlecode Issues tell us about Exceptions in Android and Java? }
 
@@
 This paper reports a repository mining study which analyzed all the issues created on
-Android projects available in GitHub (482 applications) and Google code (141 applications)
-Overall, our research set includes XXXX issues from which XXX stack traces 
+Android projects available in GitHub (482 applications) and Google code (157 applications)
+Overall, our research set includes 32,582 issues in GitHub and 41,674 issues in Googlecode from which (4,208)  (1,986) in GC stack traces 
 were extracted. Such stack traces were distilled and analyzed in 
@@
 had at least one issue including a stack trace. Then we performed further cleanup, inspecting the site of every Android
-reporting at least one stack trace, and as a result we could identify 158 
-Android projects.
+reporting at least one stack trace, and as a result we could identify 157. 
+Android projects.  From this set all issues (41,674) were analyzed and 1,986 stack traces were extracted.
 
@@
 
-\subsection{The Common Root Exceptions and What They can Tell us? }
+\subsection{What the Common Root Exceptions can tell us? }
 
@@
 part of Android, the application itself, libcore, or a third-party library as presented in Table  \ref{tab:signalers} 
- \footnote{9 stack traces contained only elements of JVM and were discarded from the analysis}.
+ \footnote{Nine exception stack traces contained only elements of JVM and were discarded from the analysis}.
 
@@
     \hline
-    \bfseries{Root Exception} &  \multicolumn{2}{c}{\bfseries{Projects}} &  \multicolumn{2}{c}{\bfseries{Occurrences}} & \textsf{Android} & \textsf{Libcore} & \textsf{App} & \textsf{Lib} \\
+    \bfseries{Root Exception} &  \multicolumn{2}{c}{\bfseries{Projects (GH/CG)}} &  \multicolumn{2}{c}{\bfseries{Occurrences (GH/CG)}} & \textsf{Android} & \textsf{Libcore} & \textsf{App} & \textsf{Lib} \\
     & \bfseries{\#} &  \bfseries{\%} & \bfseries{\# } & \bfseries{\% } &&&&\\
     \hline
-java.lang.NullPointerException            & 254 & 52,70\%  & 1225 & 30,08\% & 473 & 18 & 595 & 137 \\
-java.lang.IllegalStateException          & 99  & 20,54\%  & 234  & 5,75\%  & 165 & 12 & 36  & 20  \\
-java.lang.IllegalArgumentException          & 94  & 19,50\% & 255  & 6,26\% & 146 & 6  & 64  & 39   \\
-java.lang.RuntimeException                & 91  & 18,88\% & 232  & 5,70\%  & 167 & 1  & 47  & 17   \\
-java.lang.OutOfMemoryError                 & 56  & 11,62\% & 180  & 4,42\% & 121 & 15 & 17  & 23   \\
-java.lang.NoClassDefFoundError           & 52  & 10,79\%  & 73   & 1,79\%  & 9   & 0  & 37  & 26   \\
-java.lang.ClassCastException             & 49  & 10,17\%  & 94   & 2,31\%  & 43  & 0  & 40  & 11   \\
-java.lang.IndexOutOfBoundsException       & 47  & 9,75\%   & 127  & 3,12\% & 47  & 0  & 71  & 8    \\
-java.lang.NoSuchMethodError               & 40  & 8,30\%  & 57   & 1,40\%  & 9   & 0  & 39  & 9   \\
-java.util.ConcurrentModificationException & 38  & 7,88\% & 54   & 1,33\%   & 5   & 0  & 43  & 6   \\
+java.lang.NullPointerException            & 254 /  78  & 52,70\%  & 1225 / 439  & 30,08\% & 473 /  26  & 18 & 595 & 137 \\  %LIBORAPP=408
+java.lang.IllegalStateException          & 99 / 21  & 20,54\%  & 234 / 44  & 5,75\%  & 165 & 12 & 36  & 20  \\
+java.lang.IllegalArgumentException          & 94 / 48  & 19,50\% & 255 / 98  & 6,26\% & 146 & 6  & 64  & 39   \\
+java.lang.RuntimeException                & 91 / 31 & 18,88\% & 232 / 87  & 5,70\%  & 167 & 1  & 47  & 17   \\
+java.lang.OutOfMemoryError                 & 56  / 22  & 11,62\% & 180 / 57 & 4,42\% & 121 & 15 & 17  & 23   \\
+java.lang.NoClassDefFoundError           & 52  / 15 & 10,79\%  & 73 / 21   & 1,79\%  & 9   & 0  & 37  & 26   \\
+java.lang.ClassCastException             & 49  / 15 & 10,17\%  & 94 / 36   & 2,31\%  & 43  & 0  & 40  & 11   \\
+java.lang.IndexOutOfBoundsException       & 47  / 15 & 9,75\%   & 127 / 39  & 3,12\% & 47  & 0  & 71  & 8    \\
+java.lang.NoSuchMethodError               & 40 / 14  & 8,30\%  & 57 / 23  & 1,40\%  & 9   & 0  & 39  & 9   \\
+java.util.ConcurrentModificationException & 38 / 5 & 7,88\% & 54 / 11  & 1,33\%   & 5   & 0  & 43  & 6   \\
 
@@
     \hline
-    \bfseries{Category} &  \bfseries{Occurence on Stacks (\%)} \\
-    \hline
-      Programming logic (java.lang and util) & 2326 (57,9\%)\\
-      Resources (IO)                         &  727 (18,1\%) \\
-      Security                               &  165 (4,1\%)\\
-      Concurrency                            & 116  (2,9\%)\\
-      Backward compatibility                 &  219 (5,5\%) \\
-      Specific Exceptions               &  197 (4,9\%)\\
-      General (Error, Exception, Runtime)    &  267 (6,7\%)\\
+    \bfseries{Category} &  \bfseries{\% reported on stacks} \\
+    \hline
+      Programming logic (java.lang and util) &  57,9\%\\ %2326
+      Resources (IO)                         &   18,1\% \\ %727
+      Security                               &  4,1\%\\ %165 
+      Concurrency                            &  2,9\% \\ %116 
+      Backward compatibility                 & 5,5\% \\ %219
+      Specific Exceptions               &  4,9\%\\ %197 
+      General (Error, Exception, Runtime)    &  6,7\%\\ %267 
     \hline
@@
 management (which was expected since on mobile environment 
-resources are usually scarce). We could observe that there were 
-stack traces related to the main concerns of mobile develoment.
+resources are usually scarce). Some of such programming mistakes,
+however, may be interactly related to one of the other concerns. For 
+instance, a concurrency problem may lead to a NullPointerException.
+This classification however, was only based on the semantic embedded 
+on the exception types.
 
@@
 
-\subsection{What the Cross-Type  Exception Wrappings can Tell us?}
-
-\subsection{Exploring the Information Embedded on Exception Wrappings}
+\subsection{What the Exception Wrappings can tell us?}
+%\subsection{Exploring the Information Embedded on Exception Wrappings}
 Table~\ref{tab:wrappingandroid} presents the wrappings found in this study for all
@@
 \begin{tabular}{lll}
+    \hline
  \bfseries{Signaler class} &  \bfseries{Wrapper Type} & \bfseries{Ocurrences} \\
@@
 
+%\begin{table}
+%\centering
+%\begin{tabular}{ll}
+% \bfseries{Wrapper type - Root cause type} & \bfseries{\#} \\
+%    \hline
+%java.lang.RuntimeException - java.lang.OutOfMemoryError & 30\\
+%java.lang.RuntimeException - java.lang.NoClassDefFoundError & 6\\
+%java.lang.RuntimeException - java.lang.UnsatisfiedLinkError) & 5\\
+%java.lang.RuntimeException -  java.lang.StackOverflowError)  & 4\\
+%\hline
+%  \end{tabular}
+%\caption{Elements responsible for the wrapping.}
+%\label{tab:wrapping01}
+%\end{table}
+
 \begin{table}
 \centering
-\begin{tabular}{ll}
- \bfseries{Wrapper type - Root cause type} & \bfseries{Ocurrences} \\
-    \hline
-java.lang.RuntimeException - java.lang.OutOfMemoryError & 30\\
-java.lang.RuntimeException - java.lang.NoClassDefFoundError & 6\\
-java.lang.RuntimeException - java.lang.UnsatisfiedLinkError) & 5\\
-java.lang.RuntimeException -  java.lang.StackOverflowError)  & 4\\
+
+  \begin{tabular}{|p{7.9cm} p{0.1cm}|}
+%\begin{tabular}{ll}
+    \hline
+ \bfseries{Runtime-Error Wrapping Examples} &   \\  %\bfseries{\#}
+    \hline
+java.lang.RuntimeException - java.lang.OutOfMemoryError & \\ %30
+java.lang.RuntimeException - java.lang.NoClassDefFoundError &\\ %6
+java.lang.RuntimeException - java.lang.UnsatisfiedLinkError) & \\ %5
+java.lang.RuntimeException -  java.lang.StackOverflowError)  & \\ %4
+\hline
+\bfseries{Checked-Runtime Wrapping Examples} &  \\
+ \hline
+...android.robospice.CacheSavingException - java.lang.ClassCastException &  \\ %6
+java.lang.reflect.InvocationTargetException - java.lang.NullPointerException &   \\ %3
+java.lang.reflect.InvocationTargetException - ...ResourcesdNotFoundException &  \\ %2
+java.sql.SQLException - android.database.sqlite.SQLiteDiskIOException &  \\ %2
+ \hline
+\bfseries{Checked-Error Wrapping Examples} & \\
+ \hline
+java.lang.reflect.InvocationTargetException - java.lang.OutOfMemoryError &  \\ %5
+java.lang.Exception - java.lang.OutOfMemoryError &  \\ %1
+java.lang.reflect.InvocationTargetException - java.lang.NoSuchMethodError &	 \\ %1
+java.lang.reflect.InvocationTargetException - java.lang.UnsatisfiedLinkError&	 \\ %1
+\hline
+\bfseries{Error-Checked Wrapping Examples} & \\
+ \hline
+java.lang.NoClassDefFoundError - java.lang.ClassNotFoundException &  \\ %4
+java.lang.AssertionError - javax.crypto.ShortBufferException) &  \\ %3
+\hline
+\bfseries{Error-Runtime Wrapping Examples} &   \\
+ \hline
+java.lang.ExceptionInInitializerError - java.lang.NullPointerException &  \\ %9
+java.lang.ExceptionInInitializerError - java.lang.IllegalArgumentException &	 \\ %2
  \hline
   \end{tabular}
-\caption{Elements responsible for the wrapping.}
+\caption{Wrapping examples}
 \label{tab:wrapping01}
@@
 
-\begin{table}
-\centering
-\begin{tabular}{lll}
- \bfseries{Wrapper} & \bfseries{Root} & \bfseries{Ocurrences}\\
- \hline
-...android.robospice.CacheSavingException - java.lang.ClassCastException & 6 \\
-java.lang.reflect.InvocationTargetException - java.lang.NullPointerException &  3 \\
-java.lang.reflect.InvocationTargetException - android...ResourcesdNotFoundException & 2 \\
-java.sql.SQLException,android.database.sqlite.SQLiteDiskIOException & 2 \\
-java.lang.Exception,at.andiwand.odf2html.odf.IllegalMimeTypeException & 	2 \\
-  \end{tabular}
-\caption{Elements tresponsible for the wrapping.}
-\label{tab:wrapping01}
-\end{table}
+%\begin{table}
+%\centering
+%\begin{tabular}{ll}
+% \bfseries{Wrapper type - Root cause type} & \bfseries{\#}\\
+% \hline
+%...android.robospice.CacheSavingException - java.lang.ClassCastException & 6 \\
+%java.lang.reflect.InvocationTargetException - java.lang.NullPointerException &  3 \\
+%java.lang.reflect.InvocationTargetException - ...ResourcesdNotFoundException & 2 \\
+%java.sql.SQLException - android.database.sqlite.SQLiteDiskIOException & 2 \\
+%  \end{tabular}
+%\caption{Elements tresponsible for the wrapping.}
+%\label{tab:wrapping01}
+%\end{table}
 
@@
 
-\begin{table}
-\centering
-\begin{tabular}{lll}
- \bfseries{Wrapper} & \bfseries{Root} & \bfseries{Ocurrences}\\
- \hline
-java.lang.reflect.InvocationTargetException,java.lang.OutOfMemoryError & 5 \\
-java.lang.Exception,java.lang.OutOfMemoryError & 1 \\
-java.lang.reflect.InvocationTargetException,java.lang.NoSuchMethodError &	1 \\
-java.lang.reflect.InvocationTargetException,java.lang.UnsatisfiedLinkError&	1 \\
-  \end{tabular}
-\caption{Elements tresponsible for the wrapping.}
-\label{tab:wrapcheckerr}
-\end{table}
+%\begin{table}
+%\centering
+%\begin{tabular}{lll}
+% \bfseries{Wrapper type - Root cause type} & \bfseries{\#}\\
+% \hline
+%java.lang.reflect.InvocationTargetException - java.lang.OutOfMemoryError & 5 \\
+%java.lang.Exception - java.lang.OutOfMemoryError & 1 \\
+%java.lang.reflect.InvocationTargetException - java.lang.NoSuchMethodError &	1 \\
+%java.lang.reflect.InvocationTargetException - java.lang.UnsatisfiedLinkError&	1 \\
+%  \end{tabular}
+%\caption{Elements tresponsible for the wrapping.}
+%\label{tab:wrapcheckerr}
+%\end{table}
 
@@
 
-\begin{table}
-\centering
-\begin{tabular}{lll}
- \bfseries{Wrapper} & \bfseries{Root} & \bfseries{Ocurrences}\\
- \hline
-java.lang.NoClassDefFoundError & java.lang.ClassNotFoundException & 4 \\
-java.lang.AssertionError & javax.crypto.ShortBufferException) & 3 \\
- \end{tabular}
-\caption{Elements tresponsible for the wrapping.}
-\label{tab:wrapping01}
-\end{table}
+%\begin{table}
+%\centering
+%\begin{tabular}{lll}
+% \bfseries{Wrapper} & \bfseries{Root} & \bfseries{\#}\\
+% \hline
+%java.lang.NoClassDefFoundError & java.lang.ClassNotFoundException & 4 \\
+%java.lang.AssertionError & javax.crypto.ShortBufferException) & 3 \\
+% \end{tabular}
+%\caption{Elements tresponsible for the wrapping.}
+%\label{tab:wrapping01}
+%\end{table}
 
@@
 
-\begin{table}
-\centering
-\begin{tabular}{lll}
- \bfseries{Wrapper} & \bfseries{Root} & \bfseries{Ocurrences}\\
- \hline
-java.lang.ExceptionInInitializerError,java.lang.NullPointerException & 9 \\
-java.lang.ExceptionInInitializerError,java.lang.IllegalArgumentException &	2 \\
-java.lang.ExceptionInInitializerError,java.util.MissingResourceException) &	1 \\
-java.lang.ExceptionInInitializerError,org.andengine.util.exception.AndEngineRuntimeException &	1 \\
- \end{tabular}
-\caption{Elements tresponsible for the wrapping.}
-\label{tab:wrapping01}
-\end{table}
+%\begin{table}
+%\centering
+%\begin{tabular}{ll}
+% \bfseries{Wrapper type - Root cause type} & \bfseries{\#} \\
+% \hline
+%java.lang.ExceptionInInitializerError - java.lang.NullPointerException & 9 \\
+%java.lang.ExceptionInInitializerError - java.lang.IllegalArgumentException &	2 \\
+% \end{tabular}
+%\caption{Elements tresponsible for the wrapping.}
+%\label{tab:wrapping01}
+%\end{table}
 
@@
 
-\section{Discussions and Actionable Advices}
-
-The results of this study sheds light on 
-problems that may be reducing the robustness of current Java programs in the face of
-exceptional conditions.
-
-\subsection{Cross-Type Wrappings and Threats to Dependability}
+\section{Futher Discussions}
+
+\begin{figure} \centering \includegraphics[scale=0.4]{codeexample2.pdf}
+\caption{Code snippets illustrating fragilities on the exception handling code} \label{fig:exchier} \end{figure}
+
+\emph{"Everybody hates thinking about exceptions,                   							because they’re not supposed to happen."}
+\textbf{Brian Foote}
+
+\subsection{On the cross-type wrappings as a threat to dependability}
 The cross-type wrappings detected in this study points to the fact that: (1) exception 
@@
 
-\begin{figure} \centering \includegraphics[scale=0.4]{codeexample2.pdf}
-\caption{Code snipets illustrating fragilities on the exception handling code} \label{fig:exchier} \end{figure}
-
 Hence, when (mis)applied, the exception wrapping can make the exception handling code
- more complex and lead to what we call an \emph{exception handling confusion problem} which can lead the program to
+ more complex and lead to what we call an \emph{exception handling confusion problem}.
+Such problem can lead the program to
 an unpredictable state in the presence of exceptions. To illustrate this problem, we can use
@@
 
-\subsection{The Almost Infeasible Task of Preventing Uncaught Exceptions}
-
+\subsection{On the need of a rescue to the almost Infeasible task of preventing uncaught exceptions}
 In this study we could observe undocumented runtime exceptions thrown by third party code,
@@
 
+One may think that the solution for the uncaught exceptions 
+may be to define a general handler, 
+which is responsible for handling any exception that is not
+adequately handled inside the applications. Although this 
+solution may prevent  the system of abruptly crashing,
+ such general handler will not have enough
+contextual information to adequately handle the exception, 
+and only tasks possible are: to present a message to the user
+ and restart the application.
+
+Such handler does not replace a carefully designed exception 
+handling policy, which can hardly be designed in the absence of 
+third-party documentation on the exceptions that
+may the thrown by it.
+
 On the other hand documenting runtime exceptions is a tedious and error prone task, to help developers
@@
 
-\subsection{The high number of NullPointerExceptions}
+\subsection{On the Null Pointer Problem}
 
@@
 The high number of null pointer exceptions found in this mining study shed light into a problem
-that although known was not supported by large scale mining studies.
+that although known was not supported by large scale studies.
 
@@
 
-\subsection{General Solution: a General Handler?}
-One may think that the solution for the uncaught exceptions that lead to the system crash,
-is to define a general handler, which is responsible for handling any exception that is not
-adequately handled inside the applications. Although this solution may prevent 
-the system of abruptly crashing, such general handler will not have enough
-contextual information to adequately handle the exception, 
-and the main recovery task will be to present a message to the user
- and restart the application.
-
 %\noindent\emph{Unwritten Rules of Android Exception Handling Developement}
@@
 
-\subsection{The benefits of a cross-project crash analysis}
+\subsection{On the benefits of a cross-project exception stack trace analysis}
 
@@
 (because sometimes an exception can be thrown because the user is not invoking the method
-in the proper way
+in the proper way ....
 
 During such analysis we could also find the undocumented runtime exceptions 
-signaled by libraries.
+signaled by reused therd part code libraries.
 

### 2014-08-31 android-stacks.tex
changes all over

@@
 
-\subsection{What the Common Root Exceptions can tell us? }
+\subsection{What can the Common Root Exceptions tell us? }
 
@@
 
-\subsubsection{The high number of NullPointerExceptions}
+\emph{\textbf{The high number of NullPointerExceptions.}}
 From this set the java.lang.NullPointerException was the most reported root cause; 1225 stack traces 
@@
 (which represents 20,54\% of all projects reporting at least one stack trace).
-
  The NullPointerExceptions are mainly signaled inside Android platform and inside 
@@
 detected bugs were due to NullPointerExceptions and implicitly-signaled of Java
-environment (as the ones found in this study).
-
-In Section XX we provide further discussion on the NullPointerException problem, pointing 
-to ways of how to deal with it and mitigate it.
-
-\subsubsection{Direct instances of RuntimeExceptions being thrown.}
-
-From Tables~\ref{tab:toptenjava} and~\ref{tab:toptenandroid} we can observe that,
-in both Java and Android related issues, direct instances
-java.lang.RuntimeException were thrown  in 10,94\% of Java repositories and
-18,88\% of Android repositories respectively. 
-In the Android repositories, most of such exceptions  were  thrown by the Android 
-platform/OS (167 out of 232).
+environment (as the ones found in this study).In Section XX we provide further 
+discussion on the NullPointerException problem, pointing 
+to ways of how to mitigate it.
+
+\emph{\textbf{Direct instances of RuntimeExceptions being thrown.}}
+From Table ~\ref{tab:toptenandroid} we can observe that direct instances
+java.lang.RuntimeException were thrown in 18,88\% of Android repositories, and
+most of such exceptions  were  thrown by the Android platform (167 out of 232).
 
@@
 
-\subsubsection{Classifying Top 100 Exceptions}
-
-To get a broader view of the root causes of stack traces, we performed an empirical evaluation
+\emph{\textbf{Classifying Top 100 Exceptions.}} To get a broader view of the root causes of stack traces,
+ we performed an empirical evaluation
 whose goal was to categorized the top 100 root exceptions. The top 100 root causes correspond 
@@
     \hline
-      Programming logic (java.lang and util) &  57,9\%\\ %2326
-      Resources (IO)                         &   18,1\% \\ %727
-      Security                               &  4,1\%\\ %165 
-      Concurrency                            &  2,9\% \\ %116 
-      Backward compatibility                 & 5,5\% \\ %219
-      Specific Exceptions               &  4,9\%\\ %197 
-      General (Error, Exception, Runtime)    &  6,7\%\\ %267 
-    \hline
+      Programming logic (java.lang and util) &  52,0\%\\ 
+      Resources (IO, Memory, Batery)       &   23,9\% \\ 
+      Security                               &  4,1\%\\  
+      Concurrency                            &  2,9\% \\ 
+      Backward compatibility                 & 5,5\% \\ 
+      Specific Exceptions               &  4,9\%\\ 
+      General (Error, Exception, Runtime)    &  6,7\%\\
   \end{tabular}
@@
 
-In this second analysis the programming mistakes were also the
-most mentioned root causes, followed by exceptions related to resource 
-management (which was expected since on mobile environment 
-resources are usually scarce). Some of such programming mistakes,
-however, may be interactly related to one of the other concerns. For 
-instance, a concurrency problem may lead to a NullPointerException.
-This classification however, was only based on the semantic embedded 
-on the exception types.
+Analizing the Top 100 exceptions the programming mistakes
+were also the most reported. However, they may be interactly related 
+to one of the other concerns. For  instance, a concurrency problem may 
+lead to a NullPointerException. This classification however, 
+was only based on the semantic embedded  on the exception types.
+
+\noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
+Programming mistakes were reported as the root cause of  52\% of the analyzed stack traces. 
+In 30,08\% NullPointerExceptions was reported as the root cause.
+}
+\end{minipage}}
 
@@
 
-\subsection{What the Types of Root Exceptions can tell us?}
+\subsection{What can the Types of Root Exceptions tell us?}
 %\subsection{Exploring the Information Embedded on Exception Types}
@@
 Such undocumented runtime exceptions represent a threat to system robustness, specially
-when such exceptions are thrown then the third party code is invoked inside the application (e.g. libraries, or framework utility code).
-In such cases the client usually do not have access to the source code, and in the absence of
+when such exceptions are thrown then the third party code is invoked inside the application 
+(e.g. libraries, or framework utility code). In such cases the client usually do not have access to 
+the source code, and in the absence of
 the exception documentation it is very difficult or even impossible for the client of such third party code to 
@@
 recoverable conditions) as the root causes of reported exception stack traces.
-A checked exception can remain uncaught if it is declared by every method on the call chain or if 
-it is wrapped in a runtime exception and re-thrown. Next, section describes the kinds of 
-wrappings found in this study.
-
-\subsection{What the Exception Wrappings can tell us?}
+Two possible ways of a checked exception can remain uncaught are: the checked 
+exception is declared by every method on the call chain or if it is wrapped in a 
+runtime exception and re-thrown. However the inspection of stack traces 
+and the signatured of methods revealed the a third option: an undocumented
+checked exception thrown by a native method.
+
+The code snippet in Figure X illustrate this scenario.
+ The java/lang/Class declares a native method named getDeclaredMethods. 
+The Java-side declaration of of this method does not have any throws clause, 
+leading programmers and the compiler to think that no checked exceptions can be thrown.
+
+ However, the C-code implementation did throw a "checked exception" NoSuchMethodException, 
+violating the declaration. The Java compiler, could not detect this violation because does 
+not perform static exception  checking on native methods. This type of bugs is hard to diagnose
+because the developer usualy do not have access to the native code (written in other language).
+
+Consequently, since it was not expected by the programmer, when such method throws 
+this exception, such undocumented exception may remain
+uncaught and cause the app to crash, or be mistakenly handled by subsumption - 
+The exception stack traces reporting this scenario is actually a real bug of Android 
+Gingerbread version (which still accounts for 13.6% of devices running Android).
+
+\noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
+  0,8\% of the methods documented the explicitly signaled runtime exceptions.
+  The study also detected an undocumented checked exception signaled by
+  a native method defined in Android.
+}
+\end{minipage}}
+
+\subsection{What can the Exception Wrappings tell us?}
 %\subsection{Exploring the Information Embedded on Exception Wrappings}
@@
 
-\subsubsection{Runtime wrapping Checked Exception}
-
-Such wrapping was detected on 70 distinct projects. Inspecting the 
+\emph{\textbf{Runtime wrapping Checked Exception.}} Such wrapping was 
+detected on 70 distinct projects. Inspecting the 
 elements responsible for performing such wrapping, we could observe that
@@
 
-\subsubsection{Runtime Exception wrapping Error}
-
-We could find  57 stack traces on 38 projects on which an Error was wrapped in a runtime exception.
+\emph{\textbf{Runtime Exception wrapping Error.}} We could find  57 stack traces 
+on 38 projects on which an Error was wrapped in a runtime exception.
 Manually inspecting each exception stack trace, we could observe that 44% of the wrappings were performed 
@@
 
-  \begin{tabular}{|p{7.9cm} p{0.1cm}|}
-%\begin{tabular}{ll}
+ %\begin{tabular}{|p{7.9cm} p{0.1cm}|}
+\begin{tabular}{ll}
     \hline
@@
   \end{tabular}
-\caption{Wrapping examples}
+\caption{Examples of Cross-type wrappings}
 \label{tab:wrapping01}
@@
 
-\subsubsection{Checked wrapping Runtime Exception}
-
-Manually inspecting the stack traces we could observe that most of the wrappings were performed 
+\emph{\textbf{Checked wrapping Runtime Exception.}} Manually inspecting the stack 
+traces we could observe that most of the wrappings were performed 
 by RoboSpice (an android library that supports the creation of asynchronous long running tasks.
@@
 
-\subsubsection{Checked Exception wrapping Error}
-
-Almost all of these wrappings were also caused by the reflection library 
+\emph{\textbf{Checked Exception wrapping Error.}} Almost all of these wrappings
+ were also caused by the reflection library 
 used by Android applications. The method responsible for the wrappings
@@
 
-\subsubsection{Error wrapping Checked Exception}
-
-Some of these wrappings were performed by methods defined by java.lang.Class methods such as 
+\emph{\textbf{Error wrapping Checked Exception.}} Some of these wrappings were performed by methods defined by java.lang.Class methods such as 
 at java.lang.Class.getDeclaredFields(Native Method)
@@
 
-\subsubsection{Error wrapping Runtime}
-
-After manually inspecting the code associated to such wrappings we discovered that all of them were 
+\emph{\textbf{Error wrapping Runtime.}} After manually inspecting the code associated to such wrappings we discovered that all of them were 
 caused by static initializers. If any exception is thrown in the context of a static initializer (i.e., static block) 
@@
 
-\subsubsection{Multiple Wrappings}
-
-We could also observe that in 205 stacks found in Android repository the exception was wrapped 
-more then once. Figure ~\ref{fig:sizewrapbhists}  illustrates the number of wrappings per stack, both
- in Java and Android studied repositories. One interesting multiple-wrapping found in Android repo was the 
+\emph{\textbf{Multiple Wrappings.}} We could also observe that in 205 stacks found
+ in Android repository the exception was wrapped 
+more then once. One interesting multiple-wrapping found in Android repo was the 
 following: Runtime-Checked-Runtime-Checked-Runtime-Checked-Runtime.
+
+\noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
+Odd wrappings were detected (e.g., a checked exception wrapping an OutOfMemoryError)
+reveal fragilities on Java Exception Model.
+}
+\end{minipage}}
 
@@
 
-\emph{"Everybody hates thinking about exceptions,                   							because they’re not supposed to happen."}
+\emph{``Everybody hates thinking about exceptions, because they are not supposed to happen."}
 \textbf{Brian Foote}
+
+\subsection{On the need of a rescue to the almost infeasible task of preventing uncaught exceptions}
+In this study we could observe undocumented runtime exceptions thrown by third party code,
+and even undocumented checked exception thrown by Java supporting library.
+Such undocumented exceptions make difficult, and most of the times infeasible
+for the client code to protect against `unforeseen" situations that may happen 
+while calling a library code.
+
+One may think that the solution for the uncaught exceptions 
+may be to define a general handler, 
+which is responsible for handling any exception that is not
+adequately handled inside the applications. Although this 
+solution may prevent  the system of abruptly crashing,
+ such general handler will not have enough
+contextual information to adequately handle the exception, 
+and only tasks possible are: to present a message to the user
+ and restart the application.
+
+Such handler does not replace a carefully designed exception 
+handling policy, which can hardly be designed in the absence of 
+third-party documentation on the exceptions that
+may the thrown by it.
+
+On the other hand documenting runtime exceptions is a tedious and error prone task, to help developers
+mitigate this problem tools should be developed to automate the documentation of runtime exceptions
+scraping from library code, few solutions in this directions have been proposed so far ~\cite{van2005combining}.
 
@@
 
-\subsection{On the need of a rescue to the almost Infeasible task of preventing uncaught exceptions}
-In this study we could observe undocumented runtime exceptions thrown by third party code,
-and even undocumented checked exception thrown by Java supporting library.
-Such undocumented exceptions make difficult, and most of the times infeasible
-for the client code to protect against "unforeseen" situations that may happen 
-while calling a library code.
-
-One may think that the solution for the uncaught exceptions 
-may be to define a general handler, 
-which is responsible for handling any exception that is not
-adequately handled inside the applications. Although this 
-solution may prevent  the system of abruptly crashing,
- such general handler will not have enough
-contextual information to adequately handle the exception, 
-and only tasks possible are: to present a message to the user
- and restart the application.
-
-Such handler does not replace a carefully designed exception 
-handling policy, which can hardly be designed in the absence of 
-third-party documentation on the exceptions that
-may the thrown by it.
-
-On the other hand documenting runtime exceptions is a tedious and error prone task, to help developers
-mitigate this problem tools should be developed to automate the documentation of runtime exceptions
-scraping from library code, few solutions in this directions have been proposed so far ~\cite{van2005combining}.
-
-\subsection{On the Null Pointer Problem}
+\subsection{On the null pointer problem}
 
@@
 
-"I call it my billion-dollar mistake. It was the invention of the null reference in 1965 [...] I couldn’t resist the temptation to put 
+\emph{``I call it my billion-dollar mistake. It was the invention of the null reference in 1965 [...] I couldn’t resist the temptation to put 
 in a null reference, simply because it was so easy to implement. This has led to innumerable errors, vulnerabilities, and system 
-crashes, which have probably caused a billion dollars of pain and damage in the last forty year."
+crashes, which have probably caused a billion dollars of pain and damage in the last forty year."}
 

### 2014-08-31 android-stacks.tex
changes all over

@@
 \author{\IEEEauthorblockN{Roberta Coelho\IEEEauthorrefmark{1},
-Georgios Gousios\IEEEauthorrefmark{2}, Arie van Deursen\IEEEauthorrefmark{2},
-Lucas Almeida\IEEEauthorrefmark{1}}
+Georgios Gousios\IEEEauthorrefmark{2}, 
+Arie van Deursen\IEEEauthorrefmark{2},
+Lucas Almeida\IEEEauthorrefmark{1}
+}
 \IEEEauthorblockA{\IEEEauthorrefmark{1}Federal University of Rio Grande do Norte\\
@@
 of exception stack traces enabled to consistently reveal   
-fragilities of exception handling code: 
+fragilities of the exception-related code: 
 (1) undocumented runtime exceptions signaled by third party code,
@@
 Along the recent years, we have witnessed an astonishing increase in the number 
-mobile applications being developed and some of them being largely used. 
+mobile applications being developed and some of them becoming largely used. 
 Such increase is motivated by pervasiveness of the World Wide Web 
@@
 far beyond of the basic calls and textual messages, on the other hand
-they have to face new threats to application robustness imposed by mobile environment.
+they have to face threats to application robustness.
 Such applications have to cope with an increasing number of abnormal
-computation states, arising from: programming errors;
+computation states, arising from:
  noisy user inputs; faults in underlying middleware or hardware; 
-faults on third party services and libraries, compatibility issues [ref], 
-connection with devices e.g., as GPS, camera, memory and battery restrictions
-among others.
+faults on third party services and libraries; compatibility issues [ref]; 
+connection with devices e.g., as GPS, camera; memory and battery restrictions; 
+ programming errors, among others.
 
@@
 have shown that it is generally poorly understood and the less tested parts of the system ~\cite{coelho2011unveiling}.
-As a consequence that the exception handling code may become itself a source of faults, 
+As a consequence that the exception-related code may become itself a source of faults, 
 and may affect the system the other way around, making the system even less robust ~\cite{robillard2003static,coelho2011unveiling}.
@@
 Our conjecture is that several Android applications may be suffering from similar 
-fragilities on the exception handling code (of the applications themselves or 
+fragilities on the exception-related code (of the applications themselves or 
 the underlaying platform), and that such fragilities may be leading to
-decrease the application robustness, leading for instance to the Uncaught Exception problem
- are one of the main causes of application crashes ~\cite{jo2004uncaught} ~\cite{zhang2012amplifying}.
+decrease the application robustness, leading for instance to the Uncaught Exception problem -
+ which is one of the main causes of application crashes ~\cite{jo2004uncaught} ~\cite{zhang2012amplifying}.
 
@@
 
-The goal of this study was to extract exception stack traces embedded on issues of Android projects 
+The goal of this mining study was to extract exception stack traces embedded on issues of Android projects 
 (hosted in GitHub and GoogleCode), looking for its 
@@
 
-  \item[RQ2] \noindent\emph{Which patterns of exception (mis)-use emerge?}
-Based on a set of best practices compiled in this study, the study investigates whether characteristics on exception stack traces  
-can pinpoint best practices violations. Characteristics of exception stack traces and detected violations can reveal fragilities
- on the exception handling code, which threats the application dependability.
+%Which patterns of exception (mis)use emerge?
+  \item[RQ2] \noindent\emph{Which fragilities on the exception-related code emerge from them?}
+Based on a set of best practices compiled in this study, we investigated whether characteristics of exception stack traces  
+can pinpoint best practices violations. Such  violations reveal fragilities
+on the exception-related code, which threats the application dependability.
 
@@
 %of exception-related system crashes.
+
+%Characteristics of exception stack traces and detected violations can reveal fragilities
+% on the exception-related code, which threats the application dependability.
 
@@
 
-\noindent\emph{Repository Connectors}  This component is responsible to connect to
+\noindent\emph{Repository Connectors.}  This component is responsible to connect to
  the host repositories in order to collect project issues. The GitHub Connector connected
@@
 
-\noindent\emph{Exception Stack Trace Distiller} 
+\noindent\emph{Exception Stack Trace Distiller.} 
 Distills the information that composes a stack trace and storing the results in a
@@
 
-\noindent\emph{Exception Signaler Classifier} 
+\noindent\emph{Exception Signaler Classifier.} 
 This module is responsible for classifying each signaler according 
@@
 
-\subsubsection{Manual Inspections}
+\subsection{Manual Inspections}
 This study was also based in a set of manual inspection tasks. Besides the 
@@
 %conducted to answer these questions. The mixed-methods approach is composed by the following steps:
+
+%\begin{figure*}
+%\centering
+%\includegraphics[width=\hsize]{overview.pdf}
+%\caption{Overview of the study based on mixed approaches.}
+%\label{fig:overviewfig}
+%\end{figure*}
 
@@
  However, the C-code implementation did throw a "checked exception" NoSuchMethodException, 
-violating the declaration. The Java compiler, could not detect this violation because does 
-not perform static exception  checking on native methods. This type of bugs is hard to diagnose
-because the developer usualy do not have access to the native code (written in other language).
-
-Consequently, since it was not expected by the programmer, when such method throws 
+violating the declaration. The Java compiler, could not detect this violation because it does 
+not perform static exception checking on native methods. This type of bug is hard to diagnose
+because the developer usualy do not have access to the native implementations.
+
+Consequently, since it is not expected by the programmer, when such method throws 
 this exception, such undocumented exception may remain
-uncaught and cause the app to crash, or be mistakenly handled by subsumption - 
+uncaught and cause the app to crash, or may be mistakenly handled by subsumption - 
 The exception stack traces reporting this scenario is actually a real bug of Android 
@@
 \noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
-  0,8\% of the methods documented the explicitly signaled runtime exceptions.
+ Only 0,8\% of the methods documented the explicitly signaled runtime exceptions.
   The study also detected an undocumented checked exception signaled by
@@
 
-{\footnotesize
-\begin{verbatim}
-
-1. private Activity performLaunchActivity(...) {
-2.  try {
-3.   ...
-4.   ...callActivityOnCreate(...);
-5.  } catch (Exception e) {
-6.    ...
-7.   throw new RuntimeException(..., e);
-8.  }
-9.}
-\end{verbatim}
-}
+%MOVED TO THE PICTURE
+%{\footnotesize
+%\begin{verbatim}
+%1. private Activity performLaunchActivity(...) {
+%2.  try {
+%3.   ...
+%4.   ...callActivityOnCreate(...);
+%5.  } catch (Exception e) {
+%6.    ...
+%7.   throw new RuntimeException(..., e);
+%8.  }
+%9.}
+%\end{verbatim}
+%}
 
@@
 
-{\footnotesize
-\begin{verbatim}
-protected void done() {
- ...
-   try {
-      ...
-   } catch (InterruptedException e) {
-      android.util.Log.w(..., e);
-   } catch (ExecutionException e) {
-      throw new RuntimeException("...",e.getCause());
-   } catch (CancellationException e) {
-      ...
-   } catch (Throwable t) {
-     throw new RuntimeException("...", t);
-   }
-    ...
- }
-}
-\end{verbatim}
-}
+%{\footnotesize
+%\begin{verbatim}
+%protected void done() {
+% ...
+ %  try {
+ %     ...
+  % } catch (InterruptedException e) {
+  %    android.util.Log.w(..., e);
+  % } catch (ExecutionException e) {
+  %    throw new RuntimeException("...",e.getCause());
+  % } catch (CancellationException e) {
+  %    ...
+ %  } catch (Throwable t) {
+ %    throw new RuntimeException("...", t);
+ %  }
+ %   ...
+% }
+%}
+%\end{verbatim}
+%}
 
@@
     \hline
- \bfseries{Runtime-Error Wrapping Examples} &   \\  %\bfseries{\#}
-    \hline
-java.lang.RuntimeException - java.lang.OutOfMemoryError & \\ %30
-java.lang.RuntimeException - java.lang.NoClassDefFoundError &\\ %6
-java.lang.RuntimeException - java.lang.UnsatisfiedLinkError) & \\ %5
-java.lang.RuntimeException -  java.lang.StackOverflowError)  & \\ %4
-\hline
-\bfseries{Checked-Runtime Wrapping Examples} &  \\
- \hline
-...android.robospice.CacheSavingException - java.lang.ClassCastException &  \\ %6
-java.lang.reflect.InvocationTargetException - java.lang.NullPointerException &   \\ %3
-java.lang.reflect.InvocationTargetException - ...ResourcesdNotFoundException &  \\ %2
-java.sql.SQLException - android.database.sqlite.SQLiteDiskIOException &  \\ %2
- \hline
-\bfseries{Checked-Error Wrapping Examples} & \\
- \hline
-java.lang.reflect.InvocationTargetException - java.lang.OutOfMemoryError &  \\ %5
-java.lang.Exception - java.lang.OutOfMemoryError &  \\ %1
-java.lang.reflect.InvocationTargetException - java.lang.NoSuchMethodError &	 \\ %1
-java.lang.reflect.InvocationTargetException - java.lang.UnsatisfiedLinkError&	 \\ %1
-\hline
-\bfseries{Error-Checked Wrapping Examples} & \\
- \hline
-java.lang.NoClassDefFoundError - java.lang.ClassNotFoundException &  \\ %4
-java.lang.AssertionError - javax.crypto.ShortBufferException) &  \\ %3
-\hline
-\bfseries{Error-Runtime Wrapping Examples} &   \\
- \hline
-java.lang.ExceptionInInitializerError - java.lang.NullPointerException &  \\ %9
-java.lang.ExceptionInInitializerError - java.lang.IllegalArgumentException &	 \\ %2
+ id & \bfseries{Runtime-Error Wrapping Examples}    \\  %\bfseries{\#}
+    \hline
+1 & java.lang.RuntimeException - java.lang.OutOfMemoryError  \\ %30
+2 &java.lang.RuntimeException - java.lang.NoClassDefFoundError \\ %6
+3 &java.lang.RuntimeException - java.lang.UnsatisfiedLinkError)  \\ %5
+4& java.lang.RuntimeException -  java.lang.StackOverflowError)   \\ %4
+\hline
+ & \bfseries{Checked-Runtime Wrapping Examples}   \\
+ \hline
+5& ...android.robospice.CacheSavingException - java.lang.ClassCastException   \\ %6
+6&java.lang.reflect.InvocationTargetException - java.lang.NullPointerException    \\ %3
+7&java.lang.reflect.InvocationTargetException - ...ResourcesdNotFoundException  \\ %2
+8&java.sql.SQLException - android.database.sqlite.SQLiteDiskIOException   \\ %2
+ \hline
+& \bfseries{Checked-Error Wrapping Examples}  \\
+ \hline
+9&java.lang.reflect.InvocationTargetException - java.lang.OutOfMemoryError  \\ %5
+10&java.lang.Exception - java.lang.OutOfMemoryError   \\ %1
+11&java.lang.reflect.InvocationTargetException - java.lang.NoSuchMethodError 	 \\ %1
+12&java.lang.reflect.InvocationTargetException - java.lang.UnsatisfiedLinkError	 \\ %1
+\hline
+& \bfseries{Error-Checked Wrapping Examples}  \\
+ \hline
+13&java.lang.NoClassDefFoundError - java.lang.ClassNotFoundException   \\ %4
+14&java.lang.AssertionError - javax.crypto.ShortBufferException)   \\ %3
+\hline
+& \bfseries{Error-Runtime Wrapping Examples}    \\
+ \hline
+15&java.lang.ExceptionInInitializerError - java.lang.NullPointerException   \\ %9
+16&java.lang.ExceptionInInitializerError - java.lang.IllegalArgumentException 	 \\ %2
  \hline
@@
 
-\begin{figure} \centering \includegraphics[scale=0.4]{codeexample2.pdf}
-\caption{Code snippets illustrating fragilities on the exception handling code} \label{fig:exchier} \end{figure}
+\begin{figure*} \centering \includegraphics[scale=0.55]{codeexamples.pdf}
+\caption{Code snippets illustrating fragilities on the exception-related code} \label{fig:exchier} \end{figure*}
 
@@
 
-Hence, when (mis)applied, the exception wrapping can make the exception handling code
+Hence, when (mis)applied, the exception wrapping can make the exception-related code
  more complex and lead to what we call an \emph{exception handling confusion problem}.
@@
 
-\textit{Tools Support for Detecting Faults on Java Exception Handling Code.}
-
-Manually analyzing the exception handling code looking for defects, can easily 
-become infeasible [ref]. Hence, some tools have been proposed to detect faults 
+\textit{Tools Support for Detecting Faults on Java Exception-Related Code.}
+
+Manually analyzing the exception-related code and looking for defects, can easily 
+become infeasible ~\cite{Robil00}. Hence, some tools have been proposed to detect faults 
 on the exception handling code either statically (cabral, roberta,nelio,chang) 

### 2014-09-01 android-stacks.tex
including more data from manual inspections of CG collected by Lucas

@@
 last app version, the previous versions were analyzed. When the exception was
- not found on the previous versions, the exception was discarded.  Only 3 exceptions 
-remained undefined and were removed from this study.
+ not found on the previous versions, the exception was discarded.  Only 31 exceptions 
+remained undefined which were reported on 61 exception stack traces (as presented in Table ~\ref{tab:typeroottab}).
 
@@
 
+%SELECTS:
+%SELECT COUNT(distinct(repo)) FROM STACKTRACESUMMARY WHERE MAIN_CAUSE LIKE '%java.lang.NullPointerException%'
+%and repo !='android';
+
+%SELECT COUNT(*) FROM STACKTRACESUMMARY WHERE MAIN_CAUSE LIKE '%java.lang.IllegalStateException%'
+%and repo !='android' AND MAINSIGNALER_FROM LIKE '%ANDROID%';
+
 \begin{table*}
@@
     \hline
-java.lang.NullPointerException            & 254 /  78  & 52,70\%  & 1225 / 439  & 30,08\% & 473 /  26  & 18 & 595 & 137 \\  %LIBORAPP=408
-java.lang.IllegalStateException          & 99 / 21  & 20,54\%  & 234 / 44  & 5,75\%  & 165 & 12 & 36  & 20  \\
-java.lang.IllegalArgumentException          & 94 / 48  & 19,50\% & 255 / 98  & 6,26\% & 146 & 6  & 64  & 39   \\
-java.lang.RuntimeException                & 91 / 31 & 18,88\% & 232 / 87  & 5,70\%  & 167 & 1  & 47  & 17   \\
-java.lang.OutOfMemoryError                 & 56  / 22  & 11,62\% & 180 / 57 & 4,42\% & 121 & 15 & 17  & 23   \\
-java.lang.NoClassDefFoundError           & 52  / 15 & 10,79\%  & 73 / 21   & 1,79\%  & 9   & 0  & 37  & 26   \\
-java.lang.ClassCastException             & 49  / 15 & 10,17\%  & 94 / 36   & 2,31\%  & 43  & 0  & 40  & 11   \\
-java.lang.IndexOutOfBoundsException       & 47  / 15 & 9,75\%   & 127 / 39  & 3,12\% & 47  & 0  & 71  & 8    \\
-java.lang.NoSuchMethodError               & 40 / 14  & 8,30\%  & 57 / 23  & 1,40\%  & 9   & 0  & 39  & 9   \\
-java.util.ConcurrentModificationException & 38 / 5 & 7,88\% & 54 / 11  & 1,33\%   & 5   & 0  & 43  & 6   \\
+java.lang.NullPointerException            & 254 /  78  & 52,70\%  & 1225 / 439  & 30,08\% & 473 /  26  & 18 /  & 595 / 241 & 137 / 169 \\  %LIBORAPP=408
+java.lang.IllegalStateException          & 99 / 21  & 20,54\%  & 234 / 44  & 5,75\%  & 165 / 15 & 12  / 0 & 36  / 5 & 20 / 24  \\
+java.lang.IllegalArgumentException          & 94 / 48  & 19,50\% & 255 / 98  & 6,26\% & 146 / 41 & 6 / 0 & 64 / 31  & 39  / 26  \\
+java.lang.RuntimeException                & 91 / 31 & 18,88\% & 232 / 87  & 5,70\%  & 167 / 33 & 1 / 0 & 47 / 17 & 17 / 37   \\
+java.lang.OutOfMemoryError                 & 56  / 22  & 11,62\% & 180 / 57 & 4,42\% & 121 / 18 & 15 / 0 & 17 / 18  & 23 / 13  \\
+java.lang.NoClassDefFoundError           & 52  / 15 & 10,79\%  & 73 / 21   & 1,79\%  & 9 / 1   & 0 / 0  & 37 / 9  & 26 / 11   \\
+java.lang.ClassCastException             & 49  / 15 & 10,17\%  & 94 / 36   & 2,31\%  & 43 / 3  & 0  / 0 & 40 / 15  & 11 / 18  \\
+java.lang.IndexOutOfBoundsException       & 47  / 15 & 9,75\%   & 127 / 39  & 3,12\% & 47 / 2  & 0/0  & 71/22  & 8 / 14    \\
+java.lang.NoSuchMethodError               & 40 / 14  & 8,30\%  & 57 / 23  & 1,40\%  & 9  /  0  & 0  / 0 & 39 /  17    & 9 / 6  \\
+java.util.ConcurrentModificationException & 38 / 5 & 7,88\% & 54 / 11  & 1,33\%   & 5  / 0  & 0  / 0 & 43 / 3 & 6 / 7  \\
 
@@
 
+%SELECT COUNT(*) FROM STACKTRACESUMMARY WHERE
+% repo !='android'  AND MAIN_EXTENDS LIKE 'RUNTIME%'
+
 \begin{table}
@@
 
-Runtime	&	1075	&	56	&	1446	&	374	&	2951\\
-Error	&	144	&	34	&	168	&	105	&		451	\\
-Checked	&	110	&	230	&	139	&	145	&	624	\\
-Throwable	&	0	&	0	&	1	&	0	&	1	\\
-Undefined	&	1	&	0	&	6	&	8	&	15	\\
+Runtime	&	1075 	&	56	&	1446	&	374	&	2951 / 956\\
+Error	&	144	&	34	&	168	&	105	&		451 /  250	\\
+Checked	&	110	&	230	&	139	&	145	&	624  /  734	\\
+Throwable	&	0	&	0	&	1	&	0	&	1 / 1	\\
+Undefined	&	1	&	0	&	6	&	8	&	15 / 46	\\ %22 DISTINCT EXCEPTIONS AND 9 IN GIT
     \hline

### 2014-09-01 android-stacks.tex
including last data from manual inspections of CG collected by Lucas

@@
 
+%SELECT COUNT(*) FROM STACKTRACESUMMARY WHERE
+%repo !='android'  AND MAIN_EXTENDS LIKE 'RUNT%' AND MAINSIGNALER_FROM like 'APP%';
+
 \begin{table}
@@
 
-Runtime	&	1075 	&	56	&	1446	&	374	&	2951 / 956\\
-Error	&	144	&	34	&	168	&	105	&		451 /  250	\\
-Checked	&	110	&	230	&	139	&	145	&	624  /  734	\\
-Throwable	&	0	&	0	&	1	&	0	&	1 / 1	\\
-Undefined	&	1	&	0	&	6	&	8	&	15 / 46	\\ %22 DISTINCT EXCEPTIONS AND 9 IN GIT
+Runtime	&	1075 / 160	&	56 	&	1446 / 397	&	374	/  386  &	2951 / 943\\
+Error	       &	 144	 / 32              &	34	&	168	 / 134               &	105	 / 74           &	451 /  240	\\
+Checked	&	110	   / 39           &	230	&	139	  / 174            &	145	 / 466           &	624  /  679	\\
+Throwable	&	0 /0	       &	0	&	1 / 1	             &	0	     / 0               &	1 / 1	\\
+Undefined	&	1 / 1 	&	0	&	6 / 	12		&	8	/ 32	   &	15 / 45	\\ %22 DISTINCT EXCEPTIONS AND 9 IN GIT
     \hline
@@
 
+%SELECT COUNT(*) FROM STACKTRACESUMMARY WHERE
+%repo !='android'  AND ROOT_EXTENDS LIKE 'CHECKED%' AND EXCEPTION_EXTENDS LIKE 'RUNTIME'
+
+%SELECT COUNT(*) FROM STACKTRACESUMMARY WHERE
+%repo !='android'  AND ROOT_EXTENDS LIKE 'CHEC%' AND EXCEPTION_EXTENDS LIKE 'RUN%';
+
+%SELECT COUNT(*) FROM STACKTRACESUMMARY WHERE
+%repo !='android'  AND ROOT_EXTENDS LIKE 'CHEC%' AND EXCEPTION_EXTENDS LIKE 'RUN%'
+%and SIGNALER_FROM LIKE 'LIB%';
+
 \begin{table*}
@@
     \hline
-      Runtime &  Checked   & 70 & 114 & 66 &  0  & 19 &  29 \\
-      Runtime   &  Error   & 38  & 57 & 50 &  0  & 6 &  1 \\
-      Checked &  Runtime   & 12 & 22 & 4 &  0  & 9 &  9 \\
-      Checked & Error      & 7 &  8 & 5 &  0  & 0 &  3 \\
-      Error & Checked      & 4  &  7 & 0 &  7  & 0 &  0     \\
-      Error & Runtime     & 6  &  13   & 1 &  1  & 0 &  11    \\
+      %Runtime &  Checked   & 70  / 18 & 114 /  34 &  66 /  9  &  0 /  & 19 / 19 &  29 / 6 \\
+
+      Runtime &  Checked   & 88 & 148 &  75  &  0   & 38 &  35 \\
+
+      %Runtime   &  Error   & 38  \ 8  & 57    &  50 \ 8  &  0   & 6 \ 2  &  1   /  0\\
+      Runtime   &  Error   & 46  &  67    &  58  &  0   & 8  &  1   \\
+
+%      Checked &  Runtime   & 12 \ 5  & 22 \ 9 & 4 &  0  & 9 \7 &  9 \ 2 \\
+      Checked &  Runtime   & 17  & 31 & 4 &  0  & 16 &  11 \\
+
+%      Checked & Error      & 7 \ 1 &  8 & 5  &  0  & 0 \ 1 &  3 \0 \\
+
+      Checked & Error      & 8 &  9  & 5  &  0  &  1 &  3  \\
+
+ %     Error & Checked      & 4 \ 10  &  7 & 0  \ 6  &  7  & 0 \6 &  0   \ 8    \\
+     Error & Checked      & 14 &  27 &  6  &  7  &  6 &   8    \\
+
+%      Error & Runtime     & 6 \ 2  &  13   & 1 &  1  & 0 \ 1 &  11 \ 3    \\
+
+      Error & Runtime     & 8  &  17   & 1 &  1  & 1 &  14    \\
+
   \hline

### 2014-09-01 android-stacks.tex
improvements all over

@@
 \author{\IEEEauthorblockN{Roberta Coelho\IEEEauthorrefmark{1},
+Lucas Almeida\IEEEauthorrefmark{1},
 Georgios Gousios\IEEEauthorrefmark{2}, 
-Arie van Deursen\IEEEauthorrefmark{2},
-Lucas Almeida\IEEEauthorrefmark{1}
+Arie van Deursen\IEEEauthorrefmark{2}
 }
@@
 Android projects available in GitHub (482 applications) and Google code (157 applications)
-Overall, our research set includes 32,582 issues in GitHub and 41,674 issues in Googlecode from which (4,208)  (1,986) in GC stack traces 
-were extracted. Such stack traces were distilled and analyzed in 
+Overall, our research set includes 32,582 issues in GitHub and 41,674 issues in Googlecode 
+from which 6,005 exception stack traces were extracted. Such stack traces were distilled and analyzed in 
 combination with source code and bytecode analysis. This study 
@@
 
+%482 applications) and Google code (157 applications)
+
 To do so we perform a post mortem analysis of the exception stack traces embedded on all issues
@@
  which we develop specifically for this study, we mine stack traces from issues across all Android projects on
-GitHub and Googlecode. Overall, we analyzed the XXXX issues extracted from on XXX Android projects defined in
-GitHub and XXX projects defined in Googlecode. From these issues,  XXX stack traces were extracted and used as
+GitHub and Googlecode. Overall, we analyzed all issues extracted from on 482 Android projects defined in
+GitHub and 157 projects defined in Googlecode. Overall it corresponds to 74,256 issues (32,582 issues from GitHub projects 
+and 41,674 issues from Googlecode projects) from which  6,005 stack traces were extracted and used as
  a source of information for our analysis. The stack trace analysis was augmented by additional information
@@
 
-This set of 482 projects contained overall 32,582 issues from which 4,208 exception stack traces 
+%4,208
+
+This set of 482 projects contained overall 32,582 issues from which 4,042 exception stack traces 
 were extracted. Issues on Github are different from issues on dedicated bug tracking tools such as 
@@
 reporting at least one stack trace, and as a result we could identify 157. 
-Android projects.  From this set all issues (41,674) were analyzed and 1,986 stack traces were extracted.
+Android projects.  All issues created on such projects (i.e., 41,674 issues) were analyzed and 1,963 stack traces 
+were extracted.
 
@@
 contrast to existing issue parsing solutions such as Infozilla, the parser
-created in this work can extract all causes related to an exception on a stack,
-and stack traces embedded in logs files.
-AS PRESENTED IN FIGURE X
+created in this work can discover stack traces embedded in logs files \footnote{In several 
+exception stack traces, the exception frames were preceeded by logging information e.g., "03-01 15:55:01.609 (7924): at android.app.ActivityThread.access\$600(ActivityThread.java:127)" which could not be detected by existing tools.)}
+and identify the root causes and signalers associated to each exception stack trace.
 
@@
  not found on the previous versions, the exception was discarded.  Only 31 exceptions 
-remained undefined which were reported on 61 exception stack traces (as presented in Table ~\ref{tab:typeroottab}).
+remained undefined - such exceptions were reported on 61 exception stack traces as 
+presented in Table ~\ref{tab:typeroottab}.
 
@@
 part of Android, the application itself, libcore, or a third-party library as presented in Table  \ref{tab:signalers}
- \footnote{Nine exception stack traces contained only elements of JVM and were discarded from the analysis}.
+
+% \footnote{Nine exception stack traces contained only elements of JVM and were discarded from the analysis}.
 
@@
 
+%java.lang.NullPointerException            & 254 /  78  & 52,70\%  & 1225 / 439  & 30,08\% & 473 /  26  & 18 /  & 595 / 241 & 137 / 169 \\  %LIBORAPP=408
+%java.lang.IllegalStateException          & 99 / 21  & 20,54\%  & 234 / 44  & 5,75\%  & 165 / 15 & 12  / 0 & 36  / 5 & 20 / 24  \\
+%java.lang.IllegalArgumentException          & 94 / 48  & 19,50\% & 255 / 98  & 6,26\% & 146 / 41 & 6 / 0 & 64 / 31  & 39  / 26  \\
+%java.lang.RuntimeException                & 91 / 31 & 18,88\% & 232 / 87  & 5,70\%  & 167 / 33 & 1 / 0 & 47 / 17 & 17 / 37   \\
+%java.lang.OutOfMemoryError                 & 56  / 22  & 11,62\% & 180 / 57 & 4,42\% & 121 / 18 & 15 / 0 & 17 / 18  & 23 / 13  \\
+%java.lang.NoClassDefFoundError           & 52  / 15 & 10,79\%  & 73 / 21   & 1,79\%  & 9 / 1   & 0 / 0  & 37 / 9  & 26 / 11   \\
+%java.lang.ClassCastException             & 49  / 15 & 10,17\%  & 94 / 36   & 2,31\%  & 43 / 3  & 0  / 0 & 40 / 15  & 11 / 18  \\
+%java.lang.IndexOutOfBoundsException       & 47  / 15 & 9,75\%   & 127 / 39  & 3,12\% & 47 / 2  & 0/0  & 71/22  & 8 / 14    \\
+%java.lang.NoSuchMethodError               & 40 / 14  & 8,30\%  & 57 / 23  & 1,40\%  & 9  /  0  & 0  / 0 & 39 /  17    & 9 / 6  \\
+%java.util.ConcurrentModificationException & 38 / 5 & 7,88\% & 54 / 11  & 1,33\%   & 5  / 0  & 0  / 0 & 43 / 3 & 6 / 7  \\
+
+%java.lang.IllegalStateException          & 99 / 21  & 20,54\%  & 234 / 44  & 5,75\%  & 165 / 15 & 12  / 0 & 36  / 5 & 20 / 24  \\
+%java.lang.IllegalArgumentException          & 94 / 48  & 19,50\% & 255 / 98  & 6,26\% & 146 / 41 & 6 / 0 & 64 / 31  & 39  / 26  \\
+%java.lang.RuntimeException                & 91 / 31 & 18,88\% & 232 / 87  & 5,70\%  & 167 / 33 & 1 / 0 & 47 / 17 & 17 / 37   \\
+%java.lang.OutOfMemoryError                 & 56  / 22  & 11,62\% & 180 / 57 & 4,42\% & 121 / 18 & 15 / 0 & 17 / 18  & 23 / 13  \\
+%java.lang.NoClassDefFoundError           & 52  / 15 & 10,79\%  & 73 / 21   & 1,79\%  & 9 / 1   & 0 / 0  & 37 / 9  & 26 / 11   \\
+%java.lang.ClassCastException             & 49  / 15 & 10,17\%  & 94 / 36   & 2,31\%  & 43 / 3  & 0  / 0 & 40 / 15  & 11 / 18  \\
+%java.lang.IndexOutOfBoundsException       & 47  / 15 & 9,75\%   & 127 / 39  & 3,12\% & 47 / 2  & 0/0  & 71/22  & 8 / 14    \\
+%java.lang.NoSuchMethodError               & 40 / 14  & 8,30\%  & 57 / 23  & 1,40\%  & 9  /  0  & 0  / 0 & 39 /  17    & 9 / 6  \\
+%java.util.ConcurrentModificationException & 38 / 5 & 7,88\% & 54 / 11  & 1,33\%   & 5  / 0  & 0  / 0 & 43 / 3 & 6 / 7  \\
+%
+%java.lang.NullPointerException	&	254	/	78	&	52,70\%	&	1225	/	439	&	30,08\%	&	473	/	52	&	18	/	2	&	595	/	241	&	137	/	143	\\	%LIBORAPP=408
+%java.lang.IllegalStateException	&	99	/	21	&	20,54\%	&	234	/	44	&	5,75\%	&	165	/	20	&	12	/	19	&	36	/	5	&	20	/	19		\\
+%java.lang.IllegalArgumentException	&	94	/	48	&	19,50\%	&	255	/	98	&	6,26\%	&	146	/	49	&	6	/	6	&	64	/	31	&	18	/	26		\\
+%java.lang.RuntimeException	&	91	/	31	&	18,88\%	&	232	/	87	&	5,70\%	&	167	/	36	&	1	/	1	&	47	/	17	&	17	/	34		\\
+%java.lang.OutOfMemoryError	&	56	/	22	&	11,62\%	&	180	/	57	&	4,42\%	&	121	/	20	&	15	/	1	&	17	/	18	&	23	/	11		\\
+%java.lang.NoClassDefFoundError	&	52	/	15	&	10,79\%	&	73	/	21	&	1,79\%	&	9	/	1	&	0	/	0	&	37	/	9	&	26	/	11		\\
+%java.lang.ClassCastException	&	49	/	15	&	10,17\%	&	94	/	36	&	2,31\%	&	43	/	12	&	0	/	0	&	40	/	15	&	11	/	9		\\
+%java.lang.IndexOutOfBoundsException	&	47	/	15	&	9,75\%	&	127	/	39	&	3,12\%	&	47	/	6	&	0		0	&	71	/	22	&	8	/	10		\\
+%java.lang.NoSuchMethodError	&	40	/	14	&	8,30\%	&	57	/	23	&	1,40\%	&	9	/	1	&	0	/	0	&	39	/	17	&	9	/	5		\\
+%java.util.ConcurrentModificationException	&	38	/	5	&	7,88\%	&	54	/	11	&	1,33\%	&	5	/	0	&	0	/	0	&	43	/	3	&	6	/	7
+
 \begin{table*}
@@
     \hline
-    \bfseries{Root Exception} &  \multicolumn{2}{c}{\bfseries{Projects (GH/CG)}} &  \multicolumn{2}{c}{\bfseries{Occurrences (GH/CG)}} & \textsf{Android} & \textsf{Libcore} & \textsf{App} & \textsf{Lib} \\
+    \bfseries{Root Exception} &  \multicolumn{2}{c}{\bfseries{Projects (GH\&CG)}} &  \multicolumn{2}{c}{\bfseries{Occurrences (GH\&CG)}} & \textsf{Android} & \textsf{Libcore} & \textsf{App} & \textsf{Lib} \\
     & \bfseries{\#} &  \bfseries{\%} & \bfseries{\# } & \bfseries{\% } &&&&\\
     \hline
-java.lang.NullPointerException            & 254 /  78  & 52,70\%  & 1225 / 439  & 30,08\% & 473 /  26  & 18 /  & 595 / 241 & 137 / 169 \\  %LIBORAPP=408
-java.lang.IllegalStateException          & 99 / 21  & 20,54\%  & 234 / 44  & 5,75\%  & 165 / 15 & 12  / 0 & 36  / 5 & 20 / 24  \\
-java.lang.IllegalArgumentException          & 94 / 48  & 19,50\% & 255 / 98  & 6,26\% & 146 / 41 & 6 / 0 & 64 / 31  & 39  / 26  \\
-java.lang.RuntimeException                & 91 / 31 & 18,88\% & 232 / 87  & 5,70\%  & 167 / 33 & 1 / 0 & 47 / 17 & 17 / 37   \\
-java.lang.OutOfMemoryError                 & 56  / 22  & 11,62\% & 180 / 57 & 4,42\% & 121 / 18 & 15 / 0 & 17 / 18  & 23 / 13  \\
-java.lang.NoClassDefFoundError           & 52  / 15 & 10,79\%  & 73 / 21   & 1,79\%  & 9 / 1   & 0 / 0  & 37 / 9  & 26 / 11   \\
-java.lang.ClassCastException             & 49  / 15 & 10,17\%  & 94 / 36   & 2,31\%  & 43 / 3  & 0  / 0 & 40 / 15  & 11 / 18  \\
-java.lang.IndexOutOfBoundsException       & 47  / 15 & 9,75\%   & 127 / 39  & 3,12\% & 47 / 2  & 0/0  & 71/22  & 8 / 14    \\
-java.lang.NoSuchMethodError               & 40 / 14  & 8,30\%  & 57 / 23  & 1,40\%  & 9  /  0  & 0  / 0 & 39 /  17    & 9 / 6  \\
-java.util.ConcurrentModificationException & 38 / 5 & 7,88\% & 54 / 11  & 1,33\%   & 5  / 0  & 0  / 0 & 43 / 3 & 6 / 7  \\
+
+java.lang.NullPointerException	&	332	&	51,96\%	&	1664	&	27,71\%	&	525	&	20	&	836	&	280	\\
+java.lang.IllegalStateException	&	120	&	18,78\%	&	278	&	4,63\%	&	185	&	31	&	41	&	39	\\
+java.lang.IllegalArgumentException	&	142	&	22,22\%	&	353	&	5,88\%	&	195	&	12	&	95	&	44	\\
+java.lang.RuntimeException	&	122	&	19,09\%	&	319	&	5,31\%	&	203	&	2	&	64	&	51	\\
+java.lang.OutOfMemoryError	&	78	&	12,21\%	&	237	&	3,95\%	&	141	&	16	&	35	&	34	\\
+java.lang.NoClassDefFoundError	&	67	&	10,49\%	&	94	&	1,57\%	&	10	&	0	&	46	&	37	\\
+java.lang.ClassCastException	&	64	&	10,02\%	&	130	&	2,16\%	&	55	&	0	&	55	&	20	\\
+java.lang.IndexOutOfBoundsException	&	62	&	9,70\%	&	166	&	2,76\%	&	53	&	0	&	93	&	18	\\
+java.lang.NoSuchMethodError	&	54	&	8,45\%	&	80	&	1,33\%	&	10	&	0	&	56	&	14	\\
+java.util.ConcurrentModificationException	&	43	&	6,73\%	&	65	&	1,08\%	&	5	&	0	&	46	&	13	\\
+\\
 
@@
   \end{tabular}
-\caption{Root Exceptions occurrences and popularity.}
+\caption{Root Exceptions occurrences and popularity in repositories hosted in Googlecode (GC) and GitHub(GH).}
 \label{tab:toptenandroid}
@@
     \hline
-    \bfseries{Category} &  \bfseries{\% reported on stacks} \\
+    \bfseries{Category} &  \bfseries{\% Occurrences on stacks} \\
     \hline
@@
 
+% \bfseries{Root Cause Type} & \bfseries{Android} & \bfseries{Libcore} & \bfseries{App} & \bfseries{Lib}  & \bfseries{All}\\
+% Runtime	&	1075 / 260	&	56  / 17	&	1446 / 397	&	374	/  316  &	2951 / 943\\
+%Error	       &	 144	 / 44              &	34 / 12	&	168	 / 134              &	105	 / 62           &	451 /  240	\\
+%Checked	&	193	/ 83            &	230 / 84	&	139	  / 174          &	145 / 422	           &	624  / 734	\\
+%Throwable	&	0 /0	       &	0	&	1 / 1	             &	0	     / 0               &	1 / 1	\\
+%Undefined	&	1 / 3 	&	0	&	6 / 	12		&	8 / 30	   &	15 / 45	\\ %22 DISTINCT EXCEPTIONS AND 9 IN GIT
+
+%Runtime	&	1075 / 160	&	56 	&	1446 / 397	&	374	/  386  &	2951 / 943\\
+%Error	       &	 144	 / 32              &	34	&	168	 / 134               &	105	 / 74           &	451 /  240	\\
+%Checked	&	110	   / 39           &	230	&	139	  / 174            &	145	 / 466           &	624  /  679	\\
+%Throwable	&	0 /0	       &	0	&	1 / 1	             &	0	     / 0               &	1 / 1	\\
+%Undefined	&	1 / 1 	&	0	&	6 / 	12		&	8	/ 32	   &	15 / 45	\\ %22 DISTINCT EXCEPTIONS AND 9 IN GIT
+%All		& 1330	&	320	&	1760	&	632	&	4042	\\
+
+%Runtime	&	1235	&	56 	&	1843	&	760  &	3894\\
+%Error	       &	 176              &	34	&	302               &	179          &	691	\\
+%Checked	&	149           &	230	&	313            &	611           &	1303	\\
+%Throwable	&	0        &	0	&	2             &	0	                    &	2	\\
+%Undefined	&	2 	&	0	&	18		& 40	   &	60	\\ %22 DISTINCT EXCEPTIONS AND 9 IN GIT
+%  \hline
+%All		& 1562	& 320	& 	2478	& 1590	& 5950 \\
+
+%FILTRAR LIBCODE:
+%FILTRO ANDROID: 74 (com.android) 122(com.google)
+%SELECT REPO, main_signaler, *
+%  FROM stacktracesummary where mainsignaler_from='LIB_or_APP' AND main_signaler like 'com.google%' ;
+
 \begin{table}
@@
 
-Runtime	&	1075 / 160	&	56 	&	1446 / 397	&	374	/  386  &	2951 / 943\\
-Error	       &	 144	 / 32              &	34	&	168	 / 134               &	105	 / 74           &	451 /  240	\\
-Checked	&	110	   / 39           &	230	&	139	  / 174            &	145	 / 466           &	624  /  679	\\
-Throwable	&	0 /0	       &	0	&	1 / 1	             &	0	     / 0               &	1 / 1	\\
-Undefined	&	1 / 1 	&	0	&	6 / 	12		&	8	/ 32	   &	15 / 45	\\ %22 DISTINCT EXCEPTIONS AND 9 IN GIT
-    \hline
-All		& 1330	&	320	&	1760	&	632	&	4042	\\
+Runtime	&	1335	&	73	&	1843	&	690  &	3894\\
+Error	       &	 188              &	 46	&	302             &	167	           &	691	\\
+Checked	&	276           &	314	&	313          &	567	           &	1358	\\
+Throwable	&	0	       &	0	&	2            &	0	     / 0               &	2	\\
+Undefined	&	4	&	0	&	18		&	38	   &	60	\\
+ \hline
+All		& 1803	&	433	&	2478	&	1462	&	6005	\\
     \hline
@@
 The exception stack traces reporting this scenario is actually a real bug of Android 
-Gingerbread version (which still accounts for 13.6% of devices running Android).
+Gingerbread version (which still accounts for 13.6\% of devices running Android).
+
+Among the Android projects on which this exception stack trace was reported we can cite:
+PageTurner (206 stars, 130 forks, 1184 commits), ActionBarSherlock (6,041 stars, 
+3,436 forks, and 1479 commints), otto (1,618 stars,	342 forks, 	177 commints).
+
+%SELECT repo, *
+% FROM stacktracesummary where stack like '%getDeclaredMethods%';
 
@@
 \end{table*}
+
+TODO: ATUALIZAR DISCUSSAO COM BASE NOS NOVOS DADOS.
 

### 2014-09-01 android-stacks.tex
improvements refs

@@
 %Such increase on smartphones combined with the pervasiveness of the World Wide Web
-%have stimulated the creation of applications which extend the phones capabilities 
+%have stimulated the creation of applications which extend the phones caVALIAR MOVER ESTE STACK TRACES PARA UMA FIGURA QUE MOSTR COMO ELE QUEBRADO... PELO EXCEPTION Mpabilities 
 %far beyond of the basic calls and textual messages.
@@
 decrease the application robustness, leading for instance to the Uncaught Exception problem -
- which is one of the main causes of application crashes ~\cite{jo2004uncaught} ~\cite{zhang2012amplifying}.
+ which is one of the main causes of application crashes ~\cite{jo2004uncaught, zhang2012amplifying}.
 
@@
  a source of information for our analysis. The stack trace analysis was augmented by additional information
-extracted from JDK, Android SO and Android SDK and the applications using byte code and source code analysis.
+extracted from JDK, Android Plaform and the applications using byte code and source code analysis.
  Manual inspection was used to leverage the understanding of exception stack traces and support further discussions and insights.
@@
    \item Undocumented checked exception signaled by native Android library. This finding points to a fragility 
-on the way exceptions on JNI interface functions, which can lead to uncaught exceptions that are difficult do 
+on the way exceptions flow from JNI interface functions, which can lead to uncaught exceptions that are difficult do 
 debug.
@@
   \item  A multitude of programming mistakes were reported as the main causes of stack traces:
- XXX  of reported stack traces were caused by programming mistakes; and XXX of analyzed stack
- traces reported a java.lang.NullPointerException as root cause;
+ 27,71\% of the analyzed exception stack traces reported a java.lang.NullPointerException the root cause;
+and  51,96\% of all projects reported at least one exception stack trace on which the NullPoiterException
+was the root cause.
 
@@
 \section{Background}
+\label{sec:back}
 
@@
 higher-level services to the applications running on the platform) ~\cite{andguide}.
-Android platform is mostly implemented in Java and uses JNI[ref] to access native C/C++ code. 
+Android platform is mostly implemented in Java and uses Java Native Invocation (JNI) to access native C/C++ code. 
 And as mentioned before, it reuses the embedded Java Exception model to detect and handle 
@@
 
-<AVALIAR MOVER ESTE STACK TRACES PARA UMA FIGURA QUE MOSTR COMO ELE É QUEBRADO...
-PELO EXCEPTION MINER E AS INFORMAÇÕES VAO PARA O BANCO...>
+%<AVALIAR MOVER ESTE STACK TRACES PARA UMA FIGURA QUE MOSTR COMO ELE É QUEBRADO...
+%PELO EXCEPTION MINER E AS INFORMAÇÕES VAO PARA O BANCO...>
 
@@
 exceptions~\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,
-bloch2008effective} regardless the exception type used. Such guidelines do not focus on 
-advocating any specific type but rather propose ways to effectively use each of them.
+bloch2008effective}. Such guidelines do not focus on 
+advocating any specific exception type, but rather propose ways to effectively use each of them.
  We compiled a list of the guidelines related to exception types in Java, 
@@
 \section{Empirical Study Design}
+\label{sec:study}
 
@@
 example projects (i.e., toy projects) or tools to support Android development
-(e.g. \textsf{selendroid}, \textsf{roboeletric} --- tools to support the testing of Android apps).
+(e.g. Selendroid, Roboeletric - tools to support the testing of Android apps).
 The filtered set consisted of 482 apps.
@@
 This module analyzed the exception hierarchy of all exceptions thrown on:
- the last version of SDK version 7, and Android Platform and OS (level 19).
+  Android Platform (API level 19), and JDK (version 6).
 
@@
 to its origin (i.e. Android Plaform, Android Libcore, Aplication, Library). 
-We downloaded the source code of Android Platform and OS (level 19)
-and this module extracts the packages hierarchies (recursive list of 
-of Java packages), that composes it. To discover the packages that compose 
-each application, this module extracted the manifest files of each Android apps
+We downloaded the source code of Android Platform (API level 19)
+and this module extracts the recursive list of 
+of Java packages that composes it. To discover the packages that compose 
+the previous released of the API we performed a manual inspection on the Javadoc
+documentation available online for each API level ~\cite{apidocs}.
+To discover the packages composing the each application, this module 
+extracted the manifest files of each Android app
  - which defines the main packages that compose the applications.
@@
 on Android Platform, core libraries nor on applications.
-Table ~ref{signalers} summarizes the signaler classification.
+Table ~\ref{signalers} summarizes the signaler classification.
 
@@
 \subsection{Manual Inspections}
-This study was also based in a set of manual inspection tasks. Besides the 
- identification of Android projects details previously. 
-Manual inspections were used discover the type of exceptions thrown by 
-the applications and libraries we manually inspect the sourcecode or javadoc
- documentation of each of them. When the exception was removed from the 
-last app version, the previous versions were analyzed. When the exception was
- not found on the previous versions, the exception was discarded.  Only 31 exceptions 
+As mentioned before, this study was also based in a set of manual inspection tasks. Besides the 
+ (i) identifying of Android projects, and (ii) the finding packages that compose the 
+previous Android APIs leveles, manual inspections were also used discover the type of associated to some 
+reported on stack traces. Such exceptions could not be automaticaly identified 
+because they were not defined on the last version of the application,
+or were defined on libraries not fount among projects resources.
+In such cases we manually inspect the sourcecode or javadoc
+ documentation available online of each of them. 
+When the exception was not found on the previous versions, the exception where classified 
+in our study as "Undefined".  Only 31 exceptions 
 remained undefined - such exceptions were reported on 61 exception stack traces as 
@@
 \section{The Study Results}
+\label{sec:result}
 
@@
 the the exceptions commonly reported as the root causes of stack traces.
-Table \ref{tab:toptenandroid} illustrates the top 10 root causes found in the study,
+Table \ref{tab:topten} illustrates the top 10 root causes found in the study,
  ranked by the number of distinct project repositories on which they were reported. 
@@
 %distinct projects they were found and the whole set of projects analyzed with at
-%least one stack trace (482). Table~\ref{tab:toptenandroid} 
+%least one stack trace (482). Table~\ref{tab:topten} 
 %and Figure~\ref{fig:androidsignaler} presents the mined data.
@@
 \caption{Root Exceptions occurrences and popularity in repositories hosted in Googlecode (GC) and GitHub(GH).}
-\label{tab:toptenandroid}
+\label{tab:topten}
 \end{table*}
@@
 \emph{\textbf{The high number of NullPointerExceptions.}}
-From this set the java.lang.NullPointerException was the most reported root cause; 1225 stack traces 
-had a java.lang.NullPointerException  as root cause, which corresponds of 30,08\%  of 
-all analyzed stack traces. Considering the number of distinct project repositories, 
-the java.lang.NullPointerException, was reported on at least one issue of 99 projects 
-(which represents 20,54\% of all projects reporting at least one stack trace).
- The NullPointerExceptions are mainly signaled inside Android platform and inside 
-application code, although we also find
-NullPointerExceptions being signaled from third-party libraries. Regarding
+From this set the java.lang.NullPointerException was the most reported root cause; 27,71\% 
+of the analyzed exception stack traces reported a java.lang.NullPointerException the root cause. 
+Considering the frequencly of NullPointerException across projects, we could observe that
+ 51,96\% of all projects reported at least one exception stack trace on which the NullPoiterException
+was the root cause.
+ The NullPointerExceptions are mainly signaled inside the applicaiton code and the Android platform,
+ although we also find NullPointerExceptions being signaled from third-party libraries. Regarding
 reusable code, there is no consensus whether it is a good or bad practice to 
@@
 
-The high prevalence of NullPointerExceptions is aligned with the findings other 
+The high prevalence of NullPointerExceptions is aligned with the findings of other 
 works~\cite{kim2013predicting,fraser20131600,csallner2004jcrasher}. For instance, Sunghun et
@@
 testing~\cite{maji2012empirical,csallner2004jcrasher} showed that most of the automatically 
-detected bugs were due to NullPointerExceptions and implicitly-signaled of Java
-environment (as the ones found in this study).In Section XX we provide further 
+detected bugs were due to NullPointerExceptions 
+and exceptions implicitly-signaled by Java environment due to programming mistakes or resource limitations
+ (as the ones found in this study).In Section ~\ref{sec:disc}  we provide further 
 discussion on the NullPointerException problem, pointing 
@@
 \emph{\textbf{Direct instances of RuntimeExceptions being thrown.}}
-From Table ~\ref{tab:toptenandroid} we can observe that direct instances
+From Table ~\ref{tab:topten} we can observe that direct instances
 java.lang.RuntimeException were thrown in 18,88\% of Android repositories, and
@@
 
-Such categorization was based known sources of faults mobile development: concurrency [INCLUDEREF]
-, backward compatibility  [INCLUDEREF], security   [INCLUDEREF] and resource management (IO, Memory, Batery)  [INCLUDEREF].
-We added to this set categories to represent the exceptions caused by programming mistakes,
- and general Java exceptions (whose types not convey any specific cause of failure), and 
-exceptions that represent specific  exceptional situations of apps, frameworks or libraries.
+Such categorization was based known sources of faults mobile development: concurrency ~\cite{ama2012}
+, backward compatibility  ~\cite{McDon13}, security   ~\cite{enck2011study} and resource management (IO, Memory, Batery) ~\cite{Zhang12}.
+We added to this set of categories other three categories: one to represent the exceptions caused by programming mistakes;
+  another to represent general Java exceptions (e.g., java.lang.Exception) - whose types do not convey any specific cause of failure; and a category to
+represent specific  exceptional situations of apps, frameworks and libraries.
 The list of categories is illustrated in Table~\ref{tab:categories}.
@@
 
-The code snippet in Figure X illustrate this scenario.
+The code snippet in Figure ~\ref{fig:snippets} (d) illustrate this scenario.
  The java/lang/Class declares a native method named getDeclaredMethods. 
@@
 
-Figure X presents a code snippet of ActivityThreat which is responsible for (among other things)
+Figure ~\ref{fig:snippets} (a) presents a code snippet of ActivityThreat which is responsible for (among other things)
 calling the Activities onCreate method. Hence, as we can see from lines 4-8,
@@
 We could observe that several types of exceptions were encapsulated in 
-instances of Runtime. Table X shows some kinds of errors encapsulated a
-RuntimeException.
-
-%\begin{table}
-%\centering
-%\begin{tabular}{ll}
-% \bfseries{Wrapper type - Root cause type} & \bfseries{\#} \\
-%    \hline
-%java.lang.RuntimeException - java.lang.OutOfMemoryError & 30\\
-%java.lang.RuntimeException - java.lang.NoClassDefFoundError & 6\\
-%java.lang.RuntimeException - java.lang.UnsatisfiedLinkError) & 5\\
-%java.lang.RuntimeException -  java.lang.StackOverflowError)  & 4\\
-%\hline
-%  \end{tabular}
-%\caption{Elements responsible for the wrapping.}
-%\label{tab:wrapping01}
-%\end{table}
+instances of Runtime such as java.lang.OutOfMemoryError (30), java.lang.NoClassDefFoundError (6), 
+java.lang.UnsatisfiedLinkError (5), java.lang.StackOverflowErro (4), as illustrated in Table ~\ref{tab:exampeswrap}.
 
@@
 \caption{Examples of Cross-type wrappings}
-\label{tab:wrapping01}
+\label{tab:exampeswrap}
 \end{table}
@@
 
-\section{Futher Discussions}
+\section{Further Discussions}
+\label{sec:disc}
 
 \begin{figure*} \centering \includegraphics[scale=0.55]{codeexamples.pdf}
-\caption{Code snippets illustrating fragilities on the exception-related code} \label{fig:exchier} \end{figure*}
+\caption{Code snippets illustrating fragilities on the exception-related code} \label{fig:snippets} \end{figure*}
 
@@
 (i) lightweight intra-method null pointer analysis supported by Java8 @Nullable annotations;
-(ii) inter-method null pointer analysis tools such as the one proposed by [DANIELE-RELATED-WORK];
+(ii) inter-method null pointer analysis tools such as the one proposed by Nanda and Sinha ~\cite{nanda2009accurate};
 or (iii) whether language designs which avoid null pointers, such 
@@
 \subsection{Actionable Advices - The Unwritten rules of Android development}
+to do if we have space...
 
@@
 \section{Threats to Validity}
+\label{sec:threats}
 
@@
 \section{Related Work}
+\label{sec:rele}
 
@@
 \section{Conclusion}
+\label{sec:conc}
 
 In this paper we present an exploratory study in which we mined the stack 
-traces embedded in all issues of Java projects available GitHub.
-Furthermore, we analyzed in more detail the
-stack traces reported on a subset of 482 Android projects. In this study, the information extracted 
-from stack traces was used in combination with source code and bytecode analysis to 
-pin-point patterns of exception use and misuse in Java.
-Certain patterns of exception
-(mis)-use were consistently detected such as unexpected wrappings (e.g., Errors
-being wrapped in checked exceptions), revealing that  the Java hybrid exception
-model is not fully used according to its purpose; undocumented runtime
-exceptions signaled by third party code (which makes it almost impossible for
-library clients to protect against such exceptions); 
-and a high prevalence of
-\texttt{java.lang} exceptions reported on issues (representing approximately 50\% of the
-analyzed issues). Such (mis)uses detected in our study can negatively affect the 
-robustness of current Java software systems. Our results call for  
- developing tools support to help developers dealing with such problems and 
-improving the design of the exception handling mechanism in Java.
+traces embedded in all issues defined in 482 Android projects hosted in Github and 
+157 projects hosted in Googlecode.
+
+In this study, the information extracted from stack traces was used in combination 
+with source code and bytecode analysis to pin-point fragilities on the exception-related
+code. 
+This study  consistently identifies common characteristics of reported 
+exception stacks traces - e.g., the NullPointerException was the reported cause 
+on 30,08\% of all analyzed exception stack traces, and was mentioned in approximately 
+50\% of all projects analyzed. Moreover,  the cross-project analysis of exception stack traces 
+enabled to consistently reveal  fragilities of the exception-related code: 
+(i) undocumented runtime exceptions signaled by third party code,
+(ii) undocumented checked exceptions thrown by Android native code
+(iii) unexpected exception wrappings (e.g., Errors being wrapped in checked exceptions) 
+revealing that Java hybrid exception model is not fully used according to its purpose.
+Such fragilities detected in our study can negatively affect the 
+robustness of not only Android application but also Java-based 
+which can deal with the same fragilities pointed in thid work. 
+Our results calls the attention for the need of tool and language support to help 
+developers when dealing with such problems.
 

### 2014-09-01 android-stacks.tex
all over . K9mail

@@
 
-Different from GitHub, Googlecode does not provide an API to access the information related
- to hosted projects \footnote{Googlecode previoulsly provided a Web service but was deactivated in June 2013.}.
-However many well used open source Android application such as K9mail, is hosted in Googlecode which stores
-the issue history of this and other apps are store there ~\footnote{K9mail moved to Github but advice their 
-users to create issues on Googlecode as a way of not loosing the issue history.}.
-
-To overcome this limitation and mine the information available on such Googlecode 
-projects we implemented a Web crawler which navigated (incorporated in Exception Miner tool described next) 
- on each project and issue page, extracting the content of each issue to store on relational database.
-\footnote{Google code previously provided a Web service to extract information about each project, but this service was removed in what Google called "clean-up action" which happened in June 2013.}
-
+Differently from GitHub, Googlecode does not provide an API to access the information related
+ to hosted projects \footnote{Google code used to provided a Web service to such repository, but it was deactivated in June 2013 what Google called a "clean-up action".}.
+
+However, Googlecode still contain largely used open-source Android apps (e.g. K9Mail \footnote{K9Mail moved to Github but as a way of not loosing the project history 
+it advices their users to report bugs on Googlecode issue tracker.})  and the history of such apps are also valuable. 
 To overcome this limitation we needed to implement a Web Crawler (incorporated in Exception Miner tool described next) that navigated 
@@
 were extracted.
+
+%For instance: K9mail  was hosted in Googlecode 
+%before moving to GitHub and in order not to loose the project history it advices their 
+%users to create issues on Googlecode as a way to preserve project history.
+
+%To overcome this limitation and mine the information available on such Googlecode 
+%projects we implemented a Web crawler which navigated (incorporated in Exception Miner tool described next) 
+% on each project and issue page, extracting the content of each issue to store on relational database.
 
@@
 
-\emph{\textbf{Direct instances of RuntimeExceptions being thrown.}}
-From Table ~\ref{tab:topten} we can observe that direct instances
-java.lang.RuntimeException were thrown in 18,88\% of Android repositories, and
-most of such exceptions  were  thrown by the Android platform (167 out of 232).
-
-Throwing general exceptions, such as direct instances of RuntimeException, violates best
-practice III. It is considered a bad practice because exception type does not carry enough
- information to identify the cause of the exceptional behavior, and as a consequence developers need to relying on
-the exception message which can be neither complete nor precise~\cite{gosling2000java}.
+%\emph{\textbf{Direct instances of RuntimeExceptions being thrown.}}
+%From Table ~\ref{tab:topten} we can observe that direct instances
+%java.lang.RuntimeException were reported as the root cause of stacks in 19,09\% 
+%of the projects, and most of such exceptions  were  thrown by the Android platform (167 out of 232).
+%Throwing general exceptions, such as direct instances of RuntimeException, violates best
+%practice III. It is considered a bad practice because exception type does not carry enough
+%information to identify the cause of the exceptional behavior, and as a consequence developers need to relying on
+%the exception message which can be neither complete nor precise~\cite{gosling2000java}.
 
@@
 \noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
-Programming mistakes were reported as the root cause of  52\% of the analyzed stack traces. 
-In 30,08\% NullPointerExceptions was reported as the root cause.
+A multitude of programming mistakes were reported as the main causes of stack traces.
+51,96\% of all projects reported at least one exception stack trace on which the NullPoiterException
+was the root cause.
 }
@@
 Besides distilling the information available on stack traces (i.e., exception signalers, root causes, exception wrappings),
-we mined the exception types (i.e., RuntimeException, checked Exception or Error) related to each exception reported on stack traces 
-using source code and byte code analyses supported by Exception Miner described before.
+we mined the exception types (i.e., RuntimeException, checked Exception or Error) related to each exception reported on stack traces -
+to do so we used the source code and byte code analyses implemented on Exception Miner.
 
  Table~\ref{tab:typeroottab} presents the types of root causes of all stack traces reported on the analyzed issues. As we can see 
-from Table ~\ref{tab:typeroottab},  most of the reported exceptions were signaled by Android and libraries reused by Apps were runtime exceptions. 
-On the other hand, we could observe that most of the reported exceptions signaled by the set of 
-libraries reused by Android platform (i.e., org.apache.harmony,
-org.w3c.dom, sun.misc, org.apache.http, org.json, org.xml, javax -- referred in
-this work as libcore) were checked exceptions. It is pointed as a good practice
+from Table ~\ref{tab:typeroottab},  most of the reported exceptions signaled (explicilty or implicitly by the JVM) on methods defined
+ on the Android Plaform on on the application itself.
+
+On the other hand, most of the reported exceptions signaled by the set of 
+libraries reused by Android platform were checked exceptions. Such libraries are referred in
+this work as libcore and is composed by org.apache.harmony, org.w3c.dom, sun.misc, 
+org.apache.http, org.json, org.xml, javax. It is pointed as a good practice
 for libraries (see Section~\ref{sec:best}) because by using checked exceptions
-libraries can define a precise exception interface to its clients. Such
+a library can define a precise exception interface ~/cite{miller1997issues} to its clients. Such
 libraries are heavily used in several projects, and such precise exception
@@
 
-We could observe that most of the exceptions reported on issues were runtime, 
-mainly  signaled on the Android Platform and in the Application code. we could also find 
-runtime exceptions signaled from libraries.
-
-In order to evaluate if  best practice IV was being obeyed, we manually inspected the   
-signature of each method signaling a runtime exception, that was defined by third 
-party code (lib or Android) and reused by an application.
-In this task we filtered out all the exceptions implicitly 
-signaled by  JVM (due to programming mistakes) since these exceptions should not be documented on the 
+\emph{\textbf{Undocumented Runtime Exceptions}}.
+However, we can also observe a considetable amount of 
+runtime exception signaled on library reused by apps.
+In order to evaluate if  best practice IV was being obeyed, we inspected the   
+signature of each method signaling a runtime exception. 
+We inspected methods of the Android Plaftorm and 
+the libraries, because both are  third  party code reused 
+by the Android apps. In this task we filtered out all the exceptions implicitly 
+signaled by  JVM (due to programming mistakes) 
+since these exceptions should not be documented on the 
 method signature.
 
-We could observe that only 1 method out of 118 inspected signalers
+As a result of this inspection, we could observe that only 1 method out of 118 inspected signalers
  (i.e., 0,8\%) documented the explicitly thrown runtime exception in a Javadoc comment; and none of these methods
@@
 
-We could also observe checked exceptions (which according to Java specification represent
-recoverable conditions) as the root causes of reported exception stack traces.
-Two possible ways of a checked exception can remain uncaught are: the checked 
-exception is declared by every method on the call chain or if it is wrapped in a 
-runtime exception and re-thrown. However the inspection of stack traces 
-and the signatured of methods revealed the a third option: an undocumented
-checked exception thrown by a native method.
-
-The code snippet in Figure ~\ref{fig:snippets} (d) illustrate this scenario.
- The java/lang/Class declares a native method named getDeclaredMethods. 
-The Java-side declaration of of this method does not have any throws clause, 
-leading programmers and the compiler to think that no checked exceptions can be thrown.
-
- However, the C-code implementation did throw a "checked exception" NoSuchMethodException, 
-violating the declaration. The Java compiler, could not detect this violation because it does 
-not perform static exception checking on native methods. This type of bug is hard to diagnose
-because the developer usualy do not have access to the native implementations.
-
-Consequently, since it is not expected by the programmer, when such method throws 
-this exception, such undocumented exception may remain
-uncaught and cause the app to crash, or may be mistakenly handled by subsumption - 
-The exception stack traces reporting this scenario is actually a real bug of Android 
-Gingerbread version (which still accounts for 13.6\% of devices running Android).
-
-Among the Android projects on which this exception stack trace was reported we can cite:
-PageTurner (206 stars, 130 forks, 1184 commits), ActionBarSherlock (6,041 stars, 
-3,436 forks, and 1479 commints), otto (1,618 stars,	342 forks, 	177 commints).
-
 %SELECT repo, *
@@
  Only 0,8\% of the methods documented the explicitly signaled runtime exceptions. 
-  The study also detected an undocumented checked exception signaled by
-  a native method defined in Android.
 }
@@
 %\subsection{Exploring the Information Embedded on Exception Wrappings}
+
 Table~\ref{tab:wrappingandroid} presents the wrappings found in this study for all
@@
 
-%, while most of them were not wrapped along the stack trace.
-
 %SELECT COUNT(*) FROM STACKTRACESUMMARY WHERE
@@
 
+ %Runtime &  Checked   & 70  / 18 & 114 /  34 &  66 /  9  &  0 /  & 19 / 19 &  29 / 6 \\
+      %Runtime   &  Error   & 38  \ 8  & 57    &  50 \ 8  &  0   & 6 \ 2  &  1   /  0\\
+%      Checked &  Runtime   & 12 \ 5  & 22 \ 9 & 4 &  0  & 9 \7 &  9 \ 2 \\
+%      Checked & Error      & 7 \ 1 &  8 & 5  &  0  & 0 \ 1 &  3 \0 \\
+ %     Error & Checked      & 4 \ 10  &  7 & 0  \ 6  &  7  & 0 \6 &  0   \ 8    \\
+%      Error & Runtime     & 6 \ 2  &  13   & 1 &  1  & 0 \ 1 &  11 \ 3    \\
+
 \begin{table*}
@@
     \hline
-      %Runtime &  Checked   & 70  / 18 & 114 /  34 &  66 /  9  &  0 /  & 19 / 19 &  29 / 6 \\
 
       Runtime &  Checked   & 88 & 148 &  75  &  0   & 38 &  35 \\
-
-      %Runtime   &  Error   & 38  \ 8  & 57    &  50 \ 8  &  0   & 6 \ 2  &  1   /  0\\
       Runtime   &  Error   & 46  &  67    &  58  &  0   & 8  &  1   \\      
-
-%      Checked &  Runtime   & 12 \ 5  & 22 \ 9 & 4 &  0  & 9 \7 &  9 \ 2 \\
       Checked &  Runtime   & 17  & 31 & 4 &  0  & 16 &  11 \\
-
-%      Checked & Error      & 7 \ 1 &  8 & 5  &  0  & 0 \ 1 &  3 \0 \\
-
       Checked & Error      & 8 &  9  & 5  &  0  &  1 &  3  \\
-
- %     Error & Checked      & 4 \ 10  &  7 & 0  \ 6  &  7  & 0 \6 &  0   \ 8    \\
       Error & Checked      & 14 &  27 &  6  &  7  &  6 &   8    \\
-
-%      Error & Runtime     & 6 \ 2  &  13   & 1 &  1  & 0 \ 1 &  11 \ 3    \\
-
       Error & Runtime     & 8  &  17   & 1 &  1  & 1 &  14    \\
@@
 
-TODO: ATUALIZAR DISCUSSAO COM BASE NOS NOVOS DADOS.
-
 This analysis revealed unexpected wrappings such as: Error wrapping a Checked 
 exception; a Runtime wrapping and Error; and an Error wrapping a Runtime.
-Java is the only language that provides a hybrid exception handling mechanism
+Java is the only language that provides a hybrid exception model 
 which offers  different types to represent different exception behaviors (i.e., error, 
-runtime and checked) (see Section~\ref{sec:best}). According to Java specification Errors should not be
-handled inside the system since they usually represent unrecoverable conditions
-detected by the JVM such as OutOfMemoryError. Checked exceptions should
-represent recoverable conditions and  Runtime exceptions represent are usually
-related to programming mistakes from which the developer is not expected to recover
-from. Next sections discusses about such types of wrappings.
+runtime and checked) (see Section~\ref{sec:best}). For instance,
+according to Java specification, Errors should not be handled inside the system 
+since they usually represent unrecoverable conditions
+detected by the JVM (e.g.,  OutOfMemoryError). On the other hand, checked exceptions should
+represent recoverable conditions. Next we discuss and exemplify the cross-type wrappings
+ found in this study.
 
 \emph{\textbf{Runtime wrapping Checked Exception.}} Such wrapping was 
-detected on 70 distinct projects. Inspecting the 
+detected on 88 projects. Inspecting the 
 elements responsible for performing such wrapping, we could observe that
-most of such wrappings were performed on methods defined on Android
-Platform (approximately 58%).
-
-%\begin{table}
-%\centering
-%\begin{tabular}{ll}
-%    \hline
-%   \bfseries{signaler from} & \bfseries{Ocurrences} \\
-%    \hline
-%ANDROID	 & 66 \\
-%APP	& 29 	(HOW MANY?? AND WHICH??) \\
-%LIB  & 19 (HOW MANY AND WHICH?)\\
-%\hline
-% \end{tabular}
-%\caption{Elements responsible for the wrapping.}
-%\label{tab:wrapping01}
-%\end{table}
+approximately  50\% of such wrappings were performed on methods defined on Android
+Platform.
 
 Taking a closer look on the methods responsible for such wrappings,
- we could observe that approximately 41% of them were performed by a single class,
- named ActivityThreat. This class is responsible for managing the life cycle of an Android Activities.
+ we could observe that approximately 41\% of them were performed by a single class,
+ named ActivityThreat.
+
+This class is responsible for managing the life cycle of  Android Activities - 
+invoking the life cycle methods (i.e. onCreate(), onPause()) for every Activity 
+(window) of an Android app.
+Figure ~\ref{fig:snippets} (a) presents a code snippet of one the methods of 
+ActivityThrea,  the one responsible for invoking the Activities onCreate method. 
+As we can see from lines 4-8,
+ in case any exception is thrown in the context of such try-catch block, 
+it is wrapped in a general Runtime exception and re-thrown.
 
@@
     \hline
-android.app.ActivityThread & java.lang.RuntimeException & 47 \\
+android.app.ActivityThread & java.lang.RuntimeException & 61\\
 android.view.LayoutInflater & android.view.InflateException &  7 \\
-android.os.AsyncTask & java.lang.RuntimeException &  5 \\
-Other platform classes & java.lang.RuntimeException &   7 \\
+android.os.AsyncTask & java.lang.RuntimeException &  7\\
 \hline
   \end{tabular}
-\caption{Elements tresponsible for the wrapping.}
+\caption{Top 3 Runtime-Cheked wrappers.}
 \label{tab:wrapping01}
 \end{table}
-
-Figure ~\ref{fig:snippets} (a) presents a code snippet of ActivityThreat which is responsible for (among other things)
-calling the Activities onCreate method. Hence, as we can see from lines 4-8,
- in case any exception is thrown by an application Activity during its creation, 
-it is wrapped in a general Runtime exception and re-thrown.
 
@@
 
-The inversion of control makes it difficult to the application developer to associate 
-a specific handler to the exceptions thrown during the onCreate method (or any other Activity life cycle method).
-Since according to Java handling model the handler should me added on any method in the call chain,
-and in the inversion of control development model the developer usually does not have fully access to 
-the invocation call chain of a method in order to associate a handler, as a result the 
-exception may scape from any handling and remain uncaught.
-
-Difficulty of associating a specific handler to exceptions signaled by application methods called by the framework,
-may also be the reason behind the design decision of encapsulating any exception in a RuntimeException.
-As framework does not have enough contextual information the exception regardless its types,
-it encapsulates in a general type and re-throws. The only possible answer to such exceptions would be 
-to finish the application.
-
-Such design decision may lead the developers into believing that by just re-throwing 
-the exceptions they can forget about the exceptional situations during the development of the "happy path". 
-This "ignore-for-now" approach may turn the exception handling into a generalized "goto" mechanism~\cite{mandrioli1992advances}
- making the program more complex and even less reliable. This behavior may lead to the well known uncaught exceptions 
-problem~\cite{jo2004uncaught}, one of the main causes of application crashes.
+%SELECT SIGNALER_method, COUNT(*) AS N
+%  FROM stacktracesummary WHERE ROOT_EXTENDS='CHECKED' AND EXCEPTION_EXTENDS='RUNTIME' and
+%  ( SIGNALER_method like '%onDestroy%' or
+%   SIGNALER_method like '%onCreate%' or
+% SIGNALER_method like '%onStart%' or
+% SIGNALER_method like '%onResume%' or
+% SIGNALER_method like '%onPause%' or
+% SIGNALER_method like '%onStop%' or
+% SIGNALER_method like '%onDestroy%' )
+%   GROUP BY SIGNALER_method
+%  ORDER BY N DESC;
+
+%SELECT COUNT(*) AS N
+ % FROM stacktracesummary WHERE ROOT_EXTENDS='CHECKED' AND EXCEPTION_EXTENDS='RUNTIME' and
+ % ( stack like '%onDestroy%' or
+ %  stack like '%onCreate%' or
+% stack like '%onStart%' or
+% stack like '%onResume%' or
+% stack like '%onPause%' or
+% stack like '%onStop%' or
+% stack like '%onDestroy%' );
+
+%GC = 6
+
+However, since the Activity life clycle method (i.e., callActivityOnCreate(...)) that is invoked in this block,
+ do not include checked exceptions on its signatures. Hence they could not throw a checked exception to be handled 
+in the method illustrated above. However the code inspection performed on such wrappings revelaed an
+``undocumented checked exception" signaled by this life cycle method.
+
+\emph{\textbf{Undocumented Checked Exceptions.}} There are usually
+2 possible possibilities for a checked exception to be thrown by a method:
+it can either be specified on the method signature or it can be wrapped in a 
+runtime exception and re-thrown. However by inspectioning associated to the reported wrappings 
+revealed the a third option: a checked exception thrown by a native method, which was not declared
+on the method signature of the methods on the call chain composing the stack.
+
+The code snippet in Figure ~\ref{fig:snippets} (d) illustrate this scenario.
+ The java/lang/Class declares a native method named getDeclaredMethods. 
+The Java-side declaration of of this method does not have any throws clause, 
+leading programmers and the compiler to think that no checked exceptions can be thrown.
+ However, the C-code implementation did throw a "checked exception" NoSuchMethodException, 
+violating the declaration. The Java compiler, could not detect this violation because it does 
+not perform static exception checking on native methods. This type of bug is hard to diagnose
+because the developer usualy do not have access to the native implementations. 
+Consequently, since it is not expected by the programmer, when such method throws 
+this exception, such undocumented exception may remain
+uncaught and cause the app to crash, or may be mistakenly handled by subsumption - 
+The exception stack traces reporting this scenario is actually a real bug of Android 
+Gingerbread version (which still accounts for 13.6\% of devices running Android).
+
+Among the Android projects on which this exception stack trace was reported we can cite:
+PageTurner (206 stars, 130 forks, 1184 commits), ActionBarSherlock (6,041 stars, 
+3,436 forks, and 1479 commints), otto (1,618 stars,	342 forks, 	177 commints).
 
@@
 
-\emph{\textbf{Runtime Exception wrapping Error.}} We could find  57 stack traces 
-on 38 projects on which an Error was wrapped in a runtime exception.
+\emph{\textbf{Runtime Exception wrapping Error.}} We could find  64 exception stack traces 
+on 46 projects on which an Error was wrapped in a runtime exception.
 Manually inspecting each exception stack trace, we could observe that 44% of the wrappings were performed 
 by android.os.AsyncTask. This class is responsible for dealing with asynchronous tasks in Android.
-The code snippet below shows that any throwable signaled in the context of an asynchronous 
-task is converted in to a runtime and re-thrown.
-
-ASSNCTASKS NON ANNOTATED METHODS…….
-SOME HAVE EXCEPTION INTERFACE 
-EXCEPTIONS ON MEHTODS SIGNATURE OTHERS NOT
-INCONSISTENCY…. DEPENDS ON DEVELOPERS 
-WORK INSTEAD OF TOOLS.
+The code snippet is illustrated in Figure ~ref{wrappings} (c) shows that any throwable signaled in the context of an asynchronous 
+task is converted in to a runtime and re-thrown. Some of the Erros wrapped on instances of runtime exceptions were 
+java.lang.OutOfMemoryError (30), java.lang.NoClassDefFoundError (6), 
+java.lang.UnsatisfiedLinkError (5), java.lang.StackOverflowErro (4), as illustrated in Table ~\ref{tab:exampeswrap}
+
+%ASSNCTASKS NON ANNOTATED METHODS…….
+%SOME HAVE EXCEPTION INTERFACE 
+%EXCEPTIONS ON MEHTODS SIGNATURE OTHERS NOT
+%INCONSISTENCY…. DEPENDS ON DEVELOPERS 
+%WORK INSTEAD OF TOOLS.
 
@@
 
-We could observe that several types of exceptions were encapsulated in 
-instances of Runtime such as java.lang.OutOfMemoryError (30), java.lang.NoClassDefFoundError (6), 
-java.lang.UnsatisfiedLinkError (5), java.lang.StackOverflowErro (4), as illustrated in Table ~\ref{tab:exampeswrap}.
+.
 
@@
 traces we could observe that most of the wrappings were performed 
-by RoboSpice (an android library that supports the creation of asynchronous long running tasks.
-This library has 1.682 stars and 367 forks) and by the reflection Java library (used by Android, 
+by RoboSpice (an android library that supports the creation of asynchronous long running 
+tasks, which has 1.682 stars and 367 forks in GitHub) and by the reflection Java library (used by Android, 
 Application and library classes). Inspecting the code related to such wrappings we could observe that
@@
  were also caused by the reflection library 
-used by Android applications. The method responsible for the wrappings
-were native methods written in C. Table  ~\ref{tab:wrapcheckerr} illustrates
+used by Android applications. The methods responsible for the wrappings
+were also native methods written in C. Table ~\ref{tab:exampleswrap} illustrates
 some oth these wrappings some of them are masking an OutOfMemoryError
@@
 
-\emph{\textbf{Error wrapping Runtime.}} After manually inspecting the code associated to such wrappings we discovered that all of them were 
-caused by static initializers. If any exception is thrown in the context of a static initializer (i.e., static block) 
-it is converted into an ExceptionInitializerError on the point where the class is first used. Table X illustrates
+\emph{\textbf{Error wrapping Runtime.}} After manually inspecting the code associated to such wrappings 
+we discovered that all of them were caused by static initializers. If any exception is thrown in the context of a static initializer (i.e., static block) 
+it is converted into an ExceptionInitializerError on the point where the class is first used. Table ~\ref{exampleswrap} illustrates
 the exception types involved in such wrappings.
@@
 \emph{\textbf{Multiple Wrappings.}} We could also observe that in 205 stacks found
- in Android repository the exception was wrapped 
-more then once. One interesting multiple-wrapping found in Android repo was the 
+ in Android repository the exception was wrapped  more then once. One interesting 
+multiple-wrapping found in Android repo was the 
 following: Runtime-Checked-Runtime-Checked-Runtime-Checked-Runtime.
@@
 reveal fragilities on Java Exception Model.
+The study also detected an undocumented checked exception signaled by
+  a native method defined in Android.
 }
@@
 \textbf{Brian Foote}
+
+\subsection{Inversion of Control and Exception Handling}
+Reasoning about the design strategy adopted by the ActivityThread,
+we can observe that the inversion of control (on which this class and 
+most of the Android framework is based) can make it difficult for the application developer to 
+associate a specific handler to the exceptions thrown during the execution of the onCreate method (or any other life cycle method).
+According to Java handling model, the handler should be added on a method on the invocation chain. 
+However,  in the inversion of control development model the developer does not have fully access to 
+the invocation chain of a method. Therefore, the developer may be tempted to convert the checked
+ exception into a RuntimeException and resignal.
+
+Difficulty of associating a specific handler to exceptions signaled by application methods called by the framework,
+may also be the reason behind the design decision of encapsulating any exception in a RuntimeException.
+As framework does not have enough contextual information the exception regardless its types,
+it encapsulates in a general type and re-throws. The only possible answer to such exceptions would be 
+to finish the application.
+
+Such design decision may lead the developers into believing that by just re-throwing 
+the exceptions they can forget about the exceptional situations during the development of the "happy path". 
+This "ignore-for-now" approach may turn the exception handling into a generalized "goto" mechanism~\cite{mandrioli1992advances}
+ making the program more complex and even less reliable. This behavior may lead to the well known uncaught exceptions 
+problem~\cite{jo2004uncaught}, one of the main causes of application crashes.
 
@@
 
-When family of systems make use of a common core, or framework such systems can
-benefit from a cross project crash analysis. The bug issues reported for a group of systems
-can identify vulnerabilities which were more difficult to identify in a single project analysis.
-
-The test case explosion problem... A test suite is always incomplete,
-one application could benefit from the usage scenarios on which such framework is applied.
-Framework testing -  what are the main challenges on framework testing?
-A cross-project crash analysis based on repository mining techniques can 
-identify vulnerabilities in pieces of code that are shared.
-
-Using exception miner tool we could identify the methods responsible for most of the crashes - 
-such methods can be a target for manual inspections improvements, or better documentation
-(because sometimes an exception can be thrown because the user is not invoking the method
-in the proper way ....
+When family of systems make use of a common core, it is possible to benefit from a cross project crash analysis like the one performed in this study.
+ The not only the exception stack traces but any other bug issues reported for a group of systems can identify vulnerabilities 
+that are common to more then one element of the family and are more difficult to identify in a single project analysis.
+Therefore tools such as the one developed in this study may allow the developers to 
+benefit from the faults happening on different systems.
+
+%It can be a way of dealing with the test case explosion problem related to framework test.
+%Let the products test and report the crahes in a single repository and then analize them.
+%... A test suite is always incomplete,
+%one application could benefit from the usage scenarios on which such framework is applied.
+%Framework testing -  what are the main challenges on framework testing?
+
+%Using exception miner tool we could identify the methods responsible for most of the crashes - 
+%such methods can be a target for manual inspections improvements, or better documentation
+%(because sometimes an exception can be thrown because the user is not invoking the method
+%in the proper way ....
 
@@
 signaled by reused therd part code libraries.
-
-It allows the developers to benefit from the faults happening on different systems.
 
@@
 our sample.
-<OLHAR ARTIGO DO FSE QUE TRABALHA COM OPENSOURCE.
+\footnote{to do: OLHAR ARTIGO DO FSE QUE TRABALHA COM OPENSOURCE.
 FALAR QUE DESCOBERTAS FORAM DE LIBS JAVA. APESAR DO ESTUDO NAO SER FEITO EM PROJS JAVA.....
-PODEMOS CONCLUIR ALGUMAS COISAS QUE SE APLICAM A TODOS JAVA>
+PODEMOS CONCLUIR ALGUMAS COISAS QUE SE APLICAM A TODOS JAVA}
 
@@
 classify them in a set of predefined categories.
-(MENCIONAR QUE X% DOAS EXCECOES ANALISADAS ESTAVAM EMBEBIDAS EM ALGUM TIPO DE ARQUIVO DE LOG - 
-LOGCAT DO ANDROID....)
 
@@
 the analysis of stack traces extracted from issues in combination with and bytecode and source
-code analysis.
-
-RESSALTAR LIMITACAO DE QUANTIDADE DE FALSOS POSITIVOS... E QUE OS STACKS SAO REAIS...
-REPRESENTAM FLUXOS REAIS.... POST MOSTERN ANALISYS OF REAL EXECUTIONS....
+code analysis. \footnote{ to do: TODORESSALTAR LIMITACAO DE QUANTIDADE DE FALSOS POSITIVOS... E QUE OS STACKS SAO REAIS...
+REPRESENTAM FLUXOS REAIS.... POST MOSTERN ANALISYS OF REAL EXECUTIONS....}
 

### 2014-09-02 android-stacks.tex
all over

@@
 far beyond of the basic calls and textual messages, on the other hand
-they have to face threats to application robustness.
+they have to face threats to application robustness ~\cite{was2010}.
 Such applications have to cope with an increasing number of abnormal
@@
 
-  \item  A multitude of programming mistakes were reported as the main causes of stack traces:
+  \item  A multitude of programming mistakes were reported as the main causes of exception stack traces:
  27,71\% of the analyzed exception stack traces reported a java.lang.NullPointerException the root cause;
@@
 Such categorization was based known sources of faults mobile development: concurrency ~\cite{ama2012}
-, backward compatibility  ~\cite{McDon13}, security   ~\cite{enck2011study} and resource management (IO, Memory, Batery) ~\cite{Zhang12}.
+, backward compatibility  ~\cite{McDon13}, security   ~\cite{enck2011study,was2010} and resource management (IO, Memory, Batery) ~\cite{Zhang12}.
 We added to this set of categories other three categories: one to represent the exceptions caused by programming mistakes;
@@
 \noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
-A multitude of programming mistakes were reported as the main causes of stack traces.
+A multitude of programming mistakes were reported as the main causes of exception stack traces.
 51,96\% of all projects reported at least one exception stack trace on which the NullPoiterException
@@
 java.lang.OutOfMemoryError (30), java.lang.NoClassDefFoundError (6), 
-java.lang.UnsatisfiedLinkError (5), java.lang.StackOverflowErro (4), as illustrated in Table ~\ref{tab:exampeswrap}
+java.lang.UnsatisfiedLinkError (5), java.lang.StackOverflowErro (4), as illustrated in Table ~\ref{tab:exampeswrap}.
 
@@
 
-.
-
 \begin{table}
@@
 \noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
-Odd wrappings were detected (e.g., a checked exception wrapping an OutOfMemoryError)
-reveal fragilities on Java Exception Model.
+Odd wrappings were detected (e.g., a checked exception wrapping an OutOfMemoryError).
 The study also detected an undocumented checked exception signaled by
@@
 \subsection{Inversion of Control and Exception Handling}
+
+footnote{to do: INPORTANTE Falar que os metodos do ciclo de vida da app nao tem interface excepcional.
+desta forma ela não define: O que fazer se o metodo der errado? Nao deixa claro.}
+
 Reasoning about the design strategy adopted by the ActivityThread,

### 2014-09-03 android-stacks.tex
all over

@@
 
-\title{ What can GitHub and Googlecode Issues tell us about Exceptions in Android and Java? }
+\title{What can the Issues of GitHub and Googlecode tell us about Exception Fragilities in Android? }
 
@@
 %\title{On the Common Characteristics of Android Crashes: Results of a GitHub and Googlecode Mining Study}
+
+%\title{A first Glance on Characteristics of Android Crashes: Results of a GitHub and Googlecode Mining Study}
 
@@
 
+%GitHub 315,92
+% Googlecode: 127,456
+%All: 159,048
+
 \begin{abstract}
-
-This paper reports a repository mining study which analyzed all the issues created on
-Android projects available in GitHub (482 applications) and Google code (157 applications)
-Overall, our research set includes 32,582 issues in GitHub and 41,674 issues in Googlecode 
-from which 6,005 exception stack traces were extracted. Such stack traces were distilled and analyzed in 
-combination with source code and bytecode analysis. This study 
-consistently identifies common characteristics of reported 
-exception stacks traces - e.g., the NullPointerException was the reported cause 
-on 30,08\% of all analyzed exception stack traces, and was mentioned in approximately 
-50\% of all projects analyzed. Moreover,  the cross-project analysis 
-of exception stack traces enabled to consistently reveal   
-fragilities of the exception-related code: 
-(1) undocumented runtime exceptions signaled by third party code,
-(2) undocumented checked exceptions thrown by Android native code
-(3) unexpected exception wrappings (e.g., Errors being wrapped in checked exceptions) 
+The number of mobile apps is increasing in a daily rate. If on the one hand they
+extend phones capabilities far beyond of the basic calls, on the other hand they have to 
+cope with an increasing number of exceptinal conditions 
+(e.g., faults in underlying middleware or hardware; memory and battery restrictions). 
+Therefore, mechanisms for exception detection and handling are not an optional add-on but a 
+fundamental part of such apps. However, studies have shown that such mechanisms are the less 
+understood and tested parts of the system. As a consequence they may affect the system the 
+other way around, making the system even less robust.
+This paper reports a study which mined the stack traces embedded on 
+159,048 issues reported on Android projects hosted in GitHub (482 projs.) 
+and Googlecode (157 projs.). The goal of this study was to: identify common characteristics 
+of such traces and investigate the fragilities on the exception-related code that can be revealed
+from them. Overall 6,005 exception stack traces were extracted, and analyzed in 
+combination with source code and bytecode analysis. 
+Some fragilities consistently detected in this study include: 
+(1) undocumented runtime exceptions thrown by third party code;
+(2) undocumented checked exceptions thrown by Android native code;
+(3) unexpected exception wrappings (e.g., Errors being wrapped in checked exceptions),
 revealing that Java hybrid exception model is not fully used according to its purpose.
-Such fragilities impose challenges to application robustness not only 
-in Android environment but also to other systems based on Java exception model.
+Moreover, about 50\% of all analyzed projects reported at least one NullPointerExeption
+as the root cause of an exception stack trace.
+This study provide important insights on how such fragilities impose challenges 
+to application robustness not only in Android environment but also to other 
+systems based on Java exception model.
 
@@
 
+%GitHub 31,592
+% Googlecode: 127,456
+%All: 159,048
+
 %\IEEEpeerreviewmaketitle
@@
 Along the recent years, we have witnessed an astonishing increase in the number 
-mobile applications being developed and some of them becoming largely used. 
-Such increase is motivated by pervasiveness of the World Wide Web 
+mobile applications. Such increase is motivated by pervasiveness of the World Wide Web 
 and the unprecedented increase of global smartphone shipments -
@@
 
-In this context, Android has become a leading platform for such applications - thanks to its open-source 
+In this context, Android has become a leading platform - thanks to its open-source 
 model and hardware partners like Samsung, HTC, Motorola, Asus (besides Nexus Google's
@@
 
-In the same rate the number of a users increase, also increases 
-the number of users affected by application faults and crashes.
+%In the same rate the number of a users increase, also increases 
+%the number of users affected by application faults and crashes.
+
 If on the one hand such applications extend phones capabilities 
 far beyond of the basic calls and textual messages, on the other hand
-they have to face threats to application robustness ~\cite{was2010}.
-Such applications have to cope with an increasing number of abnormal
-computation states, arising from:
- noisy user inputs; faults in underlying middleware or hardware; 
-faults on third party services and libraries; compatibility issues [ref]; 
-connection with devices e.g., as GPS, camera; memory and battery restrictions; 
- programming errors, among others.
+they have to face an increasing number of threats to application robustness
+ arising from: faults in underlying middleware and hardware, 
+faults on third party services and libraries; compatibility issues; 
+faults on connected devices (e.g. camera, GPS); memory and 
+battery restrictions; among others.
 
@@
 %the number of users affected by such threats.
-
 %faulty apps and app crashes.
@@
 GitHub and Googlecode. Overall, we analyzed all issues extracted from on 482 Android projects defined in
-GitHub and 157 projects defined in Googlecode. Overall it corresponds to 74,256 issues (32,582 issues from GitHub projects 
-and 41,674 issues from Googlecode projects) from which  6,005 stack traces were extracted and used as
+
+%GitHub 31,592
+% Googlecode: 127,456
+%All: 159,048
+
+GitHub and 157 projects defined in Googlecode. Overall it corresponds to 159,048 issues (31,592 issues from GitHub projects 
+and 127,456 issues from Googlecode projects) from which  6,005 stack traces were extracted and used as
  a source of information for our analysis. The stack trace analysis was augmented by additional information
@@
 
-\begin{figure} \centering \includegraphics[scale=0.8]{stacktrace.png}
-\caption{Example of an Exception stack trace in Java.}
+\begin{figure} \centering \includegraphics[scale=0.6]{stacktrace_bw.png}
+\caption{Example of an exception stack trace in Java.}
 \label{fig:wrapping}

### 2014-09-03 android-stacks.tex
addressing comments

@@
 \begin{document}
-%\title{ What GitHub and Googlecode can tell us about the Causes of Crashes of Android Apps? }
-%\title{ A Mining Study the Googlecode can tell us about the Causes of Crashes of Android Apps? }
-
-\title{What can the Issues of GitHub and Googlecode tell us about Exception Fragilities in Android? }
-
-%\title{On the Common Characteristics Android-Apps Crashes: Results of a GitHub and Googlecode Mining Study}
-
-%\title{On the Common Characteristics of Android-Apps Crashes: Results of a GitHub and Googlecode Mining Study}
-
-%\title{On the Common Characteristics of Android Crashes: Results of a GitHub and Googlecode Mining Study}
-
-%\title{A first Glance on Characteristics of Android Crashes: Results of a GitHub and Googlecode Mining Study}
-
-%\title{A first Glance on Characteristics of Android Crashes: Results of a GitHub and Googlecode Mining Study}
+%\title{ What GitHub and Google Code can tell us about the Causes of Crashes of Android Apps? }
+%\title{ A Mining Study the Google Code can tell us about the Causes of Crashes of Android Apps? }
+
+%\title{What can the Issues of GitHub and Google Code tell us about Exception Fragilities in Android? }
+
+%\title{What can the Issues of GitHub and Google Code tell us about Exception Fragilities in Android? }
+
+\title{Discovering Robustness Threats by Mining the Exception Stack Traces of over 600 Android Projects}
+
+%\title{Discovering Fragilities on the Exception-Related Code through  Exception Stack Trace Mining}
+
+%\title{On the Common Characteristics Android-Apps Crashes: Results of a GitHub and Google Code Mining Study}
+
+%\title{On the Common Characteristics of Android-Apps Crashes: Results of a GitHub and Google Code Mining Study}
+
+%\title{On the Common Characteristics of Android Crashes: Results of a GitHub and Google Code Mining Study}
+
+%\title{A first Glance on Characteristics of Android Crashes: Results of a GitHub and Google Code Mining Study}
+
+%\title{A first Glance on Characteristics of Android Crashes: Results of a GitHub and Google Code Mining Study}
 
@@
 %GitHub 315,92
-% Googlecode: 127,456
+% Google Code: 127,456
 %All: 159,048
+
+%The goal of this study was to: identify common characteristics 
+%of such traces and investigate the fragilities on the exception-related code that can be revealed
+%from them. 
+%The goal of this study was to: identify common characteristics 
+%of such traces and investigate the whether they can reveal
+%the characteristics on the source code that may be favoring the introduction of faults 
+%(i.e. Uncaught Exceptions, and Unintened handler Actions), and consequently
+%decreasing the program robustness.
+%(of apps or the underlaying framework)
 
@@
 understood and tested parts of the system. As a consequence they may affect the system the 
-other way around, making the system even less robust.
+other way around, making the system even less robust, leading to failures such as uncaught exceptions 
+and unintended handler actions (i.e., exceptions being mistakenly caught by existing handlers).
 This paper reports a study which mined the stack traces embedded on 
 159,048 issues reported on Android projects hosted in GitHub (482 projs.) 
-and Googlecode (157 projs.). The goal of this study was to: identify common characteristics 
-of such traces and investigate the fragilities on the exception-related code that can be revealed
-from them. Overall 6,005 exception stack traces were extracted, and analyzed in 
+and Google Code (157 projs.). 
+The goal of this study was to: investigate the whether the exception stack traces
+ can reveal common characteristics on the exception-related code that
+may be favoring the introduction of the aforementioned failures, which would be responsible for 
+decreasing the application robustness. We call such characteristics ``fragilities" of the exception-related code.
+Overall 6,005 exception stack traces were extracted, and analyzed in 
 combination with source code and bytecode analysis. 
@@
 %GitHub 31,592
-% Googlecode: 127,456
+% Google Code: 127,456
 %All: 159,048
@@
 source projects that make their control version as well as the issue tracking systems
-publicly available in repository hosting sites such as GitHub and Googlecode.
+publicly available in repository hosting sites such as GitHub and Google Code.
 
@@
 %482 applications) and Google code (157 applications)
+%GitHub 31,592
+% Google Code: 127,456
+%All: 159,048
 
 To do so we perform a post mortem analysis of the exception stack traces embedded on all issues
-created on Android projects available in GitHub and Googlecode. Using a custom tool called ExceptionMiner,
+created on Android projects available in GitHub and Google Code. Using a custom tool called ExceptionMiner,
  which we develop specifically for this study, we mine stack traces from issues across all Android projects on
-GitHub and Googlecode. Overall, we analyzed all issues extracted from on 482 Android projects defined in
-
-%GitHub 31,592
-% Googlecode: 127,456
-%All: 159,048
-
-GitHub and 157 projects defined in Googlecode. Overall it corresponds to 159,048 issues (31,592 issues from GitHub projects 
-and 127,456 issues from Googlecode projects) from which  6,005 stack traces were extracted and used as
+GitHub and Google Code. Overall, we analyzed all issues extracted from on 482 Android projects defined in
+GitHub and 157 projects defined in Google Code. Overall it corresponds to 159,048 issues (31,592 issues from GitHub projects 
+and 127,456 issues from Google Code projects) from which  6,005 stack traces were extracted and used as
  a source of information for our analysis. The stack trace analysis was augmented by additional information
@@
 
-  \item  It performs a large scale post morten analysis of exception stack traces reported on Android projects hosted in GitHub and Googlecode.
+  \item  It performs a large scale post morten analysis of exception stack traces reported on Android projects hosted in GitHub and Google Code.
 
@@
 The goal of this mining study was to extract exception stack traces embedded on issues of Android projects 
-(hosted in GitHub and GoogleCode), looking for its 
+(hosted in GitHub and Google Code), looking for its 
 common characteristics and investigating whether
@@
 To answer the research questions the following information is needed: (i) the issues of reported on every Android project hosted on 
-GitHub and Googlecode; (ii) the stack traces embedded on those issues; (iii) the type of the exceptions
+GitHub and Google Code; (ii) the stack traces embedded on those issues; (iii) the type of the exceptions
  reported on the stack traces (e.g., error, runtime, checked); (iv) the origin of such exceptions 
 (e.g., the application, a library, the Android platform). Next sections describes how such information 
-was retrieved from GitHub and Googlecode, how the exceptions stack traces were extracted and distilled,
+was retrieved from GitHub and Google Code, how the exceptions stack traces were extracted and distilled,
  and how the type associated to each exception and its origins was identified.
@@
 
-This set of 482 projects contained overall 32,582 issues from which 4,042 exception stack traces 
+%GitHub 31,592
+% Google Code: 127,456
+%All: 159,048
+
+This set of 482 projects contained overall 31,592 issues from which 4,042 exception stack traces 
 were extracted. Issues on Github are different from issues on dedicated bug tracking tools such as 
@@
 them at will. Therefore, an issue may have none ore an arbitrary set of labels depending 
-on which repository it was created.
+on which repository it was created. Table ~\ref{tab:lables} illustrates the ocurrences of different lables 
+on the issues including exception stack traces.
+
+% \centering
+%  \begin{tabular}{|p{2cm}| p{5cm}|}
+%    \hline
+
+\begin{table}
+  \centering
+  \begin{tabular}{lr|lr}
+    \hline
+     \multicolumn{2}{c}{\bfseries{GitHub}} &  \multicolumn{2}{c}{\bfseries{Google Code}} \\
+      \bfseries{Lable} &  \bfseries{\% Occurrences} &  \bfseries{Lable} &  \bfseries{\% Occurrences} \\
+    \hline
+empty &	54,24\% & Defect &	91,96\% \\
+Defect &	39,56\%  & Enhancement  &	3,16\% \\
+Enhancement &	0,57\% & Task	& 1,37\% \\
+Support &	0,52\% & empty &	1,12\% \\
+Problem &	0,36\% & StackTrace &	0,70\% \\
+Others &	4,74\% &  Others &	1,68\% \\   
+  \hline
+  \end{tabular}
+  \caption{Lables on issues including exception stack traces.}
+  \label{tab:lables}
+\end{table}
 
@@
 
-\subsubsection{Android Apps in GoogleCode}
-
-Differently from GitHub, Googlecode does not provide an API to access the information related
+\subsubsection{Android Apps in Google Code}
+
+Differently from GitHub, Google Code does not provide an API to access the information related
  to hosted projects \footnote{Google code used to provided a Web service to such repository, but it was deactivated in June 2013 what Google called a "clean-up action".}.
 
-However, Googlecode still contain largely used open-source Android apps (e.g. K9Mail \footnote{K9Mail moved to Github but as a way of not loosing the project history 
-it advices their users to report bugs on Googlecode issue tracker.})  and the history of such apps are also valuable. 
+However, Google Code still contain largely used open-source Android apps (e.g. K9Mail \footnote{K9Mail moved to Github but as a way of not loosing the project history 
+it advices their users to report bugs on Google Code issue tracker.})  and the history of such apps are also valuable. 
 To overcome this limitation we needed to implement a Web Crawler (incorporated in Exception Miner tool described next) that navigated 
-through the web interface of Googlecode projects extracting all issues and issue comments and storing in a relational database for later analysis.
-
-To identify Android projects in Googlecode, we performed a similar heuristic, we performed a case insensitive search 
-(on Googlecode search interface) for the term \textsf{android}.  On January 2014, when we queried Googlecode, such heuristic could
+through the web interface of Google Code projects extracting all issues and issue comments and storing in a relational database for later analysis.
+
+To identify Android projects in Google Code, we performed a similar heuristic, we performed a case insensitive search 
+(on Google Code search interface) for the term \textsf{android}.  On January 2014, when we queried Google Code, such heuristic could
  find 788  projects, from which 724 defined at least 1 issue and, 183 projects 
@@
 reporting at least one stack trace, and as a result we could identify 157. 
-Android projects.  All issues created on such projects (i.e., 41,674 issues) were analyzed and 1,963 stack traces 
-were extracted.
-
-%For instance: K9mail  was hosted in Googlecode 
+Android projects.
+
+This set of 157 projects contained overall 127,456 issues from which 1,963 exception stack traces 
+were extracted. Based on the same assumption used for Github dataset selection, we did not retrict 
+the analysis for issues labled as Defect.
+
+%SELECT distinct(type), count(*) as n
+%  FROM exceptionalissue where repo !='android' group by type order by n desc;
+
+%For instance: K9mail  was hosted in Google Code 
 %before moving to GitHub and in order not to loose the project history it advices their 
-%users to create issues on Googlecode as a way to preserve project history.
-
-%To overcome this limitation and mine the information available on such Googlecode 
+%users to create issues on Google Code as a way to preserve project history.
+
+%To overcome this limitation and mine the information available on such Google Code 
 %projects we implemented a Web crawler which navigated (incorporated in Exception Miner tool described next) 
@@
  the host repositories in order to collect project issues. The GitHub Connector connected
- to GiHub through GHTorrent. The GoogleCode connector comprised of a Web Crawler,
- that traversed Googlecode Web interface in order to extract the project issues.
+ to GiHub through GHTorrent. The Google Code connector comprised of a Web Crawler,
+ that traversed Google Code Web interface in order to extract the project issues.
 
@@
 Specifically we provide: (i) all issues related to Android projects found
-in GITHub and GoogleCode used in this study; (iii) all stack traces extracted
+in GITHub and Google Code used in this study; (iii) all stack traces extracted
 from issues (ii) the results of manual inspection steps; (iv) the tool developed 
@@
   \end{tabular}
-\caption{Root Exceptions occurrences and popularity in repositories hosted in Googlecode (GC) and GitHub(GH).}
+\caption{Root Exceptions occurrences and popularity in repositories hosted in Google Code (GC) and GitHub(GH).}
 \label{tab:topten}
@@
 traces embedded in all issues defined in 482 Android projects hosted in Github and 
-157 projects hosted in Googlecode.
+157 projects hosted in Google Code.
 

### 2014-09-03 android-stacks.tex
addressing comments. introduction

@@
 have shown that it is generally poorly understood and the less tested parts of the system ~\cite{coelho2011unveiling}.
-As a consequence that the exception-related code may become itself a source of faults, 
-and may affect the system the other way around, making the system even less robust ~\cite{robillard2003static,coelho2011unveiling}.
+As a consequence they may affect the system the other way around.
+The exception-related code may favor the introduction of failures such as uncaught exceptions ~\cite{jo2004uncaught, zhang2012amplifying}
+and unintended handler actions ~/cite{miller1997issues} (i.e., exceptions being mistakenly caught by existing handlers) 
+which make the system even less robust.
 
@@
 
-Our conjecture is that several Android applications may be suffering from similar 
+Our conjecture is that several Android applications may be suffering from common 
 fragilities on the exception-related code (of the applications themselves or 
-the underlaying platform), and that such fragilities may be leading to
-decrease the application robustness, leading for instance to the Uncaught Exception problem -
- which is one of the main causes of application crashes ~\cite{jo2004uncaught, zhang2012amplifying}.
+the underlaying platform) which may be degratind the robustness of such applications.
+We call ``fragilities" characteristics on the exception-related code that favor the introduction
+of aforementioned failures.
+
+%, and that such fragilities may be leading to
+%decrease the application robustness, leading for instance to the Uncaught Exception problem -
+% which is one of the main causes of application crashes ~\cite{jo2004uncaught, zhang2012amplifying}.
+%investigate the whether the exception stack traces
+% can reveal common characteristics on the exception-related code that
+%may be favoring the introduction of the aforementioned failures, which would be responsible for 
+%decreasing the application robustness. We call such characteristics ``fragilities" of the exception-related code.
 
@@
 it automatically terminates, while the system prints a stack trace to the console, 
-or log file ~\cite{gosling2000java}.  A typical Java stack trace consists of an ordered list of 
-methods that were active on the call stack before the exception has occurred,  
- and the fully qualified name of such exception.
-
-Until recent years, the availability of issues reporting abnormal
-computation states (e.g., exception stack traces) was scarce. The
-available data comprised only few large scale projects (e.g. the Mozilla
-crash report dataset\footnote{https://code.google.com/p/promisedata/issues/detail?id=88}).
-Recently we have observed a rapid increase in the number of open
-source projects that make their control version as well as the issue tracking systems
-publicly available in repository hosting sites such as GitHub and Google Code.
+or log file ~\cite{gosling2000java}.  A typical Java stack trace consists of  the fully qualified name 
+the thrown exception and the ordered list of methods that were active on the call stack before 
+such exception has occurred.
 
@@
 
-%Stack traces are a useful source of information about system crashes, and is often used to support developers in debugging~\cite{schroter2010stack}.
-Moreover, information %from stack traces combined with crash and bug reports can be used to support bug classification and clustering~\cite{wang2013improving, kim2011crash, %dhaliwal2011classifying}, fault-proneness prediction models~\cite{kim2013predicting} and even automated bug fixing~\cite{sinha2009fault} tools.
-
-When available, exception stack traces provides a useful source of information about system crashes ~\cite{bettenburg2008makes} which 
- can enable different kinds of post-mortem analysis to support:  debugging~\cite{schroter2010stack}, 
+%Stack traces are a useful source of information about system crashes, and is often used to support developers in debugging~\cite{schroter2010stack}. Moreover, information %from stack traces 
+%combined with crash and bug reports can be used to support bug classification and clustering~\cite{wang2013improving, kim2011crash, %dhaliwal2011classifying}, fault-proneness prediction %models~\cite{kim2013predicting} and even automated bug fixing~\cite{sinha2009fault} tools.
+
+Until recent years, the availability of data reporting such failures was scarce. It comprised only few large scale projects (e.g. the Mozilla
+crash report dataset\footnote{https://code.google.com/p/promisedata/issues/detail?id=88}).
+Recently we have observed a rapid increase in the number of open
+source projects that make their version control systems as well as the issue tracking systems
+publicly available in repository hosting sites such as GitHub and Google Code.
+
+When available, the exception stack traces provide a useful source of information about system crashes ~\cite{bettenburg2008makes} which 
+ can enable different kinds of post-mortem analysis and support:  debugging~\cite{schroter2010stack}, 
  bug classification and clustering ~\cite{wang2013improving, kim2011crash, dhaliwal2011classifying},  
@@
 the applications. The assumption is impossible to mitigate without access to
-the full set of crash data per application. Services exist to collect all
-crash data from applications, so access to such a dataset will allow for
-a thorough replication of our analysis, perhaps at the individual application
-level.
+the full set of crash data per application. Some services exist to collect 
+crash data from mobile applications ~\cite{BugSe14,BugSn14,Googl14,Acra14}.
+However, such services do not provide open access to the crash reports
+of their client applications. Having access to such a dataset would allow 
+a replication of our analysis on an even wider setting.
 

### 2014-09-03 android-stacks.tex
addressing comments. introduction2

@@
 
+%\title{Results of a Mining Study on Exception Stack Traces embedded on issues of over 600 Android Projects}
+
 \title{Discovering Robustness Threats by Mining the Exception Stack Traces of over 600 Android Projects}
+
+%\title{Discovering Threats to Robustness through Exception Stack Trace Analysis: Results of a Mining Study of Android Projects}
+
+%\title{Discovering Threats to Robustness through Exception Stack Trace Mining}
+
+%\title{Discovering Threats to Robustness through Exception Stack Trace Mining of over 600 Android Projects}
+
+%\title{Discovering Threats to Robustness through a Mining Study of Exception Stack Traces}
 
@@
 the thrown exception and the ordered list of methods that were active on the call stack before 
-such exception has occurred.
+such exception has occurred ~\cite{gosling2000java,bloch2008effective}.
 
@@
 
-Most of such studies, however, focused the analysis on one or two system at a time.
-This paper reports an exploratory study which targets a family of applications 
-(sharing the common Android framework) aiming at extracting 
-the common characteristics of exception stack traces reported to such systems to provide
-insights on similar fragilities on the exception related code of the applications and underlaying framework.
+%Most of such studies, however, focused the analysis on one or two system at a time.
+%This paper reports an exploratory study which targets a family of applications 
+%(sharing the common Android framework) aiming at extracting 
+%the common characteristics of exception stack traces reported to such systems to provide
+%insights on similar fragilities on the exception related code of the applications and underlaying framework.
 
@@
 
-To do so we perform a post mortem analysis of the exception stack traces embedded on all issues
-created on Android projects available in GitHub and Google Code. Using a custom tool called ExceptionMiner,
- which we develop specifically for this study, we mine stack traces from issues across all Android projects on
-GitHub and Google Code. Overall, we analyzed all issues extracted from on 482 Android projects defined in
-GitHub and 157 projects defined in Google Code. Overall it corresponds to 159,048 issues (31,592 issues from GitHub projects 
-and 127,456 issues from Google Code projects) from which  6,005 stack traces were extracted and used as
- a source of information for our analysis. The stack trace analysis was augmented by additional information
-extracted from JDK, Android Plaform and the applications using byte code and source code analysis.
- Manual inspection was used to leverage the understanding of exception stack traces and support further discussions and insights.
+This paper reports a study which perform a post mortem analysis of the exception stack traces embedded on issues
+reported on Android projects hosted in GitHub and Google Code. 
+The goal of this study was to: investigate the whether the reported exception stack traces
+ can reveal common fragilities on the exception-related code which may be favoring 
+the introduction of failures such as uncaught exceptions - and as a consequence
+ contributing to deacrease the robustness of apps.
+To guide this investigation we compiled general guidelines on how to use
+exceptions proposed by Gosling~\cite{gosling2000java},
+Wirfs-Brock~\cite{wirfs2006toward} and Bloch~\cite{bloch2008effective}.
+Then, using a custom tool called ExceptionMiner,
+ which we develop specifically for this study, we mine stack traces from the issues reported 
+on 482 Android projects hosted in GitHub and 157 projects hosted in Google Code.
+Overall 159,048 issues were analyzed and 6,005 stack traces were extracted from them.
+The exception stack trace analysis was augmented with additional information
+extracted from JDK, Android Plaform and the applications using byte code 
+and source code analysis. Manual inspection was used to leverage the understanding 
+of exception stack traces and support further discussions and insights.
 
@@
 
-Such fragilities represent obstacles to the development of dependable java-based applications as they make difficult and 
+Such fragilities represent obstacles to the development of robust java-based applications as they make difficult and 
 sometimes almost impossible to protect the code against the Uncaught Exceptions that lead to system crashes.

### 2014-09-03 android-stacks.tex
addressing comments. introduction3

@@
 
-\title{Discovering Robustness Threats by Mining the Exception Stack Traces of over 600 Android Projects}
+\title{Discovering/Unveiling Robustness Threats by Mining the Exception Stack Traces of over 600 Android Projects}
 
@@
 to be wrapped in another one and re-thrown. Figure~\ref{fig:wrapping} presents 
-a stack trace  illustrates an exception wrapping. The bottom 
-part of the stack trace is the \emph{root cause}, which indicates the
+an exception stack trace which illustrates an exception wrapping. 
+For simplicity we will refer to exception stack trace as stack trace along the text.
+The bottom part of the stack trace is the \emph{root cause}, which indicates the
 first reason for the error to be thrown (in this case, the computer run out of
@@
 
-\emph{\textbf{Classifying Top 100 Exceptions.}} To get a broader view of the root causes of stack traces,
- we performed an empirical evaluation
-whose goal was to categorized the top 100 root exceptions. The top 100 root causes correspond 
-to approximately 95\% of all exception types reported on the exception stack-issues.
-
-Such categorization was based known sources of faults mobile development: concurrency ~\cite{ama2012}
-, backward compatibility  ~\cite{McDon13}, security   ~\cite{enck2011study,was2010} and resource management (IO, Memory, Batery) ~\cite{Zhang12}.
-We added to this set of categories other three categories: one to represent the exceptions caused by programming mistakes;
-  another to represent general Java exceptions (e.g., java.lang.Exception) - whose types do not convey any specific cause of failure; and a category to
-represent specific  exceptional situations of apps, frameworks and libraries.
-The list of categories is illustrated in Table~\ref{tab:categories}.
-
-\begin{table}
-  \centering
-  \begin{tabular}{|p{2cm}| p{5cm}|}
-    \hline
-    \bfseries{Category} & \bfseries{Description} \\
-    \hline
-      Programming logic &  Exceptions thrown by the JVM when the 
-evaluation of an expression violates the normal semantics of language (e.g., 
-out-of-bounds array index, division-by-zero, access to a null reference). \\ \hline
-      Resources (IO, Memory, Batery)  & Exception releated to the manipulation or restriction of resources (e.g., file, network, memory) \\ \hline
-      Security                               & Exception releated to security issues (e.g. password validation, criptography) \\ \hline
-      Concurrency                            & Exceptions related to multi-threaded programming \\ \hline
-      Backward compatibility                 &  A list of exceptions related to backward compatibility - a complete list can be found here [ref]  \\ \hline
-      Specific Exceptions              & Exceptions created for specific applications, frameworks or libraries \\ \hline
-      General (Exception, Error, Runtime)    & The top level exceptions in Java    \\ \hline
-    \hline
-  \end{tabular}
-  \caption{Description of categories.}
-  \label{tab:categories}
-\end{table}
-
-Each top 100 root exceptions was mapped to one of these categories - 
-this analysis was based on the inspection of source code, Javadoc, or additional 
-documentation related to each exception.  Table~\ref{tab:tophundrend} 
-illustrates the result of such classification, the number of exception stack traces
-on which the exceptions of each category was found as the root cause.
+\emph{\textbf{Classifying the Most Reported Root Exceptions.}} To get a broader view of the root causes of stack traces,
+ we performed an empirical evaluation whose goal was to identify the concerns related to the most reported root exceptions.
+
+Since it is infeasible to inspect the code responsible for throwning every exception reported in this study.
+This identification performed in this study was based on meaning that the name of the exception holds, 
+its description (i.e., Javadoc documentaion), and the Java specification. To exemplify, an instance of ArrayOutOfBoundException
+ usually refers to a programming mistake. Moreover, the Java specification list all exceptions related to backwardcompatibility 
+issues ~\cite{javaback} (e.g., InstantiationError, VerifyError, IllegalAccessError).
+
+Some concerns are known as sources of faults in mobile development are: concurrency ~\cite{ama2012} backward compatibility  
+~\cite{McDon13}, security  ~\cite{enck2011study,was2010} and resource management (IO, Memory, Batery) ~\cite{Zhang12}.
+Moreover, the programming errors are also as a source of faults, as illustrated on the previous section.
+However, there are exceptions that cannot be directly mapped to one of such concerns: either because they are too general 
+(e.g., java.lang.Exception) and its name does not hold a reason for the exception to be thrown; or because they are related to other concerns 
+ (e.g., specific to a given library).
+
+%The list of categories is illustrated in Table~\ref{tab:categories}.
+
+To perform such mapping, we selected a subset of all reported root exceptions. This subset was composed by 100 root exceptions 
+which were reported on 95\% of all stack traces analyzed in this study.
+
+Hence, based on the inspection of the source code and Javadoc related to each exception,  and Java specification, we identified the concern
+releated to each root exception. Than we could calculate the prevalence of a concern on the reported stack traces using the following metric.
+Given that: Stacks is the complete set of stacks found in this study, Lc is the list of root causes related to a given concern (as identified on the study),
+ and Stacks[Lc] the list of issues that mentions the concern. The occurence of a concern could be given by the formula: Stacks[Lc]/Stacks.
+
+Table~\ref{tab:tophundrend} illustrates the result of such mapping.
+
+% When the root exceptions that were too general, or related to other concerns 
+% were also maped to these additional categories. Table~\ref{tab:tophundrend} illustrates the result of such mapping.%
+%\begin{table}
+%  \centering
+%  \begin{tabular}{|p{2cm}| p{5cm}|}
+%    \hline
+%    \bfseries{Category} & \bfseries{Description} \\
+%    \hline
+%      Programming logic &  Exceptions thrown by the JVM when the 
+% evaluation of an expression violates the normal semantics of language (e.g., 
+% out-of-bounds array index, division-by-zero, access to a null reference). \\ \hline
+%      Resources (IO, Memory, Batery)  & Exception releated to the manipulation or restriction of resources (e.g., file, network, memory) \\ \hline
+%      Security                               & Exception releated to security issues (e.g. password validation, criptography) \\ \hline
+%      Concurrency                            & Exceptions related to multi-threaded programming \\ \hline
+%      Backward compatibility                 &  A list of exceptions related to backward compatibility - a complete list can be found here [ref]  \\ \hline
+%      Specific Exceptions              & Exceptions created for specific applications, frameworks or libraries \\ \hline
+%      General (Exception, Error, Runtime)    & The top level exceptions in Java    \\ \hline
+%    \hline
+%  \end{tabular}
+%  \caption{Description of categories.}
+%  \label{tab:categories}
+% \end{table}
 
@@
   \end{tabular}
-  \caption{Characterization of the 100 root causes.}
+  \caption{Identifying the concerns related to root exceptions}
   \label{tab:tophundrend}
@@
 
-To ensure the quality of the classification, three independent coders classified a randomly selected
-sample of 25 exception types (from the total 100) using the same categories;
+To ensure the quality of the process, three independent coders classified a randomly selected
+sample of 25 exception types (from the total 100) using the same list of concerns;
 the interater agreement was high (96\%). We could also observe that
@@
 
-Analizing the Top 100 exceptions the programming mistakes
-were also the most reported. However, they may be interactly related 
-to one of the other concerns. For  instance, a concurrency problem may 
-lead to a NullPointerException. This classification however, 
-was only based on the semantic embedded  on the exception types.
+These study showed that the programming mistakes associated to the most reported root causes. 
+However, they may be interactly related to one of the other concerns. For  instance, a concurrency problem may 
+lead to a NullPointerException.
 
@@
 
-\emph{``Everybody hates thinking about exceptions, because they are not supposed to happen."}
-\textbf{Brian Foote}
-
-\subsection{Inversion of Control and Exception Handling}
-
-footnote{to do: INPORTANTE Falar que os metodos do ciclo de vida da app nao tem interface excepcional.
-desta forma ela não define: O que fazer se o metodo der errado? Nao deixa claro.}
-
-Reasoning about the design strategy adopted by the ActivityThread,
-we can observe that the inversion of control (on which this class and 
-most of the Android framework is based) can make it difficult for the application developer to 
-associate a specific handler to the exceptions thrown during the execution of the onCreate method (or any other life cycle method).
-According to Java handling model, the handler should be added on a method on the invocation chain. 
-However,  in the inversion of control development model the developer does not have fully access to 
-the invocation chain of a method. Therefore, the developer may be tempted to convert the checked
- exception into a RuntimeException and resignal.
-
-Difficulty of associating a specific handler to exceptions signaled by application methods called by the framework,
-may also be the reason behind the design decision of encapsulating any exception in a RuntimeException.
-As framework does not have enough contextual information the exception regardless its types,
-it encapsulates in a general type and re-throws. The only possible answer to such exceptions would be 
-to finish the application.
-
-Such design decision may lead the developers into believing that by just re-throwing 
-the exceptions they can forget about the exceptional situations during the development of the "happy path". 
-This "ignore-for-now" approach may turn the exception handling into a generalized "goto" mechanism~\cite{mandrioli1992advances}
- making the program more complex and even less reliable. This behavior may lead to the well known uncaught exceptions 
-problem~\cite{jo2004uncaught}, one of the main causes of application crashes.
-
 \subsection{On the need of a rescue to the almost infeasible task of preventing uncaught exceptions}
@@
 %\noindent\emph{How can we benefit from such cross-project crash analysis?}
+
+%\subsection{Inversion of Control and Exception Handling}
+\subsection{On the absence of exception interface on hook methods of the Android Aplication Framework}
+
+In this study, the analyzed stack traces involved, in one way or another, methods from the 
+Android Application Framework: either framework utility methods inkoked 
+by the application, or framework hook methods  (e.g., OnCreate(), onPause())  
+ extended by the application and invoked by the framework (by inversion of control).
+ As discussed before, some of the utility 
+methods signaled ``undocumented" runtime exceptions. Additionaly, another characteristic of 
+Android Platform may also be imposing obstacles to application robustness: the absence 
+of exception interface (i.e., a list of exceptions that a method 
+might raise during its execution) of Android rook methods.
+
+For instance, none of the hook methods povided by the Activity class (and 
+extended by every Android app) specify an exception interface. Therefore, such methods
+do not stimulate the developer to think that ``something" may go wrong 
+during the execution of such methods, and prepare the code to deal with them.
+
+To make matters worse, according to Java handling model, the exception handler
+ should be added on the invocation chain of the signaler.  However,  on method invocations based 
+on inversion of control, the developer does not have fully access to the invocation 
+chain of a method.
+
+Such difficulties may lead the developers into believing that by just re-throwing the exceptions they can 
+forget about the exceptional situations, while focusing on the developement of the `happy path". 
+This `ignore-for-now" approach may lead to the known uncaught exceptions problem~\cite{jo2004uncaught},
+and unintened handler actions,  contributing to degrate apps robustness.
+
+Rephasing Brian Foote "Everybody hates thinking about exceptions, because they’re not supposed to happen"
+however one day they happen, and you need to prepare your code for it. However, further investigation
+ is needed to evaluate the impact of  the absence of exception interface of Android rook methods on
+app robustness.
 

### 2014-09-03 android-stacks.tex
all over

@@
 
-\title{Discovering/Unveiling Robustness Threats by Mining the Exception Stack Traces of over 600 Android Projects}
+\title{Unveiling Robustness Threats by Mining the Exception Stack Traces of over 600 Android Projects}
 
@@
 other way around, making the system even less robust, leading to failures such as uncaught exceptions 
-and unintended handler actions (i.e., exceptions being mistakenly caught by existing handlers).
+and unintended handling (i.e., exceptions being mistakenly caught by existing handlers).
 This paper reports a study which mined the stack traces embedded on 
@@
 (1) undocumented runtime exceptions thrown by third party code;
-(2) undocumented checked exceptions thrown by Android native code;
+(2) undocumented checked exceptions thrown by JNI native code;
 (3) unexpected exception wrappings (e.g., Errors being wrapped in checked exceptions),
@@
 The exception-related code may favor the introduction of failures such as uncaught exceptions ~\cite{jo2004uncaught, zhang2012amplifying}
-and unintended handler actions ~/cite{miller1997issues} (i.e., exceptions being mistakenly caught by existing handlers) 
+and unintended handling ~/cite{miller1997issues} (i.e., exceptions being mistakenly caught by existing handlers) 
 which make the system even less robust.
@@
 
-Until recent years, the availability of data reporting such failures was scarce. It comprised only few large scale projects (e.g. the Mozilla
-crash report dataset\footnote{https://code.google.com/p/promisedata/issues/detail?id=88}).
+Until recent years, the availability of data, reporting such failures, was scarce. 
+It comprised only few large scale projects (e.g. the Mozilla crash report dataset\footnote{https://code.google.com/p/promisedata/issues/detail?id=88}).
 Recently we have observed a rapid increase in the number of open
@@
 
-This paper reports a study which perform a post mortem analysis of the exception stack traces embedded on issues
+This study performs a post mortem analysis of the exception stack traces embedded on issues
 reported on Android projects hosted in GitHub and Google Code. 
@@
 
-Such fragilities represent obstacles to the development of robust java-based applications as they make difficult and 
-sometimes almost impossible to protect the code against the Uncaught Exceptions that lead to system crashes.
-Although the study focused on Android open-source applications, these findings are not restricted to this context, since 
-the above mentioned fragilities may happen in any Java-based system. 
-For instance: the unexpected wrappings were performed by Java libraries and elements (e.g. reflection library and 
-the static blocks); and the undocumented runtime exceptions were also found in Java libraries besides the Android platform.
-The study results are relevant to: (i) Android and Java developers to familiarize common characteristics and fragilities of exception-related code 
-which is the first step to help developers preventing and dealing with them; (ii) designers of languages and tools to consider ways of reducing 
-such fragilities in Android and Java environments.
-
-Hence, the main contributions of this study are as follows:
+As we can see, although the study was based on a subset of Android projects, 
+the findings are not limited to this setting. The detected fragilities make difficult
+or sometimes infeasible for a Java developer to prevent failures such as 
+uncaught exceptions (e.g., thrown silently by third-party utility code), and 
+unintended handling ( e.g., when an Error is wrapped in a checked Exception).
+Thereore, they impose obstacles not only to the developement of robust Android apps, 
+but also to the developement of any robust Java-based system. 
+Hence, the study results are relevant to: (i) Android and 
+Java developers to familiarize common characteristics and fragilities of exception-related code 
+which is the first step to help developers preventing and dealing with them; and (ii) designers of languages and tools 
+to consider ways of reducing such fragilities in Android and Java environments.
+The main contributions of this study is twofold:
 \begin{itemize}
@@
   \item  It performs a large scale post morten analysis of exception stack traces reported on Android projects hosted in GitHub and Google Code.
-
-  \item  It introduces ExceptionMiner, a tool developed to support cross-repository stack trace
+  \item  It introduces ExceptionMiner, a tool developed to support exception stack trace mining and analysis.
     analysis.
@@
 background on Android Platform and Java exception model. 
-Section 3 presents the study design. Section 4 reports study findings. 
+Section 3 presents the study design. Section 4 reports the study findings. 
 Section 5 provides further discussions and insights.
@@
 \subsection{The Android Platform} \label{sec:extypes}
-
 Android is an open source platform for mobile devices based on Linux kernel,
-which also contains  a core Java library, a set of libraries third-party libraries (e.g.  SQLite),
+which also comprises a subset of the core Java library, a set of third-party libraries (e.g.  SQLite),
 the Dalvik runtime environment, and the Application Framework (which provides 
 higher-level services to the applications running on the platform) ~\cite{andguide}.
-Android platform is mostly implemented in Java and uses Java Native Invocation (JNI) to access native C/C++ code. 
-And as mentioned before, it reuses the embedded Java Exception model to detect and handle 
-abnormal computation states.
+
+%The Android platform is mostly implemented in Java and uses Java Native Invocation (JNI) to access native C/C++ code. 
+%And as mentioned before, it reuses the embedded Java Exception model to detect and handle 
+%abnormal computation states.
 
@@
 Among the basic elements that comprise an Android application, we can cite the
-Activity. Each application usually defines one Activity per application window.
- Each Activity defines a set of hook methods  (e.g., OnCreate(), onPause()) that are dispatched
-automatically by the Android framework (by inversion of
-control), in response to infrastructure and application events.
+ Activity class. The Activity is the main class of an Android app (there is usualy one
+Activity per application window). Each Activity defines a set of hook methods  
+(e.g., OnCreate(), onPause()) that are dispatched automatically by the Android 
+framework (by inversion of control), in response to infrastructure and application events.
 
@@
 (extends Exception), the runtime exceptions (extends RuntimeException) and errors
-(extends Error)~\cite{gosling2000java}.
-
-The checked exceptions must be specified on the method signature that may throw it, 
-and the compiler statically checks if appropriate handlers are provided within the system.
-Both runtime exceptions and errors are known as unchecked exceptions, because 
-they do not need to be specified on method signatures and do not trigger any 
+(extends Error)~\cite{gosling2000java}. The checked exception received such name
+ because they must be declared on the method's exception interface (i.e., a list of exceptions that a method 
+might raise during its execution) and the compiler statically checks if
+ appropriate handlers are provided within the system.
+Both runtime exceptions and errors are also known as unchecked exceptions, because 
+they do not need to be specified on method exception interface and do not trigger any 
 compile time checking.
@@
 ~\cite{McDon13}, security  ~\cite{enck2011study,was2010} and resource management (IO, Memory, Batery) ~\cite{Zhang12}.
-Moreover, the programming errors are also as a source of faults, as illustrated on the previous section.
-However, there are exceptions that cannot be directly mapped to one of such concerns: either because they are too general 
-(e.g., java.lang.Exception) and its name does not hold a reason for the exception to be thrown; or because they are related to other concerns 
- (e.g., specific to a given library).
+
+%Moreover, the programming errors are also as a source of faults, as illustrated on the previous section.
+
+However, there are exceptions that cannot be directly mapped to one of such concerns. Either because they are
+related to programming mistakes - as illustrated on the previously. Or because they are too general (i.e.,  java.lang.Exception,
+java.lang.RuntimeException, java.lang.Error). Or also because they are related to other concerns (e.g., specific to a given library).
 
@@
 Given that: Stacks is the complete set of stacks found in this study, Lc is the list of root causes related to a given concern (as identified on the study),
- and Stacks[Lc] the list of issues that mentions the concern. The occurence of a concern could be given by the formula: Stacks[Lc]/Stacks.
+ and Stacks[Lc] the list of issues that mentions the concern. The prevalence of a concern could be given by the formula: Stacks[Lc]/Stacks.
 
@@
 Figure ~\ref{fig:snippets} (a) presents a code snippet of one the methods of 
-ActivityThrea,  the one responsible for invoking the Activities onCreate method. 
+ActivityThrea,  the one responsible for invoking the Activitiy's onCreate method. 
 As we can see from lines 4-8,
@@
 
-However, since the Activity life clycle method (i.e., callActivityOnCreate(...)) that is invoked in this block,
- do not include checked exceptions on its signatures. Hence they could not throw a checked exception to be handled 
-in the method illustrated above. However the code inspection performed on such wrappings revelaed an
-``undocumented checked exception" signaled by this life cycle method.
-
-\emph{\textbf{Undocumented Checked Exceptions.}} There are usually
-2 possible possibilities for a checked exception to be thrown by a method:
-it can either be specified on the method signature or it can be wrapped in a 
-runtime exception and re-thrown. However by inspectioning associated to the reported wrappings 
-revealed the a third option: a checked exception thrown by a native method, which was not declared
-on the method signature of the methods on the call chain composing the stack.
-
-The code snippet in Figure ~\ref{fig:snippets} (d) illustrate this scenario.
+%However, since the Activity life clycle method (i.e., callActivityOnCreate(...)) that is invoked in this block,
+% do not include checked exceptions on its signatures. Hence they could not throw a checked exception to be handled 
+%in the method illustrated above. However the code inspection performed on such wrappings revelaed an
+%``undocumented checked exception" signaled by this life cycle method.
+
+%\emph{\textbf{Undocumented Checked Exceptions.}} There are usually
+%2 possible for a checked exception to be thrown by a method:
+%it can either be specified on the method signature or it can be wrapped in a 
+%runtime exception and re-thrown.
+
+Inspectioning the source code, of the methods involved in this kind of wrapping,
+revealed another fragility: a checked exception thrown by a native method and not
+declared on the exception interface of thes methods signaling them. Such native method 
+was defined on Android Platform -  which uses Java Native Invocation (JNI) to access 
+native C/C++ code.
+
+The code snippet in Figure ~\ref{fig:snippets} (d) illustrates this fragility.
  The java/lang/Class declares a native method named getDeclaredMethods. 
@@
 traces embedded in all issues defined in 482 Android projects hosted in Github and 
-157 projects hosted in Google Code.
-
+157 projects hosted in Google Code. Overall it included 6,005 exception stack traces.
 In this study, the information extracted from stack traces was used in combination 
-with source code and bytecode analysis to pin-point fragilities on the exception-related
-code. 
-This study  consistently identifies common characteristics of reported 
-exception stacks traces - e.g., the NullPointerException was the reported cause 
-on 30,08\% of all analyzed exception stack traces, and was mentioned in approximately 
-50\% of all projects analyzed. Moreover,  the cross-project analysis of exception stack traces 
-enabled to consistently reveal  fragilities of the exception-related code: 
-(i) undocumented runtime exceptions signaled by third party code,
-(ii) undocumented checked exceptions thrown by Android native code
-(iii) unexpected exception wrappings (e.g., Errors being wrapped in checked exceptions) 
+with source code and bytecode analysis to identify common characteristics of exception 
+stack traces and pin-point fragilities on the exception-related code. 
+We call ``fragilities" characteristics on the exception-related code that favor the introduction
+of  uncaught exceptions and unintended handling - which can contribute to 
+degrade the robustness of applications.
+Some fragilities consistently detected in this study were: 
+(i) an abundance of NullPointerExceptions - it was mentioned as the root exception in approximately 
+50\% of the analyzed projects. 
+(ii) undocumented runtime exceptions thrown by third party code,
+(iii) undocumented checked exceptions thrown by JNI native code,
+(iv) unexpected exception wrappings (e.g., Errors being wrapped in checked exceptions) 
 revealing that Java hybrid exception model is not fully used according to its purpose.
-Such fragilities detected in our study can negatively affect the 
-robustness of not only Android application but also Java-based 
-which can deal with the same fragilities pointed in thid work. 
+As we can see, such fragilities can negatively affect not only the robustness of Android application 
+but also the robustness of any Java-based application. 
 Our results calls the attention for the need of tool and language support to help 
-developers when dealing with such problems.
+developers when dealing with such fragilities.
 

### 2014-09-03 android-stacks.tex
all over

@@
 
-%GitHub 315,92
+%GitHub 31,592
 % Google Code: 127,456
 %All: 159,048
+% bf: 31,592 (with exception)
 
@@
 %far beyond of the basic calls and textual messages. 
-
+315,92
 In this context, Android has become a leading platform - thanks to its open-source 
@@
 normally should not be handled inside the application. The user-defined exceptions 
-can be either checked or runtime. Although the Java specification~\cite{gosling2000java} 
-suggests that user-defined exceptions should be checked exception (because doing so 
+can be either checked or runtime. The Java specification~\cite{gosling2000java} 
+suggests that user-defined exceptions should be checked (because doing so 
 the callers of a method will know about the exceptions that a method can throw and so 
-they can decide what to do about them) there is a long lasting debate about the use of 
-checked and unchecked exceptions in Java.
+they can decide what to do about them). However, runtime exceptions can also be 
+used to represent user-defined exceptional conditions.In fact, there is a long-lasting debate 
+about the pros and cons of both approaches ~\cite{javatut,stackoverlow,debate},
+\footnote{152 questions in Stackoverlow are related to this debate}.
 
@@
 
-The runtime exceptions, however are not only used to represent user-defined
-exceptional conditions, they are used by the JVM when a program violates 
-the semantic constraints of the Java programming language (out-of-bounds array index, divide-by-zero 
-error, null pointer references) from which most of the programs are not expected to recover. 
-\footnote{Some programming languages react to such errors by peremptorily terminating the program; 
+Besides, the runtime exceptions are also implicitly thrown by the JVM when a program violates 
+the semantic constraints of the Java programming language (e.g., out-of-bounds array index, divide-by-zero 
+error, null pointer references).\footnote{Some programming languages react to such errors by peremptorily terminating the program; 
 other programming languages allow similar approaches e.g.C\#; and others let the program to continue
  its execution in some situations such as the out-of-bounds array index e.g., C++. }  
+The programs are not expected to handle such JVM runtime exceptions ~\cite{gosling2000java} s.
 
@@
 (in Java it is represented by the try-catch block), and unwinds the execution stack if necessary.
- To look for the handler on the invocation stack is claimed to increase software reusability, 
+ Looking for the handler on the invocation stack is claimed to increase software reusability, 
 since the invoker of an operation can handle it in a wider context ~/cite{miller1997issues}.
@@
  (also known as exception chaining). Exception wrapping allows one exception 
-to be wrapped in another one and re-thrown. Figure~\ref{fig:wrapping} presents 
+to be wrapped in another one and re-thrown. Figure~\ref{fig:wrapping} shows 
 an exception stack trace which illustrates an exception wrapping. 
 For simplicity we will refer to exception stack trace as stack trace along the text.
-The bottom part of the stack trace is the \emph{root cause}, which indicates the
-first reason for the error to be thrown (in this case, the computer run out of
+The bottom part of the stack trace is the \emph{root exception}, which indicates
+the first reason (root cause) for the exception be thrown (in this case, the computer run out of
 memory). The top part of the stack trace indicates the location of the exception
-manifestation (which will call \emph{top level wrapper} along this paper). The
-execution flow  between the root cause and the exception manifestation may
-include other intermediate exception wrappings. In all levels, the exception
+manifestation (which will call \emph{exception wrapper} along this paper). The
+execution flow  between the root exception and the wrapper may
+include other intermediate exception wrappers. In all levels, the exception
 \emph{signaler}, is the method that threw the exception, represented on the
@@
 
-There is a long-lasting debate about the pros and cons of checked and unchecked exceptions 
-~\cite{javatut,stackoverlow,debate},\footnote{152 questions in Stackoverlow
-are related to this debate}. Such debates, however, are mainly based on the developers own experience in
-using both approaches - since there is a lack of empirical studies on their the pros and cons.
-
-Nevertheless, several general guidelines have been proposed on how to use Java
+Several general guidelines have been proposed on how to use Java
 exceptions~\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,
@@
 mistakes} as carefully as checked exceptions. Doing so, the clients of a method
-will be aware of all the exceptions the method can throw and can protect its code
-against unforeseen exceptions.
+will be aware of all the exceptions the method can throw and can protect design the code
+to deal with them ~\cite{Robil00}.
 
 If the developer fails to do follow such practices (specially when developing 
-library/reusable code) it will be difficult or even impossible for the caller to 
+utility code) it will be difficult or even impossible for the caller to 
 make effective use of such method~\cite{wirfs2006toward, bloch2008effective},
-and protect the system against the uncaught exceptions - one of the main causes of 
-systems crashes.
-
-These best practices guided our exploratory study described next.
-Based on information extracted from stack traces (complemented by
-bytecode and source-code analysis), we investigate whether stack characteristics
-can reveal whether these practices have been obeyed.
+and protect the system against unforeseen exceptions thrown by the utility code,
+which can lead to the uncaught exceptions - one of the main causes of 
+systems crashes. These best practices guided our exploratory study dscribed next.
 
@@
 
-\section{Empirical Study Design}
+\section{Study Design}
 \label{sec:study}
@@
 Our study focused on open-source apps, since the information needed to perform our study
-cannot be retrieved from commercial apps, whose issue reports and 
-source code are not publicly available. The Android open-source apps were also 
+cannot be retrieved from commercial apps, whose issue report systems and 
+source codes are not publicly available. The Android open-source apps were also 
 target of other research studies ~\cite{Linar13,ahimed}.
@@
   \item[RQ1] \noindent\emph{What are the common characteristics of exception stack traces reported on issues?}
-Some of the common characteristics to be investigated are the exceptions reported as the root cause on
-the exception stack trace, the origin of such exceptions (e.g., the application, a library, the Android platform).
 
@@
 
+     \item
+
 %Which patterns of exception (mis)use emerge?
   \item[RQ2] \noindent\emph{Which fragilities on the exception-related code emerge from them?}
-Based on a set of best practices compiled in this study, we investigated whether characteristics of exception stack traces  
-can pinpoint best practices violations. Such  violations reveal fragilities
-on the exception-related code, which threats the application dependability.
 
@@
 % on the exception-related code, which threats the application dependability.
+
+Some of the common characteristics to be investigated are the exceptions reported as the root cause on
+the exception stack trace, the origin of such exceptions (e.g., the application, a library, the Android platform). 
+Based on a set of best practices compiled in this study, we investigated 
+whether characteristics of exception stack traces can pinpoint best practices violations. 
+Such  violations reveal fragilities
+on the exception-related code, which threats the application dependability.
 
@@
 
-To answer the research questions the following information is needed: (i) the issues of reported on every Android project hosted on 
-GitHub and Google Code; (ii) the stack traces embedded on those issues; (iii) the type of the exceptions
+To answer the research questions the following information is needed: (i) the issues reported on Android project hosted on 
+GitHub and Google Code; (ii) the stack traces embedded on these issues; (iii) the type of the exceptions
  reported on the stack traces (e.g., error, runtime, checked); (iv) the origin of such exceptions 
-(e.g., the application, a library, the Android platform). Next sections describes how such information 
-was retrieved from GitHub and Google Code, how the exceptions stack traces were extracted and distilled,
- and how the type associated to each exception and its origins was identified.
+(e.g., the application, a library, the Android platform). Next sections describe: how such information 
+was retrieved from GitHub and Google Code; how the exceptions stack traces were extracted and distilled,
+ and how the type associated to each exception and its origins was identified, and finally how manual
+inspections supported this work.
 
@@
 Up to 23 Feb 2014,  when we queried GHTorrent, such heuristic could
- find 2.542 repositories,  from which 589 apps had at least one
-issue containing a stack trace.
+ find 2.542 repositories. Running the ExceptionMiner tool (described next)
+ in this set we observed that 589 apps had at least one issue containing a stack trace.
 
@@
 (e.g. Selendroid, Roboeletric - tools to support the testing of Android apps).
-The filtered set consisted of 482 apps.
-
-%4,208
-
-%GitHub 31,592
-% Google Code: 127,456
-%All: 159,048
-
-This set of 482 projects contained overall 31,592 issues from which 4,042 exception stack traces 
+The filtered set consisted of 482 apps. This set of 482 projects contained overall 31,592 issues from which 4,042 exception stack traces 
 were extracted. Issues on Github are different from issues on dedicated bug tracking tools such as 
@@
 ``backing'' issue which are automatically generated. Therefore, we excluded such automatically generated
-issues from our analysis. Finally, the Github API allows the automated
-generation of issues for specific repositories, which automated tools often use
-to report crashes. In some cases, this led to high number of issues that
-included stack traces. We identified and filtered those cases out (e.g.,
-the~\textsf{pullwifi} project was responsible for almost 50\% of all Android issues in our dataset).
+issues from our analysis.
 
@@
 
-However, Google Code still contain largely used open-source Android apps (e.g. K9Mail \footnote{K9Mail moved to Github but as a way of not loosing the project history 
+However, Google Code still contains largely used open-source Android apps (e.g. K9Mail \footnote{K9Mail moved to Github but as a way of not loosing the project history 
 it advices their users to report bugs on Google Code issue tracker.})  and the history of such apps are also valuable. 
@@
 
-To identify Android projects in Google Code, we performed a similar heuristic, we performed a case insensitive search 
+To identify Android projects in Google Code, we performed a similar heuristic: we performed a case insensitive search 
 (on Google Code search interface) for the term \textsf{android}.  On January 2014, when we queried Google Code, such heuristic could
  find 788  projects, from which 724 defined at least 1 issue and, 183 projects 
-had at least one issue including a stack trace. Then we performed further cleanup, inspecting the site of every Android
-reporting at least one stack trace, and as a result we could identify 157. 
-Android projects.
-
-This set of 157 projects contained overall 127,456 issues from which 1,963 exception stack traces 
-were extracted. Based on the same assumption used for Github dataset selection, we did not retrict 
-the analysis for issues labled as Defect.
+had at least one issue including a stack trace. Then we performed further cleanup, inspecting the site of each project,
+ and as a result we could identify 157 Android projects.
+
+This set of 157 projects contained overall 127,456 issues, from which 1,963 exception stack traces 
+were extracted. Table ~\ref{tab:lables} illustrates the ocurrences of different lables 
+on the issues including exception stack traces. Doferently from GitHib in Google Code most of 
+the lables were labled as ``Defect" . On the study we considered all issues reporting stacks (regardless its lables)
+based on the same assumption described for the GitHub repository.
 

### 2014-09-03 android-stacks.tex
all over

@@
 
-\emph{\textbf{Classifying the Most Reported Root Exceptions.}} To get a broader view of the root causes of stack traces,
+\emph{\textbf{Identifying the Concerns Related to Root Exceptions.}} To get a broader view of the root exceptions of stack traces,
  we performed an empirical evaluation whose goal was to identify the concerns related to the most reported root exceptions. 
-
 Since it is infeasible to inspect the code responsible for throwning every exception reported in this study.
@@
 its description (i.e., Javadoc documentaion), and the Java specification. To exemplify, an instance of ArrayOutOfBoundException
- usually refers to a programming mistake. Moreover, the Java specification list all exceptions related to backwardcompatibility 
-issues ~\cite{javaback} (e.g., InstantiationError, VerifyError, IllegalAccessError).
-
-Some concerns are known as sources of faults in mobile development are: concurrency ~\cite{ama2012} backward compatibility  
+ usually refers to a programming mistake. Moreover, the Java specification list all exceptions related to backward compatibility 
+issues ~\cite{javaback} (e.g., InstantiationError, VerifyError, IllegalAccessError). Some concerns are known as sources 
+of faults in mobile development are: concurrency ~\cite{ama2012} backward compatibility  
 ~\cite{McDon13}, security  ~\cite{enck2011study,was2010} and resource management (IO, Memory, Batery) ~\cite{Zhang12}.
-
-%Moreover, the programming errors are also as a source of faults, as illustrated on the previous section.
-
 However, there are exceptions that cannot be directly mapped to one of such concerns. Either because they are
@@
 To perform such mapping, we selected a subset of all reported root exceptions. This subset was composed by 100 root exceptions 
-which were reported on 95\% of all stack traces analyzed in this study.
-
-Hence, based on the inspection of the source code and Javadoc related to each exception,  and Java specification, we identified the concern
+which were reported on 95\% of all stack traces analyzed in this study. Hence, based on the inspection of the source code and
+ Javadoc related to each exception,  and Java specification, we identified the concern
 releated to each root exception. Than we could calculate the prevalence of a concern on the reported stack traces using the following metric.
@@
  and Stacks[Lc] the list of issues that mentions the concern. The prevalence of a concern could be given by the formula: Stacks[Lc]/Stacks. 
-
-Table~\ref{tab:tophundrend} illustrates the result of such mapping.
+Table~\ref{tab:tophundrend} illustrates the result of this empirical evaluation.
 

### 2014-09-03 android-stacks.tex
all over

@@
  ranked by the number of distinct project repositories on which they were reported. 
-The table shows: the number of times each exception was reported as 
-the root cause of a stack trace, the number of distinct projects reporting such causes on stack traces, 
-and the percentages of such in relation to all reported exceptions and analyzed projects respectively.
-It also presents the whether the method responsible for signaling the exceptions were
-part of Android, the application itself, libcore, or a third-party library as presented in Table  \ref{tab:signalers}
+The table shows how many times method responsible for signaling the exceptions were
+part of Android Platform, the application itself, libcore, or a third-party library as presented in Table  \ref{tab:signalers}
 
@@
 We can observe that most of this set is composed by exceptions defined by java.lang,
-which are mostly not explicitly thrown by the developer but include
-the exceptions implicitly thrown by the JVM due to programming errors 
+which are mostly implicitly thrown by the JVM due to programming mistakes 
 (e.g., out-of-bounds array index, division-by-zero, access to a null reference)
  or resource limitation (e.g., OutOfMemoryError).
-For such exceptions, which represent programming bugs or resource limitations, 
-there is usually no proper handling besides presenting an error message to
+For such exceptions there is usually no proper handling besides presenting an error message to
  the user and restarting the application. Only high fault tolerant systems need to 
-provide solutions to handle them. Hence, the best strategy to deal with them
-is try to prevent them.
+provide solutions to handle them.
 
 \emph{\textbf{The high number of NullPointerExceptions.}}
-From this set the java.lang.NullPointerException was the most reported root cause; 27,71\% 
-of the analyzed exception stack traces reported a java.lang.NullPointerException the root cause. 
+From this set the java.lang.NullPointerException was the most reported root cause (27,71\% )
 Considering the frequencly of NullPointerException across projects, we could observe that

### 2014-09-04 android-stacks.tex
all over

@@
 
-  \item  It performs a large scale post morten analysis of exception stack traces reported on Android projects hosted in GitHub and Google Code.
+  \item  It performs a large scale post mortem analysis of exception stack traces reported on Android projects hosted in GitHub and Google Code.
   \item  It introduces ExceptionMiner, a tool developed to support exception stack trace mining and analysis.
@@
 
-\section{Study Design}
-\label{sec:study}
-
 %This study focus on the analysis of crash information (related to uncaught exceptions) of a family of Android applications
@@
 
-The goal of this mining study was to extract exception stack traces embedded on issues of Android projects 
-(hosted in GitHub and Google Code), looking for its 
-common characteristics and investigating whether
- they could reveal common fragilities on the exception code of both the 
-applications and the underlaying  framework.
-
+%To answer this question we selected 
+%The goal of this exploratory mining study was to extract exception stack traces embedded on issues of Android projects 
+%(hosted in GitHub and Google Code), looking for its \emph{common characteristics} and investigating whether
+% they could reveal \emph{common fragilities} on the exception code of both the applications and the underlaying  framework. 
+%Our study focused on open-source apps, since the information needed to perform our study
+%cannot be retrieved from commercial apps, whose issue report systems and 
+%source codes are not publicly available. The Android open-source apps were also 
+%target of other research studies ~\cite{Linar13,ahimed}.
+
+%extract exception stack traces embedded on issues of Android projects 
+%(hosted in GitHub and Google Code), looking for its \emph{common characteristics} and investigating whether
+% they could reveal \emph{common fragilities} on the exception code of both the applications and the underlaying  framework. 
+%Our study focused on open-source apps, since the information needed to perform our study
+%cannot be retrieved from commercial apps, whose issue report systems and 
+%source codes are not publicly available. The Android open-source apps were also 
+%target of other research studies ~\cite{Linar13,ahimed}.
+
+\section{Study Design}
+\label{sec:study}
+
+%This section describes the exploratory study whose goal was to investigate:
+%\begin{description}
+ % \item[RQ] \noindent\emph{Which fragilities on the exception-related code can be revealed from a post morten exception stack trace analysis of Android projects?} 
+ %    \item
+%\end{description}
+%One major decision that had to be made for our investigation was the selection of the target 
+%projects. We have selected a set of XXX open-source Android projects hosted in GitHub and 
+%Google Code. GitHub and Google Code are among the most used hosting sites for open-source
+ %projects,
+
+The goal of this exploratory mining study was to investigate whether the exception stack traces 
+embedded on issues of Android projects (hosted in GitHub and Google Code), can reveal 
+\emph{common fragilities} on the exception code of both the applications and the underlaying  framework. 
 Our study focused on open-source apps, since the information needed to perform our study
@@
 
-% Falar da assumption que stacj trá relacionada a crash?
-%This study is built on the following assumption: the stack traces reported on issues contain 
-%relevant information related to application crashes.
-Several studies and techniques have also based on this assumption %(e.g.,~\cite{sinha2009fault}~\cite{dhaliwal2011classifying}~\cite{kim2013predicting}).
-
-%\subsection{Research Questions}
-In the context of our study we formulated the following research questions:
-
-\begin{description}
-
-  \item[RQ1] \noindent\emph{What are the common characteristics of exception stack traces reported on issues?}
-
-%Which fragilities on the exception-related code emerge from the exception stack trace analysis? 
-%Can such characteristics reveal common fragilities on the exception code of applications or the underlying framework
-
-     \item
-
-%Which patterns of exception (mis)use emerge?
-  \item[RQ2] \noindent\emph{Which fragilities on the exception-related code emerge from them?}
-
-\end{description}
-
-%Based mainly on the information available on exception stack traces embedded 
-%on issues, the goal of this study was to gain a better understanding 
-%of exception-related system crashes.
-
-%Characteristics of exception stack traces and detected violations can reveal fragilities
-% on the exception-related code, which threats the application dependability.
-
-Some of the common characteristics to be investigated are the exceptions reported as the root cause on
-the exception stack trace, the origin of such exceptions (e.g., the application, a library, the Android platform). 
-Based on a set of best practices compiled in this study, we investigated 
-whether characteristics of exception stack traces can pinpoint best practices violations. 
-Such  violations reveal fragilities
-on the exception-related code, which threats the application dependability.
-
-For RQ1 and RQ2, we explore the domain quantitatively and highlight interesting cases by 
-exploring cases qualitatively. To support this investigation, source code and bytecode 
-analysis were used in combination with manual inspection to leverage the  understanding 
-of stack traces and support further discussions and insights.
+This exploration is guided by the set of best practices compiled in this study; we investigate whether characteristics 
+of exception stack traces can pinpoint best practices violations. As mentioned before, 
+call ``fragilities" characteristics on the exception-related code that favor the introduction
+of failures such as uncaught exceptions and unintended handling.
+
+To support this investigation, we developed a tool called ExceptionMiner (see Section ~\ref{sec:exceptionminer})
+which extracts the exception stack traces embeeded on issues, 
+and combines stack trace information with source code and bytecode 
+analysis. Such combination was needed to enable the identification of 
+fragilities on the exception-related code.
+Moreover, manual inspections were also used to leverage
+ the understanding of stack traces and support further discussions and insights.
+In this study we explore the domain quantitatively and highlight interesting cases by 
+exploring cases qualitatively.
 
@@
 
-To answer the research questions the following information is needed: (i) the issues reported on Android project hosted on 
+In this exploratory study the following information is needed: (i) the issues reported on Android project hosted on 
 GitHub and Google Code; (ii) the stack traces embedded on these issues; (iii) the type of the exceptions
  reported on the stack traces (e.g., error, runtime, checked); (iv) the origin of such exceptions 
-(e.g., the application, a library, the Android platform). Next sections describe: how such information 
+(e.g., the application, a library, the Android platform). Next, we describe: how such information 
 was retrieved from GitHub and Google Code; how the exceptions stack traces were extracted and distilled,
@@
 However, Google Code still contains largely used open-source Android apps (e.g. K9Mail \footnote{K9Mail moved to Github but as a way of not loosing the project history 
-it advices their users to report bugs on Google Code issue tracker.})  and the history of such apps are also valuable. 
+it advices their users to report bugs on Google Code issue tracker.})  and the issues reported of such apps are also valuable. 
 To overcome this limitation we needed to implement a Web Crawler (incorporated in Exception Miner tool described next) that navigated 
@@
 To identify Android projects in Google Code, we performed a similar heuristic: we performed a case insensitive search 
-(on Google Code search interface) for the term \textsf{android}.  On January 2014, when we queried Google Code, such heuristic could
- find 788  projects, from which 724 defined at least 1 issue and, 183 projects 
-had at least one issue including a stack trace. Then we performed further cleanup, inspecting the site of each project,
- and as a result we could identify 157 Android projects.
-
-This set of 157 projects contained overall 127,456 issues, from which 1,963 exception stack traces 
-were extracted. Table ~\ref{tab:lables} illustrates the ocurrences of different lables 
-on the issues including exception stack traces. Doferently from GitHib in Google Code most of 
-the lables were labled as ``Defect" . On the study we considered all issues reporting stacks (regardless its lables)
-based on the same assumption described for the GitHub repository.
+(on Google Code search interface) for the term ``android". On January 2014, when we queried Google Code, such heuristic could
+ returned a list of 788  projects. This list comprised the seeds sent to our Crawler.
+
+From this set, 724 defined at least 1 issue. Running the ExceptionMiner tool (described next)
+ in this set we observed that 183 apps had at least one issue containing an exception stack trace.
+ Then we performed further cleanup (similar to the one described previously) inspecting the site 
+of each project. As a result we could identify 157 Android projects.  This set contained overall 127,456 issues,
+ from which 1,963 exception stack traces were extracted. Table ~\ref{tab:lables} illustrates the ocurrences of different lables 
+on the issues including exception stack traces. Differently from GitHib, in Google Code most of 
+the issues were labled as ``Defect". However, based on the same assumption described for the GitHub repository
+ we considered all issues reporting stacks (regardless its lables).
 
@@
 
-\subsection{The ExceptionMiner Tool}
+\subsection{Manual Inspections}
+As mentioned before, this mixed-methods study was also supported by manual inspections. 
+Manual inspections were used, in the context of this study to support: (i) the identification of Android projects (described previously);
+(ii) and to identify the type of some exceptions reported on issues, that could not be automatically identified by the ExceptionMiner tool.
+Some exceptions could not be automaticaly identified because they were 
+not defined on the last version of the application (the one analyzed by ExceptionMiner),
+or were defined on libraries not found among application resources.
+In such cases we manually inspect the sourcecode or javadoc
+ documentation available online of each of them. 
+When the exception was not found on the previous versions, the exception where classified 
+in our study as "Undefined".  Only 31 exceptions 
+remained undefined - such exceptions were reported on 61 exception stack traces as 
+presented in Table ~\ref{tab:typeroottab}.
+
+%{MOSTRAR UMA FIGURA COM TODOS OS PASSOS E RESSALTAR OS PASSOS MANUAIS}
+%COMPLETE HERE.....
+%\subsubsection{Database of Exception Stack Trace Information}
+%\subsubsection{Distiling Exception Stack Trace Information}
+%\subsection{Process}
+%ESPALHAR ESTA SECAO NAS OUTRAS.
+%Figure~\ref{fig:overviewfig} presents an overview of the mixed-methods approach
+%conducted to answer these questions. The mixed-methods approach is composed by the following steps:
+
+\section{The ExceptionMiner Tool}
 \label{sec:exceptionminer}
 
-To support this study we developed ExceptionMiner, a modular mining tool able 
-to connect to repositories, extract issues, mine stack traces from
-them and distill exception stack trace information. The main components of
-ExceptionMiner are as follows:
-
-\noindent\emph{Repository Connectors.}  This component is responsible to connect to
- the host repositories in order to collect project issues. The GitHub Connector connected
- to GiHub through GHTorrent. The Google Code connector comprised of a Web Crawler,
- that traversed Google Code Web interface in order to extract the project issues.
-
-\noindent\emph{Exception Stack Trace Distiller.} 
-Distills the information that composes a stack trace and storing the results in a
-relational database. Some of the information extracted from the stack traces were:
- the exception being thrown, its signaler, the exception root cause, its corresponding signaler,
-and packages related to each one of these fields.
-The tool is based on a combination between a regular expression based parser 
-with heuristics able to identify and filter exception names and stack traces inline with text. In
-contrast to existing issue parsing solutions such as Infozilla, the parser
-created in this work can discover stack traces embedded in logs files \footnote{In several 
-exception stack traces, the exception frames were preceeded by logging information e.g., "03-01 15:55:01.609 (7924): at android.app.ActivityThread.access\$600(ActivityThread.java:127)" which could not be detected by existing tools.)}
-and identify the root causes and signalers associated to each exception stack trace.
+The ExceptionMiner is modular mining tool which is able to connect to issue repositories, 
+extract issues, mine stack traces from them, distill exception stack trace information,
+and enable the execution of different analyses by combining exception stack 
+trace information with bytecode and sourcecode analysis. The main components of ExceptionMiner are as follows:
+
+\emph{\textbf{Repository Connectors.}} This component enables the connection 
+with issue repositories. In this study two main connectors were created: one which connects to 
+GHTorrent database, and Google Code connector which is comprised of a Web Crawler,
+ that can traverse Google Code Web interface and extract the project issues. 
+Projects matadata and the issues associated to them are stored on a relational 
+database by this components.
+
+\emph{\textbf{Exception Stack Trace Distiller.}}
+This component distills the information that composes a stack trace.
+ Some of the attributes extracted from the stack trace are:
+ the root exception and its signaler, the exception wrappers and their corresponding signalers. 
+This component also distills fine grained information of each attribute such as: the classes and packages associated to them.
+This component is based on a combination between a regular expression based parser 
+with heuristics able to identify and filter exception names and stack traces inline with text. 
+In contrast to existing issue parsing solutions such as Infozilla, the parser
+created in this work: (i) can discover stack traces inlined with logs files \footnote{In several 
+exception stack traces, the exception frames were preceeded by logging information e.g., 
+"03-01 15:55:01.609 (7924): at android.app.ActivityThread.access\$600(ActivityThread.java:127) 
+"which could not be detected by existing tools.)}.
 
@@
 
-\noindent\emph{Exception Classifier} To support a deeper investigation of the stack traces 
-every exception defined on a stack trace need to be classified according to its type
-(e.g. Error, Exception or RuntimeException). The exception classifier is based on 
-bytecode analysis (based on Design Wizard~\cite{Brunet09}) that walks up the 
-type hierarchy of a Java exception until it reaches a base exception type.
-This module analyzed the exception hierarchy of all exceptions thrown on:
-  Android Platform (API level 19), and JDK (version 6).
-
-\noindent\emph{Exception Signaler Classifier.} 
+As mentioned before, the tool enables a set of exception stack traces analyses by combining exception stack 
+trace information with bytecode and sourcecode analysis. Each analysis is implemented as a different module 
+on Exception miner. The main analyses provided by the tool are presented next:
+
+\emph{\textbf{Exception Type Analysis}} To support a deeper investigation of the 
+stack traces every exception defined on a stack trace need to be classified according to its type
+(e.g. Error, Exception or RuntimeException). The module responsible for this analysis 
+uses Design Wizard~\cite{Brunet09} framework to inspect the bytecode of all exceptions reported on issues.
+The nidyke walks up the type hierarchy of a Java exception until it reaches a base exception type.
+This module analyzed the exception hierarchy of all exceptions defined on:
+Android Platform (API level 19) - which includes all basic Java exceptions that 
+can be thrown during the app execution. The exceptions thrown inside the apps
+were classified using manual inspections as well as the exceptions thrown only
+on previous versions of the Android platform.
+
+\emph{\textbf{Exception Signaler Analysis.}}
 This module is responsible for classifying each signaler according 
-to its origin (i.e. Android Plaform, Android Libcore, Aplication, Library). 
-We downloaded the source code of Android Platform (API level 19)
-and this module extracts the recursive list of 
-of Java packages that composes it. To discover the packages that compose 
-the previous released of the API we performed a manual inspection on the Javadoc
-documentation available online for each API level ~\cite{apidocs}.
-To discover the packages composing the each application, this module 
+to its origin (i.e. Android Application Framework, Android Libcore, Application, Library). 
+Table~\ref{tab:signalers} presents the heuristic adopted in this classification.
+To enable such classification, we needed to feed this module the information
+comprising all Java packages that compose: the Android Plaform;
+ the Android Libcore; and each analyzed Application. 
+To discover the packages composing the first two origins
+we could reffer to the Android Specification.
+To discover the packages composing each application, this module 
 extracted the manifest files of each Android app
  - which defines the main packages that compose the applications.
-Then based on pattern matching between the signaler name and the packages 
- this module identifies the origin of the exception signalers.
-The exceptions were considered from libraries if their packages were not defined 
+ When this file was not available we recursevely analyzed the 
+structure of source-code directories composing the app
+- filtering out the cases on which the application 
+also included the source-code of reused libraries.
+Then based on this information and using pattern matching 
+between the signaler name and the packages this module identifies 
+the origin of the exception signalers. The exceptions were considered
+ from libraries if their packages were not defined 
 on Android Platform, core libraries nor on applications.
 Table ~\ref{signalers} summarizes the signaler classification.
+
+%\noindent\emph{Exception Wrapping Analysis.} 
+%\noindent\emph{Concern Mapping.}
 
@@
     \hline
-    \bfseries{android} & If the exception is thrown by a method defined in Android Platform, or in a JDK library used by i.\\
-    \bfseries{app}     & If the exception is thrown by an application method or in a JDK library used by it.\\
-    \bfseries{libcore} & If the exception is thrown by one of the core libraries reused by Android (i.e., org.apache.harmony, org.w3c.dom, sun.misc, org.apache.http, org.json, org.xml). \\
-    \bfseries{lib}     & If the exception is thrown by a method that was not defined by any of the elements above.\\
+    \bfseries{android} & If the exception is thrown on a method defined in Android Platform.\\
+    \bfseries{app}     & If the exception is thrown on an application method.\\
+    \bfseries{libcore} & If the exception is thrown on one of the core libraries reused by Android (i.e., org.apache.harmony, org.w3c.dom, sun.misc, org.apache.http, org.json, org.xml). \\
+    \bfseries{lib}     & If the exception is thrown on a method that was not defined by any of the elements above.\\
     \hline
@@
 
-The information extracted by each module was structured in a Database,
-which contained the distilled information of stack traces, exception types,
-and signalers origin (e.g., root cause, root cause type, root cause signaler and its origin,
-exception wrappers,  types, and its signalers).
+%The information extracted by each module was structured in a Database,
+%which contained the distilled information of stack traces, exception types,
+%and signalers origin (e.g., root cause, root cause type, root cause signaler and its origin,
+%exception wrappers,  types, and its signalers).
 
@@
 % related with defect issues on this analysis.
-
-\subsection{Manual Inspections}
-As mentioned before, this study was also based in a set of manual inspection tasks. Besides the 
- (i) identifying of Android projects, and (ii) the finding packages that compose the 
-previous Android APIs leveles, manual inspections were also used discover the type of associated to some 
-reported on stack traces. Such exceptions could not be automaticaly identified 
-because they were not defined on the last version of the application,
-or were defined on libraries not fount among projects resources.
-In such cases we manually inspect the sourcecode or javadoc
- documentation available online of each of them. 
-When the exception was not found on the previous versions, the exception where classified 
-in our study as "Undefined".  Only 31 exceptions 
-remained undefined - such exceptions were reported on 61 exception stack traces as 
-presented in Table ~\ref{tab:typeroottab}.
-
-%{MOSTRAR UMA FIGURA COM TODOS OS PASSOS E RESSALTAR OS PASSOS MANUAIS}
-%COMPLETE HERE.....
-%\subsubsection{Database of Exception Stack Trace Information}
-%\subsubsection{Distiling Exception Stack Trace Information}
-%\subsection{Process}
-%ESPALHAR ESTA SECAO NAS OUTRAS.
-%Figure~\ref{fig:overviewfig} presents an overview of the mixed-methods approach
-%conducted to answer these questions. The mixed-methods approach is composed by the following steps:
 

### 2014-09-04 android-stacks.tex
results

@@
 trace information with bytecode and sourcecode analysis. Each analysis is implemented as a different module 
-on Exception miner. The main analyses provided by the tool are presented next:
+on Exception miner. Next, we present core analyses which were implemente on ExceptionMiner. Such analyses 
+provide aditional information to be used on further analyses described in Section ~\ref{sec:result}:
 
@@
  - which defines the main packages that compose the applications.
- When this file was not available we recursevely analyzed the 
+ Whenever this file was not available, we recursevely analyzed the 
 structure of source-code directories composing the app
@@
 
-%The information extracted by each module was structured in a Database,
-%which contained the distilled information of stack traces, exception types,
-%and signalers origin (e.g., root cause, root cause type, root cause signaler and its origin,
-%exception wrappers,  types, and its signalers).
-
-%\subsubsection{The ExceptionMiner Tool}
-%Using the ExceptionMiner tool, developed in this work, we could extract the exception stack traces defined 
-%on issues and issue comments. Executing ExceptionMiner tool each issue and issue comment defined on the 
-%16,836 projects considered in this study, we could observe that 3,758 projects (22,32\%) contained at 
-%least 1 issue on which a stack trace was found. Overall the number of issues on which a stack trace was
-% found was 21,013 from which 28,800 stack traces were extracted and analyzed - some issues defined 
-% more than one stack trace.
-
-%We then analyzed the labels associated to such issues, and we discovered that only 5,196 were
-% labled as defect ~\footnote{the dedect lables considered in this work were every lable containing 
-%as substring: defect, bug, crash, failure, fail, exception} (which represent only 24,7\% of the
-% issues on which a stack trace was found), most of the issues on which stack traces were found 
-%contained no lable. Such observation supports our assumption to consider not only the issues 
-% related with defect issues on this analysis.
-
-%\begin{figure*}
-%\centering
-%\includegraphics[width=\hsize]{overview.pdf}
-%\caption{Overview of the study based on mixed approaches.}
-%\label{fig:overviewfig}
-%\end{figure*}
-
 \subsection{Replication Package}
@@
 
-Next section presents the results for each of the study phases, providing both a
-quantitative and qualitative analysis of the outcomes.
-
 \section{The Study Results}
@@
 
-%To answer this question, we mined from every extracted stack trace the top level
-%exception being signaled, its signaler (the first method signature frame after
-%the exception), the root exception (the exception which initiated the stack
-%trace, defined on the bottom of the stack), and the root signaler. The
-%intermediate elements that compose a stack trace as shown in
-%Figure~\ref{fig:wrapping} (i.e., the intermediate causes and its execution
-%stacks) were not be considered in the analysis.
-%~\todo{Move this to the previous section? IMO, It is better to keep all facts
-%about how we did the research in one place
-%RSC: I WAS ABOUT TO MOOVE BUT THERE WAS ALREADY A SIMILAR TEXT ON 
-%THE EXCEPTION DISTEILLER}
-
-This section presents the results aimed at answering the research 
-questions presented before.
+This section presents the results of the study, providing both a
+quantitative and qualitative analysis of the outcomes. The structure of this section
+ is centered on questions that guied the search for 
+source-code fragilities based on exception stack trace information.
 
@@
 
-\emph{\textbf{Identifying the Concerns Related to Root Exceptions.}} To get a broader view of the root exceptions of stack traces,
+\emph{\textbf{The prevalence of programming mistakes.}} To get a broader view of the prevalence of
+programming mistakes of reported exception stack traces we performed the following analysis....
+
  we performed an empirical evaluation whose goal was to identify the concerns related to the most reported root exceptions. 

### 2014-09-04 android-stacks.tex
results - question 1

@@
 
-Besides, the runtime exceptions are also implicitly thrown by the JVM when a program violates 
+Besides, the runtime exceptions are also implicitly thrown by the runtime environment when a program violates 
 the semantic constraints of the Java programming language (e.g., out-of-bounds array index, divide-by-zero 
@@
  its execution in some situations such as the out-of-bounds array index e.g., C++. }  
-The programs are not expected to handle such JVM runtime exceptions ~\cite{gosling2000java} s.
+The programs are not expected to handle such runtime exceptions signaled by the runtime environment ~\cite{gosling2000java}.
 
@@
 
-In Java, once an exception is thrown, the JVM looks for the nearest enclosing exception handler
+In Java, once an exception is thrown, the runtime environment looks for the nearest enclosing exception handler
 (in Java it is represented by the try-catch block), and unwinds the execution stack if necessary.
@@
 
-\emph{II-Error represents an unrecoverable condition detected by the JVM which
+\emph{II-Error represents an unrecoverable condition detected by the runtime enviroment which
 should not he handled} (\cite{gosling2000java}). It results from failures detected
@@
 developers should either try to reuse the exception types already defined in the
-JVM or they should create a specific exception. Throwing general types such as a
+Java API or they should create a specific exception. Throwing general types such as a
 pure java.lang.Exception or a java.lang.RuntimeException is considered a bad practice.
@@
 also wise to document the explicitly thrown runtime exceptions\footnote{This
-excludes the implicitly signaled JVM runtime exceptions related to programming
+excludes the implicitly signaled runtime exceptions related to programming
 mistakes} as carefully as checked exceptions. Doing so, the clients of a method
@@
 % on each project and issue page, extracting the content of each issue to store on relational database.
-
-\subsection{Manual Inspections}
-As mentioned before, this mixed-methods study was also supported by manual inspections. 
-Manual inspections were used, in the context of this study to support: (i) the identification of Android projects (described previously);
-(ii) and to identify the type of some exceptions reported on issues, that could not be automatically identified by the ExceptionMiner tool.
-Some exceptions could not be automaticaly identified because they were 
-not defined on the last version of the application (the one analyzed by ExceptionMiner),
-or were defined on libraries not found among application resources.
-In such cases we manually inspect the sourcecode or javadoc
- documentation available online of each of them. 
-When the exception was not found on the previous versions, the exception where classified 
-in our study as "Undefined".  Only 31 exceptions 
-remained undefined - such exceptions were reported on 61 exception stack traces as 
-presented in Table ~\ref{tab:typeroottab}.
 
@@
 
+\subsection{Manual Inspections}
+In the context of the ExceptionMiner tool, manual inspections were used:
+to help on the identification of packages composing the Android platform, 
+libs and apps analyzed in this study (as described previously); and to 
+identify the type of some exceptions reported on issues, 
+that could not be automatically identified by the ExceptionMiner tool.
+Some exceptions could not be automaticaly identified because they were 
+not defined on the last version of the application (the one analyzed by ExceptionMiner),
+or were defined on libraries not found among application resources.
+In such cases we manually inspect the sourcecode or javadoc
+ documentation available online of each of them. 
+When the exception was not found on the previous versions, the exception where classified 
+in our study as "Undefined".  Only 31 exceptions 
+remained undefined - such exceptions were reported on 61 exception stack traces as 
+presented in Table ~\ref{tab:typeroottab}.
+
 \subsection{Replication Package}
@@
 quantitative and qualitative analysis of the outcomes. The structure of this section
- is centered on questions that guied the search for 
-source-code fragilities based on exception stack trace information.
+ is centered on questions that guied the search for fragilities (on the exception-related code)
+based on exception stack trace information.
 
@@
 the the exceptions commonly reported as the root causes of stack traces.
-Table \ref{tab:topten} illustrates the top 10 root causes found in the study,
- ranked by the number of distinct project repositories on which they were reported. 
-The table shows how many times method responsible for signaling the exceptions were
-part of Android Platform, the application itself, libcore, or a third-party library as presented in Table  \ref{tab:signalers}
-
-% \footnote{Nine exception stack traces contained only elements of JVM and were discarded from the analysis}.
-
-%For each exception we could discover the number of times it was reported as the root caused on exception stack traces,
-%the number of distinct projects on which such stacks were defined (and consequently its
-%popularity) and the of times it was signaled by different signaler types. We
-%could calculate the popularity of each root cause, considering the number of
-%distinct projects they were found and the whole set of projects analyzed with at
-%least one stack trace (482). Table~\ref{tab:topten} 
-%and Figure~\ref{fig:androidsignaler} presents the mined data.
-
-%SELECTS:
-%SELECT COUNT(distinct(repo)) FROM STACKTRACESUMMARY WHERE MAIN_CAUSE LIKE '%java.lang.NullPointerException%'
-%and repo !='android';
-
-%SELECT COUNT(*) FROM STACKTRACESUMMARY WHERE MAIN_CAUSE LIKE '%java.lang.IllegalStateException%'
-%and repo !='android' AND MAINSIGNALER_FROM LIKE '%ANDROID%';
-
-%java.lang.NullPointerException            & 254 /  78  & 52,70\%  & 1225 / 439  & 30,08\% & 473 /  26  & 18 /  & 595 / 241 & 137 / 169 \\  %LIBORAPP=408
-%java.lang.IllegalStateException          & 99 / 21  & 20,54\%  & 234 / 44  & 5,75\%  & 165 / 15 & 12  / 0 & 36  / 5 & 20 / 24  \\
-%java.lang.IllegalArgumentException          & 94 / 48  & 19,50\% & 255 / 98  & 6,26\% & 146 / 41 & 6 / 0 & 64 / 31  & 39  / 26  \\
-%java.lang.RuntimeException                & 91 / 31 & 18,88\% & 232 / 87  & 5,70\%  & 167 / 33 & 1 / 0 & 47 / 17 & 17 / 37   \\
-%java.lang.OutOfMemoryError                 & 56  / 22  & 11,62\% & 180 / 57 & 4,42\% & 121 / 18 & 15 / 0 & 17 / 18  & 23 / 13  \\
-%java.lang.NoClassDefFoundError           & 52  / 15 & 10,79\%  & 73 / 21   & 1,79\%  & 9 / 1   & 0 / 0  & 37 / 9  & 26 / 11   \\
-%java.lang.ClassCastException             & 49  / 15 & 10,17\%  & 94 / 36   & 2,31\%  & 43 / 3  & 0  / 0 & 40 / 15  & 11 / 18  \\
-%java.lang.IndexOutOfBoundsException       & 47  / 15 & 9,75\%   & 127 / 39  & 3,12\% & 47 / 2  & 0/0  & 71/22  & 8 / 14    \\
-%java.lang.NoSuchMethodError               & 40 / 14  & 8,30\%  & 57 / 23  & 1,40\%  & 9  /  0  & 0  / 0 & 39 /  17    & 9 / 6  \\
-%java.util.ConcurrentModificationException & 38 / 5 & 7,88\% & 54 / 11  & 1,33\%   & 5  / 0  & 0  / 0 & 43 / 3 & 6 / 7  \\
-
-%java.lang.IllegalStateException          & 99 / 21  & 20,54\%  & 234 / 44  & 5,75\%  & 165 / 15 & 12  / 0 & 36  / 5 & 20 / 24  \\
-%java.lang.IllegalArgumentException          & 94 / 48  & 19,50\% & 255 / 98  & 6,26\% & 146 / 41 & 6 / 0 & 64 / 31  & 39  / 26  \\
-%java.lang.RuntimeException                & 91 / 31 & 18,88\% & 232 / 87  & 5,70\%  & 167 / 33 & 1 / 0 & 47 / 17 & 17 / 37   \\
-%java.lang.OutOfMemoryError                 & 56  / 22  & 11,62\% & 180 / 57 & 4,42\% & 121 / 18 & 15 / 0 & 17 / 18  & 23 / 13  \\
-%java.lang.NoClassDefFoundError           & 52  / 15 & 10,79\%  & 73 / 21   & 1,79\%  & 9 / 1   & 0 / 0  & 37 / 9  & 26 / 11   \\
-%java.lang.ClassCastException             & 49  / 15 & 10,17\%  & 94 / 36   & 2,31\%  & 43 / 3  & 0  / 0 & 40 / 15  & 11 / 18  \\
-%java.lang.IndexOutOfBoundsException       & 47  / 15 & 9,75\%   & 127 / 39  & 3,12\% & 47 / 2  & 0/0  & 71/22  & 8 / 14    \\
-%java.lang.NoSuchMethodError               & 40 / 14  & 8,30\%  & 57 / 23  & 1,40\%  & 9  /  0  & 0  / 0 & 39 /  17    & 9 / 6  \\
-%java.util.ConcurrentModificationException & 38 / 5 & 7,88\% & 54 / 11  & 1,33\%   & 5  / 0  & 0  / 0 & 43 / 3 & 6 / 7  \\
-%
-%java.lang.NullPointerException	&	254	/	78	&	52,70\%	&	1225	/	439	&	30,08\%	&	473	/	52	&	18	/	2	&	595	/	241	&	137	/	143	\\	%LIBORAPP=408
-%java.lang.IllegalStateException	&	99	/	21	&	20,54\%	&	234	/	44	&	5,75\%	&	165	/	20	&	12	/	19	&	36	/	5	&	20	/	19		\\
-%java.lang.IllegalArgumentException	&	94	/	48	&	19,50\%	&	255	/	98	&	6,26\%	&	146	/	49	&	6	/	6	&	64	/	31	&	18	/	26		\\
-%java.lang.RuntimeException	&	91	/	31	&	18,88\%	&	232	/	87	&	5,70\%	&	167	/	36	&	1	/	1	&	47	/	17	&	17	/	34		\\
-%java.lang.OutOfMemoryError	&	56	/	22	&	11,62\%	&	180	/	57	&	4,42\%	&	121	/	20	&	15	/	1	&	17	/	18	&	23	/	11		\\
-%java.lang.NoClassDefFoundError	&	52	/	15	&	10,79\%	&	73	/	21	&	1,79\%	&	9	/	1	&	0	/	0	&	37	/	9	&	26	/	11		\\
-%java.lang.ClassCastException	&	49	/	15	&	10,17\%	&	94	/	36	&	2,31\%	&	43	/	12	&	0	/	0	&	40	/	15	&	11	/	9		\\
-%java.lang.IndexOutOfBoundsException	&	47	/	15	&	9,75\%	&	127	/	39	&	3,12\%	&	47	/	6	&	0		0	&	71	/	22	&	8	/	10		\\
-%java.lang.NoSuchMethodError	&	40	/	14	&	8,30\%	&	57	/	23	&	1,40\%	&	9	/	1	&	0	/	0	&	39	/	17	&	9	/	5		\\
-%java.util.ConcurrentModificationException	&	38	/	5	&	7,88\%	&	54	/	11	&	1,33\%	&	5	/	0	&	0	/	0	&	43	/	3	&	6	/	7
+Table \ref{tab:topten} presents a list the top 10 root exceptions found in the study - 
+ ranked by the number of distinct projects on which they were reported. 
+The table shows how many times the signaler of such exception was a method defined on
+the Android Platform, the Android libcore, the application itself or a third-party library -
+ following the classification as presented in Table  \ref{tab:signalers}.
 
@@
 
-We can observe that most of this set is composed by exceptions defined by java.lang,
-which are mostly implicitly thrown by the JVM due to programming mistakes 
-(e.g., out-of-bounds array index, division-by-zero, access to a null reference)
+We can observe that most of the exceptions on this list are implicitly thrown by the
+ runtime environment due to programming mistakes  (e.g., out-of-bounds array index, division-by-zero, access to a null reference)
  or resource limitation (e.g., OutOfMemoryError). 
-For such exceptions there is usually no proper handling besides presenting an error message to
- the user and restarting the application. Only high fault tolerant systems need to 
-provide solutions to handle them.
-
-\emph{\textbf{The high number of NullPointerExceptions.}}
-From this set the java.lang.NullPointerException was the most reported root cause (27,71\% )
-Considering the frequencly of NullPointerException across projects, we could observe that
- 51,96\% of all projects reported at least one exception stack trace on which the NullPoiterException
-was the root cause.
+From this set the java.lang.NullPointerException was the most reported root cause (27,71\% ). 
+If we consider the frequencly of NullPointerException 
+across projects, we can observe that 51,96\% of all projects reported at least one exception stack 
+trace on which the NullPoiterException was the root cause.
  The NullPointerExceptions are mainly signaled inside the applicaiton code and the Android platform,
@@
 testing~\cite{maji2012empirical,csallner2004jcrasher} showed that most of the automatically 
-detected bugs were due to NullPointerExceptions 
-and exceptions implicitly-signaled by Java environment due to programming mistakes or resource limitations
- (as the ones found in this study).In Section ~\ref{sec:disc}  we provide further 
-discussion on the NullPointerException problem, pointing 
-to ways of how to mitigate it.
-
-%\emph{\textbf{Direct instances of RuntimeExceptions being thrown.}}
-%From Table ~\ref{tab:topten} we can observe that direct instances
-%java.lang.RuntimeException were reported as the root cause of stacks in 19,09\% 
-%of the projects, and most of such exceptions  were  thrown by the Android platform (167 out of 232).
-%Throwing general exceptions, such as direct instances of RuntimeException, violates best
-%practice III. It is considered a bad practice because exception type does not carry enough
-%information to identify the cause of the exceptional behavior, and as a consequence developers need to relying on
-%the exception message which can be neither complete nor precise~\cite{gosling2000java}.
+detected bugs were due to NullPointerExceptions and exceptions implicitly-signaled by Java 
+environment due to programming mistakes or resource limitations
+ (as the ones found in this study).
 
 \emph{\textbf{The prevalence of programming mistakes.}} To get a broader view of the prevalence of
-programming mistakes of reported exception stack traces we performed the following analysis....
-
- we performed an empirical evaluation whose goal was to identify the concerns related to the most reported root exceptions. 
-Since it is infeasible to inspect the code responsible for throwning every exception reported in this study.
-This identification performed in this study was based on meaning that the name of the exception holds, 
-its description (i.e., Javadoc documentaion), and the Java specification. To exemplify, an instance of ArrayOutOfBoundException
- usually refers to a programming mistake. Moreover, the Java specification list all exceptions related to backward compatibility 
-issues ~\cite{javaback} (e.g., InstantiationError, VerifyError, IllegalAccessError). Some concerns are known as sources 
-of faults in mobile development are: concurrency ~\cite{ama2012} backward compatibility  
-~\cite{McDon13}, security  ~\cite{enck2011study,was2010} and resource management (IO, Memory, Batery) ~\cite{Zhang12}.
-However, there are exceptions that cannot be directly mapped to one of such concerns. Either because they are
-related to programming mistakes - as illustrated on the previously. Or because they are too general (i.e.,  java.lang.Exception,
-java.lang.RuntimeException, java.lang.Error). Or also because they are related to other concerns (e.g., specific to a given library).
-
-%The list of categories is illustrated in Table~\ref{tab:categories}.
-
-To perform such mapping, we selected a subset of all reported root exceptions. This subset was composed by 100 root exceptions 
-which were reported on 95\% of all stack traces analyzed in this study. Hence, based on the inspection of the source code and
- Javadoc related to each exception,  and Java specification, we identified the concern
-releated to each root exception. Than we could calculate the prevalence of a concern on the reported stack traces using the following metric.
-Given that: Stacks is the complete set of stacks found in this study, Lc is the list of root causes related to a given concern (as identified on the study),
- and Stacks[Lc] the list of issues that mentions the concern. The prevalence of a concern could be given by the formula: Stacks[Lc]/Stacks. 
-Table~\ref{tab:tophundrend} illustrates the result of this empirical evaluation.
-
-% When the root exceptions that were too general, or related to other concerns 
-% were also maped to these additional categories. Table~\ref{tab:tophundrend} illustrates the result of such mapping.%
-%\begin{table}
-%  \centering
-%  \begin{tabular}{|p{2cm}| p{5cm}|}
-%    \hline
-%    \bfseries{Category} & \bfseries{Description} \\
-%    \hline
-%      Programming logic &  Exceptions thrown by the JVM when the 
-% evaluation of an expression violates the normal semantics of language (e.g., 
-% out-of-bounds array index, division-by-zero, access to a null reference). \\ \hline
-%      Resources (IO, Memory, Batery)  & Exception releated to the manipulation or restriction of resources (e.g., file, network, memory) \\ \hline
-%      Security                               & Exception releated to security issues (e.g. password validation, criptography) \\ \hline
-%      Concurrency                            & Exceptions related to multi-threaded programming \\ \hline
-%      Backward compatibility                 &  A list of exceptions related to backward compatibility - a complete list can be found here [ref]  \\ \hline
-%      Specific Exceptions              & Exceptions created for specific applications, frameworks or libraries \\ \hline
-%      General (Exception, Error, Runtime)    & The top level exceptions in Java    \\ \hline
-%    \hline
-%  \end{tabular}
-%  \caption{Description of categories.}
-%  \label{tab:categories}
-% \end{table}
+programming mistakes of reported exception stack traces we: (i) identified the Java exceptions that can be
+implicitly thrown by the runtime environment due to a programing mistake or resource limitation; and (ii)
+ used ExceptionMiner to evaluate the prevalence of such exceptions on the reported stack traces.
+Table ~\ref{tab:causes} illustrates the main causes of exceptions reported on issues.
 
@@
     \hline
-    \bfseries{Category} &  \bfseries{\% Occurrences on stacks} \\
+    \bfseries{Cause} &  \bfseries{\% Occurrences on stacks} \\
     \hline
       Programming logic (java.lang and util) &  52,0\%\\ 
-      Resources (IO, Memory, Batery)       &   23,9\% \\ 
-      Security                               &  4,1\%\\  
-      Concurrency                            &  2,9\% \\ 
-      Backward compatibility                 & 5,5\% \\ 
-      Specific Exceptions               &  4,9\%\\ 
-      General (Error, Exception, Runtime)    &  6,7\%\\
+      Resources Restriction   &   23,9\% \\ 
+      General Exceptions (Error, Exception, Runtime)    &  6,7\%\\
+      Specific  &  xxx\%\\ 
  \end{tabular}
   \caption{Identifying the concerns related to root exceptions}
-  \label{tab:tophundrend}
+  \label{tab:causes}
 \end{table}
 
-To ensure the quality of the process, three independent coders classified a randomly selected
-sample of 25 exception types (from the total 100) using the same list of concerns;
-the interater agreement was high (96\%). We could also observe that
-We can therefore assume that the categories are descriptive 
-enough for covering the range of exception types analyzed.
-
-These study showed that the programming mistakes associated to the most reported root causes. 
-However, they may be interactly related to one of the other concerns. For  instance, a concurrency problem may 
-lead to a NullPointerException.
+The analysis on the complete set os stack traces revealed that, in fact, the programming mistakes were
+the most reported root causes.  Although there is usually no proper handling for such exceptions, 
+besides presenting an error message to the user and restarting the application ~\footnote{Only 
+high fault tolerant systems need to provide solutions to handle them.}.
+
+This finding calls the attention to the fact that: high number of programming errors reported 
+on issues may me an indication that developers are underestimating the impact that such 
+mistakes have on application robustness.
 
@@
 
-%TODO: Discuss about the interater agreement here...
-%\subsection{Sizes of Stack traces}
-%Concerning the size of the stacks (i.e. the number of frames composing them), we
-%could observe that  sometimes it exceeded 100 frames (158 stack frames).
-%Manually inspecting the corresponding execution trace was caused by one of the
-%two reasons: recursive calls or exceptions signaled by a method defined deep
-%down in a reused framework; exceptions successively wrapped. Irrespective of the
-%reason, in exception flows involving many methods, it is practically impossible
-%to handle an exception in a way other than exiting the application, as at the
-%place the exception occurs there is hardly enough contextual information to
-%perform recovery actions.
-
-%Considering only the packages of each exception we could observe that 81,28\%
-%of all analyzed set of stack traces reported exceptions defined by java.lang
-%(see Table XXX).
-
-%\begin{table}
-% \centering
-%\begin{tabular}{lrrrr}
-%    \hline
-% \bfseries{rank} & \bfseries{package} & \bfseries{occurr} &  \bfseries{\%occur} & % \bfseries{projects} \\
-%     \hline
-%1 & java.lang         & 19471 & 81,28\% & 3080 \\
-%2 & java.io           & 1222  & 5,10\%  & 509 \\
-%3 & java.net          & 740   & 3,09\%  & 311 \\
-%4 & java.util         & 689   & 2,88\%  & 287 \\
-%5 & java.lang.reflect & 115   & 0,48\%  & 78 \\
-%\hline
-%  \end{tabular}
-%\caption{Top 5 most popular packages associated to the root exceptions of stack traces}
-%\label{tab:toptenpopular2}
-%\end{table}
-
-%\noindent \fbox{
-%\begin{minipage}{0.96\columnwidth} 
-%\emph{RQ2: Most exceptions are due to programming mistakes
-%or generally irrecoverable errors.}
-%\end{minipage}}
-
 \subsection{What can the Types of Root Exceptions tell us?}
@@
  Table~\ref{tab:typeroottab} presents the types of root causes of all stack traces reported on the analyzed issues. As we can see 
-from Table ~\ref{tab:typeroottab},  most of the reported exceptions signaled (explicilty or implicitly by the JVM) on methods defined
+from Table ~\ref{tab:typeroottab},  most of the reported exceptions signaled (explicilty or implicitly by the runtime enviroment) on methods defined
  on the Android Plaform on on the application itself.
@@
 by the Android apps. In this task we filtered out all the exceptions implicitly 
-signaled by  JVM (due to programming mistakes) 
+signaled by  runtime environment (due to programming mistakes) 
 since these exceptions should not be documented on the 
@@
 since they usually represent unrecoverable conditions
-detected by the JVM (e.g.,  OutOfMemoryError). On the other hand, checked exceptions should
+detected by the runtime environment (e.g.,  OutOfMemoryError). On the other hand, checked exceptions should
 represent recoverable conditions. Next we discuss and exemplify the cross-type wrappings
@@
 checked exception wrapped an OutOfMemoryError, which represent resource deficiency detected 
-by the JVM which make impossible the program to recover.
+by the runtime environment which make impossible the program to recover.
 
@@
 stack traces reporting coding errors (out-of-bounds array index, division-by-zero, access to a null).
-Such exceptions are implicitly thrown by JVM, and among them the NullPointerException 
+Such exceptions are implicitly thrown by runtime environment, and among them the NullPointerException 
 were the reported root cause in most of the analyzed exception stack traces.
@@
 approach that uses stack traces to guide a dataflow analysis for locating and
-repairing faults that are caused by the JVM implicitly signaled exceptions. Kim
+repairing faults that are caused by the implicitly signaled exceptions. Kim
 at al.~\cite{kim2013predicting} proposed an approach to predict the

### 2014-09-04 android-stacks.tex
results - question 2

@@
 
-\emph{IV- All exceptions explicitly thrown by a method should be documented.}
+\emph{IV- All exceptions explicitly thrown by a library/framework reusable code should be documented.}
 (\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,bloch2008effective}).
@@
 excludes the implicitly signaled runtime exceptions related to programming
-mistakes} as carefully as checked exceptions. Doing so, the clients of a method
-will be aware of all the exceptions the method can throw and can protect design the code
+mistakes} as carefully as checked exceptions, specialy when developing
+ library/framework reusable code. Doing so, the clients of such ibrary/framework
+will be aware of all the exceptions a method can throw and can design the code
 to deal with them ~\cite{Robil00}.
@@
 we opted for not restricting the analysis only on defect
-issues.\footnote{We conducted the same analysis on the defect issues and the top
+issues\footnote{We conducted the same analysis on the defect issues and the top
 exceptions are similar to the ones mentioned on defect issues. Due to space
@@
 on the defect issues can be found at:
-\url{www.dimap.ufrn.br/~roberta/android_repo_mining}}
-
+\url{www.dimap.ufrn.br/~roberta/android_repo_mining}}.
  Moreover, issues and pull requests are dual on Github; all pull requests have a corresponding 
@@
 \subsubsection{Android Apps in Google Code}
-
-Differently from GitHub, Google Code does not provide an API to access the information related
+Google Code contains largely used open-source Android apps (e.g. K9Mail \footnote{K9Mail moved to Github but as a way of not loosing the project history 
+it advices their users to report bugs on Google Code issue tracker.}), however, differently from GitHub, Google Code does not provide an API to access the information related
  to hosted projects \footnote{Google code used to provided a Web service to such repository, but it was deactivated in June 2013 what Google called a "clean-up action".}.
-
-However, Google Code still contains largely used open-source Android apps (e.g. K9Mail \footnote{K9Mail moved to Github but as a way of not loosing the project history 
-it advices their users to report bugs on Google Code issue tracker.})  and the issues reported of such apps are also valuable. 
-To overcome this limitation we needed to implement a Web Crawler (incorporated in Exception Miner tool described next) that navigated 
+To overcome this limitation we needed to implement a Web Crawler (incorporated in ExceptionMiner tool described next) that navigated 
 through the web interface of Google Code projects extracting all issues and issue comments and storing in a relational database for later analysis.
@@
 %To overcome this limitation and mine the information available on such Google Code 
-%projects we implemented a Web crawler which navigated (incorporated in Exception Miner tool described next) 
+%projects we implemented a Web crawler which navigated (incorporated in ExceptionMiner tool described next) 
 % on each project and issue page, extracting the content of each issue to store on relational database.
@@
 trace information with bytecode and sourcecode analysis. Each analysis is implemented as a different module 
-on Exception miner. Next, we present core analyses which were implemente on ExceptionMiner. Such analyses 
+on ExceptionMiner. Next, we present core analyses which were implemente on ExceptionMiner. Such analyses 
 provide aditional information to be used on further analyses described in Section ~\ref{sec:result}:
@@
 
-\emph{\textbf{The prevalence of programming mistakes.}} To get a broader view of the prevalence of
-programming mistakes of reported exception stack traces we: (i) identified the Java exceptions that can be
-implicitly thrown by the runtime environment due to a programing mistake or resource limitation; and (ii)
- used ExceptionMiner to evaluate the prevalence of such exceptions on the reported stack traces.
-Table ~\ref{tab:causes} illustrates the main causes of exceptions reported on issues.
+\emph{\textbf{The prevalence of programming mistakes on stack traces.}} One way of getting a broader view of the prevalence~\footnote{The prevalence evaluates the proportion of a given characteristic over a group of subjects.}  of programming mistakes on reported stack traces would be to manually inspect the code responsible for throwning every exception reported  on issues.
+However, inspecting the code related to 6005 stack traces can easily become infeasible.
+Hence, we conducted an empirical investigation based on the following assumption: the name of the exception and its associated Javadoc 
+documentaion hold enought information to pinpont a ``programming mistake".To exemplify, an instance of ArrayOutOfBoundException
+ pinpoints to programming mistake which indicates that an array has been accessed with an illegal index.
+
+To perform this empirical investigation, we selected a subset of all reported root exceptions. 
+This subset was composed by the top 100 root exceptions (from a list ranked by the number of ocurrences on stacks).
+This subset was representative since it contained root exceptions that were reported as the root causes of 95\% of 
+all stack traces. Therefore, based on the inspection of the full names exceptions and the associated 
+ Javadocs we could identify from this subset, which exceptions were related to programming mistakes.
+ It is woith mentioning that  all exceptions identified as programming mistakes were exceptions defined on java.lang and java.util packages. 
+After doing such analysis, we could calculate the prevalence of such exceptions on issues.
+In other words, we could calculate the proportion of issues mentioning the exceptions identified in this analysis as related to ``programming mistakes".
+
+Table ~\ref{tab:causes} summarized the results. It also illustrates the number of general Java exceptions - 
+whose full name does not carry the reason for the exception to be thrown.
+To ensure the quality of the analysis, three independent coders classified a randomly selected
+sample of 25 exception types (from the total 100) using the same list of concerns;
+the interater agreement was high (96\%). We could also observe that
+We can therefore assume that the categories are descriptive 
+enough for covering the range of exception types analyzed.
+
+The results show that the programming mistakes associated to the most reported root causes. 
+Although there is usually no proper handling for such exceptions, 
+besides presenting an error message to the user and restarting the application ~\footnote{Only 
+high fault tolerant systems need to provide solutions to handle them.}.  
+This finding calls the attention to the following: the high number of programming errors reported 
+on issues are an indication that developers may be underestimating the impact that such 
+mistakes have on application robustness.
 
@@
   \centering
-  \begin{tabular}{lrr}
-    \hline
-    \bfseries{Cause} &  \bfseries{\% Occurrences on stacks} \\
-    \hline
-      Programming logic (java.lang and util) &  52,0\%\\ 
-      Resources Restriction   &   23,9\% \\ 
+  \begin{tabular}{lr}
+    \hline
+    \bfseries{Cause} &  \bfseries{\% Occurrences} \\
+    \hline
+      Programming Mistakes &  52,0\%\\ 
+      %Resources Restriction   &   23,9\% \\ 
       General Exceptions (Error, Exception, Runtime)    &  6,7\%\\
-      Specific  &  xxx\%\\ 
+    \hline
  \end{tabular}
@@
 \end{table}
-
-The analysis on the complete set os stack traces revealed that, in fact, the programming mistakes were
-the most reported root causes.  Although there is usually no proper handling for such exceptions, 
-besides presenting an error message to the user and restarting the application ~\footnote{Only 
-high fault tolerant systems need to provide solutions to handle them.}.
-
-This finding calls the attention to the fact that: high number of programming errors reported 
-on issues may me an indication that developers are underestimating the impact that such 
-mistakes have on application robustness.
 
@@
 
-Besides distilling the information available on stack traces (i.e., exception signalers, root causes, exception wrappings),
-we mined the exception types (i.e., RuntimeException, checked Exception or Error) related to each exception reported on stack traces -
-to do so we used the source code and byte code analyses implemented on Exception Miner.
-
- Table~\ref{tab:typeroottab} presents the types of root causes of all stack traces reported on the analyzed issues. As we can see 
-from Table ~\ref{tab:typeroottab},  most of the reported exceptions signaled (explicilty or implicitly by the runtime enviroment) on methods defined
- on the Android Plaform on on the application itself.
-
-On the other hand, most of the reported exceptions signaled by the set of 
-libraries reused by Android platform were checked exceptions. Such libraries are referred in
-this work as libcore and is composed by org.apache.harmony, org.w3c.dom, sun.misc, 
-org.apache.http, org.json, org.xml, javax. It is pointed as a good practice
-for libraries (see Section~\ref{sec:best}) because by using checked exceptions
-a library can define a precise exception interface ~/cite{miller1997issues} to its clients. Such
-libraries are heavily used in several projects, and such precise exception
-interface may be related to the library's maturity.
+As mentioned before, using the ExceptionMiner tool in combination with manual inspections we could identify
+the root exception types (i.e., RuntimeException, Error, checked Exception) as well as its origins - which were 
+identified based on the package names of the signalers related to it on the stack traces (see Section ~\ref{sec:exceptionminer}).
+Table~\ref{tab:typeroottab} presents the types and origins of root exceptions of all stack traces analyzed.
+
+We can see, from Table ~\ref{tab:typeroottab}, that contrating with the other origins, most of the 
+ exceptions signaled on Android Libcore (i.e., the set of libraries reused by the Android) were 
+checked exceptions. Such set comprises: org.apache.harmony, org.w3c.dom, sun.misc, 
+org.apache.http, org.json, org.xml, javax. Signaling checked exceptions is considered a 
+good practice (see best practice IV in Section~\ref{sec:best}) because by using checked exceptions a library can define a precise 
+exception interface ~/cite{miller1997issues} to its clients.
+ Since such libraries are heavily used  in several projects, such finding can be related to the library's maturity.
 
@@
 % repo !='android'  AND MAIN_EXTENDS LIKE 'RUNTIME%' 
-
 %SELECT COUNT(*) FROM STACKTRACESUMMARY WHERE
@@
 
-% \bfseries{Root Cause Type} & \bfseries{Android} & \bfseries{Libcore} & \bfseries{App} & \bfseries{Lib}  & \bfseries{All}\\
-% Runtime	&	1075 / 260	&	56  / 17	&	1446 / 397	&	374	/  316  &	2951 / 943\\
-%Error	       &	 144	 / 44              &	34 / 12	&	168	 / 134              &	105	 / 62           &	451 /  240	\\
-%Checked	&	193	/ 83            &	230 / 84	&	139	  / 174          &	145 / 422	           &	624  / 734	\\
-%Throwable	&	0 /0	       &	0	&	1 / 1	             &	0	     / 0               &	1 / 1	\\
-%Undefined	&	1 / 3 	&	0	&	6 / 	12		&	8 / 30	   &	15 / 45	\\ %22 DISTINCT EXCEPTIONS AND 9 IN GIT
-
-%Runtime	&	1075 / 160	&	56 	&	1446 / 397	&	374	/  386  &	2951 / 943\\
-%Error	       &	 144	 / 32              &	34	&	168	 / 134               &	105	 / 74           &	451 /  240	\\
-%Checked	&	110	   / 39           &	230	&	139	  / 174            &	145	 / 466           &	624  /  679	\\
-%Throwable	&	0 /0	       &	0	&	1 / 1	             &	0	     / 0               &	1 / 1	\\
-%Undefined	&	1 / 1 	&	0	&	6 / 	12		&	8	/ 32	   &	15 / 45	\\ %22 DISTINCT EXCEPTIONS AND 9 IN GIT
-%All		& 1330	&	320	&	1760	&	632	&	4042	\\
-
-%Runtime	&	1235	&	56 	&	1843	&	760  &	3894\\
-%Error	       &	 176              &	34	&	302               &	179          &	691	\\
-%Checked	&	149           &	230	&	313            &	611           &	1303	\\
-%Throwable	&	0        &	0	&	2             &	0	                    &	2	\\
-%Undefined	&	2 	&	0	&	18		& 40	   &	60	\\ %22 DISTINCT EXCEPTIONS AND 9 IN GIT
-%  \hline
-%All		& 1562	& 320	& 	2478	& 1590	& 5950 \\
-
-%FILTRAR LIBCODE:
-%FILTRO ANDROID: 74 (com.android) 122(com.google)
-%SELECT REPO, main_signaler, *
-%  FROM stacktracesummary where mainsignaler_from='LIB_or_APP' AND main_signaler like 'com.google%' ;
-
 \begin{table}
 \centering
-\begin{tabular}{lccccc}
+\begin{tabular}{lccccr}
     \hline
@@
 
-Runtime	&	1335	&	73	&	1843	&	690  &	3894\\
-Error	       &	 188              &	 46	&	302             &	167	           &	691	\\
-Checked	&	276           &	314	&	313          &	567	           &	1358	\\
-Throwable	&	0	       &	0	&	2            &	0	     / 0               &	2	\\
-Undefined	&	4	&	0	&	18		&	38	   &	60	\\
- \hline
-All		& 1803	&	433	&	2478	&	1462	&	6005	\\
+Runtime	&	1335	&	73	&	1843	&	690  &	3894 (64,85\%)\\  %2466+854
+Error	       &	 188              &	 46	&	302             &	167	           &	691 (11,51\%)	\\
+Checked	&	276           &	314	&	313          &	567	           &	1358 (22,61\%)	\\
+Throwable	&	0	       &	0	&	2            &	0         &	2 (0,03\%)	\\
+Undefined	&	4	&	0	&	18		&	38	   &	60	(1,00\%) \\
+ \hline
+All		& 1  803	&	433	&	2478	&	1462	&	6005	\\
     \hline
@@
 
-%\begin{figure}
-%\centering
-%\includegraphics[width=\hsize]{chart_exceptiontypes.pdf}
-%\includegraphics[width=\hsize]{exception_types.pdf}
-%\caption{Distribution of Checked and Unchecked exceptions.}
-%\label{fig:typeroot}
-%\end{figure}
-
-% REMOVED BECAUSE WE DO NOT HAVE SPACE TO DISCUSS IT, AND IT IS NOT WELL
-% CONNECTED WITH PREVIOUS.
-%\subsection{Uncaught Checked Exceptions}
-%Although most of the  reported exceptions were runtime, we could also find checked
-%exceptions as the root causes of stack traces reported on issues. A checked exception
-%can remain uncaught in two circumstances: if all methods on the execution trace
-%on which this exception flows explicitly specifies the exception, or if the
-%checked exception is wrapped in a runtime and than re-thrown. Since checked exceptions 
-%should be used to represent recoverable conditions, ignoring a
-%checked exception, is considered a bad practice. As mentioned before, when the developer is confronted
-%with a checked exception, the designer of the API is telling him to handle the
-%exceptional condition. To better understand what was causing checked exceptions
-%to scape from being handled we investigated the kinds of wrappings that happened
-%on stacks the stacks.  Next sections presents the wrappings found and discuss
-%about them.
-
-\emph{\textbf{Undocumented Runtime Exceptions}}.
-However, we can also observe a considetable amount of 
-runtime exception signaled on library reused by apps.
-In order to evaluate if  best practice IV was being obeyed, we inspected the   
-signature of each method signaling a runtime exception. 
-We inspected methods of the Android Plaftorm and 
-the libraries, because both are  third  party code reused 
-by the Android apps. In this task we filtered out all the exceptions implicitly 
-signaled by  runtime environment (due to programming mistakes) 
-since these exceptions should not be documented on the 
-method signature.
-
-As a result of this inspection, we could observe that only 1 method out of 118 inspected signalers
+We can also observe that most of the reported exceptions were RuntimeExceptions
+(64,85\%); and most of them were signaled on methods defined on the Application 
+the Android Platform or on libraries.Taking into account the results of empirical study on 
+the prevalende of ``programming erros" we know that a considerable ammount 
+of such exceptions, are related to the programming mistakes
+as pointed out previously.
+
+However, within this set of runtime exceptions there are also the ones that do not correspond to programming mistakes,
+and according to best practice IV should be documented as part of the exception interface of libraries/framework reusable
+methods.
+
+To investigate the obedience to such practice, we firstly filtered out all the exceptions implicitly
+ signaled by  runtime environment (due to programming mistakes) since these exceptions 
+should not be documented on the method signature. Then we inspected the code each method 
+(defined either on Android or on libraries) signaling  runtime exceptions not-related to programming mistakes.
+As a result of this inspection, we could observe that only 1 method out of 118 inspected methods
  (i.e., 0,8\%) documented the explicitly thrown runtime exception in a Javadoc comment; and none of these methods
@@
 \emph{\textbf{Runtime wrapping Checked Exception.}} Such wrapping was 
-detected on 88 projects. Inspecting the 
-elements responsible for performing such wrapping, we could observe that
+detected on 88 projects. Inspecting the elements responsible for performing such wrapping, we could observe that
 approximately  50\% of such wrappings were performed on methods defined on Android
@@
  we could observe that approximately 41\% of them were performed by a single class,
- named ActivityThreat.
-
-This class is responsible for managing the life cycle of  Android Activities - 
+ named ActivityThreat. This class is responsible for managing the life cycle of  Android Activities - 
 invoking the life cycle methods (i.e. onCreate(), onPause()) for every Activity 
@@
 Figure ~\ref{fig:snippets} (a) presents a code snippet of one the methods of 
-ActivityThrea,  the one responsible for invoking the Activitiy's onCreate method. 
-As we can see from lines 4-8,
- in case any exception is thrown in the context of such try-catch block, 
+ActivityThreat,  the one responsible for invoking the Activitiy's onCreate method. 
+As we can see from lines 4-8, in case any exception is thrown in the context of such try-catch block, 
 it is wrapped in a general Runtime exception and re-thrown.
@@
 Inspectioning the source code, of the methods involved in this kind of wrapping,
-revealed another fragility: a checked exception thrown by a native method and not
-declared on the exception interface of thes methods signaling them. Such native method 
-was defined on Android Platform -  which uses Java Native Invocation (JNI) to access 
+revealed an unexpected fragility: a checked exception thrown by a native method and not
+declared on the exception interface of these methods signaling them. Such native method 
+was defined on the Android Platform -  which uses Java Native Invocation (JNI) to access 
 native C/C++ code.
@@
  However, the C-code implementation did throw a "checked exception" NoSuchMethodException, 
-violating the declaration. The Java compiler, could not detect this violation because it does 
+violating the declaration. The Java compiler, cannot detect this violation because it does 
 not perform static exception checking on native methods. This type of bug is hard to diagnose
@@
 
+ "Everybody hates thinking about exceptions, because they’re not supposed to happen"
+  Brian Foote
+
 \subsection{On the need of a rescue to the almost infeasible task of preventing uncaught exceptions}
@@
 
-One may think that the solution for the uncaught exceptions 
-may be to define a general handler, 
+One may think that the solution for the uncaught exceptions may be to define a general handler, 
 which is responsible for handling any exception that is not
@@
 
-Such handler does not replace a carefully designed exception 
+However, such handler cannot replace a carefully designed exception 
 handling policy, which can hardly be designed in the absence of 
@@
 
-On the other hand documenting runtime exceptions is a tedious and error prone task, to help developers
+On the other hand, documenting runtime exceptions is a tedious and error prone task, to help developers
 mitigate this problem tools should be developed to automate the documentation of runtime exceptions
@@
 
-\subsection{On the cross-type wrappings as a threat to dependability}
-The cross-type wrappings detected in this study points to the fact that: (1) exception 
-wrappings may prevent exception types from being used according to its initial purpose
- (e.g, Errors should represent situations that should not be handled); and (2) may  be used
- to bypass the language restrictions imposed by checked exceptions  (e.g,
- checked exceptions may be wrapped in runtime exceptions).
-
-Hence, when (mis)applied, the exception wrapping can make the exception-related code
+\subsection{The exception handling confusion problem}
+When (mis)applied, the exception wrapping can make the exception-related code
  more complex and lead to what we call an \emph{exception handling confusion problem}.
-Such problem can lead the program to
-an unpredictable state in the presence of exceptions. To illustrate this problem, we can use
- one of the examples found in this study: when the developer is 
+Such problem can lead the program to an unpredictable state in the presence of exceptions. 
+To illustrate this problem, we can use one of the examples found in this study: when the developer is 
 confronted with a checked exception, the designer of the API is telling him 
@@
 by the runtime environment which make impossible the program to recover. 
-
 Currently there is no way of enforcing Java exception type conventions during program development.
@@
 
+%The cross-type wrappings detected in this study points to the fact that: (1) exception 
+%wrappings may prevent exception types from being used according to its initial purpose
+%(e.g, Errors should represent situations that should not be handled); and (2) may have been
+% used to bypass the language restrictions imposed by checked exceptions  (e.g,
+% checked exceptions may be wrapped in runtime exceptions).
+
 \subsection{On the null pointer problem}
 
-Besides the explicitly thrown runtime exceptions, in this work we also observe a high number of 
-stack traces reporting coding errors (out-of-bounds array index, division-by-zero, access to a null).
-Such exceptions are implicitly thrown by runtime environment, and among them the NullPointerException 
-were the reported root cause in most of the analyzed exception stack traces.
-
-The null references was firstly introduced by Tony Hoare in ALGOL W, which after some years he called his “one-billion-dollar mistake”:
+IThe null references was firstly introduced by Tony Hoare in ALGOL W, which after some years he called his “one-billion-dollar mistake” ~\cite{hoare}:
 
@@
 
-The high number of null pointer exceptions found in this mining study shed light into a problem
-that although known was not supported by large scale studies.
-
+In this sudy, the null references were, infact, responsible for several reported application errors - providing more evidences to Hoare's speach.
 This observation emphasizes the importance of solutions to avoid NullPointerExceptions, such as:
-(i) lightweight intra-method null pointer analysis supported by Java8 @Nullable annotations;
+(i) lightweight intra-method null pointer analysis supported by Java8 @Nullable annotations ~\footnote{The @Nullable annotations are used by tools such as FindBugs, Eclipse, NetBeans, IntelliJ to detect potential
+null pointer dereferences at compile time. In April/2014 the Android Studio 0.5.5 was 
+released with support to these annotations};
 (ii) inter-method null pointer analysis tools such as the one proposed by Nanda and Sinha ~\cite{nanda2009accurate};
@@
 or computations that may fail) could improve the robustness of Java programs. 
-
-The @Nullable annotations are used by analyzers like the Checker Framework, FindBugs, Eclipse, 
-NetBeans, IntelliJ, and other commercial tools which run at compile time and detect potential null 
-pointer dereferences. Android Studio 0.5.5 Released support these annotations.
-
-%(1) mostrar a utilizacao de @nullable em projetos android. rodar em alguns e mostrar.
-%Android Project Unwritten Rules (http://blog.lemberg.co.uk/android-project-unwritten-rules)
-%(2) dar foco em android
-%(3) falar do modelo do ciclo de vida da aplicacao e da necessidade de usar excecoes Runtime.
-%(4) Local onde tratar excecao - handler padrao - politica da aplicacao. tratamento difucultado multiplas threads e assyncronos.
-
-%\noindent\emph{Unwritten Rules of Android Exception Handling Developement}
-%\noindent\emph{Dependency injection - Inversion of Control -  callback and Exceptions }  
-%\subsection{Dependency injection - Inversion of Control -  callback and Exceptions}  
-%\noindent\emph{How can we benefit from such cross-project crash analysis?}
-
-%\subsection{Inversion of Control and Exception Handling}
-\subsection{On the absence of exception interface on hook methods of the Android Aplication Framework}
-
-In this study, the analyzed stack traces involved, in one way or another, methods from the 
-Android Application Framework: either framework utility methods inkoked 
-by the application, or framework hook methods  (e.g., OnCreate(), onPause())  
- extended by the application and invoked by the framework (by inversion of control).
- As discussed before, some of the utility 
-methods signaled ``undocumented" runtime exceptions. Additionaly, another characteristic of 
-Android Platform may also be imposing obstacles to application robustness: the absence 
-of exception interface (i.e., a list of exceptions that a method 
-might raise during its execution) of Android rook methods.
-
-For instance, none of the hook methods povided by the Activity class (and 
-extended by every Android app) specify an exception interface. Therefore, such methods
-do not stimulate the developer to think that ``something" may go wrong 
-during the execution of such methods, and prepare the code to deal with them.
-
-To make matters worse, according to Java handling model, the exception handler
- should be added on the invocation chain of the signaler.  However,  on method invocations based 
-on inversion of control, the developer does not have fully access to the invocation 
-chain of a method.
-
-Such difficulties may lead the developers into believing that by just re-throwing the exceptions they can 
-forget about the exceptional situations, while focusing on the developement of the `happy path". 
-This `ignore-for-now" approach may lead to the known uncaught exceptions problem~\cite{jo2004uncaught},
-and unintened handler actions,  contributing to degrate apps robustness.
-
-Rephasing Brian Foote "Everybody hates thinking about exceptions, because they’re not supposed to happen"
-however one day they happen, and you need to prepare your code for it. However, further investigation
- is needed to evaluate the impact of  the absence of exception interface of Android rook methods on
-app robustness.
-
-\subsection{On the benefits of a cross-project exception stack trace analysis}
-
-When family of systems make use of a common core, it is possible to benefit from a cross project crash analysis like the one performed in this study.
- The not only the exception stack traces but any other bug issues reported for a group of systems can identify vulnerabilities 
-that are common to more then one element of the family and are more difficult to identify in a single project analysis.
-Therefore tools such as the one developed in this study may allow the developers to 
-benefit from the faults happening on different systems.
-
-%It can be a way of dealing with the test case explosion problem related to framework test.
-%Let the products test and report the crahes in a single repository and then analize them.
-%... A test suite is always incomplete,
-%one application could benefit from the usage scenarios on which such framework is applied.
-%Framework testing -  what are the main challenges on framework testing?
-
-%Using exception miner tool we could identify the methods responsible for most of the crashes - 
-%such methods can be a target for manual inspections improvements, or better documentation
-%(because sometimes an exception can be thrown because the user is not invoking the method
-%in the proper way ....
-
-During such analysis we could also find the undocumented runtime exceptions 
-signaled by reused therd part code libraries.
-
-\subsection{Actionable Advices - The Unwritten rules of Android development}
-to do if we have space...
-
-%The results of a mining study should be a set of actionable advices (tao xie).
+.
 
@@
 \label{sec:threats}
-
 \noindent\emph{Internal Validity} We used a heuristics-based parser to mine
@@
 methods' bytecode operations.  They observed that most of the stack traces were
-related to NullPointerException and other JMV implicitly thrown exceptions had
+related to NullPointerException and other implicitly thrown exceptions had
 the higher prevalence on the analyzed set of stacks.
@@
 \end{document}
+
+http://docs.oracle.com/javase/7/docs/api/
+
+%JAVA.LANG EXCEPTIONS
+ArithmeticException
+ArrayIndexOutOfBoundsException
+ArrayStoreException
+ClassCastException
+ClassNotFoundException
+CloneNotSupportedException
+EnumConstantNotPresentException
+Exception
+IllegalAccessException
+IllegalArgumentException
+IllegalMonitorStateException
+IllegalStateException
+IllegalThreadStateException
+IndexOutOfBoundsException
+InstantiationException
+InterruptedException
+NegativeArraySizeException
+NoSuchFieldException
+NoSuchMethodException
+NullPointerException
+NumberFormatException
+ReflectiveOperationException
+RuntimeException
+SecurityException
+StringIndexOutOfBoundsException
+TypeNotPresentException
+UnsupportedOperationException
+
+JAVA UTIL:
+ConcurrentModificationException
+DuplicateFormatFlagsException
+EmptyStackException
+FormatFlagsConversionMismatchException
+FormatterClosedException
+IllegalFormatCodePointException
+IllegalFormatConversionException
+IllegalFormatException
+IllegalFormatFlagsException
+IllegalFormatPrecisionException
+IllegalFormatWidthException
+IllformedLocaleException
+InputMismatchException
+InvalidPropertiesFormatException
+MissingFormatArgumentException
+MissingFormatWidthException
+MissingResourceException
+NoSuchElementException
+TooManyListenersException
+UnknownFormatConversionException
+UnknownFormatFlagsException

### 2014-09-04 android-stacks.tex
results - question 3

@@
 
-%the mobile industry shipped over 315 million smartphones in the last quarter of 2013
-%The pervasiveness of the World Wide Web and the astonishing increase 
-%in the number of smartphones stimulated the creation of applications which extend the phones
-% capabilities far beyond of the basic calls and textual messages. 
-%f we look at global smartphone shipments the numbers are astonishng;
-%the mobile industry shipped over 315 million smartphones in the last quarter of 2013 [ref-googleIO].
-%Such increase on smartphones combined with the pervasiveness of the World Wide Web
-%have stimulated the creation of applications which extend the phones caVALIAR MOVER ESTE STACK TRACES PARA UMA FIGURA QUE MOSTR COMO ELE QUEBRADO... PELO EXCEPTION Mpabilities 
-%far beyond of the basic calls and textual messages. 
-315,92
 In this context, Android has become a leading platform - thanks to its open-source 
@@
 and the number of apps built on top of Android increases in a daily rate.
-
-%ref: http://www.gartner.com/newsroom/id/2645115
-% Gartner, Inc. Press release, 2012. www.gartner.com/newsroom/id/2237315.
-% http://www.gartner.com/newsroom/id/2645115
-
-%If on the one hand such applications are getting mode and more popular....
-%On the other hand... there is a need for reliability, as some core functionalities 
-%are moving to apps....such applications impose more challenges to software reliability 
-%Such mobile platforms are intrinsically multi-threaded and support a development model 
-% slightly (or sometimes drastically) different from those adopted on Web development [arie-icse],  
-%abnormal computation states
-
-%In the same rate the number of a users increase, also increases 
-%the number of users affected by application faults and crashes.
 
@@
 battery restrictions; among others.
-
-%Hence, in the same rate the number of mobile users increase, also increases 
-%the number of users affected by such threats.
-%faulty apps and app crashes.
 
@@
 
-% Google estimates that Android has over 1 billion active users per month, 
-% who pull out theirs phones and check it over 100 billion times.
-
-%If we google for "Android app crashes"
-% we can find 1.200.000 pages about the subject on forums and blogs. In Feb. 2014
-%Gartner issue a report on mobile technologies and capabilities 
-%[ref-https://www.gartner.com/doc/2665315/top--mobile-technologies-capabilities], 
-%which emphasized the importance of monitoring application crashes.
-
 Our conjecture is that several Android applications may be suffering from common 
 fragilities on the exception-related code (of the applications themselves or 
-the underlaying platform) which may be degratind the robustness of such applications.
+the underlaying platform) which may be degratind their robustness.
 We call ``fragilities" characteristics on the exception-related code that favor the introduction
 of aforementioned failures.
-
-%, and that such fragilities may be leading to
-%decrease the application robustness, leading for instance to the Uncaught Exception problem -
-% which is one of the main causes of application crashes ~\cite{jo2004uncaught, zhang2012amplifying}.
-%investigate the whether the exception stack traces
-% can reveal common characteristics on the exception-related code that
-%may be favoring the introduction of the aforementioned failures, which would be responsible for 
-%decreasing the application robustness. We call such characteristics ``fragilities" of the exception-related code.
-
-%[ref AST].
 
@@
 
-% The
-%concentration of open source projects and the existence of initiatives that mine
-%such information, such as GHTorrent~\cite{Gousi13}, allow the information about
-%projects to be visible within and across projects and make a full population
-%studies possible.
-
-%Among the issues defined on issue trackers, we can find issues reporting bugs, and system crashes 
-%found as a result of application executions. Such issues usually contain exception stack traces [ref][ref]. 
-%<SUCH ISSUES INCLUDE STACK TRACES. WHAT IS THE RELATION BETWEEN STACK TRACE AND CRASH>
-%<stack traces are part of good issues....>
-
-%Stack traces are a useful source of information about system crashes, and is often used to support developers in debugging~\cite{schroter2010stack}. Moreover, information %from stack traces 
-%combined with crash and bug reports can be used to support bug classification and clustering~\cite{wang2013improving, kim2011crash, %dhaliwal2011classifying}, fault-proneness prediction %models~\cite{kim2013predicting} and even automated bug fixing~\cite{sinha2009fault} tools.
-
 Until recent years, the availability of data, reporting such failures, was scarce. 
@@
 automated bug fixing ~\cite{sinha2009fault} and fault-proneness prediction models ~\cite{kim2013predicting}.
-
-%Most of such studies, however, focused the analysis on one or two system at a time.
-%This paper reports an exploratory study which targets a family of applications 
-%(sharing the common Android framework) aiming at extracting 
-%the common characteristics of exception stack traces reported to such systems to provide
-%insights on similar fragilities on the exception related code of the applications and underlaying framework.
-
-%482 applications) and Google code (157 applications)
-%GitHub 31,592
-% Google Code: 127,456
-%All: 159,048
 
@@
 
-%projects: 788	 - projects with issues - 724 - 
-%projects with issues (with exceptions): 243 
-%projects with stacks: 158  -  stacks: 3559 (defect)
-%Github: issues: 50342 - stacks: 3865 - projects: 2542 - projects with stacks: 293
-
-%To guide our analysis, we compiled a set of best practices on how to use
-%exceptions by composing relevant suggestions by Gosling~\cite{gosling2000java},
-%Wirfs-Brock~\cite{wirfs2006toward} and Bloch~\cite{bloch2008effective}.  The
-%comparison of those with the mined exceptions revealed several interesting
-%findings, such as:
-
 Some outcomes consistently detected during this study were the following:
@@
 
-\emph{II-Error represents an unrecoverable condition detected by the runtime enviroment which
-should not he handled} (\cite{gosling2000java}). It results from failures detected
-by the Java Virtual Machine which indicate resource deficiencies, invariant
-failures or other conditions that make impossible the program to recover.
+\emph{II-Error represents an unrecoverable conditions which should not he handled} 
+(\cite{gosling2000java}).  Errors should result from failures detected
+by the runtime environemt which indicate resource deficiencies, invariant
+failures or other conditions, that make impossible for the program to recover.
 
@@
 \subsection{What can the Exception Wrappings tell us?}
-%\subsection{Exploring the Information Embedded on Exception Wrappings}
-
-Table~\ref{tab:wrappingandroid} presents the wrappings found in this study for all
-stack traces found in Android repositories. As we can see, some of the checked
-exceptions were indeed wrapped in runtime exceptions or even errors.
 
@@
 \end{table*}
-
-This analysis revealed unexpected wrappings such as: Error wrapping a Checked 
-exception; a Runtime wrapping and Error; and an Error wrapping a Runtime.
 Java is the only language that provides a hybrid exception model 
-which offers  different types to represent different exception behaviors (i.e., error, 
-runtime and checked) (see Section~\ref{sec:best}). For instance,
-according to Java specification, Errors should not be handled inside the system 
-since they usually represent unrecoverable conditions
-detected by the runtime environment (e.g.,  OutOfMemoryError). On the other hand, checked exceptions should
-represent recoverable conditions. Next we discuss and exemplify the cross-type wrappings
+which offers three kinds of exceptions each one holding an inttended exception behavior (i.e., error, 
+runtime and checked). The best practices I and II states the behavior that Error an Checked exceptions hold, which are: 
+(i) Errors should not be handled inside the system since they usually represent unrecoverable conditions detected 
+by the runtime environment (e.g.,  OutOfMemoryError); (ii) checked exceptions, on the other hand, should represent recoverable conditions.
+
+Table~\ref{tab:wrappingandroid} presents all wrappings found in this study, that
+envolve different exception types (i.e., Error, Checked, Exception). We call 
+such a wrapping a ``cross-type wrapping".  Such wrappings reveals 
+violations to best practices I and II (e.g. Error wrapping a Checked exception),
+To illustrate the problem behind such wrappings, we can use one of the wrappings found in this study. 
+When the developer is confronted with a checked exception, the designer of the API is telling him 
+to handle the exceptional condition, however we could observe that in some cases the 
+checked exception wrapped an OutOfMemoryError, which represent resource deficiency detected 
+by the runtime environment which make impossible the program to recover. 
+Hence, trying to handle this exception may lead the program to an unpredictable state.
+
+Next we discuss and exemplify the cross-type wrappings
  found in this study.
@@
 approximately  50\% of such wrappings were performed on methods defined on Android
-Platform.
-
-Taking a closer look on the methods responsible for such wrappings,
+Platform. Taking a closer look on the methods responsible for such wrappings,
  we could observe that approximately 41\% of them were performed by a single class,
@@
 
-\begin{table}
-\centering
-\begin{tabular}{lll}
-    \hline
- \bfseries{Signaler class} &  \bfseries{Wrapper Type} & \bfseries{Ocurrences} \\
-    \hline
-android.app.ActivityThread & java.lang.RuntimeException & 61\\
-android.view.LayoutInflater & android.view.InflateException &  7 \\
-android.os.AsyncTask & java.lang.RuntimeException &  7\\
-\hline
-  \end{tabular}
-\caption{Top 3 Runtime-Cheked wrappers.}
-\label{tab:wrapping01}
-\end{table}
-
-%MOVED TO THE PICTURE
-%{\footnotesize
-%\begin{verbatim}
-%1. private Activity performLaunchActivity(...) {
-%2.  try {
-%3.   ...
-%4.   ...callActivityOnCreate(...);
-%5.  } catch (Exception e) {
-%6.    ...
-%7.   throw new RuntimeException(..., e);
-%8.  }
-%9.}
-%\end{verbatim}
-%}
-
-%SELECT SIGNALER_method, COUNT(*) AS N
-%  FROM stacktracesummary WHERE ROOT_EXTENDS='CHECKED' AND EXCEPTION_EXTENDS='RUNTIME' and
-%  ( SIGNALER_method like '%onDestroy%' or
-%   SIGNALER_method like '%onCreate%' or
-% SIGNALER_method like '%onStart%' or
-% SIGNALER_method like '%onResume%' or
-% SIGNALER_method like '%onPause%' or
-% SIGNALER_method like '%onStop%' or
-% SIGNALER_method like '%onDestroy%' )
-%   GROUP BY SIGNALER_method
-%  ORDER BY N DESC;
-
-%SELECT COUNT(*) AS N
- % FROM stacktracesummary WHERE ROOT_EXTENDS='CHECKED' AND EXCEPTION_EXTENDS='RUNTIME' and
- % ( stack like '%onDestroy%' or
- %  stack like '%onCreate%' or
-% stack like '%onStart%' or
-% stack like '%onResume%' or
-% stack like '%onPause%' or
-% stack like '%onStop%' or
-% stack like '%onDestroy%' );
-
-%GC = 6
-
-%However, since the Activity life clycle method (i.e., callActivityOnCreate(...)) that is invoked in this block,
-% do not include checked exceptions on its signatures. Hence they could not throw a checked exception to be handled 
-%in the method illustrated above. However the code inspection performed on such wrappings revelaed an
-%``undocumented checked exception" signaled by this life cycle method.
-
-%\emph{\textbf{Undocumented Checked Exceptions.}} There are usually
-%2 possible for a checked exception to be thrown by a method:
-%it can either be specified on the method signature or it can be wrapped in a 
-%runtime exception and re-thrown.
+%\begin{table}
+%\centering
+%\begin{tabular}{lll}
+%   \hline
+% \bfseries{Signaler class} &  \bfseries{Wrapper Type} & \bfseries{Ocurrences} \\
+%    \hline
+%android.app.ActivityThread & java.lang.RuntimeException & 61\\
+%android.view.LayoutInflater & android.view.InflateException &  7 \\
+%android.os.AsyncTask & java.lang.RuntimeException &  7\\
+%\hline
+%  \end{tabular}
+%\caption{Top 3 Runtime-Cheked wrappers.}
+%\label{tab:wrapping01}
+%\end{table}
 
@@
 
-% Several types of exceptions were  wrapped the RuntimeExceptions thrown by the 
-% ActivityThread. Table X illustrates them.
-% #\begin{table}
-% \centering
-% \begin{tabular}{lll}
-%    \hline
-%    \bfseries{wrapper} & \bfseries{root} & \bfseries{\#} \\
-%    \hline
-% java.lang.RuntimeException & java.lang.ClassNotFoundException &	32 \\
-% java.lang.RuntimeException & java.lang.NoSuchMethodException &	6 \\
-% java.lang.RuntimeException & java.lang.InstantiationException) & 5 \\
-% ...andlyticsproject.NetworkException & org.apache...HttpResponseException &	5\\
-% android.view.InflateException & java.lang.NoSuchMethodException & 4 \\
-% \hline
-%  \end{tabular}
-% \caption{Elements responsible for the wrapping.}
-% \label{tab:wrapping01}
-% \end{table}
-
-% \begin{table}
-% \centering
-% \begin{tabular}{lll}
-%   \hline
-%   \bfseries{Exception} & \bfseries{Ocurrences} \\
-%   \hline
-% java.lang.RuntimeException &	69 & 61\% \\
-% android.view.InflateException	& 10 & 8,7\% \\
-% com.github.andlyticsproject.console.NetworkException	& 7 & - \\
-% java.lang.IllegalStateException	& 6 & - \\
-% \hline
-%  \end{tabular}
-% \caption{Elements tresponsible for the wrapping.}
-% \label{tab:wrapping01}
-% \end{table}
-
 \emph{\textbf{Runtime Exception wrapping Error.}} We could find  64 exception stack traces 
@@
 by android.os.AsyncTask. This class is responsible for dealing with asynchronous tasks in Android.
-The code snippet is illustrated in Figure ~ref{wrappings} (c) shows that any throwable signaled in the context of an asynchronous 
-task is converted in to a runtime and re-thrown. Some of the Erros wrapped on instances of runtime exceptions were 
-java.lang.OutOfMemoryError (30), java.lang.NoClassDefFoundError (6), 
-java.lang.UnsatisfiedLinkError (5), java.lang.StackOverflowErro (4), as illustrated in Table ~\ref{tab:exampeswrap}.
-
-%ASSNCTASKS NON ANNOTATED METHODS…….
-%SOME HAVE EXCEPTION INTERFACE 
-%EXCEPTIONS ON MEHTODS SIGNATURE OTHERS NOT
-%INCONSISTENCY…. DEPENDS ON DEVELOPERS 
-%WORK INSTEAD OF TOOLS.
-
-%{\footnotesize
-%\begin{verbatim}
-%protected void done() {
-% ...
- %  try {
- %     ...
-  % } catch (InterruptedException e) {
-  %    android.util.Log.w(..., e);
-  % } catch (ExecutionException e) {
-  %    throw new RuntimeException("...",e.getCause());
-  % } catch (CancellationException e) {
-  %    ...
- %  } catch (Throwable t) {
- %    throw new RuntimeException("...", t);
- %  }
- %   ...
-% }
-%}
-%\end{verbatim}
-%}
+The code snippet is illustrated in Figure ~ref{wrappings} (c) shows that any instance of Throwable signaled in the context of an asynchronous 
+task is converted in to a runtime and re-thrown.  Table ~\ref{tab:exampeswrap} exemplifies some of such wrappings.
+Such wrapping masks the unrecoverable exception in a general excepion - doing so  that can be mistakenly handled
 
@@
 
-%\begin{table}
-%\centering
-%\begin{tabular}{ll}
-% \bfseries{Wrapper type - Root cause type} & \bfseries{\#}\\
-% \hline
-%...android.robospice.CacheSavingException - java.lang.ClassCastException & 6 \\
-%java.lang.reflect.InvocationTargetException - java.lang.NullPointerException &  3 \\
-%java.lang.reflect.InvocationTargetException - ...ResourcesdNotFoundException & 2 \\
-%java.sql.SQLException - android.database.sqlite.SQLiteDiskIOException & 2 \\
-%  \end{tabular}
-%\caption{Elements tresponsible for the wrapping.}
-%\label{tab:wrapping01}
-%\end{table}
-
 \emph{\textbf{Checked Exception wrapping Error.}} Almost all of these wrappings
@@
 used by Android applications. The methods responsible for the wrappings
-were also native methods written in C. Table ~\ref{tab:exampleswrap} illustrates
+were also native methods written in C. Table~\ref{tab:exampleswrap} illustrates
 some oth these wrappings some of them are masking an OutOfMemoryError
@@
 
-%\begin{table}
-%\centering
-%\begin{tabular}{lll}
-% \bfseries{Wrapper type - Root cause type} & \bfseries{\#}\\
-% \hline
-%java.lang.reflect.InvocationTargetException - java.lang.OutOfMemoryError & 5 \\
-%java.lang.Exception - java.lang.OutOfMemoryError & 1 \\
-%java.lang.reflect.InvocationTargetException - java.lang.NoSuchMethodError &	1 \\
-%java.lang.reflect.InvocationTargetException - java.lang.UnsatisfiedLinkError&	1 \\
-%  \end{tabular}
-%\caption{Elements tresponsible for the wrapping.}
-%\label{tab:wrapcheckerr}
-%\end{table}
-
-\emph{\textbf{Error wrapping Checked Exception.}} Some of these wrappings were performed by methods defined by java.lang.Class methods such as 
-at java.lang.Class.getDeclaredFields(Native Method)
-at java.lang.Class.getDeclaredConstructors(Native Method), 
-while others were performed by javax.crypto package. Table X illustrates
+\emph{\textbf{Error wrapping Checked Exception.}} These wrappings were performed
+either by native methods defined by java.lang.Class (e.g., getDeclaredFields, getDeclaredConstructors) 
+or methods from javax.crypto package. Table~\ref{tab:exampleswrap} illustrates
 the exception types involved in such wrappings.
-
-%\begin{table}
-%\centering
-%\begin{tabular}{lll}
-% \bfseries{Wrapper} & \bfseries{Root} & \bfseries{\#}\\
-% \hline
-%java.lang.NoClassDefFoundError & java.lang.ClassNotFoundException & 4 \\
-%java.lang.AssertionError & javax.crypto.ShortBufferException) & 3 \\
-% \end{tabular}
-%\caption{Elements tresponsible for the wrapping.}
-%\label{tab:wrapping01}
-%\end{table}
 
@@
 
-%\begin{table}
-%\centering
-%\begin{tabular}{ll}
-% \bfseries{Wrapper type - Root cause type} & \bfseries{\#} \\
-% \hline
-%java.lang.ExceptionInInitializerError - java.lang.NullPointerException & 9 \\
-%java.lang.ExceptionInInitializerError - java.lang.IllegalArgumentException &	2 \\
-% \end{tabular}
-%\caption{Elements tresponsible for the wrapping.}
-%\label{tab:wrapping01}
-%\end{table}
-
 \emph{\textbf{Multiple Wrappings.}} We could also observe that in 205 stacks found
@@
 
+AS WE COULD OBSERVE...
+
 \noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
-Odd wrappings were detected (e.g., a checked exception wrapping an OutOfMemoryError).
-The study also detected an undocumented checked exception signaled by
-  a native method defined in Android.
+MUDARRRR Odd wrappings were detected (e.g., a checked exception wrapping an OutOfMemoryError).
+The study also detected an undocumented checked exception signaled by a native method defined in Android.
 }
@@
 
-\subsection{On the need of a rescue to the almost infeasible task of preventing uncaught exceptions}
+\subsection{On the need of rescues to the almost infeasible task of preventing uncaught exceptions}
 In this study we could observe undocumented runtime exceptions thrown by third party code,
@@
 scraping from library code, few solutions in this directions have been proposed so far ~\cite{van2005combining}.
+
+\subsection{On the null pointer problem}
+The null references was firstly introduced by Tony Hoare in ALGOL W, which after some years he called his “one-billion-dollar mistake” ~\cite{hoare}:
+
+\emph{``I call it my billion-dollar mistake. It was the invention of the null reference in 1965 [...] I couldn’t resist the temptation to put 
+in a null reference, simply because it was so easy to implement. This has led to innumerable errors, vulnerabilities, and system 
+crashes, which have probably caused a billion dollars of pain and damage in the last forty year."}
+
+In this sudy, the null references were, infact, responsible for several reported application errors - providing more evidences to Hoare's speach.
+This observation emphasizes the importance of solutions to avoid NullPointerExceptions, such as:
+(i) lightweight intra-method null pointer analysis supported by Java8 @Nullable annotations ~\footnote{The @Nullable annotations are used by tools such as FindBugs, Eclipse, NetBeans, IntelliJ to detect potential
+null pointer dereferences at compile time. In April/2014 the Android Studio 0.5.5 was 
+released with support to these annotations};
+(ii) inter-method null pointer analysis tools such as the one proposed by Nanda and Sinha ~\cite{nanda2009accurate};
+or (iii) whether language designs which avoid null pointers, such 
+as Monads ~\cite{Walde95} (i.e., used in functional languages for values that may not be available 
+or computations that may fail) could improve the robustness of Java programs.
 
@@
 by the runtime environment which make impossible the program to recover.
+
 Currently there is no way of enforcing Java exception type conventions during program development.
@@
 % checked exceptions may be wrapped in runtime exceptions).
-
-\subsection{On the null pointer problem}
-
-IThe null references was firstly introduced by Tony Hoare in ALGOL W, which after some years he called his “one-billion-dollar mistake” ~\cite{hoare}:
-
-\emph{``I call it my billion-dollar mistake. It was the invention of the null reference in 1965 [...] I couldn’t resist the temptation to put 
-in a null reference, simply because it was so easy to implement. This has led to innumerable errors, vulnerabilities, and system 
-crashes, which have probably caused a billion dollars of pain and damage in the last forty year."}
-
-In this sudy, the null references were, infact, responsible for several reported application errors - providing more evidences to Hoare's speach.
-This observation emphasizes the importance of solutions to avoid NullPointerExceptions, such as:
-(i) lightweight intra-method null pointer analysis supported by Java8 @Nullable annotations ~\footnote{The @Nullable annotations are used by tools such as FindBugs, Eclipse, NetBeans, IntelliJ to detect potential
-null pointer dereferences at compile time. In April/2014 the Android Studio 0.5.5 was 
-released with support to these annotations};
-(ii) inter-method null pointer analysis tools such as the one proposed by Nanda and Sinha ~\cite{nanda2009accurate};
-or (iii) whether language designs which avoid null pointers, such 
-as Monads ~\cite{Walde95} (i.e., used in functional languages for values that may not be available 
-or computations that may fail) could improve the robustness of Java programs. 
-.
 

### 2014-09-05 android-stacks.tex
results - intro e wrappings

@@
 (1) undocumented runtime exceptions thrown by third party code;
-(2) undocumented checked exceptions thrown by JNI native code;
+(2) undocumented checked exceptions thrown by JNI interface;
 (3) unexpected exception wrappings (e.g., Errors being wrapped in checked exceptions),
@@
 
+%Such increase is motivated by pervasiveness of the World Wide Web 
+%and the unprecedented increase of global smartphone shipments -
+%the mobile industry shipped over 315 million smartphones in the last quarter of 2013 ~\cite{googleio}.
+
+%In this context, Android has become a leading platform - thanks to its open-source 
+%model and hardware partners like Samsung, HTC, Motorola, Asus (besides Nexus Google's
+% own device) ~\cite{gartner}.  Google estimates that there is over 1 billion Android active users per month ~\cite{googleio},
+%and the number of apps built on top of Android increases in a daily rate.
+
+%Our conjecture is that several Android applications may be suffering from common 
+%fragilities on the exception-related code (of the applications themselves or 
+%the underlaying platform) which may be degratind their robustness.
+%We call ``fragilities" characteristics on the exception-related code that favor the introduction
+%of aforementioned failures.
+
+%In Java, when an application fails due to an uncaught exception, 
+%it automatically terminates, while the system prints a stack trace to the console, 
+%or log file ~\cite{gosling2000java}.  A typical Java stack trace consists of  the fully qualified name 
+%the thrown exception and the ordered list of methods that were active on the call stack before 
+%such exception has occurred ~\cite{gosling2000java,bloch2008effective}.
+%and unintended handling~\cite{miller1997issues} (i.e., exceptions being mistakenly caught by existing handlers)
+
 \section{Introduction}
@@
 Along the recent years, we have witnessed an astonishing increase in the number 
-mobile applications. Such increase is motivated by pervasiveness of the World Wide Web 
-and the unprecedented increase of global smartphone shipments -
-the mobile industry shipped over 315 million smartphones in the last quarter of 2013 ~\cite{googleio}.
-
-In this context, Android has become a leading platform - thanks to its open-source 
-model and hardware partners like Samsung, HTC, Motorola, Asus (besides Nexus Google's
- own device) ~\cite{gartner}.  Google estimates that there is over 1 billion Android active users per month ~\cite{googleio},
-and the number of apps built on top of Android increases in a daily rate.
-
-If on the one hand such applications extend phones capabilities 
+mobile applications. If on the one hand such applications extend phones capabilities 
 far beyond of the basic calls and textual messages, on the other hand
 they have to face an increasing number of threats to application robustness
- arising from: faults in underlying middleware and hardware, 
-faults on third party services and libraries; compatibility issues; 
-faults on connected devices (e.g. camera, GPS); memory and 
-battery restrictions; among others.
+ arising from: faults on underlying middleware and hardware (e.g., GPS receiver, camera, sensors);
+faults on third party services and libraries; compatibility issues ~\cite{McDon13}; 
+ memory and battery restrictions~\cite{Zhang12}; concurrency issues ~\cite{ama2012}; among others.
 
 Therefore, techniques for error detection and handling are not  an optional add-on but a 
-fundamental part of such apps.
-% as any other modern system~\cite{bruntink2006discovering}. 
-The Android platform uses the exception handling mechanism embedded
- in Java to deal with such conditions. Although the exception handling mechanism
-~\cite{goodenough1975exception} is one of the most used schemes for
-detecting and recovering from such exceptional conditions, studies (e.g.~\cite{miller1997issues,Robil00,shah2010understanding, garcia2007extracting,garcia2001comparative,cabral2007exception,coelho2011unveiling}),
-have shown that it is generally poorly understood and the less tested parts of the system ~\cite{coelho2011unveiling}.
+fundamental part of such apps. The exception handling mechanism ~\cite{goodenough1975exception},
+embedded in many mainstream programming languages, such as Java, C++ and C\#,
+ is one of the most used techniques for detecting and recovering from such exceptional conditions. 
+Android, the leadling plaform for mobile applications~\cite{gartner} (thanks to its open-source 
+model and hardware partners like Samsung, HTC, Motorola, Asus besides Nexus Google's
+ own device)  reuses the embedded Java Exception model to detect and handle 
+ exceptional conditions.
+
+Studies have shown, however, that the exception-related code is generally poorly understood and the
+ less tested parts of the system~\cite{miller1997issues,Robil00,shah2010understanding, 
+garcia2007extracting,garcia2001comparative,cabral2007exception,coelho2011unveiling}.
 As a consequence they may affect the system the other way around.
-The exception-related code may favor the introduction of failures such as uncaught exceptions ~\cite{jo2004uncaught, zhang2012amplifying}
-and unintended handling ~/cite{miller1997issues} (i.e., exceptions being mistakenly caught by existing handlers) 
-which make the system even less robust.
-
-Our conjecture is that several Android applications may be suffering from common 
-fragilities on the exception-related code (of the applications themselves or 
-the underlaying platform) which may be degratind their robustness.
-We call ``fragilities" characteristics on the exception-related code that favor the introduction
-of aforementioned failures.
+The exception-related code may favor the introduction of failures such as 
+uncaught exceptions~\cite{jo2004uncaught, zhang2012amplifying}
+which make the system even less robust~\cite{coelho2011unveiling}.
 
@@
 publicly available in repository hosting sites such as GitHub and Google Code.
-
-When available, the exception stack traces provide a useful source of information about system crashes ~\cite{bettenburg2008makes} which 
- can enable different kinds of post-mortem analysis and support:  debugging~\cite{schroter2010stack}, 
- bug classification and clustering ~\cite{wang2013improving, kim2011crash, dhaliwal2011classifying},  
-automated bug fixing ~\cite{sinha2009fault} and fault-proneness prediction models ~\cite{kim2013predicting}.
+Developers and users of open-source projects hosted these sites 
+constantly uses the provided infra-strucute to share information across 
+the software development (including exception stack traces detected).
+
+%When available, the exception stack traces provide a useful source of information about system crashes ~\cite{bettenburg2008makes} which 
+% can enable different kinds of post-mortem analysis and support:  debugging~\cite{schroter2010stack}, 
+% bug classification and clustering~\cite{wang2013improving, kim2011crash, dhaliwal2011classifying},  
+%automated bug fixing~\cite{sinha2009fault} and fault-proneness prediction models~\cite{kim2013predicting}.
 
@@
     \hline
-    \bfseries{Root Exception} &  \multicolumn{2}{c}{\bfseries{Projects (GH\&CG)}} &  \multicolumn{2}{c}{\bfseries{Occurrences (GH\&CG)}} & \textsf{Android} & \textsf{Libcore} & \textsf{App} & \textsf{Lib} \\
+    \bfseries{Root Exception} &  \multicolumn{2}{c}{\bfseries{Projects}} &  \multicolumn{2}{c}{\bfseries{Occurrences}} & \textsf{Android} & \textsf{Libcore} & \textsf{App} & \textsf{Lib} \\
     & \bfseries{\#} &  \bfseries{\%} & \bfseries{\# } & \bfseries{\% } &&&&\\
@@
 java.util.ConcurrentModificationException	&	43	&	6,73\%	&	65	&	1,08\%	&	5	&	0	&	46	&	13	\\
-\\
-
     \hline
@@
 trace on which the NullPoiterException was the root cause.
+
  The NullPointerExceptions are mainly signaled inside the applicaiton code and the Android platform,
- although we also find NullPointerExceptions being signaled from third-party libraries. Regarding
-reusable code, there is no consensus whether it is a good or bad practice to 
+ although we could also find NullPointerExceptions being signaled by third-party libraries. 
+Regarding reusable code, there is no consensus whether it is a good or bad practice to 
 re-throw a NullPointerException. Some prefer to encapsulate such an exception on
@@
 sample of 25 exception types (from the total 100) using the same list of concerns;
-the interater agreement was high (96\%). We could also observe that
-We can therefore assume that the categories are descriptive 
-enough for covering the range of exception types analyzed.
+the resulting interrater agreement was high (96\%). We could also observe that
+We can therefore assume that the exception names and Javadoc 
+were enough for classifying the exception types analyzed.
 
@@
       Programming Mistakes &  52,0\%\\ 
-      %Resources Restriction   &   23,9\% \\ 
-      General Exceptions (Error, Exception, Runtime)    &  6,7\%\\
+      Memory Restriction   &   4,1\% \\ 
+      General Exceptions   &  6,7\%\\
     \hline
@@
 A multitude of programming mistakes were reported as the main causes of exception stack traces.
-51,96\% of all projects reported at least one exception stack trace on which the NullPoiterException
+51,96\% of all projects reported at least one exception
+ stack trace on which the NullPoiterException
 was the root cause.
@@
 envolve different exception types (i.e., Error, Checked, Exception). We call 
-such a wrapping a ``cross-type wrapping".  Such wrappings reveals 
-violations to best practices I and II (e.g. Error wrapping a Checked exception),
-To illustrate the problem behind such wrappings, we can use one of the wrappings found in this study. 
-When the developer is confronted with a checked exception, the designer of the API is telling him 
-to handle the exceptional condition, however we could observe that in some cases the 
-checked exception wrapped an OutOfMemoryError, which represent resource deficiency detected 
-by the runtime environment which make impossible the program to recover. 
-Hence, trying to handle this exception may lead the program to an unpredictable state.
-
-Next we discuss and exemplify the cross-type wrappings
- found in this study.
+such a wrapping a ``cross-type wrapping".  Some of these wrappings may reveal 
+violations to best practices I and II (e.g. Error wrapping a Checked exception) as discussed next.
 
@@
 
+\noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
+An unexpected fragility was detected: a checked exception being thown and not specified on 
+on the ``exception interface" of the signaler methods.
+}
+\end{minipage}}
+
 \emph{\textbf{Runtime Exception wrapping Error.}} We could find  64 exception stack traces 
@@
 \emph{\textbf{Checked wrapping Runtime Exception.}} Manually inspecting the stack 
-traces we could observe that most of the wrappings were performed 
+traces we could observe that such wrappings were performed either 
 by RoboSpice (an android library that supports the creation of asynchronous long running 
-tasks, which has 1.682 stars and 367 forks in GitHub) and by the reflection Java library (used by Android, 
-Application and library classes). Inspecting the code related to such wrappings we could observe that
-the reflection library convert every exception that happens 
-when invoking a reflected method into an InvoncationTargetException (this conversion is 
-performed inside native code).
+tasks, which has 1.682 stars and 367 forks in GitHub) or by the reflection Java library (used by
+methods defined in Android, Application and library classes).
 
 \emph{\textbf{Checked Exception wrapping Error.}} Almost all of these wrappings
- were also caused by the reflection library 
-used by Android applications. The methods responsible for the wrappings
+ were also caused by the reflection library used by Android applications. The methods responsible for the wrappings
 were also native methods written in C. Table~\ref{tab:exampleswrap} illustrates
@@
 
-\emph{\textbf{Multiple Wrappings.}} We could also observe that in 205 stacks found
- in Android repository the exception was wrapped  more then once. One interesting 
-multiple-wrapping found in Android repo was the 
-following: Runtime-Checked-Runtime-Checked-Runtime-Checked-Runtime.
-
-AS WE COULD OBSERVE...
+We could also observe that some stack traces included several intermediate cross-type wrappings.
+One interesting example was the following: Runtime-Checked-Runtime-Checked-Runtime-Checked-Runtime.
+Although some of these wrappings may be a result of design decision
+the mis-use of exception wrapppings may make the exception handling 
+code more complex (e.g., the multiple wrappings) and error-prone,
+ and lead to what we call ``exception confusion problem"
+To illustrate this problem we can use one of the wrappings discussed above.
+When the developer is confronted with a checked exception, the designer of the API is telling him 
+to handle the exceptional condition. However we could observe that in some cases the 
+checked exception wrapped an OutOfMemoryError, which represent resource deficiency detected 
+by the runtime environment which make impossible the program to recover. 
+Hence, trying to handle this exception may lead the program to an unpredictable state.
 
 \noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
-MUDARRRR Odd wrappings were detected (e.g., a checked exception wrapping an OutOfMemoryError).
-The study also detected an undocumented checked exception signaled by a native method defined in Android.
+Some of the detected cross-type exception wrappings points to a potential ``exception confusion problem".
+To illustrate: a checked exception wrapping an OutOfMemoryError; and the multiple wrapping Runtime-Checked-Runtime-Checked-Runtime-Checked-Runtime.
 }
@@
 
-\begin{figure*} \centering \includegraphics[scale=0.55]{codeexamples.pdf}
+\begin{figure*} \centering \includegraphics[scale=0.55]{examples_new2.png}
 \caption{Code snippets illustrating fragilities on the exception-related code} \label{fig:snippets} \end{figure*}
@@
  more complex and lead to what we call an \emph{exception handling confusion problem}.
-Such problem can lead the program to an unpredictable state in the presence of exceptions. 
-To illustrate this problem, we can use one of the examples found in this study: when the developer is 
-confronted with a checked exception, the designer of the API is telling him 
-to handle the exceptional condition, however we could observe that in some cases the 
-checked exception wrapped an OutOfMemoryError, which represent resource deficiency detected 
-by the runtime environment which make impossible the program to recover.
-
+Such problem can lead the program to an unpredictable state in the presence of exceptions,
+as illustrated by the scenario on which a checked exception wrapped an OutOfMemoryError. 
 Currently there is no way of enforcing Java exception type conventions during program development.
@@
 \label{sec:threats}
+
 \noindent\emph{Internal Validity} We used a heuristics-based parser to mine
@@
 exceptions or stack traces. Our limited manual inspection did not reveal such
-cases.
-
-\noindent\emph{External Validity} The results presented here were based on mining
+cases. Moreover, in this study we manually classified the exceptions 
+related to programming mistakes. To ensure the quality of the analysis, 
+we calculated the interrater agreement after three independent 
+coders classified a randomly selected sample (of 25 exception 
+types from the total 100); the interater agreement was high (96\%).
+
+Another threat relates to the fact that parts of our analysis 
+are based on the availability of stack traces on issues on Github and Googlecode. 
+In using these datasets, we make an underlying assumption: the stack traces reported on issues are 
+representative of valid crash information of the applications. 
+One way to mitigate such threat would be to access to the full 
+set of crash data per application. Although some services exist 
+to collect crash data from mobile applications ~\cite{BugSe14,BugSn14,Googl14,Acra14},
+they do not provide open access to the crash reports of their client applications.
+In our study, we mitigated this threat by manually inspecting
+the source code associated to a subset of reported exception stack traces.
+This subset contained the stack traces related to the main findings 
+of the study (e.g., `ùndocumented runtime and checked exceptions",
+and cross-type wrappings"). Through the manual inspection
+ we could map the stack traces information to real characteristics 
+of the exception-related code of analyzed apps.
+
+\noindent\emph{External Validity} First, we limited out analysis to open-source
+Android apps. We are aware that the reported exceptions for commercial
+apps can be different from the ones found in this study. However, 
+in this study revealed that several reported exception stack traces 
+were reporting root exceptions signaled by Android utility code and libraries. 
+Moreover, since the ``fragilities" observed in this study are due to
+characteristics of Java exception model, they impose challenges 
+the robustness not ony to Android apps but also to other systems
+ based on Java exception model.
+
+Second, the results presented here were based on mining
 issues on Github through the GHTorrent dataset. While comprehensive and
@@
 our sample.
+
 \footnote{to do: OLHAR ARTIGO DO FSE QUE TRABALHA COM OPENSOURCE.
@@
 
-\noindent\emph{Construct Validity} Parts of our analysis are based on the availability of stack traces on issues on
-Github. In using this dataset, we make an underlying assumption: the
-stack traces reported on issues are representative of real crashes in
-the applications. The assumption is impossible to mitigate without access to
-the full set of crash data per application. Some services exist to collect 
-crash data from mobile applications ~\cite{BugSe14,BugSn14,Googl14,Acra14}.
-However, such services do not provide open access to the crash reports
-of their client applications. Having access to such a dataset would allow 
-a replication of our analysis on an even wider setting.
-
+\
 \section{Related Work}
@@
 \end{document}
-
-http://docs.oracle.com/javase/7/docs/api/
-
-%JAVA.LANG EXCEPTIONS
-ArithmeticException
-ArrayIndexOutOfBoundsException
-ArrayStoreException
-ClassCastException
-ClassNotFoundException
-CloneNotSupportedException
-EnumConstantNotPresentException
-Exception
-IllegalAccessException
-IllegalArgumentException
-IllegalMonitorStateException
-IllegalStateException
-IllegalThreadStateException
-IndexOutOfBoundsException
-InstantiationException
-InterruptedException
-NegativeArraySizeException
-NoSuchFieldException
-NoSuchMethodException
-NullPointerException
-NumberFormatException
-ReflectiveOperationException
-RuntimeException
-SecurityException
-StringIndexOutOfBoundsException
-TypeNotPresentException
-UnsupportedOperationException
-
-JAVA UTIL:
-ConcurrentModificationException
-DuplicateFormatFlagsException
-EmptyStackException
-FormatFlagsConversionMismatchException
-FormatterClosedException
-IllegalFormatCodePointException
-IllegalFormatConversionException
-IllegalFormatException
-IllegalFormatFlagsException
-IllegalFormatPrecisionException
-IllegalFormatWidthException
-IllformedLocaleException
-InputMismatchException
-InvalidPropertiesFormatException
-MissingFormatArgumentException
-MissingFormatWidthException
-MissingResourceException
-NoSuchElementException
-TooManyListenersException
-UnknownFormatConversionException
-UnknownFormatFlagsException

### 2014-09-05 android-stacks.tex
activity

@@
  signaled by  runtime environment (due to programming mistakes) since these exceptions 
-should not be documented on the method signature. Then we inspected the code each method 
-(defined either on Android or on libraries) signaling  runtime exceptions not-related to programming mistakes.
+should not be documented on the method signature.
+
+Then we inspected the code each method (defined either on Android  Aplication Framework or on libraries) 
+explicitly signaling a runtime exception not-related to programming mistakes.
 As a result of this inspection, we could observe that only 1 method out of 118 inspected methods
@@
 (e.g. libraries, or framework utility code). In such cases the client usually do not have access to 
-the source code, and in the absence of
-the exception documentation it is very difficult or even impossible for the client of such third party code to 
+the source code, and in the absence of the exception documentation it is very difficult or even impossible for the client of such third party code to 
 protect the application against undocumented runtime exceptions. As a consequence, the
@@
 
-%SELECT repo, *
-% FROM stacktracesummary where stack like '%getDeclaredMethods%';
-
 \noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
  Only 0,8\% of the methods documented the explicitly signaled runtime exceptions. 
+}
+\end{minipage}}
+
+During the manual inspection, besides inspecting the code of exception signalers we also
+inspected the intermediate methods reported on the stacks.  The inspected stack traces envolved, 
+in one way or another, methods from the  Android Application Framework: either framework 
+utility methods, or framework hook methods  (e.g., OnCreate(), onPause()).
+During this inspection we could observe that none of the hook methods provided by the Activity class 
+specify an exception interface.  The``exception interface" can stimulate
+can stimulate the developer (in this case, extending Activity class) to: think that 
+``something" may go wrong during the execution of such methods; and design the 
+exception handling code accordingly (the absence of such interface
+ violates to best practice IV).
+
+\noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
+ None of the Activity hook methods define an ``exception interface".
 }

### 2014-09-05 android-stacks.tex
intro

--- 2014-09-05 android-stacks.tex
@@
 extend phones capabilities far beyond of the basic calls, on the other hand they have to 
-cope with an increasing number of exceptinal conditions 
+cope with an increasing number of exceptional conditions 
 (e.g., faults in underlying middleware or hardware; memory and battery restrictions). 
@@
 (1) undocumented runtime exceptions thrown by third party code;
-(2) undocumented checked exceptions thrown by JNI interface;
+(2) undocumented checked exceptions thrown by a JNI interface function;
 (3) unexpected exception wrappings (e.g., Errors being wrapped in checked exceptions),
@@
 as the root cause of an exception stack trace.
-This study provide important insights on how such fragilities impose challenges 
-to application robustness not only in Android environment but also to other 
-systems based on Java exception model.
+This study provides insights on such fragilities and the robustness threats  
+they impose not only to Android apps but also to other systems 
+based on Java exception model.
 
@@
 
-Along the recent years, we have witnessed an astonishing increase in the number 
+Along the recent years, we have witnessed an astonishing increase in the number of
 mobile applications. If on the one hand such applications extend phones capabilities 
@@
 faults on third party services and libraries; compatibility issues ~\cite{McDon13}; 
- memory and battery restrictions~\cite{Zhang12}; concurrency issues ~\cite{ama2012}; among others.
+ memory and battery restrictions~\cite{Zhang12}; among others.
 
@@
  is one of the most used techniques for detecting and recovering from such exceptional conditions. 
-Android, the leadling plaform for mobile applications~\cite{gartner} (thanks to its open-source 
+Android, the leading plaform for mobile applications~\cite{gartner} (thanks to its open-source 
 model and hardware partners like Samsung, HTC, Motorola, Asus besides Nexus Google's
- own device)  reuses the embedded Java Exception model to detect and handle 
+ own device)  reuses the embedded Java exception model to detect and handle 
  exceptional conditions.
@@
 The exception-related code may favor the introduction of failures such as 
-uncaught exceptions~\cite{jo2004uncaught, zhang2012amplifying}
-which make the system even less robust~\cite{coelho2011unveiling}.
+uncaught exceptions~\cite{jo2004uncaught, zhang2012amplifying} - 
+which can lead to system crashes, making the system even less robust~\cite{coelho2011unveiling}.
 
@@
 publicly available in repository hosting sites such as GitHub and Google Code.
-Developers and users of open-source projects hosted these sites 
-constantly uses the provided infra-strucute to share information across 
-the software development (including exception stack traces detected).
+Developers and users of open-source projects uses such provided
+ infra-strucute to share information (e.g., exception stack traces~\cite{bettenburg2008makes,schroter2010stack}) across 
+the software development.
 
@@
 reported on Android projects hosted in GitHub and Google Code. 
-The goal of this study was to: investigate the whether the reported exception stack traces
- can reveal common fragilities on the exception-related code which may be favoring 
-the introduction of failures such as uncaught exceptions - and as a consequence
- contributing to deacrease the robustness of apps.
+The goal of this study was to: investigate whether the reported exception stack traces
+ can reveal common fragilities on the exception-related code.  
+We call ``fragilities" characteristics on the exception-related code 
+that favor the introduction of failures such as uncaught 
+exceptions -  contributing to decrease the applicaiton robustness.
 To guide this investigation we compiled general guidelines on how to use
@@
 
-   \item  Unexpected exception wrappings (e.g., Error wrapping checked and runtime
-    exceptions and vice-versa), indicating that the purpose of Java's hybrid exception model may not have been adequately used.
-
-   \item Undocumented runtime exceptions signaled by third party code (i.e., libraries and Android platform). 
-This study found 118 distinct methods defined on third party code and  reported as the root signalers of 
-runtime exceptions. Only 1 of such methods (0,8\%) documented the explicitly thrown runtime exception 
-in a Javadoc comment; and none of these methods included the exception as part of the exception 
-interface (i.e., using throws clause).
-
-   \item Undocumented checked exception signaled by native Android library. This finding points to a fragility 
-on the way exceptions flow from JNI interface functions, which can lead to uncaught exceptions that are difficult do 
-debug.
-
  \item  A multitude of programming mistakes were reported as the main causes of exception stack traces:
@@
 
+   \item  Cross-type exception wrappings (e.g., an OutOfMemoryError wrapped in a checked exception)
+were detected on stack traces. Indicating that the purpose of Java's hybrid exception model may not have been adequately used.
+
+   \item Undocumented runtime exceptions signaled by third-party code (i.e., libraries and Android platform). 
+    In this study, none of the third-party methods explicitly signaling runtime exceptions included such exceptions
+    on ``exception specifications" (i.e., a list of exceptions that a method might raise during its execution).
+
+   \item Undocumented checked exception signaled by a  JNI interface function defined on Android Platform. 
+    Some flows reported a checked exception thown by a JNI interface function, and  such exception was not  
+    part of the ``exception interface" of JNI function nor the methods on its invokation chain - it can lead to 
+   uncaught exceptions that are difficult to debug.
+
 \end{itemize}
 
-As we can see, although the study was based on a subset of Android projects, 
-the findings are not limited to this setting. The detected fragilities make difficult
-or sometimes infeasible for a Java developer to prevent failures such as 
-uncaught exceptions (e.g., thrown silently by third-party utility code), and 
-unintended handling ( e.g., when an Error is wrapped in a checked Exception).
-Thereore, they impose obstacles not only to the developement of robust Android apps, 
-but also to the developement of any robust Java-based system. 
-Hence, the study results are relevant to: (i) Android and 
-Java developers to familiarize common characteristics and fragilities of exception-related code 
-which is the first step to help developers preventing and dealing with them; and (ii) designers of languages and tools 
-to consider ways of reducing such fragilities in Android and Java environments.
-The main contributions of this study is twofold:
-\begin{itemize}
-
-  \item  It performs a large scale post mortem analysis of exception stack traces reported on Android projects hosted in GitHub and Google Code.
-  \item  It introduces ExceptionMiner, a tool developed to support exception stack trace mining and analysis.
-    analysis.
-
-\end{itemize}
-
-% AVD: We have not looked at private issues, and the tool is described above.
-%We strongly believe that, specially in the open-source environment,
-%faults are not to be hidden in a private bug issue. Faults should be shared and
-%discussed, so that a developer can learn from other projects mistakes.
-%Currently, the search facilities of repositories are very limited, and the
-%ExceptionMiner tool is a contribution in this direction.
+Such outcomes point to threats not only to the developement of robust Android apps, 
+but also to the developement of any robust Java-based system. For instance, in the absence of the 
+``exception specification" of  a third-party code, it is difficult or 
+even enfeasible for the developer to protect the code against ``unforeseen" exceptions. 
+Since, in such cases the client usually do not have access to the source code, the undocumented exceptions may 
+remain uncaught and lead to system crashes. Moreover, the cross-type exception wrappings 
+(e.g., OutOfMemoryError is wrapped in a checked Exception) 
+may lead to unintended handling problems - trying to handle an instance of OutOfMemoryError ``hiden" in 
+a checked exception may lead the program to an unpredictable state.
+
+Hence, the study results are relevant to Android and 
+Java developers who may underestimate how likely are the
+ exceptions related to programming mistakes, and how 
+ tricky it can be to prevent the uncaught exceptions from third-party code
+ in the absence of ``exception interfaces". Moreover, results are relevant to 
+designers of languages and tools to consider ways of reducing such fragilities 
+in Android and Java environments.
 
@@
 used to represent user-defined exceptional conditions.In fact, there is a long-lasting debate 
-about the pros and cons of both approaches ~\cite{javatut,stackoverlow,debate},
+about the pros and cons of both approaches~\cite{javatut,stackoverlow,debate},
 \footnote{152 questions in Stackoverlow are related to this debate}.
@@
  Looking for the handler on the invocation stack is claimed to increase software reusability, 
-since the invoker of an operation can handle it in a wider context ~/cite{miller1997issues}.
+since the invoker of an operation can handle it in a wider context ~\cite{miller1997issues}.
 
@@
 on the defect issues can be found at:
-\url{www.dimap.ufrn.br/~roberta/android_repo_mining}}.
+\url{https://github.com/souzacoelho/exceptionminer}}.
  Moreover, issues and pull requests are dual on Github; all pull requests have a corresponding 
@@
 
-\emph{\textbf{Exception Type Analysis}} To support a deeper investigation of the 
+\emph{\textbf{Exception Type Analysis.}} To support a deeper investigation of the 
 stack traces every exception defined on a stack trace need to be classified according to its type
@@
 \subsection{Replication Package}
-All the data used in this study will be publicly available at http://.....
+All the data used in this study will be publicly available at https://github.com/souzacoelho/exceptionminer
 Specifically we provide: (i) all issues related to Android projects found
@@
     \hline
-      Programming Mistakes &  52,0\%\\ 
-      Memory Restriction   &   4,1\% \\ 
-      General Exceptions   &  6,7\%\\
+      Programming Mistakes &  51,96\%\\ 
+      Memory Restriction   &   4,10\% \\ 
+      General Exceptions   &  6,70\%\\
     \hline
@@
 good practice (see best practice IV in Section~\ref{sec:best}) because by using checked exceptions a library can define a precise 
-exception interface ~/cite{miller1997issues} to its clients.
+exception interface ~\cite{miller1997issues} to its clients.
  Since such libraries are heavily used  in several projects, such finding can be related to the library's maturity.
@@
 Such undocumented runtime exceptions represent a threat to system robustness, specially
-when such exceptions are thrown then the third party code is invoked inside the application 
+when such exceptions are thrown then the third party code invoked inside the application 
 (e.g. libraries, or framework utility code). In such cases the client usually do not have access to 
-the source code, and in the absence of
-the exception documentation it is very difficult or even impossible for the client of such third party code to 
-protect the application against undocumented runtime exceptions. As a consequence, the
+the source code. Hence in the absence of the exception documentation it is very difficult or even impossible
+ for the client to design the application to deal with ``unforeseen runtime exceptions. As a consequence, the
  undocumented runtime exception may remain uncaught and lead to system crashes.
-
-%SELECT repo, *
-% FROM stacktracesummary where stack like '%getDeclaredMethods%';
 
@@
 \subsection{What can the Exception Wrappings tell us?}
-
 %SELECT COUNT(*) FROM STACKTRACESUMMARY WHERE
@@
 In this study we could observe undocumented runtime exceptions thrown by third party code,
-and even undocumented checked exception thrown by Java supporting library.
+and even undocumented checked exception thrown by a JNI interface.
 Such undocumented exceptions make difficult, and most of the times infeasible
@@
 
-\noindent\emph{Internal Validity} We used a heuristics-based parser to mine
+\noindent\emph{Internal Validity.} We used a heuristics-based parser to mine
 exceptions from issues.  Our parsing strategy was conservative by default; for
@@
 In our study, we mitigated this threat by manually inspecting
-the source code associated to a subset of reported exception stack traces.
-This subset contained the stack traces related to the main findings 
+the source code associated to a subset of the reported exception stack traces.
+Such subset comprise the stack traces related to the main findings 
 of the study (e.g., `ùndocumented runtime and checked exceptions",
-and cross-type wrappings"). Through the manual inspection
- we could map the stack traces information to real characteristics 
-of the exception-related code of analyzed apps.
-
-\noindent\emph{External Validity} First, we limited out analysis to open-source
-Android apps. We are aware that the reported exceptions for commercial
-apps can be different from the ones found in this study. However, 
-in this study revealed that several reported exception stack traces 
-were reporting root exceptions signaled by Android utility code and libraries. 
-Moreover, since the ``fragilities" observed in this study are due to
-characteristics of Java exception model, they impose challenges 
-the robustness not ony to Android apps but also to other systems
- based on Java exception model.
-
-Second, the results presented here were based on mining
-issues on Github through the GHTorrent dataset. While comprehensive and
-extensive, GHTorrent is not an exact replica of Github, so several issues might
-be left out. Due to the way data collection works with GHTorrent, for projects
-that are relatively inactive, GHTorrent might have not collected a significant
-proportion of their issues. We did not investigate the extend of this threat on
-our sample.
-
-\footnote{to do: OLHAR ARTIGO DO FSE QUE TRABALHA COM OPENSOURCE.
-FALAR QUE DESCOBERTAS FORAM DE LIBS JAVA. APESAR DO ESTUDO NAO SER FEITO EM PROJS JAVA.....
-PODEMOS CONCLUIR ALGUMAS COISAS QUE SE APLICAM A TODOS JAVA}
-
-\
+and cross-type wrappings").
+
+\noindent\emph{External Validity.} We limited out analysis to a subset of
+existing open-source Android apps. We are aware that the reported exceptions 
+for commercial apps can be different from the ones found in this study, and that
+this subset is a small percentage of existing apps.
+Such threats to validity are similar to the ones imposed on the other empirical studies 
+which also used free or open-source Android apps ~\cite{Linar13,McDon13,ahimed}.
+Moreover,  several exception stack traces that support the findings of this study
+refered to exceptions comming from methods defined on Android Application Framework
+and third-party libraries.  Additionaly,  the ``fragilities" observed in this study are due to
+characteristics of Java exception model, which can impose challenges 
+the robustness of not ony to Android apps but also to other systems
+ based on the same exception model.
+
 \section{Related Work}
@@
 
-\textit{Tools Support for Detecting Faults on Java Exception-Related Code.}
-
-Manually analyzing the exception-related code and looking for defects, can easily 
-become infeasible ~\cite{Robil00}. Hence, some tools have been proposed to detect faults 
+\textit{Detecting Faults on Java Exception-Related Code.}  Some tools have been 
+proposed to detect faults  
 on the exception handling code either statically (cabral, roberta,nelio,chang) 
@@
 
-Robillard and Murphy~\cite{Robil00}
-employed dataflow analysis to find the propagation paths of checked and
+Robillard and Murphy~\cite{Robil00} employed dataflow analysis to find the propagation paths of checked and
 unchecked exception types. Modeled after Robillard and Murphy's work, other
@@
 (ii) undocumented runtime exceptions thrown by third party code,
-(iii) undocumented checked exceptions thrown by JNI native code,
+(iii) undocumented checked exceptions thrown by JNI interface,
 (iv) unexpected exception wrappings (e.g., Errors being wrapped in checked exceptions) 

### 2014-09-05 android-stacks.tex
related

@@
 
-\textit{Detecting Faults on Java Exception-Related Code.}  Some tools have been 
-proposed to detect faults  
-on the exception handling code either statically (cabral, roberta,nelio,chang) 
-or dynamicaly through testing (roberta). Both approaches, however, have inherent
- limitations. On the one hand, the static analysis tools needs to deal with the false 
-positives and cannot adequately deal with dynamically loaded code (e.g. reflection  
-[ref]). On the other hand the testing tools are llimited to the fact that the developer
- needs to create a test to exercise every exception, which is also costly.... (olhar tese). 
-Moreover, both methods fail to account for exceptions implicitly signaled by the runtime 
-environment (e.g. due to resource depletion which can be signaled in every statement invocation).
-
-Robillard and Murphy~\cite{Robil00} employed dataflow analysis to find the propagation paths of checked and
-unchecked exception types. Modeled after Robillard and Murphy's work, other
-tools have been proposed to support the static analysis of exception
-flows~\cite{coelho2008assessing}. The main limitation of all
-static analysis tools is the number of false positives inherent to static analysis
-solutions, which can lead to a high number of exception flows, specially if
-considering Java Environment exceptions and exceptions signaled from libraries.
-
-Additionally, Cabral and Marques~\cite{cabral2007exception} analyzed the
-exception handling code of 32 open-source systems, both for Java and .NET. They
+\textit{Detecting Faults on Java Exception-Related Code.} 
+Cabral and Marques~\cite{cabral2007exception} analyzed the
+source-code of 32 open-source systems, both for Java and .NET. They
 observed that the action handlers were very simple (e.g., logging and present a
@@
 and robustness, based on their own experience with Java enterprise applications.
-Our work differs from those two works as we tried to identify bad practices from
-the analysis of stack traces extracted from issues in combination with and bytecode and source
-code analysis. \footnote{ to do: TODORESSALTAR LIMITACAO DE QUANTIDADE DE FALSOS POSITIVOS... E QUE OS STACKS SAO REAIS...
-REPRESENTAM FLUXOS REAIS.... POST MOSTERN ANALISYS OF REAL EXECUTIONS....}
+Our work differs from those
+
+Some tools ~\cite{Robil00}~\cite{coelho2008assessing} have ben proposed 
+to employ dataflow analysis to find the propagation paths of checked and unchecked 
+exception types. The main limitation of all static analysis tools is the number of false 
+positives inherent to static analysis solutions, which can lead to extremely high number 
+of exception flows, specially if considering Java Environment exceptions and exceptions 
+signaled from libraries. Moreover, such also have limited support some libraries (e.g., Java Reflection).
+Our owrk
 

### 2014-09-05 android-stacks.tex
intro 2

@@
 types from the total 100); the interater agreement was high (96\%). 
-
 Another threat relates to the fact that parts of our analysis 
@@
 
-\noindent\emph{External Validity.} We limited out analysis to a subset of
-existing open-source Android apps. We are aware that the reported exceptions 
+\noindent\emph{External Validity.} Our work use GHTorrent dataset. Although 
+comprehensive and extensive is not an exact replica of Github. 
+However, the results of this study does not depent on the analysis of
+a complete Github dataset. Instead, we limited out analysis to a subset of
+existing open-source Android projects availabe on GitHub and Googlecode.
+Moreover, we are aware that the reported exceptions 
 for commercial apps can be different from the ones found in this study, and that
@@
 source-code of 32 open-source systems, both for Java and .NET. They
-observed that the action handlers were very simple (e.g., logging and present a
+observed that the actions inside handlers were very simple (e.g., logging and present a
 message to the user). Reimer and Srinivasan~\cite{reimer2003analyzing} listed a
 set of bad practices on exception handling that hinder software maintainability
-and robustness, based on their own experience with Java enterprise applications.
+and robustness, based on their experience with Java enterprise applications.
 Our work differs from those

### 2014-09-05 android-stacks.tex
threat

@@
 \emph{\textbf{The prevalence of programming mistakes on stack traces.}} One way of getting a broader view of the prevalence~\footnote{The prevalence evaluates the proportion of a given characteristic over a group of subjects.}  of programming mistakes on reported stack traces would be to manually inspect the code responsible for throwning every exception reported  on issues.
-However, inspecting the code related to 6005 stack traces can easily become infeasible.
+However, inspecting the source code related to 6005 stack traces can easily become infeasible.
 Hence, we conducted an empirical investigation based on the following assumption: the name of the exception and its associated Javadoc 
@@
 
-\emph{\textbf{Error wrapping Runtime.}} After manually inspecting the code associated to such wrappings 
+\emph{\textbf{Error wrapping Runtime.}} After manually inspecting the source code associated to such wrappings 
 we discovered that all of them were caused by static initializers. If any exception is thrown in the context of a static initializer (i.e., static block) 
@@
 One interesting example was the following: Runtime-Checked-Runtime-Checked-Runtime-Checked-Runtime.
-Although some of these wrappings may be a result of design decision
+
+Although some of these wrappings may be a result of design decisions,
 the mis-use of exception wrapppings may make the exception handling 
@@
  and lead to what we call ``exception confusion problem"
+
 To illustrate this problem we can use one of the wrappings discussed above.
 When the developer is confronted with a checked exception, the designer of the API is telling him 
-to handle the exceptional condition. However we could observe that in some cases the 
-checked exception wrapped an OutOfMemoryError, which represent resource deficiency detected 
-by the runtime environment which make impossible the program to recover. 
-Hence, trying to handle this exception may lead the program to an unpredictable state.
+to handle the exceptional condition (according to Java Specification and 
+best practices). However, if the developer tries to handle such exception,
+which is wrapping (or ``hiding") an OutOfMemoryError, such handling 
+may lead the program to an unpredictable state.
+
+The cross-type wrappings show us that even if the developer sticks to Java
+specification and best practices, it will be dificult for her/him to prevent the code from 
+incurring into a situation as the one described above.
 
@@
 comprehensive and extensive is not an exact replica of Github. 
-However, the results of this study does not depent on the analysis of
-a complete Github dataset. Instead, we limited out analysis to a subset of
-existing open-source Android projects availabe on GitHub and Googlecode.
-Moreover, we are aware that the reported exceptions 
+However, the result of this study does not depent on the analysis of
+a complete Github dataset. Instead, the goal of our study was to 
+pin-point fragilities on the exception-related code based on 
+exception stack trace mining of a subset of projects.
+We limited out analysis to a subset of existing open-source Android projects.
+We are aware that the exception stack traces reported 
 for commercial apps can be different from the ones found in this study, and that
 this subset is a small percentage of existing apps.
-Such threats to validity are similar to the ones imposed on the other empirical studies 
+Such threats are similar to the ones imposed to other empirical studies 
 which also used free or open-source Android apps ~\cite{Linar13,McDon13,ahimed}.

### 2014-09-05 android-stacks.tex
square content

@@
 \begin{abstract}
-The number of mobile apps is increasing in a daily rate. If on the one hand they
-extend phones capabilities far beyond of the basic calls, on the other hand they have to 
-cope with an increasing number of exceptional conditions 
-(e.g., faults in underlying middleware or hardware; memory and battery restrictions). 
-Therefore, mechanisms for exception detection and handling are not an optional add-on but a 
-fundamental part of such apps. However, studies have shown that such mechanisms are the less 
-understood and tested parts of the system. As a consequence they may affect the system the 
-other way around, making the system even less robust, leading to failures such as uncaught exceptions 
-and unintended handling (i.e., exceptions being mistakenly caught by existing handlers).
-This paper reports a study which mined the stack traces embedded on 
-159,048 issues reported on Android projects hosted in GitHub (482 projs.) 
-and Google Code (157 projs.). 
-The goal of this study was to: investigate the whether the exception stack traces
+%The number of mobile apps is increasing in a daily rate. If on the one hand they
+%extend phones capabilities far beyond of the basic calls, on the other hand they have to 
+%cope with an increasing number of exceptional conditions 
+%(e.g., faults in underlying middleware or hardware; memory and battery restrictions). 
+%Therefore, mechanisms for exception detection and handling are not an optional add-on but a 
+%fundamental part of such apps. However, studies have shown that such mechanisms are the less 
+%understood and tested parts of the system. As a consequence they may affect the system the 
+%other way around, making the system even less robust, leading to failures such as uncaught exceptions 
+%and unintended handling (i.e., exceptions being mistakenly caught by existing handlers).
+
+This paper reports a study which mined the Java stack traces included in
+159,048 issues reported on Android projects hosted in GitHub (482 projects) 
+and Google Code (157 projsects). 
+The goal of this study is to investigate  whether the exception handling logic on these stack traces
  can reveal common characteristics on the exception-related code that
-may be favoring the introduction of the aforementioned failures, which would be responsible for 
-decreasing the application robustness. We call such characteristics ``fragilities" of the exception-related code.
+are particularly fault prone and may lead to a decrease in application robustness.
 Overall 6,005 exception stack traces were extracted, and analyzed in 
 combination with source code and bytecode analysis. 
-Some fragilities consistently detected in this study include: 
+Key problems identified in this study include: 
 (1) undocumented runtime exceptions thrown by third party code;
 (2) undocumented checked exceptions thrown by a JNI interface function;
-(3) unexpected exception wrappings (e.g., Errors being wrapped in checked exceptions),
-revealing that Java hybrid exception model is not fully used according to its purpose.
-Moreover, about 50\% of all analyzed projects reported at least one NullPointerExeption
-as the root cause of an exception stack trace.
+(3) unexpected exception wrappings (such as Errors being wrapped in checked exceptions),
+indicating that Java's hybrid exception model is not fully used in accordance with its purpose.
+Moreover, about 50\% of all analyzed projects reported at least one NullPointerException
+as the root cause of an exception stack trace, suggesting the presence of programming mistakes.
 This study provides insights on such fragilities and the robustness threats  
@@
 %All: 159,048
+
+%Total page numbers then:
+%intro:       1
+%background:  1.5
+%method:      1
+%tool         0.5
+%findings     4
+%discussion   1
+%threats      .5
+%conc         .5
+%------------------+
+%total        10 pages
 
@@
 
-Along the recent years, we have witnessed an astonishing increase in the number of
-mobile applications. If on the one hand such applications extend phones capabilities 
-far beyond of the basic calls and textual messages, on the other hand
-they have to face an increasing number of threats to application robustness
- arising from: faults on underlying middleware and hardware (e.g., GPS receiver, camera, sensors);
-faults on third party services and libraries; compatibility issues ~\cite{McDon13}; 
- memory and battery restrictions~\cite{Zhang12}; among others.
+In recent years, we have witnessed an astonishing increase in the number of
+mobile applications. These applications extend phones capabilities 
+far beyond basic calls and textual messages. Such applications, however,
+have to face an increasing number of threats to application robustness
+ arising from: failures in underlying middleware and hardware (e.g., camera, sensors);
+failures in third party services and libraries; compatibility issues~\cite{McDon13}; 
+ memory and battery restrictions; and noisy external resources ~\cite{Zhang12} (e.g., wireless connections,GPS, bluetooth).
 
 Therefore, techniques for error detection and handling are not  an optional add-on but a 
-fundamental part of such apps. The exception handling mechanism ~\cite{goodenough1975exception},
+fundamental part of such apps. The exception handling mechanism~\cite{goodenough1975exception},
 embedded in many mainstream programming languages, such as Java, C++ and C\#,
  is one of the most used techniques for detecting and recovering from such exceptional conditions.
-Android, the leading plaform for mobile applications~\cite{gartner} (thanks to its open-source 
-model and hardware partners like Samsung, HTC, Motorola, Asus besides Nexus Google's
- own device)  reuses the embedded Java exception model to detect and handle 
- exceptional conditions.
-
-Studies have shown, however, that the exception-related code is generally poorly understood and the
- less tested parts of the system~\cite{miller1997issues,Robil00,shah2010understanding, 
+In this paper we will be concerned with exception handling in Android apps,
+which reuse Java's exception handling model.
+
+%Android, the leading plaform for mobile applications~\cite{gartner} (thanks to its open-source 
+%model and hardware partners like Samsung, HTC, Motorola, Asus besides Nexus Google's
+% own device)  reuses the embedded Java exception model to detect and handle 
+% exceptional conditions.
+
+Studies have shown that exception-related code is generally poorly understood and among the least tested parts of the system~\cite{miller1997issues,Robil00,shah2010understanding, 
 garcia2007extracting,garcia2001comparative,cabral2007exception,coelho2011unveiling}.
-As a consequence they may affect the system the other way around.
-The exception-related code may favor the introduction of failures such as 
-uncaught exceptions~\cite{jo2004uncaught, zhang2012amplifying} - 
+As a consequence they may inadvertently negatively affect the system: Exception-related code may introduce failures such as 
+uncaught exceptions~\cite{jo2004uncaught, Zhang12} - 
 which can lead to system crashes, making the system even less robust~\cite{coelho2011unveiling}.
@@
 
-Until recent years, the availability of data, reporting such failures, was scarce. 
-It comprised only few large scale projects (e.g. the Mozilla crash report dataset\footnote{https://code.google.com/p/promisedata/issues/detail?id=88}).
-Recently we have observed a rapid increase in the number of open
-source projects that make their version control systems as well as the issue tracking systems
-publicly available in repository hosting sites such as GitHub and Google Code.
-Developers and users of open-source projects uses such provided
- infra-strucute to share information (e.g., exception stack traces~\cite{bettenburg2008makes,schroter2010stack}) across 
-the software development.
+%% I am not so sure this data was not available. Let's omit this paragraph
+%Until recently, the availability of data reporting such failures was scarce. 
+%It comprised only few large scale projects (e.g. the Mozilla crash report dataset\footnote{https://code.google.com/p/promisedata/issues/detail?id=88}).
+%Recently we have observed a rapid increase in the number of open
+%source projects that make their version control systems as well as the issue tracking systems
+%publicly available in repository hosting sites such as GitHub and Google Code.
+%Developers and users of open-source projects uses such provided
+% infrastrucute to share information (e.g., exception stack %traces~\cite{bettenburg2008makes,schroter2010stack}) across 
+%the software development.
 
@@
 
-This study performs a post mortem analysis of the exception stack traces embedded on issues
+This study performs a post mortem analysis of the exception stack traces 
+included in issues
 reported on Android projects hosted in GitHub and Google Code. 
-The goal of this study was to: investigate whether the reported exception stack traces
- can reveal common fragilities on the exception-related code.  
-We call ``fragilities" characteristics on the exception-related code 
-that favor the introduction of failures such as uncaught 
-exceptions -  contributing to decrease the applicaiton robustness.
+The goal of this study is to investigate whether the reported exception stack traces
+ can reveal common fragilities in the exception-related code.  
+We call ``fragilities'' characteristics of the exception-related code 
+that increase the likelihood of introducing failures such as uncaught 
+exceptions -  leading to a decrease in application robustness.
+
 To guide this investigation we compiled general guidelines on how to use
@@
 Overall 159,048 issues were analyzed and 6,005 stack traces were extracted from them.
-The exception stack trace analysis was augmented with additional information
-extracted from JDK, Android Plaform and the applications using byte code 
-and source code analysis. Manual inspection was used to leverage the understanding 
-of exception stack traces and support further discussions and insights.
-
-Some outcomes consistently detected during this study were the following:
+The exception stack trace analysis was augmented by means of byte code 
+and source code analysis
+ with additional exception-related information
+extracted from the Android platform and Android applications.
+
+Outcomes of our study include:
+
+%Manual inspection was used to leverage the understanding 
+%of exception stack traces and support further discussions and insights.
+
+% Some outcomes consistently detected during this study were the following:
 
@@
 
- \item  A multitude of programming mistakes were reported as the main causes of exception stack traces:
- 27,71\% of the analyzed exception stack traces reported a java.lang.NullPointerException the root cause;
-and  51,96\% of all projects reported at least one exception stack trace on which the NullPoiterException
-was the root cause.
-
-   \item  Cross-type exception wrappings (e.g., an OutOfMemoryError wrapped in a checked exception)
-were detected on stack traces. Indicating that the purpose of Java's hybrid exception model may not have been adequately used.
-
-   \item Undocumented runtime exceptions signaled by third-party code (i.e., libraries and Android platform). 
-    In this study, none of the third-party methods explicitly signaling runtime exceptions included such exceptions
-    on ``exception specifications" (i.e., a list of exceptions that a method might raise during its execution).
-
-   \item Undocumented checked exception signaled by a  JNI interface function defined on Android Platform. 
-    Some flows reported a checked exception thown by a JNI interface function, and  such exception was not  
-    part of the ``exception interface" of JNI function nor the methods on its invokation chain - it can lead to 
+       \item Approximately 52\% of the reported stack traces can be attributed to programming mistakes. 
+In particular, 27.71\% of all stack traces reported a java.lang.NullPointerException as its root cause.
+
+       \item  4.4\% of the reported exception stack traces are ``undocumented runtime exceptions'' raised by third-party code. 
+             (35 methods from the Android platform and 44 from libraries).
+
+     \item  We detected several cross-type exception wrappings, such as an OutOfMemoryError wrapped in a checked exception,
+in stack traces. This suggests that the purpose of Java's hybrid exception model comprising checked and unchecked exceptions may not have been adequately used.
+
+   \item Undocumented checked exception signaled by a  JNI interface function defined on Android platform. 
+    Some flows reported a checked exception thrown by a JNI interface function, yet this exception was not  
+    part of the ``exception interface'' of the JNI function nor the methods on its invocation chain. This can lead to 
    uncaught exceptions that are difficult to debug.
@@
 
-Such outcomes point to threats not only to the developement of robust Android apps, 
-but also to the developement of any robust Java-based system. For instance, in the absence of the 
-``exception specification" of  a third-party code, it is difficult or 
-even enfeasible for the developer to protect the code against ``unforeseen" exceptions. 
-Since, in such cases the client usually do not have access to the source code, the undocumented exceptions may 
+Such outcomes point to threats not only to the development of robust Android apps, 
+but also to the development of any robust Java-based system. For instance, in the absence of the 
+``exception specification" of third-party code, it is difficult or 
+even infeasible for the developer to protect the code against ``unforeseen'' exceptions. 
+Since in such cases the client usually does not have access to the source code, such undocumented exceptions may 
 remain uncaught and lead to system crashes. Moreover, the cross-type exception wrappings 
 (e.g., OutOfMemoryError is wrapped in a checked Exception) 
-may lead to unintended handling problems - trying to handle an instance of OutOfMemoryError ``hiden" in 
-a checked exception may lead the program to an unpredictable state.
+may lead to unintended handling problems - e.g, trying to handle an instance of OutOfMemoryError ``hidden" in 
+a checked exception may bring the program to an unpredictable state.
 
 Hence, the study results are relevant to Android and 
-Java developers who may underestimate how likely are the
- exceptions related to programming mistakes, and how 
+Java developers who may underestimate how 
+ exceptions are related to programming mistakes, and how 
  tricky it can be to prevent the uncaught exceptions from third-party code
- in the absence of ``exception interfaces". Moreover, results are relevant to 
+ in the absence of ``exception interfaces". Moreover, the results are relevant to 
 designers of languages and tools to consider ways of reducing such fragilities 
@@
 
-The remaining of this paper is organized as follows. Section 2 presents a
-background on Android Platform and Java exception model. 
-Section 3 presents the study design. Section 4 reports the study findings. 
-Section 5 provides further discussions and insights.
-Section 6 presents the threats to validity associated to this study. Finally Section
-7 describes the related work, and Section 8 concludes the paper and outlines
+The remaining of this paper is organized as follows. 
+Section~\ref{sec:back} presents a
+background on Android platform and Java exception model. 
+Section~\ref{sec:study} presents the study design. 
+Section~\ref{sec:exceptionminer} describes the ExceptionMiner tool we developed to conduct our study.
+Section~\ref{sec:result} reports the findings of our study.
+Section~\ref{sec:disc} provides a discussion of the wider implications of our results.
+Section~\ref{sec:threats} presents the threats to validity associated to this study. 
+Finally Section~\ref{sec:rele} describes related work, 
+and Section~\ref{sec:conc} concludes the paper and outlines
 directions for future work.
@@
 
-%The Android platform is mostly implemented in Java and uses Java Native Invocation (JNI) to access native C/C++ code. 
-%And as mentioned before, it reuses the embedded Java Exception model to detect and handle 
-%abnormal computation states.
-
 The Android Application Framework supports an event-driven development model. 
 Among the basic elements that comprise an Android application, we can cite the
- Activity class. The Activity is the main class of an Android app (there is usualy one
+ Activity class. The Activity is the main class of an Android app (there is usually one
 Activity per application window). Each Activity defines a set of hook methods  
@@
 framework (by inversion of control), in response to infrastructure and application events.
-
 The Android Application Framework also provides several supporting APIs, such as 
-the APIs to support concurrent development (e.g. AsyncTask). 
-Such APIs particularly useful due to the high demand for responsiveness 
+the APIs to support concurrent development (e.g. AsyncTask) - 
+which are usually needed due to the high demand for responsiveness 
 in mobile environment (mobile apps can easily become 
@@
 
-In Java, an exception is represented according to the class hierarchy shown in
-Figure~\ref{fig:exchier}.  According to it every exception is an
-instance of the Throwable class, and can be of three kinds: the checked exceptions
+In Java, exceptions are represented according to a class hierarchy, on which
+ every exception is an instance of the Throwable class, and can be of three kinds: the checked exceptions
 (extends Exception), the runtime exceptions (extends RuntimeException) and errors
@@
 can be either checked or runtime. The Java specification~\cite{gosling2000java} 
-suggests that user-defined exceptions should be checked (because doing so 
-the callers of a method will know about the exceptions that a method can throw and so 
-they can decide what to do about them). However, runtime exceptions can also be 
-used to represent user-defined exceptional conditions.In fact, there is a long-lasting debate 
-about the pros and cons of both approaches~\cite{javatut,stackoverlow,debate},
-\footnote{152 questions in Stackoverlow are related to this debate}.
-
-\begin{figure} \centering \includegraphics[width=\hsize]{new2_hierarchy.png}
-  \caption{Exception Hierarchy in Java} \label{fig:exchier} \end{figure}
-
-Besides, the runtime exceptions are also implicitly thrown by the runtime environment when a program violates 
+suggests that user-defined exceptions should be checked, because doing so 
+the callers of a method will know about the exceptions that a method can throw, and then 
+they can decide what to do about them. However, runtime exceptions can also be 
+used to represent user-defined exceptional conditions.There is a long-lasting debate 
+about the pros and cons of both approaches~\cite{javatut,stackoverlow,debate}
+%\footnote{152 questions in Stackoverlow are related to this debate}.
+
+The runtime exceptions are also implicitly thrown by the runtime environment when a program violates 
 the semantic constraints of the Java programming language (e.g., out-of-bounds array index, divide-by-zero 
@@
 
-%In Java an exception can be thrown in one of the following
-%circumstances~\cite{gosling2000java}: 
-%(i) explicitly thrown when a throw statement is executed; 
-%(ii) implicitly thrown by the JVM when the evaluation of an expression
-% violates the normal semantics of language, also referred as a coding error
-% (e.g., out-of-bounds array index, division-by-zero, access to a null reference); or 
-%(iii) implicitly thrown by the JVM due to an internal error or resource limitation (e.g.,
-%OutofMemoryError). 
-%\todo{This sub-section is quite similar to the section introduction, shouldn't
-%we merge them?}
-
-%When an exception is thrown, it causes a transfer of control from the point
-%where the exception occurred to a point that can be specified by the programmer
-%(exception handler) - in Java it is represented by the try-catch block .
+%\begin{figure} \centering \includegraphics[width=\hsize]{new2_hierarchy.png}
+% \caption{Exception Hierarchy in Java} \label{fig:exchier} \end{figure}
 
@@
 
-\begin{figure} \centering \includegraphics[scale=0.6]{stacktrace_bw.png}
+\begin{figure} \centering \includegraphics[scale=0.5]{stacktrace_bw.png}
 \caption{Example of an exception stack trace in Java.}
@@
 \end{figure}
-
-%<AVALIAR MOVER ESTE STACK TRACES PARA UMA FIGURA QUE MOSTR COMO ELE É QUEBRADO...
-%PELO EXCEPTION MINER E AS INFORMAÇÕES VAO PARA O BANCO...>
-
-% \subsection{Public Project Repositories and Issue Trackers} 
-% ONE OF THE REVIEWERS COMPLAINED ABOUT THE ABSENCE OF EXPLANATIONS OF WHAT AN ISSU IS...
-% As mentioned before, currently more and more projects are becoming available in 
-% public project repositories such as GitHub and Google Code.
-% Such repositories provides facilities such as....
-% The issue trackers available in such projects ....
 
@@
 
-\emph{II-Error represents an unrecoverable conditions which should not he handled} 
+\emph{II-Error represents an unrecoverable conditions which should not be handled} 
 (\cite{gosling2000java}).  Errors should result from failures detected
-by the runtime environemt which indicate resource deficiencies, invariant
+by the runtime environment which indicate resource deficiencies, invariant
 failures or other conditions, that make impossible for the program to recover.
@@
 pure java.lang.Exception or a java.lang.RuntimeException is considered a bad practice.
-
-%\textbf{IV-Do not repeatedly re-throw exceptions, handle exceptions as close as
-%possible to the problem}~\cite{wirfs2006toward}. 
-%Programs that frequently throw exceptions lose in performance ~\cite{wirfs2006toward,gosling2000java}.
-%Moreover, 
-%Close to the place that the exception was throw the caller usually has
-%enough contextual information to perform  corrective actions. When the exception
-%is propagated far away from the method where it was signaled (i.e. further away
-%#from the problem) it can be difficult to take meaningful recovery actions.
-%*~\todo{RSC:this practice was not well explored, can be removed if we need space.}
 
@@
 exceptions being thrown by it. According to~\cite{bloch2008effective}, it is
-also wise to document the explicitly thrown runtime exceptions\footnote{This
+also wise to document  (either or on the exception interface or using @throws tag on Javadoc) 
+the explicitly thrown runtime exceptions\footnote{This
 excludes the implicitly signaled runtime exceptions related to programming
-mistakes} as carefully as checked exceptions, specialy when developing
- library/framework reusable code. Doing so, the clients of such ibrary/framework
+mistakes} as carefully as checked exceptions, e when developing
+ library/framework reusable code. Doing so, the clients of such library/framework
 will be aware of all the exceptions a method can throw and can design the code
@@
 
-If the developer fails to do follow such practices (specially when developing 
-utility code) it will be difficult or even impossible for the caller to 
+If the developer fails to follow such practices (specially when developing 
+library/framework code) it will be difficult or even impossible for the caller to 
 make effective use of such method~\cite{wirfs2006toward, bloch2008effective},
@@
 which can lead to the uncaught exceptions - one of the main causes of 
-systems crashes. These best practices guided our exploratory study dscribed next.
-
-%Next section discusses the
-%study procedure and the practices' misuses that emerged from this study.
-
-%This study focus on the analysis of crash information (related to uncaught exceptions) of a family of Android applications
-%aiming at  extracting their common characteristics and looking for similar vulnerabilities on applications and underlaying platform.
-%In Java when an application crashes due to an uncaught exception ... 
-%Exception Stach trace is usually (procurar por artigo).
-
-%To answer this question we selected 
-%The goal of this exploratory mining study was to extract exception stack traces embedded on issues of Android projects 
-%(hosted in GitHub and Google Code), looking for its \emph{common characteristics} and investigating whether
-% they could reveal \emph{common fragilities} on the exception code of both the applications and the underlaying  framework. 
-%Our study focused on open-source apps, since the information needed to perform our study
-%cannot be retrieved from commercial apps, whose issue report systems and 
-%source codes are not publicly available. The Android open-source apps were also 
-%target of other research studies ~\cite{Linar13,ahimed}.
-
-%extract exception stack traces embedded on issues of Android projects 
-%(hosted in GitHub and Google Code), looking for its \emph{common characteristics} and investigating whether
-% they could reveal \emph{common fragilities} on the exception code of both the applications and the underlaying  framework. 
-%Our study focused on open-source apps, since the information needed to perform our study
-%cannot be retrieved from commercial apps, whose issue report systems and 
-%source codes are not publicly available. The Android open-source apps were also 
-%target of other research studies ~\cite{Linar13,ahimed}.
+systems crashes. These best practices guided our exploratory study described next.
 
@@
 \label{sec:study}
+
+\begin{figure*} \centering \includegraphics[scale=0.3]{studyoverview3.png}
+\caption{Study overview.} \label{fig:overview} \end{figure*}
 
@@
 embedded on issues of Android projects (hosted in GitHub and Google Code), can reveal 
-\emph{common fragilities} on the exception code of both the applications and the underlaying  framework. 
+\emph{common fragilities} on the exception code of both the applications and the underlying  framework. 
 Our study focused on open-source apps, since the information needed to perform our study
@@
 This exploration is guided by the set of best practices compiled in this study; we investigate whether characteristics 
-of exception stack traces can pinpoint best practices violations. As mentioned before, 
-call ``fragilities" characteristics on the exception-related code that favor the introduction
+of exception stack traces and the related source code it can pinpoint best practices violations. As mentioned before, 
+we call  ``fragilities" characteristics on the exception-related code that favor the introduction
 of failures such as uncaught exceptions and unintended handling.
 
-To support this investigation, we developed a tool called ExceptionMiner (see Section ~\ref{sec:exceptionminer})
-which extracts the exception stack traces embeeded on issues, 
+To support this investigation, we developed a tool called ExceptionMiner (Section~\ref{sec:exceptionminer})
+which extracts the exception stack traces embedded on issues, 
 and combines stack trace information with source code and bytecode 
 analysis. Such combination was needed to enable the identification of 
-fragilities on the exception-related code.
-Moreover, manual inspections were also used to leverage
+fragilities on the exception-related code. Moreover, manual inspections were also used to leverage
  the understanding of stack traces and support further discussions and insights.
 In this study we explore the domain quantitatively and highlight interesting cases by 
-exploring cases qualitatively.
+exploring cases qualitatively.  Figure~ref{fig:overview} gives an overview of the study, whose tasks are detailed next.
 
@@
 
+\begin{table}
+  \centering
+  \begin{tabular}{lr|lr}
+    \hline
+     \multicolumn{2}{c}{\bfseries{GitHub}} &  \multicolumn{2}{c}{\bfseries{Google Code}} \\
+      \bfseries{Lable} &  \bfseries{\% Occurrences} &  \bfseries{Lable} &  \bfseries{\% Occurrences} \\
+    \hline
+empty &	54.24\% & Defect &	91.96\% \\
+Defect &	39.56\%  & Enhancement  &	3.16\% \\
+Enhancement &	0.57\% & Task	& 1.37\% \\
+Support &	0.52\% & empty &	1.12\% \\
+Problem &	0.36\% & StackTrace &	0.70\% \\
+Others &	4.74\% &  Others &	1.68\% \\   
+  \hline
+  \end{tabular}
+  \caption{Labels on issues including exception stack traces.}
+  \label{tab:lables}
+\end{table}
+
 %As mentioned before, this study mined information available on GitHub issues,
@@
 Up to 23 Feb 2014,  when we queried GHTorrent, such heuristic could
- find 2.542 repositories. Running the ExceptionMiner tool (described next)
- in this set we observed that 589 apps had at least one issue containing a stack trace.
+ find 2.542 repositories. Running the ExceptionMiner tool 
+ in this set we observed that 589 projects had at least one issue containing a stack trace.
 
@@
 repositories are offered a pre-defined set of labels, but repository owners can modify 
-them at will. Therefore, an issue may have none ore an arbitrary set of labels depending 
-on which repository it was created. Table ~\ref{tab:lables} illustrates the ocurrences of different lables 
+them at will. Therefore, an issue may have none or an arbitrary set of labels depending 
+on which repository it was created. Table~\ref{tab:lables} illustrates the ocurrences of different labels 
 on the issues including exception stack traces.
-
-% \centering
-%  \begin{tabular}{|p{2cm}| p{5cm}|}
-%    \hline
-
-\begin{table}
-  \centering
-  \begin{tabular}{lr|lr}
-    \hline
-     \multicolumn{2}{c}{\bfseries{GitHub}} &  \multicolumn{2}{c}{\bfseries{Google Code}} \\
-      \bfseries{Lable} &  \bfseries{\% Occurrences} &  \bfseries{Lable} &  \bfseries{\% Occurrences} \\
-    \hline
-empty &	54,24\% & Defect &	91,96\% \\
-Defect &	39,56\%  & Enhancement  &	3,16\% \\
-Enhancement &	0,57\% & Task	& 1,37\% \\
-Support &	0,52\% & empty &	1,12\% \\
-Problem &	0,36\% & StackTrace &	0,70\% \\
-Others &	4,74\% &  Others &	1,68\% \\   
-  \hline
-  \end{tabular}
-  \caption{Lables on issues including exception stack traces.}
-  \label{tab:lables}
-\end{table}
 
@@
  Moreover, issues and pull requests are dual on Github; all pull requests have a corresponding 
-``backing'' issue which are automatically generated. Therefore, we excluded such automatically generated
+``backing'' issue (i.e., an issue that is automatically generated per pull request). Therefore, we excluded such automatically generated
 issues from our analysis.
@@
 Google Code contains largely used open-source Android apps (e.g. K9Mail \footnote{K9Mail moved to Github but as a way of not loosing the project history 
-it advices their users to report bugs on Google Code issue tracker.}), however, differently from GitHub, Google Code does not provide an API to access the information related
- to hosted projects \footnote{Google code used to provided a Web service to such repository, but it was deactivated in June 2013 what Google called a "clean-up action".}.
+it advises their users to report bugs on Google Code issue tracker: https://github.com/k9mail/k-9/wiki/LoggingErrors.}),
+ however, differently from GitHub, Google Code does not provide an API to access the information related
+ to hosted projects \footnote{Google code used to provided a Web service to such repository, but it was deactivated in June 2013 what Google called a `clean-up action".}.
 To overcome this limitation we needed to implement a Web Crawler (incorporated in ExceptionMiner tool described next) that navigated 
@@
 (on Google Code search interface) for the term ``android". On January 2014, when we queried Google Code, such heuristic could
- returned a list of 788  projects. This list comprised the seeds sent to our Crawler.
-
-From this set, 724 defined at least 1 issue. Running the ExceptionMiner tool (described next)
- in this set we observed that 183 apps had at least one issue containing an exception stack trace.
+ returned a list of 788  projects. This list comprised the seeds sent to our Crawler .
+
+The Crawler retrieved all issues and comments for these projects.
+From this set, 724 defined at least 1 issue. Running the ExceptionMiner tool 
+ in this set we observed that 183 projects had at least one issue containing an exception stack trace.
  Then we performed further cleanup (similar to the one described previously) inspecting the site 
 of each project. As a result we could identify 157 Android projects.  This set contained overall 127,456 issues,
- from which 1,963 exception stack traces were extracted. Table ~\ref{tab:lables} illustrates the ocurrences of different lables 
-on the issues including exception stack traces. Differently from GitHib, in Google Code most of 
+ from which 1,963 exception stack traces were extracted. Table~\ref{tab:lables} illustrates the ocurrences of different labels 
+on the issues including exception stack traces. Differently from GitHub, in Google Code most of 
 the issues were labled as ``Defect". However, based on the same assumption described for the GitHub repository
- we considered all issues reporting stacks (regardless its lables).
+ we considered all issues reporting stacks (regardless its labels).
 
@@
 
-The ExceptionMiner is modular mining tool which is able to connect to issue repositories, 
+The ExceptionMiner is a mining tool which can connect to different issue repositories, 
 extract issues, mine stack traces from them, distill exception stack trace information,
@@
  that can traverse Google Code Web interface and extract the project issues. 
-Projects matadata and the issues associated to them are stored on a relational 
-database by this components.
+Projects meta-data and the issues associated to each project are stored on a relational 
+database.
 
@@
 In contrast to existing issue parsing solutions such as Infozilla, the parser
-created in this work: (i) can discover stack traces inlined with logs files \footnote{In several 
+created in this work can discover stack traces inlined with logs files \footnote{In several 
 exception stack traces, the exception frames were preceeded by logging information e.g., 
@@
 trace information with bytecode and sourcecode analysis. Each analysis is implemented as a different module 
-on ExceptionMiner. Next, we present core analyses which were implemente on ExceptionMiner. Such analyses 
-provide aditional information to be used on further analyses described in Section ~\ref{sec:result}:
+on ExceptionMiner. Next, we present core analyses implemente on ExceptionMiner. Such analyses 
+provide additional information to be used on further analyses described in Section~\ref{sec:result}:
 
@@
 (e.g. Error, Exception or RuntimeException). The module responsible for this analysis 
-uses Design Wizard~\cite{Brunet09} framework to inspect the bytecode of all exceptions reported on issues.
-The nidyke walks up the type hierarchy of a Java exception until it reaches a base exception type.
-This module analyzed the exception hierarchy of all exceptions defined on:
-Android Platform (API level 19) - which includes all basic Java exceptions that 
-can be thrown during the app execution. The exceptions thrown inside the apps
-were classified using manual inspections as well as the exceptions thrown only
-on previous versions of the Android platform.
+uses Design Wizard~\cite{Brunet09} framework to inspect the bytecode of all exceptions reported on issues,
+and a specific implementation to analise the source code of Java Exception classes.
+It walks up the type hierarchy of a Java exception until it reaches a base exception type.
+This module analyzed all exceptions defined on Android platform (API level 19) - 
+which includes all basic Java exceptions that can be thrown during the app execution,
+and exceptions thrown by Android core libraries. Moreover it also analized the 
+libraries and applications which defined exceptions reported on stack traces
+(the tool only analyzed the last version available).
 
@@
 To discover the packages composing the first two origins
-we could reffer to the Android Specification.
+we could refer to the Android Specification.
 To discover the packages composing each application, this module 
@@
  - which defines the main packages that compose the applications.
- Whenever this file was not available, we recursevely analyzed the 
+ Whenever this file was not available, the tool recursively analyzed the 
 structure of source-code directories composing the app
@@
  from libraries if their packages were not defined 
-on Android Platform, core libraries nor on applications.
-Table ~\ref{signalers} summarizes the signaler classification.
+on Android platform, or on core libraries or on applications.
+Table~\ref{tab:signalers} summarizes the signaler classification.
 
@@
     \hline
-    \bfseries{android} & If the exception is thrown on a method defined in Android Platform.\\
-    \bfseries{app}     & If the exception is thrown on an application method.\\
+    \bfseries{android} & If the exception is thrown on a method defined in Android platform.\\
+    \bfseries{app}     & If the exception is thrown on an method defined on an Android app.\\
     \bfseries{libcore} & If the exception is thrown on one of the core libraries reused by Android (i.e., org.apache.harmony, org.w3c.dom, sun.misc, org.apache.http, org.json, org.xml). \\
@@
 \subsection{Manual Inspections}
-In the context of the ExceptionMiner tool, manual inspections were used:
-to help on the identification of packages composing the Android platform, 
-libs and apps analyzed in this study (as described previously); and to 
+In the context of the ExceptionMiner tool, manual inspections were used to:
+(i) support on the identification of packages composing the Android platform, 
+libs and apps analyzed in this study (as described previously); and (ii)  
 identify the type of some exceptions reported on issues, 
-that could not be automatically identified by the ExceptionMiner tool.
-Some exceptions could not be automaticaly identified because they were 
-not defined on the last version of the application (the one analyzed by ExceptionMiner),
-or were defined on libraries not found among application resources.
-In such cases we manually inspect the sourcecode or javadoc
- documentation available online of each of them. 
-When the exception was not found on the previous versions, the exception where classified 
+that were not be automatically identified by the ExceptionMiner tool
+(because they were defined on previous versions of libraries,
+apps and Android Plaform). When the exception could not be  
+found niether automatically, nor manually (because they were related an app or lib the previous versions
+that were no more availble) the exception were classified 
 in our study as "Undefined".  Only 31 exceptions 
@@
 \subsection{Replication Package}
-All the data used in this study will be publicly available at https://github.com/souzacoelho/exceptionminer
+All the data used in this study is publicly available at https://github.com/souzacoelho/exceptionminer
 Specifically we provide: (i) all issues related to Android projects found
@@
 quantitative and qualitative analysis of the outcomes. The structure of this section
- is centered on questions that guied the search for fragilities (on the exception-related code)
+ is centered on questions that guided the search for fragilities on the exception-related code
 based on exception stack trace information.
@@
 After distilling the information available on the exception stack traces we could find 
-the the exceptions commonly reported as the root causes of stack traces.
-Table \ref{tab:topten} presents a list the top 10 root exceptions found in the study - 
+the exceptions commonly reported as the root causes of stack traces.
+Table \ref{tab:topten} presents a list of the top 10 root exceptions found in the study - 
  ranked by the number of distinct projects on which they were reported. 
-The table shows how many times the signaler of such exception was a method defined on
-the Android Platform, the Android libcore, the application itself or a third-party library -
- following the classification as presented in Table  \ref{tab:signalers}.
+This table also shows how many times the signaler of such exception was a method defined on
+the Android platform, the Android libcore, the application itself or a third-party library -
+ following the classification presented in Table  \ref{tab:signalers}.
 
@@
 
-java.lang.NullPointerException	&	332	&	51,96\%	&	1664	&	27,71\%	&	525	&	20	&	836	&	280	\\
-java.lang.IllegalStateException	&	120	&	18,78\%	&	278	&	4,63\%	&	185	&	31	&	41	&	39	\\
-java.lang.IllegalArgumentException	&	142	&	22,22\%	&	353	&	5,88\%	&	195	&	12	&	95	&	44	\\
-java.lang.RuntimeException	&	122	&	19,09\%	&	319	&	5,31\%	&	203	&	2	&	64	&	51	\\
-java.lang.OutOfMemoryError	&	78	&	12,21\%	&	237	&	3,95\%	&	141	&	16	&	35	&	34	\\
-java.lang.NoClassDefFoundError	&	67	&	10,49\%	&	94	&	1,57\%	&	10	&	0	&	46	&	37	\\
-java.lang.ClassCastException	&	64	&	10,02\%	&	130	&	2,16\%	&	55	&	0	&	55	&	20	\\
-java.lang.IndexOutOfBoundsException	&	62	&	9,70\%	&	166	&	2,76\%	&	53	&	0	&	93	&	18	\\
-java.lang.NoSuchMethodError	&	54	&	8,45\%	&	80	&	1,33\%	&	10	&	0	&	56	&	14	\\
-java.util.ConcurrentModificationException	&	43	&	6,73\%	&	65	&	1,08\%	&	5	&	0	&	46	&	13	\\
+java.lang.NullPointerException	&	332	&	51.96\%	&	1664	&	27.71\%	&	525	&	20	&	836	&	280	\\
+java.lang.IllegalStateException	&	120	&	18.78\%	&	278	&	4.63\%	&	185	&	31	&	41	&	39	\\
+java.lang.IllegalArgumentException	&	142	&	22.22\%	&	353	&	5.88\%	&	195	&	12	&	95	&	44	\\
+java.lang.RuntimeException	&	122	&	19.09\%	&	319	&	5.31\%	&	203	&	2	&	64	&	51	\\
+java.lang.OutOfMemoryError	&	78	&	12.21\%	&	237	&	3.95\%	&	141	&	16	&	35	&	34	\\
+java.lang.NoClassDefFoundError	&	67	&	10.49\%	&	94	&	1.57\%	&	10	&	0	&	46	&	37	\\
+java.lang.ClassCastException	&	64	&	10.02\%	&	130	&	2.16\%	&	55	&	0	&	55	&	20	\\
+java.lang.IndexOutOfBoundsException	&	62	&	9.70\%	&	166	&	2.76\%	&	53	&	0	&	93	&	18	\\
+java.lang.NoSuchMethodError	&	54	&	8.45\%	&	80	&	1.33\%	&	10	&	0	&	56	&	14	\\
+java.util.ConcurrentModificationException	&	43	&	6.73\%	&	65	&	1.08\%	&	5	&	0	&	46	&	13	\\
     \hline
@@
  or resource limitation (e.g., OutOfMemoryError). 
-From this set the java.lang.NullPointerException was the most reported root cause (27,71\% ). 
-If we consider the frequencly of NullPointerException 
-across projects, we can observe that 51,96\% of all projects reported at least one exception stack 
+From this set the java.lang.NullPointerException was the most reported root cause (27.71\%). 
+If we consider the frequency of NullPointerException 
+across projects, we can observe that 51.96\% of all projects reported at least one exception stack 
 trace on which the NullPoiterException was the root cause.
 
- The NullPointerExceptions are mainly signaled inside the applicaiton code and the Android platform,
- although we could also find NullPointerExceptions being signaled by third-party libraries. 
-Regarding reusable code, there is no consensus whether it is a good or bad practice to 
+ The NullPointerException was mainly signaled inside the application code (50\%) and the Android platform (31.5\%),
+ although we could also find the NullPointerException being signaled by third-party libraries (16.3\%). 
+Regarding reusable code (e.g., libraries and frameworks), there is no consensus whether it is a good or bad practice to 
 re-throw a NullPointerException. Some prefer to encapsulate such an exception on
@@
 
-The high prevalence of NullPointerExceptions is aligned with the findings of other 
+The high prevalence of NullPointerException is aligned with the findings of other research
 works~\cite{kim2013predicting,fraser20131600,csallner2004jcrasher}. For instance, Sunghun et
-al.~\cite{kim2013predicting} showed that in Eclipse bug report system 38\% the bugs 
-related to exception handling were caused by NullPointerException; other works on robustness 
+al.~\cite{kim2013predicting} showed that 38\% the bugs related to exception handling in Eclipse project
+ were caused by NullPointerException; other works on robustness 
 testing~\cite{maji2012empirical,csallner2004jcrasher} showed that most of the automatically 
-detected bugs were due to NullPointerExceptions and exceptions implicitly-signaled by Java 
+detected bugs were due to NullPointerException and exceptions implicitly-signaled by Java 
 environment due to programming mistakes or resource limitations
- (as the ones found in this study).
+ (as the ones found in our study).
 
@@
 Hence, we conducted an empirical investigation based on the following assumption: the name of the exception and its associated Javadoc 
-documentaion hold enought information to pinpont a ``programming mistake".To exemplify, an instance of ArrayOutOfBoundException
+documentation hold enough information to pinpoint a ``programming mistake".To exemplify, an instance of ArrayOutOfBoundException
  pinpoints to programming mistake which indicates that an array has been accessed with an illegal index.
@@
  Javadocs we could identify from this subset, which exceptions were related to programming mistakes.
- It is woith mentioning that  all exceptions identified as programming mistakes were exceptions defined on java.lang and java.util packages. 
+ It is worth mentioning that  all exceptions identified as programming mistakes were exceptions defined on java.lang and java.util packages. 
 After doing such analysis, we could calculate the prevalence of such exceptions on issues.
@@
 
-Table ~\ref{tab:causes} summarized the results. It also illustrates the number of general Java exceptions - 
+Table ~\ref{tab:causes} summarizes the results. It also illustrates the number of general Java exceptions - 
 whose full name does not carry the reason for the exception to be thrown.
-To ensure the quality of the analysis, three independent coders classified a randomly selected
+To ensure the quality of the analysis, three independent developers classified a randomly selected
 sample of 25 exception types (from the total 100) using the same list of concerns;
-the resulting interrater agreement was high (96\%). We could also observe that
-We can therefore assume that the exception names and Javadoc 
-were enough for classifying the exception types analyzed.
-
-The results show that the programming mistakes associated to the most reported root causes. 
-Although there is usually no proper handling for such exceptions, 
-besides presenting an error message to the user and restarting the application ~\footnote{Only 
+the resulting interrater agreement was high (96\%).
+
+The results show that the programming mistakes are associated to the most reported root causes. 
+Although there is usually no proper handling for such exceptions~\cite{wirfs2006toward}, 
+besides presenting an error message to the user and restarting the application~\footnote{Only 
 high fault tolerant systems need to provide solutions to handle them.}.  
-This finding calls the attention to the following: the high number of programming errors reported 
+This finding calls the attention to the following: the high number of programming mistakes reported 
 on issues are an indication that developers may be underestimating the impact that such 
@@
     \hline
-      Programming Mistakes &  51,96\%\\ 
-      Memory Restriction   &   4,10\% \\ 
-      General Exceptions   &  6,70\%\\
+      Programming Mistakes &  51.96\%\\ 
+      %Memory Restriction   &   4.10\% \\ 
+      General Exceptions   &  6.70\%\\
     \hline
@@
 \noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
-A multitude of programming mistakes were reported as the main causes of exception stack traces.
-51,96\% of all projects reported at least one exception
- stack trace on which the NullPoiterException
-was the root cause.
+The study revealed: approximately 52\% of the reported stack traces were associated to programming mistakes;
+and 27.71\% of all stack traces reported a java.lang.NullPointerException as the root cause.
 }
@@
 the root exception types (i.e., RuntimeException, Error, checked Exception) as well as its origins - which were 
-identified based on the package names of the signalers related to it on the stack traces (see Section ~\ref{sec:exceptionminer}).
+identified based on the package names of the signalers related to it on the stack traces (Section~\ref{sec:exceptionminer}).
 Table~\ref{tab:typeroottab} presents the types and origins of root exceptions of all stack traces analyzed.
 
-We can see, from Table ~\ref{tab:typeroottab}, that contrating with the other origins, most of the 
+We can see, from Table ~\ref{tab:typeroottab}, that contrasting with the other origins, most of the 
  exceptions signaled on Android Libcore (i.e., the set of libraries reused by the Android) were 
@@
 org.apache.http, org.json, org.xml, javax. Signaling checked exceptions is considered a 
-good practice (see best practice IV in Section~\ref{sec:best}) because by using checked exceptions a library can define a precise 
+good practice (see best practice IV in Section~\ref{sec:best}) because by using 
+checked exceptions a library can define a precise 
 exception interface ~\cite{miller1997issues} to its clients.
- Since such libraries are heavily used  in several projects, such finding can be related to the library's maturity.
+ Since such libraries are heavily used  in several projects, such finding can be related to the libraries maturity.
 
@@
 \centering
-\begin{tabular}{lccccr}
-    \hline
-    \bfseries{Root Cause Type} & \bfseries{Android} & \bfseries{Libcore} & \bfseries{App} & \bfseries{Lib}  & \bfseries{All}\\
-    \hline
-
-Runtime	&	1335	&	73	&	1843	&	690  &	3894 (64,85\%)\\  %2466+854
-Error	       &	 188              &	 46	&	302             &	167	           &	691 (11,51\%)	\\
-Checked	&	276           &	314	&	313          &	567	           &	1358 (22,61\%)	\\
-Throwable	&	0	       &	0	&	2            &	0         &	2 (0,03\%)	\\
-Undefined	&	4	&	0	&	18		&	38	   &	60	(1,00\%) \\
+\begin{tabular}{lccccrr}
+    \hline
+    \bfseries{Root Type} & \bfseries{Android} & \bfseries{Libcore} & \bfseries{App} & \bfseries{Lib}  & \bfseries{All} & \bfseries{\%} \\
+    \hline
+
+Runtime	&	1335	&	73	&	1843	&	690  &	3894 & 64.85\% \\  %2466+854
+Error	       &	 188              &	 46	&	302             &	167	           &	691 & 11.51\%	\\
+Checked	&	276           &	314	&	313          &	567	           &	1358 & 22.61\%	\\
+Throwable	&	0	       &	0	&	2            &	0         &	2 & 0.03\%	\\
+Undefined	&	4	&	0	&	18		&	38	   &	60	& 1.00\% \\
  \hline
@@
 
-We can also observe that most of the reported exceptions were RuntimeExceptions
-(64,85\%); and most of them were signaled on methods defined on the Application 
-the Android Platform or on libraries.Taking into account the results of empirical study on 
-the prevalende of ``programming erros" we know that a considerable ammount 
+We can also observe that most of the reported exceptions were runtime 
+(64.85\%); and most of them were signaled on methods defined either on the Application (47.3\%)
+or on the Android platform (34.3\%). We could also find runtime exceptions thrown by library code (17.7\%).
+ Taking into account the results of empirical study on the prevalence of ``programming mistakes", 
+we conclude that a considerable amount 
 of such exceptions, are related to the programming mistakes
-as pointed out previously.
-
-However, within this set of runtime exceptions there are also the ones that do not correspond to programming mistakes,
-and according to best practice IV should be documented as part of the exception interface of libraries/framework reusable
+as pointed out previously. However, within this set of runtime exceptions there are also the ones that do not correspond to programming mistakes,
+and according to Java specification should be documented as part of the exception interface of libraries/framework reusable
 methods.
 
-To investigate the obedience to such practice, we firstly filtered out all the exceptions implicitly
+\emph{\textbf{Inspecting Exception Interfaces.}} To investigate the obedience to such practice, we firstly filtered out all the exceptions implicitly
  signaled by  runtime environment (due to programming mistakes) since these exceptions 
-should not be documented on the method signature.
-
-Then we inspected the code each method (defined either on Android  Aplication Framework or on libraries) 
+should not be documented on the method signature.  Then we inspected the code each method 
+(defined either on Android  Aplication Framework or on libraries) 
 explicitly signaling a runtime exception not-related to programming mistakes.
-As a result of this inspection, we could observe that only 1 method out of 118 inspected methods
- (i.e., 0,8\%) documented the explicitly thrown runtime exception in a Javadoc comment; and none of these methods
-included the runtime exception  (reported on the stack) as part of the exception interface of the method (i.e., using 
-throws clause on method signature). This result is aligned with the results of other study conducted by from 
+Table~\ref{tab:runtimeinterface} presents the results of this inspection. 
+We could find 79 methods (both from libraries and Android platform) that  explicitly threw a runtime exception 
+without listing such exception on the exception interface (i.e., using 
+throws clause on method signature). From this set only one method (defined on a library)
+included an @throws tag on Javadoc - reporting that the given runtime exception
+could be thrown in some contidions. These methods were responsible for 
+267 exception stack traces mined in this study.
+
+This result is aligned with the results of other study conducted by  
 Sacramento et al ~\cite{sacramento2006unchecked} which observed that the
 runtime exceptions in .Net programs are most often not documented.
-
-Such undocumented runtime exceptions represent a threat to system robustness, specially
-when such exceptions are thrown then the third party code invoked inside the application 
-(e.g. libraries, or framework utility code). In such cases the client usually do not have access to 
+These ``undocumented runtime exceptions" represent a threat to system robustness, specially
+when such exceptions are thrown by a third party code invoked inside the application 
+(e.g. libraries, or framework utility code). In such cases the developer usually does not have access to 
 the source code. Hence in the absence of the exception documentation it is very difficult or even impossible
- for the client to design the application to deal with ``unforeseen runtime exceptions. As a consequence, the
+ for the client to design the application to deal with ``unforeseen" runtime exceptions. As a consequence, the
  undocumented runtime exception may remain uncaught and lead to system crashes.
 
+Moreover, these stack traces could also pointed out another best practice violation: the methods from Android platform
+ presented in Table~\ref{tab:runtimeinterface} signaled direct instances of RuntimeException. 
+Using such general exception, is considered a bad practice according to Java specification and compiled guidelines.
+
 \noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
- Only 0,8\% of the methods documented the explicitly signaled runtime exceptions. 
+The study revealed: 267 exception stack traces were caused by ``undocumented runtime exceptions" thrown by libraries or the Android platform; 
+several methods defined on 11 android packages signaling direct instances of RuntimeException.
 }
@@
 
+\begin{table}
+\centering
+
+ %\begin{tabular}{|p{7.9cm} p{0.1cm}|}
+\begin{tabular}{lrr}
+    \hline
+ \bfseries{Origin}   & \bfseries{methods} &  \bfseries{stacks}  \\ 
+\hline					
+Libraries	& 	44	& 	205	\\					
+\emph{android packages}   &  &   \\ 
+android.app	& 	5	& 	16	\\
+android.content	& 	2	& 	2	\\
+android.database	& 	1	& 	1	\\
+android.graphics	& 	5	& 	11	\\
+android.hardware	& 	4	& 	5	\\
+android.media	& 	3	& 	3	\\
+android.opengl	& 	1	& 	8	\\
+android.os	& 	11	& 	12	\\
+android.text	& 	1	& 	1	\\
+android.video	& 	1	& 	1	\\
+android.providers	& 	1	& 	2	\\
+\hline					
+\bfseries{All android} 	&	35	&	62	\\
+\hline					
+\bfseries{All}	&	79	&	267	\\
+\hline					
+  \end{tabular}
+\caption{Absence of exception interfaces on methods.}
+\label{tab:runtimeinterface}
+\end{table}
+
+\emph{\textbf{On the Presence of Checked Exceptions on Exception Interfaces.}} 
+In this study, the exception stack trace analysis revealed an unexpected fragility: a 
+checked exception thrown by a native method and not declared on the exception 
+interface of these methods signaling them. Such native method was defined on
+ the Android platform -  which uses Java Native Invocation (JNI) to access 
+native C/C++ code. This exception was thrown by the method getDeclaredMethods defined in java.lang.Class. 
+The Java-side declaration of this method does not have any throws clause, 
+leading programmers and the compiler to think that no checked exceptions can be thrown.
+ However, the C-code implementation did throw a ``checked exception" called NoSuchMethodException, 
+violating the declaration. The Java compiler could not detect this violation, because it does 
+not perform static exception checking on native methods. This type of bug is hard to diagnose
+because the developer usually do not have access to the native implementations. 
+Consequently, since it is not expected by the programmer, when such method throws 
+this exception, such undocumented exception may remain
+uncaught and cause the app to crash, or may be mistakenly handled by subsumption.
+The exception stack traces reporting this scenario is actually a real bug of Android 
+Gingerbread version (which still accounts for 13.6\% of devices running Android).
+%Among the Android projects on which this exception stack trace was reported we can cite:
+%PageTurner (206 stars), ActionBarSherlock (6,041 stars), otto (1,618 stars).
+
+\noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
+The study revealed: a checked exception being thown by a native method and not declared  
+on the ``exception interface" of such method and all methods on the invocation chain.
+}
+\end{minipage}}
+
 \subsection{What can the Exception Wrappings tell us?}
-%SELECT COUNT(*) FROM STACKTRACESUMMARY WHERE
-%repo !='android'  AND ROOT_EXTENDS LIKE 'CHECKED%' AND EXCEPTION_EXTENDS LIKE 'RUNTIME'
-
-%SELECT COUNT(*) FROM STACKTRACESUMMARY WHERE
-%repo !='android'  AND ROOT_EXTENDS LIKE 'CHEC%' AND EXCEPTION_EXTENDS LIKE 'RUN%';
-
-%SELECT COUNT(*) FROM STACKTRACESUMMARY WHERE
-%repo !='android'  AND ROOT_EXTENDS LIKE 'CHEC%' AND EXCEPTION_EXTENDS LIKE 'RUN%'
-%and SIGNALER_FROM LIKE 'LIB%';
-
- %Runtime &  Checked   & 70  / 18 & 114 /  34 &  66 /  9  &  0 /  & 19 / 19 &  29 / 6 \\
-      %Runtime   &  Error   & 38  \ 8  & 57    &  50 \ 8  &  0   & 6 \ 2  &  1   /  0\\
-%      Checked &  Runtime   & 12 \ 5  & 22 \ 9 & 4 &  0  & 9 \7 &  9 \ 2 \\
-%      Checked & Error      & 7 \ 1 &  8 & 5  &  0  & 0 \ 1 &  3 \0 \\
- %     Error & Checked      & 4 \ 10  &  7 & 0  \ 6  &  7  & 0 \6 &  0   \ 8    \\
-%      Error & Runtime     & 6 \ 2  &  13   & 1 &  1  & 0 \ 1 &  11 \ 3    \\
+
+\begin{table}
+\centering
+
+ %\begin{tabular}{|p{7.9cm} p{0.1cm}|}
+\begin{tabular}{ll}
+    \hline
+ id & \bfseries{Runtime wrapping Error}    \\  %\bfseries{\#}
+    \hline
+1 & java.lang.RuntimeException - java.lang.OutOfMemoryError  \\ %30
+2& java.lang.RuntimeException -  java.lang.StackOverflowError)   \\ %4
+\hline
+& \bfseries{Checked wrapping Error}  \\
+ \hline
+3&java.lang.reflect.InvocationTargetException - java.lang.OutOfMemoryError  \\ %5
+4&java.lang.Exception - java.lang.OutOfMemoryError   \\ %1
+\hline
+& \bfseries{Error wrapping Checked}  \\
+ \hline
+5&java.lang.NoClassDefFoundError - java.lang.ClassNotFoundException   \\ %4
+6&java.lang.AssertionError - javax.crypto.ShortBufferException)   \\ %3
+\hline
+& \bfseries{Error wrapping Runtime}    \\
+ \hline
+7&java.lang.ExceptionInInitializerError - java.lang.NullPointerException   \\ %9
+8&java.lang.ExceptionInInitializerError - java.lang.IllegalArgumentException 	 \\ %2
+ \hline
+  \end{tabular}
+\caption{Examples of Cross-type wrappings}
+\label{tab:exampeswrap}
+\end{table}
 
@@
 
-      Runtime &  Checked   & 88 & 148 &  75  &  0   & 38 &  35 \\
-      Runtime   &  Error   & 46  &  67    &  58  &  0   & 8  &  1   \\      
-      Checked &  Runtime   & 17  & 31 & 4 &  0  & 16 &  11 \\
+      Runtime &  Checked   & 88 & 148 &  75 &  0   & 38 &  35 \\  %(50.7\%) 
+      Runtime   &  Error      & 46  &  67    &  58  &  0   & 8  &  1   \\      %(86.5\%) 
+      Checked &  Runtime   & 17  & 31 & 4 &  0  & 16  &  11 \\  %(51.6\%) 
       Checked & Error         & 8    &  9  & 5  &  0  &  1 &  3  \\
@@
 \end{table*}
+
 Java is the only language that provides a hybrid exception model 
-which offers three kinds of exceptions each one holding an inttended exception behavior (i.e., error, 
-runtime and checked). The best practices I and II states the behavior that Error an Checked exceptions hold, which are: 
-(i) Errors should not be handled inside the system since they usually represent unrecoverable conditions detected 
-by the runtime environment (e.g.,  OutOfMemoryError); (ii) checked exceptions, on the other hand, should represent recoverable conditions.
-
-Table~\ref{tab:wrappingandroid} presents all wrappings found in this study, that
-envolve different exception types (i.e., Error, Checked, Exception). We call 
-such a wrapping a ``cross-type wrapping".  Some of these wrappings may reveal 
-violations to best practices I and II (e.g. Error wrapping a Checked exception) as discussed next.
-
-\emph{\textbf{Runtime wrapping Checked Exception.}} Such wrapping was 
-detected on 88 projects. Inspecting the elements responsible for performing such wrapping, we could observe that
-approximately  50\% of such wrappings were performed on methods defined on Android
-Platform. Taking a closer look on the methods responsible for such wrappings,
- we could observe that approximately 41\% of them were performed by a single class,
+which offers three kinds of exceptions each one holding an intended exception behavior (i.e., error, 
+runtime and checked). Table~\ref{tab:wrappingandroid} presents the wrappings found in this study that
+include different exception types (i.e., Error, checked Exception and Runtime). We call 
+such a wrapping a ``cross-type wrapping". Next we exemplify some of such
+ wrappings and discuss whether some of them may impact the system robustness.
+
+\emph{\textbf{Runtime Exception wrapping Error.}} From Table~\ref{tab:wrappingandroid}, we could observe that most of such wrappings
+were performed by Android platform (50.7\%). Inspecting each of them revealed that the android.os.AsyncTask, was responsible for 44\% 
+of these wrappings. This class is responsible for dealing with asynchronous tasks in Android.
+The code snippet below shows a general catch clause that converts any instance of Throwable (signaled during the execution
+of an asynchronous task) into an instance of RuntimeException and re-throw it.  
+Table~\ref{tab:exampeswrap} exemplifies some of the wrappings performed by this tasks, such as 
+RuntimeExceptions wrapping istances of OutOfMemoryError and StackOverflowError. 
+Such wraps mask an unrecoverable error wrapping on a general runtime exception.
+
+{\footnotesize
+\begin{verbatim}
+   try {
+      ...
+   } catch (InterruptedException e) {
+      android.util.Log.w(..., e);
+   } catch (ExecutionException e) {
+      throw new RuntimeException("...",e.getCause());
+   } catch (CancellationException e) {
+      ...
+   } catch (Throwable t) {
+     throw new RuntimeException("...", t);
+   }
+\end{verbatim}
+}
+
+\emph{\textbf{Checked Exception wrapping Error.}} Most of these wrappings
+ were also caused by the reflection library used by applications' methods. The methods responsible for the wrappings
+were also native methods written in C. Table~\ref{tab:exampeswrap} illustrates
+some of these wrappings some of them are masking an OutOfMemoryError
+into a checked exception. Such wrappings may also mask an unrecoverable error
+and may lead to the ``exception confusion problem" described next.
+
+\emph{\textbf{Error wrapping Runtime.}} The source code inspection related that
+these wrappings were performed by: native methods defined by 
+java.lang.Class (e.g., getDeclaredFields, getDeclaredConstructors), 
+methods from javax.crypto package. Table~\ref{tab:exampeswrap}  illustrates
+examples of such wrappings. In these examples, wapping a runtime exception in an error
+reflected a design decision.
+
+\emph{\textbf{Error wrapping Checked Exception.}} 
+The source code inspection related that to such wrappings were 
+performed by Java static initializers. If any exception is thrown in the context of a static initializer 
+(i.e., static block)  it is converted into an ExceptionInitializerError 
+on the point where the class is first used.  Table~\ref{tab:exampeswrap} illustrates
+examples of such wrappings. We could observe that these examples also reflected
+a design decision.
+
+\emph{\textbf{Runtime wrapping Checked Exception.}} This wrapping was responsible for 49.5\% of
+the cross-type wrappings.  From this set 50\%  were performed on methods defined on Android platform. 
+We could observe that approximately 41\% of them were performed by a single class,
  named ActivityThreat. This class is responsible for managing the life cycle of  Android Activities - 
 invoking the life cycle methods (i.e. onCreate(), onPause()) for every Activity 
-(window) of an Android app.
-Figure ~\ref{fig:snippets} (a) presents a code snippet of one the methods of 
-ActivityThreat,  the one responsible for invoking the Activitiy's onCreate method. 
-As we can see from lines 4-8, in case any exception is thrown in the context of such try-catch block, 
-it is wrapped in a general Runtime exception and re-thrown.
-
-%\begin{table}
-%\centering
-%\begin{tabular}{lll}
-%   \hline
-% \bfseries{Signaler class} &  \bfseries{Wrapper Type} & \bfseries{Ocurrences} \\
-%    \hline
-%android.app.ActivityThread & java.lang.RuntimeException & 61\\
-%android.view.LayoutInflater & android.view.InflateException &  7 \\
-%android.os.AsyncTask & java.lang.RuntimeException &  7\\
-%\hline
-%  \end{tabular}
-%\caption{Top 3 Runtime-Cheked wrappers.}
-%\label{tab:wrapping01}
-%\end{table}
-
-Inspectioning the source code, of the methods involved in this kind of wrapping,
-revealed an unexpected fragility: a checked exception thrown by a native method and not
-declared on the exception interface of these methods signaling them. Such native method 
-was defined on the Android Platform -  which uses Java Native Invocation (JNI) to access 
-native C/C++ code.
-
-The code snippet in Figure ~\ref{fig:snippets} (d) illustrates this fragility.
- The java/lang/Class declares a native method named getDeclaredMethods. 
-The Java-side declaration of of this method does not have any throws clause, 
-leading programmers and the compiler to think that no checked exceptions can be thrown.
- However, the C-code implementation did throw a "checked exception" NoSuchMethodException, 
-violating the declaration. The Java compiler, cannot detect this violation because it does 
-not perform static exception checking on native methods. This type of bug is hard to diagnose
-because the developer usualy do not have access to the native implementations. 
-Consequently, since it is not expected by the programmer, when such method throws 
-this exception, such undocumented exception may remain
-uncaught and cause the app to crash, or may be mistakenly handled by subsumption - 
-The exception stack traces reporting this scenario is actually a real bug of Android 
-Gingerbread version (which still accounts for 13.6\% of devices running Android).
-
-Among the Android projects on which this exception stack trace was reported we can cite:
-PageTurner (206 stars, 130 forks, 1184 commits), ActionBarSherlock (6,041 stars, 
-3,436 forks, and 1479 commints), otto (1,618 stars,	342 forks, 	177 commints).
+(window) of an Android app. This class wraps every instance of java.lang.Exception 
+signaled withing the life cycle methods in a general RuntimeException and re-throws. 
+We observed that it was a common design decision in other methods of Android platform
+(although usually considered a bad practice).
+
+We could also observe that some stack traces included successive cross-type wrappings, 
+such as: Runtime - Checked - Runtime - Checked - Runtime - Checked - Runtime. Hence, although some of these wrappings may be a result of design decisions,
+the mis-use of exception wrapppings may make the exception handling 
+code more complex (e.g., the multiple wrappings) and error-prone,
+ and lead to what we call ``exception confusion problem". To illustrate this problem we can use one of the wrappings discussed above.
+When the developer is confronted with a checked exception, the designer of the API is telling him/her 
+to handle the exceptional condition (according to Java Specification and 
+best practices). However, such exception may be wrapping an Error such as OutOfMemoryError (which indicate resource deficiency 
+that makes impossible for the program to recover). Hence, trying to handle such exception  
+may lead the program to an unpredictable state.
 
 \noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
-An unexpected fragility was detected: a checked exception being thown and not specified on 
-on the ``exception interface" of the signaler methods.
+The cross-type wrappings is an indication that even if the developer sticks to Java
+specification and best practices, it may be dificult for her/him to prevent the code from 
+incurring into situations that lead the program to an unpredictable state - 
+e.g., when handling checked Exceptions that are `hiding" Errors.
 }
@@
 
-\emph{\textbf{Runtime Exception wrapping Error.}} We could find  64 exception stack traces 
-on 46 projects on which an Error was wrapped in a runtime exception.
-Manually inspecting each exception stack trace, we could observe that 44% of the wrappings were performed 
-by android.os.AsyncTask. This class is responsible for dealing with asynchronous tasks in Android.
-The code snippet is illustrated in Figure ~ref{wrappings} (c) shows that any instance of Throwable signaled in the context of an asynchronous 
-task is converted in to a runtime and re-thrown.  Table ~\ref{tab:exampeswrap} exemplifies some of such wrappings.
-Such wrapping masks the unrecoverable exception in a general excepion - doing so  that can be mistakenly handled
-
-\begin{table}
-\centering
-
- %\begin{tabular}{|p{7.9cm} p{0.1cm}|}
-\begin{tabular}{ll}
-    \hline
- id & \bfseries{Runtime-Error Wrapping Examples}    \\  %\bfseries{\#}
-    \hline
-1 & java.lang.RuntimeException - java.lang.OutOfMemoryError  \\ %30
-2 &java.lang.RuntimeException - java.lang.NoClassDefFoundError \\ %6
-3 &java.lang.RuntimeException - java.lang.UnsatisfiedLinkError)  \\ %5
-4& java.lang.RuntimeException -  java.lang.StackOverflowError)   \\ %4
-\hline
- & \bfseries{Checked-Runtime Wrapping Examples}   \\
- \hline
-5& ...android.robospice.CacheSavingException - java.lang.ClassCastException   \\ %6
-6&java.lang.reflect.InvocationTargetException - java.lang.NullPointerException    \\ %3
-7&java.lang.reflect.InvocationTargetException - ...ResourcesdNotFoundException  \\ %2
-8&java.sql.SQLException - android.database.sqlite.SQLiteDiskIOException   \\ %2
- \hline
-& \bfseries{Checked-Error Wrapping Examples}  \\
- \hline
-9&java.lang.reflect.InvocationTargetException - java.lang.OutOfMemoryError  \\ %5
-10&java.lang.Exception - java.lang.OutOfMemoryError   \\ %1
-11&java.lang.reflect.InvocationTargetException - java.lang.NoSuchMethodError 	 \\ %1
-12&java.lang.reflect.InvocationTargetException - java.lang.UnsatisfiedLinkError	 \\ %1
-\hline
-& \bfseries{Error-Checked Wrapping Examples}  \\
- \hline
-13&java.lang.NoClassDefFoundError - java.lang.ClassNotFoundException   \\ %4
-14&java.lang.AssertionError - javax.crypto.ShortBufferException)   \\ %3
-\hline
-& \bfseries{Error-Runtime Wrapping Examples}    \\
- \hline
-15&java.lang.ExceptionInInitializerError - java.lang.NullPointerException   \\ %9
-16&java.lang.ExceptionInInitializerError - java.lang.IllegalArgumentException 	 \\ %2
- \hline
-  \end{tabular}
-\caption{Examples of Cross-type wrappings}
-\label{tab:exampeswrap}
-\end{table}
-
-\emph{\textbf{Checked wrapping Runtime Exception.}} Manually inspecting the stack 
-traces we could observe that such wrappings were performed either 
-by RoboSpice (an android library that supports the creation of asynchronous long running 
-tasks, which has 1.682 stars and 367 forks in GitHub) or by the reflection Java library (used by
-methods defined in Android, Application and library classes).
-
-\emph{\textbf{Checked Exception wrapping Error.}} Almost all of these wrappings
- were also caused by the reflection library used by Android applications. The methods responsible for the wrappings
-were also native methods written in C. Table~\ref{tab:exampleswrap} illustrates
-some oth these wrappings some of them are masking an OutOfMemoryError
-into a checked exception.
-
-\emph{\textbf{Error wrapping Checked Exception.}} These wrappings were performed
-either by native methods defined by java.lang.Class (e.g., getDeclaredFields, getDeclaredConstructors) 
-or methods from javax.crypto package. Table~\ref{tab:exampleswrap} illustrates
-the exception types involved in such wrappings.
-
-\emph{\textbf{Error wrapping Runtime.}} After manually inspecting the source code associated to such wrappings 
-we discovered that all of them were caused by static initializers. If any exception is thrown in the context of a static initializer (i.e., static block) 
-it is converted into an ExceptionInitializerError on the point where the class is first used. Table ~\ref{exampleswrap} illustrates
-the exception types involved in such wrappings.
-
-We could also observe that some stack traces included several intermediate cross-type wrappings.
-One interesting example was the following: Runtime-Checked-Runtime-Checked-Runtime-Checked-Runtime.
-
-Although some of these wrappings may be a result of design decisions,
-the mis-use of exception wrapppings may make the exception handling 
-code more complex (e.g., the multiple wrappings) and error-prone,
- and lead to what we call ``exception confusion problem"
-
-To illustrate this problem we can use one of the wrappings discussed above.
-When the developer is confronted with a checked exception, the designer of the API is telling him 
-to handle the exceptional condition (according to Java Specification and 
-best practices). However, if the developer tries to handle such exception,
-which is wrapping (or ``hiding") an OutOfMemoryError, such handling 
-may lead the program to an unpredictable state.
-
-The cross-type wrappings show us that even if the developer sticks to Java
-specification and best practices, it will be dificult for her/him to prevent the code from 
-incurring into a situation as the one described above.
-
-\noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
-Some of the detected cross-type exception wrappings points to a potential ``exception confusion problem".
-To illustrate: a checked exception wrapping an OutOfMemoryError; and the multiple wrapping Runtime-Checked-Runtime-Checked-Runtime-Checked-Runtime.
-}
-\end{minipage}}
-
 \section{Further Discussions}
@@
 
-\begin{figure*} \centering \includegraphics[scale=0.55]{examples_new2.png}
-\caption{Code snippets illustrating fragilities on the exception-related code} \label{fig:snippets} \end{figure*}
-
- "Everybody hates thinking about exceptions, because they’re not supposed to happen"
-  Brian Foote
+%\begin{figure*} \centering \includegraphics[scale=0.55]{examples_new2.png}
+%\caption{Code snippets illustrating fragilities on the exception-related code} \label{fig:snippets} \end{figure*}
+
+ ``Everybody hates thinking about exceptions, because they are not supposed to happen"
+  (Brian Foote)~\footnote{Brian Foote shared his opinion in a conversation with James Noble - quoted on the paper: hillside.net/plop/2008/papers/ACMVersions/coelho.pdf}
 
@@
 contextual information to adequately handle the exception, 
-and only tasks possible are: to present a message to the user
- and restart the application.
-
-However, such handler cannot replace a carefully designed exception 
-handling policy, which can hardly be designed in the absence of 
+a possible handler task is: to present a message to the user
+ and restart the application. However, such handler cannot replace a carefully designed exception 
+handling policy ~\cite{Robil00}, which can hardly be designed in the absence of 
 third-party documentation on the exceptions that
-may the thrown by it.
-
-On the other hand, documenting runtime exceptions is a tedious and error prone task, to help developers
+may the thrown by it. On the other hand, documenting runtime exceptions is a tedious and error prone task, to help developers
 mitigate this problem tools should be developed to automate the documentation of runtime exceptions
@@
 \subsection{On the null pointer problem}
-The null references was firstly introduced by Tony Hoare in ALGOL W, which after some years he called his “one-billion-dollar mistake” ~\cite{hoare}:
-
-\emph{``I call it my billion-dollar mistake. It was the invention of the null reference in 1965 [...] I couldn’t resist the temptation to put 
-in a null reference, simply because it was so easy to implement. This has led to innumerable errors, vulnerabilities, and system 
-crashes, which have probably caused a billion dollars of pain and damage in the last forty year."}
-
-In this sudy, the null references were, infact, responsible for several reported application errors - providing more evidences to Hoare's speach.
+The null references was firstly introduced by Tony Hoare in ALGOL W, which after some years he called his ``one-billion-dollar mistake"~\cite{hoare2}:
+\emph{``I call it my billion-dollar mistake. It was the invention of the null reference in 1965 [...] This has led to innumerable errors, vulnerabilities, and system 
+crashes, which have probably caused a billion dollars of pain and damage in the last forty years."}
+
+In this sudy, the null references were, infact, responsible for several reported application errors - providing more evidences to Hoare's speech.
 This observation emphasizes the importance of solutions to avoid NullPointerExceptions, such as:
-(i) lightweight intra-method null pointer analysis supported by Java8 @Nullable annotations ~\footnote{The @Nullable annotations are used by tools such as FindBugs, Eclipse, NetBeans, IntelliJ to detect potential
-null pointer dereferences at compile time. In April/2014 the Android Studio 0.5.5 was 
-released with support to these annotations};
+(i) lightweight intra-method null pointer analysis supported by Java8 @Nullable annotations~\footnote{They are used by tools e.g., Eclipse, IntelliJ, Android Studio 0.5.5 (release Apr.
+2014) to detect potential
+null pointer dereferences at compile time.};
 (ii) inter-method null pointer analysis tools such as the one proposed by Nanda and Sinha ~\cite{nanda2009accurate};
-or (iii) whether language designs which avoid null pointers, such 
+or (iii) language designs which avoid null pointers, such 
 as Monads ~\cite{Walde95} (i.e., used in functional languages for values that may not be available 
@@
  problem, either preventing odd wrappings or enabling the developer to better deal with them or even
-researches on the real usefulness of Java hybrid exception model.
+empirical studies on the real usefulness of Java hybrid exception model.
 
@@
 
-\noindent\emph{Internal Validity.} We used a heuristics-based parser to mine
+%\noindent
+\emph{Internal Validity.} We used a heuristics-based parser to mine
 exceptions from issues.  Our parsing strategy was conservative by default; for
@@
 we calculated the interrater agreement after three independent 
-coders classified a randomly selected sample (of 25 exception 
-types from the total 100); the interater agreement was high (96\%). 
+developers classified a randomly selected sample (of 25 exception 
+types from the total of 100); the interater agreement was high (96\%). 
 Another threat relates to the fact that parts of our analysis 
-are based on the availability of stack traces on issues on Github and Googlecode. 
+are based on the availability of stack traces on issues reported on Github and Googlecode projects. 
 In using these datasets, we make an underlying assumption: the stack traces reported on issues are 
@@
 the source code associated to a subset of the reported exception stack traces.
-Such subset comprise the stack traces related to the main findings 
-of the study (e.g., `ùndocumented runtime and checked exceptions",
+Such subset comprises the stack traces related to the main findings 
+of the study (e.g., ``undocumented runtime and checked exceptions",
 and cross-type wrappings").
-
-\noindent\emph{External Validity.} Our work use GHTorrent dataset. Although 
+%\noindent
+
+\emph{External Validity.} Our work uses GHTorrent dataset; although 
 comprehensive and extensive is not an exact replica of Github. 
@@
 a complete Github dataset. Instead, the goal of our study was to 
-pin-point fragilities on the exception-related code based on 
+pinpoint fragilities on the exception-related code based on 
 exception stack trace mining of a subset of projects.
-We limited out analysis to a subset of existing open-source Android projects.
+We limited our analysis to a subset of existing open-source Android projects.
 We are aware that the exception stack traces reported 
@@
 Moreover, several exception stack traces that support the findings of this study
-refered to exceptions comming from methods defined on Android Application Framework
+refered to exceptions coming from methods defined on Android Application Framework
 and third-party libraries.  Additionaly,  the ``fragilities" observed in this study are due to
 characteristics of Java exception model, which can impose challenges 
-the robustness of not ony to Android apps but also to other systems
+the robustness of not only to Android apps but also to other systems
  based on the same exception model.
@@
 
-In this section, we present work that is related to the present paper, divided into
-three categories: i) papers that use the information available on stack traces;
-ii) empirical studies on the usage of Java exceptions and its fault proneness;
-and iii) tools to extract stack traces information from natural language artifacts
-(e.g., issues and emails).
-%, and iv) empirical studies involving Android apps.
-
-\textit{Empirical studies using Android apps.} Ruiz et al.~\cite{Ruiz12}
-investigated the degree of reuse across applications in Android Market, the
-study showed that almost 23\% of the classes inherited from a base class in the
-Android API, and that 217 mobile apps were reused completely by another mobile
-app. Pathak et al.~\cite{Patha11} analyzed bug reports and developers
-discussions of Android platform and found out that approximately 20\% of
-energy-related bugs in Android occurred after an OS update. McDonnell et
-al.~\cite{McDon13} conducted a case study of the co-evolution behavior of
-Android API and 10 dependent applications using the version history data found
-in github. The study found that approximately 25\% of all methods in the client
-code used the Android API, and that the methods reusing fast-evolving APIs were
-more defect prone then others. Vásquez et al.~\cite{Linar13} analyzed
-approximately 7K free Android apps and observed that the last successful apps
-used Android APIs that were on average 300\% more change-prone than the APIs
-used by the most successful apps. Pingyu and Elbaum~\cite{Zhang12} analyzed bug
-reports of 5 Android applications an observed that 29\% had to do with poor
-exceptional handling code. Our work differs from the others as it aims at
-distilling the information of bug reports describing uncaught exceptions created
-for mobile applications in Github, in order to get a first view
-of what is causing the crashes across applications available and what the
-characteristics of the stack traces can tell us about the exception structure of
-those applications.
-
-\textit{Extracting Stack Traces from natural language artifacts.} 
-%Currently, many
-%software vendors embed automatic crash reporting tools in their software
-%systems. Moreover, third party crash collection services exist, most of them 
-%targeted for applications run on mobile phones~\cite{BugSe14,BugSn14,Googl14,Acra14}.
-Apart from bug reports, stack traces can be embedded in other forms of
-communication between developers, such as discussion logs and emails.
-Being intermixed with text makes the accurate extraction of stacktraces 
-an involved process.
-Infozilla~\cite{bettenburg2008extracting}
-is based on a set of regular expressions that extract a set of frames
-related to a stack trace. The main limitation of this solution is that it is not
-able to extract stack traces embedded on verbose log files (i.e., on which we
-can find log text mixed with exception frames). Bacchelli
-et al.~\cite{bacchelli2012content} propose a solution to recognize stack trace frames
-from development emails and relate it to code artifacts (i.e. classes) mentioned
-on the stack trace. In addition to those tools, ExceptionMiner is able to 
-both extract stack traces from natural language artifacts and to 
-classify them in a set of predefined categories.
-
-\textit{Analysis and Use of Stack Trace Information.} Several works have
+\textit{Analysis and Use of Stack Trace Information.} Several research works have
 investigated the use of stack trace information to support bug classification
@@
 
-\textit{Detecting Faults on Java Exception-Related Code.} 
+\textit{Empirical Studies on Exception Handling Defects.} 
 Cabral and Marques~\cite{cabral2007exception} analyzed the
@@
 observed that the actions inside handlers were very simple (e.g., logging and present a
-message to the user). Reimer and Srinivasan~\cite{reimer2003analyzing} listed a
-set of bad practices on exception handling that hinder software maintainability
-and robustness, based on their experience with Java enterprise applications.
-Our work differs from those
-
-Some tools ~\cite{Robil00}~\cite{coelho2008assessing} have ben proposed 
-to employ dataflow analysis to find the propagation paths of checked and unchecked 
-exception types. The main limitation of all static analysis tools is the number of false 
-positives inherent to static analysis solutions, which can lead to extremely high number 
-of exception flows, specially if considering Java Environment exceptions and exceptions 
-signaled from libraries. Moreover, such also have limited support some libraries (e.g., Java Reflection).
-Our owrk
+message to the user). Coelho et al.  ~\cite{coelho2008assessing} performed an 
+empirical study considering the fault-proneness of aspect-oriented implementations 
+for handling exceptions. Two releases of both Java and AspectJ implementation of a 
+product line were assessed as part of that study. Based on the use of an exception
+flow analysis tool, the study revealed that the AOP  refactoring increased the 
+number of uncaught exceptions, degrading the robustness of the AO version of every analyzed system.
+The main limitation of approaches based static analysis based approaches are the number of false
+positives they can generate, and the problems the faced when dealing with reflection libraries 
+and dynamic class loading. Pingyu and Elbaum~\cite{Zhang12} were the first to perform
+an empirical investigation of issues, related to exception-related bugs, on Android projects.  
+They perform a small scale study on which they manually inspected the issues of 
+5 Android applications. They observed that 29\% had to do with poor
+exceptional handling code, this empirical study was used to motivate the development of a tool
+aiming at amplifying existing tests to validate exception 
+handling code associated with external resources. This work inspired ours,
+ which automaticaly mined the exception stack traces embedded on issues 
+reported on 639 open source Android projects. The goal of our study was
+to identify common fragilities on the exception related code that can lead to 
+failures such as uncaught exceptions.
+
+\textit{Extracting Stack Traces from natural language artifacts.} 
+Apart from issues and bug reports, stack traces can be embedded in other forms of
+communication between developers, such as discussion logs and emails.
+Few tools have been proposed to mine stack traces embedded on such resources.
+ Infozilla~\cite{bettenburg2008extracting} is based on a set of regular expressions that extract a set of frames
+related to a stack trace. The main limitation of this solution is that it is not
+able to extract stack traces embedded on verbose log files (i.e., on which we
+can find log text mixed with exception frames). Bacchelli
+et al.~\cite{bacchelli2012content} propose a solution to recognize stack trace frames
+from development emails and relate it to code artifacts (i.e. classes) mentioned
+on the stack trace. In addition to those tools, ExceptionMiner is able to 
+both extract stack traces from natural language artifacts and to 
+classify them in a set of predefined categories.
 
@@
 with source code and bytecode analysis to identify common characteristics of exception 
-stack traces and pin-point fragilities on the exception-related code. 
+stack traces to pinpoint fragilities on the exception-related code. 
 We call ``fragilities" characteristics on the exception-related code that favor the introduction
@@
 Some fragilities consistently detected in this study were: 
-(i) an abundance of NullPointerExceptions - it was mentioned as the root exception in approximately 
-50\% of the analyzed projects. 
-(ii) undocumented runtime exceptions thrown by third party code,
+(i) Approximately 52\% of the reported stack traces can be attributed to programming mistakes. 
+In particular, 27.71\% of all stack traces reported a java.lang.NullPointerException as its root cause;
+(ii) 267 stack traces were caused by undocumented runtime exceptions thrown by third party code caused;
 (iii) undocumented checked exceptions thrown by JNI interface,
@@
 revealing that Java hybrid exception model is not fully used according to its purpose.
-As we can see, such fragilities can negatively affect not only the robustness of Android application 
+Such fragilities can negatively affect not only the robustness of Android application 
 but also the robustness of any Java-based application. 
-Our results calls the attention for the need of tool and language support to help 
+Our results calls the attention for the need of tools and language support to help 
 developers when dealing with such fragilities.
@@
 
-\section*{Acknowledgment} This work is partially supported by: CNPq -- Proc.
-484209/2013-2 and the NWO TestRoots project (639.022.314).
+%\section*{Acknowledgment} This work is partially supported by: CNPq -- Proc.
+%484209/2013-2 and the NWO TestRoots project (639.022.314).
 
@@
 \end{document}
+
+\textit{Empirical studies using Android apps.} Ruiz et al.~\cite{Ruiz12}
+investigated the degree of reuse across applications in Android Market, the
+study showed that almost 23\% of the classes inherited from a base class in the
+Android API, and that 217 mobile apps were reused completely by another mobile
+app. Pathak et al.~\cite{Patha11} analyzed bug reports and developers
+discussions of Android platform and found out that approximately 20\% of
+energy-related bugs in Android occurred after an OS update. McDonnell et
+al.~\cite{McDon13} conducted a case study of the co-evolution behavior of
+Android API and 10 dependent applications using the version history data found
+in github. The study found that approximately 25\% of all methods in the client
+code used the Android API, and that the methods reusing fast-evolving APIs were
+more defect prone then others. Vásquez et al.~\cite{Linar13} analyzed
+approximately 7K free Android apps and observed that the last successful apps
+used Android APIs that were on average 300\% more change-prone than the APIs
+used by the most successful apps.

### 2014-09-05 android-stacks.tex
threat

--- 2014-09-05 android-stacks.tex
@@
 
+%Total page numbers then:
+%intro:       1
+%background:  1.5
+%method:      1
+%tool         0.5
+%findings     4
+%discussion   1
+%threats      .5
+%conc         .5
+%------------------+
+%total        10 pages
+
 %\IEEEpeerreviewmaketitle
@@
 they have to face an increasing number of threats to application robustness
- arising from: faults on underlying middleware and hardware (e.g., GPS receiver, camera, sensors);
+ arising from: faults on underlying middleware and hardware (e.g., camera, sensors);
 faults on third party services and libraries; compatibility issues ~\cite{McDon13}; 
- memory and battery restrictions~\cite{Zhang12}; among others.
+ memory and battery restrictions; noisy external resources ~\cite{Zhang12} (e.g., wireless connections,GPS, bluetooth) among others.
 
@@
 The exception-related code may favor the introduction of failures such as 
-uncaught exceptions~\cite{jo2004uncaught, zhang2012amplifying} - 
+uncaught exceptions~\cite{jo2004uncaught, Zhang12} - 
 which can lead to system crashes, making the system even less robust~\cite{coelho2011unveiling}.
@@
 
- \item  A multitude of programming mistakes were reported as the main causes of exception stack traces:
- 27,71\% of the analyzed exception stack traces reported a java.lang.NullPointerException the root cause;
-and  51,96\% of all projects reported at least one exception stack trace on which the NullPoiterException
-was the root cause.
+       \item Aproximately 52\% of the reported stack traces were associated to programming mistackes. 
+And 27,71\% of all stack traces of reported a java.lang.NullPointerException the root cause.
 
@@
 even enfeasible for the developer to protect the code against ``unforeseen" exceptions. 
-Since, in such cases the client usually do not have access to the source code, the undocumented exceptions may 
+Since, in such cases the client usually do not have access to the source code, such undocumented exceptions may 
 remain uncaught and lead to system crashes. Moreover, the cross-type exception wrappings 
 (e.g., OutOfMemoryError is wrapped in a checked Exception) 
-may lead to unintended handling problems - trying to handle an instance of OutOfMemoryError ``hiden" in 
+may lead to unintended handling problems - e.g, trying to handle an instance of OutOfMemoryError ``hiden" in 
 a checked exception may lead the program to an unpredictable state.
@@
 
-%The Android platform is mostly implemented in Java and uses Java Native Invocation (JNI) to access native C/C++ code. 
-%And as mentioned before, it reuses the embedded Java Exception model to detect and handle 
-%abnormal computation states.
-
 The Android Application Framework supports an event-driven development model. 
@@
 framework (by inversion of control), in response to infrastructure and application events.
-
 The Android Application Framework also provides several supporting APIs, such as 
@@
 used to represent user-defined exceptional conditions.In fact, there is a long-lasting debate 
-about the pros and cons of both approaches~\cite{javatut,stackoverlow,debate},
-\footnote{152 questions in Stackoverlow are related to this debate}.
-
-\begin{figure} \centering \includegraphics[width=\hsize]{new2_hierarchy.png}
-  \caption{Exception Hierarchy in Java} \label{fig:exchier} \end{figure}
-
-Besides, the runtime exceptions are also implicitly thrown by the runtime environment when a program violates 
+about the pros and cons of both approaches~\cite{javatut,stackoverlow,debate}
+%\footnote{152 questions in Stackoverlow are related to this debate}.
+
+The runtime exceptions are also implicitly thrown by the runtime environment when a program violates 
 the semantic constraints of the Java programming language (e.g., out-of-bounds array index, divide-by-zero 
@@
 The programs are not expected to handle such runtime exceptions signaled by the runtime environment ~\cite{gosling2000java}.
+
+%\begin{figure} \centering \includegraphics[width=\hsize]{new2_hierarchy.png}
+% \caption{Exception Hierarchy in Java} \label{fig:exchier} \end{figure}
 
@@
 This exploration is guided by the set of best practices compiled in this study; we investigate whether characteristics 
-of exception stack traces can pinpoint best practices violations. As mentioned before, 
-call ``fragilities" characteristics on the exception-related code that favor the introduction
+of exception stack traces and the related source code it can pinpoint best practices violations. As mentioned before, 
+we call ``fragilities" characteristics on the exception-related code that favor the introduction
 of failures such as uncaught exceptions and unintended handling.
@@
 \noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
-A multitude of programming mistakes were reported as the main causes of exception stack traces.
-51,96\% of all projects reported at least one exception
- stack trace on which the NullPoiterException
-was the root cause.
+Aproximately 52\% of the reported stack traces were associated to programming mistackes. 
+And 27,71\% of all stack traces of reported a java.lang.NullPointerException the root cause.
 }
@@
 of such exceptions, are related to the programming mistakes
-as pointed out previously.
-
-However, within this set of runtime exceptions there are also the ones that do not correspond to programming mistakes,
-and according to best practice IV should be documented as part of the exception interface of libraries/framework reusable
+as pointed out previously. However, within this set of runtime exceptions there are also the ones that do not correspond to programming mistakes,
+and according to Java specification should be documented as part of the exception interface of libraries/framework reusable
 methods.
 
-To investigate the obedience to such practice, we firstly filtered out all the exceptions implicitly
+\emph{\textbf{Inspecting Exception Interfaces.}} To investigate the obedience to such practice, we firstly filtered out all the exceptions implicitly
  signaled by  runtime environment (due to programming mistakes) since these exceptions 
-should not be documented on the method signature.
-
-Then we inspected the code each method (defined either on Android  Aplication Framework or on libraries) 
+should not be documented on the method signature.  Then we inspected the code each method 
+(defined either on Android  Aplication Framework or on libraries) 
 explicitly signaling a runtime exception not-related to programming mistakes.
@@
 runtime exceptions in .Net programs are most often not documented.
-
-Such undocumented runtime exceptions represent a threat to system robustness, specially
+These ``undocumented runtime exceptions" represent a threat to system robustness, specially
 when such exceptions are thrown then the third party code invoked inside the application 
@@
 
+\emph{\textbf{On the Presence of Checked Exceptions on Exception Interfaces.}} 
+In this study, the exception stack trace analysis revealed an unexpected fragility: a 
+checked exception thrown by a native method and not declared on the exception 
+interface of these methods signaling them. Such native method was defined on
+ the Android Platform -  which uses Java Native Invocation (JNI) to access 
+native C/C++ code.
+
+This exceptions was thrown by the method getDeclaredMethods defined in java.lang.Class. 
+The Java-side declaration of this method does not have any throws clause, 
+leading programmers and the compiler to think that no checked exceptions can be thrown.
+ However, the C-code implementation did throw a "checked exception" NoSuchMethodException, 
+violating the declaration. The Java compiler, could not detect this violation, because it does 
+not perform static exception checking on native methods. This type of bug is hard to diagnose
+because the developer usualy do not have access to the native implementations. 
+Consequently, since it is not expected by the programmer, when such method throws 
+this exception, such undocumented exception may remain
+uncaught and cause the app to crash, or may be mistakenly handled by subsumption - 
+The exception stack traces reporting this scenario is actually a real bug of Android 
+Gingerbread version (which still accounts for 13.6\% of devices running Android).
+%Among the Android projects on which this exception stack trace was reported we can cite:
+%PageTurner (206 stars), ActionBarSherlock (6,041 stars), otto (1,618 stars).
+
+\noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
+An unexpected fragility was detected: a checked exception being thown and not specified  
+on the ``exception interface" of the signaler methods.
+}
+\end{minipage}}
+
 \subsection{What can the Exception Wrappings tell us?}
-%SELECT COUNT(*) FROM STACKTRACESUMMARY WHERE
-%repo !='android'  AND ROOT_EXTENDS LIKE 'CHECKED%' AND EXCEPTION_EXTENDS LIKE 'RUNTIME'
-
-%SELECT COUNT(*) FROM STACKTRACESUMMARY WHERE
-%repo !='android'  AND ROOT_EXTENDS LIKE 'CHEC%' AND EXCEPTION_EXTENDS LIKE 'RUN%';
-
-%SELECT COUNT(*) FROM STACKTRACESUMMARY WHERE
-%repo !='android'  AND ROOT_EXTENDS LIKE 'CHEC%' AND EXCEPTION_EXTENDS LIKE 'RUN%'
-%and SIGNALER_FROM LIKE 'LIB%';
-
- %Runtime &  Checked   & 70  / 18 & 114 /  34 &  66 /  9  &  0 /  & 19 / 19 &  29 / 6 \\
-      %Runtime   &  Error   & 38  \ 8  & 57    &  50 \ 8  &  0   & 6 \ 2  &  1   /  0\\
-%      Checked &  Runtime   & 12 \ 5  & 22 \ 9 & 4 &  0  & 9 \7 &  9 \ 2 \\
-%      Checked & Error      & 7 \ 1 &  8 & 5  &  0  & 0 \ 1 &  3 \0 \\
- %     Error & Checked      & 4 \ 10  &  7 & 0  \ 6  &  7  & 0 \6 &  0   \ 8    \\
-%      Error & Runtime     & 6 \ 2  &  13   & 1 &  1  & 0 \ 1 &  11 \ 3    \\
+
+\begin{table}
+\centering
+
+ %\begin{tabular}{|p{7.9cm} p{0.1cm}|}
+\begin{tabular}{ll}
+    \hline
+ id & \bfseries{Runtime wrapping Error}    \\  %\bfseries{\#}
+    \hline
+1 & java.lang.RuntimeException - java.lang.OutOfMemoryError  \\ %30
+2& java.lang.RuntimeException -  java.lang.StackOverflowError)   \\ %4
+\hline
+& \bfseries{Checked wrapping Error}  \\
+ \hline
+3&java.lang.reflect.InvocationTargetException - java.lang.OutOfMemoryError  \\ %5
+4&java.lang.Exception - java.lang.OutOfMemoryError   \\ %1
+\hline
+& \bfseries{Error wrapping Checked}  \\
+ \hline
+5&java.lang.NoClassDefFoundError - java.lang.ClassNotFoundException   \\ %4
+6&java.lang.AssertionError - javax.crypto.ShortBufferException)   \\ %3
+\hline
+& \bfseries{Error wrapping Runtime}    \\
+ \hline
+7&java.lang.ExceptionInInitializerError - java.lang.NullPointerException   \\ %9
+8&java.lang.ExceptionInInitializerError - java.lang.IllegalArgumentException 	 \\ %2
+ \hline
+  \end{tabular}
+\caption{Examples of Cross-type wrappings}
+\label{tab:exampeswrap}
+\end{table}
 
@@
 \end{table*}
+
 Java is the only language that provides a hybrid exception model 
-which offers three kinds of exceptions each one holding an inttended exception behavior (i.e., error, 
-runtime and checked). The best practices I and II states the behavior that Error an Checked exceptions hold, which are: 
-(i) Errors should not be handled inside the system since they usually represent unrecoverable conditions detected 
-by the runtime environment (e.g.,  OutOfMemoryError); (ii) checked exceptions, on the other hand, should represent recoverable conditions.
-
-Table~\ref{tab:wrappingandroid} presents all wrappings found in this study, that
+which offers three kinds of exceptions each one holding an intended exception behavior (i.e., error, 
+runtime and checked). The best practices I and II are related to such type-related-behavior: 
+(i) Errors should not be handled inside the system, since they usually represent unrecoverable conditions detected 
+by the runtime environment (e.g.,  OutOfMemoryError); (ii) checked exceptions, on the other hand, should represent 
+recoverable conditions.
+
+Table~\ref{tab:wrappingandroid} presents the wrappings found in this study, that
 envolve different exception types (i.e., Error, Checked, Exception). We call 
-such a wrapping a ``cross-type wrapping".  Some of these wrappings may reveal 
-violations to best practices I and II (e.g. Error wrapping a Checked exception) as discussed next.
-
-\emph{\textbf{Runtime wrapping Checked Exception.}} Such wrapping was 
-detected on 88 projects. Inspecting the elements responsible for performing such wrapping, we could observe that
-approximately  50\% of such wrappings were performed on methods defined on Android
-Platform. Taking a closer look on the methods responsible for such wrappings,
- we could observe that approximately 41\% of them were performed by a single class,
- named ActivityThreat. This class is responsible for managing the life cycle of  Android Activities - 
-invoking the life cycle methods (i.e. onCreate(), onPause()) for every Activity 
-(window) of an Android app.
-Figure ~\ref{fig:snippets} (a) presents a code snippet of one the methods of 
-ActivityThreat,  the one responsible for invoking the Activitiy's onCreate method. 
-As we can see from lines 4-8, in case any exception is thrown in the context of such try-catch block, 
-it is wrapped in a general Runtime exception and re-thrown.
-
-%\begin{table}
-%\centering
-%\begin{tabular}{lll}
-%   \hline
-% \bfseries{Signaler class} &  \bfseries{Wrapper Type} & \bfseries{Ocurrences} \\
-%    \hline
-%android.app.ActivityThread & java.lang.RuntimeException & 61\\
-%android.view.LayoutInflater & android.view.InflateException &  7 \\
-%android.os.AsyncTask & java.lang.RuntimeException &  7\\
-%\hline
-%  \end{tabular}
-%\caption{Top 3 Runtime-Cheked wrappers.}
-%\label{tab:wrapping01}
-%\end{table}
-
-Inspectioning the source code, of the methods involved in this kind of wrapping,
-revealed an unexpected fragility: a checked exception thrown by a native method and not
-declared on the exception interface of these methods signaling them. Such native method 
-was defined on the Android Platform -  which uses Java Native Invocation (JNI) to access 
-native C/C++ code.
-
-The code snippet in Figure ~\ref{fig:snippets} (d) illustrates this fragility.
- The java/lang/Class declares a native method named getDeclaredMethods. 
-The Java-side declaration of of this method does not have any throws clause, 
-leading programmers and the compiler to think that no checked exceptions can be thrown.
- However, the C-code implementation did throw a "checked exception" NoSuchMethodException, 
-violating the declaration. The Java compiler, cannot detect this violation because it does 
-not perform static exception checking on native methods. This type of bug is hard to diagnose
-because the developer usualy do not have access to the native implementations. 
-Consequently, since it is not expected by the programmer, when such method throws 
-this exception, such undocumented exception may remain
-uncaught and cause the app to crash, or may be mistakenly handled by subsumption - 
-The exception stack traces reporting this scenario is actually a real bug of Android 
-Gingerbread version (which still accounts for 13.6\% of devices running Android).
-
-Among the Android projects on which this exception stack trace was reported we can cite:
-PageTurner (206 stars, 130 forks, 1184 commits), ActionBarSherlock (6,041 stars, 
-3,436 forks, and 1479 commints), otto (1,618 stars,	342 forks, 	177 commints).
-
-\noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
-An unexpected fragility was detected: a checked exception being thown and not specified on 
-on the ``exception interface" of the signaler methods.
-}
-\end{minipage}}
+such a wrapping a ``cross-type wrapping". Next we exemplify some of such
+ wrappings and discuss why some of them may negatively impact the robustness.
 
@@
 on 46 projects on which an Error was wrapped in a runtime exception.
-Manually inspecting each exception stack trace, we could observe that 44% of the wrappings were performed 
+Manually inspecting each exception stack trace, we could observe that 44\% of the wrappings were performed 
 by android.os.AsyncTask. This class is responsible for dealing with asynchronous tasks in Android.
-The code snippet is illustrated in Figure ~ref{wrappings} (c) shows that any instance of Throwable signaled in the context of an asynchronous 
-task is converted in to a runtime and re-thrown.  Table ~\ref{tab:exampeswrap} exemplifies some of such wrappings.
-Such wrapping masks the unrecoverable exception in a general excepion - doing so  that can be mistakenly handled
-
-\begin{table}
-\centering
-
- %\begin{tabular}{|p{7.9cm} p{0.1cm}|}
-\begin{tabular}{ll}
-    \hline
- id & \bfseries{Runtime-Error Wrapping Examples}    \\  %\bfseries{\#}
-    \hline
-1 & java.lang.RuntimeException - java.lang.OutOfMemoryError  \\ %30
-2 &java.lang.RuntimeException - java.lang.NoClassDefFoundError \\ %6
-3 &java.lang.RuntimeException - java.lang.UnsatisfiedLinkError)  \\ %5
-4& java.lang.RuntimeException -  java.lang.StackOverflowError)   \\ %4
-\hline
- & \bfseries{Checked-Runtime Wrapping Examples}   \\
- \hline
-5& ...android.robospice.CacheSavingException - java.lang.ClassCastException   \\ %6
-6&java.lang.reflect.InvocationTargetException - java.lang.NullPointerException    \\ %3
-7&java.lang.reflect.InvocationTargetException - ...ResourcesdNotFoundException  \\ %2
-8&java.sql.SQLException - android.database.sqlite.SQLiteDiskIOException   \\ %2
- \hline
-& \bfseries{Checked-Error Wrapping Examples}  \\
- \hline
-9&java.lang.reflect.InvocationTargetException - java.lang.OutOfMemoryError  \\ %5
-10&java.lang.Exception - java.lang.OutOfMemoryError   \\ %1
-11&java.lang.reflect.InvocationTargetException - java.lang.NoSuchMethodError 	 \\ %1
-12&java.lang.reflect.InvocationTargetException - java.lang.UnsatisfiedLinkError	 \\ %1
-\hline
-& \bfseries{Error-Checked Wrapping Examples}  \\
- \hline
-13&java.lang.NoClassDefFoundError - java.lang.ClassNotFoundException   \\ %4
-14&java.lang.AssertionError - javax.crypto.ShortBufferException)   \\ %3
-\hline
-& \bfseries{Error-Runtime Wrapping Examples}    \\
- \hline
-15&java.lang.ExceptionInInitializerError - java.lang.NullPointerException   \\ %9
-16&java.lang.ExceptionInInitializerError - java.lang.IllegalArgumentException 	 \\ %2
- \hline
-  \end{tabular}
-\caption{Examples of Cross-type wrappings}
-\label{tab:exampeswrap}
-\end{table}
-
-\emph{\textbf{Checked wrapping Runtime Exception.}} Manually inspecting the stack 
-traces we could observe that such wrappings were performed either 
-by RoboSpice (an android library that supports the creation of asynchronous long running 
-tasks, which has 1.682 stars and 367 forks in GitHub) or by the reflection Java library (used by
-methods defined in Android, Application and library classes).
-
-\emph{\textbf{Checked Exception wrapping Error.}} Almost all of these wrappings
- were also caused by the reflection library used by Android applications. The methods responsible for the wrappings
+The code snippet is illustrated in Figure ~ref{wrappings} (c) shows that any instance of Throwable signaled 
+in the context of an asynchronous task is converted in to a runtime and re-thrown.  Table ~\ref{tab:exampeswrap} exemplifies some of such wrappings.
+Such wrapping masks an unrecoverable error in a general exception - doing so it that can be mistakenly handled.
+
+\emph{\textbf{Checked Exception wrapping Error.}} Most of these wrappings
+ were also caused by the reflection library used by applications' methods. The methods responsible for the wrappings
 were also native methods written in C. Table~\ref{tab:exampleswrap} illustrates
@@
 
-\emph{\textbf{Error wrapping Checked Exception.}} These wrappings were performed
-either by native methods defined by java.lang.Class (e.g., getDeclaredFields, getDeclaredConstructors) 
-or methods from javax.crypto package. Table~\ref{tab:exampleswrap} illustrates
-the exception types involved in such wrappings.
-
-\emph{\textbf{Error wrapping Runtime.}} After manually inspecting the source code associated to such wrappings 
-we discovered that all of them were caused by static initializers. If any exception is thrown in the context of a static initializer (i.e., static block) 
-it is converted into an ExceptionInitializerError on the point where the class is first used. Table ~\ref{exampleswrap} illustrates
-the exception types involved in such wrappings.
+\emph{\textbf{Error wrapping Runtime and Checked Exceptions.}} 
+These wrappings were performed by: native methods defined by 
+java.lang.Class (e.g., getDeclaredFields, getDeclaredConstructors), 
+methods from javax.crypto package, and Java static initializers. 
+If any exception is thrown in the context of a static initializer 
+(i.e., static block)  it is converted into an ExceptionInitializerError 
+on the point where the class is first used.  Table~\ref{tab:exampleswrap} illustrates
+examples of such wrappings.
 
@@
 code more complex (e.g., the multiple wrappings) and error-prone,
- and lead to what we call ``exception confusion problem"
-
-To illustrate this problem we can use one of the wrappings discussed above.
+ and lead to what we call ``exception confusion problem". To illustrate this problem we can use one of the wrappings discussed above.
 When the developer is confronted with a checked exception, the designer of the API is telling him 
@@
 
+\noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
 The cross-type wrappings show us that even if the developer sticks to Java
 specification and best practices, it will be dificult for her/him to prevent the code from 
-incurring into a situation as the one described above.
-
-\noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
-Some of the detected cross-type exception wrappings points to a potential ``exception confusion problem".
-To illustrate: a checked exception wrapping an OutOfMemoryError; and the multiple wrapping Runtime-Checked-Runtime-Checked-Runtime-Checked-Runtime.
+incurring into situations that lead the program to an unpredictable state. 
 }
@@
 and only tasks possible are: to present a message to the user
- and restart the application.
-
-However, such handler cannot replace a carefully designed exception 
-handling policy, which can hardly be designed in the absence of 
+ and restart the application. However, such handler cannot replace a carefully designed exception 
+handling policy ~\cite{Robil00}, which can hardly be designed in the absence of 
 third-party documentation on the exceptions that
@@
 \subsection{On the null pointer problem}
-The null references was firstly introduced by Tony Hoare in ALGOL W, which after some years he called his “one-billion-dollar mistake” ~\cite{hoare}:
-
-\emph{``I call it my billion-dollar mistake. It was the invention of the null reference in 1965 [...] I couldn’t resist the temptation to put 
-in a null reference, simply because it was so easy to implement. This has led to innumerable errors, vulnerabilities, and system 
+The null references was firstly introduced by Tony Hoare in ALGOL W, which after some years he called his ``one-billion-dollar mistake"~\cite{hoare2}:
+\emph{``I call it my billion-dollar mistake. It was the invention of the null reference in 1965 [...] This has led to innumerable errors, vulnerabilities, and system 
 crashes, which have probably caused a billion dollars of pain and damage in the last forty year."}
@@
 This observation emphasizes the importance of solutions to avoid NullPointerExceptions, such as:
-(i) lightweight intra-method null pointer analysis supported by Java8 @Nullable annotations ~\footnote{The @Nullable annotations are used by tools such as FindBugs, Eclipse, NetBeans, IntelliJ to detect potential
-null pointer dereferences at compile time. In April/2014 the Android Studio 0.5.5 was 
-released with support to these annotations};
+(i) lightweight intra-method null pointer analysis supported by Java8 @Nullable annotations~\footnote{Such annotations are used by tools e.g., Eclipse, IntelliJ to detect potential 
+null pointer dereferences at compile time. The Android Studio 0.5.5 released in April 2014 support to these annotations.};
 (ii) inter-method null pointer analysis tools such as the one proposed by Nanda and Sinha ~\cite{nanda2009accurate};
@@
 Such subset comprise the stack traces related to the main findings 
-of the study (e.g., `ùndocumented runtime and checked exceptions",
+of the study (e.g., ``undocumented runtime and checked exceptions",
 and cross-type wrappings").
@@
 a complete Github dataset. Instead, the goal of our study was to 
-pin-point fragilities on the exception-related code based on 
+pinpoint fragilities on the exception-related code based on 
 exception stack trace mining of a subset of projects.
@@
 \label{sec:rele}
-
-In this section, we present work that is related to the present paper, divided into
-three categories: i) papers that use the information available on stack traces;
-ii) empirical studies on the usage of Java exceptions and its fault proneness;
-and iii) tools to extract stack traces information from natural language artifacts
-(e.g., issues and emails).
-%, and iv) empirical studies involving Android apps.
-
-\textit{Empirical studies using Android apps.} Ruiz et al.~\cite{Ruiz12}
-investigated the degree of reuse across applications in Android Market, the
-study showed that almost 23\% of the classes inherited from a base class in the
-Android API, and that 217 mobile apps were reused completely by another mobile
-app. Pathak et al.~\cite{Patha11} analyzed bug reports and developers
-discussions of Android platform and found out that approximately 20\% of
-energy-related bugs in Android occurred after an OS update. McDonnell et
-al.~\cite{McDon13} conducted a case study of the co-evolution behavior of
-Android API and 10 dependent applications using the version history data found
-in github. The study found that approximately 25\% of all methods in the client
-code used the Android API, and that the methods reusing fast-evolving APIs were
-more defect prone then others. Vásquez et al.~\cite{Linar13} analyzed
-approximately 7K free Android apps and observed that the last successful apps
-used Android APIs that were on average 300\% more change-prone than the APIs
-used by the most successful apps. Pingyu and Elbaum~\cite{Zhang12} analyzed bug
-reports of 5 Android applications an observed that 29\% had to do with poor
-exceptional handling code. Our work differs from the others as it aims at
-distilling the information of bug reports describing uncaught exceptions created
-for mobile applications in Github, in order to get a first view
-of what is causing the crashes across applications available and what the
-characteristics of the stack traces can tell us about the exception structure of
-those applications.
-
-\textit{Extracting Stack Traces from natural language artifacts.} 
-%Currently, many
-%software vendors embed automatic crash reporting tools in their software
-%systems. Moreover, third party crash collection services exist, most of them 
-%targeted for applications run on mobile phones~\cite{BugSe14,BugSn14,Googl14,Acra14}.
-Apart from bug reports, stack traces can be embedded in other forms of
-communication between developers, such as discussion logs and emails.
-Being intermixed with text makes the accurate extraction of stacktraces 
-an involved process.
-Infozilla~\cite{bettenburg2008extracting}
-is based on a set of regular expressions that extract a set of frames
-related to a stack trace. The main limitation of this solution is that it is not
-able to extract stack traces embedded on verbose log files (i.e., on which we
-can find log text mixed with exception frames). Bacchelli
-et al.~\cite{bacchelli2012content} propose a solution to recognize stack trace frames
-from development emails and relate it to code artifacts (i.e. classes) mentioned
-on the stack trace. In addition to those tools, ExceptionMiner is able to 
-both extract stack traces from natural language artifacts and to 
-classify them in a set of predefined categories.
 
@@
 
-\textit{Detecting Faults on Java Exception-Related Code.} 
+\textit{Empirical Studies analizing Java exception-related code.} 
 Cabral and Marques~\cite{cabral2007exception} analyzed the
@@
 observed that the actions inside handlers were very simple (e.g., logging and present a
-message to the user). Reimer and Srinivasan~\cite{reimer2003analyzing} listed a
-set of bad practices on exception handling that hinder software maintainability
-and robustness, based on their experience with Java enterprise applications.
-Our work differs from those
-
-Some tools ~\cite{Robil00}~\cite{coelho2008assessing} have ben proposed 
-to employ dataflow analysis to find the propagation paths of checked and unchecked 
-exception types. The main limitation of all static analysis tools is the number of false 
-positives inherent to static analysis solutions, which can lead to extremely high number 
-of exception flows, specially if considering Java Environment exceptions and exceptions 
-signaled from libraries. Moreover, such also have limited support some libraries (e.g., Java Reflection).
-Our owrk
+message to the user). Pingyu and Elbaum~\cite{Zhang12} 
+analyzed the bug reports of 5 Android applications an observed that 29\% had to do with poor
+exceptional handling code, this empirical study was used to motivate the development of a tool
+aiming at amplifying existing tests to validate exception 
+handling code associated with external resources. Our work mined the exception
+stack traces embedded on issues reported 639 opensource Android projects
+to identify common fragilities on the exception related code that can lead to 
+failures such as uncaught exceptions.
+
+\textit{Empirical studies using Android apps.} Ruiz et al.~\cite{Ruiz12}
+investigated the degree of reuse across applications in Android Market, the
+study showed that almost 23\% of the classes inherited from a base class in the
+Android API, and that 217 mobile apps were reused completely by another mobile
+app. Pathak et al.~\cite{Patha11} analyzed bug reports and developers
+discussions of Android platform and found out that approximately 20\% of
+energy-related bugs in Android occurred after an OS update. McDonnell et
+al.~\cite{McDon13} conducted a case study of the co-evolution behavior of
+Android API and 10 dependent applications using the version history data found
+in github. The study found that approximately 25\% of all methods in the client
+code used the Android API, and that the methods reusing fast-evolving APIs were
+more defect prone then others. Vásquez et al.~\cite{Linar13} analyzed
+approximately 7K free Android apps and observed that the last successful apps
+used Android APIs that were on average 300\% more change-prone than the APIs
+used by the most successful apps.
+
+\textit{Extracting Stack Traces from natural language artifacts.} 
+Apart from issues and bug reports, stack traces can be embedded in other forms of
+communication between developers, such as discussion logs and emails.
+Few tools have been proposed to mine stack traces embedded on such resources.
+ Infozilla~\cite{bettenburg2008extracting} is based on a set of regular expressions that extract a set of frames
+related to a stack trace. The main limitation of this solution is that it is not
+able to extract stack traces embedded on verbose log files (i.e., on which we
+can find log text mixed with exception frames). Bacchelli
+et al.~\cite{bacchelli2012content} propose a solution to recognize stack trace frames
+from development emails and relate it to code artifacts (i.e. classes) mentioned
+on the stack trace. In addition to those tools, ExceptionMiner is able to 
+both extract stack traces from natural language artifacts and to 
+classify them in a set of predefined categories.
 
@@
 with source code and bytecode analysis to identify common characteristics of exception 
-stack traces and pin-point fragilities on the exception-related code. 
+stack traces and pinpoint fragilities on the exception-related code. 
 We call ``fragilities" characteristics on the exception-related code that favor the introduction
@@
 \end{document}
+
+\emph{\textbf{Runtime wrapping Checked Exception.}} Such wrapping was 
+detected on 88 projects. Inspecting the elements responsible for performing such wrapping, we could observe that
+approximately  50\% of such wrappings were performed on methods defined on Android
+Platform. Taking a closer look on the methods responsible for such wrappings,
+ we could observe that approximately 41\% of them were performed by a single class,
+ named ActivityThreat. This class is responsible for managing the life cycle of  Android Activities - 
+invoking the life cycle methods (i.e. onCreate(), onPause()) for every Activity 
+(window) of an Android app.
+Figure ~\ref{fig:snippets} (a) presents a code snippet of one the methods of 
+ActivityThreat,  the one responsible for invoking the Activitiy's onCreate method. 
+As we can see from lines 4-8, in case any exception is thrown in the context of such try-catch block, 
+it is wrapped in a general Runtime exception and re-thrown.
+
+%\begin{table}
+%\centering
+%\begin{tabular}{lll}
+%   \hline
+% \bfseries{Signaler class} &  \bfseries{Wrapper Type} & \bfseries{Ocurrences} \\
+%    \hline
+%android.app.ActivityThread & java.lang.RuntimeException & 61\\
+%android.view.LayoutInflater & android.view.InflateException &  7 \\
+%android.os.AsyncTask & java.lang.RuntimeException &  7\\
+%\hline
+%  \end{tabular}
+%\caption{Top 3 Runtime-Cheked wrappers.}
+%\label{tab:wrapping01}
+%\end{table}
+
+\emph{\textbf{Checked wrapping Runtime Exception.}} Manually inspecting the stack 
+traces we could observe that such wrappings were performed either 
+by RoboSpice (an android library that supports the creation of asynchronous long running 
+tasks, which has 1.682 stars and 367 forks in GitHub) or by the reflection Java library (used by
+methods defined in Android, Application and library classes).

### 2014-09-06 android-stacks.tex
threat
all over

@@
 
-       \item Aproximately 52\% of the reported stack traces were associated to programming mistackes. 
+       \item Approximately 52\% of the reported stack traces were associated to programming mistackes. 
 And 27,71\% of all stack traces of reported a java.lang.NullPointerException the root cause.
 
-     \item  Cross-type exception wrappings (e.g., an OutOfMemoryError wrapped in a checked exception)
+       \item  4,4% of the reported exception stack traces ``undocumented runtime exceptions" signaled by third-party code. 
+             (35 methods from Android platform and 44 from libraries).
+
+     \item  Few cross-type exception wrappings, such as an OutOfMemoryError wrapped in a checked exception,
 were detected on stack traces. Indicating that the purpose of Java's hybrid exception model may not have been adequately used.
-
-   \item Undocumented runtime exceptions signaled by third-party code (i.e., libraries and Android platform). 
-    In this study, none of the third-party methods explicitly signaling runtime exceptions included such exceptions
-    on ``exception specifications" (i.e., a list of exceptions that a method might raise during its execution).
 
@@
 excludes the implicitly signaled runtime exceptions related to programming
-mistakes} as carefully as checked exceptions, specialy when developing
+mistakes} as carefully as checked exceptions, especially when developing
  library/framework reusable code. Doing so, the clients of such ibrary/framework
@@
 \label{sec:study}
+
+\begin{figure*} \centering \includegraphics[scale=0.3]{studyprocess2.png}
+\caption{Study overview.} \label{fig:overview} \end{figure*}
 
@@
 analysis. Such combination was needed to enable the identification of 
-fragilities on the exception-related code.
-Moreover, manual inspections were also used to leverage
+fragilities on the exception-related code. Moreover, manual inspections were also used to leverage
  the understanding of stack traces and support further discussions and insights.
 In this study we explore the domain quantitatively and highlight interesting cases by 
-exploring cases qualitatively.
+exploring cases qualitatively.  Figure~ref{fig:overview} gives an overview of the study, whose tasks are detailed next.
 
@@
     \bfseries{android} & If the exception is thrown on a method defined in Android Platform.\\
-    \bfseries{app}     & If the exception is thrown on an application method.\\
+    \bfseries{app}     & If the exception is thrown on an method defined on an Android app.\\
     \bfseries{libcore} & If the exception is thrown on one of the core libraries reused by Android (i.e., org.apache.harmony, org.w3c.dom, sun.misc, org.apache.http, org.json, org.xml). \\
@@
 
-Table ~\ref{tab:causes} summarized the results. It also illustrates the number of general Java exceptions - 
+Table ~\ref{tab:causes} summarizes the results. It also illustrates the number of general Java exceptions - 
 whose full name does not carry the reason for the exception to be thrown.
@@
       Programming Mistakes &  51,96\%\\ 
-      Memory Restriction   &   4,10\% \\ 
+      %Memory Restriction   &   4,10\% \\ 
       General Exceptions   &  6,70\%\\
@@
 \noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
-Aproximately 52\% of the reported stack traces were associated to programming mistackes. 
+Approximately 52\% of the reported stack traces were associated to programming mistackes. 
 And 27,71\% of all stack traces of reported a java.lang.NullPointerException the root cause.
@@
 explicitly signaling a runtime exception not-related to programming mistakes.
-As a result of this inspection, we could observe that only 1 method out of 118 inspected methods
- (i.e., 0,8\%) documented the explicitly thrown runtime exception in a Javadoc comment; and none of these methods
-included the runtime exception  (reported on the stack) as part of the exception interface of the method (i.e., using 
-throws clause on method signature). This result is aligned with the results of other study conducted by from 
+Table~\ref{tab:runtimeinterface} present the result of this inspection. 
+We could find 79 methods (both from libraries and Android Platform) that  explicitly threw a runtime exception 
+without listing such exception on the exception interface (i.e., using 
+throws clause on method signature). These methods were responsible for 
+267 exception stack traces mined in this study.
+
+This result is aligned with the results of other study conducted by from 
 Sacramento et al ~\cite{sacramento2006unchecked} which observed that the
@@
 
+Moreover, these stack traces could also pointed out another best practice violation: all android methods
+ presented in Table~\ref{tab:runtimeinterface} were signaling direct instances of RuntimeException. 
+Using such general exception, is considered a bad according to Java specification and compiled guidelines.
+
 \noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
- Only 0,8\% of the methods documented the explicitly signaled runtime exceptions. 
+ 267 exception stack traces were caused by ``undocumented runtime exceptions" thrown by libraries or the Android Platform.
 }
 \end{minipage}}
+
+\begin{table}
+\centering
+
+ %\begin{tabular}{|p{7.9cm} p{0.1cm}|}
+\begin{tabular}{lrr}
+    \hline
+ \bfseries{Origin}   & \bfseries{methods} &  \bfseries{stacks}  \\ 
+\hline					
+Libraries	& 	44	& 	205	\\					
+\emph{android packages}   &  &   \\ 
+android.app	& 	5	& 	16	\\
+android.content	& 	2	& 	2	\\
+android.database	& 	1	& 	1	\\
+android.graphics	& 	5	& 	11	\\
+android.hardware	& 	4	& 	5	\\
+android.media	& 	3	& 	3	\\
+android.opengl	& 	1	& 	8	\\
+android.os	& 	11	& 	12	\\
+android.text	& 	1	& 	1	\\
+android.video	& 	1	& 	1	\\
+android.providers	& 	1	& 	2	\\
+\hline					
+\bfseries{All android} 	&	35	&	62	\\
+\hline					
+\bfseries{All}	&	79	&	267	\\
+\hline					
+  \end{tabular}
+\caption{Absence of exception interfaces on methods.}
+\label{tab:runtimeinterface}
+\end{table}
 
@@
 leading programmers and the compiler to think that no checked exceptions can be thrown.
- However, the C-code implementation did throw a "checked exception" NoSuchMethodException, 
+ However, the C-code implementation did throw a ``checked exception" called NoSuchMethodException, 
 violating the declaration. The Java compiler, could not detect this violation, because it does 
@@
 on the ``exception interface" of the signaler methods.
+Methods defined on 11 android packages signaled direct instances of RuntimeException.
 }
@@
 
+Table~\ref{tab:wrappingandroid} presents the wrappings found in this study, that
+envolve different exception types (i.e., Error, Checked, Exception). We call 
+such a wrapping a ``cross-type wrapping".
+
 Java is the only language that provides a hybrid exception model 
 which offers three kinds of exceptions each one holding an intended exception behavior (i.e., error, 
-runtime and checked). The best practices I and II are related to such type-related-behavior: 
-(i) Errors should not be handled inside the system, since they usually represent unrecoverable conditions detected 
-by the runtime environment (e.g.,  OutOfMemoryError); (ii) checked exceptions, on the other hand, should represent 
-recoverable conditions.
-
-Table~\ref{tab:wrappingandroid} presents the wrappings found in this study, that
-envolve different exception types (i.e., Error, Checked, Exception). We call 
-such a wrapping a ``cross-type wrapping". Next we exemplify some of such
+runtime and checked).
+
+Next we exemplify some of such
  wrappings and discuss why some of them may negatively impact the robustness.
@@
 by android.os.AsyncTask. This class is responsible for dealing with asynchronous tasks in Android.
-The code snippet is illustrated in Figure ~ref{wrappings} (c) shows that any instance of Throwable signaled 
-in the context of an asynchronous task is converted in to a runtime and re-thrown.  Table ~\ref{tab:exampeswrap} exemplifies some of such wrappings.
+The code snippet below show a general cach clause that convertes any Throwable in to a runtime and re-throw.  
+Table ~\ref{tab:exampeswrap} exemplifies some of such wrappings.
 Such wrapping masks an unrecoverable error in a general exception - doing so it that can be mistakenly handled.
+
+{\footnotesize
+\begin{verbatim}
+   try {
+      ...
+   } catch (InterruptedException e) {
+      android.util.Log.w(..., e);
+   } catch (ExecutionException e) {
+      throw new RuntimeException("...",e.getCause());
+   } catch (CancellationException e) {
+      ...
+   } catch (Throwable t) {
+     throw new RuntimeException("...", t);
+   }
+\end{verbatim}
+}
 
@@
 
-\begin{figure*} \centering \includegraphics[scale=0.55]{examples_new2.png}
-\caption{Code snippets illustrating fragilities on the exception-related code} \label{fig:snippets} \end{figure*}
+%\begin{figure*} \centering \includegraphics[scale=0.55]{examples_new2.png}
+%\caption{Code snippets illustrating fragilities on the exception-related code} \label{fig:snippets} \end{figure*}
 
@@
 observed that the actions inside handlers were very simple (e.g., logging and present a
-message to the user). Pingyu and Elbaum~\cite{Zhang12} 
-analyzed the bug reports of 5 Android applications an observed that 29\% had to do with poor
+message to the user). Pingyu and Elbaum~\cite{Zhang12} were the first to perform
+an empirical investigation of issues (related to exception-related bugs) on Android projects.  
+They perform a small scale study on which they manually inspected the issues of 
+5 Android applications. They observed that 29\% had to do with poor
 exceptional handling code, this empirical study was used to motivate the development of a tool
 aiming at amplifying existing tests to validate exception 
-handling code associated with external resources. Our work mined the exception
-stack traces embedded on issues reported 639 opensource Android projects
+handling code associated with external resources. This work inspired ours,
+ which automaticaly mined the exception stack traces embedded on issues 
+reported on 639 opensource Android projects. The goal of our study was
 to identify common fragilities on the exception related code that can lead to 
@@
 (window) of an Android app.
-Figure ~\ref{fig:snippets} (a) presents a code snippet of one the methods of 
+The code snippet illustrates one the methods of 
 ActivityThreat,  the one responsible for invoking the Activitiy's onCreate method. 
@@
 methods defined in Android, Application and library classes).
+
+exemplos:
+
+(java.lang.RuntimeException,android.hardware.Camera.native_autoFocus(Native Method)")"

### 2014-09-06 android-stacks.tex
all over

@@
 Developers and users of open-source projects uses such provided
- infra-strucute to share information (e.g., exception stack traces~\cite{bettenburg2008makes,schroter2010stack}) across 
+ infrastrucute to share information (e.g., exception stack traces~\cite{bettenburg2008makes,schroter2010stack}) across 
 the software development.
@@
 reported on Android projects hosted in GitHub and Google Code. 
-The goal of this study was to: investigate whether the reported exception stack traces
+The goal of this study was to investigate whether the reported exception stack traces
  can reveal common fragilities on the exception-related code.  
@@
 The Android Application Framework also provides several supporting APIs, such as 
-the APIs to support concurrent development (e.g. AsyncTask). 
-Such APIs particularly useful due to the high demand for responsiveness 
+the APIs to support concurrent development (e.g. AsyncTask) - 
+which are usually needed due to the high demand for responsiveness 
 in mobile environment (mobile apps can easily become 
@@
 can be either checked or runtime. The Java specification~\cite{gosling2000java} 
-suggests that user-defined exceptions should be checked (because doing so 
-the callers of a method will know about the exceptions that a method can throw and so 
-they can decide what to do about them). However, runtime exceptions can also be 
-used to represent user-defined exceptional conditions.In fact, there is a long-lasting debate 
+suggests that user-defined exceptions should be checked, because doing so 
+the callers of a method will know about the exceptions that a method can throw, and then 
+they can decide what to do about them. However, runtime exceptions can also be 
+used to represent user-defined exceptional conditions.There is a long-lasting debate 
 about the pros and cons of both approaches~\cite{javatut,stackoverlow,debate}
@@
 % \caption{Exception Hierarchy in Java} \label{fig:exchier} \end{figure}
-
-%In Java an exception can be thrown in one of the following
-%circumstances~\cite{gosling2000java}: 
-%(i) explicitly thrown when a throw statement is executed; 
-%(ii) implicitly thrown by the JVM when the evaluation of an expression
-% violates the normal semantics of language, also referred as a coding error
-% (e.g., out-of-bounds array index, division-by-zero, access to a null reference); or 
-%(iii) implicitly thrown by the JVM due to an internal error or resource limitation (e.g.,
-%OutofMemoryError). 
-%\todo{This sub-section is quite similar to the section introduction, shouldn't
-%we merge them?}
-
-%When an exception is thrown, it causes a transfer of control from the point
-%where the exception occurred to a point that can be specified by the programmer
-%(exception handler) - in Java it is represented by the try-catch block .
 
@@
 
-%<AVALIAR MOVER ESTE STACK TRACES PARA UMA FIGURA QUE MOSTR COMO ELE É QUEBRADO...
-%PELO EXCEPTION MINER E AS INFORMAÇÕES VAO PARA O BANCO...>
-
-% \subsection{Public Project Repositories and Issue Trackers} 
-% ONE OF THE REVIEWERS COMPLAINED ABOUT THE ABSENCE OF EXPLANATIONS OF WHAT AN ISSU IS...
-% As mentioned before, currently more and more projects are becoming available in 
-% public project repositories such as GitHub and Google Code.
-% Such repositories provides facilities such as....
-% The issue trackers available in such projects ....
-
 \subsection{Best Practices}
@@
 mistakes} as carefully as checked exceptions, especially when developing
- library/framework reusable code. Doing so, the clients of such ibrary/framework
+ library/framework reusable code. Doing so, the clients of such library/framework
 will be aware of all the exceptions a method can throw and can design the code
@@
 
-To support this investigation, we developed a tool called ExceptionMiner (see Section ~\ref{sec:exceptionminer})
-which extracts the exception stack traces embeeded on issues, 
+To support this investigation, we developed a tool called ExceptionMiner (Section~\ref{sec:exceptionminer})
+which extracts the exception stack traces embedded on issues, 
 and combines stack trace information with source code and bytecode 
@@
 repositories are offered a pre-defined set of labels, but repository owners can modify 
-them at will. Therefore, an issue may have none ore an arbitrary set of labels depending 
-on which repository it was created. Table ~\ref{tab:lables} illustrates the ocurrences of different lables 
+them at will. Therefore, an issue may have none or an arbitrary set of labels depending 
+on which repository it was created. Table~\ref{tab:lables} illustrates the ocurrences of different labels 
 on the issues including exception stack traces.
@@
   \end{tabular}
-  \caption{Lables on issues including exception stack traces.}
+  \caption{Labels on issues including exception stack traces.}
   \label{tab:lables}
@@
 Google Code contains largely used open-source Android apps (e.g. K9Mail \footnote{K9Mail moved to Github but as a way of not loosing the project history 
-it advices their users to report bugs on Google Code issue tracker.}), however, differently from GitHub, Google Code does not provide an API to access the information related
+it advises their users to report bugs on Google Code issue tracker.}), however, differently from GitHub, Google Code does not provide an API to access the information related
  to hosted projects \footnote{Google code used to provided a Web service to such repository, but it was deactivated in June 2013 what Google called a "clean-up action".}.
@@
 (e.g. Error, Exception or RuntimeException). The module responsible for this analysis 
-uses Design Wizard~\cite{Brunet09} framework to inspect the bytecode of all exceptions reported on issues.
-The nidyke walks up the type hierarchy of a Java exception until it reaches a base exception type.
-This module analyzed the exception hierarchy of all exceptions defined on:
-Android Platform (API level 19) - which includes all basic Java exceptions that 
-can be thrown during the app execution. The exceptions thrown inside the apps
-were classified using manual inspections as well as the exceptions thrown only
-on previous versions of the Android platform.
+uses Design Wizard~\cite{Brunet09} framework to inspect the bytecode of all exceptions reported on issues,
+and a specific implementation to analise the source code of Java Exception classes.
+It walks up the type hierarchy of a Java exception until it reaches a base exception type.
+This module analyzed all exceptions defined on Android Platform (API level 19) - 
+which includes all basic Java exceptions that can be thrown during the app execution,
+and exceptions thrown by Android core libraries. 
+Exceptions thrown inside the apps and libraries were classified through manual inspections 
+as well as the exceptions thrown only on previous versions of the Android platform
+(only the exceptions reported on stack traces needed to be inspected).
 
@@
  from libraries if their packages were not defined 
-on Android Platform, core libraries nor on applications.
-Table ~\ref{signalers} summarizes the signaler classification.
+on Android Platform, or on core libraries or on applications.
+Table~\ref{tab:signalers} summarizes the signaler classification.
 
@@
 \subsection{Manual Inspections}
-In the context of the ExceptionMiner tool, manual inspections were used:
-to help on the identification of packages composing the Android platform, 
-libs and apps analyzed in this study (as described previously); and to 
+In the context of the ExceptionMiner tool, manual inspections were used to:
+(i) support on the identification of packages composing the Android platform, 
+libs and apps analyzed in this study (as described previously); and (ii)  
 identify the type of some exceptions reported on issues, 
-that could not be automatically identified by the ExceptionMiner tool.
-Some exceptions could not be automaticaly identified because they were 
-not defined on the last version of the application (the one analyzed by ExceptionMiner),
-or were defined on libraries not found among application resources.
-In such cases we manually inspect the sourcecode or javadoc
- documentation available online of each of them. 
-When the exception was not found on the previous versions, the exception where classified 
+that were not be automatically identified by the ExceptionMiner tool.
+When the exception was not found automatically, nor manually (because they were related an app or lib the previous versions
+that were no more availble), the exception were classified 
 in our study as "Undefined".  Only 31 exceptions 
@@
 quantitative and qualitative analysis of the outcomes. The structure of this section
- is centered on questions that guied the search for fragilities (on the exception-related code)
+ is centered on questions that guided the search for fragilities (on the exception-related code)
 based on exception stack trace information.
@@
 the the exceptions commonly reported as the root causes of stack traces.
-Table \ref{tab:topten} presents a list the top 10 root exceptions found in the study - 
+Table \ref{tab:topten} presents a list of the top 10 root exceptions found in the study - 
  ranked by the number of distinct projects on which they were reported. 
-The table shows how many times the signaler of such exception was a method defined on
+This table also shows how many times the signaler of such exception was a method defined on
 the Android Platform, the Android libcore, the application itself or a third-party library -
@@
  or resource limitation (e.g., OutOfMemoryError). 
-From this set the java.lang.NullPointerException was the most reported root cause (27,71\% ). 
-If we consider the frequencly of NullPointerException 
+From this set the java.lang.NullPointerException was the most reported root cause (27,71\%). 
+If we consider the frequency of NullPointerException 
 across projects, we can observe that 51,96\% of all projects reported at least one exception stack 
@@
 the root exception types (i.e., RuntimeException, Error, checked Exception) as well as its origins - which were 
-identified based on the package names of the signalers related to it on the stack traces (see Section ~\ref{sec:exceptionminer}).
+identified based on the package names of the signalers related to it on the stack traces (Section~\ref{sec:exceptionminer}).
 Table~\ref{tab:typeroottab} presents the types and origins of root exceptions of all stack traces analyzed.

### 2014-09-06 android-stacks.tex
all over

@@
 revealing that Java hybrid exception model is not fully used according to its purpose.
-Moreover, about 50\% of all analyzed projects reported at least one NullPointerExeption
+Moreover, about 50\% of all analyzed projects reported at least one NullPointerException
 as the root cause of an exception stack trace.
@@
 which includes all basic Java exceptions that can be thrown during the app execution,
-and exceptions thrown by Android core libraries. 
-Exceptions thrown inside the apps and libraries were classified through manual inspections 
-as well as the exceptions thrown only on previous versions of the Android platform
-(only the exceptions reported on stack traces needed to be inspected).
+and exceptions thrown by Android core libraries. Moreover it also analized the 
+libraries and applications which defined exceptions reported on stack traces
+(the tool only analyzed the last version available).
 
@@
 identify the type of some exceptions reported on issues, 
-that were not be automatically identified by the ExceptionMiner tool.
-When the exception was not found automatically, nor manually (because they were related an app or lib the previous versions
-that were no more availble), the exception were classified 
+that were not be automatically identified by the ExceptionMiner tool
+(because they were defined on previous versions of libraries,
+apps and Android Plaform). When the exception could not be  
+found niether automatically, nor manually (because they were related an app or lib the previous versions
+that were no more availble) the exception were classified 
 in our study as "Undefined".  Only 31 exceptions 
@@
 After distilling the information available on the exception stack traces we could find 
-the the exceptions commonly reported as the root causes of stack traces.
+the exceptions commonly reported as the root causes of stack traces.
 Table \ref{tab:topten} presents a list of the top 10 root exceptions found in the study - 
@@
 the Android Platform, the Android libcore, the application itself or a third-party library -
- following the classification as presented in Table  \ref{tab:signalers}.
+ following the classification presented in Table  \ref{tab:signalers}.
 
@@
 
- The NullPointerExceptions are mainly signaled inside the applicaiton code and the Android platform,
- although we could also find NullPointerExceptions being signaled by third-party libraries. 
-Regarding reusable code, there is no consensus whether it is a good or bad practice to 
+ The NullPointerException was mainly signaled inside the applicaiton code (50\%) and the Android platform (31,5\%),
+ although we could also find the NullPointerException being signaled by third-party libraries (16,3\%). 
+Regarding reusable code (e.g., libraries and frameworks), there is no consensus whether it is a good or bad practice to 
 re-throw a NullPointerException. Some prefer to encapsulate such an exception on
@@
 
-The high prevalence of NullPointerExceptions is aligned with the findings of other 
+The high prevalence of NullPointerException is aligned with the findings of other 
 works~\cite{kim2013predicting,fraser20131600,csallner2004jcrasher}. For instance, Sunghun et
@@
 testing~\cite{maji2012empirical,csallner2004jcrasher} showed that most of the automatically 
-detected bugs were due to NullPointerExceptions and exceptions implicitly-signaled by Java 
+detected bugs were due to NullPointerException and exceptions implicitly-signaled by Java 
 environment due to programming mistakes or resource limitations
@@
 Hence, we conducted an empirical investigation based on the following assumption: the name of the exception and its associated Javadoc 
-documentaion hold enought information to pinpont a ``programming mistake".To exemplify, an instance of ArrayOutOfBoundException
+documentation hold enough information to pinpont a ``programming mistake".To exemplify, an instance of ArrayOutOfBoundException
  pinpoints to programming mistake which indicates that an array has been accessed with an illegal index.
@@
 whose full name does not carry the reason for the exception to be thrown.
-To ensure the quality of the analysis, three independent coders classified a randomly selected
+To ensure the quality of the analysis, three independent developers classified a randomly selected
 sample of 25 exception types (from the total 100) using the same list of concerns;
-the resulting interrater agreement was high (96\%). We could also observe that
-We can therefore assume that the exception names and Javadoc 
-were enough for classifying the exception types analyzed.
-
-The results show that the programming mistakes associated to the most reported root causes. 
+the resulting interrater agreement was high (96\%).
+
+The results show that the programming mistakes are associated to the most reported root causes. 
 Although there is usually no proper handling for such exceptions, 
@@
 
-We can see, from Table ~\ref{tab:typeroottab}, that contrating with the other origins, most of the 
+We can see, from Table ~\ref{tab:typeroottab}, that contrasting with the other origins, most of the 
  exceptions signaled on Android Libcore (i.e., the set of libraries reused by the Android) were 
@@
 org.apache.http, org.json, org.xml, javax. Signaling checked exceptions is considered a 
-good practice (see best practice IV in Section~\ref{sec:best}) because by using checked exceptions a library can define a precise 
+good practice (see best practice IV in Section~\ref{sec:best}) because by using 
+checked exceptions a library can define a precise 
 exception interface ~\cite{miller1997issues} to its clients.
@@
 
-This result is aligned with the results of other study conducted by from 
+This result is aligned with the results of other study conducted by  
 Sacramento et al ~\cite{sacramento2006unchecked} which observed that the
@@
  267 exception stack traces were caused by ``undocumented runtime exceptions" thrown by libraries or the Android Platform.
+Methods defined on 11 android packages signaling direct instances of RuntimeException.
 }
@@
  However, the C-code implementation did throw a ``checked exception" called NoSuchMethodException, 
-violating the declaration. The Java compiler, could not detect this violation, because it does 
+violating the declaration. The Java compiler could not detect this violation, because it does 
 not perform static exception checking on native methods. This type of bug is hard to diagnose
@@
 this exception, such undocumented exception may remain
-uncaught and cause the app to crash, or may be mistakenly handled by subsumption - 
+uncaught and cause the app to crash, or may be mistakenly handled by subsumption.
 The exception stack traces reporting this scenario is actually a real bug of Android 
@@
 \noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
-An unexpected fragility was detected: a checked exception being thown and not specified  
-on the ``exception interface" of the signaler methods.
-Methods defined on 11 android packages signaled direct instances of RuntimeException.
+An unexpected fragility was detected: a checked exception being thown by a native method and not declared  
+on the ``exception interface" of such method and all methods on the invocation chain.
 }
@@
 
-Table~\ref{tab:wrappingandroid} presents the wrappings found in this study, that
-envolve different exception types (i.e., Error, Checked, Exception). We call 
-such a wrapping a ``cross-type wrapping".
-
 Java is the only language that provides a hybrid exception model 
 which offers three kinds of exceptions each one holding an intended exception behavior (i.e., error, 
-runtime and checked).
-
-Next we exemplify some of such
+runtime and checked). Table~\ref{tab:wrappingandroid} presents the wrappings found in this study, that
+envolve different exception types (i.e., Error, Checked, Exception). We call 
+such a wrapping a ``cross-type wrapping". Next we exemplify some of such
  wrappings and discuss why some of them may negatively impact the robustness.
 
-\emph{\textbf{Runtime Exception wrapping Error.}} We could find  64 exception stack traces 
-on 46 projects on which an Error was wrapped in a runtime exception.
-Manually inspecting each exception stack trace, we could observe that 44\% of the wrappings were performed 
+\emph{\textbf{Runtime Exception wrapping Error.}} We could observe that 44\% of the wrappings were performed 
 by android.os.AsyncTask. This class is responsible for dealing with asynchronous tasks in Android.
@@
 
+\emph{\textbf{Runtime wrapping Checked Exception.}} This wrapping was responsible for 49.5\% of
+the cross-type wrappings.  From this set 50\%  were performed on methods defined on Android Platform. 
+We could observe that approximately 41\% of them were performed by a single class,
+ named ActivityThreat. This class is responsible for managing the life cycle of  Android Activities - 
+invoking the life cycle methods (i.e. onCreate(), onPause()) for every Activity 
+(window) of an Android app. This class contains a such try-catch block that
+handled any instance of java.lang.Exception signaled in the context 
+of the method in a general Runtime exception and re-throws.
+
 We could also observe that some stack traces included several intermediate cross-type wrappings.
@@
  and lead to what we call ``exception confusion problem". To illustrate this problem we can use one of the wrappings discussed above.
-When the developer is confronted with a checked exception, the designer of the API is telling him 
+When the developer is confronted with a checked exception, the designer of the API is telling him/her 
 to handle the exceptional condition (according to Java Specification and 
@@
 
- "Everybody hates thinking about exceptions, because they’re not supposed to happen"
+ "Everybody hates thinking about exceptions, because they are not supposed to happen"
   Brian Foote
@@
 contextual information to adequately handle the exception, 
-and only tasks possible are: to present a message to the user
+a possible handler task is: to present a message to the user
  and restart the application. However, such handler cannot replace a carefully designed exception 
@@
 \emph{``I call it my billion-dollar mistake. It was the invention of the null reference in 1965 [...] This has led to innumerable errors, vulnerabilities, and system 
-crashes, which have probably caused a billion dollars of pain and damage in the last forty year."}
-
-In this sudy, the null references were, infact, responsible for several reported application errors - providing more evidences to Hoare's speach.
+crashes, which have probably caused a billion dollars of pain and damage in the last forty years."}
+
+In this sudy, the null references were, infact, responsible for several reported application errors - providing more evidences to Hoare's speech.
 This observation emphasizes the importance of solutions to avoid NullPointerExceptions, such as:
@@
 (ii) inter-method null pointer analysis tools such as the one proposed by Nanda and Sinha ~\cite{nanda2009accurate};
-or (iii) whether language designs which avoid null pointers, such 
+or (iii) language designs which avoid null pointers, such 
 as Monads ~\cite{Walde95} (i.e., used in functional languages for values that may not be available 
@@
 we calculated the interrater agreement after three independent 
-coders classified a randomly selected sample (of 25 exception 
+developers classified a randomly selected sample (of 25 exception 
 types from the total 100); the interater agreement was high (96\%). 

### 2014-09-06 android-stacks.tex
Resolved language issues in intro.
And shortened it a bit.

@@
 
-Along the recent years, we have witnessed an astonishing increase in the number of
-mobile applications. If on the one hand such applications extend phones capabilities 
-far beyond of the basic calls and textual messages, on the other hand
-they have to face an increasing number of threats to application robustness
- arising from: faults on underlying middleware and hardware (e.g., camera, sensors);
-faults on third party services and libraries; compatibility issues ~\cite{McDon13}; 
- memory and battery restrictions; noisy external resources ~\cite{Zhang12} (e.g., wireless connections,GPS, bluetooth) among others.
+In recent years, we have witnessed an astonishing increase in the number of
+mobile applications. These applications extend phones capabilities 
+far beyond basic calls and textual messages. Such applications, however,
+have to face an increasing number of threats to application robustness
+ arising from: failures in underlying middleware and hardware (e.g., camera, sensors);
+failures in third party services and libraries; compatibility issues~\cite{McDon13}; 
+ memory and battery restrictions; and noisy external resources ~\cite{Zhang12} (e.g., wireless connections,GPS, bluetooth).
 
 Therefore, techniques for error detection and handling are not  an optional add-on but a 
-fundamental part of such apps. The exception handling mechanism ~\cite{goodenough1975exception},
+fundamental part of such apps. The exception handling mechanism~\cite{goodenough1975exception},
 embedded in many mainstream programming languages, such as Java, C++ and C\#,
  is one of the most used techniques for detecting and recovering from such exceptional conditions.
-Android, the leading plaform for mobile applications~\cite{gartner} (thanks to its open-source 
-model and hardware partners like Samsung, HTC, Motorola, Asus besides Nexus Google's
- own device)  reuses the embedded Java exception model to detect and handle 
- exceptional conditions.
-
-Studies have shown, however, that the exception-related code is generally poorly understood and the
- less tested parts of the system~\cite{miller1997issues,Robil00,shah2010understanding, 
+In this paper we will be concerned with exception handling in Android apps,
+which reuse Java's exception handling model.
+
+%Android, the leading plaform for mobile applications~\cite{gartner} (thanks to its open-source 
+%model and hardware partners like Samsung, HTC, Motorola, Asus besides Nexus Google's
+% own device)  reuses the embedded Java exception model to detect and handle 
+% exceptional conditions.
+
+Studies have shown that exception-related code is generally poorly understood and among the least tested parts of the system~\cite{miller1997issues,Robil00,shah2010understanding, 
 garcia2007extracting,garcia2001comparative,cabral2007exception,coelho2011unveiling}.
-As a consequence they may affect the system the other way around.
-The exception-related code may favor the introduction of failures such as 
+As a consequence they may inadvertently negatively affect the system: Exception-related code may introduce failures such as 
 uncaught exceptions~\cite{jo2004uncaught, Zhang12} - 
@@
 
-Until recent years, the availability of data, reporting such failures, was scarce. 
-It comprised only few large scale projects (e.g. the Mozilla crash report dataset\footnote{https://code.google.com/p/promisedata/issues/detail?id=88}).
-Recently we have observed a rapid increase in the number of open
-source projects that make their version control systems as well as the issue tracking systems
-publicly available in repository hosting sites such as GitHub and Google Code.
-Developers and users of open-source projects uses such provided
- infrastrucute to share information (e.g., exception stack traces~\cite{bettenburg2008makes,schroter2010stack}) across 
-the software development.
+%% I am not so sure this data was not available. Let's omit this paragraph
+%Until recently, the availability of data reporting such failures was scarce. 
+%It comprised only few large scale projects (e.g. the Mozilla crash report dataset\footnote{https://code.google.com/p/promisedata/issues/detail?id=88}).
+%Recently we have observed a rapid increase in the number of open
+%source projects that make their version control systems as well as the issue tracking systems
+%publicly available in repository hosting sites such as GitHub and Google Code.
+%Developers and users of open-source projects uses such provided
+% infrastrucute to share information (e.g., exception stack %traces~\cite{bettenburg2008makes,schroter2010stack}) across 
+%the software development.
 
@@
 
-This study performs a post mortem analysis of the exception stack traces embedded on issues
+This study performs a post mortem analysis of the exception stack traces 
+included in issues
 reported on Android projects hosted in GitHub and Google Code. 
-The goal of this study was to investigate whether the reported exception stack traces
- can reveal common fragilities on the exception-related code.  
-We call ``fragilities" characteristics on the exception-related code 
-that favor the introduction of failures such as uncaught 
-exceptions -  contributing to decrease the applicaiton robustness.
+The goal of this study is to investigate whether the reported exception stack traces
+ can reveal common fragilities in the exception-related code.  
+We call ``fragilities'' characteristics of the exception-related code 
+that increase the likelihood of introducing failures such as uncaught 
+exceptions -  leading to a decrease in application robustness.
+
 To guide this investigation we compiled general guidelines on how to use
@@
 Overall 159,048 issues were analyzed and 6,005 stack traces were extracted from them.
-The exception stack trace analysis was augmented with additional information
-extracted from JDK, Android Plaform and the applications using byte code 
-and source code analysis. Manual inspection was used to leverage the understanding 
-of exception stack traces and support further discussions and insights.
-
-Some outcomes consistently detected during this study were the following:
+The exception stack trace analysis was augmented by means of byte code 
+and source code analysis
+ with additional exception-related information
+extracted from the JDK, Android Platform and the application layer.
+%Manual inspection was used to leverage the understanding 
+%of exception stack traces and support further discussions and insights.
+
+Outcomes of our study include:
+
+% Some outcomes consistently detected during this study were the following:
 
@@
 
-       \item Approximately 52\% of the reported stack traces were associated to programming mistackes. 
-And 27,71\% of all stack traces of reported a java.lang.NullPointerException the root cause.
-
-       \item  4,4% of the reported exception stack traces ``undocumented runtime exceptions" signaled by third-party code. 
-             (35 methods from Android platform and 44 from libraries).
-
-     \item  Few cross-type exception wrappings, such as an OutOfMemoryError wrapped in a checked exception,
-were detected on stack traces. Indicating that the purpose of Java's hybrid exception model may not have been adequately used.
+       \item Approximately 52\% of the reported stack traces can be attributed to programming mistakes. 
+In particular, 27.71\% of all stack traces reported a \texttt{java.lang.NullPointerException} as its root cause.
+
+       \item  4.4\% of the reported exception stack traces are ``undocumented runtime exceptions'' raised by third-party code. 
+             (35 methods from the Android platform and 44 from libraries).
+
+     \item  We detected several cross-type exception wrappings, such as an OutOfMemoryError wrapped in a checked exception,
+in stack traces. This suggests that the purpose of Java's hybrid exception model comprising checked and unchecked exceptions may not have been adequately used.
 
    \item Undocumented checked exception signaled by a  JNI interface function defined on Android Platform. 
-    Some flows reported a checked exception thown by a JNI interface function, and  such exception was not  
-    part of the ``exception interface" of JNI function nor the methods on its invokation chain - it can lead to 
+    Some flows reported a checked exception thrown by a JNI interface function, yet this exception was not  
+    part of the ``exception interface'' of the JNI function nor the methods on its invocation chain. This can lead to 
    uncaught exceptions that are difficult to debug.
@@
 but also to the developement of any robust Java-based system. For instance, in the absence of the 
-``exception specification" of  a third-party code, it is difficult or 
-even enfeasible for the developer to protect the code against ``unforeseen" exceptions. 
-Since, in such cases the client usually do not have access to the source code, such undocumented exceptions may 
+``exception specification" of third-party code, it is difficult or 
+even infeasible for the developer to protect the code against ``unforeseen'' exceptions. 
+Since in such cases the client usually does not have access to the source code, such undocumented exceptions may 
 remain uncaught and lead to system crashes. Moreover, the cross-type exception wrappings 
@@
 may lead to unintended handling problems - e.g, trying to handle an instance of OutOfMemoryError ``hiden" in 
-a checked exception may lead the program to an unpredictable state.
+a checked exception may bring the program to an unpredictable state.
 
 Hence, the study results are relevant to Android and 
-Java developers who may underestimate how likely are the
- exceptions related to programming mistakes, and how 
+Java developers who may underestimate how 
+ exceptions are related to programming mistakes, and how 
  tricky it can be to prevent the uncaught exceptions from third-party code
- in the absence of ``exception interfaces". Moreover, results are relevant to 
+ in the absence of ``exception interfaces". Moreover, the results are relevant to 
 designers of languages and tools to consider ways of reducing such fragilities 
@@
 
-The remaining of this paper is organized as follows. Section 2 presents a
+The remaining of this paper is organized as follows. 
+Section~\ref{sec:back} presents a
 background on Android Platform and Java exception model. 
-Section 3 presents the study design. Section 4 reports the study findings. 
-Section 5 provides further discussions and insights.
-Section 6 presents the threats to validity associated to this study. Finally Section
-7 describes the related work, and Section 8 concludes the paper and outlines
+Section~\ref{sec:study} presents the study design. 
+Section~\ref{sec:exceptionminer} describes the ExceptionMiner tool we developed to conduct our study.
+Section~\ref{sec:result} reports the findings of our study.
+Section~\ref{sec:disc} provides a discussion of the wider implications of our results.
+Section~\ref{sec:threats} presents the threats to validity associated to this study. 
+Finally Section~\ref{sec:rele} describes related work, 
+and Section~\ref{sec:conc} concludes the paper and outlines
 directions for future work.
@@
 
-\begin{figure*} \centering \includegraphics[scale=0.3]{studyprocess2.png}
+\begin{figure*} \centering \includegraphics[scale=0.3]{studyoverview3.png}
 \caption{Study overview.} \label{fig:overview} \end{figure*}
@@
 Google Code contains largely used open-source Android apps (e.g. K9Mail \footnote{K9Mail moved to Github but as a way of not loosing the project history 
-it advises their users to report bugs on Google Code issue tracker.}), however, differently from GitHub, Google Code does not provide an API to access the information related
+it advises their users to report bugs on Google Code issue tracker: https://github.com/k9mail/k-9/wiki/LoggingErrors.}),
+ however, differently from GitHub, Google Code does not provide an API to access the information related
  to hosted projects \footnote{Google code used to provided a Web service to such repository, but it was deactivated in June 2013 what Google called a "clean-up action".}.
@@
 revealing that Java hybrid exception model is not fully used according to its purpose.
-As we can see, such fragilities can negatively affect not only the robustness of Android application 
+Such fragilities can negatively affect not only the robustness of Android application 
 but also the robustness of any Java-based application. 

### 2014-09-06 android-stacks.tex
Shortened abstract.
And fixed a couple of English issues.

--- 2014-09-06 android-stacks.tex
@@
 \begin{abstract}
-The number of mobile apps is increasing in a daily rate. If on the one hand they
-extend phones capabilities far beyond of the basic calls, on the other hand they have to 
-cope with an increasing number of exceptional conditions 
-(e.g., faults in underlying middleware or hardware; memory and battery restrictions). 
-Therefore, mechanisms for exception detection and handling are not an optional add-on but a 
-fundamental part of such apps. However, studies have shown that such mechanisms are the less 
-understood and tested parts of the system. As a consequence they may affect the system the 
-other way around, making the system even less robust, leading to failures such as uncaught exceptions 
-and unintended handling (i.e., exceptions being mistakenly caught by existing handlers).
-This paper reports a study which mined the stack traces embedded on 
-159,048 issues reported on Android projects hosted in GitHub (482 projs.) 
-and Google Code (157 projs.). 
-The goal of this study was to: investigate the whether the exception stack traces
+%The number of mobile apps is increasing in a daily rate. If on the one hand they
+%extend phones capabilities far beyond of the basic calls, on the other hand they have to 
+%cope with an increasing number of exceptional conditions 
+%(e.g., faults in underlying middleware or hardware; memory and battery restrictions). 
+%Therefore, mechanisms for exception detection and handling are not an optional add-on but a 
+%fundamental part of such apps. However, studies have shown that such mechanisms are the less 
+%understood and tested parts of the system. As a consequence they may affect the system the 
+%other way around, making the system even less robust, leading to failures such as uncaught exceptions 
+%and unintended handling (i.e., exceptions being mistakenly caught by existing handlers).
+
+This paper reports a study which mined the Java stack traces included in
+159,048 issues reported on Android projects hosted in GitHub (482 projects) 
+and Google Code (157 projsects). 
+The goal of this study is to investigate  whether the exception handling logic on these stack traces
  can reveal common characteristics on the exception-related code that
-may be favoring the introduction of the aforementioned failures, which would be responsible for 
-decreasing the application robustness. We call such characteristics ``fragilities" of the exception-related code.
+are particularly fault prone and may lead to a decrease in application robustness.
 Overall 6,005 exception stack traces were extracted, and analyzed in 
 combination with source code and bytecode analysis. 
-Some fragilities consistently detected in this study include: 
+Key problems identified in this study include: 
 (1) undocumented runtime exceptions thrown by third party code;
 (2) undocumented checked exceptions thrown by a JNI interface function;
-(3) unexpected exception wrappings (e.g., Errors being wrapped in checked exceptions),
-revealing that Java hybrid exception model is not fully used according to its purpose.
+(3) unexpected exception wrappings (such as Errors being wrapped in checked exceptions),
+indicating that Java's hybrid exception model is not fully used in accordance with its purpose.
 Moreover, about 50\% of all analyzed projects reported at least one NullPointerException
-as the root cause of an exception stack trace.
+as the root cause of an exception stack trace, suggesting the presence of programming mistakes.
 This study provides insights on such fragilities and the robustness threats  
@@
  with additional exception-related information
-extracted from the JDK, Android Platform and the application layer.
+extracted from the Android Platform and Android applications.
+
+Outcomes of our study include:
+
 %Manual inspection was used to leverage the understanding 
 %of exception stack traces and support further discussions and insights.
-
-Outcomes of our study include:
 
@@
 \noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
- 267 exception stack traces were caused by ``undocumented runtime exceptions" thrown by libraries or the Android Platform.
-Methods defined on 11 android packages signaling direct instances of RuntimeException.
+The study revealed: 267 exception stack traces were caused by ``undocumented runtime exceptions" thrown by libraries or the Android Platform; 
+several methods defined on 11 android packages signaling direct instances of RuntimeException.
 }

### 2014-09-06 android-stacks.tex
square content

--- 2014-09-06 android-stacks.tex
@@
 \noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
-Approximately 52\% of the reported stack traces were associated to programming mistackes. 
-And 27,71\% of all stack traces of reported a java.lang.NullPointerException the root cause.
+The study revealed: approximately 52\% of the reported stack traces were associated to programming mistakes;
+and 27,71\% of all stack traces of reported a java.lang.NullPointerException the root cause.
 }
@@
 \noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
-An unexpected fragility was detected: a checked exception being thown by a native method and not declared  
+The study revealed: an unexpected fragility - a checked exception being thown by a native method and not declared  
 on the ``exception interface" of such method and all methods on the invocation chain.
@@
 \noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
-The cross-type wrappings show us that even if the developer sticks to Java
+The cross-type wrappings is an indication that even if the developer sticks to Java
 specification and best practices, it will be dificult for her/him to prevent the code from 

### 2015-02-06 android-stacks.tex
Update my afiliation

--- 2014-09-06 android-stacks.tex
@@
 Georgios Gousios\IEEEauthorrefmark{2},
-Arie van Deursen\IEEEauthorrefmark{2}
+Arie van Deursen\IEEEauthorrefmark{3}
 }
@@
 Email: \texttt{roberta@dimap.ufrn.br,lucas.almeida@ppgsc.ufrn.br}}
-\IEEEauthorblockA{\IEEEauthorrefmark{2}Delft University of Technology\\
+\IEEEauthorblockA{\IEEEauthorrefmark{2}Radboud University Nijmegen\\
+Nijmegen, The Netherlands\\
+Email: \texttt{georgios@cs.ru.nl}}
+\IEEEauthorblockA{\IEEEauthorrefmark{3}Delft University of Technology\\
 Delft, The Netherlands\\
-Email: \texttt{\{g.gousios,arie.vandeursen\}@tudelft.nl}}
+Email: \texttt{arie.vandeursen@tudelft.nl}}
 }
@@
  with additional exception-related information
-extracted from the Android Platform and Android applications.
+extracted from the Android platform and Android applications.
 
@@
        \item Approximately 52\% of the reported stack traces can be attributed to programming mistakes. 
-In particular, 27.71\% of all stack traces reported a \texttt{java.lang.NullPointerException} as its root cause.
+In particular, 27.71\% of all stack traces reported a java.lang.NullPointerException as its root cause.
 
@@
 
-   \item Undocumented checked exception signaled by a  JNI interface function defined on Android Platform. 
+   \item Undocumented checked exception signaled by a  JNI interface function defined on Android platform. 
     Some flows reported a checked exception thrown by a JNI interface function, yet this exception was not  
@@
 
-Such outcomes point to threats not only to the developement of robust Android apps, 
-but also to the developement of any robust Java-based system. For instance, in the absence of the 
+Such outcomes point to threats not only to the development of robust Android apps, 
+but also to the development of any robust Java-based system. For instance, in the absence of the 
 ``exception specification" of third-party code, it is difficult or 
@@
 (e.g., OutOfMemoryError is wrapped in a checked Exception) 
-may lead to unintended handling problems - e.g, trying to handle an instance of OutOfMemoryError ``hiden" in 
+may lead to unintended handling problems - e.g, trying to handle an instance of OutOfMemoryError ``hidden" in 
 a checked exception may bring the program to an unpredictable state.
@@
 Section~\ref{sec:back} presents a
-background on Android Platform and Java exception model. 
+background on Android platform and Java exception model. 
 Section~\ref{sec:study} presents the study design. 
@@
 Among the basic elements that comprise an Android application, we can cite the
- Activity class. The Activity is the main class of an Android app (there is usualy one
+ Activity class. The Activity is the main class of an Android app (there is usually one
 Activity per application window). Each Activity defines a set of hook methods  
@@
 exceptions being thrown by it. According to~\cite{bloch2008effective}, it is
-also wise to document the explicitly thrown runtime exceptions\footnote{This
+also wise to document  (either or on the exception interface or using @throws tag on Javadoc) 
+the explicitly thrown runtime exceptions\footnote{This
 excludes the implicitly signaled runtime exceptions related to programming
-mistakes} as carefully as checked exceptions, especially when developing
+mistakes} as carefully as checked exceptions, e when developing
  library/framework reusable code. Doing so, the clients of such library/framework
@@
 \label{sec:git}
+
+\begin{table}
+  \centering
+  \begin{tabular}{lr|lr}
+    \hline
+     \multicolumn{2}{c}{\bfseries{GitHub}} &  \multicolumn{2}{c}{\bfseries{Google Code}} \\
+      \bfseries{Lable} &  \bfseries{\% Occurrences} &  \bfseries{Lable} &  \bfseries{\% Occurrences} \\
+    \hline
+empty &	54.24\% & Defect &	91.96\% \\
+Defect &	39.56\%  & Enhancement  &	3.16\% \\
+Enhancement &	0.57\% & Task	& 1.37\% \\
+Support &	0.52\% & empty &	1.12\% \\
+Problem &	0.36\% & StackTrace &	0.70\% \\
+Others &	4.74\% &  Others &	1.68\% \\   
+  \hline
+  \end{tabular}
+  \caption{Labels on issues including exception stack traces.}
+  \label{tab:lables}
+\end{table}
 
@@
 
-% \centering
-%  \begin{tabular}{|p{2cm}| p{5cm}|}
-%    \hline
-
-\begin{table}
-  \centering
-  \begin{tabular}{lr|lr}
-    \hline
-     \multicolumn{2}{c}{\bfseries{GitHub}} &  \multicolumn{2}{c}{\bfseries{Google Code}} \\
-      \bfseries{Lable} &  \bfseries{\% Occurrences} &  \bfseries{Lable} &  \bfseries{\% Occurrences} \\
-    \hline
-empty &	54,24\% & Defect &	91,96\% \\
-Defect &	39,56\%  & Enhancement  &	3,16\% \\
-Enhancement &	0,57\% & Task	& 1,37\% \\
-Support &	0,52\% & empty &	1,12\% \\
-Problem &	0,36\% & StackTrace &	0,70\% \\
-Others &	4,74\% &  Others &	1,68\% \\   
-  \hline
-  \end{tabular}
-  \caption{Labels on issues including exception stack traces.}
-  \label{tab:lables}
-\end{table}
-
 Based on the assumption that regardless the issue type, every exception stack
@@
  however, differently from GitHub, Google Code does not provide an API to access the information related
- to hosted projects \footnote{Google code used to provided a Web service to such repository, but it was deactivated in June 2013 what Google called a "clean-up action".}.
+ to hosted projects \footnote{Google code used to provided a Web service to such repository, but it was deactivated in June 2013 what Google called a `clean-up action".}.
 To overcome this limitation we needed to implement a Web Crawler (incorporated in ExceptionMiner tool described next) that navigated 
@@
 It walks up the type hierarchy of a Java exception until it reaches a base exception type.
-This module analyzed all exceptions defined on Android Platform (API level 19) - 
+This module analyzed all exceptions defined on Android platform (API level 19) - 
 which includes all basic Java exceptions that can be thrown during the app execution,
@@
  from libraries if their packages were not defined 
-on Android Platform, or on core libraries or on applications.
+on Android platform, or on core libraries or on applications.
 Table~\ref{tab:signalers} summarizes the signaler classification.
@@
     \hline
-    \bfseries{android} & If the exception is thrown on a method defined in Android Platform.\\
+    \bfseries{android} & If the exception is thrown on a method defined in Android platform.\\
     \bfseries{app}     & If the exception is thrown on an method defined on an Android app.\\
@@
 quantitative and qualitative analysis of the outcomes. The structure of this section
- is centered on questions that guided the search for fragilities (on the exception-related code)
+ is centered on questions that guided the search for fragilities on the exception-related code
 based on exception stack trace information.
@@
 This table also shows how many times the signaler of such exception was a method defined on
-the Android Platform, the Android libcore, the application itself or a third-party library -
+the Android platform, the Android libcore, the application itself or a third-party library -
  following the classification presented in Table  \ref{tab:signalers}.
@@
  or resource limitation (e.g., OutOfMemoryError). 
-From this set the java.lang.NullPointerException was the most reported root cause (27,71\%). 
+From this set the java.lang.NullPointerException was the most reported root cause (27.71\%). 
 If we consider the frequency of NullPointerException 
-across projects, we can observe that 51,96\% of all projects reported at least one exception stack 
+across projects, we can observe that 51.96\% of all projects reported at least one exception stack 
 trace on which the NullPoiterException was the root cause.
 
- The NullPointerException was mainly signaled inside the applicaiton code (50\%) and the Android platform (31,5\%),
- although we could also find the NullPointerException being signaled by third-party libraries (16,3\%). 
+ The NullPointerException was mainly signaled inside the application code (50\%) and the Android platform (31.5\%),
+ although we could also find the NullPointerException being signaled by third-party libraries (16.3\%). 
 Regarding reusable code (e.g., libraries and frameworks), there is no consensus whether it is a good or bad practice to 
@@
 
-The high prevalence of NullPointerException is aligned with the findings of other 
+The high prevalence of NullPointerException is aligned with the findings of other research
 works~\cite{kim2013predicting,fraser20131600,csallner2004jcrasher}. For instance, Sunghun et
-al.~\cite{kim2013predicting} showed that in Eclipse bug report system 38\% the bugs 
-related to exception handling were caused by NullPointerException; other works on robustness 
+al.~\cite{kim2013predicting} showed that 38\% the bugs related to exception handling in Eclipse project
+ were caused by NullPointerException; other works on robustness 
 testing~\cite{maji2012empirical,csallner2004jcrasher} showed that most of the automatically 
@@
 environment due to programming mistakes or resource limitations
- (as the ones found in this study).
+ (as the ones found in our study).
 
@@
 Hence, we conducted an empirical investigation based on the following assumption: the name of the exception and its associated Javadoc 
-documentation hold enough information to pinpont a ``programming mistake".To exemplify, an instance of ArrayOutOfBoundException
+documentation hold enough information to pinpoint a ``programming mistake".To exemplify, an instance of ArrayOutOfBoundException
  pinpoints to programming mistake which indicates that an array has been accessed with an illegal index.
@@
 The results show that the programming mistakes are associated to the most reported root causes. 
-Although there is usually no proper handling for such exceptions, 
-besides presenting an error message to the user and restarting the application ~\footnote{Only 
+Although there is usually no proper handling for such exceptions~\cite{wirfs2006toward}, 
+besides presenting an error message to the user and restarting the application~\footnote{Only 
 high fault tolerant systems need to provide solutions to handle them.}.  
-This finding calls the attention to the following: the high number of programming errors reported 
+This finding calls the attention to the following: the high number of programming mistakes reported 
 on issues are an indication that developers may be underestimating the impact that such 
@@
     \hline
-      Programming Mistakes &  51,96\%\\ 
-      %Memory Restriction   &   4,10\% \\ 
-      General Exceptions   &  6,70\%\\
+      Programming Mistakes &  51.96\%\\ 
+      %Memory Restriction   &   4.10\% \\ 
+      General Exceptions   &  6.70\%\\
     \hline
@@
 \noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
-Approximately 52\% of the reported stack traces were associated to programming mistackes. 
-And 27,71\% of all stack traces of reported a java.lang.NullPointerException the root cause.
+The study revealed: approximately 52\% of the reported stack traces were associated to programming mistakes;
+and 27.71\% of all stack traces reported a java.lang.NullPointerException as the root cause.
 }
@@
 exception interface ~\cite{miller1997issues} to its clients.
- Since such libraries are heavily used  in several projects, such finding can be related to the library's maturity.
+ Since such libraries are heavily used  in several projects, such finding can be related to the libraries maturity.
 
@@
 
-We can also observe that most of the reported exceptions were RuntimeExceptions
-(64,85\%); and most of them were signaled on methods defined on the Application 
-the Android Platform or on libraries.Taking into account the results of empirical study on 
-the prevalende of ``programming erros" we know that a considerable ammount 
+We can also observe that most of the reported exceptions were runtime 
+(64.85\%); and most of them were signaled on methods defined either on the Application (47.3\%)
+or on the Android platform (34.3\%). We could also find runtime exceptions thrown by library code (17.7\%).
+ Taking into account the results of empirical study on the prevalence of ``programming mistakes", 
+we conclude that a considerable amount 
 of such exceptions, are related to the programming mistakes
@@
 explicitly signaling a runtime exception not-related to programming mistakes.
-Table~\ref{tab:runtimeinterface} present the result of this inspection. 
-We could find 79 methods (both from libraries and Android Platform) that  explicitly threw a runtime exception 
+Table~\ref{tab:runtimeinterface} presents the results of this inspection. 
+We could find 79 methods (both from libraries and Android platform) that  explicitly threw a runtime exception 
 without listing such exception on the exception interface (i.e., using 
-throws clause on method signature). These methods were responsible for 
+throws clause on method signature). From this set only one method (defined on a library)
+included an @throws tag on Javadoc - reporting that the given runtime exception
+could be thrown in some contidions. These methods were responsible for 
 267 exception stack traces mined in this study.
@@
 These ``undocumented runtime exceptions" represent a threat to system robustness, specially
-when such exceptions are thrown then the third party code invoked inside the application 
-(e.g. libraries, or framework utility code). In such cases the client usually do not have access to 
+when such exceptions are thrown by a third party code invoked inside the application 
+(e.g. libraries, or framework utility code). In such cases the developer usually does not have access to 
 the source code. Hence in the absence of the exception documentation it is very difficult or even impossible
- for the client to design the application to deal with ``unforeseen runtime exceptions. As a consequence, the
+ for the client to design the application to deal with ``unforeseen" runtime exceptions. As a consequence, the
  undocumented runtime exception may remain uncaught and lead to system crashes.
 
-Moreover, these stack traces could also pointed out another best practice violation: all android methods
- presented in Table~\ref{tab:runtimeinterface} were signaling direct instances of RuntimeException. 
-Using such general exception, is considered a bad according to Java specification and compiled guidelines.
+Moreover, these stack traces could also pointed out another best practice violation: the methods from Android platform
+ presented in Table~\ref{tab:runtimeinterface} signaled direct instances of RuntimeException. 
+Using such general exception, is considered a bad practice according to Java specification and compiled guidelines.
 
 \noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
-The study revealed: 267 exception stack traces were caused by ``undocumented runtime exceptions" thrown by libraries or the Android Platform; 
+The study revealed: 267 exception stack traces were caused by ``undocumented runtime exceptions" thrown by libraries or the Android platform; 
 several methods defined on 11 android packages signaling direct instances of RuntimeException.
@@
 interface of these methods signaling them. Such native method was defined on
- the Android Platform -  which uses Java Native Invocation (JNI) to access 
-native C/C++ code.
-
-This exceptions was thrown by the method getDeclaredMethods defined in java.lang.Class. 
+ the Android platform -  which uses Java Native Invocation (JNI) to access 
+native C/C++ code. This exception was thrown by the method getDeclaredMethods defined in java.lang.Class. 
 The Java-side declaration of this method does not have any throws clause, 
@@
 not perform static exception checking on native methods. This type of bug is hard to diagnose
-because the developer usualy do not have access to the native implementations. 
+because the developer usually do not have access to the native implementations. 
 Consequently, since it is not expected by the programmer, when such method throws 
@@
 \noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
-An unexpected fragility was detected: a checked exception being thown by a native method and not declared  
+The study revealed: a checked exception being thown by a native method and not declared  
 on the ``exception interface" of such method and all methods on the invocation chain.
@@
 
-      Runtime &  Checked   & 88 & 148 &  75  &  0   & 38 &  35 \\
-      Runtime   &  Error      & 46  &  67    &  58  &  0   & 8  &  1   \\      
-      Checked &  Runtime   & 17  & 31 & 4 &  0  & 16 &  11 \\
+      Runtime &  Checked   & 88 & 148 &  75 &  0   & 38 &  35 \\  %(50.7\%) 
+      Runtime   &  Error      & 46  &  67    &  58  &  0   & 8  &  1   \\      %(86.5\%) 
+      Checked &  Runtime   & 17  & 31 & 4 &  0  & 16  &  11 \\  %(51.6\%) 
       Checked & Error         & 8    &  9  & 5  &  0  &  1 &  3  \\
@@
 which offers three kinds of exceptions each one holding an intended exception behavior (i.e., error, 
-runtime and checked). Table~\ref{tab:wrappingandroid} presents the wrappings found in this study, that
-envolve different exception types (i.e., Error, Checked, Exception). We call 
+runtime and checked). Table~\ref{tab:wrappingandroid} presents the wrappings found in this study that
+include different exception types (i.e., Error, checked Exception and Runtime). We call 
 such a wrapping a ``cross-type wrapping". Next we exemplify some of such
- wrappings and discuss why some of them may negatively impact the robustness.
-
-\emph{\textbf{Runtime Exception wrapping Error.}} We could observe that 44\% of the wrappings were performed 
-by android.os.AsyncTask. This class is responsible for dealing with asynchronous tasks in Android.
-The code snippet below show a general cach clause that convertes any Throwable in to a runtime and re-throw.  
-Table ~\ref{tab:exampeswrap} exemplifies some of such wrappings.
-Such wrapping masks an unrecoverable error in a general exception - doing so it that can be mistakenly handled.
+ wrappings and discuss whether some of them may impact the system robustness.
+
+\emph{\textbf{Runtime Exception wrapping Error.}} From Table~\ref{tab:wrappingandroid}, we could observe that most of such wrappings
+were performed by Android platform (50.7\%). Inspecting each of them revealed that the android.os.AsyncTask, was responsible for 44\% 
+of these wrappings. This class is responsible for dealing with asynchronous tasks in Android.
+The code snippet below shows a general catch clause that converts any instance of Throwable (signaled during the execution
+of an asynchronous task) into an instance of RuntimeException and re-throw it.  
+Table~\ref{tab:exampeswrap} exemplifies some of the wrappings performed by this tasks, such as 
+RuntimeExceptions wrapping istances of OutOfMemoryError and StackOverflowError. 
+Such wraps mask an unrecoverable error wrapping on a general runtime exception.
 
@@
  were also caused by the reflection library used by applications' methods. The methods responsible for the wrappings
-were also native methods written in C. Table~\ref{tab:exampleswrap} illustrates
-some oth these wrappings some of them are masking an OutOfMemoryError
-into a checked exception.
-
-\emph{\textbf{Error wrapping Runtime and Checked Exceptions.}} 
-These wrappings were performed by: native methods defined by 
+were also native methods written in C. Table~\ref{tab:exampeswrap} illustrates
+some of these wrappings some of them are masking an OutOfMemoryError
+into a checked exception. Such wrappings may also mask an unrecoverable error
+and may lead to the ``exception confusion problem" described next.
+
+\emph{\textbf{Error wrapping Runtime.}} The source code inspection related that
+these wrappings were performed by: native methods defined by 
 java.lang.Class (e.g., getDeclaredFields, getDeclaredConstructors), 
-methods from javax.crypto package, and Java static initializers. 
-If any exception is thrown in the context of a static initializer 
+methods from javax.crypto package. Table~\ref{tab:exampeswrap}  illustrates
+examples of such wrappings. In these examples, wapping a runtime exception in an error
+reflected a design decision.
+
+\emph{\textbf{Error wrapping Checked Exception.}} 
+The source code inspection related that to such wrappings were 
+performed by Java static initializers. If any exception is thrown in the context of a static initializer 
 (i.e., static block)  it is converted into an ExceptionInitializerError 
-on the point where the class is first used.  Table~\ref{tab:exampleswrap} illustrates
-examples of such wrappings.
+on the point where the class is first used.  Table~\ref{tab:exampeswrap} illustrates
+examples of such wrappings. We could observe that these examples also reflected
+a design decision.
 
 \emph{\textbf{Runtime wrapping Checked Exception.}} This wrapping was responsible for 49.5\% of
-the cross-type wrappings.  From this set 50\%  were performed on methods defined on Android Platform. 
+the cross-type wrappings.  From this set 50\%  were performed on methods defined on Android platform. 
 We could observe that approximately 41\% of them were performed by a single class,
@@
 invoking the life cycle methods (i.e. onCreate(), onPause()) for every Activity 
-(window) of an Android app. This class contains a such try-catch block that
-handled any instance of java.lang.Exception signaled in the context 
-of the method in a general Runtime exception and re-throws.
-
-We could also observe that some stack traces included several intermediate cross-type wrappings.
-One interesting example was the following: Runtime-Checked-Runtime-Checked-Runtime-Checked-Runtime.
-
-Although some of these wrappings may be a result of design decisions,
+(window) of an Android app. This class wraps every instance of java.lang.Exception 
+signaled withing the life cycle methods in a general RuntimeException and re-throws. 
+We observed that it was a common design decision in other methods of Android platform
+(although usually considered a bad practice).
+
+We could also observe that some stack traces included successive cross-type wrappings, 
+such as: Runtime - Checked - Runtime - Checked - Runtime - Checked - Runtime. Hence, although some of these wrappings may be a result of design decisions,
 the mis-use of exception wrapppings may make the exception handling 
@@
 to handle the exceptional condition (according to Java Specification and 
-best practices). However, if the developer tries to handle such exception,
-which is wrapping (or ``hiding") an OutOfMemoryError, such handling 
+best practices). However, such exception may be wrapping an Error such as OutOfMemoryError (which indicate resource deficiency 
+that makes impossible for the program to recover). Hence, trying to handle such exception  
 may lead the program to an unpredictable state.
@@
 \noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
-The cross-type wrappings show us that even if the developer sticks to Java
-specification and best practices, it will be dificult for her/him to prevent the code from 
-incurring into situations that lead the program to an unpredictable state. 
+The cross-type wrappings is an indication that even if the developer sticks to Java
+specification and best practices, it may be dificult for her/him to prevent the code from 
+incurring into situations that lead the program to an unpredictable state - 
+e.g., when handling checked Exceptions that are `hiding" Errors.
 }
@@
 
- "Everybody hates thinking about exceptions, because they are not supposed to happen"
-  Brian Foote
+ ``Everybody hates thinking about exceptions, because they are not supposed to happen"
+  (Brian Foote)~\footnote{Brian Foote shared his opinion in a conversation with James Noble - quoted on the paper: hillside.net/plop/2008/papers/ACMVersions/coelho.pdf}
 
@@
 third-party documentation on the exceptions that
-may the thrown by it.
-
-On the other hand, documenting runtime exceptions is a tedious and error prone task, to help developers
+may the thrown by it. On the other hand, documenting runtime exceptions is a tedious and error prone task, to help developers
 mitigate this problem tools should be developed to automate the documentation of runtime exceptions
@@
 This observation emphasizes the importance of solutions to avoid NullPointerExceptions, such as:
-(i) lightweight intra-method null pointer analysis supported by Java8 @Nullable annotations~\footnote{Such annotations are used by tools e.g., Eclipse, IntelliJ to detect potential 
-null pointer dereferences at compile time. The Android Studio 0.5.5 released in April 2014 support to these annotations.};
+(i) lightweight intra-method null pointer analysis supported by Java8 @Nullable annotations~\footnote{They are used by tools e.g., Eclipse, IntelliJ, Android Studio 0.5.5 (release Apr.
+2014) to detect potential
+null pointer dereferences at compile time.};
 (ii) inter-method null pointer analysis tools such as the one proposed by Nanda and Sinha ~\cite{nanda2009accurate};
@@
 developers classified a randomly selected sample (of 25 exception 
-types from the total 100); the interater agreement was high (96\%). 
+types from the total of 100); the interater agreement was high (96\%). 
 Another threat relates to the fact that parts of our analysis 
-are based on the availability of stack traces on issues on Github and Googlecode. 
+are based on the availability of stack traces on issues reported on Github and Googlecode projects. 
 In using these datasets, we make an underlying assumption: the stack traces reported on issues are 
@@
 the source code associated to a subset of the reported exception stack traces.
-Such subset comprise the stack traces related to the main findings 
+Such subset comprises the stack traces related to the main findings 
 of the study (e.g., ``undocumented runtime and checked exceptions",
 and cross-type wrappings").
-
-\noindent\emph{External Validity.} Our work use GHTorrent dataset. Although 
+%\noindent
+
+\emph{External Validity.} Our work uses GHTorrent dataset; although 
 comprehensive and extensive is not an exact replica of Github. 
@@
 exception stack trace mining of a subset of projects.
-We limited out analysis to a subset of existing open-source Android projects.
+We limited our analysis to a subset of existing open-source Android projects.
 We are aware that the exception stack traces reported 
@@
 
-\textit{Analysis and Use of Stack Trace Information.} Several works have
+\textit{Analysis and Use of Stack Trace Information.} Several research works have
 investigated the use of stack trace information to support bug classification
@@
 
-\textit{Empirical Studies analizing Java exception-related code.} 
+\textit{Empirical Studies on Exception Handling Defects.} 
 Cabral and Marques~\cite{cabral2007exception} analyzed the
@@
 observed that the actions inside handlers were very simple (e.g., logging and present a
-message to the user). Pingyu and Elbaum~\cite{Zhang12} were the first to perform
-an empirical investigation of issues (related to exception-related bugs) on Android projects.  
+message to the user). Coelho et al.  ~\cite{coelho2008assessing} performed an 
+empirical study considering the fault-proneness of aspect-oriented implementations 
+for handling exceptions. Two releases of both Java and AspectJ implementation of a 
+product line were assessed as part of that study. Based on the use of an exception
+flow analysis tool, the study revealed that the AOP  refactoring increased the 
+number of uncaught exceptions, degrading the robustness of the AO version of every analyzed system.
+The main limitation of approaches based static analysis based approaches are the number of false
+positives they can generate, and the problems the faced when dealing with reflection libraries 
+and dynamic class loading. Pingyu and Elbaum~\cite{Zhang12} were the first to perform
+an empirical investigation of issues, related to exception-related bugs, on Android projects.  
 They perform a small scale study on which they manually inspected the issues of 
@@
  which automaticaly mined the exception stack traces embedded on issues 
-reported on 639 opensource Android projects. The goal of our study was
+reported on 639 open source Android projects. The goal of our study was
 to identify common fragilities on the exception related code that can lead to 
 failures such as uncaught exceptions.
+
+\textit{Extracting Stack Traces from natural language artifacts.} 
+Apart from issues and bug reports, stack traces can be embedded in other forms of
+communication between developers, such as discussion logs and emails.
+Few tools have been proposed to mine stack traces embedded on such resources.
+ Infozilla~\cite{bettenburg2008extracting} is based on a set of regular expressions that extract a set of frames
+related to a stack trace. The main limitation of this solution is that it is not
+able to extract stack traces embedded on verbose log files (i.e., on which we
+can find log text mixed with exception frames). Bacchelli
+et al.~\cite{bacchelli2012content} propose a solution to recognize stack trace frames
+from development emails and relate it to code artifacts (i.e. classes) mentioned
+on the stack trace. In addition to those tools, ExceptionMiner is able to 
+both extract stack traces from natural language artifacts and to 
+classify them in a set of predefined categories.
+
+\enlargethispage{-2\baselineskip}
+
+\section{Conclusion}
+\label{sec:conc}
+
+In this paper we present an exploratory study in which we mined the stack 
+traces embedded in all issues defined in 482 Android projects hosted in Github and 
+157 projects hosted in Google Code. Overall it included 6,005 exception stack traces.
+In this study, the information extracted from stack traces was used in combination 
+with source code and bytecode analysis to identify common characteristics of exception 
+stack traces to pinpoint fragilities on the exception-related code. 
+We call ``fragilities" characteristics on the exception-related code that favor the introduction
+of  uncaught exceptions and unintended handling - which can contribute to 
+degrade the robustness of applications.
+Some fragilities consistently detected in this study were: 
+(i) Approximately 52\% of the reported stack traces can be attributed to programming mistakes. 
+In particular, 27.71\% of all stack traces reported a java.lang.NullPointerException as its root cause;
+(ii) 267 stack traces were caused by undocumented runtime exceptions thrown by third party code caused;
+(iii) undocumented checked exceptions thrown by JNI interface,
+(iv) unexpected exception wrappings (e.g., Errors being wrapped in checked exceptions) 
+revealing that Java hybrid exception model is not fully used according to its purpose.
+Such fragilities can negatively affect not only the robustness of Android application 
+but also the robustness of any Java-based application. 
+Our results calls the attention for the need of tools and language support to help 
+developers when dealing with such fragilities.
+
+%Nowadays many software vendors embed automatic crash reporting tools in their
+%software systems. Hence, whenever a software crashes this tool sends a detailed
+%crash report to its vendors. Moreover, we can also find third party software
+%solutions specialized in bug reporting for different kinds of systems specially
+%for the increasing marked of mobile apps~\cite{BugSe14,BugSn14,Googl14,Acra14}.
+%There is plenty of information to be mined...
+
+%\section*{Acknowledgment} This work is partially supported by: CNPq -- Proc.
+%484209/2013-2 and the NWO TestRoots project (639.022.314).
+
+\bibliographystyle{IEEEtran}
+\bibliography{android-stacks}
+
+% that's all folks
+\end{document}
 
@@
 used by the most successful apps.
-
-\textit{Extracting Stack Traces from natural language artifacts.} 
-Apart from issues and bug reports, stack traces can be embedded in other forms of
-communication between developers, such as discussion logs and emails.
-Few tools have been proposed to mine stack traces embedded on such resources.
- Infozilla~\cite{bettenburg2008extracting} is based on a set of regular expressions that extract a set of frames
-related to a stack trace. The main limitation of this solution is that it is not
-able to extract stack traces embedded on verbose log files (i.e., on which we
-can find log text mixed with exception frames). Bacchelli
-et al.~\cite{bacchelli2012content} propose a solution to recognize stack trace frames
-from development emails and relate it to code artifacts (i.e. classes) mentioned
-on the stack trace. In addition to those tools, ExceptionMiner is able to 
-both extract stack traces from natural language artifacts and to 
-classify them in a set of predefined categories.
-
-\enlargethispage{-2\baselineskip}
-
-\section{Conclusion}
-\label{sec:conc}
-
-In this paper we present an exploratory study in which we mined the stack 
-traces embedded in all issues defined in 482 Android projects hosted in Github and 
-157 projects hosted in Google Code. Overall it included 6,005 exception stack traces.
-In this study, the information extracted from stack traces was used in combination 
-with source code and bytecode analysis to identify common characteristics of exception 
-stack traces and pinpoint fragilities on the exception-related code. 
-We call ``fragilities" characteristics on the exception-related code that favor the introduction
-of  uncaught exceptions and unintended handling - which can contribute to 
-degrade the robustness of applications.
-Some fragilities consistently detected in this study were: 
-(i) an abundance of NullPointerExceptions - it was mentioned as the root exception in approximately 
-50\% of the analyzed projects. 
-(ii) undocumented runtime exceptions thrown by third party code,
-(iii) undocumented checked exceptions thrown by JNI interface,
-(iv) unexpected exception wrappings (e.g., Errors being wrapped in checked exceptions) 
-revealing that Java hybrid exception model is not fully used according to its purpose.
-Such fragilities can negatively affect not only the robustness of Android application 
-but also the robustness of any Java-based application. 
-Our results calls the attention for the need of tool and language support to help 
-developers when dealing with such fragilities.
-
-%Nowadays many software vendors embed automatic crash reporting tools in their
-%software systems. Hence, whenever a software crashes this tool sends a detailed
-%crash report to its vendors. Moreover, we can also find third party software
-%solutions specialized in bug reporting for different kinds of systems specially
-%for the increasing marked of mobile apps~\cite{BugSe14,BugSn14,Googl14,Acra14}.
-%There is plenty of information to be mined...
-
-%\section*{Acknowledgment} This work is partially supported by: CNPq -- Proc.
-%484209/2013-2 and the NWO TestRoots project (639.022.314).
-
-\bibliographystyle{IEEEtran}
-\bibliography{android-stacks}
-
-% that's all folks
-\end{document}
-
-\emph{\textbf{Runtime wrapping Checked Exception.}} Such wrapping was 
-detected on 88 projects. Inspecting the elements responsible for performing such wrapping, we could observe that
-approximately  50\% of such wrappings were performed on methods defined on Android
-Platform. Taking a closer look on the methods responsible for such wrappings,
- we could observe that approximately 41\% of them were performed by a single class,
- named ActivityThreat. This class is responsible for managing the life cycle of  Android Activities - 
-invoking the life cycle methods (i.e. onCreate(), onPause()) for every Activity 
-(window) of an Android app.
-The code snippet illustrates one the methods of 
-ActivityThreat,  the one responsible for invoking the Activitiy's onCreate method. 
-As we can see from lines 4-8, in case any exception is thrown in the context of such try-catch block, 
-it is wrapped in a general Runtime exception and re-thrown.
-
-%\begin{table}
-%\centering
-%\begin{tabular}{lll}
-%   \hline
-% \bfseries{Signaler class} &  \bfseries{Wrapper Type} & \bfseries{Ocurrences} \\
-%    \hline
-%android.app.ActivityThread & java.lang.RuntimeException & 61\\
-%android.view.LayoutInflater & android.view.InflateException &  7 \\
-%android.os.AsyncTask & java.lang.RuntimeException &  7\\
-%\hline
-%  \end{tabular}
-%\caption{Top 3 Runtime-Cheked wrappers.}
-%\label{tab:wrapping01}
-%\end{table}
-
-\emph{\textbf{Checked wrapping Runtime Exception.}} Manually inspecting the stack 
-traces we could observe that such wrappings were performed either 
-by RoboSpice (an android library that supports the creation of asynchronous long running 
-tasks, which has 1.682 stars and 367 forks in GitHub) or by the reflection Java library (used by
-methods defined in Android, Application and library classes).
-
-exemplos:
-
-(java.lang.RuntimeException,android.hardware.Camera.native_autoFocus(Native Method)")"

### 2015-02-11 android-stacks.tex
new title

@@
 
-%\title{What can the Issues of GitHub and Google Code tell us about Exception Fragilities in Android? }
+\title{What can Issues on GitHub and Google Code tell us about Exception Fragilities in Android? }
 
@@
 
-\title{Unveiling Robustness Threats by Mining the Exception Stack Traces of over 600 Android Projects}
+%\title{Unveiling Robustness Threats by Mining the Exception Stack Traces of over 600 Android Projects}
 
@@
 Georgios Gousios\IEEEauthorrefmark{2}, 
-Arie van Deursen\IEEEauthorrefmark{3}
+Arie van Deursen\IEEEauthorrefmark{2}
 }
@@
 Email: \texttt{roberta@dimap.ufrn.br,lucas.almeida@ppgsc.ufrn.br}}
-\IEEEauthorblockA{\IEEEauthorrefmark{2}Radboud University Nijmegen\\
-Nijmegen, The Netherlands\\
-Email: \texttt{georgios@cs.ru.nl}}
-\IEEEauthorblockA{\IEEEauthorrefmark{3}Delft University of Technology\\
+\IEEEauthorblockA{\IEEEauthorrefmark{2}Delft University of Technology\\
 Delft, The Netherlands\\
-Email: \texttt{arie.vandeursen@tudelft.nl}}
+Email: \texttt{\{g.gousios,arie.vandeursen\}@tudelft.nl}}
 }

### 2015-02-11 android-stacks.tex
adding related works involving Android apps

--- 2015-02-06 android-stacks.tex
@@
 
+In this section, we present work that is related to the present paper, divided into
+four categories: i) papers that use the information available on stack traces;
+ii) empirical studies on the usage of Java Exceptions and its fault proneness;
+and iii) tools to extract stack trace information from natural language artifacts
+(e.g., issues and emails) and iv) empirical studies involving Android apps.
+
 \textit{Analysis and Use of Stack Trace Information.} Several research works have
@@
 
+\textit{Empirical studies using Android apps.} Ruiz et al.~\cite{Ruiz12}
+investigated the degree of reuse across applications in Android Market, the
+study showed that almost 23\% of the classes inherited from a base class in the
+Android API, and that 217 mobile apps were reused completely by another mobile
+app. Pathak et al.~\cite{Patha11} analyzed bug reports and developers
+discussions of Android platform and found out that approximately 20\% of
+energy-related bugs in Android occurred after an OS update. McDonnell et
+al.~\cite{McDon13} conducted a case study of the co-evolution behavior of
+Android API and 10 dependent applications using the version history data found
+in github. The study found that approximately 25\% of all methods in the client
+code used the Android API, and that the methods reusing fast-evolving APIs were
+more defect prone then others. Vásquez et al.~\cite{Linar13} analyzed
+approximately 7K free Android apps and observed that the last successful apps
+used Android APIs that were on average 300\% more change-prone than the APIs
+used by the most successful apps. Pingyu and Elbaum~\cite{Zhang12} analyzed bug
+reports of 5 Android applications an observed that 29\% had to do with poor
+exceptional handling code. 
+Our work differs from the others as it aims at
+distilling the information of bug reports distilling the uncaught exceptions 
+occurring at mobile applications in Github and Google Code. Such analysis discusses about
+the causes of bugs related to exceptions across applications available and what the
+characteristics of the stack traces can tell us about the exception structure of
+those applications.
+
 \enlargethispage{-2\baselineskip}

### 2015-02-11 android-stacks.tex
more changes in related work

@@
 
-%\title{What can the Issues of GitHub and Google Code tell us about Exception Fragilities in Android? }
-
-\title{What can Issues on GitHub and Google Code tell us about Exception Fragilities in Android? }
+%\title{What can the Issues of GitHub and Google Code tell us about Exception Fragilities? }
+
+\title{What can Issues on GitHub and Google Code tell us about Exception Fragilities in Android and Java? }
 
@@
 related to NullPointerException and other implicitly thrown exceptions had
-the higher prevalence on the analyzed set of stacks.
+the higher prevalence on the analyzed set of stacks. Kechagia and Spinellis ~\cite{kechagia2014}
+examined the stack traces embedded on crash reports sent by 1,800 Android apps 
+to a crash report management service (i.e., BugSense). They found that 19% of such stack traces
+were caused by unchecked and undocumented exceptions thrown by methods defined on 
+Android API (level 15). Our work differs from Kechagia and Spinellis since it is not based on
+stack traces mined from issues reported by open source developers on GitHub and Googlecode.  
+Moreover, our work identified the type of each exception mined from issues
+(classifying them as Error, Runtime or Checked) based on the source code
+analysis of the exception hierarchy; we mapped the exception signalers (i.e., were the
+exceptions signaled from libraries, the platform or the application itself); we investigated
+the adoption of best practices based on the analysis of stack trace information.
+Such analysis revealed intrigging cross-type exception wrappings which may bring the program 
+to an unpredictable state (i.e.,  trying to handle an instance of OutOfMemoryError “hidden” in a checked exception).
 
@@
 reports of 5 Android applications an observed that 29\% had to do with poor
-exceptional handling code. 
-Our work differs from the others as it aims at
+exceptional handling code. Our work differs from the others as it aims at
 distilling the information of bug reports distilling the uncaught exceptions 

### 2015-02-12 android-stacks.tex
changes on abstract. introducing bug hazard

@@
 
-\title{What can Issues on GitHub and Google Code tell us about Exception Fragilities in Android and Java? }
+\title{What can Issues on GitHub and Google Code tell us about Bug Hazards in Android and Java? }
 
@@
 159,048 issues reported on Android projects hosted in GitHub (482 projects) 
-and Google Code (157 projsects). 
-The goal of this study is to investigate  whether the exception handling logic on these stack traces
- can reveal common characteristics on the exception-related code that
-are particularly fault prone and may lead to a decrease in application robustness.
-Overall 6,005 exception stack traces were extracted, and analyzed in 
+and Google Code (157 projsects). The goal of this study is to investigate whether the analysis of stack traces information 
+could reveal bug hazards related to the exception handling code that may lead to a 
+decrease in application robustness. A bug hazard is a circumstance that increases the 
+chance of a bug to be present in the software. Overall 6,005 exception stack traces were extracted, and analyzed in 
 combination with source code and bytecode analysis. 
-Key problems identified in this study include: 
-(1) undocumented runtime exceptions thrown by third party code;
-(2) undocumented checked exceptions thrown by a JNI interface function;
-(3) unexpected exception wrappings (such as Errors being wrapped in checked exceptions),
-indicating that Java's hybrid exception model is not fully used in accordance with its purpose.
-Moreover, about 50\% of all analyzed projects reported at least one NullPointerException
-as the root cause of an exception stack trace, suggesting the presence of programming mistakes.
-This study provides insights on such fragilities and the robustness threats  
-they impose not only to Android apps but also to other systems 
-based on Java exception model.
+Some of the outcomes of this study includes the identification of the following 
+bug hazards: (i) unexpected exception wrappings (such as Errors being wrapped in checked exceptions) -
+indicating that Java's hybrid exception model become fault-prone when it is not fully used in 
+accordance with its purpose; (ii) undocumented runtime exceptions thrown by both the Android 
+platform and third party libraries; and (iii) undocumented checked exceptions thrown by the Android Platform.
+This study provides insights on such bug hazards and the robustness threats  
+they impose not only to Android apps but also to other systems based on Java exception model.
 
@@
 
-\emph{\textbf{The prevalence of programming mistakes on stack traces.}} One way of getting a broader view of the prevalence~\footnote{The prevalence evaluates the proportion of a given characteristic over a group of subjects.}  of programming mistakes on reported stack traces would be to manually inspect the code responsible for throwning every exception reported  on issues.
-However, inspecting the source code related to 6005 stack traces can easily become infeasible.
-Hence, we conducted an empirical investigation based on the following assumption: the name of the exception and its associated Javadoc 
-documentation hold enough information to pinpoint a ``programming mistake".To exemplify, an instance of ArrayOutOfBoundException
- pinpoints to programming mistake which indicates that an array has been accessed with an illegal index.
-
-To perform this empirical investigation, we selected a subset of all reported root exceptions. 
-This subset was composed by the top 100 root exceptions (from a list ranked by the number of ocurrences on stacks).
-This subset was representative since it contained root exceptions that were reported as the root causes of 95\% of 
-all stack traces. Therefore, based on the inspection of the full names exceptions and the associated 
- Javadocs we could identify from this subset, which exceptions were related to programming mistakes.
- It is worth mentioning that  all exceptions identified as programming mistakes were exceptions defined on java.lang and java.util packages. 
-After doing such analysis, we could calculate the prevalence of such exceptions on issues.
-In other words, we could calculate the proportion of issues mentioning the exceptions identified in this analysis as related to ``programming mistakes".
-
-Table ~\ref{tab:causes} summarizes the results. It also illustrates the number of general Java exceptions - 
-whose full name does not carry the reason for the exception to be thrown.
-To ensure the quality of the analysis, three independent developers classified a randomly selected
-sample of 25 exception types (from the total 100) using the same list of concerns;
-the resulting interrater agreement was high (96\%).
-
-The results show that the programming mistakes are associated to the most reported root causes. 
-Although there is usually no proper handling for such exceptions~\cite{wirfs2006toward}, 
-besides presenting an error message to the user and restarting the application~\footnote{Only 
-high fault tolerant systems need to provide solutions to handle them.}.  
-This finding calls the attention to the following: the high number of programming mistakes reported 
-on issues are an indication that developers may be underestimating the impact that such 
-mistakes have on application robustness.
+%Commented 2015
+%\emph{\textbf{The prevalence of programming mistakes on stack traces.}} One way of getting a broader view of the prevalence~\footnote{The prevalence evaluates the proportion of a given characteristic over a group %of subjects.}  of programming mistakes on reported stack traces would be to manually inspect the code responsible for throwning every exception reported  on issues.
+However, inspecting the source code related to %6005 stack traces can easily become infeasible.
+%Hence, we conducted an empirical investigation based on the following assumption: the name of the exception and its associated Javadoc 
+%documentation hold enough information to pinpoint a ``programming mistake".To exemplify, an instance of ArrayOutOfBoundException
+%pinpoints to programming mistake which indicates that an array has been accessed with an illegal index.
+
+%To perform this empirical investigation, we selected a subset of all reported root exceptions. 
+%This subset was composed by the top 100 root exceptions (from a list ranked by the number of ocurrences on stacks).
+%This subset was representative since it contained root exceptions that were reported as the root causes of 95\% of 
+%all stack traces. Therefore, based on the inspection of the full names exceptions and the associated 
+%Javadocs we could identify from this subset, which exceptions were related to programming mistakes.
+% It is worth mentioning that  all exceptions identified as programming mistakes were exceptions defined on java.lang and java.util packages. 
+%After doing such analysis, we could calculate the prevalence of such exceptions on issues.
+In other words, we could calculate the proportion of issues mentioning the exceptions identified in this analysis as related to %``programming mistakes".
+
+%Table ~\ref{tab:causes} summarizes the results. It also illustrates the number of general Java exceptions - 
+%whose full name does not carry the reason for the exception to be thrown.
+%To ensure the quality of the analysis, three independent developers classified a randomly selected
+%sample of 25 exception types (from the total 100) using the same list of concerns;
+%the resulting interrater agreement was high (96\%).
+
+%The results show that the programming mistakes are associated to the most reported root causes. 
+%Although there is usually no proper handling for such exceptions~\cite{wirfs2006toward}, 
+%besides presenting an error message to the user and restarting the application~\footnote{Only 
+%high fault tolerant systems need to provide solutions to handle them.}.  
+%This finding calls the attention to the following: the high number of programming mistakes reported 
+%on issues are an indication that developers may be underestimating the impact that such 
+%mistakes have on application robustness.
+
+%\begin{table}
+ % \centering
+ % \begin{tabular}{lr}
+ %   \hline
+ %   \bfseries{Cause} &  \bfseries{\% Occurrences} \\
+ %   \hline
+ %     Programming Mistakes &  51.96\%\\ 
+ %     %Memory Restriction   &   4.10\% \\ 
+ %     General Exceptions   &  6.70\%\\
+ %   \hline
+% \end{tabular}
+%  \caption{Identifying the concerns related to root exceptions}
+%  \label{tab:causes}
+% \end{table}
+
+%\noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
+%The study revealed: approximately 52\% of the reported stack traces were associated to programming mistakes;
+%and 27.71\% of all stack traces reported a java.lang.NullPointerException as the root cause.
+%}
+%\end{minipage}}
+
+\emph{\textbf{Identifying the Concerns Related to Root Exceptions.}} To get a broader view of the root exceptions of stack traces,
+we performed an empirical evaluation whose goal was to identify the concerns related to the most reported root exceptions.
+Since it is infeasible to inspect the code responsible for throwning every exception reported in this study.
+This identification performed in this study was based on meaning that the name of the exception holds,
+its description (i.e., Javadoc documentaion), and the Java specification. To exemplify, an instance of ArrayOutOfBoundException
+usually refers to a programming mistake. Moreover, the Java specification list all exceptions related to backward compatibility
+issues ~\cite{javaback} (e.g., InstantiationError, VerifyError, IllegalAccessError). Some concerns are known as sources
+of faults in mobile development are: concurrency ~\cite{ama2012} backward compatibility
+~\cite{McDon13}, security ~\cite{enck2011study,was2010} and resource management (IO, Memory, Batery) ~\cite{Zhang12}.
+However, there are exceptions that cannot be directly mapped to one of such concerns. Either because they are
+related to programming mistakes - as illustrated on the previously. Or because they are too general (i.e., java.lang.Exception,
+java.lang.RuntimeException, java.lang.Error). Or also because they are related to other concerns (e.g., specific to a given library).
+
+To perform such mapping, we selected a subset of all reported root exceptions. This subset was composed by 100 root exceptions
+which were reported on 95\% of all stack traces analyzed in this study. Hence, based on the inspection of the source code and
+Javadoc related to each exception, and Java specification, we identified the concern
+releated to each root exception. Than we could calculate the prevalence of a concern on the reported stack traces using the following metric.
+Given that: Stacks is the complete set of stacks found in this study, Lc is the list of root causes related to a given concern (as identified on the study),
+and Stacks[Lc] the list of issues that mentions the concern. The prevalence of a concern could be given by the formula: Stacks[Lc]/Stacks.
+Table~\ref{tab:tophundrend} illustrates the result of this empirical evaluation.
 
@@
 \centering
-  \begin{tabular}{lr}
-    \hline
-    \bfseries{Cause} &  \bfseries{\% Occurrences} \\
-    \hline
-      Programming Mistakes &  51.96\%\\ 
-      %Memory Restriction   &   4.10\% \\ 
-      General Exceptions   &  6.70\%\\
-    \hline
+\begin{tabular}{lrr}
+\hline
+\bfseries{Category} & \bfseries{\% Occurrences on stacks} \\
+\hline
+Programming logic (java.lang and util) & 52,0\%\\
+Resources (IO, Memory, Batery) & 23,9\% \\
+Security & 4,1\%\\
+Concurrency & 2,9\% \\
+Backward compatibility & 5,5\% \\
+Specific Exceptions & 4,9\%\\
+General (Error, Exception, Runtime) & 6,7\%\\
 \end{tabular}
 \caption{Identifying the concerns related to root exceptions}
-  \label{tab:causes}
+\label{tab:tophundrend}
 \end{table}
-
+To ensure the quality of the process, three independent coders classified a randomly selected
+sample of 25 exception types (from the total 100) using the same list of concerns;
+the interater agreement was high (96\%). We could also observe that
+We can therefore assume that the categories are descriptive
+enough for covering the range of exception types analyzed.
+These study showed that the programming mistakes associated to the most reported root causes.
+However, they may be interactly related to one of the other concerns. For instance, a concurrency problem may
+lead to a NullPointerException.
 \noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
-The study revealed: approximately 52\% of the reported stack traces were associated to programming mistakes;
-and 27.71\% of all stack traces reported a java.lang.NullPointerException as the root cause.
+A multitude of programming mistakes were reported as the main causes of exception stack traces.
+51,96\% of all projects reported at least one exception stack trace on which the NullPoiterException
+was the root cause.
 }

### 2015-02-12 android-stacks.tex
changes on abstract.

@@
 159,048 issues reported on Android projects hosted in GitHub (482 projects) 
-and Google Code (157 projsects). The goal of this study is to investigate whether the analysis of stack traces information 
-could reveal bug hazards related to the exception handling code that may lead to a 
-decrease in application robustness. A bug hazard is a circumstance that increases the 
-chance of a bug to be present in the software. Overall 6,005 exception stack traces were extracted, and analyzed in 
+and Google Code (157 projsects). The goal of this study is to investigate 
+whether the analysis of stack trace information could reveal bug hazards related to the exception handling code,
+ that may lead to a decrease in application robustness. 
+ Overall 6,005 exception stack traces were extracted, and analyzed in 
 combination with source code and bytecode analysis. 
-Some of the outcomes of this study includes the identification of the following 
+Some of the outcomes of this study include the identification of the following 
 bug hazards: (i) unexpected exception wrappings (such as Errors being wrapped in checked exceptions) -
@@
 
+%A bug hazard is a circumstance that increases the chance of a bug to be present in the software.
 \end{abstract}

### 2015-02-12 android-stacks.tex
changes on abstract.

@@
 The goal of this study is to investigate whether the reported exception stack traces
- can reveal common fragilities in the exception-related code.  
-We call ``fragilities'' characteristics of the exception-related code 
-that increase the likelihood of introducing failures such as uncaught 
-exceptions -  leading to a decrease in application robustness.
+ can reveal common bug hazards in the exception-related code.  
+A bug hazard  ~\cite{binder2000testing}  is a circumstance that increases the 
+chance of a bug to be present in the software. An example of a bug hazard can 
+be a characteristic of  the exception-related code which can increase the likelihood 
+of introducing uncaught exceptions (one of the main causes of application crashes).
 
@@
  in the absence of ``exception interfaces". Moreover, the results are relevant to 
-designers of languages and tools to consider ways of reducing such fragilities 
+designers of languages and tools to consider ways of dealing with such bug hazards 
 in Android and Java environments.
@@
 embedded on issues of Android projects (hosted in GitHub and Google Code), can reveal 
-\emph{common fragilities} on the exception code of both the applications and the underlying  framework. 
+\emph{bug hazards} on the exception-related code of both the applications and the underlying framework. 
+This exploration is guided by the set of best practices compiled in this study; we investigate whether characteristics 
+of exception stack traces and the related source code it can pinpoint best practices violations and therefore to bug hazards.
+As mentioned before, in this context ``bug hazards" are the characteristics on the exception-related code 
+that favor the introduction of failures such as uncaught exceptions.
+
+To support this investigation, we developed a tool called ExceptionMiner (Section~\ref{sec:exceptionminer})
+which extracts the exception stack traces embedded on issues, 
+and combines stack trace information with source code and bytecode 
+analysis. Moreover, manual inspections were also used to leverage
+ the understanding of stack traces and support further discussions and insights.
+
 Our study focused on open-source apps, since the information needed to perform our study
@@
 
-This exploration is guided by the set of best practices compiled in this study; we investigate whether characteristics 
-of exception stack traces and the related source code it can pinpoint best practices violations. As mentioned before, 
-we call  ``fragilities" characteristics on the exception-related code that favor the introduction
-of failures such as uncaught exceptions and unintended handling.
-
-To support this investigation, we developed a tool called ExceptionMiner (Section~\ref{sec:exceptionminer})
-which extracts the exception stack traces embedded on issues, 
-and combines stack trace information with source code and bytecode 
-analysis. Such combination was needed to enable the identification of 
-fragilities on the exception-related code. Moreover, manual inspections were also used to leverage
- the understanding of stack traces and support further discussions and insights.
 In this study we explore the domain quantitatively and highlight interesting cases by 
@@
 trace contains relevant information concerning the exception structure of the
-projects analyzed - and therefore can review fragilities on the exception-related code -  
+projects analyzed - and therefore can review bug hazards on the exception-related code -  
 we opted for not restricting the analysis only on defect
@@
 quantitative and qualitative analysis of the outcomes. The structure of this section
- is centered on questions that guided the search for fragilities on the exception-related code
+ is centered on questions that guided the search for bug hazards on the exception-related code
 based on exception stack trace information.
@@
 \emph{\textbf{On the Presence of Checked Exceptions on Exception Interfaces.}} 
-In this study, the exception stack trace analysis revealed an unexpected fragility: a 
+In this study, the exception stack trace analysis revealed an unexpected bug hazard: a 
 checked exception thrown by a native method and not declared on the exception 
@@
 a complete Github dataset. Instead, the goal of our study was to 
-pinpoint fragilities on the exception-related code based on 
+pinpoint bug hazards on the exception-related code based on 
 exception stack trace mining of a subset of projects.
@@
 refered to exceptions coming from methods defined on Android Application Framework
-and third-party libraries.  Additionaly,  the ``fragilities" observed in this study are due to
+and third-party libraries.  Additionaly,  the ``bug hazards" observed in this study are due to
 characteristics of Java exception model, which can impose challenges 
@@
 reported on 639 open source Android projects. The goal of our study was
-to identify common fragilities on the exception related code that can lead to 
+to identify common bug hazards on the exception related code that can lead to 
 failures such as uncaught exceptions.
@@
 with source code and bytecode analysis to identify common characteristics of exception 
-stack traces to pinpoint fragilities on the exception-related code. 
-We call ``fragilities" characteristics on the exception-related code that favor the introduction
+stack traces to pinpoint bug hazards on the exception-related code. 
+In this study, bug hazards are characteristics on the exception-related code that favor the introduction
 of  uncaught exceptions and unintended handling - which can contribute to 
 degrade the robustness of applications.
-Some fragilities consistently detected in this study were: 
-(i) Approximately 52\% of the reported stack traces can be attributed to programming mistakes. 
-In particular, 27.71\% of all stack traces reported a java.lang.NullPointerException as its root cause;
-(ii) 267 stack traces were caused by undocumented runtime exceptions thrown by third party code caused;
-(iii) undocumented checked exceptions thrown by JNI interface,
-(iv) unexpected exception wrappings (e.g., Errors being wrapped in checked exceptions) 
-revealing that Java hybrid exception model is not fully used according to its purpose.
-Such fragilities can negatively affect not only the robustness of Android application 
+Some bug hazards consistently detected in this study were: 
+(i) unexpected exception wrappings (e.g., Errors being wrapped in checked exceptions)  - 
+revealing that Java hybrid exception model is not fully used according to its purpose;
+(ii) 267 stack traces were caused by undocumented runtime exceptions thrown by third party libraries or the Android Platform;
+and (iii) undocumented checked exceptions thrown by the Android Platform. 
+Moreover, we could also observe that approximately 52\% of the reported stack traces 
+can be attributed to programming mistakes; in particular, 27.71\% of all stack traces 
+reported a java.lang.NullPointerException as its root cause;
+The bug hazards discussed in this work can negatively affect not only the robustness of Android application 
 but also the robustness of any Java-based application. 
 Our results calls the attention for the need of tools and language support to help 
-developers when dealing with such fragilities.
+developers when dealing with them.
 

### 2015-02-12 android-stacks.tex
changes in introduction - rewriting findings

@@
 failures in third party services and libraries; compatibility issues~\cite{McDon13}; 
- memory and battery restrictions; and noisy external resources ~\cite{Zhang12} (e.g., wireless connections,GPS, bluetooth).
+ memory and battery restrictions; and noisy external resources ~\cite{Zhang12} 
+(e.g., wireless connections,GPS, bluetooth).
 
@@
 garcia2007extracting,garcia2001comparative,cabral2007exception,coelho2011unveiling}.
-As a consequence they may inadvertently negatively affect the system: Exception-related code may introduce failures such as 
+As a consequence they may inadvertently negatively affect the system: exception-related code may introduce failures such as 
 uncaught exceptions~\cite{jo2004uncaught, Zhang12} - 
@@
 be a characteristic of  the exception-related code which can increase the likelihood 
-of introducing uncaught exceptions (one of the main causes of application crashes).
+of introducing the aforementioned uncaught exceptions.
 
@@
 Overall 159,048 issues were analyzed and 6,005 stack traces were extracted from them.
-The exception stack trace analysis was augmented by means of byte code 
-and source code analysis
- with additional exception-related information
-extracted from the Android platform and Android applications.
-
-Outcomes of our study include:
-
-%Manual inspection was used to leverage the understanding 
-%of exception stack traces and support further discussions and insights.
-
-% Some outcomes consistently detected during this study were the following:
+The exception stack trace analysis was augmented by means of bytecode 
+and source code analysis on the exception-related code of the Android 
+platform and Android applications. Some bug hazards consistently detected 
+during this mining study include:
 
@@
 
-       \item Approximately 52\% of the reported stack traces can be attributed to programming mistakes. 
-In particular, 27.71\% of all stack traces reported a java.lang.NullPointerException as its root cause.
-
-       \item  4.4\% of the reported exception stack traces are ``undocumented runtime exceptions'' raised by third-party code. 
-             (35 methods from the Android platform and 44 from libraries).
-
-     \item  We detected several cross-type exception wrappings, such as an OutOfMemoryError wrapped in a checked exception,
-in stack traces. This suggests that the purpose of Java's hybrid exception model comprising checked and unchecked exceptions may not have been adequately used.
-
-   \item Undocumented checked exception signaled by a  JNI interface function defined on Android platform. 
-    Some flows reported a checked exception thrown by a JNI interface function, yet this exception was not  
-    part of the ``exception interface'' of the JNI function nor the methods on its invocation chain. This can lead to 
-   uncaught exceptions that are difficult to debug.
+   \item  Cross-type exception wrappings, such as an OutOfMemoryError wrapped in a checked exception.
+Trying to handle an instance of OutOfMemoryError ``hidden" in  a checked exception may bring the program
+ to an unpredictable state. Such wrappings suggest that, when not adequately used, the Java's hybrid exception
+ model may become fault-prone.
+
+  \item  Undocumented runtime exceptions raised by the Android platform (35 methods) and  third-party libraries (44 methods) -
+ which correspond to 4.4\% of the reported exception stack traces.
+In the absence of the ``exception specification" of third-party code, it is difficult or 
+even infeasible for the developer to protect the code against ``unforeseen'' exceptions. 
+Since in such cases the client usually does not have access to the source code, such undocumented 
+exceptions may remain uncaught and lead to system crashes.
+
+   \item Undocumented checked exceptions signaled by native C code.  Some flows reported a checked 
+exception signaled by native C code invoked by the Android Platform, yet this exception was not declared
+ of the Java Native Interface. This can lead to uncaught exceptions that are 
+difficult to debug.
+
+ \item A multitude of programming mistakes - approximately 52\% of the reported stack traces can 
+be attributed to programming mistakes. In particular, 27.71\% of all stack traces reported a java.lang.NullPointerException 
+as its root cause.
 
@@
 
+ The high prevalence of NullPointerExceptions and the undocumented runtime exceptions signaled by the Android Platform
+ is aligned with the findings of other research works ~\cite{kim2013predicting,fraser20131600,csallner2004jcrasher} and 
+~\cite{kechagia2014} respectively - which detected such bug hazards while testing or mining crash reports of Java and Android projects. 
+On the one hand, the findings our study emphasize the impact of these bug harzards on the application robustness 
+by mining a different source of information: the issues created by developers on GitHub and Google Code. On the other hand, 
+our work points to bug hazards that were not detected by previous works (i.e., cross-type wrappings, undocumented checked exceptions and 
+undocumented runtime exceptions thrown by third-party libraries) which can also negatively affect the application robustness.
+
 Such outcomes point to threats not only to the development of robust Android apps, 
-but also to the development of any robust Java-based system. For instance, in the absence of the 
-``exception specification" of third-party code, it is difficult or 
-even infeasible for the developer to protect the code against ``unforeseen'' exceptions. 
-Since in such cases the client usually does not have access to the source code, such undocumented exceptions may 
-remain uncaught and lead to system crashes. Moreover, the cross-type exception wrappings 
-(e.g., OutOfMemoryError is wrapped in a checked Exception) 
-may lead to unintended handling problems - e.g, trying to handle an instance of OutOfMemoryError ``hidden" in 
-a checked exception may bring the program to an unpredictable state.
-
-Hence, the study results are relevant to Android and 
-Java developers who may underestimate how 
- exceptions are related to programming mistakes, and how 
- tricky it can be to prevent the uncaught exceptions from third-party code
- in the absence of ``exception interfaces". Moreover, the results are relevant to 
-designers of languages and tools to consider ways of dealing with such bug hazards 
-in Android and Java environments.
+but also to the development of any robust Java-based system. 
+Hence, the study results are relevant to Android and Java developers who may underestimate the effect of such
+ bug hazards the application robustness, and how tricky it can be to prevent them. Moreover, such bug hazards 
+calls for improvements on languages and tools to deal with such bug hazards in Android and Java environments.
 
@@
 higher-level services to the applications running on the platform) ~\cite{andguide}.
-
-The Android Application Framework supports an event-driven development model. 
-Among the basic elements that comprise an Android application, we can cite the
- Activity class. The Activity is the main class of an Android app (there is usually one
-Activity per application window). Each Activity defines a set of hook methods  
-(e.g., OnCreate(), onPause()) that are dispatched automatically by the Android 
-framework (by inversion of control), in response to infrastructure and application events.
 The Android Application Framework also provides several supporting APIs, such as 
@@
 unresponsive due to resource limitation and excessive network access).
+
+% The Android Application Framework supports an event-driven development model. 
+% event driven and inversion of control.
+Among the basic elements that comprise an Android application, the Activity 
+is the main class of an app (there is usually one
+Activity per application window). Each Activity defines a set of hook methods  
+(e.g., OnCreate(), onPause()) that are dispatched automatically by the Android 
+framework, in response to infrastructure and application events.
 

### 2015-02-12 android-stacks.tex
section 2 - changes

@@
 
-In Java, exceptions are represented according to a class hierarchy, on which
+\emph{\textbf{Exception Types.}} In Java, exceptions are represented according to a class hierarchy, on which
  every exception is an instance of the Throwable class, and can be of three kinds: the checked exceptions
@@
 
-By convention an Error represents an unrecoverable condition which usually results
-from failures detected by the Java Virtual Machine, such as OutOfMemoryError and
-normally should not be handled inside the application. The user-defined exceptions 
-can be either checked or runtime. The Java specification~\cite{gosling2000java} 
-suggests that user-defined exceptions should be checked, because doing so 
-the callers of a method will know about the exceptions that a method can throw, and then 
-they can decide what to do about them. However, runtime exceptions can also be 
-used to represent user-defined exceptional conditions.There is a long-lasting debate 
+By convention, instances of Error represent  unrecoverable conditions which usually results
+from failures detected by the Java Virtual Machine due to resource limitations such as OutOfMemoryError;
+normally they cannot be handled inside the application.
+
+Instances of RuntimException are implicitly thrown by Java runtime environment when a program violates 
+the semantic constraints of the Java programming language (e.g., out-of-bounds array index, divide-by-zero 
+error, null pointer references). Some programming languages react to such errors by peremptorily terminating the program; 
+other programming languages allow similar approaches e.g.C\#; and others let the program to continue
+ its execution in some situations such as the out-of-bounds array index e.g., C++. 
+The programs are not expected to handle such runtime exceptions signaled by the runtime environment according 
+to the Java Specification ~\cite{gosling2000java}.
+
+The user-defined exceptions can be either checked or runtime. There is a long-lasting debate 
 about the pros and cons of both approaches~\cite{javatut,stackoverlow,debate}
-%\footnote{152 questions in Stackoverlow are related to this debate}.
-
-The runtime exceptions are also implicitly thrown by the runtime environment when a program violates 
-the semantic constraints of the Java programming language (e.g., out-of-bounds array index, divide-by-zero 
-error, null pointer references).\footnote{Some programming languages react to such errors by peremptorily terminating the program; 
-other programming languages allow similar approaches e.g.C\#; and others let the program to continue
- its execution in some situations such as the out-of-bounds array index e.g., C++. }  
-The programs are not expected to handle such runtime exceptions signaled by the runtime environment ~\cite{gosling2000java}.
+%\footnote{152 questions in Stackoverflow are related to this debate}. 
+Section ~\ref{sec:best} presents
+a set of best practices related to each exception type.
 
@@
 
-In Java, once an exception is thrown, the runtime environment looks for the nearest enclosing exception handler
+\emph{\textbf{Exception Propagation.}} In Java, once an exception is thrown, 
+the runtime environment looks for the nearest enclosing exception handler
 (in Java it is represented by the try-catch block), and unwinds the execution stack if necessary.
@@
 
-\emph{IV- All exceptions explicitly thrown by a library/framework reusable code should be documented.}
+\emph{IV- All exceptions explicitly thrown by reusable code should be documented.}
 (\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,bloch2008effective}).
@@
 also wise to document (either or on the exception interface or using @throws tag on Javadoc) 
-the explicitly thrown runtime exceptions\footnote{This
-excludes the implicitly signaled runtime exceptions related to programming
-mistakes} as carefully as checked exceptions, e when developing
- library/framework reusable code. Doing so, the clients of such library/framework
+the explicitly thrown runtime exceptions as carefully as checked exceptions. 
+When mentioning reffering to `` explicit" thrown exceptions
+we  exclude the implicitly signaled runtime exceptions related to programming
+mistakes. Such practice is mostly needed when developing library/framework 
+code reusable code - as the developers do not have access to the source code. 
+Doing so, the clients of such library/framework
 will be aware of all the exceptions a method can throw and can design the code
@@
 In this study we explore the domain quantitatively and highlight interesting cases by 
-exploring cases qualitatively.  Figure~ref{fig:overview} gives an overview of the study, whose tasks are detailed next.
+exploring cases qualitatively.  Figure~\ref{fig:overview} gives an overview of the study, whose tasks are detailed next.
 

### 2015-02-12 android-stacks.tex
section 4 - changes

@@
 \emph{bug hazards} on the exception-related code of both the applications and the underlying framework. 
-This exploration is guided by the set of best practices compiled in this study; we investigate whether characteristics 
-of exception stack traces and the related source code it can pinpoint best practices violations and therefore to bug hazards.
+This exploration is guided by the set of best practices compiled in this study. We investigate whether characteristics 
+of exception stack traces can pinpoint best practices violations and therefore reveal bug hazards.
 As mentioned before, in this context ``bug hazards" are the characteristics on the exception-related code 
@@
 
+Figure~\ref{fig:overview} gives an overview of the study.
+Firstly  the issues reported on Android projects hosted on GitHub and Google Code were recovered (Figure~\ref{fig:overview} - (1) and (2)).
+Then the stack traces embedded each issue were extracted and distilled (Figure~\ref{fig:overview} - (3)).
+The stack trace information were then combined with source code and bytecode analysis in order to discover 
+the type of the exceptions reported on the stack traces (e.g., error, runtime, checked)  (Figure~\ref{fig:overview} - (5))  and the origin of such exceptions 
+(e.g., the application, a library, the Android platform)  (Figure~\ref{fig:overview} - (6)). 
+Manual inspections were also used to support this mining process (Figure~\ref{fig:overview} - (4), (7) and (9)) and support
+the search for bug hazards performed in this study  (Figure~\ref{fig:overview} - (8)). Next sections detail each step of this mining process.
+
 Our study focused on open-source apps, since the information needed to perform our study
@@
 target of other research studies ~\cite{Linar13,ahimed}.   
-
 In this study we explore the domain quantitatively and highlight interesting cases by 
-exploring cases qualitatively.  Figure~\ref{fig:overview} gives an overview of the study, whose tasks are detailed next.
-
-\subsection{Data Extraction Process}
-\label{sec:miningproc}
-
-In this exploratory study the following information is needed: (i) the issues reported on Android project hosted on 
-GitHub and Google Code; (ii) the stack traces embedded on these issues; (iii) the type of the exceptions
- reported on the stack traces (e.g., error, runtime, checked); (iv) the origin of such exceptions 
-(e.g., the application, a library, the Android platform). Next, we describe: how such information 
-was retrieved from GitHub and Google Code; how the exceptions stack traces were extracted and distilled,
- and how the type associated to each exception and its origins was identified, and finally how manual
-inspections supported this work.
+exploring cases qualitatively.
+
+%\subsection{Data Extraction Process}
+%\label{sec:miningproc}
+
+%In this exploratory study the following information is needed: (i) the issues reported on Android projects hosted on 
+%GitHub and Google Code; (ii) the stack traces embedded on these issues; (iii) the type of the exceptions
+ %reported on the stack traces (e.g., error, runtime, checked); (iv) the origin of such exceptions 
+%(e.g., the application, a library, the Android platform). Next, we describe: how such information 
+%was retrieved from GitHub and Google Code.
 
@@
 
-\subsubsection{Android Apps in GitHub}
+\subsection{Android Apps in GitHub}
 \label{sec:git}
@@
 
-Then we performed further cleanup, inspecting the site of every Android
+Then we performed further cleanup, inspecting the site of every Android project
 reporting at least one stack trace, to make sure that they represented real
@@
 
-Based on the assumption that regardless the issue type, every exception stack
+Based on the assumption that regardless the issue lables, every exception stack
 trace contains relevant information concerning the exception structure of the
 projects analyzed - and therefore can review bug hazards on the exception-related code -  
-we opted for not restricting the analysis only on defect
-issues\footnote{We conducted the same analysis on the defect issues and the top
-exceptions are similar to the ones mentioned on defect issues. Due to space
-limitation we limit to present the general analysis here. More detailed analysis
-on the defect issues can be found at:
-\url{https://github.com/souzacoelho/exceptionminer}}.
- Moreover, issues and pull requests are dual on Github; all pull requests have a corresponding 
-``backing'' issue (i.e., an issue that is automatically generated per pull request). Therefore, we excluded such automatically generated
-issues from our analysis.
-
-\subsubsection{Android Apps in Google Code}
+we opted for not restricting the analysis only on defect issues\footnote{We conducted the same analysis 
+on the defect issues only and the top exceptions were quite similar.
+ Due to space limitation we limit to present the general analysis here. More detailed analysis
+on the defect issues can be found at:\url{https://github.com/souzacoelho/exceptionminer}}.
+
+\subsection{Android Apps in Google Code}
 Google Code contains largely used open-source Android apps (e.g. K9Mail \footnote{K9Mail moved to Github but as a way of not loosing the project history 
@@
 
-\section{The ExceptionMiner Tool}
+\subsection{The ExceptionMiner Tool}
 \label{sec:exceptionminer}
@@
 This component also distills fine grained information of each attribute such as: the classes and packages associated to them.
-This component is based on a combination between a regular expression based parser 
-with heuristics able to identify and filter exception names and stack traces inline with text. 
+This component is based on a combination between a parser (based on regular expressions) 
+and heuristics able to identify and filter exception names and stack traces inline with text. 
 In contrast to existing issue parsing solutions such as Infozilla, the parser
-created in this work can discover stack traces inlined with logs files \footnote{In several 
+created in this work can discover stack traces mixed with log information\footnote{In several 
 exception stack traces, the exception frames were preceeded by logging information e.g., 

### 2015-02-13 android-stacks.tex
all over

@@
 
-%\begin{tiny}
-%\begin{verbatim}
-
-%03-01 W/dalvikvm(7924): threadid=1: thread exiting with uncaught exception (group=0x40adf210)
-%03-01 15:55:01.609 (7924): FATAL EXCEPTION: main
-%03-01  (7924): java.lang.RuntimeException: ...
-%E/AndroidRuntime( 7924):  at android....performLaunchActivity(ActivityThread.java:1967)
-%03-01 15:55:01.609 (7924):  at android....handleLaunchActivity(ActivityThread.java:1992)
-%03-01 15:55:01.609:  at android.app.ActivityThread.access\$600(ActivityThread.java:127)
-%E/AndroidRuntime( 7924):  at android....handleMessage(ActivityThread.java:1158)
-%\end{verbatim}
-%\end{tiny}
-
-As mentioned before, the tool enables a set of exception stack traces analyses by combining exception stack 
-trace information with bytecode and sourcecode analysis. Each analysis is implemented as a different module 
-on ExceptionMiner. Next, we present core analyses implemente on ExceptionMiner. Such analyses 
-provide additional information to be used on further analyses described in Section~\ref{sec:result}:
-
 \emph{\textbf{Exception Type Analysis.}} To support a deeper investigation of the 
@@
 uses Design Wizard~\cite{Brunet09} framework to inspect the bytecode of all exceptions reported on issues,
-and a specific implementation to analise the source code of Java Exception classes.
-It walks up the type hierarchy of a Java exception until it reaches a base exception type.
+and a specific implementation to analyze the source code of Java Exception classes (when the
+bytecode was not available). It walks up the type hierarchy of a Java exception until it reaches a base exception type.
 This module analyzed all exceptions defined on Android platform (API level 19) - 
 which includes all basic Java exceptions that can be thrown during the app execution,
-and exceptions thrown by Android core libraries. Moreover it also analized the 
+and exceptions thrown by Android core libraries. Moreover it also analyzed the 
 libraries and applications which defined exceptions reported on stack traces
@@
 - filtering out the cases on which the application 
-also included the source-code of reused libraries.
+also included the source code of reused libraries.
 Then based on this information and using pattern matching 
@@
 apps and Android Plaform). When the exception could not be  
-found niether automatically, nor manually (because they were related an app or lib the previous versions
-that were no more availble) the exception were classified 
+found niether automatically, nor manually (because they were defined on a previous version
+of the app or lib) the exception were classified 
 in our study as "Undefined".  Only 31 exceptions 
@@
  although we could also find the NullPointerException being signaled by third-party libraries (16.3\%). 
-Regarding reusable code (e.g., libraries and frameworks), there is no consensus whether it is a good or bad practice to 
+Regarding reusable code (e.g., libraries and frameworks), there is no consensus whether it is a good or a bad practice to 
 re-throw a NullPointerException. Some prefer to encapsulate such an exception on
-IllegalArgumentException, while others~\cite{bloch2008effective} argue that the
+an instance of IllegalArgumentException, while others~\cite{bloch2008effective} argue that the
 NullPointerException makes the cause of the problem explicit and hence 
@@
 The high prevalence of NullPointerException is aligned with the findings of other research
-works~\cite{kim2013predicting,fraser20131600,csallner2004jcrasher}. For instance, Sunghun et
+works~\cite{kim2013predicting,fraser20131600,csallner2004jcrasher,kechagia2014}. For instance, Sunghun et
 al.~\cite{kim2013predicting} showed that 38\% the bugs related to exception handling in Eclipse project
- were caused by NullPointerException; other works on robustness 
-testing~\cite{maji2012empirical,csallner2004jcrasher} showed that most of the automatically 
+ were caused by NullPointerException; Kechagia and Diomidis showed that NullPointerException was the  
+most reported exception on the crash reports sent to BugSense (a bug report 
+management service for Android applications); 
+other works on robustness testing~\cite{maji2012empirical,csallner2004jcrasher} showed that most of the automatically 
 detected bugs were due to NullPointerException and exceptions implicitly-signaled by Java 
@@
 %Commented 2015
-%\emph{\textbf{The prevalence of programming mistakes on stack traces.}} One way of getting a broader view of the prevalence~\footnote{The prevalence evaluates the proportion of a given characteristic over a group %of subjects.}  of programming mistakes on reported stack traces would be to manually inspect the code responsible for throwning every exception reported  on issues.
-However, inspecting the source code related to %6005 stack traces can easily become infeasible.
+%\emph{\textbf{The prevalence of programming mistakes on stack traces.}} One way of getting a broader view of the prevalence~\footnote{The prevalence %evaluates the proportion of a given characteristic over a group %of subjects.}  of programming mistakes on reported stack traces would be to manually %inspect the code responsible for throwning every exception reported  on issues.
+However, inspecting the source code related to %6005 stack traces can %easily become infeasible.
 %Hence, we conducted an empirical investigation based on the following assumption: the name of the exception and its associated Javadoc 
@@
 %After doing such analysis, we could calculate the prevalence of such exceptions on issues.
-In other words, we could calculate the proportion of issues mentioning the exceptions identified in this analysis as related to %``programming mistakes".
+In other words, we could calculate the proportion of issues %mentioning the exceptions identified in this analysis as related to %``programming mistakes".
 
@@
 we performed an empirical evaluation whose goal was to identify the concerns related to the most reported root exceptions.
-Since it is infeasible to inspect the code responsible for throwning every exception reported in this study.
-This identification performed in this study was based on meaning that the name of the exception holds,
-its description (i.e., Javadoc documentaion), and the Java specification. To exemplify, an instance of ArrayOutOfBoundException
-usually refers to a programming mistake. Moreover, the Java specification list all exceptions related to backward compatibility
-issues ~\cite{javaback} (e.g., InstantiationError, VerifyError, IllegalAccessError). Some concerns are known as sources
-of faults in mobile development are: concurrency ~\cite{ama2012} backward compatibility
-~\cite{McDon13}, security ~\cite{enck2011study,was2010} and resource management (IO, Memory, Batery) ~\cite{Zhang12}.
-However, there are exceptions that cannot be directly mapped to one of such concerns. Either because they are
-related to programming mistakes - as illustrated on the previously. Or because they are too general (i.e., java.lang.Exception,
-java.lang.RuntimeException, java.lang.Error). Or also because they are related to other concerns (e.g., specific to a given library).
+Besides the exceptions related to programming mistakes mentined before, we also looked for exceptions related to some concerns that are known as sources of faults in mobile development: concurrency ~\cite{ama2012} backward compatibility ~\cite{McDon13}, security ~\cite{enck2011study,was2010} and resource management (IO, Memory, Batery) ~\cite{Zhang12}.
+
+Since it is infeasible to inspect the code responsible for throwning every exception reported in this study,
+the concern identification of each exception was based on meaning of the exception type defined on 
+its Javadoc documentaion and on the Java specification. To exemplify: (i) an instance of ArrayOutOfBoundException  
+refers to a programming mistake according to its Javadoc; and (ii) the Java specification list all exceptions related to backward compatibility
+%issues ~\cite{javaback} (e.g., InstantiationError, VerifyError, IllegalAccessError).
 
 To perform such mapping, we selected a subset of all reported root exceptions. This subset was composed by 100 root exceptions
-which were reported on 95\% of all stack traces analyzed in this study. Hence, based on the inspection of the source code and
-Javadoc related to each exception, and Java specification, we identified the concern
-releated to each root exception. Than we could calculate the prevalence of a concern on the reported stack traces using the following metric.
-Given that: Stacks is the complete set of stacks found in this study, Lc is the list of root causes related to a given concern (as identified on the study),
-and Stacks[Lc] the list of issues that mentions the concern. The prevalence of a concern could be given by the formula: Stacks[Lc]/Stacks.
-Table~\ref{tab:tophundrend} illustrates the result of this empirical evaluation.
+which were reported on 95\% of all stack traces analyzed in this study. Hence, based on the inspection of the 
+Javadoc related to each exception, and Java specification, we identified the concern releated to each root exception. 
+Table~\ref{tab:tophundrend} illustrates the result of this empirical evaluation. This table also illustrates the exceptions 
+that could not be directly mapped to one of the aforementioned concerns, either because they were too general (i.e., java.lang.Exception,
+java.lang.RuntimeException, java.lang.Error) or because they were related to other concerns (e.g., specific to an application or given library).
 
@@
 \end{table}
+
 To ensure the quality of the process, three independent coders classified a randomly selected
@@
 enough for covering the range of exception types analyzed.
-These study showed that the programming mistakes associated to the most reported root causes.
-However, they may be interactly related to one of the other concerns. For instance, a concurrency problem may
-lead to a NullPointerException.
+
+%These study showed that the programming mistakes associated to the most reported root causes.
+%However, they may be interactly related to one of the other concerns. For instance, a concurrency problem may
+%lead to a NullPointerException.
+
+%For instance, the Java specification list all exceptions related to backward compatibility
+%issues ~\cite{javaback} (e.g., InstantiationError, VerifyError, IllegalAccessError). 
+%were also investigated in this study
+
 \noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
 A multitude of programming mistakes were reported as the main causes of exception stack traces.
-51,96\% of all projects reported at least one exception stack trace on which the NullPoiterException
+52\% of all projects reported at least one exception stack trace on which the NullPoiterException
 was the root cause.
@@
 
-\subsection{What can the Types of Root Exceptions tell us?}
+\subsection{What can the Exception Types tell us?}
 %\subsection{Exploring the Information Embedded on Exception Types}
@@
 
-We can see, from Table ~\ref{tab:typeroottab}, that contrasting with the other origins, most of the 
+We can observe that most of the reported exceptions were runtime 
+(64.85\%); and most of them were signaled on methods defined either on the Application (47.3\%)
+or on the Android platform (34.3\%). We could also find runtime exceptions thrown by library code (17.7\%).
+
+We can also see, from Table ~\ref{tab:typeroottab}, that contrasting with the other origins, most of the 
  exceptions signaled on Android Libcore (i.e., the set of libraries reused by the Android) were 
@@
 
-We can also observe that most of the reported exceptions were runtime 
-(64.85\%); and most of them were signaled on methods defined either on the Application (47.3\%)
-or on the Android platform (34.3\%). We could also find runtime exceptions thrown by library code (17.7\%).
- Taking into account the results of empirical study on the prevalence of ``programming mistakes", 
-we conclude that a considerable amount 
-of such exceptions, are related to the programming mistakes
-as pointed out previously. However, within this set of runtime exceptions there are also the ones that do not correspond to programming mistakes,
-and according to Java specification should be documented as part of the exception interface of libraries/framework reusable
-methods.
-
-\emph{\textbf{Inspecting Exception Interfaces.}} To investigate the obedience to such practice, we firstly filtered out all the exceptions implicitly
- signaled by  runtime environment (due to programming mistakes) since these exceptions 
+\emph{\textbf{Inspecting Exception Interfaces.}}  According to the best practices mentioned before, runtime exceptions 
+should be documented as part of the exception interface of libraries/framework reusable
+methods. To investigate the obedience to such practice, we firstly filtered out all the exceptions implicitly
+ signaled by  runtime environment (due to programming mistakes) - since these exceptions 
 should not be documented on the method signature.  Then we inspected the code each method 
-(defined either on Android  Aplication Framework or on libraries) 
+(defined either on Android  Aplication Framework or on third-party libraries) 
 explicitly signaling a runtime exception not-related to programming mistakes.
@@
 This result is aligned with the results of other study conducted by  
-Sacramento et al ~\cite{sacramento2006unchecked} which observed that the
-runtime exceptions in .Net programs are most often not documented.
-These ``undocumented runtime exceptions" represent a threat to system robustness, specially
+Sacramento et al ~\cite{sacramento2006unchecked} - which observed that the
+runtime exceptions in .Net programs are most often not documented; and 
+Kechagia and Spinellis ~\cite{kechagia2014} which discovered a set of methods 
+on the Android API which do not documented runtime exceptions. One limitation 
+of the work of Kechagia and Spinellis is that it did filter out exceptions that should 
+although unchecked not be documented -  because they are implicitly signaled by the 
+JVM due to resource restrictions or violations on Java semantic constraints.
+
+The ``undocumented runtime exceptions" that are explicitly siglened represent a threat to system robustness, specially
 when such exceptions are thrown by a third party code invoked inside the application 
@@
 
-Moreover, these stack traces could also pointed out another best practice violation: the methods from Android platform
- presented in Table~\ref{tab:runtimeinterface} signaled direct instances of RuntimeException. 
-Using such general exception, is considered a bad practice according to Java specification and compiled guidelines.
-
 \noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
@@
 Libraries	& 	44	& 	205	\\
-\emph{android packages}   &  &   \\ 
-android.app	& 	5	& 	16	\\
-android.content	& 	2	& 	2	\\
-android.database	& 	1	& 	1	\\
-android.graphics	& 	5	& 	11	\\
-android.hardware	& 	4	& 	5	\\
-android.media	& 	3	& 	3	\\
-android.opengl	& 	1	& 	8	\\
-android.os	& 	11	& 	12	\\
-android.text	& 	1	& 	1	\\
-android.video	& 	1	& 	1	\\
-android.providers	& 	1	& 	2	\\
-\hline					
-\bfseries{All android} 	&	35	&	62	\\
+Android  	&	35	&	62	\\					
+%\emph{android packages}   &  &   \\ 
+%android.app	& 	5	& 	16	\\
+%android.content	& 	2	& 	2	\\
+%android.database	& 	1	& 	1	\\
+%android.graphics	& 	5	& 	11	\\
+%android.hardware	& 	4	& 	5	\\
+%android.media	& 	3	& 	3	\\
+%android.opengl	& 	1	& 	8	\\
+%android.os	& 	11	& 	12	\\
+%android.text	& 	1	& 	1	\\
+%android.video	& 	1	& 	1	\\
+%android.providers	& 	1	& 	2	\\
+%\hline					
+%\bfseries{All android} 	&	35	&	62	\\
 \hline					
@@
 of an asynchronous task) into an instance of RuntimeException and re-throw it.  
-Table~\ref{tab:exampeswrap} exemplifies some of the wrappings performed by this tasks, such as 
-RuntimeExceptions wrapping istances of OutOfMemoryError and StackOverflowError. 
+Table~\ref{tab:exampeswrap} exemplifies some of the wrappings performed in this code snipet:  
+an instance of RuntimeExceptions wrapping instances of OutOfMemoryError and StackOverflowError. 
 Such wraps mask an unrecoverable error wrapping on a general runtime exception.
+
+Moreover, such wrapping also pointed out to another best practice violation: the methods from Android platform
+signaling direct instances of RuntimeException. Using such general exception, is considered a bad practice 
+according to Java specification and compiled guidelines, because it loses contextual information about the exception.
 
@@
 
-\emph{\textbf{Error wrapping Runtime.}} The source code inspection related that
-these wrappings were performed by: native methods defined by 
-java.lang.Class (e.g., getDeclaredFields, getDeclaredConstructors), 
-methods from javax.crypto package. Table~\ref{tab:exampeswrap}  illustrates
-examples of such wrappings. In these examples, wapping a runtime exception in an error
-reflected a design decision.
-
-\emph{\textbf{Error wrapping Checked Exception.}} 
-The source code inspection related that to such wrappings were 
-performed by Java static initializers. If any exception is thrown in the context of a static initializer 
+\emph{\textbf{Error wrapping Runtime and Checked Exceptions}} Table~\ref{tab:exampeswrap}  illustrates
+examples of instances of Error wrapping instances of RuntimeException. A
+lthough such wapping mixes different exception types,
+since there is no obligation associated to handling runtime exceptions, such wrapping 
+does not violated the aforementioned best practices. On the other hand, the inspection 
+also revealed instances of Error wrapping checked exceptions. Such wrappings were 
+mostly performed by Java static initializers. If any exception is thrown in the context of a static initializer 
 (i.e., static block)  it is converted into an ExceptionInitializerError 
-on the point where the class is first used.  Table~\ref{tab:exampeswrap} illustrates
-examples of such wrappings. We could observe that these examples also reflected
-a design decision.
+on the point where the class is first used.  Table~\ref{tab:exampeswrap} also illustrates
+examples of such wrappings. Although such wapping may represent a design decision,
+it violates the best practice related to checked exceptions and errors as it mix the intended handling 
+behaviour associated to both types.
 
@@
 We could also observe that some stack traces included successive cross-type wrappings, 
-such as: Runtime - Checked - Runtime - Checked - Runtime - Checked - Runtime. Hence, although some of these wrappings may be a result of design decisions,
-the mis-use of exception wrapppings may make the exception handling 
+such as: Runtime - Checked - Runtime - Checked - Runtime - Checked - Runtime.
+Hence, although some of these wrappings may be a result of design decisions, the mis-use of exception wrapppings may make the exception handling
 code more complex (e.g., the multiple wrappings) and error-prone,
@@
 \subsection{On the null pointer problem}
-The null references was firstly introduced by Tony Hoare in ALGOL W, which after some years he called his ``one-billion-dollar mistake"~\cite{hoare2}:
-\emph{``I call it my billion-dollar mistake. It was the invention of the null reference in 1965 [...] This has led to innumerable errors, vulnerabilities, and system 
-crashes, which have probably caused a billion dollars of pain and damage in the last forty years."}
-
-In this sudy, the null references were, infact, responsible for several reported application errors - providing more evidences to Hoare's speech.
+The null references was firstly introduced by Tony Hoare in ALGOL W, which after some years he called his ``one-billion-dollar mistake"~\cite{hoare2}.
+
+%:
+%\emph{``I call it my billion-dollar mistake. It was the invention of the null reference in 1965 [...] 
+%This has led to innumerable errors, vulnerabilities, and system 
+%crashes, which have probably caused a billion dollars of pain and damage in the last forty years."}
+
+In this sudy, the null references were, infact, responsible for several reported issues - providing more evidences to Hoare's speech.
 This observation emphasizes the importance of solutions to avoid NullPointerExceptions, such as:
@@
 stack traces mined from issues reported by open source developers on GitHub and Googlecode.  
-Moreover, our work identified the type of each exception mined from issues
+Moreover, our study mapped the origin of each exception 
+(i.e., libraries, the Android platform or the application itself) and investigated
+the adoption of best practices based on the analysis of stack trace information.
+Our work also identified the type of each exception mined from issues
 (classifying them as Error, Runtime or Checked) based on the source code
-analysis of the exception hierarchy; we mapped the exception signalers (i.e., were the
-exceptions signaled from libraries, the platform or the application itself); we investigated
-the adoption of best practices based on the analysis of stack trace information.
-Such analysis revealed intrigging cross-type exception wrappings which may bring the program 
-to an unpredictable state (i.e.,  trying to handle an instance of OutOfMemoryError “hidden” in a checked exception).
+analysis of the exception hierarchy and analized the exception wrappings that can 
+happen during the exception propagation.  Such analysis revealed intrigging 
+bug hazards such as the cross-type exception wrappings not discussed in previous works.
 
@@
 Cabral and Marques~\cite{cabral2007exception} analyzed the
-source-code of 32 open-source systems, both for Java and .NET. They
+source code of 32 open-source systems, both for Java and .NET. They
 observed that the actions inside handlers were very simple (e.g., logging and present a

### 2015-02-13 android-stacks.tex
all over
premilinar version to submit

@@
 
- The high prevalence of NullPointerExceptions and the undocumented runtime exceptions signaled by the Android Platform
- is aligned with the findings of other research works ~\cite{kim2013predicting,fraser20131600,csallner2004jcrasher} and 
-~\cite{kechagia2014} respectively - which detected such bug hazards while testing or mining crash reports of Java and Android projects. 
-On the one hand, the findings our study emphasize the impact of these bug harzards on the application robustness 
+ The high prevalence of NullPointerExceptions is aligned with the findings of other research works ~\cite{kim2013predicting,fraser20131600,csallner2004jcrasher} as well as the undocumented runtime 
+exceptions signaled by the Android Platform ~\cite{kechagia2014}. If on the one hand, the findings our
+ study emphasize the impact of these bug harzards on the application robustness 
 by mining a different source of information: the issues created by developers on GitHub and Google Code. On the other hand, 
@@
 % event driven and inversion of control.
-Among the basic elements that comprise an Android application, the Activity 
-is the main class of an app (there is usually one
-Activity per application window). Each Activity defines a set of hook methods  
-(e.g., OnCreate(), onPause()) that are dispatched automatically by the Android 
-framework, in response to infrastructure and application events.
+%Among the basic elements that comprise an Android application, the Activity 
+%is the main class of an app (there is usually one
+%Activity per application window). Each Activity defines a set of hook methods  
+%(e.g., OnCreate(), onPause()) that are dispatched automatically by the Android 
+%framework, in response to infrastructure and application events.
 
@@
 from failures detected by the Java Virtual Machine due to resource limitations such as OutOfMemoryError;
-normally they cannot be handled inside the application.
-
-Instances of RuntimException are implicitly thrown by Java runtime environment when a program violates 
+normally they cannot be handled inside the application.  Instances of RuntimException are implicitly 
+thrown by Java runtime environment when a program violates 
 the semantic constraints of the Java programming language (e.g., out-of-bounds array index, divide-by-zero 
@@
 The programs are not expected to handle such runtime exceptions signaled by the runtime environment according 
-to the Java Specification ~\cite{gosling2000java}.
-
-The user-defined exceptions can be either checked or runtime. There is a long-lasting debate 
+to the Java Specification ~\cite{gosling2000java}. The user-defined exceptions can be either checked
+ or runtime. There is a long-lasting debate 
 about the pros and cons of both approaches~\cite{javatut,stackoverlow,debate}
@@
 projects analyzed - and therefore can review bug hazards on the exception-related code -  
-we opted for not restricting the analysis only on defect issues\footnote{We conducted the same analysis 
-on the defect issues only and the top exceptions were quite similar.
- Due to space limitation we limit to present the general analysis here. More detailed analysis
-on the defect issues can be found at:\url{https://github.com/souzacoelho/exceptionminer}}.
+we opted for not restricting the analysis only on defect issues.
+
+%\footnote{We conducted the same analysis 
+%on the defect issues only and the top exceptions were quite similar.
+% Due to space limitation we limit to present the general analysis here. More detailed analysis
+%on the defect issues can be found at:\url{https://github.com/souzacoelho/exceptionminer}}.
 
 \subsection{Android Apps in Google Code}
-Google Code contains largely used open-source Android apps (e.g. K9Mail \footnote{K9Mail moved to Github but as a way of not loosing the project history 
-it advises their users to report bugs on Google Code issue tracker: https://github.com/k9mail/k-9/wiki/LoggingErrors.}),
+Google Code contains largely used open-source Android apps (e.g.
+K9Mail\footnote{K9Mail moved to Github but as a way of not loosing the project history it advises their users to report bugs on Google Code issue tracker: https://github.com/k9mail/k-9/wiki/LoggingErrors.}),
  however, differently from GitHub, Google Code does not provide an API to access the information related
- to hosted projects \footnote{Google code used to provided a Web service to such repository, but it was deactivated in June 2013 what Google called a `clean-up action".}.
+ to hosted projects\footnote{Google code used to provided a Web service to such repository, but it was deactivated in June 2013 what Google called a `clean-up action".}.
 To overcome this limitation we needed to implement a Web Crawler (incorporated in ExceptionMiner tool described next) that navigated 
@@
 "03-01 15:55:01.609 (7924): at android.app.ActivityThread.access\$600(ActivityThread.java:127) 
-"which could not be detected by existing tools.)}.
+"which could not be detected by existing tools.}.
 
@@
 \hline
-\bfseries{Category} & \bfseries{\% Occurrences on stacks} \\
+\bfseries{Concern} & \bfseries{\% Occurrences on stacks} \\
 \hline
@@
 General (Error, Exception, Runtime) & 6,7\%\\
+\hline
 \end{tabular}
@@
 \emph{\textbf{Runtime Exception wrapping Error.}} From Table~\ref{tab:wrappingandroid}, we could observe that most of such wrappings
-were performed by Android platform (50.7\%). Inspecting each of them revealed that the android.os.AsyncTask, was responsible for 44\% 
-of these wrappings. This class is responsible for dealing with asynchronous tasks in Android.
-The code snippet below shows a general catch clause that converts any instance of Throwable (signaled during the execution
-of an asynchronous task) into an instance of RuntimeException and re-throw it.  
-Table~\ref{tab:exampeswrap} exemplifies some of the wrappings performed in this code snipet:  
-an instance of RuntimeExceptions wrapping instances of OutOfMemoryError and StackOverflowError. 
-Such wraps mask an unrecoverable error wrapping on a general runtime exception.
-
-Moreover, such wrapping also pointed out to another best practice violation: the methods from Android platform
-signaling direct instances of RuntimeException. Using such general exception, is considered a bad practice 
-according to Java specification and compiled guidelines, because it loses contextual information about the exception.
+were performed by Android platform (50.7\%).
+The code snippet below was extracted from Android and shows a general catch clause 
+that converts any instance of Throwable (signaled during the execution
+of an asynchronous task) into an instance of RuntimeException and re-throws it.  
+Table~\ref{tab:exampeswrap} presents exceptions wrapped in this code snipet.  
+Such wrappings mask an unrecoverable Error on a general runtime exception.
 
@@
 
+\emph{\textbf{Runtime wrapping Checked Exception.}} This wrapping was responsible for 49.5\% of
+the cross-type wrappings.  From this set 50\%  were performed on methods defined on Android platform. 
+We observed that it was a common implementation practice in the methods of Android platform.
+However, using such general exception, is considered a bad practice according to Java specification 
+and compiled guidelines, because it loses contextual information about the exception.
+
 \emph{\textbf{Checked Exception wrapping Error.}} Most of these wrappings
@@
 
-\emph{\textbf{Runtime wrapping Checked Exception.}} This wrapping was responsible for 49.5\% of
-the cross-type wrappings.  From this set 50\%  were performed on methods defined on Android platform. 
-We could observe that approximately 41\% of them were performed by a single class,
- named ActivityThreat. This class is responsible for managing the life cycle of  Android Activities - 
-invoking the life cycle methods (i.e. onCreate(), onPause()) for every Activity 
-(window) of an Android app. This class wraps every instance of java.lang.Exception 
-signaled withing the life cycle methods in a general RuntimeException and re-throws. 
-We observed that it was a common design decision in other methods of Android platform
-(although usually considered a bad practice).
-
 We could also observe that some stack traces included successive cross-type wrappings, 
@@
 
-\subsection{On the need of rescues to the almost infeasible task of preventing uncaught exceptions}
+\emph{\textbf{The exception handling confusion problem.}}
+When (mis)applied, the exception wrapping can make the exception-related code
+ more complex and lead to what we call an \emph{exception handling confusion problem}.
+Such problem can lead the program to an unpredictable state in the presence of exceptions,
+as illustrated by the scenario on which a checked exception wrapped an OutOfMemoryError. 
+Currently there is no way of enforcing Java exception type conventions during program development.
+Hence, further investigation is needed on finding ways to help developers in dealing with such
+ problem, either preventing odd wrappings or enabling the developer to better deal with them or even
+empirical studies on the real usefulness of Java hybrid exception model.
+
+%The cross-type wrappings detected in this study points to the fact that: (1) exception 
+%wrappings may prevent exception types from being used according to its initial purpose
+%(e.g, Errors should represent situations that should not be handled); and (2) may have been
+% used to bypass the language restrictions imposed by checked exceptions  (e.g,
+% checked exceptions may be wrapped in runtime exceptions).
+
+\emph{\textbf{On the null pointer problem.}}
+The null references was firstly introduced by Tony Hoare in ALGOL W, which after some years he called his ``one-billion-dollar mistake"~\cite{hoare2}.
+In this sudy, the null references were, infact, responsible for several reported issues - providing more evidences to Hoare's speech.
+This observation emphasizes the importance of solutions to avoid NullPointerExceptions, such as:
+(i) lightweight intra-method null pointer analysis supported by Java8 @Nullable annotations~\footnote{They are used by tools e.g., Eclipse, IntelliJ, Android Studio 0.5.5 (release Apr.
+2014) to detect potential
+null pointer dereferences at compile time.};
+(ii) inter-method null pointer analysis tools such as the one proposed by Nanda and Sinha ~\cite{nanda2009accurate};
+or (iii) language designs which avoid null pointers, such 
+as Monads ~\cite{Walde95} (i.e., used in functional languages for values that may not be available 
+or computations that may fail) could improve the robustness of Java programs.
+
+\emph{\textbf{On the need of rescues to the almost infeasible task of preventing uncaught exceptions}}
 In this study we could observe undocumented runtime exceptions thrown by third party code,
@@
 for the client code to protect against `unforeseen" situations that may happen 
-while calling a library code.
-
-One may think that the solution for the uncaught exceptions may be to define a general handler, 
+while calling a library code. One may think that the solution for the uncaught exceptions may be to define a general handler, 
 which is responsible for handling any exception that is not
@@
 
-\subsection{On the null pointer problem}
-The null references was firstly introduced by Tony Hoare in ALGOL W, which after some years he called his ``one-billion-dollar mistake"~\cite{hoare2}.
-
-%:
-%\emph{``I call it my billion-dollar mistake. It was the invention of the null reference in 1965 [...] 
-%This has led to innumerable errors, vulnerabilities, and system 
-%crashes, which have probably caused a billion dollars of pain and damage in the last forty years."}
-
-In this sudy, the null references were, infact, responsible for several reported issues - providing more evidences to Hoare's speech.
-This observation emphasizes the importance of solutions to avoid NullPointerExceptions, such as:
-(i) lightweight intra-method null pointer analysis supported by Java8 @Nullable annotations~\footnote{They are used by tools e.g., Eclipse, IntelliJ, Android Studio 0.5.5 (release Apr.
-2014) to detect potential
-null pointer dereferences at compile time.};
-(ii) inter-method null pointer analysis tools such as the one proposed by Nanda and Sinha ~\cite{nanda2009accurate};
-or (iii) language designs which avoid null pointers, such 
-as Monads ~\cite{Walde95} (i.e., used in functional languages for values that may not be available 
-or computations that may fail) could improve the robustness of Java programs.
-
-\subsection{The exception handling confusion problem}
-When (mis)applied, the exception wrapping can make the exception-related code
- more complex and lead to what we call an \emph{exception handling confusion problem}.
-Such problem can lead the program to an unpredictable state in the presence of exceptions,
-as illustrated by the scenario on which a checked exception wrapped an OutOfMemoryError. 
-Currently there is no way of enforcing Java exception type conventions during program development.
-Hence, further investigation is needed on finding ways to help developers in dealing with such
- problem, either preventing odd wrappings or enabling the developer to better deal with them or even
-empirical studies on the real usefulness of Java hybrid exception model.
-
-%The cross-type wrappings detected in this study points to the fact that: (1) exception 
-%wrappings may prevent exception types from being used according to its initial purpose
-%(e.g, Errors should represent situations that should not be handled); and (2) may have been
-% used to bypass the language restrictions imposed by checked exceptions  (e.g,
-% checked exceptions may be wrapped in runtime exceptions).
-
 \section{Threats to Validity}
@@
 exceptions or stack traces. Our limited manual inspection did not reveal such
-cases. Moreover, in this study we manually classified the exceptions 
-related to programming mistakes. To ensure the quality of the analysis, 
-we calculated the interrater agreement after three independent 
+cases. Moreover, in this study we manually mapped the concerns related to exceptions.
+ To ensure the quality of the analysis, we calculated the interrater agreement after three independent 
 developers classified a randomly selected sample (of 25 exception 
@@
 \textit{Analysis and Use of Stack Trace Information.} Several research works have
-investigated the use of stack trace information to support bug classification
+investigated the use of stack trace information to support: bug classification
 and clustering~\cite{wang2013improving, kim2011crash, dhaliwal2011classifying},
-fault-proneness prediction models~\cite{kim2013predicting} and even automated
-bug fixing tools~\cite{sinha2009fault}. Kim et al.~\cite{kim2011crash} use an
+fault prediction models~\cite{kim2013predicting}, automated
+bug fixing tools~\cite{sinha2009fault} and also the analysis of Android APIs~\cite{kechagia2014}. 
+Kim et al.~\cite{kim2011crash} use an
 aggregated form of multiple stack traces available in crash reports to detect
@@
 related to NullPointerException and other implicitly thrown exceptions had
-the higher prevalence on the analyzed set of stacks. Kechagia and Spinellis ~\cite{kechagia2014}
+the higher prevalence on the analyzed set of stacks. Kechagia and Spinellis~\cite{kechagia2014}
 examined the stack traces embedded on crash reports sent by 1,800 Android apps 
@@
 empirical study considering the fault-proneness of aspect-oriented implementations 
-for handling exceptions. Two releases of both Java and AspectJ implementation of a 
-product line were assessed as part of that study. Based on the use of an exception
+for handling exceptions. Two releases of both Java and AspectJ implementations were 
+assessed as part of that study. Based on the use of an exception
 flow analysis tool, the study revealed that the AOP  refactoring increased the 
@@
 Android API and 10 dependent applications using the version history data found
-in github. The study found that approximately 25\% of all methods in the client
+in GitHub. The study found that approximately 25\% of all methods in the client
 code used the Android API, and that the methods reusing fast-evolving APIs were
@@
 used Android APIs that were on average 300\% more change-prone than the APIs
-used by the most successful apps. Pingyu and Elbaum~\cite{Zhang12} analyzed bug
-reports of 5 Android applications an observed that 29\% had to do with poor
-exceptional handling code. Our work differs from the others as it aims at
-distilling the information of bug reports distilling the uncaught exceptions 
-occurring at mobile applications in Github and Google Code. Such analysis discusses about
-the causes of bugs related to exceptions across applications available and what the
-characteristics of the stack traces can tell us about the exception structure of
-those applications.
+used by the most successful apps. Our work differs from the others as it aims at
+distilling stack trace information of bug reports and combine such information
+with bytecode analysis, source code analysis and manual inspections
+to identify bug hazards on the exception handling code of Android apps.
 

### 2015-02-13 android-stacks.tex
changes introduction

@@
  to an unpredictable state. Such wrappings suggest that, when not adequately used, the Java's hybrid exception
- model may become fault-prone.
+ model may negatively impact the application robustness.
 
@@
  The high prevalence of NullPointerExceptions is aligned with the findings of other research works ~\cite{kim2013predicting,fraser20131600,csallner2004jcrasher} as well as the undocumented runtime 
-exceptions signaled by the Android Platform ~\cite{kechagia2014}. If on the one hand, the findings our
+exceptions signaled by the Android Platform ~\cite{kechagia2014}. On the one hand, the findings our
  study emphasize the impact of these bug harzards on the application robustness 
-by mining a different source of information: the issues created by developers on GitHub and Google Code. On the other hand, 
-our work points to bug hazards that were not detected by previous works (i.e., cross-type wrappings, undocumented checked exceptions and 
-undocumented runtime exceptions thrown by third-party libraries) which can also negatively affect the application robustness.
+by mining a different source of information as the ones used in previous works. The present work 
+mined issues created by developers on GitHub and Google Code, while other works analyzed crash and automated test reports.
+ On the other hand, our work points to bug hazards that were not detected by previous works (i.e., cross-type wrappings, undocumented checked exceptions and undocumented runtime exceptions thrown by third-party libraries) which can also negatively affect the application robustness.
 

### 2015-02-13 android-stacks.tex
changes introduction

@@
  The high prevalence of NullPointerExceptions is aligned with the findings of other research works ~\cite{kim2013predicting,fraser20131600,csallner2004jcrasher} as well as the undocumented runtime 
-exceptions signaled by the Android Platform ~\cite{kechagia2014}. On the one hand, the findings our
+exceptions signaled by the Android Platform ~\cite{kechagia2014}. On the one hand, some of the findings of our
  study emphasize the impact of these bug harzards on the application robustness 

### 2015-02-13 android-stacks.tex
typpos

@@
 
-This paper reports a study which mined the Java stack traces included in
+This paper reports a study which mined the exception stack traces included in
 159,048 issues reported on Android projects hosted in GitHub (482 projects) 
 and Google Code (157 projsects). The goal of this study is to investigate 
-whether the analysis of stack trace information could reveal bug hazards related to the exception handling code,
+whether the analysis of stack trace information could reveal \emph{bug hazards} related to the exception handling code,
  that may lead to a decrease in application robustness. 
@@
 Some of the outcomes of this study include the identification of the following 
-bug hazards: (i) unexpected exception wrappings (such as Errors being wrapped in checked exceptions) -
-indicating that Java's hybrid exception model become fault-prone when it is not fully used in 
-accordance with its purpose; (ii) undocumented runtime exceptions thrown by both the Android 
-platform and third party libraries; and (iii) undocumented checked exceptions thrown by the Android Platform.
-This study provides insights on such bug hazards and the robustness threats  
+\emph{bug hazards}: (i) unexpected cross-type exception wrappings (for instance, trying to handle an instance 
+of OutOfMemoryError ``hidden" in  a checked exception)  which can make the exception-related code 
+more complex and negatively impact the application robustness; (ii) undocumented runtime exceptions 
+thrown by both the Android platform and third party libraries; and (iii) undocumented checked exceptions
+ thrown by the Android Platform. Such undocumented exceptions make difficult, and most of the times infeasible
+for the client code to protect against `unforeseen" situations that may happen while calling a third-party code.
+This study provides further insights on such \emph{bug hazards} and the robustness threats  
 they impose not only to Android apps but also to other systems based on Java exception model.
@@
 In this paper we will be concerned with exception handling in Android apps,
-which reuse Java's exception handling model.
+which reuses Java's exception handling model.
 
@@
 it automatically terminates, while the system prints a stack trace to the console, 
-or log file ~\cite{gosling2000java}.  A typical Java stack trace consists of  the fully qualified name 
-the thrown exception and the ordered list of methods that were active on the call stack before 
+or on a log file ~\cite{gosling2000java}.  A typical Java stack trace consists of  the fully qualified name 
+of the thrown exception and the ordered list of methods that were active on the call stack before 
 such exception has occurred ~\cite{gosling2000java,bloch2008effective}.
@@
 The goal of this study is to investigate whether the reported exception stack traces
- can reveal common bug hazards in the exception-related code.  
-A bug hazard  ~\cite{binder2000testing}  is a circumstance that increases the 
+ can reveal common \emph{bug hazards} in the exception-related code.  
+A \emph{bug hazard}~\cite{binder2000testing}  is a circumstance that increases the 
 chance of a bug to be present in the software. An example of a bug hazard can 
@@
 
-To guide this investigation we compiled general guidelines on how to use
+To guide this investigation we compiled general guidelines on how to use Java
 exceptions proposed by Gosling~\cite{gosling2000java},
@@
 and source code analysis on the exception-related code of the Android 
-platform and Android applications. Some bug hazards consistently detected 
+platform and Android applications. Some \emph{bug hazards} consistently detected 
 during this mining study include:
@@
 Trying to handle an instance of OutOfMemoryError ``hidden" in  a checked exception may bring the program
- to an unpredictable state. Such wrappings suggest that, when not adequately used, the Java's hybrid exception
- model may negatively impact the application robustness.
+ to an unpredictable state. Such wrappings suggest that when (mis)applied, the exception wrapping can make 
+the exception-related code more complex and negatively impact the application robustness.
 
@@
 exception signaled by native C code invoked by the Android Platform, yet this exception was not declared
- of the Java Native Interface. This can lead to uncaught exceptions that are 
+ of the Java Native Interface invoking it. This can lead to uncaught exceptions that are 
 difficult to debug.
@@
 by mining a different source of information as the ones used in previous works. The present work 
-mined issues created by developers on GitHub and Google Code, while other works analyzed crash and automated test reports.
- On the other hand, our work points to bug hazards that were not detected by previous works (i.e., cross-type wrappings, undocumented checked exceptions and undocumented runtime exceptions thrown by third-party libraries) which can also negatively affect the application robustness.
+mined issues created by developers on GitHub and Google Code, while the previous works analyzed crash reports and automated test reports.
+ On the other hand, our work points to bug hazards that were not detected by previous works (i.e., cross-type wrappings, undocumented checked exceptions and undocumented runtime exceptions thrown by third-party libraries) which represent new threats to the application robustness.
 
@@
 Hence, the study results are relevant to Android and Java developers who may underestimate the effect of such
- bug hazards the application robustness, and how tricky it can be to prevent them. Moreover, such bug hazards 
-calls for improvements on languages and tools to deal with such bug hazards in Android and Java environments.
+ \emph{bug hazards} the application robustness, and how tricky it can be to prevent them. Moreover, such \emph{bug hazards} 
+calls for improvements on languages and tools to better deal with them in Android and Java environments.
 
@@
 Android is an open source platform for mobile devices based on Linux kernel,
-which also comprises a subset of the core Java library, a set of third-party libraries (e.g.  SQLite),
-the Dalvik runtime environment, and the Application Framework (which provides 
-higher-level services to the applications running on the platform) ~\cite{andguide}.
-The Android Application Framework also provides several supporting APIs, such as 
-the APIs to support concurrent development (e.g. AsyncTask) - 
-which are usually needed due to the high demand for responsiveness 
-in mobile environment (mobile apps can easily become 
-unresponsive due to resource limitation and excessive network access).
+which also comprises (i) a set of native libraries written in C/C++ 
+(e.g., WebKit, OpenGL, FreeType, SQLite, Media, C runtime library) - which 
+fulfill a wide range of functions including graphics drawing, SSL communication, 
+SQLite database management, audio and video playback etc; (ii) a set of Java Core Libraries 
+- which include a subset of Java standard libraries and mainly “wrappers” to access the set of C/C++ 
+native libraries using the Java Native Interface (JNI); (iii) the Dalvik runtime environment - 
+which was specifically designed to deal with the resource constraints of a mobile device; 
+ and (iv) the Application Framework which provides higher-level APIs to the applications
+ running on the platform.
 
@@
 (extends Error)~\cite{gosling2000java}. The checked exception received such name
- because they must be declared on the method's exception interface (i.e., a list of exceptions that a method 
+ because they must be declared on the method's \emph{exception interface} (i.e., a list of exceptions that a method 
 might raise during its execution) and the compiler statically checks if
@@
 Both runtime exceptions and errors are also known as unchecked exceptions, because 
-they do not need to be specified on method exception interface and do not trigger any 
+they do not need to be specified on method \emph{exception interface} and do not trigger any 
 compile time checking.
 
-By convention, instances of Error represent  unrecoverable conditions which usually results
+By convention, instances of Error represent  unrecoverable conditions which usually result
 from failures detected by the Java Virtual Machine due to resource limitations such as OutOfMemoryError;
@@
 error, null pointer references). Some programming languages react to such errors by peremptorily terminating the program; 
-other programming languages allow similar approaches e.g.C\#; and others let the program to continue
+other programming languages allow similar approaches e.g., C\#; and others let the program to continue
  its execution in some situations such as the out-of-bounds array index e.g., C++. 
-The programs are not expected to handle such runtime exceptions signaled by the runtime environment according 
-to the Java Specification ~\cite{gosling2000java}. The user-defined exceptions can be either checked
+According to the Java Specification~\cite{gosling2000java} the programs are not 
+expected to handle such runtime exceptions signaled by the runtime environment. 
+The user-defined exceptions can be either checked
  or runtime. There is a long-lasting debate 
@@
 Section ~\ref{sec:best} presents
-a set of best practices related to each exception type.
+a set of best practices related to each of them.
 
@@
 an exception stack trace which illustrates an exception wrapping. 
-For simplicity we will refer to exception stack trace as stack trace along the text.
+For simplicity along the text we will refer to exception stack trace as stack trace.
 The bottom part of the stack trace is the \emph{root exception}, which indicates
@@
  We compiled a list of the guidelines related to exception types in Java, 
-summarized below:\footnote{We also compiled guidelines related
-to exception handling but they are out of the scope of this paper.}
+summarized below\footnote{We also compiled aditional guidelines related
+to ways of handling exceptions but they are out of the scope of this paper.}:
 
@@
 \emph{I-Checked exceptions should be used to represent recoverable
-conditions} (\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,bloch2008effective})
+conditions} (\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,bloch2008effective}).
 The developer should use checked exceptions for conditions from which the caller
@@
 exceptions being thrown by it. According to~\cite{bloch2008effective}, it is
-also wise to document (either or on the exception interface or using @throws tag on Javadoc) 
-the explicitly thrown runtime exceptions as carefully as checked exceptions. 
-When mentioning reffering to `` explicit" thrown exceptions
-we  exclude the implicitly signaled runtime exceptions related to programming
+also wise to document (either on the \emph{exception interface} or using @throws tag on Javadoc) 
+the \emph{explicitly thrown} runtime exceptions as carefully as checked exceptions. 
+The \emph{explicitly thrown} exceptions exclude the ones implicitly 
+thrown by the runtime environment due to programming
 mistakes. Such practice is mostly needed when developing library/framework 
@@
 will be aware of all the exceptions a method can throw and can design the code
-to deal with them ~\cite{Robil00}.
-
-If the developer fails to follow such practices (specially when developing 
-library/framework code) it will be difficult or even impossible for the caller to 
+to deal with them ~\cite{Robil00}. If the developer fails to follow this practice 
+ it will be difficult or even impossible for the caller to 
 make effective use of such method~\cite{wirfs2006toward, bloch2008effective},
-and protect the system against unforeseen exceptions thrown by the utility code,
-which can lead to the uncaught exceptions - one of the main causes of 
-systems crashes. These best practices guided our exploratory study described next.
+and protect the system against unforeseen exceptions thrown by the utility code. 
+Such exceptions can lead to the uncaught exceptions - one of the main causes of 
+systems crashes.
 
@@
  %projects,
+% We investigate whether characteristics 
+%of exception stack traces can pinpoint best practices violations and therefore reveal \emph{bug hazards}.
 
@@
 \emph{bug hazards} on the exception-related code of both the applications and the underlying framework. 
-This exploration is guided by the set of best practices compiled in this study. We investigate whether characteristics 
-of exception stack traces can pinpoint best practices violations and therefore reveal bug hazards.
-As mentioned before, in this context ``bug hazards" are the characteristics on the exception-related code 
+As mentioned before, in this context \emph{bug hazards} are the characteristics on the exception-related code 
 that favor the introduction of failures such as uncaught exceptions.
 
+This exploration is guided by the set of best practices compiled in this study.
 To support this investigation, we developed a tool called ExceptionMiner (Section~\ref{sec:exceptionminer})
@@
 analysis. Moreover, manual inspections were also used to leverage
- the understanding of stack traces and support further discussions and insights.
+ the understanding of stack traces and support further discussions and insights (Section~\ref{sec:manual}).
 
@@
 (e.g., the application, a library, the Android platform)  (Figure~\ref{fig:overview} - (6)). 
-Manual inspections were also used to support this mining process (Figure~\ref{fig:overview} - (4), (7) and (9)) and support
-the search for bug hazards performed in this study  (Figure~\ref{fig:overview} - (8)). Next sections detail each step of this mining process.
+Manual inspection steps were used to support the mining process (Figure~\ref{fig:overview} - (4), (7) and (9)) and 
+the search for \emph{bug hazards}  (Figure~\ref{fig:overview} - (8)). Next sections detail each step of this mining process.
 
@@
 To identify Android projects, we performed a case insensitive search for the
-term \textsf{android} in the repository's names and short descriptions.  
+term ``android" in the repository's names and short descriptions.  
 Up to 23 Feb 2014,  when we queried GHTorrent, such heuristic could
@@
 trace contains relevant information concerning the exception structure of the
-projects analyzed - and therefore can review bug hazards on the exception-related code -  
+projects analyzed - and therefore can review \emph{bug hazards} on the exception-related code -  
 we opted for not restricting the analysis only on defect issues.
@@
 The Crawler retrieved all issues and comments for these projects.
-From this set, 724 defined at least 1 issue. Running the ExceptionMiner tool 
+From this set, 724 projects defined at least 1 issue. Running the ExceptionMiner tool 
  in this set we observed that 183 projects had at least one issue containing an exception stack trace.
@@
 The ExceptionMiner is a mining tool which can connect to different issue repositories, 
-extract issues, mine stack traces from them, distill exception stack trace information,
+extract issues, mine exception stack traces from them, distill exception stack trace information,
 and enable the execution of different analyses by combining exception stack 
@@
 GHTorrent database, and Google Code connector which is comprised of a Web Crawler,
- that can traverse Google Code Web interface and extract the project issues. 
+ that can traverse the Google Code web interface and extract the project issues. 
 Projects meta-data and the issues associated to each project are stored on a relational 
@@
 \emph{\textbf{Exception Stack Trace Distiller.}}
+This component is based on a combination between a parser (based on regular expressions) 
+and heuristics able to identify and filter exception names and stack traces inline with text. 
 This component distills the information that composes a stack trace.
@@
 This component also distills fine grained information of each attribute such as: the classes and packages associated to them.
-This component is based on a combination between a parser (based on regular expressions) 
-and heuristics able to identify and filter exception names and stack traces inline with text. 
 In contrast to existing issue parsing solutions such as Infozilla, the parser
@@
 stack traces every exception defined on a stack trace need to be classified according to its type
-(e.g. Error, Exception or RuntimeException). The module responsible for this analysis 
+(e.g. Error, checked Exception or RuntimeException). The module responsible for this analysis 
 uses Design Wizard~\cite{Brunet09} framework to inspect the bytecode of all exceptions reported on issues,
@@
 which includes all basic Java exceptions that can be thrown during the app execution,
-and exceptions thrown by Android core libraries. Moreover it also analyzed the 
-libraries and applications which defined exceptions reported on stack traces
+and exceptions thrown by Android core libraries. Moreover, it also analyzed the 
+exceptions reported on stack traces that were defined on applications and third-party libraries 
 (the tool only analyzed the last version available).
@@
 Table~\ref{tab:signalers} presents the heuristic adopted in this classification.
-To enable such classification, we needed to feed this module the information
+To enable such classification, we needed to feed this module with the information
 comprising all Java packages that compose: the Android Plaform;
@@
 also included the source code of reused libraries.
-Then based on this information and using pattern matching 
-between the signaler name and the packages this module identifies 
+Then, based on this information and using pattern matching 
+between the signaler name and the packages, this module identifies 
 the origin of the exception signalers. The exceptions were considered
- from libraries if their packages were not defined 
-on Android platform, or on core libraries or on applications.
+ from libraries if their packages were neither defined 
+on Android platform, nor on core libraries, nor on applications.
 Table~\ref{tab:signalers} summarizes the signaler classification.
@@
 \subsection{Manual Inspections}
+\label{sec:manual}
 In the context of the ExceptionMiner tool, manual inspections were used to:
@@
 apps and Android Plaform). When the exception could not be  
-found niether automatically, nor manually (because they were defined on a previous version
+found neither automatically, nor manually (because they were defined on a previous version
 of the app or lib) the exception were classified 
-in our study as "Undefined".  Only 31 exceptions 
+in our study as ``Undefined".  Only 31 exceptions 
 remained undefined - such exceptions were reported on 61 exception stack traces as 
@@
 \subsection{Replication Package}
-All the data used in this study is publicly available at https://github.com/souzacoelho/exceptionminer
+All the data used in this study is publicly available at https://github.com/souzacoelho/exceptionminer.
 Specifically we provide: (i) all issues related to Android projects found
@@
 quantitative and qualitative analysis of the outcomes. The structure of this section
- is centered on questions that guided the search for bug hazards on the exception-related code
+ is centered on questions that guided the search for \emph{bug hazards} on the exception-related code
 based on exception stack trace information.
@@
 
-After distilling the information available on the exception stack traces we could find 
+After distilling the information available on the exception stack traces, we could find 
 the exceptions commonly reported as the root causes of stack traces.
@@
 %Commented 2015
-%\emph{\textbf{The prevalence of programming mistakes on stack traces.}} One way of getting a broader view of the prevalence~\footnote{The prevalence %evaluates the proportion of a given characteristic over a group %of subjects.}  of programming mistakes on reported stack traces would be to manually %inspect the code responsible for throwning every exception reported  on issues.
-However, inspecting the source code related to %6005 stack traces can %easily become infeasible.
+%\emph{\textbf{The prevalence of programming mistakes on stack traces.}} 
+%One way of getting a broader view of the prevalence~\footnote{The prevalence
+ %evaluates the proportion of a given characteristic over a group 
+%of subjects.}  of programming mistakes on reported stack traces would be to manually 
+%inspect the code responsible for throwning every exception reported  on issues. 
+%However, inspecting the source code related to %6005 stack traces can %easily become infeasible.
 %Hence, we conducted an empirical investigation based on the following assumption: the name of the exception and its associated Javadoc 
@@
 Since it is infeasible to inspect the code responsible for throwning every exception reported in this study,
-the concern identification of each exception was based on meaning of the exception type defined on 
+the concern identification of each exception was based on meaning of the exception type - defined on 
 its Javadoc documentaion and on the Java specification. To exemplify: (i) an instance of ArrayOutOfBoundException  
-refers to a programming mistake according to its Javadoc; and (ii) the Java specification list all exceptions related to backward compatibility
-%issues ~\cite{javaback} (e.g., InstantiationError, VerifyError, IllegalAccessError).
+refers to a programming mistake according to its Javadoc; and (ii) the Java specification lists all 
+exceptions related to backward compatibility issues~\cite{javaback} (e.g., InstantiationError, VerifyError, IllegalAccessError).
 
@@
 which were reported on 95\% of all stack traces analyzed in this study. Hence, based on the inspection of the 
-Javadoc related to each exception, and Java specification, we identified the concern releated to each root exception. 
+Javadoc related to each exception and Java specification, we identified the concern releated to each root exception. 
 Table~\ref{tab:tophundrend} illustrates the result of this empirical evaluation. This table also illustrates the exceptions 
@@
 sample of 25 exception types (from the total 100) using the same list of concerns;
-the interater agreement was high (96\%). We could also observe that
+the interater agreement was high (96\%).
 We can therefore assume that the categories are descriptive
@@
 identified based on the package names of the signalers related to it on the stack traces (Section~\ref{sec:exceptionminer}).
-Table~\ref{tab:typeroottab} presents the types and origins of root exceptions of all stack traces analyzed.
+Table~\ref{tab:typeroottab} presents the types and origins of root exceptions of all analyzed stack traces.
 
@@
 checked exceptions a library can define a precise 
-exception interface ~\cite{miller1997issues} to its clients.
+\emph{exception interface} ~\cite{miller1997issues} to its clients.
  Since such libraries are heavily used  in several projects, such finding can be related to the libraries maturity.
@@
 \emph{\textbf{Inspecting Exception Interfaces.}}  According to the best practices mentioned before, runtime exceptions 
-should be documented as part of the exception interface of libraries/framework reusable
+should be documented as part of the \emph{exception interface} of libraries/framework reusable
 methods. To investigate the obedience to such practice, we firstly filtered out all the exceptions implicitly
@@
 We could find 79 methods (both from libraries and Android platform) that  explicitly threw a runtime exception 
-without listing such exception on the exception interface (i.e., using 
+without listing such exception on the \emph{exception interface} (i.e., using 
 throws clause on method signature). From this set only one method (defined on a library)
@@
 The study revealed: a checked exception being thown by a native method and not declared  
-on the ``exception interface" of such method and all methods on the invocation chain.
+on the \emph{exception interface} of such method and all methods on the invocation chain.
 }
@@
 a complete Github dataset. Instead, the goal of our study was to 
-pinpoint bug hazards on the exception-related code based on 
+pinpoint \emph{bug hazards} on the exception-related code based on 
 exception stack trace mining of a subset of projects.
@@
 refered to exceptions coming from methods defined on Android Application Framework
-and third-party libraries.  Additionaly,  the ``bug hazards" observed in this study are due to
+and third-party libraries.  Additionaly,  the \emph{bug hazards} observed in this study are due to
 characteristics of Java exception model, which can impose challenges 
@@
 happen during the exception propagation.  Such analysis revealed intrigging 
-bug hazards such as the cross-type exception wrappings not discussed in previous works.
+\emph{bug hazards} such as the cross-type exception wrappings not discussed in previous works.
 
@@
 reported on 639 open source Android projects. The goal of our study was
-to identify common bug hazards on the exception related code that can lead to 
+to identify common \emph{bug hazards} on the exception related code that can lead to 
 failures such as uncaught exceptions.
@@
 with bytecode analysis, source code analysis and manual inspections
-to identify bug hazards on the exception handling code of Android apps.
+to identify \emph{bug hazards} on the exception handling code of Android apps.
 
@@
 with source code and bytecode analysis to identify common characteristics of exception 
-stack traces to pinpoint bug hazards on the exception-related code. 
-In this study, bug hazards are characteristics on the exception-related code that favor the introduction
+stack traces to pinpoint \emph{bug hazards} on the exception-related code. 
+In this study, \emph{bug hazards} are characteristics on the exception-related code that favor the introduction
 of  uncaught exceptions and unintended handling - which can contribute to 
 degrade the robustness of applications.
-Some bug hazards consistently detected in this study were: 
+Some \emph{bug hazards} consistently detected in this study were: 
 (i) unexpected exception wrappings (e.g., Errors being wrapped in checked exceptions)  - 
@@
 reported a java.lang.NullPointerException as its root cause;
-The bug hazards discussed in this work can negatively affect not only the robustness of Android application 
+The \emph{bug hazards} discussed in this work can negatively affect not only the robustness of Android application 
 but also the robustness of any Java-based application. 

### 2015-02-13 android-stacks.tex
typpos

@@
 We can observe that most of the reported exceptions were runtime 
-(64.85\%); and most of them were signaled on methods defined either on the Application (47.3\%)
+(64.85\%); and most of them were signaled by methods defined either on the Application (47.3\%)
 or on the Android platform (34.3\%). We could also find runtime exceptions thrown by library code (17.7\%).
@@
 
-\emph{\textbf{Inspecting Exception Interfaces.}}  According to the best practices mentioned before, runtime exceptions 
+\emph{\textbf{Inspecting Exception Interfaces.}}  According to the best practices mentioned before, 
+ \emph{explicitly thrown} runtime exceptions 
 should be documented as part of the \emph{exception interface} of libraries/framework reusable
@@
 (defined either on Android  Aplication Framework or on third-party libraries) 
-explicitly signaling a runtime exception not-related to programming mistakes.
+explicitly signaling a runtime exception.
+
 Table~\ref{tab:runtimeinterface} presents the results of this inspection. 
@@
 
-This result is aligned with the results of other study conducted by  
-Sacramento et al ~\cite{sacramento2006unchecked} - which observed that the
-runtime exceptions in .Net programs are most often not documented; and 
+This result is aligned with the results of other two studies ~\cite{sacramento2006unchecked,kechagia2014}. 
+One conducted by  Sacramento et al ~\cite{sacramento2006unchecked} - which observed that the
+runtime exceptions in .Net programs are most often not documented. And other conducted by 
 Kechagia and Spinellis ~\cite{kechagia2014} which discovered a set of methods 
 on the Android API which do not documented runtime exceptions. One limitation 
-of the work of Kechagia and Spinellis is that it did filter out exceptions that should 
-although unchecked not be documented -  because they are implicitly signaled by the 
+of the work of Kechagia and Spinellis is that it did not filter out exceptions that,  
+although runtime, should not be documented because they were implicitly signaled by the 
 JVM due to resource restrictions or violations on Java semantic constraints.
 
-The ``undocumented runtime exceptions" that are explicitly siglened represent a threat to system robustness, specially
-when such exceptions are thrown by a third party code invoked inside the application 
-(e.g. libraries, or framework utility code). In such cases the developer usually does not have access to 
+When explicitly signaling a runtime exception and not documenting it, 
+the developer imposes a threat to system robustness, specially
+when such exceptions are thrown by a third party code (e.g. libraries, or framework utility code)
+ invoked inside the application. In such cases the developer usually does not have access to 
 the source code. Hence in the absence of the exception documentation it is very difficult or even impossible
@@
 \noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
-The study revealed: 267 exception stack traces were caused by ``undocumented runtime exceptions" thrown by libraries or the Android platform; 
-several methods defined on 11 android packages signaling direct instances of RuntimeException.
+The study revealed: 267 exception stack traces were caused by ``undocumented runtime exceptions" 
+thrown by libraries or the Android platform.
 }
@@
  %\begin{tabular}{|p{7.9cm} p{0.1cm}|}
-\begin{tabular}{lrr}
-    \hline
- \bfseries{Origin}   & \bfseries{methods} &  \bfseries{stacks}  \\ 
-\hline					
-Libraries	& 	44	& 	205	\\
-Android  	&	35	&	62	\\					
-%\emph{android packages}   &  &   \\ 
-%android.app	& 	5	& 	16	\\
-%android.content	& 	2	& 	2	\\
-%android.database	& 	1	& 	1	\\
-%android.graphics	& 	5	& 	11	\\
-%android.hardware	& 	4	& 	5	\\
-%android.media	& 	3	& 	3	\\
-%android.opengl	& 	1	& 	8	\\
-%android.os	& 	11	& 	12	\\
-%android.text	& 	1	& 	1	\\
-%android.video	& 	1	& 	1	\\
-%android.providers	& 	1	& 	2	\\
-%\hline					
-%\bfseries{All android} 	&	35	&	62	\\
-\hline					
-\bfseries{All}	&	79	&	267	\\
+\begin{tabular}{lrrrr}
+    \hline
+ \bfseries{Origin}   &  \bfseries{stacks}  & \bfseries{signaler methods} &  \bfseries{throws clause}  &  \bfseries{\@throws}  \\ 
+\hline					
+Libraries	& 	205	 & 	44	& 0	& 1	\\
+Android  	&	62 &	35	&0 &  0	\\					
+\hline					
+\bfseries{All}	&	267 &	79 &  0  & 1\\
 \hline					

### 2015-02-14 android-stacks.tex
Small textual changes to the abstract.
Latex quotes are ``...''

@@
 
-This paper reports a study which mined the exception stack traces included in
+This paper reports on a study mining the exception stack traces included in
 159,048 issues reported on Android projects hosted in GitHub (482 projects) 
-and Google Code (157 projsects). The goal of this study is to investigate 
-whether the analysis of stack trace information could reveal \emph{bug hazards} related to the exception handling code,
- that may lead to a decrease in application robustness. 
- Overall 6,005 exception stack traces were extracted, and analyzed in 
-combination with source code and bytecode analysis. 
-Some of the outcomes of this study include the identification of the following 
+and Google Code (157 projects). The goal of this study is to investigate 
+whether stack trace information can reveal \emph{bug hazards} related to exception handling code that may lead to a decrease in application robustness. 
+ Overall 6,005 exception stack traces were extracted, and subjected to source code and bytecode analysis. 
+The outcomes of this study include the identification of the following 
 \emph{bug hazards}: (i) unexpected cross-type exception wrappings (for instance, trying to handle an instance 
-of OutOfMemoryError ``hidden" in  a checked exception)  which can make the exception-related code 
+of OutOfMemoryError ``hidden'' in  a checked exception)  which can make the exception-related code 
 more complex and negatively impact the application robustness; (ii) undocumented runtime exceptions 
@@
  thrown by the Android Platform. Such undocumented exceptions make difficult, and most of the times infeasible
-for the client code to protect against `unforeseen" situations that may happen while calling a third-party code.
+for the client code to protect against ``unforeseen'' situations that may happen while calling third-party code.
 This study provides further insights on such \emph{bug hazards} and the robustness threats  
-they impose not only to Android apps but also to other systems based on Java exception model.
+they impose to Android apps as well as to other systems based on the Java exception model.
 

### 2015-02-14 android-stacks.tex
Various small textual changes to intro.
"research work" is normally not used in plural form.

Changed to just research.

@@
 have to face an increasing number of threats to application robustness
- arising from: failures in underlying middleware and hardware (e.g., camera, sensors);
+ arising from failures in the underlying middleware and hardware (e.g., camera, sensors);
 failures in third party services and libraries; compatibility issues~\cite{McDon13}; 
@@
 of the thrown exception and the ordered list of methods that were active on the call stack before 
-such exception has occurred ~\cite{gosling2000java,bloch2008effective}.
+the exception occurred ~\cite{gosling2000java,bloch2008effective}.
 
@@
 Then, using a custom tool called ExceptionMiner,
- which we develop specifically for this study, we mine stack traces from the issues reported 
+ which we developed specifically for this study, we mine stack traces from the issues reported 
 on 482 Android projects hosted in GitHub and 157 projects hosted in Google Code.
@@
 
-   \item Undocumented checked exceptions signaled by native C code.  Some flows reported a checked 
+   \item Undocumented checked exceptions signaled by native C code.  Some flows contained a checked 
 exception signaled by native C code invoked by the Android Platform, yet this exception was not declared
- of the Java Native Interface invoking it. This can lead to uncaught exceptions that are 
+ in the Java Native Interface invoking it. This can lead to uncaught exceptions that are 
 difficult to debug.
@@
  \item A multitude of programming mistakes - approximately 52\% of the reported stack traces can 
-be attributed to programming mistakes. In particular, 27.71\% of all stack traces reported a java.lang.NullPointerException 
-as its root cause.
+be attributed to programming mistakes. In particular, 27.71\% of all stack traces contained a java.lang.NullPointerException 
+as their root cause.
 
@@
 
- The high prevalence of NullPointerExceptions is aligned with the findings of other research works ~\cite{kim2013predicting,fraser20131600,csallner2004jcrasher} as well as the undocumented runtime 
-exceptions signaled by the Android Platform ~\cite{kechagia2014}. On the one hand, some of the findings of our
- study emphasize the impact of these bug harzards on the application robustness 
+ The high prevalence of NullPointerExceptions is in line with findings of earlier research~\cite{kim2013predicting,fraser20131600,csallner2004jcrasher}, as are the undocumented runtime 
+exceptions signaled by the Android Platform ~\cite{kechagia2014}. 
+Some of the findings of our
+ study emphasize the impact of these bug hazards on the application robustness 
 by mining a different source of information as the ones used in previous works. The present work 
-mined issues created by developers on GitHub and Google Code, while the previous works analyzed crash reports and automated test reports.
- On the other hand, our work points to bug hazards that were not detected by previous works (i.e., cross-type wrappings, undocumented checked exceptions and undocumented runtime exceptions thrown by third-party libraries) which represent new threats to the application robustness.
-
-Such outcomes point to threats not only to the development of robust Android apps, 
+mined issues created by developers on GitHub and Google Code, while  previous research analyzed crash reports and automated test reports.
+ Furthermore, our work points to bug hazards that were not detected by previous research (i.e., cross-type wrappings, undocumented checked exceptions and undocumented runtime exceptions thrown by third-party libraries) which represent new threats to application robustness.
+
+Our findings point to threats not only to the development of robust Android apps, 
 but also to the development of any robust Java-based system. 
 Hence, the study results are relevant to Android and Java developers who may underestimate the effect of such
- \emph{bug hazards} the application robustness, and how tricky it can be to prevent them. Moreover, such \emph{bug hazards} 
-calls for improvements on languages and tools to better deal with them in Android and Java environments.
-
-The remaining of this paper is organized as follows. 
-Section~\ref{sec:back} presents a
-background on Android platform and Java exception model. 
+ \emph{bug hazards} on the application robustness,
+and who have to face the difficulty of preventing them.
+Moreover, such \emph{bug hazards} 
+call for improvements on languages and tools to better support exception handling in Android and Java environments.
+
+The remainder of this paper is organized as follows. 
+Section~\ref{sec:back} provides the necessary
+background on the Android platform and Java exception model. 
 Section~\ref{sec:study} presents the study design. 

### 2015-02-14 android-stacks.tex
Added reference to paper on crashes in distributed systems.
From paper:

We found the majority of catastrophic failures could easily have been prevented by performing simple testing on error handling code

https://www.usenix.org/conference/osdi14/technical-sessions/presentation/yuan

@@
 Studies have shown that exception-related code is generally poorly understood and among the least tested parts of the system~\cite{miller1997issues,Robil00,shah2010understanding, 
-garcia2007extracting,garcia2001comparative,cabral2007exception,coelho2011unveiling}.
+garcia2007extracting,garcia2001comparative,cabral2007exception,coelho2011unveiling,yuan:2014.osdi}.
 As a consequence they may inadvertently negatively affect the system: exception-related code may introduce failures such as 

### 2015-02-14 android-stacks.tex
Addressed most bibtex warnings.
In particular year was missing in many cases,
as was the publisher.

Also expanded some of the conference acronyms.
Small changes to start of background section.

@@
 \subsection{The Android Platform} \label{sec:extypes}
-Android is an open source platform for mobile devices based on Linux kernel,
-which also comprises (i) a set of native libraries written in C/C++ 
-(e.g., WebKit, OpenGL, FreeType, SQLite, Media, C runtime library) - which 
+Android is an open source platform for mobile devices based on the Linux kernel.
+Android also comprises (i) a set of native libraries written in C/C++ 
+(e.g., WebKit, OpenGL, FreeType, SQLite, Media, C runtime library) to
 fulfill a wide range of functions including graphics drawing, SSL communication, 
 SQLite database management, audio and video playback etc; (ii) a set of Java Core Libraries 
-- which include a subset of Java standard libraries and mainly “wrappers” to access the set of C/C++ 
-native libraries using the Java Native Interface (JNI); (iii) the Dalvik runtime environment - 
-which was specifically designed to deal with the resource constraints of a mobile device; 
+including a subset of the Java standard libraries and various wrappers to access the set of C/C++ 
+native libraries using the Java Native Interface (JNI); (iii) the Dalvik runtime environment, which was specifically designed to deal with the resource constraints of a mobile device; 
  and (iv) the Application Framework which provides higher-level APIs to the applications
@@
 
-\bibliographystyle{IEEEtran}
+\bibliographystyle{plain}
 \bibliography{android-stacks}

### 2015-02-14 android-stacks.tex
Small changes in description of Java's exception model (II.B).

@@
 (extends Exception), the runtime exceptions (extends RuntimeException) and errors
-(extends Error)~\cite{gosling2000java}. The checked exception received such name
- because they must be declared on the method's \emph{exception interface} (i.e., a list of exceptions that a method 
+(extends Error)~\cite{gosling2000java}. Checked exception received their name
+ because they must be declared on the method's \emph{exception interface} (i.e., the list of exceptions that a method 
 might raise during its execution) and the compiler statically checks if
  appropriate handlers are provided within the system.
-Both runtime exceptions and errors are also known as unchecked exceptions, because 
-they do not need to be specified on method \emph{exception interface} and do not trigger any 
+Both runtime exceptions and errors are also known as ``unchecked exceptions'', as 
+they do not need to be specified on the method \emph{exception interface} and do not trigger any 
 compile time checking.
@@
 By convention, instances of Error represent  unrecoverable conditions which usually result
-from failures detected by the Java Virtual Machine due to resource limitations such as OutOfMemoryError;
-normally they cannot be handled inside the application.  Instances of RuntimException are implicitly 
+from failures detected by the Java Virtual Machine due to resource limitations, such as OutOfMemoryError.
+Normally these cannot be handled inside the application.  Instances of RuntimeException are implicitly 
 thrown by Java runtime environment when a program violates 
 the semantic constraints of the Java programming language (e.g., out-of-bounds array index, divide-by-zero 
-error, null pointer references). Some programming languages react to such errors by peremptorily terminating the program; 
-other programming languages allow similar approaches e.g., C\#; and others let the program to continue
- its execution in some situations such as the out-of-bounds array index e.g., C++. 
-According to the Java Specification~\cite{gosling2000java} the programs are not 
+error, null pointer references). Some programming languages react to such errors by immediately terminating the program, while
+other languages, such as C++, let the program  continue
+ its execution in some situations such as the out-of-bounds array index.
+According to the Java Specification~\cite{gosling2000java} programs are not 
 expected to handle such runtime exceptions signaled by the runtime environment.
-The user-defined exceptions can be either checked
- or runtime. There is a long-lasting debate 
+
+User-defined exceptions can be either checked
+ or unchecked, by extending either Exception or RuntimeException. There is a long-lasting debate 
 about the pros and cons of both approaches~\cite{javatut,stackoverlow,debate}
@@
 the runtime environment looks for the nearest enclosing exception handler
-(in Java it is represented by the try-catch block), and unwinds the execution stack if necessary.
- Looking for the handler on the invocation stack is claimed to increase software reusability, 
-since the invoker of an operation can handle it in a wider context ~\cite{miller1997issues}.
+(Java's try-catch block), and unwinds the execution stack if necessary.
+This search for the handler on the invocation stack aims at increasing software reusability, 
+since the invoker of an operation can handle the exception in a wider context ~\cite{miller1997issues}.
 
  A common way of  propagating exceptions in Java programs is through exception wrapping
- (also known as exception chaining). Exception wrapping allows one exception 
-to be wrapped in another one and re-thrown. Figure~\ref{fig:wrapping} shows 
-an exception stack trace which illustrates an exception wrapping. 
-For simplicity along the text we will refer to exception stack trace as stack trace.
+ (also known as exception chaining): One exception 
+is caught and wrapped in another which is then thrown instead. Figure~\ref{fig:wrapping} shows 
+an exception stack trace which illustrates such exception wrapping. 
+For simplicity, in this paper we will refer to ``exception stack trace'' as as just stack trace.
 The bottom part of the stack trace is the \emph{root exception}, which indicates
-the first reason (root cause) for the exception be thrown (in this case, the computer run out of
+the first reason (root cause) for the exception thrown (in this case, the computer run out of
 memory). The top part of the stack trace indicates the location of the exception
-manifestation (which will call \emph{exception wrapper} along this paper). The
+manifestation (which we will refer to as the \emph{exception wrapper} in this paper). The
 execution flow  between the root exception and the wrapper may

### 2015-02-14 android-stacks.tex
Small changes to best practices II.C

@@
 exceptions~\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,
-bloch2008effective}. Such guidelines do not focus on 
-advocating any specific exception type, but rather propose ways to effectively use each of them.
- We compiled a list of the guidelines related to exception types in Java, 
-summarized below\footnote{We also compiled aditional guidelines related
-to ways of handling exceptions but they are out of the scope of this paper.}:
+bloch2008effective}. Such guidelines do not
+advocate any specific exception type, but rather propose ways to effectively use each of them.
+Based on these, for the purpose of our analysis we compiled the following list of Java exception handling best practices.
 
@@
 
-%\textbf{I
 \emph{I-Checked exceptions should be used to represent recoverable
@@
 client can explicitly ignore the exception (swallowing, or converting it to
-other type) at the expense of the program's robustness~\cite{gosling2000java}.
-
-\emph{II-Error represents an unrecoverable conditions which should not be handled} 
+another type) at the expense of the program's robustness~\cite{gosling2000java}.
+
+\emph{II-Error represents an unrecoverable condition which should not be handled} 
 (\cite{gosling2000java}).  Errors should result from failures detected
 by the runtime environment which indicate resource deficiencies, invariant
-failures or other conditions, that make impossible for the program to recover.
+failures or other conditions, from which the program cannot possibly recover.
 
@@
 developers should either try to reuse the exception types already defined in the
-Java API or they should create a specific exception. Throwing general types such as a
-pure java.lang.Exception or a java.lang.RuntimeException is considered a bad practice.
+Java API or they should create a specific exception. Thus, throwing general types such as a
+pure java.lang.Exception or a java.lang.RuntimeException is considered bad practice.
 
@@
 (\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,bloch2008effective}).
-The exceptions thrown by a method are an important part of methods interface,
-and is required to use the method properly. The checked exceptions are already
-part of the  methods signature, and the method caller is aware of the checked
-exceptions being thrown by it. According to~\cite{bloch2008effective}, it is
-also wise to document (either on the \emph{exception interface} or using @throws tag on Javadoc) 
+The exceptions thrown by a method are an important part of its interface,
+since knowledge of them is required to use a method properly. 
+Checked exceptions are already
+part of the  method signature, making the method caller aware of the checked
+exceptions it may have to handle.
+
+According to~\cite{bloch2008effective}, it is
+also wise to document (either on the \emph{exception interface} or using the @throws tag in Javadoc) 
 the \emph{explicitly thrown} runtime exceptions as carefully as checked exceptions. 
-The \emph{explicitly thrown} exceptions exclude the ones implicitly 
+These \emph{explicitly thrown} exceptions exclude the ones implicitly 
 thrown by the runtime environment due to programming
-mistakes. Such practice is mostly needed when developing library/framework 
-code reusable code - as the developers do not have access to the source code. 
-Doing so, the clients of such library/framework
+mistakes.
+
+This practice is mostly needed when developing  working with library/framework 
+code -- as the developers do not have access to the source code. 
+Doing so, the clients of such a library/framework
 will be aware of all the exceptions a method can throw and can design the code
@@
  it will be difficult or even impossible for the caller to 
-make effective use of such method~\cite{wirfs2006toward, bloch2008effective},
+make effective use of such a method~\cite{wirfs2006toward, bloch2008effective},
 and protect the system against unforeseen exceptions thrown by the utility code. 
-Such exceptions can lead to the uncaught exceptions - one of the main causes of 
+This can lead to uncaught exceptions - one of the main causes of 
 systems crashes.

### 2015-02-14 android-stacks.tex
Rewrite of best practice IV (document all exceptions thrown).
to make it more succinct.

@@
 (\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,bloch2008effective}).
-The exceptions thrown by a method are an important part of its interface,
-since knowledge of them is required to use a method properly. 
-Checked exceptions are already
-part of the  method signature, making the method caller aware of the checked
-exceptions it may have to handle.
-
-According to~\cite{bloch2008effective}, it is
-also wise to document (either on the \emph{exception interface} or using the @throws tag in Javadoc) 
-the \emph{explicitly thrown} runtime exceptions as carefully as checked exceptions. 
-These \emph{explicitly thrown} exceptions exclude the ones implicitly 
-thrown by the runtime environment due to programming
-mistakes.
-
-This practice is mostly needed when developing  working with library/framework 
-code -- as the developers do not have access to the source code. 
-Doing so, the clients of such a library/framework
-will be aware of all the exceptions a method can throw and can design the code
-to deal with them ~\cite{Robil00}. If the developer fails to follow this practice 
- it will be difficult or even impossible for the caller to 
-make effective use of such a method~\cite{wirfs2006toward, bloch2008effective},
-and protect the system against unforeseen exceptions thrown by the utility code. 
-This can lead to uncaught exceptions - one of the main causes of 
-systems crashes.
+For checked exceptions, this is automatically the case.
+Bloch~\cite{bloch2008effective} furthermore recommends to document explicitly thrown
+run time exceptions, either using a throws declaration in the signature, or using
+the @throws tag in the Javadoc.
+Doing so, in particular for public APIs of libraries or frameworks,
+makes clients aware of all exceptions possibly thrown,
+enabling them to design the code to deal with them and use the API effectively~\cite{Robil00,wirfs2006toward}.
 

### 2015-02-14 android-stacks.tex
Small changes to the opening paragraphs of Study Design (III).

@@
 
-\subsection{The Android Platform} \label{sec:extypes}
+\subsection{The Android Platform}
 Android is an open source platform for mobile devices based on the Linux kernel.
@@
 
-\begin{figure*} \centering \includegraphics[scale=0.3]{studyoverview3.png}
-\caption{Study overview.} \label{fig:overview} \end{figure*}
+\begin{figure*} 
+\centering \includegraphics[width=\hsize]{studyoverview3.png}
+\caption{Study overview.}\label{fig:overview} 
+\end{figure*}
 
@@
 
-The goal of this exploratory mining study was to investigate whether the exception stack traces 
-embedded on issues of Android projects (hosted in GitHub and Google Code), can reveal 
-\emph{bug hazards} on the exception-related code of both the applications and the underlying framework. 
-As mentioned before, in this context \emph{bug hazards} are the characteristics on the exception-related code 
+The goal of our mining study is to explore to what extent exception stack traces 
+contained in issues of Android projects (hosted in GitHub and Google Code), can reveal 
+\emph{bug hazards} in the exception-related code of both the applications and the underlying framework. 
+As mentioned before, in this context \emph{bug hazards} are the characteristics of exception-related code 
 that favor the introduction of failures such as uncaught exceptions.
 
-This exploration is guided by the set of best practices compiled in this study.
-To support this investigation, we developed a tool called ExceptionMiner (Section~\ref{sec:exceptionminer})
+This exploration is guided by the set of best practices covered in Section~\ref{sec:best}.
+To support our investigation, we developed a tool called ExceptionMiner (Section~\ref{sec:exceptionminer})
 which extracts the exception stack traces embedded on issues, 
 and combines stack trace information with source code and bytecode 
-analysis. Moreover, manual inspections were also used to leverage
+analysis. Moreover, we use manual inspection to augment
  the understanding of stack traces and support further discussions and insights (Section~\ref{sec:manual}).
-
-Figure~\ref{fig:overview} gives an overview of the study.
-Firstly  the issues reported on Android projects hosted on GitHub and Google Code were recovered (Figure~\ref{fig:overview} - (1) and (2)).
-Then the stack traces embedded each issue were extracted and distilled (Figure~\ref{fig:overview} - (3)).
-The stack trace information were then combined with source code and bytecode analysis in order to discover 
-the type of the exceptions reported on the stack traces (e.g., error, runtime, checked)  (Figure~\ref{fig:overview} - (5))  and the origin of such exceptions 
-(e.g., the application, a library, the Android platform)  (Figure~\ref{fig:overview} - (6)). 
-Manual inspection steps were used to support the mining process (Figure~\ref{fig:overview} - (4), (7) and (9)) and 
-the search for \emph{bug hazards}  (Figure~\ref{fig:overview} - (8)). Next sections detail each step of this mining process.
-
-Our study focused on open-source apps, since the information needed to perform our study
-cannot be retrieved from commercial apps, whose issue report systems and 
-source codes are not publicly available. The Android open-source apps were also 
-target of other research studies ~\cite{Linar13,ahimed}.   
 In this study we explore the domain quantitatively and highlight interesting cases by 
 exploring cases qualitatively.
+
+Figure~\ref{fig:overview} gives an overview of our study.
+First,  the issues reported on Android projects hosted on GitHub (1) and Google Code (2) are recovered.
+Then the stack traces embedded each issue are extracted and distilled (3).
+The stack trace information is then combined with source code and bytecode analysis in order to discover 
+the type of the exceptions (5) reported on the stack traces (e.g., error, runtime, checked),
+and the origin (6) of such exceptions 
+(e.g., the application, a library, the Android platform). 
+Manual inspection steps (4, 7, 9) are used to support the mining process  and 
+the search for \emph{bug hazards}  (8). The next sections detail each step of this mining process.
+
+Our study focuses on open-source apps, since the information needed to perform our study
+cannot be retrieved from commercial apps, whose issue report systems and 
+source codes are generally not publicly available. 
+Open source Android open-source apps have also been the 
+target of other research~\cite{Linar13,Ruiz12} addressing the \emph{reuse} and API stability.
 
@@
 Such threats are similar to the ones imposed to other empirical studies 
-which also used free or open-source Android apps ~\cite{Linar13,McDon13,ahimed}.
+which also used free or open-source Android apps ~\cite{Linar13,McDon13,Ruiz12}.
 Moreover, several exception stack traces that support the findings of this study

### 2015-02-15 android-stacks.tex
Small changes to android apps on github (III.A).

@@
 
-This study used the dataset provided by the GHTorrent project~\cite{Gousi13}, 
+This study uses the dataset provided by the GHTorrent project~\cite{Gousi13}, 
 an off-line mirror of the data  offered through the Github API.  
@@
 term ``android" in the repository's names and short descriptions.  
-Up to 23 Feb 2014,  when we queried GHTorrent, such heuristic could
- find 2.542 repositories. Running the ExceptionMiner tool 
+Up to 23 February 2014,  when we queried GHTorrent, this resulted in 2,542 repositories. Running the ExceptionMiner tool 
  in this set we observed that 589 projects had at least one issue containing a stack trace.
 
-Then we performed further cleanup, inspecting the site of every Android project
+Then we performed a further clean up, inspecting the site of every Android project
 reporting at least one stack trace, to make sure that they represented real
@@
 The filtered set consisted of 482 apps. This set of 482 projects contained overall 31,592 issues from which 4,042 exception stack traces 
-were extracted. Issues on Github are different from issues on dedicated bug tracking tools such as 
+were extracted.
+
+Issues on Github are different from issues on dedicated bug tracking tools such as 
 Bugzilla and Jira. The most important difference is that there are no predefined fields
@@
 them at will. Therefore, an issue may have none or an arbitrary set of labels depending 
-on which repository it was created. Table~\ref{tab:lables} illustrates the ocurrences of different labels 
+on its repository. Table~\ref{tab:lables} illustrates the ocurrences of different labels 
 on the issues including exception stack traces.
 
-Based on the assumption that regardless the issue lables, every exception stack
-trace contains relevant information concerning the exception structure of the
-projects analyzed - and therefore can review \emph{bug hazards} on the exception-related code -  
-we opted for not restricting the analysis only on defect issues.
+Regardless of the issue labels, every exception stack
+trace may contain relevant information concerning the exception structure of the
+projects analyzed, and therefore can reveal \emph{bug hazards} on the exception-related code.
+Because of this, we opted for not restricting the analysis to just defect issues.
 

### 2015-02-15 android-stacks.tex
Small changes to Android apps in google code (III.B).

@@
 \subsection{Android Apps in Google Code}
-Google Code contains largely used open-source Android apps (e.g.
-K9Mail\footnote{K9Mail moved to Github but as a way of not loosing the project history it advises their users to report bugs on Google Code issue tracker: https://github.com/k9mail/k-9/wiki/LoggingErrors.}),
- however, differently from GitHub, Google Code does not provide an API to access the information related
- to hosted projects\footnote{Google code used to provided a Web service to such repository, but it was deactivated in June 2013 what Google called a `clean-up action".}.
-To overcome this limitation we needed to implement a Web Crawler (incorporated in ExceptionMiner tool described next) that navigated 
-through the web interface of Google Code projects extracting all issues and issue comments and storing in a relational database for later analysis.
+Google Code contains widely used open-source Android apps (e.g.
+K9Mail\footnote{K9Mail moved to Github but as a way of not loosing the project history it advises their users to report bugs on the Google Code issue tracker: \url{https://github.com/k9mail/k-9/wiki/LoggingErrors}.}).
+However, differently from GitHub, Google Code does not provide an API to access the information related
+ to hosted projects.\footnote{Google code used to provide a Web service to its repositories, but this was deactivated in June 2013 in what Google called a ``clean-up action''.}
+To overcome this limitation we needed to implement a Web Crawler (incorporated in ExceptionMiner tool described next) that navigates 
+through the web interface of Google Code projects extracting all issues and issue comments and storing them in a relational database for later analysis.
 
 To identify Android projects in Google Code, we performed a similar heuristic: we performed a case insensitive search 
-(on Google Code search interface) for the term ``android". On January 2014, when we queried Google Code, such heuristic could
- returned a list of 788  projects. This list comprised the seeds sent to our Crawler .
+(on the Google Code search interface) for the term ``android''.
+On January 2014, when we queried Google Code, this resulted in a list of 788  projects.
+This list comprised the seeds sent to our Crawler.
 
@@
  in this set we observed that 183 projects had at least one issue containing an exception stack trace.
- Then we performed further cleanup (similar to the one described previously) inspecting the site 
-of each project. As a result we could identify 157 Android projects.  This set contained overall 127,456 issues,
- from which 1,963 exception stack traces were extracted. Table~\ref{tab:lables} illustrates the ocurrences of different labels 
+ Then we performed further clean up (similar to the one described previously) inspecting the site 
+of each project. As a result we could identify 157 Android projects.  This set contained  127,456 issues in total,
+ from which 1,963 exception stack traces were extracted. Table~\ref{tab:lables} illustrates the occurrences of different labels 
 on the issues including exception stack traces. Differently from GitHub, in Google Code most of 
-the issues were labled as ``Defect". However, based on the same assumption described for the GitHub repository
- we considered all issues reporting stacks (regardless its labels).
+the issues were labeled as ``Defect''. However, based on the same assumption described for the GitHub repository
+ we considered all issues reporting stack traces (regardless its labels).
 

### 2015-02-15 android-stacks.tex
small changes in the Exception miner tool (III.C).

@@
 
-The ExceptionMiner is a mining tool which can connect to different issue repositories, 
+The ExceptionMiner is a tool which can connect to different issue repositories, 
 extract issues, mine exception stack traces from them, distill exception stack trace information,
 and enable the execution of different analyses by combining exception stack 
-trace information with bytecode and sourcecode analysis. The main components of ExceptionMiner are as follows:
+trace information with byte code and source code analysis. The main components of ExceptionMiner are the following:
 
@@
 with issue repositories. In this study two main connectors were created: one which connects to 
-GHTorrent database, and Google Code connector which is comprised of a Web Crawler,
- that can traverse the Google Code web interface and extract the project issues. 
-Projects meta-data and the issues associated to each project are stored on a relational 
+GHTorrent database, and a Google Code connector which is comprised of a Web Crawler
+ that can traverse the Google Code web interface and extract the project's issues. 
+Project meta-data and the issues associated with each project are stored in a relational 
 database.
@@
 This component distills the information that composes a stack trace.
- Some of the attributes extracted from the stack trace are:
- the root exception and its signaler, the exception wrappers and their corresponding signalers. 
-This component also distills fine grained information of each attribute such as: the classes and packages associated to them.
-In contrast to existing issue parsing solutions such as Infozilla, the parser
-created in this work can discover stack traces mixed with log information\footnote{In several 
+ Some of the attributes extracted from the stack trace are
+ the root exception and its signaler, as well as the exception wrappers and their corresponding signalers. 
+This component also distills fine grained information of each attribute such as the classes and packages associated to them.
+In contrast to existing issue parsing solutions such as Infozilla, our parser
+ can discover stack traces mixed with log file information\footnote{In several 
 exception stack traces, the exception frames were preceeded by logging information e.g., 
-"03-01 15:55:01.609 (7924): at android.app.ActivityThread.access\$600(ActivityThread.java:127) 
-"which could not be detected by existing tools.}.
+\texttt{03-01 15:55:01.609 (7924): at android.app.ActivityThread.access\$600(ActivityThread\-.java:127)} which could not be detected by existing tools.}.
 
 \emph{\textbf{Exception Type Analysis.}} To support a deeper investigation of the 
-stack traces every exception defined on a stack trace need to be classified according to its type
+stack traces every exception defined on a stack trace needs to be classified according to its type
 (e.g. Error, checked Exception or RuntimeException). The module responsible for this analysis 
-uses Design Wizard~\cite{Brunet09} framework to inspect the bytecode of all exceptions reported on issues,
+uses the Design Wizard framework \cite{Brunet09} to inspect the bytecode of all exceptions reported on issues,
 and a specific implementation to analyze the source code of Java Exception classes (when the
 bytecode was not available). It walks up the type hierarchy of a Java exception until it reaches a base exception type.
-This module analyzed all exceptions defined on Android platform (API level 19) - 
+
+With this module we analyzed all exceptions defined in the Android platform (Version 4.4, API level 19),
 which includes all basic Java exceptions that can be thrown during the app execution,
-and exceptions thrown by Android core libraries. Moreover, it also analyzed the 
+and exceptions thrown by Android core libraries. Moreover, we also analyzed the 
 exceptions reported on stack traces that were defined on applications and third-party libraries 
@@
 This module is responsible for classifying each signaler according 
-to its origin (i.e. Android Application Framework, Android Libcore, Application, Library). 
-Table~\ref{tab:signalers} presents the heuristic adopted in this classification.
-To enable such classification, we needed to feed this module with the information
+to its origin (i.e., Android Application Framework, Android Libcore, Application, Library). 
+Table~\ref{tab:signalers} presents the heuristics adopted in this classification.
+
+To conduct this classification, we provide this module with the information
 comprising all Java packages that compose: the Android Plaform;
  the Android Libcore; and each analyzed Application.
-To discover the packages composing the first two origins
-we could refer to the Android Specification.
-To discover the packages composing each application, this module 
-extracted the manifest files of each Android app
- - which defines the main packages that compose the applications.
- Whenever this file was not available, the tool recursively analyzed the 
-structure of source-code directories composing the app
-- filtering out the cases on which the application 
-also included the source code of reused libraries.
+
+To discover the packages for the first two origins
+we can use to the Android specification.
+To discover the packages for the third origin, the application itself, this module
+extracts the manifest files of each Android app, which defines the main packages that the applications consist of.
+ If this file is not available, the tool recursively analyzes the 
+structure of source-code directories composing the appm filtering out the cases in which the application also includes the source code of reused libraries.
 Then, based on this information and using pattern matching 
 between the signaler name and the packages, this module identifies 
-the origin of the exception signalers. The exceptions were considered
- from libraries if their packages were neither defined 
-on Android platform, nor on core libraries, nor on applications.
-Table~\ref{tab:signalers} summarizes the signaler classification.
+the origin of the exception signalers.
+
+The exceptions are considered to come
+ from libraries if their packages are neither defined 
+within the Android platform, nor on core libraries, nor on applications.
+Table~\ref{tab:signalers} summarizes this signaler classification.
 
@@
     \hline
-    \bfseries{android} & If the exception is thrown on a method defined in Android platform.\\
-    \bfseries{app}     & If the exception is thrown on an method defined on an Android app.\\
-    \bfseries{libcore} & If the exception is thrown on one of the core libraries reused by Android (i.e., org.apache.harmony, org.w3c.dom, sun.misc, org.apache.http, org.json, org.xml). \\
+    \bfseries{android} & If the exception is thrown in a method defined in Android platform.\\
+    \bfseries{app}     & If the exception is thrown in an method defined on an Android app.\\
+    \bfseries{libcore} & If the exception is thrown in one of the core libraries reused by Android (e.g., org.apache.harmony, org.w3c.dom, sun.misc, org.apache.http, org.json, org.xml). \\
     \bfseries{lib}     & If the exception is thrown on a method that was not defined by any of the elements above.\\

### 2015-02-15 android-stacks.tex
Pass over section III done.

@@
 \begin{figure*} 
-\centering \includegraphics[width=\hsize]{studyoverview3.png}
+\centering \includegraphics[width=.9\hsize]{studyoverview3.png}
 \caption{Study overview.}\label{fig:overview} 
@@
 \label{sec:manual}
-In the context of the ExceptionMiner tool, manual inspections were used to:
-(i) support on the identification of packages composing the Android platform, 
+In our experiments, the output of the ExceptionMiner tool was manually extended
+in order to 
+(i) support the identification of packages composing the Android platform, 
 libs and apps analyzed in this study (as described previously); and (ii)  
-identify the type of some exceptions reported on issues, 
+identify the type of some exceptions reported in issues 
 that were not be automatically identified by the ExceptionMiner tool
@@
 apps and Android Plaform). When the exception could not be  
-found neither automatically, nor manually (because they were defined on a previous version
-of the app or lib) the exception were classified 
-in our study as ``Undefined".  Only 31 exceptions 
-remained undefined - such exceptions were reported on 61 exception stack traces as 
-presented in Table ~\ref{tab:typeroottab}.
+found automatically or manually (because they were defined on a previous version
+of the app or lib), we classified the exception as ``Undefined''.  Only 31 exceptions 
+remained undefined, which occurred in 60 different exception stack traces (see Table~\ref{tab:typeroottab}).
 
 \subsection{Replication Package}
-All the data used in this study is publicly available at https://github.com/souzacoelho/exceptionminer.
+All the data used in this study is publicly available at \url{https://github.com/souzacoelho/exceptionminer}.
 Specifically we provide: (i) all issues related to Android projects found
-in GITHub and Google Code used in this study; (iii) all stack traces extracted
-from issues (ii) the results of manual inspection steps; (iv) the tool developed 
-in this work to support stack trace extraction and distilling.
+in GiTHub and Google Code used in this study; (ii) all stack traces extracted
+from issues; (iii) the results of manual inspection steps; (iv) the
+ExceptionMiner tool we developed to support stack trace extraction and distilling.
 

### 2015-02-15 android-stacks.tex
Small changes to first part of common root results.
Changed 3 'what can we learn' questions into just the
keywords as section title.

Also changed intro of IV to not refer to questions anymore.

@@
 This section presents the results of the study, providing both a
-quantitative and qualitative analysis of the outcomes. The structure of this section
- is centered on questions that guided the search for \emph{bug hazards} on the exception-related code
-based on exception stack trace information.
-
-\subsection{What can the Common Root Exceptions tell us? }
+quantitative and qualitative analysis of the outcomes. 
+We center our presentation of the results around \emph{bug hazards} we
+distilled from (1) common root exceptions; (2) exception types; and
+(3) exception wrappings.
+
+\subsection{Common Root Exceptions}
 
@@
  ranked by the number of distinct projects on which they were reported. 
-This table also shows how many times the signaler of such exception was a method defined on
+This table also shows how many times the signaler of such an exception was a method defined on
 the Android platform, the Android libcore, the application itself or a third-party library -
@@
 
-We can observe that most of the exceptions on this list are implicitly thrown by the
+We can observe that most of the exceptions in this list are implicitly thrown by the
  runtime environment due to programming mistakes  (e.g., out-of-bounds array index, division-by-zero, access to a null reference)
- or resource limitation (e.g., OutOfMemoryError). 
+ or resource limitations (e.g., OutOfMemoryError). 
 From this set the java.lang.NullPointerException was the most reported root cause (27.71\%). 
@@
 Regarding reusable code (e.g., libraries and frameworks), there is no consensus whether it is a good or a bad practice to 
-re-throw a NullPointerException. Some prefer to encapsulate such an exception on
+explicitly throw a NullPointerException. Some prefer to encapsulate such an exception on
 an instance of IllegalArgumentException, while others~\cite{bloch2008effective} argue that the
 NullPointerException makes the cause of the problem explicit and hence 
-should not be wrapped.
-
-The high prevalence of NullPointerException is aligned with the findings of other research
-works~\cite{kim2013predicting,fraser20131600,csallner2004jcrasher,kechagia2014}. For instance, Sunghun et
-al.~\cite{kim2013predicting} showed that 38\% the bugs related to exception handling in Eclipse project
- were caused by NullPointerException; Kechagia and Diomidis showed that NullPointerException was the  
+can be signalled by an API expecting a non-null argument.
+
+The high prevalence of NullPointerException is aligned with the
+findings of other 
+research~\cite{kim2013predicting,fraser20131600,csallner2004jcrasher,kechagia2014}. 
+For instance, Sunghun et
+al.~\cite{kim2013predicting} show that 38\% the bugs related to
+exception handling in the Eclipse project
+ are caused by NullPointerException.
+Furhtermore, Kechagia and Spinellis showed that the NullPointerException was the  
 most reported exception on the crash reports sent to BugSense (a bug report 
-management service for Android applications); 
-other works on robustness testing~\cite{maji2012empirical,csallner2004jcrasher} showed that most of the automatically 
-detected bugs were due to NullPointerException and exceptions implicitly-signaled by Java 
+management service for Android applications)  \cite{kechagia2014}.
+Other research on robustness testing~\cite{maji2012empirical,csallner2004jcrasher} shows that most of the automatically 
+detected bugs were due to NullPointerException and exceptions
+implicitly-signaled by the Java 
 environment due to programming mistakes or resource limitations
@@
 
-\subsection{What can the Exception Types tell us?}
+\subsection{Exception Types}
 %\subsection{Exploring the Information Embedded on Exception Types}
@@
 
-\subsection{What can the Exception Wrappings tell us?}
+\subsection{Exception Wrappings}
 

### 2015-02-15 android-stacks.tex
Finished sectcion IV.A / Root causes
Did extract-command refactoring for the boxed conclusions

Sharpened the first boxed conclusion.

Removed 'evaluation', as this is an empirical study, but not
an evaluation of results.

@@
 \usepackage{balance}
+
+\newcommand{\finding}[1]{
+\noindent\fbox{%
+\begin{minipage}{.97\columnwidth}%
+\emph{#1}%
+\end{minipage}}
+}
 
@@
 \emph{\textbf{Identifying the Concerns Related to Root Exceptions.}} To get a broader view of the root exceptions of stack traces,
-we performed an empirical evaluation whose goal was to identify the concerns related to the most reported root exceptions.
-Besides the exceptions related to programming mistakes mentined before, we also looked for exceptions related to some concerns that are known as sources of faults in mobile development: concurrency ~\cite{ama2012} backward compatibility ~\cite{McDon13}, security ~\cite{enck2011study,was2010} and resource management (IO, Memory, Batery) ~\cite{Zhang12}.
+we performed a manual inspection in order to identify the underlying
+concerns related to the most frequently reported root exceptions.
+Besides the exceptions related to programming mistakes mentioned before, we also looked for exceptions related to some concerns that are known as sources of faults in mobile development: concurrency ~\cite{ama2012} backward compatibility ~\cite{McDon13}, security ~\cite{enck2011study,was2010} and resource management (IO, Memory, Batery) ~\cite{Zhang12}.
 
 Since it is infeasible to inspect the code responsible for throwning every exception reported in this study,
-the concern identification of each exception was based on meaning of the exception type - defined on 
-its Javadoc documentaion and on the Java specification. To exemplify: (i) an instance of ArrayOutOfBoundException  
+the concern identification of each exception was based on intended
+meaning of the particular exception type, as defined in 
+its Javadoc documentaion and in the Java specification. 
+For example: (i) an instance of ArrayOutOfBoundException  
 refers to a programming mistake according to its Javadoc; and (ii) the Java specification lists all 
-exceptions related to backward compatibility issues~\cite{javaback} (e.g., InstantiationError, VerifyError, IllegalAccessError).
-
-To perform such mapping, we selected a subset of all reported root exceptions. This subset was composed by 100 root exceptions
-which were reported on 95\% of all stack traces analyzed in this study. Hence, based on the inspection of the 
-Javadoc related to each exception and Java specification, we identified the concern releated to each root exception. 
-Table~\ref{tab:tophundrend} illustrates the result of this empirical evaluation. This table also illustrates the exceptions 
+exceptions related to backward compatibility~\cite{javaback}, such as
+InstantiationError, VerifyError, and IllegalAccessError.
+
+To perform this concern analysis, we selected a subset of all reported
+root exception, consisting of 100 exceptions
+reported on 95\% of all stack traces analyzed in this study. Hence, based on the inspection of the 
+Javadoc related to each exception and the Java specification, we
+identified the underlying concern releated to each root exception. 
+Table~\ref{tab:tophundrend} contains the results of this analysis. This table also illustrates the exceptions 
 that could not be directly mapped to one of the aforementioned concerns, either because they were too general (i.e., java.lang.Exception,
-java.lang.RuntimeException, java.lang.Error) or because they were related to other concerns (e.g., specific to an application or given library).
+java.lang.RuntimeException, java.lang.Error) or because they were
+related to other concerns (e.g., specific to an application or a given library).
 
@@
 sample of 25 exception types (from the total 100) using the same list of concerns;
-the interater agreement was high (96\%).
-We can therefore assume that the categories are descriptive
-enough for covering the range of exception types analyzed.
+the inter-rater agreement was 96\%.
 
@@
 
-\noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
-A multitude of programming mistakes were reported as the main causes of exception stack traces.
-52\% of all projects reported at least one exception stack trace on which the NullPoiterException
-was the root cause.
+\finding{%
+More than 50\% of the uncaught exceptions are due to errors in
+programming logic, with the NullPointerException as most prevalent
+exception.
+For another 25\% the root cause relates to resource constraints.
 }
-\end{minipage}}
 

### 2015-02-15 android-stacks.tex
Updated first part of exception types / IV.Bo
Replace all ' ~' by just '~' (the intent of the ~ is to put
together different words).

Made all columns with numbers right aligned in tabulars.

@@
 failures in third party services and libraries; compatibility issues~\cite{McDon13}; 
- memory and battery restrictions; and noisy external resources ~\cite{Zhang12} 
+ memory and battery restrictions; and noisy external resources~\cite{Zhang12} 
 (e.g., wireless connections,GPS, bluetooth).
@@
 it automatically terminates, while the system prints a stack trace to the console, 
-or on a log file ~\cite{gosling2000java}.  A typical Java stack trace consists of  the fully qualified name 
+or on a log file~\cite{gosling2000java}.  A typical Java stack trace consists of  the fully qualified name 
 of the thrown exception and the ordered list of methods that were active on the call stack before 
-the exception occurred ~\cite{gosling2000java,bloch2008effective}.
+the exception occurred~\cite{gosling2000java,bloch2008effective}.
 
@@
  The high prevalence of NullPointerExceptions is in line with findings of earlier research~\cite{kim2013predicting,fraser20131600,csallner2004jcrasher}, as are the undocumented runtime 
-exceptions signaled by the Android Platform ~\cite{kechagia2014}. 
+exceptions signaled by the Android Platform~\cite{kechagia2014}. 
 Some of the findings of our
@@
 %\footnote{152 questions in Stackoverflow are related to this debate}. 
-Section ~\ref{sec:best} presents
+Section~\ref{sec:best} presents
 a set of best practices related to each of them.
@@
 This search for the handler on the invocation stack aims at increasing software reusability, 
-since the invoker of an operation can handle the exception in a wider context ~\cite{miller1997issues}.
+since the invoker of an operation can handle the exception in a wider context~\cite{miller1997issues}.
 
@@
   \centering
-  \begin{tabular}{rcccccccc}
+%  \begin{tabular}{rcccccccc}
+  \begin{tabular}{lrrrrrrrr}
     \hline
@@
 concerns related to the most frequently reported root exceptions.
-Besides the exceptions related to programming mistakes mentioned before, we also looked for exceptions related to some concerns that are known as sources of faults in mobile development: concurrency ~\cite{ama2012} backward compatibility ~\cite{McDon13}, security ~\cite{enck2011study,was2010} and resource management (IO, Memory, Batery) ~\cite{Zhang12}.
+Besides the exceptions related to programming mistakes mentioned before, we also looked for exceptions related to some concerns that are known as sources of faults in mobile development: concurrency~\cite{ama2012} backward compatibility~\cite{McDon13}, security~\cite{enck2011study,was2010} and resource management (IO, Memory, Batery)~\cite{Zhang12}.
 
@@
 To perform this concern analysis, we selected a subset of all reported
-root exception, consisting of 100 exceptions
-reported on 95\% of all stack traces analyzed in this study. Hence, based on the inspection of the 
+root exceptions, consisting of 100 exceptions
+reported in 95\% of all stack traces analyzed in this study. Hence, based on the inspection of the 
 Javadoc related to each exception and the Java specification, we
@@
 As mentioned before, using the ExceptionMiner tool in combination with manual inspections we could identify
-the root exception types (i.e., RuntimeException, Error, checked Exception) as well as its origins - which were 
-identified based on the package names of the signalers related to it on the stack traces (Section~\ref{sec:exceptionminer}).
+the root exception \emph{type} (i.e., RuntimeException, Error, checked Exception) as well as its \emph{origin} - which we 
+identified based on the package names of the signalers related to it in the stack traces (Section~\ref{sec:exceptionminer}).
 Table~\ref{tab:typeroottab} presents the types and origins of root exceptions of all analyzed stack traces.
 
-We can observe that most of the reported exceptions were runtime 
-(64.85\%); and most of them were signaled by methods defined either on the Application (47.3\%)
+We can observe that most of the reported exceptions are of type runtime 
+(64.85\%); and that the most common origins are methods defined either on the Application (47.3\%)
 or on the Android platform (34.3\%). We could also find runtime exceptions thrown by library code (17.7\%).
 
-We can also see, from Table ~\ref{tab:typeroottab}, that contrasting with the other origins, most of the 
- exceptions signaled on Android Libcore (i.e., the set of libraries reused by the Android) were 
-checked exceptions. Such set comprises: org.apache.harmony, org.w3c.dom, sun.misc, 
-org.apache.http, org.json, org.xml, javax. Signaling checked exceptions is considered a 
+We can also see, from Table~\ref{tab:typeroottab}, that in contrast with the other origins, most of the 
+ exceptions signaled on Android Libcore (i.e., the set of libraries reused by Android) are 
+checked exceptions. This set comprises: org.apache.harmony, org.w3c.dom, sun.misc, 
+org.apache.http, org.json, org.xml, and javax. Signaling checked exceptions is considered a 
 good practice (see best practice IV in Section~\ref{sec:best}) because by using 
 checked exceptions a library can define a precise 
-\emph{exception interface} ~\cite{miller1997issues} to its clients.
- Since such libraries are heavily used  in several projects, such finding can be related to the libraries maturity.
+\emph{exception interface}~\cite{miller1997issues} to its clients.
+ Since such libraries are widely used  in several projects, this
+ finding can be attributed to the libraries' maturity.
 
@@
 \centering
-\begin{tabular}{lccccrr}
+\begin{tabular}{lrrrrrr}
     \hline
@@
   \end{tabular}
-\caption{Types of root exceptions.}
+\caption{Types and origins of root exceptions.}
   \label{tab:typeroottab}
@@
 should be documented as part of the \emph{exception interface} of libraries/framework reusable
-methods. To investigate the obedience to such practice, we firstly filtered out all the exceptions implicitly
- signaled by  runtime environment (due to programming mistakes) - since these exceptions 
-should not be documented on the method signature.  Then we inspected the code each method 
+methods. To investigate the conformance to this practice, we firstly filtered out all the exceptions implicitly
+ signaled by  the runtime environment (due to programming mistakes) - since these exceptions 
+should not be documented on the method signature.  Then we inspected
+the code for each method 
 (defined either on Android  Aplication Framework or on third-party libraries) 
@@
 
-This result is aligned with the results of other two studies ~\cite{sacramento2006unchecked,kechagia2014}. 
-One conducted by  Sacramento et al ~\cite{sacramento2006unchecked} - which observed that the
+This result is aligned with the results of other two studies~\cite{sacramento2006unchecked,kechagia2014}. 
+One conducted by  Sacramento et al~\cite{sacramento2006unchecked} - which observed that the
 runtime exceptions in .Net programs are most often not documented. And other conducted by 
-Kechagia and Spinellis ~\cite{kechagia2014} which discovered a set of methods 
+Kechagia and Spinellis~\cite{kechagia2014} which discovered a set of methods 
 on the Android API which do not documented runtime exceptions. One limitation 
@@
 null pointer dereferences at compile time.};
-(ii) inter-method null pointer analysis tools such as the one proposed by Nanda and Sinha ~\cite{nanda2009accurate};
+(ii) inter-method null pointer analysis tools such as the one proposed by Nanda and Sinha~\cite{nanda2009accurate};
 or (iii) language designs which avoid null pointers, such 
-as Monads ~\cite{Walde95} (i.e., used in functional languages for values that may not be available 
+as Monads~\cite{Walde95} (i.e., used in functional languages for values that may not be available 
 or computations that may fail) could improve the robustness of Java programs.
@@
  and restart the application. However, such handler cannot replace a carefully designed exception 
-handling policy ~\cite{Robil00}, which can hardly be designed in the absence of 
+handling policy~\cite{Robil00}, which can hardly be designed in the absence of 
 third-party documentation on the exceptions that
@@
 mitigate this problem tools should be developed to automate the documentation of runtime exceptions
-scraping from library code, few solutions in this directions have been proposed so far ~\cite{van2005combining}.
+scraping from library code, few solutions in this directions have been proposed so far~\cite{van2005combining}.
 
@@
 set of crash data per application. Although some services exist 
-to collect crash data from mobile applications ~\cite{BugSe14,BugSn14,Googl14,Acra14},
+to collect crash data from mobile applications~\cite{BugSe14,BugSn14,Googl14,Acra14},
 they do not provide open access to the crash reports of their client applications.
@@
 Such threats are similar to the ones imposed to other empirical studies 
-which also used free or open-source Android apps ~\cite{Linar13,McDon13,Ruiz12}.
+which also used free or open-source Android apps~\cite{Linar13,McDon13,Ruiz12}.
 Moreover, several exception stack traces that support the findings of this study
@@
 observed that the actions inside handlers were very simple (e.g., logging and present a
-message to the user). Coelho et al.  ~\cite{coelho2008assessing} performed an 
+message to the user). Coelho et al.~\cite{coelho2008assessing} performed an 
 empirical study considering the fault-proneness of aspect-oriented implementations 

### 2015-02-15 android-stacks.tex
new title and bytecode explanation

@@
 
-\title{What can Issues on GitHub and Google Code tell us about Bug Hazards in Android and Java? }
+%\title{What can Issues on GitHub and Google Code tell us about Bug Hazards in Android and Java? }
+
+\title{Unveiling Exception Handling Bug Hazards in Android based on GitHub and Google Code Issues }
 
@@
 (e.g. Error, checked Exception or RuntimeException). The module responsible for this analysis 
-uses the Design Wizard framework \cite{Brunet09} to inspect the bytecode of all exceptions reported on issues,
-and a specific implementation to analyze the source code of Java Exception classes (when the
-bytecode was not available). It walks up the type hierarchy of a Java exception until it reaches a base exception type.
+uses the Design Wizard framework \cite{Brunet09} to inspect the bytecode of the exceptions reported 
+on stack traces.  It walks up the type hierarchy of a Java exception until it reaches 
+a base exception type. Hence in this study the bytecode analysis was only used to discover the type 
+of each mined exception when the jar file of such exception was available on the project or 
+on the reused core Java libraries. An specific implementation (based on source code analysis)
+ was needed to to discover the exception type when the bytecode was not available.
 
@@
 (defined either on Android  Aplication Framework or on third-party libraries) 
-explicitly signaling a runtime exception.
-
-Table~\ref{tab:runtimeinterface} presents the results of this inspection. 
+explicitly signaling a runtime exception. Table~\ref{tab:runtimeinterface} presents the results of this inspection. 
 We could find 79 methods (both from libraries and Android platform) that  explicitly threw a runtime exception 
@@
 JVM due to resource restrictions or violations on Java semantic constraints.
-
 When explicitly signaling a runtime exception and not documenting it, 

### 2015-02-15 android-stacks.tex
Sharpened conclusions on exception types (IV.B)

@@
 mobile applications. These applications extend phones capabilities 
-far beyond basic calls and textual messages. Such applications, however,
-have to face an increasing number of threats to application robustness
+far beyond basic calls and textual messages. They must, however,
+face an increasing number of threats to application robustness
  arising from failures in the underlying middleware and hardware (e.g., camera, sensors);
@@
 
- A common way of  propagating exceptions in Java programs is through exception wrapping
- (also known as exception chaining): One exception 
+ A common way of  propagating exceptions in Java programs is through
+ exception wrapping (also called chaining):
+One exception 
 is caught and wrapped in another which is then thrown instead. Figure~\ref{fig:wrapping} shows 
@@
 
+\finding{Almost two thirds of all crashes come from run time
+  exceptions. Most of these originate from the application layer.}
+
 \emph{\textbf{Inspecting Exception Interfaces.}}  According to the best practices mentioned before, 
@@
 the code for each method 
-(defined either on Android  Aplication Framework or on third-party libraries) 
-explicitly signaling a runtime exception. Table~\ref{tab:runtimeinterface} presents the results of this inspection. 
-We could find 79 methods (both from libraries and Android platform) that  explicitly threw a runtime exception 
-without listing such exception on the \emph{exception interface} (i.e., using 
-throws clause on method signature). From this set only one method (defined on a library)
-included an @throws tag on Javadoc - reporting that the given runtime exception
-could be thrown in some contidions. These methods were responsible for 
+(defined either in the Android  Aplication Framework or in third-party libraries) 
+explicitly signaling a runtime exception.
+
+Table~\ref{tab:runtimeinterface} presents the results of this inspection. 
+We found 79 methods (both from libraries and the Android platform) that  explicitly threw a runtime exception 
+without listing it on the \emph{exception interface} (i.e., using 
+throws clause in the method signature). From this set only one method (defined on a library)
+included an @throws tag in its Javadoc - reporting that the given runtime exception
+could be thrown in some conditions. These methods were responsible for 
 267 exception stack traces mined in this study.
 
-This result is aligned with the results of other two studies~\cite{sacramento2006unchecked,kechagia2014}. 
-One conducted by  Sacramento et al~\cite{sacramento2006unchecked} - which observed that the
-runtime exceptions in .Net programs are most often not documented. And other conducted by 
-Kechagia and Spinellis~\cite{kechagia2014} which discovered a set of methods 
-on the Android API which do not documented runtime exceptions. One limitation 
-of the work of Kechagia and Spinellis is that it did not filter out exceptions that,  
+This result is in line with the results of two other studies~\cite{sacramento2006unchecked,kechagia2014}. 
+Sacramento et al~\cite{sacramento2006unchecked} observed that the
+runtime exceptions in .NET programs are most often not documented. 
+Kechagia and Spinellis~\cite{kechagia2014} identified a set of methods 
+on the Android API which do not document its runtime exceptions. One limitation 
+of the latter work is that it did not filter out exceptions that,  
 although runtime, should not be documented because they were implicitly signaled by the 
@@
 When explicitly signaling a runtime exception and not documenting it, 
-the developer imposes a threat to system robustness, specially
-when such exceptions are thrown by a third party code (e.g. libraries, or framework utility code)
+the developer imposes a threat to system robustness, especially
+when such exceptions are thrown by third party code (e.g., libraries or framework utility code)
  invoked inside the application. In such cases the developer usually does not have access to 
 the source code. Hence in the absence of the exception documentation it is very difficult or even impossible
- for the client to design the application to deal with ``unforeseen" runtime exceptions. As a consequence, the
+ for the client to design the application to deal with ``unforeseen'' runtime exceptions. As a consequence, the
  undocumented runtime exception may remain uncaught and lead to system crashes.
 
-\noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
-The study revealed: 267 exception stack traces were caused by ``undocumented runtime exceptions" 
-thrown by libraries or the Android platform.
-}
-\end{minipage}}
+\finding{%
+Only a small fraction (4\%, 267 stack traces) of run time exceptions are programmatically
+thrown. Almost none (0.4\%, just one) of these were documented.}
 
@@
     \hline
- \bfseries{Origin}   &  \bfseries{stacks}  & \bfseries{signaler methods} &  \bfseries{throws clause}  &  \bfseries{\@throws}  \\ 
+ \bfseries{Origin}   &  \bfseries{stacks}  & \bfseries{signaler methods} &  \bfseries{throws clause}  &  \bfseries{@throws}  \\ 
 \hline					

### 2015-02-15 android-stacks.tex
Sharpened native method section, finishing IV.B.

@@
 
-\emph{\textbf{On the Presence of Checked Exceptions on Exception Interfaces.}} 
-In this study, the exception stack trace analysis revealed an unexpected bug hazard: a 
+\emph{\textbf{Missing Checked Exceptions on Exception Interfaces.}} 
+Our exception stack trace analysis revealed an unexpected bug hazard: a 
 checked exception thrown by a native method and not declared on the exception 
-interface of these methods signaling them. Such native method was defined on
- the Android platform -  which uses Java Native Invocation (JNI) to access 
+interface of these methods signaling them.
+
+The native method in question was defined in
+ the Android platform, which uses Java Native Invocation (JNI) to access 
 native C/C++ code. This exception was thrown by the method getDeclaredMethods defined in java.lang.Class. 
@@
 violating the declaration. The Java compiler could not detect this violation, because it does 
-not perform static exception checking on native methods. This type of bug is hard to diagnose
-because the developer usually do not have access to the native implementations. 
+not perform static exception checking on native methods.
+
+This type of bug is hard to diagnose
+because the developer usually does not have access to the native implementations. 
 Consequently, since it is not expected by the programmer, when such method throws 
-this exception, such undocumented exception may remain
-uncaught and cause the app to crash, or may be mistakenly handled by subsumption.
-The exception stack traces reporting this scenario is actually a real bug of Android 
+this exception, such an undocumented exception may remain
+uncaught and cause the app to crash, or maybe mistakenly handled by subsumption.
+The exception stack traces reporting this scenario actually correspond
+to a real bug of Android 
 Gingerbread version (which still accounts for 13.6\% of devices running Android).
 %Among the Android projects on which this exception stack trace was reported we can cite:
-%PageTurner (206 stars), ActionBarSherlock (6,041 stars), otto (1,618 stars).
-
-\noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
-The study revealed: a checked exception being thown by a native method and not declared  
-on the \emph{exception interface} of such method and all methods on the invocation chain.
+%PageTurner (206 stars), ActionBarSherlock (6,041 stars), otto (1,618
+%stars).
+
+\finding{%
+For native methods, even checked exceptions can be
+thrown without being documented on the exception interface.
 }
-\end{minipage}}
 

### 2015-02-16 android-stacks.tex
Sharpened Exception wrappings / IV.C

@@
     \hline
- id & \bfseries{Runtime wrapping Error}    \\  %\bfseries{\#}
+ id & \bfseries{Runtime Exception wrapping an Error}    \\  %\bfseries{\#}
     \hline
@@
 \hline
-& \bfseries{Checked wrapping Error}  \\
+& \bfseries{Checked Exception wrapping an Error}  \\
  \hline
@@
 \hline
-& \bfseries{Error wrapping Checked}  \\
+& \bfseries{Error wrapping a Checked Exception}  \\
  \hline
@@
 \hline
-& \bfseries{Error wrapping Runtime}    \\
+& \bfseries{Error wrapping a Runtime Exceptino}    \\
  \hline
@@
 which offers three kinds of exceptions each one holding an intended exception behavior (i.e., error, 
-runtime and checked). Table~\ref{tab:wrappingandroid} presents the wrappings found in this study that
-include different exception types (i.e., Error, checked Exception and Runtime). We call 
-such a wrapping a ``cross-type wrapping". Next we exemplify some of such
- wrappings and discuss whether some of them may impact the system robustness.
-
-\emph{\textbf{Runtime Exception wrapping Error.}} From Table~\ref{tab:wrappingandroid}, we could observe that most of such wrappings
-were performed by Android platform (50.7\%).
+runtime and checked). Table~\ref{tab:wrappingandroid} presents some wrappings found in this study that
+include different exception types (i.e., Error, checked Exception and Runtime).
+Below, we discuss the most
+important of such ``cross-type wrappings in more detail.
+
+\emph{\textbf{Runtime Exception wrapping an Error.}} 
+From Table~\ref{tab:wrappingandroid}, we see that most of these wrappings
+are performed by the Android platform (50.7\%).
 The code snippet below was extracted from Android and shows a general catch clause 
@@
 of an asynchronous task) into an instance of RuntimeException and re-throws it.  
-Table~\ref{tab:exampeswrap} presents exceptions wrapped in this code snipet.  
-Such wrappings mask an unrecoverable Error on a general runtime exception.
+Table~\ref{tab:exampeswrap} presents examples of exceptions that were actually wrapped in this code snipet.  
+Such wrappings mask an unrecoverable Error into a general runtime exception.
 
@@
 
-\emph{\textbf{Runtime wrapping Checked Exception.}} This wrapping was responsible for 49.5\% of
+\emph{\textbf{Runtime Exception wrapping a Checked Exception.}} 
+This wrapping was responsible for 49.5\% of
 the cross-type wrappings.  From this set 50\%  were performed on methods defined on Android platform. 
-We observed that it was a common implementation practice in the methods of Android platform.
-However, using such general exception, is considered a bad practice according to Java specification 
-and compiled guidelines, because it loses contextual information about the exception.
-
-\emph{\textbf{Checked Exception wrapping Error.}} Most of these wrappings
+We observe that it is a common implementation practice in the methods of Android platform.
+However, using such a general exception, is considered a bad practice
+according to the Java specification 
+and common guidelines, as it loses contextual information about the exception.
+
+\emph{\textbf{Checked Exception wrapping an Error.}} Most of these wrappings
  were also caused by the reflection library used by applications' methods. The methods responsible for the wrappings
@@
 into a checked exception. Such wrappings may also mask an unrecoverable error
-and may lead to the ``exception confusion problem" described next.
+and may lead to ``exception confusion'' described next.
 
 \emph{\textbf{Error wrapping Runtime and Checked Exceptions}} Table~\ref{tab:exampeswrap}  illustrates
-examples of instances of Error wrapping instances of RuntimeException. A
-lthough such wapping mixes different exception types,
-since there is no obligation associated to handling runtime exceptions, such wrapping 
-does not violated the aforementioned best practices. On the other hand, the inspection 
+examples of instances of Error wrapping instances of
+RuntimeException. 
+Although such a wapping mixes different exception types,
+since there is no obligation associated to handling runtime exceptions, it
+does not violate the aforementioned best practices.
+
+On the other hand, the inspection 
 also revealed instances of Error wrapping checked exceptions. Such wrappings were 
@@
 on the point where the class is first used.  Table~\ref{tab:exampeswrap} also illustrates
-examples of such wrappings. Although such wapping may represent a design decision,
-it violates the best practice related to checked exceptions and errors as it mix the intended handling 
+examples of such wrappings. Although such a wrapping may represent a design decision,
+it violates the best practice related to checked exceptions and errors as it mixes the intended handling 
 behaviour associated to both types.
 
-We could also observe that some stack traces included successive cross-type wrappings, 
-such as: Runtime - Checked - Runtime - Checked - Runtime - Checked - Runtime.
-Hence, although some of these wrappings may be a result of design decisions, the mis-use of exception wrapppings may make the exception handling
+We can also observe that some stack traces include successive cross-type wrappings, 
+such as: Runtime - Checked - Runtime - Checked - Runtime - Checked -
+Runtime. 
+Hence, although some of these wrappings may be a result from design decisions, the mis-use of exception wrapppings may make the exception handling 
 code more complex (e.g., the multiple wrappings) and error-prone,
- and lead to what we call ``exception confusion problem". To illustrate this problem we can use one of the wrappings discussed above.
+ and lead to ``exception confusion''.
+
+To illustrate this problem we can use one of the wrappings discussed above.
 When the developer is confronted with a checked exception, the designer of the API is telling him/her 
 to handle the exceptional condition (according to Java Specification and 
-best practices). However, such exception may be wrapping an Error such as OutOfMemoryError (which indicate resource deficiency 
-that makes impossible for the program to recover). Hence, trying to handle such exception  
+best practices). However, such exception may be wrapping an Error such
+as an OutOfMemoryError, which indicates a resource deficiency 
+that the program cannot possibly recover from). Hence, trying to
+handle such an exception  
 may lead the program to an unpredictable state.
 
-\noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
-The cross-type wrappings is an indication that even if the developer sticks to Java
-specification and best practices, it may be dificult for her/him to prevent the code from 
-incurring into situations that lead the program to an unpredictable state - 
-e.g., when handling checked Exceptions that are `hiding" Errors.
+\finding{%
+Cross-type exception wrappings are common.
+They violate the semantics of Java's original exception
+design (e.g., when mapping unrecoverable Errors to other types of
+exceptions),
+and may lead to lengthy exception chains.
 }
-\end{minipage}}
 

### 2015-02-16 android-stacks.tex

@@
 %\emph{\textbf{The prevalence of programming mistakes on stack traces.}} 
-%One way of getting a broader view of the prevalence~\footnote{The prevalence
+%One way of getting a broader view of the prevalence~\footnote The prevalence
  %evaluates the proportion of a given characteristic over a group 
@@
 
-\section{Further Discussions}
+\section{Discussion}
 \label{sec:disc}
@@
 
- ``Everybody hates thinking about exceptions, because they are not supposed to happen"
-  (Brian Foote)~\footnote{Brian Foote shared his opinion in a conversation with James Noble - quoted on the paper: hillside.net/plop/2008/papers/ACMVersions/coelho.pdf}
+\begin{quotation}
+\noindent
+ \emph{``Everybody hates thinking about exceptions, because they are not supposed to happen''\\
+  (Brian Foote)}\footnote{Brian Foote shared his opinion in a conversation with James Noble - quoted on the paper: hillside.net/plop/2008/papers/ACMVersions/coelho.pdf}
+\end{quotation}
 
 \emph{\textbf{The exception handling confusion problem.}}
-When (mis)applied, the exception wrapping can make the exception-related code
- more complex and lead to what we call an \emph{exception handling confusion problem}.
-Such problem can lead the program to an unpredictable state in the presence of exceptions,
-as illustrated by the scenario on which a checked exception wrapped an OutOfMemoryError. 
+When (mis)applied, exception wrapping can make the exception-related code
+ more complex and lead to what we call the \emph{exception handling confusion problem}.
+This problem can lead the program to an unpredictable state in the presence of exceptions,
+as illustrated by the scenario in which a checked exception wraps an OutOfMemoryError. 
 Currently there is no way of enforcing Java exception type conventions during program development.
-Hence, further investigation is needed on finding ways to help developers in dealing with such
- problem, either preventing odd wrappings or enabling the developer to better deal with them or even
-empirical studies on the real usefulness of Java hybrid exception model.
+Hence, further investigation is needed on finding ways to help developers in dealing with this
+ problem, either preventing odd wrappings or enabling the developer to
+ better deal with them.
+Furthermore, this calls for empirical studies on the actual usefulness of Java's hybrid exception model.
 
@@
 \emph{\textbf{On the null pointer problem.}}
-The null references was firstly introduced by Tony Hoare in ALGOL W, which after some years he called his ``one-billion-dollar mistake"~\cite{hoare2}.
-In this sudy, the null references were, infact, responsible for several reported issues - providing more evidences to Hoare's speech.
-This observation emphasizes the importance of solutions to avoid NullPointerExceptions, such as:
-(i) lightweight intra-method null pointer analysis supported by Java8 @Nullable annotations~\footnote{They are used by tools e.g., Eclipse, IntelliJ, Android Studio 0.5.5 (release Apr.
-2014) to detect potential
+The null reference was firstly introduced by Tony Hoare in ALGOL W, which after some years he called his ``one-billion-dollar mistake''~\cite{hoare2}.
+In this sudy, the null references were, infact, responsible for several reported issues - providing further evidence to Hoare's statement.
+This observation emphasizes the need for solutions to avoid NullPointerExceptions, such as:
+(i) lightweight intra-method null pointer analysis as supported by
+Java 8 @Nullable annotations\footnote{Already supported by tools such as Eclipse, IntelliJ, Android Studio 0.5.5 (release Apr. 2014) to detect potential 
 null pointer dereferences at compile time.};
@@
 or (iii) language designs which avoid null pointers, such 
-as Monads~\cite{Walde95} (i.e., used in functional languages for values that may not be available 
+as Monads~\cite{Walde95} (as used in functional languages for values that may not be available 
 or computations that may fail) could improve the robustness of Java programs.
 
-\emph{\textbf{On the need of rescues to the almost infeasible task of preventing uncaught exceptions}}
+\emph{\textbf{Preventing uncaught exceptions}}
 In this study we could observe undocumented runtime exceptions thrown by third party code,
 and even undocumented checked exception thrown by a JNI interface.
-Such undocumented exceptions make difficult, and most of the times infeasible
-for the client code to protect against `unforeseen" situations that may happen 
-while calling a library code. One may think that the solution for the uncaught exceptions may be to define a general handler, 
+Such undocumented exceptions make it difficult, and most of the times infeasible
+for the client code to protect against `unforeseen'' situations that may happen 
+while calling a library code.
+
+One may think that the solution for the uncaught exceptions may be to define a general handler, 
 which is responsible for handling any exception that is not
 adequately handled inside the applications. Although this 
-solution may prevent  the system of abruptly crashing,
- such general handler will not have enough
+solution may prevent  the system from abruptly crashing,
+ such a general handler will not have enough
 contextual information to adequately handle the exception, 
-a possible handler task is: to present a message to the user
- and restart the application. However, such handler cannot replace a carefully designed exception 
-handling policy~\cite{Robil00}, which can hardly be designed in the absence of 
+beyond storing a message in a log file and restarting the application.
+However, such a handler cannot replace a carefully designed exception 
+handling policy~\cite{Robil00}, which requires
 third-party documentation on the exceptions that
-may the thrown by it. On the other hand, documenting runtime exceptions is a tedious and error prone task, to help developers
-mitigate this problem tools should be developed to automate the documentation of runtime exceptions
-scraping from library code, few solutions in this directions have been proposed so far~\cite{van2005combining}.
+may the thrown by APIs used. 
+Since  documenting runtime exceptions is a tedious and error prone
+task, 
+this calls for tool support to automate the extraction of runtime exceptions
+from library code. Initial steps in thid direction have been proposed
+by van Doorn and Steegmans \cite{van2005combining}.
 

### 2015-02-16 android-stacks.tex

@@
 this calls for tool support to automate the extraction of runtime exceptions
-from library code. Initial steps in thid direction have been proposed
+from library code. Initial steps in this direction have been proposed
 by van Doorn and Steegmans \cite{van2005combining}.
@@
 exceptions or stack traces. Our limited manual inspection did not reveal such
-cases. Moreover, in this study we manually mapped the concerns related to exceptions.
+cases. Moreover, in this study we manually mapped the concerns related
+to exceptions.
+
  To ensure the quality of the analysis, we calculated the interrater agreement after three independent 
@@
 types from the total of 100); the interater agreement was high (96\%).
-Another threat relates to the fact that parts of our analysis 
-are based on the availability of stack traces on issues reported on Github and Googlecode projects. 
-In using these datasets, we make an underlying assumption: the stack traces reported on issues are 
-representative of valid crash information of the applications. 
-One way to mitigate such threat would be to access to the full 
-set of crash data per application. Although some services exist 
-to collect crash data from mobile applications~\cite{BugSe14,BugSn14,Googl14,Acra14},
-they do not provide open access to the crash reports of their client applications.
-In our study, we mitigated this threat by manually inspecting
-the source code associated to a subset of the reported exception stack traces.
-Such subset comprises the stack traces related to the main findings 
-of the study (e.g., ``undocumented runtime and checked exceptions",
-and cross-type wrappings").
-%\noindent
-
-\emph{External Validity.} Our work uses GHTorrent dataset; although 
+
+\emph{External Validity.} Our work uses the GHTorrent dataset, which although 
 comprehensive and extensive is not an exact replica of Github. 
-However, the result of this study does not depent on the analysis of
+However, the result of this study does not depend on the analysis of
 a complete Github dataset. Instead, the goal of our study was to 
@@
 exception stack trace mining of a subset of projects.
+
 We limited our analysis to a subset of existing open-source Android projects.
@@
 Such threats are similar to the ones imposed to other empirical studies 
-which also used free or open-source Android apps~\cite{Linar13,McDon13,Ruiz12}.
+which also use free or open-source Android apps~\cite{Linar13,McDon13,Ruiz12}.
 Moreover, several exception stack traces that support the findings of this study
@@
 
+Another threat relates to the fact that parts of our analysis 
+are based on the availability of stack traces on issues reported on Github and Googlecode projects. 
+In using these datasets, we make an underlying assumption: the stack traces reported on issues are 
+representative of valid crash information of the applications. 
+One way to mitigate this threat would be to access to the full 
+set of crash data per application. Although some services exist 
+to collect crash data from mobile applications~\cite{BugSe14,BugSn14,Googl14,Acra14},
+they do not provide open access to the crash reports of their client applications.
+In our study, we mitigated this threat by manually inspecting
+the source code associated to a subset of the reported exception stack traces.
+This subset comprises the stack traces related to the main findings 
+of the study (e.g., ``undocumented runtime and checked exceptions'',
+and ``cross-type wrappings'').
+%\noindent
+
 \section{Related Work}
@@
 
-\textit{Analysis and Use of Stack Trace Information.} Several research works have
+\textit{Analysis and Use of Stack Trace Information.} Several papers have
 investigated the use of stack trace information to support: bug classification
@@
 examined the stack traces embedded on crash reports sent by 1,800 Android apps 
-to a crash report management service (i.e., BugSense). They found that 19% of such stack traces
+to a crash report management service (i.e., BugSense). They found that 19\% of such stack traces
 were caused by unchecked and undocumented exceptions thrown by methods defined on 
@@
 
-\enlargethispage{-2\baselineskip}
+%\enlargethispage{-2\baselineskip}
 

### 2015-02-16 android-stacks.tex
Conclusions done.
Rewrite to make sure that we list the contributions and key
findings.

@@
 
-In this paper we present an exploratory study in which we mined the stack 
+The goal of this paper is to investigate 
+to what extent stack trace information can reveal \emph{bug hazards} 
+related to exception handling code that may lead to a decrease in
+application robustness. 
+To that end, we mined the stack 
 traces embedded in all issues defined in 482 Android projects hosted in Github and 
-157 projects hosted in Google Code. Overall it included 6,005 exception stack traces.
-In this study, the information extracted from stack traces was used in combination 
-with source code and bytecode analysis to identify common characteristics of exception 
-stack traces to pinpoint \emph{bug hazards} on the exception-related code. 
-In this study, \emph{bug hazards} are characteristics on the exception-related code that favor the introduction
-of  uncaught exceptions and unintended handling - which can contribute to 
-degrade the robustness of applications.
-Some \emph{bug hazards} consistently detected in this study were: 
-(i) unexpected exception wrappings (e.g., Errors being wrapped in checked exceptions)  - 
-revealing that Java hybrid exception model is not fully used according to its purpose;
-(ii) 267 stack traces were caused by undocumented runtime exceptions thrown by third party libraries or the Android Platform;
-and (iii) undocumented checked exceptions thrown by the Android Platform. 
-Moreover, we could also observe that approximately 52\% of the reported stack traces 
-can be attributed to programming mistakes; in particular, 27.71\% of all stack traces 
-reported a java.lang.NullPointerException as its root cause;
-The \emph{bug hazards} discussed in this work can negatively affect not only the robustness of Android application 
-but also the robustness of any Java-based application. 
-Our results calls the attention for the need of tools and language support to help 
-developers when dealing with them.
+157 projects hosted in Google Code. Overall it included 6,005
+exception stack traces.
+
+Our first key contribution is
+a novel approach and toolset (ExceptionMiner) for analyzing Java
+  exception stack traces as occurring on GitHub and Google Code
+  issues.
+Our second contribution is
+an empirical study of over 6000 actual stack traces,
+  demonstrating that (1) half of the system crashes are due to errors
+  in programming logic, with null pointer exceptions being most
+  prominent;
+  (2) Documentation for explicitly thrown RuntimeExceptions is almost
+  never provided; 
+  (3) Extensive use of wrapping leads to hard to understand chains
+  violating Java's exception handling principles.
+
+Our results shed light on common problems and bug hazards in Java
+exception handling code, and call for tool support to help developers
+understand their own and third party exception handling and wrapping logic.
 

### 2015-02-16 android-stacks.tex
correcting a typpo

@@
 on stack traces.  It walks up the type hierarchy of a Java exception until it reaches 
-a base exception type. Hence in this study the bytecode analysis was only used to discover the type 
+a base exception type. Hence in this study the bytecode analysis was used to discover the type 
 of each mined exception when the jar file of such exception was available on the project or 
-on the reused core Java libraries. An specific implementation (based on source code analysis)
- was needed to to discover the exception type when the bytecode was not available.
+on a reused  Java library. An specific implementation (based on source code analysis)
+ was needed to discover the exception type when the bytecode was not available.
 

### 2015-03-27 android-stacks.tex
addressing reviwers comments

@@
 %cope with an increasing number of exceptional conditions 
-%(e.g., faults in underlying middleware or hardware; memory and battery restrictions). 
+%(e.g., faults in underlying middleware or hardware; memory and batery restrictions). 
 %Therefore, mechanisms for exception detection and handling are not an optional add-on but a 
@@
 failures in third party services and libraries; compatibility issues~\cite{McDon13}; 
- memory and battery restrictions; and noisy external resources~\cite{Zhang12} 
+ memory and batery restrictions; and noisy external resources~\cite{Zhang12} 
 (e.g., wireless connections,GPS, bluetooth).
@@
 Moreover, such \emph{bug hazards} 
-call for improvements on languages and tools to better support exception handling in Android and Java environments.
+call for improvements on languages (e.g. to prevent null pointer dereferences) and tools to better support exception handling in Android and Java environments.
 
@@
 For simplicity, in this paper we will refer to ``exception stack trace'' as as just stack trace.
-The bottom part of the stack trace is the \emph{root exception}, which indicates
+The bottom part of the stack trace is the \emph{root exception} (Figure~\ref{fig:wrapping}-A), which indicates
 the first reason (root cause) for the exception thrown (in this case, the computer run out of
 memory). The top part of the stack trace indicates the location of the exception
-manifestation (which we will refer to as the \emph{exception wrapper} in this paper). The
+manifestation, which we will refer to as the \emph{exception wrapper} in this paper (Figure~\ref{fig:wrapping}-C). The
 execution flow  between the root exception and the wrapper may
-include other intermediate exception wrappers. In all levels, the exception
+include other intermediate exception wrappers (Figure~\ref{fig:wrapping}-D). In all levels, the exception
 \emph{signaler}, is the method that threw the exception, represented on the
-stack trace as the first method call below the exception declaration.
-
-\begin{figure} \centering \includegraphics[scale=0.5]{stacktrace_bw.png}
+stack trace as the first method call below the exception declaration (Figure~\ref{fig:wrapping}-B).
+
+\begin{figure} \centering \includegraphics[scale=0.55]{stack_review6.png}
 \caption{Example of an exception stack trace in Java.}
@@
 \begin{figure*} 
-\centering \includegraphics[width=.9\hsize]{studyoverview3.png}
+\centering \includegraphics[width=.9\hsize]{overview_review.png}
 \caption{Study overview.}\label{fig:overview} 
@@
 First,  the issues reported on Android projects hosted on GitHub (1) and Google Code (2) are recovered.
-Then the stack traces embedded each issue are extracted and distilled (3).
+Then the stack traces embedded in each issue are extracted and distilled (3).
 The stack trace information is then combined with source code and bytecode analysis in order to discover 
@@
 reporting at least one stack trace, to make sure that they represented real
-mobile apps. During this clean up 106 apps were removed because they were either
+mobile apps. During this clean up 107 apps were removed because they were either
 example projects (i.e., toy projects) or tools to support Android development
@@
 identify the type of some exceptions reported in issues 
-that were not be automatically identified by the ExceptionMiner tool
+that were not automatically identified by the ExceptionMiner tool
 (because they were defined on previous versions of libraries,
@@
 \finding{%
-More than 50\% of the uncaught exceptions are due to errors in
+More than 50\% of the uncaught exceptions analyzed in this study are due to errors in
 programming logic, with the NullPointerException as most prevalent

### 2015-03-27 android-stacks.tex
addressing reviwers comments

@@
 
-This section presents the results of the study, providing both a
-quantitative and qualitative analysis of the outcomes. 
-We center our presentation of the results around \emph{bug hazards} we
-distilled from (1) common root exceptions; (2) exception types; and
-(3) exception wrappings.
-
-\subsection{Common Root Exceptions}
+%This section presents the results of the study, providing both a
+%quantitative and qualitative analysis of the outcomes. 
+%We center our presentation of the results around \emph{bug hazards} we
+%distilled from (1) common root exceptions; (2) exception types; and
+%(3) exception wrappings.
+
+This section presents the results of the study, this presentation is centered around parts 
+of exception stack traces (i.e., the root exceptions, the exception types, 
+and the exception wrappings).  Each part is then analyzed in detail to 
+check whether they can reveal bug harzards on the exception handling code
+related or not to violations of the best practices presented
+ in Section~\ref{sec:best}.
+
+\subsection{Can the root exceptions reveal bug harzards?}
 
@@
 
-%Commented 2015
-%\emph{\textbf{The prevalence of programming mistakes on stack traces.}} 
-%One way of getting a broader view of the prevalence~\footnote The prevalence
- %evaluates the proportion of a given characteristic over a group 
-%of subjects.}  of programming mistakes on reported stack traces would be to manually 
-%inspect the code responsible for throwning every exception reported  on issues. 
-%However, inspecting the source code related to %6005 stack traces can %easily become infeasible.
-%Hence, we conducted an empirical investigation based on the following assumption: the name of the exception and its associated Javadoc 
-%documentation hold enough information to pinpoint a ``programming mistake".To exemplify, an instance of ArrayOutOfBoundException
-%pinpoints to programming mistake which indicates that an array has been accessed with an illegal index.
-
-%To perform this empirical investigation, we selected a subset of all reported root exceptions. 
-%This subset was composed by the top 100 root exceptions (from a list ranked by the number of ocurrences on stacks).
-%This subset was representative since it contained root exceptions that were reported as the root causes of 95\% of 
-%all stack traces. Therefore, based on the inspection of the full names exceptions and the associated 
-%Javadocs we could identify from this subset, which exceptions were related to programming mistakes.
-% It is worth mentioning that  all exceptions identified as programming mistakes were exceptions defined on java.lang and java.util packages. 
-%After doing such analysis, we could calculate the prevalence of such exceptions on issues.
-In other words, we could calculate the proportion of issues %mentioning the exceptions identified in this analysis as related to %``programming mistakes".
-
-%Table ~\ref{tab:causes} summarizes the results. It also illustrates the number of general Java exceptions - 
-%whose full name does not carry the reason for the exception to be thrown.
-%To ensure the quality of the analysis, three independent developers classified a randomly selected
-%sample of 25 exception types (from the total 100) using the same list of concerns;
-%the resulting interrater agreement was high (96\%).
-
-%The results show that the programming mistakes are associated to the most reported root causes. 
-%Although there is usually no proper handling for such exceptions~\cite{wirfs2006toward}, 
-%besides presenting an error message to the user and restarting the application~\footnote{Only 
-%high fault tolerant systems need to provide solutions to handle them.}.  
-%This finding calls the attention to the following: the high number of programming mistakes reported 
-%on issues are an indication that developers may be underestimating the impact that such 
-%mistakes have on application robustness.
-
-%\begin{table}
- % \centering
- % \begin{tabular}{lr}
- %   \hline
- %   \bfseries{Cause} &  \bfseries{\% Occurrences} \\
- %   \hline
- %     Programming Mistakes &  51.96\%\\ 
- %     %Memory Restriction   &   4.10\% \\ 
- %     General Exceptions   &  6.70\%\\
- %   \hline
-% \end{tabular}
-%  \caption{Identifying the concerns related to root exceptions}
-%  \label{tab:causes}
-% \end{table}
-
-%\noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
-%The study revealed: approximately 52\% of the reported stack traces were associated to programming mistakes;
-%and 27.71\% of all stack traces reported a java.lang.NullPointerException as the root cause.
-%}
-%\end{minipage}}
-
 \emph{\textbf{Identifying the Concerns Related to Root Exceptions.}} To get a broader view of the root exceptions of stack traces,
@@
 related to other concerns (e.g., specific to an application or a given library). 
+To ensure the quality of the process, three independent coders classified a randomly selected
+sample of 25 exception types (from the total 100) using the same list of concerns;
+the inter-rater agreement was 96\%.
 
@@
 
-To ensure the quality of the process, three independent coders classified a randomly selected
-sample of 25 exception types (from the total 100) using the same list of concerns;
-the inter-rater agreement was 96\%.
-
-%These study showed that the programming mistakes associated to the most reported root causes.
-%However, they may be interactly related to one of the other concerns. For instance, a concurrency problem may
-%lead to a NullPointerException.
-
-%For instance, the Java specification list all exceptions related to backward compatibility
-%issues ~\cite{javaback} (e.g., InstantiationError, VerifyError, IllegalAccessError). 
-%were also investigated in this study
+This analysis revealed that the exceptions implicitly thrown by runtime environment 
+due to mistakes on programming logic (e.g., out-of-bounds array index, null pointer 
+references) and resource limitations represent a major bug hazard to the exception 
+handling code of Android applications - security ..... os outros nao estao causando 
+muitos problemas.
+
+Although such exceptions do not violate the best practices described before (which are related to the 
+explicitly thrown exceptions), they represent a major threat to apps robustness.
 

### 2015-03-27 android-stacks.tex
last changes

@@
 on its repository. Table~\ref{tab:lables} illustrates the ocurrences of different labels 
-on the issues including exception stack traces.
-
-Regardless of the issue labels, every exception stack
+on the issues including exception stack traces. Regardless of the issue labels, every exception stack
 trace may contain relevant information concerning the exception structure of the
@@
 through the web interface of Google Code projects extracting all issues and issue comments and storing them in a relational database for later analysis.
-
 To identify Android projects in Google Code, we performed a similar heuristic: we performed a case insensitive search 
@@
 of each project. As a result we could identify 157 Android projects.  This set contained  127,456 issues in total,
- from which 1,963 exception stack traces were extracted. Table~\ref{tab:lables} illustrates the occurrences of different labels 
+ from which 1,963 exception stack traces were extracted. Table~\ref{txab:lables} illustrates the occurrences of different labels 
 on the issues including exception stack traces. Differently from GitHub, in Google Code most of 
@@
  was needed to discover the exception type when the bytecode was not available.
-
 With this module we analyzed all exceptions defined in the Android platform (Version 4.4, API level 19),
@@
 Table~\ref{tab:signalers} presents the heuristics adopted in this classification.
-
 To conduct this classification, we provide this module with the information
 comprising all Java packages that compose: the Android Plaform;
- the Android Libcore; and each analyzed Application.
-
-To discover the packages for the first two origins
+ the Android Libcore; and each analyzed Application. To discover the packages for the first two origins
 we can use to the Android specification.
@@
 NullPointerException makes the cause of the problem explicit and hence 
-can be signalled by an API expecting a non-null argument.
-
-The high prevalence of NullPointerException is aligned with the
+can be signalled by an API expecting a non-null argument.The high prevalence of NullPointerException is aligned with the
 findings of other 
 research~\cite{kim2013predicting,fraser20131600,csallner2004jcrasher,kechagia2014}. 
-For instance, Sunghun et
-al.~\cite{kim2013predicting} show that 38\% the bugs related to
-exception handling in the Eclipse project
- are caused by NullPointerException.
-Furhtermore, Kechagia and Spinellis showed that the NullPointerException was the  
+For instance, Kechagia and Spinellis showed that the NullPointerException was the  
 most reported exception on the crash reports sent to BugSense (a bug report 
@@
 
+%Sunghun et
+%al.~\cite{kim2013predicting} show that 38\% the bugs related to
+%exception handling in the Eclipse project
+%are caused by NullPointerException.
+%Furhtermore,
+
 \emph{\textbf{Identifying the Concerns Related to Root Exceptions.}} To get a broader view of the root exceptions of stack traces,
@@
 Besides the exceptions related to programming mistakes mentioned before, we also looked for exceptions related to some concerns that are known as sources of faults in mobile development: concurrency~\cite{ama2012} backward compatibility~\cite{McDon13}, security~\cite{enck2011study,was2010} and resource management (IO, Memory, Batery)~\cite{Zhang12}.
-
 Since it is infeasible to inspect the code responsible for throwning every exception reported in this study,
@@
 
-This analysis revealed that the exceptions implicitly thrown by runtime environment 
-due to mistakes on programming logic (e.g., out-of-bounds array index, null pointer 
-references) and resource limitations represent a major bug hazard to the exception 
-handling code of Android applications - security ..... os outros nao estao causando 
-muitos problemas.
-
-Although such exceptions do not violate the best practices described before (which are related to the 
-explicitly thrown exceptions), they represent a major threat to apps robustness.
+This analysis revealed that approximately 75\% of the exceptions
+ that caused the stack traces are implicitly thrown by runtime 
+environment  due to mistakes on programming logic (e.g., out-of-bounds array index, null pointer 
+references) and resource limitations. Although such exceptions do not directly point to violations on
+the best practices described before (which are related to the 
+explicitly thrown exceptions) they impose a major threat to apps robustness,
+and therefore represent a critical bug hazard to the exception 
+handling code of Android applications. 
+Security and concurrency, which are known to be critica issues on  Android apps, 
+raised few of the reported exceptions (less then 5\% of the anlysed stack traces).
+
+%\finding{%
+%More than 50\% of the uncaught exceptions analyzed in this study are due to errors in
+%programming logic, with the NullPointerException as most prevalent
+%exception.
+%For another 25\% the root cause relates to resource constraints.
+%}
 
 \finding{%
-More than 50\% of the uncaught exceptions analyzed in this study are due to errors in
-programming logic, with the NullPointerException as most prevalent
-exception.
-For another 25\% the root cause relates to resource constraints.
+Exceptions related to programming logic and resource limitations represent a major bug hazard to Android apps - 
+since they represent approximately 75\% of the exceptions that caused the stack traces.
+From this set the NullPointerException as most prevalent exception.
 }
 
-\subsection{Exception Types}
+\subsection{Can the exception types reveal bug hazards?}
 %\subsection{Exploring the Information Embedded on Exception Types}
@@
 or on the Android platform (34.3\%). We could also find runtime exceptions thrown by library code (17.7\%).
-
  We can also see, from Table~\ref{tab:typeroottab}, that in contrast with the other origins, most of the 
@@
 
-\finding{Almost two thirds of all crashes come from run time
-  exceptions. Most of these originate from the application layer.}
+\finding{Almost two thirds of all crashes come from runtime
+  exceptions. Most of these originate from the application layer.
+Hence, most of the bug hazards are located on this layer.}
 
@@
 explicitly signaling a runtime exception. 
-
 Table~\ref{tab:runtimeinterface} presents the results of this inspection. 
@@
 \finding{%
-Only a small fraction (4\%, 267 stack traces) of run time exceptions are programmatically
-thrown. Almost none (0.4\%, just one) of these were documented.}
+Only a small fraction (4\%, 267 stack traces) of runtime exceptions are programmatically
+thrown. Almost none (0.4\%, just one) of these were documented. Such undocumented 
+runtime exceptions violate best practices III and IV reveal a bug harzard to Java/Android apps.}
 
@@
 checked exception thrown by a native method and not declared on the exception 
-interface of these methods signaling them.
-
-The native method in question was defined in
+interface of these methods signaling them. The native method in question was defined in
  the Android platform, which uses Java Native Invocation (JNI) to access 
@@
 violating the declaration. The Java compiler could not detect this violation, because it does 
-not perform static exception checking on native methods.
-
-This type of bug is hard to diagnose
+not perform static exception checking on native methods. This type of bug is hard to diagnose
 because the developer usually does not have access to the native implementations. 
@@
 thrown without being documented on the exception interface.
+It violates best practices III and IV and represent a bug hazard
+ hard to diagnose.
 }
 
-\subsection{Exception Wrappings}
+\subsection{Can the exception wrappings reveal bug hazards?}
 
@@
 Such wrappings mask an unrecoverable Error into a general runtime exception.
+
+%{\footnotesize
+%\begin{verbatim}
+%   try {
+%      ...
+%   } catch (InterruptedException e) {
+%      android.util.Log.w(..., e);
+%   } catch (ExecutionException e) {
+%      throw new RuntimeException("...",e.getCause());
+%   } catch (CancellationException e) {
+%      ...
+%   } catch (Throwable t) {
+%     throw new RuntimeException("...", t);
+%   }
+%\end{verbatim}
+%}
 
@@
 
+\finding{%
+A common practice in Android plaform are general catch clauses masking any exception 
+into a general RuntimeException - it violates best practice III and is considered a bug 
+hazard as it loses contextual information about the exception.
+}
+
 \emph{\textbf{Checked Exception wrapping an Error.}} Most of these wrappings
@@
 code more complex (e.g., the multiple wrappings) and error-prone,
- and lead to ``exception confusion''.
-
-To illustrate this problem we can use one of the wrappings discussed above.
+ and lead to ``exception confusion''. To illustrate this problem we can use one of the wrappings discussed above.
 When the developer is confronted with a checked exception, the designer of the API is telling him/her 
@@
 Cross-type exception wrappings are common.
-They violate the semantics of Java's original exception
-design (e.g., when mapping unrecoverable Errors to other types of
-exceptions),
-and may lead to lengthy exception chains.
+They represent a bug hazard once they violate the semantics of Java's original exception
+design, detailed on best practices I and II  (e.g., when mapping unrecoverable Errors to other types of
+exceptions). 
 }
@@
 cases. Moreover, in this study we manually mapped the concerns related
-to exceptions.
-
- To ensure the quality of the analysis, we calculated the interrater agreement after three independent 
+to exceptions. To ensure the quality of the analysis, we calculated the
+ interrater agreement after three independent 
 developers classified a randomly selected sample (of 25 exception 
@@
 exception stack trace mining of a subset of projects.
-
 We limited our analysis to a subset of existing open-source Android projects.
@@
 In this section, we present work that is related to the present paper, divided into
-four categories: i) papers that use the information available on stack traces;
-ii) empirical studies on the usage of Java Exceptions and its fault proneness;
-and iii) tools to extract stack trace information from natural language artifacts
-(e.g., issues and emails) and iv) empirical studies involving Android apps.
+four categories as detailed next.
 
@@
 
+\textit{Extracting Stack Traces from natural language artifacts.} 
+Apart from issues and bug reports, stack traces can be embedded in other forms of
+communication between developers, such as discussion logs and emails.
+Few tools have been proposed to mine stack traces embedded on such resources.
+ Infozilla~\cite{bettenburg2008extracting} is based on a set of regular expressions that extract a set of frames
+related to a stack trace. The main limitation of this solution is that it is not
+able to extract stack traces embedded on verbose log files (i.e., on which we
+can find log text mixed with exception frames). Bacchelli
+et al.~\cite{bacchelli2012content} propose a solution to recognize stack trace frames
+from development emails and relate it to code artifacts (i.e. classes) mentioned
+on the stack trace. In addition to those tools, ExceptionMiner is able to 
+both extract stack traces from natural language artifacts and to 
+classify them in a set of predefined categories.
+
 \textit{Empirical Studies on Exception Handling Defects.} 
@@
 
-\textit{Extracting Stack Traces from natural language artifacts.} 
-Apart from issues and bug reports, stack traces can be embedded in other forms of
-communication between developers, such as discussion logs and emails.
-Few tools have been proposed to mine stack traces embedded on such resources.
- Infozilla~\cite{bettenburg2008extracting} is based on a set of regular expressions that extract a set of frames
-related to a stack trace. The main limitation of this solution is that it is not
-able to extract stack traces embedded on verbose log files (i.e., on which we
-can find log text mixed with exception frames). Bacchelli
-et al.~\cite{bacchelli2012content} propose a solution to recognize stack trace frames
-from development emails and relate it to code artifacts (i.e. classes) mentioned
-on the stack trace. In addition to those tools, ExceptionMiner is able to 
-both extract stack traces from natural language artifacts and to 
-classify them in a set of predefined categories.
-
 \textit{Empirical studies using Android apps.} Ruiz et al.~\cite{Ruiz12}
@@
 157 projects hosted in Google Code. Overall it included 6,005
-exception stack traces.
-
-Our first key contribution is
+exception stack traces. Our first key contribution is
 a novel approach and toolset (ExceptionMiner) for analyzing Java
@@
   violating Java's exception handling principles.
-
 Our results shed light on common problems and bug hazards in Java

### 2015-03-27 android-stacks.tex
results section

@@
 
-This section presents the results of the study, this presentation is centered around parts 
-of exception stack traces (i.e., the root exceptions, the exception types, 
-and the exception wrappings).  Each part is then analyzed in detail to 
+This section presents the results of the study, this presentation is centered around 
+characteristics distilled from stack traces: (i) the root exceptions, (ii) the exception types, 
+and (iii) the exception wrappings.  Each part is then analyzed in detail to 
 check whether they can reveal bug harzards on the exception handling code

### 2015-03-29 android-stacks.tex
minor changes on results

--- 2015-03-27 android-stacks.tex
@@
 \begin{document}
-%\title{ What GitHub and Google Code can tell us about the Causes of Crashes of Android Apps? }
-%\title{ A Mining Study the Google Code can tell us about the Causes of Crashes of Android Apps? }
-
-%\title{What can the Issues of GitHub and Google Code tell us about Exception Fragilities? }
-
-%\title{What can Issues on GitHub and Google Code tell us about Bug Hazards in Android and Java? }
 
 \title{Unveiling Exception Handling Bug Hazards in Android based on GitHub and Google Code Issues }
-
-%\title{Results of a Mining Study on Exception Stack Traces embedded on issues of over 600 Android Projects}
-
-%\title{Unveiling Robustness Threats by Mining the Exception Stack Traces of over 600 Android Projects}
-
-%\title{Discovering Threats to Robustness through Exception Stack Trace Analysis: Results of a Mining Study of Android Projects}
-
-%\title{Discovering Threats to Robustness through Exception Stack Trace Mining}
-
-%\title{Discovering Threats to Robustness through Exception Stack Trace Mining of over 600 Android Projects}
-
-%\title{Discovering Threats to Robustness through a Mining Study of Exception Stack Traces}
-
-%\title{Discovering Fragilities on the Exception-Related Code through  Exception Stack Trace Mining}
-
-%\title{On the Common Characteristics Android-Apps Crashes: Results of a GitHub and Google Code Mining Study}
-
-%\title{On the Common Characteristics of Android-Apps Crashes: Results of a GitHub and Google Code Mining Study}
-
-%\title{On the Common Characteristics of Android Crashes: Results of a GitHub and Google Code Mining Study}
-
-%\title{A first Glance on Characteristics of Android Crashes: Results of a GitHub and Google Code Mining Study}
-
-%\title{A first Glance on Characteristics of Android Crashes: Results of a GitHub and Google Code Mining Study}
 
@@
 thrown by both the Android platform and third party libraries; and (iii) undocumented checked exceptions
- thrown by the Android Platform. Such undocumented exceptions make difficult, and most of the times infeasible
+ thrown by the Android Platform. Such undocumented exceptions make it difficult, and most of the times infeasible
 for the client code to protect against ``unforeseen'' situations that may happen while calling third-party code.
@@
 
-%A bug hazard is a circumstance that increases the chance of a bug to be present in the software.
+%A bug hazard is a circumstance that increases the chance of a bug being present in the software.
 \end{abstract}
@@
 A \emph{bug hazard}~\cite{binder2000testing}  is a circumstance that increases the 
-chance of a bug to be present in the software. An example of a bug hazard can 
+chance of a bug being present in the software. An example of a bug hazard can 
 be a characteristic of  the exception-related code which can increase the likelihood 
@@
 Section~\ref{sec:back} provides the necessary
-background on the Android platform and Java exception model. 
+background on the Android platform and the Java exception model. 
 Section~\ref{sec:study} presents the study design. 
@@
 
-%\begin{figure} \centering \includegraphics[scale=0.8]{arch3.png}
-%  \caption{Android Architecture} \label{fig:exchier} \end{figure}
-
 \subsection{Exception Model in Java} \label{sec:extypes}
@@
 a set of best practices related to each of them.
-
-%\begin{figure} \centering \includegraphics[width=\hsize]{new2_hierarchy.png}
-% \caption{Exception Hierarchy in Java} \label{fig:exchier} \end{figure}
 
@@
 that favor the introduction of failures such as uncaught exceptions. 
-
-This exploration is guided by the set of best practices covered in Section~\ref{sec:best}.
 To support our investigation, we developed a tool called ExceptionMiner (Section~\ref{sec:exceptionminer})
@@
 \emph{\textbf{Exception Stack Trace Distiller.}}
-This component is based on a combination between a parser (based on regular expressions) 
+This component combines a parser (based on regular expressions) 
 and heuristics able to identify and filter exception names and stack traces inline with text. 
@@
 of each mined exception when the jar file of such exception was available on the project or 
-on a reused  Java library. An specific implementation (based on source code analysis)
+on a reused  Java library. A specific implementation (based on source code analysis)
  was needed to discover the exception type when the bytecode was not available.
@@
  the Android Libcore; and each analyzed Application. To discover the packages for the first two origins
-we can use to the Android specification.
+we can use the Android specification.
 To discover the packages for the third origin, the application itself, this module
@@
 (because they were defined on previous versions of libraries,
-apps and Android Plaform). When the exception could not be  
+apps and Android Platform). When the exception could not be  
 found automatically or manually (because they were defined on a previous version
@@
 
-This section presents the results of the study, this presentation is centered around parts 
-of exception stack traces (i.e., the root exceptions, the exception types, 
-and the exception wrappings).  Each part is then analyzed in detail to 
-check whether they can reveal bug harzards on the exception handling code
-related or not to violations of the best practices presented
- in Section~\ref{sec:best}.
+This section presents the results of the study; this presentation is centered around 
+the information distilled from stack traces: (i) the root exceptions (i.e., the exceptions that caused the stack traces);
+ (ii) the exception types (i.e, Checked, Runtime, Error, Throwable) and (iii) the exception wrappings. 
+Each piece information is then analyzed in detail to 
+check whether they can reveal bug harzards on the exception handling code -
+related to (i) specific violations of the best practices presented in Section~\ref{sec:best},
+or (ii) the general use of exception handling to suport robust development.
 
 \subsection{Can the root exceptions reveal bug harzards?}
-
 After distilling the information available on the exception stack traces, we could find 
@@
 \end{table*}
-
-%\begin{figure}
-%\centering
-%\includegraphics[width=\hsize]{top_exceptios_android_new2.pdf}
-%\includegraphics[width=\hsize]{top_exceptions_android_new}
-%\caption{Top exceptions in Android repositories.}
-%\label{fig:androidsignaler}
-%\end{figure}
 
@@
 \finding{Almost two thirds of all crashes come from runtime
-  exceptions. Most of these originate from the application layer.
-Hence, most of the bug hazards are located on this layer.}
+  exceptions. Most of these originate from the application layer.}
 
@@
 because the developer usually does not have access to the native implementations. 
-Consequently, since it is not expected by the programmer, when such method throws 
+Consequently, since it is not expected by the programmer, when such a method throws 
 this exception, such an undocumented exception may remain
@@
 However, using such a general exception, is considered a bad practice
-according to the Java specification 
-and common guidelines, as it loses contextual information about the exception.
+according to the Java specification and common guidelines, as it loses contextual information about the exception.
 
 \finding{%
-A common practice in Android plaform are general catch clauses masking any exception 
-into a general RuntimeException - it violates best practice III and is considered a bug 
-hazard as it loses contextual information about the exception.
+General catch clauses masking any exception into a general RuntimeException 
+is a common practice in the Android Platform; it violates best practice III 
+and is considered a bug hazard as it loses contextual information about the exception.
 }
@@
 examples of such wrappings. Although such a wrapping may represent a design decision,
-it violates the best practice related to checked exceptions and errors as it mixes the intended handling 
+it violates the best practices related to checked exceptions and errors as it mixes the intended handling 
 behaviour associated to both types.
@@
 \label{sec:disc}
-
-%\begin{figure*} \centering \includegraphics[scale=0.55]{examples_new2.png}
-%\caption{Code snippets illustrating fragilities on the exception-related code} \label{fig:snippets} \end{figure*}
 
@@
 and third-party libraries.  Additionaly,  the \emph{bug hazards} observed in this study are due to
-characteristics of Java exception model, which can impose challenges 
+characteristics of the Java exception model, which can impose challenges 
 the robustness of not only to Android apps but also to other systems
@@
 were caused by unchecked and undocumented exceptions thrown by methods defined on 
-Android API (level 15). Our work differs from Kechagia and Spinellis since it is not based on
+Android API (level 15). Our work differs from Kechagia and Spinellis since it is based on
 stack traces mined from issues reported by open source developers on GitHub and Googlecode.  
@@
 analysis of the exception hierarchy and analized the exception wrappings that can 
-happen during the exception propagation.  Such analysis revealed intrigging 
+happen during the exception propagation.  Such analysis revealed intriging 
 \emph{bug hazards} such as the cross-type exception wrappings not discussed in previous works.
@@
 number of uncaught exceptions, degrading the robustness of the AO version of every analyzed system.
-The main limitation of approaches based static analysis based approaches are the number of false
+The main limitation of approaches based on static analysis based approaches are the number of false
 positives they can generate, and the problems the faced when dealing with reflection libraries 
@@
 code used the Android API, and that the methods reusing fast-evolving APIs were
-more defect prone then others. Vásquez et al.~\cite{Linar13} analyzed
+more defect prone then others. Vasquez et al.~\cite{Linar13} analyzed
 approximately 7K free Android apps and observed that the last successful apps

### 2015-09-10 SPRINGER_JOURNAL/msr_journal.tex
extended paper version

--- 2015-03-29 android-stacks.tex
@@
-\documentclass[conference]{IEEEtran}
-
-\usepackage[pdftex]{graphicx}
-\usepackage[cmex10]{amsmath}
-\usepackage{algorithmic}
-\usepackage{caption}
-\usepackage{subfig}
-\usepackage{url}
-\usepackage{balance}
+%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% This is a general template file for the LaTeX package SVJour3
+% for Springer journals.          Springer Heidelberg 2010/09/16
+%
+% Copy it to a new file with a new name and use it as the basis
+% for your article. Delete % signs as needed.
+%
+% This template includes a few options for different layouts and
+% content for various journals. Please consult a previous issue of
+% your journal as needed.
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% First comes an example EPS file -- just ignore it and
+% proceed on the \documentclass line
+% your LaTeX will extract the file if required
+\begin{filecontents*}{example.eps}
+%!PS-Adobe-3.0 EPSF-3.0
+%%BoundingBox: 19 19 221 221
+%%CreationDate: Mon Sep 29 1997
+%%Creator: programmed by hand (JK)
+%%EndComments
+gsave
+newpath
+  20 20 moveto
+  20 220 lineto
+  220 220 lineto
+  220 20 lineto
+closepath
+2 setlinewidth
+gsave
+  .4 setgray fill
+grestore
+stroke
+grestore
+\end{filecontents*}
+%
+\RequirePackage{fix-cm}
+%
+%\documentclass{svjour3}                     % onecolumn (standard format)
+%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
+\documentclass[smallextended]{svjour3}       % onecolumn (second format)
+%\documentclass[twocolumn]{svjour3}          % twocolumn
+%
+\smartqed  % flush right qed marks, e.g. at end of proof
+%
+\usepackage{graphicx}
+%
+% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
+%
+% insert here the call for the packages your document requires
+\usepackage[numbers]{natbib}
+
+% etc.
+%
+% please place your own definitions here and don't use \def but
+% \newcommand{}{}
+%
+% Insert the name of "your journal" with
+% \journalname{myjournal}
+%
 
@@
 
-\title{Unveiling Exception Handling Bug Hazards in Android based on GitHub and Google Code Issues }
-
-\author{\IEEEauthorblockN{Roberta Coelho\IEEEauthorrefmark{1},
-Lucas Almeida\IEEEauthorrefmark{1},
-Georgios Gousios\IEEEauthorrefmark{2},
-Arie van Deursen\IEEEauthorrefmark{3}
+\title{Exception Handling Bug Hazards in Android\thanks{Grants or other notes
+about the article that should go on the front page should be
+placed here. General acknowledgments should be placed at the end of the article.}
 }
-\IEEEauthorblockA{\IEEEauthorrefmark{1}Federal University of Rio Grande do Norte\\
-Natal, Brazil\\
-Email: \texttt{roberta@dimap.ufrn.br,lucas.almeida@ppgsc.ufrn.br}}
-\IEEEauthorblockA{\IEEEauthorrefmark{2}Radboud University Nijmegen\\
-Nijmegen, The Netherlands\\
-Email: \texttt{georgios@cs.ru.nl}}
-\IEEEauthorblockA{\IEEEauthorrefmark{3}Delft University of Technology\\
-Delft, The Netherlands\\
-Email: \texttt{arie.vandeursen@tudelft.nl}}
+\subtitle{Results from a Mining Study and The Developers Perspective}
+
+%\titlerunning{Short form of title}        % if too long for running head
+
+\author{First Author         \and
+        Second Author %etc.
 }
 
-\newcommand{\todo}[1]{\textbf{TODO}\footnote{\textbf{TODO:} #1}}
-
-% make the title area
+%\authorrunning{Short form of author list} % if too long for running head
+
+\institute{F. Author \at
+              first address \\
+              Tel.: +123-45-678910\\
+              Fax: +123-45-678910\\
+              \email{fauthor@example.com}           %  \\
+%             \emph{Present address:} of F. Author  %  if needed
+           \and
+           S. Author \at
+              second address
+}
+
+\date{Received: date / Accepted: date}
+% The correct dates will be entered by the editor
+
 \maketitle
 
-%GitHub 31,592
-% Google Code: 127,456
-%All: 159,048
-% bf: 31,592 (with exception)
-
-%The goal of this study was to: identify common characteristics 
-%of such traces and investigate the fragilities on the exception-related code that can be revealed
-%from them. 
-%The goal of this study was to: identify common characteristics 
-%of such traces and investigate the whether they can reveal
-%the characteristics on the source code that may be favoring the introduction of faults 
-%(i.e. Uncaught Exceptions, and Unintened handler Actions), and consequently
-%decreasing the program robustness.
-%(of apps or the underlaying framework)
-
 \begin{abstract}
-%The number of mobile apps is increasing in a daily rate. If on the one hand they
-%extend phones capabilities far beyond of the basic calls, on the other hand they have to 
-%cope with an increasing number of exceptional conditions 
-%(e.g., faults in underlying middleware or hardware; memory and batery restrictions). 
-%Therefore, mechanisms for exception detection and handling are not an optional add-on but a 
-%fundamental part of such apps. However, studies have shown that such mechanisms are the less 
-%understood and tested parts of the system. As a consequence they may affect the system the 
-%other way around, making the system even less robust, leading to failures such as uncaught exceptions 
-%and unintended handling (i.e., exceptions being mistakenly caught by existing handlers).
-
-This paper reports on a study mining the exception stack traces included in
-159,048 issues reported on Android projects hosted in GitHub (482 projects) 
-and Google Code (157 projects). The goal of this study is to investigate 
-whether stack trace information can reveal \emph{bug hazards} related to exception handling code that may lead to a decrease in application robustness. 
- Overall 6,005 exception stack traces were extracted, and subjected to source code and bytecode analysis. 
-The outcomes of this study include the identification of the following 
-\emph{bug hazards}: (i) unexpected cross-type exception wrappings (for instance, trying to handle an instance 
-of OutOfMemoryError ``hidden'' in  a checked exception)  which can make the exception-related code 
-more complex and negatively impact the application robustness; (ii) undocumented runtime exceptions 
-thrown by both the Android platform and third party libraries; and (iii) undocumented checked exceptions
- thrown by the Android Platform. Such undocumented exceptions make it difficult, and most of the times infeasible
-for the client code to protect against ``unforeseen'' situations that may happen while calling third-party code.
-This study provides further insights on such \emph{bug hazards} and the robustness threats  
-they impose to Android apps as well as to other systems based on the Java exception model.
-
-%A bug hazard is a circumstance that increases the chance of a bug being present in the software.
+Insert your abstract here. Include keywords, PACS and mathematical
+subject classification numbers as needed.
+\keywords{First keyword \and Second keyword \and More}
+% \PACS{PACS code1 \and PACS code2 \and more}
+% \subclass{MSC code1 \and MSC code2 \and more}
 \end{abstract}
 
-%GitHub 31,592
-% Google Code: 127,456
-%All: 159,048
-
-%Total page numbers then:
-%intro:       1
-%background:  1.5
-%method:      1
-%tool         0.5
-%findings     4
-%discussion   1
-%threats      .5
-%conc         .5
-%------------------+
-%total        10 pages
-
-%\IEEEpeerreviewmaketitle
-
-%Such increase is motivated by pervasiveness of the World Wide Web 
-%and the unprecedented increase of global smartphone shipments -
-%the mobile industry shipped over 315 million smartphones in the last quarter of 2013 ~\cite{googleio}.
-
-%In this context, Android has become a leading platform - thanks to its open-source 
-%model and hardware partners like Samsung, HTC, Motorola, Asus (besides Nexus Google's
-% own device) ~\cite{gartner}.  Google estimates that there is over 1 billion Android active users per month ~\cite{googleio},
-%and the number of apps built on top of Android increases in a daily rate.
-
-%Our conjecture is that several Android applications may be suffering from common 
-%fragilities on the exception-related code (of the applications themselves or 
-%the underlaying platform) which may be degratind their robustness.
-%We call ``fragilities" characteristics on the exception-related code that favor the introduction
-%of aforementioned failures.
-
-%In Java, when an application fails due to an uncaught exception, 
-%it automatically terminates, while the system prints a stack trace to the console, 
-%or log file ~\cite{gosling2000java}.  A typical Java stack trace consists of  the fully qualified name 
-%the thrown exception and the ordered list of methods that were active on the call stack before 
-%such exception has occurred ~\cite{gosling2000java,bloch2008effective}.
-%and unintended handling~\cite{miller1997issues} (i.e., exceptions being mistakenly caught by existing handlers)
-
 \section{Introduction}
+\label{intro}
 
@@
 directions for future work.
-
 \section{Background}
@@
 
-\section{Study Design}
+%%%%%%%%%%%%%%%%%%%
+
+\section{The Mining Study Design}
 \label{sec:study}
@@
 % Due to space limitation we limit to present the general analysis here. More detailed analysis
-%on the defect issues can be found at:\url{https://github.com/souzacoelho/exceptionminer}}.
+%on the defect issues can be found at:https://github.com/souzacoelho/exceptionminer}.
 
@@
 Google Code contains widely used open-source Android apps (e.g.
-K9Mail\footnote{K9Mail moved to Github but as a way of not loosing the project history it advises their users to report bugs on the Google Code issue tracker: \url{https://github.com/k9mail/k-9/wiki/LoggingErrors}.}).
+K9Mail\footnote{K9Mail moved to Github but as a way of not loosing the project history it advises their users to report bugs on the Google Code issue tracker: https://github.com/k9mail/k-9/wiki/LoggingErrors.}).
 However, differently from GitHub, Google Code does not provide an API to access the information related
@@
 \subsection{Replication Package}
-All the data used in this study is publicly available at \url{https://github.com/souzacoelho/exceptionminer}.
+All the data used in this study is publicly available at tool website hosted in GitHub: https://github.com/souzacoelho/exceptionminer.
 Specifically we provide: (i) all issues related to Android projects found
@@
 
-\section{The Study Results}
+\section{The Mining Study Results}
 \label{sec:result}
@@
 
+\section{The Developers Perspective}
+\label{sec:dev}
+To better understand the developers perspective about the bug hazards detected during the mining 
+study, we set up an exploratory qualitative investigation and survey Android developers on how they use exception handling in their projects.
+Our field of study is GitHub; using our GHTorrent database [ref], we aimed our survey at developers from Android projects available in GitHub and mined during the first phase of this study.
+An explicit goal is to learn from developers perspective about the main exception handling bug hazards in Android.
+We therefore use a survey as our main research instrument.
+The survey mixed  Liket-scale with open-ended questions.
+The questions were motivated by the analysis of the existing bug hazards detected in the first phase of this study.
+Next sections detail the survey.
+
+%Our main findings reveal that developers often handle exceptions while developing Android apps and %we provide insights into the main challenges faced by them during this task. xxxxxxx
+
+\subsection{Research Questions}
+
+Although the mining study revelaed a set of bug hazards in Android exception handling no work has examined yet what are the developers perceives about them.
+With Android development rapidly rising in popularity, it is important to better understand of how developers deal with such bug hazards and what they do to overcome them.
+
+When developing Java-based applications it is inevitable to deal with exceptions.
+Consequently, our first question explores how often developers deal with the exception handling code in Android development.
+
+\noindent\emph{\textbf{RQ1:How often developers deal with the exception handling code in Android development?}} To make the analysis easier, we further refine RQ1 in the following subquestions:
+RQ1.1:\emph{How often developers handle exceptions?}
+RQ1.2:\emph{How often developers throw exceptions?}
+
+\noindent\emph{\textbf{RQ2:Do developers know about exception handling best practices?}}
+
+\noindent\emph{\textbf{RQ3:What are the developers perspectives about the exception handling bug hazards?}}
+
+\noindent\emph{RQ3.1:Have developers ever faced a crash caused by NullPointerException?}
+
+\noindent\emph{RQ3.2:Are developers aware of the robustness threats caused by cross-type wrappings?}
+
+\noindent\emph{RQ3.3:Are developers aware of the robustness threats caused by JNL non-documented checked exceptions?}
+
+\subsection{Protocol}
+
+Since our aim is to learn from a large number of developers, we use surveys which scale well.
+The survey is split into two logical sections; multiple choice or Likert-scale questions and open-ended questions.
+The multiple choice questions are intermixed with open-ended questions to further elicit the developer’s opinions.
+Moreover, the survey also contains Likert scales questions to force participants to make a choice.
+Overall, the survey includes XXXX open-ended questions, XXXXX Likert scale questions and XXXXX multiple choice questions.
+The respondents could fill in the survey in about 15 minutes.
+
+\subsection{Participants}
+
+ In previous work [REFMSR], investigated the issues reported on several Android projects hosted in GitHub and Googlecode.
+To ensure that our sample consists of repositories that were real Android apps, we had to inspect every project site and discard the toy-programs and non-Android repositories in our GHTorrent dataset [XXXXXX].
+Overall we could identify XXXX Android repositories.
+For each repository, we extract the developers whose email are registered with GitHub.
+For the pilot phase, we emailed XXXXX and received XXXX answers.
+
+The majority of our respondents have more than XXX years of Java development experience (XXXXX%) and  considerable experience in Android development (XXXXX years)  (XXXXXX%).
+
+\subsection{Data Analysis}
+
+complete!!!
+
+\subsection{Results}
+
+In this section, we present our findings per research question. 
+We also include direct quotes from developers along the text with the answer identification - which can be traced in our dtabase.
+
+\subsubsection{Q1:How often developers deal with the exception handling code in Android development}
+
+\subsubsection{Q1:How often developers deal with the exception handling code in Android development}
+
+\subsubsection{Q1:How often developers deal with the exception handling code in Android development}
+
 \section{Discussion}
@@
 %\noindent
+
+\section{Qualitative Study Design}
+\label{sec:qualitative}
+
+INCLUDE THE QUALITATIVE DESIGN....
 
@@
 
-%Nowadays many software vendors embed automatic crash reporting tools in their
-%software systems. Hence, whenever a software crashes this tool sends a detailed
-%crash report to its vendors. Moreover, we can also find third party software
-%solutions specialized in bug reporting for different kinds of systems specially
-%for the increasing marked of mobile apps~\cite{BugSe14,BugSn14,Googl14,Acra14}.
-%There is plenty of information to be mined...
-
-%\section*{Acknowledgment} This work is partially supported by: CNPq -- Proc.
-%484209/2013-2 and the NWO TestRoots project (639.022.314).
-
-\bibliographystyle{plain}
-\bibliography{android-stacks}
-
-% that's all folks
+\begin{acknowledgements}
+If you'd like to thank anyone, place your comments here
+and remove the percent signs.
+\end{acknowledgements}
+
+% BibTeX users please use one of
+\bibliographystyle{spbasic}      % basic style, author-year citations
+\bibliography{android-stacks}   % name your BibTeX data base
+
 \end{document}
-
-\textit{Empirical studies using Android apps.} Ruiz et al.~\cite{Ruiz12}
-investigated the degree of reuse across applications in Android Market, the
-study showed that almost 23\% of the classes inherited from a base class in the
-Android API, and that 217 mobile apps were reused completely by another mobile
-app. Pathak et al.~\cite{Patha11} analyzed bug reports and developers
-discussions of Android platform and found out that approximately 20\% of
-energy-related bugs in Android occurred after an OS update. McDonnell et
-al.~\cite{McDon13} conducted a case study of the co-evolution behavior of
-Android API and 10 dependent applications using the version history data found
-in github. The study found that approximately 25\% of all methods in the client
-code used the Android API, and that the methods reusing fast-evolving APIs were
-more defect prone then others. Vásquez et al.~\cite{Linar13} analyzed
-approximately 7K free Android apps and observed that the last successful apps
-used Android APIs that were on average 300\% more change-prone than the APIs
-used by the most successful apps.
+% end of file template.tex

### 2015-09-10 SPRINGER_JOURNAL/template.tex
extended paper version

--- 2015-09-10 SPRINGER_JOURNAL/msr_journal.tex
@@
 % insert here the call for the packages your document requires
-\usepackage[numbers]{natbib}
-
+%\usepackage{latexsym}
 % etc.
@@
 
-\title{Exception Handling Bug Hazards in Android\thanks{Grants or other notes
-about the article that should go on the front page should be
-placed here. General acknowledgments should be placed at the end of the article.}
+\title{Exception Handling Bug Hazards in Android%\thanks{Grants or other notes
+%about the article that should go on the front page should be
+%placed here. General acknowledgments should be placed at the end of the article.}
 }
-\subtitle{Results from a Mining Study and The Developers Perspective}
+\subtitle{Results from Mining and Qualitative Studies}
 
@@
 
-%%%%%%%%%%%%%%%%%%%
-
-\section{The Mining Study Design}
+\section{Study Design}
 \label{sec:study}
@@
 \subsection{Replication Package}
-All the data used in this study is publicly available at tool website hosted in GitHub: https://github.com/souzacoelho/exceptionminer.
+All the data used in this study is publicly available at https://github.com/souzacoelho/exceptionminer.
 Specifically we provide: (i) all issues related to Android projects found
@@
 
-\section{The Mining Study Results}
+\section{The Study Results}
 \label{sec:result}
@@
 
-\section{The Developers Perspective}
-\label{sec:dev}
-To better understand the developers perspective about the bug hazards detected during the mining 
-study, we set up an exploratory qualitative investigation and survey Android developers on how they use exception handling in their projects.
-Our field of study is GitHub; using our GHTorrent database [ref], we aimed our survey at developers from Android projects available in GitHub and mined during the first phase of this study.
-An explicit goal is to learn from developers perspective about the main exception handling bug hazards in Android.
-We therefore use a survey as our main research instrument.
-The survey mixed  Liket-scale with open-ended questions.
-The questions were motivated by the analysis of the existing bug hazards detected in the first phase of this study.
-Next sections detail the survey.
-
-%Our main findings reveal that developers often handle exceptions while developing Android apps and %we provide insights into the main challenges faced by them during this task. xxxxxxx
-
-\subsection{Research Questions}
-
-Although the mining study revelaed a set of bug hazards in Android exception handling no work has examined yet what are the developers perceives about them.
-With Android development rapidly rising in popularity, it is important to better understand of how developers deal with such bug hazards and what they do to overcome them.
-
-When developing Java-based applications it is inevitable to deal with exceptions.
-Consequently, our first question explores how often developers deal with the exception handling code in Android development.
-
-\noindent\emph{\textbf{RQ1:How often developers deal with the exception handling code in Android development?}} To make the analysis easier, we further refine RQ1 in the following subquestions:
-RQ1.1:\emph{How often developers handle exceptions?}
-RQ1.2:\emph{How often developers throw exceptions?}
-
-\noindent\emph{\textbf{RQ2:Do developers know about exception handling best practices?}}
-
-\noindent\emph{\textbf{RQ3:What are the developers perspectives about the exception handling bug hazards?}}
-
-\noindent\emph{RQ3.1:Have developers ever faced a crash caused by NullPointerException?}
-
-\noindent\emph{RQ3.2:Are developers aware of the robustness threats caused by cross-type wrappings?}
-
-\noindent\emph{RQ3.3:Are developers aware of the robustness threats caused by JNL non-documented checked exceptions?}
-
-\subsection{Protocol}
-
-Since our aim is to learn from a large number of developers, we use surveys which scale well.
-The survey is split into two logical sections; multiple choice or Likert-scale questions and open-ended questions.
-The multiple choice questions are intermixed with open-ended questions to further elicit the developer’s opinions.
-Moreover, the survey also contains Likert scales questions to force participants to make a choice.
-Overall, the survey includes XXXX open-ended questions, XXXXX Likert scale questions and XXXXX multiple choice questions.
-The respondents could fill in the survey in about 15 minutes.
-
-\subsection{Participants}
-
- In previous work [REFMSR], investigated the issues reported on several Android projects hosted in GitHub and Googlecode.
-To ensure that our sample consists of repositories that were real Android apps, we had to inspect every project site and discard the toy-programs and non-Android repositories in our GHTorrent dataset [XXXXXX].
-Overall we could identify XXXX Android repositories.
-For each repository, we extract the developers whose email are registered with GitHub.
-For the pilot phase, we emailed XXXXX and received XXXX answers.
-
-The majority of our respondents have more than XXX years of Java development experience (XXXXX%) and  considerable experience in Android development (XXXXX years)  (XXXXXX%).
-
-\subsection{Data Analysis}
-
-complete!!!
-
-\subsection{Results}
-
-In this section, we present our findings per research question. 
-We also include direct quotes from developers along the text with the answer identification - which can be traced in our dtabase.
-
-\subsubsection{Q1:How often developers deal with the exception handling code in Android development}
-
-\subsubsection{Q1:How often developers deal with the exception handling code in Android development}
-
-\subsubsection{Q1:How often developers deal with the exception handling code in Android development}
-
 \section{Discussion}
@@
 %\noindent
-
-\section{Qualitative Study Design}
-\label{sec:qualitative}
-
-INCLUDE THE QUALITATIVE DESIGN....
 

### 2015-09-24 SPRINGER_JOURNAL/msr_journal.tex
extended paper version

@@
 
-\title{Exception Handling Bug Hazards in Android\thanks{Grants or other notes
-about the article that should go on the front page should be
-placed here. General acknowledgments should be placed at the end of the article.}
-}
-\subtitle{Results from a Mining Study and The Developers Perspective}
+%\title{Exception Handling Bug Hazards in Android\thanks{Grants or other notes
+%about the article that should go on the front page should be
+%placed here. General acknowledgments should be placed at the end of the article.}
+%}
+%\subtitle{Results from a Mining Study and The Developers Perspective}
+
+\title{Exception Handling Bug Hazards in Android}
+\subtitle{Results from a Mining and Survey-based Study}
 
@@
 
-\author{First Author         \and
-        Second Author %etc.
+\author{Roberta Coelho         \and
+       Lucas Almeida \and
+       Georgios Gousios  \and
+       Arie van Deursen  \and
+      Christoph Treude
 }
@@
 \label{sec:dev}
-To better understand the developers perspective about the bug hazards detected during the mining 
-study, we set up an exploratory qualitative investigation and survey Android developers on how they use exception handling in their projects.
-Our field of study is GitHub; using our GHTorrent database [ref], we aimed our survey at developers from Android projects available in GitHub and mined during the first phase of this study.
-An explicit goal is to learn from developers perspective about the main exception handling bug hazards in Android.
+
+Although the mining study revelaed a set of bug hazards in Android exception handling no work had examined yet what are the developers perceives about them.
+With Android development rapidly rising in popularity, it is important to better understand how developers deal with such bug hazards and what they do to overcome them.
+
+To better understand the developers perspective about the bug hazards detected during the mining study, we set up an exploratory qualitative investigation and survey Android developers on how they face such bug hazards in their projects.
+Our field of study is GitHub; using our GHTorrent database~\cite{Gousi13}, we aimed our survey at developers from Android projects available in GitHub and which issues were mined during the first phase of this study.
+
+Our main goal is to learn from developers perspective about the main exception handling bug hazards in Android.
 We therefore use a survey as our main research instrument.
@@
 
-%Our main findings reveal that developers often handle exceptions while developing Android apps and %we provide insights into the main challenges faced by them during this task. xxxxxxx
+%Our main findings reveal that developers often handle exceptions while developing Android apps and 
+%we provide insights into the main challenges faced by them during this task. xxxxxxx
 
@@
 
-Although the mining study revelaed a set of bug hazards in Android exception handling no work has examined yet what are the developers perceives about them.
-With Android development rapidly rising in popularity, it is important to better understand of how developers deal with such bug hazards and what they do to overcome them.
-
 When developing Java-based applications it is inevitable to deal with exceptions.
-Consequently, our first question explores how often developers deal with the exception handling code in Android development.
-
-\noindent\emph{\textbf{RQ1:How often developers deal with the exception handling code in Android development?}} To make the analysis easier, we further refine RQ1 in the following subquestions:
-RQ1.1:\emph{How often developers handle exceptions?}
-RQ1.2:\emph{How often developers throw exceptions?}
-
-\noindent\emph{\textbf{RQ2:Do developers know about exception handling best practices?}}
-
-\noindent\emph{\textbf{RQ3:What are the developers perspectives about the exception handling bug hazards?}}
-
-\noindent\emph{RQ3.1:Have developers ever faced a crash caused by NullPointerException?}
-
-\noindent\emph{RQ3.2:Are developers aware of the robustness threats caused by cross-type wrappings?}
-
-\noindent\emph{RQ3.3:Are developers aware of the robustness threats caused by JNL non-documented checked exceptions?}
+Consequently, our first question explores how often developers deal with the exception handling code in Android development:
+
+\emph{\textbf{RQ 1:}How often developers deal with the exception handling code in Android development?} To make the analysis easier, we further refine RQ1 in the following subquestions:
+\begin{itemize}
+\item RQ1.1:\emph{How often developers handle exceptions?}
+\item RQ1.2:\emph{How often developers throw exceptions?}
+\end{itemize}
+
+Another theme that show us how developers may deal with such bug hazards is
+ their knowledge about exception handling best practices presented in Section~\ref{sec:best}.
+ This leads us to our second research question:
+
+\bigskip
+
+\emph{\textbf{RQ 2:}Do developers know about exception handling best practices?}
+
+\bigskip
+
+Then the third research question focus on how developers deal with the main bug hazards found during the mining study.
+
+\bigskip
+
+\emph{\textbf{RQ 3:}What are the developers perspectives about the exception handling bug hazards?} To answer such question we further refine RQ3 in the following subquestions:
+\begin{itemize}
+\item\emph{RQ3.1:Have developers ever faced a crash caused by NullPointerException?}
+\item\emph{RQ3.2:Are developers aware of the robustness threats caused by cross-type wrappings?}
+\item\emph{RQ3.3:Are developers aware of the robustness threats caused by JNL non-documented checked exceptions?}
+\end{itemize}
+
+Some of such questions were indirectly answered by presenting a set of code snippets on which
+the bug hazards were present and asked developers what they would do when faced with such scenarios.
 
@@
 Moreover, the survey also contains Likert scales questions to force participants to make a choice.
-Overall, the survey includes XXXX open-ended questions, XXXXX Likert scale questions and XXXXX multiple choice questions.
+Overall, the survey includes 13 open-ended questions, 5 Likert scale questions and 10 multiple choice questions.
 The respondents could fill in the survey in about 15 minutes.
@@
 
- In previous work [REFMSR], investigated the issues reported on several Android projects hosted in GitHub and Googlecode.
-To ensure that our sample consists of repositories that were real Android apps, we had to inspect every project site and discard the toy-programs and non-Android repositories in our GHTorrent dataset [XXXXXX].
-Overall we could identify XXXX Android repositories.
-For each repository, we extract the developers whose email are registered with GitHub.
-For the pilot phase, we emailed XXXXX and received XXXX answers.
+ In previous work ~\cite{coelho2015}, we mined exception stack traces available on issues reported on several Android projects hosted in GitHub and Googlecode.
+To ensure that our sample consists of repositories that were real Android apps, we had to inspect every project site and discard the toy-programs and non-Android repositories.
+For each repository, we extract the developers whose email are registered on each repository.
+We emailed the developers and received 74 answers.
 
@@
 \subsection{Data Analysis}
-
-complete!!!
 

### 2015-09-24 SPRINGER_JOURNAL/msr_journal.tex
extended paper version

@@
 \begin{table}
+  \scriptsize
   \centering
@@
   \centering
+  \scriptsize
   \begin{tabular}{rp{29em}}
@@
 \begin{table*}
+  \tiny
   \centering
@@
 \begin{table}
+\scriptsize
 \centering
@@
 %}
+
+\bigskip
 
@@
 
+\bigskip
+
 \finding{Almost two thirds of all crashes come from runtime
   exceptions. Most of these originate from the application layer.}
+
+\bigskip
 
@@
 
+\bigskip
+
 \finding{%
@@
 
+\bigskip
+
 \begin{table}
+\scriptsize
 \centering
@@
 
+\bigskip
+
 \subsection{Can the exception wrappings reveal bug hazards?}
@@
 \begin{table*}
+\scriptsize
   \centering
@@
     \hline
-    \bfseries{Wrapper Type}  &  \bfseries{Root Cause Type} &  \bfseries{Projects}  &  \bfseries{Occurrences} & \textsf{Android} & \textsf{Java/Libcore} & \textsf{Lib} & \textsf{App}  \\
+    \bfseries{Wrapper}  &  \bfseries{Root Cause} &  \bfseries{Projects}  &  \bfseries{Occurrences} & \textsf{Android} & \textsf{Java/Libcore} & \textsf{Lib} & \textsf{App}  \\
     \hline
@@
 }
+
+\bigskip
 
@@
 
+\bigskip
+
 \section{The Developers Perspective}
@@
 
- In previous work ~\cite{coelho2015}, we mined exception stack traces available on issues reported on several Android projects hosted in GitHub and Googlecode.
+ In previous work~\cite{co15}, we mined exception stack traces available on issues reported on several Android projects hosted in GitHub and Googlecode.
 To ensure that our sample consists of repositories that were real Android apps, we had to inspect every project site and discard the toy-programs and non-Android repositories.
@@
 
-The majority of our respondents have more than XXX years of Java development experience (XXXXX%) and  considerable experience in Android development (XXXXX years)  (XXXXXX%).
-
-\subsection{Data Analysis}
-
-\subsection{Results}
+The majority of our respondents have more than 2 years of Java development experience (87.3\%) and  considerable experience in Android development (more than 1 year)  (85.9\%) - see Figure~\ref{fig:devexpertise}.
+
+\begin{figure} \centering \includegraphics[scale=0.35]{charts/Q1Q2.png}
+\caption{Developers expertise in Java and Android development.}
+\label{fig:devexpertise}
+\end{figure}
+
+\subsection{Findings}
 
@@
 
-\subsubsection{Q1:How often developers deal with the exception handling code in Android development}
-
-\subsubsection{Q1:How often developers deal with the exception handling code in Android development}
-
-\subsubsection{Q1:How often developers deal with the exception handling code in Android development}
+\bigskip
+
+\noindent\emph{\textbf{Q1:How often developers deal with the exception handling code in Android development?}}
+
+Many of our survey respondents said they have to handle exceptions most of the time during Android app development (64.8\%).
+However, the frequency on which they throw exception inside apps is smaller; most of the survey respondents said they throw exceptions only some of the time or seldom (80.3\%).
+It shows that although we can try to avoid dealing with exceptions throwing few exceptions, it is almost always inevitable to handle the exceptions thrown by Android platform and reused libraries.
+
+\begin{figure} \centering \includegraphics[scale=0.35]{charts/Q3Q4.png}
+\caption{How often.}
+\label{fig:oftenhandle}
+\end{figure}
+
+\bigskip
+
+\noindent\emph{\textbf{Q2: Do developers know about exception handling best practices?}}
+
+We can observe that 62\% of surveyed Android developers follow Java exception handling best practices.
+Among the best practices cited by developers the most popular ones were: avoid generic catches and avoid empty catch blocks.
+They did not mention a any guide or reference of best practices.
+When asked about best practices specific to Android developers, most of the developers mentioned that they did no about them.
+The ones who answered positively said that they adopt the same best practices as in any Java program.
+
+\begin{figure} \centering \includegraphics[scale=0.35]{charts/Q7Q8.png}
+\caption{How often.}
+\label{fig:oftenhandle}
+\end{figure}
+
+\bigskip
+
+\noindent\emph{\textbf{Q3:What are the developers perspectives about the exception handling bug hazards?}}
 

### 2015-09-27 SPRINGER_JOURNAL/msr_journal.tex
extended paper version

@@
 
+%%%%GREY BOX
+
+\usepackage{color}
+\definecolor{lightgray}{gray}{0.85}
+
+\newcommand\greybox[1]{%
+  \vskip\baselineskip%
+  \par\noindent\colorbox{lightgray}{%
+    \begin{minipage}{.97\columnwidth}%
+    \emph{#1}%
+    \end{minipage}%
+  }%
+  \vskip\baselineskip%
+}
+%%%% END GREY BOX
+
 \begin{document}
@@
 
-\emph{\textbf{RQ 1:}How often developers deal with the exception handling code in Android development?} To make the analysis easier, we further refine RQ1 in the following subquestions:
+\emph{\textbf{RQ 1:} How often developers deal with the exception handling in Android development?} To make the analysis easier, we further refine RQ1 in the following subquestions:
 \begin{itemize}
@@
 
-Another theme that show us how developers may deal with such bug hazards is
+Another issue that show us how developers may deal with such bug hazards is
  their knowledge about exception handling best practices presented in Section~\ref{sec:best}.
@@
 
-\emph{\textbf{RQ 2:}Do developers know about exception handling best practices?}
+\emph{\textbf{RQ 2:} Do Android developers adopt exception handling best practices?}
+
+\bigskip
+
+\emph{\textbf{RQ 4:}  Do Android developers consider implementing the exception handling code of Android apps more difficult than general Java applications?}
 
@@
 
-\begin{figure} \centering \includegraphics[scale=0.35]{charts/Q1Q2.png}
+\begin{figure} \centering \includegraphics[scale=0.35]{charts/Q1Q2_new.png}
 \caption{Developers expertise in Java and Android development.}
@@
 In this section, we present our findings per research question. 
-We also include direct quotes from developers along the text with the answer identification - which can be traced in our dtabase.
-
-\bigskip
-
-\noindent\emph{\textbf{Q1:How often developers deal with the exception handling code in Android development?}}
+We also include a selection of quotes from the exploratory survey 
+respondents. Each respondent has an identification which can be traced in our database.
+
+\bigskip
+
+\noindent\emph{\textbf{Q1:How often developers deal with the exception handling in Android development?}}
 
@@
 
-\begin{figure} \centering \includegraphics[scale=0.35]{charts/Q3Q4.png}
-\caption{How often.}
+\begin{figure} \centering \includegraphics[scale=0.35]{charts/often.png}
+\caption{How often developers throw (a) and handle (b) exceptions during app development.}
 \label{fig:oftenhandle}
 \end{figure}
+
+Developers were also asked whether the exception handling code helped with the development of robust Android applications.
+Although most of the developers answered positively (78.9\%) when asked to provide one or more reasons for their answers, most of them said that if not carefully developed the exception handling code may lead to app crashes.
+Table XXX presents the top reasons why the exception handling code may or may not help de development of robust apps.
+
+\begin{table}
+\scriptsize
+\centering
+\begin{tabular}{lrr}
+\hline
+\bfseries{Top 5 Resons} & \bfseries{\#} & \bfseries{\%} \\
+\hline
+\bfseries{For Positively Affecting Robustness} &   &   \\
+Programming logic (java.lang and util)  & 52,0\% & 52,0\%\\
+Resources (IO, Memory, Batery)  & 52,0\% & 23,9\% \\
+Security  & 52,0\% & 4,1\%\\
+Concurrency  & 52,0\% & 2,9\% \\
+Backward compatibility  & 52,0\% & 5,5\% \\
+\bfseries{For Negatively Affecting Robustness} &   &   \\
+Specific Exceptions  & 52,0\% & 4,9\%\\
+General (Error, Exception, Runtime)  & 52,0\% & 6,7\%\\
+Specific Exceptions  & 52,0\% & 4,9\%\\
+General (Error, Exception, Runtime)  & 52,0\% & 6,7\%\\
+\hline
+\end{tabular}
+\caption{Top 5 reasons why exception handling helps or impairs the development of robust apps}
+\label{tab:tophundrend}
+\end{table}
 
@@
 
-\begin{figure} \centering \includegraphics[scale=0.35]{charts/Q7Q8.png}
-\caption{How often.}
+\begin{figure} \centering \includegraphics[scale=0.35]{charts/bestpractices.png}
+\caption{The developers awareness of exception handling best practices in Java (a) and Android (b). }
 \label{fig:oftenhandle}
@@
 \noindent\emph{\textbf{Q3:What are the developers perspectives about the exception handling bug hazards?}}
+
+\greybox{Sometimes when you develop multi-threaded applications, when the thread finishes you don't have any GUI (or Context) to show information, so you have to be aware of your application's state (foreground, background)}
+
+Handle life cycle of activities and fragments.
+
+Q10: What makes exception handling in Android different from exception handling in Java?
+
+The Android framework is what adds the complexity in figuring out what caused an exception because more often than not, the error is triggered from the framework as a result of something else you did
+
+Sometimes when you develop multi-threaded applications, when the thread finishes you don't have any GUI (or Context) to show information, so you have to be aware of your application's state (foreground, background).
+
+Application-Activity-Fragment lifecycle.
+
+You can wrap the main function in java in try/catch in android you can\'t but there are equivalents
+
+Though not particular to Android, most Android apps have several different entry points: Activities, Services, etc, can be started by different services; same thing with events.
+It may be hard to ensure that all this cases are properly handled.
+The logcat is a helpful way of tracking exceptions when developing and testing on real phones.
+
+Any uncatched exception will crash the app.
+There are many unknown exception thrown by the system, that are only happen once in a lifetime like IllegalStateException: eglMakeCurrent failed EGL\_BAD\_CONTEXT
+
+The lifecycle of activities\/fragements etc make it harder to work out what order things are called in and so exceptions can occur because you didn't realise that another method isn't called
 

### 2015-09-28 SPRINGER_JOURNAL/msr_journal.tex
RQ1 e RQ2

@@
 
-Our main goal is to learn from developers perspective about the main exception handling bug hazards in Android.
+Our main goal is to learn from developers perspective about the Java exception handling usage and the main exception handling bug hazards in Android.
 We therefore use a survey as our main research instrument.
@@
 
+We used grounded theory coding~\cite{charmaz2006}, iterating through our open-ended survey responses, to answer some of the research questions.
+The grounded theory coding used in this study consists of at two phases: (1) initial coding entails a close reading of the data and (2) later we use focused coding ~\cite{charmaz2006} to pinpoint and develop the most salient themes~\cite{charmaz2006} in the analyzed data.
+
 \subsection{Participants}
 
- In previous work~\cite{co15}, we mined exception stack traces available on issues reported on several Android projects hosted in GitHub and Googlecode.
+ In previous work~\cite{coelhoetal15}, we mined exception stack traces available on issues reported on several Android projects hosted in GitHub and Googlecode.
 To ensure that our sample consists of repositories that were real Android apps, we had to inspect every project site and discard the toy-programs and non-Android repositories.
@@
 Developers were also asked whether the exception handling code helped with the development of robust Android applications.
-Although most of the developers answered positively (78.9\%) when asked to provide one or more reasons for their answers, most of them said that if not carefully developed the exception handling code may lead to app crashes.
-Table XXX presents the top reasons why the exception handling code may or may not help de development of robust apps.
+Although most of the developers answered positively (78.9\%) when asked to provide one or more reasons for their answers, most of the respondents also provide reasons why the exception handling code can sometimes impairs the robustness.
+Since positive and negative reasons were provided intermixed we analyzed both set of answers and ranked the top reasons why exception handling that can negatively or positively affect robustness according to developers perceptions.
+
+ Table ~\ref{tab:topreasons} presents the top reasons why the exception handling code may or may not help de development of robust apps.
+Although most of the respondents answered that exception handling code helped with the development of robust Android applications, almost every respondent pointed a drawback associated to he exception handling code, saying that if not used with care the exception handling code may lead to app crashes.
 
@@
 \hline
-\bfseries{Top 5 Resons} & \bfseries{\#} & \bfseries{\%} \\
+\bfseries{Top Resons} & \bfseries{\#} & \bfseries{\%} \\
 \hline
 \bfseries{For Positively Affecting Robustness} &   &   \\
-Programming logic (java.lang and util)  & 52,0\% & 52,0\%\\
-Resources (IO, Memory, Batery)  & 52,0\% & 23,9\% \\
-Security  & 52,0\% & 4,1\%\\
-Concurrency  & 52,0\% & 2,9\% \\
-Backward compatibility  & 52,0\% & 5,5\% \\
+Exception message improves error diagnosis &	10	& 14\% \\
+Anticipated erroneous situation help writing more robust code & 9	 & 13\% \\
+Exceptions for unrecoverable behavior	& 7 &	10\% \\
+Checked exceptions force handling  &	6	& 9\% \\
+Useful to gracefully crash 	& 4 &	6\% \\
+
 \bfseries{For Negatively Affecting Robustness} &   &   \\
-Specific Exceptions  & 52,0\% & 4,9\%\\
-General (Error, Exception, Runtime)  & 52,0\% & 6,7\%\\
-Specific Exceptions  & 52,0\% & 4,9\%\\
-General (Error, Exception, Runtime)  & 52,0\% & 6,7\%\\
+Exceptions should be caught to improve robustness	& 7	& 10\% \\
+Creates a different control flow makes debugging harder &	2 &	3\% \\
+unchecked/runtime exceptions favor crashes	& 2	& 3\% \\
+NullPointerExceptoins are tricky to avoid	 & 2	& 3\% \\
+Many device specific exceptions prevent proper handling	& 1 & 1\% \\
+
 \hline
@@
 \caption{Top 5 reasons why exception handling helps or impairs the development of robust apps}
-\label{tab:tophundrend}
+\label{tab:topreasons}
 \end{table}

### 2015-09-29 SPRINGER_JOURNAL/msr_journal.tex
table prevent

@@
 For each repository, we extract the developers whose email are registered on each repository.
-We emailed the developers and received 74 answers.
+We emailed the developers and received 68 valid answers.
 
@@
 \bfseries{For Positively Affecting Robustness} &   &   \\
-Exception message improves error diagnosis &	10	& 14\% \\
+Exception message improves error diagnosis &	10	& 15\% \\
 Anticipated erroneous situation help writing more robust code & 9	 & 13\% \\
@@
 \bigskip
+
+\begin{table}
+\scriptsize
+\centering
+\begin{tabular}{lrr}
+\hline
+\bfseries{Top Ways} & \bfseries{\#} & \bfseries{\%} \\
+\hline
+Handling exceptions & 18 & 25\% \\
+       - handling exceptions in all entry points/all over	  &   &   \\	
+       - using catch-all clauses 		  &   &   \\	
+       - addinf try catch bocks around risky methods		  &   &   \\	
+Crash report tools	& 6 &	8\% \\
+Null checks &	11	& 15\% \\
+ - null check annotations (@Null and @NonNull)	& 3 & 	4\% \\
+Edge cases/boundaries checks  &	4 &	6\% \\
+Testing &	15 &	21\% \\
+
+Let it crash and fix afterward	& 4 & 6\% \\
+
+Many device specific exceptions prevent proper handling	& 1 & 1\% \\
+
+\hline
+\end{tabular}
+\caption{Top ways of preventing crashes }
+\label{tab:topreasons}
+\end{table}
 

### 2015-10-06 SPRINGER_JOURNAL/msr_journal.tex
including tables with tags and overall comments about them

--- 2015-09-28 SPRINGER_JOURNAL/msr_journal.tex
@@
 \bfseries{For Positively Affecting Robustness} &   &   \\
-Exception message improves error diagnosis &	10	& 14\% \\
+Improves error diagnosis  &	10	& 15\% \\
 Anticipated erroneous situation help writing more robust code  & 9	 & 13\% \\ 
-Exceptions for unrecoverable behavior	& 7 &	10\% \\
+%Anticipated erroneous situation help writing more robust code 
+Exceptions only for unrecoverable behavior	& 7 &	10\% \\
 Checked exceptions force handling  &	6	& 9\% \\
@@
 \bfseries{For Negatively Affecting Robustness} &   &   \\
-Exceptions should be caught to improve robustness	& 7	& 10\% \\
-Creates a different control flow makes debugging harder &	2 &	3\% \\
-unchecked/runtime exceptions favor crashes	& 2	& 3\% \\
-NullPointerExceptoins are tricky to avoid	 & 2	& 3\% \\
-Many device specific exceptions prevent proper handling	& 1 & 1\% \\
+Every exception should be caught to improve robustness	& 7	& 10\% \\
+Makes debugging harder &	2 &	3\% \\
+Unchecked/runtime exceptions can crash apps & 2	& 3\% \\
+NullPointerExceptoins can happen anywhere/are tricky to avoid	 & 2	& 3\% \\
+VM inefficiency 	& 1 & 1\% \\
 
@@
 \bigskip
+
+///// PREVENT CRASHES /////
+Finally developers were asked how they prevent their apps from crashing.
+Most of the respondents answered that they prevent apps from crashing by handling exceptions/errors (27\%).
+They mentioned about some strategies such as: handling exceptions in all entry points/all over; using catch-all clauses; and adding try catch blocks around risky methods.
+Different from conventional Java programs, an Android app is composed by several entry points, each window (i.e.
+Activity) is a potential entry point on which exceptions can scape leading to an app crash.
+Hence the importance of handling exceptions in all entry points, and using general cash clauses which cannot sometimes prevent crash but allows the developer to present a detailed error message to the user before crashing.
+Many respondents also said that they used testing approaches to prevent crashes -  few of them mentioned crowd testing (3\%) an emerging approach in testing apps.
+Moreover, null checks were also mentioned by 15\% of the developers as a way to prevent crashes.
+The adoption of coding styles were also pointed out by 14\% of respondents as a way to prevent craches.
+Finally, the crash report tools (e.g.
+ACRA [ref]) were reported as a way to prevent crashes.
+Once such tools notify developers about current crashes, the application can be fixed and hence future crashes (caused by what was fixed) could be prevented.
+Table ~ref{tab:topprevent} presents the themes for the strategies for crash prevention mentioned by developers.	////
+
+\begin{table}
+\scriptsize
+\centering
+\begin{tabular}{lrr}
+\hline
+\bfseries{Top Ways} & \bfseries{\#} & \bfseries{\%} \\
+\hline
+exception/error handling & 19 &	27\% \\
+testing &	14 &	20\% \\
+null check(s) &	11 &	15\% \\
+- null check annotations (@Null and @NonNull)	& 3 & 	4\% \\
+coding style(s)	& 10 &	14\% \\
+crash report / crash report tools &	8	& 11\% \\
+\hline
+\end{tabular}
+\caption{Top ways of preventing crashes - 71 non-empty respondents. }
+\label{tab:topprevent}
+\end{table}
+
+/////// JAVA BEST PRACTICES ////// 
+As mentioned before several general guidelines have been proposed on how to
+use Java exceptions.
+Section XXXXX compiled a list some exception handling best practices that guided our mining study.
+In this survey also we asked the Android developers whether they knew about any Java exception handling best practice.
+Although most of them said that knew and adopted EH practices 68\%, from this set the most mentioned best practices are presented on Table XXXX - considering that one developer mentioned more than one practice.
+Some of the practices mentioned, however, are not well known Java EH best practices but seam to be used by some Android developers to better deal with exceptions in Android app context: the crash fast approach [ref], the use of crash report tools, and to favor checked exceptions instead of runtime to represent the exceptional conditions in Android apps./////
+
+\begin{table}
+\scriptsize
+\centering
+\begin{tabular}{lrr}
+\hline
+\bfseries{Top Ways} & \bfseries{\#} & \bfseries{\%} \\
+\hline
+Use specific handlers / dont catch generic exceptions	& 9 & 23\% \\
+Don't swallow Exceptions	 & 7 & 18\% \\
+Dont throw Runtime / Favor Checked exceptions & 4 &	10\% \\
+Do not use exception for normal  flow control. &	4 &	10\% \\
+Free Resources in finally-blocks	& 4	& 10\% \\
+crash fast	 & 3 &	8\% \\
+crash report tools	&  2 &	5\% \\
+Don't catch Errors 	&  2 &	5\% \\
+\hline
+\end{tabular}
+\caption{Top mentioned Java EH best practices -  40 non-empty answers. }
+\label{tab:javapractices}
+\end{table}
+
+/////// ANDROID EH BEST PRACTICES /////
+Developers who informed that adopted EH best practices were also asked whether they knew about any EH best practice specific to Android context.
+Most of them 43\% mentioned that they applied the same best practices that they used in Java programs.
+Only some of them mentioned that adopted specific best practices such as: (i) the use of crash report tools (21\%)  - to notify the developer about the uncaught exceptions that hapen in the application code; and (ii) the use of a global exception handler (14\%) - the UncaughtExceptionHandler advised on Android tutorials [ref].	 /////
+
+\begin{table}
+\scriptsize
+\centering
+\begin{tabular}{lrr}
+\hline
+\bfseries{Top Ways} & \bfseries{\#} & \bfseries{\%} \\
+\hline
+same as java	& 6 &	43\% \\
+crash report tools	 & 3 &	21\% \\
+add a global exception handler (UncaughtExceptionHandler)	 & 2 & 14\% \\
+use checked exceptions	& 1 &	7\% \\
+use appropriate exception messages	& 1 &	7\% \\
+\hline
+\end{tabular}
+\caption{Top mentioned Android-specific EH best practices -  14 non-empty responses. }
+\label{tab:androidpractices}
+\end{table}
+
+///// HANDLING RUNTIME AND CHECKED /////
+Developers were presented to two pieces of code: one on which a RuntimeException was caught and another one where an IOException has caught; both in the contexts of Activity life-cycle methods.
+Then de developers were asked whether or not the way to handle a runtime exception should be different from the way to handle a checked exception.
+This question was motivated by the fact that during the mining study we could identify that many checked exception were just wrapped in a runtime exception and rethrown, as a way to by pass any kind of handling.
+When asked about of how to deal with a checked exception signaled by a method (invoked in the context of onPause() activity method), most respondents (62.50\%) said that they should add a try-catch block surrounding the method invocation.
+Also most of the respondents said that to deal with a method that is signaling a runtime exception we should surround such method by a try-catch block (xxxxx\%).
+However, what developers mentioned as handling actions (what is performed inside the catch clause to deal with the exception) differes in both cases.
+One of them mentioned that when the designer of an API decides to signal a checked exception s/he is telling the developer that the exception should be handled.
+Hence, some handling actions mentioned for checked exceptions were: (i) retrying the same operation; (ii) involving the user in finding a solution to the exception condition such as opening a pop-up and asking the user to define a new place to save the file (as in the example presented to them involved a IOException been signaled); and also (iii) presenting an error message to the user.
+On the other hand, the most common way of dealing with a runtime exception would be to catch it and present an error message to the user, or swallow the exception.
+Some developers also mentioned to use a Toast to support this task - a toast is an Android component that provides simple feedback about an operation in a small popup.
+(http://developer.android.com/guide/topics/ui/notifiers/toasts.html).
+Few developers mentioned that the runtime exception should flow upstream so the framework so it could gracefully crash the app.
+Although there is a long lasting debate about the pros and cons of checked and unchecked exceptions, most Android developers condireded checked exceptions a way od using exceptions that can prevent uncaught exception crashes - since in order to a checked exception to flow upstream and crash the app the developers needs to explicitly do so.
+//////
+
+\begin{table}
+\scriptsize
+\centering
+\begin{tabular}{lrr}
+\hline
+\bfseries{Top Ways} & \bfseries{\#} & \bfseries{\%} \\
+\hline
+add a try-catch block &	37 &	59\% \\
+present error message	&  14 &	22\% \\
+log the exception	& 12 &	19\% \\
+throw a checked exception	& 12 & 19\% \\
+let it crash / crash fast 	& 6 &	10\% \\
+swallow the exception	& 5 &	8\% \\
+add a throws declaration	& 5 &	8\% \\
+retry &	5 &	8\% \\
+let it crash &	4 &	6\% \\
+report crash &	3 &	5\% \\
+Use Toast & 	3 &	5\% \\
+\hline
+\end{tabular}
+\caption{Top ways of handling runtime exceptions -  63 non-empty responses. }
+\label{tab:handlingruntime}
+\end{table}
+
+\begin{table}
+\scriptsize
+\centering
+\begin{tabular}{lrr}
+\hline
+\bfseries{Top Ways} & \bfseries{\#} & \bfseries{\%} \\
+\hline
+add a try-catch block	& 25 &	63\% \\
+present error message	& 7 &	18\% \\
+involve the user in solving	& 5 &	13\% \\
+retry  &	5 &	13\% \\
+investigate the cause &	3 &	8\% \\
+same as runtime	& 3 &	8\% \\
+add a throws declaration / pass upstream &	2 &	5\% \\
+swallow	& 1 &	3\% \\
+toast	 & 1 & 3\% \\
+\hline
+\end{tabular}
+\caption{Top ways of dealing with a checked exception signaled in the context of an Activity method -  40 non-empty responses. }
+\label{tab:handlingruntime}
+\end{table}
+
+////// PREVENT NULL POINTER /////
+Developers were also asked about what they usually do to prevent NullPointerExceptions to happen when their app is in production.
+Most of them sais that include null-checks on the application code (59\%) mostly when checking the return of a method and as method guard conditions.
+Many respondents (39\%) also answered that instead of just sprinkling the code with null-checks when a NullPointerException happens they investigate the cause of the null reference and work to fix it, instead of just adding a null check.
+Some respondents also mention that use annotations such as @Nullable @NonNull, provided in Android since version XXX; such annotations guide de developer about what can and cannot be null during the application lifecycle.
+Altough catching the NullPointerExceptions is considered a bad practice some respondents said that they do it to prevent crashes caused by NullPointerExceptions.
+Such behavior was mentioned by one of the responded as the behavior adopted as as way non-expericente developers take to prevent such crashes - which is undoubtly a mistake.
+Some respondents also mentioned to replace the NullPointerException by IllegalArgumentException when a method should not receive a null as parameter.
+This is a good exception handling practice mentioned by Bloch in [ref].
+Only few developers mentioned about coding standards and static code analyzers to prevent it (2\%).////////
+
+\greybox{Determine why the object was null and attempt to fix this situation.
+In the case of external API calls which return null, then check for null (the quick-and-dirty way).
+For internal calls, use @NotNull and @Nullable annotations to provided more guidance on when an object \"may be\" and \"should never be\" null.}
+
+\greybox{Probably the most common exception, as such the reasons and fixes are super varying.
+But in general, null checks should be utilized and illegal state or argument exceptions should be used with an appropriate message.
+Which will communicate useful information without the confusion an NPE would normally cause.}
+
+\begin{table}
+\scriptsize
+\centering
+\begin{tabular}{lrr}
+\hline
+\bfseries{Top Ways} & \bfseries{\#} & \bfseries{\%} \\
+\hline
+null-checks &	36 &	59\% \\
+investigate/fix the cause	& 24 & 	39\% \\
+@Nullable @NonNull &	8 &	13\% \\
+catch null-pointer (mistake)	& 3 &	5\% \\
+initialize/use default variable	& 3 &	5\% \\
+new control-flow for null	& 3 &	5\% \\
+avoid using nulls / avoid to use methods can throw null	& 2 &	3\% \\
+static analisis	& 2 &	3\% \\
+automated testing	& 1 &	2\% \\
+make sure value is not null	& 1 &	2\% \\
+coding standards	& 1 &	2\% \\
+guard statements 	& 1 &	2\% \\
+handlers in all entry points	& 1 & 	2\% \\
+Use ART more than Dalvik	& 1 &	2\% \\
+Use IllegalArgument or IllegaState exception instead	& 1 &	2\% \\
+Verify the state of activity/fragment	& 1 &	2\% \\
+Write code in Scala, without nulls (joke)	& 1 &	2\% \\
+\hline
+\end{tabular}
+\caption{Top ways for preventing NullPointerException in Android apps. }
+\label{tab:handlingruntime}
+\end{table}
 

### 2015-10-06 SPRINGER_JOURNAL/msr_journal.tex
including tables with tags and overall comments about them

@@
 \end{tabular}
-\caption{Top ways for preventing NullPointerException in Android apps. }
+\caption{Top ways for preventing NullPointerException in Android apps - 61 non-empty responses. }
+\label{tab:handlingruntime}
+\end{table}
+
+//////NULLPONTER WORSE IN ANDROID /////
+Developers were also asked whether or not craches caused by NullPointerException were most common in Android development than in stantard Java development.
+Almost half of them (XXX\%) mentioned that Android apps were more suceptible to NullPointerException crashes.
+Some of the reasons mentioned by developers were: (i) the activity and fragment lifecycle - since both are constantly distroyed its references can frequently become null leading to NullPointerExceptions; (ii) the framework complexity - several methods od Android framework return null, and there is not enought documentation to inform the developer when using such methods.
+Moreover, few respondents also answered that such NullPointerException crashes can be caused by backward compatibility as well as memory issues as illustrated in Table X.
+The other half mentoned that NullPointerExceptions are  common in Android development as it is common in Java standard development.		//////
+
+\greybox{Honestly, there are a lot of very unskilled Java programmers out there writing Android apps.
+When they encounter NPE, they tend to null-check that variable, which just puts a bandage on the problem and causes other failures (usually also NPE's) later on in the application's lifecycle.}
+
+\greybox{This is nothing specific to Android, it's just how Java works.}
+
+\greybox{They can happen pretty much anywhere.		The Android Fragment system comes to mind in this case.
+Often, it is possible to find yourself in a state where getActivity() is null within the Fragment during certain points in the life cycle, and that is something I have to plan for.
+This might have been avoidable under a different structure.}
+
+\begin{table}
+\scriptsize
+\centering
+\begin{tabular}{lrr}
+\hline
+\bfseries{Top Ways} & \bfseries{\#} & \bfseries{\%} \\
+\hline
+activity/fragment life cycle  &	9 &	47\% \\
+framework complexity	& 6 &	32\% \\
+API / Backward compatibility	& 2 &	11\% \\
+memory issues & 	1 &	5\% \\
+no layers to catch runtimeexceptions	& 1 &	5\% \\
+\hline
+\end{tabular}
+\caption{Top reasons why NullPointerExceptions are more frequent in Android apps acording to 19 respondents . }
 \label{tab:handlingruntime}

### 2015-10-06 SPRINGER_JOURNAL/msr_journal.tex
including tables with tags and overall comments about them

@@
 
+/////// CROSS TYPE WRAPPING //////
+Developers were presented to a code snipet on which as cross-type wrapping was performed - as illustrated in Figure X.
+Then developers were asked whether such cross-type wrapping could affect the application robustness in some way.
+18\% of the respondents mentioned that the app would crash anyway, regardless the exception was wrapped or not.
+Most of the respondents, however, mentioned disadvantages of such wrapping such as: (i) it impairs proper handling since when wrapping in a general exception we loose information about the exception situation to he handled (24/%); such wrapping treat all exceptions as critical in other words, every exception will remain uncaught and will lead to an app crash; it does not allow the developer to perform a proper handling (e.g.
+retry) for exception types that are not critical.
+		//////
+
+\begin{table}
+\scriptsize
+\centering
+\begin{tabular}{lrr}
+\hline
+\bfseries{Top Ways} & \bfseries{\#} & \bfseries{\%} \\
+\hline
+
+impairs proper handling (loses exception information)	& 12 &	24\% \\
+uncaught will crash the app	& 12 &	24\% \\
+app will crash anyway	& 9	& 18\% \\
+should catch / handle properly (do local recovery)	 & 5 & 	10\% \\
+treat all exceptions as critical	& 5 &	10\% \\
+useless rethrow	& 2 &	4\% \\
+activity mehtods cannot throw exceptions &	1 &	2\% \\
+
+\hline
+\end{tabular}
+\caption{Top reasons why cross-type wrapping may impact app robustness - 49 non-empty responses. }
+\label{tab:handlingruntime}
+\end{table}
+
 \noindent\emph{\textbf{Q2: Do developers know about exception handling best practices?}}

### 2015-10-06 SPRINGER_JOURNAL/msr_journal.tex
including tables with tags and overall comments about them

@@
 18\% of the respondents mentioned that the app would crash anyway, regardless the exception was wrapped or not.
-Most of the respondents, however, mentioned disadvantages of such wrapping such as: (i) it impairs proper handling since when wrapping in a general exception we loose information about the exception situation to he handled (24/%); such wrapping treat all exceptions as critical in other words, every exception will remain uncaught and will lead to an app crash; it does not allow the developer to perform a proper handling (e.g.
+Most of the respondents, however, mentioned disadvantages of such wrapping such as: (i) it impairs proper handling since when wrapping in a general exception we loose information about the exception situation to he handled (24\%); such wrapping treat all exceptions as critical in other words, every exception will remain uncaught and will lead to an app crash; it does not allow the developer to perform a proper handling (e.g.
 retry) for exception types that are not critical.
@@
 \end{table}
+
+//////// JNI /////
+Only 3 respondents out of 71 (4\% of respondents) were aware of the fact that there are ways a method can throw a non-declared checked exception such as JNI/native code and relfection or directly changing the bytecode/dalvik code bypassing the compile-time checking.
+Most of the repondents were not aware that a method could throw a checked exception withou declaring it on its signature, regardless the circunstance.
+Most respondents mentioned that if a method throws a non-declared exception it must be a RuntimeException.
+We could observe that this exception handling bug hazard, although represent a threat to app robustness, is hardly known by the Android developers involved in this survey, and can represent.	////
+
+\greybox{I am guessing this is because the exception is thrown from native code (ie, C++ code in the JVM) where Java correct-ness semantics rules do not always apply.}
 

### 2015-10-06 SPRINGER_JOURNAL/msr_journal.tex
including tables with tags and overall comments about them,organizing comments

@@
 
-\section{The Mining Study Design}
+\section{The Repository Mining Study}
 \label{sec:study}
@@
 
-\section{The Mining Study Results}
+\subsection{The Mining Study Results}
 \label{sec:result}
@@
 
-\subsection{Can the root exceptions reveal bug harzards?}
+\subsubsection{Can the root exceptions reveal bug harzards?}
 After distilling the information available on the exception stack traces, we could find 
@@
 
-\subsection{Can the exception types reveal bug hazards?}
+\subsubsection{Can the exception types reveal bug hazards?}
 %\subsection{Exploring the Information Embedded on Exception Types}
@@
 
-\subsection{Can the exception wrappings reveal bug hazards?}
+\subsubsection{Can the exception wrappings reveal bug hazards?}
 
@@
 \bigskip
-
-\noindent\emph{\textbf{Q1:How often developers deal with the exception handling in Android development?}}
-
-Many of our survey respondents said they have to handle exceptions most of the time during Android app development (64.8\%).
+\noindent\emph{\textbf{Q1:How developers deal with the exception handling in Android development?}}
+\bigskip
+
+This first research question explores the ways in which developers deal with the exception handling code, either throwing or handling exceptions,  while developing an Android app.
+Developers were asked about the frequency they develop exception handling code and whether or not they adopt best practices while developing.
+
+%\noindent\emph{Frequency}
+
+Developers were firstly asked how often they dealt with the exception handling code.
+Many of our survey respondents recognized they have to handle exceptions most of the time during Android app development (64.8\%).
 However, the frequency on which they throw exception inside apps is smaller; most of the survey respondents said they throw exceptions only some of the time or seldom (80.3\%).
@@
 
-Developers were also asked whether the exception handling code helped with the development of robust Android applications.
-Although most of the developers answered positively (78.9\%) when asked to provide one or more reasons for their answers, most of the respondents also provide reasons why the exception handling code can sometimes impairs the robustness.
-Since positive and negative reasons were provided intermixed we analyzed both set of answers and ranked the top reasons why exception handling that can negatively or positively affect robustness according to developers perceptions.
-
- Table ~\ref{tab:topreasons} presents the top reasons why the exception handling code may or may not help de development of robust apps.
-Although most of the respondents answered that exception handling code helped with the development of robust Android applications, almost every respondent pointed a drawback associated to he exception handling code, saying that if not used with care the exception handling code may lead to app crashes.
+%\emph{EH Best Practices}
+%/////// JAVA BEST PRACTICES ////// 
+As mentioned before several general guidelines have been proposed on how to
+use Java exceptions.
+Section XXXXX compiled a list some exception handling (EH) best practices that guided our mining study.
+In this survey the developers were also asked if they knew about and used Java EH best practices.
+Although most of them said that knew and adopted Java EH practices 68\%, from this set the most mentioned best practices are presented on Table XXXX - considering that one developer mentioned more than one practice.
+Some of the practices mentioned, however, are not well known Java EH best practices but seam to be used by some Android developers to better deal with exceptions in Android app context: the crash fast approach [ref], the use of crash report tools, and to favor checked exceptions instead of runtime to represent the exceptional conditions in Android apps.
 
@@
 \hline
-\bfseries{Top Resons} & \bfseries{\#} & \bfseries{\%} \\
+\bfseries{Top Java EH Best Practices} & \bfseries{\#} & \bfseries{\%} \\
+\hline
+Use specific handlers / dont catch generic exceptions	& 9 & 23\% \\
+Don't swallow Exceptions	 & 7 & 18\% \\
+Dont throw Runtime / Favor Checked exceptions & 4 &	10\% \\
+Do not use exception for normal  flow control. &	4 &	10\% \\
+Free Resources in finally-blocks	& 4	& 10\% \\
+crash fast	 & 3 &	8\% \\
+crash report tools	&  2 &	5\% \\
+Don't catch Errors 	&  2 &	5\% \\
+\hline
+\end{tabular}
+\caption{Top mentioned Java EH best practices -  40 non-empty answers. }
+\label{tab:javapractices}
+\end{table}
+
+%/////// ANDROID EH BEST PRACTICES /////
+The developers who informed that adopted EH best practices were also asked whether they knew about any EH best practice specific to Android context.
+Most of them 43\% mentioned that they applied the same best practices that they used in Java programs.
+Only some of them mentioned that adopted specific best practices such as: (i) the use of crash report tools (21\%)  - to notify the developer about the uncaught exceptions that hapen in the application code; and (ii) the use of a global exception handler (14\%) - the UncaughtExceptionHandler advised on Android tutorials [ref].
+
+\begin{table}
+\scriptsize
+\centering
+\begin{tabular}{lrr}
+\hline
+\bfseries{Top Android EH Best Practices } & \bfseries{\#} & \bfseries{\%} \\
+\hline
+Same as java	& 6 &	43\% \\
+Crash report tools	 & 3 &	21\% \\
+Add a global exception handler (UncaughtExceptionHandler)	 & 2 & 14\% \\
+Use checked exceptions	& 1 &	7\% \\
+Use appropriate exception messages	& 1 &	7\% \\
+\hline
+\end{tabular}
+\caption{Top mentioned Android-specific EH best practices -  14 non-empty responses. }
+\label{tab:androidpractices}
+\end{table}
+
+\bigskip
+
+\noindent\emph{\textbf{Q2: How the exception handling code affect the development of robust apps and how developers prevent apps from crashing?}}
+
+\bigskip
+
+Developers were also asked whether the exception handling code helped with the development of robust Android applications.
+Although most of the developers answered positively (78.9\%) when asked to provide one or more reasons for their answers, most of the respondents also provide reasons why the exception handling code can sometimes impairs the robustness.
+Since positive and negative reasons were provided intermixed we analyzed both set of answers and ranked the top reasons why exception handling that can negatively or positively affect robustness according to developers perceptions.
+
+ Table ~\ref{tab:topreasons} presents the top reasons why the exception handling code may or may not help de development of robust apps.
+Although most of the respondents answered that exception handling code helped with the development of robust Android applications, almost every respondent pointed a drawback associated to he exception handling code, saying that if not used with care the exception handling code may lead to app crashes.
+
+\begin{table}
+\scriptsize
+\centering
+\begin{tabular}{lrr}
+\hline
+\bfseries{Top Reasons} & \bfseries{\#} & \bfseries{\%} \\
 \hline
@@
 
-\bigskip
-
-///// PREVENT CRASHES /////
+%\bigskip
+
 Finally developers were asked how they prevent their apps from crashing.
@@
 Once such tools notify developers about current crashes, the application can be fixed and hence future crashes (caused by what was fixed) could be prevented.
-Table ~ref{tab:topprevent} presents the themes for the strategies for crash prevention mentioned by developers.	////
+Table ~ref{tab:topprevent} presents the themes for the strategies for crash prevention mentioned by developers.
 
@@
 \hline
-\bfseries{Top Ways} & \bfseries{\#} & \bfseries{\%} \\
-\hline
-exception/error handling & 19 &	27\% \\
+\bfseries{Top Ways of Crash Prevention} & \bfseries{\#} & \bfseries{\%} \\
+\hline
+handling exception/error  & 19 &	27\% \\
 testing &	14 &	20\% \\
@@
 
-/////// JAVA BEST PRACTICES ////// 
-As mentioned before several general guidelines have been proposed on how to
-use Java exceptions.
-Section XXXXX compiled a list some exception handling best practices that guided our mining study.
-In this survey also we asked the Android developers whether they knew about any Java exception handling best practice.
-Although most of them said that knew and adopted EH practices 68\%, from this set the most mentioned best practices are presented on Table XXXX - considering that one developer mentioned more than one practice.
-Some of the practices mentioned, however, are not well known Java EH best practices but seam to be used by some Android developers to better deal with exceptions in Android app context: the crash fast approach [ref], the use of crash report tools, and to favor checked exceptions instead of runtime to represent the exceptional conditions in Android apps./////
+//// research
+
+\bigskip
+
+\noindent\emph{\textbf{Q3.1: How NullPointerException impact the development of robust Android apps?}}
+
+\bigskip
+
+//////NULLPONTER WORSE IN ANDROID /////
+Developers were also asked whether or not craches caused by NullPointerException were most common in Android development than in stantard Java development.
+Almost half of them (XXX\%) mentioned that Android apps were more suceptible to NullPointerException crashes.
+Some of the reasons mentioned by developers were: (i) the activity and fragment lifecycle - since both are constantly distroyed its references can frequently become null leading to NullPointerExceptions; (ii) the framework complexity - several methods od Android framework return null, and there is not enought documentation to inform the developer when using such methods.
+Moreover, few respondents also answered that such NullPointerException crashes can be caused by backward compatibility as well as memory issues as illustrated in Table X.
+The other half mentoned that NullPointerExceptions are  common in Android development as it is common in Java standard development.		//////
+
+\greybox{Honestly, there are a lot of very unskilled Java programmers out there writing Android apps.
+When they encounter NPE, they tend to null-check that variable, which just puts a bandage on the problem and causes other failures (usually also NPE's) later on in the application's lifecycle.}
+
+\greybox{This is nothing specific to Android, it's just how Java works.}
+
+\greybox{They can happen pretty much anywhere.		The Android Fragment system comes to mind in this case.
+Often, it is possible to find yourself in a state where getActivity() is null within the Fragment during certain points in the life cycle, and that is something I have to plan for.
+This might have been avoidable under a different structure.}
 
@@
 \hline
-Use specific handlers / dont catch generic exceptions	& 9 & 23\% \\
-Don't swallow Exceptions	 & 7 & 18\% \\
-Dont throw Runtime / Favor Checked exceptions & 4 &	10\% \\
-Do not use exception for normal  flow control. &	4 &	10\% \\
-Free Resources in finally-blocks	& 4	& 10\% \\
-crash fast	 & 3 &	8\% \\
-crash report tools	&  2 &	5\% \\
-Don't catch Errors 	&  2 &	5\% \\
+activity/fragment life cycle  &	9 &	47\% \\
+framework complexity	& 6 &	32\% \\
+API / Backward compatibility	& 2 &	11\% \\
+memory issues & 	1 &	5\% \\
+no layers to catch runtimeexceptions	& 1 &	5\% \\
 \hline
 \end{tabular}
-\caption{Top mentioned Java EH best practices -  40 non-empty answers. }
-\label{tab:javapractices}
-\end{table}
-
-/////// ANDROID EH BEST PRACTICES /////
-Developers who informed that adopted EH best practices were also asked whether they knew about any EH best practice specific to Android context.
-Most of them 43\% mentioned that they applied the same best practices that they used in Java programs.
-Only some of them mentioned that adopted specific best practices such as: (i) the use of crash report tools (21\%)  - to notify the developer about the uncaught exceptions that hapen in the application code; and (ii) the use of a global exception handler (14\%) - the UncaughtExceptionHandler advised on Android tutorials [ref].	 /////
-
-\begin{table}
-\scriptsize
-\centering
-\begin{tabular}{lrr}
-\hline
-\bfseries{Top Ways} & \bfseries{\#} & \bfseries{\%} \\
-\hline
-same as java	& 6 &	43\% \\
-crash report tools	 & 3 &	21\% \\
-add a global exception handler (UncaughtExceptionHandler)	 & 2 & 14\% \\
-use checked exceptions	& 1 &	7\% \\
-use appropriate exception messages	& 1 &	7\% \\
-\hline
-\end{tabular}
-\caption{Top mentioned Android-specific EH best practices -  14 non-empty responses. }
-\label{tab:androidpractices}
-\end{table}
-
-///// HANDLING RUNTIME AND CHECKED /////
-Developers were presented to two pieces of code: one on which a RuntimeException was caught and another one where an IOException has caught; both in the contexts of Activity life-cycle methods.
-Then de developers were asked whether or not the way to handle a runtime exception should be different from the way to handle a checked exception.
-This question was motivated by the fact that during the mining study we could identify that many checked exception were just wrapped in a runtime exception and rethrown, as a way to by pass any kind of handling.
-When asked about of how to deal with a checked exception signaled by a method (invoked in the context of onPause() activity method), most respondents (62.50\%) said that they should add a try-catch block surrounding the method invocation.
-Also most of the respondents said that to deal with a method that is signaling a runtime exception we should surround such method by a try-catch block (xxxxx\%).
-However, what developers mentioned as handling actions (what is performed inside the catch clause to deal with the exception) differes in both cases.
-One of them mentioned that when the designer of an API decides to signal a checked exception s/he is telling the developer that the exception should be handled.
-Hence, some handling actions mentioned for checked exceptions were: (i) retrying the same operation; (ii) involving the user in finding a solution to the exception condition such as opening a pop-up and asking the user to define a new place to save the file (as in the example presented to them involved a IOException been signaled); and also (iii) presenting an error message to the user.
-On the other hand, the most common way of dealing with a runtime exception would be to catch it and present an error message to the user, or swallow the exception.
-Some developers also mentioned to use a Toast to support this task - a toast is an Android component that provides simple feedback about an operation in a small popup.
-(http://developer.android.com/guide/topics/ui/notifiers/toasts.html).
-Few developers mentioned that the runtime exception should flow upstream so the framework so it could gracefully crash the app.
-Although there is a long lasting debate about the pros and cons of checked and unchecked exceptions, most Android developers condireded checked exceptions a way od using exceptions that can prevent uncaught exception crashes - since in order to a checked exception to flow upstream and crash the app the developers needs to explicitly do so.
-//////
-
-\begin{table}
-\scriptsize
-\centering
-\begin{tabular}{lrr}
-\hline
-\bfseries{Top Ways} & \bfseries{\#} & \bfseries{\%} \\
-\hline
-add a try-catch block &	37 &	59\% \\
-present error message	&  14 &	22\% \\
-log the exception	& 12 &	19\% \\
-throw a checked exception	& 12 & 19\% \\
-let it crash / crash fast 	& 6 &	10\% \\
-swallow the exception	& 5 &	8\% \\
-add a throws declaration	& 5 &	8\% \\
-retry &	5 &	8\% \\
-let it crash &	4 &	6\% \\
-report crash &	3 &	5\% \\
-Use Toast & 	3 &	5\% \\
-\hline
-\end{tabular}
-\caption{Top ways of handling runtime exceptions -  63 non-empty responses. }
-\label{tab:handlingruntime}
-\end{table}
-
-\begin{table}
-\scriptsize
-\centering
-\begin{tabular}{lrr}
-\hline
-\bfseries{Top Ways} & \bfseries{\#} & \bfseries{\%} \\
-\hline
-add a try-catch block	& 25 &	63\% \\
-present error message	& 7 &	18\% \\
-involve the user in solving	& 5 &	13\% \\
-retry  &	5 &	13\% \\
-investigate the cause &	3 &	8\% \\
-same as runtime	& 3 &	8\% \\
-add a throws declaration / pass upstream &	2 &	5\% \\
-swallow	& 1 &	3\% \\
-toast	 & 1 & 3\% \\
-\hline
-\end{tabular}
-\caption{Top ways of dealing with a checked exception signaled in the context of an Activity method -  40 non-empty responses. }
+\caption{Top reasons why NullPointerExceptions are more frequent in Android apps acording to 19 respondents . }
 \label{tab:handlingruntime}
@@
 
-//////NULLPONTER WORSE IN ANDROID /////
-Developers were also asked whether or not craches caused by NullPointerException were most common in Android development than in stantard Java development.
-Almost half of them (XXX\%) mentioned that Android apps were more suceptible to NullPointerException crashes.
-Some of the reasons mentioned by developers were: (i) the activity and fragment lifecycle - since both are constantly distroyed its references can frequently become null leading to NullPointerExceptions; (ii) the framework complexity - several methods od Android framework return null, and there is not enought documentation to inform the developer when using such methods.
-Moreover, few respondents also answered that such NullPointerException crashes can be caused by backward compatibility as well as memory issues as illustrated in Table X.
-The other half mentoned that NullPointerExceptions are  common in Android development as it is common in Java standard development.		//////
-
-\greybox{Honestly, there are a lot of very unskilled Java programmers out there writing Android apps.
-When they encounter NPE, they tend to null-check that variable, which just puts a bandage on the problem and causes other failures (usually also NPE's) later on in the application's lifecycle.}
-
-\greybox{This is nothing specific to Android, it's just how Java works.}
-
-\greybox{They can happen pretty much anywhere.		The Android Fragment system comes to mind in this case.
-Often, it is possible to find yourself in a state where getActivity() is null within the Fragment during certain points in the life cycle, and that is something I have to plan for.
-This might have been avoidable under a different structure.}
+\bigskip
+
+\noindent\emph{\textbf{Q3.2: How Cross-Type Wrappings impact the development of robust Android apps?}}
+
+\bigskip
+
+Some of the best practices are related to the different ways of handling checked and unchecked exceptions. When the designer of an API specifies that a method throws a checked exception, it is telling
+the caller of the API that such exception should be handled.
+
+///// HANDLING RUNTIME AND CHECKED /////
+Developers were presented to two pieces of code: one on which a RuntimeException was caught and another one where an IOException has caught; both in the contexts of Activity life-cycle methods.
+Then de developers were asked whether or not the way to handle a runtime exception should be different from the way to handle a checked exception.
+
+This question was motivated by the fact that during the mining study we could identify that many checked exception were just wrapped in a runtime exception and rethrown, as a way to by pass any kind of handling.
+When asked about of how to deal with a checked exception signaled by a method (invoked in the context of onPause() activity method), most respondents (62.50\%) said that they should add a try-catch block surrounding the method invocation.
+Also most of the respondents said that to deal with a method that is signaling a runtime exception we should surround such method by a try-catch block (xxxxx\%).
+However, what developers mentioned as handling actions (what is performed inside the catch clause to deal with the exception) differes in both cases.
+One of them mentioned that when the designer of an API decides to signal a checked exception s/he is telling the developer that the exception should be handled.
+Hence, some handling actions mentioned for checked exceptions were: (i) retrying the same operation; (ii) involving the user in finding a solution to the exception condition such as opening a pop-up and asking the user to define a new place to save the file (as in the example presented to them involved a IOException been signaled); and also (iii) presenting an error message to the user.
+On the other hand, the most common way of dealing with a runtime exception would be to catch it and present an error message to the user, or swallow the exception.
+Some developers also mentioned to use a Toast to support this task - a toast is an Android component that provides simple feedback about an operation in a small popup.
+(http://developer.android.com/guide/topics/ui/notifiers/toasts.html).
+Few developers mentioned that the runtime exception should flow upstream so the framework so it could gracefully crash the app.
+Although there is a long lasting debate about the pros and cons of checked and unchecked exceptions, most Android developers condireded checked exceptions a way od using exceptions that can prevent uncaught exception crashes - since in order to a checked exception to flow upstream and crash the app the developers needs to explicitly do so.
+//////
 
@@
 \hline
-activity/fragment life cycle  &	9 &	47\% \\
-framework complexity	& 6 &	32\% \\
-API / Backward compatibility	& 2 &	11\% \\
-memory issues & 	1 &	5\% \\
-no layers to catch runtimeexceptions	& 1 &	5\% \\
+add a try-catch block &	37 &	59\% \\
+present error message	&  14 &	22\% \\
+log the exception	& 12 &	19\% \\
+throw a checked exception	& 12 & 19\% \\
+let it crash / crash fast 	& 6 &	10\% \\
+swallow the exception	& 5 &	8\% \\
+add a throws declaration	& 5 &	8\% \\
+retry &	5 &	8\% \\
+let it crash &	4 &	6\% \\
+report crash &	3 &	5\% \\
+Use Toast & 	3 &	5\% \\
 \hline
 \end{tabular}
-\caption{Top reasons why NullPointerExceptions are more frequent in Android apps acording to 19 respondents . }
+\caption{Top ways of handling runtime exceptions -  63 non-empty responses. }
+\label{tab:handlingruntime}
+\end{table}
+
+\begin{table}
+\scriptsize
+\centering
+\begin{tabular}{lrr}
+\hline
+\bfseries{Top Ways} & \bfseries{\#} & \bfseries{\%} \\
+\hline
+add a try-catch block	& 25 &	63\% \\
+present error message	& 7 &	18\% \\
+involve the user in solving	& 5 &	13\% \\
+retry  &	5 &	13\% \\
+investigate the cause &	3 &	8\% \\
+same as runtime	& 3 &	8\% \\
+add a throws declaration / pass upstream &	2 &	5\% \\
+swallow	& 1 &	3\% \\
+toast	 & 1 & 3\% \\
+\hline
+\end{tabular}
+\caption{Top ways of dealing with a checked exception signaled in the context of an Activity method -  40 non-empty responses. }
 \label{tab:handlingruntime}
@@
 \end{table}
+
+\bigskip
+\noindent\emph{\textbf{RQ 3.3 Are developers aware the hidden dangers of non-declared checked exceptions?}}
+\bigskip
 

### 2015-10-06 SPRINGER_JOURNAL/msr_journal.tex
survey section

@@
 
-\subsection{Replication Package}
-All the data used in this study is publicly available at tool website hosted in GitHub: https://github.com/souzacoelho/exceptionminer.
-Specifically we provide: (i) all issues related to Android projects found
-in GiTHub and Google Code used in this study; (ii) all stack traces extracted
-from issues; (iii) the results of manual inspection steps; (iv) the
-ExceptionMiner tool we developed to support stack trace extraction and distilling.
-
 \subsection{The Mining Study Results}
@@
 
-When developing Java-based applications it is inevitable to deal with exceptions.
-Consequently, our first question explores how often developers deal with the exception handling code in Android development:
-
-\emph{\textbf{RQ 1:} How often developers deal with the exception handling in Android development?} To make the analysis easier, we further refine RQ1 in the following subquestions:
-\begin{itemize}
-\item RQ1.1:\emph{How often developers handle exceptions?}
-\item RQ1.2:\emph{How often developers throw exceptions?}
-\end{itemize}
-
-Another issue that show us how developers may deal with such bug hazards is
- their knowledge about exception handling best practices presented in Section~\ref{sec:best}.
- This leads us to our second research question:
-
-\bigskip
-
-\emph{\textbf{RQ 2:} Do Android developers adopt exception handling best practices?}
-
-\bigskip
-
-\emph{\textbf{RQ 4:}  Do Android developers consider implementing the exception handling code of Android apps more difficult than general Java applications?}
-
-\bigskip
-
-Then the third research question focus on how developers deal with the main bug hazards found during the mining study.
-
-\bigskip
-
-\emph{\textbf{RQ 3:}What are the developers perspectives about the exception handling bug hazards?} To answer such question we further refine RQ3 in the following subquestions:
-\begin{itemize}
-\item\emph{RQ3.1:Have developers ever faced a crash caused by NullPointerException?}
-\item\emph{RQ3.2:Are developers aware of the robustness threats caused by cross-type wrappings?}
-\item\emph{RQ3.3:Are developers aware of the robustness threats caused by JNL non-documented checked exceptions?}
-\end{itemize}
+When developing Java-based applications it is inevitable to deal with exceptions. Consequently, our first question explores how developers deal with the exception handling code in Android development:
+
+\emph{\textbf{RQ 1:} How often developers with the exception handling code while developing Android apps?}
+
+To make the analysis easier, we further refine RQ1 in subquestions, such as: \emph{How often developers handle exceptions?}; \emph{How often developers throw exceptions?}.
+Moreover, another question that show us how developers may deal with the EH code is: \emph{Do they knew about Java EH best practices and/or Android specific EH best practices?}
+
+Then following research question focus on the developers perspectives about the main bug hazards found during the mining study, more specifically:
+
+\emph{\textbf{RQ 2:} How the NullPointerExceptions impact the development of robust Android apps?}
+
+\emph{\textbf{RQ 3:} How cross-type wrappings impact the development of robust Android apps?}
+
+\emph{\textbf{RQ 4:} Are developers aware of the robustness threats caused by JNI undocumented checked exceptions?}
+
+%\emph{\textbf{RQ 4:}  Are developers aware of the non-declared checked exceptions?}
+
+%\emph{RQ 2:Have developers ever faced a crash caused by NullPointerException?}
+%\emph{RQ 3:Are developers aware of the robustness threats caused by cross-type wrappings?}; %\emph{RQ 4: Are developers aware of the robustness threats caused by JNI undocumented checked exceptions?}
 
@@
 the bug hazards were present and asked developers what they would do when faced with such scenarios.
+
+Finally developers were asked whether the exception handling code helped with the development of robust Android applications, and what they usually do to prevent apps from crashing.
+Hence, the last research question that guided this survey was the following:
+
+\emph{\textbf{RQ 5:} How the exception handling code affect the development of robust apps and how developers prevent apps from crashing?}
 
@@
 \end{figure}
+
+\begin{figure} \centering \includegraphics[scale=0.35]{charts/bestpractices.png}
+\caption{The developers awareness of exception handling best practices in Java (a) and Android (b). }
+\label{fig:oftenhandle}
+\end{figure}
+
+\bigskip
 
@@
 
-\noindent\emph{\textbf{Q2: How the exception handling code affect the development of robust apps and how developers prevent apps from crashing?}}
-
-\bigskip
-
-Developers were also asked whether the exception handling code helped with the development of robust Android applications.
-Although most of the developers answered positively (78.9\%) when asked to provide one or more reasons for their answers, most of the respondents also provide reasons why the exception handling code can sometimes impairs the robustness.
-Since positive and negative reasons were provided intermixed we analyzed both set of answers and ranked the top reasons why exception handling that can negatively or positively affect robustness according to developers perceptions.
-
- Table ~\ref{tab:topreasons} presents the top reasons why the exception handling code may or may not help de development of robust apps.
-Although most of the respondents answered that exception handling code helped with the development of robust Android applications, almost every respondent pointed a drawback associated to he exception handling code, saying that if not used with care the exception handling code may lead to app crashes.
-
-\begin{table}
-\scriptsize
-\centering
-\begin{tabular}{lrr}
-\hline
-\bfseries{Top Reasons} & \bfseries{\#} & \bfseries{\%} \\
-\hline
-\bfseries{For Positively Affecting Robustness} &   &   \\
-Improves error diagnosis  &	10	& 15\% \\
-Anticipated erroneous situation help writing more robust code  & 9	 & 13\% \\ 
-%Anticipated erroneous situation help writing more robust code 
-Exceptions only for unrecoverable behavior	& 7 &	10\% \\
-Checked exceptions force handling  &	6	& 9\% \\
-Useful to gracefully crash 	& 4 &	6\% \\
-
-\bfseries{For Negatively Affecting Robustness} &   &   \\
-Every exception should be caught to improve robustness	& 7	& 10\% \\
-Makes debugging harder &	2 &	3\% \\
-Unchecked/runtime exceptions can crash apps & 2	& 3\% \\
-NullPointerExceptoins can happen anywhere/are tricky to avoid	 & 2	& 3\% \\
-VM inefficiency 	& 1 & 1\% \\
-
-\hline
-\end{tabular}
-\caption{Top 5 reasons why exception handling helps or impairs the development of robust apps}
-\label{tab:topreasons}
-\end{table}
-
-%\bigskip
-
-Finally developers were asked how they prevent their apps from crashing.
-Most of the respondents answered that they prevent apps from crashing by handling exceptions/errors (27\%).
-They mentioned about some strategies such as: handling exceptions in all entry points/all over; using catch-all clauses; and adding try catch blocks around risky methods.
-Different from conventional Java programs, an Android app is composed by several entry points, each window (i.e.
-Activity) is a potential entry point on which exceptions can scape leading to an app crash.
-Hence the importance of handling exceptions in all entry points, and using general cash clauses which cannot sometimes prevent crash but allows the developer to present a detailed error message to the user before crashing.
-Many respondents also said that they used testing approaches to prevent crashes -  few of them mentioned crowd testing (3\%) an emerging approach in testing apps.
-Moreover, null checks were also mentioned by 15\% of the developers as a way to prevent crashes.
-The adoption of coding styles were also pointed out by 14\% of respondents as a way to prevent craches.
-Finally, the crash report tools (e.g.
-ACRA [ref]) were reported as a way to prevent crashes.
-Once such tools notify developers about current crashes, the application can be fixed and hence future crashes (caused by what was fixed) could be prevented.
-Table ~ref{tab:topprevent} presents the themes for the strategies for crash prevention mentioned by developers.
-
-\begin{table}
-\scriptsize
-\centering
-\begin{tabular}{lrr}
-\hline
-\bfseries{Top Ways of Crash Prevention} & \bfseries{\#} & \bfseries{\%} \\
-\hline
-handling exception/error  & 19 &	27\% \\
-testing &	14 &	20\% \\
-null check(s) &	11 &	15\% \\
-- null check annotations (@Null and @NonNull)	& 3 & 	4\% \\
-coding style(s)	& 10 &	14\% \\
-crash report / crash report tools &	8	& 11\% \\
-\hline
-\end{tabular}
-\caption{Top ways of preventing crashes - 71 non-empty respondents. }
-\label{tab:topprevent}
-\end{table}
-
-//// research
-
-\bigskip
-
-\noindent\emph{\textbf{Q3.1: How NullPointerException impact the development of robust Android apps?}}
-
-\bigskip
-
-//////NULLPONTER WORSE IN ANDROID /////
+\noindent\emph{\textbf{Q 2: How the NullPointerExceptions impact the development of robust Android apps?}}
+
+\bigskip
+
+%//////NULLPONTER WORSE IN ANDROID /////
 Developers were also asked whether or not craches caused by NullPointerException were most common in Android development than in stantard Java development.
@@
 Some of the reasons mentioned by developers were: (i) the activity and fragment lifecycle - since both are constantly distroyed its references can frequently become null leading to NullPointerExceptions; (ii) the framework complexity - several methods od Android framework return null, and there is not enought documentation to inform the developer when using such methods.
-Moreover, few respondents also answered that such NullPointerException crashes can be caused by backward compatibility as well as memory issues as illustrated in Table X.
-The other half mentoned that NullPointerExceptions are  common in Android development as it is common in Java standard development.		//////
-
-\greybox{Honestly, there are a lot of very unskilled Java programmers out there writing Android apps.
-When they encounter NPE, they tend to null-check that variable, which just puts a bandage on the problem and causes other failures (usually also NPE's) later on in the application's lifecycle.}
-
-\greybox{This is nothing specific to Android, it's just how Java works.}
 
@@
 This might have been avoidable under a different structure.}
+
+\greybox{The Android framework is what adds the complexity in figuring out what caused an exception because more often than not, the error is triggered from the framework as a result of something else you did.}
+
+Moreover, few respondents also answered that such NullPointerException crashes can be caused by backward compatibility as well as memory issues as illustrated in Table X.
+The other half mentoned that NullPointerExceptions are  common in Android development as it is common in Java standard development.
+
+\greybox{Honestly, there are a lot of very unskilled Java programmers out there writing Android apps.
+When they encounter NPE, they tend to null-check that variable, which just puts a bandage on the problem and causes other failures (usually also NPE's) later on in the application's lifecycle.}
+
+%\greybox{This is nothing specific to Android, it's just how Java works.}
 
@@
 
-////// PREVENT NULL POINTER /////
+%////// PREVENT NULL POINTER /////
 Developers were also asked about what they usually do to prevent NullPointerExceptions to happen when their app is in production.
@@
 This is a good exception handling practice mentioned by Bloch in [ref].
-Only few developers mentioned about coding standards and static code analyzers to prevent it (2\%).////////
+Only few developers mentioned about coding standards and static code analyzers to prevent it (2\%).
 
@@
 
-\noindent\emph{\textbf{Q3.2: How Cross-Type Wrappings impact the development of robust Android apps?}}
-
-\bigskip
-
-Some of the best practices are related to the different ways of handling checked and unchecked exceptions. When the designer of an API specifies that a method throws a checked exception, it is telling
-the caller of the API that such exception should be handled.
-
-///// HANDLING RUNTIME AND CHECKED /////
-Developers were presented to two pieces of code: one on which a RuntimeException was caught and another one where an IOException has caught; both in the contexts of Activity life-cycle methods.
-Then de developers were asked whether or not the way to handle a runtime exception should be different from the way to handle a checked exception.
+\noindent\emph{\textbf{RQ 3: How Cross-Type Wrappings impact the development of robust Android apps?}}
+
+\bigskip
+
+Some of the best practices are related to the different ways exceptions should be handled according to whether it is a checked and an unchecked exception.
+When the designer of an API specifies that a method throws a checked exception, it is telling the caller of the API that such exception should be handled.
+
+%///// HANDLING RUNTIME AND CHECKED /////		
+To acess developers perspective on whether checked and unchecked exceptions should be handled different inside the app, the developers were presented to two pieces of code: one on which a RuntimeException was caught and another one where an IOException has caught; both in the contexts of Activity life-cycle methods.
+Then the developers were asked whether or not the way to handle a runtime exception should be different from the way to handle a checked exception.
 
@@
 However, what developers mentioned as handling actions (what is performed inside the catch clause to deal with the exception) differes in both cases.
-One of them mentioned that when the designer of an API decides to signal a checked exception s/he is telling the developer that the exception should be handled.
+
+One of the respondents mentioned that when the designer of an API decides to signal a checked exception s/he is telling the developer that the exception should be handled.
 Hence, some handling actions mentioned for checked exceptions were: (i) retrying the same operation; (ii) involving the user in finding a solution to the exception condition such as opening a pop-up and asking the user to define a new place to save the file (as in the example presented to them involved a IOException been signaled); and also (iii) presenting an error message to the user.
+
 On the other hand, the most common way of dealing with a runtime exception would be to catch it and present an error message to the user, or swallow the exception.
-Some developers also mentioned to use a Toast to support this task - a toast is an Android component that provides simple feedback about an operation in a small popup.
-(http://developer.android.com/guide/topics/ui/notifiers/toasts.html).
+Some developers also mentioned to use a Toast to support this task\footnote{A toast is an Android component that provides simple feedback about an operation in a small popup - http://developer.android.com/guide/topics/ui/notifiers/toasts.html}.
 Few developers mentioned that the runtime exception should flow upstream so the framework so it could gracefully crash the app.
 Although there is a long lasting debate about the pros and cons of checked and unchecked exceptions, most Android developers condireded checked exceptions a way od using exceptions that can prevent uncaught exception crashes - since in order to a checked exception to flow upstream and crash the app the developers needs to explicitly do so.
-//////
 
@@
 \hline
-add a try-catch block &	37 &	59\% \\
-present error message	&  14 &	22\% \\
-log the exception	& 12 &	19\% \\
-throw a checked exception	& 12 & 19\% \\
-let it crash / crash fast 	& 6 &	10\% \\
-swallow the exception	& 5 &	8\% \\
-add a throws declaration	& 5 &	8\% \\
-retry &	5 &	8\% \\
-let it crash &	4 &	6\% \\
-report crash &	3 &	5\% \\
+Add a try-catch block &	37 &	59\% \\
+Present error message	&  14 &	22\% \\
+Log the exception	& 12 &	19\% \\
+Throw a checked exception	& 12 & 19\% \\
+Let it crash / crash fast 	& 6 &	10\% \\
+Swallow the exception	& 5 &	8\% \\
+Add a throws declaration	& 5 &	8\% \\
+Report crash &	3 &	5\% \\
 Use Toast & 	3 &	5\% \\
@@
 \hline
-add a try-catch block	& 25 &	63\% \\
-present error message	& 7 &	18\% \\
-involve the user in solving	& 5 &	13\% \\
-retry  &	5 &	13\% \\
-investigate the cause &	3 &	8\% \\
-same as runtime	& 3 &	8\% \\
-add a throws declaration / pass upstream &	2 &	5\% \\
-swallow	& 1 &	3\% \\
-toast	 & 1 & 3\% \\
+Add a try-catch block	& 25 &	63\% \\
+Present error message	& 7 &	18\% \\
+Involve the user in solving	& 5 &	13\% \\
+Retry  &	5 &	13\% \\
+Investigate the cause &	3 &	8\% \\
+Same way as runtime	& 3 &	8\% \\
+Add a throws declaration / pass upstream &	2 &	5\% \\
+%Swallow	& 1 &	3\% \\
+%Toast	 & 1 & 3\% \\
 \hline
@@
 
-/////// CROSS TYPE WRAPPING //////
-Developers were presented to a code snipet on which as cross-type wrapping was performed - as illustrated in Figure X.
+%/////// CROSS TYPE WRAPPING //////
+Hence, developers were presented to a code snipet on which as cross-type wrapping was performed\footnote{This cross-type wrapping was found in several applicarions during the mining study as well as in some classes of Android framework} - as illustrated in Figure X.
 Then developers were asked whether such cross-type wrapping could affect the application robustness in some way.
-18\% of the respondents mentioned that the app would crash anyway, regardless the exception was wrapped or not.
+Table XXXX illustrates the  most mentioned reasons why cross-type wrapping may affect app robustness; 18\% of the respondents mentioned that the app would crash anyway, regardless the exception was wrapped or not.
 Most of the respondents, however, mentioned disadvantages of such wrapping such as: (i) it impairs proper handling since when wrapping in a general exception we loose information about the exception situation to he handled (24\%); such wrapping treat all exceptions as critical in other words, every exception will remain uncaught and will lead to an app crash; it does not allow the developer to perform a proper handling (e.g.
 retry) for exception types that are not critical.
-		//////
+
+{\footnotesize
+\begin{verbatim}
+  @Override
+  public void onPause() {
+     try {
+        ...
+      } catch (Exception e) {
+            throw new RuntimeException("...", e);
+      }
+ }
+\end{verbatim}
+}
 
@@
 \hline
-\bfseries{Top Ways} & \bfseries{\#} & \bfseries{\%} \\
-\hline
-
+\bfseries{Top Reasons Why Cross-Type Wrapping Affect Robustness} & \bfseries{\#} & \bfseries{\%} \\
+\hline
 impairs proper handling (loses exception information)	& 12 &	24\% \\
@@
 \end{tabular}
-\caption{Top reasons why cross-type wrapping may impact app robustness - 49 non-empty responses. }
+\caption{Top reasons why cross-type wrapping may affect app robustness - 49 non-empty responses. }
 \label{tab:handlingruntime}
@@
 \bigskip
-\noindent\emph{\textbf{RQ 3.3 Are developers aware the hidden dangers of non-declared checked exceptions?}}
-\bigskip
-
-//////// JNI /////
-Only 3 respondents out of 71 (4\% of respondents) were aware of the fact that there are ways a method can throw a non-declared checked exception such as JNI/native code and relfection or directly changing the bytecode/dalvik code bypassing the compile-time checking.
-Most of the repondents were not aware that a method could throw a checked exception withou declaring it on its signature, regardless the circunstance.
+\noindent\emph{\textbf{RQ 4 Are developers aware of the non-declared checked exceptions?}}
+\bigskip
+
+%//////// JNI /////
+The respondents were presented to a piece of code on which a non-declared checked exception has scaped, and were asked if they new any reason for it to happen.
+Only 3 respondents out of 71 (4\% of respondents) were aware of the fact that there are ways a method can throw a non-declared checked exception, such as JNI/native code, relfection, and directly changing the bytecode/dalvik code which bypass the compile-time checking.
+Most of the repondents were not aware that a method could throw a checked exception withou declaring it on the method's signature.
 Most respondents mentioned that if a method throws a non-declared exception it must be a RuntimeException.
-We could observe that this exception handling bug hazard, although represent a threat to app robustness, is hardly known by the Android developers involved in this survey, and can represent.	////
+We could observe that this exception handling bug hazard detected during the mining study, although represent a threat to app robustness, is hardly known by the Android developers involved in this survey.
+Below we can find a quote from one of the 3 respondents aware of JNI non-delcared checked exceptions.
 
@@
 
-\noindent\emph{\textbf{Q2: Do developers know about exception handling best practices?}}
-
-We can observe that 62\% of surveyed Android developers follow Java exception handling best practices.
-Among the best practices cited by developers the most popular ones were: avoid generic catches and avoid empty catch blocks.
-They did not mention a any guide or reference of best practices.
-When asked about best practices specific to Android developers, most of the developers mentioned that they did no about them.
-The ones who answered positively said that they adopt the same best practices as in any Java program.
-
-\begin{figure} \centering \includegraphics[scale=0.35]{charts/bestpractices.png}
-\caption{The developers awareness of exception handling best practices in Java (a) and Android (b). }
-\label{fig:oftenhandle}
-\end{figure}
-
-\bigskip
-
-\noindent\emph{\textbf{Q3:What are the developers perspectives about the exception handling bug hazards?}}
-
-\greybox{Sometimes when you develop multi-threaded applications, when the thread finishes you don't have any GUI (or Context) to show information, so you have to be aware of your application's state (foreground, background)}
-
-Handle life cycle of activities and fragments.
-
-Q10: What makes exception handling in Android different from exception handling in Java?
-
-The Android framework is what adds the complexity in figuring out what caused an exception because more often than not, the error is triggered from the framework as a result of something else you did
-
-Sometimes when you develop multi-threaded applications, when the thread finishes you don't have any GUI (or Context) to show information, so you have to be aware of your application's state (foreground, background).
-
-Application-Activity-Fragment lifecycle.
-
-You can wrap the main function in java in try/catch in android you can\'t but there are equivalents
-
-Though not particular to Android, most Android apps have several different entry points: Activities, Services, etc, can be started by different services; same thing with events.
+%\noindent\emph{\textbf{Q2: Do developers know about exception handling best practices?}}
+%We can observe that 62\% of surveyed Android developers follow Java exception handling best 
+%practices. Among the best practices cited by developers the most popular ones were: avoid generic 
+%catches and avoid empty catch blocks. They did not mention a any guide or reference of best practices. 
+%When asked about best practices specific to Android developers, most of the developers mentioned that 
+%they did no about them. The ones who answered positively said that they adopt the same best practices 
+%as in any Java program.
+
+\noindent\emph{\textbf{RQ5: How the exception handling code affect the development of robust apps and how developers prevent apps from crashing?}}
+
+\bigskip
+
+Developers were also asked whether the exception handling code helped with the development of robust Android applications.
+Although most of the developers answered positively (78.9\%) when asked to provide one or more reasons for their answers, most of the respondents also provide reasons why the exception handling code can sometimes impairs the robustness.
+Since positive and negative reasons were provided intermixed we analyzed both set of answers and ranked the top reasons why exception handling that can negatively or positively affect robustness according to developers perceptions.
+Table ~\ref{tab:topreasons} presents the top reasons why the exception handling code may or may not help de development of robust apps.
+
+\begin{table}
+\scriptsize
+\centering
+\begin{tabular}{lrr}
+\hline
+\bfseries{Top Reasons} & \bfseries{\#} & \bfseries{\%} \\
+\hline
+\bfseries{For Positively Affecting Robustness} &   &   \\
+Improves error diagnosis  &	10	& 15\% \\
+Anticipated erroneous situation help writing more robust code  & 9	 & 13\% \\ 
+%Anticipated erroneous situation help writing more robust code 
+Exceptions only for unrecoverable behavior	& 7 &	10\% \\
+Checked exceptions force handling  &	6	& 9\% \\
+Useful to gracefully crash 	& 4 &	6\% \\
+
+\bfseries{For Negatively Affecting Robustness} &   &   \\
+Every exception should be caught to improve robustness	& 7	& 10\% \\
+Makes debugging harder &	2 &	3\% \\
+Unchecked/runtime exceptions can crash apps & 2	& 3\% \\
+NullPointerExceptoins can happen anywhere/are tricky to avoid	 & 2	& 3\% \\
+VM inefficiency 	& 1 & 1\% \\
+
+\hline
+\end{tabular}
+\caption{Top 5 reasons why exception handling helps or impairs the development of robust apps}
+\label{tab:topreasons}
+\end{table}
+
+Although most of the respondents answered that exception handling code helped with the development of robust Android applications, almost every respondent pointed a drawback associated to he exception handling code, saying that if not used with care the exception handling code may lead to app crashes.
+
+\greybox{Any uncatched exception will crash the app.
+There are many unknown exception thrown by the system, that are only happen once in a lifetime like IllegalStateException: eglMakeCurrent failed EGL\_BAD\_CONTEXT}
+
+\greybox{The lifecycle of activities\/fragements etc make it harder to work out what order things are called in and so exceptions can occur because you didn't realise that another method isn't called.}
+
+%\bigskip
+
+Finally developers were asked how they prevent their apps from crashing.
+Most of the respondents answered that they prevent apps from crashing by handling exceptions/errors (27\%).
+They mentioned about some strategies such as: handling exceptions in all entry points/all over; using catch-all clauses; and adding try catch blocks around risky methods.
+Different from conventional Java programs, an Android app is composed by several entry points, each window (i.e.
+Activity) is a potential entry point on which exceptions can scape leading to an app crash.
+Hence the importance of handling exceptions in all entry points, and using general cash clauses which cannot sometimes prevent crash but allows the developer to present a detailed error message to the user before crashing.
+Many respondents also said that they used testing approaches to prevent crashes -  few of them mentioned crowd testing (3\%) an emerging approach in testing apps.
+Moreover, null checks were also mentioned by 15\% of the developers as a way to prevent crashes.
+The adoption of coding styles were also pointed out by 14\% of respondents as a way to prevent craches.
+Finally, the crash report tools (e.g.
+ACRA [ref]) were reported as a way to prevent crashes.
+Once such tools notify developers about current crashes, the application can be fixed and hence future crashes (caused by what was fixed) could be prevented.
+Table ~ref{tab:topprevent} presents the themes for the strategies for crash prevention mentioned by developers.
+
+\greybox{Though not particular to Android, most Android apps have several different entry points: Activities, Services, etc, can be started by different services; same thing with events.
 It may be hard to ensure that all this cases are properly handled.
-The logcat is a helpful way of tracking exceptions when developing and testing on real phones.
-
-Any uncatched exception will crash the app.
-There are many unknown exception thrown by the system, that are only happen once in a lifetime like IllegalStateException: eglMakeCurrent failed EGL\_BAD\_CONTEXT
-
-The lifecycle of activities\/fragements etc make it harder to work out what order things are called in and so exceptions can occur because you didn't realise that another method isn't called
+}
+
+\greybox{You can wrap the main function in java in try/catch in android you can\'t. }
+
+\begin{table}
+\scriptsize
+\centering
+\begin{tabular}{lrr}
+\hline
+\bfseries{Top Ways of Crash Prevention} & \bfseries{\#} & \bfseries{\%} \\
+\hline
+handling exception/error  & 19 &	27\% \\
+testing &	14 &	20\% \\
+null check(s) &	11 &	15\% \\
+- null check annotations (@Null and @NonNull)	& 3 & 	4\% \\
+coding style(s)	& 10 &	14\% \\
+crash report / crash report tools &	8	& 11\% \\
+\hline
+\end{tabular}
+\caption{Top ways of preventing crashes - 71 non-empty respondents. }
+\label{tab:topprevent}
+\end{table}
+
+\bigskip
+
+\section{Replication Package}
+All the data used in the mining study and in the survey-based study is publicly available at the ExceptionMiner tool website hosted in GitHub\footnote{ https://github.com/souzacoelho/exceptionminer}.
+Specifically we provide: (i) all issues related to Android projects found
+in GiTHub and Google Code used in this study; (ii) all stack traces extracted
+from issues; (iii) the results of manual inspection steps; (iv) the
+ExceptionMiner tool we developed to support stack trace extraction and distilling; (v) the survey questions; and (iv) the survey responses and summary.
 
@@
 \noindent
- \emph{``Everybody hates thinking about exceptions, because they are not supposed to happen''\\
+ \emph{``Everybody hates thinking about exceptions, because they are not supposed to happen'' 
   (Brian Foote)}\footnote{Brian Foote shared his opinion in a conversation with James Noble - quoted on the paper: hillside.net/plop/2008/papers/ACMVersions/coelho.pdf}
@@
 
-\emph{\textbf{The exception handling confusion problem.}}
+%\emph{\textbf{The exception handling confusion problem.}}
+
+\subsection{The exception handling confusion problem}
 When (mis)applied, exception wrapping can make the exception-related code
@@
  better deal with them.
-Furthermore, this calls for empirical studies on the actual usefulness of Java's hybrid exception model.
+Furthermore, only some of the Android developers surveyed in this study were aware of the 
+rebustness threats caused by cross-type wrappings as the one cited above. This calls for empirical studies on the actual usefulness of Java's hybrid exception model.
 
@@
 
-\emph{\textbf{On the null pointer problem.}}
+%\emph{\textbf{On the null pointer problem.}}
+\subsection{On the null pointer problem.}
 The null reference was firstly introduced by Tony Hoare in ALGOL W, which after some years he called his ``one-billion-dollar mistake''~\cite{hoare2}.
 In this sudy, the null references were, infact, responsible for several reported issues - providing further evidence to Hoare's statement.
-This observation emphasizes the need for solutions to avoid NullPointerExceptions, such as:
+Moreover many of the survey respondents recognized that NullPointerExceptions was one of the main causes of application crashes since they can happen almost
+anyware in the application code, and to make thinks worse the life-cycle of Android apps on which objects are constantly recreated (i.e., Activity and Fragment classes) favor this kind of exception to happen added to the fact that many Android framework methods return null without making such
+return type explicitly on the documentation.
+Such observations emphasizes the need for solutions to avoid NullPointerExceptions, such as:
 (i) lightweight intra-method null pointer analysis as supported by
@@
 or computations that may fail) could improve the robustness of Java programs. 
-
-\emph{\textbf{Preventing uncaught exceptions}}
+Some of the Android developers mentioned how @Nullable annotations could helpful to deal with
+NullPointerExceptions and consequently prevent app crashes caused by it.
+
+%\emph{\textbf{Preventing uncaught exceptions}}
+\subsection{Preventing uncaught exceptions}
 In this study we could observe undocumented runtime exceptions thrown by third party code,
@@
 for the client code to protect against `unforeseen'' situations that may happen 
-while calling a library code.
+while calling a library code. In the survey-based study Android developers were asked about
+ways to prevent application crashes which are mainly caused by uncaught exceptions.
+Many of them emphasized the importance of handling exceptions to prevent crashes
+but also mentioned how difficult is to handle every specific exception that can happen.
+Several developers recognized to follow a more reactive behaviour against crashes:
+they advocate the use of crash report tools, and once the crash happens for the first time,
+its reported and the application can be changed to better cope with the exceptions that crashed the crashes, consequently preventing future similar crashes.
 
@@
 adequately handled inside the applications. Although this 
-solution may prevent  the system from abruptly crashing,
- such a general handler will not have enough
+solution may prevent  the system from abruptly crashing, as mentioned by some of the surveyed developers such a general handler will not have enough
 contextual information to adequately handle the exception, 
@@
 Since  documenting runtime exceptions is a tedious and error prone
-task, 
-this calls for tool support to automate the extraction of runtime exceptions
+task, this calls for tool support to automate the extraction of runtime exceptions
 from library code. Initial steps in this direction have been proposed
@@
 
-\section{Threats to Validity}
-\label{sec:threats}
-
+%\section{Threats to Validity}
+%\label{sec:threats}
 %\noindent
+
+%\emph{\textbf{Threats to Validity of the Mining Study.}} 
+\subsection{Mining Study - Threats to Validity }
+
 \emph{Internal Validity.} We used a heuristics-based parser to mine
@@
 and ``cross-type wrappings'').
+
+\subsection{Limitations of the Survey-based Study.}
+
 %\noindent

### 2015-10-07 SPRINGER_JOURNAL/msr_journal.tex
survey section changes

@@
 \title{Exception Handling Bug Hazards in Android}
-\subtitle{Results from a Mining and Survey-based Study}
+\subtitle{Results from a Mining and Exploratory Survey Study}
 
@@
 
-\institute{F. Author \at
-              first address \\
-              Tel.: +123-45-678910\\
-              Fax: +123-45-678910\\
-              \email{fauthor@example.com}           %  \\
-%             \emph{Present address:} of F. Author  %  if needed
+\institute{Roberta Coelho \at
+              Federal University of Rio Grande do Norte, CIVT/UFRN. Av. Senador Salgado Filho, 3000. 
+Lagoa Nova, CEP: 59.078-970. Natal/RN. Brazil \\
+              Tel.: +55-84-3342-2216 \\
+              \email{roberta@dimap.ufrn.br}           
            \and
-           S. Author \at
-              second address
+          Lucas Almeida \at
+             \email{lucas.almeida@ppgsc.ufrn.br} \\    
+          \and    
+         Georgios Gousios   \at
+  \email{georgios@cs.ru.nl} \\    
+ \and
+         Arie van Deursen  \at
+  \email{arie.vandeursen@tudelft.nl} \\      
+\and 
+   Christoph Treude \at
+  \email{ctreude@dimap.ufrn.br} \\    
 }
@@
 \begin{abstract}
-Insert your abstract here. Include keywords, PACS and mathematical
-subject classification numbers as needed.
-\keywords{First keyword \and Second keyword \and More}
+This paper reports two studies:  a repository mining study whose goal was to identify 
+exception handling bug hazards in Android apps; and an exploratory survey study - whose goal
+was to assess the perspectives of developers conserning such bug hazards.
+The repository mining study mined the exception stack traces embedded on 
+159,048 issues reported on Android projects hosted in GitHub (482 projs.) 
+and Googlecode (157 projs.).  Overall 6,005 exception stack traces were extracted, 
+and analyzed in combination with source code and bytecode analysis. 
+Some bug hazards were consistently detected such as:  unexpected exception wrappings (e.g., Errors being wrapped in checked exceptions), and NullPointerExeptions as the major cause of reported chashes in about 50\% of all analyzed projects.
+Finally, a survey was applied to some developers
+involved on such GitHub projects and their perspective about the bug hazards and 
+how they deal with them as assessed and discussed.
+
+\keywords{Exception handling \and Android development \and Repository mining \and Exploratory survey}
 % \PACS{PACS code1 \and PACS code2 \and more}
@@
 
-In recent years, we have witnessed an astonishing increase in the number of
-mobile applications. These applications extend phones capabilities 
-far beyond basic calls and textual messages. They must, however,
-face an increasing number of threats to application robustness
- arising from failures in the underlying middleware and hardware (e.g., camera, sensors);
-failures in third party services and libraries; compatibility issues~\cite{McDon13}; 
- memory and batery restrictions; and noisy external resources~\cite{Zhang12} 
-(e.g., wireless connections,GPS, bluetooth).
-
-Therefore, techniques for error detection and handling are not  an optional add-on but a 
+The number of mobile apps is increasing in a daily rate. If on the one hand they
+extend phones capabilities far beyond of the basic calls, on the other hand they have to 
+cope with an increasing number of exceptinal conditions 
+(e.g., faults in underlying middleware or hardware; compatibility issues~\cite{McDon13}; memory and battery restrictions; noisy external resources~\cite{Zhang12}).
+
+Therefore, mechanisms for exception detection and handling are not an optional add-on but a 
 fundamental part of such apps. The exception handling mechanism~\cite{goodenough1975exception},
@@
 
-Studies have shown that exception-related code is generally poorly understood and among the least tested parts of the system~\cite{miller1997issues,Robil00,shah2010understanding, 
+However, such mechanisms are more often then not the less 
+understood and tested parts of the system~\cite{miller1997issues,Robil00,shah2010understanding, 
 garcia2007extracting,garcia2001comparative,cabral2007exception,coelho2011unveiling,yuan:2014.osdi}.
@@
 the exception occurred~\cite{gosling2000java,bloch2008effective}.
+When available, the exception stack traces provide a useful source of information about system crashes ~\cite{bettenburg2008makes} which 
+can enable different kinds of post-mortem analysis and support:  debugging~\cite{schroter2010stack}, 
+bug classification and clustering~\cite{wang2013improving, kim2011crash, dhaliwal2011classifying},  
+automated bug fixing~\cite{sinha2009fault} and fault-proneness prediction models~\cite{kim2013predicting}.
 
@@
 
-%When available, the exception stack traces provide a useful source of information about system crashes ~\cite{bettenburg2008makes} which 
-% can enable different kinds of post-mortem analysis and support:  debugging~\cite{schroter2010stack}, 
-% bug classification and clustering~\cite{wang2013improving, kim2011crash, dhaliwal2011classifying},  
-%automated bug fixing~\cite{sinha2009fault} and fault-proneness prediction models~\cite{kim2013predicting}.
-
-This study performs a post mortem analysis of the exception stack traces 
-included in issues
-reported on Android projects hosted in GitHub and Google Code. 
+This work is conducted in two phases. Firstly, a mining study performs a post 
+mortem analysis of the exception stack traces 
+included in issues reported on Android projects hosted in GitHub and Google Code.
 The goal of this study is to investigate whether the reported exception stack traces
@@
 of introducing the aforementioned uncaught exceptions.
+Secondly, an exploratory survey to the Android developers involved on the mined projects
+to assess their perspective about the exception handling bug hazards found.
 
@@
 
- The high prevalence of NullPointerExceptions is in line with findings of earlier research~\cite{kim2013predicting,fraser20131600,csallner2004jcrasher}, as are the undocumented runtime 
-exceptions signaled by the Android Platform~\cite{kechagia2014}. 
-Some of the findings of our
- study emphasize the impact of these bug hazards on the application robustness 
+The exploratory survey was conducted with 70 Android developers involved in one or more of the 
+GitHub Android projects whose issues were mined. This survey reveaks that only few developers (3\% of respondents)
+ new about the undocumented checked exceptions signaled by native C code of Android platform. 
+Moreover, most of the developers recognized that cross-type exception wrappings may
+negatively impact the application robustness. The uncaught exceptions due to errors in programming 
+logic, with the NullPointerException was identified by most developers (68\%) as the first or second main cause 
+of application crashes.
+
+ The high prevalence of NullPointerExceptions found in the mining study and confirmed on the exploratory survey
+ is in line with findings of earlier research~\cite{kim2013predicting,fraser20131600,csallner2004jcrasher}, 
+as are the undocumented runtime exceptions signaled by the Android Platform~\cite{kechagia2014}.
+
+Some of the findings of our mining study emphasize the impact of these bug hazards on the application robustness 
 by mining a different source of information as the ones used in previous works. The present work 
@@
  Furthermore, our work points to bug hazards that were not detected by previous research (i.e., cross-type wrappings, undocumented checked exceptions and undocumented runtime exceptions thrown by third-party libraries) which represent new threats to application robustness.
+Moreover, we perform  the first exploratory survey study whose
+goal was to assess developers perspective regarding a set of exception handling bug hazards
+as well as how developers deal with exceptions and prevent crashes while developing.
 
@@
 Section~\ref{sec:result} reports the findings of our study.
-Section~\ref{sec:disc} provides a discussion of the wider implications of our results.
-Section~\ref{sec:threats} presents the threats to validity associated to this study. 
+Section~\ref{sec:disc} provides a discussion of the wider implications of our results,
+ presents the threats to validity associated to the mining study and discusses the limitations
+of the survey-based study. 
 Finally Section~\ref{sec:rele} describes related work, 
@@
 For each repository, we extract the developers whose email are registered on each repository.
-We emailed the developers and received 68 valid answers.
-
+We emailed the 1,824 developers and received 71 valid answers.
 The majority of our respondents have more than 2 years of Java development experience (87.3\%) and  considerable experience in Android development (more than 1 year)  (85.9\%) - see Figure~\ref{fig:devexpertise}.
 
-\begin{figure} \centering \includegraphics[scale=0.35]{charts/Q1Q2_new.png}
+%Q1Q2_new.png
+\begin{figure} \centering \includegraphics[scale=0.35]{charts/expertise.png}
 \caption{Developers expertise in Java and Android development.}
@@
 
-Moreover, few respondents also answered that such NullPointerException crashes can be caused by backward compatibility as well as memory issues as illustrated in Table X.
+Moreover, few respondents also answered that such NullPointerException crashes can be caused by backward compatibility as well as memory issues as illustrated in Table XXXX.
 The other half mentoned that NullPointerExceptions are  common in Android development as it is common in Java standard development.
@@
 \bigskip
-\noindent\emph{\textbf{RQ 4 Are developers aware of the non-declared checked exceptions?}}
+\noindent\emph{\textbf{RQ 4: Are developers aware of the robustness threats caused by JNI undocumented checked exceptions?}}
 \bigskip
@@
 %\bigskip
+
+Developers were asked to rank the main causes of crashes.
+Approximately 68\% of the repondents ranked the exceptions signaled by programming logic mistakes e.g.
+NullPointerExceptions as the Top 1 and Top 2 cause of crashes as illustrated in Figure ~\ref{figure:ranking}.
+In our mining study the exceptions signaled due to programming logic mistakes (java.lang and util) where the causes of 52,0\% of the exception stack traces found.
+Hence, the developers perspective is aligned with what was revealed in mining study.
+
+\begin{figure*} 
+\centering \includegraphics[width=.9\hsize]{charts/maincauses.png}
+\caption{Developers perspectives on the main causes of crashes.}\label{fig:ranking} 
+\end{figure*}
 
@@
 \subsection{Limitations of the Survey-based Study.}
+Due to the exploratory nature of the second phase of this work whose 
+goal was to identify the developers perspectives concerning the exception 
+handling bug hazards found during the repository mining 
+phase, we choose Gounded Theory techniques. The results of our
+survey-based study may not apply to every Android developer, since other populations 
+might add new insights.  The population we collected data from was comprised 
+by Android developers of the GitHub 
+Android projects whose issues were mined on the first phase of this work, and who had
+time and motivation to answer the survey questions.
+
+Altough the themes and findings that emerged in our study cannot 
+be generalized, they give a first view of developers perspectives about
+ EH bug hazards found. Hence, we believe that the survey-based study 
+have contributed with valuable insights about how Android developers 
+used the exception handling code and what are their perspectives regarding
+exception handling bug hazards.
 
@@
 
+\textit{Exploratory Survey Studies.} Exploratory surveys have been used in the software
+engineering context to discover the user perspective regarding a broad range of 
+topics such as: how developers define productivity [ref]; how GitHub developers use
+pull-requests[ref]; how the testing culture of open-shource projects can be characterized [ref]; 
+and even how developers use Twiter [ref]. This kind of study is important as a way of 
+better understanding the developers behavior and hence provide recommendations and 
+tools to help on specific development tasks. This was the first exploratory survey study whose
+goal was to assess developers perspective regarding a set of exception handling bug hazards
+as well as how developers deal with exceptions and prevent crashes while developing.
+
 %\enlargethispage{-2\baselineskip}
@@
 
-The goal of this paper is to investigate 
+The goal of this paper is two-fold: (i) to investigate 
 to what extent stack trace information can reveal \emph{bug hazards} 
 related to exception handling code that may lead to a decrease in
-application robustness. 
+application robustness; and (ii) investigate the developers perspective
+concerning the detected EH \emph{bug hazards}.
+
 To that end, we mined the stack 
 traces embedded in all issues defined in 482 Android projects hosted in Github and 
-157 projects hosted in Google Code. Overall it included 6,005
-exception stack traces. Our first key contribution is
+157 projects hosted in Google Code - overall it included 6,005
+exception stack traces. And we surveyed developers associated to some of 
+the mined GitHub projects.
+Our first key contribution is
 a novel approach and toolset (ExceptionMiner) for analyzing Java
@@
   violating Java's exception handling principles.
+Our third contribution was the first exploratory to assess the 
+developers perspective concerning the exception handling 
+bug hazards in Android development.
  Our results shed light on common problems and bug hazards in Java
@@
 
-\begin{acknowledgements}
-If you'd like to thank anyone, place your comments here
-and remove the percent signs.
-\end{acknowledgements}
+%\begin{acknowledgements}
+%If you'd like to thank anyone, place your comments here
+%and remove the percent signs.
+%\end{acknowledgements}
 

### 2015-10-07 SPRINGER_JOURNAL/msr_journal.tex
survey section

@@
 %Q1Q2_new.png
-\begin{figure} \centering \includegraphics[scale=0.35]{charts/expertise.png}
+\begin{figure} \centering \includegraphics[scale=0.38]{charts/expertise.png}
 \caption{Developers expertise in Java and Android development.}
@@
 
-\begin{figure} \centering \includegraphics[scale=0.35]{charts/often.png}
-\caption{How often developers throw (a) and handle (b) exceptions during app development.}
+\begin{figure} \centering \includegraphics[scale=0.63]{charts/often_new.png}
+\caption{How often developers throw and handle exceptions during app development.}
 \label{fig:oftenhandle}
@@
 
-\begin{figure} \centering \includegraphics[scale=0.35]{charts/bestpractices.png}
-\caption{The developers awareness of exception handling best practices in Java (a) and Android (b). }
+\begin{figure} \centering \includegraphics[scale=0.57]{charts/summary_new.png}
+\caption{Summary of some survey question answers. }
 \label{fig:oftenhandle}
@@
 Hence, the developers perspective is aligned with what was revealed in mining study.
-
+%[width=.9\hsize]
 \begin{figure*} 
-\centering \includegraphics[width=.9\hsize]{charts/maincauses.png}
+\centering \includegraphics[scale=0.42]{charts/maincauses.png}
 \caption{Developers perspectives on the main causes of crashes.}\label{fig:ranking} 

### 2015-10-07 SPRINGER_JOURNAL/msr_journal.tex
survey section

@@
 %///// HANDLING RUNTIME AND CHECKED /////		
-To acess developers perspective on whether checked and unchecked exceptions should be handled different inside the app, the developers were presented to two pieces of code: one on which a RuntimeException was caught and another one where an IOException has caught; both in the contexts of Activity life-cycle methods.
+To assess developers perspective on whether checked and unchecked exceptions should be handled different inside the app, the developers were presented to two pieces of code: one on which a RuntimeException was caught and another one where an IOException has caught; both in the contexts of Activity life-cycle methods.
 Then the developers were asked whether or not the way to handle a runtime exception should be different from the way to handle a checked exception.
@@
 related to exception handling code that may lead to a decrease in
-application robustness; and (ii) investigate the developers perspective
+application robustness; and (ii) to assess the developers perspective
 concerning the detected EH \emph{bug hazards}.

### 2015-10-07 SPRINGER_JOURNAL/msr_journal.tex
survey section - new charts

@@
 
-\begin{figure} \centering \includegraphics[scale=0.63]{charts/often_new.png}
+\begin{figure} \centering \includegraphics[scale=0.63]{charts/agree_new2.png}
 \caption{How often developers throw and handle exceptions during app development.}
@@
 
-\begin{figure} \centering \includegraphics[scale=0.57]{charts/summary_new.png}
+\begin{figure} \centering \includegraphics[scale=0.57]{charts/yes_no_new2.png}
 \caption{Summary of some survey question answers. }
@@
 %//////NULLPONTER WORSE IN ANDROID /////
-Developers were also asked whether or not craches caused by NullPointerException were most common in Android development than in stantard Java development.
+Developers were asked if their apps ever crashed  because of a NullPointerException. The vast majority of respondents (96\%) said that their apps crashed at least once due to a NullPointerException. 
+Their answers are aligned with the high prevalence of NullPointerExceptions found in the mining study - in Table XXX we can see that NullPointerExceptions were the main cause of the mined exception stack traces.
+The developers were also asked whether or not craches caused by NullPointerException were most common in Android development than in stantard Java development.
 Almost half of them (XXX\%) mentioned that Android apps were more suceptible to NullPointerException crashes.

### 2015-10-07 SPRINGER_JOURNAL/msr_journal.tex
survey section

@@
 
-\begin{figure} \centering \includegraphics[scale=0.63]{charts/agree_new2.png}
+\begin{figure} \centering \includegraphics[scale=0.60]{charts/agree_new2.png}
 \caption{How often developers throw and handle exceptions during app development.}
@@
 
-\begin{figure} \centering \includegraphics[scale=0.57]{charts/yes_no_new2.png}
+\begin{figure} \centering \includegraphics[scale=0.60]{charts/yes_no_new2.png}
 \caption{Summary of some survey question answers. }
@@
 \begin{figure*} 
-\centering \includegraphics[scale=0.42]{charts/maincauses.png}
+\centering \includegraphics[scale=0.60]{charts/ranking.png}
 \caption{Developers perspectives on the main causes of crashes.}\label{fig:ranking} 

### 2015-10-07 SPRINGER_JOURNAL/msr_journal.tex
survey section
one editorial pass on Section 4

@@
 \title{Exception Handling Bug Hazards in Android}
-\subtitle{Results from a Mining and Exploratory Survey Study}
+\subtitle{Results from a Mining Study and an Exploratory Survey}
 
@@
 
-\section{The Developers Perspective}
+\section{The Developers' Perspective}
 \label{sec:dev}
 
-Although the mining study revelaed a set of bug hazards in Android exception handling no work had examined yet what are the developers perceives about them.
+Although the mining study revealed a set of bug hazards in Android exception handling, no work has examined yet what the developers' perceptions about them are.
 With Android development rapidly rising in popularity, it is important to better understand how developers deal with such bug hazards and what they do to overcome them.
 
-To better understand the developers perspective about the bug hazards detected during the mining study, we set up an exploratory qualitative investigation and survey Android developers on how they face such bug hazards in their projects.
-Our field of study is GitHub; using our GHTorrent database~\cite{Gousi13}, we aimed our survey at developers from Android projects available in GitHub and which issues were mined during the first phase of this study.
-
-Our main goal is to learn from developers perspective about the Java exception handling usage and the main exception handling bug hazards in Android.
+To better understand the developers' perspective about the bug hazards detected during the mining study, we set up an exploratory qualitative investigation and surveyed Android developers on how they perceive such bug hazards in their projects.
+The scope of our study is GitHub---using our GHTorrent database~\cite{Gousi13}, we aimed our survey at developers from Android projects available in GitHub for which issues were mined during the first phase of this study.
+
+Our main goal is to learn from developers about the Java exception handling usage and the main exception handling bug hazards in Android.
 We therefore use a survey as our main research instrument.
@@
 The questions were motivated by the analysis of the existing bug hazards detected in the first phase of this study.
-Next sections detail the survey.
+The next sections detail the survey.
 
@@
 
-\emph{\textbf{RQ 1:} How often developers with the exception handling code while developing Android apps?}
-
-To make the analysis easier, we further refine RQ1 in subquestions, such as: \emph{How often developers handle exceptions?}; \emph{How often developers throw exceptions?}.
-Moreover, another question that show us how developers may deal with the EH code is: \emph{Do they knew about Java EH best practices and/or Android specific EH best practices?}
-
-Then following research question focus on the developers perspectives about the main bug hazards found during the mining study, more specifically:
-
-\emph{\textbf{RQ 2:} How the NullPointerExceptions impact the development of robust Android apps?}
-
-\emph{\textbf{RQ 3:} How cross-type wrappings impact the development of robust Android apps?}
+\emph{\textbf{RQ 1:} How often do developers deal with exception handling code while developing Android apps?}
+
+To make the analysis easier, we further refine RQ1 in subquestions, such as: \emph{How often do developers handle exceptions?}; \emph{How often do developers throw exceptions?}.
+Moreover, another question that shows us how developers may deal with the EH code is: \emph{Do developers know about Java EH best practices and/or Android specific EH best practices?}
+
+The subsequent research questions focus on the developers' perspectives about the main bug hazards found during the mining study, more specifically:
+
+\emph{\textbf{RQ 2:} How do NullPointerExceptions impact the development of robust Android apps?}
+
+\emph{\textbf{RQ 3:} How do cross-type wrappings impact the development of robust Android apps?}
 
@@
 
-Some of such questions were indirectly answered by presenting a set of code snippets on which
-the bug hazards were present and asked developers what they would do when faced with such scenarios.
+Some of these questions were indirectly answered by presenting a set of code snippets in which
+the bug hazards were present and asking developers what they would do when faced with such scenarios.
 
@@
 
-\emph{\textbf{RQ 5:} How the exception handling code affect the development of robust apps and how developers prevent apps from crashing?}
+\emph{\textbf{RQ 5:} How does the exception handling code affect the development of robust apps and how do developers prevent apps from crashing?}
 
@@
 The survey is split into two logical sections; multiple choice or Likert-scale questions and open-ended questions.
-The multiple choice questions are intermixed with open-ended questions to further elicit the developer’s opinions.
-Moreover, the survey also contains Likert scales questions to force participants to make a choice.
-Overall, the survey includes 13 open-ended questions, 5 Likert scale questions and 10 multiple choice questions.
-The respondents could fill in the survey in about 15 minutes.
+The multiple choice questions are intermixed with open-ended questions to further elicit the developer's opinions.
+Moreover, the survey also contains Likert-scale questions to force participants to make a choice.
+Overall, the survey includes 13 open-ended questions, 5 Likert-scale questions and 10 multiple choice questions.
+The respondents could complete the survey in about 15 minutes.
 
 We used grounded theory coding~\cite{charmaz2006}, iterating through our open-ended survey responses, to answer some of the research questions.
-The grounded theory coding used in this study consists of at two phases: (1) initial coding entails a close reading of the data and (2) later we use focused coding ~\cite{charmaz2006} to pinpoint and develop the most salient themes~\cite{charmaz2006} in the analyzed data.
+The grounded theory coding used in this study consists of two phases: (1) initial coding entails a close reading of the data and (2) later we use focused coding ~\cite{charmaz2006} to pinpoint and develop the most salient themes~\cite{charmaz2006} in the analyzed data.
 
@@
 
- In previous work~\cite{coelhoetal15}, we mined exception stack traces available on issues reported on several Android projects hosted in GitHub and Googlecode.
-To ensure that our sample consists of repositories that were real Android apps, we had to inspect every project site and discard the toy-programs and non-Android repositories.
-For each repository, we extract the developers whose email are registered on each repository.
+ In previous work~\cite{coelhoetal15}, we mined exception stack traces available in issues reported for several Android projects hosted on GitHub and Googlecode.
+To ensure that our sample consists of repositories that were real Android apps, we had to inspect every project site and discard toy-programs and non-Android repositories.
+For each repository, we extract the developers whose emails are registered.
 We emailed the 1,824 developers and received 71 valid answers.
@@
 \bigskip
-\noindent\emph{\textbf{Q1:How developers deal with the exception handling in Android development?}}
+\noindent\emph{\textbf{Q1:How do developers deal with the exception handling in Android development?}}
 \bigskip
@@
 
-Developers were firstly asked how often they dealt with the exception handling code.
+Developers were first asked how often they dealt with the exception handling code.
 Many of our survey respondents recognized they have to handle exceptions most of the time during Android app development (64.8\%).
-However, the frequency on which they throw exception inside apps is smaller; most of the survey respondents said they throw exceptions only some of the time or seldom (80.3\%).
-It shows that although we can try to avoid dealing with exceptions throwing few exceptions, it is almost always inevitable to handle the exceptions thrown by Android platform and reused libraries.
+However, the frequency in which they throw exceptions inside apps is smaller; most of the survey respondents said they throw exceptions only some of the time or seldom (80.3\%).
+It shows that although we can try to avoid dealing with exceptions by throwing few exceptions, it is almost always inevitable to handle the exceptions thrown by the Android platform and reused libraries.
 
@@
 use Java exceptions.
-Section XXXXX compiled a list some exception handling (EH) best practices that guided our mining study.
+Section XXXXX compiled a list of some exception handling (EH) best practices that guided our mining study.
 In this survey the developers were also asked if they knew about and used Java EH best practices.
-Although most of them said that knew and adopted Java EH practices 68\%, from this set the most mentioned best practices are presented on Table XXXX - considering that one developer mentioned more than one practice.
-Some of the practices mentioned, however, are not well known Java EH best practices but seam to be used by some Android developers to better deal with exceptions in Android app context: the crash fast approach [ref], the use of crash report tools, and to favor checked exceptions instead of runtime to represent the exceptional conditions in Android apps.
+Most of them said that they knew and adopted Java EH practices (68\%).
+From this set the most mentioned best practices are presented in Table XXXX -- considering that one developer mentioned more than one practice.
+Some of the practices mentioned, however, are not well known Java EH best practices but seem to be used by some Android developers to better deal with exceptions in the Android app context: the crash fast approach [ref], the use of crash report tools, and to favor checked exceptions instead of runtime to represent the exceptional conditions in Android apps.
 
@@
 %/////// ANDROID EH BEST PRACTICES /////
-The developers who informed that adopted EH best practices were also asked whether they knew about any EH best practice specific to Android context.
-Most of them 43\% mentioned that they applied the same best practices that they used in Java programs.
-Only some of them mentioned that adopted specific best practices such as: (i) the use of crash report tools (21\%)  - to notify the developer about the uncaught exceptions that hapen in the application code; and (ii) the use of a global exception handler (14\%) - the UncaughtExceptionHandler advised on Android tutorials [ref].
+The developers who informed us about their EH best practices were also asked whether they knew about any EH best practices specific to Android.
+Most of them (43\%) mentioned that they applied the same best practices that they used in Java programs.
+Only some of them mentioned that they adopted specific best practices such as: (i) the use of crash report tools (21\%) -- to notify the developer about the uncaught exceptions that happen in the application code; and (ii) the use of a global exception handler (14\%) -- the UncaughtExceptionHandler advised in Android tutorials [ref].
 
@@
 
-\noindent\emph{\textbf{Q 2: How the NullPointerExceptions impact the development of robust Android apps?}}
+\noindent\emph{\textbf{Q 2: How do the NullPointerExceptions impact the development of robust Android apps?}}
 
@@
 Developers were asked if their apps ever crashed because of a NullPointerException. The vast majority of respondents (96\%) said that their apps crashed at least once due to a NullPointerException. 
-Their answers are aligned with the high prevalence of NullPointerExceptions found in the mining study - in Table XXX we can see that NullPointerExceptions were the main cause of the mined exception stack traces.
-The developers were also asked whether or not craches caused by NullPointerException were most common in Android development than in stantard Java development.
-Almost half of them (XXX\%) mentioned that Android apps were more suceptible to NullPointerException crashes.
-Some of the reasons mentioned by developers were: (i) the activity and fragment lifecycle - since both are constantly distroyed its references can frequently become null leading to NullPointerExceptions; (ii) the framework complexity - several methods od Android framework return null, and there is not enought documentation to inform the developer when using such methods.
+Their answers are aligned with the high prevalence of NullPointerExceptions found in the mining study -- in Table XXX we can see that NullPointerExceptions were the main cause of the mined exception stack traces.
+The developers were also asked whether crashes caused by NullPointerExceptions were more common in Android development than in standard Java development.
+Almost half of them (XXX\%) mentioned that Android apps were more susceptible to NullPointerException crashes.
+Some of the reasons mentioned by developers were: (i) the activity and fragment lifecycle -- since both are constantly destroyed their references can frequently become null, leading to NullPointerExceptions; (ii) the framework complexity -- several methods of the Android framework return null, and there is not enough documentation to inform the developer when using such methods.
 
@@
 
-Moreover, few respondents also answered that such NullPointerException crashes can be caused by backward compatibility as well as memory issues as illustrated in Table XXXX.
-The other half mentoned that NullPointerExceptions are  common in Android development as it is common in Java standard development.
+Moreover, a few respondents also answered that such NullPointerException crashes can be caused by backward compatibility as well as memory issues as illustrated in Table XXXX.
+The other half mentioned that NullPointerExceptions are  common in Android development since they are common in Java standard development.
 
@@
 %////// PREVENT NULL POINTER /////
-Developers were also asked about what they usually do to prevent NullPointerExceptions to happen when their app is in production.
-Most of them sais that include null-checks on the application code (59\%) mostly when checking the return of a method and as method guard conditions.
+Developers were also asked about what they usually do to prevent NullPointerExceptions from happening when their app is in production.
+Most of them mentioned including null-checks in the application code (59\%), mostly when checking the return of a method and as method guard conditions.
 Many respondents (39\%) also answered that instead of just sprinkling the code with null-checks when a NullPointerException happens they investigate the cause of the null reference and work to fix it, instead of just adding a null check.
-Some respondents also mention that use annotations such as @Nullable @NonNull, provided in Android since version XXX; such annotations guide de developer about what can and cannot be null during the application lifecycle.
-Altough catching the NullPointerExceptions is considered a bad practice some respondents said that they do it to prevent crashes caused by NullPointerExceptions.
-Such behavior was mentioned by one of the responded as the behavior adopted as as way non-expericente developers take to prevent such crashes - which is undoubtly a mistake.
+Some respondents also mentioned using annotations such as @Nullable @NonNull, provided in Android since version XXX; such annotations guide the developer about what can and cannot be null during the application lifecycle.
+Although catching the NullPointerExceptions is considered a bad practice some respondents said that they do it to prevent crashes caused by NullPointerExceptions.
+Such behavior was mentioned by one of the respondents as the behavior adopted by non-experienced developers  to prevent such crashes -- which is undoubtedly a mistake.
 Some respondents also mentioned to replace the NullPointerException by IllegalArgumentException when a method should not receive a null as parameter.
-This is a good exception handling practice mentioned by Bloch in [ref].
-Only few developers mentioned about coding standards and static code analyzers to prevent it (2\%).
+This is a good exception handling practice mentioned by Bloch [ref].
+Only a few developers mentioned coding standards and static code analyzers to prevent it (2\%).
 
@@
 
-\noindent\emph{\textbf{RQ 3: How Cross-Type Wrappings impact the development of robust Android apps?}}
-
-\bigskip
-
-Some of the best practices are related to the different ways exceptions should be handled according to whether it is a checked and an unchecked exception.
+\noindent\emph{\textbf{RQ 3: How do Cross-Type Wrappings impact the development of robust Android apps?}}
+
+\bigskip
+
+Some of the best practices are related to the different ways an exception should be handled according to whether it is a checked and an unchecked exception.
 When the designer of an API specifies that a method throws a checked exception, it is telling the caller of the API that such exception should be handled.
@@
 %///// HANDLING RUNTIME AND CHECKED /////		
-To assess developers perspective on whether checked and unchecked exceptions should be handled different inside the app, the developers were presented to two pieces of code: one on which a RuntimeException was caught and another one where an IOException has caught; both in the contexts of Activity life-cycle methods.
-Then the developers were asked whether or not the way to handle a runtime exception should be different from the way to handle a checked exception.
-
-This question was motivated by the fact that during the mining study we could identify that many checked exception were just wrapped in a runtime exception and rethrown, as a way to by pass any kind of handling.
-When asked about of how to deal with a checked exception signaled by a method (invoked in the context of onPause() activity method), most respondents (62.50\%) said that they should add a try-catch block surrounding the method invocation.
-Also most of the respondents said that to deal with a method that is signaling a runtime exception we should surround such method by a try-catch block (xxxxx\%).
-However, what developers mentioned as handling actions (what is performed inside the catch clause to deal with the exception) differes in both cases.
+To assess the developers' perspective on whether checked and unchecked exceptions should be handled differently inside the app, the developers were presented with two pieces of code: one in which a RuntimeException was caught and another one where an IOException was caught; both in the context of Activity life-cycle methods.
+Then the developers were asked whether the way to handle a runtime exception should be different from the way to handle a checked exception.
+
+This question was motivated by the fact that during the mining study we could identify that many checked exceptions were just wrapped in a runtime exception and re-thrown, as a way to bypass any kind of handling.
+When asked about how to deal with a checked exception signaled by a method (invoked in the context of the onPause() activity method), most respondents (62.50\%) said that they should add a try-catch block surrounding the method invocation.
+Also, most of the respondents said that to deal with a method that is signaling a runtime exception we should surround such a method with a try-catch block (xxxxx\%).
+However, what developers mentioned as handling actions (what is performed inside the catch clause to deal with the exception) differs in both cases.
 
 One of the respondents mentioned that when the designer of an API decides to signal a checked exception s/he is telling the developer that the exception should be handled.
-Hence, some handling actions mentioned for checked exceptions were: (i) retrying the same operation; (ii) involving the user in finding a solution to the exception condition such as opening a pop-up and asking the user to define a new place to save the file (as in the example presented to them involved a IOException been signaled); and also (iii) presenting an error message to the user.
+Hence, some handling actions mentioned for checked exceptions were: (i) retrying the same operation; (ii) involving the user in finding a solution to the exception condition such as opening a pop-up and asking the user to define a new place to save the file (the example presented to them involved a IOException being signaled); and also (iii) presenting an error message to the user.
 
 On the other hand, the most common way of dealing with a runtime exception would be to catch it and present an error message to the user, or swallow the exception.
-Some developers also mentioned to use a Toast to support this task\footnote{A toast is an Android component that provides simple feedback about an operation in a small popup - http://developer.android.com/guide/topics/ui/notifiers/toasts.html}.
-Few developers mentioned that the runtime exception should flow upstream so the framework so it could gracefully crash the app.
-Although there is a long lasting debate about the pros and cons of checked and unchecked exceptions, most Android developers condireded checked exceptions a way od using exceptions that can prevent uncaught exception crashes - since in order to a checked exception to flow upstream and crash the app the developers needs to explicitly do so.
+Some developers also mentioned to use a toast to support this task\footnote{A toast is an Android component that provides simple feedback about an operation in a small popup - http://developer.android.com/guide/topics/ui/notifiers/toasts.html}.
+Few developers mentioned that the runtime exception should flow upstream to the framework so it could gracefully crash the app.
+Although there is a long lasting debate about the pros and cons of checked and unchecked exceptions, most Android developers considered checked exceptions as a way of using exceptions that can prevent uncaught exception crashes -- since in order for a checked exception to flow upstream and crash the app the developers need to explicitly do so.
 
@@
 %/////// CROSS TYPE WRAPPING //////
-Hence, developers were presented to a code snipet on which as cross-type wrapping was performed\footnote{This cross-type wrapping was found in several applicarions during the mining study as well as in some classes of Android framework} - as illustrated in Figure X.
+Hence, developers were presented with a code snippet in which a cross-type wrapping was performed\footnote{This cross-type wrapping was found in several applications during the mining study as well as in some classes of the Android framework.} -- as illustrated in Figure X.
 Then developers were asked whether such cross-type wrapping could affect the application robustness in some way.
-Table XXXX illustrates the  most mentioned reasons why cross-type wrapping may affect app robustness; 18\% of the respondents mentioned that the app would crash anyway, regardless the exception was wrapped or not.
-Most of the respondents, however, mentioned disadvantages of such wrapping such as: (i) it impairs proper handling since when wrapping in a general exception we loose information about the exception situation to he handled (24\%); such wrapping treat all exceptions as critical in other words, every exception will remain uncaught and will lead to an app crash; it does not allow the developer to perform a proper handling (e.g.
+Table XXXX illustrates the  most mentioned reasons of why cross-type wrapping may affect app robustness; 18\% of the respondents mentioned that the app would crash anyway, regardless of whether the exception was wrapped or not.
+Most of the respondents, however, mentioned disadvantages of such wrapping, such as: (i) it impairs proper handling since wrapping in a general exception loses information about the exception situation to be handled (24\%); such wrapping treats all exceptions as critical -- in other words, every exception will remain uncaught and will lead to an app crash -- and it does not allow the developer to perform a proper handling (e.g.
 retry) for exception types that are not critical.
@@
 %//////// JNI /////
-The respondents were presented to a piece of code on which a non-declared checked exception has scaped, and were asked if they new any reason for it to happen.
-Only 3 respondents out of 71 (4\% of respondents) were aware of the fact that there are ways a method can throw a non-declared checked exception, such as JNI/native code, relfection, and directly changing the bytecode/dalvik code which bypass the compile-time checking.
-Most of the repondents were not aware that a method could throw a checked exception withou declaring it on the method's signature.
+The respondents were presented with a piece of code in which a non-declared checked exception has appeared, and they were asked if they knew any reason for this to happen.
+Only 3 respondents out of 71 (4\% of respondents) were aware of the fact that there are ways a method can throw a non-declared checked exception, such as JNI/native code, reflection, and directly changing the bytecode/dalvik code which bypass the compile-time checking.
+Most of the respondents were not aware that a method could throw a checked exception without declaring it in the method's signature.
 Most respondents mentioned that if a method throws a non-declared exception it must be a RuntimeException.
-We could observe that this exception handling bug hazard detected during the mining study, although represent a threat to app robustness, is hardly known by the Android developers involved in this survey.
-Below we can find a quote from one of the 3 respondents aware of JNI non-delcared checked exceptions.
+We could observe that this exception handling bug hazard detected during the mining study, although representing a threat to app robustness, is hardly known by the Android developers involved in this survey.
+Below we can find a quote from one of the 3 respondents aware of JNI non-declared checked exceptions.
 
@@
 
-\noindent\emph{\textbf{RQ5: How the exception handling code affect the development of robust apps and how developers prevent apps from crashing?}}
+\noindent\emph{\textbf{RQ5: How does the exception handling code affect the development of robust apps and how do developers prevent apps from crashing?}}
 
@@
 Developers were also asked whether the exception handling code helped with the development of robust Android applications.
-Although most of the developers answered positively (78.9\%) when asked to provide one or more reasons for their answers, most of the respondents also provide reasons why the exception handling code can sometimes impairs the robustness.
-Since positive and negative reasons were provided intermixed we analyzed both set of answers and ranked the top reasons why exception handling that can negatively or positively affect robustness according to developers perceptions.
-Table ~\ref{tab:topreasons} presents the top reasons why the exception handling code may or may not help de development of robust apps.
+Although most of the developers answered positively (78.9\%), when asked to provide one or more reasons for their answers, most of the respondents also provide reasons why the exception handling code can sometimes impair the robustness.
+Since positive and negative reasons were provided intermixed we analyzed both sets of answers and ranked the top reasons why exception handling can negatively or positively affect robustness according to developers' perceptions.
+Table~\ref{tab:topreasons} presents the top reasons why the exception handling code may or may not help thee development of robust apps.
 
@@
 
-Although most of the respondents answered that exception handling code helped with the development of robust Android applications, almost every respondent pointed a drawback associated to he exception handling code, saying that if not used with care the exception handling code may lead to app crashes.
+Although most of the respondents answered that exception handling code helped with the development of robust Android applications, almost every respondent pointed out a drawback associated with the exception handling code, saying that if it is not used with care the exception handling code may lead to app crashes.
 
@@
 Developers were asked to rank the main causes of crashes.
-Approximately 68\% of the repondents ranked the exceptions signaled by programming logic mistakes e.g.
-NullPointerExceptions as the Top 1 and Top 2 cause of crashes as illustrated in Figure ~\ref{figure:ranking}.
-In our mining study the exceptions signaled due to programming logic mistakes (java.lang and util) where the causes of 52,0\% of the exception stack traces found.
+Approximately 68\% of the respondents ranked the exceptions signaled by programming logic mistakes, e.g.
+NullPointerExceptions, as the Top 1 and Top 2 cause of crashes, as illustrated in Figure~\ref{figure:ranking}.
+In our mining study the exceptions signaled due to programming logic mistakes (java.lang and util) where the causes of 52.0\% of the exception stack traces found.
 Hence, the developers perspective is aligned with what was revealed in mining study.
@@
 Most of the respondents answered that they prevent apps from crashing by handling exceptions/errors (27\%).
-They mentioned about some strategies such as: handling exceptions in all entry points/all over; using catch-all clauses; and adding try catch blocks around risky methods.
-Different from conventional Java programs, an Android app is composed by several entry points, each window (i.e.
-Activity) is a potential entry point on which exceptions can scape leading to an app crash.
-Hence the importance of handling exceptions in all entry points, and using general cash clauses which cannot sometimes prevent crash but allows the developer to present a detailed error message to the user before crashing.
-Many respondents also said that they used testing approaches to prevent crashes -  few of them mentioned crowd testing (3\%) an emerging approach in testing apps.
+They mentioned strategies such as: handling exceptions in all entry points/all over; using catch-all clauses; and adding try-catch blocks around risky methods.
+Different from conventional Java programs, an Android app is composed of several entry points, each window (i.e.
+Activity) is a potential entry point in which exceptions can arise leading to an app crash.
+This explains the importance of handling exceptions in all entry points, and using general crash clauses which sometimes cannot prevent crashes but allow the developer to present a detailed error message to the user before crashing.
+Many respondents also said that they used testing approaches to prevent crashes -- few of them mentioned crowd testing (3\%), an emerging approach in testing apps.
 Moreover, null checks were also mentioned by 15\% of the developers as a way to prevent crashes.
-The adoption of coding styles were also pointed out by 14\% of respondents as a way to prevent craches.
-Finally, the crash report tools (e.g.
+The adoption of coding styles was also pointed out by 14\% of respondents as a way to prevent crashes.
+Finally, crash report tools (e.g.
 ACRA [ref]) were reported as a way to prevent crashes.
 Once such tools notify developers about current crashes, the application can be fixed and hence future crashes (caused by what was fixed) could be prevented.
-Table ~ref{tab:topprevent} presents the themes for the strategies for crash prevention mentioned by developers.
+Table~ref{tab:topprevent} presents the themes for the strategies for crash prevention mentioned by developers.
 
@@
 
-\greybox{You can wrap the main function in java in try/catch in android you can\'t. }
+\greybox{You can wrap the main function in java in try/catch in android you can't. }
 

### 2015-10-08 SPRINGER_JOURNAL/msr_journal.tex
coders sentence
including research questions on the mining study - following Christoph's suggestions

@@
 
-The goal of our mining study is to explore to what extent exception stack traces 
-contained in issues of Android projects (hosted in GitHub and Google Code), can reveal 
-\emph{bug hazards} in the exception-related code of both the applications and the underlying framework. 
+Our mining study was guided by a general research question: \emph{RQ 1: Can the information available in exception stack traces 
+ reveal exception handling \emph{bug hazards} in both the Android applications and framework?}
 As mentioned before, in this context \emph{bug hazards} are the characteristics of exception-related code 
 that favor the introduction of failures such as uncaught exceptions.
-To support our investigation, we developed a tool called ExceptionMiner (Section~\ref{sec:exceptionminer})
+
+Our study focused on the exception stack traces contained in issues of Android projects (hosted in GitHub and Google Code).
+And to support our investigation, we developed a tool called ExceptionMiner (Section~\ref{sec:exceptionminer})
 which extracts the exception stack traces embedded on issues, 
@@
 
-This section presents the results of the study; this presentation is centered around 
-the information distilled from stack traces: (i) the root exceptions (i.e., the exceptions that caused the stack traces);
+This section presents the results of our study that was guided by the general research question:
+ \emph{RQ 1: Can the information available in exception stack traces reveal exception handling 
+bug hazards in both the Android applications and framework?}
+To make the analysis easier, we further refine this question in subquestions, each one focusing on pieces of information
+ distilled from stack traces, more specifically: (i) the root exceptions (i.e., the exceptions that caused the stack traces);
  (ii) the exception types (i.e, Checked, Runtime, Error, Throwable) and (iii) the exception wrappings. 
-Each piece information is then analyzed in detail to 
+Hence, this section is centered around the following subquestions:  \emph{RQ 1.1: Can the root exceptions reveal bug harzards?}; 
+ \emph{RQ 1.2 Can the exception types reveal bug hazards?}; and  \emph{RQ 1.3 Can the exception wrappings reveal bug hazards?}.
+
+In this section each piece information is then analyzed in detail to 
 check whether they can reveal bug harzards on the exception handling code -
@@
 
-\subsubsection{Can the root exceptions reveal bug harzards?}
+\bigskip
+
+%\subsubsection{RQ 1.1 Can the root exceptions reveal bug harzards?}
+\noindent\emph{RQ 1.1 Can the root exceptions reveal bug harzards?}
+
+\bigskip
+
 After distilling the information available on the exception stack traces, we could find 
@@
 
-\subsubsection{Can the exception types reveal bug hazards?}
+\bigskip
+
+%\subsubsection{Can the exception types reveal bug hazards?}
+\noindent\emph{RQ 1.2 Can the exception types reveal bug hazards?}
 %\subsection{Exploring the Information Embedded on Exception Types}
 %\subsection{The Exception Types and What They can Tell us? }
+
+\bigskip
 
@@
 
-\subsubsection{Can the exception wrappings reveal bug hazards?}
+%\subsubsection{Can the exception wrappings reveal bug hazards?}
+\noindent\emph{RQ 1.3 Can the exception wrappings reveal bug hazards?}
+
+\bigskip
 
@@
 
-When developing Java-based applications it is inevitable to deal with exceptions. Consequently, our first question explores how developers deal with the exception handling code in Android development:
-
-\emph{\textbf{RQ 1:} How often do developers deal with exception handling code while developing Android apps?}
-
-To make the analysis easier, we further refine RQ1 in subquestions, such as: \emph{How often do developers handle exceptions?}; \emph{How often do developers throw exceptions?}.
+The general research question guided our exploratory survey was the following: \emph{RQ 2: How developers deal with the exception handling code in Android apps and what are the developers' perspectives about the main bug hazards found during the mining study?} This general research question was broken in a set of research questions that are answered by the exploratory survey.
+
+When developing Java-based applications it is inevitable to deal with exceptions. Hence, our first question explores how developers deal with the exception handling code in Android development:
+
+\emph{\textbf{RQ 2.1:} How do developers deal with exception handling code while developing Android apps?}
+
+To make the analysis easier, we further refine this question in subquestions, such as: \emph{How often do developers handle exceptions?}; \emph{How often do developers throw exceptions?}.
 Moreover, another question that shows us how developers may deal with the EH code is: \emph{Do developers know about Java EH best practices and/or Android specific EH best practices?}
@@
 
-\emph{\textbf{RQ 2:} How do NullPointerExceptions impact the development of robust Android apps?}
-
-\emph{\textbf{RQ 3:} How do cross-type wrappings impact the development of robust Android apps?}
-
-\emph{\textbf{RQ 4:} Are developers aware of the robustness threats caused by JNI undocumented checked exceptions?}
-
-%\emph{\textbf{RQ 4:}  Are developers aware of the non-declared checked exceptions?}
+\emph{\textbf{RQ 2.2:} How do NullPointerExceptions impact the development of robust Android apps?}
+
+\emph{\textbf{RQ 2.3:} How do cross-type wrappings impact the development of robust Android apps?}
+
+\emph{\textbf{RQ 2.4:} Are developers aware of the robustness threats caused by JNI undocumented checked exceptions?}
+
+%\emph{\textbf{RQ 2.5:}  Are developers aware of the non-declared checked exceptions?}
 
@@
 
-\emph{\textbf{RQ 5:} How does the exception handling code affect the development of robust apps and how do developers prevent apps from crashing?}
+\emph{\textbf{RQ 2.5:} How does the exception handling code affect the development of robust apps and how do developers prevent apps from crashing?}
 
@@
 The grounded theory coding used in this study consists of two phases: (1) initial coding entails a close reading of the data and (2) later we use focused coding ~\cite{charmaz2006} to pinpoint and develop the most salient themes~\cite{charmaz2006} in the analyzed data.
+The coding process was performed by two coders who agreed upon the list of themes that emerged from the answers'set of each question.
 
@@
 \bigskip
-\noindent\emph{\textbf{Q1:How do developers deal with the exception handling in Android development?}}
+\noindent\emph{RQ 2.1: How do developers deal with the exception handling in Android development?}
 \bigskip
@@
 
-\noindent\emph{\textbf{Q 2: How do the NullPointerExceptions impact the development of robust Android apps?}}
+\noindent\emph{\textbf{RQ 2.2: How do the NullPointerExceptions impact the development of robust Android apps?}}
 
@@
 
-\noindent\emph{\textbf{RQ 3: How do Cross-Type Wrappings impact the development of robust Android apps?}}
+\noindent\emph{\textbf{RQ 2.3: How do Cross-Type Wrappings impact the development of robust Android apps?}}
 
@@
 \bigskip
-\noindent\emph{\textbf{RQ 4: Are developers aware of the robustness threats caused by JNI undocumented checked exceptions?}}
+\noindent\emph{\textbf{RQ 2.4: Are developers aware of the robustness threats caused by JNI undocumented checked exceptions?}}
 \bigskip
@@
 
-\noindent\emph{\textbf{RQ5: How does the exception handling code affect the development of robust apps and how do developers prevent apps from crashing?}}
+\noindent\emph{\textbf{RQ 2.5: How does the exception handling code affect the development of robust apps and how do developers prevent apps from crashing?}}
 

### 2015-10-08 SPRINGER_JOURNAL/msr_journal.tex
coders sentence

@@
 
-\noindent\emph{\textbf{RQ 2.2: How do the NullPointerExceptions impact the development of robust Android apps?}}
+\noindent\emph{RQ 2.2: How do the NullPointerExceptions impact the development of robust Android apps?}
 
@@
 \bigskip 
-
-\noindent\emph{\textbf{RQ 2.3: How do Cross-Type Wrappings impact the development of robust Android apps?}}
+%\textbf{
+\noindent\emph{RQ 2.3: How do Cross-Type Wrappings impact the development of robust Android apps?}
 
@@
 \bigskip
-\noindent\emph{\textbf{RQ 2.4: Are developers aware of the robustness threats caused by JNI undocumented checked exceptions?}}
+\noindent\emph{RQ 2.4: Are developers aware of the robustness threats caused by JNI undocumented checked exceptions?}
 \bigskip
@@
 
-%\noindent\emph{\textbf{Q2: Do developers know about exception handling best practices?}}
+%\noindent\emph{Q2: Do developers know about exception handling best practices?}
 %We can observe that 62\% of surveyed Android developers follow Java exception handling best 
@@
 
-\noindent\emph{\textbf{RQ 2.5: How does the exception handling code affect the development of robust apps and how do developers prevent apps from crashing?}}
+\noindent\emph{RQ 2.5: How does the exception handling code affect the development of robust apps and how do developers prevent apps from crashing?}
 

### 2015-10-08 SPRINGER_JOURNAL/msr_journal.tex
including more quotes

@@
 
-In this section each piece information is then analyzed in detail to 
+In this section each piece information is analyzed in detail to 
 check whether they can reveal bug harzards on the exception handling code -
@@
 In this section, we present our findings per research question. 
-We also include a selection of quotes from the exploratory survey 
-respondents. Each respondent has an identification which can be traced in our database.
+To illustrate the different aspects of each
+finding, we provide a selection of quotes from the exploratory
+survey. 
+To enable traceability, each respondent has an identification
+ which can be traced in our database using the following convention: 
+D\#. For instance, D1 corresponds to the answer provided
+ by developer 1.
 
@@
 
-\bigskip
+%\bigskip
 
@@
 
+``Don't throw RuntimeException (in most cases), don't catch OutOfMemoryException, etc." [D42]
+
+``If a method signature declares that it throws an exception, it means that the caller should handle the exception. [D29]
+
 \begin{table}
@@
 
-\bigskip
+%\bigskip
+
+\greybox{``They [NullPointerExceptions] can happen pretty much anywhere.
+The Android Fragment system comes to mind in this case.
+Often, it is possible to find yourself in a state where getActivity() is null within the Fragment during certain points in the life cycle, and that is something I have to plan for.
+This might have been avoidable under a different structure."[D56]}
+
+%\bigskip
 
@@
 
-\greybox{They can happen pretty much anywhere.	 The Android Fragment system comes to mind in this case.
-Often, it is possible to find yourself in a state where getActivity() is null within the Fragment during certain points in the life cycle, and that is something I have to plan for.
-This might have been avoidable under a different structure.}
-
-\greybox{The Android framework is what adds the complexity in figuring out what caused an exception because more often than not, the error is triggered from the framework as a result of something else you did.}
+%``android"
 
@@
 
-\greybox{Honestly, there are a lot of very unskilled Java programmers out there writing Android apps.
-When they encounter NPE, they tend to null-check that variable, which just puts a bandage on the problem and causes other failures (usually also NPE's) later on in the application's lifecycle.}
-
-%\greybox{This is nothing specific to Android, it's just how Java works.}
+\greybox{``Honestly, there are a lot of very unskilled Java programmers out there writing Android apps.
+When they encounter NPE, they tend to null-check that variable, which just puts a bandage on the problem and causes other failures (usually also NPE's) later on in the application's lifecycle.
+This is nothing specific to Android, it's just how Java works."[D42]}
+
+\greybox{``The Android framework is what adds the complexity in figuring out what caused an exception because more often than not, the error is triggered from the framework as a result of something else you did." [D58]}
+
+\greybox{``Android destroys and recreates itself all of the time ( especially during screen rotation ) and if you do not handle that it will crash on you every time.
+with the complexity of an activity with a fragment that has fragments and each of those fragments has custom objects and variables that need to be either retained ( so saved and put back ) or recreated such as views it can get complex if you don't have an understanding of how the android life cycle works." [D43]}
 
@@
 
-\greybox{Determine why the object was null and attempt to fix this situation.
+\greybox{``Determine why the object was null and attempt to fix this situation.
 In the case of external API calls which return null, then check for null (the quick-and-dirty way).
-For internal calls, use @NotNull and @Nullable annotations to provided more guidance on when an object \"may be\" and \"should never be\" null.}
-
-\greybox{Probably the most common exception, as such the reasons and fixes are super varying.
+For internal calls, use @NotNull and @Nullable annotations to provided more guidance on when an object \"may be\" and \"should never be\" null." [D42]}
+
+\greybox{``Probably the most common exception, as such the reasons and fixes are super varying.
 But in general, null checks should be utilized and illegal state or argument exceptions should be used with an appropriate message.
-Which will communicate useful information without the confusion an NPE would normally cause.}
+Which will communicate useful information without the confusion an NPE would normally cause."[D48]}
 
@@
 
+``RuntimeException-s are not particularly well suited for Android's robustness, as any uncaught one would cause a crash." [D26]
+
+``this treats all exceptions as critical failure, even though they might not all be unrecoverable" [D70]
+
+``Its very important in android that we gracefully handle an exception. Try not to crash an application."[41]
+
+``Rethrowing a checked exception as an unchecked exception is the worst choice; it throws away any value of checked exceptions, elevates the exception out of almost all handling, and does absolutely nothing helpful in the process.
+Further, catching `Exception` suggests that the author doesn't know what exceptions might occur and wrongfully assumed that this solution is somehow safer than doing nothing."[D19]
+
+``Giving the user meaningful feed back is important, so the more specific you can be about what went wrong, the better."[D66]
+
+``because it is checked exception, you should expect it to happen rather than just crash"[D70]
+
 Some of the best practices are related to the different ways an exception should be handled according to whether it is a checked and an unchecked exception.
@@
 
-\greybox{I am guessing this is because the exception is thrown from native code (ie, C++ code in the JVM) where Java correct-ness semantics rules do not always apply.}
+\greybox{``I am guessing this is because the exception is thrown from native code (ie, C++ code in the JVM) where Java correct-ness semantics rules do not always apply." [D42]}
 
@@
 
-\greybox{Any uncatched exception will crash the app.
-There are many unknown exception thrown by the system, that are only happen once in a lifetime like IllegalStateException: eglMakeCurrent failed EGL\_BAD\_CONTEXT}
-
-\greybox{The lifecycle of activities\/fragements etc make it harder to work out what order things are called in and so exceptions can occur because you didn't realise that another method isn't called.}
+\greybox{``Any uncatched exception will crash the app.
+There are many unknown exception thrown by the system, that are only happen once in a lifetime like IllegalStateException: eglMakeCurrent failed EGL\_BAD\_CONTEXT"[D32]}
+
+\greybox{``The lifecycle of activities\/fragements etc make it harder to work out what order things are called in and so exceptions can occur because you didn't realise that another method isn't called."[D70]}
 
@@
 
-\greybox{Though not particular to Android, most Android apps have several different entry points: Activities, Services, etc, can be started by different services; same thing with events.
-It may be hard to ensure that all this cases are properly handled.
-}
-
-\greybox{You can wrap the main function in java in try/catch in android you can't. }
+\greybox{``Though not particular to Android, most Android apps have several different entry points: Activities, Services, etc, can be started by different services; same thing with events.
+It may be hard to ensure that all this cases are properly handled." [D46]}
+
+\greybox{``You can wrap the main function in java in try/catch in android you can't..."[D12]}
 

### 2015-10-08 SPRINGER_JOURNAL/msr_journal.tex
including more quotes re-writting text

@@
 use Java exceptions.
-Section XXXXX compiled a list of some exception handling (EH) best practices that guided our mining study.
+Section~\ref{sec:best}  compiled a list of some exception handling (EH) best practices that guided our mining study.
 In this survey the developers were also asked if they knew about and used Java EH best practices.
 Most of them said that they knew and adopted Java EH practices (68\%).
-From this set the most mentioned best practices are presented in Table XXXX -- considering that one developer mentioned more than one practice.
-Some of the practices mentioned, however, are not well known Java EH best practices but seem to be used by some Android developers to better deal with exceptions in the Android app context: the crash fast approach [ref], the use of crash report tools, and to favor checked exceptions instead of runtime to represent the exceptional conditions in Android apps.
-
-``Don't throw RuntimeException (in most cases), don't catch OutOfMemoryException, etc." [D42]
-
-``If a method signature declares that it throws an exception, it means that the caller should handle the exception. [D29]
+From this set the most mentioned best practices are presented in Table~\ref{tab:javapractices} -- considering that one developer mentioned more than one practice.
+Most of the best practices mentioned were the ones cited in Section~\ref{sec:best}.
+For instance:\emph{``Don't throw [pure] RuntimeException (in most cases), don't catch OutOfMemoryError, etc." [D42]}.
+Some of the practices mentioned, however, are not well known Java EH best practices but seem to be used by some Android developers to better deal with exceptions in the Android app context, such as the use of crash fast approach~\footnote{http://www.slideshare.net/pyricau/crash} -- \emph{``Crash is good for hinting developer.
+Fast crash fast solve." [D51]}; and to favor checked exceptions instead of runtime to represent the exceptional conditions in Android apps -- \emph{``RuntimeException-s are not particularly well suited for Android's robustness, as any uncaught one would cause a crash." [D26]}.
+
+%``Don't throw [pure] RuntimeException (in most cases), don't catch OutOfMemoryError, etc." [D42]
 
@@
 Most of them (43\%) mentioned that they applied the same best practices that they used in Java programs.
-Only some of them mentioned that they adopted specific best practices such as: (i) the use of crash report tools (21\%) -- to notify the developer about the uncaught exceptions that happen in the application code; and (ii) the use of a global exception handler (14\%) -- the UncaughtExceptionHandler advised in Android tutorials [ref].
+Only some of them mentioned that they adopted specific best practices such as: (i) the use of crash report tools (21\%) -- to notify the developer about the uncaught exceptions that happen in the application code; and (ii) the use of a global exception handler (14\%) -- the UncaughtExceptionHandler advised in Android documentation~\footnote{http://developer.android.com/reference/java/lang/Thread.UncaughtExceptionHandler.html}.
 
@@
 
+\finding{
+Most of the respondents (64.8\%) recognize that handle exceptions most of the time during the app development. 
+Most of them (62\%) mentioned that adopt EH best practices.
+}
+
+\bigskip
+
 \noindent\emph{RQ 2.2: How do the NullPointerExceptions impact the development of robust Android apps?}
 
-%\bigskip
-
-\greybox{``They [NullPointerExceptions] can happen pretty much anywhere.
-The Android Fragment system comes to mind in this case.
-Often, it is possible to find yourself in a state where getActivity() is null within the Fragment during certain points in the life cycle, and that is something I have to plan for.
-This might have been avoidable under a different structure."[D56]}
-
-%\bigskip
-
 %//////NULLPONTER WORSE IN ANDROID /////
-Developers were asked if their apps ever crashed because of a NullPointerException. The vast majority of respondents (96\%) said that their apps crashed at least once due to a NullPointerException. 
+Developers were asked if their apps ever crashed because of a NullPointerException.
+The vast majority of respondents (96\%) said that their apps crashed at least once due to a NullPointerException.
 Their answers are aligned with the high prevalence of NullPointerExceptions found in the mining study -- in Table XXX we can see that NullPointerExceptions were the main cause of the mined exception stack traces.
@@
 Almost half of them (XXX\%) mentioned that Android apps were more susceptible to NullPointerException crashes.
-Some of the reasons mentioned by developers were: (i) the activity and fragment lifecycle -- since both are constantly destroyed their references can frequently become null, leading to NullPointerExceptions; (ii) the framework complexity -- several methods of the Android framework return null, and there is not enough documentation to inform the developer when using such methods.
+
+Some of  reasons were mentioned by developers such as: the activity and fragment lifecycle -- \emph{``Android destroys and recreates itself all of the time (especially during screen rotation) and if you do not handle that it will crash on you every time.
+with the complexity of an activity with a fragment that has fragments and each of those fragments has custom objects and variables that need to be either retained (so saved and put back) or recreated such as views it can get complex if you don't have an understanding of how the android life cycle works." [D43]}.
+Another developer also mentioned that: \emph{``They [NullPointerExceptions] can happen pretty much anywhere.
+The Android Fragment system comes to mind in this case.
+Often, it is possible to find yourself in a state where getActivity() is null within the Fragment during certain points in the life cycle, and that is something I have to plan for.
+This might have been avoidable under a different structure."[D56]}.
+
+The framework complexity was also mentioned as a major cause of null pointer crashes -- since several methods of the Android framework return null, and there is not enough documentation to inform the developer when using such methods.
+Some respondents mentioned: \emph{``The Android framework is what adds the complexity in figuring out what caused an exception because more often than not, the error is triggered from the framework as a result of something else you did." [D58]}; \emph{``Java is simple language, and the problem arises when you are using third party frameworks build on Java (or using Android which provide lots of new classes).
+And the problem is you don't always know if specific method can return null object or only valid values."[D7]}.
 
@@
 
-Moreover, a few respondents also answered that such NullPointerException crashes can be caused by backward compatibility as well as memory issues as illustrated in Table XXXX.
+Moreover, a few respondents also answered that such NullPointerException crashes can be caused by backward compatibility as well as memory issues as illustrated in Table~\ref{tab:causenull}.
 The other half mentioned that NullPointerExceptions are common in Android development since they are common in Java standard development.
-
-\greybox{``Honestly, there are a lot of very unskilled Java programmers out there writing Android apps.
+One responded even mentioned: \emph{``Honestly, there are a lot of very unskilled Java programmers out there writing Android apps.
 When they encounter NPE, they tend to null-check that variable, which just puts a bandage on the problem and causes other failures (usually also NPE's) later on in the application's lifecycle.
 This is nothing specific to Android, it's just how Java works."[D42]}
-
-\greybox{``The Android framework is what adds the complexity in figuring out what caused an exception because more often than not, the error is triggered from the framework as a result of something else you did." [D58]}
-
-\greybox{``Android destroys and recreates itself all of the time ( especially during screen rotation ) and if you do not handle that it will crash on you every time.
-with the complexity of an activity with a fragment that has fragments and each of those fragments has custom objects and variables that need to be either retained ( so saved and put back ) or recreated such as views it can get complex if you don't have an understanding of how the android life cycle works." [D43]}
 
@@
 \caption{Top reasons why NullPointerExceptions are more frequent in Android apps acording to 19 respondents . }
-\label{tab:handlingruntime}
+\label{tab:causenull}
 \end{table}
@@
 \caption{Top ways for preventing NullPointerException in Android apps - 61 non-empty responses. }
-\label{tab:handlingruntime}
+\label{tab:preventnull}
 \end{table}
@@
 \bigskip
-
-``RuntimeException-s are not particularly well suited for Android's robustness, as any uncaught one would cause a crash." [D26]
 

### 2015-10-08 SPRINGER_JOURNAL/msr_journal.tex
including more quotes re-writting text

@@
 Many respondents (39\%) also answered that instead of just sprinkling the code with null-checks when a NullPointerException happens they investigate the cause of the null reference and work to fix it, instead of just adding a null check.
-Some respondents also mentioned using annotations such as @Nullable @NonNull, provided in Android since version XXX; such annotations guide the developer about what can and cannot be null during the application lifecycle.
+\emph{Determine why the object was null and attempt to fix this situation.
+In the case of external API calls which return null, then check for null (the quick-and-dirty way).
+For internal calls, use @NotNull and @Nullable annotations to provided more guidance on when an object \"may be\" and \"should never be\" null." [D42]}
+
+Some respondents also mentioned using annotations such as @Nullable @NonNull, provided by Android Studio 0.5.5; such annotations guide the developer about what can and cannot be null during the application lifecycle.
 Although catching the NullPointerExceptions is considered a bad practice some respondents said that they do it to prevent crashes caused by NullPointerExceptions.
@@
 Some respondents also mentioned to replace the NullPointerException by IllegalArgumentException when a method should not receive a null as parameter.
-This is a good exception handling practice mentioned by Bloch [ref].
+\emph{``Probably the most common exception, as such the reasons and fixes are super varying.
+But in general, null checks should be utilized and illegal state or argument exceptions should be used with an appropriate message.
+Which will communicate useful information without the confusion an NPE would normally cause."[D48]}.This is a good exception handling practice mentioned by Bloch [ref].
 Only a few developers mentioned coding standards and static code analyzers to prevent it (2\%).
-
-\greybox{``Determine why the object was null and attempt to fix this situation.
-In the case of external API calls which return null, then check for null (the quick-and-dirty way).
-For internal calls, use @NotNull and @Nullable annotations to provided more guidance on when an object \"may be\" and \"should never be\" null." [D42]}
-
-\greybox{``Probably the most common exception, as such the reasons and fixes are super varying.
-But in general, null checks should be utilized and illegal state or argument exceptions should be used with an appropriate message.
-Which will communicate useful information without the confusion an NPE would normally cause."[D48]}
 

### 2015-10-08 SPRINGER_JOURNAL/msr_journal.tex
including more quotes re-writting text

@@
 Some of the practices mentioned, however, are not well known Java EH best practices but seem to be used by some Android developers to better deal with exceptions in the Android app context, such as the use of crash fast approach~\footnote{http://www.slideshare.net/pyricau/crash} -- \emph{``Crash is good for hinting developer.
-Fast crash fast solve." [D51]}; and to favor checked exceptions instead of runtime to represent the exceptional conditions in Android apps -- \emph{``RuntimeException-s are not particularly well suited for Android's robustness, as any uncaught one would cause a crash." [D26]}.
-
-%``Don't throw [pure] RuntimeException (in most cases), don't catch OutOfMemoryError, etc." [D42]
+Fast crash fast solve." [D51]}; and to favor checked exceptions instead of runtime to represent the exceptional conditions in Android apps -- \emph{``RuntimeException-s are not particularly well suited for Android's robustness, as any uncaught one would cause a crash." [D26]}; \emph{``Checked exception help, unchecked exceptions don't." [D25]}; \emph{``avoid unchecked exceptions except if the case is really abnormal and should not happen [...]"[D45]}.
 
@@
 \end{tabular}
-\caption{Top mentioned Java EH best practices -  40 non-empty answers. }
+\caption{Top mentioned Java EH best practices -  40 non-empty responses. }
 \label{tab:javapractices}
@@
 The vast majority of respondents (96\%) said that their apps crashed at least once due to a NullPointerException.
-Their answers are aligned with the high prevalence of NullPointerExceptions found in the mining study -- in Table XXX we can see that NullPointerExceptions were the main cause of the mined exception stack traces.
-The developers were also asked whether crashes caused by NullPointerExceptions were more common in Android development than in standard Java development.
-Almost half of them (XXX\%) mentioned that Android apps were more susceptible to NullPointerException crashes.
-
-Some of  reasons were mentioned by developers such as: the activity and fragment lifecycle -- \emph{``Android destroys and recreates itself all of the time (especially during screen rotation) and if you do not handle that it will crash on you every time.
+Their answers are aligned with the high prevalence of NullPointerExceptions found in the mining study -- in Table~\ref{tab:topten} in Section~\ref{sec:result} we can see that NullPointerExceptions were the main cause of the mined exception stack traces.
+Developers reported some reasons for the hight prevelance of crashes caused by NullPointerExceptions.
+The activity and fragment lifecycle was one of them -- \emph{``Android destroys and recreates itself all of the time (especially during screen rotation) and if you do not handle that it will crash on you every time.
 with the complexity of an activity with a fragment that has fragments and each of those fragments has custom objects and variables that need to be either retained (so saved and put back) or recreated such as views it can get complex if you don't have an understanding of how the android life cycle works." [D43]}.
@@
 And the problem is you don't always know if specific method can return null object or only valid values."[D7]}.
-
-%``android"
-
 Moreover, a few respondents also answered that such NullPointerException crashes can be caused by backward compatibility as well as memory issues as illustrated in Table~\ref{tab:causenull}.
-The other half mentioned that NullPointerExceptions are common in Android development since they are common in Java standard development.
+
+Almost half of the developers mentioned, however, that NullPointerExceptions are common in Android development since they are common in Java standard development as well.
 One responded even mentioned: \emph{``Honestly, there are a lot of very unskilled Java programmers out there writing Android apps.
@@
 Such behavior was mentioned by one of the respondents as the behavior adopted by non-experienced developers  to prevent such crashes -- which is undoubtedly a mistake.
-Some respondents also mentioned to replace the NullPointerException by IllegalArgumentException when a method should not receive a null as parameter.
+Some respondents also suggested to replace the NullPointerException by IllegalArgumentException when a method should not receive a null as parameter.
 \emph{``Probably the most common exception, as such the reasons and fixes are super varying.
 But in general, null checks should be utilized and illegal state or argument exceptions should be used with an appropriate message.
-Which will communicate useful information without the confusion an NPE would normally cause."[D48]}.This is a good exception handling practice mentioned by Bloch [ref].
+Which will communicate useful information without the confusion an NPE would normally cause."[D48]}.
+This is a good exception handling practice mentioned by Bloch~\cite{bloch2008effective}.
 Only a few developers mentioned coding standards and static code analyzers to prevent it (2\%).
@@
 automated testing	& 1 &	2\% \\
-make sure value is not null	& 1 &	2\% \\
-coding standards	& 1 &	2\% \\
-guard statements 	& 1 &	2\% \\
-handlers in all entry points	& 1 & 	2\% \\
-Use ART more than Dalvik	& 1 &	2\% \\
-Use IllegalArgument or IllegaState exception instead	& 1 &	2\% \\
-Verify the state of activity/fragment	& 1 &	2\% \\
-Write code in Scala, without nulls (joke)	& 1 &	2\% \\
+%make sure value is not null	& 1 &	2\% \\
+%coding standards	& 1 &	2\% \\
+%guard statements 	& 1 &	2\% \\
+%handlers in all entry points	& 1 & 	2\% \\
+%Use ART more than Dalvik	& 1 &	2\% \\
+%Use IllegalArgument or IllegaState exception instead	& 1 &	2\% \\
+%Verify the state of activity/fragment	& 1 &	2\% \\
+%Write code in Scala, without nulls (joke)	& 1 &	2\% \\
 \hline
@@
 \bigskip
+
+\finding{Almost 96\% of the developers recognized that their apps have already crashed due to a NullPointerException.
+They mentioned that Android application life-cycle (47\%) and the framework complexity (32\%) may favor such uncaught NullPointerExceptions to happen in Android context.
+}
+
+\bigskip 
 %\textbf{
@@
 
-``this treats all exceptions as critical failure, even though they might not all be unrecoverable" [D70]
-
-``Its very important in android that we gracefully handle an exception. Try not to crash an application."[41]
-
-``Rethrowing a checked exception as an unchecked exception is the worst choice; it throws away any value of checked exceptions, elevates the exception out of almost all handling, and does absolutely nothing helpful in the process.
-Further, catching `Exception` suggests that the author doesn't know what exceptions might occur and wrongfully assumed that this solution is somehow safer than doing nothing."[D19]
-
-``Giving the user meaningful feed back is important, so the more specific you can be about what went wrong, the better."[D66]
-
-``because it is checked exception, you should expect it to happen rather than just crash"[D70]
-
 Some of the best practices are related to the different ways an exception should be handled according to whether it is a checked and an unchecked exception.
@@
 
-%///// HANDLING RUNTIME AND CHECKED /////		
 To assess the developers' perspective on whether checked and unchecked exceptions should be handled differently inside the app, the developers were presented with two pieces of code: one in which a RuntimeException was caught and another one where an IOException was caught; both in the context of Activity life-cycle methods.
@@
 When asked about how to deal with a checked exception signaled by a method (invoked in the context of the onPause() activity method), most respondents (62.50\%) said that they should add a try-catch block surrounding the method invocation.
-Also, most of the respondents said that to deal with a method that is signaling a runtime exception we should surround such a method with a try-catch block (xxxxx\%).
+One of the respondents mentioned that \emph{``If a method signature declares that it throws an exception, it means that the caller should handle the exception."[D29]}; \emph{``because it is checked exception, you should expect it to happen rather than just crash"[D70]}.
+However, most of the respondents also said that to deal with a method that is signaling a runtime exception we should surround such a method with a try-catch block (59\%).
+
 However, what developers mentioned as handling actions (what is performed inside the catch clause to deal with the exception) differs in both cases.
-
-One of the respondents mentioned that when the designer of an API decides to signal a checked exception s/he is telling the developer that the exception should be handled.
-Hence, some handling actions mentioned for checked exceptions were: (i) retrying the same operation; (ii) involving the user in finding a solution to the exception condition such as opening a pop-up and asking the user to define a new place to save the file (the example presented to them involved a IOException being signaled); and also (iii) presenting an error message to the user.
-
-On the other hand, the most common way of dealing with a runtime exception would be to catch it and present an error message to the user, or swallow the exception.
+For runtime exceptions most of the handling actions were dedicated to present more specific error messages and prevent the app from abruptly crashing.
+\emph{``[...] They signal exceptional behavior that may not be recoverable, so they offer a useful way to log and gracefully crash.} [D19].
+\emph{``Giving the user meaningful feed back is important, so the more specific you can be about what went wrong, the better."[D66]}.
 Some developers also mentioned to use a toast to support this task\footnote{A toast is an Android component that provides simple feedback about an operation in a small popup - http://developer.android.com/guide/topics/ui/notifiers/toasts.html}.
-Few developers mentioned that the runtime exception should flow upstream to the framework so it could gracefully crash the app.
-Although there is a long lasting debate about the pros and cons of checked and unchecked exceptions, most Android developers considered checked exceptions as a way of using exceptions that can prevent uncaught exception crashes -- since in order for a checked exception to flow upstream and crash the app the developers need to explicitly do so.
+Few developers also mentioned that the runtime exception should flow upstream to the framework so it could crash the app.
+
+On the other hand, some handling actions mentioned for checked exceptions were: (i) retrying the same operation; (ii) involving the user in finding a solution to the exception condition such as opening a pop-up and asking the user to define a new place to save the file (the example presented to them involved a IOException being signaled); and also (iii) presenting an error message to the user.
+The handling actions associated to checked exceptions focused more on crash prevention than the ones related to runtime exceptions.
+\emph{``Its very important in android that we gracefully handle an exception.
+Try not to crash an application."[41]}.
+It was also interesting to observe that the developers suggested involving the user and solving the exceptional condition represented by a checked exception - \emph{``We should try alternative saving ways and if it fails we should prompt the user for a new location and always keep the user informed"[D11]}.
+\emph{``You still need to catch an IOException to prevent a crash.
+However, because it is related to saving a file, you may need to reset any data within that catch statement and either alert the user that it has failed or perform a limited retry (in the case of an HTTP upload or something prone to server-side failure)."[D5]}.
+
+ Although there is a long lasting debate about the pros and cons of checked and unchecked exceptions, the survey revealed that most Android developers considered checked exceptions as a way of using exceptions that can prevent uncaught exception crashes -- since in order for a checked exception to flow upstream and crash the app the developers need to explicitly do so.
 
@@
 %/////// CROSS TYPE WRAPPING //////
-Hence, developers were presented with a code snippet in which a cross-type wrapping was performed\footnote{This cross-type wrapping was found in several applications during the mining study as well as in some classes of the Android framework.} -- as illustrated in Figure X.
-Then developers were asked whether such cross-type wrapping could affect the application robustness in some way.
-Table XXXX illustrates the  most mentioned reasons of why cross-type wrapping may affect app robustness; 18\% of the respondents mentioned that the app would crash anyway, regardless of whether the exception was wrapped or not.
-Most of the respondents, however, mentioned disadvantages of such wrapping, such as: (i) it impairs proper handling since wrapping in a general exception loses information about the exception situation to be handled (24\%); such wrapping treats all exceptions as critical -- in other words, every exception will remain uncaught and will lead to an app crash -- and it does not allow the developer to perform a proper handling (e.g.
-retry) for exception types that are not critical.
+After presenting two different code snippets on which a checked and a runtime exception were caught, developers were presented with a code snippet in which a cross-type wrapping was performed\footnote{This cross-type wrapping was found in several applications during the mining study as well as in some classes of the Android framework.} -- as illustrated below.
 
@@
 
+Then developers were asked whether such cross-type wrapping could affect the application robustness in some way.
+Table XXXX illustrates the  most mentioned reasons of why cross-type wrapping may affect app robustness; 18\% of the respondents mentioned that the app would crash anyway, regardless of whether the exception was wrapped or not.
+Most of the respondents, however, mentioned disadvantages of such wrapping, such as: (i) it impairs proper handling since wrapping in a general exception loses information about the exception situation to be handled (24\%); such wrapping treats all exceptions as critical -- in other words, every exception will remain uncaught and will lead to an app crash -- and it does not allow the developer to perform a proper handling (e.g.
+retry) for exception types that are not critical.
+
 \begin{table}
@@
 \end{table}
+
+``this treats all exceptions as critical failure, even though they might not all be unrecoverable" [D70]
+
+``Rethrowing a checked exception as an unchecked exception is the worst choice; it throws away any value of checked exceptions, elevates the exception out of almost all handling, and does absolutely nothing helpful in the process.
+Further, catching `Exception` suggests that the author doesn't know what exceptions might occur and wrongfully assumed that this solution is somehow safer than doing nothing."[D19]
 
@@
 \end{tabular}
-\caption{Top ways of preventing crashes - 71 non-empty respondents. }
+\caption{Top ways of preventing crashes - 71 non-empty responses. }
 \label{tab:topprevent}

### 2015-10-08 SPRINGER_JOURNAL/msr_journal.tex
including more quotes re-writting text

@@
 \begin{verbatim}
+
    try {
@@
 \hline
-\bfseries{Top Ways} & \bfseries{\#} & \bfseries{\%} \\
+\bfseries{Top Reasons} & \bfseries{\#} & \bfseries{\%} \\
 \hline
@@
 \hline
-\bfseries{Top Ways} & \bfseries{\#} & \bfseries{\%} \\
+\bfseries{Top Ways of Preventing NullPointerExceptions} & \bfseries{\#} & \bfseries{\%} \\
 \hline
@@
 On the other hand, some handling actions mentioned for checked exceptions were: (i) retrying the same operation; (ii) involving the user in finding a solution to the exception condition such as opening a pop-up and asking the user to define a new place to save the file (the example presented to them involved a IOException being signaled); and also (iii) presenting an error message to the user.
-The handling actions associated to checked exceptions focused more on crash prevention than the ones related to runtime exceptions.
+
+We can observe that the handling actions associated to checked exceptions focused more on crash prevention than the ones related to runtime exceptions.
 \emph{``Its very important in android that we gracefully handle an exception.
@@
 
- Although there is a long lasting debate about the pros and cons of checked and unchecked exceptions, the survey revealed that most Android developers considered checked exceptions as a way of using exceptions that can prevent uncaught exception crashes -- since in order for a checked exception to flow upstream and crash the app the developers need to explicitly do so.
+ Although there is a long lasting debate about the pros and cons of checked and unchecked exceptions, the survey revealed that many Android developers considered checked exceptions as a way of using exceptions that can prevent uncaught exception crashes -- since in order for a checked exception to flow upstream and crash the app the developers need to explicitly do so.
 
@@
 \hline
-\bfseries{Top Ways} & \bfseries{\#} & \bfseries{\%} \\
+\bfseries{Top Ways of Handling Runtime Exceptions} & \bfseries{\#} & \bfseries{\%} \\
 \hline
@@
 \hline
-\bfseries{Top Ways} & \bfseries{\#} & \bfseries{\%} \\
+\bfseries{Top Ways of Handling Checked Exceptions} & \bfseries{\#} & \bfseries{\%} \\
 \hline
@@
 %/////// CROSS TYPE WRAPPING //////
-After presenting two different code snippets on which a checked and a runtime exception were caught, developers were presented with a code snippet in which a cross-type wrapping was performed\footnote{This cross-type wrapping was found in several applications during the mining study as well as in some classes of the Android framework.} -- as illustrated below.
+After presenting two different code snippets on which a checked and a runtime exception were thrown, developers were presented with a code snippet in which a cross-type wrapping was performed\footnote{This cross-type wrapping was found in several applications during the mining study as well as in some classes of the Android framework.} -- as illustrated below.
+Then developers were asked whether such cross-type wrapping could affect the application robustness in some way.
 
@@
 
-Then developers were asked whether such cross-type wrapping could affect the application robustness in some way.
-Table XXXX illustrates the  most mentioned reasons of why cross-type wrapping may affect app robustness; 18\% of the respondents mentioned that the app would crash anyway, regardless of whether the exception was wrapped or not.
+ Table~\ref{tab:robustness} illustrates the  most mentioned reasons of why cross-type wrapping may affect app robustness; 18\% of the respondents mentioned that the app would crash anyway, regardless of whether the exception was wrapped or not.
 Most of the respondents, however, mentioned disadvantages of such wrapping, such as: (i) it impairs proper handling since wrapping in a general exception loses information about the exception situation to be handled (24\%); such wrapping treats all exceptions as critical -- in other words, every exception will remain uncaught and will lead to an app crash -- and it does not allow the developer to perform a proper handling (e.g.
 retry) for exception types that are not critical.
+A respondent emphasized that: \emph{``this treats all exceptions as critical failure, even though they might not all be unrecoverable"[D70]}.
+Another developer alerted that: \emph{``Rethrowing a checked exception as an unchecked exception is the worst choice; it throws away any value of checked exceptions, elevates the exception out of almost all handling, and does absolutely nothing helpful in the process.
+Further, catching `Exception` suggests that the author doesn't know what exceptions might occur and wrongfully assumed that this solution is somehow safer than doing nothing."[D19]}
+
+\finding{ Wrapping a checked exception in a runtime exception was considered a bed practice by may respondents; as it loses exception information it impairs proper handling (24\%), and since the runtime exception may remain uncaught it may crash the app (24\%).
+}
 
@@
 \caption{Top reasons why cross-type wrapping may affect app robustness - 49 non-empty responses. }
-\label{tab:handlingruntime}
+\label{tab:robustness}
 \end{table}
-
-``this treats all exceptions as critical failure, even though they might not all be unrecoverable" [D70]
-
-``Rethrowing a checked exception as an unchecked exception is the worst choice; it throws away any value of checked exceptions, elevates the exception out of almost all handling, and does absolutely nothing helpful in the process.
-Further, catching `Exception` suggests that the author doesn't know what exceptions might occur and wrongfully assumed that this solution is somehow safer than doing nothing."[D19]
 
@@
 Only 3 respondents out of 71 (4\% of respondents) were aware of the fact that there are ways a method can throw a non-declared checked exception, such as JNI/native code, reflection, and directly changing the bytecode/dalvik code which bypass the compile-time checking.
-Most of the respondents were not aware that a method could throw a checked exception without declaring it in the method's signature.
+One of them said: \emph{``I am guessing this is because the exception is thrown from native code (ie, C++ code in the JVM) where Java correct-ness semantics rules do not always apply." [D42]}.
+
+Most of the respondents, however, were not aware that a method could throw a checked exception without declaring it in the method's signature.
 Most respondents mentioned that if a method throws a non-declared exception it must be a RuntimeException.
@@
 
-\greybox{``I am guessing this is because the exception is thrown from native code (ie, C++ code in the JVM) where Java correct-ness semantics rules do not always apply." [D42]}
+\bigskip
+
+\finding{Only 4\% of the developers aware of the robustness threats caused by JNI undocumented checked exceptions.}
+
+\bigskip
 

### 2015-10-08 SPRINGER_JOURNAL/msr_journal.tex
including more quotes re-writting text

@@
 Anticipated erroneous situation help writing more robust code  & 9	 & 13\% \\ 
-%Anticipated erroneous situation help writing more robust code 
 Exceptions only for unrecoverable behavior	& 7 &	10\% \\
@@
 \bfseries{For Negatively Affecting Robustness} &   &   \\
-Every exception should be caught to improve robustness	& 7	& 10\% \\
+If not handled it crashs the app	 & 7	& 10\% \\
 Makes debugging harder &	2 &	3\% \\
@@
 Although most of the respondents answered that exception handling code helped with the development of robust Android applications, almost every respondent pointed out a drawback associated with the exception handling code, saying that if it is not used with care the exception handling code may lead to app crashes.
-
-\greybox{``Any uncatched exception will crash the app.
-There are many unknown exception thrown by the system, that are only happen once in a lifetime like IllegalStateException: eglMakeCurrent failed EGL\_BAD\_CONTEXT"[D32]}
-
-\greybox{``The lifecycle of activities\/fragements etc make it harder to work out what order things are called in and so exceptions can occur because you didn't realise that another method isn't called."[D70]}
+\emph{``Any uncatched exception will crash the app.
+There are many unknown exception thrown by the system, that are only happen once in a lifetime like IllegalStateException: eglMakeCurrent failed EGL\_BAD\_CONTEXT"[D32]}.
+\emph{``The lifecycle of activities\/fragements etc make it harder to work out what order things are called in and so exceptions can occur because you didn't realise that another method isn't called."[D70]}
 
@@
 
-Developers were asked to rank the main causes of crashes.
+Developers were also asked to rank the main causes of crashes.
 Approximately 68\% of the respondents ranked the exceptions signaled by programming logic mistakes, e.g.
@@
 Activity) is a potential entry point in which exceptions can arise leading to an app crash.
-This explains the importance of handling exceptions in all entry points, and using general crash clauses which sometimes cannot prevent crashes but allow the developer to present a detailed error message to the user before crashing.
+This explains the importance of handling exceptions in all entry points, and using general catch clauses which sometimes cannot prevent crashes but allow the developer to present a detailed error message to the user before crashing.
+\emph{``You can wrap the main function in java in try/catch in android you can't..."[D12]}.
+\emph{``Though not particular to Android, most Android apps have several different entry points: Activities, Services, etc, can be started by different services; same thing with events.
+It may be hard to ensure that all this cases are properly handled." [D46]}
+
 Many respondents also said that they used testing approaches to prevent crashes -- few of them mentioned crowd testing (3\%), an emerging approach in testing apps.
@@
 Finally, crash report tools (e.g.
-ACRA [ref]) were reported as a way to prevent crashes.
+ACRA~\cite{Acra14}) were reported as a way to prevent crashes.
 Once such tools notify developers about current crashes, the application can be fixed and hence future crashes (caused by what was fixed) could be prevented.
@@
 
-\greybox{``Though not particular to Android, most Android apps have several different entry points: Activities, Services, etc, can be started by different services; same thing with events.
-It may be hard to ensure that all this cases are properly handled." [D46]}
-
-\greybox{``You can wrap the main function in java in try/catch in android you can't..."[D12]}
+\bigskip
+
+\finding{Developers recognized that exception handling affects the application robustness in two different ways.
+On the one hand,  anticipated erroneous situation help writing more robust code.
+On the other hand, any uncaught exception will lead the an app crash.}
+
+\bigskip
 

### 2015-10-08 SPRINGER_JOURNAL/msr_journal.tex
including more quotes re-writting text
fixing refs

@@
 %Q1Q2_new.png
-\begin{figure} \centering \includegraphics[scale=0.38]{charts/expertise.png}
+\begin{figure} \centering \includegraphics[scale=0.38]{charts/expertise_new.png}
 \caption{Developers expertise in Java and Android development.}
@@
 engineering context to discover the user perspective regarding a broad range of 
-topics such as: how developers define productivity [ref]; how GitHub developers use
-pull-requests[ref]; how the testing culture of open-shource projects can be characterized [ref]; 
-and even how developers use Twiter [ref]. This kind of study is important as a way of 
+topics such as: assess the developers perceptions on productivity ~\cite{meyer2014}; 
+how GitHub developers use pull-requests ~\cite{gousios2015}; 
+how the testing culture of open-source projects can be characterized ~\cite{pham2013}; 
+and even how developers use Twiter ~\cite{singer2014}. This kind of study is important as a way of 
 better understanding the developers behavior and hence provide recommendations and 

### 2015-10-08 SPRINGER_JOURNAL/msr_journal.tex
fixing refs
typo

@@
 cope with an increasing number of exceptinal conditions 
-(e.g., faults in underlying middleware or hardware; compatibility issues~\cite{McDon13}; memory and battery restrictions; noisy external resources~\cite{Zhang12}).
+(e.g., faults in underlying middleware or hardware; compatibility issues~\cite{McDon13}; memory and batery restrictions; noisy external resources~\cite{Zhang12}).
 
@@
 
-\begin{figure} \centering \includegraphics[scale=0.60]{charts/agree_new2.png}
+\begin{figure} \centering \includegraphics[scale=0.58]{charts/agree_new2.png}
 \caption{How often developers throw and handle exceptions during app development.}

### 2015-10-09 SPRINGER_JOURNAL/msr_journal.tex
typo
typos

--- 2015-10-08 SPRINGER_JOURNAL/msr_journal.tex
@@
 cope with an increasing number of exceptinal conditions 
-(e.g., faults in underlying middleware or hardware; compatibility issues~\cite{McDon13}; memory and battery restrictions; noisy external resources~\cite{Zhang12}).
+(e.g., faults in underlying middleware or hardware; compatibility issues~\cite{McDon13}; 
+memory and battery restrictions; noisy external resources~\cite{Zhang12}).
 
@@
 
-%Android, the leading plaform for mobile applications~\cite{gartner} (thanks to its open-source 
+%Android, the leading platform for mobile applications~\cite{gartner} (thanks to its open-source 
 %model and hardware partners like Samsung, HTC, Motorola, Asus besides Nexus Google's
@@
 Open source Android open-source apps have also been the 
-target of other research~\cite{Linar13,Ruiz12} addressing the \emph{reuse} and API stability.
+target of other research~\cite{Linar13,Ruiz12} addressing \emph{reuse} and API stability.
 
@@
 To conduct this classification, we provide this module with the information
-comprising all Java packages that compose: the Android Plaform;
+comprising all Java packages that compose: the Android Platform;
  the Android Libcore; and each analyzed Application. To discover the packages for the first two origins
@@
 
-\finding{Almost two thirds of all crashes come from runtime
-  exceptions. Most of these originate from the application layer.}
+\finding{Almost 65\% of all crashes come from runtime
+  exceptions. Approximately 47\% of these originate from the application layer.}
 
@@
 Approximately 68\% of the respondents ranked the exceptions signaled by programming logic mistakes, e.g.
-NullPointerExceptions, as the Top 1 and Top 2 cause of crashes, as illustrated in Figure~\ref{figure:ranking}.
+NullPointerExceptions, as the Top 1 and Top 2 cause of crashes, as illustrated in Figure~\ref{fig:ranking}.
 In our mining study the exceptions signaled due to programming logic mistakes (java.lang and util) where the causes of 52.0\% of the exception stack traces found.
@@
 analysis of the exception hierarchy and analized the exception wrappings that can 
-happen during the exception propagation.  Such analysis revealed intriging 
+happen during the exception propagation.  Such analysis revealed intriguing 
 \emph{bug hazards} such as the cross-type exception wrappings not discussed in previous works.

### 2015-10-09 SPRINGER_JOURNAL/msr_journal.tex
fixing typos

@@
 exception handling bug hazards in Android apps; and an exploratory survey study - whose goal
-was to assess the perspectives of developers conserning such bug hazards.
+was to assess the perspectives of developers concerning such bug hazards.
 The repository mining study mined the exception stack traces embedded on 
@@
 and analyzed in combination with source code and bytecode analysis. 
-Some bug hazards were consistently detected such as:  unexpected exception wrappings (e.g., Errors being wrapped in checked exceptions), and NullPointerExeptions as the major cause of reported chashes in about 50\% of all analyzed projects.
+Some bug hazards were consistently detected such as:  unexpected exception wrappings 
+(e.g., Errors being wrapped in checked exceptions), and NullPointerExeptions 
+as the major cause of reported crashes - reported in 
+28\% of all mined stack traces found in about 50\% of all analyzed projects. 
 Finally, a survey was applied to some developers 
-involved on such GitHub projects and their perspective about the bug hazards and 
-how they deal with them as assessed and discussed.
+involved on some of the GitHub Android projects analyzed in our study 
+and their perspective about the bug hazards and 
+how they deal with them was assessed and discussed.
 
@@
 extend phones capabilities far beyond of the basic calls, on the other hand they have to 
-cope with an increasing number of exceptinal conditions 
+cope with an increasing number of exceptional conditions 
 (e.g., faults in underlying middleware or hardware; compatibility issues~\cite{McDon13}; 

### 2015-10-09 SPRINGER_JOURNAL/msr_journal.tex
fixing typos

@@
 The exploratory survey was conducted with 70 Android developers involved in one or more of the 
-GitHub Android projects whose issues were mined. This survey reveaks that only few developers (3\% of respondents)
+GitHub Android projects whose issues were mined. This survey reveals that only few developers (3\% of respondents)
  new about the undocumented checked exceptions signaled by native C code of Android platform. 
@@
  If this file is not available, the tool recursively analyzes the 
-structure of source-code directories composing the appm filtering out the cases in which the application also includes the source code of reused libraries.
+structure of source-code directories composing the application, and filters out the cases in which the application also includes the source code of reused libraries.
 Then, based on this information and using pattern matching 
@@
 bug hazards in both the Android applications and framework?}
-To make the analysis easier, we further refine this question in subquestions, each one focusing on pieces of information
+To make the analysis easier, we further refine this question in sub-questions, each one focusing on pieces of information
  distilled from stack traces, more specifically: (i) the root exceptions (i.e., the exceptions that caused the stack traces);
  (ii) the exception types (i.e, Checked, Runtime, Error, Throwable) and (iii) the exception wrappings. 
-Hence, this section is centered around the following subquestions:  \emph{RQ 1.1: Can the root exceptions reveal bug harzards?}; 
+Hence, this section is centered around the following sub-questions:  \emph{RQ 1.1: Can the root exceptions reveal bug harzards?}; 
  \emph{RQ 1.2 Can the exception types reveal bug hazards?}; and  \emph{RQ 1.3 Can the exception wrappings reveal bug hazards?}.
@@
 it violates the best practices related to checked exceptions and errors as it mixes the intended handling 
-behaviour associated to both types.
+behavior associated to both types.
 
@@
 
-To make the analysis easier, we further refine this question in subquestions, such as: \emph{How often do developers handle exceptions?}; \emph{How often do developers throw exceptions?}.
+To make the analysis easier, we further refine this question in sub-questions, such as: \emph{How often do developers handle exceptions?}; \emph{How often do developers throw exceptions?}.
 Moreover, another question that shows us how developers may deal with the EH code is: \emph{Do developers know about Java EH best practices and/or Android specific EH best practices?}
@@
 The grounded theory coding used in this study consists of two phases: (1) initial coding entails a close reading of the data and (2) later we use focused coding ~\cite{charmaz2006} to pinpoint and develop the most salient themes~\cite{charmaz2006} in the analyzed data.
-The coding process was performed by two coders who agreed upon the list of themes that emerged from the answers'set of each question.
+The coding process was performed by two coders who agreed upon the list of themes that emerged from the answers' set of each question.
 
@@
 Their answers are aligned with the high prevalence of NullPointerExceptions found in the mining study -- in Table~\ref{tab:topten} in Section~\ref{sec:result} we can see that NullPointerExceptions were the main cause of the mined exception stack traces.
-Developers reported some reasons for the hight prevelance of crashes caused by NullPointerExceptions.
+Developers reported some reasons for the high prevalence of crashes caused by NullPointerExceptions.
 The activity and fragment lifecycle was one of them -- \emph{``Android destroys and recreates itself all of the time (especially during screen rotation) and if you do not handle that it will crash on you every time.
@@
 \bfseries{For Negatively Affecting Robustness} &   &   \\
-If not handled it crashs the app	 & 7	& 10\% \\
+Crashes the app if not handled	 & 7	& 10\% \\
 Makes debugging harder &	2 &	3\% \\
-Unchecked/runtime exceptions can crash apps & 2	& 3\% \\
+Unchecked/runtime exceptions may crash the app & 2	& 3\% \\
 NullPointerExceptoins can happen anywhere/are tricky to avoid	 & 2	& 3\% \\
@@
 Furthermore, only some of the Android developers surveyed in this study were aware of the 
-rebustness threats caused by cross-type wrappings as the one cited above. This calls for empirical studies on the actual usefulness of Java's hybrid exception model.
+robustness threats caused by cross-type wrappings as the one cited above. This calls for empirical studies on the actual usefulness of Java's hybrid exception model.
 
@@
 The null reference was firstly introduced by Tony Hoare in ALGOL W, which after some years he called his ``one-billion-dollar mistake''~\cite{hoare2}.
-In this sudy, the null references were, infact, responsible for several reported issues - providing further evidence to Hoare's statement.
+In this study, the null references were, in fact, responsible for several reported issues - providing further evidence to Hoare's statement.
 Moreover many of the survey respondents recognized that NullPointerExceptions was one of the main causes of application crashes since they can happen almost
-anyware in the application code, and to make thinks worse the life-cycle of Android apps on which objects are constantly recreated (i.e., Activity and Fragment classes) favor this kind of exception to happen added to the fact that many Android framework methods return null without making such
+any ware in the application code, and to make thinks worse the life-cycle of Android apps on which objects are constantly recreated (i.e., Activity and Fragment classes) favor this kind of exception to happen added to the fact that many Android framework methods return null without making such
 return type explicitly on the documentation.
@@
 but also mentioned how difficult is to handle every specific exception that can happen.
-Several developers recognized to follow a more reactive behaviour against crashes:
+Several developers recognized to follow a more reactive behavior against crashes:
 they advocate the use of crash report tools, and once the crash happens for the first time,
@@
 Moreover, several exception stack traces that support the findings of this study
-refered to exceptions coming from methods defined on Android Application Framework
-and third-party libraries.  Additionaly,  the \emph{bug hazards} observed in this study are due to
+referred to exceptions coming from methods defined on Android Application Framework
+and third-party libraries.  Additionally,  the \emph{bug hazards} observed in this study are due to
 characteristics of the Java exception model, which can impose challenges 
@@
 handling bug hazards found during the repository mining 
-phase, we choose Gounded Theory techniques. The results of our
+phase, we choose Grounded Theory techniques. The results of our
 survey-based study may not apply to every Android developer, since other populations 
@@
 
-Altough the themes and findings that emerged in our study cannot 
+Although the themes and findings that emerged in our study cannot 
 be generalized, they give a first view of developers perspectives about
@@
 (classifying them as Error, Runtime or Checked) based on the source code
-analysis of the exception hierarchy and analized the exception wrappings that can 
+analysis of the exception hierarchy and analyzed the exception wrappings that can 
 happen during the exception propagation.  Such analysis revealed intriguing 
@@
 handling code associated with external resources. This work inspired ours,
- which automaticaly mined the exception stack traces embedded on issues 
+ which automatically mined the exception stack traces embedded on issues 
 reported on 639 open source Android projects. The goal of our study was

### 2015-10-09 SPRINGER_JOURNAL/msr_journal.tex
fixing typos

@@
 of each project. As a result we could identify 157 Android projects.  This set contained  127,456 issues in total,
- from which 1,963 exception stack traces were extracted. Table~\ref{txab:lables} illustrates the occurrences of different labels 
+ from which 1,963 exception stack traces were extracted. Table~\ref{tab:lables} illustrates the occurrences of different labels 
 on the issues including exception stack traces. Differently from GitHub, in Google Code most of 
@@
 
-Developers were first asked how often they dealt with the exception handling code.
+Developers were first asked how often they dealt with the exception handling code - see Figure~\ref{fig:often}.
 Many of our survey respondents recognized they have to handle exceptions most of the time during Android app development (64.8\%).
 However, the frequency in which they throw exceptions inside apps is smaller; most of the survey respondents said they throw exceptions only some of the time or seldom (80.3\%).
-It shows that although we can try to avoid dealing with exceptions by throwing few exceptions, it is almost always inevitable to handle the exceptions thrown by the Android platform and reused libraries.
+It shows that although we can try to avoid throwing few exceptions, it is almost always inevitable to handle the exceptions thrown by the Android platform and reused libraries.
 
@@
 \caption{How often developers throw and handle exceptions during app development.}
-\label{fig:oftenhandle}
+\label{fig:often}
 \end{figure}
@@
 \caption{Summary of some survey question answers. }
-\label{fig:oftenhandle}
+\label{fig:allquest}
 \end{figure}
@@
 Section~\ref{sec:best}  compiled a list of some exception handling (EH) best practices that guided our mining study.
-In this survey the developers were also asked if they knew about and used Java EH best practices.
+In this survey the developers were also asked if they knew about and used Java EH best practices - see Figure~\ref{fig:allquest}.
 Most of them said that they knew and adopted Java EH practices (68\%).
@@
 %/////// ANDROID EH BEST PRACTICES /////
-The developers who informed us about their EH best practices were also asked whether they knew about any EH best practices specific to Android.
+The developers who informed us about their EH best practices were also asked whether they knew about any EH best practices specific to Android - see Figure~\ref{fig:allquest}.
 Most of them (43\%) mentioned that they applied the same best practices that they used in Java programs.
@@
 %//////NULLPONTER WORSE IN ANDROID /////
-Developers were asked if their apps ever crashed because of a NullPointerException.
+Developers were asked if their apps ever crashed because of a NullPointerException - see Figure~\ref{fig:allquest}.
 The vast majority of respondents (96\%) said that their apps crashed at least once due to a NullPointerException.
@@
 
-Developers were also asked whether the exception handling code helped with the development of robust Android applications.
+Developers were also asked whether the exception handling code helped with the development of robust Android applications - see Figure~\ref{fig:allquest}.
 Although most of the developers answered positively (78.9\%), when asked to provide one or more reasons for their answers, most of the respondents also provide reasons why the exception handling code can sometimes impair the robustness.

### 2015-10-09 SPRINGER_JOURNAL/msr_journal.tex
fixing typos

@@
 This paper reports two studies:  a repository mining study whose goal was to identify 
-exception handling bug hazards in Android apps; and an exploratory survey study - whose goal
+exception handling bug hazards in Android apps; and an survey-based study whose goal
 was to assess the perspectives of developers concerning such bug hazards.
@@
 %Q1Q2_new.png
-\begin{figure} \centering \includegraphics[scale=0.38]{charts/expertise_new.png}
+\begin{figure} \centering \includegraphics[scale=0.38]{expertise_new.png}
 \caption{Developers expertise in Java and Android development.}
@@
 
-\begin{figure} \centering \includegraphics[scale=0.58]{charts/agree_new2.png}
+\begin{figure} \centering \includegraphics[scale=0.58]{agree_new2.png}
 \caption{How often developers throw and handle exceptions during app development.}
@@
 
-\begin{figure} \centering \includegraphics[scale=0.60]{charts/yes_no_new2.png}
+\begin{figure} \centering \includegraphics[scale=0.60]{yes_no_new2.png}
 \caption{Summary of some survey question answers. }
@@
 \begin{figure*} 
-\centering \includegraphics[scale=0.60]{charts/ranking.png}
+\centering \includegraphics[scale=0.60]{ranking.png}
 \caption{Developers perspectives on the main causes of crashes.}\label{fig:ranking} 
@@
 
-%\begin{acknowledgements}
-%If you'd like to thank anyone, place your comments here
-%and remove the percent signs.
-%\end{acknowledgements}
+\begin{acknowledgements}
+This work is partially supported by INES, funded by CNPq and
+FACEPE, grants 573964/2008-4 and APQ-1037-1.03/08.
+\end{acknowledgements}
 

### 2015-10-09 SPRINGER_JOURNAL/msr_journal.tex
fixing typos

@@
 This paper reports two studies:  a repository mining study whose goal was to identify 
-exception handling bug hazards in Android apps; and an survey-based study whose goal
+exception handling bug hazards in Android apps; and a survey-based study whose goal
 was to assess the perspectives of developers concerning such bug hazards.

### 2015-10-11 SPRINGER_JOURNAL/msr_journal.tex
fixing typos
grammar and spelling

@@
 \begin{abstract}
-This paper reports two studies:  a repository mining study whose goal was to identify 
-exception handling bug hazards in Android apps; and a survey-based study whose goal
-was to assess the perspectives of developers concerning such bug hazards.
-The repository mining study mined the exception stack traces embedded on 
-159,048 issues reported on Android projects hosted in GitHub (482 projs.) 
-and Googlecode (157 projs.).  Overall 6,005 exception stack traces were extracted, 
-and analyzed in combination with source code and bytecode analysis. 
-Some bug hazards were consistently detected such as:  unexpected exception wrappings 
-(e.g., Errors being wrapped in checked exceptions), and NullPointerExeptions 
-as the major cause of reported crashes - reported in 
-28\% of all mined stack traces found in about 50\% of all analyzed projects. 
-Finally, a survey was applied to some developers 
-involved on some of the GitHub Android projects analyzed in our study 
-and their perspective about the bug hazards and 
-how they deal with them was assessed and discussed.
+This paper reports two studies:  a repository mining study whose goal was to identify exception handling bug hazards in Android apps; and a survey-based study whose goal was to assess the perspectives of developers concerning such bug hazards.
+The repository mining study mined the exception stack traces embedded in 159,048 issues reported in Android projects hosted on GitHub (482 projs.)  and Google Code (157 projs.).
+Overall 6,005 exception stack traces were extracted and analyzed in combination with source code and bytecode analysis.
+Some bug hazards were consistently detected, such as:  unexpected exception wrappings (e.g., errors being wrapped in checked exceptions) and NullPointerExceptions as the major cause of reported crashes -- reported in 28\% of all mined stack traces found in about 50\% of all analyzed projects.
+Finally, a survey was applied to some developers involved in some of the GitHub Android projects analyzed in our study and their perspective about the bug hazards and how they deal with them was assessed and discussed.
 
@@
 
-The number of mobile apps is increasing in a daily rate. If on the one hand they
-extend phones capabilities far beyond of the basic calls, on the other hand they have to 
+The number of mobile apps is increasing at a daily rate. If on the one hand they
+extend phones' capabilities far beyond the basic calls, on the other hand they have to 
 cope with an increasing number of exceptional conditions 
@@
 
-However, such mechanisms are more often then not the less 
+However, such mechanisms are more often than not the least 
 understood and tested parts of the system~\cite{miller1997issues,Robil00,shah2010understanding, 
@@
 As a consequence they may inadvertently negatively affect the system: exception-related code may introduce failures such as 
-uncaught exceptions~\cite{jo2004uncaught, Zhang12} - 
+uncaught exceptions~\cite{jo2004uncaught, Zhang12} -- 
 which can lead to system crashes, making the system even less robust~\cite{coelho2011unveiling}.
@@
 it automatically terminates, while the system prints a stack trace to the console, 
-or on a log file~\cite{gosling2000java}.  A typical Java stack trace consists of  the fully qualified name 
+or to a log file~\cite{gosling2000java}.  A typical Java stack trace consists of  the fully qualified name 
 of the thrown exception and the ordered list of methods that were active on the call stack before 
@@
 
-This work is conducted in two phases. Firstly, a mining study performs a post 
+This work is conducted in two phases. First, a mining study performs a post 
 mortem analysis of the exception stack traces 
-included in issues reported on Android projects hosted in GitHub and Google Code.
+included in issues reported on Android projects hosted on GitHub and Google Code.
 The goal of this study is to investigate whether the reported exception stack traces
@@
 of introducing the aforementioned uncaught exceptions.
-Secondly, an exploratory survey to the Android developers involved on the mined projects
+Second, we conducted an exploratory survey with the Android developers involved in the mined projects
 to assess their perspective about the exception handling bug hazards found.
@@
  which we developed specifically for this study, we mine stack traces from the issues reported 
-on 482 Android projects hosted in GitHub and 157 projects hosted in Google Code.
+in 482 Android projects hosted on GitHub and 157 projects hosted on Google Code.
 Overall 159,048 issues were analyzed and 6,005 stack traces were extracted from them.
 The exception stack trace analysis was augmented by means of bytecode 
-and source code analysis on the exception-related code of the Android 
+and source code analysis for the exception-related code of the Android 
 platform and Android applications. Some \emph{bug hazards} consistently detected 
@@
 Trying to handle an instance of OutOfMemoryError ``hidden" in  a checked exception may bring the program
- to an unpredictable state. Such wrappings suggest that when (mis)applied, the exception wrapping can make 
+ to an unpredictable state. Such wrappings suggest that, when (mis)applied, the exception wrapping can make 
 the exception-related code more complex and negatively impact the application robustness.
 
-  \item  Undocumented runtime exceptions raised by the Android platform (35 methods) and  third-party libraries (44 methods) -
+  \item  Undocumented runtime exceptions raised by the Android platform (35 methods) and  third-party libraries (44 methods) --
  which correspond to 4.4\% of the reported exception stack traces.
@@
 
- \item A multitude of programming mistakes - approximately 52\% of the reported stack traces can 
+ \item A multitude of programming mistakes -- approximately 52\% of the reported stack traces can 
 be attributed to programming mistakes. In particular, 27.71\% of all stack traces contained a java.lang.NullPointerException 
@@
 GitHub Android projects whose issues were mined. This survey reveals that only few developers (3\% of respondents)
- new about the undocumented checked exceptions signaled by native C code of Android platform. 
+ knew about the undocumented checked exceptions signaled by native C code of the Android platform. 
 Moreover, most of the developers recognized that cross-type exception wrappings may
 negatively impact the application robustness. The uncaught exceptions due to errors in programming 
-logic, with the NullPointerException was identified by most developers (68\%) as the first or second main cause 
+logic, e.g. the NullPointerException, were identified by most developers (68\%) as the first or second main cause 
 of application crashes.
 
- The high prevalence of NullPointerExceptions found in the mining study and confirmed on the exploratory survey
+ The high prevalence of NullPointerExceptions found in the mining study and confirmed in the exploratory survey
  is in line with findings of earlier research~\cite{kim2013predicting,fraser20131600,csallner2004jcrasher}, 
@@
 Moreover, we perform  the first exploratory survey study whose
-goal was to assess developers perspective regarding a set of exception handling bug hazards
+goal was to assess developers' perspective regarding a set of exception handling bug hazards
 as well as how developers deal with exceptions and prevent crashes while developing.
@@
 Moreover, such \emph{bug hazards} 
-call for improvements on languages (e.g. to prevent null pointer dereferences) and tools to better support exception handling in Android and Java environments.
+call for improvements of languages (e.g. to prevent null pointer dereferences) and tools to better support exception handling in Android and Java environments.
 
@@
 Section~\ref{sec:disc} provides a discussion of the wider implications of our results,
- presents the threats to validity associated to the mining study and discusses the limitations
+ presents the threats to validity associated with the mining study and discusses the limitations
 of the survey-based study. 
@@
 
-\emph{\textbf{Exception Types.}} In Java, exceptions are represented according to a class hierarchy, on which
+\emph{\textbf{Exception Types.}} In Java, exceptions are represented according to a class hierarchy, in which
  every exception is an instance of the Throwable class, and can be of three kinds: the checked exceptions
 (extends Exception), the runtime exceptions (extends RuntimeException) and errors
-(extends Error)~\cite{gosling2000java}. Checked exception received their name
- because they must be declared on the method's \emph{exception interface} (i.e., the list of exceptions that a method 
+(extends Error)~\cite{gosling2000java}. Checked exceptions received their name
+ because they must be declared in the method's \emph{exception interface} (i.e., the list of exceptions that a method 
 might raise during its execution) and the compiler statically checks if
@@
 Both runtime exceptions and errors are also known as ``unchecked exceptions'', as 
-they do not need to be specified on the method \emph{exception interface} and do not trigger any 
+they do not need to be specified in the method \emph{exception interface} and do not trigger any 
 compile time checking.
@@
 Normally these cannot be handled inside the application.  Instances of RuntimeException are implicitly 
-thrown by Java runtime environment when a program violates 
+thrown by the Java runtime environment when a program violates 
 the semantic constraints of the Java programming language (e.g., out-of-bounds array index, divide-by-zero 
@@
  or unchecked, by extending either Exception or RuntimeException. There is a long-lasting debate 
-about the pros and cons of both approaches~\cite{javatut,stackoverlow,debate}
+about the pros and cons of both approaches~\cite{javatut,stackoverlow,debate}.
 %\footnote{152 questions in Stackoverflow are related to this debate}. 
@@
 One exception 
-is caught and wrapped in another which is then thrown instead. Figure~\ref{fig:wrapping} shows 
+is caught and wrapped in another one which is then thrown instead. Figure~\ref{fig:wrapping} shows 
 an exception stack trace which illustrates such exception wrapping. 
-For simplicity, in this paper we will refer to ``exception stack trace'' as as just stack trace.
+For simplicity, in this paper we will refer to ``exception stack trace'' as just stack trace.
 The bottom part of the stack trace is the \emph{root exception} (Figure~\ref{fig:wrapping}-A), which indicates
-the first reason (root cause) for the exception thrown (in this case, the computer run out of
+the first reason (root cause) for the exception thrown (in this case, the computer ran out of
 memory). The top part of the stack trace indicates the location of the exception
@@
 execution flow  between the root exception and the wrapper may
-include other intermediate exception wrappers (Figure~\ref{fig:wrapping}-D). In all levels, the exception
-\emph{signaler}, is the method that threw the exception, represented on the
+include other intermediate exception wrappers (Figure~\ref{fig:wrapping}-D). At all levels, the exception
+\emph{signaler} is the method that threw the exception, represented in the
 stack trace as the first method call below the exception declaration (Figure~\ref{fig:wrapping}-B).
@@
 
-Our study focused on the exception stack traces contained in issues of Android projects (hosted in GitHub and Google Code).
-And to support our investigation, we developed a tool called ExceptionMiner (Section~\ref{sec:exceptionminer})
-which extracts the exception stack traces embedded on issues, 
+Our study focused on the exception stack traces contained in issues of Android projects (hosted on GitHub and Google Code).
+To support our investigation, we developed a tool called ExceptionMiner (Section~\ref{sec:exceptionminer})
+which extracts the exception stack traces embedded in issues 
 and combines stack trace information with source code and bytecode 
@@
 Figure~\ref{fig:overview} gives an overview of our study.
-First,  the issues reported on Android projects hosted on GitHub (1) and Google Code (2) are recovered.
+First,  the issues reported in Android projects hosted on GitHub (1) and Google Code (2) are recovered.
 Then the stack traces embedded in each issue are extracted and distilled (3).
 The stack trace information is then combined with source code and bytecode analysis in order to discover 
-the type of the exceptions (5) reported on the stack traces (e.g., error, runtime, checked),
+the type of the exceptions (5) reported in the stack traces (e.g., error, runtime, checked),
 and the origin (6) of such exceptions 
@@
 cannot be retrieved from commercial apps, whose issue report systems and 
-source codes are generally not publicly available. 
-Open source Android open-source apps have also been the 
+source code are generally not publicly available. 
+Open source Android apps have also been the 
 target of other research~\cite{Linar13,Ruiz12} addressing \emph{reuse} and API stability.
@@
 
-\subsection{Android Apps in GitHub}
+\subsection{Android Apps on GitHub}
 \label{sec:git}
@@
 This study uses the dataset provided by the GHTorrent project~\cite{Gousi13}, 
-an off-line mirror of the data  offered through the Github API.  
+an off-line mirror of the data  offered through the GitHub API.  
 To identify Android projects, we performed a case insensitive search for the
-term ``android" in the repository's names and short descriptions.  
+term ``android'' in the repositories' names and short descriptions.  
 Up to 23 February 2014,  when we queried GHTorrent, this resulted in 2,542 repositories. Running the ExceptionMiner tool 
- in this set we observed that 589 projects had at least one issue containing a stack trace.
+ on this set we observed that 589 projects had at least one issue containing a stack trace.
 
@@
 example projects (i.e., toy projects) or tools to support Android development
-(e.g. Selendroid, Roboeletric - tools to support the testing of Android apps).
-The filtered set consisted of 482 apps. This set of 482 projects contained overall 31,592 issues from which 4,042 exception stack traces 
+(e.g. Selendroid, Roboeletric -- tools to support the testing of Android apps).
+The filtered set consisted of 482 apps. This set of 482 projects contained a total of 31,592 issues from which 4,042 exception stack traces 
 were extracted.
 
-Issues on Github are different from issues on dedicated bug tracking tools such as 
+Issues on GitHub are different from issues in dedicated bug tracking tools such as 
 Bugzilla and Jira. The most important difference is that there are no predefined fields
-  (e.g. severity and priority). Instead, Github uses a more open ended tagging system, on which
+  (e.g. severity and priority). Instead, GitHub uses a more open ended tagging system, where
 repositories are offered a pre-defined set of labels, but repository owners can modify 
 them at will. Therefore, an issue may have none or an arbitrary set of labels depending 
-on its repository. Table~\ref{tab:lables} illustrates the ocurrences of different labels 
+on its repository. Table~\ref{tab:lables} illustrates the occurrences of different labels 
 on the issues including exception stack traces. Regardless of the issue labels, every exception stack
 trace may contain relevant information concerning the exception structure of the
-projects analyzed, and therefore can reveal \emph{bug hazards} on the exception-related code.
+projects analyzed, and therefore can reveal \emph{bug hazards} in the exception-related code.
 Because of this, we opted for not restricting the analysis to just defect issues.
@@
 Google Code contains widely used open-source Android apps (e.g.
-K9Mail\footnote{K9Mail moved to Github but as a way of not loosing the project history it advises their users to report bugs on the Google Code issue tracker: https://github.com/k9mail/k-9/wiki/LoggingErrors.}).
+K9Mail\footnote{K9Mail moved to GitHub but as a way of not loosing the project history it advises their users to report bugs in the Google Code issue tracker: https://github.com/k9mail/k-9/wiki/LoggingErrors.}).
 However, differently from GitHub, Google Code does not provide an API to access the information related
- to hosted projects.\footnote{Google code used to provide a Web service to its repositories, but this was deactivated in June 2013 in what Google called a ``clean-up action''.}
-To overcome this limitation we needed to implement a Web Crawler (incorporated in ExceptionMiner tool described next) that navigates 
-through the web interface of Google Code projects extracting all issues and issue comments and storing them in a relational database for later analysis.
-To identify Android projects in Google Code, we performed a similar heuristic: we performed a case insensitive search 
+ to hosted projects.\footnote{Google Code used to provide a Web service to its repositories, but this was deactivated in June 2013 in what Google called a ``clean-up action''.}
+To overcome this limitation we needed to implement a Web Crawler (incorporated in the ExceptionMiner tool described next) that navigates 
+ the web interface of Google Code projects extracting all issues and issue comments and storing them in a relational database for later analysis.
+To identify Android projects on Google Code, we performed a similar heuristic: we performed a case insensitive search 
 (on the Google Code search interface) for the term ``android''.
-On January 2014, when we queried Google Code, this resulted in a list of 788  projects.
+In January 2014, when we queried Google Code, this resulted in a list of 788  projects.
 This list comprised the seeds sent to our Crawler.
@@
 From this set, 724 projects defined at least 1 issue. Running the ExceptionMiner tool 
- in this set we observed that 183 projects had at least one issue containing an exception stack trace.
+ on this set we observed that 183 projects had at least one issue containing an exception stack trace.
  Then we performed further clean up (similar to the one described previously) inspecting the site 
@@
  from which 1,963 exception stack traces were extracted. Table~\ref{tab:lables} illustrates the occurrences of different labels 
-on the issues including exception stack traces. Differently from GitHub, in Google Code most of 
+on the issues including exception stack traces. Differently from GitHub, on Google Code most of 
 the issues were labeled as ``Defect''. However, based on the same assumption described for the GitHub repository
- we considered all issues reporting stack traces (regardless its labels).
+ we considered all issues reporting stack traces (regardless of their labels).
 
@@
 with issue repositories. In this study two main connectors were created: one which connects to 
-GHTorrent database, and a Google Code connector which is comprised of a Web Crawler
- that can traverse the Google Code web interface and extract the project's issues. 
+the GHTorrent database, and a Google Code connector which is comprised of a Web Crawler
+ that can traverse the Google Code web interface and extract a project's issues. 
 Project meta-data and the issues associated with each project are stored in a relational 
@@
  the root exception and its signaler, as well as the exception wrappers and their corresponding signalers. 
-This component also distills fine grained information of each attribute such as the classes and packages associated to them.
+This component also distills fine grained information of each attribute such as the classes and packages associated with them.
 In contrast to existing issue parsing solutions such as Infozilla, our parser
  can discover stack traces mixed with log file information\footnote{In several 
-exception stack traces, the exception frames were preceeded by logging information e.g., 
+exception stack traces, the exception frames were preceded by logging information e.g., 
 \texttt{03-01 15:55:01.609 (7924): at android.app.ActivityThread.access\$600(ActivityThread\-.java:127)} which could not be detected by existing tools.}.
@@
 \emph{\textbf{Exception Type Analysis.}} To support a deeper investigation of the 
-stack traces every exception defined on a stack trace needs to be classified according to its type
+stack traces every exception defined in a stack trace needs to be classified according to its type
 (e.g. Error, checked Exception or RuntimeException). The module responsible for this analysis 
 uses the Design Wizard framework \cite{Brunet09} to inspect the bytecode of the exceptions reported 
-on stack traces.  It walks up the type hierarchy of a Java exception until it reaches 
+in stack traces.  It walks up the type hierarchy of a Java exception until it reaches 
 a base exception type. Hence in this study the bytecode analysis was used to discover the type 
-of each mined exception when the jar file of such exception was available on the project or 
-on a reused  Java library. A specific implementation (based on source code analysis)
+of each mined exception when the jar file of such an exception was available in the project or 
+in a reused  Java library. A specific implementation (based on source code analysis)
  was needed to discover the exception type when the bytecode was not available.
@@
 and exceptions thrown by Android core libraries. Moreover, we also analyzed the 
-exceptions reported on stack traces that were defined on applications and third-party libraries 
+exceptions reported in stack traces that were defined by applications and third-party libraries 
 (the tool only analyzed the last version available).
@@
  If this file is not available, the tool recursively analyzes the 
-structure of source-code directories composing the application, and filters out the cases in which the application also includes the source code of reused libraries.
+structure of source code directories composing the application, and filters out the cases in which the application also includes the source code of reused libraries.
 Then, based on this information and using pattern matching 
@@
  from libraries if their packages are neither defined 
-within the Android platform, nor on core libraries, nor on applications.
+within the Android platform, nor in core libraries, nor in the applications.
 Table~\ref{tab:signalers} summarizes this signaler classification.
@@
     \hline
-    \bfseries{android} & If the exception is thrown in a method defined in Android platform.\\
-    \bfseries{app}     & If the exception is thrown in an method defined on an Android app.\\
+    \bfseries{android} & If the exception is thrown in a method defined in the Android platform.\\
+    \bfseries{app}     & If the exception is thrown in a method defined in an Android app.\\
     \bfseries{libcore} & If the exception is thrown in one of the core libraries reused by Android (e.g., org.apache.harmony, org.w3c.dom, sun.misc, org.apache.http, org.json, org.xml). \\
-    \bfseries{lib}     & If the exception is thrown on a method that was not defined by any of the elements above.\\
+    \bfseries{lib}     & If the exception is thrown in a method that was not defined by any of the elements above.\\
     \hline
@@
 that were not automatically identified by the ExceptionMiner tool
-(because they were defined on previous versions of libraries,
+(because they were defined in previous versions of libraries,
 apps and Android Platform). When the exception could not be  
-found automatically or manually (because they were defined on a previous version
+found automatically or manually (because they were defined in a previous version
 of the app or lib), we classified the exception as ``Undefined''.  Only 31 exceptions 
@@
 bug hazards in both the Android applications and framework?}
-To make the analysis easier, we further refine this question in sub-questions, each one focusing on pieces of information
+To make the analysis easier, we further refine this question into sub-questions, each one focusing on pieces of information
  distilled from stack traces, more specifically: (i) the root exceptions (i.e., the exceptions that caused the stack traces);
  (ii) the exception types (i.e, Checked, Runtime, Error, Throwable) and (iii) the exception wrappings. 
-Hence, this section is centered around the following sub-questions:  \emph{RQ 1.1: Can the root exceptions reveal bug harzards?}; 
+Hence, this section is centered around the following sub-questions:  \emph{RQ 1.1: Can the root exceptions reveal bug hazards?}; 
  \emph{RQ 1.2 Can the exception types reveal bug hazards?}; and  \emph{RQ 1.3 Can the exception wrappings reveal bug hazards?}.
 
-In this section each piece information is analyzed in detail to 
-check whether they can reveal bug harzards on the exception handling code -
+In this section each piece of information is analyzed in detail to 
+check whether it can reveal bug hazards in the exception handling code --
 related to (i) specific violations of the best practices presented in Section~\ref{sec:best},
-or (ii) the general use of exception handling to suport robust development.
+or (ii) the general use of exception handling to support robust development.
 
@@
 
-After distilling the information available on the exception stack traces, we could find 
+After distilling the information available in the exception stack traces, we could find 
 the exceptions commonly reported as the root causes of stack traces.
-Table \ref{tab:topten} presents a list of the top 10 root exceptions found in the study - 
- ranked by the number of distinct projects on which they were reported. 
-This table also shows how many times the signaler of such an exception was a method defined on
-the Android platform, the Android libcore, the application itself or a third-party library -
+Table \ref{tab:topten} presents a list of the top 10 root exceptions found in the study -- 
+ ranked by the number of distinct projects in which they were reported. 
+This table also shows how many times the signaler of such an exception was a method defined by
+the Android platform, the Android Libcore, the application itself or a third-party library --
  following the classification presented in Table  \ref{tab:signalers}.
@@
   \end{tabular}
-\caption{Root Exceptions occurrences and popularity in repositories hosted in Google Code (GC) and GitHub(GH).}
+\caption{Root Exceptions occurrences and popularity in repositories hosted on Google Code (GC) and GitHub(GH).}
 \label{tab:topten}
@@
 across projects, we can observe that 51.96\% of all projects reported at least one exception stack 
-trace on which the NullPoiterException was the root cause.
+trace in which the NullPointerException was the root cause.
 
@@
 Regarding reusable code (e.g., libraries and frameworks), there is no consensus whether it is a good or a bad practice to 
-explicitly throw a NullPointerException. Some prefer to encapsulate such an exception on
+explicitly throw a NullPointerException. Some prefer to encapsulate such an exception in
 an instance of IllegalArgumentException, while others~\cite{bloch2008effective} argue that the
 NullPointerException makes the cause of the problem explicit and hence 
-can be signalled by an API expecting a non-null argument.The high prevalence of NullPointerException is aligned with the
+can be signaled by an API expecting a non-null argument.The high prevalence of NullPointerException is aligned with the
 findings of other 
@@
 For instance, Kechagia and Spinellis showed that the NullPointerException was the  
-most reported exception on the crash reports sent to BugSense (a bug report 
+most reported exception in the crash reports sent to BugSense (a bug report 
 management service for Android applications)  \cite{kechagia2014}.
@@
 detected bugs were due to NullPointerException and exceptions
-implicitly-signaled by the Java 
+implicitly signaled by the Java 
 environment due to programming mistakes or resource limitations
@@
 concerns related to the most frequently reported root exceptions.
-Besides the exceptions related to programming mistakes mentioned before, we also looked for exceptions related to some concerns that are known as sources of faults in mobile development: concurrency~\cite{ama2012} backward compatibility~\cite{McDon13}, security~\cite{enck2011study,was2010} and resource management (IO, Memory, Batery)~\cite{Zhang12}.
-Since it is infeasible to inspect the code responsible for throwning every exception reported in this study,
-the concern identification of each exception was based on intended
+Besides the exceptions related to programming mistakes mentioned before, we also looked for exceptions related to concerns that are known as sources of faults in mobile development: concurrency~\cite{ama2012} backward compatibility~\cite{McDon13}, security~\cite{enck2011study,was2010} and resource management (IO, Memory, Battery)~\cite{Zhang12}.
+Since it is infeasible to inspect the code responsible for throwing every exception reported in this study,
+the concern identification for each exception was based on intended
 meaning of the particular exception type, as defined in 
-its Javadoc documentaion and in the Java specification. 
+its Javadoc documentation and in the Java specification. 
 For example: (i) an instance of ArrayOutOfBoundException  
@@
 Javadoc related to each exception and the Java specification, we
-identified the underlying concern releated to each root exception. 
+identified the underlying concern related to each root exception. 
 Table~\ref{tab:tophundrend} contains the results of this analysis. This table also illustrates the exceptions 
@@
 Programming logic (java.lang and util) & 52,0\%\\
-Resources (IO, Memory, Batery) & 23,9\% \\
+Resources (IO, Memory, Battery) & 23,9\% \\
 Security & 4,1\%\\
@@
 This analysis revealed that approximately 75\% of the exceptions
- that caused the stack traces are implicitly thrown by runtime 
-environment  due to mistakes on programming logic (e.g., out-of-bounds array index, null pointer 
-references) and resource limitations. Although such exceptions do not directly point to violations on
+ that caused the stack traces are implicitly thrown by the runtime 
+environment  due to mistakes in the programming logic (e.g., out-of-bounds array index, null pointer 
+references) and resource limitations. Although such exceptions do not directly point to violations of
 the best practices described before (which are related to the 
-explicitly thrown exceptions) they impose a major threat to apps robustness,
+explicitly thrown exceptions) they impose a major threat to app robustness,
 and therefore represent a critical bug hazard to the exception 
 handling code of Android applications. 
-Security and concurrency, which are known to be critica issues on  Android apps, 
-raised few of the reported exceptions (less then 5\% of the anlysed stack traces).
+Security and concurrency, which are known to be critical issues for Android apps, 
+raised few of the reported exceptions (less than 5\% of the analysed stack traces).
 
@@
 \finding{%
-Exceptions related to programming logic and resource limitations represent a major bug hazard to Android apps - 
+Exceptions related to programming logic and resource limitations represent a major bug hazard to Android apps -- 
 since they represent approximately 75\% of the exceptions that caused the stack traces.
-From this set the NullPointerException as most prevalent exception.
+From this set the NullPointerException is the most prevalent exception.
 }
@@
 As mentioned before, using the ExceptionMiner tool in combination with manual inspections we could identify
-the root exception \emph{type} (i.e., RuntimeException, Error, checked Exception) as well as its \emph{origin} - which we 
-identified based on the package names of the signalers related to it in the stack traces (Section~\ref{sec:exceptionminer}).
+the root exception \emph{type} (i.e., RuntimeException, Error, checked Exception) as well as its \emph{origin} -- which we 
+identified based on the package names of the signalers in the stack traces (Section~\ref{sec:exceptionminer}).
 Table~\ref{tab:typeroottab} presents the types and origins of root exceptions of all analyzed stack traces.
@@
 We can observe that most of the reported exceptions are of type runtime 
-(64.85\%); and that the most common origins are methods defined either on the Application (47.3\%)
-or on the Android platform (34.3\%). We could also find runtime exceptions thrown by library code (17.7\%).
- We can also see, from Table~\ref{tab:typeroottab}, that in contrast with the other origins, most of the 
- exceptions signaled on Android Libcore (i.e., the set of libraries reused by Android) are 
+(64.85\%); and that the most common origins are methods defined either in the Application (47.3\%)
+or in the Android platform (34.3\%). We could also find runtime exceptions thrown by library code (17.7\%).
+ We can also see, from Table~\ref{tab:typeroottab}, that in contrast to the other origins, most of the 
+ exceptions signaled by Android Libcore (i.e., the set of libraries reused by Android) are 
 checked exceptions. This set comprises: org.apache.harmony, org.w3c.dom, sun.misc, 
@@
  \emph{explicitly thrown} runtime exceptions 
-should be documented as part of the \emph{exception interface} of libraries/framework reusable
-methods. To investigate the conformance to this practice, we firstly filtered out all the exceptions implicitly
- signaled by  the runtime environment (due to programming mistakes) - since these exceptions 
-should not be documented on the method signature.  Then we inspected
+should be documented as part of the \emph{exception interface} of the reusable
+methods of libraries/frameworks. To investigate the conformance to this practice, we first filtered out all the exceptions implicitly
+ signaled by  the runtime environment (due to programming mistakes) -- since these exceptions 
+should not be documented in the method signature.  Then we inspected
 the code for each method 
-(defined either in the Android  Aplication Framework or in third-party libraries) 
+(defined either in the Android  Application Framework or in third-party libraries) 
 explicitly signaling a runtime exception. 
@@
 We found 79 methods (both from libraries and the Android platform) that  explicitly threw a runtime exception 
-without listing it on the \emph{exception interface} (i.e., using 
-throws clause in the method signature). From this set only one method (defined on a library)
-included an @throws tag in its Javadoc - reporting that the given runtime exception
+without listing it in the \emph{exception interface} (i.e., using 
+a throws clause in the method signature). From this set only one method (defined in a library)
+included an @throws tag in its Javadoc -- reporting that the given runtime exception
 could be thrown in some conditions. These methods were responsible for 
@@
 This result is in line with the results of two other studies~\cite{sacramento2006unchecked,kechagia2014}. 
-Sacramento et al~\cite{sacramento2006unchecked} observed that the
+Sacramento et al.~\cite{sacramento2006unchecked} observed that the
 runtime exceptions in .NET programs are most often not documented. 
 Kechagia and Spinellis~\cite{kechagia2014} identified a set of methods 
-on the Android API which do not document its runtime exceptions. One limitation 
+in the Android API which do not document their runtime exceptions. One limitation 
 of the latter work is that it did not filter out exceptions that,  
 although runtime, should not be documented because they were implicitly signaled by the 
-JVM due to resource restrictions or violations on Java semantic constraints.
+JVM due to resource restrictions or violations of semantic Java constraints.
 When explicitly signaling a runtime exception and not documenting it, 
@@
 thrown. Almost none (0.4\%, just one) of these were documented. Such undocumented 
-runtime exceptions violate best practices III and IV reveal a bug harzard to Java/Android apps.}
+runtime exceptions violate best practices III and IV and reveal a bug hazard to Java/Android apps.}
 
@@
   \end{tabular}
-\caption{Absence of exception interfaces on methods.}
+\caption{Absence of exception interfaces in methods.}
 \label{tab:runtimeinterface}
@@
 
-\emph{\textbf{Missing Checked Exceptions on Exception Interfaces.}} 
+\emph{\textbf{Missing Checked Exceptions in Exception Interfaces.}} 
 Our exception stack trace analysis revealed an unexpected bug hazard: a 
-checked exception thrown by a native method and not declared on the exception 
-interface of these methods signaling them. The native method in question was defined in
+checked exception thrown by a native method and not declared in the exception 
+interface of the methods signaling them. The native method in question was defined in
  the Android platform, which uses Java Native Invocation (JNI) to access 
@@
 Consequently, since it is not expected by the programmer, when such a method throws 
-this exception, such an undocumented exception may remain
-uncaught and cause the app to crash, or maybe mistakenly handled by subsumption.
+this exception, the undocumented exception may remain
+uncaught and cause the app to crash, or may be mistakenly handled by subsumption.
 The exception stack traces reporting this scenario actually correspond
-to a real bug of Android 
+to a real bug of the Android 
 Gingerbread version (which still accounts for 13.6\% of devices running Android).
@@
 For native methods, even checked exceptions can be
-thrown without being documented on the exception interface.
-It violates best practices III and IV and represent a bug hazard
+thrown without being documented in the exception interface.
+This violates best practices III and IV and represents a bug hazard
  hard to diagnose.
@@
 1 & java.lang.RuntimeException - java.lang.OutOfMemoryError  \\ %30
-2& java.lang.RuntimeException -  java.lang.StackOverflowError)   \\ %4
+2& java.lang.RuntimeException -  java.lang.StackOverflowError   \\ %4
 \hline
@@
 5&java.lang.NoClassDefFoundError - java.lang.ClassNotFoundException   \\ %4
-6&java.lang.AssertionError - javax.crypto.ShortBufferException)   \\ %3
-\hline
-& \bfseries{Error wrapping a Runtime Exceptino}    \\
+6&java.lang.AssertionError - javax.crypto.ShortBufferException   \\ %3
+\hline
+& \bfseries{Error wrapping a Runtime Exception}    \\
  \hline
@@
 Below, we discuss the most
-important of such ``cross-type wrappings in more detail.
+important of such ``cross-type wrappings'' in more detail.
 
@@
 of an asynchronous task) into an instance of RuntimeException and re-throws it.  
-Table~\ref{tab:exampeswrap} presents examples of exceptions that were actually wrapped in this code snipet.  
+Table~\ref{tab:exampeswrap} presents examples of exceptions that were actually wrapped in this code snippet.  
 Such wrappings mask an unrecoverable Error into a general runtime exception.
@@
 This wrapping was responsible for 49.5\% of
-the cross-type wrappings.  From this set 50\%  were performed on methods defined on Android platform. 
-We observe that it is a common implementation practice in the methods of Android platform.
-However, using such a general exception, is considered a bad practice
+the cross-type wrappings.  From this set 50\%  were performed by methods defined by the Android platform. 
+We observe that this is a common implementation practice in the methods of the Android platform.
+However, using such a general exception is considered a bad practice
 according to the Java specification and common guidelines, as it loses contextual information about the exception.
@@
 were also native methods written in C. Table~\ref{tab:exampeswrap} illustrates
-some of these wrappings some of them are masking an OutOfMemoryError
+some of these wrappings -- some of them are masking an OutOfMemoryError
 into a checked exception. Such wrappings may also mask an unrecoverable error
-and may lead to ``exception confusion'' described next.
-
-\emph{\textbf{Error wrapping Runtime and Checked Exceptions}} Table~\ref{tab:exampeswrap}  illustrates
+and may lead to the ``exception confusion'' described next.
+
+\emph{\textbf{Error wrapping Runtime and Checked Exceptions.}} Table~\ref{tab:exampeswrap}  illustrates
 examples of instances of Error wrapping instances of
 RuntimeException. 
-Although such a wapping mixes different exception types,
-since there is no obligation associated to handling runtime exceptions, it
+Although such a wrapping mixes different exception types,
+since there is no obligation associated with handling runtime exceptions, it
 does not violate the aforementioned best practices.
@@
 (i.e., static block)  it is converted into an ExceptionInitializerError 
-on the point where the class is first used.  Table~\ref{tab:exampeswrap} also illustrates
+at the point where the class is first used.  Table~\ref{tab:exampeswrap} also illustrates
 examples of such wrappings. Although such a wrapping may represent a design decision,
 it violates the best practices related to checked exceptions and errors as it mixes the intended handling 
-behavior associated to both types.
+behavior associated with both types.
 
@@
 Runtime. 
-Hence, although some of these wrappings may be a result from design decisions, the mis-use of exception wrapppings may make the exception handling 
+Hence, although some of these wrappings may be a result of design decisions, the mis-use of exception wrappings may make the exception handling 
 code more complex (e.g., the multiple wrappings) and error-prone,
@@
 When the developer is confronted with a checked exception, the designer of the API is telling him/her 
-to handle the exceptional condition (according to Java Specification and 
-best practices). However, such exception may be wrapping an Error such
+to handle the exceptional condition (according to the Java Specification and 
+best practices). However, such an exception may be wrapping an Error such
 as an OutOfMemoryError, which indicates a resource deficiency 
-that the program cannot possibly recover from). Hence, trying to
+that the program cannot possibly recover from. Hence, trying to
 handle such an exception  
@@
 They represent a bug hazard once they violate the semantics of Java's original exception
-design, detailed on best practices I and II  (e.g., when mapping unrecoverable Errors to other types of
+design, detailed in best practices I and II  (e.g., when mapping unrecoverable Errors to other types of
 exceptions). 
@@
 
-The general research question guided our exploratory survey was the following: \emph{RQ 2: How developers deal with the exception handling code in Android apps and what are the developers' perspectives about the main bug hazards found during the mining study?} This general research question was broken in a set of research questions that are answered by the exploratory survey.
+The general research question guiding our exploratory survey was the following: \emph{RQ 2: How do developers deal with the exception handling code in Android apps and what are the developers' perspectives about the main bug hazards found during the mining study?} This general research question was broken into a set of research questions that are answered by the exploratory survey.
 
@@
 
-To make the analysis easier, we further refine this question in sub-questions, such as: \emph{How often do developers handle exceptions?}; \emph{How often do developers throw exceptions?}.
-Moreover, another question that shows us how developers may deal with the EH code is: \emph{Do developers know about Java EH best practices and/or Android specific EH best practices?}
+To make the analysis easier, we further refine this question into sub-questions, such as: \emph{How often do developers handle exceptions?}; \emph{How often do developers throw exceptions?}.
+Moreover, another question that shows us how developers may deal with the exception handling (EH) code is: \emph{Do developers know about Java EH best practices and/or Android specific EH best practices?}
 
@@
 We used grounded theory coding~\cite{charmaz2006}, iterating through our open-ended survey responses, to answer some of the research questions.
-The grounded theory coding used in this study consists of two phases: (1) initial coding entails a close reading of the data and (2) later we use focused coding ~\cite{charmaz2006} to pinpoint and develop the most salient themes~\cite{charmaz2006} in the analyzed data.
-The coding process was performed by two coders who agreed upon the list of themes that emerged from the answers' set of each question.
+The grounded theory coding used in this study consists of two phases: (1) initial coding entails a close reading of the data and (2) later we use focused coding~\cite{charmaz2006} to pinpoint and develop the most salient themes~\cite{charmaz2006} in the analyzed data.
+The coding process was performed by two coders who agreed upon the list of themes that emerged from the answer set of each question.
 
@@
 
- In previous work~\cite{coelhoetal15}, we mined exception stack traces available in issues reported for several Android projects hosted on GitHub and Googlecode.
+ In previous work~\cite{coelhoetal15}, we mined exception stack traces available in issues reported for several Android projects hosted on GitHub and Google Code.
 To ensure that our sample consists of repositories that were real Android apps, we had to inspect every project site and discard toy-programs and non-Android repositories.
@@
 We emailed the 1,824 developers and received 71 valid answers.
-The majority of our respondents have more than 2 years of Java development experience (87.3\%) and  considerable experience in Android development (more than 1 year)  (85.9\%) - see Figure~\ref{fig:devexpertise}.
+The majority of our respondents have more than 2 years of Java development experience (87.3\%) and  considerable experience in Android development (more than 1 year)  (85.9\%) -- see Figure~\ref{fig:devexpertise}.
 
@@
  which can be traced in our database using the following convention: 
-D\#. For instance, D1 corresponds to the answer provided
+D\#. For instance, D1 corresponds to an answer provided
  by developer 1.
@@
 This first research question explores the ways in which developers deal with the exception handling code, either throwing or handling exceptions,  while developing an Android app.
-Developers were asked about the frequency they develop exception handling code and whether or not they adopt best practices while developing.
+Developers were asked about the frequency at which they develop exception handling code and whether or not they adopt best practices while developing.
 
@@
 
-Developers were first asked how often they dealt with the exception handling code - see Figure~\ref{fig:often}.
+Developers were first asked how often they dealt with the exception handling code -- see Figure~\ref{fig:often}.
 Many of our survey respondents recognized they have to handle exceptions most of the time during Android app development (64.8\%).
-However, the frequency in which they throw exceptions inside apps is smaller; most of the survey respondents said they throw exceptions only some of the time or seldom (80.3\%).
-It shows that although we can try to avoid throwing few exceptions, it is almost always inevitable to handle the exceptions thrown by the Android platform and reused libraries.
+However, the frequency at which they throw exceptions inside apps is smaller; most of the survey respondents said they throw exceptions only some of the time or seldom (80.3\%).
+It shows that although we can try to avoid throwing exceptions, it is almost always inevitable to handle the exceptions thrown by the Android platform and reused libraries.
 
@@
 Section~\ref{sec:best}  compiled a list of some exception handling (EH) best practices that guided our mining study.
-In this survey the developers were also asked if they knew about and used Java EH best practices - see Figure~\ref{fig:allquest}.
+In this survey the developers were also asked if they knew about and used Java EH best practices -- see Figure~\ref{fig:allquest}.
 Most of them said that they knew and adopted Java EH practices (68\%).
@@
 For instance:\emph{``Don't throw [pure] RuntimeException (in most cases), don't catch OutOfMemoryError, etc." [D42]}.
-Some of the practices mentioned, however, are not well known Java EH best practices but seem to be used by some Android developers to better deal with exceptions in the Android app context, such as the use of crash fast approach~\footnote{http://www.slideshare.net/pyricau/crash} -- \emph{``Crash is good for hinting developer.
-Fast crash fast solve." [D51]}; and to favor checked exceptions instead of runtime to represent the exceptional conditions in Android apps -- \emph{``RuntimeException-s are not particularly well suited for Android's robustness, as any uncaught one would cause a crash." [D26]}; \emph{``Checked exception help, unchecked exceptions don't." [D25]}; \emph{``avoid unchecked exceptions except if the case is really abnormal and should not happen [...]"[D45]}.
+Some of the practices mentioned, however, are not well known Java EH best practices but seem to be used by some Android developers to better deal with exceptions in the Android app context, such as the use of a crash fast approach\footnote{http://www.slideshare.net/pyricau/crash} -- \emph{``Crash is good for hinting developer.
+Fast crash fast solve." [D51]}; and to favor checked exceptions instead of runtime to represent the exceptional conditions in Android apps -- \emph{``RuntimeExceptions are not particularly well suited for Android's robustness, as any uncaught one would cause a crash." [D26]}; \emph{``Checked exception help, unchecked exceptions don't." [D25]}; \emph{``avoid unchecked exceptions except if the case is really abnormal and should not happen [...]"[D45]}.
 
@@
 \hline
-Use specific handlers / dont catch generic exceptions	& 9 & 23\% \\
+Use specific handlers / don't catch generic exceptions	& 9 & 23\% \\
 Don't swallow Exceptions	 & 7 & 18\% \\
-Dont throw Runtime / Favor Checked exceptions & 4 &	10\% \\
-Do not use exception for normal  flow control. &	4 &	10\% \\
+Don't throw Runtime / Favor Checked exceptions & 4 &	10\% \\
+Do not use exception for normal  flow control &	4 &	10\% \\
 Free Resources in finally-blocks	& 4	& 10\% \\
@@
 \end{tabular}
-\caption{Top mentioned Java EH best practices -  40 non-empty responses. }
+\caption{Top mentioned Java EH best practices -- 40 non-empty responses. }
 \label{tab:javapractices}
@@
 %/////// ANDROID EH BEST PRACTICES /////
-The developers who informed us about their EH best practices were also asked whether they knew about any EH best practices specific to Android - see Figure~\ref{fig:allquest}.
+The developers who informed us about their EH best practices were also asked whether they knew about any EH best practices specific to Android -- see Figure~\ref{fig:allquest}.
 Most of them (43\%) mentioned that they applied the same best practices that they used in Java programs.
-Only some of them mentioned that they adopted specific best practices such as: (i) the use of crash report tools (21\%) -- to notify the developer about the uncaught exceptions that happen in the application code; and (ii) the use of a global exception handler (14\%) -- the UncaughtExceptionHandler advised in Android documentation~\footnote{http://developer.android.com/reference/java/lang/Thread.UncaughtExceptionHandler.html}.
+Only some of them mentioned that they adopted specific best practices such as: (i) the use of crash report tools (21\%) -- to notify the developer about the uncaught exceptions that happen in the application code; and (ii) the use of a global exception handler (14\%) -- the UncaughtExceptionHandler advised in the Android documentation\footnote{http://developer.android.com/reference/java/lang/Thread.UncaughtExceptionHandler.html}.
 
@@
 \hline
-Same as java	& 6 &	43\% \\
+Same as Java	& 6 &	43\% \\
 Crash report tools	 & 3 &	21\% \\
@@
 \end{tabular}
-\caption{Top mentioned Android-specific EH best practices -  14 non-empty responses. }
+\caption{Top mentioned Android-specific EH best practices -- 14 non-empty responses. }
 \label{tab:androidpractices}
@@
 \finding{
-Most of the respondents (64.8\%) recognize that handle exceptions most of the time during the app development. 
-Most of them (62\%) mentioned that adopt EH best practices.
+Most of the respondents (64.8\%) recognize that they handle exceptions most of the time during the app development. 
+Most of them (62\%) mentioned that they adopt EH best practices.
 }
@@
 %//////NULLPONTER WORSE IN ANDROID /////
-Developers were asked if their apps ever crashed because of a NullPointerException - see Figure~\ref{fig:allquest}.
+Developers were asked if their apps ever crashed because of a NullPointerException -- see Figure~\ref{fig:allquest}.
 The vast majority of respondents (96\%) said that their apps crashed at least once due to a NullPointerException.
@@
 \end{tabular}
-\caption{Top reasons why NullPointerExceptions are more frequent in Android apps acording to 19 respondents . }
+\caption{Top reasons why NullPointerExceptions are more frequent in Android apps according to 19 respondents.}
 \label{tab:causenull}
@@
 In the case of external API calls which return null, then check for null (the quick-and-dirty way).
-For internal calls, use @NotNull and @Nullable annotations to provided more guidance on when an object \"may be\" and \"should never be\" null." [D42]}
-
-Some respondents also mentioned using annotations such as @Nullable @NonNull, provided by Android Studio 0.5.5; such annotations guide the developer about what can and cannot be null during the application lifecycle.
+For internal calls, use @NotNull and @Nullable annotations to provided more guidance on when an object ``may be'' and ``should never be'' null." [D42]}
+
+Some respondents also mentioned using annotations such as @Nullable and @NotNull, provided by Android Studio 0.5.5; such annotations guide the developer about what can and cannot be null during the application lifecycle.
 Although catching the NullPointerExceptions is considered a bad practice some respondents said that they do it to prevent crashes caused by NullPointerExceptions.
@@
 investigate/fix the cause	& 24 & 	39\% \\
-@Nullable @NonNull &	8 &	13\% \\
+@Nullable @NotNull &	8 &	13\% \\
 catch null-pointer (mistake)	& 3 &	5\% \\
@@
 avoid using nulls / avoid to use methods can throw null	& 2 &	3\% \\
-static analisis	& 2 &	3\% \\
+static analysis	& 2 &	3\% \\
 automated testing	& 1 &	2\% \\
@@
 \end{tabular}
-\caption{Top ways for preventing NullPointerException in Android apps - 61 non-empty responses. }
+\caption{Top ways for preventing NullPointerException in Android apps -- 61 non-empty responses. }
 \label{tab:preventnull}
@@
 \finding{Almost 96\% of the developers recognized that their apps have already crashed due to a NullPointerException.
-They mentioned that Android application life-cycle (47\%) and the framework complexity (32\%) may favor such uncaught NullPointerExceptions to happen in Android context.
+They mentioned that the Android application life-cycle (47\%) and the framework complexity (32\%) may favor such uncaught NullPointerExceptions to happen in the Android context.
 }
@@
 
-Some of the best practices are related to the different ways an exception should be handled according to whether it is a checked and an unchecked exception.
-When the designer of an API specifies that a method throws a checked exception, it is telling the caller of the API that such exception should be handled.
+Some of the best practices are related to the different ways an exception should be handled according to whether it is a checked or an unchecked exception.
+When the designer of an API specifies that a method throws a checked exception, it is telling the caller of the API that such an exception should be handled.
 
@@
 This question was motivated by the fact that during the mining study we could identify that many checked exceptions were just wrapped in a runtime exception and re-thrown, as a way to bypass any kind of handling.
-When asked about how to deal with a checked exception signaled by a method (invoked in the context of the onPause() activity method), most respondents (62.50\%) said that they should add a try-catch block surrounding the method invocation.
+When asked about how to deal with a checked exception signaled by a method (invoked in the context of the onPause() activity method), most respondents (63\%) said that they should add a try-catch block surrounding the method invocation.
 One of the respondents mentioned that \emph{``If a method signature declares that it throws an exception, it means that the caller should handle the exception."[D29]}; \emph{``because it is checked exception, you should expect it to happen rather than just crash"[D70]}.
@@
 For runtime exceptions most of the handling actions were dedicated to present more specific error messages and prevent the app from abruptly crashing.
-\emph{``[...] They signal exceptional behavior that may not be recoverable, so they offer a useful way to log and gracefully crash.} [D19].
+\emph{``[...] They signal exceptional behavior that may not be recoverable, so they offer a useful way to log and gracefully crash'' [D19]}.
 \emph{``Giving the user meaningful feed back is important, so the more specific you can be about what went wrong, the better."[D66]}.
-Some developers also mentioned to use a toast to support this task\footnote{A toast is an Android component that provides simple feedback about an operation in a small popup - http://developer.android.com/guide/topics/ui/notifiers/toasts.html}.
+Some developers also mentioned to use a toast to support this task\footnote{A toast is an Android component that provides simple feedback about an operation in a small popup -- http://developer.android.com/guide/topics/ui/notifiers/toasts.html}.
 Few developers also mentioned that the runtime exception should flow upstream to the framework so it could crash the app.
@@
 
-We can observe that the handling actions associated to checked exceptions focused more on crash prevention than the ones related to runtime exceptions.
+We can observe that the handling actions associated with checked exceptions focused more on crash prevention than the ones related to runtime exceptions.
 \emph{``Its very important in android that we gracefully handle an exception.
 Try not to crash an application."[41]}.
-It was also interesting to observe that the developers suggested involving the user and solving the exceptional condition represented by a checked exception - \emph{``We should try alternative saving ways and if it fails we should prompt the user for a new location and always keep the user informed"[D11]}.
+It was also interesting to observe that the developers suggested involving the user and solving the exceptional condition represented by a checked exception -- \emph{``We should try alternative saving ways and if it fails we should prompt the user for a new location and always keep the user informed"[D11]}.
 \emph{``You still need to catch an IOException to prevent a crash.
@@
 \end{tabular}
-\caption{Top ways of handling runtime exceptions -  63 non-empty responses. }
+\caption{Top ways of handling runtime exceptions -- 63 non-empty responses. }
 \label{tab:handlingruntime}
@@
 \end{tabular}
-\caption{Top ways of dealing with a checked exception signaled in the context of an Activity method -  40 non-empty responses. }
+\caption{Top ways of dealing with a checked exception signaled in the context of an Activity method -- 40 non-empty responses. }
 \label{tab:handlingruntime}
@@
 %/////// CROSS TYPE WRAPPING //////
-After presenting two different code snippets on which a checked and a runtime exception were thrown, developers were presented with a code snippet in which a cross-type wrapping was performed\footnote{This cross-type wrapping was found in several applications during the mining study as well as in some classes of the Android framework.} -- as illustrated below.
+After presenting two different code snippets in which a checked and a runtime exception were thrown, developers were presented with a code snippet in which a cross-type wrapping was performed\footnote{This cross-type wrapping was found in several applications during the mining study as well as in some classes of the Android framework.} -- as illustrated below.
 Then developers were asked whether such cross-type wrapping could affect the application robustness in some way.
@@
 
-\finding{ Wrapping a checked exception in a runtime exception was considered a bed practice by may respondents; as it loses exception information it impairs proper handling (24\%), and since the runtime exception may remain uncaught it may crash the app (24\%).
+\bigskip
+
+\finding{ Wrapping a checked exception in a runtime exception was considered a bad practice by many respondents; since it loses exception information it impairs proper handling (24\%), and since the runtime exception may remain uncaught it may crash the app (24\%).
 }
@@
 useless rethrow	& 2 &	4\% \\
-activity mehtods cannot throw exceptions &	1 &	2\% \\
+activity methods cannot throw exceptions &	1 &	2\% \\
 
@@
 \end{tabular}
-\caption{Top reasons why cross-type wrapping may affect app robustness - 49 non-empty responses. }
+\caption{Top reasons why cross-type wrapping may affect app robustness -- 49 non-empty responses. }
 \label{tab:robustness}
@@
 The respondents were presented with a piece of code in which a non-declared checked exception has appeared, and they were asked if they knew any reason for this to happen.
-Only 3 respondents out of 71 (4\% of respondents) were aware of the fact that there are ways a method can throw a non-declared checked exception, such as JNI/native code, reflection, and directly changing the bytecode/dalvik code which bypass the compile-time checking.
-One of them said: \emph{``I am guessing this is because the exception is thrown from native code (ie, C++ code in the JVM) where Java correct-ness semantics rules do not always apply." [D42]}.
+Only 3 respondents out of 71 (4\% of respondents) were aware of the fact that there are ways a method can throw a non-declared checked exception, such as JNI/native code, reflection, and directly changing the bytecode/dalvik code, which bypass the compile-time checking.
+One of them said: \emph{``I am guessing this is because the exception is thrown from native code (i.e., C++ code in the JVM) where Java correct-ness semantics rules do not always apply." [D42]}.
 
@@
 We could observe that this exception handling bug hazard detected during the mining study, although representing a threat to app robustness, is hardly known by the Android developers involved in this survey.
-Below we can find a quote from one of the 3 respondents aware of JNI non-declared checked exceptions.
-
-\bigskip
-
-\finding{Only 4\% of the developers aware of the robustness threats caused by JNI undocumented checked exceptions.}
+%Below we can find a quote from one of the 3 respondents aware of JNI non-declared checked exceptions.
+
+\bigskip
+
+\finding{Only 4\% of the developers were aware of the robustness threats caused by JNI undocumented checked exceptions.}
 
@@
 
-Developers were also asked whether the exception handling code helped with the development of robust Android applications - see Figure~\ref{fig:allquest}.
-Although most of the developers answered positively (78.9\%), when asked to provide one or more reasons for their answers, most of the respondents also provide reasons why the exception handling code can sometimes impair the robustness.
+Developers were also asked whether the exception handling code helped with the development of robust Android applications -- see Figure~\ref{fig:allquest}.
+Although most of the developers answered positively (78.9\%), when asked to provide one or more reasons for their answers, most of the respondents also provided reasons why the exception handling code can sometimes impair the robustness.
 Since positive and negative reasons were provided intermixed we analyzed both sets of answers and ranked the top reasons why exception handling can negatively or positively affect robustness according to developers' perceptions.
-Table~\ref{tab:topreasons} presents the top reasons why the exception handling code may or may not help thee development of robust apps.
+Table~\ref{tab:topreasons} presents the top reasons why the exception handling code may or may not help the development of robust apps.
 
@@
 Unchecked/runtime exceptions may crash the app & 2	& 3\% \\
-NullPointerExceptoins can happen anywhere/are tricky to avoid	 & 2	& 3\% \\
+NullPointerExceptions can happen anywhere/are tricky to avoid	 & 2	& 3\% \\
 VM inefficiency 	& 1 & 1\% \\
@@
 There are many unknown exception thrown by the system, that are only happen once in a lifetime like IllegalStateException: eglMakeCurrent failed EGL\_BAD\_CONTEXT"[D32]}.
-\emph{``The lifecycle of activities\/fragements etc make it harder to work out what order things are called in and so exceptions can occur because you didn't realise that another method isn't called."[D70]}
+\emph{``The lifecycle of activities\/fragments etc make it harder to work out what order things are called in and so exceptions can occur because you didn't realise that another method isn't called."[D70]}
 
@@
 NullPointerExceptions, as the Top 1 and Top 2 cause of crashes, as illustrated in Figure~\ref{fig:ranking}.
-In our mining study the exceptions signaled due to programming logic mistakes (java.lang and util) where the causes of 52.0\% of the exception stack traces found.
-Hence, the developers perspective is aligned with what was revealed in mining study.
+In our mining study the exceptions signaled due to programming logic mistakes (java.lang and util) were the causes of 52\% of the exception stack traces found.
+Hence, the developers' perspective is aligned with what was revealed in the mining study.
 %[width=.9\hsize]
@@
 \centering \includegraphics[scale=0.60]{ranking.png}
-\caption{Developers perspectives on the main causes of crashes.}\label{fig:ranking} 
+\caption{Developers' perspectives on the main causes of crashes.}\label{fig:ranking} 
 \end{figure*}
@@
 Once such tools notify developers about current crashes, the application can be fixed and hence future crashes (caused by what was fixed) could be prevented.
-Table~ref{tab:topprevent} presents the themes for the strategies for crash prevention mentioned by developers.
+Table~\ref{tab:topprevent} presents the themes for the strategies for crash prevention mentioned by developers.
 
@@
 \finding{Developers recognized that exception handling affects the application robustness in two different ways.
-On the one hand,  anticipated erroneous situation help writing more robust code.
-On the other hand, any uncaught exception will lead the an app crash.}
+On the one hand,  anticipated erroneous situations help writing more robust code.
+On the other hand, any uncaught exception will lead to an app crash.}
 
@@
 null check(s) &	11 &	15\% \\
-- null check annotations (@Null and @NonNull)	& 3 & 	4\% \\
+- null check annotations (@Null and @NotNull)	& 3 & 	4\% \\
 coding style(s)	& 10 &	14\% \\
@@
 \end{tabular}
-\caption{Top ways of preventing crashes - 71 non-empty responses. }
+\caption{Top ways of preventing crashes -- 71 non-empty responses. }
 \label{tab:topprevent}
@@
 \section{Replication Package}
-All the data used in the mining study and in the survey-based study is publicly available at the ExceptionMiner tool website hosted in GitHub\footnote{ https://github.com/souzacoelho/exceptionminer}.
+All the data used in the mining study and in the survey-based study is publicly available at the ExceptionMiner tool website hosted on GitHub\footnote{ https://github.com/souzacoelho/exceptionminer}.
 Specifically we provide: (i) all issues related to Android projects found
-in GiTHub and Google Code used in this study; (ii) all stack traces extracted
+on GitHub and Google Code used in this study; (ii) all stack traces extracted
 from issues; (iii) the results of manual inspection steps; (iv) the
@@
  \emph{``Everybody hates thinking about exceptions, because they are not supposed to happen'' 
-  (Brian Foote)}\footnote{Brian Foote shared his opinion in a conversation with James Noble - quoted on the paper: hillside.net/plop/2008/papers/ACMVersions/coelho.pdf}
+  (Brian Foote)}\footnote{Brian Foote shared his opinion in a conversation with James Noble -- quoted in the paper: hillside.net/plop/2008/papers/ACMVersions/coelho.pdf}
 \end{quotation}
@@
  more complex and lead to what we call the \emph{exception handling confusion problem}.
-This problem can lead the program to an unpredictable state in the presence of exceptions,
+This problem can lead a program to an unpredictable state in the presence of exceptions,
 as illustrated by the scenario in which a checked exception wraps an OutOfMemoryError. 
@@
 \subsection{On the null pointer problem.}
-The null reference was firstly introduced by Tony Hoare in ALGOL W, which after some years he called his ``one-billion-dollar mistake''~\cite{hoare2}.
-In this study, the null references were, in fact, responsible for several reported issues - providing further evidence to Hoare's statement.
-Moreover many of the survey respondents recognized that NullPointerExceptions was one of the main causes of application crashes since they can happen almost
-any ware in the application code, and to make thinks worse the life-cycle of Android apps on which objects are constantly recreated (i.e., Activity and Fragment classes) favor this kind of exception to happen added to the fact that many Android framework methods return null without making such
-return type explicitly on the documentation.
+The null reference was first introduced by Tony Hoare in ALGOL W, which after some years he called his ``one-billion-dollar mistake''~\cite{hoare2}.
+In this study, the null references were, in fact, responsible for several reported issues -- providing further evidence to Hoare's statement.
+Moreover many of the survey respondents recognized that NullPointerExceptions were one of the main causes of application crashes since they can happen almost
+anywhere in the application code, and to make things worse the life-cycle of Android apps in which objects are constantly recreated (i.e., Activity and Fragment classes) favors this kind of exception to happen added to the fact that many Android framework methods return null without making this
+return type explicit in the documentation.
 Such observations emphasizes the need for solutions to avoid NullPointerExceptions, such as:
@@
 as Monads~\cite{Walde95} (as used in functional languages for values that may not be available 
-or computations that may fail) could improve the robustness of Java programs. 
-Some of the Android developers mentioned how @Nullable annotations could helpful to deal with
-NullPointerExceptions and consequently prevent app crashes caused by it.
+or computations that may fail), to improve the robustness of Java programs. 
+Some of the Android developers mentioned that @Nullable annotations could be helpful to deal with
+NullPointerExceptions and consequently prevent app crashes caused by them.
 
@@
 In this study we could observe undocumented runtime exceptions thrown by third party code,
-and even undocumented checked exception thrown by a JNI interface.
-Such undocumented exceptions make it difficult, and most of the times infeasible
-for the client code to protect against `unforeseen'' situations that may happen 
-while calling a library code. In the survey-based study Android developers were asked about
+and even undocumented checked exceptions thrown by a JNI interface.
+Such undocumented exceptions make it difficult, and most of the times infeasible,
+for the client code to protect against ``unforeseen'' situations that may happen 
+while calling library code. In the survey-based study Android developers were asked about
 ways to prevent application crashes which are mainly caused by uncaught exceptions.
 Many of them emphasized the importance of handling exceptions to prevent crashes
-but also mentioned how difficult is to handle every specific exception that can happen.
-Several developers recognized to follow a more reactive behavior against crashes:
+but also mentioned how difficult it is to handle every specific exception that can happen.
+Several developers follow a more reactive behavior against crashes:
 they advocate the use of crash report tools, and once the crash happens for the first time,
-its reported and the application can be changed to better cope with the exceptions that crashed the crashes, consequently preventing future similar crashes.
+it is reported and the application can be changed to better cope with the exceptions that caused the crashes, consequently preventing future similar crashes.
 
@@
 solution may prevent  the system from abruptly crashing, as mentioned by some of the surveyed developers such a general handler will not have enough
-contextual information to adequately handle the exception, 
+contextual information to adequately handle the exceptions, 
 beyond storing a message in a log file and restarting the application.
-However, such a handler cannot replace a carefully designed exception 
+Such a handler cannot replace a carefully designed exception 
 handling policy~\cite{Robil00}, which requires
@@
 %\emph{\textbf{Threats to Validity of the Mining Study.}} 
-\subsection{Mining Study - Threats to Validity }
+\subsection{Mining Study -- Threats to Validity }
 
@@
 as valid exception identifiers, while, in many cases, developers use the
-exception name in issue description. Conservative parsing may minimize false
+exception name in the issue description. Conservative parsing may minimize false
 positives, which was our initial target, but also tends to increase false
@@
  interrater agreement after three independent 
-developers classified a randomly selected sample (of 25 exception 
-types from the total of 100); the interater agreement was high (96\%).
+raters classified a randomly selected sample (of 25 exception 
+types from the total of 100); the interrater agreement was high (96\%).
 
 \emph{External Validity.} Our work uses the GHTorrent dataset, which although 
-comprehensive and extensive is not an exact replica of Github. 
+comprehensive and extensive is not an exact replica of GitHub. 
 However, the result of this study does not depend on the analysis of
-a complete Github dataset. Instead, the goal of our study was to 
-pinpoint \emph{bug hazards} on the exception-related code based on 
+a complete GitHub dataset. Instead, the goal of our study was to 
+pinpoint \emph{bug hazards} in the exception-related code based on 
 exception stack trace mining of a subset of projects.
@@
 this subset is a small percentage of existing apps.
-Such threats are similar to the ones imposed to other empirical studies 
+Such threats are similar to the ones of other empirical studies 
 which also use free or open-source Android apps~\cite{Linar13,McDon13,Ruiz12}.
 Moreover, several exception stack traces that support the findings of this study
-referred to exceptions coming from methods defined on Android Application Framework
+referred to exceptions coming from methods defined in the Android Application Framework
 and third-party libraries.  Additionally,  the \emph{bug hazards} observed in this study are due to
-characteristics of the Java exception model, which can impose challenges 
-the robustness of not only to Android apps but also to other systems
+characteristics of the Java exception model, which can impose challenges to 
+the robustness of not only Android apps but also to other systems
  based on the same exception model.
@@
 Another threat relates to the fact that parts of our analysis 
-are based on the availability of stack traces on issues reported on Github and Googlecode projects. 
-In using these datasets, we make an underlying assumption: the stack traces reported on issues are 
+are based on the availability of stack traces in issues reported on GitHub and Google Code projects. 
+In using these datasets, we make an underlying assumption: the stack traces reported in issues are 
 representative of valid crash information of the applications. 
-One way to mitigate this threat would be to access to the full 
+One way to mitigate this threat would be to access the full 
 set of crash data per application. Although some services exist 
@@
 In our study, we mitigated this threat by manually inspecting
-the source code associated to a subset of the reported exception stack traces.
+the source code associated with a subset of the reported exception stack traces.
 This subset comprises the stack traces related to the main findings 
@@
 
-\subsection{Limitations of the Survey-based Study.}
+\subsection{Limitations of the Survey-based Study}
 Due to the exploratory nature of the second phase of this work whose 
-goal was to identify the developers perspectives concerning the exception 
+goal was to identify the developers' perspectives concerning the exception 
 handling bug hazards found during the repository mining 
-phase, we choose Grounded Theory techniques. The results of our
+phase, we chose Grounded Theory techniques. The results of our
 survey-based study may not apply to every Android developer, since other populations 
 might add new insights.  The population we collected data from was comprised 
-by Android developers of the GitHub 
-Android projects whose issues were mined on the first phase of this work, and who had
+of Android developers of the GitHub 
+Android projects whose issues were mined in the first phase of this work, and who had
 time and motivation to answer the survey questions.
@@
 Although the themes and findings that emerged in our study cannot 
-be generalized, they give a first view of developers perspectives about
+be generalized, they give a first view of developers' perspectives about
  EH bug hazards found. Hence, we believe that the survey-based study 
-have contributed with valuable insights about how Android developers 
-used the exception handling code and what are their perspectives regarding
-exception handling bug hazards.
+has contributed with valuable insights about how Android developers 
+used the exception handling code and what their perspectives regarding
+exception handling bug hazards are.
 
@@
 et al.~\cite{dhaliwal2011classifying} proposed a crash grouping approach that
-can reduce bug fixing time in approximately 5\%. Wang et
+can reduce bug fixing time by approximately 5\%. Wang et
 al.~\cite{wang2013improving} propose an approach to identify correlated crash
 types and describe a fault localization method to locate and rank files related
-to the bug described on a stack trace. Schroter et al.~\cite{schroter2010stack}
+to the bug described in a stack trace. Schr\"oter et al.~\cite{schroter2010stack}
 conducted an empirical study on the usefulness of stack traces for bug fixing
 and showed that developers fixed the bugs faster when failing stack traces were
-included on bug issues.  In a similar study, Bettenburg et
+included in bug issues.  In a similar study, Bettenburg et
 al.~\cite{bettenburg2008makes} identify stack traces as the second most stack
@@
 related to NullPointerException and other implicitly thrown exceptions had
-the higher prevalence on the analyzed set of stacks. Kechagia and Spinellis~\cite{kechagia2014}
-examined the stack traces embedded on crash reports sent by 1,800 Android apps 
+the higher prevalence in the analyzed set of stacks. Kechagia and Spinellis~\cite{kechagia2014}
+examined the stack traces embedded in crash reports sent by 1,800 Android apps 
 to a crash report management service (i.e., BugSense). They found that 19\% of such stack traces
-were caused by unchecked and undocumented exceptions thrown by methods defined on 
+were caused by unchecked and undocumented exceptions thrown by methods defined in the 
 Android API (level 15). Our work differs from Kechagia and Spinellis since it is based on
-stack traces mined from issues reported by open source developers on GitHub and Googlecode.  
+stack traces mined from issues reported by open source developers on GitHub and Google Code.  
 Moreover, our study mapped the origin of each exception 
@@
 related to a stack trace. The main limitation of this solution is that it is not
-able to extract stack traces embedded on verbose log files (i.e., on which we
+able to extract stack traces embedded in verbose log files (i.e., in which we
 can find log text mixed with exception frames). Bacchelli
 et al.~\cite{bacchelli2012content} propose a solution to recognize stack trace frames
-from development emails and relate it to code artifacts (i.e. classes) mentioned
-on the stack trace. In addition to those tools, ExceptionMiner is able to 
+from development emails and relate them to code artifacts (i.e. classes) mentioned
+in the stack trace. In addition to those tools, ExceptionMiner is able to 
 both extract stack traces from natural language artifacts and to 
-classify them in a set of predefined categories.
+classify them into a set of predefined categories.
 
@@
 source code of 32 open-source systems, both for Java and .NET. They
-observed that the actions inside handlers were very simple (e.g., logging and present a
+observed that the actions inside handlers were very simple (e.g., logging and presenting a
 message to the user). Coelho et al.~\cite{coelho2008assessing} performed an 
@@
 number of uncaught exceptions, degrading the robustness of the AO version of every analyzed system.
-The main limitation of approaches based on static analysis based approaches are the number of false
-positives they can generate, and the problems the faced when dealing with reflection libraries 
+The main limitation of approaches based on static analysis approaches are the number of false
+positives they can generate, and the problems faced when dealing with reflection libraries 
 and dynamic class loading. Pingyu and Elbaum~\cite{Zhang12} were the first to perform
-an empirical investigation of issues, related to exception-related bugs, on Android projects.  
-They perform a small scale study on which they manually inspected the issues of 
+an empirical investigation of issues, related to exception-related bugs, in Android projects.  
+They perform a small scale study in which they manually inspected the issues of 
 5 Android applications. They observed that 29\% had to do with poor
-exceptional handling code, this empirical study was used to motivate the development of a tool
+exceptional handling code, and this empirical study was used to motivate the development of a tool
 aiming at amplifying existing tests to validate exception 
 handling code associated with external resources. This work inspired ours,
- which automatically mined the exception stack traces embedded on issues 
-reported on 639 open source Android projects. The goal of our study was
-to identify common \emph{bug hazards} on the exception related code that can lead to 
+ which automatically mined the exception stack traces embedded in issues 
+reported in 639 open source Android projects. The goal of our study was
+to identify common \emph{bug hazards} in the exception related code that can lead to 
 failures such as uncaught exceptions.
@@
 \textit{Empirical studies using Android apps.} Ruiz et al.~\cite{Ruiz12}
-investigated the degree of reuse across applications in Android Market, the
+investigated the degree of reuse across applications in the Android Market; the
 study showed that almost 23\% of the classes inherited from a base class in the
 Android API, and that 217 mobile apps were reused completely by another mobile
-app. Pathak et al.~\cite{Patha11} analyzed bug reports and developers
-discussions of Android platform and found out that approximately 20\% of
+app. Pathak et al.~\cite{Patha11} analyzed bug reports and developers'
+discussions of the Android platform and found that approximately 20\% of
 energy-related bugs in Android occurred after an OS update. McDonnell et
 al.~\cite{McDon13} conducted a case study of the co-evolution behavior of
-Android API and 10 dependent applications using the version history data found
+the Android API and 10 dependent applications using the version history data found
 in GitHub. The study found that approximately 25\% of all methods in the client
 code used the Android API, and that the methods reusing fast-evolving APIs were
-more defect prone then others. Vasquez et al.~\cite{Linar13} analyzed
-approximately 7K free Android apps and observed that the last successful apps
+more defect prone than others. Vasquez et al.~\cite{Linar13} analyzed
+approximately 7K free Android apps and observed that the least successful apps
 used Android APIs that were on average 300\% more change-prone than the APIs
 used by the most successful apps. Our work differs from the others as it aims at
-distilling stack trace information of bug reports and combine such information
+distilling stack trace information from bug reports and combining such information
 with bytecode analysis, source code analysis and manual inspections
-to identify \emph{bug hazards} on the exception handling code of Android apps.
+to identify \emph{bug hazards} in the exception handling code of Android apps.
 
@@
 engineering context to discover the user perspective regarding a broad range of 
-topics such as: assess the developers perceptions on productivity ~\cite{meyer2014}; 
-how GitHub developers use pull-requests ~\cite{gousios2015}; 
-how the testing culture of open-source projects can be characterized ~\cite{pham2013}; 
-and even how developers use Twiter ~\cite{singer2014}. This kind of study is important as a way of 
-better understanding the developers behavior and hence provide recommendations and 
-tools to help on specific development tasks. This was the first exploratory survey study whose
-goal was to assess developers perspective regarding a set of exception handling bug hazards
+topics such as: assessing the developers' perceptions on productivity~\cite{meyer2014}; 
+how GitHub developers use pull-requests~\cite{gousios2015}; 
+how the testing culture of open-source projects can be characterized~\cite{pham2013}; 
+and even how developers use Twitter~\cite{singer2014}. This kind of study is important as a way of 
+better understanding the developers' behavior and hence providing recommendations and 
+tools to help with specific development tasks. This was the first exploratory survey study whose
+goal was to assess developers' perspective regarding a set of exception handling bug hazards
 as well as how developers deal with exceptions and prevent crashes while developing.
@@
 related to exception handling code that may lead to a decrease in
-application robustness; and (ii) to assess the developers perspective
+application robustness; and (ii) to assess the developers' perspective
 concerning the detected EH \emph{bug hazards}.
@@
 To that end, we mined the stack 
-traces embedded in all issues defined in 482 Android projects hosted in Github and 
-157 projects hosted in Google Code - overall it included 6,005
-exception stack traces. And we surveyed developers associated to some of 
+traces embedded in all issues defined in 482 Android projects hosted on GitHub and 
+157 projects hosted on Google Code -- overall considering 6,005
+exception stack traces. We also surveyed developers associated with some of 
 the mined GitHub projects.
@@
 a novel approach and toolset (ExceptionMiner) for analyzing Java
-  exception stack traces as occurring on GitHub and Google Code
+  exception stack traces as occurring in GitHub and Google Code
   issues.
@@
   prominent;
-  (2) Documentation for explicitly thrown RuntimeExceptions is almost
+  (2) documentation for explicitly thrown RuntimeExceptions is almost
   never provided; 
-  (3) Extensive use of wrapping leads to hard to understand chains
+  (3) extensive use of wrapping leads to hard-to-understand chains
   violating Java's exception handling principles.
-Our third contribution was the first exploratory to assess the 
-developers perspective concerning the exception handling 
+Our third contribution was the first exploratory study to assess the 
+developers' perspective concerning the exception handling 
 bug hazards in Android development.

### 2015-10-11 SPRINGER_JOURNAL/msr_journal.tex
Included some qualitative findings in the conclusion.
Sharpened the conclusion a bit, most notably to establish the
link between the empirical and qualitative studies.

Also included main *outcomes* of the qualitative study.

@@
 application robustness; and (ii) to assess the developers' perspective
-concerning the detected EH \emph{bug hazards}.
-
-To that end, we mined the stack 
+concerning the detected exception handling bug hazards.
+
+To realize this goal, we mined the stack 
 traces embedded in all issues defined in 482 Android projects hosted on GitHub and 
 157 projects hosted on Google Code -- overall considering 6,005
-exception stack traces. We also surveyed developers associated with some of 
+exception stack traces. We subsequently surveyed developers associated with 
+a selection of 
 the mined GitHub projects.
+
 Our first key contribution is
@@
   issues.
+
 Our second contribution is
 an empirical study of over 6000 actual stack traces,
-  demonstrating that (1) half of the system crashes are due to errors
+  demonstrating that:
+\begin{enumerate} 
+  \item Half of the system crashes are due to errors
   in programming logic, with null pointer exceptions being most
   prominent;
-  (2) documentation for explicitly thrown RuntimeExceptions is almost
+  \item Documentation for explicitly thrown RuntimeExceptions is almost
   never provided; 
-  (3) extensive use of wrapping leads to hard-to-understand chains
+  \item Extensive use of wrapping leads to hard-to-understand chains
   violating Java's exception handling principles.
-Our third contribution was the first exploratory study to assess the 
+\end{enumerate}
+
+Our third contribution is a qualitative study to assess the 
 developers' perspective concerning the exception handling 
 bug hazards in Android development.
- Our results shed light on common problems and bug hazards in Java
+This study corroborates the findings of our stack trace analysis, most notably the prevalence
+of null pointer exceptions and the reliability implications of (in particular cross-type) wrappings.
+Furthermore, we found that few developers are aware of the undocumented checked exceptions signaled by native C code of the Android platform.
+
+In conclusion, our findings shed light on common problems and bug hazards in Java
 exception handling code, and call for tool support to help developers

### 2015-10-11 SPRINGER_JOURNAL/msr_journal.tex
Using 'url' package and \url{...} for urls.
So that they are broken nicely on new lines.

@@
 \usepackage[numbers]{natbib}
+\usepackage{url}
 
@@
 Google Code contains widely used open-source Android apps (e.g.
-K9Mail\footnote{K9Mail moved to GitHub but as a way of not loosing the project history it advises their users to report bugs in the Google Code issue tracker: https://github.com/k9mail/k-9/wiki/LoggingErrors.}).
+K9Mail\footnote{K9Mail moved to GitHub but as a way of not loosing the project history it advises their users to report bugs in the Google Code issue tracker: \url{https://github.com/k9mail/k-9/wiki/LoggingErrors.}}).
 However, differently from GitHub, Google Code does not provide an API to access the information related
@@
 For instance:\emph{``Don't throw [pure] RuntimeException (in most cases), don't catch OutOfMemoryError, etc." [D42]}.
-Some of the practices mentioned, however, are not well known Java EH best practices but seem to be used by some Android developers to better deal with exceptions in the Android app context, such as the use of a crash fast approach\footnote{http://www.slideshare.net/pyricau/crash} -- \emph{``Crash is good for hinting developer.
+Some of the practices mentioned, however, are not well known Java EH best practices but seem to be used by some Android developers to better deal with exceptions in the Android app context, such as the use of a crash fast approach\footnote{\url{http://www.slideshare.net/pyricau/crash}} -- \emph{``Crash is good for hinting developer.
 Fast crash fast solve." [D51]}; and to favor checked exceptions instead of runtime to represent the exceptional conditions in Android apps -- \emph{``RuntimeExceptions are not particularly well suited for Android's robustness, as any uncaught one would cause a crash." [D26]}; \emph{``Checked exception help, unchecked exceptions don't." [D25]}; \emph{``avoid unchecked exceptions except if the case is really abnormal and should not happen [...]"[D45]}.
@@
 Most of them (43\%) mentioned that they applied the same best practices that they used in Java programs.
-Only some of them mentioned that they adopted specific best practices such as: (i) the use of crash report tools (21\%) -- to notify the developer about the uncaught exceptions that happen in the application code; and (ii) the use of a global exception handler (14\%) -- the UncaughtExceptionHandler advised in the Android documentation\footnote{http://developer.android.com/reference/java/lang/Thread.UncaughtExceptionHandler.html}.
+Only some of them mentioned that they adopted specific best practices such as: (i) the use of crash report tools (21\%) -- to notify the developer about the uncaught exceptions that happen in the application code; and (ii) the use of a global exception handler (14\%) -- the UncaughtExceptionHandler advised in the Android documentation\footnote{\url{http://developer.android.com/reference/java/lang/Thread.UncaughtExceptionHandler.html}}.
 
@@
 \emph{``Giving the user meaningful feed back is important, so the more specific you can be about what went wrong, the better."[D66]}.
-Some developers also mentioned to use a toast to support this task\footnote{A toast is an Android component that provides simple feedback about an operation in a small popup -- http://developer.android.com/guide/topics/ui/notifiers/toasts.html}.
+Some developers also mentioned to use a toast to support this task\footnote{A toast is an Android component that provides simple feedback about an operation in a small popup -- \url{http://developer.android.com/guide/topics/ui/notifiers/toasts.html}}.
 Few developers also mentioned that the runtime exception should flow upstream to the framework so it could crash the app.
@@
 \section{Replication Package}
-All the data used in the mining study and in the survey-based study is publicly available at the ExceptionMiner tool website hosted on GitHub\footnote{ https://github.com/souzacoelho/exceptionminer}.
+All the data used in the mining study and in the survey-based study is publicly available at the ExceptionMiner tool website hosted on GitHub\footnote{\url{https://github.com/souzacoelho/exceptionminer}}.
 Specifically we provide: (i) all issues related to Android projects found

### 2015-10-11 SPRINGER_JOURNAL/msr_journal.tex
Rewrite of abstract.

@@
 \begin{abstract}
-This paper reports two studies:  a repository mining study whose goal was to identify exception handling bug hazards in Android apps; and a survey-based study whose goal was to assess the perspectives of developers concerning such bug hazards.
+% Abstract somewhat based on Nature guidelines, 
+% https://twitter.com/TrevorABranch/status/620699527486373888/photo/1?ref_src=twsrc%5Etfw
+%
+% Basic intro (1-2 sentences)
+Adequate handling of exceptions has proven difficult for many software engineers.
+% More detailed background (2-3 sentences)
+Mobile app developers in particular, have to cope with 
+compatibility, middleware, memory constraints, and battery restrictions.
+% Clear sentence with general problem studied (1 sentence)
+The goal of this paper is to obtain a thorough understanding of common exception handling
+\emph{bug hazards} that app developers face.
+% (method 1)
+To that end, we first provide a detailed empirical study of over 6,000 
+Java exception stack traces 
+we extracted from over 600 open source Android projects.
+% Main result 1 / new knowledge
+Key insights from this study include common causes for system crashes, and
+common chains of wrappings between checked and unchecked exceptions.
+% (method 2)
+Furthermore, we provide a survey with 70 developers involved
+in at least one of the projects analyzed.
+The results corroborate the stack trace findings, and indicate
+that developers are unaware of frequently occurring undocumented exception handling behavior.
+% General context and implications
+Overall, the findings of our study call for tool support to help developers
+understand their own and third party exception handling and wrapping logic.
+
+%This paper reports two studies:  a repository mining study whose goal was to identify exception handling bug hazards in Android apps; and a survey-based study whose goal was to assess the perspectives of developers concerning such bug hazards.
 The repository mining study mined the exception stack traces embedded in 159,048 issues reported in Android projects hosted on GitHub (482 projs.)  and Google Code (157 projs.).

### 2015-10-12 SPRINGER_JOURNAL/msr_journal.tex
addressing Christpph and Arie's suggestions

@@
 Moreover, such \emph{bug hazards} 
-call for improvements of languages (e.g. to prevent null pointer dereferences) and tools to better support exception handling in Android and Java environments.
+call for improvements of languages (e.g. to prevent null pointer dereferences) and tools
+ to better support exception handling in Android and Java environments.
+
+%Section~\ref{sec:exceptionminer} 
+%Section~\ref{sec:result}
 
@@
 background on the Android platform and the Java exception model. 
-Section~\ref{sec:study} presents the study design. 
-Section~\ref{sec:exceptionminer} describes the ExceptionMiner tool we developed to conduct our study.
-Section~\ref{sec:result} reports the findings of our study.
+Section~\ref{sec:study} presents the mining study design, describes the ExceptionMiner tool 
+we developed to conduct our study, and reports the study results.
+Section~\ref{sec:dev} details the exploratory survey.
 Section~\ref{sec:disc} provides a discussion of the wider implications of our results,
  presents the threats to validity associated with the mining study and discusses the limitations
-of the survey-based study. 
+of the survey-based study, and points to the replication. 
 Finally Section~\ref{sec:rele} describes related work, 
@@
   \end{tabular}
-\caption{Root Exceptions occurrences and popularity in repositories hosted on Google Code (GC) and GitHub(GH).}
+\caption{Root Exceptions occurrences and popularity in analyzed repositories.}
 \label{tab:topten}
@@
 We emailed the 1,824 developers and received 71 valid answers.
-The majority of our respondents have more than 2 years of Java development experience (87.3\%) and  considerable experience in Android development (more than 1 year)  (85.9\%) -- see Figure~\ref{fig:devexpertise}.
+The majority of our respondents have more than 2 years of Java development experience (87.3\%) and of Android development (85.9\%) -- see Figure~\ref{fig:devexpertise}.
 
@@
 \bigskip
-
-\section{Replication Package}
-All the data used in the mining study and in the survey-based study is publicly available at the ExceptionMiner tool website hosted on GitHub\footnote{\url{https://github.com/souzacoelho/exceptionminer}}.
-Specifically we provide: (i) all issues related to Android projects found
-on GitHub and Google Code used in this study; (ii) all stack traces extracted
-from issues; (iii) the results of manual inspection steps; (iv) the
-ExceptionMiner tool we developed to support stack trace extraction and distilling; (v) the survey questions; and (iv) the survey responses and summary.
 
@@
 
-%\noindent
+\subsection{Replication Package}
+All the data used in the mining study and in the survey-based study is publicly 
+available at the ExceptionMiner tool website hosted on GitHub\footnote{\url{https://github.com/souzacoelho/exceptionminer}}.
+Specifically we provide: (i) all issues related to Android projects found
+on GitHub and Google Code used in this study; (ii) all stack traces extracted
+from issues; (iii) the results of manual inspection steps; (iv) the
+ExceptionMiner tool we developed to support stack trace extraction and distilling; 
+(v) the survey questions; and (iv) the survey responses and summary.
 

### 2015-10-12 SPRINGER_JOURNAL/msr_journal.tex
addressing Christpph and Arie's suggestions

@@
 % (method 2)
-Furthermore, we provide a survey with 70 developers involved
+Furthermore, we provide a survey with 71 developers involved
 in at least one of the projects analyzed.
@@
 
-The exploratory survey was conducted with 70 Android developers involved in one or more of the 
+The exploratory survey was conducted with 71 Android developers involved in one or more of the 
 GitHub Android projects whose issues were mined. This survey reveals that only few developers (3\% of respondents)
@@
 For instance, Kechagia and Spinellis showed that the NullPointerException was the  
-most reported exception in the crash reports sent to BugSense (a bug report 
-management service for Android applications)  \cite{kechagia2014}.
+most reported exception in the crash reports sent to BugSense~\footnote{https://www.bugsense.com/} (a bug report 
+management service for Android applications)~\cite{kechagia2014}.
 Other research on robustness testing~\cite{maji2012empirical,csallner2004jcrasher} shows that most of the automatically 
@@
 refers to a programming mistake according to its Javadoc; and (ii) the Java specification lists all 
-exceptions related to backward compatibility~\cite{javaback}, such as
+exceptions related to backward compatibility~\footnote{http://docs. oracle.com/javase/specs/jls/se7/html/jls-13.html}, such as
 InstantiationError, VerifyError, and IllegalAccessError.
@@
 Developers were also asked about what they usually do to prevent NullPointerExceptions from happening when their app is in production.
-Most of them mentioned including null-checks in the application code (59\%), mostly when checking the return of a method and as method guard conditions.
+Table~ref{tab:preventnull} presents the ways for preventing NullPointerException in Android apps most cited by developers.
+Most of the developers mentioned including null-checks in the application code (59\%), mostly when checking the return of a method and as method guard conditions.
 Many respondents (39\%) also answered that instead of just sprinkling the code with null-checks when a NullPointerException happens they investigate the cause of the null reference and work to fix it, instead of just adding a null check.
@@
 This question was motivated by the fact that during the mining study we could identify that many checked exceptions were just wrapped in a runtime exception and re-thrown, as a way to bypass any kind of handling.
-When asked about how to deal with a checked exception signaled by a method (invoked in the context of the onPause() activity method), most respondents (63\%) said that they should add a try-catch block surrounding the method invocation.
+When asked about how to deal with a checked exception signaled by a method (invoked in the context of the onPause() activity method), most respondents (63\%) said that they should add a try-catch block surrounding the method invocation - see Table~\ref{tab:handlingchecked}.
 One of the respondents mentioned that \emph{``If a method signature declares that it throws an exception, it means that the caller should handle the exception."[D29]}; \emph{``because it is checked exception, you should expect it to happen rather than just crash"[D70]}.
-However, most of the respondents also said that to deal with a method that is signaling a runtime exception we should surround such a method with a try-catch block (59\%).
+However, most of the respondents also said that to deal with a method that is signaling a runtime exception we should surround such a method with a try-catch block (59\%) - see Table~\ref{tab:handlingruntime}.
 
@@
  Although there is a long lasting debate about the pros and cons of checked and unchecked exceptions, the survey revealed that many Android developers considered checked exceptions as a way of using exceptions that can prevent uncaught exception crashes -- since in order for a checked exception to flow upstream and crash the app the developers need to explicitly do so.
+
+\begin{table}
+\scriptsize
+\centering
+\begin{tabular}{lrr}
+\hline
+\bfseries{Top Ways of Handling Checked Exceptions} & \bfseries{\#} & \bfseries{\%} \\
+\hline
+Add a try-catch block	& 25 &	63\% \\
+Present error message	& 7 &	18\% \\
+Involve the user in solving	& 5 &	13\% \\
+Retry  &	5 &	13\% \\
+Investigate the cause &	3 &	8\% \\
+Same way as runtime	& 3 &	8\% \\
+Add a throws declaration / pass upstream &	2 &	5\% \\
+%Swallow	& 1 &	3\% \\
+%Toast	 & 1 & 3\% \\
+\hline
+\end{tabular}
+\caption{Top ways of dealing with a checked exception signaled in the context of an Activity method -- 40 non-empty responses. }
+\label{tab:handlingchecked}
+\end{table}
 
@@
 \caption{Top ways of handling runtime exceptions -- 63 non-empty responses. }
-\label{tab:handlingruntime}
-\end{table}
-
-\begin{table}
-\scriptsize
-\centering
-\begin{tabular}{lrr}
-\hline
-\bfseries{Top Ways of Handling Checked Exceptions} & \bfseries{\#} & \bfseries{\%} \\
-\hline
-Add a try-catch block	& 25 &	63\% \\
-Present error message	& 7 &	18\% \\
-Involve the user in solving	& 5 &	13\% \\
-Retry  &	5 &	13\% \\
-Investigate the cause &	3 &	8\% \\
-Same way as runtime	& 3 &	8\% \\
-Add a throws declaration / pass upstream &	2 &	5\% \\
-%Swallow	& 1 &	3\% \\
-%Toast	 & 1 & 3\% \\
-\hline
-\end{tabular}
-\caption{Top ways of dealing with a checked exception signaled in the context of an Activity method -- 40 non-empty responses. }
 \label{tab:handlingruntime}
@@
 Finally, crash report tools (e.g.
-ACRA~\cite{Acra14}) were reported as a way to prevent crashes.
+ACRA~\footnote{code.google.com/p/acra/}) were reported as a way to prevent crashes.
 Once such tools notify developers about current crashes, the application can be fixed and hence future crashes (caused by what was fixed) could be prevented.
@@
 
+%~\cite{Acra14}
 \bigskip
@@
 set of crash data per application. Although some services exist 
-to collect crash data from mobile applications~\cite{BugSe14,BugSn14,Googl14,Acra14},
+to collect crash data from mobile applications (e.g., ACRA~\footnote{code.google.com/p/acra/},  Google analytics~\footnote{https://www.google.com/analytics/}, Bugsense, Bugsnag~\footnote{https://bugsnag.com/}),
 they do not provide open access to the crash reports of their client applications.
@@
 included in bug issues.  In a similar study, Bettenburg et
-al.~\cite{bettenburg2008makes} identify stack traces as the second most stack
-trace feature for developers.  Sinha et al.~\cite{sinha2009fault} proposed an
+al.~\cite{bettenburg2008makes} identify stack traces as the second most relevant
+ feature of good bug reports.  Sinha et al.~\cite{sinha2009fault} proposed an
 approach that uses stack traces to guide a dataflow analysis for locating and

### 2015-10-14 SPRINGER_JOURNAL/msr_journal.tex
typo
Small sentence refactorings.
Mostly to make the text somewhat shorter / more to the point.

@@
 
-Although the mining study revealed a set of bug hazards in Android exception handling, no work has examined yet what the developers' perceptions about them are.
+%Although the mining study revealed a set of bug hazards in Android exception handling, no work has examined yet what the developers' perceptions about them are.
 With Android development rapidly rising in popularity, it is important to better understand how developers deal with such bug hazards and what they do to overcome them.
@@
 The questions were motivated by the analysis of the existing bug hazards detected in the first phase of this study.
-The next sections detail the survey.
+%The next sections detail the survey.
 
@@
 
-The general research question guiding our exploratory survey was the following: \emph{RQ 2: How do developers deal with the exception handling code in Android apps and what are the developers' perspectives about the main bug hazards found during the mining study?} This general research question was broken into a set of research questions that are answered by the exploratory survey.
+The general research question guiding our exploratory survey is
+\emph{RQ 2: How do developers deal with the exception handling code in Android apps and what are the developers' perspectives about the main bug hazards found during the mining study?} This general research question has been broken into a set of research questions that are answered by the exploratory survey.
 
@@
 
-To make the analysis easier, we further refine this question into sub-questions, such as: \emph{How often do developers handle exceptions?}; \emph{How often do developers throw exceptions?}.
-Moreover, another question that shows us how developers may deal with the exception handling (EH) code is: \emph{Do developers know about Java EH best practices and/or Android specific EH best practices?}
+To make the analysis easier, we further refine this question into sub-questions, such as: \emph{How often do developers handle exceptions?} \emph{How often do developers throw exceptions?} And, to show us how developers may deal with the exception handling (EH) code: \emph{Do developers know about Java EH best practices and/or Android specific EH best practices?}
 
@@
 
-Some of these questions were indirectly answered by presenting a set of code snippets in which
-the bug hazards were present and asking developers what they would do when faced with such scenarios.
-
-Finally developers were asked whether the exception handling code helped with the development of robust Android applications, and what they usually do to prevent apps from crashing.
-Hence, the last research question that guided this survey was the following:
+We address these questions by presenting a set of code snippets in which
+the bug hazards are present, asking developers what they would do when faced with such scenarios.
+
+Our last research question addresses
+whether the exception handling code helps with the development of robust Android applications, and what developers usually do to prevent apps from crashing:
 
@@
 Since our aim is to learn from a large number of developers, we use surveys which scale well.
-The survey is split into two logical sections; multiple choice or Likert-scale questions and open-ended questions.
+The survey consists of multiple choice or Likert-scale questions (to force participants to make a choice) intermixed with open-ended questions to further elicit the developer's opinions.
+%The survey is split into two logical sections; multiple choice or Likert-scale questions and open-ended questions.
 The multiple choice questions are intermixed with open-ended questions to further elicit the developer's opinions.
 Moreover, the survey also contains Likert-scale questions to force participants to make a choice.
-Overall, the survey includes 13 open-ended questions, 5 Likert-scale questions and 10 multiple choice questions.
-The respondents could complete the survey in about 15 minutes.
+Overall, the survey includes 13 open-ended questions, 5 Likert-scale questions and 10 multiple choice questions. The respondents could complete the survey in about 15 minutes.
 
 We used grounded theory coding~\cite{charmaz2006}, iterating through our open-ended survey responses, to answer some of the research questions.
-The grounded theory coding used in this study consists of two phases: (1) initial coding entails a close reading of the data and (2) later we use focused coding~\cite{charmaz2006} to pinpoint and develop the most salient themes~\cite{charmaz2006} in the analyzed data.
+The grounded theory coding used in this study consists of two phases: (1) initial coding entailing a close reading of the data and (2) focused coding~\cite{charmaz2006} to pinpoint and develop the most salient themes~\cite{charmaz2006} in the analyzed data.
 The coding process was performed by two coders who agreed upon the list of themes that emerged from the answer set of each question.
@@
 Developers were also asked about what they usually do to prevent NullPointerExceptions from happening when their app is in production.
-Table~ref{tab:preventnull} presents the ways for preventing NullPointerException in Android apps most cited by developers.
+Table~\ref{tab:preventnull} presents the ways for preventing NullPointerException in Android apps most cited by developers.
 Most of the developers mentioned including null-checks in the application code (59\%), mostly when checking the return of a method and as method guard conditions.

### 2016-04-13 SPRINGER_JOURNAL/msr_journal.tex
addressing journal reviewers comments.

@@
 
-To better understand the developers' perspective about the bug hazards detected during the mining study, we set up an exploratory qualitative investigation and surveyed Android developers on how they perceive such bug hazards in their projects.
-The scope of our study is GitHub---using our GHTorrent database~\cite{Gousi13}, we aimed our survey at developers from Android projects available in GitHub for which issues were mined during the first phase of this study.
-
-Our main goal is to learn from developers about the Java exception handling usage and the main exception handling bug hazards in Android.
-We therefore use a survey as our main research instrument.
-The survey mixed  Liket-scale with open-ended questions.
-The questions were motivated by the analysis of the existing bug hazards detected in the first phase of this study.
+\textcolor{red}{As previously menitoned, the goal of this work is to obtain a thorough understanding of common exception handling \emph{bug hazards} that app developers face.
+In the first phase of this work, we conducted a mining study which identified common exception handling bug hazards in app development.
+ In the second phase of this work, we set up an exploratory qualitative investigation and surveyed Android developers on how they perceive the bug hazards detected in the mining study.
+The scope of our study is GitHub---using our GHTorrent database~\cite{Gousi13}, we aimed our survey at developers from Android projects available in GitHub for which issues were mined during the first phase of this study.}
+
+\textcolor{red}{The survey mixed Liket-scale with open-ended questions, and the motivations behind the survey questions are twofold: (i) to learn from developers about the usage of the exception handling code in app development;  and (ii) to get developers percerptions about the bug hazards detected in the first phase of this study.}
+
+%Our main goal is to learn from developers about the Java exception handling usage and the main 
+%exception handling bug hazards in Android. We therefore use a survey as our main research 
+%instrument. The survey mixed  Liket-scale with open-ended questions. The questions were motivated 
+%by the analysis of the existing bug hazards detected in the first phase of this study. 
 %The next sections detail the survey.

### 2016-04-13 SPRINGER_JOURNAL/msr_journal.tex
addressing journal reviewers comments on protocol

@@
 %The next sections detail the survey.
-
 %Our main findings reveal that developers often handle exceptions while developing Android apps and 
@@
 
-The general research question guiding our exploratory survey is
-\emph{RQ 2: How do developers deal with the exception handling code in Android apps and what are the developers' perspectives about the main bug hazards found during the mining study?} This general research question has been broken into a set of research questions that are answered by the exploratory survey.
+\textcolor{red}{Hence, the general research question guiding our exploratory survey is twofold \emph{RQ 2: How do developers deal with the exception handling code in Android apps and what are the developers' perspectives about the main bug hazards found during the mining study?} This general research question has been broken into a set of research questions that are answered by the exploratory survey.}
 
@@
 
-We used grounded theory coding~\cite{charmaz2006}, iterating through our open-ended survey responses, to answer some of the research questions.
-The grounded theory coding used in this study consists of two phases: (1) initial coding entailing a close reading of the data and (2) focused coding~\cite{charmaz2006} to pinpoint and develop the most salient themes~\cite{charmaz2006} in the analyzed data.
-The coding process was performed by two coders who agreed upon the list of themes that emerged from the answer set of each question.
+\textcolor{red}{Given the exploratory nature of our research, we used methods from Grounded Theory ~\cite{charmaz2006} to answer some of our research questions by iterating through the open-ended survey responses.
+The grounded theory coding used in this study consists of at two phases: (1) initial coding entails a close reading of the data and (2) later we used focused coding ~\cite{charmaz2006} to pinpoint and develop the most salient themes ~\cite{charmaz2006} in the analyzed data.
+Answers to different questions in the survey were coded separately.
+For all open-ended questions in the survey, coding was done by two researchers until saturation was reached.
+Disagreements were discussed and resolved as part of the coding, thus, we are unable to report level of agreement.
+In all cases, the first author was one of the two coders, and one of the other authors coded the data as well in close collaboration with the first author until saturation was reached.
+The first author then went through the remaining responses to assign codes to responses.}
+
+%We used grounded theory coding~\cite{charmaz2006}, iterating through our open-ended survey re% sponses, to answer some of the research questions. The 
+%grounded theory coding used in this study consists of two phases: (1) initial coding entailing a close reading of the data and (2) focused 
+%coding~\cite{charmaz2006} to pinpoint and develop the most salient themes~\cite{charmaz2006} in the analyzed data. The coding process was performed 
+%by two coders who agreed upon the list of themes that emerged from the answer set of each question.
 

### 2016-04-14 SPRINGER_JOURNAL/msr_journal.tex
addressing journal reviewers comments - REVIWER 1

@@
 
-\textcolor{red}{The survey mixed Liket-scale with open-ended questions, and the motivations behind the survey questions are twofold: (i) to learn from developers about the usage of the exception handling code in app development;  and (ii) to get developers percerptions about the bug hazards detected in the first phase of this study.}
+%\textcolor{red}{The survey mixed Liket-scale with open-ended questions, and the motivations behind the survey questions are twofold: (i) to learn from 
+%developers about the usage of the exception handling code in app development;  and (ii) to get developers percerptions about the bug hazards detected in 
+%the first phase of this study.}
 
@@
 
-\textcolor{red}{Hence, the general research question guiding our exploratory survey is twofold \emph{RQ 2: How do developers deal with the exception handling code in Android apps and what are the developers' perspectives about the main bug hazards found during the mining study?} This general research question has been broken into a set of research questions that are answered by the exploratory survey.}
+\textcolor{red}{The general research question guiding our exploratory survey is twofold \emph{RQ 2: How do developers deal with the exception handling code in Android apps and what are the developers' perspectives about the main bug hazards found during the mining study?} This general research question has been broken into a set of research questions that are answered by the exploratory survey.}
 
@@
 
-Since our aim is to learn from a large number of developers, we use surveys which scale well.
-The survey consists of multiple choice or Likert-scale questions (to force participants to make a choice) intermixed with open-ended questions to further elicit the developer's opinions.
-%The survey is split into two logical sections; multiple choice or Likert-scale questions and open-ended questions.
+\textcolor{red}{Since our aim is to learn from a large number of developers, we use surveys which scale well.}
+\textcolor{red}{The survey is split into two logical sections; the motaivations behind each logical section are: 
+(i) questions aiming at learning from developers about the usage of the exception handling code in app development;  and 
+(ii) questions focusing on getting developers percerptionds about the bug hazards detected in the first phase of this study.
+The survey comprises multiple choice or Likert-scale questions and open-ended questions.
 The multiple choice questions are intermixed with open-ended questions to further elicit the developer's opinions.
 Moreover, the survey also contains Likert-scale questions to force participants to make a choice.
-Overall, the survey includes 13 open-ended questions, 5 Likert-scale questions and 10 multiple choice questions. The respondents could complete the survey in about 15 minutes.
+Overall, the survey includes 13 open-ended questions, 5 Likert-scale questions and 10 multiple choice questions.
+The respondents could complete the survey in about 15 minutes.}
 
@@
 The first author then went through the remaining responses to assign codes to responses.}
-
-%We used grounded theory coding~\cite{charmaz2006}, iterating through our open-ended survey re% sponses, to answer some of the research questions. The 
-%grounded theory coding used in this study consists of two phases: (1) initial coding entailing a close reading of the data and (2) focused 
-%coding~\cite{charmaz2006} to pinpoint and develop the most salient themes~\cite{charmaz2006} in the analyzed data. The coding process was performed 
-%by two coders who agreed upon the list of themes that emerged from the answer set of each question.
 
@@
 For instance:\emph{``Don't throw [pure] RuntimeException (in most cases), don't catch OutOfMemoryError, etc." [D42]}.
-Some of the practices mentioned, however, are not well known Java EH best practices but seem to be used by some Android developers to better deal with exceptions in the Android app context, such as the use of a crash fast approach\footnote{\url{http://www.slideshare.net/pyricau/crash}} -- \emph{``Crash is good for hinting developer.
-Fast crash fast solve." [D51]}; and to favor checked exceptions instead of runtime to represent the exceptional conditions in Android apps -- \emph{``RuntimeExceptions are not particularly well suited for Android's robustness, as any uncaught one would cause a crash." [D26]}; \emph{``Checked exception help, unchecked exceptions don't." [D25]}; \emph{``avoid unchecked exceptions except if the case is really abnormal and should not happen [...]"[D45]}.
+Some of the practices mentioned, however, are not well known Java EH best practices but seem to be used by some Android developers to better deal with exceptions in the Android app context, such as the use of a crash fast approach\footnote{\url{http://www.slideshare.net/pyricau/crash-fast}} -- \emph{``Crash is good for hinting developer.
+Fast crash fast solve." [D51]}.
+\textcolor{red}{According to this approach, developers should let the application to crash as soon as an unexpected situation happens.
+The motivation behind this approach is that the developer will receive the crash information and fix the potential bug that caused it.
+On the other hand, if the developer lets the app to continue runing, after an exception had happened, the aplication may enter an inconsistent state.
+In such scenario, the time between the fault being exercised and the failure (the fault manifestation) increases, which may impair finding and fixing the potential bug that caused the failure.} Moreover, some respondents favor checked exceptions instead of runtime to represent the exceptional conditions in Android apps -- \emph{``RuntimeExceptions are not particularly well suited for Android's robustness, as any uncaught one would cause a crash." [D26]}; \emph{``Checked exception help, unchecked exceptions don't." [D25]}; \emph{``avoid unchecked exceptions except if the case is really abnormal and should not happen [...]"[D45]}.
 
@@
 
+\textcolor{red}{This section discusses lessons learned from the mining and the survey studies. The mining study revealed a set of bug 
+hazards -- such as (i) the cross-type wrappings; (ii) the abundance of null pointer problems; and (iii) the undocumented runtime 
+exceptions signaled by third-party code -- which were confirmed by developers during the survey.
+Next subsections discuss about each of such bug hazards, presenting how the developer's perceived them and pointing to what can be done to help developers
+to deal with them.}
+
 %\emph{\textbf{The exception handling confusion problem.}}
@@
 types from the total of 100); the interrater agreement was high (96\%).
+
+\textcolor{red}{Another internal threat to validity of this study regards to the process adopted to 
+identify the type of each exception reported on stack traces. 
+To perform such type analysis we chose a version of the souce-code or bytecode
+ of each reported exception, as follows: (i) for all exceptions signaled by the Android Platform 
+and the Java Environment the analysis was based on Version 4.4 of Android platform (API level 19);  (ii) 
+for the exceptions signaled by applications the analysis considered
+the last version of the application source-code available on the Github/Google code; and (iii) 
+for all exceptions signaled by third-party libraries the type analysis
+considered the last bytecode version available on the app repository.
+When the exception could not be found automatically or manually (based on the chosen version), 
+we classified the exception as ``Undefined''.  Only 31 exceptions 
+remained undefined, which occurred in 60 different exception stack traces out of 6,005 mined stack traces.
+Hence, the exception type could not be identified only in 1\% of the mined flows.
+There should be situations where the exception type (i.e., checked or runtime) changes from version to another. 
+Altough we believe that it will not happen very often, we did not investigate such issue.
+ One way to mitigate such threat would be to consider the issue date to select the most 
+suitable source-code/bytecode version to perform the analysis.
+However, the number of applications analyzed (approx. 6000) 
+would increase the time and complexity needed to follow such improved process.}
 

### 2016-04-16 SPRINGER_JOURNAL/msr_journal.tex
addressing journal reviewers comments - REVIWER 1 more

@@
 
-\begin{figure} \centering \includegraphics[scale=0.60]{yes_no_new2.png}
-\caption{Summary of some survey question answers. }
+%\begin{figure} \centering \includegraphics[scale=0.60]{yes_no_new2.png}
+\begin{figure} \centering \includegraphics[scale=0.60]{yes_no_best_new.png}
+\caption{Summary of the survey questions related to EH best practices. }
 \label{fig:allquest}
@@
 %//////NULLPONTER WORSE IN ANDROID /////
-Developers were asked if their apps ever crashed because of a NullPointerException -- see Figure~\ref{fig:allquest}.
+Developers were asked if their apps ever crashed because of a NullPointerException -- see Figure~\ref{fig:questnullrobust}.
 The vast majority of respondents (96\%) said that their apps crashed at least once due to a NullPointerException.
 Their answers are aligned with the high prevalence of NullPointerExceptions found in the mining study -- in Table~\ref{tab:topten} in Section~\ref{sec:result} we can see that NullPointerExceptions were the main cause of the mined exception stack traces.
+
+%\begin{figure} \centering \includegraphics[scale=0.60]{yes_no_new2.png}
+\begin{figure} \centering \includegraphics[scale=0.60]{yes_no_null_robust_new.png}
+\caption{Summary of the survey questions relates to the null pointer problem (RQ 2.2) and the effect of EH on app robustness (RQ 2.5). }
+\label{fig:questnullrobust}
+\end{figure}
+
 Developers reported some reasons for the high prevalence of crashes caused by NullPointerExceptions.
@@
 
-Developers were also asked whether the exception handling code helped with the development of robust Android applications -- see Figure~\ref{fig:allquest}.
+Developers were also asked whether the exception handling code helped with the development of robust Android applications -- see Figure~\ref{fig:questnullrobust}.
 Although most of the developers answered positively (78.9\%), when asked to provide one or more reasons for their answers, most of the respondents also provided reasons why the exception handling code can sometimes impair the robustness.
+
 Since positive and negative reasons were provided intermixed we analyzed both sets of answers and ranked the top reasons why exception handling can negatively or positively affect robustness according to developers' perceptions.
@@
 better understanding the developers' behavior and hence providing recommendations and 
-tools to help with specific development tasks. This was the first exploratory survey study whose
+tools to help with specific development tasks.
+
+Some surveys focused on Android developers such as xxxx.
+
+- P.S.Kochhar, F.Thung, N.Nagappan,T.
+Zimmermann,and D.
+Lo.
+Understanding the test automation culture of app developers.
+In 2015 IEEE 8th International Conference on Software Testing, Verification and Validation (ICST), pages 1-10, 2015.
+
+Kochhar et al conducted a survey-based study whose goal was to discover the commonly used tools for mobile app testing and the problems faced by developers while testing the apps.
+The survey was sent to 3,905 emails of Android developers and received 83 responses (response rate of 2.13%).
+
+- M. E. Joorabchi, A. Mesbah, and P. Kruchten. Real challenges in mobile apps. In
+ACM/IEEE International Symposium on Empirical Software Engineering and Measurement (ESEM'13), pages 15-24, 2013.
+
+  Joorabchi et al performed a study whose goal was to better understand the the main challenges developers face when building apps for different mobile devices.
+They interviewed 12 senior mobile developers from 9 different companies, and performed a semi-structured survey, with 188 respondents from the mobile development community.
+The survey link was sent to popular Mobile Development Meetup groups, LinkedIn groups related to native mobile development and shared through authors Twitter accounts.
+
+- M. Linares-Vásquez, C. Vendome, Q. Luo, and D. Poshyvanyk. How developers detect and fix performance bottlenecks in Android apps. In Proc. ICSME, pages 352-361. IEEE, 2015.
+
+ Linares-Vásquez surveyed 485 open source Android app and library developers (hosted on GitHub) to understand developers’ practices for detecting and fixing performance bottlenecks in mobile apps.
+This work emailed the survey to 24,340 emails and received 628 survey responses - response rate 2.6%.
+
+- Gabriele Bavota, Mario Linares Vásquez, Carlos Eduardo Bernal-Cárdenas, Massimiliano Di Penta, Rocco Oliveto, Denys Poshyvanyk: The Impact of API Change- and Fault-Proneness on the User Ratings of Android Apps.
+IEEE Trans.
+Software Eng.
+41(4): 384-407 (2015.
+
+ Bavota at all investigated the impact of API change-proness and fault-proneness on the user ratings of Android apps.
+This work surveyed 45 professional Android developers to assess developers perspective on whether such problems could be the cause for unfavorable user ratings.
+
+1,800 e-mail
+45 responses
+the response rate is 4 percent
+
+Even if the response rate achieved in our study is quite low it is similar to other surveys reported in the literature (e.g., [20], [21]).
+
+[20] A. J. Ko, R. DeLine, and G. Venolia, “Information needs in collocated
+software development teams,” in Proc. 29th Int. Conf. Softw.
+Eng., 2007, pp. 344–353.
+
+[21] A. Hindle, C. Bird, T. Zimmermann, and N. Nagappan, “Do topics
+make sense to managers and developers?” Empirical Softw. Eng.,
+pp. 1–37, 2014, http://dx.doi.org/10.1007/s10664-014-9312-1
+
+This was the first exploratory survey study whose
 goal was to assess developers' perspective regarding a set of exception handling bug hazards

### 2016-04-18 SPRINGER_JOURNAL/msr_journal.tex
addressing journal reviewers comments - REVIWER ***2***

@@
  signaled by  the runtime environment (due to programming mistakes) -- since these exceptions 
-should not be documented in the method signature.  Then we inspected
+should not be documented in the method signature\footnote{The filtering was performed in two steps:
+firstly, we analyzed the bytecode of the JVM and identified all runtime exceptions defined by it (e.g.,
+java.lang.NullPointerExceptoin,java.lang.ArrayIndexOutOfBounds), then
+if the libraries/framework method was signaling one of such exceptions it was filtered out from the analysis}.
+ Then we inspected
 the code for each method 
@@
 
+\emph{\textbf{Runtime Exception wrapping a Checked Exception.}} 
+This wrapping was responsible for 49.5\% of
+the cross-type wrappings.  From this set 50\%  were performed by methods defined by the Android platform. 
+We observe that this is a common implementation practice in the methods of the Android platform.
+According to Java best practices checked exceptions represent conditions from which the caller
+is expected to recover. By converting a checked exception in a general Runtime class,
+besides loosing contextual information about the exception,
+the client can simply ignore such exception - which can negatively impact the app robustness. 
+\textcolor{red}{Altough it is considered a bad practice, when such wrapping is accompained with a well defined 
+exception handling policy (which assures that a proper handling will be provided within the app)
+the impact on app robustness can be mitigated.}
+
+\bigskip
+
+\finding{%
+General catch clauses masking any exception into a general RuntimeException 
+is a common practice in the Android Platform; it violates best practice III 
+and is considered a bug hazard as it loses contextual information about the exception.
+}
+
 \emph{\textbf{Runtime Exception wrapping an Error.}} 
@@
 of an asynchronous task) into an instance of RuntimeException and re-throws it.  
-Table~\ref{tab:exampeswrap} presents examples of exceptions that were actually wrapped in this code snippet.  
-Such wrappings mask an unrecoverable Error into a general runtime exception.
+Table~\ref{tab:exampeswrap} presents examples of exceptions that 
+were actually wrapped in this code snippet: java.lang.RuntimeException wrapped
+an java.lang.OutOfMemoryError and a java.lang.StackOverflowError.
+
+\textcolor{red}{Such Errors result from failures detected by the runtime environment which indicate 
+resource deficiencies from which the program cannot possibly recover.
+Hence, `hiding" an unrecoverable condition into a runtime exception
+can lead to a scenario where the developer tries to handle such exception, leading the program
+to na unpredictable state. This kind of wrapping should be avoided, since represent a
+serious bug hazard.}
 
@@
 
-\emph{\textbf{Runtime Exception wrapping a Checked Exception.}} 
-This wrapping was responsible for 49.5\% of
-the cross-type wrappings.  From this set 50\%  were performed by methods defined by the Android platform. 
-We observe that this is a common implementation practice in the methods of the Android platform.
-However, using such a general exception is considered a bad practice
-according to the Java specification and common guidelines, as it loses contextual information about the exception.
-
-\bigskip
-
-\finding{%
-General catch clauses masking any exception into a general RuntimeException 
-is a common practice in the Android Platform; it violates best practice III 
-and is considered a bug hazard as it loses contextual information about the exception.
-}
-
 \bigskip
@@
 some of these wrappings -- some of them are masking an OutOfMemoryError
-into a checked exception. Such wrappings may also mask an unrecoverable error
-and may lead to the ``exception confusion'' described next.
+into a checked exception. 
+\textcolor{red}{By confronting the API user with a checked exception,
+the API designer is forcing the client to handle the exceptional condition. 
+But in this case the exception condition in and unrecoverable error,
+which makes infeasible the task of providing a proper handler. 
+This kind of wrapping is even more dangerous the previous one
+and may lead to the ``exception confusion'' described next.}
 
@@
 
-Some surveys focused on Android developers such as xxxx.
+Some of such exploratory surveys were applied to Android developers  [][][][]. Despite the low response rate
+of all these surverys (between 4\% and 2\%), these studies could get interesting insights about the state 
+of practice of Android development. Kochhar et al conducted a survey-based study to discover the 
+commonly used tools for mobile app testing as well as the problems 
+faced by developers while testing the apps. The survey was sent to 3,905 emails of Android developers 
+and received 83 responses (response rate of 2.13%). 
+Joorabchi et al. performed a survey-based study whose goal was to better understand the the main challenges 
+developers face when building apps for different mobile devices. 
+They interviewed 12 senior mobile developers and performed a semi-structured survey, with 188 respondents
+ from the mobile development community - since the survey was shared via e-mail groups and social media websites
+response rate of the survey could not be calculated. Linares-Vásquez et al. surveyed 485 open source Android 
+app and library developers (hosted on GitHub)  to understand developers’ practices for detecting and fixing 
+performance bottlenecks in mobile apps.  This work emailed the survey to 24,340 developers and received 
+628 responses - the response rate was 2.6\%.  
+Bavota at al. investigated the impact of API change-proness and fault-proneness on the user ratings of Android apps. 
+They surveyed developers to assess their perspective on whether such problems could be the cause for unfavorable user ratings. 
+The response rate was 4\%.
+None of these surveys, however, assed the how developers deal with exceptions in Android nor 
+the developers' perspective regarding a set of exception handling bug hazards. Our survey
+was the first exploratory study focusing on exception handling issues in Android development.
+
+Even if the response rate achieved in our study is quite low it is similar to other surveys reported in the literature (e.g., [20], [21]).
 
@@
 
-Kochhar et al conducted a survey-based study whose goal was to discover the commonly used tools for mobile app testing and the problems faced by developers while testing the apps.
+Kochhar et al conducted a survey-based study whose goal was to discover the commonly used tools for mobile app testing as well as the problems faced by developers while testing the apps.
 The survey was sent to 3,905 emails of Android developers and received 83 responses (response rate of 2.13%).

### 2016-04-19 SPRINGER_JOURNAL/msr_journal.tex
addressing journal reviewers comments - REVIWER ***2***

@@
 \hline
-Programming logic (java.lang and util) & 52,0\%\\
-Resources (IO, Memory, Battery) & 23,9\% \\
-Security & 4,1\%\\
-Concurrency & 2,9\% \\
-Backward compatibility & 5,5\% \\
-Specific Exceptions & 4,9\%\\
-General (Error, Exception, Runtime) & 6,7\%\\
+Programming logic (java.lang and util) & 52.0\%\\
+Resources (IO, Memory, Battery) & 23.9\% \\
+Security & 4.1\%\\
+Concurrency & 2.9\% \\
+Backward compatibility & 5.5\% \\
+Specific Exceptions & 4.9\%\\
+General (Error, Exception, Runtime) & 6.7\%\\
 \hline
@@
 
+\bigskip
+
 \emph{\textbf{Runtime Exception wrapping an Error.}} 
@@
 of an asynchronous task) into an instance of RuntimeException and re-throws it.
-Table~\ref{tab:exampeswrap} presents examples of exceptions that 
-were actually wrapped in this code snippet: java.lang.RuntimeException wrapped
-an java.lang.OutOfMemoryError and a java.lang.StackOverflowError.
-
-\textcolor{red}{Such Errors result from failures detected by the runtime environment which indicate 
-resource deficiencies from which the program cannot possibly recover.
-Hence, `hiding" an unrecoverable condition into a runtime exception
-can lead to a scenario where the developer tries to handle such exception, leading the program
-to na unpredictable state. This kind of wrapping should be avoided, since represent a
-serious bug hazard.}
 
@@
 %}
-
 {\footnotesize
@@
 
+Table~\ref{tab:exampeswrap} presents examples of exceptions that 
+were actually wrapped in this code snippet: java.lang.RuntimeException wrapped
+an java.lang.OutOfMemoryError and a java.lang.StackOverflowError. 
+\textcolor{red}{Such Errors result from failures detected by the runtime environment which indicate 
+resource deficiencies from which the program cannot possibly recover.
+Hence, `hiding" an unrecoverable condition into a runtime exception
+can lead to a scenario where the developer tries to handle such exception, leading the program
+to na unpredictable state. This kind of wrapping should be avoided, since represent a
+serious bug hazard.}
+
 \bigskip
@@
 
- In previous work~\cite{coelhoetal15}, we mined exception stack traces available in issues reported for several Android projects hosted on GitHub and Google Code.
+%~\cite{coelhoetal15}
+ In the first phase of this study, described previously, we mined exception stack traces available in issues reported for several Android projects hosted on GitHub and Google Code.
 To ensure that our sample consists of repositories that were real Android apps, we had to inspect every project site and discard toy-programs and non-Android repositories.
@@
 We emailed the 1,824 developers and received 71 valid answers.
+\textcolor{red}{The response rate was 3.9\% --
+although the response rate achieved in our study is low, it is inlined to similar surveys reported in the literature (e.g., ~\cite{kochhar2015understanding,joorabchi2013real,bavota2015impact,ko2007information,hindle2015topics}) whose response rates mostly varies between 2\% and 4\%.}
 The majority of our respondents have more than 2 years of Java development experience (87.3\%) and of Android development (85.9\%) -- see Figure~\ref{fig:devexpertise}.
@@
 %Q1Q2_new.png
-\begin{figure} \centering \includegraphics[scale=0.38]{expertise_new.png}
+\begin{figure} \centering \includegraphics[scale=0.38]{expertise_new_point.png}
 \caption{Developers expertise in Java and Android development.}
@@
 
-\begin{figure} \centering \includegraphics[scale=0.58]{agree_new2.png}
+%\begin{figure} \centering \includegraphics[scale=0.58]{agree_new2_point.png}
+\begin{figure} \centering \includegraphics[scale=0.58]{agree_new4_point.png}
 \caption{How often developers throw and handle exceptions during app development.}
@@
 tools to help with specific development tasks. 
-
-Some of such exploratory surveys were applied to Android developers  [][][][]. Despite the low response rate
-of all these surverys (between 4\% and 2\%), these studies could get interesting insights about the state 
-of practice of Android development. Kochhar et al conducted a survey-based study to discover the 
+ \textcolor{red}{Some of such exploratory surveys were applied to Android developers~\cite{kochhar2015understanding,joorabchi2013real,bavota2015impact,linares2015developers}. 
+Kochhar et al. ~\cite{kochhar2015understanding} conducted a survey-based study to discover the 
 commonly used tools for mobile app testing as well as the problems 
@@
 and received 83 responses (response rate of 2.13%). 
-Joorabchi et al. performed a survey-based study whose goal was to better understand the the main challenges 
+Joorabchi et al. ~\cite{joorabchi2013real} performed a survey-based study whose goal was to better understand the the main challenges 
 developers face when building apps for different mobile devices. 
@@
  from the mobile development community - since the survey was shared via e-mail groups and social media websites
-response rate of the survey could not be calculated. Linares-Vásquez et al. surveyed 485 open source Android 
+response rate of the survey could not be calculated. Linares-Vásquez et al. ~\cite{linares2015developers} surveyed 485 open source Android 
 app and library developers (hosted on GitHub)  to understand developers’ practices for detecting and fixing 
@@
 628 responses - the response rate was 2.6\%.  
-Bavota at al. investigated the impact of API change-proness and fault-proneness on the user ratings of Android apps. 
+Bavota at al. ~\cite{bavota2015impact} investigated the impact of API change-proness and fault-proneness on the user ratings of Android apps. 
 They surveyed developers to assess their perspective on whether such problems could be the cause for unfavorable user ratings. 
 The response rate was 4\%.
+A common characteristic among such works and our work is the low response rate of the surveys (between 4\% and 2\%).
 None of these surveys, however, assed the how developers deal with exceptions in Android nor 
-the developers' perspective regarding a set of exception handling bug hazards. Our survey
-was the first exploratory study focusing on exception handling issues in Android development.
-
-Even if the response rate achieved in our study is quite low it is similar to other surveys reported in the literature (e.g., [20], [21]).
-
-- P.S.Kochhar, F.Thung, N.Nagappan,T.
-Zimmermann,and D.
-Lo.
-Understanding the test automation culture of app developers.
-In 2015 IEEE 8th International Conference on Software Testing, Verification and Validation (ICST), pages 1-10, 2015.
-
-Kochhar et al conducted a survey-based study whose goal was to discover the commonly used tools for mobile app testing as well as the problems faced by developers while testing the apps.
-The survey was sent to 3,905 emails of Android developers and received 83 responses (response rate of 2.13%).
-
-- M. E. Joorabchi, A. Mesbah, and P. Kruchten. Real challenges in mobile apps. In
-ACM/IEEE International Symposium on Empirical Software Engineering and Measurement (ESEM'13), pages 15-24, 2013.
-
-  Joorabchi et al performed a study whose goal was to better understand the the main challenges developers face when building apps for different mobile devices.
-They interviewed 12 senior mobile developers from 9 different companies, and performed a semi-structured survey, with 188 respondents from the mobile development community.
-The survey link was sent to popular Mobile Development Meetup groups, LinkedIn groups related to native mobile development and shared through authors Twitter accounts.
-
-- M. Linares-Vásquez, C. Vendome, Q. Luo, and D. Poshyvanyk. How developers detect and fix performance bottlenecks in Android apps. In Proc. ICSME, pages 352-361. IEEE, 2015.
-
- Linares-Vásquez surveyed 485 open source Android app and library developers (hosted on GitHub) to understand developers’ practices for detecting and fixing performance bottlenecks in mobile apps.
-This work emailed the survey to 24,340 emails and received 628 survey responses - response rate 2.6%.
-
-- Gabriele Bavota, Mario Linares Vásquez, Carlos Eduardo Bernal-Cárdenas, Massimiliano Di Penta, Rocco Oliveto, Denys Poshyvanyk: The Impact of API Change- and Fault-Proneness on the User Ratings of Android Apps.
-IEEE Trans.
-Software Eng.
-41(4): 384-407 (2015.
-
- Bavota at all investigated the impact of API change-proness and fault-proneness on the user ratings of Android apps.
-This work surveyed 45 professional Android developers to assess developers perspective on whether such problems could be the cause for unfavorable user ratings.
-
-1,800 e-mail
-45 responses
-the response rate is 4 percent
-
-Even if the response rate achieved in our study is quite low it is similar to other surveys reported in the literature (e.g., [20], [21]).
-
-[20] A. J. Ko, R. DeLine, and G. Venolia, “Information needs in collocated
-software development teams,” in Proc. 29th Int. Conf. Softw.
-Eng., 2007, pp. 344–353.
-
-[21] A. Hindle, C. Bird, T. Zimmermann, and N. Nagappan, “Do topics
-make sense to managers and developers?” Empirical Softw. Eng.,
-pp. 1–37, 2014, http://dx.doi.org/10.1007/s10664-014-9312-1
-
-This was the first exploratory survey study whose
-goal was to assess developers' perspective regarding a set of exception handling bug hazards
-as well as how developers deal with exceptions and prevent crashes while developing.
+the developers' perspective regarding a set of exception handling bug hazards. In our study
+we conduced the first exploratory survey focusing on exception handling issues in Android development.}
 
@@
 \bibliographystyle{spbasic}      % basic style, author-year citations
-\bibliography{android-stacks}   % name your BibTeX data base
+\bibliography{msr_journal.bib}   % name your BibTeX data base
 

### 2016-04-19 SPRINGER_JOURNAL/msr_journal.tex
addressing journal reviewers comments - REVIWER 3
Remove build artifacts, add Makefile
A few corrections

@@
          Georgios Gousios   \at
-  \email{georgios@cs.ru.nl} \\    
+  \email{g.gousios@cs.ru.nl} \\
  \and
@@
 \begin{figure*}
-\centering \includegraphics[width=.9\hsize]{overview_review.png}
+%\centering \includegraphics[width=.9\hsize]{overview_review.png}
+\centering \includegraphics[width=1\hsize]{overview_review.png}
 \caption{Study overview.}\label{fig:overview}
@@
 
+%SELECT COUNT(*) FROM STACKTRACESUMMARY WHERE
+% repo !='android'  AND MAIN_EXTENDS LIKE 'RUNTIME%'
+%SELECT COUNT(*) FROM STACKTRACESUMMARY WHERE
+%repo !='android'  AND MAIN_EXTENDS LIKE 'RUNT%' AND MAINSIGNALER_FROM like 'APP%';
+
+\begin{table}
+\scriptsize
+\centering
+\begin{tabular}{lrrrrrr}
+    \hline
+    \bfseries{Root Type} & \bfseries{Android} & \bfseries{Libcore} & \bfseries{App} & \bfseries{Lib}  & \bfseries{All} & \bfseries{\%} \\
+    \hline
+
+Runtime	&	1335	&	73	&	1843	&	690  &	3894 & 64.85\% \\  %2466+854
+Error	       &	 188              &	 46	&	302             &	167	           &	691 & 11.51\%	\\
+Checked	&	276           &	314	&	313          &	567	           &	1358 & 22.61\%	\\
+Throwable	&	0	       &	0	&	2            &	0         &	2 & 0.03\%	\\
+Undefined	&	4	&	0	&	18		&	38	   &	60	& 1.00\% \\
+ \hline
+All		& 1  803	&	433	&	2478	&	1462	&	6005 	\\
+    \hline
+  \end{tabular}
+\caption{Types and origins of root exceptions.}
+  \label{tab:typeroottab}
+\end{table}
+
+\bigskip
+
 We can observe that most of the reported exceptions are of type runtime
@@
 
-%SELECT COUNT(*) FROM STACKTRACESUMMARY WHERE
-% repo !='android'  AND MAIN_EXTENDS LIKE 'RUNTIME%' 
-%SELECT COUNT(*) FROM STACKTRACESUMMARY WHERE
-%repo !='android'  AND MAIN_EXTENDS LIKE 'RUNT%' AND MAINSIGNALER_FROM like 'APP%';
-
-\begin{table}
-\scriptsize
-\centering
-\begin{tabular}{lrrrrrr}
-    \hline
-    \bfseries{Root Type} & \bfseries{Android} & \bfseries{Libcore} & \bfseries{App} & \bfseries{Lib}  & \bfseries{All} & \bfseries{\%} \\
-    \hline
-
-Runtime	&	1335	&	73	&	1843	&	690  &	3894 & 64.85\% \\  %2466+854
-Error	       &	 188              &	 46	&	302             &	167	           &	691 & 11.51\%	\\
-Checked	&	276           &	314	&	313          &	567	           &	1358 & 22.61\%	\\
-Throwable	&	0	       &	0	&	2            &	0         &	2 & 0.03\%	\\
-Undefined	&	4	&	0	&	18		&	38	   &	60	& 1.00\% \\
- \hline
-All		& 1  803	&	433	&	2478	&	1462	&	6005 	\\
-    \hline
-  \end{tabular}
-\caption{Types and origins of root exceptions.}
-  \label{tab:typeroottab}
-\end{table}
-
-\bigskip
+%TABLE 5 WAS HERE
 
@@
 resource deficiencies from which the program cannot possibly recover.
-Hence, `hiding" an unrecoverable condition into a runtime exception
-can lead to a scenario where the developer tries to handle such exception, leading the program
-to na unpredictable state. This kind of wrapping should be avoided, since represent a
+Hence, ``hiding" an unrecoverable condition into a runtime exception
+can lead to a scenario where the developer tries to handle such an exception, leading the program
+to an unpredictable state. This kind of wrapping should be avoided, since represent a
 serious bug hazard.}
@@
 were also native methods written in C. Table~\ref{tab:exampeswrap} illustrates
-some of these wrappings -- some of them are masking an OutOfMemoryError
+some of these wrappings --- some of them are masking an OutOfMemoryError
 into a checked exception.
@@
 the API designer is forcing the client to handle the exceptional condition.
-But in this case the exception condition in and unrecoverable error,
-which makes infeasible the task of providing a proper handler. 
+In this case, the exception condition masks an unrecoverable error,
+which makes the task of providing a proper handler
+infeasible.\footnote{TODO: Please note that Errors are not supposed to be
+caught. The main problem here is that the program continues with e.g. no
+memory available.}
 This kind of wrapping is even more dangerous the previous one
@@
 
-\textcolor{red}{As previously menitoned, the goal of this work is to obtain a thorough understanding of common exception handling \emph{bug hazards} that app developers face.
+\textcolor{red}{As previously mentioned, the goal of this work is to obtain a thorough understanding of common exception handling \emph{bug hazards} that app developers face.
 In the first phase of this work, we conducted a mining study which identified common exception handling bug hazards in app development.
@@
 
-\textcolor{red}{The general research question guiding our exploratory survey is twofold \emph{RQ 2: How do developers deal with the exception handling code in Android apps and what are the developers' perspectives about the main bug hazards found during the mining study?} This general research question has been broken into a set of research questions that are answered by the exploratory survey.}
+\textcolor{red}{The overall research question guiding our exploratory survey is
+the following:
+\emph{RQ 2: How do developers deal with the exception handling code in Android apps and what are the developers' perspectives about the main bug hazards found during the mining study?} This general research question has been broken into a set of research questions that are answered by the exploratory survey.}
 
@@
 
-\textcolor{red}{Since our aim is to learn from a large number of developers, we use surveys which scale well.}
-\textcolor{red}{The survey is split into two logical sections; the motaivations behind each logical section are: 
+\textcolor{red}{Since our aim is to learn from a large number of developers, we
+use surveys which are known to scale well.}
+\textcolor{red}{The survey is split into two logical sections; the motivations behind each logical section are:
 (i) questions aiming at learning from developers about the usage of the exception handling code in app development;  and
-(ii) questions focusing on getting developers percerptionds about the bug hazards detected in the first phase of this study.
+(ii) questions focusing on getting developers perceptions about the bug hazards detected in the first phase of this study.
 The survey comprises multiple choice or Likert-scale questions and open-ended questions.
@@
 
-\textcolor{red}{Given the exploratory nature of our research, we used methods from Grounded Theory ~\cite{charmaz2006} to answer some of our research questions by iterating through the open-ended survey responses.
+\textcolor{red}{Given~\footnote{TODO: I (GG) would put this paragraph before
+the previous one} the exploratory nature of our research, we used methods from Grounded Theory ~\cite{charmaz2006} to answer some of our research questions by iterating through the open-ended survey responses.
 The grounded theory coding used in this study consists of at two phases: (1) initial coding entails a close reading of the data and (2) later we used focused coding ~\cite{charmaz2006} to pinpoint and develop the most salient themes ~\cite{charmaz2006} in the analyzed data.
@@
 \textcolor{red}{The response rate was 3.9\% --
-although the response rate achieved in our study is low, it is inlined to similar surveys reported in the literature (e.g., ~\cite{kochhar2015understanding,joorabchi2013real,bavota2015impact,ko2007information,hindle2015topics}) whose response rates mostly varies between 2\% and 4\%.}
+although the response rate achieved in our study is low, it is in line with similar surveys reported in the literature (e.g., ~\cite{kochhar2015understanding,joorabchi2013real,bavota2015impact,ko2007information,hindle2015topics}) whose response rates mostly varies between 2\% and 4\%.}
 The majority of our respondents have more than 2 years of Java development experience (87.3\%) and of Android development (85.9\%) -- see Figure~\ref{fig:devexpertise}.
@@
 As mentioned before several general guidelines have been proposed on how to
-use Java exceptions.
-Section~\ref{sec:best}  compiled a list of some exception handling (EH) best practices that guided our mining study.
-In this survey the developers were also asked if they knew about and used Java EH best practices -- see Figure~\ref{fig:allquest}.
-Most of them said that they knew and adopted Java EH practices (68\%).
-From this set the most mentioned best practices are presented in Table~\ref{tab:javapractices} -- considering that one developer mentioned more than one practice.
-Most of the best practices mentioned were the ones cited in Section~\ref{sec:best}.
-For instance:\emph{``Don't throw [pure] RuntimeException (in most cases), don't catch OutOfMemoryError, etc." [D42]}.
-Some of the practices mentioned, however, are not well known Java EH best practices but seem to be used by some Android developers to better deal with exceptions in the Android app context, such as the use of a crash fast approach\footnote{\url{http://www.slideshare.net/pyricau/crash-fast}} -- \emph{``Crash is good for hinting developer.
-Fast crash fast solve." [D51]}.
-\textcolor{red}{According to this approach, developers should let the application to crash as soon as an unexpected situation happens.
-The motivation behind this approach is that the developer will receive the crash information and fix the potential bug that caused it.
-On the other hand, if the developer lets the app to continue runing, after an exception had happened, the aplication may enter an inconsistent state.
+use Java exceptions. Section~\ref{sec:best}  compiled a list of some exception
+handling (EH) best practices that guided our mining study. In this survey the
+developers were also asked if they knew about and used Java EH best practices --
+see Figure~\ref{fig:allquest}. Most of them said that they knew and adopted Java
+EH practices (68\%). From this set the most mentioned best practices are
+presented in Table~\ref{tab:javapractices} -- considering that one developer
+mentioned more than one practice. Most of the best practices mentioned were the
+ones cited in Section~\ref{sec:best}. For instance:\emph{``Don't throw [pure]
+RuntimeException (in most cases), don't catch OutOfMemoryError, etc." [D42]}.
+Some of the practices mentioned, however, are not well known Java EH best
+practices but seem to be used by some Android developers to better deal with
+exceptions in the Android app context, such as the use of a crash fast
+approach\footnote{\url{http://www.slideshare.net/pyricau/crash-fast}} --
+\emph{``Crash is good for hinting developer. Fast crash fast solve." [D51]}.
+\textcolor{red}{According to this approach, developers should let the
+application to crash as soon as an unexpected situation happens. The motivation
+behind this approach is that the developer will receive the crash information
+and fix the potential bug that caused it. On the other hand, if the developer
+lets the app to continue running, after an exception had happened, the
+application may enter an inconsistent state.
 In such scenario, the time between the fault being exercised and the failure (the fault manifestation) increases, which may impair finding and fixing the potential bug that caused the failure.} Moreover, some respondents favor checked exceptions instead of runtime to represent the exceptional conditions in Android apps -- \emph{``RuntimeExceptions are not particularly well suited for Android's robustness, as any uncaught one would cause a crash." [D26]}; \emph{``Checked exception help, unchecked exceptions don't." [D25]}; \emph{``avoid unchecked exceptions except if the case is really abnormal and should not happen [...]"[D45]}.
@@
 
-\textcolor{red}{This section discusses lessons learned from the mining and the survey studies. The mining study revealed a set of bug 
+\textcolor{red}{This section discusses the lessons learned from the mining and the survey studies. The mining study revealed a set of bug
 hazards -- such as (i) the cross-type wrappings; (ii) the abundance of null pointer problems; and (iii) the undocumented runtime
-exceptions signaled by third-party code -- which were confirmed by developers during the survey.
-Next subsections discuss about each of such bug hazards, presenting how the developer's perceived them and pointing to what can be done to help developers
-to deal with them.}
+exceptions signaled by third-party code, which were confirmed by developers
+during the survey. Here, we discuss the identified bug hazards, we present how
+the developers perceived them and point out at what developers can do in
+order to deal with them.}
 
@@
 
-\textcolor{red}{Another internal threat to validity of this study regards to the process adopted to 
-identify the type of each exception reported on stack traces. 
-To perform such type analysis we chose a version of the souce-code or bytecode
- of each reported exception, as follows: (i) for all exceptions signaled by the Android Platform 
+\textcolor{red}{The process to
+identify the type of exceptions reported in stack traces may not be completely
+accurate.
+Specifically, we performed the selection of the version of the exception source code or bytecode
+to analyze, as follows: (i) for all exceptions signaled by the Android Platform
 and the Java Environment the analysis was based on Version 4.4 of Android platform (API level 19);  (ii)
-for the exceptions signaled by applications the analysis considered
+for the exceptions signaled by applications, the analysis considered
 the last version of the application source-code available on the Github/Google code; and (iii)
 for all exceptions signaled by third-party libraries the type analysis
-considered the last bytecode version available on the app repository.
+considered the latest bytecode version available on the app repository.
 When the exception could not be found automatically or manually (based on the chosen version),
@@
 remained undefined, which occurred in 60 different exception stack traces out of 6,005 mined stack traces.
-Hence, the exception type could not be identified only in 1\% of the mined flows.
+Hence, the exception type could not be accurately identified in 1\% of the mined
+exception flows.
 There should be situations where the exception type (i.e., checked or runtime) changes from version to another.
-Altough we believe that it will not happen very often, we did not investigate such issue.
- One way to mitigate such threat would be to consider the issue date to select the most 
-suitable source-code/bytecode version to perform the analysis.
-However, the number of applications analyzed (approx. 6000) 
-would increase the time and complexity needed to follow such improved process.}
+Although we believe that it will not happen very often, we did not investigate
+this issue.}
+
+% One way to mitigate such a threat would be to consider the issue date to select the most
+%suitable source-code/bytecode version to perform the analysis.
+%However, the number of applications analyzed (approx. 6000)
+%would increase the time and complexity needed to follow such improved process.}
 
@@
 tools to help with specific development tasks.
- \textcolor{red}{Some of such exploratory surveys were applied to Android developers~\cite{kochhar2015understanding,joorabchi2013real,bavota2015impact,linares2015developers}. 
+
+\textcolor{red}{Exploratory surveys have also targeted Android developers~\cite{kochhar2015understanding,joorabchi2013real,bavota2015impact,linares2015developers}.
 Kochhar et al. ~\cite{kochhar2015understanding} conducted a survey-based study to discover the
@@
 faced by developers while testing the apps. The survey was sent to 3,905 emails of Android developers
-and received 83 responses (response rate of 2.13%). 
-Joorabchi et al. ~\cite{joorabchi2013real} performed a survey-based study whose goal was to better understand the the main challenges 
+and received 83 responses (response rate of 2.13\%).
+Joorabchi et al. ~\cite{joorabchi2013real} performed a survey-based study whose goal was to better understand the main challenges
 developers face when building apps for different mobile devices.
 They interviewed 12 senior mobile developers and performed a semi-structured survey, with 188 respondents
- from the mobile development community - since the survey was shared via e-mail groups and social media websites
+ from the mobile development community, since the survey was shared via e-mail groups and social media websites
 response rate of the survey could not be calculated. Linares-Vásquez et al. ~\cite{linares2015developers} surveyed 485 open source Android
@@
 A common characteristic among such works and our work is the low response rate of the surveys (between 4\% and 2\%).
-None of these surveys, however, assed the how developers deal with exceptions in Android nor 
-the developers' perspective regarding a set of exception handling bug hazards. In our study
+None of these surveys, however, assessed how developers deal with exceptions in Android nor
+the developers' perspective regarding a set of exception handling bug hazards.
+In our study,
 we conduced the first exploratory survey focusing on exception handling issues in Android development.}

### 2016-04-20 SPRINGER_JOURNAL/msr_journal.tex
addressing ref 3 and Georgios comments

@@
 refers to a programming mistake according to its Javadoc; and (ii) the Java specification lists all
-exceptions related to backward compatibility~\footnote{http://docs. oracle.com/javase/specs/jls/se7/html/jls-13.html}, such as
+exceptions related to backward compatibility
+~\footnote{http://docs.oracle.com/javase/specs/jls/se7/html/jls-13.html}, such as
 InstantiationError, VerifyError, and IllegalAccessError.
@@
 into a checked exception.
-\textcolor{red}{By confronting the API user with a checked exception,
+\textcolor{red}{On the one hand, by confronting the API user with a checked exception,
 the API designer is forcing the client to handle the exceptional condition.
-In this case, the exception condition masks an unrecoverable error,
-which makes the task of providing a proper handler
-infeasible.\footnote{TODO: Please note that Errors are not supposed to be
-caught. The main problem here is that the program continues with e.g. no
-memory available.}
-This kind of wrapping is even more dangerous the previous one
+On the other hand, according to Java specification Errors are not supposed to be
+caught. In this case, even if the exception is caught by a handler, the problem 
+that trigged the Error remains: there is not enought memory to execute the app.
+This kind of wrapping is even more dangerous than the previous one
 and may lead to the ``exception confusion'' described next.}
@@
 
-To make the analysis easier, we further refine this question into sub-questions, such as: \emph{How often do developers handle exceptions?} \emph{How often do developers throw exceptions?} And, to show us how developers may deal with the exception handling (EH) code: \emph{Do developers know about Java EH best practices and/or Android specific EH best practices?}
+To make the analysis easier, we further refine this question into sub-questions, as follows.
+To investigate whether the development of exception handling code is a daily concern for developers or it is something that developers face seldom and for that reason do not care much, the following subquestions are added: \emph{How often do developers handle exceptions?} \emph{How often do developers throw exceptions?} And, to check if the Java EH best practices, available in literature, are used in practice and to investigate if any Android-specific EH best practice are used by developers, the following subquestion is added  subquestion: \emph{Do developers know about Java EH best practices and/or Android specific EH best practices?}
 
@@
 
-%\emph{\textbf{RQ 2.5:}  Are developers aware of the non-declared checked exceptions?}
-
-%\emph{RQ 2:Have developers ever faced a crash caused by NullPointerException?}
-%\emph{RQ 3:Are developers aware of the robustness threats caused by cross-type wrappings?}; %\emph{RQ 4: Are developers aware of the robustness threats caused by JNI undocumented checked exceptions?}
-
-We address these questions by presenting a set of code snippets in which
-the bug hazards are present, asking developers what they would do when faced with such scenarios.
+We address these questions by presenting a set of code snippets in which the bug hazards are present, asking developers what they would do when faced with such scenarios.
 
@@
 
-\textcolor{red}{Since our aim is to learn from a large number of developers, we
-use surveys which are known to scale well.}
-\textcolor{red}{The survey is split into two logical sections; the motivations behind each logical section are:
-(i) questions aiming at learning from developers about the usage of the exception handling code in app development;  and
-(ii) questions focusing on getting developers perceptions about the bug hazards detected in the first phase of this study.
-The survey comprises multiple choice or Likert-scale questions and open-ended questions.
+%~\footnote{TODO: I (GG) would put this paragraph before the previous one} 
+%~\footnote{TODO: I (RSC) merged the beginning of both paragraphs. }
+
+\textcolor{red}{Given the exploratory nature of our research, we used methods from Grounded Theory ~\cite{charmaz2006} to answer some of our research questions.
+Since our aim is to learn from a large number of developers, we use surveys which are known to scale well.
+The survey is split into two logical sections; the motivations behind each logical section are: (i) questions aiming at learning from developers about the usage of the exception handling code in app development;  and (ii) questions focusing on getting developers perceptions about the bug hazards detected in the first phase of this study.}
+
+\textcolor{red}{The survey comprises multiple choice or Likert-scale questions and open-ended questions.
 The multiple choice questions are intermixed with open-ended questions to further elicit the developer's opinions.
@@
 
-\textcolor{red}{Given~\footnote{TODO: I (GG) would put this paragraph before
-the previous one} the exploratory nature of our research, we used methods from Grounded Theory ~\cite{charmaz2006} to answer some of our research questions by iterating through the open-ended survey responses.
+\textcolor{red}{We used grounded theory coding to iterate through the open-ended survey responses.
 The grounded theory coding used in this study consists of at two phases: (1) initial coding entails a close reading of the data and (2) later we used focused coding ~\cite{charmaz2006} to pinpoint and develop the most salient themes ~\cite{charmaz2006} in the analyzed data.
@@
 
-\textcolor{red}{The process to
-identify the type of exceptions reported in stack traces may not be completely
+\textcolor{red}{The process for
+identifing the type of exceptions reported in stack traces may not be completely
 accurate.
@@
 \begin{acknowledgements}
-This work is partially supported by INES, funded by CNPq and
-FACEPE, grants 573964/2008-4 and APQ-1037-1.03/08.
+This work is partially supported by the National Institute of Science and Technology for Software Engineering (INES), CNPq and FACEPE, grants 573964/2008-4, 552645/2011-7, and APQ-1037-1.03/08, CNPq Universal grant 484209/2013-2, and CAPES/PROAP.
 \end{acknowledgements}

### 2016-04-20 SPRINGER_JOURNAL/msr_journal.tex
Minor grammar/spelling updates in red part of document

@@
    Christoph Treude \at
-  \email{ctreude@dimap.ufrn.br} \\
+  \email{christoph.treude@adelaide.edu.au} \\
 }
@@
 the client can simply ignore such exception - which can negatively impact the app robustness.
-\textcolor{red}{Altough it is considered a bad practice, when such wrapping is accompained with a well defined
-exception handling policy (which assures that a proper handling will be provided within the app)
+\textcolor{red}{Although it is considered a bad practice, when such wrapping is accompained with a well defined
+exception handling policy (which assures that proper handling will be provided within the app)
 the impact on app robustness can be mitigated.}
@@
 can lead to a scenario where the developer tries to handle such an exception, leading the program
-to an unpredictable state. This kind of wrapping should be avoided, since represent a
+to an unpredictable state. This kind of wrapping should be avoided, since it represents a
 serious bug hazard.}
@@
 the API designer is forcing the client to handle the exceptional condition.
-On the other hand, according to Java specification Errors are not supposed to be
+On the other hand, according to the Java specification Errors are not supposed to be
 caught. In this case, even if the exception is caught by a handler, the problem 
-that trigged the Error remains: there is not enought memory to execute the app.
+that trigged the Error remains: there is not enough memory to execute the app.
 This kind of wrapping is even more dangerous than the previous one
@@
 Since our aim is to learn from a large number of developers, we use surveys which are known to scale well.
-The survey is split into two logical sections; the motivations behind each logical section are: (i) questions aiming at learning from developers about the usage of the exception handling code in app development;  and (ii) questions focusing on getting developers perceptions about the bug hazards detected in the first phase of this study.}
+The survey is split into two logical sections; the motivations behind each logical section are: (i) questions aiming at learning from developers about the usage of the exception handling code in app development;  and (ii) questions focusing on getting developers' perceptions about the bug hazards detected in the first phase of this study.}
 
@@
 \textcolor{red}{We used grounded theory coding to iterate through the open-ended survey responses.
-The grounded theory coding used in this study consists of at two phases: (1) initial coding entails a close reading of the data and (2) later we used focused coding ~\cite{charmaz2006} to pinpoint and develop the most salient themes ~\cite{charmaz2006} in the analyzed data.
+The grounded theory coding used in this study consists of two phases: (1) initial coding entails a close reading of the data and (2) later we used focused coding ~\cite{charmaz2006} to pinpoint and develop the most salient themes ~\cite{charmaz2006} in the analyzed data.
 Answers to different questions in the survey were coded separately.
@@
 \textcolor{red}{The response rate was 3.9\% --
-although the response rate achieved in our study is low, it is in line with similar surveys reported in the literature (e.g., ~\cite{kochhar2015understanding,joorabchi2013real,bavota2015impact,ko2007information,hindle2015topics}) whose response rates mostly varies between 2\% and 4\%.}
+although the response rate achieved in our study is low, it is in line with similar surveys reported in the literature (e.g., ~\cite{kochhar2015understanding,joorabchi2013real,bavota2015impact,ko2007information,hindle2015topics}) whose response rates mostly vary between 2\% and 4\%.}
 The majority of our respondents have more than 2 years of Java development experience (87.3\%) and of Android development (85.9\%) -- see Figure~\ref{fig:devexpertise}.
@@
 \textcolor{red}{According to this approach, developers should let the
-application to crash as soon as an unexpected situation happens. The motivation
+application crash as soon as an unexpected situation happens. The motivation
 behind this approach is that the developer will receive the crash information
 and fix the potential bug that caused it. On the other hand, if the developer
-lets the app to continue running, after an exception had happened, the
+lets the app continue to run after an exception had happened, the
 application may enter an inconsistent state.
@@
 \textcolor{red}{The process for
-identifing the type of exceptions reported in stack traces may not be completely
+identifying the type of exceptions reported in stack traces may not be completely
 accurate.
@@
 for the exceptions signaled by applications, the analysis considered
-the last version of the application source-code available on the Github/Google code; and (iii)
+the last version of the application source-code available on Github/Google code; and (iii)
 for all exceptions signaled by third-party libraries the type analysis
@@
 exception flows.
-There should be situations where the exception type (i.e., checked or runtime) changes from version to another.
+There should be situations where the exception type (i.e., checked or runtime) changes from one version to another.
 Although we believe that it will not happen very often, we did not investigate
@@
  from the mobile development community, since the survey was shared via e-mail groups and social media websites
-response rate of the survey could not be calculated. Linares-Vásquez et al. ~\cite{linares2015developers} surveyed 485 open source Android
-app and library developers (hosted on GitHub)  to understand developers’ practices for detecting and fixing
+response rate of the survey could not be calculated. Linares-V\'{a}squez et al. ~\cite{linares2015developers} surveyed 485 open source Android
+app and library developers (for projects hosted on GitHub)  to understand developers' practices for detecting and fixing
 performance bottlenecks in mobile apps.  This work emailed the survey to 24,340 developers and received
 628 responses - the response rate was 2.6\%.
-Bavota at al. ~\cite{bavota2015impact} investigated the impact of API change-proness and fault-proneness on the user ratings of Android apps.
+Bavota et al. ~\cite{bavota2015impact} investigated the impact of API change-proness and fault-proneness on the user ratings of Android apps.
 They surveyed developers to assess their perspective on whether such problems could be the cause for unfavorable user ratings.

### 2016-04-20 SPRINGER_JOURNAL/msr_journal.tex
addressing the motivations behind research questions

@@
  In the second phase of this work, we set up an exploratory qualitative investigation and surveyed Android developers on how they perceive the bug hazards detected in the mining study.
-The scope of our study is GitHub---using our GHTorrent database~\cite{Gousi13}, we aimed our survey at developers from Android projects available in GitHub for which issues were mined during the first phase of this study.}
+The scope of our study is GitHub---using our GHTorrent database~\cite{Gousi13}, we aimed our survey at developers from Android projects available in GitHub for which issues were mined during the first phase of this study.
+Next sections detail this exploratory qualitative investigation.}
 
@@
 
-To make the analysis easier, we further refine this question into sub-questions, as follows.
-To investigate whether the development of exception handling code is a daily concern for developers or it is something that developers face seldom and for that reason do not care much, the following subquestions are added: \emph{How often do developers handle exceptions?} \emph{How often do developers throw exceptions?} And, to check if the Java EH best practices, available in literature, are used in practice and to investigate if any Android-specific EH best practice are used by developers, the following subquestion is added  subquestion: \emph{Do developers know about Java EH best practices and/or Android specific EH best practices?}
-
-The subsequent research questions focus on the developers' perspectives about the main bug hazards found during the mining study, more specifically:
+\textcolor{red}{To make the analysis easier, we further refine this question into sub-questions, as follows.
+To investigate whether the development of exception handling code is a daily concern for developers or it is something that developers rarely face and for that reason do not care much, the following subquestions are added: \emph{How often do developers handle exceptions?} \emph{How often do developers throw exceptions?} And, to evaluate developers knowledge concerning EH best practices, we also investigated: \emph{Do developers know about Java EH best practices and/or Android specific EH best practices?}}
+
+The subsequent research questions focus on the developers' perspectives about the main bug hazards found during the mining study.
+We address these questions by presenting a set of code snippets in which the bug hazards are present, asking developers what they would do when faced with such scenarios.
+The questions related to bug hazards are as follows:
 
@@
 
-We address these questions by presenting a set of code snippets in which the bug hazards are present, asking developers what they would do when faced with such scenarios.
-
-Our last research question addresses
-whether the exception handling code helps with the development of robust Android applications, and what developers usually do to prevent apps from crashing:
+\textcolor{red}{Our last research question addresses whether the exception handling code helps with the development of robust Android applications, and what developers usually do to prevent apps from crashing.
+The motivation behind this question was to discover common practices for dealing with uncaught exceptions and to acess developers opinion about the role of the exception handling mechanism in the development of robust apps.
+Hence, the last question is as follows:}
 

### 2016-04-20 SPRINGER_JOURNAL/msr_journal.tex
including another limitation for the exploratory survey

@@
 the client can simply ignore such exception - which can negatively impact the app robustness.
-\textcolor{red}{Although it is considered a bad practice, when such wrapping is accompained with a well defined
+Although it is considered a bad practice, when such wrapping is accompained with a well defined
 exception handling policy (which assures that proper handling will be provided within the app)
-the impact on app robustness can be mitigated.}
+the impact on app robustness can be mitigated.
+%\textcolor{red}
 
@@
 an java.lang.OutOfMemoryError and a java.lang.StackOverflowError.
-\textcolor{red}{Such Errors result from failures detected by the runtime environment which indicate
+Such Errors result from failures detected by the runtime environment which indicate
 resource deficiencies from which the program cannot possibly recover.
@@
 to an unpredictable state. This kind of wrapping should be avoided, since it represents a
-serious bug hazard.}
+serious bug hazard.
+%\textcolor{red}
 
@@
 into a checked exception.
-\textcolor{red}{On the one hand, by confronting the API user with a checked exception,
+{On the one hand, by confronting the API user with a checked exception,
 the API designer is forcing the client to handle the exceptional condition.
@@
 This kind of wrapping is even more dangerous than the previous one
-and may lead to the ``exception confusion'' described next.}
+and may lead to the ``exception confusion'' described next.
+%\textcolor{red}
 
@@
 
-\textcolor{red}{As previously mentioned, the goal of this work is to obtain a thorough understanding of common exception handling \emph{bug hazards} that app developers face.
+As previously mentioned, the goal of this work is to obtain a thorough understanding of common exception handling \emph{bug hazards} that app developers face.
 In the first phase of this work, we conducted a mining study which identified common exception handling bug hazards in app development.
@@
 The scope of our study is GitHub---using our GHTorrent database~\cite{Gousi13}, we aimed our survey at developers from Android projects available in GitHub for which issues were mined during the first phase of this study.
-Next sections detail this exploratory qualitative investigation.}
-
-%\textcolor{red}{The survey mixed Liket-scale with open-ended questions, and the motivations behind the survey questions are twofold: (i) to learn from
-%developers about the usage of the exception handling code in app development;  and (ii) to get developers percerptions about the bug hazards detected in
-%the first phase of this study.}
-
-%Our main goal is to learn from developers about the Java exception handling usage and the main
-%exception handling bug hazards in Android. We therefore use a survey as our main research
-%instrument. The survey mixed  Liket-scale with open-ended questions. The questions were motivated
-%by the analysis of the existing bug hazards detected in the first phase of this study.
-%The next sections detail the survey.
-%Our main findings reveal that developers often handle exceptions while developing Android apps and
-%we provide insights into the main challenges faced by them during this task. xxxxxxx
+Next sections detail this exploratory qualitative investigation.
+%\textcolor{red}
 
 \subsection{Research Questions}
-
-\textcolor{red}{The overall research question guiding our exploratory survey is
+%\textcolor{red}
+The overall research question guiding our exploratory survey is
 the following:
-\emph{RQ 2: How do developers deal with the exception handling code in Android apps and what are the developers' perspectives about the main bug hazards found during the mining study?} This general research question has been broken into a set of research questions that are answered by the exploratory survey.}
+\emph{RQ 2: How do developers deal with the exception handling code in Android apps and what are the developers' perspectives about the main bug hazards found during the mining study?} This general research question has been broken into a set of research questions that are answered by the exploratory survey.
 
@@
 
-\textcolor{red}{To make the analysis easier, we further refine this question into sub-questions, as follows.
-To investigate whether the development of exception handling code is a daily concern for developers or it is something that developers rarely face and for that reason do not care much, the following subquestions are added: \emph{How often do developers handle exceptions?} \emph{How often do developers throw exceptions?} And, to evaluate developers knowledge concerning EH best practices, we also investigated: \emph{Do developers know about Java EH best practices and/or Android specific EH best practices?}}
+%\textcolor{red}
+To make the analysis easier, we further refine this question into sub-questions, as follows.
+To investigate whether the development of exception handling code is a daily concern for developers or it is something that developers rarely face and for that reason do not care much, the following subquestions are added: \emph{How often do developers handle exceptions?} \emph{How often do developers throw exceptions?} And, to evaluate developers knowledge concerning EH best practices, we also investigated: \emph{Do developers know about Java EH best practices and/or Android specific EH best practices?}
 
@@
 
-\textcolor{red}{Our last research question addresses whether the exception handling code helps with the development of robust Android applications, and what developers usually do to prevent apps from crashing.
+%\textcolor{red}
+Our last research question addresses whether the exception handling code helps with the development of robust Android applications, and what developers usually do to prevent apps from crashing.
 The motivation behind this question was to discover common practices for dealing with uncaught exceptions and to acess developers opinion about the role of the exception handling mechanism in the development of robust apps.
-Hence, the last question is as follows:}
+Hence, the last question is as follows:
 
@@
 
-\textcolor{red}{Given the exploratory nature of our research, we used methods from Grounded Theory ~\cite{charmaz2006} to answer some of our research questions.
+%\textcolor{red}
+Given the exploratory nature of our research, we used methods from Grounded Theory ~\cite{charmaz2006} to answer some of our research questions.
 Since our aim is to learn from a large number of developers, we use surveys which are known to scale well.
-The survey is split into two logical sections; the motivations behind each logical section are: (i) questions aiming at learning from developers about the usage of the exception handling code in app development;  and (ii) questions focusing on getting developers' perceptions about the bug hazards detected in the first phase of this study.}
-
-\textcolor{red}{The survey comprises multiple choice or Likert-scale questions and open-ended questions.
+The survey is split into two logical sections; the motivations behind each logical section are: (i) questions aiming at learning from developers about the usage of the exception handling code in app development;  and (ii) questions focusing on getting developers' perceptions about the bug hazards detected in the first phase of this study.
+
+%\textcolor{red}
+The survey comprises multiple choice or Likert-scale questions and open-ended questions.
 The multiple choice questions are intermixed with open-ended questions to further elicit the developer's opinions.
@@
 Overall, the survey includes 13 open-ended questions, 5 Likert-scale questions and 10 multiple choice questions.
-The respondents could complete the survey in about 15 minutes.}
-
-\textcolor{red}{We used grounded theory coding to iterate through the open-ended survey responses.
+The respondents could complete the survey in about 15 minutes.
+
+%\textcolor{red}
+We used grounded theory coding to iterate through the open-ended survey responses.
 The grounded theory coding used in this study consists of two phases: (1) initial coding entails a close reading of the data and (2) later we used focused coding ~\cite{charmaz2006} to pinpoint and develop the most salient themes ~\cite{charmaz2006} in the analyzed data.
@@
 In all cases, the first author was one of the two coders, and one of the other authors coded the data as well in close collaboration with the first author until saturation was reached.
-The first author then went through the remaining responses to assign codes to responses.}
+The first author then went through the remaining responses to assign codes to responses.
 
@@
 %~\cite{coelhoetal15}
+%\textcolor{red}{
  In the first phase of this study, described previously, we mined exception stack traces available in issues reported for several Android projects hosted on GitHub and Google Code.
@@
 We emailed the 1,824 developers and received 71 valid answers.
-\textcolor{red}{The response rate was 3.9\% --
-although the response rate achieved in our study is low, it is in line with similar surveys reported in the literature (e.g., ~\cite{kochhar2015understanding,joorabchi2013real,bavota2015impact,ko2007information,hindle2015topics}) whose response rates mostly vary between 2\% and 4\%.}
+The response rate was 3.9\% --
+although the response rate achieved in our study is low, it is in line with similar surveys reported in the literature (e.g., ~\cite{kochhar2015understanding,joorabchi2013real,bavota2015impact,ko2007information,hindle2015topics}) whose response rates mostly vary between 2\% and 4\%.
 The majority of our respondents have more than 2 years of Java development experience (87.3\%) and of Android development (85.9\%) -- see Figure~\ref{fig:devexpertise}.
@@
 \emph{``Crash is good for hinting developer. Fast crash fast solve." [D51]}.
-\textcolor{red}{According to this approach, developers should let the
+According to this approach, developers should let the
 application crash as soon as an unexpected situation happens. The motivation
@@
 application may enter an inconsistent state.
-In such scenario, the time between the fault being exercised and the failure (the fault manifestation) increases, which may impair finding and fixing the potential bug that caused the failure.} Moreover, some respondents favor checked exceptions instead of runtime to represent the exceptional conditions in Android apps -- \emph{``RuntimeExceptions are not particularly well suited for Android's robustness, as any uncaught one would cause a crash." [D26]}; \emph{``Checked exception help, unchecked exceptions don't." [D25]}; \emph{``avoid unchecked exceptions except if the case is really abnormal and should not happen [...]"[D45]}.
+In such scenario, the time between the fault being exercised and the failure (the fault manifestation) increases, which may impair finding and fixing the potential bug that caused the failure.
+Moreover, some respondents favor checked exceptions instead of runtime to represent the exceptional conditions in Android apps -- \emph{``RuntimeExceptions are not particularly well suited for Android's robustness, as any uncaught one would cause a crash." [D26]}; \emph{``Checked exception help, unchecked exceptions don't." [D25]}; \emph{``avoid unchecked exceptions except if the case is really abnormal and should not happen [...]"[D45]}.
+%\textcolor{red}
 
@@
 
-\textcolor{red}{This section discusses the lessons learned from the mining and the survey studies. The mining study revealed a set of bug
+This section discusses the lessons learned from the mining and the survey studies. The mining study revealed a set of bug
 hazards -- such as (i) the cross-type wrappings; (ii) the abundance of null pointer problems; and (iii) the undocumented runtime
@@
 the developers perceived them and point out at what developers can do in
-order to deal with them.}
+order to deal with them.
+%\textcolor{red}{
 
@@
 
-\textcolor{red}{The process for
+The process for
 identifying the type of exceptions reported in stack traces may not be completely
@@
 Although we believe that it will not happen very often, we did not investigate
-this issue.}
+this issue.
+%\textcolor{red}
 
@@
 \subsection{Limitations of the Survey-based Study}
-Due to the exploratory nature of the second phase of this work whose
+\emph{Results Generalization.} Due to the exploratory nature of the second phase of this work whose
 goal was to identify the developers' perspectives concerning the exception
@@
 time and motivation to answer the survey questions.
-
 Although the themes and findings that emerged in our study cannot
@@
 exception handling bug hazards are.
+
+%I found strange that the authors did not exploit this opportunity for asking their opinion about the specific issues you identified in *their* apps.
+IMHO, this would provide much more qualitative feedback (while of course requiring the customisation of the survey for each specific developer).
+Can you please justify this choice in the paper?
+
+\emph{Survey Customization.} Although the survey was applied to developers who had contributed to at least one of the GitHub apps, analyzed in the mining study, the survey was not customized to each developer.
+In other words, it did not contain questions focusing specific bug hazards identified in their apps.
+That customization of the survey (for each specific developer/app) could give more insights about the exception handling usage and bug hazards, and could even have improved the response rate for the survey.
+However, we did not followed this approach because the high number of projects analyzed (482 Android apps) and their contributors (1,824 developers) would increase the complexity and time needed (i) to customize the survey and specially (ii) to analyze the survey responses - since themes could emerge from specific contexts.
+This approach can be very useful in a guided interview involving some of the respondents to refine the findings of this exploratory survey, in a future work.
+
+%TODO INCLUDE THE RESON BEHINF THE GENERAL SURVEY.
 
@@
 
-\textcolor{red}{Exploratory surveys have also targeted Android developers~\cite{kochhar2015understanding,joorabchi2013real,bavota2015impact,linares2015developers}.
+Exploratory surveys have also targeted Android developers~\cite{kochhar2015understanding,joorabchi2013real,bavota2015impact,linares2015developers}.
 Kochhar et al. ~\cite{kochhar2015understanding} conducted a survey-based study to discover the
@@
 In our study,
-we conduced the first exploratory survey focusing on exception handling issues in Android development.}
+we conduced the first exploratory survey focusing on exception handling issues in Android development.
+%\textcolor{red}
 

### 2016-04-20 SPRINGER_JOURNAL/msr_journal.tex
fixing typos

@@
 Security and concurrency, which are known to be critical issues for Android apps,
-raised few of the reported exceptions (less than 5\% of the analysed stack traces).
+raised few of the reported exceptions (less than 5\% of the analyzed stack traces).
 
@@
 the client can simply ignore such exception - which can negatively impact the app robustness.
-Although it is considered a bad practice, when such wrapping is accompained with a well defined
+Although it is considered a bad practice, when such wrapping is accompanied by a well defined
 exception handling policy (which assures that proper handling will be provided within the app)
@@
 Our last research question addresses whether the exception handling code helps with the development of robust Android applications, and what developers usually do to prevent apps from crashing.
-The motivation behind this question was to discover common practices for dealing with uncaught exceptions and to acess developers opinion about the role of the exception handling mechanism in the development of robust apps.
+The motivation behind this question was to discover common practices for dealing with uncaught exceptions and to assess developers perspective about the role of the exception handling mechanism in the development of robust apps.
 Hence, the last question is as follows:
@@
 \emph{Survey Customization.} Although the survey was applied to developers who had contributed to at least one of the GitHub apps, analyzed in the mining study, the survey was not customized to each developer.
-In other words, it did not contain questions focusing specific bug hazards identified in their apps.
+In other words, it did not contain questions focusing on specific bug hazards identified in their apps.
 That customization of the survey (for each specific developer/app) could give more insights about the exception handling usage and bug hazards, and could even have improved the response rate for the survey.
-However, we did not followed this approach because the high number of projects analyzed (482 Android apps) and their contributors (1,824 developers) would increase the complexity and time needed (i) to customize the survey and specially (ii) to analyze the survey responses - since themes could emerge from specific contexts.
-This approach can be very useful in a guided interview involving some of the respondents to refine the findings of this exploratory survey, in a future work.
+However, we did not followed this approach because the high number of projects analyzed (482 Android apps) and their contributors (1,824 developers) would increase the complexity and time needed (i) to prepare the survey and specially (ii) to analyze the survey responses - since themes could emerge from specific contexts.
+This approach can be very useful in a guided interview involving some of the respondents to refine the findings of this exploratory survey in a future work.
 

### 2016-04-20 SPRINGER_JOURNAL/msr_journal.tex
formating issues

@@
 NullPointerException makes the cause of the problem explicit and hence
-can be signaled by an API expecting a non-null argument.The high prevalence of NullPointerException is aligned with the
+can be signaled by an API expecting a non-null argument.
+
+The high prevalence of NullPointerException is aligned with the
 findings of other
@@
 Javadoc related to each exception and the Java specification, we
-identified the underlying concern related to each root exception.
-Table~\ref{tab:tophundrend} contains the results of this analysis. This table also illustrates the exceptions
+identified the underlying concern related to each root exception. Table~\ref{tab:tophundrend} contains the results of this analysis. This table also illustrates the exceptions
 that could not be directly mapped to one of the aforementioned concerns, either because they were too general (i.e., java.lang.Exception,
@@
 \bigskip
-
+\bigskip
+\bigskip
+\bigskip
+\bigskip
+\bigskip
 %\subsubsection{Can the exception types reveal bug hazards?}
@@
 %TABLE 5 WAS HERE
+\bigskip
 
@@
 
-%\textcolor{red}
 The survey comprises multiple choice or Likert-scale questions and open-ended questions.
@@
 
-%\textcolor{red}
 We used grounded theory coding to iterate through the open-ended survey responses.

### 2016-07-20 SPRINGER_JOURNAL/techreport/TUD-SERG-2016-018.tex
Create techreport for accepted version.
For upload to TUD-SERG-2016-018 to TU Delft preprint server.

\documentclass[letterpaper,twoside,10pt]{book}
\usepackage{se-report}

%---------------------------------------------------------------------%
% User definable options                                              %    
%---------------------------------------------------------------------%

\title{Exception Handling Bug Hazards in Android: 
Results from a Mining Study and an Exploratory Survey}

\author{Roberta Coelho, Lucas Almeida, Georgios Gousios,\\ Arie van Deursen and Christoph Treude}

\headertitle{Exception Handling Bug Hazards in Android}

\year{2016} % for report id and copyright notice
\report{018}

%\note{Accepted for publication in the Proceedings of the Working Conference
%on Reverse Engineering (WCRE), 2010, IEEE Computer Society.}

\note{This paper has been accepted for publication in \emph{Empirical Software Engineering}, 2016.\\
It is a substantially revised and extended version of 
Roberta Coelho, Lucas Almeida, Georgios Gousios, Arie van Deursen:
Unveiling Exception Handling Bug Hazards in Android Based on GitHub and Google Code Issues.
\emph{Working Conference on Mining Software Repositories} (MSR) 2015: 134-145.
}

%\copyrightmessage{%
%\copyright{} 2016 ACM. Personal use of this material is permitted. Permission from ACM must be obtained for all other uses, in any current or future media, including reprinting/republishing this material for advertising or promotional purposes, creating new collective works, for resale or redistribution to servers or lists, or reuse of any copyrighted component of this work in other works.}

\begin{document}
\maketitle

%\includepdf[trim=0 2cm 0 2cm,clip=true]{../build/pullreqs.pdf}
\includepdf[scale=0.9,clip=true]{msr_journal.pdf}

\makebackpage
\end{document}

### 2016-07-21 SPRINGER_JOURNAL/techreport/TUD-SERG-2016-018.tex
Add DOI
http://dx.doi.org/10.1007/s10664-016-9443-7

@@
 
-\note{This paper has been accepted for publication in \emph{Empirical Software Engineering}, 2016.\\
-It is a substantially revised and extended version of 
+\note{This paper has been accepted for publication in \emph{Empirical Software Engineering}, 2016, where it will have DOI \url{http://dx.doi.org/10.1007/s10664-016-9443-7}.
+\\[1em]
+This paper is a substantially revised and extended version of 
 Roberta Coelho, Lucas Almeida, Georgios Gousios, Arie van Deursen:
